### 2013-07-17 GodProof.tex
Transfer all files to new repository

\documentclass{article}

\usepackage{latexsym}
\usepackage{bussproofs}
\EnableBpAbbreviations
\newcommand{\rl}[1]{\RightLabel{#1}}

% Logical symbols
\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\all}{\forall}
\newcommand{\ex}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

\author{Bruno Woltzenlogel Paleo}

\title{G\"{o}del's Ontological Proof of God's Existence }

\begin{document}

\maketitle

\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}

\noindent
``There is a scientific (exact) philosophy and theology, 
which deals with concepts of the highest abstractness; and this is also most highly fruitful for science. [\ldots] Religions are, for the most part, bad; but religion is not.'' - Kurt G\"{o}del

\section{Possible witnessing of positive properties}

\textbf{Axioms:}
\begin{itemize}
\item \textbf{(1)} Properties necessarily entailed by \emph{positive} properties are also positive: 
$$
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
$$
%
\item \textbf{(2)} A property's negation is positive iff the property is not positive: 
$$ 
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
$$
\end{itemize}

\noindent
\textbf{Theorem 1:} Positive properties possibly have a witness:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$

\noindent
\textbf{Formal proof:}

\begin{prooftree}
        \AXC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$}
        \UIC{$ \all \psi.[(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \psi(x)]) \imp P(\psi)]$}
        \UIC{$(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \neg \varphi'(x)]) \imp P(\neg \varphi')$} \doubleLine
        \UIC{$(P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp P(\neg \varphi')$}
                        \AXC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$}
                        \UIC{$ P(\neg \varphi') \biimp \neg P(\varphi') $} \doubleLine
                 \BIC{$ (P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp \neg P(\varphi') $} \doubleLine
                 \UIC{$ P(\varphi') \imp \pos \ex x.\varphi'(x) $}
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
\end{prooftree}

\section{Possible existence of a God}

\textbf{Axioms:}
\begin{itemize}
\item \textbf{(3)} Being God is a positive property: 
$$
P(G)
$$
\end{itemize}

\noindent
\textbf{Theorem 2:} It is possible that a God exists:
$$
\pos \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

\begin{prooftree}
\AXC{$P(G)$}
                 \AXC{$ $} \dashedLine \RightLabel{Th. 1}
                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$}
                 \UIC{$ P(G) \imp \pos \ex x.G(x) $}
    \BIC{$\pos \ex x. G(x)$}
\end{prooftree}

\section{Essentiality of being God}

\textbf{Definitions:}
\begin{itemize}
\item A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))$
\end{itemize}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(4)} Positive properties are necessarily positive: 
$$
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
$$
\end{itemize}

\noindent
\textbf{Theorem 3}: If an individual is a God, then being God is an essential property for that individual:
$$
\all y.[G(y) \imp \ess{G}{y}]
$$

\noindent
\textbf{Formal proof:}

\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$ G(x) $}
        \AXC{$ $} \RightLabel{2}
        \UIC{$\psi(x)$}
        \UIC{$G(x) \imp \psi(x)$}
        \UIC{$\all x.[G(x) \imp \psi(x)]$} \RightLabel{Necessitation}
        \UIC{$\nec \all x.[G(x) \imp \psi(x)]$} \RightLabel{2}
        \UIC{$\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]$}
        \UIC{$\all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$}
    \BIC{$G(x) \wedge \all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$} \dottedLine
    \UIC{$\ess{G}{x}$} \RightLabel{1}
    \UIC{$G(x) \imp \ess{G}{x}$}
    \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
\end{prooftree}

\noindent
\textbf{Note:} the formal proof above uses the necessitation rule of the basic modal logic \textbf{K}, instead of using Axiom 4, which corresponds to a restricted form of necessitation. I currently do not know how to formally prove theorem 3 using axiom 4 and not relying on necessitation.

\section{Necessity of God's existence}

\textbf{Definitions:}
\begin{itemize}
\item An individual is a \emph{God} if and only if he possesses all positive properties: 
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\item An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed: 
$$ 
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
$$
\end{itemize}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(5)} Necessary existence is a positive property: 
$$
P(E)
$$
\end{itemize}

\noindent
\textbf{Theorem A:} If there is a God, then there necessarily exists a God:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\ex z. G(z)$}
\UIC{$G(g)$}
\end{prooftree}

\begin{prooftree}
\AXC{$ $} \dashedLine
\UIC{$G(g)$}
        \AXC{$ $} \dashedLine \RightLabel{Th. 3}
        \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
        \UIC{$G(g) \imp \ess{G}{g}$}
    \BIC{$\ess{G}{g}$}
                \AXC{$P(E)$}
                        \AXC{$ $} \dashedLine
                        \UIC{$G(g)$} \dottedLine
                        \UIC{$\all \varphi.[P(\varphi) \imp \varphi(g)]$}
                        \UIC{$P(E) \imp E(g)$}
                     \BIC{$E(g)$} \dottedLine
                     \UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}
                     \UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}
        \BIC{$\nec \ex x. G(x)$} \RightLabel{1}
        \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}

\noindent
\textbf{Note:} Theorem A could be proved more quickly using the necessitation rule. Interestingly, the proof above shows that, by using the given axioms and definitions of god and necessary existence, theorem A can be derived even without the necessitation rule.

\section{Necessary existence of a God}

\noindent
\textbf{Theorem 4:} The existence of a God is necessary:
$$
\nec \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

\begin{small}
\begin{prooftree}
\AXC{$ $} \dashedLine \RightLabel{\textbf{S5}}
\UIC{$ \all \varphi. [\pos \ldots \pos \nec \varphi \biimp \nec \varphi] $}
\UIC{$ \pos \nec  \ex x. G(x) \biimp \nec \ex x. G(x) $}
        \AXC{$ $} \dashedLine  \RightLabel{Th. 2}
        \UIC{$\pos \ex x. G(x)$}
                \AXC{$ $} \dashedLine \RightLabel{Th. A}
                \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
            \BIC{$\pos \nec \ex x. G(x)$}
    \BIC{$\nec \ex x. G(x)$}
\end{prooftree}
\end{small}

\noindent
\textbf{Note:} The proof above relies on a theorem of the modal logic \textbf{S5}, which is a quite strong modal logic. It would be interesting to try to derive theorem 4 with weaker modal axioms.

\section{God's existence}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(M)} What is necessary is the case:
$$
\all \varphi. [\nec \varphi \imp \varphi]
$$
\end{itemize}

\noindent
\textbf{Theorem:} There exists a God:
$$
\ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

\begin{prooftree}
\AXC{$ $} \dashedLine \RightLabel{Th. 4}
\UIC{$ \nec \ex x. G(x)$}
        \AXC{$ \all \varphi. [\nec \varphi \imp \varphi]$}
        \UIC{$\nec \ex x. G(x) \imp \ex x. G(x)$}
    \BIC{$\ex x. G(x)$}
\end{prooftree}

\end{document}

### 2013-07-17 TPTP THF Formalization/call_tptp.sh

@@ -1,0 +1,28 @@
+tptp4X=./tptp4X
+call=./RemoteSOT.pl
+# if the above programs don't work for you download the SystemOnTPTP package,
+# install it and change the above paths accordingly
+
+proverlist='LEO-II---1.6.0 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 cocATP---0.1.8'
+timeout=20
+
+echo
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+echo
+
+for file in "$@"
+do
+    for prover in $proverlist 
+    do
+    function_to_fork () { 
+        echo $prover : $file ++++++ `$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover | grep RESULT` 
+    }
+    function_to_fork &
+    done
+    echo 
+done
+
+
+
+
+

### 2013-07-17 TPTP THF Formalization/call_tptp_detail.sh

@@ -1,0 +1,28 @@
+tptp4X=./tptp4X
+call=./RemoteSOT.pl
+# if the above programs don't work for you download the SystemOnTPTP package,
+# install it and change the above paths accordingly
+
+proverlist='LEO-II---1.6.0 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 cocATP---0.1.8'
+timeout=20
+
+echo
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+echo
+
+for file in "$@"
+do
+    for prover in $proverlist 
+    do
+    function_to_fork () { 
+        echo $prover : $file ++++++ `$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover -q01`
+    }
+    function_to_fork &
+    done
+    echo 
+done
+
+
+
+
+

### 2013-08-15 Talks/FU-Berlin/GödelGod-Talk.tex
Add stub for slides of talk to be given at FU-Berlin

\documentclass[9pt]{beamer}
\mode<presentation>
{
  \usetheme{Warsaw}
  \usecolortheme{crane}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{verbatim}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{pxfonts}
\usepackage{amsthm}
\usepackage{graphicx}

\usepackage{bussproofs}
\usepackage{proof}

\usepackage{fancybox}
\usepackage{fancyvrb}

% Sequent Calculus Proof Settings
\EnableBpAbbreviations
\def\fCenter{\mbox{\ $\vdash$\ }}

\newcommand{\mypause}{\pause}
%\newcommand{\mypause}{}

\title[Formalizations of G\"{o}del's Proof of God's Existence]{Formalizations of G\"{o}del's Proof of God's Existence}

\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}

\institute[] {
  \inst{}%
}

\date[29.10.13]
{2013 October ??}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\end{document}

### 2013-08-15 Abstract/GodProof-Abstract.tex
Add abstract draft

\documentclass{article}

\usepackage{latexsym}
\usepackage{bussproofs}
\EnableBpAbbreviations
\newcommand{\rl}[1]{\RightLabel{#1}}

% Logical symbols
\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\all}{\forall}
\newcommand{\ex}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}

\title{Formalizations of G\"{o}del's Ontological Proof of God's Existence}

\begin{document}

\maketitle

Das Logikgenie Kurt Gödel (1906-1978) hat vor seinem Tod ein
Argumentationsformalismus für den Beweis der Existenz Gottes
vorgeschlagen. Nun ist es Christoph Benzmüller, Wissenschaftler an der FU
Berlin im Dahlem Center for Intelligent Systems, in Kooperation mit
Bruno Woltzenlogel-Paleo von der TU Wien gelungen, Kurt Gödel's
Gottesbeweis auf dem Computer abzubilden, den Formalismus zu
verifizieren und zu automatisieren. Der Computer hat dann von Gödels
Prämissen ausgehend bewiesen: es gäbe Gott. "Theorembeweisen" heisst
die mathematische Disziplin bei der solche automatisierten Systemen
entwickelt werden.

Versuche sich der Existenz (oder Nichtexistenz) Gottes mithilfe
abstrakter, ontologischer Argumentationen zu nähern haben eine lange
Tradition in der Philosophie. Vor Gödel haben schon St. Anselm,
Descartes und Leibniz ihre jeweilige Beweise vorgelegt. Was Gödel als
Logiker antrieb war die Frage, ob sich ausgehend von einer kleinen
Anzahl (debattierbarer) grundlegender Axiome und Begriffsdefinitionen
die notwendige Existenz Gottes mithilfe einer mathematisch präzisen,
formal-logischen Argumentationskette nachweisen lässt.

In dem Computer wurde die Kernaussage, die notwendige Existenz Gottes,
wie von Gödel vorgeschlagen in vier Einzelargumente zerlegt und die
Beweisführung und Verifikation jedes dieser vier Einzelargumente wurde
dann mithilfe von Theorembeweisern automatisch erzeugt. Auch die
Widerspruchsfreiheit der grundlegenden Axiome und Definitionen konnte
mithilfe des Computers analysiert werden.

In der theoretischen Philosophie erfolgt die formal-logische
Auseinandersetzung mit solchen Gottesbeweisen bisher ausschließlich
mit Papier und Bleistift. Der Einsatz von Computern scheiterte bisher,
weil es den "Logiken" der verfügbaren Systeme noch an Ausdruckstärke
fehlte um die abstrakten Begriffe adäquat zu erfassen. Gödels
Gottesbeweis verwendet beispielsweise eine komplexe höherstufige
sogenannte Modallogik, die mit Begriffen wie "möglich" und
"notwendig" arbeitet.

Aktuelle Arbeiten von Christoph Benzmüller und Larry Paulson
(Cambridge University, UK) zeigen jedoch, dass sich viele
ausdrucksstarke Logiken, einschliesslich höherstufige Modallogiken, in
die klassische höherstufige Logik einbetten lassen. Letztere kann
also, mit einigen Einschränkungen, als eine universelle Logik
angesehen werden. Für die klassische höherstufige Logik wurden aber in
den vergangenen Jahren leistungsfähige
Theorembeweiser entwickelt und solche Systeme kommen nun in den
Arbeiten von Benzmüller und Woltzenlogel-Paleo zum Einsatz.

Die von den Wissenschaftlern gewählte Herangehensweise eröffnet damit
neue Perspektiven für eine computer-assistierte theoretische
Philosophie. Dabei bleibt die kritische Auseinandersetzung mit
grundlegenden Begriffen und Axiomen dem Menschen vorbehalten. Der
Computer kann aber helfen, Fehler in den formal-logischen Anteilen zu
minimieren: der Computer kann die Argumentationskette überprüfen und so
teilweise Leibniz Diktum bei logischen Streiterien einsetzen:
Calculemus --- Lasst es uns berechnen!

\end{document}

### 2013-08-16 Abstract/GodProof-Abstract.tex
Write first draft of one-page abstract.

@@
-\documentclass{article}
+\documentclass{llncs}
 
@@
 
-\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}
+\title{
+  Formalizations of G\"{o}del's Ontological Proof of \\
+  God's Existence
+  %\thanks{Supported by ... }
+}
 
-\title{Formalizations of G\"{o}del's Ontological Proof of God's Existence}
+\author{
+  Christoph Benzm\"{u}ller\inst{1} 
+  \and 
+  Bruno Woltzenlogel Paleo\inst{2}
+}
+
+\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}
+
+\institute{
+  Theory and Logic Group, Vienna University of Technology, Austria \\
+  \email{bruno@logic.at}
+  \and
+  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
+  \email{c.benzmueller@gmail.com}
+}
 
@@
 
-Das Logikgenie Kurt Gödel (1906-1978) hat vor seinem Tod ein
-Argumentationsformalismus für den Beweis der Existenz Gottes
-vorgeschlagen. Nun ist es Christoph Benzmüller, Wissenschaftler an der FU
-Berlin im Dahlem Center for Intelligent Systems, in Kooperation mit
-Bruno Woltzenlogel-Paleo von der TU Wien gelungen, Kurt Gödel's
-Gottesbeweis auf dem Computer abzubilden, den Formalismus zu
-verifizieren und zu automatisieren. Der Computer hat dann von Gödels
-Prämissen ausgehend bewiesen: es gäbe Gott. "Theorembeweisen" heisst
-die mathematische Disziplin bei der solche automatisierten Systemen
-entwickelt werden.
+Attempts to prove the existence (or non-existence) of God by means of 
+abstract ontological arguments are an old tradition in philosophy and theology,
+G\"{o}del's proof is a modern culmination of this tradition, following particularly 
+the footsteps of Leibniz.
+%
+G\"{o}del defines God as a being who possesses all \emph{positive} properties.
+He does not extensively discuss what positive properties are, 
+but instead he states a few reasonable but debatable axioms that they should satisfy.
+From these axioms, he claims (without detailed proof) that:
+\begin{enumerate}
+\item If it is possible that God exists, then it is necessary that God exists.
+\item It is possible that God exists.
+\end{enumerate}
 
-Versuche sich der Existenz (oder Nichtexistenz) Gottes mithilfe
-abstrakter, ontologischer Argumentationen zu nähern haben eine lange
-Tradition in der Philosophie. Vor Gödel haben schon St. Anselm,
-Descartes und Leibniz ihre jeweilige Beweise vorgelegt. Was Gödel als
-Logiker antrieb war die Frage, ob sich ausgehend von einer kleinen
-Anzahl (debattierbarer) grundlegender Axiome und Begriffsdefinitionen
-die notwendige Existenz Gottes mithilfe einer mathematisch präzisen,
-formal-logischen Argumentationskette nachweisen lässt.
+In the literature \cite{ToDo} slightly different versions of the axioms exist 
+and derivations of the above mentioned claims are presented with various degrees
+of detail and formality. Here \cite{ToDo: GitHub Repository} (a version of) G\"{o}del's proof is 
+constructed for the first-time
+with the utmost degree of detail and formality. The following has been done:
+\begin{itemize}
+\item A detailed natural deduction proof.
+%
+\item A formalization of the axioms and theorems in the TPTP THF format \cite{ToDo}.
+%
+\item Automatic verification of the consistency of the axioms and 
+definitions with ToDo \cite{ToDo}.
+%
+\item Automatic demonstration of the theorems with the provers LEO-II \cite{ToDo} and Satallax \cite{ToDo}.
 
-In dem Computer wurde die Kernaussage, die notwendige Existenz Gottes,
-wie von Gödel vorgeschlagen in vier Einzelargumente zerlegt und die
-Beweisführung und Verifikation jedes dieser vier Einzelargumente wurde
-dann mithilfe von Theorembeweisern automatisch erzeugt. Auch die
-Widerspruchsfreiheit der grundlegenden Axiome und Definitionen konnte
-mithilfe des Computers analysiert werden.
+\item A step-by-step formalization using the Coq proof assistant \cite{ToDo}.
 
-In der theoretischen Philosophie erfolgt die formal-logische
-Auseinandersetzung mit solchen Gottesbeweisen bisher ausschließlich
-mit Papier und Bleistift. Der Einsatz von Computern scheiterte bisher,
-weil es den "Logiken" der verfügbaren Systeme noch an Ausdruckstärke
-fehlte um die abstrakten Begriffe adäquat zu erfassen. Gödels
-Gottesbeweis verwendet beispielsweise eine komplexe höherstufige
-sogenannte Modallogik, die mit Begriffen wie "möglich" und
-"notwendig" arbeitet.
+\item A formalization using the Isabelle proof assistant \cite{ToDo} partially automated with Sledgehammer \cite{ToDo} and Metis \cite{ToDo}.
+\end{itemize}
 
-Aktuelle Arbeiten von Christoph Benzmüller und Larry Paulson
-(Cambridge University, UK) zeigen jedoch, dass sich viele
-ausdrucksstarke Logiken, einschliesslich höherstufige Modallogiken, in
-die klassische höherstufige Logik einbetten lassen. Letztere kann
-also, mit einigen Einschränkungen, als eine universelle Logik
-angesehen werden. Für die klassische höherstufige Logik wurden aber in
-den vergangenen Jahren leistungsfähige
-Theorembeweiser entwickelt und solche Systeme kommen nun in den
-Arbeiten von Benzmüller und Woltzenlogel-Paleo zum Einsatz.
+G\"{o}del's proof is challenging to formalize and automatically verify 
+because it requires very expressive logical languages with 
+modal operators (\emph{possible} and \emph{necessary}) and higher-order quantifiers. 
+The computer-assisted formalizations rely on an embedding of the modal logic S5 
+into classical higher-order logic with Henkin 
+semantics \cite{ToDo: Current works of Christoph Benzmüller and Larry Paulson} and 
+employed recently developed interactive and automated deduction tools designed for this logic.
 
-Die von den Wissenschaftlern gewählte Herangehensweise eröffnet damit
-neue Perspektiven für eine computer-assistierte theoretische
-Philosophie. Dabei bleibt die kritische Auseinandersetzung mit
-grundlegenden Begriffen und Axiomen dem Menschen vorbehalten. Der
-Computer kann aber helfen, Fehler in den formal-logischen Anteilen zu
-minimieren: der Computer kann die Argumentationskette überprüfen und so
-teilweise Leibniz Diktum bei logischen Streiterien einsetzen:
-Calculemus --- Lasst es uns berechnen!
-
+This work attests the maturity of such automated deduction tools and 
+opens new perspectives for a computer-assisted theoretical philosophy, 
+where the critical discussion of the underlying 
+concepts, definitions and axioms remains a human responsibility, but the computer can assist 
+in building and checking rigorously correct logical arguments. And in case of 
+logico-philosophical disputes, the computer can check the disputing arguments
+and partially fulfill Leibniz' dictum:
+Calculemus --- Let us calculate!
 \end{document}

### 2013-08-16 Abstract/GodProof-Abstract.tex
Improve abstract

@@
 
-This work attests the maturity of such automated deduction tools and 
-opens new perspectives for a computer-assisted theoretical philosophy, 
-where the critical discussion of the underlying 
+This work attests the maturity of contemporary interactive and 
+automated deduction tools and 
+opens new perspectives for a computer-assisted theoretical philosophy. 
+The critical discussion of the underlying 
 concepts, definitions and axioms remains a human responsibility, but the computer can assist 
-in building and checking rigorously correct logical arguments. And in case of 
+in building and checking rigorously correct logical arguments. In case of 
 logico-philosophical disputes, the computer can check the disputing arguments

### 2013-08-16 Abstract/GodProof-Abstract.tex
added Bibliography.bib and slightly modified abstract

@@
 \title{
-  Formalizations of G\"{o}del's Ontological Proof of \\
+  Formalization and Automation of Variants of G\"{o}del's Ontological Proof of 
   God's Existence
@@
 \institute{
+  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
+  \email{c.benzmueller@gmail.com}
+  \and 
   Theory and Logic Group, Vienna University of Technology, Austria \\
   \email{bruno@logic.at}
-  \and
-  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
 }
@@
 Attempts to prove the existence (or non-existence) of God by means of
-abstract ontological arguments are an old tradition in philosophy and theology,
-G\"{o}del's proof is a modern culmination of this tradition, following particularly 
-the footsteps of Leibniz.
+abstract ontological arguments are an old tradition in philosophy and
+theology.  G\"{o}del's proof \cite{Goedel} is a modern culmination of
+this tradition, following particularly the footsteps of Leibniz.
 %
@@
 but instead he states a few reasonable but debatable axioms that they should satisfy.
-From these axioms, he claims (without detailed proof) that:
-\begin{enumerate}
-\item If it is possible that God exists, then it is necessary that God exists.
-\item It is possible that God exists.
-\end{enumerate}
+%
+In our recent work we did employ the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property strictly implied by a positive property is positive,
+(2) A property is positive if and only if its negation is not positive, and (3) the property of being God-like is positive. Moreover, the following definitions are employed:
+(A) x is God-like if and only if x incorporates all positive properties, (B) a property P is essential for x (and essence of x) if and only if P is a property of x and every property Q that x has is strictly implied by P, (C) x necessarily exists if and only if every essence of x is necessarily exemplified.
 
-In the literature \cite{ToDo} slightly different versions of the axioms exist 
+From these axioms and definitions we then infer:
+\begin{itemize}
+\item Positive properties are eventually exemplified.
+\item Eventually God exists.
+\item If x is God-like, then the property of being 
+God-like is an essence of x.
+\item Necessarily God exists.
+\end{itemize}
+
+% From these axioms, he claims (without detailed proof) that:
+% \begin{enumerate}
+% \item If it is possible that God exists, then it is necessary that God exists.
+% \item It is possible that God exists.
+% \end{enumerate}
+
+In the literature \cite{ToDo} slightly different versions of the axioms and definitions exist 
 and derivations of the above mentioned claims are presented with various degrees
-of detail and formality. Here \cite{ToDo: GitHub Repository} (a version of) G\"{o}del's proof is 
+of detail and formality. The above variant of G\"{o}del's proof has now been 
 constructed for the first-time
-with the utmost degree of detail and formality. The following has been done:
+with the utmost degree of detail and formality; cf.~\cite{ToDo:GitHubRepository}. The following has been done:
 \begin{itemize}
@@
 %
-\item A formalization of the axioms and theorems in the TPTP THF format \cite{ToDo}.
+\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
 %
 \item Automatic verification of the consistency of the axioms and 
-definitions with ToDo \cite{ToDo}.
+definitions with Nitpick \cite{Nitpick}.
 %
-\item Automatic demonstration of the theorems with the provers LEO-II \cite{ToDo} and Satallax \cite{ToDo}.
+\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.
 
@@
 
-\item A formalization using the Isabelle proof assistant \cite{ToDo} partially automated with Sledgehammer \cite{ToDo} and Metis \cite{ToDo}.
+\item A formalization using the Isabelle proof assistant \cite{Isabelle} partially automated with Sledgehammer \cite{ToDo} and Metis \cite{ToDo}.
 \end{itemize}
 
-G\"{o}del's proof is challenging to formalize and automatically verify 
+G\"{o}del's proof is challenging to formalize and verify 
 because it requires very expressive logical languages with 
 modal operators (\emph{possible} and \emph{necessary}) and higher-order quantifiers. 
-The computer-assisted formalizations rely on an embedding of the modal logic S5 
+Our computer-assisted formalizations rely on an embedding of the modal logic S5 
 into classical higher-order logic with Henkin 
-semantics \cite{ToDo: Current works of Christoph Benzmüller and Larry Paulson} and 
+semantics \cite{J23} and 
 employed recently developed interactive and automated deduction tools designed for this logic.
@@
 This work attests the maturity of contemporary interactive and
-automated deduction tools and 
-opens new perspectives for a computer-assisted theoretical philosophy. 
-The critical discussion of the underlying 
-concepts, definitions and axioms remains a human responsibility, but the computer can assist 
-in building and checking rigorously correct logical arguments. In case of 
-logico-philosophical disputes, the computer can check the disputing arguments
-and partially fulfill Leibniz' dictum:
-Calculemus --- Let us calculate!
+automated deduction tools and opens new perspectives for a
+computer-assisted theoretical philosophy.  
+The critical discussion of the underlying concepts, definitions and
+axioms remains a human responsibility, but the computer can assist in
+building and checking rigorously correct logical arguments. In case of
+logico-philosophical disputes, the computer can check the disputing
+arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
+calculate!
+
+Future work includes an
+extensive study of other formalizations of ontological arguments with
+our machinery.
+
+\bibliographystyle{plain}
+\bibliography{Bibliography}
+
 \end{document}

### 2013-08-16 Abstract/GodProof-Abstract.tex
Improve abstract

@@
 but instead he states a few reasonable but debatable axioms that they should satisfy.
-%
-In our recent work we did employ the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property strictly implied by a positive property is positive,
-(2) A property is positive if and only if its negation is not positive, and (3) the property of being God-like is positive. Moreover, the following definitions are employed:
-(A) x is God-like if and only if x incorporates all positive properties, (B) a property P is essential for x (and essence of x) if and only if P is a property of x and every property Q that x has is strictly implied by P, (C) x necessarily exists if and only if every essence of x is necessarily exemplified.
+Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof.
+Our formalization employs the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property necessarily implied by a positive property is positive;
+(2) A property is positive if and only if its negation is not positive; (3) the property of being God-like is positive; (4) positive properties are necessarily positive; and (5) necessary existence is a positive property.
+Moreover, the following definitions are employed:
+(D1) x is God-like if and only if x has every positive property, (D2) a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, (D3) x necessarily exists if and only if every essence of x is necessarily exemplified.
 
 From these axioms and definitions we then infer:
-\begin{itemize}
-\item Positive properties are eventually exemplified.
-\item Eventually God exists.
+\begin{enumerate}
+\item Positive properties are possibly exemplified.
+\item Possibly God exists.
 \item If x is God-like, then the property of being 
@@
 \item Necessarily God exists.
-\end{itemize}
+\end{enumerate}
 
@@
 
-In the literature \cite{ToDo} slightly different versions of the axioms and definitions exist 
-and derivations of the above mentioned claims are presented with various degrees
-of detail and formality. The above variant of G\"{o}del's proof has now been 
+The above variant of G\"{o}del's proof has now been 
 constructed for the first-time
@@
 because it requires very expressive logical languages with 
-modal operators (\emph{possible} and \emph{necessary}) and higher-order quantifiers. 
+modal operators (\emph{possibily} and \emph{necessarily}) and higher-order quantifiers. 
 Our computer-assisted formalizations rely on an embedding of the modal logic S5 

### 2013-08-16 Abstract/GodProof-Abstract.tex
Number the lemmas and theorems
some further modifications

@@
 From these axioms and definitions we then infer:
-\begin{enumerate}
-\item Positive properties are possibly exemplified.
-\item Possibly God exists.
-\item If x is God-like, then the property of being 
+
+\begin{description}
+\item {(L1)} Positive properties are possibly exemplified.
+\item {(L2)} Possibly God exists.
+\item {(L3)} If x is God-like, then the property of being 
 God-like is an essence of x.
-\item Necessarily God exists.
-\end{enumerate}
+\item {(T)} \ Necessarily God exists.
+\end{description}
 
@@
 
-\item A formalization using the Isabelle proof assistant \cite{Isabelle} partially automated with Sledgehammer \cite{ToDo} and Metis \cite{ToDo}.
+\item A formalization using the Isabelle proof assistant \cite{Isabelle} partially automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
 \end{itemize}
@@
 into classical higher-order logic with Henkin 
-semantics \cite{J23} and 
+semantics \cite{J23,B9} and 
 employed recently developed interactive and automated deduction tools designed for this logic.
@@
 
-Future work includes an
-extensive study of other formalizations of ontological arguments with
-our machinery.
+Future work includes an extensive study of other formalizations of
+ontological arguments with our machinery. Variations of the
+distinctive features of the base logic S5 (e.g. non-rigid symbols, 
+varying domains, etc.) are enabled in these studies due
+to the flexibility of the embeddings based approach.
 

### 2013-08-17 Abstract/GodProof-Abstract.tex
Add bibliographical reference for Coq
Add citations and compare axioms and definitions with Gödel's originals

@@
 
-\usepackage{latexsym}
-\usepackage{bussproofs}
-\EnableBpAbbreviations
-\newcommand{\rl}[1]{\RightLabel{#1}}
+\usepackage{url}
 
@@
 abstract ontological arguments are an old tradition in philosophy and
-theology.  G\"{o}del's proof \cite{Goedel} is a modern culmination of
+theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
 this tradition, following particularly the footsteps of Leibniz.
@@
 but instead he states a few reasonable but debatable axioms that they should satisfy.
-Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof.
-Our formalization employs the following axioms\footnote{Say where they come from and how they relate to G\"odel.}: (1) Any property necessarily implied by a positive property is positive;
-(2) A property is positive if and only if its negation is not positive; (3) the property of being God-like is positive; (4) positive properties are necessarily positive; and (5) necessary existence is a positive property.
-Moreover, the following definitions are employed:
-(D1) x is God-like if and only if x has every positive property, (D2) a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, (D3) x necessarily exists if and only if every essence of x is necessarily exemplified.
+Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g.
+\cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+Our formalization employs the following axioms and definitions\footnote{
+(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}.
+(A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction.
+(A4) is a weaker form of G\"odel's axiom 3.
+(D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
+}: 
+ToDo: Include formulas too??
+\begin{description}
+\item {(A1)} Any property necessarily implied by a positive property is positive;
+\item {(A2)} A property is positive if and only if its negation is not positive; 
+\item {(A3)} the property of being God-like is positive; 
+\item {(A4)} positive properties are necessarily positive; 
+\item {(A5)} necessary existence is a positive property.
+\item {(D1)} x is God-like if and only if x has every positive property, 
+\item {(D2)} a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
+\item {(D3)} x necessarily exists if and only if every essence of x is necessarily exemplified.
+\end{description}
 
+\noindent
 From these axioms and definitions we then infer:
-
+%
+ToDo: Include formulas too??
 \begin{description}
@@
 constructed for the first-time
-with the utmost degree of detail and formality; cf.~\cite{ToDo:GitHubRepository}. The following has been done:
+with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done:
 \begin{itemize}
@@
 
-\item A step-by-step formalization using the Coq proof assistant \cite{ToDo}.
+\item A step-by-step formalization using the Coq proof assistant \cite{Coq}.
 

### 2013-08-17 Abstract/GodProof-Abstract.tex
achieved full automation with metis in Isabelle; some lemmata were introduced for this

@@
 
-\usepackage{url}
+\usepackage{url,amsmath,amssymb,a4wide}
 
@@
 \title{
-  Formalization and Automation of Variants of G\"{o}del's Ontological Proof of 
-  God's Existence
-  %\thanks{Supported by ... }
+  Formalization, Mechanization and Automation of Variants of G\"{o}del's Proof of 
+  God's Existence\thanks{This work has been supported by the German Research Foundation under grant BE2501/9-1. Bruno ????}
 }
@@
 (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
-}: 
-ToDo: Include formulas too??
+}.
+
+\begin{align}
+
+\emd{align}
+
 \begin{description}
-\item {(A1)} Any property necessarily implied by a positive property is positive;
-\item {(A2)} A property is positive if and only if its negation is not positive; 
-\item {(A3)} the property of being God-like is positive; 
-\item {(A4)} positive properties are necessarily positive; 
-\item {(A5)} necessary existence is a positive property.
-\item {(D1)} x is God-like if and only if x has every positive property, 
-\item {(D2)} a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
-\item {(D3)} x necessarily exists if and only if every essence of x is necessarily exemplified.
+\item[(A1)] Any property necessarily implied by a positive property is positive: \\
+       $\forall P \forall Q (positive P \wedge \Box \forall x (P x \Rightarrow Q x)) \Rightarrow positive Q)$
+\item[(A2)] A property is positive if and only if its negation is not positive\\
+       $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
+\item[(A3)] the property of being God-like is positive; 
+\item[(A4)] positive properties are necessarily positive; 
+\item[(A5)] necessary existence is a positive property.
+\item[(D1)] x is God-like if and only if x has every positive property, 
+\item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
+\item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
 \end{description}
@@
 \begin{description}
-\item {(L1)} Positive properties are possibly exemplified.
-\item {(L2)} Possibly God exists.
-\item {(L3)} If x is God-like, then the property of being 
+\item[(L1)] Positive properties are possibly exemplified.
+\item[(L2)] Possibly God exists.
+\item[(L3)] If x is God-like, then the property of being 
 God-like is an essence of x.
-\item {(T)} \ Necessarily God exists.
+\item[(T)] \ Necessarily God exists.
 \end{description}
@@
 \begin{itemize}
-\item A detailed natural deduction proof.
+\item A detailed natural deduction proof; this proof has guided our work.
 %
-\item A formalization of the axioms and theorems in the TPTP THF format \cite{J22}.
+\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
 %
@@
 
-\item A formalization using the Isabelle proof assistant \cite{Isabelle} partially automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
+\item A formalization using the Isabelle proof assistant \cite{Isabelle} where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
 \end{itemize}

### 2013-08-18 Abstract/GodProof-Abstract.tex
syntax improvement

@@
 \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-Our formalization employs the following axioms and definitions\footnote{
+Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
 (A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}.
@@
 (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
-}.
+}
 
 \begin{align}
-
-\emd{align}
+& \text{Any property necessarily implied by a positive property is positive:} \notag \\
+& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
+& \text{A property is positive if and only if its negation is not positive:} \notag \\
+& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
+& \text{The property of being God-like is positive:} \notag \\
+& \quad pos\,god \tag{A3} \\
+& \text{Positive properties are necessarily positive:} \notag \\
+& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
+& \text{Necessary existence is a positive property:} \notag \\
+& \quad pos\,nec\_exists \tag{A5} \\
+& \text{x is God-like if and only if x has every positive property:} \notag \\
+& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
+& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2} 
+\end{align}
 
@@
 \item[(A1)] Any property necessarily implied by a positive property is positive: \\
-       $\forall P \forall Q (positive P \wedge \Box \forall x (P x \Rightarrow Q x)) \Rightarrow positive Q)$
-\item[(A2)] A property is positive if and only if its negation is not positive\\
+       $\forall P \forall Q ((positive P \wedge (\Box \forall x (P x \Rightarrow Q x))) \Rightarrow positive Q)$
+\item[(A2)] A property is positive if and only if its negation is not positive:\\
        $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
-\item[(A3)] the property of being God-like is positive; 
+\item[(A3)] The property of being God-like is positive:
+
 \item[(A4)] positive properties are necessarily positive; 
@@
 constructed for the first-time
-with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done:
+with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done (and in this order):
 \begin{itemize}

### 2013-08-18 Natural Deduction/GodProof-ND.tex
Remove notes from draft
They were mostly based on an incorrect notion of Necessitation

--- 2013-07-17 GodProof.tex
@@
 
-\author{Bruno Woltzenlogel Paleo}
+\author{Bruno Woltzenlogel Paleo, Annika Siders}
 
@@
 
+Let the following derivation with the open assumption $G(x)$ be $\Pi_1$:
+
 \begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$ G(x) $}
-        \AXC{$ $} \RightLabel{2}
-        \UIC{$\psi(x)$}
-        \UIC{$G(x) \imp \psi(x)$}
-        \UIC{$\all x.[G(x) \imp \psi(x)]$} \RightLabel{Necessitation}
-        \UIC{$\nec \all x.[G(x) \imp \psi(x)]$} \RightLabel{2}
-        \UIC{$\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]$}
-        \UIC{$\all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$}
-    \BIC{$G(x) \wedge \all \psi.[\psi(x) \imp \nec \all x.[G(x) \imp \psi(x)]]$} \dottedLine
-    \UIC{$\ess{G}{x}$} \RightLabel{1}
-    \UIC{$G(x) \imp \ess{G}{x}$}
+\AXC{$ \neg P(\psi)^1$} 
+       \AXC{$ $} \RightLabel{Ax. 2}
+       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall$E}  
+       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp$E}   
+ \BIC{$P(\neg\psi)$}\RightLabel{} 
+           	\AXC{$G(x)$} \dottedLine\RightLabel{Definition of G}
+		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
+              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp$ E}
+           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp$E}  
+            \AXC{$ \psi(x)^2$} \RightLabel{$\imp$E}  
+             \BIC{$\bot $}\RightLabel{RAA, 1}  
+              \UIC{$ P(\psi)$}\RightLabel{$\imp$ I, 2}
+              \UIC{$ \psi(x)\imp P(\psi)$}
+\end{prooftree}
+
+Let the following derivation with the open assumption $G(x)$ be $\Pi_2$: 
+\begin{prooftree}
+       \AXC{$\psi(x)^1$} \RightLabel{}
+             	\AXC{$\Pi_1 $}  \dashedLine
+		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp$E}
+       		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
+		\AXC{$$} \RightLabel{Ax. 4}
+       		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
+		\UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp$E}  
+          \BIC{$\Box P(\psi)$}\RightLabel{$\imp$I, 1} 
+           \UIC{$\psi(x)\imp\Box P(\psi)$} 
+\end{prooftree}
+
+Let the following derivation without open assumptions be $\Pi_3$:
+
+\begin{prooftree}
+       \AXC{$P(\psi)^1$} \RightLabel{}
+             	\AXC{$G(x)^2$} \dottedLine\RightLabel{Definition of G}
+		\UIC{$\forall\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
+		\UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp$E}  
+       		\BIC{$\psi(x)$}\RightLabel{$\imp$I, 2}  
+       		\UIC{$G(x)\imp \psi(x)$}\RightLabel{$\forall$I} 
+		\UIC{$\forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$I, 1}  
+		\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}
+\end{prooftree}
+
+Let the following derivation with the open assumption $G(x)$ be $\Pi_4$: 
+\begin{prooftree}
+       \AXC{$\psi(x)^1 $} \RightLabel{}
+             	\AXC{$\Pi_2$}  \dashedLine
+		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp$E}
+		\BIC{$\Box P(\psi)$}
+		  	\AXC{$\Box P(\psi)^2 $}\RightLabel{$\Box$E}  
+       			\UIC{$P(\psi)$}
+				\AXC{$\Pi_3$}  \dashedLine\RightLabel{}
+       				\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\imp$E}  
+       			\BIC{$\forall x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
+			\UIC{$\Box \forall x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp$I, 2} 
+			\UIC{$\Box P(\psi)\imp \Box \forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$E} 
+         	 \BIC{$\Box \forall x .(G(x)\imp \psi(x)$}\RightLabel{$\imp$I, 1}
+           \UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}  
+\end{prooftree}
+
+The use of the necessitation rule is valid, because the only open assumption $\Box P(\psi)$ is boxed.
+
+We construct a derivation of theorem 3 with a subderivation $\Pi_4[G(x)^1]$, which means that the open assumption $G(x)$ in $\Pi_4$ is discharged with the rule labeled 1.
+
+\begin{prooftree}
+       \AXC{$G(x)^1 $} \RightLabel{}
+             	\AXC{$\Pi_4[G(x)^1]$}  \dashedLine
+		\UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\forall $I}
+		\UIC{$\forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\RightLabel{$\&$I}
+		\BIC{$G(x)\& \forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{Definition of ess} 
+       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp$I, 1}
+			\UIC{$G(x)\imp \ess{G}{x}$}
 			  \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
 \end{prooftree}
-
-\noindent
-\textbf{Note:} the formal proof above uses the necessitation rule of the basic modal logic \textbf{K}, instead of using Axiom 4, which corresponds to a restricted form of necessitation.
-I currently do not know how to formally prove theorem 3 using axiom 4 and not relying on necessitation.
 
@@
 
-\noindent
-\textbf{Note:} Theorem A could be proved more quickly using the necessitation rule.
-Interestingly, the proof above shows that, by using the given axioms and definitions of god and necessary existence, theorem A can be derived even without the necessitation rule.
-
 \section{Necessary existence of a God}
@@
 \end{small}
-
-\noindent
-\textbf{Note:} The proof above relies on a theorem of the modal logic \textbf{S5}, which is a quite strong modal logic. It would be interesting to try to derive theorem 4 with weaker modal axioms.
 

### 2013-08-18 Abstract/GodProof-Abstract.tex
Make the numbering of axioms, definitions and lemmas agree between abstract and natural deduction formalization

@@
 & \quad pos\,nec\_exists \tag{A5} \\
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
+& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D1} \\
 & \text{x is God-like if and only if x has every positive property:} \notag \\
-& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
-& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2} 
+& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D2}
 \end{align}
@@
 \item[(A5)] necessary existence is a positive property.
-\item[(D1)] x is God-like if and only if x has every positive property, 
-\item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
+\item[(D1)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
+\item[(D2)] x is God-like if and only if x has every positive property, 
 \item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.

### 2013-08-18 Natural Deduction/GodProof-ND.tex
Make the numbering of axioms, definitions and lemmas agree between abstract and natural deduction formalization

@@
 
-\title{G\"{o}del's Ontological Proof of God's Existence }
+\title{G\"{o}del's Ontological Proof of God's Existence (Draft)}
 
@@
 \begin{itemize}
-\item \textbf{(1)} Properties necessarily entailed by \emph{positive} properties are also positive:
+\item \textbf{(A1)} Properties necessarily entailed by \emph{positive} properties are also positive:
 $$
@@
 %
-\item \textbf{(2)} A property's negation is positive iff the property is not positive:
+\item \textbf{(A2)} A property's negation is positive iff the property is not positive:
 $$
@@
 \noindent
-\textbf{Theorem 1:} Positive properties possibly have a witness:
+\textbf{Lemma 1:} Positive properties possibly have a witness:
 $$
@@
 \begin{itemize}
-\item \textbf{(3)} Being God is a positive property:
+\item \textbf{(A3)} Being God is a positive property:
 $$
@@
 \noindent
-\textbf{Theorem 2:} It is possible that a God exists:
+\textbf{Lemma 2:} It is possible that a God exists:
 $$
@@
 \begin{itemize}
-\item A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
+\item \textbf{(D1)} A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
 (\psi(x) \imp \nec \all x.
@@
 \begin{itemize}
-\item \textbf{(4)} Positive properties are necessarily positive:
+\item \textbf{(A4)} Positive properties are necessarily positive:
 $$
@@
 \noindent
-\textbf{Theorem 3}: If an individual is a God, then being God is an essential property for that individual:
+\textbf{Lemma 3}: If an individual is a God, then being God is an essential property for that individual:
 $$
@@
 \begin{itemize}
-\item An individual is a \emph{God} if and only if he possesses all positive properties:
+\item \textbf{(D2)} An individual is a \emph{God} if and only if he possesses all positive properties:
 $$
@@
 $$
-\item An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed:
+\item \textbf{(D3)} An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed:
 $$
@@
 \begin{itemize}
-\item \textbf{(5)} Necessary existence is a positive property:
+\item \textbf{(A5)} Necessary existence is a positive property:
 $$
@@
 \noindent
-\textbf{Theorem A:} If there is a God, then there necessarily exists a God:
+\textbf{Auxiliary Lemma:} If there is a God, then there necessarily exists a God:
 $$
@@
 \noindent
-\textbf{Theorem 4:} The existence of a God is necessary:
+\textbf{Theorem:} The existence of a God is necessary:
 $$
@@
 \noindent
-\textbf{Theorem:} There exists a God:
+\textbf{Corollary:} There exists a God:
 $$

### 2013-08-18 Abstract/GodProof-Abstract.tex
improvents on the text and presentation of the abstract

--- 2013-08-17 Abstract/GodProof-Abstract.tex
@@
 
-\usepackage{url,amsmath,amssymb,a4wide}
+\usepackage{url,amsmath,amssymb}
 
@@
 
-\title{
-  Formalization, Mechanization and Automation of Variants of G\"{o}del's Proof of 
-  God's Existence\thanks{This work has been supported by the German Research Foundation under grant BE2501/9-1. Bruno ????}
-}
+\title{ Formalization, Mechanization and Automation % of Variants
+  of G\"{o}del's Proof of God's Existence\thanks{This work has been
+    supported by the German Research Foundation under grant
+    BE2501/9-1. Bruno ????}  }
 
@@
 (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
-}.
+}
+\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this sources.}
+\allowdisplaybreaks[1] 
+\begin{align}
+& \text{Any property necessarily implied by a positive property is positive.} \notag \\
+& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
+& \text{A property is positive if and only if its negation is not positive.} \notag \\
+& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
+& \text{The property of being God-like is positive.} \notag \\
+& \quad pos\,god \tag{A3} \\
+& \text{Positive properties are necessarily positive.} \notag \\
+& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
+& \text{Necessary existence is a positive property.} \notag \\
+& \quad pos\,nec\_exists \tag{A5} 
+\end{align}
+\begin{align}
+& \text{x is God-like if and only if x has every positive property.} \notag \\
+& \quad god\,x := \forall P (pos\,P \Rightarrow P\,x) \tag{D1} \\ 
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
+& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box \forall y (P\,y \Rightarrow Q\,y)) \tag{D2} \\
+& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
+& \quad nec\_exists\,x := \forall P (ess\,P\,x \Rightarrow \Box \exists y (P\,y)) \tag{D3} 
+\end{align}
 
-\begin{align}
+% \begin{description}
+% \item[(A1)] Any property necessarily implied by a positive property is positive: \\
+%        $\forall P \forall Q ((positive P \wedge (\Box \forall x (P x \Rightarrow Q x))) \Rightarrow positive Q)$
+% \item[(A2)] A property is positive if and only if its negation is not positive:\\
+%        $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
+% \item[(A3)] The property of being God-like is positive:
 
-\emd{align}
-
-\begin{description}
-\item[(A1)] Any property necessarily implied by a positive property is positive: \\
-       $\forall P \forall Q (positive P \wedge \Box \forall x (P x \Rightarrow Q x)) \Rightarrow positive Q)$
-\item[(A2)] A property is positive if and only if its negation is not positive\\
-       $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
-\item[(A3)] the property of being God-like is positive; 
-\item[(A4)] positive properties are necessarily positive; 
-\item[(A5)] necessary existence is a positive property.
-\item[(D1)] x is God-like if and only if x has every positive property, 
-\item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
-\item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
-\end{description}
+% \item[(A4)] positive properties are necessarily positive; 
+% \item[(A5)] necessary existence is a positive property.
+% \item[(D1)] x is God-like if and only if x has every positive property, 
+% \item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
+% \item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
+% \end{description}
 
@@
 From these axioms and definitions we then infer:
+\begin{align}
+& \text{Positive properties are possibly exemplified.} \notag \\
+& \quad \forall P (pos\,P \Rightarrow \Diamond \exists x (P\,x)) \tag{L1} \\
+& \text{Possibly God exists.} \notag \\
+& \quad \Diamond \exists x (god\,x) \tag{L2} \\
+& \parbox[c][2em][c]{0.8\textwidth}{If x is God-like, then the property of being God-like is an essence of x.} \notag \\
+& \quad \forall x (god\,x \Rightarrow ess\,god\,x) \tag{L3} \\
+& \text{Necessarily God exists.} \notag \\
+& \quad \Box \exists x (god\,x) \tag{T}
+\end{align}
+
 %
-ToDo: Include formulas too??
-\begin{description}
-\item[(L1)] Positive properties are possibly exemplified.
-\item[(L2)] Possibly God exists.
-\item[(L3)] If x is God-like, then the property of being 
-God-like is an essence of x.
-\item[(T)] \ Necessarily God exists.
-\end{description}
+%ToDo: Include formulas too??
+% \begin{description}
+% \item[(L1)] Positive properties are possibly exemplified.
+% \item[(L2)] Possibly God exists.
+% \item[(L3)] If x is God-like, then the property of being 
+% God-like is an essence of x.
+% \item[(T)] \ Necessarily God exists.
+% \end{description}
 
@@
 constructed for the first-time
-with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done:
+with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done (and in this order):
 \begin{itemize}
-\item A detailed natural deduction proof; this proof has guided our work.
+\item A detailed natural deduction proof.\marginpar{\scriptsize Chris: We need to make sure that this ND proof is based on a sound calculus.}
 %
@@
 
-G\"{o}del's proof is challenging to formalize and verify 
-because it requires very expressive logical languages with 
-modal operators (\emph{possibily} and \emph{necessarily}) and higher-order quantifiers. 
-Our computer-assisted formalizations rely on an embedding of the modal logic S5 
-into classical higher-order logic with Henkin 
-semantics \cite{J23,B9} and 
-employed recently developed interactive and automated deduction tools designed for this logic.
+G\"{o}del's proof is challenging to formalize and verify because it
+requires an expressive logical language with modal operators
+(\emph{possibily} and \emph{necessarily}) and with
+quantififiers for individuals and sets of individuals (properties).  Our computer-assisted formalizations rely on an
+embedding of the modal logic S5 into classical higher-order logic with
+Henkin semantics \cite{J23,B9}. The formalization is thus essentially
+done in classical higher-order logic where quantified S5 is emulated.
 
 This work attests the maturity of contemporary interactive and
-automated deduction tools and opens new perspectives for a
-computer-assisted theoretical philosophy.  
-The critical discussion of the underlying concepts, definitions and
-axioms remains a human responsibility, but the computer can assist in
-building and checking rigorously correct logical arguments. In case of
-logico-philosophical disputes, the computer can check the disputing
-arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
-calculate!
+automated deduction tools for classical higher-order logic and it
+demonstrates the elegance and practical relevance of the embeddings
+based approach.  Most importantly, our work opens new perspectives for
+a computer-assisted theoretical philosophy.  The critical discussion
+of the underlying concepts, definitions and axioms remains a human
+responsibility, but the computer can assist in building and checking
+rigorously correct logical arguments. In case of logico-philosophical
+disputes, the computer can check the disputing arguments and partially
+fulfill Leibniz' dictum: Calculemus --- Let us calculate!
 

### 2013-08-18 Natural Deduction/GodProof-ND.tex
Make definition numbers agree in Coq, ND and abstract

@@
 \begin{itemize}
-\item \textbf{(D1)} A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
+\item \textbf{(D1)} An individual is a \emph{God} if and only if he possesses all positive properties:
+$$
+G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+$$
+\item \textbf{(D2)} A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
 (\psi(x) \imp \nec \all x.
@@
 \begin{itemize}
-\item \textbf{(D2)} An individual is a \emph{God} if and only if he possesses all positive properties:
-$$
-G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
-$$
 \item \textbf{(D3)} An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed:

### 2013-08-18 Abstract/GodProof-Abstract.tex
Make definition numbers agree in Coq, ND and abstract

@@
 & \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
-& \text{Necessary existence is a positive property.} \notag \\
-& \quad pos\,nec\_exists \tag{A5} 
-\end{align}
-\begin{align}
-& \text{x is God-like if and only if x has every positive property.} \notag \\
-& \quad god\,x := \forall P (pos\,P \Rightarrow P\,x) \tag{D1} \\ 
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
-& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box \forall y (P\,y \Rightarrow Q\,y)) \tag{D2} \\
-& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
-& \quad nec\_exists\,x := \forall P (ess\,P\,x \Rightarrow \Box \exists y (P\,y)) \tag{D3} 
+& \text{Necessary existence is a positive property:} \notag \\
+& \quad pos\,nec\_exists \tag{A5} \\
+& \text{x is God-like if and only if x has every positive property:} \notag \\
+& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
+& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2}
 \end{align}
@@
 % \item[(A3)] The property of being God-like is positive:
-
 % \item[(A4)] positive properties are necessarily positive; 
@@
 % \item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
-% \end{description}
 

### 2013-08-18 Abstract/GodProof-Abstract.tex
minor modifications in abstract

@@
 & \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
-& \text{Necessary existence is a positive property:} \notag \\
+& \text{Necessary existence is a positive property.} \notag \\
 & \quad pos\,nec\_exists \tag{A5} \\
-& \text{x is God-like if and only if x has every positive property:} \notag \\
-& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
-& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2}
+%\end{align}
+%\begin{align}
+& \text{x is God-like if and only if x has every positive property.} \notag \\
+& \quad god\,x := \forall P (pos\,P \Rightarrow P\,x) \tag{D1} \\ 
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
+& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box \forall y (P\,y \Rightarrow Q\,y)) \tag{D2} \\
+& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
+& \quad nec\_exists\,x := \forall P (ess\,P\,x \Rightarrow \Box \exists y (P\,y)) \tag{D3} 
 \end{align}

### 2013-08-18 Abstract/GodProof-Abstract.tex
Remove deprecated comments
Replace formulas by the formulas used in the ND proof

--- 2013-08-18 Abstract/GodProof-Abstract.tex
@@
 \newcommand{\pos}{\Diamond} % possibly
+\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
 
@@
 }
-\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this sources.}
+\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
+\marginpar{\scriptsize Bruno: I think the numbers (A1, A2,.., D1,..D3, L1,..,L3,T) should be in the left with a description environment, as they were before.
+I think it looked nicer like that.
+If you agree, I can change back.}
 \allowdisplaybreaks[1] 
@@
 & \text{Any property necessarily implied by a positive property is positive.} \notag \\
-& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
+& \quad 
+\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
 & \text{A property is positive if and only if its negation is not positive.} \notag \\
-& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
+& \quad 
+\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+\tag{A2} \\
 & \text{The property of being God-like is positive.} \notag \\
-& \quad pos\,god \tag{A3} \\
+& \quad 
+P(G) 
+\tag{A3} \\
 & \text{Positive properties are necessarily positive.} \notag \\
-& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
-& \text{Necessary existence is a positive property:} \notag \\
-& \quad pos\,nec\_exists \tag{A5} \\
-& \text{x is God-like if and only if x has every positive property:} \notag \\
-& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
-& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2}
+& \quad 
+\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+\tag{A4} \\
+& \text{Necessary existence is a positive property.} \notag \\
+& \quad 
+P(E)
+\tag{A5} \\
+& \text{x is God-like if and only if x has every positive property.} \notag \\
+& \quad 
+G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+\tag{D1} \\ 
+& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
+& \quad 
+\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
+\tag{D2} \\
+& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
+& \quad 
+E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
+\tag{D3} 
 \end{align}
-
-% \begin{description}
-% \item[(A1)] Any property necessarily implied by a positive property is positive: \\
-%        $\forall P \forall Q ((positive P \wedge (\Box \forall x (P x \Rightarrow Q x))) \Rightarrow positive Q)$
-% \item[(A2)] A property is positive if and only if its negation is not positive:\\
-%        $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
-% \item[(A3)] The property of being God-like is positive:
-% \item[(A4)] positive properties are necessarily positive; 
-% \item[(A5)] necessary existence is a positive property.
-% \item[(D1)] x is God-like if and only if x has every positive property, 
-% \item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
-% \item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
 
@@
 & \text{Positive properties are possibly exemplified.} \notag \\
-& \quad \forall P (pos\,P \Rightarrow \Diamond \exists x (P\,x)) \tag{L1} \\
+& \quad 
+\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
+\tag{L1} \\
 & \text{Possibly God exists.} \notag \\
-& \quad \Diamond \exists x (god\,x) \tag{L2} \\
+& \quad 
+\pos \ex x. G(x)
+\tag{L2} \\
 & \parbox[c][2em][c]{0.8\textwidth}{If x is God-like, then the property of being God-like is an essence of x.} \notag \\
-& \quad \forall x (god\,x \Rightarrow ess\,god\,x) \tag{L3} \\
+& \quad 
+\all y.[G(y) \imp \ess{G}{y}]
+\tag{L3} \\
 & \text{Necessarily God exists.} \notag \\
-& \quad \Box \exists x (god\,x) \tag{T}
+& \quad 
+\nec \ex x. G(x)
+\tag{T}
 \end{align}
-
-%
-%ToDo: Include formulas too??
-% \begin{description}
-% \item[(L1)] Positive properties are possibly exemplified.
-% \item[(L2)] Possibly God exists.
-% \item[(L3)] If x is God-like, then the property of being 
-% God-like is an essence of x.
-% \item[(T)] \ Necessarily God exists.
-% \end{description}
-
-% From these axioms, he claims (without detailed proof) that:
-% \begin{enumerate}
-% \item If it is possible that God exists, then it is necessary that God exists.
-% \item It is possible that God exists.
-% \end{enumerate}
 
@@
 requires an expressive logical language with modal operators
-(\emph{possibily} and \emph{necessarily}) and with
+(\emph{possibly} and \emph{necessarily}) and with
 quantififiers for individuals and sets of individuals (properties).  Our computer-assisted formalizations rely on an

### 2013-08-18 Natural Deduction/GodProof-ND.tex
Add ToDo reminder
Fix typo

@@
 $$
-\item \textbf{(D2)} A property is \emph{essential} for an individual if and only if it holds for that inidividual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
+\item \textbf{(D2)} A property is \emph{essential} for an individual if and only if it holds for that individual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
 (\psi(x) \imp \nec \all x.
@@
 
+ToDo: This section still needs more details. See Coq formalization for more details.
+
 \noindent
@@
                 \AXC{$ $} \dashedLine \RightLabel{Th. A}
-                \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
+                \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$} \doubleLine
             \BIC{$\pos \nec \ex x. G(x)$}

### 2013-08-19 Abstract/GodProof-Abstract.tex

@@
 \marginpar{\scriptsize Bruno: I think the numbers (A1, A2,.., D1,..D3, L1,..,L3,T) should be in the left with a description environment, as they were before.
-I think it looked nicer like that.
-If you agree, I can change back.}
+I think it looked nicer before.
+Can I change it back?}
 \allowdisplaybreaks[1] 
@@
 \tag{A5} \\
-& \text{x is God-like if and only if x has every positive property.} \notag \\
+& \text{Being God-like is having every positive property.} \notag \\
 & \quad 
@@
 \tag{D1} \\ 
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
+& \parbox[c][2em][c]{0.8\textwidth}{A property is an essence of an individual if and only if the individual has it and every property  that the individual has is necessarily implied by it.}\notag \\
 & \quad 
-\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
+\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all y. (\varphi(y) \imp \psi(y)))
 \tag{D2} \\
-& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
+& \parbox[c][2em][c]{0.8\textwidth}{An individual necessarily exists if and only if all its essences are necessarily exemplified.} \notag \\
 & \quad 
@@
 \tag{L1} \\
-& \text{Possibly God exists.} \notag \\
+& \text{Possibly, God exists.} \notag \\
 & \quad 
@@
 \tag{L2} \\
-& \parbox[c][2em][c]{0.8\textwidth}{If x is God-like, then the property of being God-like is an essence of x.} \notag \\
+& \parbox[c][2em][c]{0.8\textwidth}{The property of being God-like is an essence of any God-like being.} \notag \\
 & \quad 
@@
 \tag{L3} \\
-& \text{Necessarily God exists.} \notag \\
+& \text{Necessarily, God exists.} \notag \\
 & \quad 
@@
 \end{align}
+\marginpar{\scriptsize Bruno: I think it might be possible to improve the natural language statements further.
+After our improvements converge here, we should update the statements in all formalizations (THF, Coq, Isabelle, ND).}
 

### 2013-08-19 TPTP THF Formalization/call_tptp.sh
modified formalization of axiom2 in THF and Isabelle, coincides now with Scott's version

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Nitrox---2013'
-timeout=180
+timeout=60
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)

### 2013-08-19 Abstract/GodProof-Abstract.tex
improved text of abstract

--- 2013-08-18 Abstract/GodProof-Abstract.tex
@@
 
-\usepackage{url,amsmath,amssymb}
+\usepackage{url,amsmath,amssymb,a4wide}
 
-% Logical symbols
 \newcommand{\imp}{\rightarrow}
@@
 \newcommand{\pos}{\Diamond} % possibly
-\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
+\newcommand{\ess}[2]{#1 \ \mathit{Ess.} \ #2}
+\newcommand{\NE}{\mathit{NE}}
 
@@
     supported by the German Research Foundation under grant
-    BE2501/9-1. Bruno ????}  }
+    BE2501/9-1.}  }
 
@@
 \institute{
-  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
+  Dahlem Center for Intelligent Sstems, Freie Universit\"{a}t Berlin, Germany\\
   \email{c.benzmueller@gmail.com}
@@
 abstract ontological arguments are an old tradition in philosophy and
-theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
+theology.  G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
 this tradition, following particularly the footsteps of Leibniz.
@@
 He does not extensively discuss what positive properties are, 
-but instead he states a few reasonable but debatable axioms that they should satisfy.
-Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g.
-\cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
-(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}.
-(A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction.
-(A4) is a weaker form of G\"odel's axiom 3.
-(D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
-}
-\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
-\marginpar{\scriptsize Bruno: I think the numbers (A1, A2,.., D1,..D3, L1,..,L3,T) should be in the left with a description environment, as they were before.
-I think it looked nicer like that.
-If you agree, I can change back.}
+but instead he states a few reasonable (but debatable) axioms that they should satisfy.
+Various slightly different versions of axioms and definitions have
+been considered by G\"{o}del and by several philosophers who commented
+on his proof
+(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+
+Dana Scott's version of G\"odel's proof \cite{ScottNotes} employs the
+following axioms (\textbf{A}), definitions (\textbf{D}), corollaries
+(\textbf{C}) and theorems (\textbf{T}), and it proceeds in the
+following order:\footnote{ (A1), (A2), (A5), (D1), (D3) are logically
+  equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and
+  3 in G\"odel's notes \cite{Goedel1970,GoedellNotes}. (A3) was introduced by
+  Scott \cite{ScottNotes} and could be derived from G\"odel's axiom 1 and
+  (D1) in a logic with infinitary conjunction. (A4) is a weaker form
+  of G\"odel's axiom 3. (D2) has an extra conjunct lacking in
+  G\"odel's definition 2; this lack is believed to have been an
+  oversight by G\"odel \cite{Hazen}. }
+% \marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
+% \marginpar{\scriptsize Bruno: I think the numbers (A1, A2,..,
+% D1,..D3, L1,..,L3,T) should be in the left with a description
+% environment, as they were before. I think it looked nicer
+% before. Can I change it back?}
+
 \allowdisplaybreaks[1] 
-\begin{align}
-& \text{Any property necessarily implied by a positive property is positive.} \notag \\
-& \quad 
-\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
-& \text{A property is positive if and only if its negation is not positive.} \notag \\
-& \quad 
-\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
-\tag{A2} \\
-& \text{The property of being God-like is positive.} \notag \\
-& \quad 
-P(G) 
-\tag{A3} \\
-& \text{Positive properties are necessarily positive.} \notag \\
-& \quad 
-\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
-\tag{A4} \\
-& \text{Necessary existence is a positive property.} \notag \\
-& \quad 
-P(E)
-\tag{A5} \\
-& \text{x is God-like if and only if x has every positive property.} \notag \\
-& \quad 
-G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
-\tag{D1} \\ 
-& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
-& \quad 
-\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
-\tag{D2} \\
-& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
-& \quad 
-E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
-\tag{D3} 
-\end{align}
+\begin{description}
+\item[A1] Either the property or its negation is positive, but not
+  both:  \hfill 
+  $\all \phi. [P(\neg \phi) \biimp \neg P(\phi)]$
+\item[A2] A property is positive if it necessarily  contains a
+  positive property:  \\ \phantom{b} \hfill 
+  $\all \phi. \all \psi.[(P(\phi) \wedge \nec \all x.[\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]$ 
+\item[T1] Positive properties are possibly exemplified. \hfill $\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]$
+\item[D1] x is God-like if it possesses all positive properties: \hfill
+  $G(x) \biimp \forall \phi. [P(\phi) \to \phi(x)]$
+\item[A3]  The property of being God-like is positive: \hfill   $P(G)$
+\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x. G(x)$
+\item[A4]  Being a positive property is logical, hence,
+  necessary: \hfill 
+  $\all \phi.[P(\phi) \to \Box \; P(\phi)]$
+\item[D2] $\phi$ is the essence of $x$ iff $x$ has $\phi$ and this
+  property is necessarily minimal.:\\
+  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
+  \psi. (\psi(x) \imp \nec \all y. (\phi(y) \imp \psi(y)))$
+\item[T2]  The property of being God-like is an essence of any
+  God-like being: \hfill $\all y.[G(y) \imp \ess{G}{y}]$ 
+\item[D3] $\NE(x)$ means that $x$ necessarily exists if it has an
+  essential property: \\
+  \phantom{b} \hfill $\NE(x) \biimp \all \phi.[\ess{\phi}{x} \imp \nec \ex x.\phi(x)]$
+\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
+\item[T3] Necessarily, God exists: \hfill $\nec \ex x. G(x)$ 
+\end{description}
 
-\noindent
-From these axioms and definitions we then infer:
-\begin{align}
-& \text{Positive properties are possibly exemplified.} \notag \\
-& \quad 
-\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
-\tag{L1} \\
-& \text{Possibly God exists.} \notag \\
-& \quad 
-\pos \ex x. G(x)
-\tag{L2} \\
-& \parbox[c][2em][c]{0.8\textwidth}{If x is God-like, then the property of being God-like is an essence of x.} \notag \\
-& \quad 
-\all y.[G(y) \imp \ess{G}{y}]
-\tag{L3} \\
-& \text{Necessarily God exists.} \notag \\
-& \quad 
-\nec \ex x. G(x)
-\tag{T}
-\end{align}
+% \begin{align}
+% & \text{Any property necessarily implied by a positive property is positive.} \notag \\
+% & \quad 
+% \all \varphi.\all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
+% & \text{A property is positive if and only if its negation is not positive.} \notag \\
+% & \quad 
+% \all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+% \tag{A2} \\
+% & \text{The property of being God-like is positive.} \notag \\
+% & \quad 
+% P(G) 
+% \tag{A3} \\
+% & \text{Positive properties are necessarily positive.} \notag \\
+% & \quad 
+% \all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+% \tag{A4} \\
+% & \text{Necessary existence is a positive property.} \notag \\
+% & \quad 
+% P(E)
+% \tag{A5} \\
+% & \text{Being God-like is having every positive property.} \notag \\
+% & \quad 
+% G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+% \tag{D1} \\ 
+% & \parbox[c][2em][c]{0.8\textwidth}{A property is an essence of an individual if and only if the individual has it and every property  that the individual has is necessarily implied by it.}\notag \\
+% & \quad 
+% \ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all y. (\varphi(y) \imp \psi(y)))
+% \tag{D2} \\
+% & \parbox[c][2em][c]{0.8\textwidth}{An individual necessarily exists if and only if all its essences are necessarily exemplified.} \notag \\
+% & \quad 
+% E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
+% \tag{D3} 
+% \end{align}
 
-The above variant of G\"{o}del's proof has now been 
-constructed for the first-time
+% \noindent
+% From these axioms and definitions we then infer:
+% \begin{align}
+% & \text{Positive properties are possibly exemplified.} \notag \\
+% & \quad 
+% \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
+% \tag{L1} \\
+% & \text{Possibly, God exists.} \notag \\
+% & \quad 
+% \pos \ex x. G(x)
+% \tag{L2} \\
+% & \parbox[c][2em][c]{0.8\textwidth}{The property of being God-like is an essence of any God-like being.} \notag \\
+% & \quad 
+% \all y.[G(y) \imp \ess{G}{y}]
+% \tag{L3} \\
+% & \text{Necessarily, God exists.} \notag \\
+% & \quad 
+% \nec \ex x. G(x)
+% \tag{T}
+% \end{align}
+% \marginpar{\scriptsize Bruno: I think it might be possible to improve the natural language statements further.
+After our improvements converge here, we should update the statements in all formalizations (THF, Coq, Isabelle, ND).}
+
+The above version of G\"{o}del's proof has now been 
+analysed for the first-time
 with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done (and in this order):
 \begin{itemize}
-\item A detailed natural deduction proof.\marginpar{\scriptsize Chris: We need to make sure that this ND proof is based on a sound calculus.}
+\item A detailed natural deduction proof.
 %
@@
 (\emph{possibly} and \emph{necessarily}) and with
-quantififiers for individuals and sets of individuals (properties).  Our computer-assisted formalizations rely on an
-embedding of the modal logic S5 into classical higher-order logic with
+quantifiers for individuals and properties.  Our computer-assisted formalizations rely on an
+embedding of the modal logic into classical higher-order logic with
 Henkin semantics \cite{J23,B9}. The formalization is thus essentially
-done in classical higher-order logic where quantified S5 is emulated.
+done in classical higher-order logic where  quantified modal logic is
+emulated.
+
+In our ongoing computer-assisted study of G\"odel's proof we have
+obtained the following results:
+\begin{itemize}
+\item Logic S5 is not needed for obtaining the results, logic KB is
+  sufficient.
+\item For proving theorem T1, only the left to right direction of
+  axiom A1 is needed. However, the backward direction of A1 is
+  required for showing T2.
+\item The first conjunct $\phi(x)$ in D2, introduced by Scott, seems important; without it the set of axioms and definitions becomesx inconsistent.
+\end{itemize}
 
@@
 
-Future work includes an extensive study of other formalizations of
-ontological arguments with our machinery. Variations of the
-distinctive features of the base logic S5 (e.g. non-rigid symbols, 
-varying domains, etc.) are enabled in these studies due
-to the flexibility of the embeddings based approach.
+Future work includes a substantial widening of our computer-assisted
+study of ontological arguments. Due to the flexibility of the
+embeddings based approach, variations of the distinctive features of
+the base logic (e.g. non-rigid symbols, varying domains, etc.)  will
+be enabled in these studies
 
+%\footnotesize
 \bibliographystyle{plain}

### 2013-08-20 Abstract/GodProof-Abstract.tex
Fix typo
Improve natural language statements of axioms and definitions.
Justifications for improvements:

A2 - "contains" presupposes a set-theoretic interpretration for properties

A4 - it is intuitively strange to talk about properties "being" (existing), even though technically that's what we do with Henkin semantics. I belive my reformulation is less strange in this respect.

Natural language text is more "natural" when it doesn't contain logical variables and it can also be more easily appreciated by readers who are not formal logicians.

@@
   $\all \phi. [P(\neg \phi) \biimp \neg P(\phi)]$
-\item[A2] A property is positive if it necessarily  contains a
+\item[A2] A property is positive if it is necessarily implied by a
   positive property:  \\ \phantom{b} \hfill 
@@
 \item[T1] Positive properties are possibly exemplified. \hfill $\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]$
-\item[D1] x is God-like if it possesses all positive properties: \hfill
+\item[D1] A God-like being possesses all positive properties: \hfill
   $G(x) \biimp \forall \phi. [P(\phi) \to \phi(x)]$
@@
 \item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x. G(x)$
-\item[A4]  Being a positive property is logical, hence,
-  necessary: \hfill 
+\item[A4]  Positive properties are necessarily positive: \hfill 
   $\all \phi.[P(\phi) \to \Box \; P(\phi)]$
-\item[D2] $\phi$ is the essence of $x$ iff $x$ has $\phi$ and this
-  property is necessarily minimal.:\\
+\item[D2] An essence of an individual is a property possessed by it and necessarily implying any of its properties:\\
   \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
   \psi. (\psi(x) \imp \nec \all y. (\phi(y) \imp \psi(y)))$
-\item[T2]  The property of being God-like is an essence of any
+\item[T2]  Being God-like is an essence of any
   God-like being: \hfill $\all y.[G(y) \imp \ess{G}{y}]$ 
-\item[D3] $\NE(x)$ means that $x$ necessarily exists if it has an
-  essential property: \\
+\item[D3] Necessary existance of an individual is the necessary exemplification of all its essences: \\
   \phantom{b} \hfill $\NE(x) \biimp \all \phi.[\ess{\phi}{x} \imp \nec \ex x.\phi(x)]$
@@
   required for showing T2.
-\item The first conjunct $\phi(x)$ in D2, introduced by Scott, seems important; without it the set of axioms and definitions becomesx inconsistent.
+\item The first conjunct $\phi(x)$ in D2, introduced by Scott, is important; without it the set of axioms and definitions becomes inconsistent.
 \end{itemize}

### 2013-08-20 TPTP THF Formalization/call_tptp.sh
changed syntax in Isabelle formalization; it corresponds now closely to Scott's version of Goedel's proof

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Nitrox---2013'
-timeout=60
+timeout=180
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)

### 2013-08-20 Abstract/GodProof-Abstract.tex

@@
 \newcommand{\pos}{\Diamond} % possibly
-\newcommand{\ess}[2]{#1 \ \mathit{Ess.} \ #2}
+\newcommand{\ess}[2]{#1 \ \mathit{ess.} \ #2}
 \newcommand{\NE}{\mathit{NE}}
@@
 \title{ Formalization, Mechanization and Automation % of Variants
-  of G\"{o}del's Proof of God's Existence\thanks{This work has been
+  of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
     supported by the German Research Foundation under grant
@@
 (\textbf{C}) and theorems (\textbf{T}), and it proceeds in the
-following order:\footnote{ (A1), (A2), (A5), (D1), (D3) are logically
+following order:\footnote{ A1, A2, A5, D1, D3 are logically
   equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and
-  3 in G\"odel's notes \cite{Goedel1970,GoedellNotes}. (A3) was introduced by
-  Scott \cite{ScottNotes} and could be derived from G\"odel's axiom 1 and
-  (D1) in a logic with infinitary conjunction. (A4) is a weaker form
-  of G\"odel's axiom 3. (D2) has an extra conjunct lacking in
-  G\"odel's definition 2; this lack is believed to have been an
-  oversight by G\"odel \cite{Hazen}. }
-% \marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
-% \marginpar{\scriptsize Bruno: I think the numbers (A1, A2,..,
-% D1,..D3, L1,..,L3,T) should be in the left with a description
-% environment, as they were before. I think it looked nicer
-% before. Can I change it back?}
+  3 in G\"odel's notes \cite{Goedel1970,GoedelNotes}. 
+  A3 was introduced by Scott \cite{ScottNotes} and 
+  could be derived from G\"odel's axiom 1 and
+  D1 in a logic with infinitary conjunction. 
+  A4 is a weaker formof G\"odel's axiom 3. 
+  D2 has an extra conjunct $\phi(x)$ lacking in G\"odel's definition 2; 
+  this is believed to have been 
+  an oversight by G\"odel \cite{Hazen}.}
 
@@
 \begin{description}
-\item[A1] Either the property or its negation is positive, but not
+\item[A1] Either a property or its negation is positive, but not
   both:  \hfill 
   $\all \phi. [P(\neg \phi) \biimp \neg P(\phi)]$
-\item[A2] A property is positive if it is necessarily implied by a
-  positive property:  \\ \phantom{b} \hfill 
+\item[A2] A property necessarily implied \\ by a
+  positive property is positive: \phantom{b} \hfill 
   $\all \phi. \all \psi.[(P(\phi) \wedge \nec \all x.[\phi(x)
   \imp \psi(x)]) \imp P(\psi)]$ 
-\item[T1] Positive properties are possibly exemplified. \hfill $\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]$
-\item[D1] A God-like being possesses all positive properties: \hfill
+\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]$
+\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
   $G(x) \biimp \forall \phi. [P(\phi) \to \phi(x)]$
@@
   $\all \phi.[P(\phi) \to \Box \; P(\phi)]$
-\item[D2] An essence of an individual is a property possessed by it and necessarily implying any of its properties:\\
+\item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
   \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
@@
 \item[T2]  Being God-like is an essence of any
-  God-like being: \hfill $\all y.[G(y) \imp \ess{G}{y}]$ 
-\item[D3] Necessary existance of an individual is the necessary exemplification of all its essences: \\
-  \phantom{b} \hfill $\NE(x) \biimp \all \phi.[\ess{\phi}{x} \imp \nec \ex x.\phi(x)]$
+  God-like being: \hfill $\all x.[G(x) \imp \ess{G}{x}]$ 
+\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
+  \phantom{b} \hfill $\NE(x) \biimp \all \phi.[\ess{\phi}{x} \imp \nec \ex y.\phi(y)]$
 \item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
@@
 
-% \begin{align}
-% & \text{Any property necessarily implied by a positive property is positive.} \notag \\
-% & \quad 
-% \all \varphi.\all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
-% & \text{A property is positive if and only if its negation is not positive.} \notag \\
-% & \quad 
-% \all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
-% \tag{A2} \\
-% & \text{The property of being God-like is positive.} \notag \\
-% & \quad 
-% P(G) 
-% \tag{A3} \\
-% & \text{Positive properties are necessarily positive.} \notag \\
-% & \quad 
-% \all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
-% \tag{A4} \\
-% & \text{Necessary existence is a positive property.} \notag \\
-% & \quad 
-% P(E)
-% \tag{A5} \\
-% & \text{Being God-like is having every positive property.} \notag \\
-% & \quad 
-% G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
-% \tag{D1} \\ 
-% & \parbox[c][2em][c]{0.8\textwidth}{A property is an essence of an individual if and only if the individual has it and every property  that the individual has is necessarily implied by it.}\notag \\
-% & \quad 
-% \ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all y. (\varphi(y) \imp \psi(y)))
-% \tag{D2} \\
-% & \parbox[c][2em][c]{0.8\textwidth}{An individual necessarily exists if and only if all its essences are necessarily exemplified.} \notag \\
-% & \quad 
-% E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
-% \tag{D3} 
-% \end{align}
-
-% \noindent
-% From these axioms and definitions we then infer:
-% \begin{align}
-% & \text{Positive properties are possibly exemplified.} \notag \\
-% & \quad 
-% \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
-% \tag{L1} \\
-% & \text{Possibly, God exists.} \notag \\
-% & \quad 
-% \pos \ex x. G(x)
-% \tag{L2} \\
-% & \parbox[c][2em][c]{0.8\textwidth}{The property of being God-like is an essence of any God-like being.} \notag \\
-% & \quad 
-% \all y.[G(y) \imp \ess{G}{y}]
-% \tag{L3} \\
-% & \text{Necessarily, God exists.} \notag \\
-% & \quad 
-% \nec \ex x. G(x)
-% \tag{T}
-% \end{align}
-% \marginpar{\scriptsize Bruno: I think it might be possible to improve the natural language statements further.
-After our improvements converge here, we should update the statements in all formalizations (THF, Coq, Isabelle, ND).}
-
+\noindent
 The above version of G\"{o}del's proof has now been 
 analysed for the first-time
-with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done (and in this order):
+with an unprecedent degree of detail 
+and formality; cf.~\cite{FormalTheologyRepository}. 
+The following has been done (and in this order):
 \begin{itemize}
@@
 \begin{itemize}
-\item Logic S5 is not needed for obtaining the results, logic KB is
-  sufficient.
+\item The basic modal logic K is sufficient for proving T1, C and T2. \marginpar{\scriptsize{Bruno: check this.}}
+\item Modal logic S5 is not needed for proving T3; the logic KB is
+  sufficient. \marginpar{\scriptsize{Bruno: Maybe cite a reference for S5 and KB?}}
+\item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
+  and definitions would be inconsistent.
 \item For proving theorem T1, only the left to right direction of
   axiom A1 is needed. However, the backward direction of A1 is
-  required for showing T2.
-\item The first conjunct $\phi(x)$ in D2, introduced by Scott, is important; without it the set of axioms and definitions becomes inconsistent.
+  required for proving T2.
+
 \end{itemize}
@@
 This work attests the maturity of contemporary interactive and
-automated deduction tools for classical higher-order logic and it
-demonstrates the elegance and practical relevance of the embeddings
-based approach.  Most importantly, our work opens new perspectives for
+automated deduction tools for classical higher-order logic and
+demonstrates the elegance and practical relevance of the embeddings-based approach.  Most importantly, our work opens new perspectives for
 a computer-assisted theoretical philosophy.  The critical discussion
@@
 
-Future work includes a substantial widening of our computer-assisted
-study of ontological arguments. Due to the flexibility of the
-embeddings based approach, variations of the distinctive features of
-the base logic (e.g. non-rigid symbols, varying domains, etc.)  will
-be enabled in these studies
+\marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
+We could try to improve it, but I think simply removing it would also be satisfactory.
+The sentence about Leibniz' dictum already provides a very nice ``climax'' for the end of the paper.
+Anything that we say after it feels like an ``anticlimax''...}}
+Future work shall widen our computer-assisted approach to the
+study of other ontological arguments. Variations of the distinctive features of
+the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
+embeddings-based approach, will
+be considered.
 

### 2013-08-21 Abstract/GodProof-Abstract.tex
update of THF README and asbtract

@@
   both:  \hfill 
-  $\all \phi. [P(\neg \phi) \biimp \neg P(\phi)]$
+  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$
 \item[A2] A property necessarily implied \\ by a
   positive property is positive: \phantom{b} \hfill 
-  $\all \phi. \all \psi.[(P(\phi) \wedge \nec \all x.[\phi(x)
+  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
   \imp \psi(x)]) \imp P(\psi)]$ 
-\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]$
+\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$
 \item[D1] A \emph{God-like} being possesses all positive properties: \hfill
-  $G(x) \biimp \forall \phi. [P(\phi) \to \phi(x)]$
+  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$
 \item[A3]  The property of being God-like is positive: \hfill   $P(G)$
-\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x. G(x)$
+\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$
 \item[A4]  Positive properties are necessarily positive: \hfill 
-  $\all \phi.[P(\phi) \to \Box \; P(\phi)]$
+  $\all \phi [P(\phi) \to \Box \; P(\phi)]$
 \item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
   \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
-  \psi. (\psi(x) \imp \nec \all y. (\phi(y) \imp \psi(y)))$
+  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$
 \item[T2]  Being God-like is an essence of any
-  God-like being: \hfill $\all x.[G(x) \imp \ess{G}{x}]$ 
+  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ 
 \item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
-  \phantom{b} \hfill $\NE(x) \biimp \all \phi.[\ess{\phi}{x} \imp \nec \ex y.\phi(y)]$
+  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$
 \item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
-\item[T3] Necessarily, God exists: \hfill $\nec \ex x. G(x)$ 
+\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
 \end{description}
@@
 
-\marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
+% \marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
 We could try to improve it, but I think simply removing it would also be satisfactory.
@@
 Anything that we say after it feels like an ``anticlimax''...}}
-Future work shall widen our computer-assisted approach to the
-study of other ontological arguments. Variations of the distinctive features of
-the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
-embeddings-based approach, will
-be considered.
+% Future work shall widen our computer-assisted approach to the
+% study of other ontological arguments. Variations of the distinctive features of
+% the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
+% embeddings-based approach, will
+% be considered.
 

### 2013-08-21 Abstract/GodProof-Abstract.tex
finalize asbtract

@@
 \noindent
-The above version of G\"{o}del's proof has now been 
+Scott's version of G\"{o}del's proof has now been 
 analysed for the first-time
 with an unprecedent degree of detail 
-and formality; cf.~\cite{FormalTheologyRepository}. 
+and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
 The following has been done (and in this order):
@@
 
-\item A formalization using the Isabelle proof assistant \cite{Isabelle} where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
+\item A formalization using the Isabelle proof assistant \cite{Isabelle}, where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
 \end{itemize}
@@
 \begin{itemize}
-\item The basic modal logic K is sufficient for proving T1, C and T2. \marginpar{\scriptsize{Bruno: check this.}}
+\item The basic modal logic K is sufficient for proving T1, C and T2. 
+%\marginpar{\scriptsize{Bruno: check this. Chris: Done.}}
 \item Modal logic S5 is not needed for proving T3; the logic KB is
-  sufficient. \marginpar{\scriptsize{Bruno: Maybe cite a reference for S5 and KB?}}
+  sufficient. 
+% \marginpar{\scriptsize{Bruno: Maybe cite a reference for S5 and KB?}}
 \item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
@@
 automated deduction tools for classical higher-order logic and
-demonstrates the elegance and practical relevance of the embeddings-based approach.  Most importantly, our work opens new perspectives for
-a computer-assisted theoretical philosophy.  The critical discussion
-of the underlying concepts, definitions and axioms remains a human
-responsibility, but the computer can assist in building and checking
-rigorously correct logical arguments. In case of logico-philosophical
-disputes, the computer can check the disputing arguments and partially
-fulfill Leibniz' dictum: Calculemus --- Let us calculate!
+demonstrates the elegance and practical relevance of the
+embeddings-based approach.  Most importantly, our work opens new
+perspectives for a computer-assisted theoretical philosophy.  The
+critical discussion of the underlying concepts, definitions and axioms
+remains a human responsibility, but the computer can assist in
+building and checking rigorously correct logical arguments. In case of
+logico-philosophical disputes, the computer can check the disputing
+arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
+calculate!
 
@@
 \bibliographystyle{plain}
-\bibliography{Bibliography}
+%\bibliography{Bibliography}
+\begin{thebibliography}{10}
+
+\bibitem{Adams}
+R.M. Adams.
+\newblock Introductory note to *1970.
+\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+  Letters}}. Oxford University Press, 1995.
+
+\bibitem{AndersonGettings}
+A.C. Anderson and M.~Gettings.
+\newblock G\"odel ontological proof revisited.
+\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
+  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
+
+\bibitem{B9}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Exploring properties of normal multimodal logics in simple type
+  theory with {LEO-II}.
+\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
+  Birthday}}, pages 386--406. College Publications.
+
+\bibitem{J23}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Quantified multimodal logics in simple type theory.
+\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
+  7(1):7--20, 2013.
+
+\bibitem{LEO-II}
+C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
+\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
+  logic.
+\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
+  162--170. Springer, 2008.
+
+\bibitem{Coq}
+Y.~Bertot and P.~Casteran.
+\newblock {\em {Interactive Theorem Proving and Program Development}}.
+\newblock Springer, 2004.
+
+\bibitem{Sledgehammer}
+J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+\newblock Extending {Sledgehammer} with {SMT} solvers.
+\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
+
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for higher-order logic based on a
+  relational model finder.
+\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+  Springer, 2010.
+
+\bibitem{Satallax}
+C.E. Brown.
+\newblock Satallax: An automated higher-order prover.
+\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+  Springer, 2012.
+
+\bibitem{ContemporaryBibliography}
+R.~Corazzon.
+\newblock Contemporary bibligraphy on the ontological proof
+  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
+
+\bibitem{Fitting}
+M.~Fitting.
+\newblock {\em {Types, Tableaux and G\"odel's God}}.
+\newblock Kluver Academic Press, 2002.
+
+\bibitem{Goedel1970}
+K.~G\"odel.
+\newblock Ontological proof.
+\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+  Letters}}. Oxford University Press, 1970.
+
+\bibitem{GoedelNotes}
+K.~G\"odel.
+\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{Hazen}
+A.P. Hazen.
+\newblock On g\"odel's ontological proof.
+\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+
+\bibitem{Hurd03first-orderproof}
+J.~Hurd.
+\newblock First-order proof tactics in higher-order logic theorem provers.
+\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
+  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
+
+\bibitem{Isabelle}
+T.~Nipkow, L.C. Paulson, and M.~Wenzel.
+\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
+\newblock Number 2283 in LNCS. Springer, 2002.
+
+\bibitem{FormalTheologyRepository}
+B.~Woltzenlogel Paleo and C.~Benzm\"uller.
+\newblock Formal theology repository
+  (\url{https://github.com/FormalTheology/GoedelGod}).
+
+\bibitem{ScottNotes}
+D.~Scott.
+\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{sobel2004logic}
+J.H. Sobel.
+\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
+\newblock Cambridge U. Press, 2004.
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller.
+\newblock Automated reasoning in higher-order logic using the {TPTP THF}
+  infrastructure.
+\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
+
+\end{thebibliography}
 

### 2013-08-22 TPTP THF Formalization/call_tptp.sh
update of comments in Isabelle file; making clear that sledgehammer can automate all theorems and lemmata right away

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Nitrox---2013'
-timeout=60
+timeout=600
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)

### 2013-08-22 TPTP THF Formalization/call_tptp.sh
further update of comments and structure of Isabelle proof

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Nitrox---2013'
-timeout=600
+timeout=60
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)

### 2013-08-25 Natural Deduction/GodProof-ND.tex
Make the natural language statements in the ND proof agree with those in the abstract

@@
 \newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
+\newcommand{\NE}[1]{\mathit{NE}(#1)}
 
@@
 
-\section{Possible witnessing of positive properties}
+\section{Introduction}
 
-\textbf{Axioms:}
-\begin{itemize}
-\item \textbf{(A1)} Properties necessarily entailed by \emph{positive} properties are also positive:
+\section{Natural Deduction}
+
+\section{Possibly, God Exists}
+
+\begin{description}
+\item[A1] Either a property or its negation is positive, but not both:
+$$
+\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+$$
+%
+\item[A2] A property necessarily implied by a positive property is positive:
 $$
@@
 $$
-%
-\item \textbf{(A2)} A property's negation is positive iff the property is not positive:
-$$
-\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
-$$
-\end{itemize}
+\end{description}
 
-\noindent
-\textbf{Lemma 1:} Positive properties possibly have a witness:
+\begin{description}
+\item[T1] Positive properties are possibly exemplified:
 $$
@@
 $$
-
-\noindent
-\textbf{Formal proof:}
+\end{description}
 
@@
 
-\section{Possible existence of a God}
-
-\textbf{Axioms:}
-\begin{itemize}
-\item \textbf{(A3)} Being God is a positive property:
+\begin{description}
+\item[D1] A \emph{God-like} being possesses all positive properties:
+$$
+G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+$$
+\item[A3] The property of being God-like is positive:
 $$
@@
 $$
-\end{itemize}
+\end{description}
 
-\noindent
-\textbf{Lemma 2:} It is possible that a God exists:
+\begin{description}
+\item[C1] Possibly, God exists:
 $$
@@
 
-\textbf{Definitions:}
-\begin{itemize}
-\item \textbf{(D1)} An individual is a \emph{God} if and only if he possesses all positive properties:
-$$
-G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
-$$
-\item \textbf{(D2)} A property is \emph{essential} for an individual if and only if it holds for that individual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
-(\psi(x) \imp \nec \all x.
-(\varphi(x) \imp \psi(x)))$
-\end{itemize}
-
-\noindent
-\textbf{Axioms:}
-\begin{itemize}
-\item \textbf{(A4)} Positive properties are necessarily positive:
+\begin{description}
+\item[A4] Positive properties are necessarily positive:
 $$
@@
 $$
-\end{itemize}
+%
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
+(\psi(x) \imp \nec \all x.
+(\varphi(x) \imp \psi(x)))$
+\end{description}
 
-\noindent
-\textbf{Lemma 3}: If an individual is a God, then being God is an essential property for that individual:
+\begin{description}
+\item[T2]: Being God-like is an essence of any
+  God-like being:
 $$
@@
        		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
-		\AXC{$$} \RightLabel{Ax. 4}
+		\AXC{$ $} \RightLabel{Ax. 4}
        		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
@@
 
-\textbf{Definitions:}
-\begin{itemize}
-\item \textbf{(D3)} An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed:
+\begin{description}
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
 $$
@@
 $$
-\end{itemize}
-
-\noindent
-\textbf{Axioms:}
-\begin{itemize}
-\item \textbf{(A5)} Necessary existence is a positive property:
+%
+\item[A5] Necessary existence is a positive property:
 $$
@@
 $$
-\end{itemize}
+\end{description}
 
-\noindent
-\textbf{Auxiliary Lemma:} If there is a God, then there necessarily exists a God:
+\begin{description}
+\item[L1] If there is a God, then there necessarily exists a God:
 $$
@@
 
-\section{Necessary existence of a God}
+\section{Necessarily, God exists}
 
@@
 
-\noindent
-\textbf{Theorem:} The existence of a God is necessary:
+\begin{description}
+\item[T3] The existence of a God is necessary:
 $$
@@
 
-\section{God's existence}
+\section{God exists}
 
-\noindent
-\textbf{Axioms:}
-\begin{itemize}
-\item \textbf{(M)} What is necessary is the case:
+\begin{description}
+\item[M\phantom{0}] What is necessary is the case:
 $$
@@
 $$
-\end{itemize}
+\end{description}
 
-\noindent
-\textbf{Corollary:} There exists a God:
+\begin{description}
+\item[C2] There exists a God:
 $$

### 2013-08-25 Natural Deduction/GodProof-ND.tex
Improve ND paper

@@
 \newcommand{\rl}[1]{\RightLabel{#1}}
+
+\newtheorem{axiom}{Axiom}
+\newtheorem{definition}{Definition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+
+\newenvironment{proof}[1][Proof]{\begin{trivlist}
+\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
 
@@
 \newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
-\newcommand{\NE}[1]{\mathit{NE}(#1)}
+\newcommand{\NE}{E}
 
@@
 
+ToDo: Show and explain here the rules of the calculus we are using.
+
+ToDo: We should use a calculus for the basic modal logic K. Everything else should be stated as axioms.
+
+ToDo: cite a paper that proves soundness and completeness for this calculus.
+
 \section{Possibly, God Exists}
 
-\begin{description}
-\item[A1] Either a property or its negation is positive, but not both:
+\begin{axiom}
+\label{A1}
+Either a property or its negation is positive, but not both:
 $$
@@
 $$
-%
-\item[A2] A property necessarily implied by a positive property is positive:
+\end{axiom}
+
+\begin{axiom}
+\label{A2}
+A property necessarily implied by a positive property is positive:
 $$
@@
 $$
-\end{description}
+\end{axiom}
 
-\begin{description}
-\item[T1] Positive properties are possibly exemplified:
+\begin{theorem}
+\label{T1}
+Positive properties are possibly exemplified:
 $$
@@
 $$
-\end{description}
-
+\end{theorem}
+\begin{proof} \hfill
 \begin{prooftree}
-        \AXC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$}
-        \UIC{$ \all \psi.[(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \psi(x)]) \imp P(\psi)]$}
+        \AXC{Axiom \ref{A2}} \dashedLine
+        \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
+        \UIC{$ \all \psi.[(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
         \UIC{$(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \neg \varphi'(x)]) \imp P(\neg \varphi')$} \doubleLine
         \UIC{$(P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp P(\neg \varphi')$}
-                        \AXC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$}
+                        \AXC{Axiom \ref{A1}} \dashedLine
+                        \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
                         \UIC{$ P(\neg \varphi') \biimp \neg P(\varphi') $} \doubleLine
                  \BIC{$ (P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp \neg P(\varphi') $} \doubleLine
-                 \UIC{$ P(\varphi') \imp \pos \ex x.\varphi'(x) $}
+                 \UIC{$ P(\varphi') \imp \pos \ex x.\varphi'(x) $} \RightLabel{$\all_I$}
                  \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
 \end{prooftree}
+\end{proof}
 
-\begin{description}
-\item[D1] A \emph{God-like} being possesses all positive properties:
+\begin{definition}
+\label{D1}
+A \emph{God-like} being possesses all positive properties:
 $$
@@
 $$
-\item[A3] The property of being God-like is positive:
+\end{definition}
+
+\begin{axiom}
+\label{A3}
+The property of being God-like is positive:
 $$
@@
 $$
-\end{description}
+\end{axiom}
 
-\begin{description}
-\item[C1] Possibly, God exists:
+\begin{corollary}
+\label{C1}
+Possibly, God exists:
 $$
@@
 $$
-\end{description}
-
+\end{corollary}
+\begin{proof} \hfill
 \begin{prooftree}
-\AXC{$P(G)$}
-                 \AXC{$ $} \dashedLine \RightLabel{Th. 1}
-                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$}
-                 \UIC{$ P(G) \imp \pos \ex x.G(x) $}
+\AXC{Axiom \ref{A3}} \dashedLine
+\UIC{$P(G)$}
+                 \AXC{Theorem \ref{T1}} \dashedLine
+                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$} \RightLabel{$\all_E $}
+                 \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
     \BIC{$\pos \ex x. G(x)$}
 \end{prooftree}
+\end{proof}
 
-\section{Essentiality of being God}
+\section{Being God is an essence of any God}
 
-\begin{description}
-\item[A4] Positive properties are necessarily positive:
+\begin{axiom}
+\label{A4}
+Positive properties are necessarily positive:
 $$
@@
 $$
-%
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi.
-(\psi(x) \imp \nec \all x.
-(\varphi(x) \imp \psi(x)))$
-\end{description}
+\end{axiom}
 
-\begin{description}
-\item[T2]: Being God-like is an essence of any
-  God-like being:
+\begin{definition}
+\label{D2}
+An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: 
+$$
+\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
+$$
+\end{definition}
+
+ToDo: instead of using the $\nec_E$ rule, we should use the $M$ axiom.
+
+\begin{theorem}
+\label{T2}
+Being God-like is an essence of any God-like being:
 $$
@@
 $$
-\end{description}
-
-Let the following derivation with the open assumption $G(x)$ be $\Pi_1$:
+\end{theorem}
+\begin{proof}
+Let the following derivation with the open assumption $G(x)$ be $\Pi_1[G(x)]$:
 
@@
 \AXC{$ \neg P(\psi)^1$} 
-       \AXC{$ $} \RightLabel{Ax. 2}
-       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall$E}  
-       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp$E}   
+       \AXC{Axiom \ref{A1}} \dashedLine
+       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall_E$}  
+       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}   
  \BIC{$P(\neg\psi)$}\RightLabel{} 
-           	\AXC{$G(x)$} \dottedLine\RightLabel{Definition of G}
-		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
-              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp$ E}
-           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp$E}  
-            \AXC{$ \psi(x)^2$} \RightLabel{$\imp$E}  
-             \BIC{$\bot $}\RightLabel{RAA, 1}  
-              \UIC{$ P(\psi)$}\RightLabel{$\imp$ I, 2}
+           	\AXC{$G(x)$} \dottedLine\RightLabel{D\ref{D1}}
+		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
+              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp_E$}
+           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}  
+            \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}  
+             \BIC{$\bot $}\RightLabel{$\mathit{RAA}^1$}  
+              \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
               \UIC{$ \psi(x)\imp P(\psi)$}
@@
 
-Let the following derivation with the open assumption $G(x)$ be $\Pi_2$: 
+\noindent
+Let the following derivation with the open assumption $G(x)$ be $\Pi_2[G(x)]$: 
 \begin{prooftree}
-       \AXC{$\psi(x)^1$} \RightLabel{}
-             	\AXC{$\Pi_1 $}  \dashedLine
-		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp$E}
-       		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
-		\AXC{$ $} \RightLabel{Ax. 4}
-       		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
-		\UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp$E}  
-          \BIC{$\Box P(\psi)$}\RightLabel{$\imp$I, 1} 
+       \AXC{$\psi(x)^3$}
+             	\AXC{$\Pi_1[G(x)] $}  \dashedLine
+		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp_E$}
+       		\BIC{$P(\psi)$} 
+		              \AXC{Axiom \ref{A4}} \dashedLine
+       	          \UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\all_E$}  
+		              \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}  
+           \BIC{$\Box P(\psi)$}\RightLabel{$\imp_I^3$} 
            \UIC{$\psi(x)\imp\Box P(\psi)$} 
@@
 
+\noindent
 Let the following derivation without open assumptions be $\Pi_3$:
@@
 \begin{prooftree}
-       \AXC{$P(\psi)^1$} \RightLabel{}
-             	\AXC{$G(x)^2$} \dottedLine\RightLabel{Definition of G}
-		\UIC{$\forall\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall$E}
-		\UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp$E}  
-       		\BIC{$\psi(x)$}\RightLabel{$\imp$I, 2}  
-       		\UIC{$G(x)\imp \psi(x)$}\RightLabel{$\forall$I} 
-		\UIC{$\forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$I, 1}  
+       \AXC{$P(\psi)^4$}
+             	\AXC{$G(x)^5$} \dottedLine\RightLabel{D\ref{D1}}
+		\UIC{$\all\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\all_E$}
+		\UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp_E$}  
+       		\BIC{$\psi(x)$}\RightLabel{$\imp_I^5$}  
+       		\UIC{$G(x)\imp \psi(x)$}\RightLabel{$\all_I$} 
+		\UIC{$\all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^4$}  
 		\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}
@@
 
-Let the following derivation with the open assumption $G(x)$ be $\Pi_4$: 
+\noindent
+Let the following derivation with the open assumption $G(x)$ be $\Pi_4[G(x)]$: 
 \begin{prooftree}
-       \AXC{$\psi(x)^1 $} \RightLabel{}
+       \AXC{$\psi(x)^6 $} \RightLabel{}
              	\AXC{$\Pi_2$}  \dashedLine
-		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp$E}
+		\UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
 		\BIC{$\Box P(\psi)$}
-		  	\AXC{$\Box P(\psi)^2 $}\RightLabel{$\Box$E}  
+		  	\AXC{$\Box P(\psi)^7 $}\RightLabel{$\Box_E$}  
        			\UIC{$P(\psi)$}
 				\AXC{$\Pi_3$}  \dashedLine\RightLabel{}
-       				\UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\imp$E}  
-       			\BIC{$\forall x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
-			\UIC{$\Box \forall x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp$I, 2} 
-			\UIC{$\Box P(\psi)\imp \Box \forall x .(G(x)\imp \psi(x))$} \RightLabel{$\imp$E} 
-         	 \BIC{$\Box \forall x .(G(x)\imp \psi(x)$}\RightLabel{$\imp$I, 1}
-           \UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}  
+       				\UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}  
+       			\BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
+			\UIC{$\Box \all x .(G(x)\imp \psi(x))$}  \RightLabel{$\imp_I^7$} 
+			\UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$} 
+         	 \BIC{$\Box \all x .(G(x)\imp \psi(x)$}\RightLabel{$\imp_I^6$}
+           \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}  
 \end{prooftree}
 
-The use of the necessitation rule is valid, because the only open assumption $\Box P(\psi)$ is boxed.
-
-We construct a derivation of theorem 3 with a subderivation $\Pi_4[G(x)^1]$, which means that the open assumption $G(x)$ in $\Pi_4$ is discharged with the rule labeled 1.
+\noindent
+The use of the necessitation rule above is correct, because the only open assumption $\Box P(\psi)$ is boxed.
+In the derivation of Theorem \ref{T2} below, the assumption $G(x)$ in the subderivation $\Pi_4[G(x)^8]$ is discharged by the rule labeled $8$.
 
 \begin{prooftree}
-       \AXC{$G(x)^1 $} \RightLabel{}
-             	\AXC{$\Pi_4[G(x)^1]$}  \dashedLine
-		\UIC{$\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x))$}\RightLabel{$\forall $I}
-		\UIC{$\forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\RightLabel{$\&$I}
-		\BIC{$G(x)\& \forall \psi .(\psi(x)\imp\Box \forall x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{Definition of ess} 
-       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp$I, 1}
+       \AXC{$G(x)^8 $} \RightLabel{}
+             	\AXC{$\Pi_4[G(x)^8]$}  \dashedLine
+		\UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\all_I$}
+		\UIC{$\all \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\RightLabel{$\wedge_I$}
+		\BIC{$G(x)\wedge \forall \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{D\ref{D2}} 
+       			\UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
 			\UIC{$G(x)\imp \ess{G}{x}$}
@@
 \end{prooftree}
+\end{proof}
 
-\section{Necessity of God's existence}
+\section{If God's existence is possible, it is necessary}
 
-\begin{description}
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+\begin{definition}
+\label{D3}
+\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
 $$
@@
 $$
-%
-\item[A5] Necessary existence is a positive property:
+\end{definition}
+
+\begin{axiom}
+\label{A5}
+Necessary existence is a positive property:
 $$
@@
 $$
-\end{description}
+\end{axiom}
 
-\begin{description}
-\item[L1] If there is a God, then there necessarily exists a God:
+\begin{lemma}
+\label{L1}
+If there is a God, then necessarily there exists a God:
 $$
@@
 $$
-\end{description}
-
+\end{lemma}
+\begin{proof} \hfill
 \begin{prooftree}
@@
 \UIC{$G(g)$}
-        \AXC{$ $} \dashedLine \RightLabel{Th. 3}
+        \AXC{Theorem \ref{T2}} \dashedLine
         \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
@@
     \BIC{$\ess{G}{g}$}
-                \AXC{$P(E)$}
+                \AXC{Axiom \ref{A5}} \dashedLine
+                \UIC{$P(E)$}
                         \AXC{$ $} \dashedLine
@@
 \end{prooftree}
+\end{proof}
 
@@
 
-\begin{description}
-\item[T3] The existence of a God is necessary:
+ToDo: this is proven in a way that is slightly different from G\"odel's 1970.
+
+\begin{theorem}
+\label{T3}
+Necessarily, God exists:
 $$
@@
 $$
-\end{description}
-
+\end{theorem}
+\begin{proof}\hfill
 \begin{small}
 \begin{prooftree}
-\AXC{$ $} \dashedLine \RightLabel{\textbf{S5}}
+\AXC{\textbf{S5}} \dashedLine
 \UIC{$ \all \varphi. [\pos \ldots \pos \nec \varphi \biimp \nec \varphi] $}
 \UIC{$ \pos \nec \ex x. G(x) \biimp \nec \ex x. G(x) $}
-        \AXC{$ $} \dashedLine \RightLabel{Th. 2}
+        \AXC{Corollary \ref{C1}} \dashedLine
         \UIC{$\pos \ex x. G(x)$}
-                \AXC{$ $} \dashedLine \RightLabel{Th. A}
+                \AXC{Lemma \ref{L1}} \dashedLine
                 \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$} \doubleLine
@@
 \end{small}
+\end{proof}
 
@@
 
-\begin{description}
-\item[M\phantom{0}] What is necessary is the case:
+\begin{axiom}[M]
+\label{M} What is necessary is the case:
 $$
@@
 $$
-\end{description}
+\end{axiom}
 
-\begin{description}
-\item[C2] There exists a God:
+\begin{corollary}
+\label{C2} 
+There exists a God:
 $$
@@
 $$
-\end{description}
-
+\end{corollary}
+\begin{proof}\hfill
 \begin{prooftree}
-\AXC{$ $} \dashedLine \RightLabel{Th. 4}
+\AXC{Theorem \ref{T3}} \dashedLine
 \UIC{$ \nec \ex x. G(x)$}

### 2013-08-25 Natural Deduction/GodProof-ND.tex
Add remarks about box elimination
Improve ND proof

@@
 
+ToDo: Investigate the relationship between the $\nec_E$ rule and the $M$ axiom (reflexivity). We don't want $M$ to be provable in our calculus.
+
 ToDo: cite a paper that proves soundness and completeness for this calculus.
@@
         \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
-        \UIC{$ \all \psi.[(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
-        \UIC{$(P(\varphi') \wedge \nec \all x.[\varphi'(x) \imp \neg \varphi'(x)]) \imp P(\neg \varphi')$} \doubleLine
-        \UIC{$(P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp P(\neg \varphi')$}
+        \UIC{$ \all \psi.[(P(\rho) \wedge \nec \all x.[\rho(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
+        \UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \doubleLine
+        \UIC{$(P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp P(\neg \rho)$}
                         \AXC{Axiom \ref{A1}} \dashedLine
                         \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
-                        \UIC{$ P(\neg \varphi') \biimp \neg P(\varphi') $} \doubleLine
-                 \BIC{$ (P(\varphi') \wedge \nec \all x.[\neg \varphi'(x)]) \imp \neg P(\varphi') $} \doubleLine
-                 \UIC{$ P(\varphi') \imp \pos \ex x.\varphi'(x) $} \RightLabel{$\all_I$}
+                        \UIC{$ P(\neg \rho) \biimp \neg P(\rho) $} \doubleLine
+                 \BIC{$ (P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp \neg P(\rho) $} \doubleLine
+                 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
                  \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
@@
 
-ToDo: instead of using the $\nec_E$ rule, we should use the $M$ axiom.
-
 \begin{theorem}
@@
 		              \AXC{Axiom \ref{A4}} \dashedLine
-       	          \UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\all_E$}  
+       	          \UIC{$\all \varphi .(P(\varphi)\imp \Box P(\varphi))$}\RightLabel{$\all_E$}  
 		              \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}  
@@
 			\UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$} 
-         	 \BIC{$\Box \all x .(G(x)\imp \psi(x)$}\RightLabel{$\imp_I^6$}
+         	 \BIC{$\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_I^6$}
            \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}  
@@
        			\UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
-			\UIC{$G(x)\imp \ess{G}{x}$}
+			\UIC{$G(x)\imp \ess{G}{x}$} \RightLabel{$\all_I$}
 			  \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
@@
 $$
-E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)]
+E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
 $$

### 2013-09-07 Abstract/Journal/GodProof-Abstract-JournalVersion.tex
Move abstract versions to corresponding subfolders

--- 2013-08-21 Abstract/GodProof-Abstract.tex
@@
-\documentclass{llncs}
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is a general template file for the LaTeX package SVJour3
+% for Springer journals.          Springer Heidelberg 2006/03/15
+%
+% Copy it to a new file with a new name and use it as the basis
+% for your article. Delete % signs as needed.
+%
+% This template includes a few options for different layouts and
+% content for various journals. Please consult a previous issue of
+% your journal as needed.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+%!PS-Adobe-3.0 EPSF-3.0
+%%BoundingBox: 19 19 221 221
+%%CreationDate: Mon Sep 29 1997
+%%Creator: programmed by hand (JK)
+%%EndComments
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}         % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+% \usepackage{url,graphicx,wrapfig,proof,natbib,stmaryrd,xspace}
+
+% \usepackage{amssymb,amsmath,amscd} 
+% \usepackage{verbatim,listings}
+
+% \usepackage{mathpartir}
+% \usepackage{prooftree}
 
@@
 
-\title{ Formalization, Mechanization and Automation % of Variants
+\begin{document}
+
+\title{Formalization, Mechanization and Automation 
   of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
@@
 \author{
-  Christoph Benzm\"{u}ller\inst{1} 
+  Christoph Benzm\"{u}ller
   \and 
-  Bruno Woltzenlogel Paleo\inst{2}
+  Bruno Woltzenlogel Paleo
 }
 
-\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}
+%\authorrunning{Short form of author list} % if too long for running head
 
-\institute{
-  Dahlem Center for Intelligent Sstems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
+\institute{Christoph Benzm\"uller \at
+             Dahlem Center for Intelligent Systems, Freie
+             Universit\"at Berlin, Germany,
+              \email{c.benzmueller@fu-berlin.de}           %  \\
+%             \emph{Present address:} of F. Author  %  if needed
            \and
-  Theory and Logic Group, Vienna University of Technology, Austria \\
+           Bruno Woltzenlogel Paleo \at  Theory and Logic Group, Vienna University of Technology, Austria, 
   \email{bruno@logic.at}
@@
 
-\begin{document}
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 
+% \begin{abstract}
+%  G\"{o}del's ontological proof has been verified and automated with proof assistants and automated theorem provers.
+% \end{abstract}
+
+% \section*{}
 Attempts to prove the existence (or non-existence) of God by means of
@@
 following order:\footnote{ A1, A2, A5, D1, D3 are logically
-  equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and
+  equivalent to, respectively, axioms 2, 5 and 4 and definitions 1 and
   3 in G\"odel's notes \cite{Goedel1970,GoedelNotes}. 
@@
   D1 in a logic with infinitary conjunction. 
-  A4 is a weaker formof G\"odel's axiom 3. 
+  A4 is a weaker form of G\"odel's axiom 3. 
   D2 has an extra conjunct $\phi(x)$ lacking in G\"odel's definition 2; 
@@
 with an unprecedent degree of detail 
-and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
+and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository} for all sources. 
 The following has been done (and in this order):
@@
 \item The basic modal logic K is sufficient for proving T1, C and T2. 
-%\marginpar{\scriptsize{Bruno: check this. Chris: Done.}}
 \item Modal logic S5 is not needed for proving T3; the logic KB is
   sufficient. 
-% \marginpar{\scriptsize{Bruno: Maybe cite a reference for S5 and KB?}}
 \item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
@@
 
-% \marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
-We could try to improve it, but I think simply removing it would also be satisfactory.
-The sentence about Leibniz' dictum already provides a very nice ``climax'' for the end of the paper.
-Anything that we say after it feels like an ``anticlimax''...}}
-% Future work shall widen our computer-assisted approach to the
-% study of other ontological arguments. Variations of the distinctive features of
-% the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
-% embeddings-based approach, will
-% be considered.
-
 %\footnotesize
+%\bibliographystyle{spbasic}
 \bibliographystyle{plain}
 %\bibliography{Bibliography}
+
 \begin{thebibliography}{10}

### 2013-09-07 Abstract/Square/GodProof-Abstract-SquareVersion.tex
Move abstract versions to corresponding subfolders

--- 2013-08-21 Abstract/GodProof-Abstract.tex
@@
 calculate!
-
-% \marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
-We could try to improve it, but I think simply removing it would also be satisfactory.
-The sentence about Leibniz' dictum already provides a very nice ``climax'' for the end of the paper.
-Anything that we say after it feels like an ``anticlimax''...}}
-% Future work shall widen our computer-assisted approach to the
-% study of other ontological arguments. Variations of the distinctive features of
-% the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
-% embeddings-based approach, will
-% be considered.
 

### 2013-09-10 Abstract/ArXiv/GodProof-Abstract.tex
corrected typo

--- 2013-09-07 Abstract/Square/GodProof-Abstract-SquareVersion.tex
@@
 
-\title{ Formalization, Mechanization and Automation % of Variants
+\title{Formalization, Mechanization and Automation % of Variants
   of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
     supported by the German Research Foundation under grant
-    BE2501/9-1.}  }
+    BE2501/9-1.}}
 
@@
 \institute{
-  Dahlem Center for Intelligent Sstems, Freie Universit\"{a}t Berlin, Germany\\
+  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
   \email{c.benzmueller@gmail.com}

### 2013-09-12 Natural Deduction/GodProof-ND.tex
Add ToDo
Fix mistake in Pi_1
Add Annika's KB proof

@@
 \section{Introduction}
+
+ToDo: Do also Scott's and Gödel's proofs.
 
@@
        \AXC{Axiom \ref{A1}} \dashedLine
-       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall_E$}  
+       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\all_E$}  
        \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}   
@@
 		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
-              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp_E$}
+              	\UIC{$P(\neg \psi)\imp \neg \psi(x)$}\RightLabel{$\imp_E$}
            \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}  
@@
 Necessarily, God exists:
+
+\noindent
+\textbf{Theorem:} The existence of a God is necessary:
 $$
@@
 \end{theorem}
-\begin{proof}\hfill
+
+\noindent
+\textbf{Formal proof:}
+Let the following derivation be $ \Pi$:
+
 \begin{small}
 \begin{prooftree}
-\AXC{\textbf{S5}} \dashedLine
-\UIC{$ \all \varphi. [\pos \ldots \pos \nec \varphi \biimp \nec \varphi] $}
-\UIC{$ \pos \nec \ex x. G(x) \biimp \nec \ex x. G(x) $}
-        \AXC{Corollary \ref{C1}} \dashedLine
-        \UIC{$\pos \ex x. G(x)$}
-                \AXC{Lemma \ref{L1}} \dashedLine
-                \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$} \doubleLine
-            \BIC{$\pos \nec \ex x. G(x)$}
-    \BIC{$\nec \ex x. G(x)$}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\nec \ex x. G(x)$} \RightLabel{$\nec$ I}
+\UIC{$\nec\nec \ex x. G(x)$}\RightLabel{$\imp$ I, 1}
+\UIC{$\nec \ex x. G(x)\imp\nec\nec \ex x. G(x)$} \RightLabel{propositional logic}
+\UIC{$\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x)$}\RightLabel{$\nec$ I}
+\UIC{$\nec(\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x))$}
+\AXC{Axiom K} \dashedLine 
+\BIC{$\nec\neg\nec\nec \ex x. G(x)\imp\nec\neg  \nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
+ \UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
+
 \end{prooftree}
 \end{small}
-\end{proof}
 
-\section{God exists}
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\nec\pos\neg\nec \ex x. G(x)$}
+\AXC{$ \Pi$} \dashedLine 
+\UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
+ \BIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
+\UIC{$\neg\nec \ex x. G(x)$} 
+\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
+\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
+ \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
+\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
+\UIC{$\neg \pos\ex x. G(x)$}
+
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $}\dashedLine
+\UIC{$\neg \pos\ex x. G(x)$} 
+\AXC{ Corollary \ref{C1}} \dashedLine 
+\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
+ \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
+ \UIC{$\neg\nec\pos\neg\nec \ex x. G(x)$}
+\AXC{ Axiom B} \dashedLine 
+\UIC{$ \neg\nec \ex x. G(x)\imp \nec\pos\neg\nec \ex x. G(x)$} \RightLabel{propositional logic}
+\UIC{$ \neg \nec\pos\neg\nec \ex x. G(x)\imp  \nec \ex x. G(x)$} \RightLabel{$\imp$ E}
+ \BIC{$ \nec \ex x. G(x)$}
+\end{prooftree}
+\end{small}
+
+\textbf{Better proof, that directly proves $\ex x. G(x)$:}
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+
+ \UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
+\UIC{$\neg\nec \ex x. G(x)$} 
+\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
+\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
+ \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
+\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
+\UIC{$\neg \pos\ex x. G(x)$} 
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $}\dashedLine
+\UIC{$\neg \pos\ex x. G(x)$} 
+\AXC{ Corollary \ref{C1}} \dashedLine 
+\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
+ \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
+ \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
+ \UIC{$\neg\nec\pos \neg \ex x. G(x)$}
+\AXC{ Axiom B} \dashedLine 
+\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \RightLabel{propositional logic}
+\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp  \neg  \neg \ex x. G(x)$} \RightLabel{$\imp$ E}
+ \BIC{$ \neg  \neg \ex x. G(x)$}\RightLabel{$ \neg  \neg$ E}
+ \UIC{$  \ex x. G(x)$}
+\end{prooftree}
+\end{small}
+
+Note that the last step is classical and we do not prove the existential statement by providing an object for which the statement holds. 
+This proof makes section \ref{GodsE} superflous and the use of axiom "M" unnecessary.
+
+The system used contains the $\nec$E-rule with the restriction that we have a $\nec$I below it.
+This is equivalent to modal sytem K that contains axiom K and necessitation rule N.
+We aslo use axiom B ($ A\imp\nec\pos A$).
+No other modal axioms are needed.
+
+\section{God exists}\label{GodsE}
 
@@
 
+\section{Proof system equivalent to system K}
+
+We show that the proof system with boxed parts of derivations is equivalent to the system K. The modal system K consists of the axiom K and the necessitation rule N.
+
+\begin{axiom}
+\label{AK}
+The transitivity axiom K. 
+$$
+\nec(A\imp B)\imp  (\nec A\imp \nec B)
+$$
+\end{axiom}
+
+\begin{axiom}
+\label{N}
+The necessitation rule N.
+
+If $A$ is a theorem, then $\nec A$ is a theorem. 
+\end{axiom}
+
+\begin{lemma}
+\label{systemK}
+The axiom K is derivable in the system. 
+\end{lemma}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec$ E}
+\UIC{$A\imp B$} 
+\AXC{$\nec A ^1 $}\RightLabel{$\nec$ E}
+\UIC{$A$} \RightLabel{$\imp$ E}
+ \BIC{$ B$} \RightLabel{$\nec$I}
+\UIC{$ \nec B$} \RightLabel{$\imp$ I, 1}
+ \UIC{$\nec A\imp \nec B$}  \RightLabel{$\imp$ I, 2}
+ \UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$}
+\end{prooftree}
+\end{small}
+
+\begin{lemma}
+\label{systemK}
+Assuming the axiom K and the necessitation rule N, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
+\end{lemma}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$A^1 $}\dashedLine
+\UIC{$ B$} \RightLabel{$\imp$ I, 1}
+\UIC{$A\imp B$} \RightLabel{rule N}
+\UIC{$\nec(A\imp B)$} 
+\AXC{Axiom K}\dashedLine
+\UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$} \RightLabel{$\imp$ E}
+\BIC{$\nec A\imp \nec B$} 
+\AXC{$\nec A ^1 $}\RightLabel{$\imp$ E}
+\BIC{$\nec B$}
+\end{prooftree}
+\end{small}
+
 \end{document}

### 2013-09-12 Abstract/Journal/GodProof-Abstract-JournalVersion.tex

@@
 
-\institute{Christoph Benzm\"uller \at
-             Dahlem Center for Intelligent Systems, Freie
-             Universit\"at Berlin, Germany,
-              \email{c.benzmueller@fu-berlin.de}           %  \\
+\institute{Christoph Benzm\"uller,
+             %Dahlem Center for Intelligent Systems, 
+             Freie Universit\"at Berlin, Germany, 
+             \email{c.benzmueller@fu-berlin.de}          \\
 %             \emph{Present address:} of F. Author  %  if needed
-           \and
-           Bruno Woltzenlogel Paleo \at  Theory and Logic Group, Vienna University of Technology, Austria, 
+           %\and
+           Bruno Woltzenlogel Paleo, 
+           % Theory and Logic Group, 
+           Vienna University of Technology, Austria,
            \email{bruno@logic.at}
@@
   $\all \phi [P(\phi) \to \Box \; P(\phi)]$
-\item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
+\item[D2] An \emph{essence} of an individual is a property possessed by it \\ and necessarily implying any of its properties:
   \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
@@
 \newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
+  Birthday}}, pp. 386--406. College Publications.
 
@@
   logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
+\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pp.
   162--170. Springer, 2008.
@@
   relational model finder.
-\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pp. 131--146.
   Springer, 2010.
@@
 \newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117.
   Springer, 2012.
@@
 K.~G\"odel.
-\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pp. 144--145.
 \newblock In  \cite{sobel2004logic}, 2004.
@@
 D.~Scott.
-\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
+\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pp. 145--146.
 \newblock In  \cite{sobel2004logic}, 2004.

### 2013-09-12 Natural Deduction/GodProof-ND.tex
Add draft definition of package

@@
 \usepackage{latexsym}
+\usepackage{proof}
 \usepackage{bussproofs}
@@
 
-\newtheorem{axiom}{Axiom}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-
-\newenvironment{proof}[1][Proof]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+\usepackage{calculi}
+\usepackage{theorems}
 
@@
 
-ToDo: Show and explain here the rules of the calculus we are using.
+\newcommand{\s}{\qquad}
 
-ToDo: We should use a calculus for the basic modal logic K. Everything else should be stated as axioms.
+\begin{calculus}
+{The natural deduction calculus \ND}
+{fig:ND}
 
-ToDo: Investigate the relationship between the $\nec_E$ rule and the $M$ axiom (reflexivity). We don't want $M$ to be provable in our calculus.
+\vspace{1em}
 
-ToDo: cite a paper that proves soundness and completeness for this calculus.
+\s\s\s
+\infer[\imp_I^n]{A \imp B}{ \infer*{B}{A^n} }
+\s\s\s
+\infer[\imp_E]{B}{A & A \imp B}
+
+\vspace{2em}
+
+\s\s
+\infer[\wedge_I]{A \wedge B}{A & B}
+\s\s
+\infer[\wedge_{E_1}]{A}{A \wedge B}
+\s\s
+\infer[\wedge_{E_2}]{B}{A \wedge B}
+
+\vspace{2em}
+
+\s
+\infer[\all_I]{\all x. A[x]}{ A[\alpha] }
+\s
+\infer[\all_E]{A[t]}{ \all x. A[x] }
+\s\s
+\infer[\ex_I]{\ex x. A[x]}{ A[t] }
+\s
+\infer[\ex_E]{A[\alpha]}{ \ex x. A[x] }
+
+\vspace{1em}
+
+ToDo: State eigen-variable conditions
+\end{calculus}
+
+\begin{calculus}
+{\NDK: Rules for Modal Operators}
+{fig:NDK}
+\s\s\s
+\infer[\nec_I]{\nec A}{A}
+\s\s
+\infer[\nec_E]{A}{\nec A}
+
+\vspace{1em}
+
+ToDo: state conditions
+\end{calculus}
 

### 2013-09-12 Natural Deduction/GodProof-ND.tex
Add draft definition of the natural deduction calculus

@@
 
+\usepackage{fancybox}
 \usepackage{latexsym}
@@
 \begin{calculus}
-{The natural deduction calculus \ND}
+{The intuitionistic natural deduction calculus \ND}
 {fig:ND}
@@
 
-\s\s\s
-\infer[\imp_I^n]{A \imp B}{ \infer*{B}{A^n} }
-\s\s\s
+\s\s
+\infer[\imp_I]{A \imp B}{ B }
+\s\s
+\infer[\imp_I^n]{A \imp B}{ \infer*{B}{\infer[n]{A}{}} }
+\s\s
 \infer[\imp_E]{B}{A & A \imp B}
@@
 
+\s\s
+\infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
+\s\s
+\infer[\vee_{I_1}]{A \vee B}{A}
+\s\s
+\infer[\vee_{I_2}]{A \vee B}{B}
+
+\vspace{2em}
+
 \s
@@
 \s
-\infer[\ex_E]{A[\alpha]}{ \ex x. A[x] }
+\infer[\ex_E]{A[\beta]}{ \ex x. A[x] }
 
@@
 
-ToDo: State eigen-variable conditions
+ToDo: add intuitionistic rule for negation/bottom
 \end{calculus}
@@
 \begin{calculus}
-{\NDK: Rules for Modal Operators}
+{Classical Rules/Axioms}
+{fig:Classical}
+
+ToDo: add classical rules/axioms that we need
+\end{calculus}
+
+\begin{calculus}
+{Rules for Modal Operators}
 {fig:NDK}
-\s\s\s
-\infer[\nec_I]{\nec A}{A}
-\s\s
-\infer[\nec_E]{A}{\nec A}
 
@@
 
-ToDo: state conditions
+\s\s\s\s
+\infer[\nec_I]{\nec A}{\fbox{\infer*{A}{} }}
+\s\s\s
+\infer[\nec_E]{\fbox{ \infer*{}{A} }  }{\nec A}
+
+\vspace{1em}
 \end{calculus}
+
+\noindent
+A \emph{derivation} is a directed acyclich graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures \ref{fig:ND} and \ref{fig:NDK}.
+Parts of a derivation may be surrounded by boxes.
+A \emph{proof} is a derivation that additionally satisfies the following conditions:
+
+\begin{itemize}
+\item \textbf{eigen-variable conditions:}
+if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
+if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+%
+\item \textbf{boxed assumption condition:} any assumption should be discharged within the box where it is made.
+%
+\item \textbf{unboxed root condition:} the proof's root should not be inside any box.
+\end{itemize}
+
+\noindent
+Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
+Dashed lines are used to refer to a proof shown elsewhere.
+Dotted lines are used to indicate folding and unfolding of definitions.
 
@@
 
-ToDo: This section still needs more details. See Coq formalization for more details.
-
 ToDo: this is proven in a way that is slightly different from G\"odel's 1970.
@@
 Necessarily, God exists:
-
-\noindent
-\textbf{Theorem:} The existence of a God is necessary:
 $$
@@
 \AXC{$ $} \RightLabel{1}
-\UIC{$\nec \ex x. G(x)$} \RightLabel{$\nec$ I}
-\UIC{$\nec\nec \ex x. G(x)$}\RightLabel{$\imp$ I, 1}
+\UIC{$\nec \ex x. G(x)$} \RightLabel{$\nec_I$}
+\UIC{$\nec\nec \ex x. G(x)$}\RightLabel{$\imp_I^1$}
 \UIC{$\nec \ex x. G(x)\imp\nec\nec \ex x. G(x)$} \RightLabel{propositional logic}
-\UIC{$\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x)$}\RightLabel{$\nec$ I}
+\UIC{$\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
 \UIC{$\nec(\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x))$}
 \AXC{Axiom K} \dashedLine 
-\BIC{$\nec\neg\nec\nec \ex x. G(x)\imp\nec\neg  \nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
+\BIC{$\nec\neg\nec\nec \ex x. G(x)\imp\nec\neg  \nec \ex x. G(x)$} \dottedLine \RightLabel{$\pos$}
  \UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
@@
 \UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
- \BIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
+ \BIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec_E$, axiom T}
 \UIC{$\neg\nec \ex x. G(x)$} 
 \AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
-\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
- \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
-\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
+\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp_E$}
+ \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec_I$}
+\UIC{$\nec \neg\ex x. G(x)$}  \dottedLine
 \UIC{$\neg \pos\ex x. G(x)$}
@@
 \AXC{ Corollary \ref{C1}} \dashedLine 
-\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
- \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
+\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp_E$}
+ \BIC{$ \bot$} \RightLabel{$\imp_I^1$}
  \UIC{$\neg\nec\pos\neg\nec \ex x. G(x)$}
 \AXC{ Axiom B} \dashedLine 
-\UIC{$ \neg\nec \ex x. G(x)\imp \nec\pos\neg\nec \ex x. G(x)$} \RightLabel{propositional logic}
-\UIC{$ \neg \nec\pos\neg\nec \ex x. G(x)\imp  \nec \ex x. G(x)$} \RightLabel{$\imp$ E}
+\UIC{$ \neg\nec \ex x. G(x)\imp \nec\pos\neg\nec \ex x. G(x)$} \doubleLine
+\UIC{$ \neg \nec\pos\neg\nec \ex x. G(x)\imp  \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
  \BIC{$ \nec \ex x. G(x)$}
@@
 \AXC{$ $} \RightLabel{1}
-
- \UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
+\UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec_E$, axiom T}
 \UIC{$\neg\nec \ex x. G(x)$} 
-\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
-\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
- \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
-\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
+\AXC{ Lemma \ref{L1} } \dashedLine
+\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp_E$}
+ \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec_I$}
+\UIC{$\nec \neg\ex x. G(x)$}  \dottedLine
 \UIC{$\neg \pos\ex x. G(x)$} 
@@
 \AXC{ Corollary \ref{C1}} \dashedLine 
-\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
- \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
- \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
+\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp_E$}
+ \BIC{$ \bot$} \RightLabel{$\imp_I^1$}
+ \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \dottedLine
  \UIC{$\neg\nec\pos \neg \ex x. G(x)$}
 \AXC{ Axiom B} \dashedLine 
-\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \RightLabel{propositional logic}
-\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp  \neg  \neg \ex x. G(x)$} \RightLabel{$\imp$ E}
+\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \doubleLine
+\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp  \neg  \neg \ex x. G(x)$} \RightLabel{$\imp_E$}
  \BIC{$ \neg  \neg \ex x. G(x)$}\RightLabel{$ \neg  \neg$ E}
@@
 
-The system used contains the $\nec$E-rule with the restriction that we have a $\nec$I below it.
+The system used contains the $\nec_E$-rule with the restriction that we have a $\nec_I$ below it.
 This is equivalent to modal sytem K that contains axiom K and necessitation rule N.
@@
 
-\begin{axiom}
-\label{AK}
-The transitivity axiom K. 
+\begin{axiom}[The transitivity axiom K]
+\label{K}
 $$
@@
 
-\begin{axiom}
-\label{N}
-The necessitation rule N.
-
+\begin{axiom}[The necessitation rule]
+\label{Necessitation}
 If $A$ is a theorem, then $\nec A$ is a theorem. 
@@
 \begin{prooftree}
-\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec$ E}
+\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec_E$}
 \UIC{$A\imp B$} 
-\AXC{$\nec A ^1 $}\RightLabel{$\nec$ E}
-\UIC{$A$} \RightLabel{$\imp$ E}
- \BIC{$ B$} \RightLabel{$\nec$I}
-\UIC{$ \nec B$} \RightLabel{$\imp$ I, 1}
- \UIC{$\nec A\imp \nec B$}  \RightLabel{$\imp$ I, 2}
+      \AXC{$\nec A ^1 $}\RightLabel{$\nec_E$}
+      \UIC{$A$} \RightLabel{$\imp_E$}
+   \BIC{$ B$} \RightLabel{$\nec_I$}
+   \UIC{$ \nec B$} \RightLabel{$\imp_I^1$}
+   \UIC{$\nec A\imp \nec B$}  \RightLabel{$\imp_I^2$}
    \UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$}
@@
 \begin{prooftree}
-\AXC{$A^1 $}\dashedLine
-\UIC{$ B$} \RightLabel{$\imp$ I, 1}
-\UIC{$A\imp B$} \RightLabel{rule N}
+\AXC{$A^1 $}\noLine
+\UIC{$\vdots$}\noLine
+\UIC{$ B$} \RightLabel{$\imp_I^1$}
+\UIC{$A\imp B$} \RightLabel{Necessitation}
 \UIC{$\nec(A\imp B)$} 
       \AXC{Axiom K}\dashedLine
-\UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$} \RightLabel{$\imp$ E}
+      \UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$} \RightLabel{$\imp_E$}
   \BIC{$\nec A\imp \nec B$} 
-\AXC{$\nec A ^1 $}\RightLabel{$\imp$ E}
+        \AXC{$\nec A ^1 $}\RightLabel{$\imp_E$}
     \BIC{$\nec B$}

### 2013-09-13 Abstract/Square/GodProof-Abstract-SquareVersion.tex
Reduce size of Square version

@@
 
-\title{ Formalization, Mechanization and Automation % of Variants
-  of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
+\title{ Formalization and Automated Verification of % Variants of
+    \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
     supported by the German Research Foundation under grant
@@
 on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-
-Dana Scott's version of G\"odel's proof \cite{ScottNotes} employs the
-following axioms (\textbf{A}), definitions (\textbf{D}), corollaries
-(\textbf{C}) and theorems (\textbf{T}), and it proceeds in the
-following order:\footnote{ A1, A2, A5, D1, D3 are logically
-  equivalent to, respectively, axioms 2, 5 and 4 and definitions 1 and
-  3 in G\"odel's notes \cite{Goedel1970,GoedelNotes}. 
-  A3 was introduced by Scott \cite{ScottNotes} and 
-  could be derived from G\"odel's axiom 1 and
-  D1 in a logic with infinitary conjunction. 
-  A4 is a weaker form of G\"odel's axiom 3. 
-  D2 has an extra conjunct $\phi(x)$ lacking in G\"odel's definition 2; 
-  this is believed to have been 
-  an oversight by G\"odel \cite{Hazen}.}
-
-\allowdisplaybreaks[1] 
-\begin{description}
-\item[A1] Either a property or its negation is positive, but not
-  both:  \hfill 
-  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$
-\item[A2] A property necessarily implied \\ by a
-  positive property is positive: \phantom{b} \hfill 
-  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]$ 
-\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$
-\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
-  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$
-\item[A3]  The property of being God-like is positive: \hfill   $P(G)$
-\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$
-\item[A4]  Positive properties are necessarily positive: \hfill 
-  $\all \phi [P(\phi) \to \Box \; P(\phi)]$
-\item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
-  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
-  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$
-\item[T2]  Being God-like is an essence of any
-  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ 
-\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
-  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$
-\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
-\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
-\end{description}
-
-\noindent
-Scott's version of G\"{o}del's proof has now been 
-analysed for the first-time
+(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). We have analyzed Scott's version of G\"{o}del's proof for the first-time
 with an unprecedent degree of detail 
@@
 calculate!
+
+\clearpage
 

### 2013-09-13 Abstract/Square/GodProof-Abstract-SquareVersion.tex
further reduction to 1 page

@@
 
-\usepackage{url,amsmath,amssymb,a4wide}
+\usepackage{amsmath,amssymb,a4wide}
 
@@
   Dahlem Center for Intelligent Sstems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
+ % \email{c.benzmueller@gmail.com}
   \and 
   Theory and Logic Group, Vienna University of Technology, Austria \\
-  \email{bruno@logic.at}
+ %  \email{bruno@logic.at}
 }
@@
 been considered by G\"{o}del and by several philosophers who commented
-on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). We have analyzed Scott's version of G\"{o}del's proof for the first-time
+on his proof. %
+%(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography})
+We have analyzed Scott's version of G\"{o}del's proof \cite{ScottNotes} for the first-time
 with an unprecedent degree of detail 
-and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
-The following has been done (and in this order):
-\begin{itemize}
-\item A detailed natural deduction proof.
+and formality with the help of theorem provers; cf.~\url{https://github.com/FormalTheology/GoedelGod}.
+
+The following has been done (and in this order):\footnote{System URLs:
+ TPTP---\url{http://tptp.org};
+ Nitpick---\url{http://www4.in.tum.de/\~{}blanchet/nitpick.html};
+ LEO-II---\url{http://leoprover.org};
+ Satallax---\url{http://www.ps.uni-saarland.de/~cebrown/satallax/}; 
+ Coq---\url{http://coq.inria.fr};
+ Isabelle---\url{http://isabelle.in.tum.de}.
+}
+(i) a detailed natural deduction proof;
 %
-\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
+(ii) a formalization in TPTP THF syntax;
 %
-\item Automatic verification of the consistency of the axioms and 
-definitions with Nitpick \cite{Nitpick}.
+(iii) an automatic verification of the consistency of the axioms and 
+definitions with Nitpick;
 %
-\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.
-
-\item A step-by-step formalization using the Coq proof assistant \cite{Coq}.
-
-\item A formalization using the Isabelle proof assistant \cite{Isabelle}, where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
-\end{itemize}
+(iv) an automatic demonstration of the theorems with the provers LEO-II and Satallax;
+%
+(v) a step-by-step formalization using the Coq proof assistant;
+%
+(vi) a formalization using the Isabelle proof assistant, where the
+theorems (and some additional lemmata) have been automated with the
+Isabelle tools Sledgehammer and Metis.
 
@@
 embedding of the modal logic into classical higher-order logic with
-Henkin semantics \cite{J23,B9}. The formalization is thus essentially
+Henkin semantics. The formalization is thus essentially
 done in classical higher-order logic where  quantified modal logic is
@@
 
-In our ongoing computer-assisted study of G\"odel's proof we have
-obtained the following results:
-\begin{itemize}
-\item The basic modal logic K is sufficient for proving T1, C and T2. 
-\item Modal logic S5 is not needed for proving T3; the logic KB is
+In our ongoing computer-assisted study of G\"odel's proof our proof
+tools have made
+some interesting observations, including:
+(a) The basic modal logic K is sufficient for proving the first three
+theorems (T1, Coro and T2) as outlined in Scott's notes.
+(b)  For proving the final theorem (T3), logic KB is
   sufficient.
-\item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
-  and definitions would be inconsistent.
-\item For proving theorem T1, only the left to right direction of
-  axiom A1 is needed. However, the backward direction of A1 is
-  required for proving T2.
-
-\end{itemize}
+(c) G\"odel's original  version of the proof \cite{GoedelNotes}, which omits
+conjunct $\phi(x)$ in definition of \textit{essence}, seems inconsistent.
+% (d) for proving theorem T1, only the left to right direction of
+%   axiom A1 is needed. However, the backward direction of A1 is
+%   required for proving T2.
+% \end{itemize}
 
@@
 
-\clearpage
+\small
 
@@
 
-\bibitem{Adams}
-R.M. Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1995.
+% \bibitem{Adams}
+% R.M. Adams.
+% \newblock Introductory note to *1970.
+% \newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+%   Letters}}. Oxford University Press, 1995.
 
-\bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings.
-\newblock G\"odel ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
+% \bibitem{AndersonGettings}
+% A.C. Anderson and M.~Gettings.
+% \newblock G\"odel ontological proof revisited.
+% \newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
+%   Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
 
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C. Paulson.
-\newblock Exploring properties of normal multimodal logics in simple type
-  theory with {LEO-II}.
-\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
+% \bibitem{B9}
+% C.~Benzm{\"u}ller and L.C. Paulson.
+% \newblock Exploring properties of normal multimodal logics in simple type
+%   theory with {LEO-II}.
+% \newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
+%   Birthday}}, pages 386--406. College Publications.
 
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C. Paulson.
-\newblock Quantified multimodal logics in simple type theory.
-\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
-  7(1):7--20, 2013.
+% \bibitem{J23}
+% C.~Benzm{\"u}ller and L.C. Paulson.
+% \newblock Quantified multimodal logics in simple type theory.
+% \newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
+%   7(1):7--20, 2013.
 
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
-\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
-  logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
-  162--170. Springer, 2008.
+% \bibitem{LEO-II}
+% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
+% \newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
+%   logic.
+% \newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
+%   162--170. Springer, 2008.
 
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran.
-\newblock {\em {Interactive Theorem Proving and Program Development}}.
-\newblock Springer, 2004.
+% \bibitem{Coq}
+% Y.~Bertot and P.~Casteran.
+% \newblock {\em {Interactive Theorem Proving and Program Development}}.
+% \newblock Springer, 2004.
 
-\bibitem{Sledgehammer}
-J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
-\newblock Extending {Sledgehammer} with {SMT} solvers.
-\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
+% \bibitem{Sledgehammer}
+% J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+% \newblock Extending {Sledgehammer} with {SMT} solvers.
+% \newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
 
-\bibitem{Nitpick}
-J.C. Blanchette and T.~Nipkow.
-\newblock Nitpick: A counterexample generator for higher-order logic based on a
-  relational model finder.
-\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
-  Springer, 2010.
+% \bibitem{Nitpick}
+% J.C. Blanchette and T.~Nipkow.
+% \newblock Nitpick: A counterexample generator for higher-order logic based on a
+%   relational model finder.
+% \newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+%   Springer, 2010.
 
-\bibitem{Satallax}
-C.E. Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
-  Springer, 2012.
+% \bibitem{Satallax}
+% C.E. Brown.
+% \newblock Satallax: An automated higher-order prover.
+% \newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+%   Springer, 2012.
 
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary bibligraphy on the ontological proof
-  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
+% \bibitem{ContemporaryBibliography}
+% R.~Corazzon.
+% \newblock Contemporary bibligraphy on the ontological proof
+%   (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
 
-\bibitem{Fitting}
-M.~Fitting.
-\newblock {\em {Types, Tableaux and G\"odel's God}}.
-\newblock Kluver Academic Press, 2002.
+% \bibitem{Fitting}
+% M.~Fitting.
+% \newblock {\em {Types, Tableaux and G\"odel's God}}.
+% \newblock Kluver Academic Press, 2002.
 
@@
 
-\bibitem{Hazen}
-A.P. Hazen.
-\newblock On g\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+% \bibitem{Hazen}
+% A.P. Hazen.
+% \newblock On g\"odel's ontological proof.
+% \newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
 
-\bibitem{Hurd03first-orderproof}
-J.~Hurd.
-\newblock First-order proof tactics in higher-order logic theorem provers.
-\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
-  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
+% \bibitem{Hurd03first-orderproof}
+% J.~Hurd.
+% \newblock First-order proof tactics in higher-order logic theorem provers.
+% \newblock In {\em Design and Application of Strategies/Tactics in Higher Order
+%   Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
 
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel.
-\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
+% \bibitem{Isabelle}
+% T.~Nipkow, L.C. Paulson, and M.~Wenzel.
+% \newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
+% \newblock Number 2283 in LNCS. Springer, 2002.
 
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
+% \bibitem{FormalTheologyRepository}
+% B.~Woltzenlogel Paleo and C.~Benzm\"uller.
+% \newblock Formal theology repository
+%   (\url{https://github.com/FormalTheology/GoedelGod}).
 
@@
 
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller.
-\newblock Automated reasoning in higher-order logic using the {TPTP THF}
-  infrastructure.
-\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
+% \bibitem{J22}
+% G.~Sutcliffe and C.~Benzm{\"u}ller.
+% \newblock Automated reasoning in higher-order logic using the {TPTP THF}
+%   infrastructure.
+% \newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
 

### 2013-09-13 Abstract/Square/GodProof-Abstract-SquareVersion.tex
further reduction to 1 page

@@
  LEO-II---\url{http://leoprover.org};
- Satallax---\url{http://www.ps.uni-saarland.de/~cebrown/satallax/}; 
+ Satallax---\url{http://www.ps.uni-saarland.de/\~{}cebrown/satallax}; 
  Coq---\url{http://coq.inria.fr};
@@
 (c) G\"odel's original  version of the proof \cite{GoedelNotes}, which omits
-conjunct $\phi(x)$ in definition of \textit{essence}, seems inconsistent.
+conjunct $\phi(x)$ in definition of \emph{essence}, seems inconsistent.
 % (d) for proving theorem T1, only the left to right direction of

### 2013-09-17 Natural Deduction/GodProof-ND.tex
Add Annika's introduction and ND derivation of Modal Collapse

@@
 
-ToDo: Do also Scott's and Gödel's proofs.
+ToDo: Do also Scott's and G\"{o}del's proofs.
+
+The first ontological argument for the existence of God was written by St.
+Anselm (1033-1109).
+The argument considers a maximally concievable being.
+This being must exist, because if it did not have the property of existence, then we could concieve of a greater being that apart from the other properties also has the property of existence.
+The main critique of this argument is that we do not know whether the concept maximal concievable being in fact designates anything or if it is inconsistent, like a round square.
+The question is if the argument is sound.
+
+As Bertrand Russell has pointed out the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have.
+Kant, on the other hand argued against the ontological argument on the basis that existence is not an analytic property.
+This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
+All that we can say is that if God exists, then he necessarily exists.
+
+St.
+Anselm's proof was processed by Descartes and Leibniz.
+Leibniz identified the critical point of the argument as establishing the possible existence of God.
+Leibniz gave an argument for that the properties of God, the perfections, are compatible.
+This implies that it is possible to have all perfections at once and therefore the existence of a maximal being with all these properties is possible.
+
+G\"odel continued this argument by defining the properties that must hold for these perfections and giving axioms from which the existence of God can be derived.
+The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
+The formal argument of G\"odel is based on Leibniz proof, which in turn is based on Descartes proof.
+These proofs have two parts; a proof that if God's existence is possible, then it is necessary and a proof that God's existence is in fact possible.
+The deductive system used in the formal proof is a system for modal logic extended with second-order quantification (or possibly third-order quantification), because the axioms and definitions required for the concept of positive properties or perfections involve quantification over properties.
+
+G\"odel treated positive properties as not just atomic properties, like Leibniz, but also consisting of collections of these properties, which by Leibniz argument are compatible or possible.
+(Fitting, p.139) G\"odel fomulated an axiom stating that the conjunction of any set of positive properties is positive.
+The property of God-likeness, which defines what a God is, must therefore be an infinite conjunction of all the positive properties.
+This axiom of the positiveness of conjunctions can be given a third order fomulation which is the reason for the claim that third-order quantification may be required.
+(Fitting, 2002, p.
+148).
+However, this axiom is oly used in the proof to establish that the property of God-likeness is itself positive that led Scott to simply assume the positiveness of  God-likeness as an axiom instead.
+
+Nevertheless, both axioms have the unfortunate property of being equivalent to the possibility of God's existence, given the other axioms of G\"odel.
+The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself.
+Thus, we have an axiom which seems to be equivalent to the conclusion.
+
+If one assumes extensionality, that two objects with the same properties are identical, then the definitions of G\"odel's proof give monotheism.
+Namely, two Gods would both be God-like and by the definition of god-like, they would have the same positive properties and be identical.
+
+It may be so that for all formulas the formula implies its boxed version.
+That is all that is is necessarily and modality collapses.
+Fitting provides two solutions to the problem by modifying the axioms.
 
@@
 
-\section{Necessarily, God exists}
+\section{God exists}
 
@@
 \begin{theorem}
-\label{T3}
-Necessarily, God exists:
+\label{T4}
+God exists:
 $$
-\nec \ex x. G(x)
+\ex x. G(x)
 $$
@@
 \textbf{Formal proof:}
-Let the following derivation be $ \Pi$:
 
-\begin{small}
-\begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$\nec \ex x. G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec\nec \ex x. G(x)$}\RightLabel{$\imp_I^1$}
-\UIC{$\nec \ex x. G(x)\imp\nec\nec \ex x. G(x)$} \RightLabel{propositional logic}
-\UIC{$\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
-\UIC{$\nec(\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x))$}
-\AXC{Axiom K} \dashedLine 
-\BIC{$\nec\neg\nec\nec \ex x. G(x)\imp\nec\neg  \nec \ex x. G(x)$} \dottedLine \RightLabel{$\pos$}
- \UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
-
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$\nec\pos\neg\nec \ex x. G(x)$}
-\AXC{$ \Pi$} \dashedLine 
-\UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
- \BIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec_E$, axiom T}
-\UIC{$\neg\nec \ex x. G(x)$} 
-\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
-\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp_E$}
- \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec \neg\ex x. G(x)$}  \dottedLine
-\UIC{$\neg \pos\ex x. G(x)$}
-
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ $}\dashedLine
-\UIC{$\neg \pos\ex x. G(x)$} 
-\AXC{ Corollary \ref{C1}} \dashedLine 
-\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp_E$}
- \BIC{$ \bot$} \RightLabel{$\imp_I^1$}
- \UIC{$\neg\nec\pos\neg\nec \ex x. G(x)$}
-\AXC{ Axiom B} \dashedLine 
-\UIC{$ \neg\nec \ex x. G(x)\imp \nec\pos\neg\nec \ex x. G(x)$} \doubleLine
-\UIC{$ \neg \nec\pos\neg\nec \ex x. G(x)\imp  \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
- \BIC{$ \nec \ex x. G(x)$}
-\end{prooftree}
-\end{small}
-
-\textbf{Better proof, that directly proves $\ex x. G(x)$:}
 \begin{small}
@@
 
+\section{Necessarily, God exists}
+
+We can also prove that god exists necessarily in our system by simply introducing box on a theorem by rule N in modal sytem K. So we get:
+
+\begin{corollary}
+\label{T3}
+Necessarily, God exists:
+$$
+\nec \ex x. G(x)
+$$
+\end{corollary}
+
 \section{God exists}\label{GodsE}
+This section is superfluous.
 
@@
 
+\section{Crisis in the calculus}
+
+The axioms of the calculus imply the following collapse of modalities:
+
+\begin{theorem}
+\label{T5}
+For all constant fomulas (without free variables), $A$,  we have:
+$$
+A \imp \nec A
+$$
+\end{theorem}
+
+Note that in intuitionistic predicate logic we have $\all y.[B \imp C]  \biimp [\ex y.B \imp C]$ if $y$ is not free in $C$.
+
+\begin{small}
+\begin{prooftree}
+\AXC{Theorem \ref{T2}}\dashedLine
+\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }
+\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \RightLabel{Prop. logic }
+\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \RightLabel{Second order quantifier elimination and logical steps }
+\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
+\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \RightLabel{intuitionistic predicate logic}
+\UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+\begin{small}
+\begin{prooftree}
+\AXC{}\dashedLine
+\UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
+\AXC{Theorem \ref{T4}}\dashedLine
+\UIC{$\ex y.G(y)$} \RightLabel{$\imp_E$}
+\BIC{$ A \imp \nec \all x. (G(x) \imp A)$} \RightLabel{intuitionistic predicate logic}
+\UIC{$ A \imp \nec (\ex x. G(x) \imp A)$}
+\AXC{$ $} \RightLabel{1}
+\UIC{$A$} \RightLabel{$\imp_E$}
+\BIC{$ \nec (\ex x. G(x) \imp A)$} \RightLabel{$\nec_E$}
+\UIC{$ \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
+\AXC{Theorem \ref{T4}}\dashedLine
+\UIC{$\ex x.G(x)$} \RightLabel{$\imp_E$}
+\BIC{$A$}  \RightLabel{$\nec_I$}
+\UIC{$ \nec A$}  \RightLabel{$\imp_I^1$}
+   \UIC{$A\imp \nec A$}
+\end{prooftree}
+\end{small}
+
+This collapse of the sytem is proved in (Fitting, 2002) and he provides two solutions by the modifications of the axioms, see chapter 11 section 8-11.
+The argument may require that one doesn't instantiate the second order quantifier with $A$ directly but with $\lambda x.A$ to make $A$ a property.
+But still we have $(\lambda x.A)(y)\equiv A$ because $A$ is constant.
+
+The question arises what kind of second order quantification is allowed.
+The instantiations that are used in our proof are instantiated with predicate variables that later work as eigenvariables for the second order universal introduction.
+In the proof of corollary \ref{C1} we instantiate with $G$ and in the proof of lemma \ref{L1}.
+we instantiate with $G$ and $E$.
+
+In Sobel on G\"{o}del's ontological proof by Robert E. Koons this question is raised.
+
+One can also discuss if a second order modal logic is sound and complete.
+With Henkin models, which restrict the second order quantification with comprehension, we have that normal second order logic is sound and complete.
+There is a completeness proof for a second order modal logic found in A completeness theorem in second order modal logic by Cocchiarella.
+
 \end{document}

### 2013-10-28 Talks/FU-Berlin/GödelGod-Talk.tex
minor modifications
talk preparation and more press links

@@
 
+%\usepackage{rotating}
+
 % Sequent Calculus Proof Settings
@@
 
-\title[Formalizations of G\"{o}del's Proof of God's Existence]{Formalizations of G\"{o}del's Proof of God's Existence}
+\title[On G\"{o}del's Proof of God's
+Existence]{Formalization, Mechanization and Automation of \\ G\"{o}del's
+  Proof of God's Existence}
 
@@
 
-\institute[] {
+\institute[]{
   \inst{}%
@@
 
-\date[29.10.13]
-{2013 October ??}
+\date[1.11.2013]
+{November 1, 2013}
 
@@
 
+\begin{frame}{Introduction} \small
+\begin{minipage}{.56\textwidth} 
+\includegraphics[width=\textwidth]{spiegel1} 
+\vskip1em
+Germany \\
+- Telepolis \& Heise \\
+- Spiegel Online \\
+- FAZ \\
+- Die Welt \\
+- Berliner Morgenpost \\
+- Hamburger Abendpost \\
+- \ldots \\
+\end{minipage} \hfill
+%
+\begin{minipage}{.3\textwidth}
+Austria \\
+- myscience.at \\
+- Wiener Zeitung \\
+- ORF \\
+- \ldots \\
+
+Italy \\
+- Repubblica \\
+- Today.it \\
+- Ilsussidario \\
+- \ldots \\
+
+% Russia \\
+% - \ldots \\
+
+India \\
+- DNA India \\
+- Delhi Daily News \\
+- Indoa Today \\
+- \ldots \\
+
+International \\
+- Spiegel International \\
+- Yahoo Finance \\
+- CNET \\
+- United Press Intl. \\
+- \ldots
+\end{minipage}
+\end{frame}
+
+\begin{frame}{Introduction}\large
+%Ontological argument: Conclude that God exists from premises by pure, a priori reasoning.
+\begin{block}{Def: \textcolor{blue}{Ontological Argument/Proof}}
+* deductive argument \\
+
+* for the existence of god \\
+
+* starting from premises, which are justified by pure reasoning,
+i.e. they do not depend on observation in the world.
+\end{block}
+
+\vfill \pause
+Existence of God: different types of arguments/proofs\\[.2em]
+\begin{itemize}
+\item[---]a posteriori (use experience/observation in the world)
+  \begin{itemize}
+  \item[------]teleological
+  \item[------]cosmological
+  \item[------]moral
+  \item[------] \ldots
+  \end{itemize}  
+\item[---]a priori (based on pure reasoning, independent)
+  \begin{itemize}
+  \item[------]\textcolor{blue}{ontological argument}
+    \begin{itemize}
+    \item[------]definitional 
+    \item[------]modal 
+    \item[------] \ldots
+    \end{itemize}
+  \item[------]other a priori arguments
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Introduction}
+
+\emph{\huge Wohl eine jede Philosophie kreist um den ontologischen
+  Gottesbeweis} \\[2em]
+(Adorno, Th. W.: Negative Dialektik. Frankfurt a. M. 1966, p.378)
+\vfill
+\ldots \hfill
+\includegraphics[height=2cm]{buch3.jpg} \hfill
+\includegraphics[height=2cm]{buch2.jpg} \hfill 
+\includegraphics[height=2cm]{buch4.jpg} \hfill
+\includegraphics[height=2cm]{buch5.jpg} \hfill
+\includegraphics[height=2cm]{buch6.jpg} \hfill
+\includegraphics[height=2cm]{buch1.jpg} \hfill
+\ldots
+\end{frame}
+
+\begin{frame}{Introduction} \Large
+{Rich history on ontological arguments (\textcolor{blue}{pros} and \textcolor{red}{cons})}\\[1em]
+
+\ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
+\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
+\ldots\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
+                \rotatebox[origin = bl]{65}{\textcolor{blue}{Leibniz}}  \hskip-1.2em
+\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Hume}}  \hskip-1em
+          \rotatebox[origin = bl]{65}{\textcolor{red}{Kant}}  \hskip-.8em
+\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hegel}}  \hskip-1.3em
+\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Frege}}  \hskip-1.3em
+\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hartshorne}} \hskip-1.9em
+          \rotatebox[origin = bl]{65}{\textcolor{blue}{Malcolm}}  \hskip-1.4em
+          \rotatebox[origin = bl]{65}{\textcolor{red}{Lewis}}  \hskip-1em
+          \rotatebox[origin = bl]{65}{\textcolor{blue}{Plantinga}}  \hskip-1.6em
+          \rotatebox[origin = bl]{65}{\textcolor{blue}{G\"odel}}   \hskip-1.2em
+\ldots \\[1em]
+
+\pause
+\vfill
+Anselm's notion of God:\\
+\,\hfill \emph{``God is that, than which nothing greater can be
+  conceived.''} \\[1em]
+
+G\"odel's notion of God:\\
+\,\hfill \emph{``A God-like being possesses all `positive' properties.''} \\[1em]
+
+To show by logical reasoning: \\
+\,\hfill \emph{``(Necessarily) God exists.''} \\[1em]
+
+% \rnode{n2}{}\emph{}
+% \ncline[nodesep=2pt,linecolor=black,linewidth=1pt]{->}{n1}{n2}
+\end{frame}
+
+\begin{frame}{Introduction} \Large
+Different Interests in Ontological Arguments: \\[1em]
+\begin{itemize}
+\item \textcolor{blue}{Philosophical:} Boundaries of Mataphysics \& Epistemology
+  \begin{itemize}
+  \item We talk about a metaphysical concept (God), 
+  \item but we want to draw
+      a conclusion for the real world. \\[2em]
+  \end{itemize} 
+\item \textcolor{blue}{Theistic:} Successful argument should convince atheists. \\[2em]
+\item \textcolor{red}{Our:} Can computers (theorem provers) be used
+  \begin{itemize}
+  \item to formalize the definitions and axioms?
+  \item to verify the arguments step-by-step?
+  \item to fully automate (sub-)arguments? \\[1em]
+  \end{itemize}
+  \textcolor{red}{\emph{``Computer-assisted Theoretical Philosophy''}}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Introduction} \large
+Main Challenge: \hfill No theorem provers for \emph{Second-order Modal
+  Logic} \\[1em]
+
+Our Idea: \hfill Exploit an embedding in \emph{Classical Higher-order
+  Logic} \\
+\,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
+
+What we did \textcolor{blue}{(rough outline for remaining
+  presentation!)}: \\
+
+\begin{itemize}
+\item[A:] pen and paper: detailed natural deduction proof \hfill Bruno
+\item[B:] formalization: axioms, defs, thms  in \textsc{TPTP THF}
+\item[C:] consistency: automatic verification with \textsc{Nitpick} 
+\item[D:] proof automation: theorems provers Leo-II and \textsc{Satallax} 
+\item[E:] step-by-step verification: proof assistant \textsc{Coq} 
+\item[F:] automation \& verification: proof assistant
+  \textsc{Isabelle}
+\item 
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2013-10-28 Talks/FU-Berlin/talk-part-1.tex
FU talk

\begin{frame}{Introduction} \small
\begin{minipage}{.56\textwidth} 
\includegraphics[width=\textwidth]{spiegel1} 
\vskip1em
Germany \\
- Telepolis \& Heise \\
- Spiegel Online \\
- FAZ \\
- Die Welt \\
- Berliner Morgenpost \\
- Hamburger Abendpost \\
- \ldots \\
\end{minipage} \hfill
%
\begin{minipage}{.3\textwidth}
Austria \\
- myscience.at \\
- Wiener Zeitung \\
- ORF \\
- \ldots \\

Italy \\
- Repubblica \\
- Today.it \\
- Ilsussidario \\
- \ldots \\

% Russia \\
% - \ldots \\

India \\
- DNA India \\
- Delhi Daily News \\
- Indoa Today \\
- \ldots \\

International \\
- Spiegel International \\
- Yahoo Finance \\
- CNET \\
- United Press Intl. \\
- \ldots
\end{minipage}
\end{frame}

\begin{frame}{Introduction}\large
%Ontological argument: Conclude that God exists from premises by pure, a priori reasoning.
\begin{block}{Def: \textcolor{blue}{Ontological Argument/Proof}}
* deductive argument \\

* for the existence of god \\

* starting from premises, which are justified by pure reasoning,
i.e. they do not depend on observation in the world.
\end{block}

\vfill \pause
Existence of God: different types of arguments/proofs\\[.2em]
\begin{itemize}
\item[---]a posteriori (use experience/observation in the world)
  \begin{itemize}
  \item[------]teleological
  \item[------]cosmological
  \item[------]moral
  \item[------] \ldots
  \end{itemize}  
\item[---]a priori (based on pure reasoning, independent)
  \begin{itemize}
  \item[------]\textcolor{blue}{ontological argument}
    \begin{itemize}
    \item[------]definitional 
    \item[------]modal 
    \item[------] \ldots
    \end{itemize}
  \item[------]other a priori arguments
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Introduction}

\emph{\huge Wohl eine jede Philosophie kreist um den ontologischen
  Gottesbeweis} \\[2em]
(Adorno, Th. W.: Negative Dialektik. Frankfurt a. M. 1966, p.378)
\vfill
\ldots \hfill
\includegraphics[height=2cm]{buch3.jpg} \hfill
\includegraphics[height=2cm]{buch2.jpg} \hfill 
\includegraphics[height=2cm]{buch4.jpg} \hfill
\includegraphics[height=2cm]{buch5.jpg} \hfill
\includegraphics[height=2cm]{buch6.jpg} \hfill
\includegraphics[height=2cm]{buch1.jpg} \hfill
\ldots
\end{frame}

\begin{frame}{Introduction} \Large
{Rich history on ontological arguments (\textcolor{blue}{pros} and \textcolor{red}{cons})}\\[1em]

\ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
\ldots\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
                \rotatebox[origin = bl]{65}{\textcolor{blue}{Leibniz}}  \hskip-1.2em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Hume}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Kant}}  \hskip-.8em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hegel}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Frege}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hartshorne}} \hskip-1.9em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Malcolm}}  \hskip-1.4em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Lewis}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Plantinga}}  \hskip-1.6em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{G\"odel}}   \hskip-1.2em
\ldots \\[1em]

\pause
\vfill
Anselm's notion of God:\\
\,\hfill \emph{``God is that, than which nothing greater can be
  conceived.''} \\[1em]

G\"odel's notion of God:\\
\,\hfill \emph{``A God-like being possesses all `positive' properties.''} \\[1em]

To show by logical reasoning: \\
\,\hfill \emph{``(Necessarily) God exists.''} \\[1em]

% \rnode{n2}{}\emph{}
% \ncline[nodesep=2pt,linecolor=black,linewidth=1pt]{->}{n1}{n2}
\end{frame}

\begin{frame}{Introduction} \Large
Different Interests in Ontological Arguments: \\[1em]
\begin{itemize}
\item \textcolor{blue}{Philosophical:} Boundaries of Mataphysics \& Epistemology
  \begin{itemize}
  \item We talk about a metaphysical concept (God), 
  \item but we want to draw
      a conclusion for the real world. \\[2em]
  \end{itemize} 
\item \textcolor{blue}{Theistic:} Successful argument should convince atheists. \\[2em]
\item \textcolor{red}{Our:} Can computers (theorem provers) be used
  \begin{itemize}
  \item to formalize the definitions and axioms?
  \item to verify the arguments step-by-step?
  \item to fully automate (sub-)arguments? \\[1em]
  \end{itemize}
  \textcolor{red}{\emph{``Computer-assisted Theoretical Philosophy''}}
\end{itemize}
\end{frame}

\begin{frame}{Introduction} \large
Main Challenge: \hfill No theorem provers for \emph{Second-order Modal
  Logic} \\[1em]

Our Idea: \hfill Exploit an embedding in \emph{Classical Higher-order
  Logic} \\
\,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]

What we did \textcolor{blue}{(rough outline for remaining
  presentation!)}: \\

\begin{itemize}
\item[A:] pen and paper: detailed natural deduction proof 
\item[B:] formalization: axioms, defs, thms  in \textsc{TPTP THF}
\item[C:] consistency: automatic verification with \textsc{Nitpick} 
\item[D:] proof automation: theorems provers Leo-II and \textsc{Satallax} 
\item[E:] step-by-step verification: proof assistant \textsc{Coq} 
\item[F:] automation \& verification: proof assistant
  \textsc{Isabelle}
\item[ ] Conclusion 
\end{itemize}
Did we get anything new? \ldots Probably yes --- let's discuss afterwards!
\end{frame}

### 2013-10-28 Talks/FU-Berlin/talk-part-6.tex
FU talk

\begin{frame}{Conclusion} \small
todo
\end{frame}

### 2013-10-29 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
added Isabelle Session

%
\begin{isabellebody}%
\def\isabellecontext{GoedelGod}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
A formalization and (partial) automation of Dana Scott's version \cite{ScottNotes}
 of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is presented. 
 QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
 Thus, the formalization is essentially a formalization in HOL. The employed embedding 
 of QML KB in HOL is adapting the work of Benzm\"uller and Paulson \cite{J23,B9}.
 Note that our QML KB formalization employs quantification over individuals and 
 quantification over sets of individuals (poperties).

 The formalization presented here has carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system we refer to the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at \url{http://isabelle.in.tum.de}.

 Some further notes:
 \begin{enumerate}
 \item This LaTeX text document has been produced automatically from the Isabelle source
 code document at \url{??? todo ???} with the Isabelle build tool.
 \item The formalization presented here follows the THF \cite{J22} formalization available at: 
    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/THF}
 \item The formalization is also related to the Coq \cite{Coq} formalization at:  
    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Coq}
 \item All reasoning gaps in Scott's proof script have been automated 
    with Sledgehammer \cite{Sledgehammer} performing remote calls to Satallax 
    \cite{Satallax} and LEO-II \cite{LEO-II}. These calls then suggested respective 
    Metis \cite{Metis} calls as given below. Metis proofs are verified in Isabelle/HOL.
 \item The reconstruction with Metis still fails for Theorem T3, but when some further
    lemmata are introduced we get success; these lemmata are called help1-4.
 \item For consistency checking the Nitpick model finder \cite{Nitpick} has been employed.
 \end{enumerate}%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsection{An Embedding of QML KB in HOL%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
Types for possible worlds and for individuals.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{typedecl}\isamarkupfalse%
\ i\ \ \ \ \ %
\isamarkupcmt{the type for possible worlds%
}
\ \isanewline
\ \ \isacommand{typedecl}\isamarkupfalse%
\ mu\ \ \ \ %
\isamarkupcmt{the type for indiviuals%
}
%
\begin{isamarkuptext}%
An accessibility relation on worlds.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{consts}\isamarkupfalse%
\ r\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i\ {\isasymRightarrow}\ bool{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}r{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{0}}{\isacharparenright}\ \ \ \ %
\isamarkupcmt{accessibility relation r%
}
%
\begin{isamarkuptext}%
The B axiom (symmetry) for relation r; B is only needed for proving theorem T3.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ sym{\isacharcolon}\ {\isachardoublequoteopen}x\ r\ y\ {\isasymlongrightarrow}\ y\ r\ x{\isachardoublequoteclose}%
\begin{isamarkuptext}%
QML formulas are identified with certain HOL terms of type \isa{i\ {\isasymRightarrow}\ bool}. 
 This type will be abbreviated in the remainder as \isa{{\isasymsigma}}%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ {\isasymsigma}\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}i\ {\isasymRightarrow}\ bool{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The classical connectives $\neg, \wedge, \Rightarrow$, and $\forall$
 (for individuals and over sets of individuals) and $\exists$ (over individuals) are
 lifted to type $\sigma$. Further connectives could be introduced analogously.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{definition}\isamarkupfalse%
\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ p\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymand}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{definition}\isamarkupfalse%
\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
For the grounding of lifted formulas the meta-predicate \isa{valid} is 
 introduced.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{definition}\isamarkupfalse%
\ valid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ bool{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}\ {\isasymequiv}\ {\isasymforall}w{\isachardot}\ p\ w{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The model finder Nitpick can now be used to confirm that the axioms and 
 definitions above are consistent. Unfortunately, the respective call-command for 
 Nitpick is not very intuitive.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{lemma}\isamarkupfalse%
\ True\ \isacommand{nitpick}\isamarkupfalse%
\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{The Formalization of G\"odel's Definitions and Axioms%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
We introduce constant symbol \isa{P} (G\"odel's "Positive").%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{consts}\isamarkupfalse%
\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The meaning of \isa{P} is constrained by axioms 
 \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or 
 its negation is positive, but not both) and 
 \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
 \imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\isanewline
\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}X{\isachardot}\ {\isasymPhi}\ X\ m{\isasymRightarrow}\ {\isasympsi}\ X{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ 
         (Positive properties are possibly exemplified).  
         T1 is proved directly by Sledghammer with command 
         \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}. This successful 
         attempt then suggest to instead try the Metis call below.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{theorem}\isamarkupfalse%
\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{unfolding}\isamarkupfalse%
\ mand{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mnot{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Next, we introduce the symbol \isa{G}, G\"odel's "God-like", which is defined 
 as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
 all positive properties:).%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{definition}\isamarkupfalse%
\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
We add axiom \isa{A{\isadigit{3}}}: $P(G)$ (The property of being God-like is positive) 
 and prove corollary \isa{C}: $\pos \ex x G(x)$ (Possibly, God exists). This
 this is directly proved by Sledgehammer, which then suggests the Metis call below.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ A{\isadigit{3}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\isanewline
\ \ \isacommand{corollary}\isamarkupfalse%
\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{unfolding}\isamarkupfalse%
\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
We add axiom \isa{A{\isadigit{4}}}: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 
 (Positive properties are necessarily positive).%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
We introduce symbol \isa{ess}, for "Essence", which is defined as 
 $\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
 \imp \psi(y)))$ (An \emph{essence} of an individual is a property possessed by it 
 and necessarily implying any of its properties).%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{definition}\isamarkupfalse%
\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
We now prove theorem \isa{T{\isadigit{2}}}: $\all x [G(x) \imp \ess{G}{x}]$ (Being 
 God-like is an essence of any God-like being). Again, Sledgehammer is used first, 
 which then suggests the Metis call below.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{theorem}\isamarkupfalse%
\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ A{\isadigit{1}}a\ A{\isadigit{1}}b\ A{\isadigit{4}}\ \isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mnot{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mand{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ G{\isacharunderscore}def\ ess{\isacharunderscore}def\ \isanewline
\ \ \ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Now, symbol \isa{NE}, for "Necessary Existence", is introduced. It is 
 defined as $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$ (Necessary 
 existence of an individual is the necessary exemplification of all its essences).%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{definition}\isamarkupfalse%
\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
We add axiom \isa{A{\isadigit{5}}}: $P(\NE)$ (Necessary existence is a positive property)%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ A{\isadigit{5}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ NE{\isacharbrackright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The main theorem T3: $\nec \ex x G(x)$ (Necessarily, God exists) can be proven
 directly with Sledgehammer: 
 \isa{theorem\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequote}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequote}\ sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}} 
 , but the suggested Metis call then unfortunately still fails. In order to 
 obtain a full verification of the proof we therefore introduce some further lemmata.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{lemma}\isamarkupfalse%
\ help{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymbox}\ p{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ sym\ \isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ help{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ p{\isacharbrackright}\ {\isasymand}\ {\isacharbrackleft}p\ m{\isasymRightarrow}\ {\isasymbox}\ p{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ help{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ {\isacharbrackleft}G\ x\ m{\isasymRightarrow}\ G\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ A{\isadigit{3}}\ A{\isadigit{5}}\ \isacommand{unfolding}\isamarkupfalse%
\ G{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ NE{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharcomma}\ mono{\isacharunderscore}tags{\isacharparenright}{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\ \isanewline
%
\endisadelimproof
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ help{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ help{\isadigit{3}}\ T{\isadigit{2}}\ \isacommand{unfolding}\isamarkupfalse%
\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Now, also Metis succeeds with proving theorem \isa{T{\isadigit{3}}}, and we also get
 \isa{T{\isadigit{4}}}:%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{theorem}\isamarkupfalse%
\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ help{\isadigit{1}}\ help{\isadigit{2}}\ C\ help{\isadigit{4}}\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Finally, we verify the consistency of our entire theory with Nitpick.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{lemma}\isamarkupfalse%
\ True\ \isacommand{nitpick}\isamarkupfalse%
\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided very important 
 help wrt consistency checking in Isabelle. Jasmin Blanchette also showed us how to produce latex documents from Isabelle files.%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\ \end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End:

### 2013-10-29 Formalizations/Isabelle/GoedelGodSession/output/document/root.tex
added Isabelle Session

--- 2013-08-21 Abstract/GodProof-Abstract.tex
@@
-\documentclass{llncs}
-
-\usepackage{url,amsmath,amssymb,a4wide}
+\documentclass[11pt,a4paper]{article}
+\usepackage{isabelle,isabellesym,amsmath,amssymb,a4wide}
 
@@
 
-\title{ Formalization, Mechanization and Automation % of Variants
-  of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
-    supported by the German Research Foundation under grant
-    BE2501/9-1.}  }
+% further packages required for unusual symbols (see also
+% isabellesym.sty), use only when needed
 
-\author{
-  Christoph Benzm\"{u}ller\inst{1} 
-  \and 
-  Bruno Woltzenlogel Paleo\inst{2}
-}
+%\usepackage{amssymb}
+  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
+  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
+  %\<triangleq>, \<yen>, \<lozenge>
 
-\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}
+%\usepackage{eurosym}
+  %for \<euro>
 
-\institute{
-  Dahlem Center for Intelligent Sstems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
-  \and 
-  Theory and Logic Group, Vienna University of Technology, Austria \\
-  \email{bruno@logic.at}
-}
+%\usepackage[only,bigsqcap]{stmaryrd}
+  %for \<Sqinter>
+
+%\usepackage{eufrak}
+  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)
+
+%\usepackage{textcomp}
+  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
+  %\<currency>
+
+% this should be the last package used
+\usepackage{pdfsetup}
+
+% urls in roman style, theory text in math-similar italics
+\urlstyle{rm}
+\isabellestyle{it}
+
+% for uniform font size
+%\renewcommand{\isastyle}{\isastyleminor}
 
@@
 
+\title{G\"odel's God in Isabelle/HOL}
+\author{Christoph Benzm\"uller and Bruno Woltzenlogel Paleo}
 \maketitle
 
-Attempts to prove the existence (or non-existence) of God by means of
-abstract ontological arguments are an old tradition in philosophy and
-theology.  G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
-this tradition, following particularly the footsteps of Leibniz.
-%
-G\"{o}del defines God as a being who possesses all \emph{positive} properties.
-He does not extensively discuss what positive properties are, 
-but instead he states a few reasonable (but debatable) axioms that they should satisfy.
-Various slightly different versions of axioms and definitions have
-been considered by G\"{o}del and by several philosophers who commented
-on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+%\tableofcontents
 
-Dana Scott's version of G\"odel's proof \cite{ScottNotes} employs the
-following axioms (\textbf{A}), definitions (\textbf{D}), corollaries
-(\textbf{C}) and theorems (\textbf{T}), and it proceeds in the
-following order:\footnote{ A1, A2, A5, D1, D3 are logically
-  equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and
-  3 in G\"odel's notes \cite{Goedel1970,GoedelNotes}. 
-  A3 was introduced by Scott \cite{ScottNotes} and 
-  could be derived from G\"odel's axiom 1 and
-  D1 in a logic with infinitary conjunction. 
-  A4 is a weaker formof G\"odel's axiom 3. 
-  D2 has an extra conjunct $\phi(x)$ lacking in G\"odel's definition 2; 
-  this is believed to have been 
-  an oversight by G\"odel \cite{Hazen}.}
+% sane default for proof documents
+\parindent 0pt\parskip 0.5ex
 
-\allowdisplaybreaks[1] 
-\begin{description}
-\item[A1] Either a property or its negation is positive, but not
-  both:  \hfill 
-  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$
-\item[A2] A property necessarily implied \\ by a
-  positive property is positive: \phantom{b} \hfill 
-  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]$ 
-\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$
-\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
-  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$
-\item[A3]  The property of being God-like is positive: \hfill   $P(G)$
-\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$
-\item[A4]  Positive properties are necessarily positive: \hfill 
-  $\all \phi [P(\phi) \to \Box \; P(\phi)]$
-\item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
-  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
-  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$
-\item[T2]  Being God-like is an essence of any
-  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ 
-\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
-  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$
-\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
-\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
-\end{description}
+% generated text of all theories
+\input{session}
 
-\noindent
-Scott's version of G\"{o}del's proof has now been 
-analysed for the first-time
-with an unprecedent degree of detail 
-and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
-The following has been done (and in this order):
-\begin{itemize}
-\item A detailed natural deduction proof.
-%
-\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
-%
-\item Automatic verification of the consistency of the axioms and 
-definitions with Nitpick \cite{Nitpick}.
-%
-\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.
+% optional bibliography
+%\bibliographystyle{abbrv}
+%\bibliography{root}
 
-\item A step-by-step formalization using the Coq proof assistant \cite{Coq}.
-
-\item A formalization using the Isabelle proof assistant \cite{Isabelle}, where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
-\end{itemize}
-
-G\"{o}del's proof is challenging to formalize and verify because it
-requires an expressive logical language with modal operators
-(\emph{possibly} and \emph{necessarily}) and with
-quantifiers for individuals and properties.  Our computer-assisted formalizations rely on an
-embedding of the modal logic into classical higher-order logic with
-Henkin semantics \cite{J23,B9}. The formalization is thus essentially
-done in classical higher-order logic where  quantified modal logic is
-emulated.
-
-In our ongoing computer-assisted study of G\"odel's proof we have
-obtained the following results:
-\begin{itemize}
-\item The basic modal logic K is sufficient for proving T1, C and T2. 
-%\marginpar{\scriptsize{Bruno: check this. Chris: Done.}}
-\item Modal logic S5 is not needed for proving T3; the logic KB is
-  sufficient. 
-% \marginpar{\scriptsize{Bruno: Maybe cite a reference for S5 and KB?}}
-\item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
-  and definitions would be inconsistent.
-\item For proving theorem T1, only the left to right direction of
-  axiom A1 is needed. However, the backward direction of A1 is
-  required for proving T2.
-
-\end{itemize}
-
-This work attests the maturity of contemporary interactive and
-automated deduction tools for classical higher-order logic and
-demonstrates the elegance and practical relevance of the
-embeddings-based approach.  Most importantly, our work opens new
-perspectives for a computer-assisted theoretical philosophy.  The
-critical discussion of the underlying concepts, definitions and axioms
-remains a human responsibility, but the computer can assist in
-building and checking rigorously correct logical arguments. In case of
-logico-philosophical disputes, the computer can check the disputing
-arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
-calculate!
-
-% \marginpar{\scriptsize{Bruno: I find this last paragraph a bit vague and not as self-contained as the rest of the abstract.
-We could try to improve it, but I think simply removing it would also be satisfactory.
-The sentence about Leibniz' dictum already provides a very nice ``climax'' for the end of the paper.
-Anything that we say after it feels like an ``anticlimax''...}}
-% Future work shall widen our computer-assisted approach to the
-% study of other ontological arguments. Variations of the distinctive features of
-% the base logic (e.g. non-rigid symbols, varying domains, etc.), made easier by the flexibility of the
-% embeddings-based approach, will
-% be considered.
-
-%\footnotesize
-\bibliographystyle{plain}
-%\bibliography{Bibliography}
+\small
 \begin{thebibliography}{10}
-
-\bibitem{Adams}
-R.M. Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1995.
-
-\bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings.
-\newblock G\"odel ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
 
@@
 \newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117.
   Springer, 2012.
-
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary bibligraphy on the ontological proof
-  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
-
-\bibitem{Fitting}
-M.~Fitting.
-\newblock {\em {Types, Tableaux and G\"odel's God}}.
-\newblock Kluver Academic Press, 2002.
-
-\bibitem{Goedel1970}
-K.~G\"odel.
-\newblock Ontological proof.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1970.
 
@@
 K.~G\"odel.
-\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pp. 144--145.
 \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{Hazen}
-A.P. Hazen.
-\newblock On g\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
-
-\bibitem{Hurd03first-orderproof}
+\bibitem{Metis}
 J.~Hurd.
@@
 
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
-
 \bibitem{ScottNotes}
 D.~Scott.
-\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
+\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pp. 145--146.
 \newblock In  \cite{sobel2004logic}, 2004.
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2013-10-29 Talks/FU-Berlin/GödelGod-Talk.tex
FU talk

--- 2013-08-15 Talks/FU-Berlin/GödelGod-Talk.tex
@@
 
-\title[Formalizations of G\"{o}del's Proof of God's Existence]{Formalizations of G\"{o}del's Proof of God's Existence}
+\title[On G\"{o}del's Proof of God's
+Existence]{Formalization, Mechanization and Automation of \\ G\"{o}del's
+  Proof of God's Existence}
 
-\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}
+\author{\textbf{Christoph Benzm\"{u}ller} and \textbf{Bruno Woltzenlogel Paleo}}
 
-\institute[] {
-  \inst{}%
-}
+% \institute[]{
+%   \inst{}%
+% }
 
-\date[29.10.13]
-{2013 October ??}
+\date[1.11.2013]{November 1, 2013}
 
@@
   \titlepage
+\includegraphics[height=2.5cm]{buch7.jpg} 
+\hfill
+\includegraphics[height=2.5cm]{nd.jpg}
 \end{frame}
 
+\input{talk-part-1.tex}
+
+\input{talk-part-2.tex}
+
+\input{talk-part-3.tex}
+
+\input{talk-part-4.tex}
+
+\input{talk-part-5.tex}
+
+\input{talk-part-6.tex}
+
 \end{document}

### 2013-10-29 Talks/FU-Berlin/talk-part-1.tex

@@
 
+\hskip-.5em
 \ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
 \ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
-\ldots\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
+\ldots\ldots   \rotatebox[origin = bl]{65}{\textcolor{blue}{Descartes}} \hskip-1.7em
+               \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
                \rotatebox[origin = bl]{65}{\textcolor{blue}{Leibniz}}  \hskip-1.2em
@@
 \begin{itemize}
-\item \textcolor{blue}{Philosophical:} Boundaries of Mataphysics \& Epistemology
+\item \textcolor{blue}{Philosophical:} Boundaries of Metaphysics \& Epistemology
   \begin{itemize}
@@
   \item but we want to draw
-      a conclusion for the real world. \\[2em]
+      a conclusion for the real world. \\[1em]
+  \item Necessary Existence (NE): metaphysical NE vs. logical NE  vs. modal NE \\[2em]
   \end{itemize} 
@@
 \begin{frame}{Introduction} \large
-Main Challenge: \hfill No theorem provers for \emph{Second-order Modal
-  Logic} \\[1em]
+Main challenge: \hfill No theorem provers for \emph{Higher-order Modal Logic} \\[1em]
 
-Our Idea: \hfill Exploit an embedding in \emph{Classical Higher-order
-  Logic} \\
+Our idea: \hfill Exploit an embedding in \emph{Higher-order Classical Logic} \\
 \,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
@@
 \item[F:] automation \& verification: proof assistant 
-  \textsc{Isabelle}
-\item[ ] Conclusion 
+  \textsc{Isabelle} \\[2em]
+%\item[ ] Conclusion \\[2em]
 \end{itemize}
-Did we get anything new? \ldots Probably yes --- let's discuss afterwards!
+Did we get new results? \hfill  \textcolor{red}{Yes --- let's discuss later!}
 \end{frame}

### 2013-10-29 Talks/FU-Berlin/talk-part-4.tex

\begin{frame}{Coq Proof} \small
todo
\end{frame}

### 2013-10-29 Talks/FU-Berlin/talk-part-5.tex

\begin{frame}{Isabelle Proof} \small
todo
\end{frame}

### 2013-10-29 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
new Isabelle/HOL version of the proof

@@
  of QML KB in HOL is adapting the work of Benzm\"uller and Paulson \cite{J23,B9}.
- Note that our QML KB formalization employs quantification over individuals and 
+ Note that the QML KB formalization employs quantification over individuals and 
  quantification over sets of individuals (poperties).
 
- The formalization presented here has carried and formally verified in the Isabelle/HOL 
- proof assistant; for more information on this system we refer to the textbook by Nipkow, 
+ The formalization presented here has been carried and formally verified in the Isabelle/HOL 
+ proof assistant; for more information on this system see the textbook by Nipkow, 
  Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
- at \url{http://isabelle.in.tum.de}.
+ at the Isabelle homepage: \url{http://isabelle.in.tum.de}.
 
- Some further notes:
+ Some further notes: \sloppy
  \begin{enumerate}
  \item This LaTeX text document has been produced automatically from the Isabelle source
- code document at \url{??? todo ???} with the Isabelle build tool.
- \item The formalization presented here follows the THF \cite{J22} formalization available at: 
-    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/THF}
- \item The formalization is also related to the Coq \cite{Coq} formalization at:  
-    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Coq}
+ code document at 
+ \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle/GoedelGodSession} 
+ with the Isabelle build tool.
+ \item The formalization presented here is related to the THF \cite{J22} and 
+    Coq \cite{Coq} formalizations available at
+    \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/}.
  \item All reasoning gaps in Scott's proof script have been automated 
-    with Sledgehammer \cite{Sledgehammer} performing remote calls to Satallax 
-    \cite{Satallax} and LEO-II \cite{LEO-II}. These calls then suggested respective 
-    Metis \cite{Metis} calls as given below. Metis proofs are verified in Isabelle/HOL.
- \item The reconstruction with Metis still fails for Theorem T3, but when some further
-    lemmata are introduced we get success; these lemmata are called help1-4.
- \item For consistency checking the Nitpick model finder \cite{Nitpick} has been employed.
+    with Sledgehammer \cite{Sledgehammer} performing remote calls to the higher-order automated
+    theorem prover LEO-II \cite{LEO-II}. These calls then suggested respective 
+    Metis \cite{Metis} calls as given below. The Metis proofs are then verified in Isabelle/HOL.
+ \item For consistency checking, the Nitpick model finder \cite{Nitpick} has been employed.
  \end{enumerate}%
@@
 \begin{isamarkuptext}%
-Types for possible worlds and for individuals.%
+The types \isa{i} for possible worlds (or states) and \isa{mu} for individuals 
+are introduced.%
 \end{isamarkuptext}%
@@
 \begin{isamarkuptext}%
-An accessibility relation on worlds.%
+Possible worlds are connected by an accessibility relation .%
 \end{isamarkuptext}%
@@
 \begin{isamarkuptext}%
-The B axiom (symmetry) for relation r; B is only needed for proving theorem T3.%
+The \isa{B} axiom (symmetry) for relation r is stated. \isa{B} is needed only 
+for proving theorem T3.%
 \end{isamarkuptext}%
@@
 (for individuals and over sets of individuals) and $\exists$ (over individuals) are
- lifted to type $\sigma$. Further connectives could be introduced analogously.%
+lifted to type $\sigma$. Further connectives could be introduced analogously. \isa{definition} 
+could be used instead of \isa{abbreviation}; the latter are always fully expanded/rewritten,
+which is fine here, where the focus has been on proof automation, but which would lead to 
+unnessarily complex proof tasks in a purely interactive session.%
 \end{isamarkuptext}%
 \isamarkuptrue%
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ p\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymand}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
-\ \ \isacommand{definition}\isamarkupfalse%
-\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymPhi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPhi}\ m{\isasymand}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymPhi}\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
-For the grounding of lifted formulas the meta-predicate \isa{valid} is 
- introduced.%
+For the grounding of lifted formulas the meta-predicate \isa{valid} is introduced.%
 \end{isamarkuptext}%
 \isamarkuptrue%
-\ \ \isacommand{definition}\isamarkupfalse%
+\ \ \isacommand{abbreviation}\isamarkupfalse%
 \ valid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ bool{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}\ {\isasymequiv}\ {\isasymforall}w{\isachardot}\ p\ w{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
-The model finder Nitpick can now be used to confirm that the axioms and 
- definitions above are consistent. Unfortunately, the respective call-command for 
- Nitpick is not very intuitive.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ True\ \isacommand{nitpick}\isamarkupfalse%
-\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
-\isadelimproof
-\ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{oops}\isamarkupfalse%
-%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\isamarkupsection{The Formalization of G\"odel's Definitions and Axioms%
-}
-\isamarkuptrue%
-%
-\begin{isamarkuptext}%
-We introduce constant symbol \isa{P} (G\"odel's "Positive").%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{consts}\isamarkupfalse%
-\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-The meaning of \isa{P} is constrained by axioms 
- \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or 
- its negation is positive, but not both) and 
- \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
- \imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\isanewline
-\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}X{\isachardot}\ {\isasymPhi}\ X\ m{\isasymRightarrow}\ {\isasympsi}\ X{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ 
-         (Positive properties are possibly exemplified).  
-         T1 is proved directly by Sledghammer with command 
-         \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}. This successful 
-         attempt then suggest to instead try the Metis call below.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{unfolding}\isamarkupfalse%
-\ mand{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ \isanewline
-\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mnot{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
-\ \ \ \ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-Next, we introduce the symbol \isa{G}, G\"odel's "God-like", which is defined 
- as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
- all positive properties:).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{definition}\isamarkupfalse%
-\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We add axiom \isa{A{\isadigit{3}}}: $P(G)$ (The property of being God-like is positive) 
- and prove corollary \isa{C}: $\pos \ex x G(x)$ (Possibly, God exists). This
- this is directly proved by Sledgehammer, which then suggests the Metis call below.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\ A{\isadigit{3}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\isanewline
-\ \ \isacommand{corollary}\isamarkupfalse%
-\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{unfolding}\isamarkupfalse%
-\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-We add axiom \isa{A{\isadigit{4}}}: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 
- (Positive properties are necessarily positive).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We introduce symbol \isa{ess}, for "Essence", which is defined as 
- $\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
- \imp \psi(y)))$ (An \emph{essence} of an individual is a property possessed by it 
- and necessarily implying any of its properties).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{definition}\isamarkupfalse%
-\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
-\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We now prove theorem \isa{T{\isadigit{2}}}: $\all x [G(x) \imp \ess{G}{x}]$ (Being 
- God-like is an essence of any God-like being). Again, Sledgehammer is used first, 
- which then suggests the Metis call below.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{1}}a\ A{\isadigit{1}}b\ A{\isadigit{4}}\ \isacommand{unfolding}\isamarkupfalse%
-\ valid{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mnot{\isacharunderscore}def\isanewline
-\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ mand{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ G{\isacharunderscore}def\ ess{\isacharunderscore}def\ \isanewline
-\ \ \ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-Now, symbol \isa{NE}, for "Necessary Existence", is introduced. It is 
- defined as $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$ (Necessary 
- existence of an individual is the necessary exemplification of all its essences).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{definition}\isamarkupfalse%
-\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We add axiom \isa{A{\isadigit{5}}}: $P(\NE)$ (Necessary existence is a positive property)%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\ A{\isadigit{5}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ NE{\isacharbrackright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-The main theorem T3: $\nec \ex x G(x)$ (Necessarily, God exists) can be proven
- directly with Sledgehammer: 
- \isa{theorem\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequote}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequote}\ sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}
- , but the suggested Metis call then unfortunately still fails. In order to 
- obtain a full verification of the proof we therefore introduce some further lemmata.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ help{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymbox}\ p{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ sym\ \isacommand{unfolding}\isamarkupfalse%
-\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-\isanewline
-%
-\endisadelimproof
-\isanewline
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ help{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ p{\isacharbrackright}\ {\isasymand}\ {\isacharbrackleft}p\ m{\isasymRightarrow}\ {\isasymbox}\ p{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{unfolding}\isamarkupfalse%
-\ valid{\isacharunderscore}def\ mdia{\isacharunderscore}def\ mbox{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-\isanewline
-%
-\endisadelimproof
-\ \ \isanewline
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ help{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ {\isacharbrackleft}G\ x\ m{\isasymRightarrow}\ G\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{3}}\ A{\isadigit{5}}\ \isacommand{unfolding}\isamarkupfalse%
-\ G{\isacharunderscore}def\ mforall{\isacharunderscore}indset{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ NE{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
-\ \ \ \ \isacommand{by}\isamarkupfalse%
-\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharcomma}\ mono{\isacharunderscore}tags{\isacharparenright}{\isacharparenright}%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-\ \isanewline
-%
-\endisadelimproof
-\isanewline
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ help{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
-%
-\isadelimproof
-\ \ \ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ help{\isadigit{3}}\ T{\isadigit{2}}\ \isacommand{unfolding}\isamarkupfalse%
-\ mexists{\isacharunderscore}ind{\isacharunderscore}def\ mforall{\isacharunderscore}ind{\isacharunderscore}def\ mimplies{\isacharunderscore}def\ valid{\isacharunderscore}def\isanewline
-\ \ \ \ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-Now, also Metis succeeds with proving theorem \isa{T{\isadigit{3}}}, and we also get
- \isa{T{\isadigit{4}}}:%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
-\isadelimproof
-\ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ help{\isadigit{1}}\ help{\isadigit{2}}\ C\ help{\isadigit{4}}\ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-Finally, we verify the consistency of our entire theory with Nitpick.%
+The model finder Nitpick confirms that the axioms and definitions above are consistent. 
+Unfortunately, the respective command syntax for Nitpick is not very intuitive.%
 \end{isamarkuptext}%
@@
 \begin{isamarkuptext}%
-\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided very important 
- help wrt consistency checking in Isabelle. Jasmin Blanchette also showed us how to produce latex documents from Isabelle files.%
+Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{consts}\isamarkupfalse%
+\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi 
+[P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or its negation is positive, but not both.) 
+and \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
+\imp P(\psi)]$ (A property necessarily implied by a positive property is positive.).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{axiomatization}\isamarkupfalse%
+\ \isakeyword{where}\isanewline
+\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}X{\isachardot}\ {\isasymPhi}\ X\ m{\isasymRightarrow}\ {\isasympsi}\ X{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ (Positive 
+properties are possibly exemplified).
+T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}.
+This successful attempt then suggest to
+instead try the Metis call in the line below. This Metis call generates a proof object that is 
+verified in Isabelle/HOL's kernel.%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{theorem}\isamarkupfalse%
+\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{using}\isamarkupfalse%
+\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{by}\isamarkupfalse%
+\ metis%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+Next, the symbol \isa{G}, for "God-like", is introduced and defined 
+as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
+all positive properties:).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{definition}\isamarkupfalse%
+\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive.).
+Sledgehammer and Metis then prove corollary \isa{C}: $\pos \ex x G(x)$ 
+(Possibly, God exists.).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{axiomatization}\isamarkupfalse%
+\ \isakeyword{where}\ A{\isadigit{3}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\isanewline
+\ \ \isacommand{corollary}\isamarkupfalse%
+\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{using}\isamarkupfalse%
+\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
+\ metis%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+We add axiom \isa{A{\isadigit{4}}}: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 
+(Positive properties are necessarily positive).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{axiomatization}\isamarkupfalse%
+\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+Symbol \isa{ess}, for "Essence", is introduced and defined as 
+$\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
+\imp \psi(y)))$ (An \emph{essence} of an individual is a property possessed by it 
+and necessarily implying any of its properties.).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{definition}\isamarkupfalse%
+\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
+\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+Next, Sledgehammer and Metis prove theorem \isa{T{\isadigit{2}}}: $\all x [G(x) \imp \ess{G}{x}]$ 
+(Being God-like is an essence of any God-like being).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{theorem}\isamarkupfalse%
+\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{by}\isamarkupfalse%
+\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharparenright}\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+Symbol \isa{NE}, for "Necessary Existence", is introduced and
+defined as $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$ (Necessary 
+existence of an individual is the necessary exemplification of all its essences.).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{definition}\isamarkupfalse%
+\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+Moreover, axiom \isa{A{\isadigit{5}}} is added: $P(\NE)$ (Necessary existence is a positive 
+property.).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{axiomatization}\isamarkupfalse%
+\ \isakeyword{where}\ A{\isadigit{5}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}P\ NE{\isacharbrackright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+Finally, Sledgehammer and Metis prove the main theorem \isa{T{\isadigit{3}}}: $\nec \ex x G(x)$ 
+(Necessarily, God exists).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{theorem}\isamarkupfalse%
+\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{using}\isamarkupfalse%
+\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
+\ metis%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+\isanewline
+%
+\endisadelimproof
+\isanewline
+\ \ \isacommand{corollary}\isamarkupfalse%
+\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{using}\isamarkupfalse%
+\ T{\isadigit{1}}\ T{\isadigit{3}}\ sym\ G{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
+\ metis%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+Finally, the consistency of our entire theory with Nitpick.%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{lemma}\isamarkupfalse%
+\ True\ \isacommand{nitpick}\isamarkupfalse%
+\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
+\isadelimproof
+\ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{oops}\isamarkupfalse%
+%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided 
+very important help wrt consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
+produce latex documents from Isabelle sources, and he pointed to many useful tricks in Isabelle.%
 \end{isamarkuptext}%

### 2013-10-30 Talks/FU-Berlin/talk-part-1.tex
new Isabelle/HOL version of the proof
update of FU talk
Add "Die Presse"

@@
 Austria \\
-- myscience.at \\
+- Die Presse \\
 - Wiener Zeitung \\
@@
 - Delhi Daily News \\
-- Indoa Today \\
+- India Today \\
 - \ldots \\

### 2013-10-30 Talks/FU-Berlin/talk-part-1.tex
update of FU talk

@@
 \begin{frame}{Introduction} \large
-Main challenge: \hfill No theorem provers for \emph{Higher-order Modal Logic} \\[1em]
+Main challenge: \hfill No provers for \emph{Higher-order Modal Logic} \textcolor{red}{HML} \\[1em]
 
-Our idea: \hfill Exploit an embedding in \emph{Higher-order Classical Logic} \\
+Our idea: \hfill Embedding in \emph{Higher-order Classical Logic} \textcolor{blue}{HOL} \\
 \,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
@@
 \item[A:] pen and paper: detailed natural deduction proof 
-\item[B:] formalization: axioms, defs, thms  in \textsc{TPTP THF}
-\item[C:] consistency: automatic verification with \textsc{Nitpick} 
-\item[D:] proof automation: theorems provers Leo-II and \textsc{Satallax} 
-\item[E:] step-by-step verification: proof assistant \textsc{Coq} 
-\item[F:] automation \& verification: proof assistant 
+\item[B:] formalization: axioms, defs, thms  in \textcolor{blue}{HOL}
+\item[] proof automation: theorems provers \textsc{Leo-II} and \textsc{Satallax} 
+\item[] consistency: automatic verification with \textsc{Nitpick} 
+\item[C:] step-by-step verification: proof assistant \textsc{Coq} 
+\item[D:] automation \& verification: proof assistant 
   \textsc{Isabelle} \\[2em]

### 2013-10-30 Talks/FU-Berlin/talk-part-1.tex
update of talk

@@
-\begin{frame}{Introduction} \small
+\begin{frame}{} \small
+\vskip1em
 \begin{minipage}{.56\textwidth} 
-\includegraphics[width=\textwidth]{spiegel1} 
+\colorbox{gray}{\includegraphics[width=\textwidth]{spiegel1}} 
 \vskip1em
@@
 - Repubblica \\
-- Today.it \\
 - Ilsussidario \\
@@
 
+US \\
+- ABC News \\
+- \ldots \\
+
 International \\
@@
 - Yahoo Finance \\
-- CNET \\
+% - CNET \\
 - United Press Intl. \\
-- \ldots
+- \ldots \\
 \end{minipage}
+\end{frame}
+
+\begin{frame}{Introduction --- Quick answers to your most pressing questions!} \large
+\colorbox{gray}{\includegraphics[width=\textwidth]{MacBookGrab}} 
+\pause
+\vfill
+Are we in contact with Steve Jobs? \hfill No \\[2em]
+Do you really need a MacBook to obtain the results? \hfill No \\[2em]
+Is Apple sending us money? \hfill No \\
+\, \hfill (but maybe they should)
 \end{frame}
@@
 \begin{frame}{Introduction} \large
-Main challenge: \hfill No provers for \emph{Higher-order Modal Logic} \textcolor{red}{HML} \\[1em]
+Main challenge: \hfill No provers for \emph{Higher-order Modal Logic\/} (\textcolor{red}{HML}) \\[1em]
 
-Our idea: \hfill Embedding in \emph{Higher-order Classical Logic} \textcolor{blue}{HOL} \\
+Our solution: \hfill Embedding in \emph{Higher-order Classical Logic\/} (\textcolor{blue}{HOL}) \\
 \,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
@@
 \item[A:] pen and paper: detailed natural deduction proof 
-\item[B:] formalization: axioms, defs, thms  in \textcolor{blue}{HOL}
-\item[] proof automation: theorems provers \textsc{Leo-II} and \textsc{Satallax} 
-\item[] consistency: automatic verification with \textsc{Nitpick} 
+\item[B:] formalization: in classical higher-order logic (\textcolor{blue}{HOL})
+\item[] proof automation: theorem provers \textsc{Leo-II} and \textsc{Satallax} 
+\item[] consistency: model finder \textsc{Nitpick (Nitrox)} 
 \item[C:] step-by-step verification: proof assistant \textsc{Coq} 

### 2013-10-30 Formalizations/THF/call_tptp.sh

@@ -8,7 +8,6 @@ timeout=90
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
-echo
 
 for file in "$@"
 do

### 2013-10-30 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
new Isabelle/HOL version of the proof
updates for FU talk

@@
  Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
- at the Isabelle homepage: \url{http://isabelle.in.tum.de}.
+ at: \url{http://isabelle.in.tum.de}.
 
@@
  \item The formalization presented here is related to the THF \cite{J22} and 
-    Coq \cite{Coq} formalizations available at
+    Coq \cite{Coq} formalizations at
     \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/}.
  \item All reasoning gaps in Scott's proof script have been automated 
-    with Sledgehammer \cite{Sledgehammer} performing remote calls to the higher-order automated
-    theorem prover LEO-II \cite{LEO-II}. These calls then suggested respective 
-    Metis \cite{Metis} calls as given below. The Metis proofs are then verified in Isabelle/HOL.
- \item For consistency checking, the Nitpick model finder \cite{Nitpick} has been employed.
+    with Sledgehammer \cite{Sledgehammer}, performing remote calls to the higher-order automated
+    theorem prover LEO-II \cite{LEO-II}. These calls suggest the 
+    Metis \cite{Metis} calls as given below. The Metis proofs are verified in Isabelle/HOL.
+ \item For consistency checking, the model finder \cite{Nitpick} has been employed.
  \end{enumerate}%
@@
 \begin{isamarkuptext}%
-The types \isa{i} for possible worlds (or states) and \isa{mu} for individuals 
+The types \isa{i} for possible worlds (or states) and $\mu$ for individuals 
 are introduced.%
@@
 \ \ \isacommand{typedecl}\isamarkupfalse%
-\ i\ \ \ \ \ %
+\ i\ \ \ \ %
 \isamarkupcmt{the type for possible worlds%
@@
 \ \ \isacommand{typedecl}\isamarkupfalse%
-\ mu\ \ \ \ %
+\ {\isasymmu}\ \ \ \ %
 \isamarkupcmt{the type for indiviuals%
@@
 \begin{isamarkuptext}%
-Possible worlds are connected by an accessibility relation .%
+Possible worlds are connected by an accessibility relation \isa{r}.%
 \end{isamarkuptext}%
@@
 QML formulas are identified with certain HOL terms of type \isa{i\ {\isasymRightarrow}\ bool}. 
-This type will be abbreviated in the remainder as \isa{{\isasymsigma}}%
+This type will be abbreviated in the remainder as \isa{{\isasymsigma}}.%
 \end{isamarkuptext}%
@@
 \begin{isamarkuptext}%
-The classical connectives $\neg, \wedge, \Rightarrow$, and $\forall$
-(for individuals and over sets of individuals) and $\exists$ (over individuals) are
-lifted to type $\sigma$. Further connectives could be introduced analogously. \isa{definition} 
-could be used instead of \isa{abbreviation}; the latter are always fully expanded/rewritten,
-which is fine here, where the focus has been on proof automation, but which would lead to 
-unnessarily complex proof tasks in a purely interactive session.%
+The classical connectives $\neg, \wedge, \rightarrow$, and $\forall$
+(over individuals and over sets of individuals) and $\exists$ (over individuals) are
+lifted to type $\sigma$. Further connectives could be introduced analogously. Definitions 
+could be used instead of abbreviations.%
 \end{isamarkuptext}%
@@
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymPhi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
+\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymphi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPhi}\ m{\isasymand}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymPhi}\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
+\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ m{\isasymand}\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymphi}\ w\ {\isasymand}\ {\isasympsi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
+\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ m{\isasymRightarrow}\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymphi}\ w\ {\isasymlongrightarrow}\ {\isasympsi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
+\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%
+\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
-For the grounding of lifted formulas the meta-predicate \isa{valid} is introduced.%
+For grounding lifted formulas, the meta-predicate \isa{valid} is introduced.%
 \end{isamarkuptext}%
@@
 \ valid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ bool{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}\ {\isasymequiv}\ {\isasymforall}w{\isachardot}\ p\ w{\isachardoublequoteclose}%
+\isamarkupsection{G\"odel's Ontological Argument%
+}
+\isamarkuptrue%
+%
 \begin{isamarkuptext}%
-The model finder Nitpick confirms that the axioms and definitions above are consistent. 
-Unfortunately, the respective command syntax for Nitpick is not very intuitive.%
+Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%
 \end{isamarkuptext}%
 \isamarkuptrue%
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ True\ \isacommand{nitpick}\isamarkupfalse%
-\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
+\ \ \isacommand{consts}\isamarkupfalse%
+\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi 
+[P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or its negation is positive, but not both.) 
+and \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
+\imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{axiomatization}\isamarkupfalse%
+\ \isakeyword{where}\isanewline
+\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPhi}\ x\ m{\isasymRightarrow}\ {\isasympsi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\begin{isamarkuptext}%
+We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ (Positive 
+properties are possibly exemplified).
+T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}}.
+This successful attempt then suggests to
+instead try the Metis call in the line below. This Metis call generates a proof object 
+that is verified in Isabelle/HOL's kernel.%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{theorem}\isamarkupfalse%
+\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
+\ \ \isanewline
+%
 \isadelimproof
-\ %
+\ \ %
 \endisadelimproof
@@
 \isatagproof
-\isacommand{oops}\isamarkupfalse%
-%
+\isacommand{by}\isamarkupfalse%
+\ {\isacharparenleft}metis\ A{\isadigit{1}}a\ A{\isadigit{2}}{\isacharparenright}%
 \endisatagproof
@@
 \begin{isamarkuptext}%
-Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{consts}\isamarkupfalse%
-\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi 
-[P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or its negation is positive, but not both.) 
-and \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
-\imp P(\psi)]$ (A property necessarily implied by a positive property is positive.).%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\isanewline
-\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}X{\isachardot}\ {\isasymPhi}\ X\ m{\isasymRightarrow}\ {\isasympsi}\ X{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
-\begin{isamarkuptext}%
-We prove theorem T1: $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ (Positive 
-properties are possibly exemplified).
-T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}.
-This successful attempt then suggest to
-instead try the Metis call in the line below. This Metis call generates a proof object that is 
-verified in Isabelle/HOL's kernel.%
-\end{isamarkuptext}%
-\isamarkuptrue%
-\ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
-\ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
-%
-\isadelimproof
-\ \ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{by}\isamarkupfalse%
-\ metis%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
-%
-\begin{isamarkuptext}%
-Next, the symbol \isa{G}, for "God-like", is introduced and defined 
+Next, the symbol \isa{G} for "God-like" is introduced and defined 
 as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
@@
 \ \ \isacommand{definition}\isamarkupfalse%
-\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
-Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive.).
+Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive).
 Sledgehammer and Metis then prove corollary \isa{C}: $\pos \ex x G(x)$ 
-(Possibly, God exists.).%
+(Possibly, God exists).%
 \end{isamarkuptext}%
@@
 \ \ \isacommand{corollary}\isamarkupfalse%
-\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isanewline
 %
@@
 \begin{isamarkuptext}%
-We add axiom \isa{A{\isadigit{4}}}: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 
+Axiom \isa{A{\isadigit{4}}} is added: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 
 (Positive properties are necessarily positive).%
@@
 \ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
-Symbol \isa{ess}, for "Essence", is introduced and defined as 
+Symbol \isa{ess} for "Essence" is introduced and defined as 
 $\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
@@
 \ \ \isacommand{definition}\isamarkupfalse%
-\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
-\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
+\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
@@
 \ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
-\ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
+\ \ \isanewline
 %
@@
 \ \ \isacommand{definition}\isamarkupfalse%
-\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
 Moreover, axiom \isa{A{\isadigit{5}}} is added: $P(\NE)$ (Necessary existence is a positive 
-property.).%
+property).%
 \end{isamarkuptext}%
@@
 \ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isanewline
 %
@@
 \ \ \isacommand{corollary}\isamarkupfalse%
-\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
+\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isanewline
 %
@@
 \begin{isamarkuptext}%
-Finally, the consistency of our entire theory with Nitpick.%
+The consistency of the entire theory is checked with Nitpick.%
 \end{isamarkuptext}%
 \isamarkuptrue%
-\ \ \isacommand{lemma}\isamarkupfalse%
-\ True\ \isacommand{nitpick}\isamarkupfalse%
-\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
-\isadelimproof
-\ %
-\endisadelimproof
-%
-\isatagproof
-\isacommand{oops}\isamarkupfalse%
-%
-\endisatagproof
-{\isafoldproof}%
-%
-\isadelimproof
-%
-\endisadelimproof
 %
@@
 very important help wrt consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
-produce latex documents from Isabelle sources, and he pointed to many useful tricks in Isabelle.%
+produce latex documents from Isabelle sources, and he showed us useful tricks in Isabelle.%
 \end{isamarkuptext}%

### 2013-10-30 Talks/FU-Berlin/talk-part-6.tex
Add empty slides for the criticism section
Make last slide plain
Add a new result
Add an item to conclusion

@@
-\begin{frame}{Conclusion} \small
-todo
+\begin{transitionframe}
+\textbf{Part E:}
+
+Criticisms
+\end{transitionframe}
+
+\begin{frame}{Criticisms}{S5}
+
 \end{frame}
+
+\begin{frame}{Criticisms}{Modal Collapse}
+
+\end{frame}
+
+\begin{frame}{Criticisms}{No Neutral Properties}
+
+\end{frame}
+
+\begin{transitionframe}
+\textbf{Part F:}
+
+Conclusions
+\end{transitionframe}
+
+\begin{frame}{Summary of Results about the Ontological Proof}
+\begin{itemize}[<+->]
+\item K sufficient for T1, C1 and T2 
+\item S5 not needed for T3
+\item KB sufficient for T3 
+\item A simpler new proof of C1
+\item G\"odel's original axioms (without conjunct $\phi(x)$ in D2) are inconsistent
+\item Scott's axioms are consistent
+\item For T1, only half of A1 (A1a) is needed 
+\item For T2, the other half (A1b) is needed
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Summary of Results for Logic} \small
+\begin{itemize}[<+->]
+\item Infra-structure for reasoning with modal logic using existing proof assistants and higher-order automated theorem provers
+\item A new natural deduction calculus for higher-order modal logic
+\end{itemize}
+\end{frame}
+
+\begin{frame}[plain]
+\includegraphics[width=\textwidth]{TUWien-GodComputerC} 
+\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-5.tex
Fix transition frames
update of FU talk

@@
-\begin{frame}{Isabelle Proof} \small
-todo
+\begin{transitionframe}{} \Large \centering
+\textbf{Part D:} \\
+\quad automation \& verification: proof assistant \textsc{Isabelle} \\[2em]
+\end{transitionframe}
+
+\begin{frame}{} \Large \centering
+\vskip.5em
+\colorbox{gray}{\includegraphics[width=\textwidth]{IsabelleGrab}} 
 \end{frame}
+
+\begin{frame}{Automation \& Verification in Proof Assistant \textsc{Isabelle/HOL}} \Large
+Isabelle/HOL   (Cambridge University/TU Munich)
+\begin{itemize}
+\item HOL instance of the generic \textsc{Isabelle} proof assistant
+\item User interaction and proof automation 
+\item Automation is supported by \textsc{Sledgehammer} tool
+\item Verification of the proofs in \textsc{Isabelle/HOL}'s small proof kernel
+\end{itemize}
+\vfill
+What have we done?
+\begin{itemize}
+\item Proof automation of G\"odel's proof script (Scott version)
+\item \textsc{Sledghammer} makes calls to remote THF provers in Miami
+\item These calls the suggest respective calls to the \textsc{Metis} prover
+\item \textsc{Metis} proofs are verified in \textsc{Isabelle/HOL}'s proof kernel
+\end{itemize}
+\vfill
+See the handout (generated from the Isabelle source file).
+\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-6.tex
Add another result
Add slides for criticisms

@@
 
-\begin{frame}{Criticisms}{S5}
+\begin{frame}{Criticisms}{S5} \centering
+$$
+\all P.[ \pos \nec P \imp \nec P ] 
+$$
+
+\medskip
+
+If something is possibly necessary, then it is necessary.
+
+\pause
+
+\bigskip
+
+$
+\pos\only<8->{_c} \nec\only<8->{_c} (A \vee \neg A)
+$
+\pause
+\qquad 
+$
+\nec\only<8->{_c} (A \vee \neg A)
+$
+
+\pause
+
+\bigskip
+
+logical necessity $\sim$ validity
+%\qquad\qquad
+\hfill
+logical possibility $\sim$ satisfiability
+
+\pause
+
+\medskip
+
+$ 
+\textrm{for all } M, M \models F 
+\quad \longrightarrow \quad
+\nec F
+%\qquad\qquad\quad
+\hfill
+\textrm{exists } M, M \models F 
+\quad \longrightarrow \quad
+\pos F
+$
+
+\pause
+
+\bigskip
+
+What about iterations?
+$$
+\pos \nec \pos \pos F
+$$
+
+\medskip
+
+\pause
+
+Weak intuitions $\Rightarrow$ dozens of modal logics
+
+\medskip
+
+\pause
+
+\alert{S5 is considered adequate}
 
@@
 
-\begin{frame}{Criticisms}{Modal Collapse}
+\begin{frame}{Criticisms}{Modal Collapse} \centering
+$$
+\all P.[ P \imp \nec P ] 
+$$
+
+\medskip
+
+Everything that is the case is so necessarily.
+
+\pause
+
+\medskip
+
+Follows from T2, T3 and D2.
+
+\pause
+
+\medskip
+
+There are no contingent ``truths''. \\ \pause
+Everything is determined. \\ \pause
+There is no free will. \\ \pause
+God's existence makes no difference.
+
+\pause
+\bigskip
+
+Many proposed solutions: Anderson, Fitting, H\'ajek, \ldots
 
@@
 
-\begin{frame}{Criticisms}{No Neutral Properties}
+\begin{frame}{Criticisms}{No Neutral Properties} \centering
 
+$$\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$$
+
+Either a property is positive or its negation is (never both)
+
+\pause
+\bigskip
+
+Are the following properties positive or negative?
+
+$$
+\lambda x. G(x) \qquad \pause \lambda x. human(x) \qquad \pause \lambda x. foreigner(x) \qquad \lambda x. \neg foreigner(x), \ldots
+$$
+
+\pause
+\bigskip
+
+Solution: \\
+``\ldots positive in the moral aesthetic sense (independently of the accidental structure of the world). Only then the ax. true. [sic] \ldots''
+\\ \hfill - G\"odel, 1970
 \end{frame}
@@
 \item A new natural deduction calculus for higher-order modal logic
+\item Difficult benchmarks for higher-order automated theorem provers
 \end{itemize}
@@
 \begin{frame}[plain]
-\includegraphics[width=\textwidth]{TUWien-GodComputerC} 
+\colorbox{black}{\includegraphics[width=\textwidth]{TUWien-GodComputerC}}
 \end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-1.tex
several updates for FU talk

@@
 
-* for the existence of god \\
+* for the existence of God \\
 
@@
 \begin{frame}{Introduction}
-
+\vskip1em
 \emph{\huge Wohl eine jede Philosophie kreist um den ontologischen
-  Gottesbeweis} \\[2em]
+  Gottesbeweis} \\[1.5em]
 (Adorno, Th. W.: Negative Dialektik. Frankfurt a. M. 1966, p.378)
 \vfill
+\begin{center}
+\fcolorbox{gray}{white}{
+\begin{minipage}{.9\textwidth}
 \ldots \hfill
@@
 \ldots
+\end{minipage}
+}
+\end{center}
 \end{frame}
@@
 \begin{itemize}
-\item[A:] pen and paper: detailed natural deduction proof 
-\item[B:] formalization: in classical higher-order logic (\textcolor{blue}{HOL})
-\item[] proof automation: theorem provers \textsc{Leo-II} and \textsc{Satallax} 
-\item[] consistency: model finder \textsc{Nitpick (Nitrox)} 
-\item[C:] step-by-step verification: proof assistant \textsc{Coq} 
-\item[D:] automation \& verification: proof assistant 
-  \textsc{Isabelle} \\[2em]
+\item[A:] Pen and paper: \hfill detailed natural deduction proof 
+\item[B:] Formalization: \hfill in classical higher-order logic (\textcolor{blue}{HOL})
+\item[] Automation: \hfill theorem provers \textsc{Leo-II} and \textsc{Satallax} 
+\item[] Consistency: \hfill model finder \textsc{Nitpick (Nitrox)} 
+\item[C:] Step-by-step verification: \hfill proof assistant \textsc{Coq} 
+\item[D:] Automation \& verification: \hfill proof assistant \textsc{Isabelle} \\[2em]
 %\item[ ] Conclusion \\[2em]
 \end{itemize}
-Did we get new results? \hfill  \textcolor{red}{Yes --- let's discuss later!}
+Did we get new results? \hfill  \textcolor{red}{Yes --- let's discuss this later!}
 \end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-1.tex
Add Gaunilo
update for FU talk

@@
 \ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
+          \rotatebox[origin = bl]{65}{\textcolor{red}{Gaunilo}} \hskip-1.3em
 \ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
@@
 \begin{frame}{Introduction} \large
-Main challenge: \hfill No provers for \emph{Higher-order Modal Logic\/} (\textcolor{red}{HML}) \\[1em]
+\hskip-1em Challenge: \hfill No provers for \emph{Higher-order Quantified Modal Logic\/} (\textcolor{red}{QML}) \\[1em]
 
-Our solution: \hfill Embedding in \emph{Higher-order Classical Logic\/} (\textcolor{blue}{HOL}) \\
+\hskip-1em Our solution: \hfill Embedding in \emph{Higher-order Classical Logic\/} (\textcolor{blue}{HOL}) \\
 \,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
 
-What we did \textcolor{blue}{(rough outline for remaining
+\hskip-1em What we did \textcolor{blue}{(rough outline for remaining
   presentation!)}: \\
@@
 \end{itemize}
-Did we get new results? \hfill  \textcolor{red}{Yes --- let's discuss this later!}
+\hskip-1em  Did we get any new results? \hfill  \textcolor{red}{Yes --- let's discuss this later!}
 \end{frame}

### 2013-10-31 Formalizations/Isabelle/GoedelGodSession/document/root.tex
updates for FU talk

--- 2013-10-29 Formalizations/Isabelle/GoedelGodSession/output/document/root.tex
@@
 \usepackage{isabelle,isabellesym,amsmath,amssymb,a4wide}
+\usepackage{graphicx,xcolor}
 
@@
 \author{Christoph Benzm\"uller and Bruno Woltzenlogel Paleo}
+\date{November 1, 2013}
 \maketitle
+
+%\noindent\colorbox{gray}{\includegraphics[width=.99\textwidth]{$HOME/GoedelGod/Talks/FU-Berlin/ScottsScriptGrab}} %$
+
+\noindent\fcolorbox{gray}{white}{
+\begin{minipage}{.96\textwidth}\small
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:  \hfill 
+  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$ \\[-1.5em]
+\item[A2] A property necessarily implied \\ by a
+  positive property is positive: \phantom{b} \hfill 
+  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]$ \\[-1.5em]
+\item[T1] Positive properties are possibly exemplified: \hfill $\all
+  \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ \\[-1.5em]
+\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
+  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ \\[-1.5em]
+\item[A3]  The property of being God-like is positive: \hfill   $P(G)$ \\[-1.5em]
+\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$ \\[-1.5em]
+\item[A4]  Positive properties are necessarily positive: \hfill 
+  $\all \phi [P(\phi) \to \Box \; P(\phi)]$ \\[-1.5em]
+\item[D2] An \emph{essence} of an individual is a property possessed by it \\ and necessarily implying any of its properties:
+  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
+  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$ \\[-1.5em]
+\item[T2]  Being God-like is an essence of any
+  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ \\[-1.5em]
+\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
+  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec
+  \ex y \phi(y)]$ \\[-1.5em]
+\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$ \\[-1.5em]
+\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
+\end{itemize}
+\end{minipage}
+}
+\vskip1em
 
@@
   relational model finder.
-\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pp. 131--146.
+\newblock In {\em Proc. of ITP 2010}, LNCS 6172, pp. 131--146.
   Springer, 2010.
@@
 \newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117.
+\newblock In {\em Proc. of IJCAR 2012}, LNAI 7364, pp. 111 -- 117.
   Springer, 2012.
@@
 \newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
+\newblock LNCS 2283. Springer, 2002.
 
@@
 \newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambridge U. Press, 2004.
+\newblock Cambr. U. Press, 2004.
 

### 2013-10-31 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex

@@
 A formalization and (partial) automation of Dana Scott's version \cite{ScottNotes}
- of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is presented. 
- QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
+ of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is
+ presented. QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
  Thus, the formalization is essentially a formalization in HOL. The employed embedding 
@@
 
- The formalization presented here has been carried and formally verified in the Isabelle/HOL 
+ The formalization presented here has been carried out and verified in the Isabelle/HOL 
  proof assistant; for more information on this system see the textbook by Nipkow, 
@@
     Metis \cite{Metis} calls as given below. The Metis proofs are verified in Isabelle/HOL.
- \item For consistency checking, the model finder \cite{Nitpick} has been employed.
+ \item For consistency checking, the model finder Nitpick \cite{Nitpick} has been employed.
  \end{enumerate}%
@@
 (over individuals and over sets of individuals) and $\exists$ (over individuals) are
-lifted to type $\sigma$. Further connectives could be introduced analogously. Definitions 
-could be used instead of abbreviations.%
+lifted to type $\sigma$. The lifted connectives are \isa{m{\isasymnot}}, \isa{m{\isasymand}}, \isa{m{\isasymRightarrow}},
+\isa{{\isasymforall}}, \isa{{\isasymPi}}, and \isa{{\isasymexists}}. Further connectives could be introduced analogously. 
+Definitions could be used instead of abbreviations.%
 \end{isamarkuptext}%
@@
 \ \ \isacommand{abbreviation}\isamarkupfalse%
+\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ \ \isacommand{abbreviation}\isamarkupfalse%
 \ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
-\ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
@@
 \begin{isamarkuptext}%
-Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%
+Constant symbol \isa{P} (G\"odel's `Positive') is declared.%
 \end{isamarkuptext}%
@@
 \ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
-\ \ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\ \isanewline
 %
@@
 \begin{isamarkuptext}%
-Next, the symbol \isa{G} for "God-like" is introduced and defined 
+Next, the symbol \isa{G} for `God-like'  is introduced and defined 
 as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
-all positive properties:).%
+all positive properties).%
 \end{isamarkuptext}%
@@
 \ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isanewline
-%
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%
 \isadelimproof
-\ \ %
+\ %
 \endisadelimproof
@@
 \isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
-\ metis%
+\isacommand{by}\isamarkupfalse%
+\ {\isacharparenleft}metis\ A{\isadigit{3}}\ T{\isadigit{1}}{\isacharparenright}%
 \endisatagproof
@@
 \begin{isamarkuptext}%
-Symbol \isa{ess} for "Essence" is introduced and defined as 
+Symbol \isa{ess} for `Essence' is introduced and defined as 
 $\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
 \imp \psi(y)))$ (An \emph{essence} of an individual is a property possessed by it 
-and necessarily implying any of its properties.).%
+and necessarily implying any of its properties).%
 \end{isamarkuptext}%
@@
 \ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
-\ \ \isanewline
-%
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%
 \isadelimproof
-\ \ %
+\ %
 \endisadelimproof
@@
 \isacommand{by}\isamarkupfalse%
-\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharparenright}\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%
+\ {\isacharparenleft}metis\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%
 \endisatagproof
@@
 \begin{isamarkuptext}%
-Symbol \isa{NE}, for "Necessary Existence", is introduced and
+Symbol \isa{NE}, for `Necessary Existence', is introduced and
 defined as $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$ (Necessary 
-existence of an individual is the necessary exemplification of all its essences.).%
+existence of an individual is the necessary exemplification of all its essences).%
 \end{isamarkuptext}%
@@
 \ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isanewline
-%
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%
 \isadelimproof
-\ \ %
+\ %
 \endisadelimproof
@@
 \isatagproof
-\isacommand{using}\isamarkupfalse%
-\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
-\ metis%
+\isacommand{by}\isamarkupfalse%
+\ {\isacharparenleft}metis\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def{\isacharparenright}%
 \endisatagproof
@@
 \isadelimproof
-\isanewline
 %
@@
 \isanewline
+\isanewline
 \ \ \isacommand{corollary}\isamarkupfalse%
-\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
-\ \ \isanewline
-%
+\ C{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
+\ \ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%
 \isadelimproof
-\ \ %
+\ %
 \endisadelimproof
@@
 \isatagproof
-\isacommand{using}\isamarkupfalse%
-\ T{\isadigit{1}}\ T{\isadigit{3}}\ sym\ G{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
-\ metis%
+\isacommand{by}\isamarkupfalse%
+\ {\isacharparenleft}metis\ T{\isadigit{1}}\ T{\isadigit{3}}\ G{\isacharunderscore}def\ sym{\isacharparenright}%
 \endisatagproof
@@
 \isamarkuptrue%
+\ \ \isacommand{lemma}\isamarkupfalse%
+\ True\ \isacommand{nitpick}\isamarkupfalse%
+\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
+\isadelimproof
+\ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{oops}\isamarkupfalse%
+%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+%
+\begin{isamarkuptext}%
+It has been critisized that G\"odel's ontological argument implies what is called the 
+modal collapse. The prover Satallax \cite{Satallax} can indeed show this, but verification with 
+Metis still fails.%
+\end{isamarkuptext}%
+\isamarkuptrue%
+\ \ \isacommand{lemma}\isamarkupfalse%
+\ MC{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}p\ m{\isasymRightarrow}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
+%
+\isadelimproof
+\ \ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{using}\isamarkupfalse%
+\ T{\isadigit{2}}\ T{\isadigit{3}}\ ess{\isacharunderscore}def\ sym%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
+\ \isacommand{sledgehammer}\isamarkupfalse%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}satallax{\isacharbrackright}%
+\isadelimproof
+\ %
+\endisadelimproof
+%
+\isatagproof
+\isacommand{oops}\isamarkupfalse%
+%
+\endisatagproof
+{\isafoldproof}%
+%
+\isadelimproof
+%
+\endisadelimproof
 %

### 2013-10-31 Talks/FU-Berlin/talk-part-1.tex
update for FU talk

@@
   \end{itemize} 
-\item \textcolor{blue}{Theistic:} Successful argument should convince atheists. \\[2em]
-\item \textcolor{red}{Our:} Can computers (theorem provers) be used
+\item \textcolor{blue}{Theistic:} Successful argument should convince atheists \\[2em]
+\item \textcolor{red}{Our:} Can computers (theorem provers) be used \ldots
   \begin{itemize}
-  \item to formalize the definitions and axioms?
-  \item to verify the arguments step-by-step?
-  \item to fully automate (sub-)arguments? \\[1em]
+  \item \ldots to formalize the definitions, axioms and theorems?
+  \item \ldots to verify the arguments step-by-step?
+  \item \ldots to fully automate (sub-)arguments? \\[1em]
   \end{itemize}

### 2013-10-31 Talks/FU-Berlin/talk-part-5.tex
update for FU talk

@@
 \begin{transitionframe}{} \Large \centering
-\textbf{Part D:} \\
+\textbf{Part D:} \\[.5em]
 \quad automation \& verification: proof assistant \textsc{Isabelle} \\[2em]
@@
 
-\begin{frame}{Automation \& Verification in Proof Assistant \textsc{Isabelle/HOL}} \Large
+\begin{frame}{Automation \& Verification in Proof Assistant \textsc{Isabelle/HOL}} \large
 Isabelle/HOL   (Cambridge University/TU Munich)
@@
 \vfill
-What have we done?
+What we did?
 \begin{itemize}
 \item Proof automation of G\"odel's proof script (Scott version)
-\item \textsc{Sledghammer} makes calls to remote THF provers in Miami
+\item \textsc{Sledgehammer} makes calls to remote THF provers in Miami
 \item These calls the suggest respective calls to the \textsc{Metis} prover
@@
 \end{frame}
+
+\begin{frame}{Automation \& Verification in Proof Assistant
+    \textsc{Isabelle/HOL}} \large
+
+\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-4.tex
Add Coq slide
Add transition frames
Fix font size

@@
-\begin{frame}{Coq Proof} \small
-todo
+\begin{transitionframe}
+\textbf{Part C:}
+
+Formalization and Verification in Coq
+\end{transitionframe}
+
+\begin{frame}{Coq Proof}{Demo}
+\begin{itemize}
+\item \begin{LARGE} Goal: verification of the natural deduction proof \end{LARGE}
+\begin{itemize}
+\item \begin{large} Step-by-step formalization \end{large}
+\pause
+\item \begin{large} Almost no automation (intentionally!) \end{large}
+\end{itemize}
+%
+\pause
+\item \begin{LARGE} Interesting facts: \end{LARGE}
+\begin{itemize}
+\item \begin{large} Embedding is transparent to the user \end{large}
+\pause
+\item \begin{large} Embedding gives labeled calculus for free \end{large}
+\end{itemize}
+\end{itemize}
 \end{frame}
+
+\begin{frame}{Coq Proof}
+ToDo: Screenshot
+\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-6.tex

@@
 \end{frame}
-
-\begin{transitionframe}
-\textbf{Part F:}
-
-Conclusions
-\end{transitionframe}
-
-\begin{frame}{Summary of Results about the Ontological Proof}
-\begin{itemize}[<+->]
-\item K sufficient for T1, C1 and T2 
-\item S5 not needed for T3
-\item KB sufficient for T3 
-\item A simpler new proof of C1
-\item G\"odel's original axioms (without conjunct $\phi(x)$ in D2) are inconsistent
-\item Scott's axioms are consistent
-\item For T1, only half of A1 (A1a) is needed 
-\item For T2, the other half (A1b) is needed
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Summary of Results for Logic} \small
-\begin{itemize}[<+->]
-\item Infra-structure for reasoning with modal logic using existing proof assistants and higher-order automated theorem provers
-\item A new natural deduction calculus for higher-order modal logic
-\item Difficult benchmarks for higher-order automated theorem provers
-\end{itemize}
-\end{frame}
-
-\begin{frame}[plain]
-\colorbox{black}{\includegraphics[width=\textwidth]{TUWien-GodComputerC}}
-\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-7.tex

\begin{transitionframe}
\textbf{Part F:}

Conclusions
\end{transitionframe}

\begin{frame}{Summary of Results about the Ontological Proof}
\begin{itemize}[<+->]
\item K sufficient for T1, C1 and T2 
\item S5 not needed for T3
\item KB sufficient for T3 
\item A simpler new proof of C1
\item G\"odel's original axioms (without conjunct $\phi(x)$ in D2) are inconsistent
\item Scott's axioms are consistent
\item For T1, only half of A1 (A1a) is needed 
\item For T2, the other half (A1b) is needed
\end{itemize}
\end{frame}

\begin{frame}{Summary of Results for Logic} \small
\begin{itemize}[<+->]
\item Infra-structure for reasoning with modal logic using existing proof assistants and higher-order automated theorem provers
\item A new natural deduction calculus for higher-order modal logic
\item Difficult benchmarks for higher-order automated theorem provers
\end{itemize}
\end{frame}

\begin{frame}{Conclusion} \large
What have we achieved \\[.5em]
\begin{itemize}
\item Verification of G\"odel's ontological argument with HOL provers
  \begin{itemize}
  \item exact parameters known: constant domain quantification, Henkin Semantics
  \item parameters can be varied and experiments can repeated
  \end{itemize}
\item Gained some novel results and insights
\item Major  step towards \alert{Computer-assisted Theoretical Philosophy}
 \begin{itemize}
  \item see also Ed Zalta's \emph{Computational Metaphysics} project at Stanford University
  \item remember Leibniz' dictum --- \emph{Calculemus!}
  \end{itemize}
\item Highly fascinating bridge between CS, Philosophy and Theology
\item Major public interest 
\end{itemize}
\vfill
Future Work
\begin{itemize}
\item 
\end{itemize}
\end{frame}

\begin{frame}[plain]
\colorbox{black}{\includegraphics[width=\textwidth]{TUWien-GodComputerC}}
\end{frame}

### 2013-10-31 Talks/FU-Berlin/talk-part-7.tex
update for FU talk

@@
 
-\begin{frame}{Summary of Results about the Ontological Proof}
-\begin{itemize}[<+->]
-\item K sufficient for T1, C1 and T2 
-\item S5 not needed for T3
-\item KB sufficient for T3 
-\item A simpler new proof of C1
-\item G\"odel's original axioms (without conjunct $\phi(x)$ in D2) are inconsistent
+\begin{frame}{Summary of Results} \large
+
+The (\alert{new}) insights we gained from experiments include:\\[.5em]
+\begin{itemize}
+\item Logic K sufficient for T1, C and T2 
+\item Logic S5 not needed for T3
+\item \alert{Logic KB sufficient for T3 (not well known)}
+\item \alert{We found a simpler new proof of C}
+\item \alert{G\"odel's axioms (without conjunct $\phi(x)$ in D2) are inconsistent}
 \item Scott's axioms are consistent
@@
 
-\begin{frame}{Summary of Results for Logic} \small
-\begin{itemize}[<+->]
-\item Infra-structure for reasoning with modal logic using existing proof assistants and higher-order automated theorem provers
+\begin{frame}{Summary of Results} \large
+
+Our novel contributions the  theorem proving community includes \\[.5em]
+\begin{itemize}
+\item Powerful infrastructure for reasoning with higher-order modal logic using existing proof assistants and HOL provers
 \item A new natural deduction calculus for higher-order modal logic
-\item Difficult benchmarks for higher-order automated theorem provers
+\item Difficult new benchmarks problems for HOL provers
+\item Huge media attention
 \end{itemize}
@@
 \begin{frame}{Conclusion} \large
-What have we achieved \\[.5em]
+\vskip-1em What have we achieved \\[.5em]
 \begin{itemize}
@@
   \end{itemize}
-\item Highly fascinating bridge between CS, Philosophy and Theology
-\item Major public interest 
+\item Interesting bridge between CS, Philosophy and Theology
 \end{itemize}
+
+\pause
+
 \vfill
-Future Work
+\vskip-1em Ongoing and future work \\[.5em]
 \begin{itemize}
-\item 
+\item Formalize and verify literature on ontological arguments
+  \begin{itemize}
+  \item \ldots in particular the criticism and improvements to G\"odel 
   \end{itemize}
+\item Own contributions --- supported by theorem provers
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Some Comments and Reactions}
+\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment1}}\\[.7em]
+
+\, \hfill \colorbox{gray}{\includegraphics[width=.7\textwidth]{Comment2}}\\[.7em]
+
+\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment3}}\\[1em]
+
+\, \hfill \ldots find more on the internet \ldots
 \end{frame}

### 2013-10-31 Formalizations/THF/call_tptp.sh
Improve call_tptp script.

@@ -4,17 +4,18 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-timeout=90
+timeout=60
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
 
 for file in "$@"
 do
+echo
     for prover in $proverlist 
     do
     function_to_fork () { 
-        echo $prover : $file ++++++ `$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover | grep RESULT` 
+        echo $file ++++++ `$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover | grep RESULT` 
     }
     function_to_fork &
     done

### 2013-10-31 Talks/FU-Berlin/talk-part-1.tex
Organize images

@@
 \begin{minipage}{.56\textwidth} 
-\colorbox{gray}{\includegraphics[width=\textwidth]{spiegel1}} 
+\colorbox{gray}{\includegraphics[width=\textwidth]{Images/News/spiegel1}} 
 \vskip1em
@@
 \begin{frame}{Introduction --- Quick answers to your most pressing questions!} \large
-\colorbox{gray}{\includegraphics[width=\textwidth]{MacBookGrab}} 
+\colorbox{gray}{\includegraphics[width=\textwidth]{Images/News/MacBookGrab}} 
 \pause
@@
 \ldots \hfill
-\includegraphics[height=2cm]{buch3.jpg} \hfill
-\includegraphics[height=2cm]{buch2.jpg} \hfill 
-\includegraphics[height=2cm]{buch4.jpg} \hfill
-\includegraphics[height=2cm]{buch5.jpg} \hfill
-\includegraphics[height=2cm]{buch6.jpg} \hfill
-\includegraphics[height=2cm]{buch1.jpg} \hfill
+\includegraphics[height=2cm]{Images/Books/buch3.jpg} \hfill
+\includegraphics[height=2cm]{Images/Books/buch2.jpg} \hfill 
+\includegraphics[height=2cm]{Images/Books/buch4.jpg} \hfill
+\includegraphics[height=2cm]{Images/Books/buch5.jpg} \hfill
+\includegraphics[height=2cm]{Images/Books/buch6.jpg} \hfill
+\includegraphics[height=2cm]{Images/Books/buch1.jpg} \hfill
 \ldots
@@
 \item \textcolor{blue}{Theistic:} Successful argument should convince atheists \\[2em]
-\item \textcolor{red}{Our:} Can computers (theorem provers) be used \ldots
+\item \textcolor{red}{Ours:} Can computers (theorem provers) be used \ldots
   \begin{itemize}

### 2013-10-31 Talks/FU-Berlin/talk-part-7.tex
Organize images

@@
-\begin{transitionframe}
+\begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}
 \textbf{Part F:}
@@
 \begin{frame}{Some Comments and Reactions}
-\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment1}}\\[.7em]
+\colorbox{gray}{\includegraphics[width=.8\textwidth]{Images/Comments/Comment1}}\\[.7em]
 
-\, \hfill \colorbox{gray}{\includegraphics[width=.7\textwidth]{Comment2}}\\[.7em]
+\, \hfill \colorbox{gray}{\includegraphics[width=.7\textwidth]{Images/Comments/Comment2}}\\[.7em]
 
-\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment3}}\\[1em]
+\colorbox{gray}{\includegraphics[width=.8\textwidth]{Images/Comments/Comment3}}\\[1em]
 
@@
 \begin{frame}[plain]
-\colorbox{black}{\includegraphics[width=\textwidth]{TUWien-GodComputerC}}
+\colorbox{black}{\includegraphics[width=\textwidth]{Images/Transitions/GodComputerC}}
 \end{frame}

### 2013-11-01 Talks/FU-Berlin/talk-part-7.tex
Fix images
Change transition images

@@
-\begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}
+\begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}{black}
 \textbf{Part F:}
@@
 
-Our novel contributions the  theorem proving community includes \\[.5em]
+Our novel contributions to the theorem proving community include \\[.5em]
 \begin{itemize}
@@
 \end{frame}
+
+\begin{frame}{Licenses} \centering
+\includegraphics[scale=0.5]{Images/CC-BY-SA.png}
+
+\bigskip
+\bigskip
+
+\begin{center}
+The following images used in these slides were obtained in commons.wikimedia.org and are licensed as follows:
+
+\bigskip
+
+CC-BY-SA:
+
+ReligiousSymbols, PaganReligiousSymbols, NoGod.
+
+\bigskip
+
+Public Domain:
+
+TrifidNebula
+\end{center}
+\end{frame}

### 2013-11-04 Abstract/IWIL-2013/abstract.tex
Abstract for IWIL invited paper

--- 2013-09-10 Abstract/ArXiv/GodProof-Abstract.tex
@@
-\documentclass{llncs}
+\documentclass{easychair}
 
-\usepackage{url,amsmath,amssymb,a4wide}
+\usepackage{amsmath,amssymb}
 
@@
 
-\title{Formalization, Mechanization and Automation % of Variants
-  of \\ G\"{o}del's Proof of God's Existence\thanks{This work has been
+\title{G\"{o}del's God on the Computer\thanks{This work has been
     supported by the German Research Foundation under grant
@@
 \authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}
+\titlerunning{G\"{o}del's God on the Computer}
 
 \institute{
-  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
+ Freie Universit\"{a}t Berlin, Germany \\
+ \email{c.benzmueller@fu-berlin.de} \\
   \and 
-  Theory and Logic Group, Vienna University of Technology, Austria \\
-  \email{bruno@logic.at}
+ Vienna University of Technology, Austria \\
+  \email{bruno@logic.at} \\
 }
@@
 been considered by G\"{o}del and by several philosophers who commented
-on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+on his proof; cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}.
+We have analyzed Scott's version of G\"{o}del's proof \cite{ScottNotes} for the first-time
+with an unprecedent degree of detail 
+and formality with the help of theorem provers; cf.~\url{https://github.com/FormalTheology/GoedelGod}.
 
-Dana Scott's version of G\"odel's proof \cite{ScottNotes} employs the
-following axioms (\textbf{A}), definitions (\textbf{D}), corollaries
-(\textbf{C}) and theorems (\textbf{T}), and it proceeds in the
-following order:\footnote{ A1, A2, A5, D1, D3 are logically
-  equivalent to, respectively, axioms 2, 5 and 4 and definitions 1 and
-  3 in G\"odel's notes \cite{Goedel1970,GoedelNotes}. 
-  A3 was introduced by Scott \cite{ScottNotes} and 
-  could be derived from G\"odel's axiom 1 and
-  D1 in a logic with infinitary conjunction. 
-  A4 is a weaker form of G\"odel's axiom 3. 
-  D2 has an extra conjunct $\phi(x)$ lacking in G\"odel's definition 2; 
-  this is believed to have been 
-  an oversight by G\"odel \cite{Hazen}.}
+The following has been done (and in this order):
+(i) a detailed natural deduction proof;
+%
+(ii) a formalization in TPTP THF syntax \cite{THF};
+%
+(iii) an automatic verification of the consistency of the axioms and 
+definitions with Nitpick \cite{Nitpick};
+%
+(iv) an automatic demonstration of the theorems with the provers LEO-II \cite{LEO} and Satallax \cite{Satallax};
+%
+(v) a step-by-step formalization using the Coq proof assistant \cite{Coq};
+%
+(vi) a formalization using the Isabelle proof assistant \cite{Isabelle}, where the
+theorems have been automated with the Isabelle tools Sledgehammer \cite{Sledgehammer} and Metis \cite{Metis}.
 
-\allowdisplaybreaks[1] 
-\begin{description}
-\item[A1] Either a property or its negation is positive, but not
-  both:  \hfill 
-  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$
-\item[A2] A property necessarily implied \\ by a
-  positive property is positive: \phantom{b} \hfill 
-  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]$ 
-\item[T1] Positive properties are possibly exemplified: \hfill $\all \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$
-\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
-  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$
-\item[A3]  The property of being God-like is positive: \hfill   $P(G)$
-\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$
-\item[A4]  Positive properties are necessarily positive: \hfill 
-  $\all \phi [P(\phi) \to \Box \; P(\phi)]$
-\item[D2] An \emph{essence} of an individual is  \\ a property possessed by it and \\ necessarily implying any of its properties:
-  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
-  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$
-\item[T2]  Being God-like is an essence of any
-  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ 
-\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
-  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$
-\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$
-\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
-\end{description}
-
-\noindent
-Scott's version of G\"{o}del's proof has now been 
-analysed for the first-time
-with an unprecedent degree of detail 
-and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
-The following has been done (and in this order):
-\begin{itemize}
-\item A detailed natural deduction proof.
-%
-\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
-%
-\item Automatic verification of the consistency of the axioms and 
-definitions with Nitpick \cite{Nitpick}.
-%
-\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.
-
-\item A step-by-step formalization using the Coq proof assistant \cite{Coq}.
-
-\item A formalization using the Isabelle proof assistant \cite{Isabelle}, where the theorems (and some additional lemmata) have been automated with Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof}.
-\end{itemize}
+Furthermore, the multi-step THF development of the proof has
+recently been encoded in Sutcliffe's new TPI scripting
+language~\cite{}. This scripting language offers important features
+that were missing in the TPTP world so far. Our collaboration with
+Sutcliffe has also led to some improvements of the TPI langauge.
 
@@
 
-In our ongoing computer-assisted study of G\"odel's proof we have
-obtained the following results:
-\begin{itemize}
-\item The basic modal logic K is sufficient for proving T1, C and T2. 
-\item Modal logic S5 is not needed for proving T3; the logic KB is
-  sufficient. 
-\item Without the first conjunct $\phi(x)$ in D2 the set of axioms 
-  and definitions would be inconsistent.
-\item For proving theorem T1, only the left to right direction of
-  axiom A1 is needed. However, the backward direction of A1 is
-  required for proving T2.
-
-\end{itemize}
+In our ongoing computer-assisted study of G\"odel's proof the proof
+tools have made some interesting observations, some of which were 
+unknown so far.
 
@@
 calculate!
+
+\small
 
@@
 
-\bibitem{LEO-II}
+\bibitem{LEO}
 C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
@@
 
-\bibitem{Hazen}
-A.P. Hazen.
-\newblock On g\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+% \bibitem{Hazen}
+% A.P. Hazen.
+% \newblock On g\"odel's ontological proof.
+% \newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
 
-\bibitem{Hurd03first-orderproof}
+\bibitem{Metis}
 J.~Hurd.
@@
 
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
+% \bibitem{FormalTheologyRepository}
+% B.~Woltzenlogel Paleo and C.~Benzm\"uller.
+% \newblock Formal theology repository
+%   (\url{https://github.com/FormalTheology/GoedelGod}).
 
@@
 
-\bibitem{J22}
+\bibitem{THF}
 G.~Sutcliffe and C.~Benzm{\"u}ller.

### 2013-11-04 Abstract/IWIL-2013/abstract.tex
Abstract for IWIL

@@
 recently been encoded in Sutcliffe's new TPI scripting
-language~\cite{}. This scripting language offers important features
+language~\cite{TPI}. This scripting language offers important features
 that were missing in the TPTP world so far. Our collaboration with
-Sutcliffe has also led to some improvements of the TPI langauge.
+Sutcliffe has also led to some improvements of the TPI language.
 
@@
 
+\bibitem{THF}
+G.~Sutcliffe.
+\newblock The TPTP TPI Scripting Language (check!).
+\newblock {\em Proc. of IWIL-10}, 2013.
+
 \end{thebibliography}

### 2013-11-04 Abstract/IWIL-2013/abstract.tex
Abstract for IWIL

@@
 We have analyzed Scott's version of G\"{o}del's proof \cite{ScottNotes} for the first-time
-with an unprecedent degree of detail 
+with an unprecedented degree of detail 
 and formality with the help of theorem provers; cf.~\url{https://github.com/FormalTheology/GoedelGod}.
@@
 
-\bibitem{THF}
+\bibitem{TPI}
 G.~Sutcliffe.

### 2013-11-04 Abstract/IWIL-2013/abstract.tex
Abstract for IWIL
more press links

@@
 
-\title{G\"{o}del's God on the Computer\thanks{This work has been
+\title{G\"{o}del's God on the Computer (Abstract)\thanks{This work has been
     supported by the German Research Foundation under grant
@@
 
-In our ongoing computer-assisted study of G\"odel's proof the proof
-tools have made some interesting observations, some of which were 
-unknown so far.
-
-This work attests the maturity of contemporary interactive and
-automated deduction tools for classical higher-order logic and
-demonstrates the elegance and practical relevance of the
-embeddings-based approach.  Most importantly, our work opens new
-perspectives for a computer-assisted theoretical philosophy.  The
-critical discussion of the underlying concepts, definitions and axioms
-remains a human responsibility, but the computer can assist in
-building and checking rigorously correct logical arguments. In case of
-logico-philosophical disputes, the computer can check the disputing
-arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
-calculate!
+In our ongoing computer-assisted study of G\"odel's proof, the
+automated reasoning tools have made some interesting observations,
+some of which were unknown so far. This attests the maturity of
+contemporary interactive and automated deduction tools for classical
+higher-order logic and demonstrates the elegance and practical
+relevance of the embeddings-based approach.  Most importantly, our
+work opens new perspectives for a computer-assisted theoretical
+philosophy.  The critical discussion of the underlying concepts,
+definitions and axioms remains a human responsibility, but the
+computer can assist in building and checking rigorously correct
+logical arguments. In case of logico-philosophical disputes, the
+computer can check the disputing arguments and partially fulfill
+Leibniz' dictum: Calculemus --- Let us calculate!
 

### 2013-11-04 Abstract/IWIL-2013/abstract.tex
more press links
update of IWIL abstract
several updates wrt to tpi script and Isabelle session

@@
 
-\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}
+\authorrunning{C. Benzm\"{u}ller and B. Woltzenlogel Paleo}
 \titlerunning{G\"{o}del's God on the Computer}
@@
 embedding of the modal logic into classical higher-order logic with
-Henkin semantics. The formalization is thus essentially
+Henkin semantics  \cite{J23,B9}. The formalization is thus essentially
 done in classical higher-order logic where  quantified modal logic is
@@
 automated reasoning tools have made some interesting observations,
-some of which were unknown so far. This attests the maturity of
+some of which were unknown so far. Our work attests the maturity of
 contemporary interactive and automated deduction tools for classical
@@
 Leibniz' dictum: Calculemus --- Let us calculate!
-
-\small
 
@@
 \newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
+  Science, and Physics: Lecture Notes in Logic 6}}, pages 167--172. {Springer},
+  1996.
 
@@
 \newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
+  Birthday}}, pages 386--406. College Publications, 2008.
 
@@
 \bibitem{LEO}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
+C.~Benzm{\"u}ller, F.~Theiss, L.C. Paulson, and A.~Fietzke.
 \newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
   logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
-  162--170. Springer, 2008.
+\newblock In A.~Armando, P.~Baumgartner, and G.~Dowek, editors, {\em Automated
+  Reasoning, 4th International Joint Conference, IJCAR 2008, Sydney, Australia,
+  August 12-15, 2008, Proceedings}, volume 5195 of {\em LNCS}, pages 162--170.
+  Springer, 2008.
 
@@
 C.E. Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+\newblock Satallax: An automatic higher-order prover.
+\newblock In B.~Gramlich, D.~Miller, and U.~Sattler, editors, {\em Automated
+  Reasoning, 6th International Joint Conference, IJCAR 2012, Manchester, UK,
+  June 26-29, 2012. Proceedings}, volume 7364 of {\em LNCS}, pages 111--117.
   Springer, 2012.
@@
 
-% \bibitem{Hazen}
-% A.P. Hazen.
-% \newblock On g\"odel's ontological proof.
-% \newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
-
 \bibitem{Metis}
@@
 \newblock In {\em Design and Application of Strategies/Tactics in Higher Order
-  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
+  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, pages 56--68, 2003.
 
@@
 \newblock Number 2283 in LNCS. Springer, 2002.
-
-% \bibitem{FormalTheologyRepository}
-% B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-% \newblock Formal theology repository
-%   (\url{https://github.com/FormalTheology/GoedelGod}).
 
@@
 
+\bibitem{TPI}
+G.~Sutcliffe.
+\newblock {The TPTP Process Instruction Language}.
+\newblock In S.~Schulz, G.~Sutcliffe, and B.~Konev, editors, {\em {Proceedings
+  of the 10th International Workshop on the Implementation of Logics}},
+  EasyChair Proceedings in Computing, 2013.
+\newblock To appear.
+
 \bibitem{THF}
@@
 
-\bibitem{TPI}
-G.~Sutcliffe.
-\newblock The TPTP TPI Scripting Language (check!).
-\newblock {\em Proc. of IWIL-10}, 2013.
-
 \end{thebibliography}

### 2013-11-05 Formalizations/THF/call_tptp.sh
minor updates

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-timeout=60
+timeout=120
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)

### 2013-11-06 Talks/TU-Wien/talk-part-7.tex
Add TU-Wien Talk

--- 2013-11-01 Talks/FU-Berlin/talk-part-7.tex
@@
 \begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}{black}
-\textbf{Part F:}
-
-Conclusions
+\textbf{Conclusions}
 \end{transitionframe}
@@
 \item Logic K sufficient for T1, C and T2 
-\item Logic S5 not needed for T3
+\item Logic S5 \emph{not} needed for T3
 \item \alert{Logic KB sufficient for T3 (not well known)}
@@
 
-Our novel contributions to the theorem proving community include \\[.5em]
 \begin{itemize}
-\item Powerful infrastructure for reasoning with higher-order modal logic using existing proof assistants and HOL provers
-\item A new natural deduction calculus for higher-order modal logic
-\item Difficult new benchmarks problems for HOL provers
+\item Contributions to the theorem proving community include \\[.5em]
+\begin{itemize}
+\item Powerful infrastructure for reasoning with QML
+\item A (new?) natural deduction calculus for \\ higher-order (intuitionistic) modal logic
+\item Difficult new benchmark problems for HOL provers
 \item Huge media attention
 \end{itemize}
-\end{frame}
 
-\begin{frame}{Conclusion} \large
-\vskip-1em What have we achieved \\[.5em]
-\begin{itemize}
-\item Verification of G\"odel's ontological argument with HOL provers
-  \begin{itemize}
-  \item exact parameters known: constant domain quantification, Henkin Semantics
-  \item parameters can be varied and experiments can repeated
-  \end{itemize}
-\item Gained some novel results and insights
+\item Interesting bridge between CS, Philosophy and Theology
+
 \item Major step towards \alert{Computer-assisted Theoretical Philosophy}
@@
   \end{itemize}
-\item Interesting bridge between CS, Philosophy and Theology
 \end{itemize}
+\end{frame}
 
-\pause
-
-\vfill
-\vskip-1em Ongoing and future work \\[.5em]
+\begin{frame}{Ongoing and Future Work} \large
 \begin{itemize}
-\item Formalize and verify literature on ontological arguments
+\item Formalize and verify other ontological arguments
   \begin{itemize}
-  \item \ldots in particular the criticism and improvements to G\"odel 
+  \item \ldots particularly the criticisms and improvements
   \end{itemize}
-\item Own contributions --- supported by theorem provers
+\item Experiment with other embeddings (e.g. varying domains)
+\item Eliminate and introduce cuts
 \end{itemize}
@@
 
-ReligiousSymbols, PaganReligiousSymbols, NoGod.
+ReligiousSymbols, PaganReligiousSymbols.
 

### 2013-11-08 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
several updates wrt to tpi script and Isabelle session
Improve LaTeX

@@
 \begin{isamarkuptext}%
-A formalization and (partial) automation of Dana Scott's version \cite{ScottNotes}
- of Goedel's ontological argument \cite{GoedelNotes} in quantified modal logic KB (QML KB) is
- presented. QML KB is in turn modeled as a fragment of classical higher-order logic (HOL). 
- Thus, the formalization is essentially a formalization in HOL. The employed embedding 
+Dana Scott's version \cite{ScottNotes}
+ of Goedel's ontological argument \cite{GoedelNotes} for God's existence is here
+ formalized in quantified modal logic KB (QML KB) within the proof assistant Isabelle/HOL. 
+ QML KB is  modeled as a fragment of classical higher-order logic (HOL); 
+ thus, the formalization is essentially a formalization in HOL. The employed embedding 
  of QML KB in HOL is adapting the work of Benzm\"uller and Paulson \cite{J23,B9}.
  Note that the QML KB formalization employs quantification over individuals and 
- quantification over sets of individuals (poperties).
+ quantification over sets of individuals (properties).
 
- The formalization presented here has been carried out and verified in the Isabelle/HOL 
- proof assistant; for more information on this system see the textbook by Nipkow, 
- Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
+ The gaps in Scott's proof have been automated 
+ with Sledgehammer \cite{Sledgehammer}, performing remote calls to the higher-order automated
+ theorem prover LEO-II \cite{LEO-II}. Sledgehammer then suggests the 
+ Metis \cite{Metis} calls. The Metis proofs are verified by Isabelle/HOL.
+ For consistency checking, the model finder Nitpick \cite{Nitpick} has been employed.
+
+ Isabelle is described in the textbook by Nipkow, 
+ Paulson, and Wenzel \cite{Isabelle} and in tutorials available 
  at: \url{http://isabelle.in.tum.de}.
 
- Some further notes: \sloppy
- \begin{enumerate}
- \item This LaTeX text document has been produced automatically from the Isabelle source
- code document at 
- \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle/GoedelGodSession} 
- with the Isabelle build tool.
- \item The formalization presented here is related to the THF \cite{J22} and 
+\subsection{Related Work}
+
+ The formalization presented here is related to the THF \cite{J22} and 
  Coq \cite{Coq} formalizations at 
  \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/}.
- \item All reasoning gaps in Scott's proof script have been automated 
-    with Sledgehammer \cite{Sledgehammer}, performing remote calls to the higher-order automated
-    theorem prover LEO-II \cite{LEO-II}. These calls suggest the 
-    Metis \cite{Metis} calls as given below. The Metis proofs are verified in Isabelle/HOL.
- \item For consistency checking, the model finder Nitpick \cite{Nitpick} has been employed.
+
+ A medieval ontological argument by Anselm was formalized in PVS by John Rushby \cite{ToDo}.
  \end{enumerate}%
@@
 \begin{isamarkuptext}%
-The types \isa{i} for possible worlds (or states) and $\mu$ for individuals 
+The types \isa{i} for possible worlds and $\mu$ for individuals 
 are introduced.%
@@
 \begin{isamarkuptext}%
-QML formulas are identified with certain HOL terms of type \isa{i\ {\isasymRightarrow}\ bool}. 
-This type will be abbreviated in the remainder as \isa{{\isasymsigma}}.%
+QML formulas are translated as HOL terms of type \isa{i\ {\isasymRightarrow}\ bool}. 
+This type is abbreviated as \isa{{\isasymsigma}}.%
 \end{isamarkuptext}%
@@
 lifted to type $\sigma$. The lifted connectives are \isa{m{\isasymnot}}, \isa{m{\isasymand}}, \isa{m{\isasymRightarrow}},
-\isa{{\isasymforall}}, \isa{{\isasymPi}}, and \isa{{\isasymexists}}. Further connectives could be introduced analogously. 
+\isa{{\isasymforall}}, \isa{{\isasymPi}}, and \isa{{\isasymexists}}. Other connectives could be introduced analogously. 
 Definitions could be used instead of abbreviations.%
@@
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
@@
 \ \ \isacommand{abbreviation}\isamarkupfalse%
-\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
+\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ \ w\ r\ v\ {\isasymlongrightarrow}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline
 \ \ \isacommand{abbreviation}\isamarkupfalse%
@@
 \begin{isamarkuptext}%
-The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \phi 
-[P(\neg \phi) \biimp \neg P(\phi)]$ (Either a property or its negation is positive, but not both.) 
-and \isa{A{\isadigit{2}}}: $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x) \imp \psi(x)]) 
+The meaning of \isa{P} is restricted by axioms \isa{A{\isadigit{1}}{\isacharparenleft}a{\isacharslash}b{\isacharparenright}}: $\all \varphi 
+[P(\neg \varphi) \biimp \neg P(\varphi)]$ (Either a property or its negation is positive, but not both.) 
+and \isa{A{\isadigit{2}}}: $\all \varphi \all \psi [(P(\varphi) \wedge \nec \all x [\varphi(x) \imp \psi(x)]) 
 \imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
@@
 \ \isakeyword{where}\isanewline
-\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
-\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPhi}\ x\ m{\isasymRightarrow}\ {\isasympsi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymphi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymphi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymphi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
+\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymphi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymphi}\ x\ m{\isasymRightarrow}\ {\isasympsi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
@@
 properties are possibly exemplified).
-T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}}.
-This successful attempt then suggests to
-instead try the Metis call in the line below. This Metis call generates a proof object 
+T1 is proved directly by Sledgehammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}}.
+Sledgehammer suggests to call Metis with axioms A1a and A2. 
+Metis sucesfully generates a proof object 
 that is verified in Isabelle/HOL's kernel.%
@@
 \ \ \isacommand{theorem}\isamarkupfalse%
-\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
+\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ P\ {\isasymphi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
 \ \ \isacommand{sledgehammer}\isamarkupfalse%
@@
 Next, the symbol \isa{G} for `God-like'  is introduced and defined 
-as $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ (A God-like being possesses 
+as $G(x) \biimp \forall \varphi [P(\phi) \to \varphi(x)]$ \\ (A God-like being possesses 
 all positive properties).%
@@
 \ \ \isacommand{definition}\isamarkupfalse%
-\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ P\ {\isasymphi}\ m{\isasymRightarrow}\ {\isasymphi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
@@
 \ \ \isacommand{axiomatization}\isamarkupfalse%
-\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
+\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ P\ {\isasymphi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
 Symbol \isa{ess} for `Essence' is introduced and defined as 
-$\ess{\phi}{x} \biimp \phi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\phi(y) 
+$\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi (\psi(x) \imp \nec \all y (\varphi(y) 
 \imp \psi(y)))$ (An \emph{essence} of an individual is a property possessed by it 
@@
 \ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
-\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ ess\ x\ {\isacharequal}\ {\isasymphi}\ x\ m{\isasymand}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymphi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
@@
 Symbol \isa{NE}, for `Necessary Existence', is introduced and
-defined as $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec \ex y \phi(y)]$ (Necessary 
+defined as $\NE(x) \biimp \all \varphi [\ess{\varphi}{x} \imp \nec \ex y \varphi(y)]$ (Necessary 
 existence of an individual is the necessary exemplification of all its essences).%
@@
 \ \ \isacommand{definition}\isamarkupfalse%
-\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
+\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isasymphi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
 \begin{isamarkuptext}%
@@
 \ \ \isacommand{sledgehammer}\isamarkupfalse%
-\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%
+\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}{\isacharparenleft}T{\isadigit{1}}\ T{\isadigit{3}}\ G{\isacharunderscore}def\ sym{\isacharparenright}%
 \isadelimproof
@@
 \isacommand{oops}\isamarkupfalse%
+\isanewline
 %
@@
 \endisadelimproof
-%
-\begin{isamarkuptext}%
-\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided 
-very important help wrt consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
-produce latex documents from Isabelle sources, and he showed us useful tricks in Isabelle.%
-\end{isamarkuptext}%
-\isamarkuptrue%
 %

### 2013-11-08 Formalizations/Isabelle/ScottsVersion/output/document/root.tex
Reorganize folders

--- 2013-10-31 Formalizations/Isabelle/GoedelGodSession/document/root.tex
@@
 \author{Christoph Benzm\"uller and Bruno Woltzenlogel Paleo}
-\date{November 1, 2013}
+%\date{November 1, 2013}
 \maketitle
@@
   $\all \phi [P(\phi) \to \Box \; P(\phi)]$ \\[-1.5em]
-\item[D2] An \emph{essence} of an individual is a property possessed by it \\ and necessarily implying any of its properties:
+\item[D2] An \emph{essence} of an individual is a property possessed by it \\ and necessarily implying any of its properties: \\
   \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
@@
 
-\small
+\paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided 
+very important help on issues related to consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
+produce latex documents from Isabelle sources and showed us useful tricks in Isabelle.
+
+%\small
 \begin{thebibliography}{10}

### 2013-11-11 Formalizations/Isabelle/ScottsVersion/output/document/GoedelGodModalCollapse.tex
Improve Isabelle paper

%
\begin{isabellebody}%
\def\isabellecontext{GoedelGodModalCollapse}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Modal Collapse%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
G\"odel's axioms have been criticized for entailing the so-called 
modal collapse. The prover Satallax \cite{Satallax} confirms this. 
However, sledgehammer does not seem to be able to determine which axioms, 
definitions and previous theorems are used by Satallax;
hence it suggests to call Metis using everything, but this (unsurprisingly) fails.
Attempting to use `Sledegehammer min' to minimize Sledgehammer's suggestion does not work.
Nevertheless, calling Metis with \isa{T{\isadigit{2}}}, \isa{T{\isadigit{3}}} and \isa{ess{\isacharunderscore}def} does work.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{lemma}\isamarkupfalse%
\ MC{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}{\isacharparenleft}{\isasymPhi}\ m{\isasymrightarrow}\ {\isacharparenleft}{\isasymbox}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}satallax{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Chris: \isa{{\isacharparenleft}metis\ T{\isadigit{2}}\ T{\isadigit{3}}\ ess{\isacharunderscore}def{\isacharparenright}} still fails on my iMac, will try on my new MacBook tomorrow and then submit.%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End:

### 2013-11-11 Formalizations/Isabelle/ScottsVersion/output/document/GoedelGodModalCollapse.tex
Improve Isabelle paper

@@
 modal collapse. The prover Satallax \cite{Satallax} confirms this. 
-However, sledgehammer does not seem to be able to determine which axioms, 
+However, sledgehammer is not able to determine which axioms, 
 definitions and previous theorems are used by Satallax;
-hence it suggests to call Metis using everything, but this (unsurprisingly) fails.
+hence it suggests to call Metis using everything, but this (unsurprinsingly) fails.
 Attempting to use `Sledegehammer min' to minimize Sledgehammer's suggestion does not work.
-Nevertheless, calling Metis with \isa{T{\isadigit{2}}}, \isa{T{\isadigit{3}}} and \isa{ess{\isacharunderscore}def} does work.%
+Calling Metis with \isa{T{\isadigit{2}}}, \isa{T{\isadigit{3}}} and \isa{ess{\isacharunderscore}def} also does not work.%
 \end{isamarkuptext}%
@@
 \isacommand{oops}\isamarkupfalse%
+\isanewline
+\ \ \isanewline
+\isanewline
 %
@@
 \endisadelimproof
-%
-\begin{isamarkuptext}%
-Chris: \isa{{\isacharparenleft}metis\ T{\isadigit{2}}\ T{\isadigit{3}}\ ess{\isacharunderscore}def{\isacharparenright}} still fails on my iMac, will try on my new MacBook tomorrow and then submit.%
-\end{isamarkuptext}%
-\isamarkuptrue%
 %

### 2013-12-13 Natural Deduction/GodProof-ND.tex
Add Annika's improvements on the natural deduction proof

@@
 \usepackage{theorems}
+\usepackage{natbib}
 
@@
 Anselm (1033-1109).
-The argument considers a maximally concievable being.
-This being must exist, because if it did not have the property of existence, then we could concieve of a greater being that apart from the other properties also has the property of existence.
-The main critique of this argument is that we do not know whether the concept maximal concievable being in fact designates anything or if it is inconsistent, like a round square.
+The argument considers a maximally conceivable being.
+This being must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
+The main critique of this argument is that we do not know whether the concept maximal conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
 The question is if the argument is sound.
@@
 As Bertrand Russell has pointed out the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have.
-Kant, on the other hand argued against the ontological argument on the basis that existence is not an analytic property.
+Kant, on the other hand argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
 This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
-All that we can say is that if God exists, then he necessarily exists.
+All that we can say is that if God exists, then God necessarily exists.
 
@@
 
-G\"odel treated positive properties as not just atomic properties, like Leibniz, but also consisting of collections of these properties, which by Leibniz argument are compatible or possible.
-(Fitting, p.139) G\"odel fomulated an axiom stating that the conjunction of any set of positive properties is positive.
+G\"odel treated positive properties as not just atomic properties, like Leibniz, but also consisting of collections of these properties, which by Leibniz argument are compatible or possible.\citep{fitting}[p.139.] G\"odel formulated an axiom stating that the conjunction of any set of positive properties is positive.
 The property of God-likeness, which defines what a God is, must therefore be an infinite conjunction of all the positive properties.
-This axiom of the positiveness of conjunctions can be given a third order fomulation which is the reason for the claim that third-order quantification may be required.
-(Fitting, 2002, p.
-148).
-However, this axiom is oly used in the proof to establish that the property of God-likeness is itself positive that led Scott to simply assume the positiveness of  God-likeness as an axiom instead.
+This axiom of the positiveness of conjunctions can be given a third order formulation which is the reason for the claim that third-order quantification may be required \citep{and}.
+%\citep{fitting}[p.
+148.]
+However, this axiom is only used in the proof to establish that the property of God-likeness is itself positive.
+G\"odel's student Scott, who wrote down a version of G\"odel's proof after G\"odel had confided his work in him, simply chose to assume the positiveness of  God-likeness as an axiom instead.
 
-Nevertheless, both axioms have the unfortunate property of being equivalent to the possibility of God's existence, given the other axioms of G\"odel.
-The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself.
-Thus, we have an axiom which seems to be equivalent to the conclusion.
+Nevertheless, both axioms have the unfortunate property of being equivalent to the possibility of God's existence, given the other axioms of G\"odel \citep{sobel2}.
+%\citep{fitting}[Section 11.4.
+Objections.]
+The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself. Thus, we have an axiom which seems to be equivalent to the conclusion.
 
-If one assumes extensionality, that two objects with the same properties are identical, then the definitions of G\"odel's proof give monotheism.
-Namely, two Gods would both be God-like and by the definition of god-like, they would have the same positive properties and be identical.
+Another, possibly undesirable, consequence of the formalism is that there can only be one God, which is defined by the axioms.
+Namely, if one assumes that the theory contains the equality relation then a God has the property of being identical to itself.
+Since any other god-like being would have at least the same properties it shares the property of being identical to the one God.
+Thus, we have proved monotheism \citep{sobel2}[Ch.
+4, section 3.3.2].
+%\citep{fitting}[The proof is Exercise 7.1, p.163].
 
-It may be so that for all formulas the formula implies its boxed version.
-That is all that is is necessarily and modality collapses.
-Fitting provides two solutions to the problem by modifying the axioms.
+In \citep{sobel} a concern is raised about the whole system of proof, which is further elaborated on in \citep{sobel2}[Ch.
+4,  section 6].
+Depending on the properties of the rules for second order quantification allowed in the system, the system may have modal collapse.
+That is, if no restrictions are placed on the properties over which quantification is allowed, then the system says that for all formulas the formula implies its boxed version.
+However, there are solutions to this problem by modification of the axioms.
+Sobel's own solution to the problem is to treat properties as extensional instead of intentional.
+Another solution by Anderson %(1990)
+is to require that Godlike beings have positive properties necessarily. A summary of these arguments can be found in Fitting \citep{fitting}[Sections 11.9 and 11.10].
 
@@
 
-ToDo: add intuitionistic rule for negation/bottom
+We let negation be a defined concept by $\neg A\equiv A\imp \bot$.
+The rules for negation introduction and elimination are special cases of the implication rules.
+Equivalence is also a defined concept with  $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
 \end{calculus}
@@
 \begin{calculus}
-{Classical Rules/Axioms}
+{Classical Rule: Double negation elimination}
 {fig:Classical}
-
-ToDo: add classical rules/axioms that we need
+\infer[\neg\neg_E]{A}{ \neg\neg A }
 \end{calculus}
@@
 \s\s\s\s
-\infer[\nec_I]{\nec A}{\fbox{\infer*{A}{} }}
-\s\s\s
-\infer[\nec_E]{\fbox{ \infer*{}{A} }  }{\nec A}
+\infer[\nec_I]{\nec A}{\eta: \fbox{\infer*{A}{}} }
+\s\s\s\s\s
+\infer[\nec_E]{n: \fbox{ \infer*{}{A} } }{\nec A}
+
+\vspace{2em}
+
+\s\s\s\s
+\infer[\pos_I]{\pos A}{n: \fbox{\infer*{A}{}} }
+\s\s\s\s\s
+\infer[\pos_E]{\eta: \fbox{ \infer*{}{A} } }{\pos A}
 
@@
 if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+%
+\item \textbf{eigen-box conditions:} $\eta$ must be a fresh name for a box.
 %
@@
             \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}
-             \BIC{$\bot $}\RightLabel{$\mathit{RAA}^1$}  
+             \BIC{$\bot $}\RightLabel{$\imp_I^1$}
+             \UIC{$\neg \neg P(\psi)$}\RightLabel{$\neg\neg_E$}
               \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
@@
                                        \UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}
-       			\BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{Necessitation}
+                               \BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{$\Box_I$}
                         \UIC{$\Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^7$}
@@
 \AXC{$ $} \RightLabel{1}
-\UIC{$\ex z. G(z)$}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
 \UIC{$G(g)$}
@@
         \AXC{Theorem \ref{T2}} \dashedLine
-        \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
-        \UIC{$G(g) \imp \ess{G}{g}$}
+        \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}\RightLabel{$\all_E$}
+        \UIC{$G(g) \imp \ess{G}{g}$}\RightLabel{$\imp_E$}
     \BIC{$\ess{G}{g}$}
@@
                         \AXC{$ $} \dashedLine
-                        \UIC{$G(g)$} \dottedLine
-                        \UIC{$\all \varphi.[P(\varphi) \imp \varphi(g)]$}
-                        \UIC{$P(E) \imp E(g)$}
-                     \BIC{$E(g)$} \dottedLine
-                     \UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}
-                     \UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}
-        \BIC{$\nec \ex x. G(x)$} \RightLabel{1}
+                        \UIC{$G(g)$} \dottedLine \RightLabel{D\ref{D1}}
+                        \UIC{$\all \varphi.[P(\varphi) \imp \varphi(g)]$}\RightLabel{$\all_E$}
+                        \UIC{$P(E) \imp E(g)$}\RightLabel{$\imp_E$}
+                     \BIC{$E(g)$} \dottedLine\RightLabel{D\ref{D3}}
+                     \UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
+                     \UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
+        \BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
         \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
@@
 
-\begin{axiom}[The necessitation rule]
+\begin{axiom}[The necessitation rule, $\Box_I$]
 \label{Necessitation}
@@
 \begin{lemma}
-\label{systemK}
-Assuming the axiom K and the necessitation rule N, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
+\label{systemK2}
+Assuming the axiom K and the necessitation rule $\Box_I$, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
 \end{lemma}
@@
 \UIC{$ B$} \RightLabel{$\imp_I^1$}
-\UIC{$A\imp B$} \RightLabel{Necessitation}
+\UIC{$A\imp B$} \RightLabel{$\Box_I$}
 \UIC{$\nec(A\imp B)$}
@@
 
-\section{Crisis in the calculus}
+\section{Modal collapse}
 
@@
 
-This collapse of the sytem is proved in (Fitting, 2002) and he provides two solutions by the modifications of the axioms, see chapter 11 section 8-11.
+This collapse of the system was first proved in \citep{sobel} and the argument is also given in \citep{fitting}[ch.
+11, sections 8-11] who provides summaries of two solutions by the modifications of the axioms.
 The argument may require that one doesn't instantiate the second order quantifier with $A$ directly but with $\lambda x.A$ to make $A$ a property.
@@
 we instantiate with $G$ and $E$.
-
-In Sobel on G\"{o}del's ontological proof by Robert E. Koons this question is raised.
+%In Sobel on G\"{o}del's ontological proof by Robert E.
+Koons this question is raised.
 
@@
 
+\begin{thebibliography}{9}
+
+\bibitem[{\itshape Anderson \& Gettings(1996)}]{and}
+Anderson, C. A.\& Gettings, M. 1996.  {\itshape G\"odel's ontological proof revisited}. In: edited by Hajek P. {\itshape G\"odel '96},  Springer.
+
+\bibitem[{\itshape Fitting(2002)}]{fitting}
+Fitting, M. 2002.  {\itshape Types, Tableaus, and G\"odel's God}, Kluwer Academic Publishers.
+
+\bibitem[{\itshape Kant(1781)}]{kant}
+Kant, I.  original 1781.   {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
+
+\bibitem[{\itshape Sobel(1987)}]{sobel}
+Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: edited by J. J. Thompson. {\itshape On being and saying : essays for Richard Cartwright},  MIT Press.
+
+\bibitem[{\itshape Sobel(2001)}]{sobel2}
+Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
+
+\end{thebibliography}
+
 \end{document}

### 2014-01-22 Papers/ITP/GodProof-ITP.tex
Rename folder of papers

\documentclass{llncs}

\usepackage{url,amsmath,amssymb,a4wide}

\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\all}{\forall}
\newcommand{\ex}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly
\newcommand{\ess}[2]{#1 \ \mathit{ess.} \ #2}
\newcommand{\NE}{\mathit{NE}}

\title{
God in Coq: \\
Modal Logic in Coq \\ 
and Scott's Version of G\"{o}del's Proof of God's Existence \\
{\small (\emph{Rough Diamond Proof Pearl})}
%   \thanks{
%   This work has been supported by the 
%   German Research Foundation under grant
%   BE2501/9-1.}
}

\author{
  Christoph Benzm\"{u}ller\inst{1} 
  \and 
  Bruno Woltzenlogel Paleo\inst{2}
}

\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}

\institute{
  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
  \email{c.benzmueller@gmail.com}
  \and 
  Theory and Logic Group, Vienna University of Technology, Austria \\
  \email{bruno@logic.at}
}

\begin{document}

\maketitle

\section{Introduction}

G\"{o}del's proof of God's existence is challenging 
to formalize and verify because it requires an 
expressive logical language with modal operators
($\Diamond$ and $\Box$) and
higher-order quantifiers. Therefore, it can serve as a good benchmark to measure and compare progress in the development of automated and interactive reasonging tools for higher-order modal logics. To fulfill this goal, formalizations \cite{FormalTheologyRepository} in TPTP THF \cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been completed so far, in this order. This paper focuses on the formalization in Coq.

An embedding of modal logics 
into higher-order logic with
Henkin semantics \cite{J23,B9} provides the necessary theoretical foundation that enables the use of interactive theorem provers for modal logics as well. In the particular case of Coq, the implementation of this embedding is discussed in detail in Section \ref{sec:modal} and Scott's version of G\"odel's proof is shown in Section \ref{sec:proof}.

\section{Modal Logic in Coq}
\label{sec:modal}

A crucial aspect of modal logics \cite{ToDo} is that the so-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$ is a theorem, but $A \imp \Box A$ is not necessarily a theorem. Naive attempts to define the modal operators $\Box$ and $\Diamond$ may easily be unsound in this respect. To avoid this issue, the \emph{possible world semantics} of modal logics can be explicitly embedded into higher-order logics \cite{J23,B9}. To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type that depends on possible worlds. Possible worlds are connected by an accessibility relation.

\begin{verbatim}
(* Type for worlds *)
Parameter i: Type.

(* Type for individuals *)
Parameter u: Type.

(* Type of modal propositions *)
Definition o := i -> Prop.

(* Acessibility relation for worlds *)
Parameter r: i -> i -> Prop.
\end{verbatim}

\noindent
All modal connectives are simply lifted versions of the usual logical connectives. Notations are used to allow the modal connectives to be used as similarly as possible as the usual connectives. The prefix ``\texttt{m}'' is used to distinguish the modal connectives: if $\odot$ is a connective on type \text{Prop}, $\texttt{m}\odot$ is a connective on the lifted type $\texttt{o}$ of modal propositions.

\begin{verbatim}
Definition mnot (p: o)(w: i) := ~ (p w).
Notation "m~  p" := (mnot p) (at level 74, right associativity).

Definition mand (p q:o)(w: i) := (p w) /\ (q w).
Notation "p m/\ q" := (mand p q) (at level 79, right associativity).

Definition mor (p q:o)(w: i) := (p w) \/ (q w).
Notation "p m\/ q" := (mor p q) (at level 79, right associativity).

Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).

Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
\end{verbatim}

\noindent
Likewise, modal quantifiers are lifted versions of the usual quantifiers.

\begin{verbatim}
Definition A {t: Type}(p: t -> o) := fun w => forall x, p x w.
Notation "'mforall'  x , p" := (A (fun x => p))
  (at level 200, x ident, right associativity) : type_scope.
Notation "'mforall' x : t , p" := (A (fun x:t => p))
  (at level 200, x ident, right associativity, 
    format "'[' 'mforall' '/ '  x  :  t , '/ '  p ']'")
  : type_scope.

Definition E {t: Type}(p: t -> o) := fun w => exists x, p x w.
Notation "'mexists' x , p" := (E (fun x => p))
  (at level 200, x ident, right associativity) : type_scope.
Notation "'mexists' x : t , p" := (E (fun x:t => p))
  (at level 200, x ident, right associativity, 
    format "'[' 'mexists' '/ '  x  :  t , '/ '  p ']'")
  : type_scope.
\end{verbatim}

\noindent
The modal operators $\Diamond$ (\emph{possibly}) and $\Box$ (\emph{necessarily}) are defined according to their meanings in the possible world semantics.

\begin{verbatim}
(* Modal operator for 'necessarily' *)
Definition box (p: o) := fun w => forall w1, (r w w1) -> (p w1).

(* Modal operator for 'possibly' *)
Definition dia (p: o) := fun w => exists w1, (r w w1) /\ (p w1).
\end{verbatim}

\noindent
Validity of a modal proposition is validity of the grounded proposition on any world.

\noindent
Although

\begin{verbatim}
Ltac box_intro w H := intros w H.

Ltac box_elim H w1 R1 Hn := 
  let P := match type of H with
             (* forall w0:i, r ?w w0 -> ?p w0 => p *)
             ((box ?p) _) => p
           end
  in cut (P w1); [intros Hn | apply (H w1 R1)].

Ltac dia_elim H w R newH := destruct H as [w [R newH]].

Ltac dia_intro w := (exists w; split; [assumption | idtac]).
\end{verbatim}

\begin{verbatim}
(* Modal validity of lifted propositions *)
Definition V (p: o) := forall w, p w.
\end{verbatim}

\begin{verbatim}
Lemma modus_ponens_inside_dia: V (mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)).
Proof.
intro.
intros p q.
intro H1.
intro H2.
dia_elim H1 w1 R1 H1.
exists w1.
split.
  exact R1.

  apply H2. 
    exact R1.

    exact H1.
Qed.

Lemma not_dia_box_not: V (mforall p, ((m~ (dia p)) m-> (box (m~ p)) )).
Proof.
intro.
intro p.
intro H.
intros w1 H1.
intro H2.
apply H.
exists w1; split. exact H1.
exact H2.
Qed.
\end{verbatim}

\begin{verbatim}
(* Modal accessibility axioms *)

Axiom reflexivity: forall w, r w w.

Axiom transitivity: forall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).

Axiom symmetry: forall w1 w2, (r w1 w2) -> (r w2 w1).
\end{verbatim}

\begin{verbatim}
(* Modal axioms *)
(* Here we show how they can be derived from the accessibility axioms *)

Theorem K: (V (mforall p, mforall q, (box (p m-> q)) m-> ( (box p) m-> (box q) ) )).
Proof.
intros w p q.
intros H1 H2.
intros w1 R1.
apply H1.
  exact R1.

  apply H2.
  exact R1.
Qed.

Theorem T: (V (mforall p, (box p) m-> p)).
Proof.
intros w p.
intro H.
apply H.
apply reflexivity.
Qed.  
\end{verbatim}

\begin{verbatim}
(* In strong modal logics, such as S5, iterations of modal operators can be collapsed *)
Theorem dia_box_to_box: V (mforall p, (dia (box p)) m-> (box p)).
Proof.
intro.
intro p.
intro H1.
destruct H1 as [w1 [R1 H1]].
intro. intro R0.
apply H1.
apply transitivity with (w2 := w).
  apply symmetry.
  exact R1.

  exact R0.
Qed.
\end{verbatim}

\section{G\"odel's Proof of God's Existence}
\label{sec:proof}

Moreover, automated reasoning tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II}, Satallax \cite{Satallax}, Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof} have been used and a few interesting new facts have been discovered about this proof \cite{TodO:ArXiv or Journal abstract}. The formalization in Isabelle \cite{ToDo:ArchiveFormalProofs}.

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable (but debatable) axioms that they should satisfy.
Various slightly different versions of axioms and definitions have
been considered by G\"{o}del and by several philosophers who commented
on his proof
(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).

Dana Scott's version of G\"odel's proof \cite{ScottNotes}. This formalization aims at
being as similar as possible to Dana Scott's version of the proof

The numbering of axioms, definitions and theorems is exactly the same as in Scott's notes

The formal proofs follow the same structure of Scott's proof sketches and fill their gaps

Whenever a `cut' or `assert' uses a lemma mentioned in Scott's sketches,
this is emphasized with a comment

\begin{verbatim}
Require Import Coq.Logic.Classical.
Require Import Coq.Logic.Classical_Pred_Type.

Require Import Modal.

Ltac proof_by_contradiction H := apply NNPP; intro H.

(* Constant predicate that distinguishes positive properties *)
Parameter Positive: (u -> o) -> o.

(* Axiom A1: either a property or its negation is positive, but not both *)
Axiom axiom1a : V (mforall p, (Positive (fun x: u => m~(p x))) m-> (m~ (Positive p))).
Axiom axiom1b : V (mforall p, (m~ (Positive p)) m-> (Positive (fun x: u => m~ (p x))) ).

(* Axiom A2: a property necessarily implied by a positive property is positive *)
Axiom axiom2: V (mforall p, mforall q, Positive p m/\ (box (mforall x, (p x) m-> (q x) )) m-> Positive q).

(* Theorem T1: positive properties are possibly exemplified *)
Theorem theorem1: V (mforall p, (Positive p) m-> dia (mexists x, p x) ).
Proof.
intro.
intro p.
intro H1.
proof_by_contradiction H2.
apply not_dia_box_not in H2.
assert (H3: ((box (mforall x, m~ (p x))) w)). (* Lemma from Scott's notes *)
  box_intro w1 R1.
  intro x.
  assert (H4: ((m~ (mexists x : u, p x)) w1)).
    box_elim H2 w1 R1 G2. 
    exact G2.

    clear H2 R1 H1 w.
    intro H5.
    apply H4.
    exists x.
    exact H5.

  assert (H6: ((box (mforall x, (p x) m-> m~ (x m= x))) w)). (* Lemma from Scott's notes *)    
    box_intro w1 R1.
    intro x.
    intro H7.
    intro H8.
    box_elim H3 w1 R1 G3.
    apply G3 with (x := x).
    exact H7.

    assert (H9: ((Positive (fun x => m~ (x m= x))) w)). (* Lemma from Scott's notes *)
    apply (axiom2 w p (fun x => m~ (x m= x))).
    split.
      exact H1.

      exact H6.
    assert (H10: ((box (mforall x, (p x) m-> (x m= x))) w)). (* Lemma from Scott's notes *)
      box_intro w1 R1.
      intros x H11.     
      reflexivity.

      assert (H11 : ((Positive (fun x => (x m= x))) w)). (* Lemma from Scott's notes *)
        apply (axiom2 w p (fun x => x m= x )).
        split.
          exact H1.

          exact H10.

        clear H1 H2 H3 H6 H10 p.
        apply axiom1a in H9.
        contradiction.
Qed.

(* Definition D1: God: a God-like being possesses all positive properties *)
Definition G(x: u) := mforall p, (Positive p) m-> (p x).

(* Axiom A3: the property of being God-like is positive *)
Axiom axiom3: V (Positive G).

(* Corollary C1: possibly, God exists *)
Theorem corollary1: V (dia (mexists x, G x)). 
Proof.
intro. 
apply theorem1.
apply axiom3.
Qed.

(* Axiom A4: positive properties are necessarily positive *)
Axiom axiom4: V (mforall p, (Positive p) m-> box (Positive p)).

(* Definition D2: essence: an essence of an individual is a property possessed by it and necessarily implying any of its properties *)
Definition Essence(p: u -> o)(x: u) := (p x) m/\ mforall q, ((q x) m-> box (mforall y, (p y) m-> (q y))).
Notation "p 'ess' x" := (Essence p x) (at level 69).

(* Theorem T2: being God-like is an essence of any God-like being *)
Theorem theorem2: V (mforall x, (G x) m-> (G ess x)).
Proof.
intro.
intro g.
intro H1.
unfold Essence.
split.
  exact H1.

  intro q.
  intro H2.
  assert (H3: ((Positive q) w)).
    proof_by_contradiction H4.
    unfold G in H1.
    apply axiom1b in H4.
    apply H1 in H4.
    contradiction.

    cut (box (Positive q) w). (* Lemma from Scott's notes *)
      apply K.
      box_intro w1 R1.
      intro H5.
      intro y.
      intro H6.
      unfold G in H6.
      apply (H6 q).
      exact H5.

      apply axiom4.
      exact H3.
Qed.

(* At this point in Scott's notes there are two notes that are not necessary for the proof, *)
(* but it would be interesting to formalize them anyway *)

(* Definition D3: necessary existence: necessary existence of an individual is the necessary exemplification of all its essences *)
Definition NE(x: u) := mforall p, (p ess x) m-> box (mexists y, (p y)).

(* Axiom A5: necessary existence is a positive property *)
Axiom axiom5: V (Positive NE).

Lemma lemma1: V ((mexists z, (G z)) m-> box (mexists x, (G x))).
Proof.
intro.
intro H1.
destruct H1 as [g H2].
cut ((G ess g) w).      (* Lemma from Scott's notes *)
  assert (H3: (NE g w)).       (* Lemma from Scott's notes *)
    unfold G in H2.
    apply (H2 NE).
    apply axiom5.

    unfold NE in H3.
    apply H3.

  apply theorem2.
  exact H2.
Qed.

Lemma lemma2: V (dia (mexists z, (G z)) m-> box (mexists x, (G x))).
Proof.
intro.
intro H.
cut (dia (box (mexists x, G x)) w).  (* Lemma from Scott's notes *)
  apply dia_box_to_box.

  apply (modus_ponens_inside_dia w (mexists z, G z)).
    exact H.

    box_intro w1 R1.
    apply lemma1.
Qed.

(* Theorem T3: necessarily, a God exists *)
Theorem theorem3: V (box (mexists x, (G x))).
Proof.
intro.
apply lemma2.
apply corollary1.
Qed.

(* Corollary C2: There exists a god *)
Theorem corollary2: V (mexists x, (G x)).
Proof.
intro.
apply T.
apply theorem3.
Qed.
\end{verbatim}

\section{Conclusions}
\label{sec:conclusions}

cite related work about modal logics in Coq.

no need to bother about the embedding.

This work attests the maturity of contemporary interactive and
automated deduction tools for classical higher-order logic and
demonstrates the elegance and practical relevance of the
embeddings-based approach.  Most importantly, our work opens new
perspectives for a computer-assisted theoretical philosophy.  The
critical discussion of the underlying concepts, definitions and axioms
remains a human responsibility, but the computer can assist in
building and checking rigorously correct logical arguments. In case of
logico-philosophical disputes, the computer can check the disputing
arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
calculate!
The philosophical and theological importance of the correctness of this proof is beyond the scope of this paper.

We welcome feedback on

\paragraph{Acknowledgements:} ToDo: thank Coq-Club members.

%\footnotesize
\bibliographystyle{plain}
%\bibliography{Bibliography}
\begin{thebibliography}{10}

\bibitem{Adams}
R.M. Adams.
\newblock Introductory note to *1970.
\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
  Letters}}. Oxford University Press, 1995.

\bibitem{AndersonGettings}
A.C. Anderson and M.~Gettings.
\newblock G\"odel ontological proof revisited.
\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.

\bibitem{B9}
C.~Benzm{\"u}ller and L.C. Paulson.
\newblock Exploring properties of normal multimodal logics in simple type
  theory with {LEO-II}.
\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
  Birthday}}, pages 386--406. College Publications.

\bibitem{J23}
C.~Benzm{\"u}ller and L.C. Paulson.
\newblock Quantified multimodal logics in simple type theory.
\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
  7(1):7--20, 2013.

\bibitem{LEO-II}
C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
  logic.
\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
  162--170. Springer, 2008.

\bibitem{Coq}
Y.~Bertot and P.~Casteran.
\newblock {\em {Interactive Theorem Proving and Program Development}}.
\newblock Springer, 2004.

\bibitem{Sledgehammer}
J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
\newblock Extending {Sledgehammer} with {SMT} solvers.
\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.

\bibitem{Nitpick}
J.C. Blanchette and T.~Nipkow.
\newblock Nitpick: A counterexample generator for higher-order logic based on a
  relational model finder.
\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
  Springer, 2010.

\bibitem{Satallax}
C.E. Brown.
\newblock Satallax: An automated higher-order prover.
\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
  Springer, 2012.

\bibitem{ContemporaryBibliography}
R.~Corazzon.
\newblock Contemporary bibligraphy on the ontological proof
  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).

\bibitem{Fitting}
M.~Fitting.
\newblock {\em {Types, Tableaux and G\"odel's God}}.
\newblock Kluver Academic Press, 2002.

\bibitem{Goedel1970}
K.~G\"odel.
\newblock Ontological proof.
\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
  Letters}}. Oxford University Press, 1970.

\bibitem{GoedelNotes}
K.~G\"odel.
\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{Hazen}
A.P. Hazen.
\newblock On g\"odel's ontological proof.
\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.

\bibitem{Hurd03first-orderproof}
J.~Hurd.
\newblock First-order proof tactics in higher-order logic theorem provers.
\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel.
\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
\newblock Number 2283 in LNCS. Springer, 2002.

\bibitem{FormalTheologyRepository}
B.~Woltzenlogel Paleo and C.~Benzm\"uller.
\newblock Formal theology repository
  (\url{https://github.com/FormalTheology/GoedelGod}).

\bibitem{ScottNotes}
D.~Scott.
\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{sobel2004logic}
J.H. Sobel.
\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
\newblock Cambridge U. Press, 2004.

\bibitem{J22}
G.~Sutcliffe and C.~Benzm{\"u}ller.
\newblock Automated reasoning in higher-order logic using the {TPTP THF}
  infrastructure.
\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.

\end{thebibliography}

\end{document}

### 2014-01-23 Papers/ITP/GodProof-ITP.tex
Update paper

@@
 \begin{verbatim}
+Definition mequal (x y: u) := fun w: i => x = y.
+Notation "x m= y" := (mequal x y) (at level 99, right associativity).
+
 Definition mnot (p: o)(w: i) := ~ (p w).
@@
 \noindent
-Validity of a modal proposition is validity of the grounded proposition on any world.
-
-\noindent
-Although
+Although the embedding via the possible world semantics is necessary, it should be as transparent as possible to the user.
+Fortunately, basic Coq tactics such \texttt{intro}, \texttt{apply} and \texttt{split} automatically unfold the modal definitions and thus can be used with modal connectives and quantifiers as they are used with the usual connectives and quantifiers.
+Nothing else needs to be done.
+For the modal operators, however, the following simple tactics have been implemented, in order to allow the user to work with the concepts of \emph{necessity} and \emph{possibility} without having to unfold the definitions of modal operators and think in terms of possible worlds.
+These tactics remind rules of labelled calculi for modal logics.
 
@@
 
+\noindent
+Validity of a modal proposition is validity of its grounding on any world.
+
 \begin{verbatim}
@@
 \end{verbatim}
+
+\noindent
+The following lemmas are convenient
 
@@
 
+Note: possible worlds machinery is hidden from the user.
+
 \section{Conclusions}

### 2014-01-25 Papers/ITP/GodProof-ITP.tex
Start using fancy verbatim

@@
 
-\usepackage{url,amsmath,amssymb,a4wide}
+\usepackage{url,amsmath,amssymb,a4wide,fancyvrb,color}
 
 \newcommand{\imp}{\rightarrow}
-\newcommand{\biimp}{\leftrightarrow}
-\newcommand{\all}{\forall}
-\newcommand{\ex}{\exists}
-\newcommand{\seq}{\vdash}
-\newcommand{\nec}{\Box} % necessarily
-\newcommand{\pos}{\Diamond} % possibly
-\newcommand{\ess}[2]{#1 \ \mathit{ess.} \ #2}
-\newcommand{\NE}{\mathit{NE}}
+
+\newcommand{\red}[1]{\textcolor[rgb]{1,0,0}{#1}}
+\newcommand{\blue}[1]{\textcolor[rgb]{0,0,1}{#1}}
+\newcommand{\brown}[1]{\textcolor[rgb]{0.8,0.6,0.4}{#1}}
+
+\newcommand{\Parameter}{\red{Parameter}}
+\newcommand{\Lemma}{\red{Lemma}}
+\newcommand{\Theorem}{\red{Theorem}}
+\newcommand{\Definition}{\red{Definition}}
+\newcommand{\Notation}{\blue{Notation}}
+\newcommand{\Prop}{\blue{Prop}}
+\newcommand{\Type}{\blue{Type}}
+\newcommand{\iin}{\blue{in}}
+\newcommand{\as}{\blue{as}}
+\newcommand{\fforall}{\blue{forall}}
+\newcommand{\fun}{\blue{fun}}
+\newcommand{\Proof}{\blue{Proof}}
+\newcommand{\Qed}{\blue{Qed}}
+\newcommand{\Hint}{\blue{Hint}}
+\newcommand{\com}[1]{\brown{#1}}
 
@@
 higher-order quantifiers. 
-Therefore, it can serve as a good benchmark to measure and compare progress in the development of automated and interactive reasonging tools for higher-order modal logics.
-To fulfill this goal, formalizations \cite{FormalTheologyRepository} in TPTP THF \cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been completed so far, in this order.
+Specialized modal logic tools lack higher-order capabilities, 
+while higher-order provers lack built-in support for modalities. 
+However, by embedding modal logics into higher-order logic, 
+it is in principle possible to use standard 
+higher-order automated and interactive theorem provers 
+for reasoning about and within modal logics \cite{J23,B9}.
+To realize this potential, formalizations \cite{FormalTheologyRepository} of G\"odel's proof in TPTP THF \cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been completed so far, in this order.
 This paper focuses on the formalization in Coq.
-
-An embedding of modal logics 
-into higher-order logic with
-Henkin semantics \cite{J23,B9} provides the necessary theoretical foundation that enables the use of interactive theorem provers for modal logics as well.
-In the particular case of Coq, the implementation of this embedding is discussed in detail in Section \ref{sec:modal} and Scott's version of G\"odel's proof is shown in Section \ref{sec:proof}.
+The implementation of the embedding is discussed in detail in Section \ref{sec:modal} and Scott's version of G\"odel's proof is shown in Section \ref{sec:proof}.
 
@@
 
-\begin{verbatim}
-(* Type for worlds *)
-Parameter i: Type.
-
-(* Type for individuals *)
-Parameter u: Type.
-
-(* Type of modal propositions *)
-Definition o := i -> Prop.
-
-(* Acessibility relation for worlds *)
-Parameter r: i -> i -> Prop.
-\end{verbatim}
+\begin{Verbatim}[commandchars=\\\{\}]
+\Parameter i: \Type. \com{(* Type for worlds *)}
+\Parameter u: \Type. \com{(* Type for individuals *)}
+\Definition o := i -> \Prop. \com{(* Type of modal propositions *)}
+\Parameter r: i -> i -> \Prop. \com{(* Acessibility relation for worlds *)}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
-Definition mequal (x y: u) := fun w: i => x = y.
-Notation "x m= y" := (mequal x y) (at level 99, right associativity).
+\begin{Verbatim}[commandchars=\\\{\}]
+\Definition mequal (x y: u) := \fun w: i => x = y.
+\Notation "x m= y" := (mequal x y) (at level 99, right associativity).
 
-Definition mnot (p: o)(w: i) := ~ (p w).
-Notation "m~  p" := (mnot p) (at level 74, right associativity).
+\Definition mnot (p: o)(w: i) := ~ (p w).
+\Notation "m~  p" := (mnot p) (at level 74, right associativity).
 
-Definition mand (p q:o)(w: i) := (p w) /\ (q w).
-Notation "p m/\ q" := (mand p q) (at level 79, right associativity).
+\Definition mand (p q:o)(w: i) := (p w) /\ (q w).
+\Notation "p m/\ q" := (mand p q) (at level 79, right associativity).
 
-Definition mor (p q:o)(w: i) := (p w) \/ (q w).
-Notation "p m\/ q" := (mor p q) (at level 79, right associativity).
+\Definition mor (p q:o)(w: i) := (p w) \/ (q w).
+\Notation "p m\/ q" := (mor p q) (at level 79, right associativity).
 
-Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
-Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).
+\Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
+\Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).
 
-Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
-Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
-\end{verbatim}
+\Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
+\Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
-Definition A {t: Type}(p: t -> o) := fun w => forall x, p x w.
-Notation "'mforall'  x , p" := (A (fun x => p))
+\begin{Verbatim}[commandchars=\\\{\}]
+\Definition A {t: Type}(p: t -> o) := \fun w => \fforall x, p x w.
+\Notation "'mforall'  x , p" := (A (fun x => p))
   (at level 200, x ident, right associativity) : type_scope.
-Notation "'mforall' x : t , p" := (A (fun x:t => p))
+\Notation "'mforall' x : t , p" := (A (\fun x:t => p))
   (at level 200, x ident, right associativity, 
@@
 
-Definition E {t: Type}(p: t -> o) := fun w => exists x, p x w.
-Notation "'mexists' x , p" := (E (fun x => p))
+\Definition E {t: Type}(p: t -> o) := \fun w => exists x, p x w.
+\Notation "'mexists' x , p" := (E (\fun x => p))
   (at level 200, x ident, right associativity) : type_scope.
-Notation "'mexists' x : t , p" := (E (fun x:t => p))
+\Notation "'mexists' x : t , p" := (E (\fun x:t => p))
   (at level 200, x ident, right associativity, 
@@
   : type_scope.
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{Verbatim}[commandchars=\\\{\}]
 (* Modal operator for 'necessarily' *)
@@
 Definition dia (p: o) := fun w => exists w1, (r w w1) /\ (p w1).
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-\paragraph{Acknowledgements:} ToDo: thank Coq-Club members.
+\paragraph{Acknowledgements:} ToDo: thank Coq-Club members. Cedric Auger
 

### 2014-01-25 Papers/ITP/GodProof-ITP.tex
Write conclusion

@@
 \newcommand{\Parameter}{\red{Parameter}}
+\newcommand{\Ltac}{\red{Ltac}}
+\newcommand{\Axiom}{\red{Axiom}}
 \newcommand{\Lemma}{\red{Lemma}}
@@
 \newcommand{\Type}{\blue{Type}}
+\newcommand{\llet}{\blue{let}}
+\newcommand{\match}{\blue{match}}
+\newcommand{\with}{\blue{with}}
+\newcommand{\eend}{\blue{end}}
 \newcommand{\iin}{\blue{in}}
 \newcommand{\as}{\blue{as}}
+\newcommand{\Require}{\blue{Require}}
+\newcommand{\Import}{\blue{Import}}
 \newcommand{\fforall}{\blue{forall}}
@@
 \newcommand{\com}[1]{\brown{#1}}
+\newcommand{\bslash}{\symbol{92}}
 
@@
 
-\begin{Verbatim}[commandchars=\\\{\}]
+\newcommand{\verbsize}{\small}
+
+\begin{small}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Parameter i: \Type. \com{(* Type for worlds *)}
@@
 \end{Verbatim}
+\end{small}
 
@@
 
-\begin{Verbatim}[commandchars=\\\{\}]
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Definition mequal (x y: u) := \fun w: i => x = y.
@@
 
-\Definition mand (p q:o)(w: i) := (p w) /\ (q w).
-\Notation "p m/\ q" := (mand p q) (at level 79, right associativity).
+\Definition mand (p q:o)(w: i) := (p w) /\bslash (q w).
+\Notation "p m/\bslash q" := (mand p q) (at level 79, right associativity).
 
-\Definition mor (p q:o)(w: i) := (p w) \/ (q w).
-\Notation "p m\/ q" := (mor p q) (at level 79, right associativity).
+\Definition mor (p q:o)(w: i) := (p w) \bslash\slash (q w).
+\Notation "p m\bslash\slash q" := (mor p q) (at level 79, right associativity).
 
@@
 
-\begin{Verbatim}[commandchars=\\\{\}]
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Definition A {t: Type}(p: t -> o) := \fun w => \fforall x, p x w.
@@
 \noindent
-The modal operators $\Diamond$ (\emph{possibly}) and $\Box$ (\emph{necessarily}) are defined according to their meanings in the possible world semantics.
+The modal operators $\Diamond$ (\emph{possibly}) and $\Box$ (\emph{necessarily}) are defined accordingly to their meanings in the possible world semantics.
 
-\begin{Verbatim}[commandchars=\\\{\}]
-(* Modal operator for 'necessarily' *)
-Definition box (p: o) := fun w => forall w1, (r w w1) -> (p w1).
-
-(* Modal operator for 'possibly' *)
-Definition dia (p: o) := fun w => exists w1, (r w w1) /\ (p w1).
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Definition box (p: o) := \fun w => \fforall w1, (r w w1) -> (p w1).
+\Definition dia (p: o) := \fun w => exists w1, (r w w1) /\bslash (p w1).
 \end{Verbatim}
@@
 For the modal operators, however, the following simple tactics have been implemented, in order to allow the user to work with the concepts of \emph{necessity} and \emph{possibility} without having to unfold the definitions of modal operators and think in terms of possible worlds.
-These tactics remind rules of labelled calculi for modal logics.
 
-\begin{verbatim}
-Ltac box_intro w H := intros w H.
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac box_i := \llet w := fresh "w" \iin \llet R := fresh "R" 
+               \iin (intro w at top; intro R at top).
 
-Ltac box_elim H w1 R1 Hn := 
-  let P := match type of H with
-             (* forall w0:i, r ?w w0 -> ?p w0 => p *)
-             ((box ?p) _) => p
-           end
-  in cut (P w1); [intros Hn | apply (H w1 R1)].
+\Ltac box_elim H w1 H1 := \match type of H \with 
+  ((box ?p) ?w) => cut (p w1); [intros H1 | (apply (H w1); try assumption)] \eend.
 
-Ltac dia_elim H w R newH := destruct H as [w [R newH]].
+\Ltac box_e H H1:= \match goal \with | [ |- (_ ?w) ] => box_elim H w H1 \eend.
 
-Ltac dia_intro w := (exists w; split; [assumption | idtac]).
-\end{verbatim}
+\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" 
+                 \iin (destruct H \as [w [R H]]; move w at top; move R at top).
+
+\Ltac dia_i w := (exists w; split; [assumption | idtac]).
+\end{Verbatim}
 
 \noindent
-Validity of a modal proposition is validity of its grounding on any world.
+Validity of a modal proposition is validity of its grounding on any world. A non-intrusive notation and a tactic that grounds the proposition to an arbitrary world are provided.
 
-\begin{verbatim}
-(* Modal validity of lifted propositions *)
-Definition V (p: o) := forall w, p w.
-\end{verbatim}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Definition V (p: o) := \fforall w, p w.
+\Notation "[ p ]" := (V p).
+\Ltac mv := \match goal \with [|- (V _)] => intro \eend.
+\end{Verbatim}
 
 \noindent
-The following lemmas are convenient
+The following lemmas are quite convenient. The first is a kind of modus ponens applicable to formulas under the scope of a modal operator. The second allows pushing negations inside modalities.
 
-\begin{verbatim}
-Lemma modus_ponens_inside_dia: V (mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)).
-Proof.
-intro.
-intros p q.
-intro H1.
-intro H2.
-dia_elim H1 w1 R1 H1.
-exists w1.
-split.
-  exact R1.
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Lemma mp_dia: [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
+\Proof. mv. intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. \Qed.
 
-  apply H2. 
-    exact R1.
+\Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
+\Proof. mv. intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. \Qed.
+\end{Verbatim}
 
-    exact H1.
-Qed.
+\noindent
+The most well-known modal logics can be distinguished by their acceptance of the following reachability axioms. \textbf{S5} is the modal logic that accepts all three.
 
-Lemma not_dia_box_not: V (mforall p, ((m~ (dia p)) m-> (box (m~ p)) )).
-Proof.
-intro.
-intro p.
-intro H.
-intros w1 H1.
-intro H2.
-apply H.
-exists w1; split. exact H1.
-exact H2.
-Qed.
-\end{verbatim}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Axiom reflexivity: \fforall w, r w w.
+\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
+\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
+\end{Verbatim}
 
-\begin{verbatim}
-(* Modal accessibility axioms *)
+\noindent
+Modal logic ``axioms'' can be derived from the reachability axioms.
 
-Axiom reflexivity: forall w, r w w.
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Theorem K: [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
+\Proof. mv. intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. \Qed.
 
-Axiom transitivity: forall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
+\Theorem T: [ mforall p, (box p) m-> p ].
+\Proof. mv. intro p. intro H. box_e H H1. exact H1. apply reflexivity. \Qed.
+\end{Verbatim}
 
-Axiom symmetry: forall w1 w2, (r w1 w2) -> (r w2 w1).
-\end{verbatim}
+\noindent
+In strong modal logics, such as S5, iterations of modal operators can be collapsed. This is a controversial principle used in G\"odel's and Scott's versions of the proof.
 
-\begin{verbatim}
-(* Modal axioms *)
-(* Here we show how they can be derived from the accessibility axioms *)
-
-Theorem K: (V (mforall p, mforall q, (box (p m-> q)) m-> ( (box p) m-> (box q) ) )).
-Proof.
-intros w p q.
-intros H1 H2.
-intros w1 R1.
-apply H1.
-  exact R1.
-
-  apply H2.
-  exact R1.
-Qed.
-
-Theorem T: (V (mforall p, (box p) m-> p)).
-Proof.
-intros w p.
-intro H.
-apply H.
-apply reflexivity.
-Qed.  
-\end{verbatim}
-
-\begin{verbatim}
-(* In strong modal logics, such as S5, iterations of modal operators can be collapsed *)
-Theorem dia_box_to_box: V (mforall p, (dia (box p)) m-> (box p)).
-Proof.
-intro.
-intro p.
-intro H1.
-destruct H1 as [w1 [R1 H1]].
-intro. intro R0.
-apply H1.
-apply transitivity with (w2 := w).
-  apply symmetry.
-  exact R1.
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Theorem dia_box_to_box: [ mforall p, (dia (box p)) m-> (box p) ].
+\Proof. mv. intro p. intro H1. dia_e H1. box_i. box_e H1 H2. exact H2. eapply transitivity.
+  apply symmetry. exact R.
 
   exact R0.
-Qed.
-\end{verbatim}
+\Qed.
+\end{Verbatim}
 
@@
 
-Moreover, automated reasoning tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II}, Satallax \cite{Satallax}, Sledgehammer \cite{Sledgehammer} and Metis \cite{Hurd03first-orderproof} have been used and a few interesting new facts have been discovered about this proof \cite{TodO:ArXiv or Journal abstract}.
-The formalization in Isabelle \cite{ToDo:ArchiveFormalProofs}.
-
 Attempts to prove the existence (or non-existence) of God by means of
 abstract ontological arguments are an old tradition in philosophy and
-theology.  G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
+theology.  
+G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
 this tradition, following particularly the footsteps of Leibniz.
-%
-G\"{o}del defines God as a being who possesses all \emph{positive} properties.
-He does not extensively discuss what positive properties are, 
-but instead he states a few reasonable (but debatable) axioms that they should satisfy.
 Various slightly different versions of axioms and definitions have
@@
 on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). The formalization shown in this section aims at
+being as similar as possible to Dana Scott's version of the proof \cite{ScottNotes}.
+The formulation and numbering of axioms, definitions and theorems is the same as in Scott's notes.
+Even the Coq proof scripts follow all the steps in Scott's notes.
+Scott's assertions are emphasized with comments.
+In contrast to the formalization in Isabelle \cite{ToDo:our isabelle formalization in ArchiveFormalProofs}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
+This was a deliberate choice, mainly because it allowed a qualitative evaluation of the convenience of the possible world semantic embedding approach for \emph{interactive} theorem proving.
+Moreover, in order to formalize precisely Scott's version and not another automatically found version, automation would have to be heavily limited anyway.
+Furthermore, the deliberate preference for simple tactics (mostly \emph{intro}, \emph{apply} and the modal tactics described in the previous section) results in proof scripts that could be read as natural deduction proofs.
+This hopefully makes the formalization more accessible to those who are not experts in Coq's tactics but are nevertheless interested in G\"odel's proof.
 
-Dana Scott's version of G\"odel's proof \cite{ScottNotes}. This formalization aims at
-being as similar as possible to Dana Scott's version of the proof
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Require \Import Coq.Logic.Classical Coq.Logic.Classical_Pred_Type Modal.
 
-The numbering of axioms, definitions and theorems is exactly the same as in Scott's notes
+\Ltac proof_by_contradiction H := apply NNPP; intro H.
 
-The formal proofs follow the same structure of Scott's proof sketches and fill their gaps
+\com{(* Constant predicate that distinguishes positive properties *)}
+\Parameter Positive: (u -> o) -> o.
 
-Whenever a `cut' or `assert' uses a lemma mentioned in Scott's sketches,
-this is emphasized with a comment
+\com{(* Axiom A1: either a property or its negation is positive, but not both *)}
+\Axiom axiom1a : [ mforall p, (Positive (\fun x: u => m~(p x))) m-> (m~ (Positive p)) ].
+\Axiom axiom1b : [ mforall p, (m~ (Positive p)) m-> (Positive (\fun x: u => m~ (p x))) ].
 
-\begin{verbatim}
-Require Import Coq.Logic.Classical.
-Require Import Coq.Logic.Classical_Pred_Type.
+\com{(* Axiom A2: a property necessarily implied by a positive property is positive *)}
+\Axiom axiom2: [ mforall p, mforall q, 
+  Positive p m/\bslash (box (mforall x, (p x) m-> (q x) )) m-> Positive q ].
 
-Require Import Modal.
+\com{(* Theorem T1: positive properties are possibly exemplified *)}
+\Theorem theorem1: [ mforall p, (Positive p) m-> dia (mexists x, p x) ].
+\Proof. mv.
+intro p. intro H1. proof_by_contradiction H2. apply not_dia_box_not in H2.
+assert (H3: ((box (mforall x, m~ (p x))) w)). \com{(* Scott *)}
+  box_i. intro x. assert (H4: ((m~ (mexists x : u, p x)) w0)).
+    box_e H2 G2. exact G2.
 
-Ltac proof_by_contradiction H := apply NNPP; intro H.
+    clear H2 R H1 w. intro H5. apply H4. exists x. exact H5.
 
-(* Constant predicate that distinguishes positive properties *)
-Parameter Positive: (u -> o) -> o.
+  assert (H6: ((box (mforall x, (p x) m-> m~ (x m= x))) w)). \com{(* Scott *)}   
+    box_i. intro x. intros H7 H8. box_elim H3 w0 G3. eapply G3. exact H7.
 
-(* Axiom A1: either a property or its negation is positive, but not both *)
-Axiom axiom1a : V (mforall p, (Positive (fun x: u => m~(p x))) m-> (m~ (Positive p))).
-Axiom axiom1b : V (mforall p, (m~ (Positive p)) m-> (Positive (fun x: u => m~ (p x))) ).
-
-(* Axiom A2: a property necessarily implied by a positive property is positive *)
-Axiom axiom2: V (mforall p, mforall q, Positive p m/\ (box (mforall x, (p x) m-> (q x) )) m-> Positive q).
-
-(* Theorem T1: positive properties are possibly exemplified *)
-Theorem theorem1: V (mforall p, (Positive p) m-> dia (mexists x, p x) ).
-Proof.
-intro.
-intro p.
-intro H1.
-proof_by_contradiction H2.
-apply not_dia_box_not in H2.
-assert (H3: ((box (mforall x, m~ (p x))) w)). (* Lemma from Scott's notes *)
-  box_intro w1 R1.
-  intro x.
-  assert (H4: ((m~ (mexists x : u, p x)) w1)).
-    box_elim H2 w1 R1 G2. 
-    exact G2.
-
-    clear H2 R1 H1 w.
-    intro H5.
-    apply H4.
-    exists x.
-    exact H5.
-
-  assert (H6: ((box (mforall x, (p x) m-> m~ (x m= x))) w)). (* Lemma from Scott's notes *)    
-    box_intro w1 R1.
-    intro x.
-    intro H7.
-    intro H8.
-    box_elim H3 w1 R1 G3.
-    apply G3 with (x := x).
-    exact H7.
-
-    assert (H9: ((Positive (fun x => m~ (x m= x))) w)). (* Lemma from Scott's notes *)
-    apply (axiom2 w p (fun x => m~ (x m= x))).
-    split.
+    assert (H9: ((Positive (fun x => m~ (x m= x))) w)). \com{(* Scott *)}
+      apply (axiom2 w p (fun x => m~ (x m= x))). split.
         exact H1.
@@
         exact H6.
-    assert (H10: ((box (mforall x, (p x) m-> (x m= x))) w)). (* Lemma from Scott's notes *)
-      box_intro w1 R1.
-      intros x H11.     
-      reflexivity.
+      assert (H10: ((box (mforall x, (p x) m-> (x m= x))) w)). \com{(* Scott *)}
+        box_i. intros x H11. reflexivity.
 
-      assert (H11 : ((Positive (fun x => (x m= x))) w)). (* Lemma from Scott's notes *)
-        apply (axiom2 w p (fun x => x m= x )).
-        split.
+        assert (H11 : ((Positive (fun x => (x m= x))) w)). \com{(* Scott *)}
+          apply (axiom2 w p (fun x => x m= x )). split.
             exact H1.
@@
 
-        clear H1 H2 H3 H6 H10 p.
-        apply axiom1a in H9.
-        contradiction.
-Qed.
+          apply axiom1a in H9. contradiction.
+\Qed.
 
-(* Definition D1: God: a God-like being possesses all positive properties *)
-Definition G(x: u) := mforall p, (Positive p) m-> (p x).
+\com{(* Definition D1: God: a God-like being possesses all positive properties *)}
+\Definition G(x: u) := mforall p, (Positive p) m-> (p x).
 
-(* Axiom A3: the property of being God-like is positive *)
-Axiom axiom3: V (Positive G).
+\com{(* Axiom A3: the property of being God-like is positive *)}
+\Axiom axiom3: [ Positive G ].
 
-(* Corollary C1: possibly, God exists *)
-Theorem corollary1: V (dia (mexists x, G x)). 
-Proof.
-intro. 
-apply theorem1.
-apply axiom3.
-Qed.
+\com{(* Corollary C1: possibly, God exists *)}
+\Theorem corollary1: [ dia (mexists x, G x) ]. 
+\Proof. mv. apply theorem1. apply axiom3. \Qed.
 
-(* Axiom A4: positive properties are necessarily positive *)
-Axiom axiom4: V (mforall p, (Positive p) m-> box (Positive p)).
+\com{(* Axiom A4: positive properties are necessarily positive *)}
+\Axiom axiom4: [ mforall p, (Positive p) m-> box (Positive p) ].
 
-(* Definition D2: essence: an essence of an individual is a property possessed by it and necessarily implying any of its properties *)
-Definition Essence(p: u -> o)(x: u) := (p x) m/\ mforall q, ((q x) m-> box (mforall y, (p y) m-> (q y))).
-Notation "p 'ess' x" := (Essence p x) (at level 69).
+\com{(* Definition D2: essence: an essence of an individual is a property possessed by it} 
+\com{and necessarily implying any of its properties *)}
+\Definition Essence(p: u -> o)(x: u) := (p x) m/\bslash mforall q, 
+                                        ((q x) m-> box (mforall y, (p y) m-> (q y))).
+\Notation "p 'ess' x" := (Essence p x) (at level 69).
 
-(* Theorem T2: being God-like is an essence of any God-like being *)
-Theorem theorem2: V (mforall x, (G x) m-> (G ess x)).
-Proof.
-intro.
-intro g.
-intro H1.
-unfold Essence.
-split.
+\com{(* Theorem T2: being God-like is an essence of any God-like being *)}
+\Theorem theorem2: [ mforall x, (G x) m-> (G ess x) ].
+\Proof. mv. intro g. intro H1. unfold Essence. split.
   exact H1.
 
-  intro q.
-  intro H2.
-  assert (H3: ((Positive q) w)).
-    proof_by_contradiction H4.
-    unfold G in H1.
-    apply axiom1b in H4.
-    apply H1 in H4.
+  intro q. intro H2. assert (H3: ((Positive q) w)).
+    proof_by_contradiction H4. unfold G in H1. apply axiom1b in H4. apply H1 in H4. 
     contradiction.
 
-    cut (box (Positive q) w). (* Lemma from Scott's notes *)
-      apply K.
-      box_intro w1 R1.
-      intro H5.
-      intro y.
-      intro H6.
-      unfold G in H6.
-      apply (H6 q).
-      exact H5.
+    cut (box (Positive q) w). \com{(* Scott *)}
+      apply K. box_i. intro H5. intro y. intro H6. unfold G in H6. apply (H6 q). exact H5.
 
-      apply axiom4.
-      exact H3.
-Qed.
+      apply axiom4. exact H3.
+\Qed.
 
-(* At this point in Scott's notes there are two notes that are not necessary for the proof, *)
-(* but it would be interesting to formalize them anyway *)
+\com{(* Definition D3: necessary existence: necessary existence of an individual }
+\com{is the necessary exemplification of all its essences *)}
+\Definition NE(x: u) := mforall p, (p ess x) m-> box (mexists y, (p y)).
 
-(* Definition D3: necessary existence: necessary existence of an individual is the necessary exemplification of all its essences *)
-Definition NE(x: u) := mforall p, (p ess x) m-> box (mexists y, (p y)).
+\com{(* Axiom A5: necessary existence is a positive property *)}
+\Axiom axiom5: [ Positive NE ].
 
-(* Axiom A5: necessary existence is a positive property *)
-Axiom axiom5: V (Positive NE).
+\Lemma lemma1: [ (mexists z, (G z)) m-> box (mexists x, (G x)) ].
+\Proof. mv. intro H1. destruct H1 as [g H2]. cut ((G ess g) w). \com{(* Scott *)}
+  assert (H3: (NE g w)).       \com{(* Scott *)}
+    unfold G in H2. apply (H2 NE). apply axiom5.
 
-Lemma lemma1: V ((mexists z, (G z)) m-> box (mexists x, (G x))).
-Proof.
-intro.
-intro H1.
-destruct H1 as [g H2].
-cut ((G ess g) w).      (* Lemma from Scott's notes *)
-  assert (H3: (NE g w)).       (* Lemma from Scott's notes *)
-    unfold G in H2.
-    apply (H2 NE).
-    apply axiom5.
+    unfold NE in H3. apply H3.
 
-    unfold NE in H3.
-    apply H3.
+  apply theorem2. exact H2.
+\Qed.
 
-  apply theorem2.
-  exact H2.
-Qed.
-
-Lemma lemma2: V (dia (mexists z, (G z)) m-> box (mexists x, (G x))).
-Proof.
-intro.
-intro H.
-cut (dia (box (mexists x, G x)) w).  (* Lemma from Scott's notes *)
+\Lemma lemma2: [ dia (mexists z, (G z)) m-> box (mexists x, (G x)) ].
+\Proof. mv. intro H. cut (dia (box (mexists x, G x)) w).  \com{(* Scott *)}
   apply dia_box_to_box.
 
-  apply (modus_ponens_inside_dia w (mexists z, G z)).
+  apply (mp_dia w (mexists z, G z)).
     exact H.
 
-    box_intro w1 R1.
-    apply lemma1.
-Qed.
+    box_i. apply lemma1.
+\Qed.
 
-(* Theorem T3: necessarily, a God exists *)
-Theorem theorem3: V (box (mexists x, (G x))).
-Proof.
-intro.
-apply lemma2.
-apply corollary1.
-Qed.
+\com{(* Theorem T3: necessarily, a God exists *)}
+\Theorem theorem3: [ box (mexists x, (G x)) ].
+\Proof. mv. apply lemma2. apply corollary1. \Qed.
 
-(* Corollary C2: There exists a god *)
-Theorem corollary2: V (mexists x, (G x)).
-Proof.
-intro.
-apply T.
-apply theorem3.
-Qed.
-\end{verbatim}
-
-Note: possible worlds machinery is hidden from the user.
+\com{(* Corollary C2: There exists a god *)}
+\Theorem corollary2: [ mexists x, (G x) ].
+\Proof. mv. apply T. apply theorem3. \Qed.
+\end{Verbatim}
 
@@
 
-cite related work about modal logics in Coq.
+The successful formalization of Scott's version of G\"odel's proof of 
+God's existence indicates that the 
+\emph{embedding} of modal logics 
+into higher-order logics via the \emph{possible world semantics} is a viable approach for interactive 
+theorem proving within modal logics.
+The minimalistic implementation of the embedding 
+(Section \ref{sec:modal}) takes special care to hide the 
+underlying possible world machinery from the user. 
+An inspection of the proof scripts in Section \ref{sec:proof} shows 
+that this goal has been achieved. The user does not have to 
+explicitly bother about worlds and their mutual reachability; 
+the provided tactics for modalities do the job for him. 
+And for subgoals that do not involve modalities, 
+the user has all the usual interactive tactics at his/her disposal. 
+%The implementation of hints to allow automation tactics to fully cope with the embedding remains for future work.
+Nevertheless, at this stage a tactic such as \texttt{firstorder} is able to prove all lemmas and theorems in Section \ref{sec:modal} fully automatically.
 
-no need to bother about the embedding.
+The theological implications of the verified correctness of 
+this proof certainly depend on a critical discussion of the 
+underlying concepts, definitions and axioms, 
+which is beyond the scope of this paper. Nevertheless, the application
+of theorem proving technology to a philosophical domain 
+has been fruitful for both areas. New results of philosophical 
+interest have been obtained \cite{ToDo:our ArXiv or Journal abstract} 
+and basic infra-structure for interactive and automated reasoning 
+in higher-order modal logics has been created.
 
-This work attests the maturity of contemporary interactive and
-automated deduction tools for classical higher-order logic and
-demonstrates the elegance and practical relevance of the
-embeddings-based approach.  Most importantly, our work opens new
-perspectives for a computer-assisted theoretical philosophy.  The
-critical discussion of the underlying concepts, definitions and axioms
-remains a human responsibility, but the computer can assist in
-building and checking rigorously correct logical arguments. In case of
-logico-philosophical disputes, the computer can check the disputing
-arguments and partially fulfill Leibniz' dictum: Calculemus --- Let us
-calculate!
-The philosophical and theological importance of the correctness of this proof is beyond the scope of this paper.
+\marginpar{cite John Rushby's formalization of anselm's argument?}
 
-We welcome feedback on
-
-\paragraph{Acknowledgements:} ToDo: thank Coq-Club members. Cedric Auger
+\paragraph{Acknowledgements:} Thanks to Cedric Auger and Laurent Th\'ery, for their answers to our newbie questions about Ltac in the Coq-Club mailing-list.
 

### 2014-01-30 Papers/ITP/GodProof-ITP.tex
update of ITP paper

@@
 \author{
-  Christoph Benzm\"{u}ller\inst{1} 
+  Christoph Benzm\"{u}ller\inst{1}\thanks{This work has been supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
   \and 
@@
 
-G\"{o}del's proof of God's existence is challenging 
-to formalize and verify because it requires an 
-expressive logical language with modal operators
-($\Diamond$ and $\Box$) and
-higher-order quantifiers. 
-Specialized modal logic tools lack higher-order capabilities, 
-while higher-order provers lack built-in support for modalities. 
-However, by embedding modal logics into higher-order logic, 
-it is in principle possible to use standard 
-higher-order automated and interactive theorem provers 
-for reasoning about and within modal logics \cite{J23,B9}.
-To realize this potential, formalizations \cite{FormalTheologyRepository} of G\"odel's proof in TPTP THF \cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been completed so far, in this order.
-This paper focuses on the formalization in Coq.
-The implementation of the embedding is discussed in detail in Section \ref{sec:modal} and Scott's version of G\"odel's proof is shown in Section \ref{sec:proof}.
+G\"{o}del's proof of God's existence is challenging to formalize and
+verify because it requires an expressive logical language with modal
+operators ($\Diamond$ and $\Box$) and higher-order quantifiers.
+Specialized modal logic tools lack higher-order capabilities, while
+higher-order provers lack built-in support for modalities.  However,
+by embedding modal logics into higher-order logic, it is in principle
+possible to use standard higher-order automated and interactive
+theorem provers for reasoning about and within modal logics
+\cite{J23,B9}.  To realize this potential, formalizations
+\cite{FormalTheologyRepository} of G\"odel's proof in TPTP THF
+\cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been
+completed so far, in this order. This paper focuses on the
+formalization in Coq. The implementation of the embedding is discussed
+in detail in Section \ref{sec:modal} and Scott's version of G\"odel's
+proof is shown in Section \ref{sec:proof}.
 
@@
 
-A crucial aspect of modal logics \cite{ToDo} is that the so-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$ is a theorem, but $A \imp \Box A$ is not necessarily a theorem.
+A crucial aspect of modal logics \cite{ModalLogic} is that the so-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$ is a theorem, but $A \imp \Box A$ is not necessarily a theorem.
 Naive attempts to define the modal operators $\Box$ and $\Diamond$ may easily be unsound in this respect.
@@
 Possible worlds are connected by an accessibility relation.
+
+The embedding is related to labeling techniques
+\cite{Labels}. However, the expressiveness of higher-order logic is
+exploited here to encode the labels within the logic language itself.
 
@@
 \Definition o := i -> \Prop. \com{(* Type of modal propositions *)}
-\Parameter r: i -> i -> \Prop. \com{(* Acessibility relation for worlds *)}
+\Parameter r: i -> i -> \Prop. \com{(* Accessibility relation for worlds *)}
 \end{Verbatim}
@@
 \noindent
-Although the embedding via the possible world semantics is necessary, it should be as transparent as possible to the user.
-Fortunately, basic Coq tactics such \texttt{intro}, \texttt{apply} and \texttt{split} automatically unfold the modal definitions and thus can be used with modal connectives and quantifiers as they are used with the usual connectives and quantifiers.
-Nothing else needs to be done.
-For the modal operators, however, the following simple tactics have been implemented, in order to allow the user to work with the concepts of \emph{necessity} and \emph{possibility} without having to unfold the definitions of modal operators and think in terms of possible worlds.
+To maximally preserve user intuition in interactive modal logic theorem
+proving, the embedding via the possible world semantics should be as
+transparent as possible to the user. Fortunately, basic Coq tactics
+such \texttt{intro}, \texttt{apply} and \texttt{split} automatically
+unfold the modal definitions. They can thus be used with modal
+connectives and quantifiers just as they are used with the usual
+connectives and quantifiers.  For the modal operators, however, the
+following simple tactics have been implemented, in order to allow the
+user to work with the concepts of \emph{necessity} and
+\emph{possibility} without having to unfold the definitions of modal
+operators and think in terms of possible worlds.
 
@@
   apply symmetry. exact R.
-
   exact R0.
@@
 Scott's assertions are emphasized with comments.
-In contrast to the formalization in Isabelle \cite{ToDo:our isabelle formalization in ArchiveFormalProofs}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
+In contrast to the formalization in Isabelle \cite{Archiv}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
 This was a deliberate choice, mainly because it allowed a qualitative evaluation of the convenience of the possible world semantic embedding approach for \emph{interactive} theorem proving.
@@
     box_e H2 G2. exact G2.
-
     clear H2 R H1 w. intro H5. apply H4. exists x. exact H5.
-
   assert (H6: ((box (mforall x, (p x) m-> m~ (x m= x))) w)). \com{(* Scott *)}   
     box_i. intro x. intros H7 H8. box_elim H3 w0 G3. eapply G3. exact H7.
-
     assert (H9: ((Positive (fun x => m~ (x m= x))) w)). \com{(* Scott *)}
@@
         exact H1.
-
         exact H6.
@@
         box_i. intros x H11. reflexivity.
-
         assert (H11 : ((Positive (fun x => (x m= x))) w)). \com{(* Scott *)}
@@
             exact H1.
-
             exact H10.
-
           apply axiom1a in H9. contradiction.
@@
   exact H1.
-
   intro q. intro H2. assert (H3: ((Positive q) w)).
@@
     contradiction. 
-
     cut (box (Positive q) w). \com{(* Scott *)}
       apply K. box_i. intro H5. intro y. intro H6. unfold G in H6. apply (H6 q). exact H5.
-
       apply axiom4. exact H3.
@@
     unfold G in H2. apply (H2 NE). apply axiom5.
-
     unfold NE in H3. apply H3.
-
   apply theorem2. exact H2.
@@
   apply dia_box_to_box.
-
   apply (mp_dia w (mexists z, G z)).
     exact H.
-
     box_i. apply lemma1.
@@
 The successful formalization of Scott's version of G\"odel's proof of
-God's existence indicates that the 
-\emph{embedding} of modal logics 
-into higher-order logics via the \emph{possible world semantics} is a viable approach for interactive 
-theorem proving within modal logics.
-The minimalistic implementation of the embedding 
-(Section \ref{sec:modal}) takes special care to hide the 
-underlying possible world machinery from the user. 
-An inspection of the proof scripts in Section \ref{sec:proof} shows 
-that this goal has been achieved. The user does not have to 
-explicitly bother about worlds and their mutual reachability; 
-the provided tactics for modalities do the job for him. 
-And for subgoals that do not involve modalities, 
-the user has all the usual interactive tactics at his/her disposal. 
+God's existence indicates that the \emph{embedding} of modal logics
+into higher-order logics via the \emph{possible world semantics} is a
+viable approach for interactive theorem proving within modal logics.
+The minimalistic implementation of the embedding (Section
+\ref{sec:modal}) takes special care to hide the underlying possible
+world machinery from the user.  An inspection of the proof scripts in
+Section \ref{sec:proof} shows that this goal has been achieved. The
+user does not have to explicitly bother about worlds and their mutual
+reachability; the provided tactics for modalities do the job for him.
+And for subgoals that do not involve modalities, the user has all the
+usual interactive tactics at his/her disposal.
 %The implementation of hints to allow automation tactics to fully cope with the embedding remains for future work.
@@
 
-The theological implications of the verified correctness of 
-this proof certainly depend on a critical discussion of the 
-underlying concepts, definitions and axioms, 
-which is beyond the scope of this paper. Nevertheless, the application
-of theorem proving technology to a philosophical domain 
-has been fruitful for both areas. New results of philosophical 
-interest have been obtained \cite{ToDo:our ArXiv or Journal abstract} 
-and basic infra-structure for interactive and automated reasoning 
-in higher-order modal logics has been created.
+The theological implications of the verified correctness of this proof
+certainly depend on a critical discussion of the underlying concepts,
+definitions and axioms, which is beyond the scope of this paper. What
+can be said though is that our work presents further evidence that
+theism is at least not (obviously) irrational.
 
-\marginpar{cite John Rushby's formalization of anselm's argument?}
+Related work includes formalizations of Anselm's argument by Rushby
+\cite{Rushby}, and by Oppenheimer and Zalta \cite{Zalta}.  Obviously,
+the application of theorem proving technology to the domain of
+theoretical philosophy can --- as already pictured by Leibniz ---
+be very fruitful for both areas. In fact, new results of
+philosophical interest have been obtained by our
+work~\cite{arXiv}. Moreover a first, basic infrastructure for
+interactive and automated reasoning in higher-order modal logics has
+been created and will be further improved in the future.
 
@@
 \bibitem{Adams}
-R.M. Adams.
+R.M.~Adams.
 \newblock Introductory note to *1970.
@@
 \bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings.
+A.C.~Anderson and M.~Gettings.
 \newblock G\"odel ontological proof revisited.
@@
 \bibitem{B9}
-C.~Benzm{\"u}ller and L.C. Paulson.
+C.~Benzm{\"u}ller and L.C.~Paulson.
 \newblock Exploring properties of normal multimodal logics in simple type
@@
 \bibitem{J23}
-C.~Benzm{\"u}ller and L.C. Paulson.
+C.~Benzm{\"u}ller and L.C.~Paulson.
 \newblock Quantified multimodal logics in simple type theory.
@@
 
+\bibitem{arXiv}
+C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
+\newblock Formalization, Mechanization and Automation of G{\"o}del's Proof of God's Existence.
+\newblock {\em {arXiv}}:1308.4526, 2013.
+
+\bibitem{arXiv}
+C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
+\newblock G{\"o}del's God in Isabelle/HOL. 
+\newblock {\em Archive of Formal Proofs}, 2013.
+
 \bibitem{Coq}
@@
 
+\bibitem{ModalLogic}
+P.~Blackburn, J.v.~Benthem, and F.~Wolter (eds.), \newblock {\em {Handbook of Modal Logic}}. Elsevier, 2006.
+
 \bibitem{Sledgehammer}
-J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+J.C.~Blanchette, S.~B\"ohme, and L.C.~Paulson.
 \newblock Extending {Sledgehammer} with {SMT} solvers.
@@
 \bibitem{Nitpick}
-J.C. Blanchette and T.~Nipkow.
+J.C.~Blanchette and T.~Nipkow.
 \newblock Nitpick: A counterexample generator for higher-order logic based on a
   relational model finder.
-\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+\newblock In {\em Proc. of ITP 2010}, no.~6172 in LNCS, pages 131--146.
   Springer, 2010.
 
+\bibitem{Zalta}
+P.E.~Oppenheimera and E.N.~Zalta.
+\newblock A Computationally-Discovered Simplification of the Ontological Argument.
+\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.
+
 \bibitem{Satallax}
-C.E. Brown.
+C.E.~Brown.
 \newblock Satallax: An automated higher-order prover.
@@
 
+\bibitem{Labels}
+D.M.~Gabbay. {\em Labelled Deductive Systems}. Clarendon Press, 1996.
+
 \bibitem{Goedel1970}
@@
 \bibitem{Hazen}
-A.P. Hazen.
+A.P.~Hazen.
 \newblock On g\"odel's ontological proof.
@@
 
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
+\bibitem{Rushby}
+J.~Rushby.
+\newblock The Ontological Argument in PVS. 
+\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St. Petersburg, Russia, 2013.
 
@@
 \bibitem{sobel2004logic}
-J.H. Sobel.
+J.H.~Sobel.
 \newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
@@
 
+\bibitem{FormalTheologyRepository}
+B.~Woltzenlogel Paleo and C.~Benzm\"uller.
+\newblock Formal theology repository
+  (\url{https://github.com/FormalTheology/GoedelGod}).
 \end{thebibliography}

### 2014-01-31 Papers/ITP/GodProof-ITP.tex
Improve paper

@@
 To avoid this issue, the \emph{possible world semantics} of modal logics can be explicitly embedded into higher-order logics \cite{J23,B9}.
-To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type that depends on possible worlds.
-Possible worlds are connected by an accessibility relation.
-
 The embedding is related to labeling techniques
@@
 exploited here to encode the labels within the logic language itself.
+To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type that depends on possible worlds.
+Possible worlds are connected by an accessibility relation.
 
@@
 transparent as possible to the user. Fortunately, basic Coq tactics
-such \texttt{intro}, \texttt{apply} and \texttt{split} automatically
+such as \texttt{intro}, \texttt{apply} and \texttt{split} automatically
 unfold the modal definitions. They can thus be used with modal
@@
 reachability; the provided tactics for modalities do the job for him.
-And for subgoals that do not involve modalities, the user has all the
+Moreover, for subgoals that do not involve modalities, the user has all the
 usual interactive tactics at his/her disposal.
@@
 Related work includes formalizations of Anselm's argument by Rushby
-\cite{Rushby}, and by Oppenheimer and Zalta \cite{Zalta}.  Obviously,
+\cite{Rushby}, and by Oppenheimer and Zalta \cite{Zalta}.  Clearly,
 the application of theorem proving technology to the domain of

### 2014-02-01 Formalizations/THF/call_tptp.sh
improved script

@@ -4,7 +4,7 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 proverlist='LEO-II---1.6.0 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-timeout=120
+timeout=10
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
@@ -15,11 +15,16 @@ echo
     for prover in $proverlist 
     do
     function_to_fork () { 
-        echo $file ++++++ `$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover | grep RESULT` 
+      res=""
+      while ! (echo $res | grep --quiet RESULT); do
+          res=`$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover |grep RESULT` 
+	  echo ${file}---${prover}---${res}
+      done   
     }
     function_to_fork &
     done
-    echo 
+    wait
+    echo Done 
 done
 
 

### 2014-05-19 Papers/2014-ECAI/ecai2014.tex
Add ECAI papers

\documentclass{ecai2014}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}

%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

\begin{document}

\title{Guidelines for Preparing a Paper for the \\
European Conference on Artificial Intelligence}

\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3\institute{University of Leipzig,
Germany, email: somename@informatik.uni-leipzig.de} }

\maketitle
\bibliographystyle{ecai2014}

\begin{abstract}
  The purpose of this paper is to show a contributor the required
  style for a paper for ECAI-2014 and PAIS-2014. The specifications
  for layout are described so that non-\LaTeX\ users can create their
  own style sheet to achieve the same layout. The source for the
  sample file is available for \LaTeX\ users. The PostScript and the
  PDF file is available for all. A Microsoft Word template is also available.
 The layout is identical to ECAI'12
  papers. The publisher (IOS Press) will insert a footer for each
  page.
\end{abstract}

\section{PAGE LIMIT}
The page limit for ECAI-2014 and PAIS-2014 long papers is \textbf{6
  (six)} pages in the required format. The page limit for short
submissions is 2 pages. This is a strict limit. Overlength papers will
not be reviewed.

\section{GENERAL SPECIFICATIONS}
The following details should allow contributors to set up the general
page description for their paper:

\begin{enumerate}
\item The paper is set in two columns each 20.5 picas (86 mm) wide
  with a column separator of 1.5 picas (6 mm).

\item The typeface is Times Modern Roman.

\item The body text size is 9 point (3.15 mm) on a body of 11
point (3.85 mm) (i.e., 61 lines of text).

\item The effective text height for each page is 56 picas (237 mm).
The first page has less text height. It requires an additional footer
space of 3.5 picas (14.8 mm) for the copyright inserted by the publisher
and 1.5 picas (6 mm) of space before the title.
The effective text height of the first page is 51 picas (216 mm).

\item There are no running feet for the final camera-ready version of the
paper. The submission paper should have page numbers in the running feet.

\end {enumerate}

\section{TITLE, AUTHOR, AFFILIATION, COPYRIGHT
AND RUNNING FEET}
\subsection{Title}
The title is set in 20 point (7 mm) bold with leading of 22 point (7.7
mm), centered over the full text measure, with 1.5 picas (6 mm) of
space before and after.

\subsection{Author}
The author's name is set in 11 point (3.85 mm) bold with leading of 12
point (4.2 mm), centered over full text measure, with 1.5 picas (6 mm)
of space below. A footnote indicator is set in 11 point (3.85 mm)
medium and positioned as a superscript character.

\subsection{Affiliation}
The affiliation is set as a footnote to the first column. This is set
in 8 point (2.8 mm) medium with leading of 8.6 point (3.1 mm), with a
1 point (0.35 mm) footnote rule to column width.

\subsection{Copyright}
The copyright details will be inserted by the publisher.

\subsection{Running feet}
The running feet are inserted by the publisher. For submission you may
insert page numbers in the middle of the running feet. Do not,
however, insert page numbers for the camera-ready version of the
paper.

\section{ABSTRACT}
The abstract for the paper is set in 9 point (3.15 mm) medium, on a
body of 10 point (3.5 mm). The word Abstract is set in bold, followed
by a full point and a 0.5 pica space.

\section{HEADINGS}\label{heads}
Three heading levels have been specified:

\begin{enumerate}
\item{A level headings}

\begin{itemize}
\item The first level of heading is set is 11 point (3.85 mm) bold, on
  a body of 12 point (4.2 mm), 1.5 lines of space above and 0.5 lines
  of space below.

\item The heading is numbered to one digit with a 1 pica space
  separating it from the text.

\item The text is keyed in capitals and is unjustified.
\end{itemize}

\item{B level headings}
\begin{itemize}
\item The second level of heading is set is 11 point (3.85 mm) bold,
  on a body of 12 point (4.2 mm), 1.5 lines of space above and 0.5
  lines of space below.

\item The heading is numbered to two digits separated with a full
  point, with a 1 pica space separating it from the text.

\item The text is keyed in upper and lower case with an initial
  capital for first word only, and is unjustified.
\end{itemize}

\item{C level headings}
\begin{itemize}
\item The third level of heading is set is 10 point (3.5 mm) italic,
  on a body of 11 point (3.85 mm), 1.5 lines of space above and 0.5
  lines of space below.

\item The heading is numbered to three digits separated with a full
  point, with a 1 pica space separating it from the text.

\item The text is keyed in upper and lower case with an initial
  capital for first word only, and is unjustified.
\end{itemize}

\item{Acknowledgements}

  This heading is the same style as an A level heading but is not
  numbered.
\end{enumerate}

\section{TEXT}
The first paragraph of text following any heading is set to the
complete measure (i.e., do not indent the first line).

Subsequent paragraphs are set with the first line indented
by 1 pica (3.85 mm).

There isn't any inter-paragraph spacing.

\section{LISTS}
The list identifier may be an arabic number, a bullet, an em
rule or a roman numeral.

The items in a list are set in text size and indented by 1
pica (4.2 mm) from the left margin. Half a line of space is
set above and below the list to separate it from surrounding
text.

See layout of Section \ref{heads} on headings to see the results of the list macros.

\section{TABLES}
Tables are set in 8 point (2.8 mm) on a body of 10 point (3.5 mm).
The table caption is set centered at the start of the table, with
the word Table and the number in bold. The caption is set in medium
with a 1 pica (4.2 mm) space separating it from the table number.

A one line space separates the table from surrounding text.

\begin{table}
\begin{center}
{\caption{The table caption is centered on the table measure. If it
extends to two lines each is centered.}\label{table1}}
\begin{tabular}{lccccccc}
\hline
\rule{0pt}{12pt}
&\multicolumn{7}{c}{Processors}\\
&1&\multicolumn{3}{c}{2}&\multicolumn{3}{c}{4}\\
\cline{2-8}
\rule{0pt}{12pt}
Window&$\Diamond$&$\Diamond$&$\Box$&$\bigtriangleup$&$\Diamond$&$\Box$&$\bigtriangleup$
\\
\hline
\\[-6pt]
\quad1&1273&110&21.79&89\%&6717&22.42&61\%\\
\quad2&2145&116&10.99&50\%&5386&10.77&19\%\\
\quad3&3014&117&41.77&89\%&7783&42.31&58\%\\
\quad4&4753&151&71.55&77\%&7477&61.97&49\%\\
\quad5&5576&148&61.60&80\%&7551&91.80&45\%
\\
\hline
\\[-6pt]
\multicolumn{8}{l}{$\Diamond$ execution time in ticks\ \
$\Box$ speed-up values\ \
$\bigtriangleup$ efficiency values}
\end{tabular}
\end{center}
\end{table}

\section{FIGURES}
A figure caption is set centered in 8 point (2.8 mm) medium on a
leading of 10 point (3.5 mm).  It is set under the figure, with the
word Figure and the number in bold and with a 1 pica (4.2 mm) space
separating the caption text from the figure number.

One line of space separates the figure from the caption. A one line
space separates the figure from surrounding text.

\begin{figure}
\centerline{\includegraphics[height=3.5in]{ecaif01}}
\caption{Network of transputers and the structure of individual
processes } \label{procstructfig}
\end{figure}

\section{EQUATIONS}
A display equation is numbered, using arabic numbers in parentheses.
It is centered and set with one line of space above and below to
separate it from surrounding text. The following example is a simple
single line equation:

\begin{equation}
Ax=b
\label{thesystem}
\end{equation}

The next example is a multi-line equation:

\begin{eqnarray}
(x+y)(x-y) & = & x^2-xy+xy-y^2\\
(x+y)^2    & = & x^2+2xy+y^2
\end{eqnarray}

The equal signs are aligned in a multi-line equation.

\section{PROGRAM LISTINGS}
Program listings are set in 9 point (3.15 mm) Courier on a
leading of 11 point (3.85 mm). That is to say, a non-proportional
font is used to ensure the correct alignment.

A one line space separates the program listing from surrounding text.

\begin{verbatim}
void inc(x)
int* x;
{
    *x++;
}

int y = 1;
inc(&y);
printf("%d\n",y);
\end{verbatim}

\section{THEOREMS}
The text of a theorem is set in 9 point (3.15 mm) italic on a
leading of 11 point (3.85 mm). The word Theorem and its number
are set in 9 point (3.15 mm) bold.

A one line space separates the theorem from surrounding text.

\begin{theorem}
Let us assume this is a valid theorem. In reality it is a piece
of text set in the theorem environment.
\end{theorem}

\section{FOOTNOTES}
Footnotes are set in 8 point (2.8 mm) medium with leading of 8.6 point (3.1
mm), with a 1 point (0.35 mm) footnote rule to column
width\footnote{This is an example of a footnote that occurs in
the text. If the text runs to two lines the second line aligns
with the start of text in the first line.} .

\section{REFERENCES}
The reference identifier in the text is set as a sequential number in
square brackets. The reference entry itself is set in 8 point
(2.8 mm) with a leading of 10 point (3.5 mm), and appears
in the sequence in which it is cited in the paper.

\section{SAMPLE CODING}
The remainder of this paper contains examples of the specifications
detailed above and can be used for reference if required.

\section{PROGRAMMING MODEL}
\label{par}
Our algorithms were implemented using the \emph{single program,
  multiple data} model (SPMD). SPMD involves writing a single code
that will run on all the processors co-operating on a task. The data
are partitioned among the processors which know what portions of the
data they will work on \cite{kn:Golub89}.

\subsection{Structure of processes and processors}
\label{procstruct}
The grid has $P=P_{\rm{r}}\times P_{\rm{c}}$ processors, where
$P_{\rm{r}}$ is the number of rows of processors and $P_{\rm{c}}$ is
the number of columns of processors.

\subsubsection{Routing information on the grid}
\label{routing} A message may be either \emph{broadcast} or {\em
  specific}. A broadcast message originates on a processor and is
relayed through the network until it reaches all other processors. A
specific message is one that is directed to a particular target
processor.

Broadcast messages originate from a processor called \emph{central}
which is situated in the `middle' of the grid. This processor has
co-ordinates $(\lfloor P_{\rm{r}}/2 \rfloor,\lfloor P_{\rm{c}}/2
\rfloor)$.  Messages are broadcast using the \emph{row--column
  broadcast} algorithm (RCB), which uses the following strategy.  The
number of steps required to complete the RCB algorithm (i.e., until
all processors have received the broadcast value) is given by $\lfloor
P_{\rm{r}}/2\rfloor+\lfloor P_{\rm{c}}/2\rfloor$.

A specific message is routed through the processors using the
\emph{find-row--find-column} algorithm (FRFC) detailed in
\cite{kn:deCarlini91}.  The message is sent from the \emph{originator}
processor vertically until it reaches a processor sitting in the same
row as the \emph{target} processor. The message is then moved
horizontally across the processors in that row until it reaches the
target processor. An accumulation based on the recursive doubling
technique \cite[pp. 56--61]{kn:Modi88}, would require the same number
of steps as the RCB requires. If either the row or column of the
originator and target processors are the same then the message will
travel only in a horizontal or vertical direction, respectively (see
\cite{kn:Smith85}).

\section{DATA PARTITIONING}
\label{datapart}
We use \emph{data partitioning by contiguity}, defined in the
following way. To partition the data (i.e., vectors and matrices)
among the processors, we divide the set of variables
\mbox{$V=\{\,i\,\}_{i=1}^{N}$} into $P$ subsets
$\{\,W_{p}\,\}_{p=1}^{P}$ of $s=N/P$ elements each. We assume without
loss of generality that $N$ is an integer multiple of $P$. We define
each subset as $W_{p}=\{(p-1)s+j\}_{j=1}^{s}$ (see
\cite{kn:Schofield89}, \cite{kn:daCunha92a} and \cite{kn:Atkin} for
details).

Each processor $p$ is responsible for performing the computations over
the variables contained in $W_{p}$. In the case of vector operations,
each processor will hold segments of $s$ variables. The data
partitioning for operations involving matrices is discussed in Section
\ref{mv}.
\section{LINEAR ALGEBRA OPERATIONS}
\label{blas}
\subsection{Saxpy}
\label{saxpy}
The saxpy $w=u+\alpha v$ operation, where $u$, $v$ and $w$ are vectors
and $\alpha$ is a scalar value, has the characteristic that its
computation is \emph{disjoint elementwise} with respect to $u, v$ and
$w$. This means that we can compute a saxpy without any communication
between processors; the resulting vector $w$ does not need to be
distributed among the processors. Parallelism is exploited in the
saxpy by the fact that $P$ processors will compute the same operation
with a substantially smaller amount of data. The saxpy is computed as
\begin{equation}
w_{i}=u_{i}+\alpha v_{i},\quad \forall i\in\{W_{p}\}_{p=1}^{P}
\label{distsaxpy}
\end{equation}
\subsection{Inner-product and vector 2-norm}
\label{utv}
The inner-product $\alpha=u^{T}v=\sum_{i=1}^{N}{u_{i}v_{i}}$ is an
operation that involves accumulation of data, implying a high level of
communication between all processors. The mesh topology and the
processes architecture used allowed a more efficient use of the
processors than, for instance, a ring topology, reducing the time that
processors are idle waiting for the computed inner-product value to
arrive, but the problem still remains. The use of the SPMD paradigm
also implies the global broadcast of the final computed value to all
processors.

The inner-product is computed in three distinct phases. Phase 1 is the
computation of partial sums of the form
\begin{equation}
\alpha_{p}=\sum_{\forall i\in\{W_{p}\}}{u_{i}\times v_{i}},\quad p=1,\ldots,P
\label{distutv}
\end{equation}

The accumulation phase of the inner-product using the RCA algorithm is
completed in the same number of steps as the RCB algorithm (Section
\ref{routing}). This is because of the need to relay partial values
between processors without any accumulation taking place, owing to the
connectivity of the grid topology.

The vector 2-norm $\alpha=||\,u\,||_{2}=\sqrt{u^{T}u}$ is computed
using the inner-product algorithm described above. Once the
inner-product value is received by a processor during the final
broadcast phase, it computes the square root of that value giving the
required 2-norm value.

\subsection{Matrix--vector product}
\label{mv}
For the matrix--vector product $v=Au$, we use a \emph{column
  partitioning} of $A$.  Each processor holds a set $W_{p}$ (see
Section \ref{datapart}) of $s$ columns each of $N$ elements of $A$ and
$s$ elements of $u$.  The $s$ elements of $u$ stored locally have a
one-to-one correspondence to the $s$ columns of $A$ (e.g. a processor
holding element $u_{j}$ also holds the $j$-th column of $A$). Note
that whereas we have $A$ partitioned by columns among the processors,
the matrix--vector product is to be computed by \emph{rows}.

The algorithm for computing the matrix--vector product using column
partitioning is a generalization of the inner-product algorithm
described in Section \ref{utv} (without the need for a final broadcast
phase). At a given time during the execution of the algorithm, each
one of $P-1$ processors is computing a vector $w$ of $s$ elements
containing partial sums required for the segment of the vector $v$ in
the remaining `target' processor.  After this computation is complete,
each of the $P$ processors stores a vector $w$.  The resulting segment
of the matrix--vector product vector which is to be stored in the
target processor is obtained by summing together the $P$ vectors $w$,
as described below.

Each processor other than the target processor sends its $w$ vector to
one of its neighboring processors. A processor decides whether to send
the vector in either the row or column direction to reach the target
processor based on the FRFC algorithm (see Section \ref{routing}). If
a vector passes through further processors in its route to the target
processor the $w$ vectors are accumulated. Thus the target processor
will receive at most four $w$ vectors which, when summed to its own
$w$ vector, yield the desired set of $s$ elements of $v$.

\subsection{Matrix--vector product---finite-difference approximation}
We now consider a preconditioned version of the conjugate-gradients
method \cite{kn:Golub89}. Note that we do not need to form $A$
explicitly.  This implies a very low degree of information exchange
between the processors which can be effectively exploited with
transputers, since the required values of $u$ can be exchanged
independently through each link.

The preconditioning used in our implementations is the polynomial
preconditioning (see \cite{kn:Saad85}, \cite{kn:Eisenstat81},
\cite{kn:Adams85} and \cite{kn:Johnson83}), which can be implemented
very efficiently in a parallel architecture since it is expressed as a
sequence of saxpys and matrix--vector products.

We have $l$ rows and columns in the discretization grid, which we want
to partition among a $P_{\rm{r}}\times P_{\rm{c}}$ mesh of
processors. Each processor will then carry out the computations
associated with a block of $\lfloor
l/P_{\rm{r}}\rfloor+\hbox{sign}\left( l\bmod P_{\rm{r}}\right)$ rows
and $\lfloor l/P_{\rm{c}}\rfloor+\hbox{sign}\left( l\bmod
  P_{\rm{c}}\right)$ columns of the interior points of the grid.

The matrix--vector product using the column partitioning is highly
parallel. Since there is no broadcast operation involved, as soon as a
processor on the boundary of the grid (either rows or columns) has
computed and sent a $w_{p}$ vector destined to a processor `A', it can
compute and (possibly) send a $w_{p}$ vector to processor `B', at
which time its neighboring processors may also have started computing
and sending their own $w$ vectors to processor `B'.

At a given point in the matrix--vector product computation, the
processors are computing $w$ vectors destined to processor A. When
these vectors have been accumulated in the row of that processor (step
1), the processors in the top and bottom rows compute and send the $w$
vectors for processor B, while the processors on the left and right
columns of the row of processor A send the accumulated $r$ vectors to
processor A (step 2). Processor A now stores its set of the resulting
$v$ vector (which is the accumulation of the $w$ vectors). In step 3,
the processors in the bottom row compute and send the $w$ vectors for
processor C while the processor at the left-hand end of the row of
processor B sends the accumulated $w$ vectors of that column towards
processor B. The next steps are similar to the above.

In our implementation, we exploit the geometry associated with the
regular grid of points used to approximate the PDE. A geometric
partitioning is used to match the topology and connectivity present in
the grid of transputers (Section \ref{procstruct}).

The discretization of the PDE is obtained by specifying a grid size
$l$ defining an associated grid of $N=l^2$ interior points (note that
this is the order of the linear system to be solved). With each
interior point, we associate a set of values, namely the coefficients
$C, N, S, E\,$ and $W$.

\section{CONCLUSION}
We have shown that an iterative method such as the preconditioned
conjugate-gradients method may be successfully parallelized by using
highly efficient parallel implementations of the linear algebra
operations involved. We have used the same approach to parallelize
other iterative methods with similar degrees of efficiency (see
\cite{kn:daCunha92a} and \cite {kn:daCunha92b}).

\ack We would like to thank the referees for their comments, which
helped improve this paper considerably

\bibliography{ecai2014}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-19 Papers/2014-ECAI/paper.tex
Add ECAI papers

\documentclass{ecai2014}
\usepackage{times}
\usepackage{graphicx}

\usepackage{verbatim}
\usepackage{wrapfig}
\usepackage{amsmath,latexsym}
\usepackage{xspace}
\usepackage{stmaryrd}
\usepackage{txgreeks}
\usepackage[hyphens]{url}

%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

%%%%%%%%%%%% some personal definitions  %%%%%%%%%%%%%

% -----------------------------------------------
\newtheorem{definition}{Def.}
\newtheorem{example}{Ex.}
% -----------------------------------------------

\def\lambdot{\rule{0.6mm}{0.6mm}\hspace{0.4ex}} 
\def\all#1{\forall #1\lambdot}
\def\exi#1{\exists #1\lambdot}
\def\lam#1{\lambda #1\lambdot}

\def\modal#1{#1}
\def\mfalse{\modal\bot}
\def\mtrue{\modal\top}
\def\mnot{\modal\neg\,}
\def\mor{\,\modal\vee\,}
\def\mand{\,\modal\wedge\,}
\def\mimpl{\,\modal\supset\,}
\def\miff{\,\modal\Leftrightarrow\,}
\def\mball#1{\modal\Box_{#1}\,}
\def\mdexi#1{\modal\Diamond_{#1}\,}
\def\mall#1{\modal{\forall}{#1}\lambdot\,}
\def\mexi#1{\modal{\exists}{#1}\lambdot\,}
\def\mpi{\modal{\Pi}\,}
\def\mvalid{\modal{\texttt{valid}}}

\def\Metaeq{=}
\def\bnormform#1{\left.#1\hspace*{-.4ex}\right\downarrow_\beta}
\def\benormform#1{\left.#1\hspace*{-.4ex}\right\downarrow_{\beta\eta}}
\def\Bnormform#1{\left.{#1}\hspace*{-.4ex}\right\downarrow_\beta}
\def\Benormform#1{\left.{#1}\hspace*{-.4ex}\right\downarrow_{\beta\eta}}
\def\ambnormform#1{{#1}\hspace*{-1.1ex}\downarrow_{\kern-.2em\scriptscriptstyle *}} % ``ambiguous'' normal form
\def\eqb{\Metaeq_{\beta}}
\def\eqe{\Metaeq_{\eta}}
\def\eqbe{\Metaeq_{\beta\eta}}
\def\convarrow{\rightarrow}

\newcommand\entity[1]{\text{\textrm{#1}}}
\def\QHL{\entity{QHL}}
\def\QML{\entity{QML}}
\def\NOM{\entity{NOM}}
\def\SVAR{\entity{SVAR}}
\def\CON{\entity{CON}}
\def\FVAR{\entity{FVAR}}
%\def\IC{\entity{IC}}
\def\FSYM{\entity{FSYM}}
\def\RSYM{\entity{RSYM}}
\def\QHLSTT{\entity{QHLSTT}}
\def\QMLSTT{\entity{QMLSTT}}
\def\IV{\entity{IV}}
\def\PV{\entity{PV}}
\def\SYM{\entity{SYM}}
\def\IVSTT{\entity{IVSTT}}
\def\PVSTT{\entity{PVSTT}}
\def\SYMSTT{\entity{SYMSTT}}
\def\SSTT{\entity{SSTT}}
\def\AR{\entity{AR}}
\def\STT{\entity{STT}\xspace}

\def\QKPIm{\entity{QK}\pi^-\xspace}
\def\QKPI{\entity{QK}\pi\xspace}
\def\QKPIp{\entity{QK}\pi^+\xspace}
\def\QSFPIm{\entity{QS5}\pi^-\xspace}
\def\QSFPI{\entity{QS5}\pi\xspace}
\def\QSFPIp{\entity{QS5}\pi^+\xspace}

\def\stt{\entity{STT}\xspace}
\def\tt{\entity{STT}}
\def\lm{\entity{MM}\xspace}
\def\ipl{\entity{IPL}\xspace}
\def\HOML{\entity{HOML}\xspace}
\def\HOL{\entity{HOL}\xspace}

\def\worldtype{\mu}
\def\indtype{\iota}
\def\mutype{\mu}
\def\boola{\omicron}
\def\boolb{\hat{\omicron}}

\def\ar{\shortrightarrow}

\newcommand\hol[1]{\boldsymbol{#1}}
\newcommand\lift[1]{\lceil #1 \rceil}
\newcommand\llift[1]{\dot{#1}}

\newcommand{\imp}{\supset}
\newcommand{\biimp}{\equiv}
\newcommand{\allq}{\forall}
\newcommand{\exq}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly
\newcommand{\ess}[2]{#1 \ \mathit{ess.} \ #2}
\newcommand{\NE}{\mathit{NE}}

%%%%%%%%%%

\begin{document}

\title{Automating G\"odel's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers}

\author{Christoph Benzm\"uller\institute{Freie Universit\"at Berlin,
    Germany, email: c.benzmueller@fu-berlin.de}
  \and Bruno Woltzenlogel Paleo\institute{Technical University
    Vienna, Austria,  email: bruno@logic.at}  
}

\maketitle

\begin{abstract}
  Kurt G{\"o}del's ontological argument for God's existence has been
  formalized and automated on a computer with higher-order automated
  theorem provers. From G{\"o}del's premises, the computer proved:
  necessarily, there exists God. On the other hand, the theorem
  provers have also confirmed prominent criticism on G{\"o}del's
  ontological argument, and they found some new results about it.

  The background theory of the work presented here offers a novel
  perspective towards a \emph{computational theoretical philosophy}.
\end{abstract}

\section{INTRODUCTION}
Kurt G{\"o}del proposed an argumentation formalism to prove the
existence of God \cite{GoedelNotes,ScottNotes}. Attempts to prove the
existence (or non-existence) of God by means of abstract, ontological
arguments are an old tradition in western philosophy.  Before
G{\"o}del, several prominent philosophers, including St. Anselm of
Canterbury, Descartes and Leibniz, have presented similar
arguments. Moreover, there is an impressive body of recent and ongoing
work (cf.~\cite{sobel2004logic,Fitting,ContemporaryBibliography} and the references therein).
Ontological arguments, for or against the existence of God,
illustrate well an essential aspect of metaphysics: some (necessary) facts
for our existing world are deduced by purely a priori, analytical means from some
abstract definitions and axioms. % Contingent truths are to be
% distinguished from necessary truths.

What motivated G{\"o}del as a logician was the question, whether it is
possible to deduce the existence of God from a small number of
foundational (but debatable) axioms and definitions, with a mathematically precise,
formal argumentation chain in a well defined logic.

In theoretical philosophy, formal logical confrontations with such
ontological arguments had been so far (mainly) limited to paper
and pen.  Up to now, the use of computers was prevented, because the
logics of the available theorem proving systems were not expressive
enough to formalize the abstract concepts adequately. G{\"o}del's proof
uses, for example, a complex higher-order modal logic (\HOML) to handle
concepts such as \emph{possible} and \emph{necessary} and to support
quantification over individuals and properties.

Current works \cite{J23,B9} of the first author and Paulson illustrate that many
expressive logics, including quantified (multi-)modal logics, can be
embedded into the classical higher-order logic (\HOL), which can thus be seen
as a universal logic \cite{C36}. For this universal logic, efficient automated
theorem provers have been developed in recent years, and these systems
were now employed in our work.

G\"{o}del defines God (see Fig.~\ref{fig1}) as a being who possesses all \emph{positive}
properties.  He does not extensively discuss what positive properties
are, but instead he states a few reasonable (but debatable) axioms
that they should satisfy.  Various slightly different versions of
axioms and definitions have been considered by G\"{o}del and by
several philosophers who commented on his proof
(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}%\small
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both: \\ \phantom{b} \hfill 
  $\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$ 
\item[A2] A property necessarily implied by a
  positive property is positive:  \hfill 
  $\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
  \imp \psi(x)]) \imp P(\psi)]}$ 
\item[T1] Positive properties are possibly exemplified: \\ \phantom{b} \hfill $\hol{\allq
  \phi [P(\phi) \imp \pos \exq x \phi(x)]}$ 
\item[D1] A \emph{God-like} being possesses all positive properties: \\ \phantom{b} \hfill
  $\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$ 
\item[A3]  The property of being God-like is positive: \hfill   $\hol{P(G)}$ 
\item[C\phantom{1}] Possibly, God exists: \hfill $\hol{\pos \exq x G(x)}$ 
\item[A4]  Positive properties are necessarily positive: \\  \phantom{b} \hfill 
  $\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$ 
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: \\
  \phantom{b} \hfill $\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$ 
\item[T2]  Being God-like is an essence of any
  God-like being:  \\ \phantom{b}  \hfill $\hol{\allq x [G(x) \imp \ess{G}{x}]}$ 
\item[D3] \emph{Necessary existence} of an individ.~is the necessary exemplification of all its essences: 
  \phantom{b} \hfill $\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
  \exq y \phi(y)]}$
\item[A5] Necessary existence is a positive property: \hfill $\hol{P(\NE)}$ 
\item[T3] Necessarily, God exists: \hfill $\hol{\nec \exq x G(x)}$ 
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Scott's version of G\"odel's ontological argument \cite{ScottNotes}.\label{fig1}} 
\end{figure}

The overall idea of G{\"o}del's proof is in the tradition of Anselm's
argument, who defined God as some entity of which nothing greater can be
conceived. Anselm argued that existence in the actual world would
make such an assumed being even greater; hence, by definition God must
exist. G{\"o}del's ontological argument is clearly related to this
reasoning pattern. However, it also tries to fix some fundamental
weaknesses in Anselm's work. For example, G{\"o}del explicitly proves
that God's existence is possible, which has been a basic assumption of
Anselm. Because of this, Anselm's argument has been
criticized as incomplete by Leibniz. Leibniz instead claimed that the
assumption should be derivable from the definition of God as a perfect
being and from the notion of perfection.  G{\"o}del's proof 
addresses this critique, and it also addresses the critique of others,
including Kant's objection that existence should not be treated as a predicate. On the other hand, G{\"o}del's work
still leaves room for criticism, in particular, his axioms are so
strong that they imply \emph{modal collapse}, that is, a situation where
contingent truths and necessary
truths coincide. More information on the philosophical debate on
G{\"o}del's proof is provided in~\cite{sobel2004logic}.

We have analyzed Dana Scott's version of G\"{o}del's proof
\cite{ScottNotes} (cf.~Fig.~\ref{fig1}) for the first-time with an
unprecedented degree of detail and formality with the help of
higher-order automated theorem provers (HOL-ATPs).\footnote{All
  sources of our formalization are publicly available
  at~\url{https://github.com/FormalTheology/GoedelGod}. Our work has
  attracted major public interest, and leading media institutions
  worldwide have reported on it; some exemplary links to respective
  media reports and interviews are available at the above URL (see
  `Press' subfolder).}  The following has been done (and in this
order):
% \footnote{System URLs:
%  TPTP---\url{http://tptp.org};
%  Nitpick---\url{http://www4.in.tum.de/\~{}blanchet/nitpick.html};
%  LEO-II---\url{http://leoprover.org};
%  Satallax---\url{http://www.ps.uni-saarland.de/\~{}cebrown/satallax}; 
%  Coq---\url{http://coq.inria.fr};
%  Isabelle---\url{http://isabelle.in.tum.de}.
% }
(i) a detailed natural deduction proof;
%
(ii) a formalization in TPTP THF syntax \cite{J22};
%
(iii) an automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick};
%
(iv) an automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax};
%
(v) a step-by-step formalization using the Coq proof assistant \cite{Coq};
%
(vi) a formalization using the Isabelle proof assistant \cite{Isabelle}, where the
theorems (and some additional lemmata) have been automated with the
Isabelle tools Sledgehammer %~\cite{Sledgehammer} 
and Metis. %~\cite{Metis}.
Subsequently, we have studied additional consequences
of G\"odel's axioms, including modal collapse and monotheism, and we
have investigated variations of the proof, for example, by switching
from constant domain semantics to varying domain semantics.

In this paper we focus on the core aspect of our work related to AI: proof automation with HOL-ATPs (cf.~aspects (ii)--(iv)
above). % In addition, we present the detailed technical
% background, i.e., the embedding of \HOML in \HOL.  None of these
% aspects have been adressed in detail yet in any of our other existing 
% (short and partly non-reviewed) papers on the subject \cite{J30,W50,J29,J28}.
The particular contributions of this paper are as follows: 
% In Sec.~\ref{sec1} we present Scott's script \cite{ScottNotes} of
% G\"odel's proof, and we summarize the results of our
% recent automation of this script with HOL-ATPs. 
In Sec.~\ref{sec2} we present an elegant embedding of 
\HOML \cite{Gallin75,homl} in \HOL \cite{andrewsSEP,B5}. 
This background theory extends and
adapts the work as presented in \cite{J23,B9}.
In Sec.~\ref{sec3}, we present details on the
encoding of this embedding and of G{\"o}del's argument in the
concrete THF syntax \cite{J22} for \HOL, and we report on the
experiments we have conducted with HOL-ATPs.  The main findings of
these experiments are summarized in Sec.~\ref{sec4},
% : in addition to
% automating G{\"o}del's proof script we have automated also some further
% results (e.g.~modal collapse), and we have studied
% the argument under modified logical settings (varying vs.~constant domain semantics).
% (by switching from constant
% domain semantics to varying domain semantics)
Related and future work is addressed in Sec.~\ref{sec5}, and the paper
is concluded in Sec.~\ref{sec6}.
%  where we also discuss potential
% implications towards a \emph{computational theoretical
%   philosophy}.
None of the above aspects have been addressed (at least
not in depth) in any of our other existing (short and partly
non-reviewed) publications on the subject \cite{J30,W50,J29,J28}.

\section{THEORY FRAMEWORK}\label{sec2}
An embedding of quantified modal logic (with first-order and
propositional quantifiers) in HOL has been presented in
\cite{J23}. The theory below extends this work: quantifiers for all
types are now supported, and nested uninterpreted predicate and
function symbols of arbitrary types
are allowed as opposed to allowing top-level uninterpreted
predicate symbols over individual variables only.

\subsection{Higher-order modal logic}
A notion of \HOML is introduced that extends \HOL with a modal
operator $\Box$. An appropriate notion of semantics for \HOML is
obtained by adapting Henkin semantics for \HOL (cf.~\cite{Henkin50}
and~\cite{Gallin75}). The presentation in this section is adapted
from~\cite{homl} and~\cite{andrewsSEP}.

% First-order quantification can be constant domain or varying domain.  Below we only consider the constant domain case: every possible world has the same domain. We adapt the presentation of syntax and
%  semantics of quantified modal logic from Fitting \cite{Fitting02} to
%  the higher-order case.

% Let $\IV$ be a set of first-order (individual) variables, $\PV$ a
% set of propositional variables, and $\SYM$ a set of predicate
% symbols of any arity. Like Fitting, we keep our definitions simple by not having function or constant symbols; our language has no terms other than variables.
%   While Fitting \cite{Fitting02} studies quantified monomodal logic, we
%   are interested in quantified multimodal logic. Hence, we introduce
%   multiple $\mball{r}$ operators for symbols $r$ from an index set
%   $S$.

\begin{definition}\label{homltypes} The set $\entity{T}$ of \emph{simple
    types} is freely generated from the set of basic types $\{o,
  \mu\}$ ($o$ stands for Booleans and $\mu$ for individuals) using the
  function type constructor $\ar$. We may avoid parentheses, and
  $\alpha\ar\alpha\ar\alpha$ then stands for
  $(\alpha\ar(\alpha\ar\alpha))$, that is, function types associate to
  the right.
\end{definition}

\begin{definition}\label{homlgrammar}
The \emph{grammar} for \HOML is:
\begin{align*} 
  s,t \quad ::= \quad & p_\alpha \mid X_\alpha \mid (\lam{X_\alpha}
  s_\beta)_{\alpha\ar\beta} \mid (s_{\alpha\ar\beta}\, t_\alpha)_\beta
  \mid (\neg_{o\ar o}\, s_o)_o \mid \\
  & ((\vee_{o\ar o\ar o} s_o)\, t_o)_o  
  \mid (\forall_{(\alpha\ar
    o)\ar o}(\lam{X_\alpha} s_o))_o  \mid (\Box_{o\ar o}\, s_o)_o
%P \mid k(X^1,\ldots,X^n) \mid \mnot s \mid s \mor t \mid \all{X} s  \mid \all{P} s  \mid \mball{r} s
\end{align*}
where $\alpha,\beta\in T$.  $p_\alpha$ denotes typed constants and
$X_\alpha$ typed variables (distinct from $p_\alpha$).  Complex typed
terms are constructed via abstraction and application. The type of
each term is given as a subscript.  Terms $s_o$ of type $o$ are called
formulas.  The logical connectives of choice are
$\neg_{o\ar o}$, $\vee_{o\ar
  o\ar o}$, $\forall_{(\alpha\ar
  o)\ar o}$ (for $\alpha\in T$), and
$\Box_{o\ar o}$.  Type subscripts may be dropped if
irrelevant or obvious. Similarly, parentheses may be avoided.  Binder
notation $\all{X_\alpha} s_o$ is used as shorthand for
$\forall_{(\alpha\ar o)\ar o}
(\lam{X_\alpha} s_o)$, and infix notation $s \vee t$ is employed
instead of $((\vee s)\, t)$. From the above connectives, other logical
connectives, such as $\top$, $\bot$, $\wedge$, $\supset$, $\equiv$, $\exists$,
and $\Diamond$, can be defined in the usual way.
\end{definition}

\begin{definition}\label{homlsubstitution}
  \emph{Substitution} of a term $A_\alpha$ for a variable $X_\alpha$
  in a term $B_\beta$ is denoted by $[A/X]B$.  Since we consider
  $\alpha$-conversion implicitly, we assume the bound variables of $B$
  avoid variable capture.
\end{definition}

\begin{definition}\label{homlbetaeta} %\footnote{\textbf{Chris:} Is Def.4 needed at all?}
  Two common relations on terms are given by $\beta$-reduction and
  $\eta$-reduction.  A $\beta$-redex has the form $(\lam{X}s)t$ and
  $\beta$-reduces to $[t/X]s$.  An $\eta$-redex has the form
  $(\lam{X}s X)$ where variable $X$ is not free in $s$; it
  $\eta$-reduces to $s$.  We write $s\eqb t$ to mean $s$ can be
  converted to $t$ by a series of $\beta$-reductions and expansions.
  Similarly, $s\eqbe t$ means $s$ can be converted to $t$ using both
  $\beta$ and $\eta$.  For each $s_\alpha\in \HOML$ there is a unique
  \emph{$\beta$-normal form} and a unique \emph{$\beta\eta$-normal
    form}.
% From this fact we know $s\eqb t$ ($s\eqbe t$) iff
% $\Bnormform s\Metaeq\Bnormform t$
% ($\Benormform s\Metaeq\Benormform t$).
% The semantics of $\stt$ is well understood and thoroughly documented
% in the literature \cite{Henkin50,Andrews72b,Andrews72a,BBK04}; our
% summary below is adapted from Andrews~\cite{sep-type-theory-church}.
\end{definition}

\begin{definition}\label{homlframe}
A \emph{frame} $D$ is a collection $\{D_\alpha\}_{\alpha\in\entity{T}}$
of nonempty sets $D_\alpha$, such that $D_o = \{T,F\}$
(for truth and falsehood).  The
$D_{\alpha\ar\beta}$ are collections of functions mapping
$D_\alpha$ into $D_\beta$. 
\end{definition}

\begin{definition}\label{homlassignment}
A \emph{variable assignment} $g$ maps
variables $X_\alpha$ to elements in $D_\alpha$. $g[d/W]$ denotes the
assignment that is identical to $g$, except for variable $W$, which is
now mapped to $d$.
\end{definition}

\begin{definition}\label{homlmodel}
  A \emph{model} for \HOML is a quadruple $M=\langle W, R, D,
  \{I_w\}_{w\in W} \rangle$, where $W$ is a set of worlds (or states),
  $R$ is an accessibility relation between the worlds in $W$, $D$ is a
  frame, and for each $w\in W$, $\{I_w\}_{w\in W}$ is a family of
  typed interpretation functions mapping constant symbols $p_\alpha$
  to appropriate elements of $D_\alpha$, called the \emph{denotation
    of $p_\alpha$ in world $w$} (the logical connectives $\neg$,
  $\vee$, $\forall$, and $\Box$ are always given the standard
  denotations, see below).  Moreover, it is assumed that the domains
  $D_{\alpha\ar\alpha\ar o}$ contain the respective identity relations
  on objects of type $\alpha$ (to overcome the extensionality
  issue discussed in \cite{Andrews:gmae72}).
\end{definition}

% \emph{Validity} of a formula $s$ for a model $M=\langle  W, R, \{D_\alpha\}_{\alpha\in\entity{T}}, \{I_\alpha\}_{\alpha\in\entity{T}} \rangle$,
% a world $w\in W$, and a variable assignment $g$ is denoted as $M,g,w\models s$ and defined as
% follows, where $[a/Z]g$ denotes the assignment identical to $g$ except that 
% $([a/Z]g)(Z) = a$:
% \begin{align*}
% M,g,w &\models k(X^1,\ldots,X^n) &\text{if and only if}& \quad \langle g^{iv}(X^1),\ldots,g^{iv}(X^n)\rangle\in I_w(k) \\
% M,g,w &\models P &\text{if and only if}& \quad  w\in g^{pv}(P) \\
% M,g,w &\models \mnot s &\text{if and only if}& \quad  M,g,w \not\models s \\
% M,g,w &\models s \mor t &\text{if and only if}& \quad  M,g,w \models s \text{ or } M,g,w \models t \\
% M,g,w &\models \mall{X}s &\text{if and only if}& \quad  M,([d/X]g^{iv},g^{pv}),w \models s \\ 
% & & & \quad \text{for all } d\in D \\
% M,g,w &\models \mall{Q}s &\text{if and only if}& \quad  M,(g^{iv},[p/Q]g^{pv}),w \models s \\ 
% & & & \quad \text{for all } p\in P \\
% M,g,w &\models \mball{r}s &\text{if and only if}& \quad  M,g,v \models s \text{ for all } v\in W \\
% & & & \quad \text{with } \langle w,v\rangle\in R_r
% \end{align*}

\begin{definition}\label{homlvalue}
The \emph{value} $\| s_\alpha\|^{M,g,w}$ of a \HOML term $s_\alpha$ on a model $M=\langle W, R, D, \{I_w\}_{w\in W}
\rangle$ in a world $w\in W$ under variable assignment $g$ is an element $d\in D_\alpha$ defined in the following way:
\begin{enumerate}
\item $\|p_\alpha\|^{M,g,w} = I_w(p_\alpha)$ and $\|X_\alpha\|^{M,g,w} = g(X_\alpha)$
\item $\|(s_{\alpha\ar\beta}\, t_\alpha)_\beta\|^{M,g,w} = \|s_{\alpha\ar\beta}\|^{M,g,w}(\|t_\alpha\|^{M,g,w})$
\item $\|(\lam{X_\alpha}s_\beta)_{\alpha\ar\beta}\|^{M,g,w} = $ the
  function $f$ from $D_\alpha$ to $D_\beta$ such that $f(d) = \|s_\beta\|^{M,g[d/X_\alpha],w}$ for all $d\in D_\alpha$
\item $\|(\neg_{o\ar o}\, s_o)_o\|^{M,g,w} = T$ iff $\|s_o\|^{M,g,w} = F$
\item $\|(({\vee_{o\ar o\ar o}}\,s_o)\,t_o)_o\|^{M,g,w} = T$ iff \,$\|s_o\|^{M,g,w} = T$ or $\|t_o\|^{M,g,w} = T$
\item $\|(\forall_{(\alpha\ar o)\ar o}(\lam{X_\alpha} s_o))_o\|^{M,g,w} = T$ iff for all $d\in D_\alpha$ we have  $\|s_o\|^{M,g[d/X_\alpha],w}  = T$
\item $\|(\Box_{o\ar o}\, s_o)_o\|^{M,g,w} = T$ iff for all $v \in W$ with $w R v$ we have  $\|s_o\|^{M,g,v}  = T$
\end{enumerate}
\end{definition}

\begin{definition}\label{homlhenkinmodel}
 A model $M=\langle W, R, D, \{I_w\}_{w\in W}\rangle$ is called a
 \emph{standard model} iff for all $\alpha,\beta\in T$ we have
 $D_{\alpha\ar\beta} = \{ f \mid f : D_\alpha \longrightarrow D_\beta
 \}$. In a \emph{Henkin model} function spaces are not necessarily
 full. Instead it is only required that $D_{\alpha\ar\beta}
 \subseteq \{ f \mid f : D_\alpha \longrightarrow D_\beta \}$ (for all
 $\alpha,\beta\in T$) and that the valuation function 
 $\|\cdot\|^{M,g,w}$ from above is total (i.e., every term denotes). Any standard model is obviously
 also a Henkin model. We consider Henkin models in the remainder.
\end{definition}

\begin{definition}\label{homlvalid}
A formula $s_o$ is \emph{true} in model $M$ for world $w$
under assignment $g$ iff $\|s_o\|^{M,g,w} = T$; this is also denoted
as $M,g,w \models s_o$.  A formula $s_o$ is called \emph{valid} in
$M$ iff $M,g,w \models s_o$ for all $w\in W$ and all
assignments $g$. Finally, a formula $s_o$ is called
\emph{valid}, which we denote by $\models s_o$, iff $s_o$ is valid for
all $M$. Moreover, we write $\Gamma\models \Delta$ (for sets of
 formulas $\Gamma$ and $\Delta$) iff there is a model $M=\langle W, R, D, \{I_w\}_{w\in W}
\rangle$, an assignment $g$, and a world $w\in W$, such that  
 $M,g,w \models s_o$ for all $s_o\in \Gamma$ and $M,g,w \models t_o$ for at least one $t_o\in \Delta$.
\end{definition}

The above definitions introduce higher-order modal logic K. In order
to obtain logics KB and S5 respective conditions on accessibility
relation $R$ are postulated: $R$ is a symmetric relation in logic KB,
and it is an equivalence relation in logic S5. If these restriction apply,
we use the notations $\models^{KB}$ and $\models^{S5}$.  G{\"o}del's
argument has been developed and studied in the context of logic
S5 (and logic S5 has subsequently been criticized). However, the
HOL-ATPs discovered (cf.~Sec.~\ref{sec4}) that logic KB
is sufficient.

An important issue for quantified modal logics is whether constant
domain or varying domain semantics is considered. The theory above
introduces constant domains. 
Terms (other than those of Boolean type) are modeled as rigid, that
is, their denotation is fixed for all worlds. 
%
An adaptation to varying or cumulative
domains is straightforward (cf.~\cite{fitting98}). Moreover, non-rigid
terms could be modeled; that is, terms whose denotation may switch
from world to world. 
%
The respective assumptions of 
G\"odel are not obvious to us.

\subsection{Classical higher-order logic}
\HOL is easily obtained from \HOML by removing the modal operator
$\Box$ from the grammar, and by dropping the set of possible worlds
$W$ and the accessibility relation $R$ from the definition of a model.
Nevertheless, we explicitly state the most relevant definitions for
the particular notion of \HOL as employed in this paper. One reason is
that we do want to carefully distinguish the \HOL and \HOML languages
in the remainder (we use boldface fonts for \HOL and standard fonts
for \HOML). There is also a subtle, but harmless, difference in the
\HOL language as employed here  in comparison to the standard
presentation: here three base types are employed, whereas
usually only two base types are considered. The third base type plays
a crucial role in our embedding of \HOML in \HOL.

\begin{definition}\label{holtypes}
  The set $\hol{T}$ of simple types freely generated from a set of
  basic types $\{\hol{o}, \hol{\mu}, \hol{\iota}\}$ using the function type
  constructor $\hol{\ar}$. $\hol{o}$ is the type of Booleans,
  $\hol{\mu}$ is the type of individuals, and type
  $\hol{\iota}$ is employed as the type of possible worlds below. As
  before we may avoid parentheses.
\end{definition}

\begin{definition}\label{holgrammar}
The grammar for higher-order logic $\HOL$ is:
\begin{align*} 
  \hol{s},\hol{t} \quad ::= \quad & \hol{p_\alpha} \mid \hol{X_\alpha} \mid \hol{(\lam{X_\alpha}
  s_\beta)_{\alpha\ar\beta}} \mid \hol{(s_{\alpha\ar\beta}\, t_\alpha)_\beta} \mid  \hol{\neg_{o\ar o}\, s_o} \mid \\
  & \hol{(({\vee_{o\ar o\ar o}}\,s_o)\,t_o)} \mid \hol{\forall_{(\alpha\ar
    o)\ar o}(\lam{X_\alpha} s_o)} 
\end{align*}
where $\hol{\alpha},\hol{\beta}\in \hol{T}$. The text from
Def.~\ref{homlgrammar} analogously applies, except that we do not
consider the modal  connectives $\hol{\Box}$ and $\hol{\Diamond}$.
\end{definition}

The definitions for substitution (Def.~\ref{homlsubstitution}), $\beta$- and $\eta$-reduction (Def.~\ref{homlbetaeta}), frame (Def.~\ref{homlframe}), and assignment (Def.~\ref{homlassignment}) remain unchanged.

\begin{definition}\label{holmodel}
  A \emph{model} for \HOL is a tuple $\hol{M}=\hol{\langle D, I \rangle}$, where
  $\hol{D}$ is a frame, and $\hol{I}$ is a family of typed interpretation
  functions mapping constant symbols $\hol{p_\alpha}$ to appropriate
  elements of $\hol{D_\alpha}$, called the \emph{denotation of $\hol{p_\alpha}$}
  (the logical connectives $\hol{\neg}$, $\hol{\vee}$, and $\hol{\forall}$ are always
  given the standard denotations, see below).  Moreover, we assume that the domains
  $\hol{D_{\alpha\ar\alpha\ar o}}$ contain the respective identity relations.
\end{definition}

\begin{definition}\label{holvalue}
  The \emph{value} $\hol{\| s_\alpha\|^{M,g}}$ of a \HOML term
  $\hol{s_\alpha}$ on a model $\hol{M}=\hol{\langle W, R, D, I_{w\in
      W} \rangle}$ in a world $\hol{w}\in \hol{W}$ under assignment $\hol{g}$ is an element $\hol{d}\in \hol{D_\alpha}$
  defined in the following way:
\begin{enumerate}
\item $\hol{\|p_\alpha\|^{M,g}} = \hol{I(p_\alpha)}$ and $\hol{\|X_\alpha\|^{M,g}} = \hol{g(X_\alpha)}$
\item $\hol{\|(s_{\alpha\ar\beta}\, t_\alpha)_\beta\|^{M,g}} = \hol{\|s_{\alpha\ar\beta}\|^{M,g}(\|t_\alpha\|^{M,g})}$
\item $\hol{\|(\lam{X_\alpha}s_\beta)_{\alpha\ar\beta}\|^{M,g}} = $
  the function $\hol{f}$ from $\hol{D_\alpha}$ to $\hol{D_\beta}$ such
  that $\hol{f(d)} = \hol{\|s_\beta\|^{M,g[d/X_\alpha]}}$ for all
  $\hol{d}\in \hol{D_\alpha}$
\item $\hol{\|(\neg_{o\ar o}\, s_o)_o\|^{M,g}} = \hol{T}$ iff $\hol{\|s_o\|^{M,g}} = \hol{F}$
\item $\hol{\|(({\vee_{o\ar o\ar o}}\,s_o)\,t_o)_o\|^{M,g}} =
  \hol{T}$ iff\, $\hol{\|s_o\|^{M,g}} = \hol{T}$ or $\hol{\|t_o\|^{M,g}}
  = \hol{T}$
\item $\hol{\|(\forall_{(\alpha\ar o)\ar o}(\lam{X_\alpha}
    s_o))_o\|^{M,g}} = \hol{T}$ iff for all $\hol{d}\in
  \hol{D_\alpha}$ we have $\hol{\|s_o\|^{M,g[d/X_\alpha]}} = \hol{T}$
%\item $\|(\Box_{o\ar o}\, s_o)_o\|^{M,g,w} = T$ iff for all $v \in W$ with $w R v$ we have  $\|s_o\|^{M,g,v}  = T$
\end{enumerate}
\end{definition}

The definition for standard and Henkin models
(Def.~\ref{homlhenkinmodel}), and for truth in a model, validity,
etc. (Def.~\ref{homlvalid}) are adapted in the obvious way, and we use
the notation $\hol{M,g \models s_o}$, $\hol{\models s_o}$, and
$\hol{\Gamma \models \Delta}$. As for \HOML, we assume Henkin
semantics in the remainder.

\subsection{\HOML as a fragment of \HOL}
The encoding of \HOML in \HOL is simple: we identify \HOML formulas of
type $o$ with certain \HOL formulas of type $\hol{\iota \ar o}$. The \HOL
type $\hol{\iota \ar o}$ is abbreviated as $\hol{\sigma}$ in the
remainder. More generally, we define for each \HOML type $\alpha\in T$
the associated raised \HOL type $\lift{\alpha}$ as follows: $\lift{\mu}
= \hol{\mu}$, $\lift{o} = \hol{\sigma} = \hol{\iota \ar o}$, and
$\lift{\alpha\ar\beta} = \lift{\alpha} \hol{\ar} \lift{\beta}$.
Hence, all \HOML terms are rigid, except for those of type $o$.
% \footnote{This choice could be modified in an obvious way to
  % obtain non-rigid terms.}

\begin{definition}\label{operators}
\HOML terms $s_\alpha$ are associated with type-raised \HOL terms
$\lift{s_\alpha}$ in the following way:
\begin{align*} {\lift{p_\alpha}} & = \hol{p}_{\lift{\alpha}}
  \\
  {\lift{X_\alpha}} & = \hol{X}_{\lift{\alpha}}
  \\
  {\lift{(s_{\alpha\ar\beta}\,t_{\alpha})}} & =
  {(\lift{s_{\alpha\ar\beta}}\,\lift{ t_{\alpha}})}
  \\
  {\lift{(\lam{X_\alpha} s_\beta)}} & = {(\hol{\lambda}
    \lift{X_\alpha} \lambdot \lift{s_\beta})}
  \\
  {\lift{(\neg_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\neg}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
  \\
  {\lift{(({\vee}_{o\ar o\ar o}\,s_{o})\,t_{o})}} & =
  ((\hol{\llift{\vee}_{\sigma\ar\sigma\ar\sigma}}\,\lift{
    s_{\alpha}})\,\lift{ t_{\alpha}})
  \\
  {\lift{((\forall_{(\alpha\ar o)\ar o}\,(\lam{X_\alpha} s_\beta)}} & =
  (\hol{\llift{\forall}_{(\alpha\ar\sigma)\ar\sigma}}\,(\hol{\lambda}
  \lift{X_\alpha} \lambdot \lift{s_\beta})
  \\
  {\lift{(\Box_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\Box}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
\end{align*}

$\hol{\llift{\neg}},\hol{\llift{\vee}},\hol{\llift{\forall}}$, and
$\hol{\llift{\Box}}$ are the \emph{type-raised modal \HOL connectives}
associated with the corresponding modal \HOML connectives. They are defined as follows (where $\hol{r_{\iota\ar
    \iota\ar o}}$ is a new constant symbol in \HOL associated with the
accessibility relation $R$ of \HOML):
\begin{align*}
\hol{\llift{\neg}_{\sigma\ar\sigma}} & = \hol{\lam{s_\sigma}\lam{W_\iota}\neg\,(s\,W)}
\\
\hol{\llift{\vee}_{\sigma\ar\sigma\ar\sigma}} & = \hol{\lam{s_\sigma} \lam{t_\sigma} \lam{W_\iota} s\,W \vee t\,W}
\\
\hol{\llift{\forall}_{(\alpha\ar \sigma)\ar
\sigma}} & = \hol{\lam{s_{\alpha\ar \sigma}} \lam{W_\iota}
\all{X_\alpha} s\,X\,W} 
\\
\hol{\llift{\Box}_{\sigma\ar\sigma}} & =  
\hol{\lam{s_\sigma} \lam{W_{\iota}} \all{V_{\iota}} \neg (r_{\iota\ar \iota\ar o}\,W\,V) \vee s\,V}
\end{align*}
As before, we write $\hol{\llift{\forall} {X_\alpha} \lambdot
  s_\sigma}$ as shorthand for $\hol{\llift{\forall}_{(\alpha\ar
    \sigma)\ar \sigma}(\lam{X_\alpha} s_\sigma)}$. Further operators,
such as $\hol{\llift{\top}}$, $\hol{\llift{\bot}}$,
$\hol{\llift{\wedge}}$, $\hol{\llift{\supset}}$, $\hol{\llift{\equiv}}$,
$\hol{\llift{\Diamond}}$, and $\hol{\llift{\exists}}$
($\hol{\llift{\exists} {X_\alpha} \lambdot s_\sigma}$ is used as
shorthand for $\hol{\llift{\exists}_{(\alpha\ar \sigma)\ar \sigma}
  (\lam{X_\alpha} s_\sigma)}$) can now be easily defined.\footnote{We
  could introduce further modal operators, such as the difference
  modality $\hol{D}$, the global modality $\hol{E}$, nominals with $\hol{!}$, and the $\hol{@}$
  operator (cf.~\cite{J23}).} The above equations can be treated as
abbreviations in \HOL theorem provers. Alternatively, they can be
 stated as axioms where $=$ is either Leibniz equality or
primitive equality (if additionally provided in the \HOL grammar, as
is the case for most modern \HOL provers).
\end{definition}
% \begin{align*}
% D_{(\iota\ar o)\ar(\iota\ar o)} & = \lam{s_{\iota\ar o}} \lam{W_\iota} \exi{V_\iota} W \not= V \wedge s V 
% \\
% E_{(\iota\ar o)\ar(\iota\ar o)} & = \lam{s_{\iota\ar o}} s \mor D\, s 
% \\
% !_{(\iota\ar o)\ar(\iota\ar o)} & = \lam{s_{\iota\ar o}} E\, (s \mand \mnotinpit{D\,s}) 
% \\
% @_{\iota\ar(\iota\ar o)\ar(\iota\ar o)} & = \lam{W_\iota} \lam{s_{\iota\ar o}} s\, W 
% \end{align*}
% % This illustrates the potential of our embedding for encoding quantified
% hybrid logic, an issue that we might explore in future work.

%  In
% particular, the \HOML connectives $\neg_{o\ar o}$, $\vee_{o\ar o\ar
%   o}$, $\Pi_{(\alpha\ar o)\ar o}$, and $\Box_{o\ar o}$ are associated
% with ``type-raised'' \HOL terms of
% $\hol{\lift{\neg}_{\sigma\ar\sigma}}$,
% $\hol{\lift{\vee}_{\sigma\ar\sigma\ar\sigma}}$,
% $\hol{\lift{\Pi}_{(\alpha\ar \sigma)\ar \sigma}}$, and
% $\hol{\lift{\Box}_{\sigma\ar\sigma}}$, respectively.

As a consequence of the above embedding we can express \HOML proof
problems elegantly in the type-raised syntax of \HOL. Using rewriting
or definition expanding, we can reduce these representations to
corresponding statements containing only the basic \HOL connectives
$\hol{\neg_{o\ar o}}$, $\hol{\vee_{o\ar o\ar o}}$, and
$\hol{\forall_{(\alpha\ar o)\ar o}}$.

\begin{example}\label{ex1} %\footnote{Maybe we should instead use an axiom or theorem from G{\"o}del's proof?}
  The \HOML formula $\Box\,\exi{P_{\mu \ar o}} P\,a_\mu$ is
  associated with the type raised \HOL formula
  $\hol{\llift{\Box}\,\llift{\exists} {P_{\mu\ar
        \sigma}} \lambdot P\ a_\mu}$, which rewrites into the following
  $\beta\eta$-normal \HOL term of type $\sigma$
\[\hol{\lam{W_\iota} \all{V_\iota} \neg (r\,W\,V) \vee \neg \all{P_{\mu\ar \sigma}} \neg (P\, a_\mu\, V)}\]
\end{example}

Next, we define validity of type-raised modal \HOL propositions 
$\hol{s_\sigma}$ in the obvious way: 
$\hol{s_\sigma}$ is valid iff for all possible worlds
$\hol{w_\iota}$ we have 
$\hol{w_\iota\in s_\sigma}$,
that is, iff
$\hol{(s_\sigma\,w_\iota)}$ holds.

\begin{definition}\label{validity}
  \emph{Validity} is modeled as an abbreviation for the following
  $\lambda$-term: $\textbf{valid} = \hol{\lam{s_{\iota\ar o}} \all{W_\iota}
    s\,W}$ (alternatively, we could define validity simply as
  $\hol{\forall_{(\iota\ar o)\ar o}}$). Instead of 
  $\hol{\textbf{valid}\,s_\sigma}$ we also use the notation
  $\hol{[s_\sigma]}$.
\end{definition}

\begin{example} \label{ex2} \sloppy We analyze whether the type-raised
  modal \HOL formula $\hol{\llift{\Box}\,\llift{\exists} {P_{\mu\ar
        \sigma}} \lambdot (P\ a_\mu)}$ is valid or not. For this, we
  formalize the \HOL proof problem
  $\hol{[{\llift{\Box}\,\llift{\exists} {P_{\mu-\ar \sigma}} \lambdot (P\
      a_\mu)}]}$, which expands into $\hol{\all{W_\iota} \all{V_\iota} \neg
    (r\,W\,V) \vee \neg \all{P_{\mu\ar \sigma}} \neg (P\, a_\mu\, V)}$.
  It is easy to check that this term is valid in Henkin semantics: put
  $\hol{P} = \hol{\lam{X_\mu} \lam{Y_\iota} \top}$.
%  Proving the statement automatically is non-trivial though as we shall shortly see.
\end{example}

\begin{theorem}[Soundness and Completeness] \label{thm1} 
For all \HOML formulas $s_o$ we have: 
\[\models s_o \quad \text{iff} \quad \hol{\models} \hol{[}\lceil s_o \rceil\hol{]}\]
\textbf{Proof sketch:} The proof adapts the ideas presented
in \cite{J23}. By contraposition it is sufficient to show $\not\models
s_o \,\,\text{iff}\,\hol{\not\models} \hol{[}\lceil s_o \rceil\hol{]}$,
that is, $\| s_o \|^{M,g,w}$ (for some \HOML model $M$, assignment
$g$, and $w$) iff $\hol{\| \all{W_\iota} \lceil s_o \rceil\,W\|^{M,g}}$
(for some \HOL model $\hol{M}$ and assignment $\hol{g}$) iff $\hol{\|
  \lceil s_o \rceil\,W\|^{M,g[w/W]}}$ (for some $\hol{M}$, $\hol{g}$,
and $w$). We easily get the proof by choosing the obvious
correspondences between $D$ and $\hol{D}$, $W$ and $\hol{D_\iota}$, $I$
and $\hol{I}$, $g$ and $\hol{g}$, $R$ and $r_{\iota\ar \iota\ar o}$,  and $w$
and $\hol{w}$. \hfill $\Box$
\end{theorem}

From Theorem~\ref{thm1} we get the following corollaries:
\[\models^{KB} s_o  \quad \text{iff}  \quad \hol{(\textbf{symmetric}\,r_{\iota\ar \iota\ar o})} \hol{\models} \hol{[}\lceil s_o \rceil\hol{]}\]
\[\models^{S5} s_o  \quad \text{iff}  \quad
\hol{(\textbf{equiv-rel}\,r_{\iota\ar \iota\ar o})} \hol{\models}
\hol{[}\lceil s_o \rceil\hol{]}\]
where \textbf{symmetric} and \textbf{equiv-rel} are defined in an
obvious way.

Constant domain quantification is
addressed above. Techniques for handling varying domain and cumulative domain
quantification in the embedding of first-order modal logics in \HOL
have been outlined in \cite{C34}. These techniques, which have also been
adapted  for the theory above, cannot be presented here for space limitations.

Note that also non-rigid terms can easily be modeled by
type-raising. For example, a non-rigid \HOML constant symbol
$\mathrm{kingOfFrance}_\mu$ would be mapped to a type-raised
(and thus world-depended) \HOL constant symbol $\mathbf{kingOfFrance_{\iota\ar \mu}}$.

\section{EXPERIMENTS} \label{sec3} The above embedding 
%from Sec.~\ref{sec2} 
has been encoded in the concrete THF0 syntax \cite{J22} for \HOL; cf.~the
files \texttt{Quantified\_K/\_KB/\_S5.ax}\footnote{The formalization in
  these files slightly varies from the above theory w.r.t. technical
  details. For example, a generic $\Box$-operator is introduced that
  can be instantiated for different accessibility relations as
  e.g. required for multi-modal logic applications
  (cf.~\cite{J23}). Moreover, since THF0 does not support
  polymorphism, copies of the $\hol{\llift{\forall}_{(\alpha\ar
      \sigma)\ar \sigma}}$ and $\hol{\llift{\exists}_{(\alpha\ar
      \sigma)\ar \sigma}}$ connectives are provided only for the
  selected types ${(\mu \ar \sigma)\ar \sigma}$ and ${((\mu \ar
    \sigma)\ar \sigma)\ar \sigma}$ as precisely required in
  G{\"o}dels's proof. The Isabelle version~\cite{J28} and the Coq
  version of the encoding instead provide respective polymorphic
  definitions.} available at
\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/THF}
(all files mentioned below are provided under this URL). The
definition for quantifier
$\hol{\llift{\forall}_{((\mu \ar \sigma)\ar \sigma)\ar \sigma}}$, for
example, is given as\footnote{\texttt{\$i}, \texttt{\$o}, and
  \texttt{mu} represent the HOL base types $\hol{i}$, $\hol{o}$, and
  $\hol{\mu}$. \texttt{\$i>\$o} encodes a function (predicate) type.
  Function application is represented by \texttt{@}, and for universal
  quantification, existential quantification and $\lambda$-abstraction
  the symbols \texttt{!}, \texttt{?}  and \texttt{\^{}} are
  employed. $\hol{\neg}$, $\hol{\vee}$, $\hol{\wedge}$, and
  $\hol{\supset}$ are written as \texttt{\url{~}}, \texttt{|},
  \texttt{\&}, and \texttt{=>}, respectively. The type-raised modal
  connectives are called \texttt{mforall\_*}, \texttt{mexists\_*},
  \texttt{mnot}, \texttt{mor}, \texttt{mand}, \texttt{mimplies}, etc.}

{\footnotesize
\begin{verbatim}
  thf(mforall_indset,definition,
      ( mforall_indset
      = ( ^ [S: ( mu > $i > $o ) > $i > $o,W: $i] :
          ! [X: mu > $i > $o] :
            ( S @ X @ W ) ) )).
\end{verbatim}
}

Subsequently the axioms, definitions, and theorems from
Fig.~\ref{fig1} and some further, related problems have been encoded
in THF0. Then the THF0 compliant HOL-ATPs LEO-II \cite{LEO-II},
Satallax \cite{Satallax}, and Nitpick \cite{Nitpick} have been
employed to automate the proof problems. LEO-II, which internally
cooperates with the first-order prover E~\cite{Schulz:AICOM-2002}, was
used exclusively in the initial round of experiments, that is, it was the
first prover to automate G{\"o}del's ontological argument.

%\subsection{Automation of Scott's version of G{\"o}del's proof}
Theorem T1 from Fig.~\ref{fig1}, for example, is formalized as

{\footnotesize
\begin{verbatim}
  thf(thmT1,conjecture,
     ( v
      @ ( mforall_indset
        @ ^ [Phi: mu > $i > $o] :
            ( mimplies @ ( p @ Phi )
            @ ( mdia
              @ ( mexists_ind
                @ ^ [X: mu] :
                    ( Phi @ X ) ) ) ) ) )).
\end{verbatim}
}

\noindent This encodes the \HOL formula
$$\hol{[\llift{\forall}\phi_{\mu\ar \sigma} \lambdot}\hol{p_{(\mu\ar
    \sigma)\ar \sigma}\phi\,\llift{\supset}\, \llift{\Diamond}
  \llift{\exists} X_\mu \lambdot \phi X]}$$ (\texttt{v} in the THF0
encoding stands for \textbf{valid} and $p$ corresponds to the
uppercase $P$, for `positive', from Fig.~\ref{fig1}).  The
respective encodings and the results of a series of recent experiments with
LEO-II (version 1.6.2), Satallax (version 2.7), and Nitpick (version
2013) are provided in Fig.~\ref{fig2}.  The first row marked with T1,
for example, shows that this theorem follows from axioms A2 and A1
(where only the $\supset$-direction is needed); LEO-II and Satallax
confirm this in 0.1 second.  The experiments have been carried
out w.r.t.~the logics K and/or KB, and w.r.t.~constant (const) and
varying (vary) domain semantics for the domains of individuals. The
exact dependencies (available axioms and definitions) are displayed for each
single problem. The results of the prover calls are given in
seconds. `---' means timeout. `THM', `CSA', `SAT', and `UNS'
are the reported result statuses; they stand for `Theorem',
`CounterSatisfiable', `Satisfiable', and `Unsatisfiable',
respectively.  {The experiments can be easily reproduced: all relevant
  files have been provided at the above URL.  For example, the two
  THF0 problem files associated with the first table row for T1 are
  \texttt{T1\_K\_const\_min.p} and \texttt{T1\_K\_vary\_min.p}, and
  those associated with the second row for T1 are
  \texttt{T1\_K\_const\_max.p} and
  \texttt{T1\_K\_vary\_max.p}, respectively. Moreover, a simple shell script
  \texttt{call\_tptp.sh} is provided, which can be used to make remote
  calls to LEO-II, Satallax, and Nitpick installed at Sutcliffe's
  SystemOnTPTP infrastructure \cite{sutcliffe2009tptp} at the
  University of Miami. The experiments used standard 2.80GHz
  computers with 1GB memory remotely located in Miami. 
\begin{figure*}[t] 
\noindent \framebox[\textwidth][r]{
\begin{minipage}{.99\textwidth}
\small
$\begin{array}{llllllll}
   & \text{\HOL encoding} & \text{dependencies} & \text{logic} & \text{status} & \text{LEO-II} & \text{Satallax} & \text{Nitpick} \\  
   &                      &                     &              &               & \text{const/vary} & \text{const/vary} & \text{const/vary} 
\\
  \text{A1} &  {\hol{[\llift{\forall}\phi_{\mu\ar \sigma} \lambdot p_{(\mu\ar \sigma)\ar \sigma} (\lambda X_\mu \lambdot \llift{\neg} (\phi X)) \, \llift{\equiv}\,  \llift{\neg} (p \phi) ]}} \\
  % \text{A2} &  \allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x) \imp \psi(x)]) \imp P(\psi)] \\
  \text{A2} &  \multicolumn{4}{l}{\hol{[\llift{\forall}\phi_{\mu\ar \sigma}\lambdot \llift{\forall}\psi_{\mu\ar \sigma}\lambdot (p_{(\mu\ar \sigma)\ar \sigma} \phi \,\llift{\wedge}\, \llift{\Box} \llift{\forall}X_\mu\lambdot (\phi X \, \llift{\supset}\, \psi X)) \, \llift{\supset}\, p \psi]}}\\
  \text{T1} & \hol{[\llift{\forall}\phi_{\mu\ar \sigma} \lambdot p_{(\mu\ar \sigma)\ar \sigma}\phi\, \llift{\supset}\, \llift{\Diamond} \llift{\exists} X_\mu \lambdot \phi X]} & \text{A1}(\supset), \text{A2} & \text{K} & \text{THM} 
  & 0.1/0.1 & 0.0/0.0 & \text{---/---}\\
  & & \text{A1}, \text{A2} & \text{K} & \text{THM} 
  & 0.1/0.1 & 0.0/5.2 & \text{---/---}\\
%\text{D1} & G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)] \\
\text{D1} & {\hol{g_{\mu\ar \sigma} = \lambda X_\mu \lambdot  \llift{\forall}\phi_{\mu\ar \sigma} \lambdot p_{(\mu\ar \sigma)\ar \sigma}\phi\, \llift{\supset}\, \phi X}} \\
\text{A3} &  \hol{[p_{(\mu\ar \sigma)\ar \sigma} g_{\mu\ar \sigma}]} \\
\text{C} &  \hol{[\llift{\Diamond} \llift{\exists} X_\mu \lambdot g_{\mu\ar \sigma} X]} & \text{T1}, \text{D1}, \text{A3} & \text{K} & \text{THM} & 0.0/0.0 & 0.0/0.0 & \text{---/---}\\
 & & \text{A1}, \text{A2}, \text{D1}, \text{A3} & \text{K} & \text{THM} 
& 0.0/0.0 & 5.2/31.3 & \text{---/---}\\
%\text{A4} &  \allq \phi [P(\phi) \imp \Box \; P(\phi)] \\
\text{A4} &  \hol{[\llift{\forall}\phi_{\mu\ar \sigma} \lambdot p_{(\mu\ar \sigma)\ar \sigma}\phi\,\llift{\supset}\, \llift{\Box} p \phi]} \\
%\text{D2} & \ess{\phi}{x} \biimp \phi(x) \wedge \allq \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))   \\
\text{D2} & \multicolumn{4}{l}{\hol{\mathbf{ess}_{(\mu\ar \sigma)\ar \mu \ar \sigma} = \lambda \phi_{\mu\ar \sigma} \lambdot \lambda X_\mu \lambdot \phi X \,\llift{\wedge}\, \llift{\forall}\psi_{\mu\ar \sigma} \lambdot (\psi X \,\llift{\supset}\,\llift{\Box}   \llift{\forall} Y_\mu \lambdot (\phi Y \,\llift{\supset}\, \psi Y))}} \\
%\text{T2} &  \allq x [G(x) \imp \ess{G}{x}] \\
\text{T2} &  \hol{[\llift{\forall} X_\mu \lambdot g_{\mu\ar \sigma} X \,\llift{\supset}\, (\mathbf{ess}_{(\mu\ar \sigma)\ar \mu \ar \sigma} g X)]} & \text{A1}, \text{D1}, \text{A4}, \text{D2} & \text{K} & \text{THM} 
& 19.1/18.3 & 0.0/0.0 & \text{---/---}\\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2} & \text{K} & \text{THM} 
& 12.9/14.0 & 0.0/0.0 & \text{---/---}\\
%\text{D3} & NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec  \exq y \phi(y)] \\
\text{D3} & \multicolumn{4}{l}{\hol{\mathbf{NE}_{\mu\ar \sigma} = \lambda X_\mu \lambdot  \llift{\forall}\phi_{\mu\ar \sigma} \lambdot (\mathbf{ess}\, \phi X \,\llift{\supset}\, \llift{\Box}  \llift{\exists} Y_\mu \lambdot \phi Y ) }} \\
\text{A5} & \hol{[p_{(\mu\ar \sigma)\ar \sigma} \mathbf{NE}_{\mu\ar \sigma}]} \\
\text{T3} & \hol{[\llift{\Box}  \llift{\exists} X_\mu \lambdot g_{\mu\ar \sigma} X]} 
  &  \text{D1}, \text{C}, \text{T2}, \text{D3}, \text{A5}  & \text{K} & \text{CSA} 
& \text{---/---} & \text{---/---} & 3.8/6.2\\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5} & \text{K} & \text{CSA} 
& \text{---/---} & \text{---/---} & 8.2/7.5\\
 & &  \text{D1}, \text{C}, \text{T2}, \text{D3}, \text{A5}  & \text{KB} & \text{THM} & 0.0/0.1 & 0.1/5.3 & \text{---/---}\\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5} & \text{KB} & \text{THM} & \text{---/---} & \text{---/---} & \text{---/---}\\ \\
\text{MC} & \hol{[s_\sigma  \,\llift{\supset}\, \llift{\Box}  s_\sigma]} 
   & \text{D2}, \text{T2}, \text{T3} & \text{KB} & \text{THM} & 17.9/\text{---} & 3.3/3.2 & \text{---/---}\\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5} & \text{KB} & \text{THM} & \text{---/---} & \text{---/---} & \text{---/---}\\
\text{FG} & \multicolumn{2}{l}{\hol{[\llift{\forall}\phi_{\mu\ar \sigma} \lambdot  \llift{\forall}X_\mu\lambdot  (g_{\mu\ar \sigma} X \,\llift{\supset}\,  (\llift{\neg} (p_{(\mu\ar \sigma)\ar \sigma} \phi) \,\llift{\supset}\, \llift{\neg} (\phi X))) ]} \quad \text{A1}, \text{D1}} & \text{KB} & \text{THM} & 16.5/\text{---} & 0.0/0.0 & \text{---/---} \\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5}  
  & \text{KB} & \text{THM} & 12.8/15.1 & 0.0/5.4 & \text{---/---} \\
\text{MT} & \hol{[\llift{\forall}X_\mu\lambdot  \llift{\forall}Y_\mu\lambdot (g_{\mu\ar \sigma} X \,\llift{\supset}\,  (g_{\mu\ar \sigma} Y \,\llift{\supset}\,  X\,\llift{=}\,Y))]} & \text{D1}, \text{FG} & \text{KB} & \text{THM} & \text{---/---}  &  0.0/3.3 & \text{---/---} \\
  & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5}  
  & \text{KB} & \text{THM} & \text{---/---}  &  \text{---/---} & \text{---/---} \\ \\
\text{CO} & \emptyset\, \text{(no goal, check for consistency)} & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5} & \text{KB} & \text{SAT} & \text{---/---} & \text{---/---} & 7.3/7.4 \\ 
% & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2}, \text{D3}, \text{A5} & \text{KB} & \text{SAT} & \text{---/---} & \text{---/---} & \text{---/---} \\ 
\text{D2'} & \multicolumn{4}{l}{\hol{\mathbf{ess}_{(\mu\ar \sigma)\ar \mu \ar \sigma} = \lambda \phi_{\mu\ar \sigma} \lambdot \lambda X_\mu \lambdot \llift{\forall}\psi_{\mu\ar \sigma} \lambdot (\psi X \,\llift{\supset}\,\llift{\Box}   \llift{\forall} Y_\mu \lambdot (\phi Y \,\llift{\supset}\, \psi Y))}} \\
\text{CO'} & \emptyset\, \text{(no goal, check for consistency)} & \text{A1}(\supset), \text{A2}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & 7.5/7.8 & \text{---/---} & \text{---/---} \\ 
 & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & \text{---/---} & \text{---/---} & \text{---/---} \\ 
\end{array}$
\end{minipage}
}
\caption{\HOL encodings and experiment results for Scott's version of G{\"o}del's ontological argument from Fig.~\ref{fig1}. \label{fig2}}
\end{figure*}

%  Figure~\ref{fig2} shows the entire results
% of this experiment. The provers were given a timeout of 60 seconds. The mentioned dependencies for T1 are encoded in
% this table in the
% third column. We also present the results for maximal sets of axioms
% (and theorems) as being accumulated along the proof script. This
% demonstrates the provers robustness and effectiveness when respective
% (ir-)relevance information is still missing and has to be 
% determined first. It was in fact the latter kind of proof problems 
% we started with in our experiments. Then the former dependency information was
% extracted from the proof objects of LEO-II.

% We have additionally encoded some further
% theorems. Moreover, we have added some variations of the embedding in
% which we e.g. consider varying domain semantics as opposed to const
% domain semantics. These encodings are available at
% \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/THF}.
% A subset of those has also been submitted to the public TPTP problem
% library at \url{www.tptp.org}, (they will be released in summer after
% the CASC-J7 competition).  Our contribution has in fact led to the set
% up of the new TPTP problem category \texttt{PHI} for philosophy
% problems.

\section{MAIN FINDINGS} \label{sec4}
Several interesting and partly novel findings have been contributed by the HOL-ATPs, including:
\begin{enumerate}
\item The axioms and definitions from Fig.~\ref{fig1} are consistent (cf.~CO in Fig.~\ref{fig2}). 
\item Logic K is sufficient for proving T1, C and T2.
\item For proving the final theorem T3, logic KB is sufficient (and
  for K a countermodel is reported). This is highly relevant since
  several philosophers have criticized G{\"o}del's argument for the
  use of logic S5. This criticism is thus provably pointless.
\item Only for T3 the HOL-ATPs still fail to produce a proof directly
  from the axioms; thus, T3 remains an interesting benchmark problem; T1, C, and T2 are rather trivial for HOL-ATPs.
\item G\"odel's original version of the proof \cite{GoedelNotes},
  which omits conjunct $\phi(x)$ in the definition of \emph{essence} (cf.~D2'),
  seems inconsistent (cf.~the failed consistency check for CO' in
  Fig.~\ref{fig2}). As far as we are aware of, this is a new result.
\item G{\"o}del's axioms imply what is called the modal collapse (cf.~MC
  in Fig.~\ref{fig2}) $\phi\supset\Box\phi$, that is, contingent truth
  implies necessary truth (which can even be interpreted as an
  argument against free will; cf.~\cite{sobel2004logic}). MC is probably the most fundamental
  criticism put forward against G{\"o}del's argument.
\item For proving T1, only the $\supset$-direction of A1 is
  needed. However, the $\subset$-direction of A1 is required for
  proving T2. Some philosophers (e.g.~\cite{anderson90:_some_emend_of_goedel_ontol_proof}) try to avoid MC by
  eluding/replacing the $\supset$-direction of A1.
\item G{\"o}del's axioms imply a `flawless God', that is, an entity
  that can only have `positive' properties (cf.~FG in
  Fig.~\ref{fig2}). However, a comment by G{\"o}del in \cite{GoedelNotes}
  explains that `positive' is to be interpreted in a moral aesthetic
  sense only.
% \footnote{In order to better understand G\"odel's notion of
%     `positive' properties, we have reformulated G{\"o}del's theory
%     and used `divine' instead of `positive'. Then we introduced
%     orthogonal predicates `positive' and `negative' and we showed that
%     God-like beings may well have positive and negative properties as
%     long as all these properties are classified as divine properties. A respective
%     formalization in Isabelle can be found at
%     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/DivineVersion/GoedelGodDivine.thy}.}
\item Another implication of G{\"o}del's axioms is monotheism (see MT
  in Fig.~\ref{fig2}). MT can easily be proved by Satallax from FG and
  D1. It remains non-trivial to prove it directly from
  G{\"o}del's axioms.
\item \label{vary} All of the above findings hold for both constant domain semantics and varying domain semantics (for the domain of individuals).
\end{enumerate}

The above findings, in particular \eqref{vary}, well illustrate that
the theory framework from Sec.~\ref{sec2} has a great potential
towards a flexible support system for \emph{computational theoretical
  philosophy}. In fact, G\"odel's ontological argument has been
verified and even automated not only for one particular setting of
logic parameters, but these logic parameters have been varied and the
validity of the argument has been reconfirmed (or falsified, cf.~D2'
and CO') for the modified setting.  Moreover, our framework is not
restricted to a particular theorem proving system, but has been
fruitfully employed with some of the most prominent automated and
interactive theorem provers available to date.

\section{RELATED AND FUTURE WORK}\label{sec5}
We are pioneering the computer-supported automation of modern versions
of the ontological argument. There are two related papers
\cite{oppenheimera11,rushby13}. Both focus on the comparably simpler
argument by Anselm. \cite{oppenheimera11} encodes (a variant) of
Anselm's argument in first-order logic and employs the theorem prover
PROVER9 in experiments; this work has been criticized in
\cite{garbacz12:_prover_simpl_expal_away}. The work in
\cite{rushby13}, which has evolved in parallel to ours, interactively
verifies Anselm's argument in the higher-order proof assistant
PVS. Note in particular, that both formalizations do not achieve the
close correspondence between the original formulations and the formal
encodings that can be found in our approach.

A particular strength of our universal logic framework is that it can
be easily adapted for logic variations and even supports flexible
combinations of logics (cf.~\cite{C36}). In ongoing and future work we
will therefore investigate further logic parameters for G{\"o}del's
argument, including varying domains at higher types and non-rigid
terms. We plan to make the entire landscape of results available to
the interested theoretical philosophy and theology communities. This
is relevant, since philosophers are sometimes imprecise about the
very details of the logics they employ.

% the main activity of our project will be the use of state-of-the-art automated deduction tools and proof assistants to represent, verify and strengthen arguments about God’s existence.

\section{CONCLUSION} \label{sec6} While computers can now calculate,
play games, translate, plan, learn and classify data much better than
we humans do, tasks involving philosophical and theological inquiries
have remained mostly untouched by our technological progress up to
now. Due to the abstract and sophisticated types of reasoning they
require, they can be considered a challenging frontier for automated
reasoning.

We accepted this challenge and decided to tackle, with automated
reasoning techniques, a philosophical problem that is almost 1000
years old: the ontological argument for God's existence, firstly
proposed by St. Anselm of Canterbury and greatly improved by
Descartes, Leibniz, G{\"o}del and many others throughout the
centuries.  % G{\"o}del's modern version of this argument can be
% formalized in a very expressive higher-order modal logic.
So far,
there was no AI system capable of dealing with
such complex problems. We created a prototypical infrastructure
extending widely used systems such as LEO-II, Satallax, and Nitpick
(and Isabelle and Coq) to allow them to cope with modalities; and
using the extended systems we were able to automatically reconstruct
and verify G{\"o}del's argument, as well as discover new facts and
confirm controversial claims about it.  This is a landmark result,
with media repercussion in a global scale, and yet it is only a
glimpse of what can be achieved by combining computer science,
philosophy and theology.

Our work, in this sense, offers new perspectives for a
computational theoretical philosophy. The critical discussion of
the underlying concepts, definitions and axioms remains a human
responsibility, but the computer can assist in building and checking
rigorously correct logical arguments. In case of logico-philosophical
disputes, the computer can check the disputing arguments and partially
fulfill Leibniz' dictum: Calculemus --- Let us calculate!

\paragraph{Acknowledgments} This work has been supported by the German Research Foundation under grants BE 2501/9-1 and BE 2501/11-1. %We thank the anonymous reviewers for their comments and suggestions.

\bibliographystyle{ecai2014}
\bibliography{Bibliography}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-19 Papers/2014-ECAI/paper.tex
added minor modifications

@@
 experiments we have conducted with HOL-ATPs.  The main findings of
-these experiments are summarized in Sec.~\ref{sec4},
+these experiments are summarized in Sec.~\ref{sec4}.
 % : in addition to
@@
 not in depth) in any of our other existing (short and partly
-non-reviewed) publications on the subject \cite{J30,W50,J29,J28}.
+non-reviewed) publications on the subject \cite{J30,J29,W50,J28}.
 
@@
     \sigma)\ar \sigma}\phi\,\llift{\supset}\, \llift{\Diamond}
-  \llift{\exists} X_\mu \lambdot \phi X]}$$ (\texttt{v} in the THF0
+  \llift{\exists} X_\mu \lambdot \phi X]}$$ \texttt{v} in the THF0
 encoding stands for \textbf{valid} and $p$ corresponds to the
-uppercase $P$, for `positive', from Fig.~\ref{fig1}).  The
+uppercase $P$, for `positive', from Fig.~\ref{fig1}.  The
 respective encodings and the results of a series of recent experiments with
@@
 2013) are provided in Fig.~\ref{fig2}.  The first row marked with T1,
-for example, shows that this theorem follows from axioms A2 and A1
+for example, shows that theorem T1 follows from axioms A2 and A1
 (where only the $\supset$-direction is needed); LEO-II and Satallax
@@
 \small
-$\begin{array}{llllllll}
+$\begin{array}{llllllll} \\[-.5em]
    & \text{\HOL encoding} & \text{dependencies} & \text{logic} & \text{status} & \text{LEO-II} & \text{Satallax} & \text{Nitpick} \\  
@@
 \text{CO'} & \emptyset\, \text{(no goal, check for consistency)} & \text{A1}(\supset), \text{A2}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & 7.5/7.8 & \text{---/---} & \text{---/---} \\
- & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & \text{---/---} & \text{---/---} & \text{---/---} \\ 
+ & & \text{A1}, \text{A2}, \text{D1}, \text{A3}, \text{A4}, \text{D2'}, \text{D3}, \text{A5} & \text{KB} & \text{UNS} & \text{---/---} & \text{---/---} & \text{---/---} \\ \\[-.5em]
 \end{array}$

### 2014-05-19 Papers/2014-ECAI/ECAI-51.tex
added further minor modifications for the final version

--- 2014-05-19 Papers/2014-ECAI/paper.tex
@@
 terms. We plan to make the entire landscape of results available to
-the interested theoretical philosophy and theology communities. This
-is relevant, since philosophers are sometimes imprecise about the
-very details of the logics they employ.
+the interested
+%theoretical philosophy and theology 
+communities. This is relevant, since philosophers are sometimes
+imprecise about the very details of the logics they employ.
 
@@
 
-\paragraph{Acknowledgments} This work has been supported by the German Research Foundation under grants BE 2501/9-1 and BE 2501/11-1.
-%We thank the anonymous reviewers for their comments and suggestions.
+%\paragraph{Acknowledgments}
+\section*{ACKNOWLEDGEMENTS}
+ This work has been supported by the German Research Foundation under grants BE 2501/9-1 and BE 2501/11-1. %We thank the anonymous reviewers for their comments and suggestions.
 

### 2014-05-19 Papers/2014-ECAI/ECAI-51.tex
added further minor modifications for the final version

@@
 This background theory extends and
-adapts the work as presented in \cite{J23,B9}.
+adapts the work as presented in \cite{B9,J23}.
 In Sec.~\ref{sec3}, we present details on the
@@
 not in depth) in any of our other existing (short and partly
-non-reviewed) publications on the subject \cite{J30,J29,W50,J28}.
+non-reviewed) publications on the subject \cite{J30,J28,W50,J29}.
 

### 2014-05-25 Papers/2014-ECAI/ECAI-51.tex
further improvements for final version

@@
 \author{Christoph Benzm\"uller\institute{Freie Universit\"at Berlin,
-    Germany, email: c.benzmueller@fu-berlin.de}
+    Germany, email: c.benzmueller@fu-berlin.de; this author has been
+    supported by the German National Research Foundation (DFG) under
+ grants BE 2501/9-1 and BE 2501/11-1. }
   \and Bruno Woltzenlogel Paleo\institute{Technical University
@@
 uses, for example, a complex higher-order modal logic (\HOML) to handle
-concepts such as \emph{possible} and \emph{necessary} and to support
+concepts such as \emph{possibility} and \emph{necessity} and to support
 quantification over individuals and properties.
@@
 embedded into the classical higher-order logic (\HOL), which can thus be seen
-as a universal logic \cite{C36}. For this universal logic, efficient automated
+as an universal logic \cite{C36}. For this universal logic, efficient automated
 theorem provers have been developed in recent years, and these systems
@@
 \begin{definition}\label{holvalue}
-  The \emph{value} $\hol{\| s_\alpha\|^{M,g}}$ of a \HOML term
-  $\hol{s_\alpha}$ on a model $\hol{M}=\hol{\langle W, R, D, I_{w\in
-      W} \rangle}$ in a world $\hol{w}\in \hol{W}$ under assignment $\hol{g}$ is an element $\hol{d}\in \hol{D_\alpha}$
+  The \emph{value} $\hol{\| s_\alpha\|^{M,g}}$ of a \HOL term
+  $\hol{s_\alpha}$ on a model $\hol{M}=\hol{\langle D, I \rangle}$ under assignment $\hol{g}$ is an element $\hol{d}\in \hol{D_\alpha}$
   defined in the following way:
@@
         \sigma}} \lambdot P\ a_\mu}$, which rewrites into the following
-  $\beta\eta$-normal \HOL term of type $\sigma$
+  $\beta\eta$-normal \HOL term of type $\hol{\sigma}$
 \[\hol{\lam{W_\iota} \all{V_\iota} \neg (r\,W\,V) \vee \neg \all{P_{\mu\ar \sigma}} \neg (P\, a_\mu\, V)}\]
@@
 \section*{ACKNOWLEDGEMENTS}
- This work has been supported by the German Research Foundation under grants BE 2501/9-1 and BE 2501/11-1. %We thank the anonymous reviewers for their comments and suggestions.
+ We thank Alexander Steen,
+ Max Wisniewski, and the anonymous reviewers for their comments and suggestions.
 

### 2014-05-25 Papers/2014-ECAI/ECAI-51.tex
Fix typo

@@
 embedded into the classical higher-order logic (\HOL), which can thus be seen
-as an universal logic \cite{C36}. For this universal logic, efficient automated
+as a universal logic \cite{C36}. For this universal logic, efficient automated
 theorem provers have been developed in recent years, and these systems

### 2014-05-26 Papers/2014-ECAI/ECAI-51.tex
new final version

@@
   {\lift{(\Box_{o\ar o}\,s_{o})}} & =
-  (\hol{\llift{\Box}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
+  (\hol{\llift{\Box}_{\sigma\ar\sigma}}\,\lift{ s_{o}})
 \end{align*}
@@
 \bibliographystyle{ecai2014}
-\bibliography{Bibliography}
+%\bibliography{Bibliography}
+\begin{thebibliography}{10}
+
+\bibitem{Adams}
+R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
+  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
+
+\bibitem{AndersonGettings}
+A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
+  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
+  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
+
+\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
+C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
+  Faith and Philosophy}, {\bf 7}(3), (1990).
+
+\bibitem{Andrews:gmae72}
+P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
+  Logic}, {\bf 37}(2),  395--397, (1972).
+
+\bibitem{andrewsSEP}
+P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
+  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
+
+\bibitem{C36}
+C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
+  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
+  pp. 232--233, Rio de Janeiro, Brazil, (2013).
+
+\bibitem{B5}
+C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
+  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
+  Elsevier, (2014).
+\newblock Forthcoming; preliminary version available at
+  {http://christoph-benzmueller.de/papers/B5.pdf}.
+
+\bibitem{C34}
+C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
+  provers for first-order modal logics', in {\em Proc. of the 20th European
+  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
+
+\bibitem{B9}
+C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
+  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
+  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
+  386--406, College Publications, (2008).
+
+\bibitem{J23}
+C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
+  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
+
+\bibitem{LEO-II}
+C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
+  cooperative automatic theorem prover for higher-order logic', in {\em
+  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
+
+\bibitem{J30}
+C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
+  automation of {G{\"o}del's} proof of {God's} existence', {\em
+  arXiv:1308.4526}, (2013).
+
+\bibitem{J28}
+C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
+  {\em Archive of Formal Proofs}, (2013).
+
+\bibitem{W50}
+C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
+  in {\em Proceedings of the 10th International Workshop on the Implementation
+  of Logics}, EPiC Series. EasyChair, (2013).
+\newblock Invited abstract.
+
+\bibitem{Coq}
+Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
+  Development}}, Springer, 2004.
+
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
+  higher-order logic based on a relational model finder', in {\em Proc. of ITP
+  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
+
+\bibitem{Satallax}
+C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
+  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
+
+\bibitem{ContemporaryBibliography}
+R.~Corazzon.
+\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
+  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
+
+\bibitem{Fitting}
+M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
+
+\bibitem{fitting98}
+M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
+  {\em Synthese Library}, Kluwer, 1998.
+
+\bibitem{Gallin75}
+D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
+
+\bibitem{garbacz12:_prover_simpl_expal_away}
+P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
+  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
+
+\bibitem{GoedelNotes}
+K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{Henkin50}
+L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
+  Logic}, {\bf 15}(2),  81--91, (1950).
+
+\bibitem{homl}
+R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
+  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
+
+\bibitem{Isabelle}
+T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
+  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
+
+\bibitem{oppenheimera11}
+P.E. Oppenheimera and E.N. Zalta, `A computationally-discovered simplification
+  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
+  89}(2),  333--349, (2011).
+
+\bibitem{rushby13}
+J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
+  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
+
+\bibitem{Schulz:AICOM-2002}
+S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
+  15}(2),  111--126, (2002).
+
+\bibitem{ScottNotes}
+D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{sobel2004logic}
+J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
+  Cambridge U. Press, 2004.
+
+\bibitem{sutcliffe2009tptp}
+G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
+  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
+  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
+  {\bf 3}(1),  1--27, (2010).
+
+\bibitem{J29}
+B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
+  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
+  (2013).
+
+\end{thebibliography}
+
 \end{document}

### 2014-07-20 Talks/2013/FU-Berlin/talk-part-7.tex
Add NCProofs talk and reorganize folders

--- 2013-11-01 Talks/FU-Berlin/talk-part-7.tex
@@
 \begin{itemize}
-\item Powerful infrastructure for reasoning with higher-order modal logic using existing proof assistants and HOL provers
+\item Powerful infrastructure for reasoning with QML
 \item A new natural deduction calculus for higher-order modal logic
@@
   \item exact parameters known: constant domain quantification, Henkin Semantics
-  \item parameters can be varied and experiments can repeated
+  \item experiments with different parameters could be performed
   \end{itemize}
@@
   \item see also Ed Zalta's \emph{Computational Metaphysics} project at Stanford University
+  \item see also John Rushby's recent verification of Anselm's proof in PVS
   \item remember Leibniz' dictum --- \emph{Calculemus!}

### 2014-07-20 Talks/2013/TU-Wien/talk-part-1.tex
Add NCProofs talk and reorganize folders

--- 2013-10-31 Talks/FU-Berlin/talk-part-1.tex
@@
 - Hamburger Abendpost \\
-- \ldots \\
+
 \end{minipage} \hfill
 %
-\begin{minipage}{.3\textwidth}
+\begin{minipage}{.35\textwidth}
 Austria \\
@@
 - ORF \\
-- \ldots \\
-
-Italy \\
-- Repubblica \\
-- Ilsussidario \\
-- \ldots \\
-
-% Russia \\
-% - \ldots \\
-
-India \\
-- DNA India \\
-- Delhi Daily News \\
-- India Today \\
-- \ldots \\
-
-US \\
-- ABC News \\
-- \ldots \\
 
@@
 - United Press Intl. \\
-- \ldots \\
+
+India \\
+- DNA India \\
+- Delhi Daily News \\
+- India Today \\
+
+US \\
+- ABC News \\
+
+Italy \\
+- Repubblica \\
+
+Spain, Russia, Brazil, Bulgaria \\
+
+\ldots
+
 \end{minipage}
@@
 
-\begin{frame}{Introduction --- Quick answers to your most pressing questions!} \large
-\colorbox{gray}{\includegraphics[width=\textwidth]{Images/News/MacBookGrab}} 
-\pause
-\vfill
-Are we in contact with Steve Jobs? \hfill No \\[2em]
-Do you really need a MacBook to obtain the results? \hfill No \\[2em]
-Is Apple sending us money? \hfill No \\
-\, \hfill (but maybe they should)
-\end{frame}
-
-\begin{frame}{Introduction}\large
-%Ontological argument: Conclude that God exists from premises by pure, a priori reasoning.
-\begin{block}{Def: \textcolor{blue}{Ontological Argument/Proof}}
-* deductive argument \\
-
-* for the existence of God \\
-
-* starting from premises, which are justified by pure reasoning,
-i.e. they do not depend on observation in the world.
-\end{block}
-
-\vfill \pause
-Existence of God: different types of arguments/proofs\\[.2em]
-\begin{itemize}
-\item[---]a posteriori (use experience/observation in the world)
-  \begin{itemize}
-  \item[------]teleological
-  \item[------]cosmological
-  \item[------]moral
-  \item[------] \ldots
-  \end{itemize}  
-\item[---]a priori (based on pure reasoning, independent)
-  \begin{itemize}
-  \item[------]\textcolor{blue}{ontological argument}
-    \begin{itemize}
-    \item[------]definitional 
-    \item[------]modal 
-    \item[------] \ldots
-    \end{itemize}
-  \item[------]other a priori arguments
-  \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Introduction}
-\vskip1em
-\emph{\huge Wohl eine jede Philosophie kreist um den ontologischen
-  Gottesbeweis} \\[1.5em]
-(Adorno, Th. W.: Negative Dialektik. Frankfurt a. M. 1966, p.378)
-\vfill
-\begin{center}
-\fcolorbox{gray}{white}{
-\begin{minipage}{.9\textwidth}
-\ldots \hfill
-\includegraphics[height=2cm]{Images/Books/buch3.jpg} \hfill
-\includegraphics[height=2cm]{Images/Books/buch2.jpg} \hfill 
-\includegraphics[height=2cm]{Images/Books/buch4.jpg} \hfill
-\includegraphics[height=2cm]{Images/Books/buch5.jpg} \hfill
-\includegraphics[height=2cm]{Images/Books/buch6.jpg} \hfill
-\includegraphics[height=2cm]{Images/Books/buch1.jpg} \hfill
-\ldots
-\end{minipage}
-}
-\end{center}
-\end{frame}
-
-\begin{frame}{Introduction} \Large
-{Rich history on ontological arguments (\textcolor{blue}{pros} and \textcolor{red}{cons})}\\[1em]
+\begin{frame}{A Long History}{\textcolor{blue}{pros} and \textcolor{red}{cons}} \Large
 
@@
 
-% \rnode{n2}{}\emph{}
-% \ncline[nodesep=2pt,linecolor=black,linewidth=1pt]{->}{n1}{n2}
 \end{frame}
 
-\begin{frame}{Introduction} \Large
-Different Interests in Ontological Arguments: \\[1em]
+\begin{frame}{The Ontological Proof Today}
+\vskip1em
+\emph{\huge Wohl eine jede Philosophie kreist um den ontologischen
+  Gottesbeweis} \\[1.5em]
+(Adorno, Th. W.: Negative Dialektik. Frankfurt a. M. 1966, p.378)
+\vfill
+\begin{center}
+\fcolorbox{gray}{white}{
+\begin{minipage}{0.9\textwidth}
+\hfill
+\includegraphics[height=2.2cm]{Images/Books/buch3.jpg} \hfill
+\includegraphics[height=2.2cm]{Images/Books/buch2.jpg} \hfill 
+\includegraphics[height=2.2cm]{Images/Books/buch4.jpg} \hfill
+\hfill
+
+\vspace{0.5 cm}
+
+\hfill
+\includegraphics[height=2.2cm]{Images/Books/buch7.jpg} \hfill
+\includegraphics[height=2.2cm]{Images/Books/buch5.jpg} \hfill
+\includegraphics[height=2.2cm]{Images/Books/buch6.jpg} \hfill
+\includegraphics[height=2.2cm]{Images/Books/buch1.jpg} \hfill
+\hfill
+\end{minipage}
+}
+
+\end{center}
+\end{frame}
+
+\begin{frame}{Motivations} \Large
 \begin{itemize}
-\item \textcolor{blue}{Philosophical:} Boundaries of Metaphysics \& Epistemology
+\item \textcolor{blue}{Philosophical:} 
   \begin{itemize}
-  \item We talk about a metaphysical concept (God), 
-  \item but we want to draw
-      a conclusion for the real world. \\[1em]
-  \item Necessary Existence (NE): metaphysical NE vs. logical NE  vs. modal NE \\[2em]
+  \item Limits of metaphysics \& epistemology
+  \item \emph{Metaphysical} versus \emph{logical} necessary existence \\[2em]
   \end{itemize} 
-\item \textcolor{blue}{Theistic:} Successful argument should convince atheists \\[2em]
-\item \textcolor{red}{Ours:} Can computers (theorem provers) be used \ldots
+\item \textcolor{blue}{Theological:} 
+  \begin{itemize}
+  \item Investigations of the nature of God
+  \item Arguments to convince atheists \\[2em]
+  \end{itemize}
+\item \textcolor{red}{Computational:} can automated reasoners be used \ldots
   \begin{itemize}
@@
   \item \ldots to verify the arguments step-by-step?
-  \item \ldots to fully automate (sub-)arguments? \\[1em]
+  \item \ldots to fully automate (sub-)arguments? \\[2em]
   \end{itemize}
+\end{itemize}
+
+\begin{center}
   \textcolor{red}{\emph{``Computer-assisted Theoretical Philosophy''}}
-\end{itemize}
+\end{center}
 \end{frame}
-
-\begin{frame}{Introduction} \large
-\hskip-1em Challenge: \hfill No provers for \emph{Higher-order Quantified Modal Logic\/} (\textcolor{red}{QML}) \\[1em]
-
-\hskip-1em Our solution: \hfill Embedding in \emph{Higher-order Classical Logic\/} (\textcolor{blue}{HOL}) \\
-\,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]
-
-\hskip-1em What we did \textcolor{blue}{(rough outline for remaining
-  presentation!)}: \\
-
-\begin{itemize}
-\item[A:] Pen and paper: \hfill detailed natural deduction proof 
-\item[B:] Formalization: \hfill in classical higher-order logic (\textcolor{blue}{HOL})
-\item[] Automation: \hfill theorem provers \textsc{Leo-II} and \textsc{Satallax} 
-\item[] Consistency: \hfill model finder \textsc{Nitpick (Nitrox)} 
-\item[C:] Step-by-step verification: \hfill proof assistant \textsc{Coq} 
-\item[D:] Automation \& verification: \hfill proof assistant \textsc{Isabelle} \\[2em]
-%\item[ ] Conclusion \\[2em]
-\end{itemize}
-\hskip-1em  Did we get any new results? \hfill  \textcolor{red}{Yes --- let's discuss this later!}
-\end{frame}

### 2014-07-20 Talks/2014/NCProofs/talk-part-7.tex
Add NCProofs talk and reorganize folders

\begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}{black}
\textbf{Conclusions and Future Work}
\end{transitionframe}

% \begin{frame}{Summary of Results} \large

% The (\alert{new}) insights we gained from experiments include:\\[.5em]
% \begin{itemize}
% \item Logic K sufficient for T1, C and T2 
% \item Logic S5 \emph{not} needed for T3
% \item \alert{Logic KB sufficient for T3 (not well known)}
% \item \alert{We found a simpler new proof of C}
% \item \alert{G\"odel's axioms (without conjunct $\phi(x)$ in D2) are inconsistent}
% \item Scott's axioms are consistent
% \item For T1, only half of A1 (A1a) is needed 
% \item For T2, the other half (A1b) is needed
% \end{itemize}
% \end{frame}

\begin{frame}{Conclusions} \large

\begin{itemize}
\item Contributions: \\[.5em]
\begin{itemize}
\item Powerful infrastructure for reasoning with QML
\item A (new?) natural deduction calculus for \\ higher-order (intuitionistic) modal logic
\item Non-trivial new benchmark problems for HOL provers
\item Verification of existing results about G\"odel's proof (e.g. Modal Collapse)
\item New results about G\"odel's proof (e.g. Inconsistency)
%\item Huge media attention
\end{itemize}

%\item Interesting bridge between CS, Philosophy and Theology

\item Major step towards \alert{Computer-assisted Theoretical Philosophy}
 \begin{itemize}
  \item see also Ed Zalta's \emph{Computational Metaphysics} project at Stanford University and John Rushby's formalization of Anselm's proof using PVS
  %\item remember Leibniz' dictum --- \emph{Calculemus!}
  \end{itemize}
\end{itemize}
\end{frame}

\newcommand\hol[1]{\boldsymbol{#1}}
\newcommand\lift[1]{\lceil #1 \rceil}
\newcommand\llift[1]{\dot{#1}}

\begin{frame}{Future Work}{From a Non-Classical Proof-Theoretic Perspective} \large

\begin{itemize}
\item Embedding of other non-classical logics
\begin{itemize}
\item Quantified Conditional Logics, \\
      Modal Logics based on Neighbourhood Semantics, \ldots
\pause
\item Other Suggestions?
\pause
\item Human Resources?
\end{itemize}

\pause

\item New non-classical calculi via HOL embedding: \\
(when does this approach work?)
\pause
\begin{enumerate}
  \item Embed your favorite logic $L$ into HOL
  \pause
  \item Play with Coq and invent new tactics to encapsulate the embedding and make it transparent to the user
  \pause
  \item Interpret the tactics as new rules for a new natural deduction calculus $C$
  \pause
  \item Prove soundness and completeness indirectly: \\
  \[
  \models_L s_o \quad \text{iff} \quad \hol{\models} \hol{\textrm{valid}(}\lceil s_o \rceil\hol{)}
  \qquad\qquad
  \hol{\vdash} \hol{\textrm{valid}(}\lceil s_o \rceil\hol{)} \quad \text{iff} \quad \vdash_C s_o
  \]

\end{enumerate}

%\pause

%\item Why is it so hard to prove T3 without intermediary steps?

%\pause

%\item Provocative question: \pause life is short! So why take the long structural proof-theoretical road, if the ``lazier'' HOL embedding approach gives us efficient theorem provers in less than 100 lines of declarative code?

\end{itemize}
\end{frame}

% \begin{frame}{Ongoing and Future Work} \large
% \begin{itemize}
% \item Formalize and verify other ontological arguments
%   \begin{itemize}
%   \item \ldots particularly the criticisms and improvements
%   \end{itemize}
% \item Experiment with other embeddings (e.g. varying domains)
% \item Eliminate and introduce cuts
% \end{itemize}
% \end{frame}

% \begin{frame}{Some Comments and Reactions}
% \colorbox{gray}{\includegraphics[width=.8\textwidth]{Images/Comments/Comment1}}\\[.7em]

% \, \hfill \colorbox{gray}{\includegraphics[width=.7\textwidth]{Images/Comments/Comment2}}\\[.7em]

% \colorbox{gray}{\includegraphics[width=.8\textwidth]{Images/Comments/Comment3}}\\[1em]

% \, \hfill \ldots find more on the internet \ldots
% \end{frame}

\begin{frame}[plain]
\colorbox{black}{\includegraphics[width=\textwidth]{Images/Transitions/GodComputerC}}
\end{frame}

% \begin{frame}{Licenses} \centering
% \includegraphics[scale=0.5]{Images/CC-BY-SA.png}

% \bigskip
% \bigskip

% \begin{center}
% The following images used in these slides were obtained in commons.wikimedia.org and are licensed as follows:

% \bigskip

% CC-BY-SA:

% ReligiousSymbols, PaganReligiousSymbols.

% \bigskip

% Public Domain:

% TrifidNebula
% \end{center}
% \end{frame}

### 2014-08-04 Formalizations/THF-FrodeBjordal/call_tptp.sh
Added THF formalization and automation of Frode Bjordal's 98 paper

@@ -1,0 +1,36 @@
+tptp4X=./tptp4X
+call=./RemoteSOT.pl
+# if the above programs don't work for you download the SystemOnTPTP package,
+# install it and change the above paths accordingly
+
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='LEO-II---1.6.2 Nitrox---2013'
+
+timeout=200
+
+echo
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+
+for file in "$@"
+do
+echo
+    for prover in $proverlist 
+    do
+    function_to_fork () { 
+      res=""
+      while ! (echo $res | grep --quiet RESULT); do
+          res=`$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover |grep RESULT` 
+	  echo ${file}---${prover}---${res}
+      done   
+    }
+    function_to_fork &
+    done
+    wait
+    echo Done 
+done
+
+
+
+
+

### 2014-08-04 Formalizations/THF-FrodeBjordal/run-experiments.sh
added experiments and results

@@ -1,0 +1,13 @@
+touch ExperimentResults.txt
+rm ExperimentResults.txt
+touch ExperimentResults.txt
+./call_tptp.sh D_EntailsInS4_A2.p >> ExperimentResults.txt			
+./call_tptp.sh D_EntailsInS4_A3.p >> ExperimentResults.txt		
+./call_tptp.sh D_EntailsInS4_A4.p >> ExperimentResults.txt		
+./call_tptp.sh D_EntailsInS4_Def1.p >> ExperimentResults.txt		
+./call_tptp.sh Def1+A2+A3+A4_Entails_D.p >> ExperimentResults.txt
+./call_tptp.sh Def1+A2+A3+A4_Entails_D-Case2.p >> ExperimentResults.txt
+./call_tptp.sh Def1+A2+A3+A4_Entails_D-Case1.p >> ExperimentResults.txt	
+./call_tptp.sh Def1+A2+A3+A4_Entails_D-CasesCombined.p >> ExperimentResults.txt
+./call_tptp.sh GodTheorem.p >> ExperimentResults.txt
+./call_tptp.sh ModalCollapse.p >> ExperimentResults.txt

### 2014-08-04 Formalizations/THF-FrodeBjordal/call_tptp.sh
added experiments and results

@@ -5,7 +5,9 @@ call=./RemoteSOT.pl
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
-proverlist='LEO-II---1.6.2 Nitrox---2013'
+#proverlist='LEO-II---1.6.2 Nitrox---2013'
+proverlist='LEO-II---1.6.2'
+#proverlist='Nitrox---2013'
 
 timeout=200
 

### 2014-08-04 Formalizations/THF-FrodeBjordal/call_tptp.sh
added experiments and results

@@ -6,8 +6,8 @@ call=./RemoteSOT.pl
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
 #proverlist='LEO-II---1.6.2 Nitrox---2013'
-proverlist='LEO-II---1.6.2'
-#proverlist='Nitrox---2013'
+#proverlist='LEO-II---1.6.2'
+proverlist='Nitrox---2013'
 
 timeout=200
 

### 2014-08-05 Papers/NaturalDeduction/GodProof-ND.tex
Improve introduction and conclusion of ND paper

--- 2013-12-13 Natural Deduction/GodProof-ND.tex
@@
 
+\title{A Variant of G\"{o}del's Ontological Proof \\
+in a Natural Deduction Calculus \\
+(Draft)}
+
 \author{Bruno Woltzenlogel Paleo, Annika Siders}
-
-\title{G\"{o}del's Ontological Proof of God's Existence (Draft)}
 
@@
 
-ToDo: Do also Scott's and G\"{o}del's proofs.
+Ontological arguments for the existence of God can be traced back at least to St.
+Anselm (1033-1109).
+His argument considered a maximally conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
 
-The first ontological argument for the existence of God was written by St.
-Anselm (1033-1109).
-The argument considers a maximally conceivable being.
-This being must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
-The main critique of this argument is that we do not know whether the concept maximal conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
-The question is if the argument is sound.
+A major critique of this argument is that we do not know whether the concept maximal conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
+%The question is if the argument is sound. 
+As Bertrand Russell pointed out, the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have \citep{citation needed}.
 
-As Bertrand Russell has pointed out the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have.
-Kant, on the other hand argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
+Kant argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
 This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
-All that we can say is that if God exists, then God necessarily exists.
+All that we can say is that if God exists, then God necessarily exists \citep{citation needed}.
 
 St.
-Anselm's proof was processed by Descartes and Leibniz.
-Leibniz identified the critical point of the argument as establishing the possible existence of God.
-Leibniz gave an argument for that the properties of God, the perfections, are compatible.
+Anselm's argument was elaborated further by Descartes and Leibniz.
+Leibniz identified that establishing the possible existence of God is a critical missing step in St.
+Anselm's argument.
+To fill this gap, he argued that the properties of God, the perfections, are compatible.
 This implies that it is possible to have all perfections at once and therefore the existence of a maximal being with all these properties is possible.
 
-G\"odel continued this argument by defining the properties that must hold for these perfections and giving axioms from which the existence of God can be derived.
-The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
-The formal argument of G\"odel is based on Leibniz proof, which in turn is based on Descartes proof.
-These proofs have two parts; a proof that if God's existence is possible, then it is necessary and a proof that God's existence is in fact possible.
-The deductive system used in the formal proof is a system for modal logic extended with second-order quantification (or possibly third-order quantification), because the axioms and definitions required for the concept of positive properties or perfections involve quantification over properties.
+G\"odel studied Leibniz's work \citep{Adams} and brought his ontological argument to a modern form using a modal logic with higher-order quantification over properties.
+In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being having all positive properties.
 
-G\"odel treated positive properties as not just atomic properties, like Leibniz, but also consisting of collections of these properties, which by Leibniz argument are compatible or possible.\citep{fitting}[p.139.] G\"odel formulated an axiom stating that the conjunction of any set of positive properties is positive.
-The property of God-likeness, which defines what a God is, must therefore be an infinite conjunction of all the positive properties.
-This axiom of the positiveness of conjunctions can be given a third order formulation which is the reason for the claim that third-order quantification may be required \citep{and}.
+G\"odel's notion of positive property and Leibniz's notion of perfection are not exactly the same \citep{Adams}.
+On a technical level, the main distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim).
+In particular, one of G\"odel's axioms states that the conjunction of any set of positive properties is positive.
+And from this axiom, it follows immediately that the property of being God-like is positive.
+While this step is intuitively and informally clear, it is not easily formalizable in a typical logical calculus, because it requires inferring that being god-like is a (possibly infinite) conjuntion of positive properties, while it has only been defined as the property holding for individuals who have all positive properties.
+This interplay between universal quantification (in the definition of a God-like being) and conjunction (in G\"odel's mentioned axiom) is a technical difficulty that probably explains why, starting with Scott \cite{scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another one that simply assumes the positivity of the property of being God-like.
+%This axiom is unfortunately equivalent to the possibility of God's existence given the other axioms, leading to criticisms that it might be too strong \citep{sobel2}\citep{fitting}[Section 11.4.
+Objections.].
+
+%The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself. Thus, we have an axiom which seems to be equivalent to the conclusion.
+
+ % This axiom of the positiveness of conjunctions can be given a third order formulation which is the reason for the claim that third-order quantification may be required \citep{and}.
 %\citep{fitting}[p.
 148.]
-However, this axiom is only used in the proof to establish that the property of God-likeness is itself positive.
-G\"odel's student Scott, who wrote down a version of G\"odel's proof after G\"odel had confided his work in him, simply chose to assume the positiveness of  God-likeness as an axiom instead.
 
-Nevertheless, both axioms have the unfortunate property of being equivalent to the possibility of God's existence, given the other axioms of G\"odel \citep{sobel2}.
-%\citep{fitting}[Section 11.4.
-Objections.]
-The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself. Thus, we have an axiom which seems to be equivalent to the conclusion.
-
-Another, possibly undesirable, consequence of the formalism is that there can only be one God, which is defined by the axioms.
+% Another, possibly undesirable, consequence of the formalism is that there can only be one God, which is defined by the axioms.
 Namely, if one assumes that the theory contains the equality relation then a God has the property of being identical to itself.
@@
 
-In \citep{sobel} a concern is raised about the whole system of proof, which is further elaborated on in \citep{sobel2}[Ch.
-4,  section 6].
-Depending on the properties of the rules for second order quantification allowed in the system, the system may have modal collapse.
-That is, if no restrictions are placed on the properties over which quantification is allowed, then the system says that for all formulas the formula implies its boxed version.
-However, there are solutions to this problem by modification of the axioms.
-Sobel's own solution to the problem is to treat properties as extensional instead of intentional.
-Another solution by Anderson %(1990)
-is to require that Godlike beings have positive properties necessarily. A summary of these arguments can be found in Fitting \citep{fitting}[Sections 11.9 and 11.10].
+This paper presents new natural deduction proofs of (Scott's version of) the lemmas and theorems in G\"odel's ontological argument \cite{Goedel}.
+In contrast to Scott's version \cite{Scott}, also shown here in natural deduction form, the version presented here is characterized by shorter and more direct derivations.
+
+Section \ref{sec:calculus} defines the higher-order modal natural deduction calculus used to construct the formal proofs.
+Section \ref{sec:scott} shows Scott's proofs, translated to natural deduction.
+And section \ref{sec:proof} presents the new proofs.
 
 \section{Natural Deduction}
+\label{sec:calculus}
 
@@
 
-\section{Possibly, God Exists}
+\section{Scott's Proof}
+
+The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
+The formal argument of G\"odel is based on Leibniz proof, which in turn is based on Descartes proof.
+These proofs have two parts; a proof that if God's existence is possible, then it is necessary and a proof that God's existence is in fact possible.
+
+\subsection{Possibly, God Exists}
 
@@
 
-\section{Being God is an essence of any God}
+\subsection{Being God is an essence of any God}
 
@@
 
-\section{If God's existence is possible, it is necessary}
+\subsection{If God's existence is possible, it is necessary}
 
@@
 
-\section{God exists}
+\subsection{God exists}
 
@@
 
-\section{Necessarily, God exists}
+\subsection{Necessarily, God exists}
 
@@
 
-\section{God exists}\label{GodsE}
+\subsection{God exists}\label{GodsE}
 This section is superfluous.
@@
 
-The axioms of the calculus imply the following collapse of modalities:
+A major criticism against G\"odel's proof is that its axioms lead to the so-called \emph{modal collapse} \citep{sobel}: it is possible to prove that everything that is the case is so necessarily, and hence actuality, possibility and necessity coincide \citep{sobel2}[Ch.
+4,  section 6, theorems 9 and 10].
+As this is an undesirable consequence, many solutions to the problem of the modal collapse have been proposed.
+
+Anderson's solution \citep{citation needed} modifies the definitions of God-like being and essence, and eliminates half of an axiom.
+This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
+Another solution involving more substantial modifications is that of Bj{\o}rdal \citep{bjordal,fuhrmann}.
+
+On another track, Fitting has argued that greater care has to be taken with the semantics of higher-order modal logics.
+Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional.
+Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
+
+Anderson \citep{anderson1990}[p.
+292] and Sobel \citep{sobel2}[p.
+133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
+
+%In his manuscript, G\"odel warned that ``positive means positive in the moral aesthetic sense [\ldots]. Only then [are] the axioms true''.
+
+It is beyond the scope of this paper to analyze these solutions in detail or propose new solutions.
+The purpose of this section is simply to show natural deduction derivations of the modal collapse, thus confirming that it holds for the axioms used in the previous sections.
 
@@
 
-This collapse of the system was first proved in \citep{sobel} and the argument is also given in \citep{fitting}[ch.
+% This collapse of the system was first proved in \citep{sobel} and the argument is also given in \citep{fitting}[ch.
 11, sections 8-11] who provides summaries of two solutions by the modifications of the axioms.
@@
 
-The question arises what kind of second order quantification is allowed.
+% The question arises what kind of second order quantification is allowed.
 The instantiations that are used in our proof are instantiated with predicate variables that later work as eigenvariables for the second order universal introduction.
@@
 
-One can also discuss if a second order modal logic is sound and complete.
+% One can also discuss if a second order modal logic is sound and complete.
 With Henkin models, which restrict the second order quantification with comprehension, we have that normal second order logic is sound and complete.
 There is a completeness proof for a second order modal logic found in A completeness theorem in second order modal logic by Cocchiarella.
+
+\section{Conclusions}
+
+The proofs of theorem T1 and corollary C1 ($\pos \ex x.
+G(x)$) do not rely on equality and use axiom A2 ($\all \varphi.
+\all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$) only once.
+Corollary C2 ($\ex x.
+G(x)$), which is usually regarded as a trivial corollary of main theorem T3 ($\nec \ex x.
+G(x)$) using the modal logic axiom T, is here derived directly from lemma L1 and corollary C1, not relying on T.
+The main theorem T3 becomes derivable from C2 by a single application of the necessitation rule.
+Furthermore, all proofs are done in the modal logic K, except for the proof of corollary C2, which requires one use of the axiom B of modal logic KB.
+
+ToDo: Sobel Anderson on KB: Sobel page 152
 

### 2014-08-06 Papers/NaturalDeduction/GodProof-ND.tex
Improve ND paper

@@
 \usepackage{bussproofs}
+\usepackage{amsmath}
 \EnableBpAbbreviations
@@
 
+The language of higher-order modal logic used here is inspired by that of Church's simple type theory \citep{church}.
+
+\begin{definition} \emph{Simple types} are given by the following grammar:
+$$
+  \theta,\tau \quad ::= \quad \mu \ \mid \ o \ \mid \ \theta \imp \tau
+$$
+where $\mu$ is the atomic type for individuals, $o$ is the atomic type for propositions and $\theta \imp \tau$ is the type for functions taking an argument of type $\theta$ and returning something of type $\tau$.
+`$\imp$' is assumed to be right associative.
+\end{definition}
+
+\begin{definition} \emph{Terms and formulas} are given by the following grammar:
+\begin{align*}
+ s,t \quad ::= \quad & 
+  p_\tau \ \mid \ 
+  X_\tau \ \mid \
+  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
+  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
+& \bot_o \ \mid \
+  \imp_{o\imp o\imp o} \ \mid \ 
+  \wedge_{o\imp o\imp o} \ \mid \
+  \vee_{o\imp o\imp o} \ \mid \\
+& \all_{(\tau\imp o)\imp o} \ \mid \ 
+  \ex_{(\tau\imp o)\imp o} \ \mid \
+  \nec_{o\imp o} \ \mid \
+  \pos_{o\imp o}
+\end{align*}
+where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$.
+Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed.
+Furthermore, subscript types are omitted when they are clear from the context.
+Negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
+\end{definition}
+
+For Natural deduction
+
+beta reduction
+
 \newcommand{\s}{\qquad}
@@
 \begin{calculus}
-{The intuitionistic natural deduction calculus \ND}
+{propositional rules}
 {fig:ND}
@@
 \s\s
+\infer[\bot_E]{A}{ \bot }
+\s
 \infer[\imp_I]{A \imp B}{ B }
-\s\s
+\s
 \infer[\imp_I^n]{A \imp B}{ \infer*{B}{\infer[n]{A}{}} }
-\s\s
+\s
 \infer[\imp_E]{B}{A & A \imp B}
@@
 
-\vspace{2em}
+\vspace{1em}
+\end{calculus}
+
+\begin{calculus}
+{double negation elimination}
+{fig:Classical}
+\infer[\neg\neg_E]{A}{ \neg\neg A }
+\end{calculus}
+
+\begin{calculus}
+{quantifier rules}
+{fig:QuantifierRules}
+
+\vspace{1em}
 
@@
 
-We let negation be a defined concept by $\neg A\equiv A\imp \bot$.
-The rules for negation introduction and elimination are special cases of the implication rules.
-Equivalence is also a defined concept with  $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
-\end{calculus}
-
-\begin{calculus}
-{Classical Rule: Double negation elimination}
-{fig:Classical}
-\infer[\neg\neg_E]{A}{ \neg\neg A }
 \end{calculus}
@@
 \s\s\s\s
-\infer[\nec_I]{\nec A}{\eta: \fbox{\infer*{A}{}} }
-\s\s\s\s\s
-\infer[\nec_E]{n: \fbox{ \infer*{}{A} } }{\nec A}
+\infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
+\s\s\s\s
+\infer[\nec_E]{w: \fbox{ \infer*{}{A} } }{\nec A}
 
-\vspace{2em}
+% \vspace{2em}
 
-\s\s\s\s
-\infer[\pos_I]{\pos A}{n: \fbox{\infer*{A}{}} }
-\s\s\s\s\s
-\infer[\pos_E]{\eta: \fbox{ \infer*{}{A} } }{\pos A}
+% \s\s\s\s
+% \infer[\pos_I]{\pos A}{n: \fbox{\infer*{A}{}} }
+% \s\s\s\s\s
+% \infer[\pos_E]{\eta: \fbox{ \infer*{}{A} } }{\pos A}
 
@@
 \noindent
-A \emph{derivation} is a directed acyclich graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures \ref{fig:ND} and \ref{fig:NDK}.
+A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures \ref{fig:ND}, \ref{fig:Classical} and \ref{fig:NDK}.
 Parts of a derivation may be surrounded by boxes.
@@
 %
-\item \textbf{eigen-box conditions:} $\eta$ must be a fresh name for a box.
+\item \textbf{eigen-box conditions:} $\omega$ must be a fresh name for a box.
 %
@@
 
-\section{Scott's Proof}
+\subsection{Suitability for K}
+
+We show that the proof system with boxed parts of derivations is equivalent to the system K. The modal system K consists of the axiom K and the necessitation rule N.
+
+\begin{axiom}[The transitivity axiom K]
+\label{K}
+$$
+\nec(A\imp B)\imp (\nec A\imp \nec B)
+$$
+\end{axiom}
+
+\begin{axiom}[The necessitation rule, $\Box_I$]
+\label{Necessitation}
+If $A$ is a theorem, then $\nec A$ is a theorem.
+\end{axiom}
+
+\begin{lemma}
+\label{systemK}
+The axiom K is derivable in the system.
+\end{lemma}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec_E$}
+\UIC{$A\imp B$}
+      \AXC{$\nec A ^1 $}\RightLabel{$\nec_E$}
+      \UIC{$A$} \RightLabel{$\imp_E$}
+   \BIC{$ B$} \RightLabel{$\nec_I$}
+   \UIC{$ \nec B$} \RightLabel{$\imp_I^1$}
+   \UIC{$\nec A\imp \nec B$} \RightLabel{$\imp_I^2$}
+   \UIC{$\nec(A\imp B)\imp (\nec A\imp \nec B)$}
+\end{prooftree}
+\end{small}
+
+\begin{lemma}
+\label{systemK2}
+Assuming the axiom K and the necessitation rule $\Box_I$, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
+\end{lemma}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$A^1 $}\noLine
+\UIC{$\vdots$}\noLine
+\UIC{$ B$} \RightLabel{$\imp_I^1$}
+\UIC{$A\imp B$} \RightLabel{$\Box_I$}
+\UIC{$\nec(A\imp B)$}
+      \AXC{Axiom K}\dashedLine
+      \UIC{$\nec(A\imp B)\imp (\nec A\imp \nec B)$} \RightLabel{$\imp_E$}
+  \BIC{$\nec A\imp \nec B$}
+        \AXC{$\nec A ^1 $}\RightLabel{$\imp_E$}
+    \BIC{$\nec B$}
+\end{prooftree}
+\end{small}
+
+\section{Scott's Proof in Natural Deduction}
 
@@
 
+\begin{axiom}
+\label{A1}
+Either a property or its negation is positive, but not both:
+$$
+\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+$$
+\end{axiom}
+
+\begin{axiom}
+\label{A2}
+A property necessarily implied by a positive property is positive:
+$$
+\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
+$$
+\end{axiom}
+
+\section{New Proof}
+
 \subsection{Possibly, God Exists}
+
+\setcounter{axiom}{0}
 
@@
 
-\section{Proof system equivalent to system K}
-
-We show that the proof system with boxed parts of derivations is equivalent to the system K. The modal system K consists of the axiom K and the necessitation rule N.
-
-\begin{axiom}[The transitivity axiom K]
-\label{K}
-$$
-\nec(A\imp B)\imp (\nec A\imp \nec B)
-$$
-\end{axiom}
-
-\begin{axiom}[The necessitation rule, $\Box_I$]
-\label{Necessitation}
-If $A$ is a theorem, then $\nec A$ is a theorem.
-\end{axiom}
-
-\begin{lemma}
-\label{systemK}
-The axiom K is derivable in the system.
-\end{lemma}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec_E$}
-\UIC{$A\imp B$}
-      \AXC{$\nec A ^1 $}\RightLabel{$\nec_E$}
-      \UIC{$A$} \RightLabel{$\imp_E$}
-   \BIC{$ B$} \RightLabel{$\nec_I$}
-   \UIC{$ \nec B$} \RightLabel{$\imp_I^1$}
-   \UIC{$\nec A\imp \nec B$} \RightLabel{$\imp_I^2$}
-   \UIC{$\nec(A\imp B)\imp (\nec A\imp \nec B)$}
-\end{prooftree}
-\end{small}
-
-\begin{lemma}
-\label{systemK2}
-Assuming the axiom K and the necessitation rule $\Box_I$, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
-\end{lemma}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$A^1 $}\noLine
-\UIC{$\vdots$}\noLine
-\UIC{$ B$} \RightLabel{$\imp_I^1$}
-\UIC{$A\imp B$} \RightLabel{$\Box_I$}
-\UIC{$\nec(A\imp B)$}
-      \AXC{Axiom K}\dashedLine
-      \UIC{$\nec(A\imp B)\imp (\nec A\imp \nec B)$} \RightLabel{$\imp_E$}
-  \BIC{$\nec A\imp \nec B$}
-        \AXC{$\nec A ^1 $}\RightLabel{$\imp_E$}
-    \BIC{$\nec B$}
-\end{prooftree}
-\end{small}
-
 \section{Modal collapse}

### 2014-08-06 Papers/NaturalDeduction/GodProof-ND.tex
Improve introduction of ND paper

@@
 This interplay between universal quantification (in the definition of a God-like being) and conjunction (in G\"odel's mentioned axiom) is a technical difficulty that probably explains why, starting with Scott \cite{scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another one that simply assumes the positivity of the property of being God-like.
+
+The main criticism against G\"odel's ontological argument is the \emph{modal collapse}, an undesirable consequence of the stipulated axioms.
+Most recent works \cite{todo} on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
+This is discussed in more detail in Section \ref{sec:collapse}.
+
 %This axiom is unfortunately equivalent to the possibility of God's existence given the other axioms, leading to criticisms that it might be too strong \citep{sobel2}\citep{fitting}[Section 11.4.
@@
 
-This paper presents new natural deduction proofs of (Scott's version of) the lemmas and theorems in G\"odel's ontological argument \cite{Goedel}.
-In contrast to Scott's version \cite{Scott}, also shown here in natural deduction form, the version presented here is characterized by shorter and more direct derivations.
+The first contribution of this paper is the detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
+A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons.
+Firstly, presentations of G\"oedel's proof are typically either informal or formalized in other styles of calculi (e.g.
+Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}).
+Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
+Secondly, natural deduction is a familiar style for many people, who can now benefit from a natural deduction presentation of G\"odel's ontological argument.
+Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.
 
-Section \ref{sec:calculus} defines the higher-order modal natural deduction calculus used to construct the formal proofs.
-Section \ref{sec:scott} shows Scott's proofs, translated to natural deduction.
-And section \ref{sec:proof} presents the new proofs.
+The main contribution of the paper, however, is the presentation of new proofs (also in natural deduction style) of the lemmas, theorems and corollaries in G\"odel's manuscript.
+In contrast to Scott's proofs \cite{Scott}, the proofs presented here are simpler and shorter, as discussed in Section \ref{sec:conclusion}.
 
@@
 
-For Natural deduction
+The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}. 
+To deal with classicality \ref{fig:Classical} and \ref{fig:NDK}. These rules
+
+A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures .
+Parts of a derivation may be surrounded by boxes.
+A \emph{proof} is a derivation that additionally satisfies the following conditions:
 
@@
 
+ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
+
 \begin{calculus}
 {propositional rules}
-{fig:ND}
+{fig:PropositionalRules}
 
@@
 
+\textbf{eigen-variable conditions:}
+if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
+if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+
+\vspace{1em}
+
 \end{calculus}
@@
 
-% \vspace{2em}
+\vspace{2em}
 
-% \s\s\s\s
-% \infer[\pos_I]{\pos A}{n: \fbox{\infer*{A}{}} }
-% \s\s\s\s\s
-% \infer[\pos_E]{\eta: \fbox{ \infer*{}{A} } }{\pos A}
+\s\s\s\s
+\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
+\s\s\s\s\s
+\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
 
@@
 
-\noindent
-A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures \ref{fig:ND}, \ref{fig:Classical} and \ref{fig:NDK}.
-Parts of a derivation may be surrounded by boxes.
-A \emph{proof} is a derivation that additionally satisfies the following conditions:
-
 \begin{itemize}
-\item \textbf{eigen-variable conditions:}
-if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
-if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+\item 
 %

### 2014-08-15 Formalizations/THF/Run_ECAI2014_Experiments.sh
added file

@@ -1,0 +1,1 @@
+./call_tptp.sh T1_K_const_max.p T1_K_const_min.p T1_K_vary_max.p T1_K_vary_min.p T1_varying.p C_K_const_max.p  C_K_const_min.p  C_K_vary_max.p   C_K_vary_min.p T2_K_const_max.p  T2_K_const_min.p  T2_K_vary_max.p   T2_K_vary_min.p  T3_KB_const_max.p  T3_KB_const_min.p  T3_KB_vary_max.p   T3_KB_vary_min.p   T3_K_const_max.p   T3_K_const_min.p   T3_K_vary_max.p    T3_K_vary_min.p   MC_KB_const_max.p  MC_KB_const_min.p  MC_KB_vary_max.p   MC_KB_vary_min.p   FG_KB_const_max.p  FG_KB_const_min.p  FG_KB_vary_max.p   FG_KB_vary_min.p  MT_KB_const_max.p  MT_KB_const_min.p  MT_KB_vary_max.p   MT_KB_vary_min.p   CO_KB_const_max.p  CO_KB_vary_max.p  CO'_KB_const_max.p  CO'_KB_const_min.p  CO'_KB_vary_max.p   CO'_KB_vary_min.p  

### 2014-08-15 Papers/NaturalDeduction/GodProof-ND.tex

@@
 \usepackage{theorems}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 
@@
 
-\author{Bruno Woltzenlogel Paleo, Annika Siders}
+\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}}
+
+\authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
+
+\institute{
+  Helsinki, Finland\\
+  \email{annika.siders@helsinki.fi}
+  \and 
+  Vienna, Austria \\
+  \email{bruno@logic.at}
+}
 
@@
 
-The first contribution of this paper is the detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
+The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
 A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons.
@@
 Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
-Secondly, natural deduction is a familiar style for many people, who can now benefit from a natural deduction presentation of G\"odel's ontological argument.
+Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
 Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.
@@
 The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
-To deal with classicality \ref{fig:Classical} and \ref{fig:NDK}. These rules
+The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}.
+Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
+Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}.
+Beta-reduction is implicit; all rules are assumed modulo beta-reduction.
+A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules.
+A \emph{proof} of a formula $F$ is a derivation without open assumptions and having $F$ as root not inside any box.
 
-A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules shown in Figures .
-Parts of a derivation may be surrounded by boxes.
-A \emph{proof} is a derivation that additionally satisfies the following conditions:
-
-beta reduction
+Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
+Dashed lines are used to refer to an axiom or theorem with proof shown elsewhere.
+Dotted lines are used to indicate folding and unfolding of definitions.
+Furthermore, as it is inconvenient to draw boxes around large derivations in \LaTeX, formulas inside boxes are labeled with the names of the boxes surrounding them.
+Therefore, the boxes can be omitted without loss of information.
 
@@
 
-ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
+% ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
 
@@
 
-\vspace{2em}
+% \vspace{2em}
 
-\s\s
-\infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
-\s\s
-\infer[\vee_{I_1}]{A \vee B}{A}
-\s\s
-\infer[\vee_{I_2}]{A \vee B}{B}
+% \s\s
+% \infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
+% \s\s
+% \infer[\vee_{I_1}]{A \vee B}{A}
+% \s\s
+% \infer[\vee_{I_2}]{A \vee B}{B}
 
@@
 \s
-\infer[\all_I]{\all x. A[x]}{ A[\alpha] }
+\infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] }
 \s
-\infer[\all_E]{A[t]}{ \all x. A[x] }
+\infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }
 \s\s
-\infer[\ex_I]{\ex x. A[x]}{ A[t] }
+\infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
 \s
-\infer[\ex_E]{A[\beta]}{ \ex x. A[x] }
+\infer[\ex_E]{A[\beta]}{ \ex x_{\tau}. A[x] }
 
@@
 
-\textbf{eigen-variable conditions:}
-if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$;
+\begin{center}
+\textbf{eigen-variable conditions:} \\
+if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$; \\
 if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+\end{center}
 
@@
 
-\vspace{2em}
+% \vspace{2em}
 
-\s\s\s\s
-\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-\s\s\s\s\s
-\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
+% \s\s\s\s
+% \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
+% \s\s\s\s\s
+% \infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
 
 \vspace{1em}
+
+\begin{center}
+\textbf{eigen-box condition:}\\ 
+$\omega$ must be a fresh name for a box. \\
+\vspace{0.5em}
+\textbf{boxed assumption condition:} \\
+assumptions should be discharged within the box where they are made.
+\end{center}
+
+\vspace{1em}
+
 \end{calculus}
 
-\begin{itemize}
-\item 
-%
-\item \textbf{eigen-box conditions:} $\omega$ must be a fresh name for a box.
-%
-\item \textbf{boxed assumption condition:} any assumption should be discharged within the box where it is made.
-%
-\item \textbf{unboxed root condition:} the proof's root should not be inside any box.
-\end{itemize}
+%\clearpage
 
 \noindent
-Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
-Dashed lines are used to refer to a proof shown elsewhere.
-Dotted lines are used to indicate folding and unfolding of definitions.
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND.
+The calculus with the additional rules shown in Figure \ref{fig:NDK} is called \NDK.
 
-\subsection{Suitability for K}
+%\clearpage
 
-We show that the proof system with boxed parts of derivations is equivalent to the system K. The modal system K consists of the axiom K and the necessitation rule N.
+\subsection{Suitability for Rigid Higher-Order Modal Logic K}
 
-\begin{axiom}[The transitivity axiom K]
-\label{K}
-$$
-\nec(A\imp B)\imp (\nec A\imp \nec B)
-$$
-\end{axiom}
+Adding the modal rules results in a calculus that is suitable for the basic normal modal logic K.
+In other words, {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
 
-\begin{axiom}[The necessitation rule, $\Box_I$]
-\label{Necessitation}
-If $A$ is a theorem, then $\nec A$ is a theorem.
-\end{axiom}
+The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
+From the point of view of a \emph{possible worlds} semantics, rigidity means that their interpretation is independent of the world in which they are being interpreted.
+Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
+Nevertheless, it should be noted that its adequacy has already been contested \cite{fitting}.
 
-\begin{lemma}
-\label{systemK}
-The axiom K is derivable in the system.
-\end{lemma}
+\begin{theorem}
+\label{theorem:completeness}
+{\NDK} is complete, relative to {\ND} extended with axiom K and the necessitation rule.
+\end{theorem}
+\begin{proof}
+The necessitation rule can be immediately simulated with the $\nec_I$ rule. Axiom K can be derived in {\NDK} as shown below:
 
@@
 \AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec_E$}
-\UIC{$A\imp B$}
+\UIC{$\omega: A\imp B$}
       \AXC{$\nec A ^1 $}\RightLabel{$\nec_E$}
-      \UIC{$A$} \RightLabel{$\imp_E$}
-   \BIC{$ B$} \RightLabel{$\nec_I$}
+      \UIC{$\omega: A$} \RightLabel{$\imp_E$}
+   \BIC{$ \omega: B$} \RightLabel{$\nec_I$}
    \UIC{$ \nec B$} \RightLabel{$\imp_I^1$}
@@
 
-\begin{lemma}
-\label{systemK2}
-Assuming the axiom K and the necessitation rule $\Box_I$, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
-\end{lemma}
+\end{proof}
+
+\begin{theorem}
+\label{theorem:soundness}
+{\NDK} is sound, relative to {\ND} extended with axiom K and the necessitation rule.
+\end{theorem}
+\begin{proof}
+It is necessary to show that {\NDK} proofs of the following form can be translated to proofs in {\ND} extended with the axiom K and the necessitation rule.
+
+\begin{small}
+$$
+\infer[\nec_I]{\nec B}{
+\infer{\omega: B}{
+  \infer{\vdots}{\infer[\nec_E]{\omega: A_1}{\nec A_1}} &
+  \ldots & 
+  \infer{\vdots}{\infer[\nec_E]{\omega: A_n}{\nec A_n}}}
+}
+$$
+\end{small}
+
+\noindent
+A translation to {\ND} extended with axiom K and necessitation is shown below for the case when $n=1$:
+
+ % Assuming the axiom K and the necessitation rule $\Box_I$, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
 
@@
 \begin{prooftree}
-\AXC{$A^1 $}\noLine
+\AXC{$ $}\RightLabel{1}
+\UIC{$A_1$}\noLine
 \UIC{$\vdots$}\noLine
 \UIC{$ B$} \RightLabel{$\imp_I^1$}
-\UIC{$A\imp B$} \RightLabel{$\Box_I$}
-\UIC{$\nec(A\imp B)$}
+\UIC{$A_1\imp B$} \RightLabel{necessitation}
+\UIC{$\nec(A_1\imp B)$}
       \AXC{Axiom K}\dashedLine
-      \UIC{$\nec(A\imp B)\imp (\nec A\imp \nec B)$} \RightLabel{$\imp_E$}
-  \BIC{$\nec A\imp \nec B$}
-        \AXC{$\nec A ^1 $}\RightLabel{$\imp_E$}
+      \UIC{$\nec(A_1\imp B)\imp (\nec A_1\imp \nec B)$} \RightLabel{$\imp_E$}
+  \BIC{$\nec A_1\imp \nec B$}
+        \AXC{$\nec A_1 $}\RightLabel{$\imp_E$}
     \BIC{$\nec B$}
@@
 \end{small}
+
+\noindent
+For $n > 1$, the translation is a straightforward generalization:
+
+\begin{scriptsize}
+\begin{prooftree}
+\AXC{$ $}\RightLabel{1}
+\UIC{$A_1$}\noLine
+\UIC{$\vdots$}
+%
+\AXC{$\ldots$}
+%
+\AXC{$ $}\RightLabel{n}
+\UIC{$A_n$}\noLine
+\UIC{$\vdots$} 
+%
+\TIC{$ B$} \doubleLine \RightLabel{$\imp_I^*$}
+\UIC{$A_1\imp \ldots \imp A_n\imp B$} \RightLabel{nec.}
+\UIC{$\nec(A_1\imp \ldots \imp A_n\imp B)$}
+      \AXC{Axiom K, iterated}\doubleLine\dashedLine
+      \UIC{$\nec(A_1\imp \ldots \imp A_n\imp B)\imp (\nec A_1\imp \ldots \imp \nec A_n\imp \nec B)$} \RightLabel{$\imp_E$}
+  \BIC{$\nec A_1\imp \ldots \imp \nec A_n\imp \nec B$}
+\end{prooftree}
+\end{scriptsize}
+
+\begin{scriptsize}
+\begin{prooftree}
+  \AXC{$ $} \dashedLine
+  \UIC{$\nec A_1\imp \ldots \imp \nec A_n\imp \nec B$}
+        \AXC{$\nec A_1 \quad \ldots \quad \nec A_n$} \doubleLine \RightLabel{$\imp_E$}
+    \BIC{$\nec B$}
+\end{prooftree}
+\end{scriptsize}
+
+\end{proof}
 

### 2014-08-15 Formalizations/THF/call_tptp.sh
small update

@@ -4,15 +4,25 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
+
 timeout=60
+
+suffix=$(date "+%Y-%m-%d--%H:%M:%S")
+resultsfile=RESULTS-$suffix
+
+touch $resultsfile
+rm $resultsfile
 
 echo
 echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\) >> $resultsfile
 
 for file in "$@"
 do
 echo
+echo >> $resultsfile
     for prover in $proverlist 
     do
     function_to_fork () { 
@@ -20,12 +30,14 @@ echo
       while ! (echo $res | grep --quiet RESULT); do
           res=`$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover |grep RESULT` 
 	  echo ${file}---${prover}---${res}
+	  echo ${file}---${prover}---${res} >> $resultsfile
       done   
     }
     function_to_fork &
     done
     wait
     echo Done 
+    echo Done >> $resultsfile
 done
 
 

### 2014-08-15 Formalizations/THF/Run_ECAI2014_Experiments.sh
small update

@@ -1,1 +1,1 @@
-./call_tptp.sh T1_K_const_max.p T1_K_const_min.p T1_K_vary_max.p T1_K_vary_min.p T1_varying.p C_K_const_max.p  C_K_const_min.p  C_K_vary_max.p   C_K_vary_min.p T2_K_const_max.p  T2_K_const_min.p  T2_K_vary_max.p   T2_K_vary_min.p  T3_KB_const_max.p  T3_KB_const_min.p  T3_KB_vary_max.p   T3_KB_vary_min.p   T3_K_const_max.p   T3_K_const_min.p   T3_K_vary_max.p    T3_K_vary_min.p   MC_KB_const_max.p  MC_KB_const_min.p  MC_KB_vary_max.p   MC_KB_vary_min.p   FG_KB_const_max.p  FG_KB_const_min.p  FG_KB_vary_max.p   FG_KB_vary_min.p  MT_KB_const_max.p  MT_KB_const_min.p  MT_KB_vary_max.p   MT_KB_vary_min.p   CO_KB_const_max.p  CO_KB_vary_max.p  CO'_KB_const_max.p  CO'_KB_const_min.p  CO'_KB_vary_max.p   CO'_KB_vary_min.p  
+./call_tptp.sh T1_K_const_max.p T1_K_const_min.p T1_K_vary_max.p T1_K_vary_min.p T1_varying.p C_K_const_max.p  C_K_const_min.p  C_K_vary_max.p   C_K_vary_min.p T2_K_const_max.p  T2_K_const_min.p  T2_K_vary_max.p   T2_K_vary_min.p  T3_KB_const_max.p  T3_KB_const_min.p  T3_KB_vary_max.p   T3_KB_vary_min.p   T3_K_const_max.p   T3_K_const_min.p   T3_K_vary_max.p    T3_K_vary_min.p   MC_KB_const_max.p  MC_KB_const_min.p  MC_KB_vary_max.p   MC_KB_vary_min.p   FG_KB_const_max.p  FG_KB_const_min.p  FG_KB_vary_max.p   FG_KB_vary_min.p  MT_KB_const_max.p  MT_KB_const_min.p  MT_KB_vary_max.p   MT_KB_vary_min.p   CO_KB_const_max.p  CO_KB_vary_max.p  CO\'_KB_const_max.p  CO\'_KB_const_min.p  CO\'_KB_vary_max.p   CO\'_KB_vary_min.p  

### 2014-08-19 Formalizations/THF/call_tptp.sh
use Satallax instead of Satallax-MaLeS

@@ -4,8 +4,8 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
-proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
+proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+#proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
 
 timeout=60
 

### 2014-08-21 Formalizations/THF-FrodeBjordal/call_tptp.sh
did improve files and re-run the Bjordal experiment

@@ -4,10 +4,10 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
 #proverlist='LEO-II---1.6.2 Nitrox---2013'
 #proverlist='LEO-II---1.6.2'
-proverlist='Nitrox---2013'
+#proverlist='Nitrox---2013'
 
 timeout=200
 

### 2014-08-25 Papers/NaturalDeduction/GodProof-ND.tex
Add ToDo

@@
 
+ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
+
 \begin{theorem}
@@
 
+\setcounter{axiom}{0}
+\setcounter{lemma}{0}
+\setcounter{theorem}{0}
+\setcounter{corollary}{0}
+
+ToDo: natural language explanation of the proof in Anderson.
+
 The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
@@
 
+\setcounter{axiom}{0}
+\setcounter{lemma}{0}
+\setcounter{theorem}{0}
+\setcounter{corollary}{0}
+
 \subsection{Possibly, God Exists}
-
-\setcounter{axiom}{0}
 
@@
 
+Anderson's footnote 5 on KB
+
+ToDo: talk about anderson's footnote 2, where he argues that an anonymous referee knew the simpler proof of T1.
+
+Hajek (magari and others) on non-provability of ex x. G(x) in KD45.
+
 \begin{thebibliography}{9}

### 2014-08-25 Papers/Universalis/ModalCollapse.tex
Add stub for Logica Universalis paper

% ------------------------------------------------------------------------
% bjourdoc.tex for birkjour.cls*******************************************
% ------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{birkjour}
%
%
% THEOREM Environments (Examples)-----------------------------------------
%
 \newtheorem{thm}{Theorem}[section]
 \newtheorem{cor}[thm]{Corollary}
 \newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
 \theoremstyle{remark}
 \newtheorem{rem}[thm]{Remark}
 \newtheorem*{ex}{Example}
 \numberwithin{equation}{section}

\begin{document}

%-------------------------------------------------------------------------
% editorial commands: to be inserted by the editorial office
%
%\firstpage{1} \volume{228} \Copyrightyear{2004} \DOI{003-0001}
%
%
%\seriesextra{Just an add-on}
%\seriesextraline{This is the Concrete Title of this Book\br H.E. R and S.T.C. W, Eds.}
%
% for journals:
%
%\firstpage{1}
%\issuenumber{1}
%\Volumeandyear{1 (2004)}
%\Copyrightyear{2004}
%\DOI{003-xxxx-y}
%\Signet
%\commby{inhouse}
%\submitted{March 14, 2003}
%\received{March 16, 2000}
%\revised{June 1, 2000}
%\accepted{July 22, 2000}
%
%
%
%---------------------------------------------------------------------------
%Insert here the title, affiliations and abstract:
%

\title[Modal Collapse]
 {The Ontological Modal Colapse \\ 
 as a Collapse of the Square of Opposition}

%----------Author 1
\author[Benzm\"uller]{Christoph Benzm\"uller}

\address{%
Viaduktstr. 42\\
P.O. Box 133\\
CH 4010 Basel\\
Switzerland}

\email{c.benzmueller@gmail.com}

\thanks{This work was completed with the support of ToDo}

%----------Author 2
\author[Woltzenlogel-Paleo]{Bruno Woltzenlogel Paleo}
\address{ }
\email{bruno.wp@gmail.com}
%----------classification, keywords, date
\subjclass{
Primary 03A02;  % Philosophical aspects of logic and foundations
Secondary 68T02 % Artificial Intelligence
}

\keywords{Modal Logics, Higher-Order Modal Logics}

\date{August 21, 2014}
%----------additions
\dedicatory{ }
%%% ----------------------------------------------------------------------

\begin{abstract}
This work presents.
\end{abstract}

%%% ----------------------------------------------------------------------
\maketitle
%%% ----------------------------------------------------------------------
%\tableofcontents
\section{Introduction}

\begin{defn}
This serves as environment for definitions. Note that the text
appears not in italics.
\end{defn}

\begin{equation}\label{testequation}
\text{This is a sample equation: } c^2=a^2+b^2
\end{equation}

\begin{thm}[Main Theorem]
In contrast to definitions, theorems appear typeset in italics as
it has become more or less standard in most textbooks and
monographs. Equations can be cited using the \verb+\eqref+ command which
automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
\end{thm}

\begin{proof}
A special environment is predefined: the \textit{proof} environment. Please use
\begin{verbatim}\begin{proof}\end{verbatim}
proof of the statement
\begin{verbatim}\end{proof}\end{verbatim}
for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
\end{proof}

There are two known problems with the placement of the end-of-proof sign:

\begin{enumerate}
  \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
be placed in the line below; if you want to avoid this, write your line in the form
\begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
which results in

\begin{proof}
$$displayed math line \eqno\qedhere$$
\end{proof}
\item if your proof ends with an aligned displayed environment, the command
\verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
\begin{verbatim}
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{verbatim}
results in
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{enumerate}
Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.

\begin{rem}
Additional comments are being typeset without boldfaced entrance
word as they may be minor important.
\end{rem}

\begin{ex}
For some constructs, even no number is required.
\end{ex}

Displayed equations may be numbered like the following one:
\begin{equation}
\sqrt{1-\sin^2(x)}=|\cos(x)|.
\end{equation}

\subsection{Conclusions}
% ------------------------------------------------------------------------

% \subsection*{Acknowledgment}
% Many thanks to ...

\begin{thebibliography}{1}
\bibitem{test} A. B. C. Test, \textit{On a Test.} J. of Testing
\textbf{88} (2000), 100--120.
\bibitem{latex} G. Gr\"atzer, \textit{Math into \LaTeX.} 3rd Edition,
Birkh\"auser, 2000.
\end{thebibliography}

% ------------------------------------------------------------------------
\end{document}
% ------------------------------------------------------------------------

### 2014-08-28 Papers/Universalis/ModalCollapse.tex
Structure paper and write introduction

@@
 \documentclass{birkjour}
+
+\usepackage{url}
+\usepackage{commands}
 %
@@
  \numberwithin{equation}{section}
+
+\def\HOML{\entity{HOML}\xspace}
+\def\HOL{\entity{HOL}\xspace}
 
@@
 \title[Modal Collapse]
- {The Ontological Modal Colapse \\ 
+ {The Ontological Modal Collapse \\ 
  as a Collapse of the Square of Opposition}
@@
 \address{%
-Viaduktstr. 42\\
-P.O. Box 133\\
-CH 4010 Basel\\
-Switzerland}
+Department of Mathematics and Computer Science\\
+Arnimallee 7 \\
+Room 115 \\
+14195 Berlin \\
+Germany
+}
 
@@
 %----------Author 2
-\author[Woltzenlogel-Paleo]{Bruno Woltzenlogel Paleo}
-\address{ }
+\author[Weber]{\\ Leon Weber}
+\address{ 
+ToDo \\
+ToDo \\
+ToDo Berlin \\
+Germany
+}
+\email{leon.weber@fu-berlin.de}
+
+%----------Author 3
+\author[Woltzenlogel-Paleo]{\\ Bruno Woltzenlogel Paleo}
+\address{ 
+Favoritenstra{\ss}e 9 \\
+Room HA0402 \\
+1040 Wien \\
+Austria
+}
 \email{bruno.wp@gmail.com}
+
 %----------classification, keywords, date
 \subjclass{
-Primary 03A02;  % Philosophical aspects of logic and foundations
-Secondary 68T02 % Artificial Intelligence
+Prim. 03A02;  % Philosophical aspects of logic and foundations
+Sec. 68T02 % Artificial Intelligence
 }
 
-\keywords{Modal Logics, Higher-Order Modal Logics}
+\keywords{Modal Logics, Higher-Order Logics, Ontological Argument}
 
-\date{August 21, 2014}
+\date{August 30, 2014}
 %----------additions
@@
 \begin{abstract}
-This work presents.
+The \emph{modal collapse} that afflicts G\"odel's modal ontological 
+argument for God's existence is discussed from the perspective of the 
+modal square of opposition. Furthermore, a computer-assisted verification 
+of the claims that the emendations by Anderson and by Frode are immune to 
+the modal collapse is presented.
 \end{abstract}
@@
 
-\begin{defn}
-This serves as environment for definitions. Note that the text
-appears not in italics.
-\end{defn}
+Attempts to prove the
+existence (or non-existence) of God by means of abstract, ontological
+arguments are an old tradition in western philosophy, with contributions by several prominent philosophers, including St. Anselm of
+Canterbury, Descartes and Leibniz.
+Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as in Fig.
+\ref{fig:scott}.
 
-\begin{equation}\label{testequation}
-\text{This is a sample equation: } c^2=a^2+b^2
-\end{equation}
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:
+  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
+\item[D1] A \emph{God-like} being possesses all positive properties: 
+  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
+\item[A3]  The property of being God-like is positive: 
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+  \exq y \phi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Scott's version of G\"odel's ontological argument \cite{ScottNotes}.\label{fig:scott}} 
+\end{figure}
 
-\begin{thm}[Main Theorem]
-In contrast to definitions, theorems appear typeset in italics as
-it has become more or less standard in most textbooks and
-monographs. Equations can be cited using the \verb+\eqref+ command which
-automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
-\end{thm}
+% Ontological arguments, for or against the existence of God,
+% illustrate well an essential aspect of metaphysics: some (necessary) facts
+% for our existing world are deduced by purely a priori, analytical means from some
+% abstract definitions and axioms. % Contingent truths are to be
+% distinguished from necessary truths.
 
-\begin{proof}
-A special environment is predefined: the \textit{proof} environment. Please use
-\begin{verbatim}\begin{proof}\end{verbatim}
-proof of the statement
-\begin{verbatim}\end{proof}\end{verbatim}
-for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
-\end{proof}
+G\"{o}del defines God (see Fig.~\ref{fig1}) as a being who possesses all \emph{positive}
+properties.  He does not extensively discuss what positive properties
+are, but instead he states a few reasonable (but debatable) axioms
+that they should satisfy.  
+The overall idea of G{\"o}del's proof is in the tradition of Anselm's
+argument, who defined God as some entity of which nothing greater can be
+conceived. Anselm argued that existence in the actual world would
+make such an assumed being even greater; hence, by definition God must
+exist. G{\"o}del's ontological argument is clearly related to this
+reasoning pattern. However, it also tries to fix some fundamental
+weaknesses in Anselm's work. For example, G{\"o}del explicitly proves
+that God's existence is possible, which has been a basic assumption for
+Anselm. Because of this, Anselm's argument has been
+criticized as incomplete by Leibniz. Leibniz claimed that the
+assumption should be derivable from the definition of God as a perfect
+being and from the notion of perfection.  G{\"o}del's proof 
+addresses this critique, and it also addresses Kant's objection that existence should not be treated as a predicate.
 
-There are two known problems with the placement of the end-of-proof sign:
+Nevertheless, G{\"o}del's work
+still leaves room for criticism. In particular, his axioms are so
+strong that they entail a \emph{modal collapse} 
+\cite{ToDo:sobel1987?,sobel2004logic}: 
+everything that is the case is so necessarily. 
+There has been an impressive body of recent and ongoing
+work (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,ToDo:FrodeAndHajek,ContemporaryBibliography} and the references therein)
+proposing solutions for the modal collapse. 
+The goal of this paper is to present a computer-assisted analysis of 
+the solutions proposed by Anderson and by Frode.
+The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}.
+This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.
 
-\begin{enumerate}
-  \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
-be placed in the line below; if you want to avoid this, write your line in the form
-\begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
-which results in
+\section{The Modal Collapse}
 
-\begin{proof}
-$$displayed math line \eqno\qedhere$$
-\end{proof}
-\item if your proof ends with an aligned displayed environment, the command
-\verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
-\begin{verbatim}
-\begin{align*}
-\alpha&=\beta+\gamma\\
-&=\delta+\epsilon\tag*{\qed}
-\end{align*}
-\end{verbatim}
-results in
-\begin{align*}
-\alpha&=\beta+\gamma\\
-&=\delta+\epsilon\tag*{\qed}
-\end{align*}
-\end{enumerate}
-Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
-and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[MC] Everything that is the case is so necessarily:
+  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
+\item[MCC] Modalities collapse:
+  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
+\end{figure}
 
-\begin{rem}
-Additional comments are being typeset without boldfaced entrance
-word as they may be minor important.
-\end{rem}
+\clearpage
 
-\begin{ex}
-For some constructs, even no number is required.
-\end{ex}
+\section{Anderson's Emendation}
 
-Displayed equations may be numbered like the following one:
-\begin{equation}
-\sqrt{1-\sin^2(x)}=|\cos(x)|.
-\end{equation}
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \phi [P(\phi) \imp \neg P(\neg \phi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being possesses all positive properties: 
+  $$\hol{G_A(x) \biimp \forall \phi [P(\phi) \biimp \nec \phi(x)]}$$ 
+\item[A3]  The property of being God-like is positive: 
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\essA{\phi}{x} \biimp \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\phi(y) \imp \psi(y))]}$$ 
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+  \exq y \phi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
+\end{figure}
 
-\subsection{Conclusions}
+\clearpage
+
+\section{Frode's Alternative}
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:
+  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
+\item[D1] A \emph{God-like} being possesses all positive properties: 
+  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
+\item[A3]  The property of being God-like is positive: 
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+  \exq y \phi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
+\end{figure}
+
+\clearpage
+
+\section{Other Solutions}
+
+ToDo: Fitting
+
+\section{Conclusions}
+
+Various slightly different versions of
+axioms and definitions have been considered by G\"{o}del and by
+several philosophers who commented on his proof
+(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+
+In theoretical philosophy, formal logical confrontations with such
+ontological arguments had been so far (mainly) limited to paper
+and pen.  Up to now, the use of computers was prevented, because the
+logics of the available theorem proving systems were not expressive
+enough to formalize the abstract concepts adequately. G{\"o}del's proof
+uses, for example, a complex higher-order modal logic (\HOML) to handle
+concepts such as \emph{possibility} and \emph{necessity} and to support
+quantification over individuals and properties.
+
+ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
+
+% \begin{defn}
+% This serves as environment for definitions. Note that the text
+% appears not in italics.
+% \end{defn}
+
+% \begin{equation}\label{testequation}
+% \text{This is a sample equation: } c^2=a^2+b^2
+% \end{equation}
+
+% \begin{thm}[Main Theorem]
+% In contrast to definitions, theorems appear typeset in italics as
+% it has become more or less standard in most textbooks and
+% monographs. Equations can be cited using the \verb+\eqref+ command which
+% automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
+% \end{thm}
+
+% \begin{proof}
+% A special environment is predefined: the \textit{proof} environment. Please use
+% \begin{verbatim}\begin{proof}\end{verbatim}
+% proof of the statement
+% \begin{verbatim}\end{proof}\end{verbatim}
+% for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
+% \end{proof}
+
+% There are two known problems with the placement of the end-of-proof sign:
+
+% \begin{enumerate}
+%   \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
+% be placed in the line below; if you want to avoid this, write your line in the form
+% \begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
+% which results in
+
+% \begin{proof}
+% $$displayed math line \eqno\qedhere$$
+% \end{proof}
+% \item if your proof ends with an aligned displayed environment, the command
+% \verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
+% \begin{verbatim}
+% \begin{align*}
+% \alpha&=\beta+\gamma\\
+% &=\delta+\epsilon\tag*{\qed}
+% \end{align*}
+% \end{verbatim}
+% results in
+% \begin{align*}
+% \alpha&=\beta+\gamma\\
+% &=\delta+\epsilon\tag*{\qed}
+% \end{align*}
+% \end{enumerate}
+% Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
+% and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.
+
+% \begin{rem}
+% Additional comments are being typeset without boldfaced entrance
+% word as they may be minor important.
+% \end{rem}
+
+% \begin{ex}
+% For some constructs, even no number is required.
+% \end{ex}
+
+% Displayed equations may be numbered like the following one:
+% \begin{equation}
+% \sqrt{1-\sin^2(x)}=|\cos(x)|.
+% \end{equation}
+
 % ------------------------------------------------------------------------
@@
 \begin{thebibliography}{1}
-\bibitem{test} A. B. C. Test, \textit{On a Test.} J. of Testing
-\textbf{88} (2000), 100--120.
-\bibitem{latex} G. Gr\"atzer, \textit{Math into \LaTeX.} 3rd Edition,
-Birkh\"auser, 2000.
+
+\bibitem{Adams}
+R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
+  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
+
+\bibitem{AndersonGettings}
+A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
+  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
+  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
+
+\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
+C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
+  Faith and Philosophy}, {\bf 7}(3), (1990).
+
+\bibitem{Andrews:gmae72}
+P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
+  Logic}, {\bf 37}(2),  395--397, (1972).
+
+\bibitem{andrewsSEP}
+P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
+  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
+
+\bibitem{C36}
+C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
+  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
+  pp. 232--233, Rio de Janeiro, Brazil, (2013).
+
+\bibitem{B5}
+C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
+  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
+  Elsevier, (2014).
+\newblock Forthcoming; preliminary version available at
+  {http://christoph-benzmueller.de/papers/B5.pdf}.
+
+\bibitem{C34}
+C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
+  provers for first-order modal logics', in {\em Proc. of the 20th European
+  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
+
+\bibitem{B9}
+C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
+  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
+  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
+  386--406, College Publications, (2008).
+
+\bibitem{J23}
+C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
+  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
+
+\bibitem{LEO-II}
+C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
+  cooperative automatic theorem prover for higher-order logic', in {\em
+  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
+
+\bibitem{J30}
+C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
+  automation of {G{\"o}del's} proof of {God's} existence', {\em
+  arXiv:1308.4526}, (2013).
+
+\bibitem{J28}
+C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
+  {\em Archive of Formal Proofs}, (2013).
+
+\bibitem{W50}
+C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
+  in {\em Proceedings of the 10th International Workshop on the Implementation
+  of Logics}, EPiC Series. EasyChair, (2013).
+\newblock Invited abstract.
+
+\bibitem{Coq}
+Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
+  Development}}, Springer, 2004.
+
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
+  higher-order logic based on a relational model finder', in {\em Proc. of ITP
+  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
+
+\bibitem{Satallax}
+C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
+  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
+
+\bibitem{ContemporaryBibliography}
+R.~Corazzon.
+\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
+  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
+
+\bibitem{Fitting}
+M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
+
+\bibitem{fitting98}
+M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
+  {\em Synthese Library}, Kluwer, 1998.
+
+\bibitem{Gallin75}
+D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
+
+\bibitem{garbacz12:_prover_simpl_expal_away}
+P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
+  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
+
+\bibitem{GoedelNotes}
+K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{Henkin50}
+L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
+  Logic}, {\bf 15}(2),  81--91, (1950).
+
+\bibitem{homl}
+R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
+  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
+
+\bibitem{Isabelle}
+T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
+  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
+
+\bibitem{oppenheimera11}
+P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
+  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
+  89}(2),  333--349, (2011).
+
+\bibitem{rushby13}
+J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
+  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
+
+\bibitem{Schulz:AICOM-2002}
+S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
+  15}(2),  111--126, (2002).
+
+\bibitem{ScottNotes}
+D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{sobel2004logic}
+J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
+  Cambridge U. Press, 2004.
+
+\bibitem{sutcliffe2009tptp}
+G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
+  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
+  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
+  {\bf 3}(1),  1--27, (2010).
+
+\bibitem{J29}
+B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
+  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
+  (2013).
+
 \end{thebibliography}

### 2014-08-30 Papers/Universalis/ModalCollapse.tex
Complete formal description of Anderson's Emendation

@@
 \item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
+  $$\hol{G(x) \equiv \forall \phi [P(\phi) \imp \phi(x)]}$$ 
 \item[A3]  The property of being God-like is positive: 
@@
   $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \equiv \phi(x) \wedge \allq
   \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
@@
 \item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+  $$\hol{\NE(x) \equiv \allq \phi [\ess{\phi}{x} \imp \nec
   \exq y \phi(y)]}$$
@@
   $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G_A(x) \biimp \forall \phi [P(\phi) \biimp \nec \phi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \phi [P(\phi) \biimp \nec \phi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
 \item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
@@
   $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\essA{\phi}{x} \biimp \allq
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\phi}{x} \equiv \allq
   \psi [\nec \psi(x) \biimp \nec \allq y (\phi(y) \imp \psi(y))]}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \phi [\essA{\phi}{x} \imp \nec
   \exq y \phi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}

### 2014-08-31 Papers/Universalis/ModalCollapse.tex
Write paragraph about Modal Collapse

--- 2014-08-28 Papers/Universalis/ModalCollapse.tex
@@
 Canterbury, Descartes and Leibniz.
-Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as in Fig.
+Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as shown in Fig.
 \ref{fig:scott}.
@@
   both:
-  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
+  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
 \item[A2] A property necessarily implied by a
   positive property is positive:
-  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
   \imp \psi(x)]) \imp P(\psi)]}$$
 \item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
 \item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
+  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
 \item[A3]  The property of being God-like is positive: 
@@
 \item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
 \item[T2]  Being God-like is an essence of any
@@
 \item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$$
+  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
 \item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
+\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
+%
 \item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
@@
 
+A crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary.
+This is Lemma L2 in G\"odel's proof.
+In the modal square of opposition (Fig.
+\ref{ToDo}), this is an unusual situation in which the \textbf{I} corner implies and entails the \textbf{A} corner, for the particular case when $\phi$ is $\exq x G(x)$.
+G\"odel's proof shows that his axioms are strong enough to invert the direction of entailment for the particular $\phi$ at issue.
+The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$.
+That's essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively.
+The modal collapse is a collapse of the modal square: it shows that assuming G\"odel's axioms, the subcontraries entail the contraries.
+
+Various ontological...
+
+ToDo: include figure of Modal Square of Opposition.
+
+ToDo: discuss what the arrows mean (entailment or implication?) in the modal square. This is not as obvious as in the non-modal square, because the deduction theorem does not hold for modal logics.
+
+ToDo: the following should always hold (check this in Isabelle!): 
+$$
+\allq \phi [\exq x G(x) \imp \phi] \imp [\phi \imp \nec \phi]
+$$
+$$
+\allq \phi [\exq x G(x) \imp \phi] \imp [\pos \phi \imp \nec \phi]
+$$
+
 \begin{figure}[t]
@@
   $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
-\item[MCC] Modalities collapse:
+\item[MC'] Everything that is possible is necessary:
+  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
+\item[MC''] Modalities collapse:
   $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
@@
 \item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \phi [P(\phi) \imp \neg P(\neg \phi)]}$$ 
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
 \item[A2] A property necessarily implied by a
   positive property is positive:
-  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
   \imp \psi(x)]) \imp P(\psi)]}$$
 \item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G_A(x) \biimp \forall \phi [P(\phi) \biimp \nec \phi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
 \item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
 \item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\essA{\phi}{x} \biimp \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\phi(y) \imp \psi(y))]}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}
@@
 \end{figure}
+
+I have meanwhile added another file ``Anderson\_var\_partial.thy''.
+In this version the actualist e-Quantifier is used exclusively in the definition of NE and everywhere else ithe possibilist Quantifiers are used.
+It is still unclear to me, what is actually meant.
+Reading Anderson 1990, last sentence on p.
+302, it could mean that ``Anderson\_var\_partial.thy'' is the correct version; but when reading AndersonGettings, then one could think that ``Anderson\_var.thy'' is meant.
+But ok we have now both versions.
+Interestingly in ``Anderson\_var\_partial.thy'' A5 is not inferable anymore, and this also holds for ``anderson\_implies\_fuhrmann".
+Nitpick finds countermodels.
 
@@
 quantification over individuals and properties.
+
+controversies, care with parameters
 

### 2014-08-31 Papers/Universalis/ModalCollapse.tex
Remove wrong claims

@@
 The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$.
-That's essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively.
-The modal collapse is a collapse of the modal square: it shows that assuming G\"odel's axioms, the subcontraries entail the contraries.
+That is essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively.
+The modal collapse is a collapse of the modal square: it shows that, assuming G\"odel's axioms, the subcontraries entail the contraries.
 
@@
 ToDo: discuss what the arrows mean (entailment or implication?) in the modal square. This is not as obvious as in the non-modal square, because the deduction theorem does not hold for modal logics.
-
-ToDo: the following should always hold (check this in Isabelle!): 
-$$
-\allq \phi [\exq x G(x) \imp \phi] \imp [\phi \imp \nec \phi]
-$$
-$$
-\allq \phi [\exq x G(x) \imp \phi] \imp [\pos \phi \imp \nec \phi]
-$$
 
@@
 \item[MC'] Everything that is possible is necessary:
-  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
+  $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
 \item[MC''] Modalities collapse:

### 2014-09-05 Papers/Universalis/ModalCollapse.tex
Remove incomplete paragraph
Remove discussion of the proofs by Anderson and Frode

@@
 
-%----------Author 2
-\author[Weber]{\\ Leon Weber}
-\address{ 
-ToDo \\
-ToDo \\
-ToDo Berlin \\
-Germany
-}
-\email{leon.weber@fu-berlin.de}
-
 %----------Author 3
-\author[Woltzenlogel-Paleo]{\\ Bruno Woltzenlogel Paleo}
+\author[Woltzenlogel-Paleo]{Bruno Woltzenlogel Paleo}
 \address{ 
@@
 Canterbury, Descartes and Leibniz.
-Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as shown in Fig.
+Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, shown in Fig.
 \ref{fig:scott}.
@@
 
-Various ontological...
-
 ToDo: include figure of Modal Square of Opposition.
-
-ToDo: discuss what the arrows mean (entailment or implication?) in the modal square. This is not as obvious as in the non-modal square, because the deduction theorem does not hold for modal logics.
 
@@
 \clearpage
-
-\section{Anderson's Emendation}
-
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
-%
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-\end{figure}
-
-I have meanwhile added another file ``Anderson\_var\_partial.thy''.
-In this version the actualist e-Quantifier is used exclusively in the definition of NE and everywhere else ithe possibilist Quantifiers are used.
-It is still unclear to me, what is actually meant.
-Reading Anderson 1990, last sentence on p.
-302, it could mean that ``Anderson\_var\_partial.thy'' is the correct version; but when reading AndersonGettings, then one could think that ``Anderson\_var.thy'' is meant.
-But ok we have now both versions.
-Interestingly in ``Anderson\_var\_partial.thy'' A5 is not inferable anymore, and this also holds for ``anderson\_implies\_fuhrmann".
-Nitpick finds countermodels.
-
-\clearpage
-
-\section{Frode's Alternative}
-
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
-  both:
-  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
-\end{figure}
-
-\clearpage
-
-\section{Other Solutions}
-
-ToDo: Fitting
 

### 2014-09-06 Papers/Universalis/ModalCollapse.tex
Make small improvements in the paper

@@
 
-\thanks{This work was completed with the support of ToDo}
+%\thanks{This work was completed with the support of ToDo}
 
@@
 argument for God's existence is discussed from the perspective of the 
-modal square of opposition. Furthermore, a computer-assisted verification 
-of the claims that the emendations by Anderson and by Frode are immune to 
-the modal collapse is presented.
+modal square of opposition.
 \end{abstract}
@@
 
-G\"{o}del defines God (see Fig.~\ref{fig1}) as a being who possesses all \emph{positive}
-properties.  He does not extensively discuss what positive properties
-are, but instead he states a few reasonable (but debatable) axioms
-that they should satisfy.  
+G\"{o}del defines God as a being who possesses all \emph{positive}
+properties and states a few reasonable (but debatable) axioms
+that such properties should satisfy.  
 The overall idea of G{\"o}del's proof is in the tradition of Anselm's
@@
 conceived. Anselm argued that existence in the actual world would
-make such an assumed being even greater; hence, by definition God must
-exist. G{\"o}del's ontological argument is clearly related to this
-reasoning pattern. However, it also tries to fix some fundamental
-weaknesses in Anselm's work. For example, G{\"o}del explicitly proves
-that God's existence is possible, which has been a basic assumption for
-Anselm. Because of this, Anselm's argument has been
-criticized as incomplete by Leibniz. Leibniz claimed that the
-assumption should be derivable from the definition of God as a perfect
-being and from the notion of perfection.  G{\"o}del's proof 
-addresses this critique, and it also addresses Kant's objection that existence should not be treated as a predicate.
+make such an assumed being even greater; hence, by definition, God must
+exist.
+However, for Anselm existence was treated as a predicate and the possibility of God's existence was assumed as granted.
+These issues were criticized by Kant and Leibniz, respectively, and succesfully addressed by G\"odel.
 
@@
 proposing solutions for the modal collapse. 
-The goal of this paper is to present a computer-assisted analysis of 
-the solutions proposed by Anderson and by Frode.
-The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}.
-This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.
+The goal of this short note is merely to remark that the modal collapse is a collapse of the modal square of opposition.
 
-\section{The Modal Collapse}
+\section{The Collapse of the Modal Square}
 
@@
 ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
+
+The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}.
+This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.
 

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Improve text about modal collapse

@@
 
-\section{The Collapse of the Modal Square}
+\section{A Collapse of the Modal Square}
 
-A crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary.
+A crucial step of most ontological arguments is the claim that 
+if God's existence is possible, then it is necessary. 
 This is Lemma L2 in G\"odel's proof. 
-In the modal square of opposition (Fig.
-\ref{ToDo}), this is an unusual situation in which the \textbf{I} corner implies and entails the \textbf{A} corner, for the particular case when $\phi$ is $\exq x G(x)$.
-G\"odel's proof shows that his axioms are strong enough to invert the direction of entailment for the particular $\phi$ at issue.
-The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$.
-That is essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively.
-The modal collapse is a collapse of the modal square: it shows that, assuming G\"odel's axioms, the subcontraries entail the contraries.
+In the modal square of opposition (Fig. \ref{ToDo}), 
+this is an unusual situation in which the \textbf{I} corner 
+must imply and entail the \textbf{A} corner, 
+in the particular case when $\phi$ is $\exq x G(x)$. 
+G\"odel's proof shows that his axioms are strong enough 
+to invert the direction of entailment for the sentence at issue. 
+The question, however, is whether the axioms are not too strong, 
+also allowing the inverted entailment for arbitrary $\phi$. 
+That is essentially the question asked by Sobel \cite{Sobel1987}; 
+and his proof of the modal collapse (MC) provides an affirmative 
+answer. It is possible to show that this form of the modal 
+collapse entails (in modal logic K) a collapse of the modal 
+square (MC'): the subcontraries entail (and even imply) the 
+contraries. Furthermore, in any modal logic where where the 
+axiom T holds (i.e. the accessibility relation is reflexive), 
+a total collapse of the modalities (MC'') occurs.
 
@@
   $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
-\item[MC''] Modalities collapse:
+\item[T] Everything that is necessary is the case:
+  $$\hol{\allq \phi [\nec \phi \imp \phi]}$$ 
+\item[EI] (Modal Existential Import):
+  $$\hol{\pos \top}$$ 
+\item[AI] Everything that is necessary is possible:
+  $$\hol{\allq \phi [\nec \phi \imp \pos \phi]}$$ 
+\item[MC''] Modalities collapse completely:
   $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Add Square Image

@@
 proposing solutions for the modal collapse. 
-The goal of this short note is merely to remark that the modal collapse is a collapse of the modal square of opposition.
+The goal of this short note is to discuss the modal collapse from the point of view of the modal square of opposition.
 
@@
 This is Lemma L2 in G\"odel's proof. 
-In the modal square of opposition (Fig. \ref{ToDo}), 
+In the modal square of opposition (Fig. \ref{fig:square}), 
 this is an unusual situation in which the \textbf{I} corner 
@@
 collapse entails (in modal logic K) a collapse of the modal 
-square (MC'): the subcontraries entail (and even imply) the 
-contraries. Furthermore, in any modal logic where where the 
-axiom T holds (i.e. the accessibility relation is reflexive), 
-a total collapse of the modalities (MC'') occurs.
+square (MC'), causing the subcontraries to entail (and even imply) their respective contraries.
+Normally, as shown in Fig.
+\ref{fig:square} in the modal square of opposition, only the other direction of entailment holds: the contraries entail their subcontraries, assuming the \emph{modal existential import} ExImp.
+
+Furthermore, in any modal logic where the 
+axiom T holds (i.e.
+where the accessibility relation is reflexive), a total collapse of the modalities (MC'') occurs.
+In particular, under this stronger form of modal collapse, the contraries entail their subcontraries even without the existential import.
+
+\begin{figure}[t]
+\includegraphics[width=0.9\textwidth]{./SquarePics/ModalSquare.png}
+\caption{Modal Square of Opposition.}
+\label{fig:square}
+\end{figure}
 
@@
   $$\hol{\allq \phi [\nec \phi \imp \phi]}$$ 
-\item[EI] (Modal Existential Import):
+\item[ExImp] (Modal Existential Import):
   $$\hol{\pos \top}$$ 
@@
 \end{figure}
+
+\clearpage
+
+\begin{Huge}
+
+$$
+\nec \ \phi
+$$
+
+$$
+\pos \ \phi
+$$
+
+$$
+\neg \ \nec \ \phi
+$$
+
+$$
+\neg \ \pos \ \phi
+$$
+
+$$
+\pos \ \exists x \ G(x)
+$$
+
+$$
+\nec \ \exists x \ G(x)
+$$
+
+\end{Huge}
 

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Boldify names of axioms and theorems

@@
 \begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
+\item[\textbf{A1}] Either a property or its negation is positive, but not
   both:
   $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
-\item[A2] A property necessarily implied by a
+\item[\textbf{A2}] A property necessarily implied by a
   positive property is positive:
@@
   \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
+\item[\textbf{T1}] Positive properties are possibly exemplified: 
   $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
+\item[\textbf{D1}] A \emph{God-like} being possesses all positive properties: 
   $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
+\item[\textbf{A3}]  The property of being God-like is positive: 
   $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
+\item[\textbf{C\phantom{1}}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[\textbf{A4}]  Positive properties are necessarily positive: 
   $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
+\item[\textbf{D2}] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
   \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
+\item[\textbf{T2}]  Being God-like is an essence of any
   God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+\item[\textbf{D3}] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
   $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
   \exq y \varphi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
+\item[\textbf{A5}] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[\textbf{L1}] If a god-like being exists, then necessarily a god-like being exists: 
   $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
-\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
+\item[\textbf{L2}] If possibly a god-like being exists, then necessarily a god-like being exists: 
   $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
 %
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\item[\textbf{T3}] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
 \end{itemize}
@@
 if God's existence is possible, then it is necessary. 
-This is Lemma L2 in G\"odel's proof. 
+This is Lemma \textbf{L2} in G\"odel's proof. 
 In the modal square of opposition (Fig. \ref{fig:square}), 
@@
 That is essentially the question asked by Sobel \cite{Sobel1987}; 
-and his proof of the modal collapse (MC) provides an affirmative 
+and his proof of the modal collapse (\textbf{MC}) provides an affirmative 
 answer. It is possible to show that this form of the modal 
-collapse entails (in modal logic K) a collapse of the modal 
-square (MC'), causing the subcontraries to entail (and even imply) their respective contraries.
+collapse entails (in modal logic \textbf{\emph{K}}) a collapse of the modal 
+square (\textbf{MC'}), causing the subcontraries to entail (and even imply) their respective contraries.
 Normally, as shown in Fig.
-\ref{fig:square} in the modal square of opposition, only the other direction of entailment holds: the contraries entail their subcontraries, assuming the \emph{modal existential import} ExImp.
+\ref{fig:square}, in the modal square of opposition only the other direction of entailment holds: the contraries entail their subcontraries, assuming the \emph{modal existential import} \textbf{ExImp}.
 
-Furthermore, in any modal logic where the 
-axiom T holds (i.e.
-where the accessibility relation is reflexive), a total collapse of the modalities (MC'') occurs.
-In particular, under this stronger form of modal collapse, the contraries entail their subcontraries even without the existential import.
+Moreover, in any modal logic where the 
+axiom \textbf{T} holds (i.e.
+where the accessibility relation is reflexive), a total collapse of the modalities (\textbf{MC''}) occurs.
+Interestingly, under this stronger form of modal collapse, the contraries entail their subcontraries even without the existential import.
 
@@
 
-ToDo: include figure of Modal Square of Opposition.
-
 \begin{figure}[t]
@@
 \begin{itemize}
-\item[MC] Everything that is the case is so necessarily:
+\item[\textbf{MC}] Everything that is the case is so necessarily:
   $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
-\item[MC'] Everything that is possible is necessary:
+\item[\textbf{MC'}] Everything that is possible is necessary:
   $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
-\item[T] Everything that is necessary is the case:
+\item[\textbf{T}] Everything that is necessary is the case:
   $$\hol{\allq \phi [\nec \phi \imp \phi]}$$ 
-\item[ExImp] (Modal Existential Import):
+\item[\textbf{ExImp}] (Modal Existential Import):
   $$\hol{\pos \top}$$ 
-\item[AI] Everything that is necessary is possible:
+\item[\textbf{AI}] Everything that is necessary is possible:
   $$\hol{\allq \phi [\nec \phi \imp \pos \phi]}$$ 
-\item[MC''] Modalities collapse completely:
+\item[\textbf{MC''}] Modalities collapse completely:
   $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
@@
 } \vskip-.5em
-\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
+\caption{Modal Collapse}
+\label{fig:collapse} 
 \end{figure}
 
-\clearpage
+\newcommand{\collapse}[1]{\mathit{collapse}(#1)}
 
-\begin{Huge}
+The
 
-$$
-\nec \ \phi
-$$
-
-$$
-\pos \ \phi
-$$
-
-$$
-\neg \ \nec \ \phi
-$$
-
-$$
-\neg \ \pos \ \phi
-$$
-
-$$
-\pos \ \exists x \ G(x)
-$$
-
-$$
-\nec \ \exists x \ G(x)
-$$
-
-\end{Huge}
+Let $collapse{\phi} \equiv \phi \imp \nec \phi$. It is known that  Variants of the ontological argument
 

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Add paragraph about the class of collapsing propositions

@@
 \item[\textbf{MC}] Everything that is the case is so necessarily:
-  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
+  $\hol{\allq \phi [\phi \imp \nec \phi]}$ \\
 \item[\textbf{MC'}] Everything that is possible is necessary:
-  $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
+  $\hol{\allq \phi [\pos \phi \imp \nec \phi]}$  \\
 \item[\textbf{T}] Everything that is necessary is the case:
-  $$\hol{\allq \phi [\nec \phi \imp \phi]}$$ 
+  $\hol{\allq \phi [\nec \phi \imp \phi]}$ \\
 \item[\textbf{ExImp}] (Modal Existential Import):
-  $$\hol{\pos \top}$$ 
+  $\hol{\pos \top}$ \\
 \item[\textbf{AI}] Everything that is necessary is possible:
-  $$\hol{\allq \phi [\nec \phi \imp \pos \phi]}$$ 
+  $\hol{\allq \phi [\nec \phi \imp \pos \phi]}$ \\
 \item[\textbf{MC''}] Modalities collapse completely:
-  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
+  $\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$
 \end{itemize}
@@
 
-The
+Although G\"odel's axioms lead to modal collapse, 
+there are several variants (e.g. \cite{anderson,bjordal}) 
+that are known to be immune to the modal collapse. 
+This means there must be at least one proposition $\phi$ 
+such that $\phi \imp \nec \phi$ 
+(from now on abbreviated $\collapse{\phi}$) is not valid 
+under the axioms and definitions used by the variant. 
+But if the variant is sufficiently similar to G\"odel's argument, 
+following the path deriving Lemmas \textbf{L1} and 
+\textbf{L2}, then 
+$\collapse{\exq x G(x)}$ must be valid. 
+Therefore, one may wonder how strong is their immunity to 
+the modal collapse: is there any other proposition $\phi$ 
+for which $\collapse{\phi}$ is also valid?
 
-Let $collapse{\phi} \equiv \phi \imp \nec \phi$. It is known that  Variants of the ontological argument
+For Anderson's emendation \cite{anderson}, for example, a form of the modal collapse (\textbf{A:MC}), restricted to positive properties applied to god-like beings, can be derived.
+The proof, under the modal logic \textbf{\emph{K}}, depends only on Anderson's alternative definition of god-like being (\textbf{A:D1}).
+This class of propositions for which the collapse occurs is tight: weaker restrictions (\textbf{A:MC1} and \textbf{A:MC2}), which could lead to larger classes, are counter-satisfiable.
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+% \item[A:A1] If a property is positive, its negation is not positive:
+%   $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+% \item[A2] A property necessarily implied by a
+%   positive property is positive:
+%   $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+%   \imp \psi(x)]) \imp P(\psi)]}$$
+% \item[T1] Positive properties are possibly exemplified: 
+%   $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[\textbf{A:D1}] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[\textbf{A:MC}] The modal collapse happens for any positive properties applied to any god-like being:
+  $$
+  \hol{\allq \varphi \allq x [(P(\varphi) \wedge G_A(x)) \imp \collapse{\varphi(x)}]}
+  $$
+\item[\textbf{A:MC1}] The modal collapse does \emph{not} happen for positive properties applied to arbitrary individuals (\emph{counter-satisfiable}):
+  $$
+  \hol{\allq \varphi \allq x [P(\varphi) \imp \collapse{\varphi(x)}]}
+  $$
+\item[\textbf{A:MC2}] The modal collapse does \emph{not} happen for an arbitrary properties applied to a god-like being (\emph{counter-satisfiable}):
+  $$
+  \hol{\allq \varphi \allq x [G_A(x) \imp \collapse{\varphi(x)}]}
+  $$
+% \item[A3']  The property of being God-like is positive: 
+%   $$\hol{P(G_A)}$$
+% \item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+% \item[A4]  Positive properties are necessarily positive: 
+%   $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+% \item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+%   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+% \item[T2']  Being God-like is an essence of any
+%   God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+% \item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+%   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+%   \exq y \varphi(y)]}$$
+% \item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+% \item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+%   $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+% \item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+%   $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+% %
+% \item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Restricted Collapse for Anderson's Emendation \cite{ToDo:Anderson}}
+\label{fig:anderson}
+\end{figure}
+
+ToDo: write a similar paragraph about Bjordal's alternative??
+
+Independently of the variant of the ontological argument under consideration, the following can be said about the class of collapsing propositions:
+\begin{enumerate}
+\item Valid propositions are collapsing: if $\phi$ is valid, then $\collapse{\phi}$ is valid.
+%
+\item The class of collapsing propositions is closed under logical equivalence: if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, then $\collapse{\phi'}$ is valid.
+%
+\item The class of collapsing propositions is not generally closed under equi-validity: even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ are equi-valid, $\collapse{\phi'}$ may not be valid.
+%
+\item The class of collapsing propositions is not generally closed under implication: even if $\collapse{\phi}$ is valid and $\phi \imp \phi'$ is valid, $\collapse{\phi'}$ may not be valid.
+\end{enumerate}
 
@@
 
-\section{Conclusions}
+\section{Final Remarks}
+
+All results announced in this note have been obtained experimentally\footnote{The source code of the experiments are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.} using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}.
 
@@
 
-The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}.
+The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist .
 This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Write conclusion

@@
 
-All results announced in this note have been obtained experimentally\footnote{The source code of the experiments are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.} using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}.
+All results announced in this note have been obtained experimentally using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}.
+The source codes of the experiments, as well as the resulting proofs and counter-models, are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
 
-Various slightly different versions of
-axioms and definitions have been considered by G\"{o}del and by
-several philosophers who commented on his proof
-(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+The technique enabling these experiments is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist.
+This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description is available in \cite{todo:ECAI}.
 
-In theoretical philosophy, formal logical confrontations with such
-ontological arguments had been so far (mainly) limited to paper
-and pen.  Up to now, the use of computers was prevented, because the
-logics of the available theorem proving systems were not expressive
-enough to formalize the abstract concepts adequately. G{\"o}del's proof
-uses, for example, a complex higher-order modal logic (\HOML) to handle
-concepts such as \emph{possibility} and \emph{necessity} and to support
-quantification over individuals and properties.
+The modal collapse is an interesting example of philosophical controversy and dispute, to which Leibniz's idea of a \emph{calculus ratiocinator} in the form of contemporary automated theorem provers can be applied.
+Current technology is increasingly ready to be embraced by those willing to practice computer-assisted theoretical philosophy \cite{oppenheimera11,rushby13}.
 
-controversies, care with parameters
+% Various slightly different versions of
+% axioms and definitions have been considered by G\"{o}del and by
+% several philosophers who commented on his proof
+% (cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
 
-ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
+% In theoretical philosophy, formal logical confrontations with such
+% ontological arguments had been so far (mainly) limited to paper
+% and pen.  Up to now, the use of computers was prevented, because the
+% logics of the available theorem proving systems were not expressive
+% enough to formalize the abstract concepts adequately. G{\"o}del's proof
+% uses, for example, a complex higher-order modal logic (\HOML) to handle
+% concepts such as \emph{possibility} and \emph{necessity} and to support
+% quantification over individuals and properties.
 
-The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist .
-This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.
+% controversies, care with parameters
+
+% ToDo: Leibniz calculemus, Rushby, Zalta
 

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Add acknowledgment and note about authorship

@@
 such that $\phi \imp \nec \phi$ 
-(from now on abbreviated $\collapse{\phi}$) is not valid 
+(from now on abbreviated as $\collapse{\phi}$) is not valid 
 under the axioms and definitions used by the variant. 
@@
 This class of propositions for which the collapse occurs is tight: weaker restrictions (\textbf{A:MC1} and \textbf{A:MC2}), which could lead to larger classes, are counter-satisfiable.
+These results hold under both constant and varying domain quantification, with possibilist and actualist quantifiers.
 
@@
 
-All results announced in this note have been obtained experimentally using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}.
-The source codes of the experiments, as well as the resulting proofs and counter-models, are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
+All results announced in this note have been obtained 
+experimentally using interactive and automated theorem provers 
+and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}. 
+The source codes of the experiments, as well as the resulting 
+proofs and counter-models, are available in 
+\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
 
-The technique enabling these experiments is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist.
-This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description is available in \cite{todo:ECAI}.
+The technique enabling these experiments is the embedding of 
+quantified modal logics into higher-order logics 
+\cite{J23,B9,C36}, for which automated theorem provers exist. 
+This technique has already been successfully employed in the 
+verification and reconstruction of G\"odel's proof 
+\cite{todo:all our previous papers}, and a detailed 
+mathematical description is available in \cite{todo:ECAI}.
 
-The modal collapse is an interesting example of philosophical controversy and dispute, to which Leibniz's idea of a \emph{calculus ratiocinator} in the form of contemporary automated theorem provers can be applied.
-Current technology is increasingly ready to be embraced by those willing to practice computer-assisted theoretical philosophy \cite{oppenheimera11,rushby13}.
+The modal collapse is an interesting example of philosophical 
+controversy and dispute, to which we can apply Leibniz's idea 
+of a \emph{calculus ratiocinator} brought to reality in the 
+form of contemporary automated theorem provers. 
+A significant advantage provided by the use of computers is 
+that all conditions (e.g. modal logic, domain conditions, 
+semantics) under which the announced results hold must be 
+explicitly written in the source code. This reduces the need 
+for interpretation and the danger of misunderstandings. 
+Current technology is increasingly ready to be embraced by 
+those willing to practice computer-assisted theoretical 
+philosophy \cite{oppenheimera11,rushby13}.
 
@@
 
-% ToDo: Leibniz calculemus, Rushby, Zalta
+\subsection*{Acknowledgment}
 
-% \begin{defn}
-% This serves as environment for definitions. Note that the text
-% appears not in italics.
-% \end{defn}
+We would like to thank Paul Weingartner, Andr\'{e} Fuhrmann and Melvin Fitting for several discussions about G\"odel's ontological argument.
 
-% \begin{equation}\label{testequation}
-% \text{This is a sample equation: } c^2=a^2+b^2
-% \end{equation}
+ToDo: who else?
 
-% \begin{thm}[Main Theorem]
-% In contrast to definitions, theorems appear typeset in italics as
-% it has become more or less standard in most textbooks and
-% monographs. Equations can be cited using the \verb+\eqref+ command which
-% automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
-% \end{thm}
+\subsection*{Note about authorship}
 
-% \begin{proof}
-% A special environment is predefined: the \textit{proof} environment. Please use
-% \begin{verbatim}\begin{proof}\end{verbatim}
-% proof of the statement
-% \begin{verbatim}\end{proof}\end{verbatim}
-% for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
-% \end{proof}
-
-% There are two known problems with the placement of the end-of-proof sign:
-
-% \begin{enumerate}
-%   \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
-% be placed in the line below; if you want to avoid this, write your line in the form
-% \begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
-% which results in
-
-% \begin{proof}
-% $$displayed math line \eqno\qedhere$$
-% \end{proof}
-% \item if your proof ends with an aligned displayed environment, the command
-% \verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
-% \begin{verbatim}
-% \begin{align*}
-% \alpha&=\beta+\gamma\\
-% &=\delta+\epsilon\tag*{\qed}
-% \end{align*}
-% \end{verbatim}
-% results in
-% \begin{align*}
-% \alpha&=\beta+\gamma\\
-% &=\delta+\epsilon\tag*{\qed}
-% \end{align*}
-% \end{enumerate}
-% Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
-% and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.
-
-% \begin{rem}
-% Additional comments are being typeset without boldfaced entrance
-% word as they may be minor important.
-% \end{rem}
-
-% \begin{ex}
-% For some constructs, even no number is required.
-% \end{ex}
-
-% Displayed equations may be numbered like the following one:
-% \begin{equation}
-% \sqrt{1-\sin^2(x)}=|\cos(x)|.
-% \end{equation}
-
-% ------------------------------------------------------------------------
-
-% \subsection*{Acknowledgment}
-% Many thanks to ...
+Alphabetic order has been used for the authors' names. The extent and kind of contribution of each author cannot be inferred from the order.
 

### 2014-09-09 Papers/Universalis/ModalCollapse.tex
contributions to the paper

@@
 \documentclass{birkjour}
+
+\usepackage[utf8]{inputenc}
 
@@
 Canterbury, Descartes and Leibniz.
-Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, shown in Fig.
+Kurt G{\"o}del and Dana Scott studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a second-order modal logic \cite{GoedelNotes,ScottNotes}, shown in Fig.
 \ref{fig:scott}.
@@
 G\"{o}del defines God as a being who possesses all \emph{positive}
-properties and states a few reasonable (but debatable) axioms
-that such properties should satisfy.  
-The overall idea of G{\"o}del's proof is in the tradition of Anselm's
-argument, who defined God as some entity of which nothing greater can be
-conceived. Anselm argued that existence in the actual world would
-make such an assumed being even greater; hence, by definition, God must
-exist.
-However, for Anselm existence was treated as a predicate and the possibility of God's existence was assumed as granted.
-These issues were criticized by Kant and Leibniz, respectively, and succesfully addressed by G\"odel.
+properties and states a few reasonable (but debatable) axioms that
+such properties should satisfy.  The overall idea of G{\"o}del's proof
+is in the tradition of Anselm's argument, who defined God as some
+entity of which nothing greater can be conceived. Anselm argued that
+existence in the actual world would make such an assumed being even
+greater (more perfect); hence, by definition, God must exist. However,
+for Anselm existence was treated as a predicate and the possibility of
+God's existence was assumed as granted. These issues were criticized
+by Kant and Leibniz, respectively, and they were addressed in the work
+of G\"odel.\footnote{C.: I deleted "successfully", since I am still
+  not totally convinced yet about the existence as predicate issue in
+  G\"odel's work; I think we should not be conclusive at this point.}
 
-Nevertheless, G{\"o}del's work
-still leaves room for criticism. In particular, his axioms are so
-strong that they entail a \emph{modal collapse} 
-\cite{ToDo:sobel1987?,sobel2004logic}: 
-everything that is the case is so necessarily. 
-There has been an impressive body of recent and ongoing
-work (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,ToDo:FrodeAndHajek,ContemporaryBibliography} and the references therein)
-proposing solutions for the modal collapse. 
-The goal of this short note is to discuss the modal collapse from the point of view of the modal square of opposition.
+Nevertheless, G{\"o}del's work still leaves room for criticism. In
+particular, his axioms are so strong that they entail a \emph{modal
+  collapse} \cite{Sobel1987,sobel2004logic}: everything that is the
+case is so necessarily.  There has been an impressive body of recent
+and ongoing work
+(cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
+and the references therein) proposing solutions for the modal
+collapse.  The goal of this contribution is to discuss the modal
+collapse from the point of view of the modal square of opposition.
 
@@
 
-A crucial step of most ontological arguments is the claim that 
-if God's existence is possible, then it is necessary. 
-This is Lemma \textbf{L2} in G\"odel's proof. 
-In the modal square of opposition (Fig. \ref{fig:square}), 
-this is an unusual situation in which the \textbf{I} corner 
-must imply and entail the \textbf{A} corner, 
-in the particular case when $\phi$ is $\exq x G(x)$. 
-G\"odel's proof shows that his axioms are strong enough 
-to invert the direction of entailment for the sentence at issue. 
-The question, however, is whether the axioms are not too strong, 
-also allowing the inverted entailment for arbitrary $\phi$. 
-That is essentially the question asked by Sobel \cite{Sobel1987}; 
-and his proof of the modal collapse (\textbf{MC}) provides an affirmative 
-answer. It is possible to show that this form of the modal 
-collapse entails (in modal logic \textbf{\emph{K}}) a collapse of the modal 
-square (\textbf{MC'}), causing the subcontraries to entail (and even imply) their respective contraries.
-Normally, as shown in Fig.
-\ref{fig:square}, in the modal square of opposition only the other direction of entailment holds: the contraries entail their subcontraries, assuming the \emph{modal existential import} \textbf{ExImp}.
+A crucial step of most ontological arguments is the claim that if
+God's existence is possible, then it is necessary.  This is Lemma
+\textbf{L2} in G\"odel's proof.  In the modal square of opposition
+(Fig. \ref{fig:square}), this is an unusual situation in which the
+\textbf{I} corner must imply and entail the \textbf{A} corner, in the
+particular case when $\phi$ is $\exq x G(x)$.  G\"odel's proof shows
+that his axioms are indeed strong enough to invert the direction of
+entailment this choice of $\phi$.
+% for the sentence at issue.  
+This observation, however, immediately brings up the question whether
+the axioms are even strong enough to enable the inverted entailment
+for arbitrary sentences $\phi$.  That is essentially the question
+asked by Sobel \cite{Sobel1987}, and his proof of the modal collapse
+(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) provides an affirmative answer. It is possible to show
+that this form of the modal collapse entails (in modal logic
+\textbf{\emph{K}}) a collapse of the modal square (\textbf{MC'}),
+causing the subcontraries to entail (and even imply) their respective
+contraries. Normally, as shown in Fig. \ref{fig:square}, in the modal
+square of opposition only the other direction of entailment holds: the
+contraries entail their subcontraries, assuming the \emph{modal
+  existential import} \textbf{ExImp}
+\cite{WhatToCiteHere}.\footnote{C.: We need a reference here; I am still
+  searching for good papers; this here is also interesting:
+  \url{http://www.tandfonline.com/doi/pdf/10.1080/01445340.2013.764962}}
 
-Moreover, in any modal logic where the 
-axiom \textbf{T} holds (i.e.
-where the accessibility relation is reflexive), a total collapse of the modalities (\textbf{MC''}) occurs.
-Interestingly, under this stronger form of modal collapse, the contraries entail their subcontraries even without the existential import.
+Moreover, in any modal logic where the axiom \textbf{T} holds
+(i.e. where the accessibility relation is reflexive), a total collapse
+of the modalities (\textbf{MC''}) occurs.\footnote{C.: I am a bit lost here. Do you want to
+  say that any logic that includes T strongly collapses? This would be 
+  odd. Where is the theorem in the Isabelle file(s) that proves this? Also: MC'' should
+  be stated more precisely, I mean like: $((\varphi \biimp \nec \varphi) \wedge (\nec \varphi
+  \biimp \pos \varphi))$.}  Interestingly, under this stronger form of modal
+  collapse, the contraries entail their subcontraries even without the
+  existential import.
 
@@
 Although G\"odel's axioms lead to modal collapse, 
-there are several variants (e.g. \cite{anderson,bjordal}) 
-that are known to be immune to the modal collapse. 
+there are several variants (e.g. \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99}) 
+that are known to be immune to it. 
 This means there must be at least one proposition $\phi$ 
-such that $\phi \imp \nec \phi$ 
+such that the implication $\phi \imp \nec \phi$ 
 (from now on abbreviated as $\collapse{\phi}$) is not valid 
@@
 But if the variant is sufficiently similar to G\"odel's argument, 
-following the path deriving Lemmas \textbf{L1} and 
+following the path deriving\footnote{C.: I don't know the phrase "path deriving", please check.} Lemmas \textbf{L1} and 
 \textbf{L2}, then 
@@
 
-For Anderson's emendation \cite{anderson}, for example, a form of the modal collapse (\textbf{A:MC}), restricted to positive properties applied to god-like beings, can be derived.
+For Anderson's emendation \cite{anderson90:_some_emend_of_goedel_ontol_proof}, for example, a form of the modal collapse (\textbf{A:MC}), restricted to positive properties applied to god-like beings, can be derived.
 The proof, under the modal logic \textbf{\emph{K}}, depends only on Anderson's alternative definition of god-like being (\textbf{A:D1}).
@@
 } \vskip-.5em
-\caption{Restricted Collapse for Anderson's Emendation \cite{ToDo:Anderson}}
+\caption{Restricted Collapse for Anderson's Emendation \cite{anderson90:_some_emend_of_goedel_ontol_proof}}
 \label{fig:anderson}
@@
 
-ToDo: write a similar paragraph about Bjordal's alternative??
+%ToDo: write a similar paragraph about Bjordal's alternative??
 
-Independently of the variant of the ontological argument under consideration, the following can be said about the class of collapsing propositions:
+Independently of the variant of the ontological argument under
+consideration, the following can be said about the class of collapsing
+propositions:
 \begin{enumerate}
@@
 
-\clearpage
+%\clearpage
 
@@
 
-All results announced in this note have been obtained 
-experimentally using interactive and automated theorem provers 
-and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}. 
-The source codes of the experiments, as well as the resulting 
-proofs and counter-models, are available in 
-\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
+All results announced in this note have been obtained experimentally
+using interactive and automated theorem provers and model finders
+\cite{LEO,Satallax,Isabelle,Coq,Nitpick}.  The source codes of the
+experiments, as well as the resulting proofs and counter-models, are
+available in
+\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.\footnote{Here
+  we should point to the exact files (and put them in a relatively
+  stable subdirectory?). We also should also use corresponding names in
+  these files, so that the reader easily finds the related 
+  parst. Right now this is still quite difficult.}
 
@@
 verification and reconstruction of G\"odel's proof 
-\cite{todo:all our previous papers}, and a detailed 
-mathematical description is available in \cite{todo:ECAI}.
+\cite{J28,J30,W50,J29}, and a detailed 
+mathematical description is available in \cite{C40}.
 
@@
 semantics) under which the announced results hold must be 
-explicitly written in the source code. This reduces the need 
+explicitly specified in the source code. This reduces the need 
 for interpretation and the danger of misunderstandings. 
@@
 philosophy \cite{oppenheimera11,rushby13}.
+
+Ongoing and future work includes the study of modal collapse in
+further variants of the ontological argument, e.g.~\cite{bjordal99,fuhrmann05:_exist_notwen}.
 
@@
 
-Alphabetic order has been used for the authors' names. The extent and kind of contribution of each author cannot be inferred from the order.
+Alphabetic order has been used for the authors' names. The extent and
+kind of contribution of each author cannot be inferred from the order.
 
-\begin{thebibliography}{1}
+\bibliographystyle{plain}
+\bibliography{Bibliography}
 
-\bibitem{Adams}
-R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
-  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
+%\begin{thebibliography}{1}
 
-\bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
-  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
-  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
+% \bibitem{Adams}
+% R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
+%   Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
 
-\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
-C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
-  Faith and Philosophy}, {\bf 7}(3), (1990).
+% \bibitem{AndersonGettings}
+% A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
+%   {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
+%   Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
 
-\bibitem{Andrews:gmae72}
-P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
-  Logic}, {\bf 37}(2),  395--397, (1972).
+% \bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
+% C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
+%   Faith and Philosophy}, {\bf 7}(3), (1990).
 
-\bibitem{andrewsSEP}
-P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
-  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
+% \bibitem{Andrews:gmae72}
+% P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
+%   Logic}, {\bf 37}(2),  395--397, (1972).
 
-\bibitem{C36}
-C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
-  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
-  pp. 232--233, Rio de Janeiro, Brazil, (2013).
+% \bibitem{andrewsSEP}
+% P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
+%   Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
 
-\bibitem{B5}
-C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
-  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
-  Elsevier, (2014).
-\newblock Forthcoming; preliminary version available at
-  {http://christoph-benzmueller.de/papers/B5.pdf}.
+% \bibitem{C36}
+% C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
+%   4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
+%   pp. 232--233, Rio de Janeiro, Brazil, (2013).
 
-\bibitem{C34}
-C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
-  provers for first-order modal logics', in {\em Proc. of the 20th European
-  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
+% \bibitem{B5}
+% C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
+%   Handbook of the History of Logic, Volume 9 --- Logic and Computation},
+%   Elsevier, (2014).
+% \newblock Forthcoming; preliminary version available at
+%   {http://christoph-benzmueller.de/papers/B5.pdf}.
 
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
-  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
-  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
-  386--406, College Publications, (2008).
+% \bibitem{C34}
+% C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
+%   provers for first-order modal logics', in {\em Proc. of the 20th European
+%   Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
 
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
-  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
+% \bibitem{B9}
+% C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
+%   logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
+%   {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
+%   386--406, College Publications, (2008).
 
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
-  cooperative automatic theorem prover for higher-order logic', in {\em
-  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
+% \bibitem{J23}
+% C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
+%   type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
 
-\bibitem{J30}
-C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
-  automation of {G{\"o}del's} proof of {God's} existence', {\em
-  arXiv:1308.4526}, (2013).
+% \bibitem{LEO-II}
+% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
+%   cooperative automatic theorem prover for higher-order logic', in {\em
+%   Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
 
-\bibitem{J28}
-C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
-  {\em Archive of Formal Proofs}, (2013).
+% \bibitem{J30}
+% C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
+%   automation of {G{\"o}del's} proof of {God's} existence', {\em
+%   arXiv:1308.4526}, (2013).
 
-\bibitem{W50}
-C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
-  in {\em Proceedings of the 10th International Workshop on the Implementation
-  of Logics}, EPiC Series. EasyChair, (2013).
-\newblock Invited abstract.
+% \bibitem{J28}
+% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
+%   {\em Archive of Formal Proofs}, (2013).
 
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
-  Development}}, Springer, 2004.
+% \bibitem{W50}
+% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
+%   in {\em Proceedings of the 10th International Workshop on the Implementation
+%   of Logics}, EPiC Series. EasyChair, (2013).
+% \newblock Invited abstract.
 
-\bibitem{Nitpick}
-J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
-  higher-order logic based on a relational model finder', in {\em Proc. of ITP
-  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
+% \bibitem{Coq}
+% Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
+%   Development}}, Springer, 2004.
 
-\bibitem{Satallax}
-C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
-  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
+% \bibitem{Nitpick}
+% J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
+%   higher-order logic based on a relational model finder', in {\em Proc. of ITP
+%   2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
 
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
-  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
+% \bibitem{Satallax}
+% C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
+%   IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
 
-\bibitem{Fitting}
-M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
+% \bibitem{ContemporaryBibliography}
+% R.~Corazzon.
+% \newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
+%   \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
 
-\bibitem{fitting98}
-M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
-  {\em Synthese Library}, Kluwer, 1998.
+% \bibitem{Fitting}
+% M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
 
-\bibitem{Gallin75}
-D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
+% \bibitem{fitting98}
+% M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
+%   {\em Synthese Library}, Kluwer, 1998.
 
-\bibitem{garbacz12:_prover_simpl_expal_away}
-P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
-  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
+% \bibitem{Gallin75}
+% D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
 
-\bibitem{GoedelNotes}
-K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{garbacz12:_prover_simpl_expal_away}
+% P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
+%   Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
 
-\bibitem{Henkin50}
-L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
-  Logic}, {\bf 15}(2),  81--91, (1950).
+% \bibitem{GoedelNotes}
+% K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{homl}
-R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
-  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
+% \bibitem{Henkin50}
+% L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
+%   Logic}, {\bf 15}(2),  81--91, (1950).
 
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
-  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
+% \bibitem{homl}
+% R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
+%   ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
 
-\bibitem{oppenheimera11}
-P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
-  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
-  89}(2),  333--349, (2011).
+% \bibitem{Isabelle}
+% T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
+%   for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
 
-\bibitem{rushby13}
-J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
-  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
+% \bibitem{oppenheimera11}
+% P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
+%   of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
+%   89}(2),  333--349, (2011).
 
-\bibitem{Schulz:AICOM-2002}
-S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
-  15}(2),  111--126, (2002).
+% \bibitem{rushby13}
+% J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
+%   ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
 
-\bibitem{ScottNotes}
-D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{Schulz:AICOM-2002}
+% S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
+%   15}(2),  111--126, (2002).
 
-\bibitem{sobel2004logic}
-J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
-  Cambridge U. Press, 2004.
+% \bibitem{ScottNotes}
+% D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{sutcliffe2009tptp}
-G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
-  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
+% \bibitem{sobel2004logic}
+% J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
+%   Cambridge U. Press, 2004.
 
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
-  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
-  {\bf 3}(1),  1--27, (2010).
+% \bibitem{sutcliffe2009tptp}
+% G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
+%   Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
 
-\bibitem{J29}
-B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
-  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
-  (2013).
+% \bibitem{J22}
+% G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
+%   using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
+%   {\bf 3}(1),  1--27, (2010).
 
-\end{thebibliography}
+% \bibitem{J29}
+% B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
+%   reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
+%   (2013).
+
+%\end{thebibliography}
 

### 2014-09-09 Papers/Universalis/ModalCollapse.tex
contributions to the paper

@@
 % for the sentence at issue.  
-This observation, however, immediately brings up the question whether
-the axioms are even strong enough to enable the inverted entailment
+This observation, however, immediately leads to the question whether
+the axioms are eventually even strong enough to enable the inverted entailment
 for arbitrary sentences $\phi$.  That is essentially the question
@@
 available in
-\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.\footnote{Here
+\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.\footnote{C.: Here
   we should point to the exact files (and put them in a relatively

### 2014-09-09 Formalizations/THF/Run_ECAI2014_Experiments.sh
improvement of scripts

@@ -1,1 +1,7 @@
+echo "Recomputation of the experiments from the ECAI-2014 paper *Automating Goedel\'s Ontological Proof of God\'s Existence with Higher-order Automated Theorem Provers* by Christoph Benzmueller and Bruno Woltzenlogel Paleo."
+echo ""
+echo "The problems tested below corresponf to the results mentioned in Figure 2 of that paper (min/max refers to the sets of dependencies tested for each conjecture statement, and the meaning of vary/const is obvious)."
+echo ""
 ./call_tptp.sh T1_K_const_max.p T1_K_const_min.p T1_K_vary_max.p T1_K_vary_min.p T1_varying.p C_K_const_max.p  C_K_const_min.p  C_K_vary_max.p   C_K_vary_min.p T2_K_const_max.p  T2_K_const_min.p  T2_K_vary_max.p   T2_K_vary_min.p  T3_KB_const_max.p  T3_KB_const_min.p  T3_KB_vary_max.p   T3_KB_vary_min.p   T3_K_const_max.p   T3_K_const_min.p   T3_K_vary_max.p    T3_K_vary_min.p   MC_KB_const_max.p  MC_KB_const_min.p  MC_KB_vary_max.p   MC_KB_vary_min.p   FG_KB_const_max.p  FG_KB_const_min.p  FG_KB_vary_max.p   FG_KB_vary_min.p  MT_KB_const_max.p  MT_KB_const_min.p  MT_KB_vary_max.p   MT_KB_vary_min.p   CO_KB_const_max.p  CO_KB_vary_max.p  CO\'_KB_const_max.p  CO\'_KB_const_min.p  CO\'_KB_vary_max.p   CO\'_KB_vary_min.p  
+echo ""
+echo "Done with the experiments."

### 2014-09-09 Formalizations/THF/call_tptp.sh

@@ -4,7 +4,8 @@ call=./RemoteSOT.pl
 # install it and change the above paths accordingly
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
-proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='LEO-II--- Satallax--- Nitrox--- '
 #proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
 
 timeout=60

### 2014-09-10 Papers/Universalis/ModalCollapse.tex
Fix issue described in footnote 3

@@
 Moreover, in any modal logic where the axiom \textbf{T} holds
-(i.e. where the accessibility relation is reflexive), a total collapse
-of the modalities (\textbf{MC''}) occurs.\footnote{C.: I am a bit lost here. Do you want to
-  say that any logic that includes T strongly collapses? This would be 
-  odd. Where is the theorem in the Isabelle file(s) that proves this? Also: MC'' should
-  be stated more precisely, I mean like: $((\varphi \biimp \nec \varphi) \wedge (\nec \varphi
-  \biimp \pos \varphi))$.}  Interestingly, under this stronger form of modal
-  collapse, the contraries entail their subcontraries even without the
-  existential import.
+(i.e. where the accessibility relation is reflexive), 
+even a total collapse of the modalities (\textbf{MC''}) 
+is entailed by \textbf{MC}.
+Interestingly, under this stronger form of modal
+collapse, the contraries entail their subcontraries even 
+without the existential import.
 
@@
 \item[\textbf{MC''}] Modalities collapse completely:
-  $\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$
+  $\hol{\allq \phi [(\phi \biimp \nec \phi) \wedge (\pos \phi \biimp \nec \phi)]}$
 \end{itemize}

### 2014-09-10 Papers/Universalis/ModalCollapse.tex
Clean the source

@@
-% ------------------------------------------------------------------------
-% bjourdoc.tex for birkjour.cls*******************************************
-% ------------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass{birkjour}
@@
 \usepackage{commands}
-%
-%
-% THEOREM Environments (Examples)-----------------------------------------
-%
+
  \newtheorem{thm}{Theorem}[section]
@@
 
-%----------Author 1
 \author[Benzm\"uller]{Christoph Benzm\"uller}
@@
 
-%\thanks{This work was completed with the support of ToDo}
-
-%----------Author 3
 \author[Woltzenlogel-Paleo]{Bruno Woltzenlogel Paleo}
@@
 \date{August 30, 2014}
-%----------additions
+
 \dedicatory{ }
-%%% ----------------------------------------------------------------------
 
@@
 
-%%% ----------------------------------------------------------------------
 \maketitle
-%%% ----------------------------------------------------------------------
+
 %\tableofcontents
+
 \section{Introduction}
@@
 
-% Ontological arguments, for or against the existence of God,
-% illustrate well an essential aspect of metaphysics: some (necessary) facts
-% for our existing world are deduced by purely a priori, analytical means from some
-% abstract definitions and axioms. % Contingent truths are to be
-% distinguished from necessary truths.
-
 G\"{o}del defines God as a being who possesses all \emph{positive}
 properties and states a few reasonable (but debatable) axioms that
-such properties should satisfy.  The overall idea of G{\"o}del's proof
+such properties should satisfy.  
+The overall idea of G{\"o}del's proof
 is in the tradition of Anselm's argument, who defined God as some
-entity of which nothing greater can be conceived. Anselm argued that
-existence in the actual world would make such an assumed being even
-greater (more perfect); hence, by definition, God must exist. However,
-for Anselm existence was treated as a predicate and the possibility of
-God's existence was assumed as granted. These issues were criticized
-by Kant and Leibniz, respectively, and they were addressed in the work
-of G\"odel.\footnote{C.: I deleted "successfully", since I am still
-  not totally convinced yet about the existence as predicate issue in
-  G\"odel's work; I think we should not be conclusive at this point.}
+entity of which nothing greater can be conceived. 
+Anselm argued that
+existence in the actual world would make 
+such an assumed being even
+greater (more perfect); 
+hence, by definition, God must exist. However,
+for Anselm existence was treated as a predicate and 
+the possibility of God's existence was assumed as granted. 
+These issues were criticized by Kant and Leibniz, 
+respectively, and they were addressed in the work of G\"odel.
 
-Nevertheless, G{\"o}del's work still leaves room for criticism. In
-particular, his axioms are so strong that they entail a \emph{modal
-  collapse} \cite{Sobel1987,sobel2004logic}: everything that is the
-case is so necessarily.  There has been an impressive body of recent
-and ongoing work
+Nevertheless, G{\"o}del's work still leaves room for criticism. 
+In particular, his axioms are so strong that they entail 
+a \emph{modal collapse} 
+\cite{Sobel1987,sobel2004logic}: 
+everything that is the case is so necessarily.  
+There has been an impressive body of recent and ongoing work
 (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
@@
 (Fig. \ref{fig:square}), this is an unusual situation in which the
-\textbf{I} corner must imply and entail the \textbf{A} corner, in the
-particular case when $\phi$ is $\exq x G(x)$.  G\"odel's proof shows
-that his axioms are indeed strong enough to invert the direction of
-entailment this choice of $\phi$.
-% for the sentence at issue.  
-This observation, however, immediately leads to the question whether
-the axioms are eventually even strong enough to enable the inverted entailment
-for arbitrary sentences $\phi$.  That is essentially the question
-asked by Sobel \cite{Sobel1987}, and his proof of the modal collapse
-(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) provides an affirmative answer. It is possible to show
+\textbf{I} corner must imply and entail the \textbf{A} corner, 
+in the particular case when $\phi$ is $\exq x G(x)$.  
+G\"odel's proof shows that his axioms are indeed 
+strong enough to invert the direction of
+entailment for this choice of $\phi$.  
+This observation, however, immediately leads 
+to the question whether the axioms are eventually 
+even strong enough to enable the inverted entailment
+for any arbitrary sentence $\phi$.  
+That is essentially the question asked by 
+Sobel \cite{Sobel1987}, and his proof of the modal collapse
+(\textbf{MC}, cf. Fig.~\ref{fig:collapse}) 
+provides an affirmative answer. It is possible to show
 that this form of the modal collapse entails (in modal logic
 \textbf{\emph{K}}) a collapse of the modal square (\textbf{MC'}),
-causing the subcontraries to entail (and even imply) their respective
-contraries. Normally, as shown in Fig. \ref{fig:square}, in the modal
-square of opposition only the other direction of entailment holds: the
-contraries entail their subcontraries, assuming the \emph{modal
-  existential import} \textbf{ExImp}
-\cite{WhatToCiteHere}.\footnote{C.: We need a reference here; I am still
+causing the subcontraries to entail 
+(and even imply) their respective
+contraries. Normally, as shown in Fig. \ref{fig:square}, 
+in the modal square of opposition only the 
+other direction of entailment holds: the
+contraries entail their subcontraries, 
+assuming the \emph{modal existential import} \textbf{ExImp}
+\cite{WhatToCiteHere}.
+\footnote{C.: We need a reference here; I am still
   searching for good papers; this here is also interesting:
@@
 \begin{figure}[t]
-\includegraphics[width=0.9\textwidth]{./SquarePics/ModalSquare.png}
+\includegraphics[width=0.9\textwidth]{SquarePics/ModalSquare.png}
 \caption{Modal Square of Opposition.}
@@
 Although G\"odel's axioms lead to modal collapse, 
-there are several variants (e.g. \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99}) 
+there are several variants 
+(e.g. \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99}) 
 that are known to be immune to it. 
@@
 But if the variant is sufficiently similar to G\"odel's argument, 
-following the path deriving\footnote{C.: I don't know the phrase "path deriving", please check.} Lemmas \textbf{L1} and 
+also deriving Lemmas \textbf{L1} and 
 \textbf{L2}, then 
@@
 
-For Anderson's emendation \cite{anderson90:_some_emend_of_goedel_ontol_proof}, for example, a form of the modal collapse (\textbf{A:MC}), restricted to positive properties applied to god-like beings, can be derived.
-The proof, under the modal logic \textbf{\emph{K}}, depends only on Anderson's alternative definition of god-like being (\textbf{A:D1}).
-This class of propositions for which the collapse occurs is tight: weaker restrictions (\textbf{A:MC1} and \textbf{A:MC2}), which could lead to larger classes, are counter-satisfiable.
-These results hold under both constant and varying domain quantification, with possibilist and actualist quantifiers.
+For Anderson's emendation 
+\cite{anderson90:_some_emend_of_goedel_ontol_proof}, 
+for example, a form of the modal collapse (\textbf{A:MC}), 
+restricted to positive properties applied to god-like beings, 
+can be derived. The proof, under the modal logic 
+\textbf{\emph{K}}, depends only on Anderson's 
+alternative definition of god-like being (\textbf{A:D1}). 
+This class of propositions for which the collapse occurs 
+is tight: weaker restrictions 
+(\textbf{A:MC1} and \textbf{A:MC2}), 
+which could lead to larger classes, are counter-satisfiable. 
+These results hold under both constant and varying domain 
+quantification, with possibilist and actualist quantifiers.
 
@@
 \begin{itemize}
-% \item[A:A1] If a property is positive, its negation is not positive:
-%   $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-% \item[A2] A property necessarily implied by a
-%   positive property is positive:
-%   $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-%   \imp \psi(x)]) \imp P(\psi)]}$$
-% \item[T1] Positive properties are possibly exemplified: 
-%   $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
 \item[\textbf{A:D1}] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
@@
   $$
-% \item[A3']  The property of being God-like is positive: 
-%   $$\hol{P(G_A)}$$
-% \item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-% \item[A4]  Positive properties are necessarily positive: 
-%   $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-% \item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-%   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-% \item[T2']  Being God-like is an essence of any
-%   God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-% \item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-%   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-%   \exq y \varphi(y)]}$$
-% \item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-% \item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-%   $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-% \item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-%   $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
-% %
-% \item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}
@@
 
-%ToDo: write a similar paragraph about Bjordal's alternative??
-
-Independently of the variant of the ontological argument under
-consideration, the following can be said about the class of collapsing
-propositions:
+Independently of the variant of the ontological argument 
+under consideration, the following can be said about 
+classes of collapsing propositions:
 \begin{enumerate}
-\item Valid propositions are collapsing: if $\phi$ is valid, then $\collapse{\phi}$ is valid.
+\item Valid propositions are collapsing: 
+if $\phi$ is valid, then $\collapse{\phi}$ is valid.
 %
-\item The class of collapsing propositions is closed under logical equivalence: if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, then $\collapse{\phi'}$ is valid.
+\item The class of collapsing propositions 
+is closed under logical equivalence: 
+if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, 
+then $\collapse{\phi'}$ is valid.
 %
-\item The class of collapsing propositions is not generally closed under equi-validity: even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ are equi-valid, $\collapse{\phi'}$ may not be valid.
+\item The class of collapsing propositions is 
+not generally closed under equi-validity: 
+even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
+are equi-valid, $\collapse{\phi'}$ may not be valid.
 %
-\item The class of collapsing propositions is not generally closed under implication: even if $\collapse{\phi}$ is valid and $\phi \imp \phi'$ is valid, $\collapse{\phi'}$ may not be valid.
+\item The class of collapsing propositions is 
+not generally closed under implication: 
+even if $\collapse{\phi}$ is valid and 
+$\phi \imp \phi'$ is valid, 
+$\collapse{\phi'}$ may not be valid.
 \end{enumerate}
-
-%\clearpage
 
@@
 
-All results announced in this note have been obtained experimentally
-using interactive and automated theorem provers and model finders
-\cite{LEO,Satallax,Isabelle,Coq,Nitpick}.  The source codes of the
-experiments, as well as the resulting proofs and counter-models, are
-available in
-\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.\footnote{C.: Here
+All results announced in this note have been 
+obtained experimentally using interactive and 
+automated theorem provers and model finders
+\cite{LEO,Satallax,Isabelle,Coq,Nitpick}.  
+The source codes of the experiments, 
+as well as the resulting proofs and counter-models, 
+are available in
+\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
+\footnote{C.: Here
   we should point to the exact files (and put them in a relatively
@@
 A significant advantage provided by the use of computers is 
-that all conditions (e.g. modal logic, domain conditions, 
+that all parameters (e.g. modal logic, domain conditions, 
 semantics) under which the announced results hold must be 
@@
 Ongoing and future work includes the study of modal collapse in
-further variants of the ontological argument, e.g.~\cite{bjordal99,fuhrmann05:_exist_notwen}.
-
-% Various slightly different versions of
-% axioms and definitions have been considered by G\"{o}del and by
-% several philosophers who commented on his proof
-% (cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-
-% In theoretical philosophy, formal logical confrontations with such
-% ontological arguments had been so far (mainly) limited to paper
-% and pen.  Up to now, the use of computers was prevented, because the
-% logics of the available theorem proving systems were not expressive
-% enough to formalize the abstract concepts adequately. G{\"o}del's proof
-% uses, for example, a complex higher-order modal logic (\HOML) to handle
-% concepts such as \emph{possibility} and \emph{necessity} and to support
-% quantification over individuals and properties.
-
-% controversies, care with parameters
+other variants of the ontological argument (e.g.~\cite{bjordal99,fuhrmann05:_exist_notwen}).
 
@@
 
-We would like to thank Paul Weingartner, Andr\'{e} Fuhrmann and Melvin Fitting for several discussions about G\"odel's ontological argument.
-
-ToDo: who else?
+We would like to thank Paul Weingartner, Andr\'{e} Fuhrmann 
+and Melvin Fitting for several discussions about 
+G\"odel's ontological argument.
 
@@
 
-Alphabetic order has been used for the authors' names. The extent and
-kind of contribution of each author cannot be inferred from the order.
+Alphabetic order has been used for the authors' names. 
+The extent and kind of contribution of each author 
+cannot be inferred from the order.
 
@@
 
-%\begin{thebibliography}{1}
-
-% \bibitem{Adams}
-% R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
-%   Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
-
-% \bibitem{AndersonGettings}
-% A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
-%   {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
-%   Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
-
-% \bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
-% C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
-%   Faith and Philosophy}, {\bf 7}(3), (1990).
-
-% \bibitem{Andrews:gmae72}
-% P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
-%   Logic}, {\bf 37}(2),  395--397, (1972).
-
-% \bibitem{andrewsSEP}
-% P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
-%   Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
-
-% \bibitem{C36}
-% C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
-%   4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
-%   pp. 232--233, Rio de Janeiro, Brazil, (2013).
-
-% \bibitem{B5}
-% C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
-%   Handbook of the History of Logic, Volume 9 --- Logic and Computation},
-%   Elsevier, (2014).
-% \newblock Forthcoming; preliminary version available at
-%   {http://christoph-benzmueller.de/papers/B5.pdf}.
-
-% \bibitem{C34}
-% C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
-%   provers for first-order modal logics', in {\em Proc. of the 20th European
-%   Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
-
-% \bibitem{B9}
-% C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
-%   logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
-%   {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
-%   386--406, College Publications, (2008).
-
-% \bibitem{J23}
-% C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
-%   type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
-
-% \bibitem{LEO-II}
-% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
-%   cooperative automatic theorem prover for higher-order logic', in {\em
-%   Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
-
-% \bibitem{J30}
-% C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
-%   automation of {G{\"o}del's} proof of {God's} existence', {\em
-%   arXiv:1308.4526}, (2013).
-
-% \bibitem{J28}
-% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
-%   {\em Archive of Formal Proofs}, (2013).
-
-% \bibitem{W50}
-% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
-%   in {\em Proceedings of the 10th International Workshop on the Implementation
-%   of Logics}, EPiC Series. EasyChair, (2013).
-% \newblock Invited abstract.
-
-% \bibitem{Coq}
-% Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
-%   Development}}, Springer, 2004.
-
-% \bibitem{Nitpick}
-% J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
-%   higher-order logic based on a relational model finder', in {\em Proc. of ITP
-%   2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
-
-% \bibitem{Satallax}
-% C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
-%   IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
-
-% \bibitem{ContemporaryBibliography}
-% R.~Corazzon.
-% \newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
-%   \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
-
-% \bibitem{Fitting}
-% M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
-
-% \bibitem{fitting98}
-% M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
-%   {\em Synthese Library}, Kluwer, 1998.
-
-% \bibitem{Gallin75}
-% D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
-
-% \bibitem{garbacz12:_prover_simpl_expal_away}
-% P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
-%   Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
-
-% \bibitem{GoedelNotes}
-% K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
-% \newblock In  \cite{sobel2004logic}, 2004.
-
-% \bibitem{Henkin50}
-% L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
-%   Logic}, {\bf 15}(2),  81--91, (1950).
-
-% \bibitem{homl}
-% R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
-%   ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
-
-% \bibitem{Isabelle}
-% T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
-%   for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
-
-% \bibitem{oppenheimera11}
-% P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
-%   of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
-%   89}(2),  333--349, (2011).
-
-% \bibitem{rushby13}
-% J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
-%   ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
-
-% \bibitem{Schulz:AICOM-2002}
-% S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
-%   15}(2),  111--126, (2002).
-
-% \bibitem{ScottNotes}
-% D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
-% \newblock In  \cite{sobel2004logic}, 2004.
-
-% \bibitem{sobel2004logic}
-% J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
-%   Cambridge U. Press, 2004.
-
-% \bibitem{sutcliffe2009tptp}
-% G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
-%   Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
-
-% \bibitem{J22}
-% G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
-%   using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
-%   {\bf 3}(1),  1--27, (2010).
-
-% \bibitem{J29}
-% B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
-%   reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
-%   (2013).
-
-%\end{thebibliography}
-
-% ------------------------------------------------------------------------
 \end{document}
-% ------------------------------------------------------------------------

### 2014-09-10 Formalizations/Isabelle/Scott/document/root.tex
Organize Isabelle files

--- 2013-11-08 Formalizations/Isabelle/ScottsVersion/output/document/root.tex
@@
 
+\begin{figure}[h]
 \noindent\fcolorbox{gray}{white}{
@@
 \item[T1] Positive properties are possibly exemplified: \hfill $\all
-  \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ \\[-1.5em]
+  \phi [P(\phi) \imp \pos \ex x \phi(x)]$ \\[-1.5em]
 \item[D1] A \emph{God-like} being possesses all positive properties: \hfill
@@
 }
+\caption{Scott's version of G\"odel's ontological argument \cite{ScottNotes}.} 
+\end{figure}
 \vskip1em
@@
 \paragraph{Acknowledgments:} Nik Sultana, Jasmin Blanchette and Larry Paulson provided 
-very important help on issues related to consistency checking in Isabelle. Jasmin Blanchette instructed us on how to 
-produce latex documents from Isabelle sources and showed us useful tricks in Isabelle.
+very important help on issues related to consistency checking in Isabelle. Jasmin Blanchette instructed us on 
+producing Isabelle sessions and he showed us some useful tricks in Isabelle.
 
@@
 
+\bibitem{rushby}
+J.~Rushby.
+\newblock The Ontological Argument in PVS. 
+\newblock {\em CAV Workshop ``Fun With Formal Methods'}, St. Petersburg, Russia, 13th of July 2013.
+
 \bibitem{ScottNotes}
@@
 \newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambr. U. Press, 2004.
+\newblock Cambridge University Press, 2004.
 

### 2014-09-10 Formalizations/THF/Run_ECAI2014_Experiments.sh
minor correction

@@ -2,6 +2,6 @@ echo "Recomputation of the experiments from the ECAI-2014 paper *Automating Goed
 echo ""
 echo "The problems tested below corresponf to the results mentioned in Figure 2 of that paper (min/max refers to the sets of dependencies tested for each conjecture statement, and the meaning of vary/const is obvious)."
 echo ""
-./call_tptp.sh T1_K_const_max.p T1_K_const_min.p T1_K_vary_max.p T1_K_vary_min.p T1_varying.p C_K_const_max.p  C_K_const_min.p  C_K_vary_max.p   C_K_vary_min.p T2_K_const_max.p  T2_K_const_min.p  T2_K_vary_max.p   T2_K_vary_min.p  T3_KB_const_max.p  T3_KB_const_min.p  T3_KB_vary_max.p   T3_KB_vary_min.p   T3_K_const_max.p   T3_K_const_min.p   T3_K_vary_max.p    T3_K_vary_min.p   MC_KB_const_max.p  MC_KB_const_min.p  MC_KB_vary_max.p   MC_KB_vary_min.p   FG_KB_const_max.p  FG_KB_const_min.p  FG_KB_vary_max.p   FG_KB_vary_min.p  MT_KB_const_max.p  MT_KB_const_min.p  MT_KB_vary_max.p   MT_KB_vary_min.p   CO_KB_const_max.p  CO_KB_vary_max.p  CO\'_KB_const_max.p  CO\'_KB_const_min.p  CO\'_KB_vary_max.p   CO\'_KB_vary_min.p  
[…]

### 2014-09-10 Papers/Universalis/ModalCollapse.tex

@@
 are available in
-\url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.
-\footnote{C.: Here
-  we should point to the exact files (and put them in a relatively
-  stable subdirectory?). We also should also use corresponding names in
-  these files, so that the reader easily finds the related 
-  parst. Right now this is still quite difficult.}
+\url{github.com/FormalTheology/GoedelGod/}
+in the files \texttt{ModalCollapse.thy} and \texttt{ModalSquareOfOpposition.thy} inside the folder \texttt{Formalizations/Isabelle/Meta} as well as in files inside the folder \texttt{Formalizations/Isabelle/Anderson}.
 

### 2014-09-10 Papers/Universalis/ModalCollapse.tex
Add paragraph about weakness in the integration

@@
 
-Ongoing and future work includes the study of modal collapse in
+Ongoing and future work includes the computer-assisted 
+study of the modal collapse in
 other variants of the ontological argument (e.g.~\cite{bjordal99,fuhrmann05:_exist_notwen}).
+Furthermore, our experiments in Isabelle revealed a weakness of the current integration of the HOL-ATPs LEO-II and Satallax via Sledgehammer: most of the problems in our study solved by the two HOL-ATPs were still too hard to be reconstructed and verified by Isabelle's internal prover Metis.
+This points to relevant future work regarding the integration of HOL-ATPs in Isabelle.
 

### 2014-09-10 Papers/Universalis/ModalCollapse.tex
Remove apostrophe from variant MC names

@@
 that this form of the modal collapse entails (in modal logic
-\textbf{\emph{K}}) a collapse of the modal square (\textbf{MC'}),
+\textbf{\emph{K}}) a collapse of the modal square (\textbf{MCs}),
 causing the subcontraries to entail 
@@
 (i.e. where the accessibility relation is reflexive), 
-even a total collapse of the modalities (\textbf{MC''}) 
+even a total collapse of the modalities (\textbf{MCt}) 
 is entailed by \textbf{MC}.
@@
   $\hol{\allq \phi [\phi \imp \nec \phi]}$ \\
-\item[\textbf{MC'}] Everything that is possible is necessary:
+\item[\textbf{MCs}] Everything that is possible is necessary:
   $\hol{\allq \phi [\pos \phi \imp \nec \phi]}$  \\
@@
   $\hol{\allq \phi [\nec \phi \imp \pos \phi]}$ \\
-\item[\textbf{MC''}] Modalities collapse completely:
+\item[\textbf{MCt}] Modalities collapse completely:
   $\hol{\allq \phi [(\phi \biimp \nec \phi) \wedge (\pos \phi \biimp \nec \phi)]}$

### 2014-09-17 Papers/2014-WADT/paper.tex
paper added

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%

\usepackage[utf8]{inputenc}

\begin{document}

\title{On Logic Embeddings and G\"odel's God}

\author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
    the German Research Foundation DFG under grants BE2501/9-1 \&
    BE2501/11-1.} \and Bruno Woltzenlogel Paleo\inst{2}
}

\institute{%Department of Mathematics and Computer Science \\ 
 Freie Universit\"at Berlin, Germany, \url{c.benzmueller@fu-berlin.de}
\and
 Vienna Technical University, Austria, \url{bruno@logic.at}}

\maketitle

% \begin{abstract}
%   Logic embeddings provide an elegant means to formalize sophisticated
%   non-classical logics in classical higher-order logic. In recent work
%   we have applied the logic embeddings approach to verify and automate
%   a prominent philosophical argument that has fascinated philosophers
%   and theologists for about 1000 years: the ontological argument for
%   the existence of God. In our work we have concentrated on Kurt
%   G\"odel's (respectively, Dana Scott's) modern version of this
%   argument, which employs a second-order modal logic.  In our ongoing
%   computer-assisted study of G\"odel's proof script, the automated
%   reasoning tools have made some interesting observations, some of
%   which were unknown so far.

%   Our work attests the maturity of contemporary interactive and
%   automated deduction tools for classical higher-order logic and
%   demonstrates the elegance and practical relevance of the
%   embeddings-based approach. Most importantly, our work opens new
%   perspectives towards a computational metaphysics.

%   In addition to the above, I will also briefly discuss the relation
%   of the logic embeddings approach to axiomatic and algebraic
%   specifications

%   The study of G\"odel's ontological proof of God's existence is joint
%   work with Bruno Woltzenlogel Paleo.
% \end{abstract}
%
% \section{Introduction}
%

Logic embeddings provide an elegant means to formalize sophisticated
non-classical logics in classical higher-order logic (HOL, Church's
simple type theory \cite{Church40}). In previous work (cf.~\cite{C35}
and the references therein) the embeddings approach has been
successfully applied to automate object-level and meta-level reasoning
for a range of logics and logic combinations with off-the-shelf HOL
theorem provers. This also includes quantified modal logics (QML)
\cite{J23} and quantified conditional logics (QCL) \cite{C37}.  For
many of the embedded logics few or none automated theorem provers did
exist before. HOL is exploited in this approach to encode the
semantics of the logics to be embedded, for example, Kripke semantics
for QMLs \cite{fitting98} or selection function semantics for QCLs
\cite{Stalnaker68}.  The embeddings approach is related to labelled
deductive systems \cite{gabbay96}, which employ meta-level
(world-)labeling techniques for the modeling and implementation of
non-classical proof systems. In our embeddings approach such labels
are instead encoded in the HOL logic.

In recent work \cite{C40,J28} we have applied the embeddings
approach to verify and automate a philosophical argument that has
fascinated philosophers and theologists for about 1000 years: the
ontological argument for the existence of
God~\cite{sobel2004logic}. We have thereby concentrated on
G\"odel's \cite{GoedelNotes}, respectively  Scott's \cite{ScottNotes},
modern
version of this argument, which employs a second-order modal logic, for
which, until now, no theorem provers were available.  In our
computer-assisted study of the argument, the HOL provers LEO-II
\cite{C26} and Satallax \cite{Satallax} have made some interesting
observations, some of which were unknown so far.

Ongoing and future work concentrates on the systematic study of
G\"odel's and Scott's proofs. We have also begun to study more recent
variants of the argument
\cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,fitting02:_types_tableaus_god,Hajek2002,Hajek2008},
which claim to remedy some fundamental problem of G\"odel's and Scott's proofs, known as the modal collapse. The long-term goal
is to work out a  landscape of the detailled logic parameters (e.g., constant
vs. varying domains, rigid vs. non-rigid terms, logics KB vs.
S4 vs.  S5, etc.) under which the proposed variants of the 
modern ontological argument hold or fail.

There is little related work~\cite{oppenheimera11,rushby13}, and this focuses solely on the comparably simpler,
original ontological argument by Anselm of Canterbury.  % These works do not achieve the close
% correspondence between the original formulations and the formal
% encodings that can be found in our approach and they do also not reach
% the same of degree of proof automation.

Our work attests the maturity of contemporary interactive and
automated deduction tools for HOL and
demonstrates the elegance and practical relevance of the
embeddings-based approach. Most importantly, our work opens new
perspectives towards a computational metaphysics.

% \paragraph{Acknowledgement:}
%   The study of G\"odel's ontological proof of God's existence is joint
%   work with Bruno Woltzenlogel Paleo.

\bibliographystyle{plain}
%\bibliography{chris,Bibliography}

\begin{thebibliography}{10}

\bibitem{AndersonGettings}
A.C. Anderson and M.~Gettings.
\newblock G\"odel ontological proof revisited.
\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
  Science, and Physics: Lecture Notes in Logic 6}}, pages 167--172. {Springer},
  1996.

\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
C.A.~Anderson.
\newblock Some emendations of {G{\"o}del's} ontological proof.
\newblock {\em Faith and Philosophy}, 7(3), 1990.

\bibitem{C37}
C. Benzm{\"u}ller.
\newblock Automating quantified conditional logics in {HOL}.
\newblock In F.~Rossi, editor, Proc. of {\em IJCAI 2013}, pages 746--753, Beijing, China, 2013.

\bibitem{C35}
C. Benzm{\"u}ller.
\newblock A top-down approach to combining logics.
\newblock In {\em Proc. of ICAART 2013}, pages 346--351, Barcelona, Spain, 2013.
  SciTePress Digital Library.

\bibitem{J28}
C. Benzm\"uller and B.~Woltzenlogel Paleo.
\newblock {G{\"o}del's God in Isabelle/HOL}.
\newblock {\em Archive of Formal Proofs}, 2013.

\bibitem{C40}
C. Benzm{\"u}ller and B.~Woltzenlogel Paleo.
\newblock Automating {G\"{o}del's} ontological proof of god's existence with
  higher-order automated theorem provers.
\newblock In 
  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
  Applications}, pages 163 -- 168. IOS Press, 2014.

\bibitem{J23}
C. Benzm{\"u}ller and L. Paulson.
\newblock Quantified multimodal logics in simple type theory.
\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
  7(1):7--20, 2013.

\bibitem{C26}
C. Benzm{\"u}ller, F. Theiss, L. Paulson, and A. Fietzke.
\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
  logic (system description).
\newblock In 
  {\em Proc. of IJCAR 2008}, volume 5195 of {\em
  LNCS}, pages 162--170. Springer, 2008.

\bibitem{bjordal99}
F. Bjørdal.
\newblock Understanding gödel’s ontological argument.
\newblock In T. Childers, editor, {\em The Logica Yearbook 1998}.
  Filosofia, 1999.

\bibitem{Satallax}
C.E. Brown.
\newblock Satallax: An automated higher-order prover.
\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
  Springer, 2012.

\bibitem{Church40}
A.~Church.
\newblock A formulation of the simple theory of types.
\newblock {\em Journal of Symbolic Logic}, 5:56--68, 1940.

\bibitem{fitting98}
M.~Fitting and R.L. Mendelsohn.
\newblock {\em First-Order Modal Logic}, volume 277 of {\em Synthese Library}.
\newblock Kluwer, 1998.

\bibitem{fitting02:_types_tableaus_god}
M.~Fitting.
\newblock {\em Types, Tableaus, and {G}{\"o}del's God}.
\newblock Kluwer, 2002.

\bibitem{fuhrmann05:_exist_notwen}
A.~Fuhrmann.
\newblock {Existenz und Notwendigkeit --- Kurt G\"odels axiomatische
  Theologie}.
\newblock In W.~Spohn et~al., editor, {\em Logik in der Philosophie}.
  Heidelberg (Synchron), 2005.

\bibitem{gabbay96}
D.M. Gabbay.
\newblock {\em Labelled Deductive Systems}.
\newblock Clarendon Press, 1996.

\bibitem{GoedelNotes}
K.~G\"odel.
\newblock {\em Appx.A: Notes in Kurt G\"odel's Hand}, pages 144--145.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{Hajek2002}
P.~Hajek.
\newblock A new small emendation of g\"odel's ontological proof.
\newblock {\em Studia Logica: An International Journal for Symbolic Logic},
  71(2):pp. 149--164, 2002.

\bibitem{Hajek2008}
P.~Hajek.
\newblock Ontological proofs of existence and non-existence.
\newblock {\em Studia Logica: An International Journal for Symbolic Logic},
  90(2):pp. 257--262, 2008.

\bibitem{oppenheimera11}
P.E. Oppenheimera and E.N. Zalta.
\newblock A computationally-discovered simplification of the ontological
  argument.
\newblock {\em Australasian J. of Philosophy}, 89(2):333--349, 2011.

\bibitem{rushby13}
J.~Rushby.
\newblock The ontological argument in {PVS}.
\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St.
  Petersburg, Russia,, 2013.

\bibitem{ScottNotes}
D.~Scott.
\newblock {\em Appx.B: Notes in Dana Scott's Hand}, pages 145--146.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{sobel2004logic}
J.H. Sobel.
\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
\newblock Cambridge U. Press, 2004.

\bibitem{Stalnaker68}
R.C.~Stalnaker.
\newblock A theory of conditionals.
\newblock In {\em Studies in Logical Theory}, pages 98--112. Blackwell, 1968.

\end{thebibliography}

\end{document}

### 2014-09-19 Papers/2014/Universalis/ModalCollapse.tex
Reorganize folders

--- 2014-09-10 Papers/Universalis/ModalCollapse.tex
@@
 assuming the \emph{modal existential import} \textbf{ExImp}
-\cite{WhatToCiteHere}.
-\footnote{C.: We need a reference here; I am still
-  searching for good papers; this here is also interesting:
-  \url{http://www.tandfonline.com/doi/pdf/10.1080/01445340.2013.764962}}
+\cite{ExImport}.
 
@@
 even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
-are equi-valid, $\collapse{\phi'}$ may not be valid.
+are equi-valid, \\ $\collapse{\phi'}$ may not be valid.
 %

### 2014-09-19 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reorganize folders

--- 2014-08-25 Papers/NaturalDeduction/GodProof-ND.tex
@@
 Anselm (1033-1109).
-His argument considered a maximally conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
+His argument considered a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
 
-A major critique of this argument is that we do not know whether the concept maximal conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
+A major critique of this argument is that we do not know whether the concept greatest conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
 %The question is if the argument is sound. 
-As Bertrand Russell pointed out, the definition of maximal allows us to define properties, like having boots, which the maximal being then also must have \citep{citation needed}.
+As Bertrand Russell pointed out, the definition of greatest allows us to define properties, like having boots, which the greatest being then also must have \citep{citation needed}.
 
@@
 To fill this gap, he argued that the properties of God, the perfections, are compatible.
-This implies that it is possible to have all perfections at once and therefore the existence of a maximal being with all these properties is possible.
+This implies that it is possible to have all perfections at once and therefore the existence of a greatest being with all these properties is possible.
 
@@
 A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons.
-Firstly, presentations of G\"oedel's proof are typically either informal or formalized in other styles of calculi (e.g.
+Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of calculi (e.g.
 Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}).
@@
 \end{definition}
+
+ToDo: decide which connectives and corresponding inference rules will be primitive.
 
@@
 \end{definition}
+
+ToDo: decide if rules for diamond should be added.
 

### 2014-10-15 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
file added

--- 2014-08-31 Papers/Universalis/ModalCollapse.tex
@@
 \title[Modal Collapse]
- {The Ontological Modal Collapse \\ 
- as a Collapse of the Square of Opposition}
+ {Computer-supported Analysis of Modern Variants of G\"odel's Ontological Argument}
 
-%----------Author 1
+%----------AUTHOR 1
 \author[Benzm\"uller]{Christoph Benzm\"uller}
@@
 %----------classification, keywords, date
-\subjclass{
-Prim. 03A02;  % Philosophical aspects of logic and foundations
-Sec. 68T02 % Artificial Intelligence
-}
+% \subjclass{
+% % Prim. 03A02;  % Philosophical aspects of logic and foundations
+% % Sec. 68T02 % Artificial Intelligence 
+% }
 
-\keywords{Modal Logics, Higher-Order Logics, Ontological Argument}
+%\keywords{Ontological Argument, Modal Collapse}
 
-\date{August 30, 2014}
+\date{\today}
 %----------additions
@@
 
-\begin{abstract}
-The \emph{modal collapse} that afflicts G\"odel's modal ontological 
-argument for God's existence is discussed from the perspective of the 
-modal square of opposition. Furthermore, a computer-assisted verification 
-of the claims that the emendations by Anderson and by Frode are immune to 
-the modal collapse is presented.
-\end{abstract}
+% \begin{abstract}
+% The \emph{modal collapse} that afflicts G\"odel's modal ontological 
+% argument for God's existence is discussed from the perspective of the 
+% modal square of opposition. Furthermore, a computer-assisted verification 
+% of the claims that the emendations by Anderson and by Frode are immune to 
+% the modal collapse is presented.
+% \end{abstract}
 
@@
 %\tableofcontents
-\section{Introduction}
-
-Attempts to prove the
-existence (or non-existence) of God by means of abstract, ontological
-arguments are an old tradition in western philosophy, with contributions by several prominent philosophers, including St. Anselm of
-Canterbury, Descartes and Leibniz.
-Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as shown in Fig.
-\ref{fig:scott}.
-
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
-  both:
-  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
-\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
-%
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Scott's version of G\"odel's ontological argument \cite{ScottNotes}.\label{fig:scott}} 
-\end{figure}
-
-% Ontological arguments, for or against the existence of God,
-% illustrate well an essential aspect of metaphysics: some (necessary) facts
-% for our existing world are deduced by purely a priori, analytical means from some
-% abstract definitions and axioms. % Contingent truths are to be
-% distinguished from necessary truths.
-
-G\"{o}del defines God (see Fig.~\ref{fig1}) as a being who possesses all \emph{positive}
-properties.  He does not extensively discuss what positive properties
-are, but instead he states a few reasonable (but debatable) axioms
-that they should satisfy.  
-The overall idea of G{\"o}del's proof is in the tradition of Anselm's
-argument, who defined God as some entity of which nothing greater can be
-conceived. Anselm argued that existence in the actual world would
-make such an assumed being even greater; hence, by definition God must
-exist. G{\"o}del's ontological argument is clearly related to this
-reasoning pattern. However, it also tries to fix some fundamental
-weaknesses in Anselm's work. For example, G{\"o}del explicitly proves
-that God's existence is possible, which has been a basic assumption for
-Anselm. Because of this, Anselm's argument has been
-criticized as incomplete by Leibniz. Leibniz claimed that the
-assumption should be derivable from the definition of God as a perfect
-being and from the notion of perfection.  G{\"o}del's proof 
-addresses this critique, and it also addresses Kant's objection that existence should not be treated as a predicate.
-
-Nevertheless, G{\"o}del's work
-still leaves room for criticism. In particular, his axioms are so
-strong that they entail a \emph{modal collapse} 
-\cite{ToDo:sobel1987?,sobel2004logic}: 
-everything that is the case is so necessarily. 
-There has been an impressive body of recent and ongoing
-work (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,ToDo:FrodeAndHajek,ContemporaryBibliography} and the references therein)
-proposing solutions for the modal collapse. 
-The goal of this paper is to present a computer-assisted analysis of 
-the solutions proposed by Anderson and by Frode.
-The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}.
-This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.
+%\section{Introduction}
 
 \section{The Modal Collapse}
+The axioms in G\"odel's ontological argument \cite{GoedelNotes,ScottNotes} imply what is
+called the modal collapse \cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box \varphi$
+holds for any formula $\varphi$ and not just for $\exists x. God(x)$
+as intended). This fact, which has recently been confirmed with
+higher-order automated theorem provers \cite{}, has led to strong
+criticism in the research community, and it has stimulated further
+adaptations of G\''odel's argument with the aim to remedy the situation
+\cite{}.
 
-A crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary.
+In recent work we have formalized and automated the proposed adaptions
+of Anderson \cite{}, Bjordahl \cite{} and Fuhrmann \cite{}.
+
+\pagebreak
+
+ crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary.
 This is Lemma L2 in G\"odel's proof.

### 2014-10-15 Formalizations/THF/call_tptp.sh
minor modifications

@@ -5,10 +5,10 @@ call=./RemoteSOT.pl
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
-proverlist='LEO-II--- Satallax--- Nitrox--- '
+proverlist='TPS---3.120601S1b LEO-II--- Satallax--- Nitrox--- '
 #proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
 
-timeout=60
+timeout=200
 
 suffix=$(date "+%Y-%m-%d--%H:%M:%S")
 resultsfile=RESULTS-$suffix

### 2014-10-17 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
update of early draft  paper

@@
 %----------AUTHOR 1
-\author[Benzm\"uller]{Christoph Benzm\"uller}
+\author[Benzm\"uller]{C. Benzm\"uller}
 
@@
 
-\thanks{This work was completed with the support of ToDo}
+\thanks{This work was supported by German National Research Foundation (DFG) under
+ grants BE 2501/9-1 and BE 2501/11-1 and by ????}
 
 %----------Author 2
-\author[Weber]{\\ Leon Weber}
+\author[Weber]{L. Weber}
 \address{ 
@@
 %----------Author 3
-\author[Woltzenlogel-Paleo]{\\ Bruno Woltzenlogel Paleo}
+\author[Woltzenlogel-Paleo]{B. Woltzenlogel Paleo}
 \address{ 
@@
 
-\section{The Modal Collapse}
-The axioms in G\"odel's ontological argument \cite{GoedelNotes,ScottNotes} imply what is
-called the modal collapse \cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box \varphi$
+%\vspace*{-5em}
+
+%\section{The Modal Collapse}
+The axioms in G\"odel's ontological argument
+\cite{GoedelNotes,ScottNotes} imply what is called the modal collapse
+\cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box \varphi$
 holds for any formula $\varphi$ and not just for $\exists x. God(x)$
 as intended). This fact, which has recently been confirmed with
-higher-order automated theorem provers \cite{}, has led to strong
-criticism in the research community, and it has stimulated further
-adaptations of G\''odel's argument with the aim to remedy the situation
-\cite{}.
+higher-order automated theorem provers \cite{J30,ECAI-2014}, has led
+to strong criticism of the argument, and it has therefore stimulated
+attempts to remedy the problem. Some authors, including Hajek
+\cite{Hajek_der_Mathematiker,Hajek_Magari_and_others}, suggest the use
+of cautious instead of full comprehension principles in order to
+escape the modal collapse. Others, including Anderson
+\cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
+and Bjordahl \cite{}, have proposed  to adapt G\"odel's axioms (in
+particular, controversial axiom A1) and definitions.
 
-In recent work we have formalized and automated the proposed adaptions
-of Anderson \cite{}, Bjordahl \cite{} and Fuhrmann \cite{}.
+We have studied --- with computer-support --- the emendations proposed
+by Anderson and Bjordahl.  More precisely, we have formalized their
+emendations within the proof assistant Isabelle/HOL \cite{Isabelle},
+and we have then automated (different variations of) their
+emendations with the higher-order reasoners Leo-II \cite{}, Satallax
+\cite{} and Nitpick \cite{} available in Isabelle/HOL.\footnote{The
+  formalizations are available in the subdirectories \url{Anderson}
+  and \url{Bjordahl} at
+  \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.}
+Our formalizations employ the embedding of higher-order modal logic
+(HOML) in classical higher-order logic (HOL) as introduced and
+employed in our previous work \cite{J30,ECAI-2014}. In this approach
+full comprehension is naturally ``built-in'' since the underlying HOL
+supports $\lambda$-abstraction. We summarize the main results of our
+experiments.
 
-\pagebreak
+Anderson's emendation (cf. Appendix \ref{apx:Anderson}) has been
+verified for different domain conditions. These variations were
+motivated by different authors who commented on Anderson's work. Anderson himself \ldots
+\begin{itemize}
+\item Constant domain semantics \footnote{Cf. Isabelle/HOL source file
+    \url{Anderson/Anderson_constant_domain.thy}}, that is, the domains of
+  objects for all types (individuals, predicates, etc.) stays the same
+  in all worlds. We consider constant domain semantics as the standard
+  setting.
+\item Varying domain semantics\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_varying_domain.thy}} for the type of indivduals individuals
+  (and constant domain for all other types). Different authors have pointed out that 
+  Ande
+\item The foll
+and constant 
+has been verified for constant domain\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_const.thy}} and varying
+  domain semantics
+\item 
+\end{itemize}
 
- crucial step of the ontological argument is the claim that if God's existence is possible, then it is necessary.
-This is Lemma L2 in G\"odel's proof.
-In the modal square of opposition (Fig.
-\ref{ToDo}), this is an unusual situation in which the \textbf{I} corner implies and entails the \textbf{A} corner, for the particular case when $\phi$ is $\exq x G(x)$.
-G\"odel's proof shows that his axioms are strong enough to invert the direction of entailment for the particular $\phi$ at issue.
-The question is whether the axioms are then not too strong, also allowing the inverted entailment for arbitrary $\phi$.
-That is essentially the question asked by Sobel \cite{Sobel1987}; and his proof of the modal collapse (MC) answers it affirmatively.
-The modal collapse is a collapse of the modal square: it shows that, assuming G\"odel's axioms, the subcontraries entail the contraries.
+ that is, the following can be proved by the HOL-ATPs:
 
-ToDo: include figure of Modal Square of Opposition.
+G\"odel's and Scott's versions of the proof are not explicit about the
+particular comprehension principles they assume.  Some authors,
+including Hajek \cite{Hajek_der_Mathematiker,Hajek_Magari_and_others},
+have suggested to consider cautious conprehension principles in order
+to escape the modal collapse.
 
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[MC] Everything that is the case is so necessarily:
-  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
-\item[MC'] Everything that is possible is necessary:
-  $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
-\item[MC''] Modalities collapse:
-  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
-\end{figure}
+% \section{Other Solutions}
 
-\clearpage
+% ToDo: Fitting
 
-\section{Anderson's Emendation}
+% \section{Conclusions}
 
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
-%
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-\end{figure}
+% Various slightly different versions of
+% axioms and definitions have been considered by G\"{o}del and by
+% several philosophers who commented on his proof
+% (cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
 
-I have meanwhile added another file ``Anderson\_var\_partial.thy''.
-In this version the actualist e-Quantifier is used exclusively in the definition of NE and everywhere else ithe possibilist Quantifiers are used.
-It is still unclear to me, what is actually meant.
-Reading Anderson 1990, last sentence on p.
-302, it could mean that ``Anderson\_var\_partial.thy'' is the correct version; but when reading AndersonGettings, then one could think that ``Anderson\_var.thy'' is meant.
-But ok we have now both versions.
-Interestingly in ``Anderson\_var\_partial.thy'' A5 is not inferable anymore, and this also holds for ``anderson\_implies\_fuhrmann".
-Nitpick finds countermodels.
+% In theoretical philosophy, formal logical confrontations with such
+% ontological arguments had been so far (mainly) limited to paper
+% and pen.  Up to now, the use of computers was prevented, because the
+% logics of the available theorem proving systems were not expressive
+% enough to formalize the abstract concepts adequately. G{\"o}del's proof
+% uses, for example, a complex higher-order modal logic (\HOML) to handle
+% concepts such as \emph{possibility} and \emph{necessity} and to support
+% quantification over individuals and properties.
 
-\clearpage
+% controversies, care with parameters
 
-\section{Frode's Alternative}
-
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
-  both:
-  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
-\end{figure}
-
-\clearpage
-
-\section{Other Solutions}
-
-ToDo: Fitting
-
-\section{Conclusions}
-
-Various slightly different versions of
-axioms and definitions have been considered by G\"{o}del and by
-several philosophers who commented on his proof
-(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-
-In theoretical philosophy, formal logical confrontations with such
-ontological arguments had been so far (mainly) limited to paper
-and pen.  Up to now, the use of computers was prevented, because the
-logics of the available theorem proving systems were not expressive
-enough to formalize the abstract concepts adequately. G{\"o}del's proof
-uses, for example, a complex higher-order modal logic (\HOML) to handle
-concepts such as \emph{possibility} and \emph{necessity} and to support
-quantification over individuals and properties.
-
-controversies, care with parameters
-
-ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
+% ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
 
@@
 
+\begin{appendix}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+% \begin{itemize}
+% \item[MC] Everything that is the case is so necessarily:
+%   $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
+% \item[MC'] Everything that is possible is necessary:
+%   $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
+% \item[MC''] Modalities collapse:
+%   $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
+% \end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
+% \end{figure}
+
+%\clearpage
+
+\section{Anderson's Emendation} \label{apx:Anderson}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
+\end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
+% \end{figure}
+
+% I have meanwhile added another file ``Anderson\_var\_partial.thy''.
+In this version the actualist e-Quantifier is used exclusively in the definition of NE and everywhere else ithe possibilist Quantifiers are used.
+It is still unclear to me, what is actually meant.
+Reading Anderson 1990, last sentence on p.
+302, it could mean that ``Anderson\_var\_partial.thy'' is the correct version; but when reading AndersonGettings, then one could think that ``Anderson\_var.thy'' is meant.
+But ok we have now both versions.
+Interestingly in ``Anderson\_var\_partial.thy'' A5 is not inferable anymore, and this also holds for ``anderson\_implies\_fuhrmann".
+Nitpick finds countermodels.
+
+%\clearpage
+
+\section{Bjordahl's Alternative} \label{apx:Bjordahl}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:
+  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
+\item[D1] A \emph{God-like} being possesses all positive properties: 
+  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
+\item[A3]  The property of being God-like is positive: 
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
+  \exq y \phi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
+% \end{figure}
+
+%\clearpage
+\end{appendix}
+
 % ------------------------------------------------------------------------

### 2014-10-17 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
draft update

@@
 \cite{GoedelNotes,ScottNotes} imply what is called the modal collapse
-\cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box \varphi$
-holds for any formula $\varphi$ and not just for $\exists x. God(x)$
-as intended). This fact, which has recently been confirmed with
-higher-order automated theorem provers \cite{J30,ECAI-2014}, has led
-to strong criticism of the argument, and it has therefore stimulated
-attempts to remedy the problem. Some authors, including Hajek
+\cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box
+\varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
+just for $\exists x. God(x)$ as intended). This fact, which has
+recently been confirmed with higher-order automated theorem provers
+\cite{J30,ECAI-2014}, has led to strong criticism of the argument, and
+it has therefore stimulated attempts to remedy the problem. Some
+authors, including Hajek
 \cite{Hajek_der_Mathematiker,Hajek_Magari_and_others}, suggest the use
@@
 emendations within the proof assistant Isabelle/HOL \cite{Isabelle},
-and we have then automated (different variations of) their
-emendations with the higher-order reasoners Leo-II \cite{}, Satallax
-\cite{} and Nitpick \cite{} available in Isabelle/HOL.\footnote{The
-  formalizations are available in the subdirectories \url{Anderson}
-  and \url{Bjordahl} at
+and we have then automated (different variations of) their emendations
+with the higher-order reasoners Leo-II \cite{}, Satallax \cite{},
+Metis \cite{} and Nitpick \cite{}; these systems are available in
+Isabelle/HOL.  Our formalizations\footnote{The formalizations are
+  available in the subdirectories \url{Anderson} and \url{Bjordahl} at
   \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.}
-Our formalizations employ the embedding of higher-order modal logic
-(HOML) in classical higher-order logic (HOL) as introduced and
-employed in our previous work \cite{J30,ECAI-2014}. In this approach
-full comprehension is naturally ``built-in'' since the underlying HOL
-supports $\lambda$-abstraction. We summarize the main results of our
-experiments.
+employ the embedding of higher-order modal logic (HOML) in classical
+higher-order logic (HOL) as introduced and employed in our previous
+work \cite{J30,ECAI-2014}. In this approach full comprehension is
+naturally ``built-in'' since the underlying HOL supports
+$\lambda$-abstraction.
 
-Anderson's emendation (cf. Appendix \ref{apx:Anderson}) has been
-verified for different domain conditions. These variations were
-motivated by different authors who commented on Anderson's work. Anderson himself \ldots
+We summarize the main results of our experiments, and we start with
+Anderson's emendation (cf. Appendix \ref{apx:Anderson}), which we have
+analysed for different domain conditions. These variations were
+motivated by various comments on Anderson's work in the literature. 
 \begin{itemize}
-\item Constant domain semantics \footnote{Cf. Isabelle/HOL source file
-    \url{Anderson/Anderson_constant_domain.thy}}, that is, the domains of
-  objects for all types (individuals, predicates, etc.) stays the same
-  in all worlds. We consider constant domain semantics as the standard
-  setting.
-\item Varying domain semantics\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_varying_domain.thy}} for the type of indivduals individuals
-  (and constant domain for all other types). Different authors have pointed out that 
-  Ande
-\item The foll
-and constant 
-has been verified for constant domain\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_const.thy}} and varying
-  domain semantics
-\item 
+\item Constant domain semantics\footnote{Cf. Isabelle/HOL source file
+    \url{Anderson/Anderson_constant_domain.thy}} (the individual
+  domains remain the same in all worlds): G\"odel's theorems C1, and
+  T3 can easily be automated (in logics K and KB, respectively) ,
+  G\"odel's axioms A4 and A5 are proven redundant (the former one in
+  logic K4B, the latter already in K); a trivial countermodel (two
+  worlds, one individual; this is simpler than
+          the countermodel with two individuals and two worlds
+          presented by Anderson, 1990) to MC generated by Nitpick (for all mentioned
+  logics); this model also proves consistency.
+\item Varying domain semantics\footnote{Cf. Isabelle/HOL source file
+    \url{Anderson/Anderson_varying_domain.thy}} (the domains for
+  indivduals may vary from world to world, all other domains remain
+  constant; different authors have pointed out that varying domain
+  quantifiers should be used in Anderson's work; here we consistently
+  used varying domain for all individual quantifiers; the varying
+  domain quantifiers have been modeled by using an explicit
+  ``Existence'' relation between indiviudals and worlds): we obtain
+  the same results as  above
+\item Mixed variant\footnote{Cf. Isabelle/HOL source file
+    \url{Anderson/Anderson_mixed_domain.thy}} (varying domain
+  quantifiers are used exclusively in the definitions of essence and
+  NE; see the comments by Fitting): G\"odel's axioms A4 and A5 remain
+  redundant (the former one in
+  logic K4B, the latter already in K).
 \end{itemize}
-
- that is, the following can be proved by the HOL-ATPs:
-
-G\"odel's and Scott's versions of the proof are not explicit about the
-particular comprehension principles they assume.  Some authors,
-including Hajek \cite{Hajek_der_Mathematiker,Hajek_Magari_and_others},
-have suggested to consider cautious conprehension principles in order
-to escape the modal collapse.
 

### 2014-10-17 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
draft update

@@
   the same results as  above
-\item Mixed variant\footnote{Cf. Isabelle/HOL source file
-    \url{Anderson/Anderson_mixed_domain.thy}} (varying domain
+\item Mixed variant\footnote{Cf. Isabelle/HOL source files
+    \url{Anderson/Anderson_mixed_domain_1/2.thy}} (varying domain
   quantifiers are used exclusively in the definitions of essence and
@@
   redundant (the former one in
-  logic K4B, the latter already in K).
+  logic K4B, the latter already in K). The situation changes when only
+  NE is modeled with a varying domain quantifier. Now, Nitpick reports
+  a countermodel. 
 \end{itemize}

### 2014-10-18 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
minor modifications

@@
 \cite{GoedelNotes,ScottNotes} imply what is called the modal collapse
-\cite{Sobel1987,SobelBook} (formula $\varphi \rightarrow \Box
+\cite{Sobel1987,SobelBook2004} (formula $\varphi \rightarrow \Box
 \varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
@@
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
-and Bjordahl \cite{}, have proposed to adapt G\"odel's axioms (in
+and Bjordahl \cite{Bjordahl}, have proposed to adapt G\"odel's axioms (in
 particular, controversial axiom A1) and definitions.
@@
 and we have then automated (different variations of) their emendations
-with the higher-order reasoners Leo-II \cite{}, Satallax \cite{},
-Metis \cite{} and Nitpick \cite{}; these systems are available in
-Isabelle/HOL.  Our formalizations\footnote{The formalizations are
-  available in the subdirectories \url{Anderson} and \url{Bjordahl} at
+with the higher-order reasoners Leo-II \cite{C26}, Satallax
+\cite{brown2012satallax}, Metis \cite{Hurd03first-orderproof} and
+Nitpick \cite{BN10}; these systems are available in Isabelle/HOL.  Our
+formalizations\footnote{The formalizations are available in the
+  subdirectories \url{Anderson} and \url{Bjordahl} at
   \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.}
@@
 naturally ``built-in'' since the underlying HOL supports
-$\lambda$-abstraction.
+$\lambda$-abstraction. The assumption of full comprehension, up to our
+understanding, is
+in-line with Anderson and Bjordahl (?).
 
@@
 analysed for different domain conditions. These variations were
-motivated by various comments on Anderson's work in the literature. 
+motivated by various comments on Anderson's work in the literature
+(maybe we should be more explicit).
 \begin{itemize}
@@
 
+Need to do Bjordahl \ldots.
+
 % \section{Other Solutions}
@@
 % several philosophers who commented on his proof
-% (cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+% (cf{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
 
@@
 
-\begin{thebibliography}{1}
+\bibliographystyle{plain}
+\bibliography{Bibliography}
 
-\bibitem{Adams}
-R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
-  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
+% \begin{thebibliography}{1}
 
-\bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
-  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
-  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
+% \bibitem{Adams}
+% R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
+%   Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
 
-\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
-C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
-  Faith and Philosophy}, {\bf 7}(3), (1990).
+% \bibitem{AndersonGettings}
+% A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
+%   {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
+%   Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
 
-\bibitem{Andrews:gmae72}
-P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
-  Logic}, {\bf 37}(2),  395--397, (1972).
+% \bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
+% C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
+%   Faith and Philosophy}, {\bf 7}(3), (1990).
 
-\bibitem{andrewsSEP}
-P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
-  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
+% \bibitem{Andrews:gmae72}
+% P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
+%   Logic}, {\bf 37}(2),  395--397, (1972).
 
-\bibitem{C36}
-C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
-  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
-  pp. 232--233, Rio de Janeiro, Brazil, (2013).
+% \bibitem{andrewsSEP}
+% P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
+%   Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
 
-\bibitem{B5}
-C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
-  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
-  Elsevier, (2014).
-\newblock Forthcoming; preliminary version available at
-  {http://christoph-benzmueller.de/papers/B5.pdf}.
+% \bibitem{C36}
+% C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
+%   4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
+%   pp. 232--233, Rio de Janeiro, Brazil, (2013).
 
-\bibitem{C34}
-C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
-  provers for first-order modal logics', in {\em Proc. of the 20th European
-  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
+% \bibitem{B5}
+% C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
+%   Handbook of the History of Logic, Volume 9 --- Logic and Computation},
+%   Elsevier, (2014).
+% \newblock Forthcoming; preliminary version available at
+%   {http://christoph-benzmueller.de/papers/B5.pdf}.
 
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
-  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
-  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
-  386--406, College Publications, (2008).
+% \bibitem{C34}
+% C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
+%   provers for first-order modal logics', in {\em Proc. of the 20th European
+%   Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
 
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
-  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
+% \bibitem{B9}
+% C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
+%   logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
+%   {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
+%   386--406, College Publications, (2008).
 
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
-  cooperative automatic theorem prover for higher-order logic', in {\em
-  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
+% \bibitem{J23}
+% C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
+%   type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
 
-\bibitem{J30}
-C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
-  automation of {G{\"o}del's} proof of {God's} existence', {\em
-  arXiv:1308.4526}, (2013).
+% \bibitem{LEO-II}
+% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
+%   cooperative automatic theorem prover for higher-order logic', in {\em
+%   Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
 
-\bibitem{J28}
-C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
-  {\em Archive of Formal Proofs}, (2013).
+% \bibitem{J30}
+% C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
+%   automation of {G{\"o}del's} proof of {God's} existence', {\em
+%   arXiv:1308.4526}, (2013).
 
-\bibitem{W50}
-C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
-  in {\em Proceedings of the 10th International Workshop on the Implementation
-  of Logics}, EPiC Series. EasyChair, (2013).
-\newblock Invited abstract.
+% \bibitem{J28}
+% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
+%   {\em Archive of Formal Proofs}, (2013).
 
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
-  Development}}, Springer, 2004.
+% \bibitem{W50}
+% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
+%   in {\em Proceedings of the 10th International Workshop on the Implementation
+%   of Logics}, EPiC Series. EasyChair, (2013).
+% \newblock Invited abstract.
 
-\bibitem{Nitpick}
-J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
-  higher-order logic based on a relational model finder', in {\em Proc. of ITP
-  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
+% \bibitem{Coq}
+% Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
+%   Development}}, Springer, 2004.
 
-\bibitem{Satallax}
-C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
-  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
+% \bibitem{Nitpick}
+% J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
+%   higher-order logic based on a relational model finder', in {\em Proc. of ITP
+%   2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
 
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
-  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
+% \bibitem{Satallax}
+% C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
+%   IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
 
-\bibitem{Fitting}
-M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
+% \bibitem{ContemporaryBibliography}
+% R.~Corazzon.
+% \newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
+%   \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
 
-\bibitem{fitting98}
-M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
-  {\em Synthese Library}, Kluwer, 1998.
+% \bibitem{Fitting}
+% M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
 
-\bibitem{Gallin75}
-D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
+% \bibitem{fitting98}
+% M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
+%   {\em Synthese Library}, Kluwer, 1998.
 
-\bibitem{garbacz12:_prover_simpl_expal_away}
-P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
-  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
+% \bibitem{Gallin75}
+% D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
 
-\bibitem{GoedelNotes}
-K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{garbacz12:_prover_simpl_expal_away}
+% P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
+%   Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
 
-\bibitem{Henkin50}
-L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
-  Logic}, {\bf 15}(2),  81--91, (1950).
+% \bibitem{GoedelNotes}
+% K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{homl}
-R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
-  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
+% \bibitem{Henkin50}
+% L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
+%   Logic}, {\bf 15}(2),  81--91, (1950).
 
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
-  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
+% \bibitem{homl}
+% R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
+%   ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
 
-\bibitem{oppenheimera11}
-P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
-  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
-  89}(2),  333--349, (2011).
+% \bibitem{Isabelle}
+% T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
+%   for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
 
-\bibitem{rushby13}
-J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
-  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
+% \bibitem{oppenheimera11}
+% P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
+%   of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
+%   89}(2),  333--349, (2011).
 
-\bibitem{Schulz:AICOM-2002}
-S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
-  15}(2),  111--126, (2002).
+% \bibitem{rushby13}
+% J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
+%   ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
 
-\bibitem{ScottNotes}
-D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{Schulz:AICOM-2002}
+% S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
+%   15}(2),  111--126, (2002).
 
-\bibitem{sobel2004logic}
-J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
-  Cambridge U. Press, 2004.
+% \bibitem{ScottNotes}
+% D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{sutcliffe2009tptp}
-G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
-  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
+% \bibitem{sobel2004logic}
+% J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
+%   Cambridge U. Press, 2004.
 
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
-  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
-  {\bf 3}(1),  1--27, (2010).
+% \bibitem{sutcliffe2009tptp}
+% G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
+%   Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
 
-\bibitem{J29}
-B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
-  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
-  (2013).
+% \bibitem{J22}
+% G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
+%   using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
+%   {\bf 3}(1),  1--27, (2010).
 
-\end{thebibliography}
+% \bibitem{J29}
+% B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
+%   reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
+%   (2013).
+
+%\end{thebibliography}
 

### 2014-10-18 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
updates for paper

@@
 \usepackage{commands}
+
+\usepackage[utf8]{inputenc}
+
 %
@@
 
-\title[Modal Collapse]
+\title[Analysis of Modern Variants of G\"odel's Ontological Argument]
  {Computer-supported Analysis of Modern Variants of G\"odel's Ontological Argument}
@@
 \address{%
-Department of Mathematics and Computer Science\\
-Arnimallee 7 \\
-Room 115 \\
-14195 Berlin \\
-Germany
+Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
 }
-
-\email{c.benzmueller@gmail.com}
+\email{c.benzmueller@fu-berlin.com}
 
@@
 \address{ 
-ToDo \\
-ToDo \\
-ToDo Berlin \\
-Germany
+Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
 }
@@
 
-%\vspace*{-5em}
+\vspace*{-5em}
 
@@
 The axioms in G\"odel's ontological argument 
-\cite{GoedelNotes,ScottNotes} imply what is called the modal collapse
+\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) imply what is called the modal collapse
 \cite{Sobel1987,SobelBook2004} (formula $\varphi \rightarrow \Box
@@
 recently been confirmed with higher-order automated theorem provers
-\cite{J30,ECAI-2014}, has led to strong criticism of the argument, and
+\cite{C40,J30}, has led to strong criticism of the argument, and
 it has therefore stimulated attempts to remedy the problem. Some
 authors, including Hajek
-\cite{Hajek_der_Mathematiker,Hajek_Magari_and_others}, suggest the use
+\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996}, suggest the use
 of cautious instead of full comprehension principles in order to
@@
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
-and Bjordahl \cite{Bjordahl}, have proposed to adapt G\"odel's axioms (in
-particular, controversial axiom A1) and definitions.
+and Bjordal \cite{bjordal99}, propose to slightly modify
+G\"odel's axioms and definitions.
 
 We have studied --- with computer-support --- the emendations proposed
-by Anderson and Bjordahl.  More precisely, we have formalized their
-emendations within the proof assistant Isabelle/HOL \cite{Isabelle},
-and we have then automated (different variations of) their emendations
-with the higher-order reasoners Leo-II \cite{C26}, Satallax
-\cite{brown2012satallax}, Metis \cite{Hurd03first-orderproof} and
-Nitpick \cite{BN10}; these systems are available in Isabelle/HOL.  Our
-formalizations\footnote{The formalizations are available in the
-  subdirectories \url{Anderson} and \url{Bjordahl} at
+by Anderson and Bjordal (thereby we have retained full comprehension).
+More precisely, we have formalized their emendations within the proof
+assistant Isabelle/HOL \cite{Isabelle}, and we have then automated
+(different variations of) their emendations with the higher-order
+reasoners Leo-II \cite{C26}, Satallax \cite{brown2012satallax}, Metis
+\cite{Hurd03first-orderproof} and Nitpick \cite{Nitpick}; these systems
+are available in Isabelle/HOL.  Our formalizations\footnote{The
+  formalizations are available in the subdirectories \url{Anderson}
+  and \url{Bjordal} at
   \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.}
 employ the embedding of higher-order modal logic (HOML) in classical
-higher-order logic (HOL) as introduced and employed in our previous
-work \cite{J30,ECAI-2014}. In this approach full comprehension is
+higher-order logic (HOL) as introduced and employed in previous
+work \cite{C40,J30,J23}. In this approach full comprehension is
 naturally ``built-in'' since the underlying HOL supports
-$\lambda$-abstraction. The assumption of full comprehension, up to our
-understanding, is
-in-line with Anderson and Bjordahl (?).
+$\lambda$-abstraction.
 
 We summarize the main results of our experiments, and we start with
-Anderson's emendation (cf. Appendix \ref{apx:Anderson}), which we have
-analysed for different domain conditions. These variations were
-motivated by various comments on Anderson's work in the literature
-(maybe we should be more explicit).
+Anderson's emendation (cf. Appendix \ref{apx:Anderson}; @Bruno: I do
+not yet address T2', L1' and L2' below; we could eventually include those as well?),
+which we have analysed for different domain conditions. These
+variations were motivated by various comments on Anderson's work in
+the literature (@Bruno: maybe we should be more explicit about these
+comments).
 \begin{itemize}
@@
     \url{Anderson/Anderson_constant_domain.thy}} (the individual
-  domains remain the same in all worlds): G\"odel's theorems C1, and
-  T3 can easily be automated (in logics K and KB, respectively) ,
+  domains remain unchanged for all worlds): G\"odel's theorems T1, C
+  and T3' can be quickly automated (in logics K and KB, respectively),
   G\"odel's axioms A4 and A5 are proven redundant (the former one in
-  logic K4B, the latter already in K); a trivial countermodel (two
-  worlds, one individual; this is simpler than
-          the countermodel with two individuals and two worlds
-          presented by Anderson, 1990) to MC generated by Nitpick (for all mentioned
-  logics); this model also proves consistency.
+  logic K4B and the latter one already in K). Moreover, a trivial
+  countermodel (consisting of two worlds and two individuals) to MC
+  generated by Nitpick (for all mentioned logics); this model also
+  proves consistency. 
 \item Varying domain semantics\footnote{Cf. Isabelle/HOL source file
@@
   indivduals may vary from world to world, all other domains remain
-  constant; different authors have pointed out that varying domain
+  constant; various authors have pointed out that varying domain
   quantifiers should be used in Anderson's work; here we consistently
-  used varying domain for all individual quantifiers; the varying
-  domain quantifiers have been modeled by using an explicit
-  ``Existence'' relation between indiviudals and worlds): we obtain
-  the same results as  above
+  used varying domain for all individual quantifiers; the modeling of
+  the varying domain quantifiers employs an explicit ``Existence''
+  relation between indiviudals and worlds as guard): In this setting we obtain the
+  same results as above.
 \item Mixed variant\footnote{Cf. Isabelle/HOL source files
     \url{Anderson/Anderson_mixed_domain_1/2.thy}} (varying domain
-  quantifiers are used exclusively in the definitions of essence and
-  NE; see the comments by Fitting): G\"odel's axioms A4 and A5 remain
-  redundant (the former one in
-  logic K4B, the latter already in K). The situation changes when only
-  NE is modeled with a varying domain quantifier. Now, Nitpick reports
-  a countermodel. 
+  quantifiers are used only in the definitions of essence and NE;
+  cf.~Fitting comments to Anderson in
+  \cite{fitting02:_types_tableaus_god}): Also in this setting we
+  obtain the same results as above. However, if a varying domain
+  quantifier is used only in the definition of NE, then the
+  situation slightly changes. Now axiom A5 is no longer provable and
+  a countermodel is reported by Nitpick. The remaining results are as before.
+
+%   G\"odel's axioms A4 and A5
+%   remain redundant (the former one in logic K4B, the latter already in
+%   K). The situation changes when only NE is modeled with a varying
+%   domain quantifier. Now, Nitpick reports a countermodel.
 \end{itemize}
 
-Need to do Bjordahl \ldots.
+Our analysis of Bjordal emendation of G\"odel's argument (cf. Appendix \ref{apx:Bjordal}) has produced the following results:
+\begin{itemize}
+\item Constant domain semantics\footnote{Cf. Isabelle/HOL source files
+    \url{Bjordal/Bjordal_A/B/C_constant_domain.thy}}: G\"odel's axiom
+  A2, A3 can be quickly automatically derived in logic K from
+  Bjordal's definition B:D. A4 can be proved in logic T
+  (reflexivity). Proving G\"odel's D1 from B:D is possible in logic
+  K4. Vice versa, for the proof that B:D follows from D1, A2, A3 and
+  A4 the povers succeed already in logic K. Hence, Bjordal's lemma
+  B:L1 holds in logic S4. The provers also show that theorem T3
+  follows from B:D, B:A1 and B:A2 already in logic KB. Modal collapse
+  does not follow in Bjordal's setting as Nitpick demonstrates with a
+  countermodel (consisting of two worlds and one individual). @Bruno: we could eventually also
+  also say that for MC1 we need a bigger countermodel?
+\item Varying domain semantics\footnote{Cf. Isabelle/HOL source files
+    \url{Bjordal/Bjordal_A/B/C_varying_domain.thy}}: The results are
+  the same as above. 
+\item @Bruno: we could also check the mixed case her; but there doesn't seem to be strong reason for doing so.
+\end{itemize}
+
+Summary (what else can we say here, feel free to add): Using our approach, the formalization and (partly) automated
+analysis of different variants of Anderson's and Bjordal's emendations
+of G\"odel's ontological argument has been surprisingly
+straightforward. The provers confirmed the claimed results and in a
+few cases they have even contributed some novel insights. The
+weakening of the comprehension principles would clearly constitute
+another interesting parameter for further experiments. However, this
+seems hard to achieve in our approach, since full comprehension is
+naturally built-in.
 
@@
 
+\newpage
 \begin{appendix}
-
+\noindent @Bruno: I think we need more compact presentation of the proofs.
 % \begin{figure}[t]
@@
 %\clearpage
+
+\small
+
+\section{Scott's version of G\"odel's ontological argment} \label{apx:Goedel} 
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:
+  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[D1] A \emph{God-like} being possesses all positive properties: 
+  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
+\item[A3]  The property of being God-like is positive: 
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
+\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
+%
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\end{itemize}
 
@@
 
-\section{Bjordahl's Alternative} \label{apx:Bjordahl}
+\section{Bjordal's Alternative} \label{apx:Bjordal}
 
@@
 % \begin{minipage}{.94\columnwidth}\small
+In Bjordal's emendation $G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).
 \begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
-  both:
-  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
+\item[B:D] A formulas $\phi$ is positive iff it is necessarily the case
+  that anything which is God-like has the property $\phi$.
+  $$\hol{P(\phi) \equiv \nec \allq x (G(x) \imp \phi(x))}$$ 
+\item[B:L1] D is logically equivalent in S4 with the union of G\"odel's definition D1 and axioms A2, A3 and A4.
+  $$\hol{D \biimp D1 \wedge A2 \wedge A3 \wedge A4}$$
+  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
+\item[B:D2] $\phi$ is a maximal composite of object $x$'s positive properties iff $x$ has $\phi$ and $\phi$ is positive and all positive properties $\psi$ which $x$ has are such that is necessarily the case that all objects which have $\phi$ also have $\psi$.
+  $$\hol{MCP(\phi,x) \equiv (\phi(x) \wedge P(\phi)) \wedge \allq \psi ((\psi(x) \wedge P(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
+\item[B:D3] $x$ has the $N$-property iff x is such that if $\phi$ is a maximal composite of $x$'s positive properties then it is necessary that some object $y$ has the property $\phi$.
+  $$\hol{N(x) \equiv \allq \phi (MCP(\phi,x) \imp \nec \allq y \phi(y))}$$
+\item[B:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+\item[B:A2] The $N$-property is positive.
+ $$\hol{P(N)}$$
 \item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 

### 2014-10-20 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Improve abstract

@@
 \usepackage{commands}
+\usepackage{xcolor}
+\usepackage{comment}
 
@@
 
-\title[Analysis of Modern Variants of G\"odel's Ontological Argument]
- {Computer-supported Analysis of Modern Variants of G\"odel's Ontological Argument}
+\title[Analysis of Emendations of G\"odel's Ontological Argument]
+ {Computer-Assisted Analysis of 
+ Emendations of G\"odel's Ontological Proof}
 
@@
 \thanks{This work was supported by German National Research Foundation (DFG) under
- grants BE 2501/9-1 and BE 2501/11-1 and by ????}
+ grants BE 2501/9-1 and BE 2501/11-1.}
 
@@
 \address{ 
-Favoritenstra{\ss}e 9 \\
-Room HA0402 \\
-1040 Wien \\
-Austria
+Room HA0402, Favoritenstra{\ss}e 9, 1040 Wien, Austria
 }
@@
 The axioms in G\"odel's ontological argument 
-\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) imply what is called the modal collapse
-\cite{Sobel1987,SobelBook2004} (formula $\varphi \rightarrow \Box
+\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail what is called the \emph{modal collapse}
+\cite{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
 \varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
-just for $\exists x. God(x)$ as intended). This fact, which has
+just for $\exists x. \mathit{God}(x)$ as intended. This fact, which has
 recently been confirmed with higher-order automated theorem provers
-\cite{C40,J30}, has led to strong criticism of the argument, and
-it has therefore stimulated attempts to remedy the problem. Some
-authors, including Hajek
-\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996}, suggest the use
-of cautious instead of full comprehension principles in order to
-escape the modal collapse. Others, including Anderson
+\cite{C40,J30}, has led to strong criticism of the argument and stimulated attempts to remedy the problem. Hájek
+\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996} proposed the use
+of cautious instead of full comprehension principles, and Fitting \cite{fitting02:_types_tableaus_god} suggested that greater care is necessary in the semantics of higher-order quantifiers in the presence of modalities.
+Others, such as
+Anderson
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
-and Bjordal \cite{bjordal99}, propose to slightly modify
+and Bjordal \cite{bjordal99}, proposed slight emendations of
 G\"odel's axioms and definitions.
-
-We have studied --- with computer-support --- the emendations proposed
-by Anderson and Bjordal (thereby we have retained full comprehension).
-More precisely, we have formalized their emendations within the proof
-assistant Isabelle/HOL \cite{Isabelle}, and we have then automated
-(different variations of) their emendations with the higher-order
+They require neither comprehension restrictions nor more complex semantics.
+Therefore, they are technically simpler to analyze with computer support.
+We have formalized them using the proof
+assistant Isabelle/HOL \cite{Isabelle} together with the automated higher-order
 reasoners Leo-II \cite{C26}, Satallax \cite{brown2012satallax}, Metis
-\cite{Hurd03first-orderproof} and Nitpick \cite{Nitpick}; these systems
-are available in Isabelle/HOL.  Our formalizations\footnote{The
+\cite{Hurd03first-orderproof}, and Nitpick \cite{Nitpick}.  
+Our formalizations\footnote{The
   formalizations are available in the subdirectories \url{Anderson}
@@
 employ the embedding of higher-order modal logic (HOML) in classical
-higher-order logic (HOL) as introduced and employed in previous
-work \cite{C40,J30,J23}. In this approach full comprehension is
-naturally ``built-in'' since the underlying HOL supports
-$\lambda$-abstraction.
+higher-order logic (HOL) as introduced in previous
+work \cite{C40,J30,J23}. We explored the effect of different domain conditions
+ on the provability of lemmas, theorems and even axioms.
+This was motivated by a controversy between Hájek and Anderson regarding the redundancy of some axioms in Anderson's emendation.
+In \emph{constant domain semantics}, the individual domains are the same in all possible worlds.
+In \emph{varying domain semantics}, the domains may vary from world to world.
+This variation is technically encoded with the help of an existence relation expressing which individuals actually exist in each world.
+Quantifiers are then uniformly formalized as \emph{actualist quantifiers} (i.e.
+guarded by the existence relation).
+Our main results are summarized in the following sections.
 
-We summarize the main results of our experiments, and we start with
-Anderson's emendation (cf. Appendix \ref{apx:Anderson}; @Bruno: I do
-not yet address T2', L1' and L2' below; we could eventually include those as well?),
-which we have analysed for different domain conditions. These
-variations were motivated by various comments on Anderson's work in
-the literature (@Bruno: maybe we should be more explicit about these
-comments).
-\begin{itemize}
-\item Constant domain semantics\footnote{Cf. Isabelle/HOL source file
-    \url{Anderson/Anderson_constant_domain.thy}} (the individual
-  domains remain unchanged for all worlds): G\"odel's theorems T1, C
+\section{Anderson's Emendation (cf. Appendix \ref{apx:Anderson})}
+
+\BWP{Bruno: I do
+not yet address T2', L1' and L2' below; we could eventually include those as well?}
+
+\BWP{Bruno: maybe we should be more explicit about these
+comments}
+
+\paragraph{Constant domain semantics}
+%\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_constant_domain.thy}} 
+     G\"odel's theorems T1, C
   and T3' can be quickly automated (in logics K and KB, respectively),
@@
   generated by Nitpick (for all mentioned logics); this model also
-  proves consistency. 
-\item Varying domain semantics\footnote{Cf. Isabelle/HOL source file
-    \url{Anderson/Anderson_varying_domain.thy}} (the domains for
-  indivduals may vary from world to world, all other domains remain
-  constant; various authors have pointed out that varying domain
-  quantifiers should be used in Anderson's work; here we consistently
-  used varying domain for all individual quantifiers; the modeling of
-  the varying domain quantifiers employs an explicit ``Existence''
-  relation between indiviudals and worlds as guard): In this setting we obtain the
+  shows the consistency of the axioms.
+
+\paragraph{Varying domain semantics}
+%\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_varying_domain.thy}} 
+: In this setting we obtain the
   same results as above.
-\item Mixed variant\footnote{Cf. Isabelle/HOL source files
-    \url{Anderson/Anderson_mixed_domain_1/2.thy}} (varying domain
+
+\paragraph{Mixed variant}
+%\footnote{Cf. Isabelle/HOL source files \url{Anderson/Anderson_mixed_domain_1/2.thy}} 
+(varying domain
   quantifiers are used only in the definitions of essence and NE;
-  cf.~Fitting comments to Anderson in
+  cf.~Fitting's comments to Anderson in
   \cite{fitting02:_types_tableaus_god}): Also in this setting we
@@
   quantifier is used only in the definition of NE, then the
-  situation slightly changes. Now axiom A5 is no longer provable and
+  situation changes slightly. Now axiom A5 is no longer provable and
   a countermodel is reported by Nitpick. The remaining results are as before.
@@
 %   domain quantifier. Now, Nitpick reports a countermodel.
-\end{itemize}
 
-Our analysis of Bjordal emendation of G\"odel's argument (cf. Appendix \ref{apx:Bjordal}) has produced the following results:
-\begin{itemize}
-\item Constant domain semantics\footnote{Cf. Isabelle/HOL source files
-    \url{Bjordal/Bjordal_A/B/C_constant_domain.thy}}: G\"odel's axiom
+\section{Bjordal's Emendation (cf. Appendix \ref{apx:Bjordal})}
+
+\paragraph{Constant domain semantics}
+%\footnote{Cf. Isabelle/HOL source files \url{Bjordal/Bjordal_A/B/C_constant_domain.thy}}: 
+    G\"odel's axiom
   A2, A3 can be quickly automatically derived in logic K from
@@
   (reflexivity). Proving G\"odel's D1 from B:D is possible in logic
-  K4. Vice versa, for the proof that B:D follows from D1, A2, A3 and
-  A4 the povers succeed already in logic K. Hence, Bjordal's lemma
+  K4. Conversely, the proof that B:D follows from D1, A2, A3 and
+  A4 is possible already in logic K. Hence, Bjordal's lemma
   B:L1 holds in logic S4. The provers also show that theorem T3
@@
   does not follow in Bjordal's setting as Nitpick demonstrates with a
-  countermodel (consisting of two worlds and one individual). @Bruno: we could eventually also
-  also say that for MC1 we need a bigger countermodel?
-\item Varying domain semantics\footnote{Cf. Isabelle/HOL source files
-    \url{Bjordal/Bjordal_A/B/C_varying_domain.thy}}: The results are
+  countermodel (consisting of two worlds and one individual).
+
+  \BWP{@Bruno: we could eventually also
+  also say that for MC1 we need a bigger countermodel?}
+\paragraph{Varying domain semantics}
+%\footnote{Cf. Isabelle/HOL source files \url{Bjordal/Bjordal_A/B/C_varying_domain.thy}}
+: The results are
   the same as above.
-\item @Bruno: we could also check the mixed case her; but there doesn't seem to be strong reason for doing so.
-\end{itemize}
+
+\section{Conclusions}
+
+Anderson's emendation (cf. Appendix \ref{apx:Anderson}; ,
+which we have analysed for different domain conditions. These
+variations were motivated by various comments on Anderson's work in
+the literature.
+
+In this approach full comprehension is
+naturally ``built-in'' since the underlying HOL supports
+$\lambda$-abstraction.
 

### 2014-10-20 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Reduce redundancy and boldface logic names

@@
 
+\usepackage{xspace}
 \usepackage{url}
@@
 \usepackage{comment}
+\usepackage{modallogics}
 
@@
 %\section{The Modal Collapse}
+\noindent
 The axioms in G\"odel's ontological argument 
@@
 
-\BWP{Bruno: maybe we should be more explicit about these
-comments}
+For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold: 
+%
+G\"odel's theorems T1, C and T3' can be quickly automated (in logics \K and \KB, respectively); \BWP{3->2 respectively}
+%
+G\"odel's axioms A4 and A5 are proven redundant (the former one in
+  logic \KFourB and the latter one already in \K).
+%
+Moreover, a trivial countermodel (consisting of two worlds and 
+two individuals) to MC generated by Nitpick (for all mentioned logics); this model also shows the consistency of the axioms.
 
-\paragraph{Constant domain semantics}
-%\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_constant_domain.thy}} 
-     G\"odel's theorems T1, C
-  and T3' can be quickly automated (in logics K and KB, respectively),
-  G\"odel's axioms A4 and A5 are proven redundant (the former one in
-  logic K4B and the latter one already in K). Moreover, a trivial
-  countermodel (consisting of two worlds and two individuals) to MC
-  generated by Nitpick (for all mentioned logics); this model also
-  shows the consistency of the axioms.
-
-\paragraph{Varying domain semantics}
-%\footnote{Cf. Isabelle/HOL source file \url{Anderson/Anderson_varying_domain.thy}} 
-: In this setting we obtain the
-  same results as above.
+ToDo: tell the story of the controversy here, and then link with the mixed variant.
 
@@
 
-\paragraph{Constant domain semantics}
-%\footnote{Cf. Isabelle/HOL source files \url{Bjordal/Bjordal_A/B/C_constant_domain.thy}}: 
+For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold:
+
     G\"odel's axiom
-  A2, A3 can be quickly automatically derived in logic K from
-  Bjordal's definition B:D. A4 can be proved in logic T
+  A2, A3 can be quickly automatically derived in logic \K from
+  Bjordal's definition B:D. A4 can be proved in logic \KT
   (reflexivity). Proving G\"odel's D1 from B:D is possible in logic
-  K4. Conversely, the proof that B:D follows from D1, A2, A3 and
-  A4 is possible already in logic K. Hence, Bjordal's lemma
-  B:L1 holds in logic S4. The provers also show that theorem T3
-  follows from B:D, B:A1 and B:A2 already in logic KB. Modal collapse
+  \KFour. Conversely, the proof that B:D follows from D1, A2, A3 and
+  A4 is possible already in logic \K. Hence, Bjordal's lemma
+  B:L1 holds in logic \SFour. The provers also show that theorem T3
+  follows from B:D, B:A1 and B:A2 already in logic \KB. Modal collapse
   does not follow in Bjordal's setting as Nitpick demonstrates with a
@@
   also say that for MC1 we need a bigger countermodel?}
-\paragraph{Varying domain semantics}
-%\footnote{Cf. Isabelle/HOL source files \url{Bjordal/Bjordal_A/B/C_varying_domain.thy}}
-: The results are
-  the same as above.
+
+\BWP{ToDo: discuss the fact that Bjordal's proof is more resistant to MC than Anderson's}
 

### 2014-10-21 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Explain constant and varying domains

@@
 % \begin{abstract}
-% The \emph{modal collapse} that afflicts G\"odel's modal ontological 
-% argument for God's existence is discussed from the perspective of the 
-% modal square of opposition. Furthermore, a computer-assisted verification 
-% of the claims that the emendations by Anderson and by Frode are immune to 
-% the modal collapse is presented.
 % \end{abstract}
@@
 %\tableofcontents
-%\section{Introduction}
 
-\vspace*{-5em}
-
-%\section{The Modal Collapse}
-\noindent
-The axioms in G\"odel's ontological argument 
-\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail what is called the \emph{modal collapse}
+\noindent The axioms in G\"odel's ontological argument
+\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail
+what is called \emph{modal collapse}
 \cite{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
 \varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
-just for $\exists x. \mathit{God}(x)$ as intended. This fact, which has
-recently been confirmed with higher-order automated theorem provers
-\cite{C40,J30}, has led to strong criticism of the argument and stimulated attempts to remedy the problem. Hájek
-\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996} proposed the use
-of cautious instead of full comprehension principles, and Fitting \cite{fitting02:_types_tableaus_god} suggested that greater care is necessary in the semantics of higher-order quantifiers in the presence of modalities.
-Others, such as
-Anderson
+just for $\exists x. \mathit{God}(x)$ as intended. This fact, which
+has recently been confirmed with higher-order automated theorem
+provers \cite{C40,J30}, has led to strong criticism of the argument
+and stimulated attempts to remedy the problem. Hájek
+\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996}
+proposed the use of cautious instead of full comprehension principles,
+and Fitting \cite{fitting02:_types_tableaus_god} suggested that
+greater care is necessary in the semantics of higher-order quantifiers
+in the presence of modalities. Others, such as Anderson
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
-and Bjordal \cite{bjordal99}, proposed slight emendations of
-G\"odel's axioms and definitions.
-They require neither comprehension restrictions nor more complex semantics.
-Therefore, they are technically simpler to analyze with computer support.
-We have formalized them using the proof
-assistant Isabelle/HOL \cite{Isabelle} together with the automated higher-order
-reasoners Leo-II \cite{C26}, Satallax \cite{brown2012satallax}, Metis
-\cite{Hurd03first-orderproof}, and Nitpick \cite{Nitpick}.  
-Our formalizations\footnote{The
-  formalizations are available in the subdirectories \url{Anderson}
-  and \url{Bjordal} at
-  \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.}
-employ the embedding of higher-order modal logic (HOML) in classical
-higher-order logic (HOL) as introduced in previous
-work \cite{C40,J30,J23}. We explored the effect of different domain conditions
- on the provability of lemmas, theorems and even axioms.
-This was motivated by a controversy between Hájek and Anderson regarding the redundancy of some axioms in Anderson's emendation.
-In \emph{constant domain semantics}, the individual domains are the same in all possible worlds.
-In \emph{varying domain semantics}, the domains may vary from world to world.
-This variation is technically encoded with the help of an existence relation expressing which individuals actually exist in each world.
-Quantifiers are then uniformly formalized as \emph{actualist quantifiers} (i.e.
-guarded by the existence relation).
-Our main results are summarized in the following sections.
+and Bjordal \cite{bjordal99}, proposed slight emendations of G\"odel's
+axioms and definitions. They require neither comprehension
+restrictions nor more complex semantics. Therefore, they are
+technically simpler to analyze with computer support. We have
+formalized them using the proof assistant Isabelle/HOL \cite{Isabelle}
+together with the automated higher-order reasoners Leo-II \cite{C26},
+Satallax \cite{brown2012satallax}, Metis 
+\cite{Hurd03first-orderproof}, and Nitpick \cite{Nitpick}.   Our
+formalizations\footnote{The   formalizations are available in the
+subdirectories \url{Anderson}   and \url{Bjordal} at   \url{https://gi
+thub.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/
+}.} employ the embedding of higher-order modal logic (HOML) in
+classical higher-order logic (HOL) as introduced in previous work
+\cite{C40,J30,J23}. We explored the effect of different domain
+conditions  on the provability of lemmas, theorems and even axioms.
+This was motivated by a controversy between Hájek and Anderson
+regarding the redundancy of some axioms in Anderson's emendation. In
+\emph{constant domain semantics}, the individual domains are the same
+in all possible worlds. In \emph{varying domain semantics}, the
+domains may vary from world to world. This variation is technically
+encoded with the help of an existence relation expressing which
+individuals actually exist in each world. Quantifiers are then
+uniformly formalized as \emph{actualistic quantifiers} (i.e. guarded by
+the existence relation). Our main results are summarized in the
+following sections.
 
@@
 
-\BWP{Bruno: I do
-not yet address T2', L1' and L2' below; we could eventually include those as well?}
+%\BWP{Bruno: I do not yet address T2', L1' and L2' below; we could eventually include those as well?}
 
@@
 %
-G\"odel's theorems T1, C and T3' can be quickly automated (in logics \K and \KB, respectively); \BWP{3->2 respectively}
+the theorems T1, C and T3' can be quickly automated (in logics \K, \K and \KB, respectively);
 %
-G\"odel's axioms A4 and A5 are proven redundant (the former one in
-  logic \KFourB and the latter one already in \K).
+the axioms A4 and A5 are proven redundant (the former in
+logic \KFourB and the latter already in \K);
 %
-Moreover, a trivial countermodel (consisting of two worlds and 
-two individuals) to MC generated by Nitpick (for all mentioned logics); this model also shows the consistency of the axioms.
+a trivial countermodel (consisting of two worlds and 
+two individuals) to MC generated by Nitpick (for all mentioned logics), which also shows the consistency of the axioms.
+
+The redundancy of A4 and A5 is particularly controversial.
+Magari \cite{TODO} claimed that the redundancy occurs already with Gödel's original axioms and definitions.
+Hájek \cite{Hajek 1996 TODO} investigated this further and claimed that Magari's claim is invalid, but true under the assumption of an additional axiom (PEP).
+Moreover, Hájek \cite{Hajek 1996 TODO} claimed that the redundancy occurs for Anderson's emended axioms and definitions \cite{Anderson 1990}.
+Anderson and Gettings \cite{AG1996, footnote 1, page 1} rebutted Hájek's claim, arguing that it holds only under constant domain semantics, while the emended argument by Anderson ought to be taken under Cocchiarella's semantics \cite{Cocchiarella} (a particular kind of varying domain semantics).
+Hájek \cite{Hajek 2002, page 7} acknowledges this rebuttal, and apparently accepts it, as evidenced by his use of A4 and A5 in his new small emendation (replacing A:A1 and A2 by a new axiom H:A12) of Anderson's variant with varying domains \cite{Hajek 2002, section 4}.
+Nevertheless, he does show yet another emended version where A4 and A5 are redundant, but A3 is replaced by a stronger axiom additionally stating that the property of actual existence is positive \cite{Hajek 2002, section 5}.
+Our results show that Hájek was originally right, under both constant and varying domain semantics with no need to strengthen A3 (though this need might exist when H:A12 is used instead of A:A1 and A2).
+It should be noted, however, that \cite{Anderson, footnote 14} vaguely remarks that only the quantifiers in T3' and in A:D2  need to be interpreted as actualistic quantifiers, while other may be taken as possibilistic quantifiers.
+We have inv  mixed variant.
 
@@
 \paragraph{Mixed variant}
-%\footnote{Cf. Isabelle/HOL source files \url{Anderson/Anderson_mixed_domain_1/2.thy}} 
 (varying domain
@@
 
-% \section{Other Solutions}
-
-% ToDo: Fitting
-
-% \section{Conclusions}
-
-% Various slightly different versions of
-% axioms and definitions have been considered by G\"{o}del and by
-% several philosophers who commented on his proof
-% (cf{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-
-% In theoretical philosophy, formal logical confrontations with such
-% ontological arguments had been so far (mainly) limited to paper
-% and pen.  Up to now, the use of computers was prevented, because the
-% logics of the available theorem proving systems were not expressive
-% enough to formalize the abstract concepts adequately. G{\"o}del's proof
-% uses, for example, a complex higher-order modal logic (\HOML) to handle
-% concepts such as \emph{possibility} and \emph{necessity} and to support
-% quantification over individuals and properties.
-
-% controversies, care with parameters
-
-% ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}
-
-% \begin{defn}
-% This serves as environment for definitions. Note that the text
-% appears not in italics.
-% \end{defn}
-
-% \begin{equation}\label{testequation}
-% \text{This is a sample equation: } c^2=a^2+b^2
-% \end{equation}
-
-% \begin{thm}[Main Theorem]
-% In contrast to definitions, theorems appear typeset in italics as
-% it has become more or less standard in most textbooks and
-% monographs. Equations can be cited using the \verb+\eqref+ command which
-% automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
-% \end{thm}
-
-% \begin{proof}
-% A special environment is predefined: the \textit{proof} environment. Please use
-% \begin{verbatim}\begin{proof}\end{verbatim}
-% proof of the statement
-% \begin{verbatim}\end{proof}\end{verbatim}
-% for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
-% \end{proof}
-
-% There are two known problems with the placement of the end-of-proof sign:
-
-% \begin{enumerate}
-%   \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
-% be placed in the line below; if you want to avoid this, write your line in the form
-% \begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
-% which results in
-
-% \begin{proof}
-% $$displayed math line \eqno\qedhere$$
-% \end{proof}
-% \item if your proof ends with an aligned displayed environment, the command
-% \verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
-% \begin{verbatim}
-% \begin{align*}
-% \alpha&=\beta+\gamma\\
-% &=\delta+\epsilon\tag*{\qed}
-% \end{align*}
-% \end{verbatim}
-% results in
-% \begin{align*}
-% \alpha&=\beta+\gamma\\
-% &=\delta+\epsilon\tag*{\qed}
-% \end{align*}
-% \end{enumerate}
-% Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
-% and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.
-
-% \begin{rem}
-% Additional comments are being typeset without boldfaced entrance
-% word as they may be minor important.
-% \end{rem}
-
-% \begin{ex}
-% For some constructs, even no number is required.
-% \end{ex}
-
-% Displayed equations may be numbered like the following one:
-% \begin{equation}
-% \sqrt{1-\sin^2(x)}=|\cos(x)|.
-% \end{equation}
-
-% ------------------------------------------------------------------------
-
-% \subsection*{Acknowledgment}
-% Many thanks to ...
-
 \bibliographystyle{plain}
@@
 
-% \begin{thebibliography}{1}
-
-% \bibitem{Adams}
-% R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
-%   Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).
-
-% \bibitem{AndersonGettings}
-% A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
-%   {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
-%   Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).
-
-% \bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
-% C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
-%   Faith and Philosophy}, {\bf 7}(3), (1990).
-
-% \bibitem{Andrews:gmae72}
-% P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
-%   Logic}, {\bf 37}(2),  395--397, (1972).
-
-% \bibitem{andrewsSEP}
-% P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
-%   Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).
-
-% \bibitem{C36}
-% C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
-%   4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
-%   pp. 232--233, Rio de Janeiro, Brazil, (2013).
-
-% \bibitem{B5}
-% C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
-%   Handbook of the History of Logic, Volume 9 --- Logic and Computation},
-%   Elsevier, (2014).
-% \newblock Forthcoming; preliminary version available at
-%   {http://christoph-benzmueller.de/papers/B5.pdf}.
-
-% \bibitem{C34}
-% C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
-%   provers for first-order modal logics', in {\em Proc. of the 20th European
-%   Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).
-
-% \bibitem{B9}
-% C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
-%   logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
-%   {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
-%   386--406, College Publications, (2008).
-
-% \bibitem{J23}
-% C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
-%   type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).
-
-% \bibitem{LEO-II}
-% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
-%   cooperative automatic theorem prover for higher-order logic', in {\em
-%   Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).
-
-% \bibitem{J30}
-% C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
-%   automation of {G{\"o}del's} proof of {God's} existence', {\em
-%   arXiv:1308.4526}, (2013).
-
-% \bibitem{J28}
-% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
-%   {\em Archive of Formal Proofs}, (2013).
-
-% \bibitem{W50}
-% C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
-%   in {\em Proceedings of the 10th International Workshop on the Implementation
-%   of Logics}, EPiC Series. EasyChair, (2013).
-% \newblock Invited abstract.
-
-% \bibitem{Coq}
-% Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
-%   Development}}, Springer, 2004.
-
-% \bibitem{Nitpick}
-% J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
-%   higher-order logic based on a relational model finder', in {\em Proc. of ITP
-%   2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).
-
-% \bibitem{Satallax}
-% C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
-%   IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).
-
-% \bibitem{ContemporaryBibliography}
-% R.~Corazzon.
-% \newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
-%   \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.
-
-% \bibitem{Fitting}
-% M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.
-
-% \bibitem{fitting98}
-% M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
-%   {\em Synthese Library}, Kluwer, 1998.
-
-% \bibitem{Gallin75}
-% D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.
-
-% \bibitem{garbacz12:_prover_simpl_expal_away}
-% P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
-%   Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).
-
-% \bibitem{GoedelNotes}
-% K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
-% \newblock In  \cite{sobel2004logic}, 2004.
-
-% \bibitem{Henkin50}
-% L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
-%   Logic}, {\bf 15}(2),  81--91, (1950).
-
-% \bibitem{homl}
-% R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
-%   ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).
-
-% \bibitem{Isabelle}
-% T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
-%   for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.
-
-% \bibitem{oppenheimera11}
-% P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
-%   of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
-%   89}(2),  333--349, (2011).
-
-% \bibitem{rushby13}
-% J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
-%   ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).
-
-% \bibitem{Schulz:AICOM-2002}
-% S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
-%   15}(2),  111--126, (2002).
-
-% \bibitem{ScottNotes}
-% D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
-% \newblock In  \cite{sobel2004logic}, 2004.
-
-% \bibitem{sobel2004logic}
-% J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
-%   Cambridge U. Press, 2004.
-
-% \bibitem{sutcliffe2009tptp}
-% G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
-%   Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).
-
-% \bibitem{J22}
-% G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
-%   using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
-%   {\bf 3}(1),  1--27, (2010).
-
-% \bibitem{J29}
-% B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
-%   reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
-%   (2013).
-
-%\end{thebibliography}
-
 \newpage
 \begin{appendix}
-\noindent @Bruno: I think we need more compact presentation of the proofs.
-% \begin{figure}[t]
-% \noindent \framebox[\columnwidth][r]{
-% \begin{minipage}{.94\columnwidth}\small
-% \begin{itemize}
-% \item[MC] Everything that is the case is so necessarily:
-%   $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
-% \item[MC'] Everything that is possible is necessary:
-%   $$\hol{\allq \phi [\pos \phi \imp \nec \phi]}$$ 
-% \item[MC''] Modalities collapse:
-%   $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
-% \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
-% \end{figure}
+\noindent
 
-%\clearpage
+\BWP{I think we need more compact presentation of the proofs.}
 

### 2014-10-23 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Correct reference to Cocchiarella's semantics

@@
 Moreover, Hájek \cite{Hajek 1996 TODO} claimed that the redundancy occurs for Anderson's emended axioms and definitions \cite{Anderson 1990}.
-Anderson and Gettings \cite{AG1996, footnote 1, page 1} rebutted Hájek's claim, arguing that it holds only under constant domain semantics, while the emended argument by Anderson ought to be taken under Cocchiarella's semantics \cite{Cocchiarella} (a particular kind of varying domain semantics).
+Anderson and Gettings \cite{AG1996, footnote 1, page 1} rebutted Hájek's claim, arguing that it holds only under constant domain semantics, while the emended argument by Anderson ought to be taken under Cocchiarella's semantics \cite{Cocchiarella} (a varying domain semantics).
 Hájek \cite{Hajek 2002, page 7} acknowledges this rebuttal, and apparently accepts it, as evidenced by his use of A4 and A5 in his new small emendation (replacing A:A1 and A2 by a new axiom H:A12) of Anderson's variant with varying domains \cite{Hajek 2002, section 4}.
-Nevertheless, he does show yet another emended version where A4 and A5 are redundant, but A3 is replaced by a stronger axiom additionally stating that the property of actual existence is positive \cite{Hajek 2002, section 5}.
+Nevertheless, he does show yet another emended version where A4 and A5 are redundant, if A3 is replaced by a stronger axiom additionally stating that the property of actual existence is positive \cite{Hajek 2002, section 5}.
 Our results show that Hájek was originally right, under both constant and varying domain semantics with no need to strengthen A3 (though this need might exist when H:A12 is used instead of A:A1 and A2).

### 2014-10-24 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Write many paragraphs about Hajek's proofs and switch Bibligraphy management from Bibtex to Biblatex

@@
 
+\usepackage[utf8]{inputenc}
+
 \usepackage{xspace}
@@
 \usepackage{modallogics}
+\usepackage[backend=bibtex,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
+\addbibresource[label=bibs]{Bibliography.bib}
 
-\usepackage[utf8]{inputenc}
+\newcommand{\AOE}{$\mathcal{AOE}$}
+\newcommand{\AOEH}{$\mathcal{AOE}'$}
+\newcommand{\AOEHH}{$\mathcal{AOE}'_0$}
+\newcommand{\AOEHHH}{$\mathcal{AOE}''$}
 
@@
 
-\title[Analysis of Emendations of G\"odel's Ontological Argument]
- {Computer-Assisted Analysis of 
- Emendations of G\"odel's Ontological Proof}
+\title[The Anderson-Hájek Ontological Controversy]
+{Computer-Assisted Analysis of the \\ 
+Anderson-Hájek Ontological Controversy}
 
@@
 
-\noindent The axioms in G\"odel's ontological argument
-\cite{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail
+\noindent The axioms in G\"odel's ontological proof
+\citep{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail
 what is called \emph{modal collapse}
-\cite{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
+\citep{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
 \varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
@@
 has recently been confirmed with higher-order automated theorem
-provers \cite{C40,J30}, has led to strong criticism of the argument
-and stimulated attempts to remedy the problem. Hájek
-\cite{Hajek_der_Mathematiker_2002,Hajek_Magari_and_others_1996}
+provers \citep{C40,J30}, has led to strong criticism of the argument
+and stimulated attempts to remedy the problem.
+\citet{Hajek_Magari_and_others_1996,Hajek_der_Mathematiker_2001}
 proposed the use of cautious instead of full comprehension principles,
-and Fitting \cite{fitting02:_types_tableaus_god} suggested that
+and \citet{fitting02:_types_tableaus_god} suggested that
 greater care is necessary in the semantics of higher-order quantifiers
-in the presence of modalities. Others, such as Anderson
-\cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}
-and Bjordal \cite{bjordal99}, proposed slight emendations of G\"odel's
+in the presence of modalities. Others, such as
+\citet{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}, \citet{Hajek2002}
+and Bjordal \citet{bjordal99}, proposed slight emendations of G\"odel's
 axioms and definitions. They require neither comprehension
@@
 technically simpler to analyze with computer support. We have
-formalized them using the proof assistant Isabelle/HOL \cite{Isabelle}
-together with the automated higher-order reasoners Leo-II \cite{C26},
-Satallax \cite{brown2012satallax}, Metis 
-\cite{Hurd03first-orderproof}, and Nitpick \cite{Nitpick}.   Our
+formalized them using the proof assistant Isabelle/HOL \citep{Isabelle}
+together with the automated higher-order reasoners Leo-II \citep{C26},
+Satallax \citep{brown2012satallax}, Metis 
+\citep{Hurd03first-orderproof}, and Nitpick \citep{Nitpick}.   Our
 formalizations\footnote{The   formalizations are available in the
-subdirectories \url{Anderson}   and \url{Bjordal} at   \url{https://gi
+subdirectories \url{Anderson}, \url{Hájek} and \url{Bjordal} at \url{gi
 thub.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/
@@
 classical higher-order logic (HOL) as introduced in previous work
-\cite{C40,J30,J23}. We explored the effect of different domain
+\citep{C40,J30,J23}. We explored the effect of different domain
 conditions on the provability of lemmas, theorems and even axioms.
@@
 uniformly formalized as \emph{actualistic quantifiers} (i.e. guarded by
-the existence relation). Our main results are summarized in the
-following sections.
+the existence relation). Our main results are summarized here.
 
-\section{Anderson's Emendation (cf. Appendix \ref{apx:Anderson})}
+%\section{Anderson's Emendation (cf. Appendix \ref{apx:Anderson})}
 
-%\BWP{Bruno: I do not yet address T2', L1' and L2' below; we could eventually include those as well?}
-
-For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold: 
+For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold for \emph{Anderson's Emendation} (cf. Appendix \ref{apx:Anderson}): 
 %
@@
 %
-the axioms A4 and A5 are proven redundant (the former in
+the axioms A4 and A5 are proven redundant\footnote{
+  An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is entailed by $S$. 
+} (the former in
 logic \KFourB and the latter already in \K);
@@
 a trivial countermodel (consisting of two worlds and 
-two individuals) to MC generated by Nitpick (for all mentioned logics), which also shows the consistency of the axioms.
+two individuals) for MC is generated by Nitpick (for all mentioned logics); all axioms and definitions are shown to be mutually consistent.
 
-The redundancy of A4 and A5 is particularly controversial.
-Magari \cite{TODO} claimed that the redundancy occurs already with Gödel's original axioms and definitions.
-Hájek \cite{Hajek 1996 TODO} investigated this further and claimed that Magari's claim is invalid, but true under the assumption of an additional axiom (PEP).
-Moreover, Hájek \cite{Hajek 1996 TODO} claimed that the redundancy occurs for Anderson's emended axioms and definitions \cite{Anderson 1990}.
-Anderson and Gettings \cite{AG1996, footnote 1, page 1} rebutted Hájek's claim, arguing that it holds only under constant domain semantics, while the emended argument by Anderson ought to be taken under Cocchiarella's semantics \cite{Cocchiarella} (a varying domain semantics).
-Hájek \cite{Hajek 2002, page 7} acknowledges this rebuttal, and apparently accepts it, as evidenced by his use of A4 and A5 in his new small emendation (replacing A:A1 and A2 by a new axiom H:A12) of Anderson's variant with varying domains \cite{Hajek 2002, section 4}.
-Nevertheless, he does show yet another emended version where A4 and A5 are redundant, if A3 is replaced by a stronger axiom additionally stating that the property of actual existence is positive \cite{Hajek 2002, section 5}.
-Our results show that Hájek was originally right, under both constant and varying domain semantics with no need to strengthen A3 (though this need might exist when H:A12 is used instead of A:A1 and A2).
-It should be noted, however, that \cite{Anderson, footnote 14} vaguely remarks that only the quantifiers in T3' and in A:D2  need to be interpreted as actualistic quantifiers, while other may be taken as possibilistic quantifiers.
-We have inv  mixed variant.
+The redundancy of A4 and A5 is particularly controversial. Magari \cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
+  An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is entailed by $S \setminus \{ A \}$.
+}, arguing that T3 is true in all models of the other axioms and definitions by Gödel.
+\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, refuted Magari's claim, but claimed that A4 and A5 are indeed superfluous under the assumption of an additional axiom (PEP, pointwise equality for positiveness), which Magari probably assumed silently.
+Moreover, Hájek \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
+\citet[footnote 1 in p.~1]{AG1996, footnote 1, page 1}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics).
+Our results show that Hájek was originally right, under both constant and varying domain semantics.
+
+Nevertheless, \citet[p.~7]{Hajek2002} (in page 7) acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
+  A4 and A5 are used by \citet[p.~11]{Hajek2002} (page 11) in the proofs of, respectively, Lemma 4 and Theorem 4.
+} 
+of A4 and A5, as well as varying domain semantics, in his new emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf.
+Appendix \ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler axiom H:A12.
+Surprisingly, the computer-assisted formalization of \AOEH shows that A4 and A5 are superfluous (not needed for proving T3).
+Moreover, A4 and A5 are independent\footnote{
+  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
+} of the other axioms and definitions. Therefore, A4 and A5 are not redundant, despite their superfluousness.
+
+Although, Hájek did not notice the superfluousness of A4 and A5 in his \AOEH, he did describe yet another emendation (his \AOEHH, cf.
+Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no claim is made w.r.t.
+to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive \cite[sec.~5]{Hajek2002}.
+Formalization of \AOEHH shows that A4 is not only superflous, but also redundant.
+For A5, no conclusive results were achieved; neither a proof nor a countermodel could be automatically generated.
+Surprisingly, a countermodel for the weaker A3 was succesfully generated.
+This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH does not entail the positiveness of being God-like.
+
+Nevertheless, \AOEHH is explicitly regarded by \citet[p.~12]{Hajek2002} (page 12) as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
+That is his final emendation (\AOEHHH, cf.
+\ref{apx:Hajek3}), which restores A3 and does use A4 and A5, albeit in a modified form (i.e.
+H:A4 and H:A5).
+The formalization of \AOEHHH shows that H:A4 is independent.
+The old A5 is independent as well, and both H:A4 and H:A5 are superfluous, but no conclusive results were achieved regarding independence or redundancy of H:A5.
+
+Additionally, Anderson's 14th footnote \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers.
+Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant).
+However, T3 becomes unprovable.
 
@@
 
-\bibliographystyle{plain}
-\bibliography{Bibliography}
+%\bibliographystyle{plain}
+%\bibliography{Bibliography}
+
+\printbibliography
 
@@
 
-% I have meanwhile added another file ``Anderson\_var\_partial.thy''.
-In this version the actualist e-Quantifier is used exclusively in the definition of NE and everywhere else ithe possibilist Quantifiers are used.
-It is still unclear to me, what is actually meant.
-Reading Anderson 1990, last sentence on p.
-302, it could mean that ``Anderson\_var\_partial.thy'' is the correct version; but when reading AndersonGettings, then one could think that ``Anderson\_var.thy'' is meant.
-But ok we have now both versions.
-Interestingly in ``Anderson\_var\_partial.thy'' A5 is not inferable anymore, and this also holds for ``anderson\_implies\_fuhrmann".
-Nitpick finds countermodels.
+\section{Hájek's First Emendation \AOEH} \label{apx:Hajek1}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
+\end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
+% \end{figure}
+
+\section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
+\end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
+% \end{figure}
+
+\section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}
+
+% \begin{figure}[t]
+% \noindent \framebox[\columnwidth][r]{
+% \begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+\item[A3']  The property of being God-like is positive: 
+  $$\hol{P(G_A)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive: 
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
+\end{itemize}
+% \end{minipage}
+% } \vskip-.5em
+% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
+% \end{figure}
 

### 2014-10-24 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Clean the paper slightly

@@
 However, T3 becomes unprovable.
-
-ToDo: tell the story of the controversy here, and then link with the mixed variant.
-
-\paragraph{Mixed variant}
-(varying domain
-  quantifiers are used only in the definitions of essence and NE;
-  cf.~Fitting's comments to Anderson in
-  \cite{fitting02:_types_tableaus_god}): Also in this setting we
-  obtain the same results as above. However, if a varying domain
-  quantifier is used only in the definition of NE, then the
-  situation changes slightly. Now axiom A5 is no longer provable and
-  a countermodel is reported by Nitpick. The remaining results are as before.
-
-%   G\"odel's axioms A4 and A5
-%   remain redundant (the former one in logic K4B, the latter already in
-%   K). The situation changes when only NE is modeled with a varying
-%   domain quantifier. Now, Nitpick reports a countermodel.
 

### 2014-10-24 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Add ToDo
Add missing bibliography items

@@
 proposed the use of cautious instead of full comprehension principles,
-and \citet{fitting02:_types_tableaus_god} suggested that
-greater care is necessary in the semantics of higher-order quantifiers
+and \citet{fitting02:_types_tableaus_god} took
+greater care of the semantics of higher-order quantifiers
 in the presence of modalities. Others, such as
-\citet{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings}, \citet{Hajek2002}
-and Bjordal \citet{bjordal99}, proposed slight emendations of G\"odel's
+\citet{anderson90:_some_emend_of_goedel_ontol_proof}, \citet{Hajek2002}
+and \citet{bjordal99}, proposed emendations of G\"odel's
 axioms and definitions. They require neither comprehension
@@
 
+For all emendations and variants discussed here, the axioms and definitions have been shown to be consistent and 
+not to entail modal collapse.
+
 For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold for \emph{Anderson's Emendation} (cf. Appendix \ref{apx:Anderson}): 
@@
 Moreover, Hájek \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
-\citet[footnote 1 in p.~1]{AG1996, footnote 1, page 1}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics).
+\citet[footnote 1 in p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics).
 Our results show that Hájek was originally right, under both constant and varying domain semantics.
@@
 Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant).
-However, T3 becomes unprovable.
+Unfortunately, T3 then becomes unprovable.
 
-\section{Bjordal's Emendation (cf. Appendix \ref{apx:Bjordal})}
+The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
+already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable.
+These claims have been confirmed by the automated theorem provers.
+Within the same trend, the alternative proposed by \citet{bjordal99} (cf.
+Appendix \ref{apx:Bjordal}) achieves a high level of minimality.
+He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being.
+He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D, A2, A3 and A4.
+This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K.
+A4 can be proved in logic \KT (i.e.
+assuming reflexivity of the accessibility relation).
+Proving D1 is possible in logic \KFour.
+Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K.
+The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB.
+Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with certain additional presuppositions, to eliminate A5 from his theory as well.
+Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous.
+All these results hold for both constant and varying domain semantics.
 
-For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold:
+Using our approach, the formalization and (partly) automated
+analysis of several variants of G\"odel's ontological argument has been surprisingly
+straightforward. The provers not only confirmed many claimed results, but also exposed a few mistakes and novel insights in a long standing controversy. 
+We believe the technology employed in this work is ready to be 
+fruitfully adopted in larger scale by philosophers.
 
-    G\"odel's axiom
-  A2, A3 can be quickly automatically derived in logic \K from
-  Bjordal's definition B:D. A4 can be proved in logic \KT
-  (reflexivity). Proving G\"odel's D1 from B:D is possible in logic
-  \KFour. Conversely, the proof that B:D follows from D1, A2, A3 and
-  A4 is possible already in logic \K. Hence, Bjordal's lemma
-  B:L1 holds in logic \SFour. The provers also show that theorem T3
-  follows from B:D, B:A1 and B:A2 already in logic \KB. Modal collapse
-  does not follow in Bjordal's setting as Nitpick demonstrates with a
-  countermodel (consisting of two worlds and one individual).
-
-  \BWP{@Bruno: we could eventually also
-  also say that for MC1 we need a bigger countermodel?}
-
-\BWP{ToDo: discuss the fact that Bjordal's proof is more resistant to MC than Anderson's}
-
-\section{Conclusions}
-
-Anderson's emendation (cf. Appendix \ref{apx:Anderson}; ,
-which we have analysed for different domain conditions. These
-variations were motivated by various comments on Anderson's work in
-the literature.
-
-In this approach full comprehension is
-naturally ``built-in'' since the underlying HOL supports
-$\lambda$-abstraction.
-
-Summary (what else can we say here, feel free to add): Using our approach, the formalization and (partly) automated
-analysis of different variants of Anderson's and Bjordal's emendations
-of G\"odel's ontological argument has been surprisingly
-straightforward. The provers confirmed the claimed results and in a
-few cases they have even contributed some novel insights. The
-weakening of the comprehension principles would clearly constitute
-another interesting parameter for further experiments. However, this
-seems hard to achieve in our approach, since full comprehension is
-naturally built-in.
-
-%\bibliographystyle{plain}
-%\bibliography{Bibliography}
+% ToDo: \\
+% 3 - Fix Hajek's proofs in the appendices. (they are just copy-pasted from Anderson's so far)
+% 4 - Check apostrophes in axiom names that depend on different definitions.
+% 5 - Add subscripts to alternative definitions (e.g. G_A)
+% 6 - Fix citet redundancy
+% 7 - Add apendix for Anderson's simplification
 
@@
 \noindent
-
-\BWP{I think we need more compact presentation of the proofs.}
 
@@
 \begin{itemize}
-\item[B:D] A formulas $\phi$ is positive iff it is necessarily the case
+\item[B:D1] A formulas $\phi$ is positive iff it is necessarily the case
   that anything which is God-like has the property $\phi$.
@@
   $$\hol{N(x) \equiv \allq \phi (MCP(\phi,x) \imp \nec \allq y \phi(y))}$$
-\item[B:A1] If a property is positive, its negation is not positive:
+\item[A:A1] If a property is positive, its negation is not positive:
   $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[B:A2] The $N$-property is positive.
+\item[A5'] The $N$-property is positive.
  $$\hol{P(N)}$$
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
 \end{itemize}

### 2014-10-26 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
fixed some typos

@@
 formalizations\footnote{The   formalizations are available in the
-subdirectories \url{Anderson}, \url{Hájek} and \url{Bjordal} at \url{gi
+subdirectories \url{Anderson}, \url{Hajek} and \url{Bjordal} at \url{gi
 thub.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/
@@
 \citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, refuted Magari's claim, but claimed that A4 and A5 are indeed superfluous under the assumption of an additional axiom (PEP, pointwise equality for positiveness), which Magari probably assumed silently.
-Moreover, Hájek \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
+Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
 \citet[footnote 1 in p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics).
@@
 } 
-of A4 and A5, as well as varying domain semantics, in his new emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf.
-Appendix \ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler axiom H:A12.
-Surprisingly, the computer-assisted formalization of \AOEH shows that A4 and A5 are superfluous (not needed for proving T3).
-Moreover, A4 and A5 are independent\footnote{
+of A4 and A5, as well as varying domain semantics, in his new
+emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
+\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
+axiom H:A12. Surprisingly, the computer-assisted formalization of
+\AOEH\ shows that A4 and A5 are still superfluous (not needed for proving T3). Moreover, A4 and A5 are independent\footnote{
   An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
@@
 to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive \cite[sec.~5]{Hajek2002}.
-Formalization of \AOEHH shows that A4 is not only superflous, but also redundant.
+Formalization of \AOEHH\ shows that A4 is not only superflous, but also redundant.
 For A5, no conclusive results were achieved; neither a proof nor a countermodel could be automatically generated.
 Surprisingly, a countermodel for the weaker A3 was succesfully generated.
-This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH does not entail the positiveness of being God-like.
+This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH\ does not entail the positiveness of being God-like.
 
-Nevertheless, \AOEHH is explicitly regarded by \citet[p.~12]{Hajek2002} (page 12) as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
+Nevertheless, \AOEHH\ is explicitly regarded by \citet[p.~12]{Hajek2002} (page 12) as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
 That is his final emendation (\AOEHHH, cf.
@@
 H:A4 and H:A5).
-The formalization of \AOEHHH shows that H:A4 is independent.
+The formalization of \AOEHHH\ shows that H:A4 is independent.
 The old A5 is independent as well, and both H:A4 and H:A5 are superfluous, but no conclusive results were achieved regarding independence or redundancy of H:A5.
@@
 
+\sloppy 
 \printbibliography

### 2014-10-26 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Minor cosmetic changes

@@
 
-Nevertheless, \citet[p.~7]{Hajek2002} (in page 7) acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
+Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
   A4 and A5 are used by \citet[p.~11]{Hajek2002} (page 11) in the proofs of, respectively, Lemma 4 and Theorem 4.
@@
 
-Nevertheless, \AOEHH\ is explicitly regarded by \citet[p.~12]{Hajek2002} (page 12) as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
+Nevertheless, \AOEHH\ is explicitly regarded by \citet[p.~12]{Hajek2002} as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
 That is his final emendation (\AOEHHH, cf.

### 2014-10-27 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Fix minor errors

@@
 Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no claim is made w.r.t.
-to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive \cite[sec.~5]{Hajek2002}.
+to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive when it comes to God-like beings \cite[sec.~5]{Hajek2002}.
 Formalization of \AOEHH\ shows that A4 is not only superflous, but also redundant.
@@
 Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant).
-Unfortunately, T3 then becomes unprovable.
+Unfortunately then, a countermodel for T3 can be found.
 
@@
 already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable.
-These claims have been confirmed by the automated theorem provers.
+These claims have been confirmed by the automated theorem provers (in logic \KFourB).
 Within the same trend, the alternative proposed by \citet{bjordal99} (cf.
@@
 He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being.
-He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D, A2, A3 and A4.
+He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1, A2, A3 and A4.
 This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K.
@@
 assuming reflexivity of the accessibility relation).
-Proving D1 is possible in logic \KFour.
+For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive.
 Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K.
@@
 Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous.
-All these results hold for both constant and varying domain semantics.
+All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.
 

### 2014-10-29 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Add formalizations of Hajek's three emendations

@@
 \begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
   $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
+  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
 \item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+  $$\hol{P(G_H)}$$
 \item[A4]  Positive properties are necessarily positive: 
@@
   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
 \item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
+  \exq y [\varphi(y)]]}$$
 \item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+
+\item[L3] 
+  \begin{itemize}
+  \item[(1)] The negation of a positive property is not positive:
+$$\hol{\allq \varphi[P(\varphi) \imp \neg P(\neg \varphi)] }$$
+  \item[(2)] Positive properties are possibly exemplified:
+$$\hol{\allq \varphi[P(\varphi) \imp \pos \exq x \varphi(x)] }$$
+  \item[(3)] If a god-like being exists, then necessarily a god-like being exists:
+$$\hol{\allq x[G_H(x) \imp \nec G_H(x)))] }$$  
+  \item[(4)] All positive properties are necessarily implied by the property of being god-like:
+$$\hol{\allq \varphi[P(\varphi) \imp \nec \allq x[G_H(x) \imp \varphi(x)]] }$$
+  \end{itemize}
+\item[L4]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_H(x) \imp \ess{G_H}{x}]}$$
+
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
 % 
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}
@@
 \section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}
-
 % \begin{figure}[t]
@@
 \begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
   $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
 \item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
   $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+\item[H:A3]  The property of being God-like and existing actually is positive: 
+  $$\hol{P(G_A \wedge E)}$$
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
 % 
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}
@@
 \section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}
-
 % \begin{figure}[t]
@@
 \begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
   $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
+  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
+\item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property: 
+  $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
 \item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
+  $$\hol{P(G_H)}$$
+\item[H:A4]  Positive\textsuperscript{\#} properties are necessarily positive\textsuperscript{\#}: 
+  $$\hol{\allq \varphi [P^\#(\varphi) \imp \Box \; P^\#(\varphi)]}$$ 
 \item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
 \item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+  \exq y [\varphi(y)]]}$$
+\item[H:A5] Necessary existence is a positive\textsuperscript{\#} property: $$\hol{P^\#(\NE_A)}$$ 
 % 
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
 \end{itemize}

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Make cosmetic changes

@@
 %
-a trivial countermodel (consisting of two worlds and 
+a trivial countermodel (with two worlds and 
 two individuals) for MC is generated by Nitpick (for all mentioned logics); all axioms and definitions are shown to be mutually consistent.
@@
 }, arguing that T3 is true in all models of the other axioms and definitions by Gödel.
-\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, refuted Magari's claim, but claimed that A4 and A5 are indeed superfluous under the assumption of an additional axiom (PEP, pointwise equality for positiveness), which Magari probably assumed silently.
+\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, and claimed that Magari's claim is not valid, but is nevertheless true  under additional silent assumptions by Magari.
 Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
@@
 Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
-  A4 and A5 are used by \citet[p.~11]{Hajek2002} (page 11) in the proofs of, respectively, Lemma 4 and Theorem 4.
+  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, respectively, Lemma 4 and Theorem 4.
 } 
@@
 axiom H:A12. Surprisingly, the computer-assisted formalization of
-\AOEH\ shows that A4 and A5 are still superfluous (not needed for proving T3). Moreover, A4 and A5 are independent\footnote{
+\AOEH\ shows that A4 and A5 are still superfluous. Moreover, A4 and A5 are independent\footnote{
   An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
@@
 
-Additionally, Anderson's 14th footnote \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers.
+Additionally, Anderson \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers.
 Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant).
-Unfortunately then, a countermodel for T3 can be found.
+Unfortunately, a countermodel for T3 can then be found.
 
@@
 The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB.
-Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with certain additional presuppositions, to eliminate A5 from his theory as well.
-Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous.
+Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with (unclear) additional modifications of the definitions, to eliminate A5 from his theory as well.
+Without any additional modification, the automated reasoners show that A5 is actually not superfluous.
 All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.
@@
 
-\section{Scott's version of G\"odel's ontological argment} \label{apx:Goedel} 
+\section{Scott's version of G\"odel's ontological argument} \label{apx:Goedel}
+
+\framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
 \begin{itemize}
@@
 \end{itemize}
+\end{minipage}
+} %\vskip-.5em
+
+\clearpage
 
@@
 
-% \begin{figure}[t]
-% \noindent \framebox[\columnwidth][r]{
-% \begin{minipage}{.94\columnwidth}\small
+\framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
 \begin{itemize}
@@
 \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-% \end{figure}
+\end{minipage}
+} %\vskip-.5em
+
+\clearpage
 
@@
 \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-% \end{figure}
+\end{minipage}
+} %\vskip-.5em
+
+\clearpage
 
 \section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}
-% \begin{figure}[t]
-% \noindent \framebox[\columnwidth][r]{
-% \begin{minipage}{.94\columnwidth}\small
+
+\framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
 \begin{itemize}
@@
 \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-% \end{figure}
+\end{minipage}
+} %\vskip-.5em
 
 \section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}
-% \begin{figure}[t]
-% \noindent \framebox[\columnwidth][r]{
-% \begin{minipage}{.94\columnwidth}\small
+
+\framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
 \begin{itemize}
@@
 \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
-% \end{figure}
+\end{minipage}
+} %\vskip-.5em
 
-%\clearpage
+\clearpage
 
@@
 
-% \begin{figure}[t]
-% \noindent \framebox[\columnwidth][r]{
-% \begin{minipage}{.94\columnwidth}\small
-In Bjordal's emendation $G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).
+\framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+$G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).
+
+\medskip
+
 \begin{itemize}
@@
 \end{itemize}
-% \end{minipage}
-% } \vskip-.5em
-% \caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
-% \end{figure}
+\end{minipage}
+} %\vskip-.5em
 
-%\clearpage
+\vskip 8em
+
 \end{appendix}

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Improve Bjordal's appendix

@@
 
+\usepackage{amsmath}
 \usepackage{xspace}
@@
 He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being.
-He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1, A2, A3 and A4.
-This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K.
-A4 can be proved in logic \KT (i.e.
+He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1', A2', A3' and A4'.
+This has been confirmed in the computer-assisted formalization: A2' and A3' can be quickly automatically derived in logic \K.
+A4' can be proved in logic \KT (i.e.
 assuming reflexivity of the accessibility relation).
 For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive.
-Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K.
-The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB.
-Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with (unclear) additional modifications of the definitions, to eliminate A5 from his theory as well.
-Without any additional modification, the automated reasoners show that A5 is actually not superfluous.
-All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.
+Conversely, the proof that B:D1 is entailed by D1', A2', A3' and A4' is possible already in logic \K.
+The provers also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and A5' already in logic \KB.
+Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5' and claims that it is possible, with (unclear) additional modifications of the definitions, to eliminate A5' from his theory as well.
+Without any additional modification, the automated reasoners show that A5' is actually not superfluous.
+All these results, with the exception of the aforementioned countermodel for D1', hold for both constant and varying domain semantics.
 
@@
 analysis of several variants of G\"odel's ontological argument has been surprisingly
-straightforward. The provers not only confirmed many claimed results, but also exposed a few mistakes and novel insights in a long standing controversy. 
+straightforward. The provers not only confirmed many claimed results, 
+but also exposed a few mistakes and novel insights in a long standing controversy. 
 We believe the technology employed in this work is ready to be 
@@
 \begin{minipage}{.94\columnwidth}\small
-$G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).
+$G_B$ (God-like) is taken as primitive and $P_B$ (Positive) is defined.
 
-\medskip
+\bigskip
 
 \begin{itemize}
-\item[B:D1] A formulas $\phi$ is positive iff it is necessarily the case
-  that anything which is God-like has the property $\phi$.
-  $$\hol{P(\phi) \equiv \nec \allq x (G(x) \imp \phi(x))}$$ 
-\item[B:L1] D is logically equivalent in S4 with the union of G\"odel's definition D1 and axioms A2, A3 and A4.
-  $$\hol{D \biimp D1 \wedge A2 \wedge A3 \wedge A4}$$
-  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
-\item[B:D2] $\phi$ is a maximal composite of object $x$'s positive properties iff $x$ has $\phi$ and $\phi$ is positive and all positive properties $\psi$ which $x$ has are such that is necessarily the case that all objects which have $\phi$ also have $\psi$.
-  $$\hol{MCP(\phi,x) \equiv (\phi(x) \wedge P(\phi)) \wedge \allq \psi ((\psi(x) \wedge P(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
-\item[B:D3] $x$ has the $N$-property iff x is such that if $\phi$ is a maximal composite of $x$'s positive properties then it is necessary that some object $y$ has the property $\phi$.
-  $$\hol{N(x) \equiv \allq \phi (MCP(\phi,x) \imp \nec \allq y \phi(y))}$$
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A5'] The $N$-property is positive.
- $$\hol{P(N)}$$
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
+\item[B:D1] A property is positive iff it is necessarily possessed by every God-like being.
+  $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$ 
+\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
+  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
+%  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
+\item[B:D2] a \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
+  $$\hol{\MCP(\phi,x) \equiv (\phi(x) \wedge P_B(\phi)) \wedge \allq \psi ((\psi(x) \wedge P_B(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
+\item[B:D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
+  $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
+\item[A:A1'] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$ 
+\item[A5'] Necessary existence is a positive property.
+ $$\hol{P_B(\NE_B)}$$
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$ 
 \end{itemize}

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Wrap paragraphs

@@
 
-For all emendations and variants discussed here, the axioms and definitions have been shown to be consistent and 
-not to entail modal collapse.
+For all emendations and variants discussed here, the axioms and
+definitions have been shown to be consistent and  not to entail modal
+collapse.
 
-For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold for \emph{Anderson's Emendation} (cf. Appendix \ref{apx:Anderson}): 
-%
-the theorems T1, C and T3' can be quickly automated (in logics \K, \K and \KB, respectively);
-%
-the axioms A4 and A5 are proven redundant\footnote{
-  An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is entailed by $S$. 
-} (the former in
-logic \KFourB and the latter already in \K);
-%
-a trivial countermodel (with two worlds and 
-two individuals) for MC is generated by Nitpick (for all mentioned logics); all axioms and definitions are shown to be mutually consistent.
+For both \textbf{constant domain semantics} and \textbf{varying domain
+semantics}, the following results hold for \emph{Anderson's
+Emendation} (cf. Appendix \ref{apx:Anderson}):  % the theorems T1, C
+and T3' can be quickly automated (in logics \K, \K and \KB,
+respectively); % the axioms A4 and A5 are proven redundant\footnote{
+An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
+entailed by $S$.  } (the former in logic \KFourB and the latter
+already in \K); % a trivial countermodel (with two worlds and  two
+individuals) for MC is generated by Nitpick (for all mentioned
+logics); all axioms and definitions are shown to be mutually
+consistent.
 
-The redundancy of A4 and A5 is particularly controversial. Magari \cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
-  An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is entailed by $S \setminus \{ A \}$.
-}, arguing that T3 is true in all models of the other axioms and definitions by Gödel.
-\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, and claimed that Magari's claim is not valid, but is nevertheless true  under additional silent assumptions by Magari.
-Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant.
-\citet[footnote 1 in p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics).
-Our results show that Hájek was originally right, under both constant and varying domain semantics.
+The redundancy of A4 and A5 is particularly controversial. Magari
+\cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
+An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is
+entailed by $S \setminus \{ A \}$. }, arguing that T3 is true in all
+models of the other axioms and definitions by Gödel.
+\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this
+further, and claimed that Magari's claim is not valid, but is
+nevertheless true  under additional silent assumptions by Magari.
+Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier
+work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes
+\citep{Hajek_Magari_and_others_1996} in writing, it was published only
+5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where
+he claims (in Theorem 5.3) that for Anderson's emended theory
+\citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are
+not only superfluous, but also redundant. \citet[footnote 1 in
+p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim,
+arguing that the redundancy of A4 and A5 holds only under constant
+domain semantics, while Anderson's emended theory ought to be taken
+under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain
+semantics). Our results show that Hájek was originally right, under
+both constant and varying domain semantics.
 
-Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
+Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, 
+and apparently accepts it, as evidenced by his use\footnote{
   A4 and A5 are used by \citet[p.~11]{Hajek2002} in, respectively, Lemma 4 and Theorem 4.
@@
 axiom H:A12. Surprisingly, the computer-assisted formalization of
-\AOEH\ shows that A4 and A5 are still superfluous. Moreover, A4 and A5 are independent\footnote{
-  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
-} of the other axioms and definitions. Therefore, A4 and A5 are not redundant, despite their superfluousness.
+\AOEH\ shows that A4 and A5 are still superfluous. Moreover, 
+A4 and A5 are independent\footnote{
+  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff 
+  there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
+} 
+of the other axioms and definitions. 
+Therefore, A4 and A5 are not redundant, despite their superfluousness.
 
-Although, Hájek did not notice the superfluousness of A4 and A5 in his \AOEH, he did describe yet another emendation (his \AOEHH, cf.
-Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no claim is made w.r.t.
-to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive when it comes to God-like beings \cite[sec.~5]{Hajek2002}.
-Formalization of \AOEHH\ shows that A4 is not only superflous, but also redundant.
-For A5, no conclusive results were achieved; neither a proof nor a countermodel could be automatically generated.
-Surprisingly, a countermodel for the weaker A3 was succesfully generated.
-This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH\ does not entail the positiveness of being God-like.
+Although, Hájek did not notice the superfluousness of A4 and A5 in his
+\AOEH, he did describe yet another emendation (his \AOEHH, cf.
+Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no
+claim is made w.r.t. to their redundancy), if A3 is replaced by a
+stronger axiom (H:A3) additionally stating that the property of actual
+existence is positive when it comes to God-like beings
+\cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
+not only superflous, but also redundant. For A5, no conclusive results
+were achieved; neither a proof nor a countermodel could be
+automatically generated. Surprisingly, a countermodel for the weaker
+A3 was succesfully generated. This is somewhat unsatisfactory (for
+theistic goals), because it shows that \AOEHH\ does not entail the
+positiveness of being God-like.
 
-Nevertheless, \AOEHH\ is explicitly regarded by \citet[p.~12]{Hajek2002} as just an intermediary step towards a more natural theory, based on a more sophisticated notion of positiveness.
-That is his final emendation (\AOEHHH, cf.
-\ref{apx:Hajek3}), which restores A3 and does use A4 and A5, albeit in a modified form (i.e.
-H:A4 and H:A5).
-The formalization of \AOEHHH\ shows that H:A4 is independent.
-The old A5 is independent as well, and both H:A4 and H:A5 are superfluous, but no conclusive results were achieved regarding independence or redundancy of H:A5.
+Nevertheless, \AOEHH\ is explicitly regarded by
+\citet[p.~12]{Hajek2002} as just an intermediary step towards a more
+natural theory, based on a more sophisticated notion of positiveness.
+That is his final emendation (\AOEHHH, cf. \ref{apx:Hajek3}), which
+restores A3 and does use A4 and A5, albeit in a modified form (i.e.
+H:A4 and H:A5). The formalization of \AOEHHH\ shows that H:A4 is
+independent. The old A5 is independent as well, and both H:A4 and H:A5
+are superfluous, but no conclusive results were achieved regarding
+independence or redundancy of H:A5.
 
-Additionally, Anderson \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers.
-Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant).
-Unfortunately, a countermodel for T3 can then be found.
+Additionally, Anderson \citep[footnote
+14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
+the quantifiers in T3' and in A:D2 need to be interpreted as
+actualistic quantifiers, while others may be taken as possibilistic
+quantifiers. Our computer-assisted study of this mixed variant shows
+that A4 is still redundant, but A5 becomes independent (hence not
+redundant). Unfortunately, a countermodel for T3 can then be found.
 
-The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
-already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable.
-These claims have been confirmed by the automated theorem provers (in logic \KFourB).
-Within the same trend, the alternative proposed by \citet{bjordal99} (cf.
-Appendix \ref{apx:Bjordal}) achieves a high level of minimality.
-He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being.
-He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1', A2', A3' and A4'.
-This has been confirmed in the computer-assisted formalization: A2' and A3' can be quickly automatically derived in logic \K.
-A4' can be proved in logic \KT (i.e.
-assuming reflexivity of the accessibility relation).
-For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive.
-Conversely, the proof that B:D1 is entailed by D1', A2', A3' and A4' is possible already in logic \K.
-The provers also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and A5' already in logic \KB.
-Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5' and claims that it is possible, with (unclear) additional modifications of the definitions, to eliminate A5' from his theory as well.
-Without any additional modification, the automated reasoners show that A5' is actually not superfluous.
-All these results, with the exception of the aforementioned countermodel for D1', hold for both constant and varying domain semantics.
+The controversy over the superfluousness of A4 and A5 indicates a
+trend to reduce the ontological argument to its bare essentials. In
+this regard,  already
+\citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates
+that, by taking a notion of \emph{defective} as primitive and defining
+the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
+derivable. These claims have been confirmed by the automated theorem
+provers (in logic \KFourB). Within the same trend, the alternative
+proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal})
+achieves a high level of minimality. He takes the property of being
+God-like as a primitive and defines (B:D1) the positive properties as
+those properties necessarily possessed by every God-like being. He
+then briefly indicates (B:L1) that B:D1 is logically equivalent, under
+modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
+has been confirmed in the computer-assisted formalization: A2' and A3'
+can be quickly automatically derived in logic \K. A4' can be proved in
+logic \KT (i.e. assuming reflexivity of the accessibility relation).
+For constant domain semantics, proving D1 is possible in logic \KFour,
+whereas for varying domain semantics, a counter model can be found
+even in logic \SFive. Conversely, the proof that B:D1 is entailed by
+D1', A2', A3' and A4' is possible already in logic \K. The provers
+also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
+A5' already in logic \KB. Bjordal's last paragraph briefly mentions
+Hájek's ideas about the superfluousness of A5' and claims that it is
+possible, with (unclear) additional modifications of the definitions,
+to eliminate A5' from his theory as well. Without any additional
+modification, the automated reasoners show that A5' is actually not
+superfluous. All these results, with the exception of the
+aforementioned countermodel for D1', hold for both constant and
+varying domain semantics.
 
-Using our approach, the formalization and (partly) automated
-analysis of several variants of G\"odel's ontological argument has been surprisingly
-straightforward. The provers not only confirmed many claimed results, 
-but also exposed a few mistakes and novel insights in a long standing controversy. 
-We believe the technology employed in this work is ready to be 
-fruitfully adopted in larger scale by philosophers.
+Using our approach, the formalization and (partly) automated analysis
+of several variants of G\"odel's ontological argument has been
+surprisingly straightforward. The provers not only confirmed many
+claimed results,  but also exposed a few mistakes and novel insights
+in a long standing controversy.  We believe the technology employed in
+this work is ready to be  fruitfully adopted in larger scale by
+philosophers.
 
@@
 % 5 - Add subscripts to alternative definitions (e.g. G_A)
-% 6 - Fix citet redundancy
-% 7 - Add apendix for Anderson's simplification
+
+% - Investigate Magari's proof
+% - Add apendix for Anderson's simplification
 

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Improve Hajek's appendix

@@
 
-%\section{Anderson's Emendation (cf. Appendix \ref{apx:Anderson})}
-
 For all emendations and variants discussed here, the axioms and
@@
 semantics}, the following results hold for \emph{Anderson's
-Emendation} (cf. Appendix \ref{apx:Anderson}):  % the theorems T1, C
+Emendation} (cf. Appendix \ref{apx:Anderson}): T1, C
 and T3' can be quickly automated (in logics \K, \K and \KB,
-respectively); % the axioms A4 and A5 are proven redundant\footnote{
+respectively); the axioms A4 and A5 are proven redundant\footnote{
 An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
 entailed by $S$.  } (the former in logic \KFourB and the latter
-already in \K); % a trivial countermodel (with two worlds and  two
+already in \K); a trivial countermodel (with two worlds and  two
 individuals) for MC is generated by Nitpick (for all mentioned
@@
 and apparently accepts it, as evidenced by his use\footnote{
-  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, respectively, Lemma 4 and Theorem 4.
+  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, 
+  respectively, Lemma 4 and Theorem 4.
 } 
@@
   An axiom $A$ is \emph{independent} of a set of axioms $S$ iff 
-  there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
+  there are models of $S$ where $A$ is true and other models of 
+  $S$ where $A$ if false. 
 } 
@@
   \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
-\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
-  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
 \item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property: 
   $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive\textsuperscript{\#}: 
+  $$\hol{G_H(x) \equiv \forall \varphi [P^\#(\varphi) \biimp \nec \varphi(x)]}$$ 
 \item[A3']  The property of being God-like is positive: 

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Add missing apostrophes

@@
 } 
-of A4 and A5, as well as varying domain semantics, in his new
+of A4 and A5', as well as varying domain semantics, in his new
 emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
@@
 axiom H:A12. Surprisingly, the computer-assisted formalization of
-\AOEH\ shows that A4 and A5 are still superfluous. Moreover, 
-A4 and A5 are independent\footnote{
+\AOEH\ shows that A4 and A5' are still superfluous. Moreover, 
+A4 and A5' are independent\footnote{
   An axiom $A$ is \emph{independent} of a set of axioms $S$ iff 
@@
 of the other axioms and definitions. 
-Therefore, A4 and A5 are not redundant, despite their superfluousness.
+Therefore, A4 and A5' are not redundant, despite their superfluousness.
 
-Although, Hájek did not notice the superfluousness of A4 and A5 in his
+Although Hájek did not notice the superfluousness of A4 and A5' in his
 \AOEH, he did describe yet another emendation (his \AOEHH, cf.
-Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no
-claim is made w.r.t. to their redundancy), if A3 is replaced by a
+Appendix \ref{apx:Hajek2}) where A4 and A5' are superfluous (though no
+claim is made w.r.t. to their redundancy), if A3' is replaced by a
 stronger axiom (H:A3) additionally stating that the property of actual
@@
 \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
-not only superflous, but also redundant. For A5, no conclusive results
+not only superflous, but also redundant. For A5', no conclusive results
 were achieved; neither a proof nor a countermodel could be
 automatically generated. Surprisingly, a countermodel for the weaker
-A3 was succesfully generated. This is somewhat unsatisfactory (for
+A3' was succesfully generated. This is somewhat unsatisfactory (for
 theistic goals), because it shows that \AOEHH\ does not entail the
@@
 That is his final emendation (\AOEHHH, cf. \ref{apx:Hajek3}), which
-restores A3 and does use A4 and A5, albeit in a modified form (i.e.
+restores A3' and does use A4 and A5', albeit in a modified form (i.e.
 H:A4 and H:A5). The formalization of \AOEHHH\ shows that H:A4 is
-independent. The old A5 is independent as well, and both H:A4 and H:A5
+independent. The old A5' is independent as well, and both H:A4 and H:A5
 are superfluous, but no conclusive results were achieved regarding
@@
 quantifiers. Our computer-assisted study of this mixed variant shows
-that A4 is still redundant, but A5 becomes independent (hence not
+that A4 is still redundant, but A5' becomes independent (hence not
 redundant). Unfortunately, a countermodel for T3 can then be found.
@@
 % ToDo: \\
-% 3 - Fix Hajek's proofs in the appendices. (they are just copy-pasted from Anderson's so far)
-% 4 - Check apostrophes in axiom names that depend on different definitions.
-% 5 - Add subscripts to alternative definitions (e.g. G_A)
 

### 2014-12-08 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Add abstract and introduction

--- 2014-01-31 Papers/ITP/GodProof-ITP.tex
@@
 \documentclass{llncs}
+\pagestyle{plain}
 
-\usepackage{url,amsmath,amssymb,a4wide,fancyvrb,color}
+\usepackage{fancybox} % Must be before "fancyvrb"
+\usepackage[utf8]{inputenc}
+\usepackage{url,amsmath,amssymb,fancyvrb,color}
+\usepackage{proof}
+\usepackage{xspace}
+
+\usepackage{calculi}
 
 \newcommand{\imp}{\rightarrow}
+\newcommand{\biimp}{\leftrightarrow}
+\newcommand{\all}{\forall}
+\newcommand{\ex}{\exists}
+\newcommand{\seq}{\vdash}
+\newcommand{\nec}{\Box} % necessarily
+\newcommand{\pos}{\Diamond} % possibly
 
@@
 
+\newcommand{\Coq}{\texttt{Coq}\xspace}
+\newcommand{\Isabelle}{\texttt{Isabelle}\xspace}
+
 \title{
-God in Coq: \\
-Modal Logic in Coq \\ 
-and Scott's Version of G\"{o}del's Proof of God's Existence \\
-{\small (\emph{Rough Diamond Proof Pearl})}
-%   \thanks{
-%   This work has been supported by the 
-%   German Research Foundation under grant
-%   BE2501/9-1.}
+Modal Logics in the Coq Proof Assistant
 }
@@
 
+\begin{abstract} 
+This paper describes an embedding of higher-order modal
+logics in the \Coq proof assistant. {\Coq}'s capabilities are thus extended in a
+minimalistic manner, which is nevertheless sufficient for the formalization
+of significant modal proofs. The elegance and convenience of this approach, 
+from a user perspective, are illustrated here with the successful formalization of Gödel's ontological argument. 
+\end{abstract}
+
 \section{Introduction}
 
-G\"{o}del's proof of God's existence is challenging to formalize and
-verify because it requires an expressive logical language with modal
-operators ($\Diamond$ and $\Box$) and higher-order quantifiers.
-Specialized modal logic tools lack higher-order capabilities, while
-higher-order provers lack built-in support for modalities.  However,
-by embedding modal logics into higher-order logic, it is in principle
-possible to use standard higher-order automated and interactive
-theorem provers for reasoning about and within modal logics
-\cite{J23,B9}.  To realize this potential, formalizations
-\cite{FormalTheologyRepository} of G\"odel's proof in TPTP THF
-\cite{J22}, Coq \cite{Coq} and Isabelle \cite{Isabelle} have been
-completed so far, in this order. This paper focuses on the
-formalization in Coq. The implementation of the embedding is discussed
-in detail in Section \ref{sec:modal} and Scott's version of G\"odel's
-proof is shown in Section \ref{sec:proof}.
+Modal logics extend the usual formal logic language with modal operators ($\Box$ and $\Diamond$) and are characterized by the \emph{necessitation rule}, according to which $\Box A$ is a theorem if $A$ is a theorem, although $A \imp \Box A$ is not necessarily a theorem.
+Various informal concepts, such as \emph{necessity and possibility}, \emph{knowledge and belief}, \emph and \emph{temporal globality and eventuality}, have been formalized with the help of modal operators.
+Therefore, modal logics \cite{ModalLogic} are well-established for a wide variety of application domains, ranging from philosophy \cite{ToDo} to formal software verification \cite{ToDo:TLAPlus}.
+
+However, general automated reasoning support for modal logics is still not as well-developed as for classical logics.
+Current deduction tools for modal logics are often limited to propositional, quantifier-free, fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular modal logics and their applications \cite{ToDo:StephanMerz}.
+Recently, first-order automated deduction techniques based on tableaux, sequent calculi and connection calculi have been generalized to cope with modalities directly and a few new provers have been implemented \cite{JensOtten}.
+
+Another recently explored possibility \cite{J23,B9,J22} is the embedding of first-order and higher-order modal logics into classical higher-order logics, for which existing higher-order automated theorem provers \cite{ToDo:Leo,Satallax} exist.
+Embedding is flexible, because various modal logics can be easily supported by adding their characteristic axioms.
+Furthermore, it is relatively simple to implement, because it does not require any modification in the source code of the higher-order prover.
+The prover can be used as is, and only the input files provided to the prover must be especially encoded.
+Furthermore, the efficacy and efficiency of the embedding approach has been confirmed in philosophically interesting benchmarks \cite{FormalTheologyRepository}.
+These qualities make embedding a convenient approach for \emph{fully automated} reasoning.
+
+However, one may wonder whether the embedding approach is adequate for \emph{interactive} reasoning, when the user proves theorems by interacting with a proof assistant.
+The main goal of this paper is to answer this question positively.
+The \Coq proof assistant was chosen because of the authors' greater familiarity with the tactic language of this system.
+
+One major initial concern was whether the embedding could be a disturbance to the user.
+Fortunately, by using \Coq's \texttt{Ltac} tactic language, we were able to  define intuitive new tactics that hide the technical details of the embedding from the user.
+The resulting infra-structure for modal reasoning within \Coq (as described in Section \ref{sec:modal}) provides a user experience where modalities can be handled transparently and straightforwardly.
+Therefore, a user with basic knowledge of modal logics and \Coq's tactics should be able to use (and extend) our implementation with no excessive overhead.
+
+In order to illustrate the use of the implemented embedding, we show here the formalization of Scott's version \cite{Scott} of Gödel's ontological argument for God's existence (in Section \ref{sec:proof}).
+This proof was chosen mainly for two reasons.
+Firstly, it requires not only modal operators, but also higher-order quantification.
+Therefore, it is beyond reach for specialized propositional and first-order theorem provers.
+Secondly, this argument addresses an ancient problem in Philosophy and Metaphysics, which has nevertheless received a lot of attention in the last 15 years, because of the discovery of the modal collapse \cite{ToDo}.
+This proof lies in the center of a vast and largely unexplored application domain for automated and interactive theorem provers.
+
+The simpler ontological argument of Anselm has been automatically verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-order theorem by Oppenheimer and Zalta \cite{Zalta}.
+Gödel's argument was automatically verified in our previous work on embedding-based fully automated modal theorem proving \cite{ArXiv,AFP}.
+But this paper presents the first fully interactive and detailed formalization of this proof in a proof assistant.
 
@@
 To avoid this issue, the \emph{possible world semantics} of modal logics can be explicitly embedded into higher-order logics \cite{J23,B9}.
+
 The embedding is related to labeling techniques
@@
 exploited here to encode the labels within the logic language itself.
-To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type that depends on possible worlds.
-Possible worlds are connected by an accessibility relation.
+To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type \texttt{o} that depends on possible worlds:
 
@@
 \Definition o := i -> \Prop. \com{(* Type of modal propositions *)}
+\end{Verbatim}
+\end{small}
+
+\noindent
+Possible worlds are connected by an accessibility relation, 
+which can be represented in \Coq by a parameter \texttt{r}, as follows:
+
+\begin{small}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Parameter r: i -> i -> \Prop. \com{(* Accessibility relation for worlds *)}
@@
 
+\newcommand{\s}{\qquad}
+
+\begin{calculus}
+{propositional rules}
+{fig:PropositionalRules}
+
+\vspace{1em}
+
+\s\s
+\infer[\bot_E]{A}{ \bot }
+\s
+\infer[\imp_I]{A \imp B}{ B }
+\s
+\infer[\imp_I^n]{A \imp B}{ \infer*{B}{\infer[n]{A}{}} }
+\s
+\infer[\imp_E]{B}{A & A \imp B}
+
+\vspace{2em}
+
+\s\s
+\infer[\wedge_I]{A \wedge B}{A & B}
+\s\s
+\infer[\wedge_{E_1}]{A}{A \wedge B}
+\s\s
+\infer[\wedge_{E_2}]{B}{A \wedge B}
+
+% \vspace{2em}
+
+% \s\s
+% \infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
+% \s\s
+% \infer[\vee_{I_1}]{A \vee B}{A}
+% \s\s
+% \infer[\vee_{I_2}]{A \vee B}{B}
+
+\vspace{1em}
+\end{calculus}
+
+\begin{calculus}
+{double negation elimination}
+{fig:Classical}
+\infer[\neg\neg_E]{A}{ \neg\neg A }
+\end{calculus}
+
+\begin{calculus}
+{quantifier rules}
+{fig:QuantifierRules}
+
+\vspace{1em}
+
+\s
+\infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] }
+\s
+\infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }
+\s\s
+\infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
+\s
+\infer[\ex_E]{A[\beta]}{ \ex x_{\tau}. A[x] }
+
+\vspace{1em}
+
+\begin{center}
+\textbf{eigen-variable conditions:} \\
+if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$; \\
+if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+\end{center}
+
+\vspace{1em}
+
+\end{calculus}
+
+\begin{calculus}
+{Rules for Modal Operators}
+{fig:NDK}
+
+\vspace{1em}
+
+\s\s\s\s
+\infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
+\s\s\s\s
+\infer[\nec_E]{w: \fbox{ \infer*{}{A} } }{\nec A}
+
+\vspace{2em}
+
+\s\s\s\s
+\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
+\s\s\s\s\s
+\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
+
+\vspace{1em}
+
+\begin{center}
+\textbf{eigen-box condition:}\\ 
+$\omega$ must be a fresh name for a box. \\
+\vspace{0.5em}
+\textbf{boxed assumption condition:} \\
+assumptions should be discharged within the box where they are made.
+\end{center}
+
+\vspace{1em}
+
+\end{calculus}
+
 \noindent
@@
 \Ltac box_elim H w1 H1 := \match type of H \with 
-  ((box ?p) ?w) => cut (p w1); [intros H1 | (apply (H w1); try assumption)] \eend.
+      ((box ?p) ?w) => cut (p w1); 
+                       [intros H1 | (apply (H w1); try assumption)] \eend.
 
@@
 
-\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" 
-                 \iin (destruct H \as [w [R H]]; move w at top; move R at top).
+\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" \iin 
+                (destruct H \as [w [R H]]; move w at top; move R at top).
 
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Lemma mp_dia: [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
-\Proof. mv. intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. \Qed.
+\Lemma mp_dia: 
+  [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
+\Proof. mv. 
+intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. 
+\Qed.
 
 \Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
-\Proof. mv. intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. \Qed.
+\Proof. mv. 
+intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. 
+\Qed.
 \end{Verbatim}
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Theorem K: [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
-\Proof. mv. intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. \Qed.
+\Theorem K: 
+  [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
+\Proof. mv. 
+intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. 
+\Qed.
 
 \Theorem T: [ mforall p, (box p) m-> p ].
-\Proof. mv. intro p. intro H. box_e H H1. exact H1. apply reflexivity. \Qed.
+\Proof. mv. 
+intro p. intro H. box_e H H1. exact H1. apply reflexivity. 
+\Qed.
 \end{Verbatim}
@@
 \Theorem dia_box_to_box: [ mforall p, (dia (box p)) m-> (box p) ].
-\Proof. mv. intro p. intro H1. dia_e H1. box_i. box_e H1 H2. exact H2. eapply transitivity.
+\Proof. mv. 
+intros p H1. dia_e H1. box_i. box_e H1 H2. exact H2. eapply transitivity. 
   apply symmetry. exact R. 
@@
 Scott's assertions are emphasized with comments.
-In contrast to the formalization in Isabelle \cite{Archiv}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
+In contrast to the formalization in Isabelle \cite{AFP}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
 This was a deliberate choice, mainly because it allowed a qualitative evaluation of the convenience of the possible world semantic embedding approach for \emph{interactive} theorem proving.
@@
 
-\com{(* Axiom A1: either a property or its negation is positive, but not both *)}
-\Axiom axiom1a : [ mforall p, (Positive (\fun x: u => m~(p x))) m-> (m~ (Positive p)) ].
-\Axiom axiom1b : [ mforall p, (m~ (Positive p)) m-> (Positive (\fun x: u => m~ (p x))) ].
+\com{(* Axiom A1:} 
+\com{   either a property or its negation is positive, but not both *)}
+\Axiom axiom1a : 
+  [ mforall p, (Positive (\fun x: u => m~(p x))) m-> (m~ (Positive p)) ].
 
-\com{(* Axiom A2: a property necessarily implied by a positive property is positive *)}
+\Axiom axiom1b : 
+  [ mforall p, (m~ (Positive p)) m-> (Positive (\fun x: u => m~ (p x))) ].
+
+\com{(* Axiom A2: }
+\com{   a property necessarily implied by a positive property is positive *)}
 \Axiom axiom2: [ mforall p, mforall q, 
@@
 
-\com{(* Definition D1: God: a God-like being possesses all positive properties *)}
+\com{(* Definition D1: }
+\com{   God: a God-like being possesses all positive properties *)}
 \Definition G(x: u) := mforall p, (Positive p) m-> (p x).
@@
 
-\com{(* Definition D2: essence: an essence of an individual is a property possessed by it} 
-\com{and necessarily implying any of its properties *)}
-\Definition Essence(p: u -> o)(x: u) := (p x) m/\bslash mforall q, 
-                                        ((q x) m-> box (mforall y, (p y) m-> (q y))).
+\com{(* Definition D2: }
+\com{   essence: an essence of an individual is a property possessed by it} 
+\com{   and necessarily implying any of its properties *)}
+\Definition Essence(p: u -> o)(x: u) := 
+    (p x) m/\bslash mforall q, ((q x) m-> box (mforall y, (p y) m-> (q y))).
 \Notation "p 'ess' x" := (Essence p x) (at level 69).
@@
   intro q. intro H2. assert (H3: ((Positive q) w)).
-    proof_by_contradiction H4. unfold G in H1. apply axiom1b in H4. apply H1 in H4. 
-                                                                    contradiction. 
+    proof_by_contradiction H4. unfold G in H1. apply axiom1b in H4. 
+    apply H1 in H4. contradiction.
+
     cut (box (Positive q) w). \com{(* Scott *)}
-      apply K. box_i. intro H5. intro y. intro H6. unfold G in H6. apply (H6 q). exact H5.
+      apply K. box_i. intro H5. intro y. intro H6. 
+      unfold G in H6. apply (H6 q). exact H5.
+
       apply axiom4. exact H3.
@@
 
-\com{(* Definition D3: necessary existence: necessary existence of an individual }
-\com{is the necessary exemplification of all its essences *)}
+\com{(* Definition D3: }
+\com{   necessary existence: necessary existence of an individual }
+\com{   is the necessary exemplification of all its essences *)}
 \Definition NE(x: u) := mforall p, (p ess x) m-> box (mexists y, (p y)).
@@
 \Lemma lemma1: [ (mexists z, (G z)) m-> box (mexists x, (G x)) ].
-\Proof. mv. intro H1. destruct H1 as [g H2]. cut ((G ess g) w). \com{(* Scott *)}
+\Proof. mv. 
+intro H1. destruct H1 as [g H2]. cut ((G ess g) w). \com{(* Scott *)}
   assert (H3: (NE g w)).       \com{(* Scott *)}
@@
 \Lemma lemma2: [ dia (mexists z, (G z)) m-> box (mexists x, (G x)) ].
-\Proof. mv. intro H. cut (dia (box (mexists x, G x)) w).  \com{(* Scott *)}
+\Proof. mv. 
+intro H. cut (dia (box (mexists x, G x)) w).  \com{(* Scott *)}
   apply dia_box_to_box.
@@
 user does not have to explicitly bother about worlds and their mutual
-reachability; the provided tactics for modalities do the job for him.
+reachability; the provided tactics for modalities do the job for him/her.
 Moreover, for subgoals that do not involve modalities, the user has all the
@@
 
+HOL Light, Isabelle
+
 The theological implications of the verified correctness of this proof
 certainly depend on a critical discussion of the underlying concepts,
-definitions and axioms, which is beyond the scope of this paper. What
-can be said though is that our work presents further evidence that
-theism is at least not (obviously) irrational.
-
-Related work includes formalizations of Anselm's argument by Rushby
-\cite{Rushby}, and by Oppenheimer and Zalta \cite{Zalta}.  Clearly,
-the application of theorem proving technology to the domain of
+definitions and axioms, which is beyond the scope of this paper. 
+%What
+%can be said though is that our work presents further evidence that
+%theism is at least not (obviously) irrational. 
+Clearly, the application of theorem proving technology to the domain of
 theoretical philosophy can --- as already pictured by Leibniz ---
@@
 
-\paragraph{Acknowledgements:} Thanks to Cedric Auger and Laurent Th\'ery, for their answers to our newbie questions about Ltac in the Coq-Club mailing-list.
+\paragraph{Acknowledgements:} we thanks Cedric Auger and Laurent Th\'ery, for their answers to our questions about Ltac in the Coq-Club mailing-list.
 
@@
 
-\bibitem{arXiv}
+\bibitem{AFP}
 C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
some small modifications wrt abstract and intro

@@
 minimalistic manner, which is nevertheless sufficient for the formalization
-of significant modal proofs. The elegance and convenience of this approach, 
+of significant modal proofs. The elegance, flexibility and convenience of this approach, 
 from a user perspective, are illustrated here with the successful formalization of Gödel's ontological argument. 
@@
 
-Another recently explored possibility \cite{J23,B9,J22} is the embedding of first-order and higher-order modal logics into classical higher-order logics, for which existing higher-order automated theorem provers \cite{ToDo:Leo,Satallax} exist.
-Embedding is flexible, because various modal logics can be easily supported by adding their characteristic axioms.
-Furthermore, it is relatively simple to implement, because it does not require any modification in the source code of the higher-order prover.
-The prover can be used as is, and only the input files provided to the prover must be especially encoded.
-Furthermore, the efficacy and efficiency of the embedding approach has been confirmed in philosophically interesting benchmarks \cite{FormalTheologyRepository}.
-These qualities make embedding a convenient approach for \emph{fully automated} reasoning.
+Another recently explored possibility \cite{J23,B9,J22} is the
+embedding of first-order and higher-order modal logics into classical
+higher-order logics, for which existing higher-order automated theorem
+provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
+because various modal logics can be easily supported by adding their
+characteristic axioms.  The embedding approach can also be easily
+adapted to support multiple modalities, and it is possible to replace
+constant domain quantifiers by varying or cumulative domain
+quantifiers.  Moreover, the approach is relatively simple to
+implement, because it does not require any modification in the source
+code of the higher-order prover. The prover can be used as is, and
+only the input files provided to the prover must be especially
+encoded. Furthermore, the efficacy and efficiency of the embedding
+approach has been confirmed in philosophically interesting and
+relevant benchmarks \cite{FormalTheologyRepository}. These qualities
+make embedding a convenient approach for \emph{fully automated}
+reasoning.
 
-However, one may wonder whether the embedding approach is adequate for \emph{interactive} reasoning, when the user proves theorems by interacting with a proof assistant.
-The main goal of this paper is to answer this question positively.
-The \Coq proof assistant was chosen because of the authors' greater familiarity with the tactic language of this system.
+However, one may wonder whether the embedding approach is adequate
+also for \emph{interactive} reasoning, when the user proves theorems
+by interacting with a proof assistant. The main goal of this paper is
+to strudy this question. Our answer is positive. The \Coq proof
+assistant was chosen because of the authors' greater familiarity with
+the tactic language of this system.
 
@@
 
-In order to illustrate the use of the implemented embedding, we show here the formalization of Scott's version \cite{Scott} of Gödel's ontological argument for God's existence (in Section \ref{sec:proof}).
-This proof was chosen mainly for two reasons.
-Firstly, it requires not only modal operators, but also higher-order quantification.
-Therefore, it is beyond reach for specialized propositional and first-order theorem provers.
+In order to illustrate the use of the implemented embedding, we show
+here the formalization of Scott's version \cite{Scott} of Gödel's
+ontological argument for God's existence (in Section
+\ref{sec:proof}). This proof was chosen mainly for two
+reasons. Firstly, it requires not only modal operators, but also
+higher-order quantification. Therefore, it is beyond reach of
+specialized propositional and first-order (modal) theorem provers.
 Secondly, this argument addresses an ancient problem in Philosophy and Metaphysics, which has nevertheless received a lot of attention in the last 15 years, because of the discovery of the modal collapse \cite{ToDo}.

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Wrap paragraphs

@@
 \begin{abstract}  
-This paper describes an embedding of higher-order modal
-logics in the \Coq proof assistant. {\Coq}'s capabilities are thus extended in a
-minimalistic manner, which is nevertheless sufficient for the formalization
-of significant modal proofs. The elegance, flexibility and convenience of this approach, 
-from a user perspective, are illustrated here with the successful formalization of Gödel's ontological argument. 
+This paper describes an embedding of higher-order
+modal logics in the \Coq proof assistant. {\Coq}'s capabilities are
+thus extended in a minimalistic manner, which is nevertheless
+sufficient for the formalization of significant modal proofs. The
+elegance, flexibility and convenience of this approach, from a user
+perspective, are illustrated here with the successful formalization of
+Gödel's ontological argument.  
 \end{abstract}
@@
 
-Modal logics extend the usual formal logic language with modal operators ($\Box$ and $\Diamond$) and are characterized by the \emph{necessitation rule}, according to which $\Box A$ is a theorem if $A$ is a theorem, although $A \imp \Box A$ is not necessarily a theorem.
-Various informal concepts, such as \emph{necessity and possibility}, \emph{knowledge and belief}, \emph and \emph{temporal globality and eventuality}, have been formalized with the help of modal operators.
-Therefore, modal logics \cite{ModalLogic} are well-established for a wide variety of application domains, ranging from philosophy \cite{ToDo} to formal software verification \cite{ToDo:TLAPlus}.
+Modal logics extend the usual formal logic language with modal
+operators ($\Box$ and $\Diamond$) and are characterized by the
+\emph{necessitation rule}, according to which $\Box A$ is a theorem if
+$A$ is a theorem, although $A \imp \Box A$ is not necessarily a
+theorem. Various informal concepts, such as \emph{necessity and
+possibility}, \emph{knowledge and belief}, \emph and \emph{temporal
+globality and eventuality}, have been formalized with the help of
+modal operators. Therefore, modal logics \cite{ModalLogic} are well-
+established for a wide variety of application domains, ranging from
+philosophy \cite{ToDo} to formal software verification
+\cite{ToDo:TLAPlus}.
 
-However, general automated reasoning support for modal logics is still not as well-developed as for classical logics.
-Current deduction tools for modal logics are often limited to propositional, quantifier-free, fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular modal logics and their applications \cite{ToDo:StephanMerz}.
-Recently, first-order automated deduction techniques based on tableaux, sequent calculi and connection calculi have been generalized to cope with modalities directly and a few new provers have been implemented \cite{JensOtten}.
+However, general automated reasoning support for modal logics is still
+not as well-developed as for classical logics. Current deduction tools
+for modal logics are often limited to propositional, quantifier-free,
+fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
+modal logics and their applications \cite{ToDo:StephanMerz}. Recently,
+first-order automated deduction techniques based on tableaux, sequent
+calculi and connection calculi have been generalized to cope with
+modalities directly and a few new provers have been implemented
+\cite{JensOtten}.
 
@@
 by interacting with a proof assistant. The main goal of this paper is
-to strudy this question. Our answer is positive. The \Coq proof
+to study this question. Our answer is positive. The \Coq proof
 assistant was chosen because of the authors' greater familiarity with
@@
 
-One major initial concern was whether the embedding could be a disturbance to the user.
-Fortunately, by using \Coq's \texttt{Ltac} tactic language, we were able to  define intuitive new tactics that hide the technical details of the embedding from the user.
-The resulting infra-structure for modal reasoning within \Coq (as described in Section \ref{sec:modal}) provides a user experience where modalities can be handled transparently and straightforwardly.
-Therefore, a user with basic knowledge of modal logics and \Coq's tactics should be able to use (and extend) our implementation with no excessive overhead.
+One major initial concern was whether the embedding could be a
+disturbance to the user. Fortunately, by using \Coq's \texttt{Ltac}
+tactic language, we were able to  define intuitive new tactics that
+hide the technical details of the embedding from the user. The
+resulting infra-structure for modal reasoning within \Coq (as
+described in Section \ref{sec:modal}) provides a user experience where
+modalities can be handled transparently and straightforwardly.
+Therefore, a user with basic knowledge of modal logics and \Coq's
+tactics should be able to use (and extend) our implementation with no
+excessive overhead.
 
@@
 here the formalization of Scott's version \cite{Scott} of Gödel's
-ontological argument for God's existence (in Section
-\ref{sec:proof}). This proof was chosen mainly for two
-reasons. Firstly, it requires not only modal operators, but also
-higher-order quantification. Therefore, it is beyond reach of
-specialized propositional and first-order (modal) theorem provers.
-Secondly, this argument addresses an ancient problem in Philosophy and Metaphysics, which has nevertheless received a lot of attention in the last 15 years, because of the discovery of the modal collapse \cite{ToDo}.
-This proof lies in the center of a vast and largely unexplored application domain for automated and interactive theorem provers.
+ontological argument for God's existence (in Section \ref{sec:proof}).
+This proof was chosen mainly for two reasons. Firstly, it requires not
+only modal operators, but also higher-order quantification. Therefore,
+it is beyond reach of specialized propositional and first-order
+(modal) theorem provers. Secondly, this argument addresses an ancient
+problem in Philosophy and Metaphysics, which has nevertheless received
+a lot of attention in the last 15 years, because of the discovery of
+the modal collapse \cite{ToDo}. This proof lies in the center of a
+vast and largely unexplored application domain for automated and
+interactive theorem provers.
 
-The simpler ontological argument of Anselm has been automatically verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-order theorem by Oppenheimer and Zalta \cite{Zalta}.
-Gödel's argument was automatically verified in our previous work on embedding-based fully automated modal theorem proving \cite{ArXiv,AFP}.
-But this paper presents the first fully interactive and detailed formalization of this proof in a proof assistant.
+The simpler ontological argument of Anselm has been automatically
+verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-
+order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
+argument was automatically verified in our previous work on embedding-
+based fully automated modal theorem proving \cite{ArXiv,AFP}. But this
+paper presents the first fully interactive and detailed formalization
+of this proof in a proof assistant.
 

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve introduction

@@
 However, general automated reasoning support for modal logics is still
-not as well-developed as for classical logics. Current deduction tools
+not as well-developed as for classical logics. Deduction tools
 for modal logics are often limited to propositional, quantifier-free,
 fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
-modal logics and their applications \cite{ToDo:StephanMerz}. Recently,
+modal logics and their applications \cite{ToDo:StephanMerz}; 
 first-order automated deduction techniques based on tableaux, sequent
-calculi and connection calculi have been generalized to cope with
-modalities directly and a few new provers have been implemented
-\cite{JensOtten}.
+calculi and connection calculi have only recently been generalized 
+and implemented in a few new provers able to directly cope with
+modalities \cite{JensOtten}.
 
 Another recently explored possibility \cite{J23,B9,J22} is the
-embedding of first-order and higher-order modal logics into classical
+embedding of first-order and even higher-order modal logics into classical
 higher-order logics, for which existing higher-order automated theorem
@@
 also for \emph{interactive} reasoning, when the user proves theorems
-by interacting with a proof assistant. The main goal of this paper is
-to study this question. Our answer is positive. The \Coq proof
-assistant was chosen because of the authors' greater familiarity with
-the tactic language of this system.
+by interacting with a proof assistant such as \Coq\footnote{The \Coq
+proof assistant was chosen merely because of the authors' greater
+familiarity with the tactic language of this system. Nevertheless, the
+techniques presented and discussed here are likely to be useful for
+other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
+\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal of this
+paper is to study this question. Our answer is positive.
 
@@
 only modal operators, but also higher-order quantification. Therefore,
-it is beyond reach of specialized propositional and first-order
+it is beyond the reach of specialized propositional and first-order
 (modal) theorem provers. Secondly, this argument addresses an ancient

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section 2 (The Embedding in Coq)

@@
 
-\section{Modal Logic in Coq}
+\section{The Embedding of Modal Logics in Coq}
 \label{sec:modal}
 
-A crucial aspect of modal logics \cite{ModalLogic} is that the so-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$ is a theorem, but $A \imp \Box A$ is not necessarily a theorem.
-Naive attempts to define the modal operators $\Box$ and $\Diamond$ may easily be unsound in this respect.
-To avoid this issue, the \emph{possible world semantics} of modal logics can be explicitly embedded into higher-order logics \cite{J23,B9}.
+A crucial aspect of modal logics \cite{ModalLogic} is that the so-
+called \emph{necessitation rule} allows $\Box A$ to be derived if $A$
+is a theorem, but $A \imp \Box A$ is not necessarily a theorem. Naive
+attempts to define the modal operators $\Box$ and $\Diamond$ may
+easily be unsound in this respect. To avoid this issue, the
+\emph{possible world semantics} of modal logics can be explicitly
+embedded into higher-order logics \cite{J23,B9}.
 
-The embedding is related to labeling techniques
-\cite{Labels}. However, the expressiveness of higher-order logic is
-exploited here to encode the labels within the logic language itself.
-To this aim, a type for worlds must be declared and modal propositions should be not of type \texttt{Prop} but of a lifted type \texttt{o} that depends on possible worlds:
+The embedding is related to labeling techniques \cite{Labels}.
+However, the expressiveness of higher-order logic is exploited here to
+encode the labels within the logical language. To this aim, a type for
+worlds must be declared and modal propositions should be not of type
+\texttt{Prop} but of a lifted type \texttt{o} that depends on possible
+worlds:
 
@@
 \noindent 
-All modal connectives are simply lifted versions of the usual logical connectives.
-Notations are used to allow the modal connectives to be used as similarly as possible as the usual connectives.
-The prefix ``\texttt{m}'' is used to distinguish the modal connectives: if $\odot$ is a connective on type \text{Prop}, $\texttt{m}\odot$ is a connective on the lifted type $\texttt{o}$ of modal propositions.
+All modal connectives are simply lifted versions of the
+usual logical connectives. Notations are used to allow the modal
+connectives to be used as similarly as possible as the usual
+connectives. The prefix ``\texttt{m}'' is used to distinguish the
+modal connectives: if $\odot$ is a connective on type \text{Prop},
+$\texttt{m}\odot$ is a connective on the lifted type $\texttt{o}$ of
+modal propositions.
 
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Definition mequal (x y: u) := \fun w: i => x = y.
+\Definition mequal (x y: u)(w: i) := x = y.
 \Notation "x m= y" := (mequal x y) (at level 99, right associativity).
@@
 \noindent 
-Likewise, modal quantifiers are lifted versions of the usual quantifiers.
+Likewise, modal quantifiers are lifted versions of the usual
+quantifiers. \Coq's type system with dependent types is particularly
+helpful here. The modal quantifiers \texttt{A} and \texttt{E} are
+defined as depending on a type \texttt{t}. Therefore, they can
+quantify over variables of any type. Moreover, the curly brackets
+indicate that \texttt{t} is an implicit argument that can be inferred
+by \Coq's type inference mechanism. This allows notations\footnote{The
+keyword \texttt{\fun} indicates a lambda abstraction: \texttt{\fun \ x
+=> p} (or \texttt{\fun \ x:t => p}) denotes the function 
+$\lambda x:t.p$, which takes an argument $x$ (of type $t$) and returns $p$.}
+(i.e. \texttt{mforall} and \texttt{mexists}) that mimic the notations
+for \Coq's usual quantifiers (i.e. \texttt{forall} and
+\texttt{exists}).
 
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Definition A {t: Type}(p: t -> o) := \fun w => \fforall x, p x w.
-\Notation "'mforall'  x , p" := (A (fun x => p))
+\Definition A \{t: Type\}(p: t -> o)(w: i) := \fforall x, p x w.
+\Notation "'mforall'  x , p" := (A (\fun x => p))
   (at level 200, x ident, right associativity) : type_scope.
@@
 
-\Definition E {t: Type}(p: t -> o) := \fun w => exists x, p x w.
+\Definition E \{t: Type\}(p: t -> o)(w: i) := exists x, p x w.
 \Notation "'mexists' x , p" := (E (\fun x => p))
@@
 \noindent 
-The modal operators $\Diamond$ (\emph{possibly}) and $\Box$ (\emph{necessarily}) are defined accordingly to their meanings in the possible world semantics.
+The modal operators $\Diamond$ (\emph{possibly}) and $\Box$
+(\emph{necessarily}) are defined accordingly to their meanings in the
+possible world semantics. $\nec p$ holds at a world $w$ iff $p$ holds
+in every world $w_1$ reachable from $w$. $\pos p$ holds at world $w$
+iff $p$ holds in some world $w_1$ reachable from $w$.
 

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section 3 (tactics for modalities)

@@
 
+\noindent
+A modal proposition is valid iff it holds in every possible world. 
+This notion of modal validity is encoded by the following defined predicate:
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Definition V (p: o) := \fforall w, p w.
+\end{Verbatim}
+
+\noindent 
+To prove a modal proposition \texttt{p} (of type \texttt{o})
+within \Coq, the proposition \texttt{(V p)} (of type \texttt{Prop})
+should be proved instead. To increase the transparency of the
+embedding to the user, the following notation is provided, allowing
+\texttt{[ p ]} to be written instead of \texttt{(V p)}.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Notation "[ p ]" := (V p).
+\end{Verbatim}
+
+\section{Tactics for Modalities}
+
+Interactive theorem proving in \Coq is usually done with tactics,
+imperative commands that reduce the theorem to be proven (i.e. the
+goal) to simpler subgoals, in a bottom-up manner. The simplest tactics
+can be regarded as rules of a natural deduction calculus\footnote{The
+underlying proof system of \Coq (the Calculus of Inductive
+Constructions \cite{Paulin}) is actually more sophisticated and
+minimalistic than the calculus shown in Figure
+\ref{fig:NaturalDeduction}. But the calculus shown here suffices for
+the purposes of this paper.} (e.g. as those shown in Figure
+\ref{fig:NaturalDeduction}).  For example: the \texttt{intro} tactic
+can be used to apply the introduction rules for implication and for
+the universal quantifier; the \texttt{apply} tactic corresponds to the
+elimination rules for implication and for the universal quantifier;
+\texttt{split} performs conjunction introduction; \texttt{exists} can
+be used for existential quantifier introduction and \texttt{destruct}
+for its elimination.
+
 \newcommand{\s}{\qquad}
@@
 \begin{calculus}
-{propositional rules}
-{fig:PropositionalRules}
+{An arbitrary natural deduction calculus}
+{fig:NaturalDeduction}
+\begin{small}
 
@@
 
+\s
+\infer[\neg\neg_E]{A}{ \neg\neg A }
 \s\s
@@
 
-% \vspace{2em}
+\vspace{2em}
 
-% \s\s
-% \infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
-% \s\s
-% \infer[\vee_{I_1}]{A \vee B}{A}
-% \s\s
-% \infer[\vee_{I_2}]{A \vee B}{B}
+\s\s
+\infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
+\s\s
+\infer[\vee_{I_1}]{A \vee B}{A}
+\s\s
+\infer[\vee_{I_2}]{A \vee B}{B}
+
+\vspace{2em}
+
+%\begin{scriptsize}
+
+\s\s\s
+\infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] } 
+\s\s
+\infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }
+
+\vspace{1em}
+
+\s\s\s
+\infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
+\s\s
+\infer[\ex_E]{C}{ \ex x_{\tau}. A[x] & \infer*{C}{A[\alpha]} }
+
+%\end{scriptsize}
+
+\vspace{2em}
+
+\begin{center}
+$\alpha$ must respect the usual \emph{eigen-variable conditions}.
+\end{center}
+
+\end{small}
 
@@
 
-\begin{calculus}
-{double negation elimination}
-{fig:Classical}
-\infer[\neg\neg_E]{A}{ \neg\neg A }
-\end{calculus}
+To maximally preserve user intuition in interactive modal logic
+theorem proving, the embedding via the possible world semantics should
+be as transparent as possible to the user. Fortunately, the basic Coq
+tactics described above automatically unfold the shallowest modal
+definition in the goal.  Therefore, they can be used with modal
+connectives and quantifiers just as they are used with the usual
+connectives and quantifiers. The situation for the new modal
+operators, on the other hand, is not as simple, unfortunately.
 
-\begin{calculus}
-{quantifier rules}
-{fig:QuantifierRules}
+Since the modal operators are, in our embedding, essentially just
+abbreviations for quantifiers guarded by reachability conditions, the
+typical tactics for quantifiers can be used, in principle.  However,
+this exposes the user to the technicalities of the embedding,
+requiring him to deal with possible worlds and their reachability
+explicitly. In order to obtain transparency also for the modal
+operators, we have implemented a few specialized tactics using \Coq's
+Ltac language.
 
-\vspace{1em}
+When applied to a goal of the form \texttt{((box p) w0)}, the tactic
+\texttt{box\_i} will introduce a fresh new world \texttt{w} and then
+introduce the assumption that \texttt{w} is reachable from
+\texttt{w0}. The new goal will be \texttt{(p w)}.
 
-\s
-\infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] }
-\s
-\infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }
-\s\s
-\infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
-\s
-\infer[\ex_E]{A[\beta]}{ \ex x_{\tau}. A[x] }
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac box_i := \llet w := fresh "w" \iin \llet R := fresh "R" 
+               \iin (intro w at top; intro R at top).
+\end{Verbatim}
 
-\vspace{1em}
+\noindent 
+If the hypothesis \texttt{H} is of the form \texttt{((box p)
+w0)} and the goal is of the form \texttt{(q w)}, the tactic
+\texttt{box\_e H H1} creates a new hypothesis \texttt{H1: (p w)}. The
+tactic \texttt{box\_elim H w1 H1} is an auxiliary tactic for
+\texttt{box\_e}. It creates a new hypothesis \texttt{H1: (p w1)}, for
+any given world \texttt{w1}, not necessarily the goal's world
+\texttt{w}. It is also responsible for automatically trying (by
+\texttt{assumption}) to solve the reachability guard conditions,
+releasing the user from this burden.
 
-\begin{center}
-\textbf{eigen-variable conditions:} \\
-if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$; \\
-if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
-\end{center}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac box_elim H w1 H1 := \match type of H \with 
+      ((box ?p) ?w) => cut (p w1); 
+                       [intros H1 | (apply (H w1); try assumption)] \eend.
+\end{Verbatim}
 
-\vspace{1em}
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac box_e H H1:= \match goal \with | [ |- (_ ?w) ] => box_elim H w H1 \eend.
+\end{Verbatim}
 
-\end{calculus}
+\noindent 
+If the hypothesis \texttt{H} is of the form \texttt{((dia p)
+w0)}, the tactic \texttt{dia\_e H} generates a new hypothesis
+\texttt{H: (p w)} for a fresh new world \texttt{w} reachable from
+\texttt{w0}.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" \iin 
+                (destruct H \as [w [R H]]; move w at top; move R at top).
+\end{Verbatim}
+
+\noindent
+The tactic \texttt{dia\_i w} transforms a goal of the form 
+\texttt{((dia p) w0)} into the simpler goal \texttt{(p w)} 
+and automatically tries to solve the guard condition that 
+\texttt{w} must be reachable from \texttt{w0}.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Ltac dia_i w := (exists w; split; [assumption | idtac]).
+\end{Verbatim}
+
+\noindent If the new modal tactics above are regarded from a natural
+deduction point of view, they give rise to the inference
+rules\footnote{The natural deduction calculus with the rules from
+Figures \ref{fig:NaturalDeduction} and \ref{fig:NDK} is sound and
+complete relatively to the calculus of Figure
+\ref{fig:NaturalDeduction} extended with a necessitation rule and the
+modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Therefore,
+assuming that the calculus from \ref{fig:NaturalDeduction} is sound
+and Henkin-complete for classical higher-order logic, the additional
+modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete
+for modal higher-order logic \textbf{K}.} shown in Figure
+\ref{fig:NDK}. The labels that name boxes in those inference rules are
+precisely the worlds that annotate goals and hypotheses in \Coq with
+the modal embedding. A hypothesis of the form \texttt{(p w)}, where
+\texttt{p} is a modal proposition of type \texttt{o} and \texttt{w} is
+a world of type \texttt{i} indicates that \texttt{p} is an assumption
+created inside a box with name \texttt{w}.
 
@@
 \textbf{boxed assumption condition:} \\
-assumptions should be discharged within the box where they are made.
+assumptions should be discharged within the box where they are created.
 \end{center}
@@
 \vspace{1em}
-
 \end{calculus}
 
-\noindent
-To maximally preserve user intuition in interactive modal logic theorem
-proving, the embedding via the possible world semantics should be as
-transparent as possible to the user. Fortunately, basic Coq tactics
-such as \texttt{intro}, \texttt{apply} and \texttt{split} automatically
-unfold the modal definitions. They can thus be used with modal
-connectives and quantifiers just as they are used with the usual
-connectives and quantifiers.  For the modal operators, however, the
-following simple tactics have been implemented, in order to allow the
-user to work with the concepts of \emph{necessity} and
-\emph{possibility} without having to unfold the definitions of modal
-operators and think in terms of possible worlds.
+Finally, our implementation also provides the tactic \texttt{mv},
+standing for \emph{modal validity}, which replaces a goal of the form
+\texttt{[ p ]} (or equivalently \texttt{(V p)}) by a goal of the form
+\texttt{(p w)} for a fresh arbitrary world \texttt{w}.
 
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Ltac box_i := \llet w := fresh "w" \iin \llet R := fresh "R" 
-               \iin (intro w at top; intro R at top).
-
-\Ltac box_elim H w1 H1 := \match type of H \with 
-      ((box ?p) ?w) => cut (p w1); 
-                       [intros H1 | (apply (H w1); try assumption)] \eend.
-
-\Ltac box_e H H1:= \match goal \with | [ |- (_ ?w) ] => box_elim H w H1 \eend.
-
-\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" \iin 
-                (destruct H \as [w [R H]]; move w at top; move R at top).
-
-\Ltac dia_i w := (exists w; split; [assumption | idtac]).
-\end{Verbatim}
-
-\noindent
-Validity of a modal proposition is validity of its grounding on any world. A non-intrusive notation and a tactic that grounds the proposition to an arbitrary world are provided.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Definition V (p: o) := \fforall w, p w.
-\Notation "[ p ]" := (V p).
 \Ltac mv := \match goal \with [|- (V _)] => intro \eend.
@@
 
+\section{Convenient Modal Lemmas}
+
 \noindent 
-The following lemmas are quite convenient. The first is a kind of modus ponens applicable to formulas under the scope of a modal operator. The second allows pushing negations inside modalities.
+The following lemmas are quite convenient. The first is a
+kind of modus ponens applicable to formulas under the scope of a modal
+operator. The second allows pushing negations inside modalities.
 

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Add image for natural deduction proof of a lemma

@@
 \usepackage[utf8]{inputenc}
-\usepackage{url,amsmath,amssymb,fancyvrb,color}
+\usepackage{url,amsmath,amssymb,fancyvrb,color,graphicx}
 \usepackage{proof}
@@
 
-\section{Convenient Modal Lemmas}
+\section{Two Simple Modal Lemmas}
 
-\noindent 
-The following lemmas are quite convenient. The first is a
-kind of modus ponens applicable to formulas under the scope of a modal
-operator. The second allows pushing negations inside modalities.
+In order to illustrate the tactics described above, 
+we show \Coq proofs for two simple but useful modal lemmas. 
+The first lemma is a kind of modus ponens with formulas 
+under the scope of modal operators.
 
@@
 \Qed.
+\end{Verbatim}
 
+% \[
+% \infer[\all_I,\all_I]{w: \all p. \all q.  \pos p \imp (\nec (p \imp q)) \imp \pos q}{
+%   \infer[\imp_I^1,\imp_I^2]{w: \pos p \imp (\nec (p \imp q)) \imp (\pos q)}{
+%     \infer[\pos_I]{w: \pos q}{
+%       \infer[\imp_E]{w_0: q}{
+%         \infer[\pos_E]{w_0: p}{ 
+%           \infer[1]{w: \pos p}{}
+%         } 
+%         & 
+%         \infer[\nec_E]{w_0: p \imp q}{
+%           \infer[2]{w: \nec (p \imp q)}{}
+%         }
+%       }
+%     }
+%   }
+% }
+% \]
+
+% \newcommand{\down}[1]{\begin{array}{c}\phantom{p} \\ #1 \end{array} }
+% \newcommand{\up}[1]{\begin{array}{c}#1 \\ \phantom{p} \end{array} }
+
+% \[
+% \infer[\all_I,\all_I]{\all p. \all q.  \pos p \imp (\nec (p \imp q)) \imp \pos q}{
+%   \infer[\imp_I^1,\imp_I^2]{\pos p \imp (\nec (p \imp q)) \imp (\pos q)}{
+%     \infer[\pos_I]{\pos q}{
+%       \infer[\imp_E]{\up{q}}{
+%         \infer[\pos_E]{\down{p} }{ 
+%           \infer[1]{\pos p}{}
+%         } 
+%         & 
+%         \infer[\nec_E]{ \down{p \imp q} }{
+%           \infer[2]{\nec (p \imp q)}{}
+%         }
+%       }
+%     }
+%   }
+% }
+% \]
+
+\begin{figure}
+\centering
+\label{fig:mp_dia}
+\includegraphics[width=0.5\textwidth]{ND-(mp_dia).png}
+\caption{Natural deduction proof of \texttt{mp\_dia}}
+\end{figure}
+
+\noindent
+The second lemma allows negations to be pushed inside modalities.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
@@
 \end{Verbatim}
+
+\section{Modal Logics beyond K}
 
@@
 
+todo: automation, firstorder
+
 The theological implications of the verified correctness of this proof

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Fix labels and references, and improve the text

@@
 \usepackage{xspace}
+\usepackage{float}
 
@@
 
-Modal logics extend the usual formal logic language with modal
+Modal logics extend usual formal logic languages by adding modal
 operators ($\Box$ and $\Diamond$) and are characterized by the
@@
 resulting infra-structure for modal reasoning within \Coq (as
-described in Section \ref{sec:modal}) provides a user experience where
+described in Section \ref{sec:Embedding}) provides a user experience where
 modalities can be handled transparently and straightforwardly.
@@
 \section{The Embedding of Modal Logics in Coq}
-\label{sec:modal}
+\label{sec:Embedding}
 
@@
 \section{Tactics for Modalities}
+\label{sec:Tactics}
 
@@
 
+\begin{center}
+$\neg A$ is an abbreviation for $A \imp \bot$.
+\end{center}
+
 \end{small}
@@
 
+\noindent
 To maximally preserve user intuition in interactive modal logic
@@
 \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-\s\s\s\s\s
+\s\s\s\s
 \infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
@@
 \textbf{eigen-box condition:}\\ 
-$\omega$ must be a fresh name for a box. \\
+$\nec_I$ and $\pos_E$ are \emph{strong} modal rules: \\
+$\omega$ must be a fresh name for the box they access \\ 
+(in analogy to the eigen-variable condition for strong quantifier rules). \\
+Every box must be accessed by \emph{exactly one} strong modal inference. \\
 \vspace{0.5em}
@@
 \section{Two Simple Modal Lemmas}
+\label{sec:SimpleExamples}
 
@@
 we show \Coq proofs for two simple but useful modal lemmas. 
-The first lemma is a kind of modus ponens with formulas 
+The first lemma resembles modus ponens, but with formulas 
 under the scope of modal operators.

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Add explanation of the correspondence between the script and the natural deduction proof

@@
 
-\begin{figure}
+\noindent 
+The proof of this lemma is displayed as a natural deduction
+proof in  Figure \ref{fig:mp_dia}. As expected, \Coq's basic tactics
+(e.g. \texttt{intros} and \texttt{apply}) work without modification.
+The \texttt{intros p q H1 H2} tactic application corresponds to the
+universal quantifier  and implication introduction inferences in the
+bottom of the proof.  The \texttt{apply H3} tactic application
+corresponds to the implication  elimination inference. The $\pos_E$,
+$\pos_I$ and $\nec_E$ inferences correspond, respectively, to the
+\texttt{dia\_e H1},  \texttt{dia\_i w0} and \texttt{box\_e H2 H3}
+tactic applications.  The internal box named $w_0$ is accessed by
+exactly one strong  modal inference, namely $\pos_E$.
+
+\begin{figure}[H]
 \centering
@@
 \noindent
-The second lemma allows negations to be pushed inside modalities.
+The same lemma could be proved without the new modal tactics, as shown below.
+But this is clearly disadvantageous, for several reasons: 
+the proof script becomes longer;
+the definitions of modal operators must be unfolded, 
+either explicitly (as done below) or implicitly in the user's mind; 
+tactic applications dealing with modal operators cannot be easily distinguished
+from tactic applications dealing with quantifiers; and  
+hypotheses about the reachability of worlds (e.g. \texttt{R1} below) 
+must be handled explicitly. In summary, without the modal tactics, 
+a convenient and intuitive correspondence between proof scripts and 
+modal natural deduction proofs would be missing.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Lemma mp_dia_alternative: 
+  [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
+\Proof. mv. 
+intros p q H1 H2. unfold dia. unfold dia in H1. unfold box in H2.
+destruct H1 as [w0 [R1 H1]]. exists w0. split.
+  exact R1.
+  apply H2.
+    exact R1.
+    exact H1.
+\Qed.
+\end{Verbatim}
+
+\noindent
+The second useful lemma allows negations to be pushed inside modalities, 
+and again the modal tactics allow this to be proved conveniently and elegantly.
 

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section about modal logics beyond K

@@
 \section{Modal Logics beyond K}
-
-\noindent
-The most well-known modal logics can be distinguished by their acceptance of the following reachability axioms. \textbf{S5} is the modal logic that accepts all three.
-
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Axiom reflexivity: \fforall w, r w w.
-\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
-\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
-\end{Verbatim}
-
-\noindent
-Modal logic ``axioms'' can be derived from the reachability axioms.
+\label{sec:BeyondK}
+
+The embedding described in Section \ref{sec:Embedding} and the 
+new tactics described in Section \ref{sec:Tactics} allow convenient 
+interactive reasoning for modal logic \textbf{K} within \Coq. 
+The axiom K is easily derivable:
 
@@
 \Qed.
-
+\end{Verbatim}
+
+\noindent
+For other modal logics beyond \textbf{K}, their frame conditions constraining 
+the reachability relation must be stated as \Coq axioms.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Axiom reflexivity: \fforall w, r w w.
+
+\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
+
+\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
+\end{Verbatim}
+
+\noindent
+Hilbert-style modal logic axioms, such as for example T, 
+can be easily derived from their corresponding frame conditions:
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Theorem T: [ mforall p, (box p) m-> p ].
@@
 \noindent
-In strong modal logics, such as S5, iterations of modal operators can be collapsed. This is a controversial principle used in G\"odel's and Scott's versions of the proof.
+In strong modal logics, such as S5 (which enforces all three frame conditions specified above), iterations of modal operators can be collapsed.
+This is a controversial principle used in G\"odel's proof shown in Section \ref{sec:proof}.
 

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section about Gödel's proof

@@
 
-\section{G\"odel's Proof of God's Existence}
+\section{G\"odel's Ontological Argument for God's Existence}
 \label{sec:proof}
+
+In order to demonstrate the efficacy and convenience of the modal
+embedding approach not only for proving simple lemmas and theorems,
+but also for larger developments, we include here a full and detailed
+formalization of Gödel's ontological argument.
 
@@
 abstract ontological arguments are an old tradition in philosophy and
-theology.  
-G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
-this tradition, following particularly the footsteps of Leibniz.
-Various slightly different versions of axioms and definitions have
-been considered by G\"{o}del and by several philosophers who commented
-on his proof
-(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). The formalization shown in this section aims at
-being as similar as possible to Dana Scott's version of the proof \cite{ScottNotes}.
-The formulation and numbering of axioms, definitions and theorems is the same as in Scott's notes.
-Even the Coq proof scripts follow all the steps in Scott's notes.
-Scott's assertions are emphasized with comments.
-In contrast to the formalization in Isabelle \cite{AFP}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation.
-This was a deliberate choice, mainly because it allowed a qualitative evaluation of the convenience of the possible world semantic embedding approach for \emph{interactive} theorem proving.
-Moreover, in order to formalize precisely Scott's version and not another automatically found version, automation would have to be heavily limited anyway.
-Furthermore, the deliberate preference for simple tactics (mostly \emph{intro}, \emph{apply} and the modal tactics described in the previous section) results in proof scripts that could be read as natural deduction proofs.
-This hopefully makes the formalization more accessible to those who are not experts in Coq's tactics but are nevertheless interested in G\"odel's proof.
+theology. G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a
+modern culmination of this tradition, following particularly the
+footsteps of Leibniz. Various slightly different versions of axioms
+and definitions have been considered by G\"{o}del and by several
+philosophers who commented on his proof (cf. 
+\cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+The formalization shown in this section aims at being as similar as
+possible to Dana Scott's version of the proof \cite{ScottNotes}. The
+formulation and numbering of axioms, definitions and theorems is the
+same as in Scott's notes. Even the \Coq proof scripts follow all the
+steps in Scott's notes. Scott's assertions are emphasized with
+comments. In contrast to the formalization in \Isabelle \cite{AFP},
+where automation via \texttt{Metis} \cite{Hurd03first-orderproof} and
+\texttt{Sledgehammer} \cite{Sledgehammer} using tools such as
+\texttt{Nitpick} \cite{Nitpick}, \texttt{LEO-II} \cite{LEO-II} and
+\texttt{Satallax} \cite{Satallax} has been successfully employed, the
+formalization in \Coq used no automation. This was a deliberate
+choice, mainly because it allowed a qualitative evaluation of the
+convenience of the possible world semantic embedding approach for
+\emph{interactive} theorem proving. Moreover, in order to formalize
+precisely Scott's version and not another automatically found version,
+automation would have to be heavily limited anyway. Furthermore, the
+deliberate preference for simple tactics (mostly \emph{intro},
+\emph{apply} and the modal tactics described in Section
+\ref{sec:Tactics}) results in proof scripts that closely correspond to
+natural deduction proofs. This hopefully makes the formalization more
+accessible to those who are not experts in \Coq's tactics but are
+nevertheless interested in G\"odel's proof.
+
+Gödel's proof requires \Coq's classical logic libraries as well as the
+\texttt{Modal} library developed by us and described in sections
+\ref{sec:Embedding} and \ref{sec:Tactics}.
 
@@
 \Require \Import Coq.Logic.Classical Coq.Logic.Classical_Pred_Type Modal.
-
+\end{Verbatim}
+
+\noindent 
+In Scott's notes, classicality occurs in uses of the
+principle of proof by contradiction. Therefore, in order to clearly
+indicate where classical logic is needed in the proof scripts, a
+simple tactic that simulates proof by contradiction can be created:
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Ltac proof_by_contradiction H := apply NNPP; intro H.
-
+\end{Verbatim}
+
+\noindent
+Gödel's theory has a single higher-order constant, \texttt{Positive}, 
+which ought to hold for properties considered \emph{positive} in a moral sense.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \com{(* Constant predicate that distinguishes positive properties *)}
 \Parameter Positive: (u -> o) -> o.
-
+\end{Verbatim}
+
+\noindent
+Five axioms are stated by Gödel to characterize positivity. 
+The first three axioms are needed to establish the first theorem and 
+its corollary stating the possibility of God's existence, 
+with God defined as a being possessing all positive properties.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \com{(* Axiom A1:} 
@@
 \Proof. mv. apply theorem1. apply axiom3. \Qed.
-
+\end{Verbatim}
+
+\noindent 
+The second part of the proof consists in showing that if
+God's existence is possible then it must be necessary
+(\texttt{lemma2}), for which the \textbf{S5} principle
+\texttt{dia\_box\_to\_box} is used.
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \com{(* Axiom A4: positive properties are necessarily positive *)}

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write conclusion

@@
 
-The successful formalization of Scott's version of G\"odel's proof of
-God's existence indicates that the \emph{embedding} of modal logics
-into higher-order logics via the \emph{possible world semantics} is a
-viable approach for interactive theorem proving within modal logics.
-The minimalistic implementation of the embedding (Section
-\ref{sec:modal}) takes special care to hide the underlying possible
-world machinery from the user.  An inspection of the proof scripts in
-Section \ref{sec:proof} shows that this goal has been achieved. The
-user does not have to explicitly bother about worlds and their mutual
-reachability; the provided tactics for modalities do the job for him/her.
-Moreover, for subgoals that do not involve modalities, the user has all the
-usual interactive tactics at his/her disposal.
-%The implementation of hints to allow automation tactics to fully cope with the embedding remains for future work.
-Nevertheless, at this stage a tactic such as \texttt{firstorder} is able to prove all lemmas and theorems in Section \ref{sec:modal} fully automatically.
-
-HOL Light, Isabelle
-
-todo: automation, firstorder
-
-The theological implications of the verified correctness of this proof
-certainly depend on a critical discussion of the underlying concepts,
-definitions and axioms, which is beyond the scope of this paper. 
-%What
-%can be said though is that our work presents further evidence that
-%theism is at least not (obviously) irrational. 
-Clearly, the application of theorem proving technology to the domain of
-theoretical philosophy can --- as already pictured by Leibniz ---
-be very fruitful for both areas. In fact, new results of
-philosophical interest have been obtained by our
-work~\cite{arXiv}. Moreover a first, basic infrastructure for
-interactive and automated reasoning in higher-order modal logics has
-been created and will be further improved in the future.
-
-\paragraph{Acknowledgements:} we thanks Cedric Auger and Laurent Th\'ery, for their answers to our questions about Ltac in the Coq-Club mailing-list.
+The successful formalization of Scott's version of  G\"odel's
+ontological argument indicates that the \emph{embedding} of  modal
+logics into higher-order logics via the \emph{possible world
+semantics} is a viable approach for fully interactive theorem proving
+within modal logics. Our lightweight implementation of the embedding
+(described in sections \ref{sec:Embedding} and \ref{sec:Tactics})
+takes special care to hide the underlying possible world machinery
+from the user. An inspection of the proof scripts in Section
+\ref{sec:proof} shows that this goal has been achieved. The user does
+not have to explicitly bother about worlds and their mutual
+reachability; the provided tactics for modalities do the job for
+him/her. Moreover, for subgoals that do not involve modalities, the
+user has all the usual interactive tactics at his/her disposal.
+
+Although fully automated (as opposed to interactive) theorem proving
+is beyond the scope of this paper, it is worth mentioning that all
+lemmas and theorems in sections \ref{sec:Embedding},
+\ref{sec:SimpleExamples} and \ref{sec:BeyondK} could have been proven
+automatically using \Coq's \texttt{firstorder} tactic. The
+implementation of hints to allow \Coq's automatic tactics to take full
+advantage of the embedding and the modal axioms still remains for
+future work.
+
+The theological implications of the verified correctness of Gödel's
+proof certainly depend on a critical discussion of the underlying
+concepts, definitions and axioms, which is beyond the scope of this
+paper.  Clearly, the application of theorem proving technology to the
+domain of theoretical philosophy can --- as already pictured by
+Leibniz --- be very fruitful for both areas. We hope that the
+infrastructure  that we have implemented for interactive and automated
+reasoning in higher-order modal logics will be useful outside
+philosophy as well.
+
+\paragraph{Acknowledgements:} we thank Cedric Auger and Laurent Théry, 
+for their answers to our questions about Ltac in the Coq-Club mailing-list.
 

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
modified draft paper

@@
 \title{
-Modal Logics in the Coq Proof Assistant
+(Interacting with)? \\ Modal Logics in the Coq Proof Assistant
 }
@@
 \institute{
-  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
-  \email{c.benzmueller@gmail.com}
+  % Dept. of Mathematics and Computer Science, 
+  Freie Universit\"{a}t Berlin, Germany,  %\\
+  \email{c.benzmueller@fu-berlin.de}
   \and 
-  Theory and Logic Group, Vienna University of Technology, Austria \\
+  % Theory and Logic Group, 
+  Vienna University of Technology, Austria, %\\
    \email{bruno@logic.at}
@@
 \begin{abstract}  
-This paper describes an embedding of higher-order
-modal logics in the \Coq proof assistant. {\Coq}'s capabilities are
-thus extended in a minimalistic manner, which is nevertheless
-sufficient for the formalization of significant modal proofs. The
+  This paper describes an embedding of higher-order modal logics in
+  the \Coq proof assistant. {\Coq}'s capabilities are extended in a
+  minimalistic manner, which is nevertheless sufficient for the
+  formalization of significant, non-trivial modal logic proofs. The
   elegance, flexibility and convenience of this approach, from a user
-perspective, are illustrated here with the successful formalization of
-Gödel's ontological argument.  
+  perspective, are illustrated here with the successful formalization
+  of Gödel's ontological argument.
 \end{abstract}
@@
 philosophy \cite{ToDo} to formal software verification
-\cite{ToDo:TLAPlus}.
+\cite{ToDo:TLAPlus}.\footnote{ChrisToBruno: \cite{ModalLogic} should
+  suffice as reference, applications are addressed in Part 4.}
 
@@
 higher-order logics, for which existing higher-order automated theorem
-provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
+provers \cite{LEO-II,Satallax} exist.  Embedding is flexible,
 because various modal logics can be easily supported by adding their
@@
 
+Propositional and quantified modal logics have applications in various
+fields, including, for example, verification, agent technologies,
+artificial intelligence, computational linguistics, and computational
+philosophy. Our contribution may thus serve as a starting point for
+many interesting further projects.
+
 In order to illustrate the use of the implemented embedding, we show
-here the formalization of Scott's version \cite{Scott} of Gödel's
+here the formalization of Scott's version \cite{ScottNotes} of Gödel's
 ontological argument for God's existence (in Section \ref{sec:proof}).
@@
 argument was automatically verified in our previous work on embedding-
-based fully automated modal theorem proving \cite{ArXiv,AFP}. But this
+based fully automated modal theorem proving \cite{arXiv,AFP}. This
 paper presents the first fully interactive and detailed formalization
-of this proof in a proof assistant.
+of this proof in a proof assistant. Hence we here reveal the proof
+structure on a cognitively adequate level of
+detail. Such detailed proof structure has been hidden in our
+other papers on the subject due to the use of automated theorem
+provers. In comparison to \cite{Rushby,Zalta}, however, our
+contribution stands out with its surprising technical 
+simplicity and elegance.
 
@@
 \ref{fig:NaturalDeduction}. But the calculus shown here suffices for
-the purposes of this paper.} (e.g. as those shown in Figure
+the purposes of this paper.} \footnote{ChrisToBruno: We need to briefly address here
+    the fact that this calculus is classical while Coq pure is
+    constructive; one or two sentences should be fine.} 
+(e.g. as those shown in Figure
 \ref{fig:NaturalDeduction}).  For example: the \texttt{intro} tactic
@@
 \begin{calculus}
-{An arbitrary natural deduction calculus}
+{A (classical)  The natural deduction calculus}
 {fig:NaturalDeduction}
@@
 operators, we have implemented a few specialized tactics using \Coq's
-Ltac language.
+Ltac language. We present these tactics next.
 
@@
 abstract ontological arguments are an old tradition in philosophy and
-theology. G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a
-modern culmination of this tradition, following particularly the
-footsteps of Leibniz. Various slightly different versions of axioms
-and definitions have been considered by G\"{o}del and by several
-philosophers who commented on his proof (cf. 
+theology. G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern
+culmination of this tradition, following particularly the footsteps of
+Leibniz. Various slightly different versions of axioms and definitions
+have been considered by G\"{o}del and by several philosophers who
+commented on his proof (cf.
 \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
-The formalization shown in this section aims at being as similar as
+The formalization shown in this Section aims at being as similar as
 possible to Dana Scott's version of the proof \cite{ScottNotes}. The
 formulation and numbering of axioms, definitions and theorems is the
-same as in Scott's notes. Even the \Coq proof scripts follow all the
-steps in Scott's notes. Scott's assertions are emphasized with
-comments. In contrast to the formalization in \Isabelle \cite{AFP},
-where automation via \texttt{Metis} \cite{Hurd03first-orderproof} and
-\texttt{Sledgehammer} \cite{Sledgehammer} using tools such as
-\texttt{Nitpick} \cite{Nitpick}, \texttt{LEO-II} \cite{LEO-II} and
-\texttt{Satallax} \cite{Satallax} has been successfully employed, the
-formalization in \Coq used no automation. This was a deliberate
-choice, mainly because it allowed a qualitative evaluation of the
-convenience of the possible world semantic embedding approach for
-\emph{interactive} theorem proving. Moreover, in order to formalize
-precisely Scott's version and not another automatically found version,
-automation would have to be heavily limited anyway. Furthermore, the
-deliberate preference for simple tactics (mostly \emph{intro},
-\emph{apply} and the modal tactics described in Section
-\ref{sec:Tactics}) results in proof scripts that closely correspond to
-natural deduction proofs. This hopefully makes the formalization more
-accessible to those who are not experts in \Coq's tactics but are
-nevertheless interested in G\"odel's proof.
+same as in Scott's notes. Even the \Coq proof scripts follow precisely
+all the steps in Scott's notes. Scott's assertions are emphasized
+below with comments. In contrast to the formalization in \Isabelle
+\cite{AFP}, where automation via \texttt{Metis}
+\cite{Hurd03first-orderproof} and \texttt{Sledgehammer}
+\cite{Sledgehammer} using tools such as \texttt{Nitpick}
+\cite{Nitpick}, \texttt{LEO-II} \cite{LEO-II} and \texttt{Satallax}
+\cite{Satallax} has been successfully employed, the formalization in
+\Coq used no automation. This was a deliberate choice, mainly because
+it allowed a qualitative evaluation of the convenience of the
+embedding approach for \emph{interactive} theorem proving. Moreover,
+in order to formalize precisely Scott's version and not another
+automatically found version, automation would have to be heavily
+limited anyway. Furthermore, the deliberate preference for simple
+tactics (mostly \emph{intro}, \emph{apply} and the modal tactics
+described in Section \ref{sec:Tactics}) results in proof scripts that
+closely correspond to natural deduction proofs. This hopefully makes
+the formalization more accessible to those who are not experts in
+\Coq's tactics but are nevertheless interested in G\"odel's proof.
 
 Gödel's proof requires \Coq's classical logic libraries as well as the
-\texttt{Modal} library developed by us and described in sections
+\texttt{Modal} library developed by us and described in Sections
 \ref{sec:Embedding} and \ref{sec:Tactics}.
@@
 The successful formalization of Scott's version of  G\"odel's
-ontological argument indicates that the \emph{embedding} of  modal
+ontological argument indicates that the \emph{embedding} of
+higher-order modal
 logics into higher-order logics via the \emph{possible world
 semantics} is a viable approach for fully interactive theorem proving
-within modal logics. Our lightweight implementation of the embedding
-(described in sections \ref{sec:Embedding} and \ref{sec:Tactics})
+within (even non-trivial) modal logics. Our lightweight implementation of the embedding
+(described in Sections \ref{sec:Embedding} and \ref{sec:Tactics})
 takes special care to hide the underlying possible world machinery
@@
 is beyond the scope of this paper, it is worth mentioning that all
-lemmas and theorems in sections \ref{sec:Embedding},
+lemmas and theorems in Sections \ref{sec:Embedding},
 \ref{sec:SimpleExamples} and \ref{sec:BeyondK} could have been proven
@@
 domain of theoretical philosophy can --- as already pictured by
-Leibniz --- be very fruitful for both areas. We hope that the
-infrastructure  that we have implemented for interactive and automated
-reasoning in higher-order modal logics will be useful outside
-philosophy as well.
-
-\paragraph{Acknowledgements:} we thank Cedric Auger and Laurent Théry, 
+Leibniz --- be very fruitful for both areas.
+
+The infrastructure that we have implemented for interactive and
+automated reasoning in higher-order modal logics will clearly be
+useful also outside philosophy; the range of potential applications is
+in fact 
+very wide.
+
+\paragraph{Acknowledgements:} We thank Cedric Auger and Laurent Théry, 
 for their answers to our questions about Ltac in the Coq-Club mailing-list.
@@
   Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
+
+\bibitem{JensOtten}
+C.~Benzm{\"u}lller, J. Otten, and Th. Raths.
+\newblock Implementing and Evaluating Provers for First-order Modal
+Logics. 
+\newblock In {\em ECAI 2012}, pages 163-168. IOS Press, 2012.
 

### 2014-12-11 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve language of sections 1 to 5

@@
 \title{
-(Interacting with)? \\ Modal Logics in the Coq Proof Assistant
+Interacting with Modal Logics \\
+in the Coq Proof Assistant
 }
@@
 \author{
-  Christoph Benzm\"{u}ller\inst{1}\thanks{This work has been supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
+  Christoph Benzm\"{u}ller\inst{1}\thanks{Supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
   \and 
@@
 
-Modal logics extend usual formal logic languages by adding modal
+Modal logics \cite{ModalLogic} extend usual formal logic languages by adding modal
 operators ($\Box$ and $\Diamond$) and are characterized by the
 \emph{necessitation rule}, according to which $\Box A$ is a theorem if
-$A$ is a theorem, although $A \imp \Box A$ is not necessarily a
-theorem. Various informal concepts, such as \emph{necessity and
-possibility}, \emph{knowledge and belief}, \emph and \emph{temporal
-globality and eventuality}, have been formalized with the help of
-modal operators. Therefore, modal logics \cite{ModalLogic} are well-
-established for a wide variety of application domains, ranging from
-philosophy \cite{ToDo} to formal software verification
-\cite{ToDo:TLAPlus}.\footnote{ChrisToBruno: \cite{ModalLogic} should
-  suffice as reference, applications are addressed in Part 4.}
-
-However, general automated reasoning support for modal logics is still
+$A$ is a theorem, even though $A \imp \Box A$ is not necessarily a
+theorem. Various notions, such as \emph{necessity and
+possibility}, \emph{obligation and permission}, 
+\emph{knowledge and belief}, \emph and \emph{temporal
+globality and eventuality}, which are ubiquitous in various application domains,
+have been formalized with the help of modal operators.
+
+Nevertheless, general automated reasoning support for modal logics is still
 not as well-developed as for classical logics. Deduction tools
@@
 
-Another recently explored possibility \cite{J23,B9,J22} is the
+Another recently explored possibility is the
 embedding of first-order and even higher-order modal logics into classical
-higher-order logics, for which existing higher-order automated theorem
-provers \cite{LEO-II,Satallax} exist.  Embedding is flexible,
-because various modal logics can be easily supported by adding their
-characteristic axioms.  The embedding approach can also be easily
-adapted to support multiple modalities, and it is possible to replace
-constant domain quantifiers by varying or cumulative domain
-quantifiers.  Moreover, the approach is relatively simple to
+higher-order logics \cite{J23,B9,J22}, 
+for which existing higher-order automated theorem
+provers \cite{LEO-II,Satallax} exist. The embedding approach is flexible,
+because various modal logics (even with multiple modalities or varying/cumulative domain quantifiers) can be easily supported by stating their
+characteristic axioms. Moreover, the approach is relatively simple to
 implement, because it does not require any modification in the source
@@
 other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
-\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal of this
-paper is to study this question. Our answer is positive.
+\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal and novelty
+of this paper is to study this question. Our answer is positive.
 
@@
 excessive overhead.
-
-Propositional and quantified modal logics have applications in various
-fields, including, for example, verification, agent technologies,
-artificial intelligence, computational linguistics, and computational
-philosophy. Our contribution may thus serve as a starting point for
-many interesting further projects.
 
@@
 The simpler ontological argument of Anselm has been automatically
-verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-
-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
+verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
+first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
 argument was automatically verified in our previous work on embedding-
@@
 
+In addition to philosophy, propositional and quantified modal logics 
+have (potential) applications in various other fields as well, including, for instance, 
+verification, artificial intelligence
+agent technologies, law and linguistics. \marginpar{BrunoToChris: Do we need citations for these applications?} 
+Therefore, the main contribution described in this paper -- convenient techniques for leveraging a powerful proof assistant such as \Coq for interactive reasoning for modal logics -- may serve as a starting point for
+many interesting projects.
+
 \section{The Embedding of Modal Logics in Coq}
@@
 
-A crucial aspect of modal logics \cite{ModalLogic} is that the so-
-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$
+A crucial aspect of modal logics \cite{ModalLogic} is that the 
+so-called \emph{necessitation rule} allows $\Box A$ to be derived if $A$
 is a theorem, but $A \imp \Box A$ is not necessarily a theorem. Naive
@@
 
-The embedding is related to labeling techniques \cite{Labels}.
-However, the expressiveness of higher-order logic is exploited here to
-encode the labels within the logical language. To this aim, a type for
+The embedding technique described in this section is 
+related to labeling techniques \cite{Labels}.
+However, the expressiveness of higher-order logic can be exploited in order to
+encode the labels within the logical language itself. To this aim, a type for
 worlds must be declared and modal propositions should be not of type
@@
 underlying proof system of \Coq (the Calculus of Inductive
-Constructions \cite{Paulin}) is actually more sophisticated and
+Constructions (CIC) \cite{Paulin}) is actually more sophisticated and
 minimalistic than the calculus shown in Figure
 \ref{fig:NaturalDeduction}. But the calculus shown here suffices for
-the purposes of this paper.} \footnote{ChrisToBruno: We need to briefly address here
-    the fact that this calculus is classical while Coq pure is
-    constructive; one or two sentences should be fine.} 
+the purposes of this paper.
+This calculus is classical, because of the double negation elimination rule.
+Although CIC is intuitionistic, it can be made classical by importing \Coq's classical library, which adds the axiom of the \emph{excluded middle} and the double negation elimination lemma.}
 (e.g. as those shown in Figure
@@
 explicitly. In order to obtain transparency also for the modal
-operators, we have implemented a few specialized tactics using \Coq's
-Ltac language. We present these tactics next.
+operators, we have implemented specialized tactics using \Coq's
+Ltac language. These tactics are among our main contributions and they are described in the remainder of this section.
 
@@
 
+\noindent
 Finally, our implementation also provides the tactic \texttt{mv},
@@
 \noindent
-For other modal logics beyond \textbf{K}, their frame conditions constraining 
-the reachability relation must be stated as \Coq axioms.
+For other modal logics beyond \textbf{K}, their frame conditions, which constrain 
+the reachability relation, must be stated as \Coq axioms.
 
@@
 \noindent
-In strong modal logics, such as S5 (which enforces all three frame conditions specified above), iterations of modal operators can be collapsed.
-This is a controversial principle used in G\"odel's proof shown in Section \ref{sec:proof}.
+In a strong modal logic such as \textbf{S5} (which requires all three frame conditions specified above), sequences of modal operators can be collapsed to a single modal operator.
+One such collapsing principle is specified and proven below.
+By applying it iteratively, any sequence $\pos \ldots \pos \nec p$ could be collapsed to $\nec p$.
 
@@
 God's existence is possible then it must be necessary
-(\texttt{lemma2}), for which the \textbf{S5} principle
+(\texttt{lemma2}), for which the controversial \textbf{S5} principle
 \texttt{dia\_box\_to\_box} is used.

### 2014-12-11 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve language of sections 6 and 7

@@
 embedding approach for \emph{interactive} theorem proving. Moreover,
-in order to formalize precisely Scott's version and not another
-automatically found version, automation would have to be heavily
+in order to formalize exactly Scott's version and not some arbitrary 
+version found automatically, automation would have to be heavily
 limited anyway. Furthermore, the deliberate preference for simple
@@
 described in Section \ref{sec:Tactics}) results in proof scripts that
-closely correspond to natural deduction proofs. This hopefully makes
+closely correspond to common natural deduction proofs. This hopefully makes
 the formalization more accessible to those who are not experts in
@@
 In Scott's notes, classicality occurs in uses of the
-principle of proof by contradiction. Therefore, in order to clearly
+principle of proof by contradiction. In order to clearly
 indicate where classical logic is needed in the proof scripts, a
-simple tactic that simulates proof by contradiction can be created:
+simple tactic that simulates proof by contradiction was created:
 
@@
 \noindent
-Five axioms are stated by Gödel to characterize positivity. 
-The first three axioms are needed to establish the first theorem and 
-its corollary stating the possibility of God's existence, 
-with God defined as a being possessing all positive properties.
+God is defined as a being possessing all positive properties, and
+five axioms are stated to characterize positivity. 
+The first part of the proof culminates in \texttt{corollary1} and establishes
+that God's existence is possible.
 
@@
 God's existence is possible then it must be necessary
-(\texttt{lemma2}), for which the controversial \textbf{S5} principle
+(\texttt{lemma2}). The controversial \textbf{S5} principle
 \texttt{dia\_box\_to\_box} is used.
@@
 semantics} is a viable approach for fully interactive theorem proving
-within (even non-trivial) modal logics. Our lightweight implementation of the embedding
+within modal logics. Our lightweight implementation of the embedding
 (described in Sections \ref{sec:Embedding} and \ref{sec:Tactics})
@@
 useful also outside philosophy; the range of potential applications is
-in fact 
-very wide.
+in fact very wide.
 

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Use bibtex for bibliography

@@
 for modal logics are often limited to propositional, quantifier-free,
-fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
+fragments \cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers} or tailored to particular
 modal logics and their applications \cite{ToDo:StephanMerz}; 
@@
 other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
-\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal and novelty
+\texttt{HOL-Light} \cite{HOLLight}, as well.} \cite{Coq}. The main goal and novelty
 of this paper is to study this question. Our answer is positive.
@@
 within modal logics. Our lightweight implementation of the embedding
-(described in Sections \ref{sec:Embedding} and \ref{sec:Tactics})
+(available in \cite{FormalTheologyRepository} and described 
+in Sections \ref{sec:Embedding} and \ref{sec:Tactics})
 takes special care to hide the underlying possible world machinery
@@
 
-%\footnotesize
 \bibliographystyle{plain}
-%\bibliography{Bibliography}
-\begin{thebibliography}{10}
-
-\bibitem{Adams}
-R.M.~Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1995.
-
-\bibitem{AndersonGettings}
-A.C.~Anderson and M.~Gettings.
-\newblock G\"odel ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
-
-\bibitem{JensOtten}
-C.~Benzm{\"u}lller, J. Otten, and Th. Raths.
-\newblock Implementing and Evaluating Provers for First-order Modal
-Logics. 
-\newblock In {\em ECAI 2012}, pages 163-168. IOS Press, 2012.
-
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C.~Paulson.
-\newblock Exploring properties of normal multimodal logics in simple type
-  theory with {LEO-II}.
-\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
-
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C.~Paulson.
-\newblock Quantified multimodal logics in simple type theory.
-\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
-  7(1):7--20, 2013.
-
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
-\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
-  logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
-  162--170. Springer, 2008.
-
-\bibitem{arXiv}
-C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
-\newblock Formalization, Mechanization and Automation of G{\"o}del's Proof of God's Existence.
-\newblock {\em {arXiv}}:1308.4526, 2013.
-
-\bibitem{AFP}
-C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
-\newblock G{\"o}del's God in Isabelle/HOL. 
-\newblock {\em Archive of Formal Proofs}, 2013.
-
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran.
-\newblock {\em {Interactive Theorem Proving and Program Development}}.
-\newblock Springer, 2004.
-
-\bibitem{ModalLogic}
-P.~Blackburn, J.v.~Benthem, and F.~Wolter (eds.), \newblock {\em {Handbook of Modal Logic}}. Elsevier, 2006.
-
-\bibitem{Sledgehammer}
-J.C.~Blanchette, S.~B\"ohme, and L.C.~Paulson.
-\newblock Extending {Sledgehammer} with {SMT} solvers.
-\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
-
-\bibitem{Nitpick}
-J.C.~Blanchette and T.~Nipkow.
-\newblock Nitpick: A counterexample generator for higher-order logic based on a
-  relational model finder.
-\newblock In {\em Proc. of ITP 2010}, no.~6172 in LNCS, pages 131--146.
-  Springer, 2010.
-
-\bibitem{Zalta}
-P.E.~Oppenheimera and E.N.~Zalta.
-\newblock A Computationally-Discovered Simplification of the Ontological Argument.
-\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.
-
-\bibitem{Satallax}
-C.E.~Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
-  Springer, 2012.
-
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary bibligraphy on the ontological proof
-  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
-
-\bibitem{Fitting}
-M.~Fitting.
-\newblock {\em {Types, Tableaux and G\"odel's God}}.
-\newblock Kluver Academic Press, 2002.
-
-\bibitem{Labels}
-D.M.~Gabbay. {\em Labelled Deductive Systems}. Clarendon Press, 1996.
-
-\bibitem{Goedel1970}
-K.~G\"odel.
-\newblock Ontological proof.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1970.
-
-\bibitem{GoedelNotes}
-K.~G\"odel.
-\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
-\newblock In  \cite{sobel2004logic}, 2004.
-
-\bibitem{Hazen}
-A.P.~Hazen.
-\newblock On g\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
-
-\bibitem{Hurd03first-orderproof}
-J.~Hurd.
-\newblock First-order proof tactics in higher-order logic theorem provers.
-\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
-  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
-
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel.
-\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
-
-\bibitem{Rushby}
-J.~Rushby.
-\newblock The Ontological Argument in PVS. 
-\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St. Petersburg, Russia, 2013.
-
-\bibitem{ScottNotes}
-D.~Scott.
-\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
-\newblock In  \cite{sobel2004logic}, 2004.
-
-\bibitem{sobel2004logic}
-J.H.~Sobel.
-\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambridge U. Press, 2004.
-
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller.
-\newblock Automated reasoning in higher-order logic using the {TPTP THF}
-  infrastructure.
-\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
-
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
-\end{thebibliography}
+\bibliography{Bibliography}
 

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
update of draft paper

--- 2014-12-11 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
 \author{
-  Christoph Benzm\"{u}ller\inst{1}\thanks{Supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
+  Christoph Benzm\"{u}ller\inst{1}\thanks{Supported by the German Research Foundation (DFG) under grant BE 2501/9-1.} 
   \and 
@@
   perspective, are illustrated here with the successful formalization
-  of Gödel's ontological argument.
+  of G\"odel's ontological argument.
 \end{abstract}
@@
 
-Another recently explored possibility is the
-embedding of first-order and even higher-order modal logics into classical
-higher-order logics \cite{J23,B9,J22}, 
-for which existing higher-order automated theorem
-provers \cite{LEO-II,Satallax} exist. The embedding approach is flexible,
-because various modal logics (even with multiple modalities or varying/cumulative domain quantifiers) can be easily supported by stating their
-characteristic axioms. Moreover, the approach is relatively simple to
-implement, because it does not require any modification in the source
-code of the higher-order prover. The prover can be used as is, and
-only the input files provided to the prover must be especially
-encoded. Furthermore, the efficacy and efficiency of the embedding
-approach has been confirmed in philosophically interesting and
+Another recently explored possibility is the embedding of first-order
+and even higher-order modal logics into classical higher-order logics
+\cite{J23,B9,J22}, for which existing higher-order automated theorem
+provers \cite{LEO-II,Satallax} exist. The embedding approach is
+flexible, because various modal logics (even with multiple modalities
+or varying/cumulative domain quantifiers) can be easily supported by
+stating their characteristic axioms. Moreover, the approach is
+relatively simple to implement, because it does not require any
+modification in the source code of the higher-order prover. The prover
+can be used as is, and only the input files provided to the prover
+must be especially encoded. Furthermore, the efficacy and efficiency
+of the embedding approach has been confirmed in philosophically
 relevant benchmarks \cite{FormalTheologyRepository}. These qualities
@@
 other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
-\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal and novelty
+\texttt{HOL Light} \cite{HOLLight}.}. The main goal and novelty
 of this paper is to study this question. Our answer is positive.
@@
 In order to illustrate the use of the implemented embedding, we show
-here the formalization of Scott's version \cite{ScottNotes} of Gödel's
+here the formalization of Scott's version \cite{ScottNotes} of G\"odel's
 ontological argument for God's existence (in Section \ref{sec:proof}).
@@
 a lot of attention in the last 15 years, because of the discovery of
-the modal collapse \cite{ToDo}. This proof lies in the center of a
+the modal collapse \cite{sobel,sobel2004logic}. This proof lies in the center of a
 vast and largely unexplored application domain for automated and
@@
 verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
-first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
+first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. G\"odel's
 argument was automatically verified in our previous work on embedding-
@@
 verification, artificial intelligence
-agent technologies, law and linguistics. \marginpar{BrunoToChris: Do we need citations for these applications?} 
+agent technologies, law and linguistics (cf.~\cite{ModalLogic} and the references therein).
 Therefore, the main contribution described in this paper -- convenient techniques for leveraging a powerful proof assistant such as \Coq for interactive reasoning for modal logics -- may serve as a starting point for
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Definition mnot (p: o)(w: i) := ~ (p w).
+\Notation "m~  p" := (mnot p) (at level 74, right associativity).
+
+\Definition mand (p q:o)(w: i) := (p w) /\bslash (q w).
+\Notation "p m/\bslash q" := (mand p q) (at level 79, right associativity).
+
+\Definition mor (p q:o)(w: i) := (p w) \bslash\slash (q w).
+\Notation "p m\bslash\slash q" := (mor p q) (at level 79, right associativity).
+
+\Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
+\Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).
+
+\Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
+\Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
+
 \Definition mequal (x y: u)(w: i) := x = y.
 \Notation "x m= y" := (mequal x y) (at level 99, right associativity).
-
-\Definition mnot (p: o)(w: i) := ~ (p w).
-\Notation "m~  p" := (mnot p) (at level 74, right associativity).
-
-\Definition mand (p q:o)(w: i) := (p w) /\bslash (q w).
-\Notation "p m/\bslash q" := (mand p q) (at level 79, right associativity).
-
-\Definition mor (p q:o)(w: i) := (p w) \bslash\slash (q w).
-\Notation "p m\bslash\slash q" := (mor p q) (at level 79, right associativity).
-
-\Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
-\Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).
-
-\Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
-\Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
 \end{Verbatim}
@@
 goal) to simpler subgoals, in a bottom-up manner. The simplest tactics
-can be regarded as rules of a natural deduction calculus\footnote{The
+can be regarded as rules of a natural deduction (ND) calculus\footnote{The
 underlying proof system of \Coq (the Calculus of Inductive
@@
 \begin{calculus}
-{A (classical)  The natural deduction calculus}
+{Rules of a (classical) ND calculus}
 {fig:NaturalDeduction}
@@
 
-\vspace{1em}
+%\vspace{1em}
 
@@
 
-\vspace{2em}
+\vspace{1em}
 
@@
 
+Rules for $\alpha\beta\eta$-equality and axioms (or rules) for extensionality are omitted here since they are not important for the rest of the paper.
+For a full, sound and Henkin-complete, classical higher-order ND calculus see \cite{HOLND}.
 \end{small}
 
-\vspace{1em}
+%\vspace{1em}
 \end{calculus}
@@
 deduction point of view, they give rise to the inference
-rules\footnote{The natural deduction calculus with the rules from
+rules\footnote{The ND calculus with the rules from
 Figures \ref{fig:NaturalDeduction} and \ref{fig:NDK} is sound and
@@
 \ref{fig:NaturalDeduction} extended with a necessitation rule and the
-modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Therefore,
-assuming that the calculus from \ref{fig:NaturalDeduction} is sound
-and Henkin-complete for classical higher-order logic, the additional
+modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Starting from a sound and Henkin-complete ND calculus for classical higher-order logic (cf. Figure~\ref{fig:NaturalDeduction}), the additional
 modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete
@@
 
-\s\s\s\s
+\s\s
 \infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
-\s\s\s\s
+\s
 \infer[\nec_E]{w: \fbox{ \infer*{}{A} } }{\nec A}
-
-\vspace{2em}
-
-\s\s\s\s
+%\vspace{2em}
+\s
 \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-\s\s\s\s
+\s
 \infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
@@
 \noindent 
-The proof of this lemma is displayed as a natural deduction
+The proof of this lemma is displayed as a ND
 proof in  Figure \ref{fig:mp_dia}. As expected, \Coq's basic tactics
@@
 \centering
-\label{fig:mp_dia}
 \includegraphics[width=0.5\textwidth]{ND-(mp_dia).png}
-\caption{Natural deduction proof of \texttt{mp\_dia}}
+\caption{ND proof of \texttt{mp\_dia}\label{fig:mp_dia}}
 \end{figure}
@@
 a convenient and intuitive correspondence between proof scripts and 
-modal natural deduction proofs would be missing.
+modal ND proofs would be missing.
 
@@
 but also for larger developments, we include here a full and detailed
-formalization of Gödel's ontological argument.
+formalization of G\"odel's ontological argument.
 
@@
 in order to formalize exactly Scott's version and not some arbitrary
-version found automatically, automation would have to be heavily
-limited anyway. Furthermore, the deliberate preference for simple
-tactics (mostly \emph{intro}, \emph{apply} and the modal tactics
-described in Section \ref{sec:Tactics}) results in proof scripts that
-closely correspond to common natural deduction proofs. This hopefully makes
-the formalization more accessible to those who are not experts in
-\Coq's tactics but are nevertheless interested in G\"odel's proof.
-
-Gödel's proof requires \Coq's classical logic libraries as well as the
+version found automatically\footnote{The proofs found automatically
+  by the above provers indeed differ from the one presented here:
+  e.g., the strong S5 principle used below (and by Scott) is not needed; the
+  ATP proofs only rely on axiom B.}, automation would have to be
+heavily limited anyway. Furthermore, the deliberate preference for
+simple tactics (mostly \emph{intro}, \emph{apply} and the modal
+tactics described in Section \ref{sec:Tactics}) results in proof
+scripts that closely correspond to common ND
+proofs. This hopefully makes the formalization more accessible to
+those who are not experts in \Coq's tactics but are nevertheless
+interested in G\"odel's proof.
+
+G\"odel's proof requires \Coq's classical logic libraries as well as the
 \texttt{Modal} library developed by us and described in Sections
@@
 \noindent
-Gödel's theory has a single higher-order constant, \texttt{Positive}, 
+G\"odel's theory has a single higher-order constant, \texttt{Positive}, 
 which ought to hold for properties considered \emph{positive} in a moral sense.
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\com{(* Axiom A1:} 
+\com{(* Axiom A1 (divided into two directions):} 
 \com{   either a property or its negation is positive, but not both *)}
@@
 lemmas and theorems in Sections \ref{sec:Embedding},
-\ref{sec:SimpleExamples} and \ref{sec:BeyondK} could have been proven
+\ref{sec:SimpleExamples} and \ref{sec:BeyondK} (but not \ref{sec:proof}) could have been proven
 automatically using \Coq's \texttt{firstorder} tactic. The
@@
 
-The theological implications of the verified correctness of Gödel's
+The theological implications of the verified correctness of G\"odel's
 proof certainly depend on a critical discussion of the underlying
@@
 
-\paragraph{Acknowledgements:} We thank Cedric Auger and Laurent Théry, 
+\paragraph{Acknowledgements:} We thank Cedric Auger and Laurent Th\'ery, 
 for their answers to our questions about Ltac in the Coq-Club mailing-list.
@@
 
+\bibitem{HOLND}
+C.~Benzm{\"u}ller, C.~Brown, and M.~Kohlhase.
+\newblock Higher-Order Semantics and Extensionality. 
+\newblock {\em Journal of Symbolic Logic} 69(4):1027-1088, 2004.
+
 \bibitem{JensOtten}
-C.~Benzm{\"u}lller, J. Otten, and Th. Raths.
+C.~Benzm{\"u}ller, J. Otten, and Th. Raths.
 \newblock Implementing and Evaluating Provers for First-order Modal
@@
 
-\bibitem{Zalta}
-P.E.~Oppenheimera and E.N.~Zalta.
-\newblock A Computationally-Discovered Simplification of the Ontological Argument.
-\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.
-
 \bibitem{Satallax}
@@
 
+\bibitem{CIC}
+Th.~Coquand and G.~Huet. 
+\newblock The Calculus of Constructions. 
+\newblock {\em Information and Computation} 76 (2-3). 1988.
+
 \bibitem{ContemporaryBibliography}
@@
 
+\bibitem{HOLLight}
+J.~Harrison.
+\newblock HOL Light: An overview.
+\newblock In {\em Proceedings of TPHOLs 2009}. Number 5674 in LNCS, pp. 60-66. Springer, 2009.
+
 \bibitem{Hazen}
 A.P.~Hazen.
-\newblock On g\"odel's ontological proof.
+\newblock On G\"odel's ontological proof.
 \newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
@@
 
+\bibitem{Zalta}
+P.E.~Oppenheimera and E.N.~Zalta.
+\newblock A Computationally-Discovered Simplification of the Ontological Argument.
+\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.
+
 \bibitem{Rushby}
@@
 \newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{sobel}
+J.H.~Sobel. 
+\newblock G\"odel's ontological proof. 
+\newblock {\em In On Being and Saying. Essays for Richard Cartwright}, pages 241-261. MIT Press, 1987.
 

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
update of draft paper

@@
 
-%\footnotesize
 \bibliographystyle{plain}

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
update of draft paper

@@
 for modal logics are often limited to propositional, quantifier-free,
-fragments \cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers} or tailored to particular
-modal logics and their applications \cite{ToDo:StephanMerz}; 
+fragments or tailored to particular
+modal logics and their applications \cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers,Coalescing}; 
 first-order automated deduction techniques based on tableaux, sequent
@@
 automated reasoning in higher-order modal logics will clearly be
-useful also outside philosophy; the range of potential applications is
-in fact very wide.
+useful also outside philosophy; the range of potential applications is very wide.
 

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Add missing references

--- 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
 \author{
-  Christoph Benzm\"{u}ller\inst{1}\thanks{Supported by the German Research Foundation (DFG) under grant BE 2501/9-1.} 
+  Christoph Benzm\"{u}ller\inst{1}\thanks{Supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
   \and 
@@
   perspective, are illustrated here with the successful formalization
-  of G\"odel's ontological argument.
+  of Gödel's ontological argument.
 \end{abstract}
@@
 
-Another recently explored possibility is the embedding of first-order
-and even higher-order modal logics into classical higher-order logics
-\cite{J23,B9,J22}, for which existing higher-order automated theorem
-provers \cite{LEO-II,Satallax} exist. The embedding approach is
-flexible, because various modal logics (even with multiple modalities
-or varying/cumulative domain quantifiers) can be easily supported by
-stating their characteristic axioms. Moreover, the approach is
-relatively simple to implement, because it does not require any
-modification in the source code of the higher-order prover. The prover
-can be used as is, and only the input files provided to the prover
-must be especially encoded. Furthermore, the efficacy and efficiency
-of the embedding approach has been confirmed in philosophically
+Another recently explored possibility is the
+embedding of first-order and even higher-order modal logics into classical
+higher-order logics \cite{J23,B9,J22}, 
+for which existing higher-order automated theorem
+provers \cite{LEO-II,Satallax} exist. The embedding approach is flexible,
+because various modal logics (even with multiple modalities or varying/cumulative domain quantifiers) can be easily supported by stating their
+characteristic axioms. Moreover, the approach is relatively simple to
+implement, because it does not require any modification in the source
+code of the higher-order prover. The prover can be used as is, and
+only the input files provided to the prover must be especially
+encoded. Furthermore, the efficacy and efficiency of the embedding
+approach has been confirmed in philosophically interesting and
 relevant benchmarks \cite{FormalTheologyRepository}. These qualities
@@
 other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
-\texttt{HOL Light} \cite{HOLLight}.}. The main goal and novelty
+\texttt{HOL-Light} \cite{HOLLight}, as well.} \cite{Coq}. The main goal and novelty
 of this paper is to study this question. Our answer is positive.
@@
 In order to illustrate the use of the implemented embedding, we show
-here the formalization of Scott's version \cite{ScottNotes} of G\"odel's
+here the formalization of Scott's version \cite{ScottNotes} of Gödel's
 ontological argument for God's existence (in Section \ref{sec:proof}).
@@
 a lot of attention in the last 15 years, because of the discovery of
-the modal collapse \cite{sobel,sobel2004logic}. This proof lies in the center of a
+the modal collapse \cite{Sobel}. This proof lies in the center of a
 vast and largely unexplored application domain for automated and
@@
 verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
-first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. G\"odel's
+first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
 argument was automatically verified in our previous work on embedding-
@@
 verification, artificial intelligence
-agent technologies, law and linguistics (cf.~\cite{ModalLogic} and the references therein).
+agent technologies, law and linguistics. \marginpar{BrunoToChris: Do we need citations for these applications?} 
 Therefore, the main contribution described in this paper -- convenient techniques for leveraging a powerful proof assistant such as \Coq for interactive reasoning for modal logics -- may serve as a starting point for
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Definition mequal (x y: u)(w: i) := x = y.
+\Notation "x m= y" := (mequal x y) (at level 99, right associativity).
+
 \Definition mnot (p: o)(w: i) := ~ (p w).
@@
 \Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
-
-\Definition mequal (x y: u)(w: i) := x = y.
-\Notation "x m= y" := (mequal x y) (at level 99, right associativity).
 \end{Verbatim}
@@
 goal) to simpler subgoals, in a bottom-up manner. The simplest tactics
-can be regarded as rules of a natural deduction (ND) calculus\footnote{The
+can be regarded as rules of a natural deduction calculus\footnote{The
 underlying proof system of \Coq (the Calculus of Inductive
@@
 \begin{calculus}
-{Rules of a (classical) ND calculus}
+{A (classical)  The natural deduction calculus}
 {fig:NaturalDeduction}
@@
 
-%\vspace{1em}
+\vspace{1em}
 
@@
 
-\vspace{1em}
+\vspace{2em}
 
@@
 
-Rules for $\alpha\beta\eta$-equality and axioms (or rules) for extensionality are omitted here since they are not important for the rest of the paper.
-For a full, sound and Henkin-complete, classical higher-order ND calculus see \cite{HOLND}.
 \end{small}
 
-%\vspace{1em}
+\vspace{1em}
 \end{calculus}
@@
 deduction point of view, they give rise to the inference
-rules\footnote{The ND calculus with the rules from
+rules\footnote{The natural deduction calculus with the rules from
 Figures \ref{fig:NaturalDeduction} and \ref{fig:NDK} is sound and
@@
 \ref{fig:NaturalDeduction} extended with a necessitation rule and the
-modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Starting from a sound and Henkin-complete ND calculus for classical higher-order logic (cf. Figure~\ref{fig:NaturalDeduction}), the additional
-modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete
-for modal higher-order logic \textbf{K}.} shown in Figure
+modal axiom K \cite{ModalND}. Therefore,
+assuming that the calculus from \ref{fig:NaturalDeduction} is sound
+and Henkin-complete for classical higher-order logic \cite{NDHenkinCompleteness}, 
+the additional modal rules in Figure \ref{fig:NDK} make it sound and 
+Henkin-complete for modal higher-order logic \textbf{K}.} shown in Figure
 \ref{fig:NDK}. The labels that name boxes in those inference rules are
@@
 
-\s\s
+\s\s\s\s
 \infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
-\s
+\s\s\s\s
 \infer[\nec_E]{w: \fbox{ \infer*{}{A} } }{\nec A}
-%\vspace{2em}
-\s
+
+\vspace{2em}
+
+\s\s\s\s
 \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-\s
+\s\s\s\s
 \infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
@@
 \noindent 
-The proof of this lemma is displayed as a ND
+The proof of this lemma is displayed as a natural deduction
 proof in  Figure \ref{fig:mp_dia}. As expected, \Coq's basic tactics
@@
 \centering
+\label{fig:mp_dia}
 \includegraphics[width=0.5\textwidth]{ND-(mp_dia).png}
-\caption{ND proof of \texttt{mp\_dia}\label{fig:mp_dia}}
+\caption{Natural deduction proof of \texttt{mp\_dia}}
 \end{figure}
@@
 a convenient and intuitive correspondence between proof scripts and 
-modal ND proofs would be missing.
+modal natural deduction proofs would be missing.
 
@@
 but also for larger developments, we include here a full and detailed
-formalization of G\"odel's ontological argument.
+formalization of Gödel's ontological argument.
 
@@
 in order to formalize exactly Scott's version and not some arbitrary 
-version found automatically\footnote{The proofs found automatically
-  by the above provers indeed differ from the one presented here:
-  e.g., the strong S5 principle used below (and by Scott) is not needed; the
-  ATP proofs only rely on axiom B.}, automation would have to be
-heavily limited anyway. Furthermore, the deliberate preference for
-simple tactics (mostly \emph{intro}, \emph{apply} and the modal
-tactics described in Section \ref{sec:Tactics}) results in proof
-scripts that closely correspond to common ND
-proofs. This hopefully makes the formalization more accessible to
-those who are not experts in \Coq's tactics but are nevertheless
-interested in G\"odel's proof.
-
-G\"odel's proof requires \Coq's classical logic libraries as well as the
+version found automatically, automation would have to be heavily
+limited anyway. Furthermore, the deliberate preference for simple
+tactics (mostly \emph{intro}, \emph{apply} and the modal tactics
+described in Section \ref{sec:Tactics}) results in proof scripts that
+closely correspond to common natural deduction proofs. This hopefully makes
+the formalization more accessible to those who are not experts in
+\Coq's tactics but are nevertheless interested in G\"odel's proof.
+
+Gödel's proof requires \Coq's classical logic libraries as well as the
 \texttt{Modal} library developed by us and described in Sections
@@
 \noindent
-G\"odel's theory has a single higher-order constant, \texttt{Positive}, 
+Gödel's theory has a single higher-order constant, \texttt{Positive}, 
 which ought to hold for properties considered \emph{positive} in a moral sense.
@@
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\com{(* Axiom A1 (divided into two directions):} 
+\com{(* Axiom A1:} 
 \com{   either a property or its negation is positive, but not both *)}
@@
 lemmas and theorems in Sections \ref{sec:Embedding},
-\ref{sec:SimpleExamples} and \ref{sec:BeyondK} (but not \ref{sec:proof}) could have been proven
+\ref{sec:SimpleExamples} and \ref{sec:BeyondK} could have been proven
 automatically using \Coq's \texttt{firstorder} tactic. The
@@
 
-The theological implications of the verified correctness of G\"odel's
+The theological implications of the verified correctness of Gödel's
 proof certainly depend on a critical discussion of the underlying
@@
 automated reasoning in higher-order modal logics will clearly be
-useful also outside philosophy; the range of potential applications is very wide.
-
-\paragraph{Acknowledgements:} We thank Cedric Auger and Laurent Th\'ery, 
+useful also outside philosophy; the range of potential applications is
+in fact very wide.
+
+\paragraph{Acknowledgements:} We thank Cedric Auger and Laurent Théry, 
 for their answers to our questions about Ltac in the Coq-Club mailing-list.
@@
 \bibliographystyle{plain}
-%\bibliography{Bibliography}
-\begin{thebibliography}{10}
-
-\bibitem{Adams}
-R.M.~Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1995.
-
-\bibitem{AndersonGettings}
-A.C.~Anderson and M.~Gettings.
-\newblock G\"odel ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
-
-\bibitem{HOLND}
-C.~Benzm{\"u}ller, C.~Brown, and M.~Kohlhase.
-\newblock Higher-Order Semantics and Extensionality. 
-\newblock {\em Journal of Symbolic Logic} 69(4):1027-1088, 2004.
-
-\bibitem{JensOtten}
-C.~Benzm{\"u}ller, J. Otten, and Th. Raths.
-\newblock Implementing and Evaluating Provers for First-order Modal
-Logics. 
-\newblock In {\em ECAI 2012}, pages 163-168. IOS Press, 2012.
-
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C.~Paulson.
-\newblock Exploring properties of normal multimodal logics in simple type
-  theory with {LEO-II}.
-\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
-
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C.~Paulson.
-\newblock Quantified multimodal logics in simple type theory.
-\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
-  7(1):7--20, 2013.
-
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
-\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
-  logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
-  162--170. Springer, 2008.
-
-\bibitem{arXiv}
-C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
-\newblock Formalization, Mechanization and Automation of G{\"o}del's Proof of God's Existence.
-\newblock {\em {arXiv}}:1308.4526, 2013.
-
-\bibitem{AFP}
-C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
-\newblock G{\"o}del's God in Isabelle/HOL. 
-\newblock {\em Archive of Formal Proofs}, 2013.
-
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran.
-\newblock {\em {Interactive Theorem Proving and Program Development}}.
-\newblock Springer, 2004.
-
-\bibitem{ModalLogic}
-P.~Blackburn, J.v.~Benthem, and F.~Wolter (eds.), \newblock {\em {Handbook of Modal Logic}}. Elsevier, 2006.
-
-\bibitem{Sledgehammer}
-J.C.~Blanchette, S.~B\"ohme, and L.C.~Paulson.
-\newblock Extending {Sledgehammer} with {SMT} solvers.
-\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
-
-\bibitem{Nitpick}
-J.C.~Blanchette and T.~Nipkow.
-\newblock Nitpick: A counterexample generator for higher-order logic based on a
-  relational model finder.
-\newblock In {\em Proc. of ITP 2010}, no.~6172 in LNCS, pages 131--146.
-  Springer, 2010.
-
-\bibitem{Satallax}
-C.E.~Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
-  Springer, 2012.
-
-\bibitem{CIC}
-Th.~Coquand and G.~Huet. 
-\newblock The Calculus of Constructions. 
-\newblock {\em Information and Computation} 76 (2-3). 1988.
-
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary bibligraphy on the ontological proof
-  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
-
-\bibitem{Fitting}
-M.~Fitting.
-\newblock {\em {Types, Tableaux and G\"odel's God}}.
-\newblock Kluver Academic Press, 2002.
-
-\bibitem{Labels}
-D.M.~Gabbay. {\em Labelled Deductive Systems}. Clarendon Press, 1996.
-
-\bibitem{Goedel1970}
-K.~G\"odel.
-\newblock Ontological proof.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1970.
-
-\bibitem{GoedelNotes}
-K.~G\"odel.
-\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
-\newblock In  \cite{sobel2004logic}, 2004.
-
-\bibitem{HOLLight}
-J.~Harrison.
-\newblock HOL Light: An overview.
-\newblock In {\em Proceedings of TPHOLs 2009}. Number 5674 in LNCS, pp. 60-66. Springer, 2009.
-
-\bibitem{Hazen}
-A.P.~Hazen.
-\newblock On G\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
-
-\bibitem{Hurd03first-orderproof}
-J.~Hurd.
-\newblock First-order proof tactics in higher-order logic theorem provers.
-\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
-  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
-
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel.
-\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
-
-\bibitem{Zalta}
-P.E.~Oppenheimera and E.N.~Zalta.
-\newblock A Computationally-Discovered Simplification of the Ontological Argument.
-\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.
-
-\bibitem{Rushby}
-J.~Rushby.
-\newblock The Ontological Argument in PVS. 
-\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St. Petersburg, Russia, 2013.
-
-\bibitem{ScottNotes}
-D.~Scott.
-\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
-\newblock In  \cite{sobel2004logic}, 2004.
-
-\bibitem{sobel}
-J.H.~Sobel. 
-\newblock G\"odel's ontological proof. 
-\newblock {\em In On Being and Saying. Essays for Richard Cartwright}, pages 241-261. MIT Press, 1987.
-
-\bibitem{sobel2004logic}
-J.H.~Sobel.
-\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambridge U. Press, 2004.
-
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller.
-\newblock Automated reasoning in higher-order logic using the {TPTP THF}
-  infrastructure.
-\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
-
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
-\end{thebibliography}
+\bibliography{Bibliography}
 

### 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Merge conflicting hunks

--- 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
 a lot of attention in the last 15 years, because of the discovery of
-the modal collapse \cite{sobel,sobel2004logic}. This proof lies in the center of a
+the modal collapse \cite{Sobel,sobel2004logic}. This proof lies in the center of a
 vast and largely unexplored application domain for automated and
@@
 Rules for $\alpha\beta\eta$-equality and axioms (or rules) for extensionality are omitted here since they are not important for the rest of the paper.
-For a full, sound and Henkin-complete, classical higher-order ND calculus see \cite{HOLND}.
+For a full, sound and Henkin-complete, classical higher-order ND calculus, see \cite{HOLND}.
 \end{small}
@@
 \ref{fig:NaturalDeduction} extended with a necessitation rule and the
-modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Starting from a sound and Henkin-complete ND calculus for classical higher-order logic (cf. Figure~\ref{fig:NaturalDeduction}), the additional
+modal axiom K \cite{ModalND}. Starting from a sound and Henkin-complete ND calculus for classical higher-order logic (cf. Figure~\ref{fig:NaturalDeduction}), the additional
 modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete

### 2015-01-02 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Rename folder with paper about Hajek and Anderson controversy
Fix calculus

@@
 
-% \vspace{2em}
+\vspace{2em}
 
-% \s\s\s\s
-% \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-% \s\s\s\s\s
-% \infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
+\s\s\s\s
+\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
+\s\s\s\s
+\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
 
@@
 \textbf{eigen-box condition:}\\ 
-$\omega$ must be a fresh name for a box. \\
+$\nec_I$ and $\pos_E$ are \emph{strong} modal rules: \\
+$\omega$ must be a fresh name for the box they access \\ 
+(in analogy to the eigen-variable condition for strong quantifier rules). \\
+Every box must be accessed by \emph{exactly one} strong modal inference. \\
 \vspace{0.5em}
 \textbf{boxed assumption condition:} \\
-assumptions should be discharged within the box where they are made.
+assumptions should be discharged within the box where they are created.
 \end{center}
@@
 
-Adding the modal rules results in a calculus that is suitable for the basic normal modal logic K.
+Adding the modal rules results in a calculus that is suitable for the basic normal modal logic \textbf{K}.
 In other words, {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
-
-The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
-From the point of view of a \emph{possible worlds} semantics, rigidity means that their interpretation is independent of the world in which they are being interpreted.
-Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
-Nevertheless, it should be noted that its adequacy has already been contested \cite{fitting}.
-
-ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
 
@@
 \end{proof}
+
+Without the restriction that every box must be accessed by exactly one strong modal inference, the calculus would be unsound for the modal logic \textbf{K}, because the formula $\all \psi.
+(\nec \psi \imp \pos \psi)$ would be derivable although it is not valid in \textbf{K}.
+
+ToDo: show example of unsound derivation.
+
+The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
+From the point of view of a \emph{possible worlds} semantics, rigidity means that their interpretation is independent of the world in which they are being interpreted.
+Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
+Nevertheless, it should be noted that its adequacy has already been contested \cite{fitting}.
+
+ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
 

### 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
update of draft paper
Conflicts:
	Papers/InProgress/ModalLogicCoq/ModalCoq.pdf
	Papers/InProgress/ModalLogicCoq/ModalCoq.tex

--- 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
 automated reasoning in higher-order modal logics will clearly be
-useful also outside philosophy; the range of potential applications is very wide.
+useful also outside philosophy; the range of potential applications is
+in fact very wide.
 

### 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve and compress language
Conflicts:
	Papers/InProgress/ModalLogicCoq/ModalCoq.pdf

--- 2014-12-12 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
 by interacting with a proof assistant such as \Coq\footnote{The \Coq
-proof assistant was chosen merely because of the authors' greater
+proof assistant was chosen because of the authors' greater
 familiarity with the tactic language of this system. Nevertheless, the
-techniques presented and discussed here are likely to be useful for
-other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
-\texttt{HOL Light} \cite{HOLLight}.}. The main goal and novelty
+techniques presented here are likely to be useful for
+other proof assistants (e.g \texttt{Isabelle} \cite{Isabelle},
+\texttt{HOL-Light} \cite{HOLLight}).}. The main goal and novelty
 of this paper is to study this question. Our answer is positive.
@@
 resulting infra-structure for modal reasoning within \Coq (as
-described in Section \ref{sec:Embedding}) provides a user experience where
+described in Sections \ref{sec:Embedding} and \ref{sec:Tactics}) provides a user experience where
 modalities can be handled transparently and straightforwardly.
@@
 
-The simpler ontological argument of Anselm has been automatically
+The ontological argument of Anselm has been automatically
 verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
-first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. G\"odel's
-argument was automatically verified in our previous work on embedding-
-based fully automated modal theorem proving \cite{arXiv,AFP}. This
-paper presents the first fully interactive and detailed formalization
-of this proof in a proof assistant. Hence we here reveal the proof
-structure on a cognitively adequate level of
-detail. Such detailed proof structure has been hidden in our
+first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. 
+In comparison, our
+contribution stands out with its surprising technical 
+simplicity and elegance, despite the greater complexity
+of G\"odel's argument.
+
+G\"odel's argument was automatically verified in our previous work on fully automated modal theorem proving based on embedding \cite{arXiv,AFP}. 
+This paper presents the first fully interactive and detailed formalization
+of this proof in a proof assistant. The proof
+structure, which has been hidden in our
 other papers on the subject due to the use of automated theorem
-provers. In comparison to \cite{Rushby,Zalta}, however, our
-contribution stands out with its surprising technical 
-simplicity and elegance.
+provers, is revealed here on a cognitively adequate level of
+detail.
 
 In addition to philosophy, propositional and quantified modal logics 
-have (potential) applications in various other fields as well, including, for instance, 
+have (potential) applications in various other fields, including, for instance, 
 verification, artificial intelligence

### 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Compress bibliography
Conflicts:
	Papers/InProgress/ModalLogicCoq/ModalCoq.pdf

@@
 fragments or tailored to particular
-modal logics and their applications \cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers,Coalescing}; 
+modal logics and their applications 
+%\cite{CoLoSS1,CoLoSS2,ConditionalTableaux,COOL,ModalProvers,Coalescing}; 
+\cite{ModalProvers}; 
 first-order automated deduction techniques based on tableaux, sequent
@@
 and even higher-order modal logics into classical higher-order logics
-\cite{J23,B9,J22}, for which existing higher-order automated theorem
+\cite{J23,B9}, for which existing higher-order automated theorem
 provers \cite{LEO-II,Satallax} exist. The embedding approach is
@@
 abstract ontological arguments are an old tradition in philosophy and
-theology. G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern
+theology. G\"{o}del's proof 
+%\cite{Goedel1970,GoedelNotes} 
+\cite{Goedel1970} 
+is a modern
 culmination of this tradition, following particularly the footsteps of
@@
 commented on his proof (cf.
-\cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+%\cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).
+\cite{sobel2004logic,Fitting,ContemporaryBibliography}).
 The formalization shown in this Section aims at being as similar as
@@
 \cite{AFP}, where automation via \texttt{Metis}
-\cite{Hurd03first-orderproof} and \texttt{Sledgehammer}
-\cite{Sledgehammer} using tools such as \texttt{Nitpick}
-\cite{Nitpick}, \texttt{LEO-II} \cite{LEO-II} and \texttt{Satallax}
+%\cite{Hurd03first-orderproof} 
+and \texttt{Sledgehammer}
+%\cite{Sledgehammer} 
+using tools such \texttt{LEO-II} \cite{LEO-II} and \texttt{Satallax}
 \cite{Satallax} has been successfully employed, the formalization in
@@
 The infrastructure that we have implemented for interactive and
-automated reasoning in higher-order modal logics will clearly be
+automated reasoning in higher-order modal logics is clearly
 useful also outside philosophy; the range of potential applications is very wide.
@@
 
+\footnotesize
 \bibliographystyle{plain}

### 2015-01-26 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix calculus
Update GodProof-ND.tex
updated proof

@@
 
+\section{A distribution principle and axiom B}
+
+To prove G\"odel's ontological proof in natural deduction we will prove that a distribution principle is derivable and state Brouwer's reduction theorem.
+
+\begin{lemma}\label{DP}
+The distribution principle 
+$$\nec (A\imp B)\imp(\pos A\imp \pos B)$$
+is provable in our system K of modal logic. 
+\end{lemma}
+
+\begin{proof}\hfill
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \dashedLine\RightLabel{2}
+\UIC{$\nec (A\imp B)$}\RightLabel{$\nec_E$}
+\UIC{$\omega: A\imp B$}
+\AXC{$ $} \dashedLine\RightLabel{1}
+\UIC{$\pos A$}\RightLabel{$\pos_E$}
+\UIC{$\omega: A$}\RightLabel{$\imp_E$}
+\BIC{$\omega: B$} \RightLabel{$\pos_I$}
+\UIC{$\pos B$}\RightLabel{$\imp_I^1$}
+\UIC{$\pos A\imp \pos B$}\RightLabel{$\imp_I^2$}
+\UIC{$\nec (A\imp B)\imp(\pos A\imp \pos B)$}
+\end{prooftree}
+\end{small}
+
+\end{proof}
+
+Brouwer's reduction theorem is a version of axiom B for modal logic and a property that enforces symmetry as a geometric property for the frame of a Kripke sematics for modal logic. \citep{negri}
+%See the following for a discussion about axiom B: http://plato.stanford.edu/entries/logic-modal/
+
 \section{Scott's Proof in Natural Deduction}
@@
 \section{New Proof}
-
 \setcounter{axiom}{0}
@@
 \setcounter{corollary}{0}
-
 \subsection{Possibly, God Exists}
-
 \begin{axiom}
@@
 Either a property or its negation is positive, but not both:
+Axiom schema, for all $\varphi$:
 $$
-\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+P(\neg \varphi) \biimp \neg P(\varphi)
 $$
 \end{axiom}
-
 \begin{axiom}
@@
 A property necessarily implied by a positive property is positive:
+Axiom schema, for all $\varphi$ and $\psi$: 
 $$
-\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
+(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)
 $$
 \end{axiom}
-
 \begin{theorem}
@@
 Positive properties are possibly exemplified:
+Axiom schema, for all $\varphi$:
 $$
-\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
+P(\varphi) \imp \pos \ex x.\varphi(x)
 $$
@@
 \begin{proof} \hfill
+
 \begin{prooftree}
-        \AXC{Axiom \ref{A2}} \dashedLine
-        \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
-        \UIC{$ \all \psi.[(P(\rho) \wedge \nec \all x.[\rho(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
-        \UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \doubleLine
-        \UIC{$(P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp P(\neg \rho)$}
-                        \AXC{Axiom \ref{A1}} \dashedLine
-                        \UIC{$\all \varphi.[ P(\neg \varphi) \biimp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
-                        \UIC{$ P(\neg \rho) \biimp \neg P(\rho) $} \doubleLine
-                 \BIC{$ (P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp \neg P(\rho) $} \doubleLine
-                 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
-                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
+
+\AXC{$[P(\rho)]^3$} 
+\AXC{$[\nec \neg \ex x.\rho(x) ]^2$} \RightLabel{$\nec_E$}
+\UIC{$\neg \ex x.\rho(x) $} 
+\AXC{$[\rho(x)]^1$} \RightLabel{$\ex_I$}
+\UIC{$\ex x.\rho(x) $} \RightLabel{$\imp_E$}
+\BIC{$ \bot $}\RightLabel{$\imp_I, 1$}
+\UIC{$ \neg \rho(x) $} \RightLabel{$\imp_I$}
+\UIC{$ \rho(x) \imp \neg \rho(x) $} \RightLabel{$\all_I$}
+\UIC{$ \all x.(\rho(x) \imp \neg \rho(x)) $}  \RightLabel{$\nec_I$}
+\UIC{$ \nec \all x.(\rho(x) \imp \neg \rho(x)) $}\RightLabel{$\wedge_I$}
+\BIC{$ P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]$}
+\AXC{Axiom \ref{A2} for $\rho$ and $\neg\rho$} \dashedLine
+\UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \RightLabel{$\imp_E$}
+\BIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
 \end{prooftree}
+
+\begin{prooftree}
+\AXC{$ $} \dashedLine
+\UIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
+\AXC{Half of axiom \ref{A1}} \dashedLine
+\UIC{$ P(\neg \rho) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
+\BIC{$\neg P(\rho) $} 
+\AXC{$[P(\rho)]^3$} 
+\BIC{$\bot$} \RightLabel{$\imp_I, 2$}
+\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \dashedLine \RightLabel{definition of $\pos$}
+\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I, 3$}
+\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
+
+\end{prooftree}
+
 \end{proof}
-
 \begin{definition}
@@
 \end{axiom}
-
 \begin{corollary}
@@
 \UIC{$P(G)$}
-                 \AXC{Theorem \ref{T1}} \dashedLine
-                 \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$} \RightLabel{$\all_E $}
+\AXC{Theorem \ref{T1} for G} \dashedLine
 \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
@@
 
-\subsection{Being God is an essence of any God}
-
-\begin{axiom}
-\label{A4}
-Positive properties are necessarily positive:
-$$
-\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
-$$
-\end{axiom}
+\subsection{If God exists, then God necessarily exists}
 
@@
 
-\begin{theorem}
-\label{T2}
-Being God-like is an essence of any God-like being:
-$$
-\all y.[G(y) \imp \ess{G}{y}]
-$$
-\end{theorem}
-\begin{proof}
-Let the following derivation with the open assumption $G(x)$ be $\Pi_1[G(x)]$:
-
-\begin{prooftree}
-\AXC{$ \neg P(\psi)^1$}
-       \AXC{Axiom \ref{A1}} \dashedLine
-       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\all_E$}
-       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}
- \BIC{$P(\neg\psi)$}\RightLabel{}
-                   \AXC{$G(x)$} \dottedLine\RightLabel{D\ref{D1}}
-                 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
-                      \UIC{$P(\neg \psi)\imp \neg \psi(x)$}\RightLabel{$\imp_E$}
-           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}
-            \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}
-             \BIC{$\bot $}\RightLabel{$\imp_I^1$}
-             \UIC{$\neg \neg P(\psi)$}\RightLabel{$\neg\neg_E$}
-              \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
-              \UIC{$ \psi(x)\imp P(\psi)$}
-\end{prooftree}
-
-\noindent
-Let the following derivation with the open assumption $G(x)$ be $\Pi_2[G(x)]$:
-\begin{prooftree}
-       \AXC{$\psi(x)^3$}
-                     \AXC{$\Pi_1[G(x)] $} \dashedLine
-                \UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp_E$}
-                       \BIC{$P(\psi)$}
-                 \AXC{Axiom \ref{A4}} \dashedLine
-                \UIC{$\all \varphi .(P(\varphi)\imp \Box P(\varphi))$}\RightLabel{$\all_E$}
-                 \UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp_E$}
-           \BIC{$\Box P(\psi)$}\RightLabel{$\imp_I^3$}
-           \UIC{$\psi(x)\imp\Box P(\psi)$}
-\end{prooftree}
-
-\noindent
-Let the following derivation without open assumptions be $\Pi_3$:
-
-\begin{prooftree}
-       \AXC{$P(\psi)^4$}
-                     \AXC{$G(x)^5$} \dottedLine\RightLabel{D\ref{D1}}
-                \UIC{$\all\varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\all_E$}
-                \UIC{$P(\psi)\imp \psi(x)$}\RightLabel{$\imp_E$}
-                       \BIC{$\psi(x)$}\RightLabel{$\imp_I^5$}
-                       \UIC{$G(x)\imp \psi(x)$}\RightLabel{$\all_I$}
-                \UIC{$\all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^4$}
-                \UIC{$P(\psi)\imp \forall x .(G(x)\imp \psi(x))$}
-\end{prooftree}
-
-\noindent
-Let the following derivation with the open assumption $G(x)$ be $\Pi_4[G(x)]$:
-\begin{prooftree}
-       \AXC{$\psi(x)^6 $} \RightLabel{}
-                     \AXC{$\Pi_2$} \dashedLine
-                \UIC{$\psi(x)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
-                \BIC{$\Box P(\psi)$}
-                         \AXC{$\Box P(\psi)^7 $}\RightLabel{$\Box_E$}
-                               \UIC{$P(\psi)$}
-                                \AXC{$\Pi_3$} \dashedLine\RightLabel{}
-                                       \UIC{$P(\psi)\imp \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_E$}
-                               \BIC{$\all x .(G(x)\imp \psi(x))$}\RightLabel{$\Box_I$}
-                        \UIC{$\Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_I^7$}
-                        \UIC{$\Box P(\psi)\imp \Box \all x .(G(x)\imp \psi(x))$} \RightLabel{$\imp_E$}
-                  \BIC{$\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\imp_I^6$}
-           \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}
-\end{prooftree}
-
-\noindent
-The use of the necessitation rule above is correct, because the only open assumption $\Box P(\psi)$ is boxed.
-In the derivation of Theorem \ref{T2} below, the assumption $G(x)$ in the subderivation $\Pi_4[G(x)^8]$ is discharged by the rule labeled $8$.
-
-\begin{prooftree}
-       \AXC{$G(x)^8 $} \RightLabel{}
-                     \AXC{$\Pi_4[G(x)^8]$} \dashedLine
-                \UIC{$\psi(x)\imp\Box \all x .(G(x)\imp \psi(x))$}\RightLabel{$\all_I$}
-                \UIC{$\all \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\RightLabel{$\wedge_I$}
-                \BIC{$G(x)\wedge \forall \psi .(\psi(x)\imp\Box \all x .(G(x)\imp \psi(x)))$}\dottedLine\RightLabel{D\ref{D2}}
-                               \UIC{$\ess{G}{x}$} \RightLabel{$\imp_I^8$}
-                        \UIC{$G(x)\imp \ess{G}{x}$} \RightLabel{$\all_I$}
-                         \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}
-\end{prooftree}
-\end{proof}
-
-\subsection{If God's existence is possible, it is necessary}
-
 \begin{definition}
@@
 \end{definition}
-
 \begin{axiom}
@@
 \end{axiom}
-
 \begin{lemma}
@@
 \end{lemma}
+
 \begin{proof} \hfill
@@
 \begin{prooftree}
-\AXC{$ $} \dashedLine
-\UIC{$G(g)$}
-        \AXC{Theorem \ref{T2}} \dashedLine
-        \UIC{$\all y.[G(y) \imp \ess{G}{y}]$}\RightLabel{$\all_E$}
-        \UIC{$G(g) \imp \ess{G}{g}$}\RightLabel{$\imp_E$}
-    \BIC{$\ess{G}{g}$}
+
+\AXC{$ $} \RightLabel{2}
+\UIC{$\neg P(\psi)$} 
+\AXC{Axiom \ref{A1}} \dashedLine
+\UIC{$\neg P(\psi) \imp P(\neg \psi)$}\RightLabel{$\imp_E$}
+\BIC{$P(\neg \psi)$}
+\AXC{$G(g)$}  \dashedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
+\UIC{$P(\neg \psi) \imp \neg \psi(g)$}\RightLabel{$\imp_E$}
+\BIC{$\neg \psi(g)$}
+\AXC{$ $} \RightLabel{4}
+\UIC{$\psi(g)$} \RightLabel{$\imp_E$}
+\BIC{$\bot$}\RightLabel{$\imp_I^2$}
+\UIC{$\neg\neg P(\psi)$}\RightLabel{$\neg\neg_E$}
+\UIC{$P(\psi)$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$P(\psi)$}
+\AXC{$ $} \RightLabel{3}
+\UIC{$G(y)$}  \dashedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
+\UIC{$P(\psi) \imp \psi(y)$}\RightLabel{$\imp_E$}
+\BIC{$\psi(y)$} \RightLabel{$\imp_I^3$}
+\UIC{$G(y) \imp \psi(y)$}\RightLabel{$\forall_I$}
+\UIC{$\forall y.(G(y) \imp \psi(y))$}\RightLabel{$\nec_I$}
+\UIC{$\nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\imp_I^4$}
+\UIC{$\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\forall_I$}
+\UIC{$\forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\RightLabel{$\wedge_I$}
+\AXC{$G(g)$}  
+\BIC{$G(g) \wedge \forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dashedLine \RightLabel{D\ref{D2}}
+\UIC{$\ess{G}{g}$}
 \AXC{Axiom \ref{A5}} \dashedLine
-                \UIC{$P(E)$}
-                        \AXC{$ $} \dashedLine
-                        \UIC{$G(g)$} \dottedLine \RightLabel{D\ref{D1}}
-                        \UIC{$\all \varphi.[P(\varphi) \imp \varphi(g)]$}\RightLabel{$\all_E$}
+\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
+\AXC{$G(g)$}\dashedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
 \UIC{$P(E) \imp E(g)$}\RightLabel{$\imp_E$}
-                     \BIC{$E(g)$} \dottedLine\RightLabel{D\ref{D3}}
+\BIC{$E(g)$}\dottedLine \RightLabel{D\ref{D3}}
 \UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
@@
 
-\subsection{God exists}
+\subsection{God necessarily exists}
 
 ToDo: this is proven in a way that is slightly different from G\"odel's 1970.
-
 \begin{theorem}
 \label{T4}
-God exists:
+God necessarily exists:
 $$
-\ex x. G(x)
+\nec \ex x. G(x)
 $$
 \end{theorem}
-
-\noindent
-\textbf{Formal proof:}
+\begin{proof}
 
@@
 \begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec_E$, axiom T}
-\UIC{$\neg\nec \ex x. G(x)$}
+\AXC{ Distribution principle \ref{DP} } \dashedLine
+\UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
 \AXC{ Lemma \ref{L1} } \dashedLine
-\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp_E$}
- \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec \neg\ex x. G(x)$} \dottedLine
-\UIC{$\neg \pos\ex x. G(x)$}
+\UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
+\UIC{$\nec[\ex x. G(x)\imp \nec \ex x. G(x)]$}\RightLabel{$\imp_E$}
+\BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
+\end{prooftree}
+\end{small}
+\begin{small}
+\begin{prooftree}
+\AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
+\AXC{ Theorem \ref{T1} } \dashedLine
+\UIC{$\pos \ex x. G(x)$}\RightLabel{$\imp_E$}
+\BIC{$\pos\nec \ex x. G(x)$}
+\AXC{ Brouwer's reduction theorem \ref{***} } \dashedLine
+\UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
+\BIC{$ \ex x. G(x)$} \RightLabel{$\nec_I$}
+\UIC{$\nec \ex x. G(x)$} 
 \end{prooftree}
@@
 
-\begin{small}
-\begin{prooftree}
-\AXC{$ $}\dashedLine
-\UIC{$\neg \pos\ex x. G(x)$}
-\AXC{ Corollary \ref{C1}} \dashedLine
-\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp_E$}
- \BIC{$ \bot$} \RightLabel{$\imp_I^1$}
- \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \dottedLine
- \UIC{$\neg\nec\pos \neg \ex x. G(x)$}
-\AXC{ Axiom B} \dashedLine
-\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \doubleLine
-\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp \neg \neg \ex x. G(x)$} \RightLabel{$\imp_E$}
- \BIC{$ \neg \neg \ex x. G(x)$}\RightLabel{$ \neg \neg$ E}
- \UIC{$ \ex x. G(x)$}
-\end{prooftree}
-\end{small}
-
-Note that the last step is classical and we do not prove the existential statement by providing an object for which the statement holds.
-This proof makes section \ref{GodsE} superflous and the use of axiom "M" unnecessary.
-
-The system used contains the $\nec_E$-rule with the restriction that we have a $\nec_I$ below it.
-This is equivalent to modal sytem K that contains axiom K and necessitation rule N.
-We aslo use axiom B ($ A\imp\nec\pos A$).
-No other modal axioms are needed.
-
-\subsection{Necessarily, God exists}
-
-We can also prove that god exists necessarily in our system by simply introducing box on a theorem by rule N in modal sytem K. So we get:
-
-\begin{corollary}
-\label{T3}
-Necessarily, God exists:
-$$
-\nec \ex x. G(x)
-$$
-\end{corollary}
-
-\subsection{God exists}\label{GodsE}
-This section is superfluous.
-
-\begin{axiom}[M]
-\label{M} What is necessary is the case:
-$$
-\all \varphi. [\nec \varphi \imp \varphi]
-$$
-\end{axiom}
-
-\begin{corollary}
-\label{C2}
-There exists a God:
-$$
-\ex x. G(x)
-$$
-\end{corollary}
-\begin{proof}\hfill
-\begin{prooftree}
-\AXC{Theorem \ref{T3}} \dashedLine
-\UIC{$ \nec \ex x. G(x)$}
-        \AXC{$ \all \varphi. [\nec \varphi \imp \varphi]$}
-        \UIC{$\nec \ex x. G(x) \imp \ex x. G(x)$}
-    \BIC{$\ex x. G(x)$}
-\end{prooftree}
 \end{proof}
@@
 
+\bibitem[{\itshape Anderson (1990)}]{and90}
+Anderson, C. A. 1990. {\itshape Some Emendations of G\"odel's Ontological Proof}. Faith and Philosophy, Vol. 7, Issue 3, pp. 291-303.
+
 \bibitem[{\itshape Anderson \& Gettings(1996)}]{and}
@@
 Kant, I.  original 1781.   {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
+
+\bibitem[{\itshape Negri (2005)}]{negri}
+Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
 

### 2015-02-25 Papers/2015/ModalLogicCoq/ModalCoq.tex
Move Coq paper

--- 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
@@
   This paper describes an embedding of higher-order modal logics in
-  the \Coq proof assistant. {\Coq}'s capabilities are extended in a
+  the \Coq proof assistant. {\Coq}'s capabilities are used to implement modal logics in a
   minimalistic manner, which is nevertheless sufficient for the
@@
 can be used as is, and only the input files provided to the prover
-must be especially encoded. Furthermore, the efficacy and efficiency
-of the embedding approach has been confirmed in philosophically
-relevant benchmarks \cite{FormalTheologyRepository}. These qualities
+must be specially encoded (using lifted versions of connectives and logical constants instead of the usual ones). Furthermore, the efficacy and efficiency
+of the embedding approach has been confirmed in benchmarks stemming from philosophy \cite{FormalTheologyRepository}. These qualities
 make embedding a convenient approach for \emph{fully automated}
@@
 usual logical connectives. Notations are used to allow the modal
-connectives to be used as similarly as possible as the usual
+connectives to be used as similarly as possible to the usual
 connectives. The prefix ``\texttt{m}'' is used to distinguish the
@@
 
-\noindent
+%\noindent
 To maximally preserve user intuition in interactive modal logic
@@
 \noindent If the new modal tactics above are regarded from a natural
-deduction point of view, they give rise to the inference
-rules\footnote{The ND calculus with the rules from
+deduction point of view, they correspond to the inference
+rules shown in Figure
+\ref{fig:NDK}. Because of this correspondence and the Henkin-completeness of the modal natural deduction calculus\footnote{
+  The ND calculus with the rules from
   Figures \ref{fig:NaturalDeduction} and \ref{fig:NDK} is sound and
@@
   \ref{fig:NaturalDeduction} extended with a necessitation rule and the
-modal axiom K \cite{ModalND}. Starting from a sound and Henkin-complete ND calculus for classical higher-order logic (cf. Figure~\ref{fig:NaturalDeduction}), the additional
+  modal axiom K \cite{ModalND}. Starting from a sound and 
+  Henkin-complete ND calculus for classical higher-order logic 
+  (cf. Figure~\ref{fig:NaturalDeduction}), the additional
   modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete
-for modal higher-order logic \textbf{K}.} shown in Figure
-\ref{fig:NDK}. The labels that name boxes in those inference rules are
+  for the rigid higher-order modal logic \textbf{K}.}, 
+the tactics allow the user to prove any valid modal formula without having 
+to unfold the definitions of the modal operators.
+
+The labels that name boxes in the inference rules of Figure \ref{fig:NDK} are
 precisely the worlds that annotate goals and hypotheses in \Coq with
@@
 \s
+\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
+\s
 \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
-\s
-\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}
 
@@
 but also for larger developments, we include here a full and detailed
-formalization of G\"odel's ontological argument.
+formalization of G\"odel's ontological argument. which has been verified in Coq 8.4pl5.
 
@@
 
-The theological implications of the verified correctness of G\"odel's
-proof certainly depend on a critical discussion of the underlying
-concepts, definitions and axioms, which is beyond the scope of this
-paper.  Clearly, the application of theorem proving technology to the
-domain of theoretical philosophy can --- as already pictured by
-Leibniz --- be very fruitful for both areas.
+%The theological implications of the verified correctness of G\"odel's
+% proof certainly depend on a critical discussion of the underlying
+% concepts, definitions and axioms, which is beyond the scope of this
+% paper.  Clearly, the application of theorem proving technology to the
+% domain of theoretical philosophy can --- as already pictured by
+% Leibniz --- be very fruitful for both areas.
 

### 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
added 4 page version of the invited abstract

--- 2014-09-17 Papers/2014-WADT/paper.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{fancyvrb,amssymb,graphicx}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 \title{On Logic Embeddings and G\"odel's God}
-
 \author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
-    the German Research Foundation DFG under grants BE2501/9-1 \&
+    the German Research Foundation DFG under grants BE2501/9-1,2 and
     BE2501/11-1.} \and Bruno Woltzenlogel Paleo\inst{2} 
@@
 \institute{%Department of Mathematics and Computer Science \\ 
- Freie Universit\"at Berlin, Germany, \url{c.benzmueller@fu-berlin.de}
+ Freie Universit\"at Berlin, Germany \\ \url{c.benzmueller@fu-berlin.de}
 \and
- Vienna Technical University, Austria, \url{bruno@logic.at}}
+ Vienna Technical University, Austria \\\url{bruno@logic.at}}
 
@@
 
-% \begin{abstract}
-%   Logic embeddings provide an elegant means to formalize sophisticated
-%   non-classical logics in classical higher-order logic. In recent work
-%   we have applied the logic embeddings approach to verify and automate
-%   a prominent philosophical argument that has fascinated philosophers
-%   and theologists for about 1000 years: the ontological argument for
-%   the existence of God. In our work we have concentrated on Kurt
-%   G\"odel's (respectively, Dana Scott's) modern version of this
-%   argument, which employs a second-order modal logic.  In our ongoing
-%   computer-assisted study of G\"odel's proof script, the automated
-%   reasoning tools have made some interesting observations, some of
-%   which were unknown so far.
+\begin{abstract}
+  We have applied an elegant and flexible  logic embeddings
+  approach to verify and automate a prominent philosophical argument:
+  the ontological argument for the existence of God. In our ongoing
+  computer-assisted study, higher-order automated reasoning tools have
+  made some interesting observations, some of which were unknown so
+  far.
+\end{abstract}
 
@@
 for QMLs \cite{fitting98} or selection function semantics for QCLs
-\cite{Stalnaker68}.  The embeddings approach is related to labelled
-deductive systems \cite{gabbay96}, which employ meta-level
-(world-)labeling techniques for the modeling and implementation of
-non-classical proof systems. In our embeddings approach such labels
-are instead encoded in the HOL logic.
+\cite{Stalnaker68}.
 
-In recent work \cite{C40,J28} we have applied the embeddings
-approach to verify and automate a philosophical argument that has
-fascinated philosophers and theologists for about 1000 years: the
-ontological argument for the existence of
-God~\cite{sobel2004logic}. We have thereby concentrated on
-G\"odel's \cite{GoedelNotes}, respectively  Scott's \cite{ScottNotes},
-modern
-version of this argument, which employs a second-order modal logic, for
-which, until now, no theorem provers were available.  In our
-computer-assisted study of the argument, the HOL provers LEO-II
-\cite{C26} and Satallax \cite{Satallax} have made some interesting
-observations, some of which were unknown so far.
+The embeddings approach is related to labelled deductive systems
+\cite{gabbay96}, which employ meta-level (world-)labeling techniques
+for the modeling and implementation of non-classical proof systems. In
+our embeddings approach such labels are instead encoded in the HOL
+logic.
+
+The embeddings approach is flexible, because various modal logics (even
+with multiple modalities or a mix of varying/cumulative domain
+quantifiers) can be easily supported by stating their characteristic
+axioms. Moreover, it is relatively simple to implement, because it
+does not require any modification in the source code of the
+higher-order prover. A minimal encoding of
+second-order modal logic KB in TPTP THF syntax \cite{J22} --- this syntax is
+accepted by a range of HOL automated theorem provers (ATPs) --- is
+exemplarily provided in Fig.~\ref{fig1}.\footnote{Some Notes on THF, which is a concrete syntax for HOL: \texttt{\$i} and
+                   \texttt{\$o} represent the HOL base types $i$ and
+                   $o$ (Booleans). \texttt{\$i>\$o} encodes a function
+                   (predicate) type.  Predicate
+                   application as in $A(X,W)$ is encoded as
+                   \texttt{((A@X)@W)} or simply as
+                   \texttt{(A@X@W)}, i.e., function/predicate application
+                   is represented by \texttt{@}; universal
+                   quantification and $\lambda$-abstraction as in
+                   $\lambda {A}_{i\rightarrow o} \forall {W_i} (A\,W)$
+                   and are represented as in
+                   \texttt{\^{}[X:\$i>\$o]:![W:\$i]:(A@W)}; comments begin with
+                   \texttt{\%}.} The given set of axioms turns
+any TPTP THF compliant HOL-ATP in a reasoning tool for second-order modal logic. A
+Henkin-style semantics is thereby assumed for both sides, the side of HOL and
+for second-order modal logic.
+
+\begin{figure}[h]
+\begin{flushright}
+\begin{minipage}{.97\textwidth}
+\VerbatimInput[frame=single,%
+               commandchars=\\\{\},%
+               fontfamily=courier,
+               fontseries=b,%
+               fontsize=\scriptsize,%
+               %rulecolor=green,%
+               %fillcolor=yellow,%
+               %formatcom=blue,%
+               framerule=1pt,%
+               framesep=5pt,
+               numbersep=5pt,
+               numbers=left]%
+               {Quantified_KB.ax}
+\end{minipage}
+\end{flushright}
+                \vskip-.5em
+               \caption{\label{fig1}HOL encoding of second-order modal
+                 logic KB in THF syntax. 
+                 Modal formulas are mapped to HOL 
+                 predicates (with type \texttt{\$i>\$o}); 
+                 type \texttt{\$i} now stands for possible worlds. 
+                 The modal connectives $\neg$ (\texttt{mnot}), $\vee$ (\texttt{mor}) and
+                 $\square$ (\texttt{mbox}), universal quantification for
+                 individuals (\texttt{mall\_ind}) and for sets of 
+                 individuals  (\texttt{mall\_indset}) are introduced in 
+                 lines 7-18. 
+                 Validity of lifted modal formulas is
+                 defined in an obvious ways (lines 20-21). Symmetry
+                 of accessibility relation $r$ is postulated in lines
+                 23-26. Hence, second-order KB is realized here; 
+                 for logic K the symmetry axiom can be dropped.
+                 % \texttt{?}
+                   % is the existential quantifier, and
+                   % $\neg,\vee,\wedge, and \Rightarrow$ (mat. impl.)
+                   % are written as \texttt{\url{~}}, \texttt{|},
+                   % \texttt{\&}, and \texttt{=>}. 
+                   % Better formatted and easier readable
+                   % presentations of our THF code can easily be
+                   % generated with the TPTP tools of
+                   % Sutcliffe;
+                   % here we did minimize space.
+                 }
+\end{figure}
+
+In recent work \cite{C40,J28} we have applied the embeddings approach
+to verify and automate a philosophical argument that has fascinated
+philosophers and theologists for about 1000 years: the ontological
+argument for the existence of God~\cite{sobel2004logic}. We have
+thereby concentrated on G\"odel's \cite{GoedelNotes}, respectively
+Scott's \cite{ScottNotes}, modern version of this argument, which
+employs a second-order modal logic (S5), for which, until now, no
+theorem provers were available.  In our computer-assisted study of the
+argument, the HOL provers LEO-II \cite{C26}, Satallax \cite{Satallax}
+and Nitpick \cite{Nitpick} have made some interesting observations,
+some of which were unknown so far. This is a landmark result, with
+media repercussion in a global scale, and yet it is only a glimpse of
+what can be achieved by combining computer science, philosophy and
+theology.
+
+We briefly summarize some of these observations: Nitpick confirms that
+Scott's axioms are consistent, while LEO-II and Satallax demonstrate that
+Gödel's original, slightly different axioms are inconsistent. As far
+as we are aware of, this is a new result. As experiments with LEO-II
+revealed, the problem lies in a subtle difference in the definitions
+of the predicate \textit{essence} (characterizing the essential
+properties of an entity) between Gödel and Scott. In several philosophy papers, however, this
+difference has been commented as a simple cosmetic change.  Moreover,
+instead of using modal logic S5, 
+LEO-II and Satallax can prove the final theorem (that is,
+$\square \exists x . G(x)$, necessarily there exists God) already for
+modal logic KB.  This is highly relevant since several philosophers
+have criticized G{\"o}del's argument for the use of logic S5.  Axiom
+B (symmetry), however, cannot be dropped, which in turn is confirmed by Nitpick.
+LEO-II and Stallax can also show that G{\"o}del's and Scott's axioms
+imply what is called the modal collapse: $\phi\supset\Box\phi$. This
+formula expresses that contingent truth implies necessary truth (which
+can even be interpreted as an argument against free will;
+cf.~\cite{sobel2004logic}). The modal collapse is probably the most
+fundamental criticism put forward against G{\"o}del's resp. Scott's
+versions of the argument. Further theorems that can be shown by LEO-II
+and Satallax include flawlessness of God and monotheism.
+
+% \item The axioms and definitions from Fig.~\ref{fig1} are consistent (cf.~CO in Fig.~\ref{fig2}). 
+% \item Logic K is sufficient for proving T1, C and T2.
+% \item For proving the final theorem T3, logic KB is sufficient (and
+%   for K a countermodel is reported). This is highly relevant since
+%   several philosophers have criticized G{\"o}del's argument for the
+%   use of logic S5. This criticism is thus provably pointless.
+% \item Only for T3 the HOL-ATPs still fail to produce a proof directly
+%   from the axioms; thus, T3 remains an interesting benchmark problem; T1, C, and T2 are rather trivial for HOL-ATPs.
+% \item G\"odel's original version of the proof \cite{GoedelNotes},
+%   which omits conjunct $\phi(x)$ in the definition of \emph{essence} (cf.~D2'),
+%   seems inconsistent (cf.~the failed consistency check for CO' in
+%   Fig.~\ref{fig2}). As far as we are aware of, this is a new result.
+% \item G{\"o}del's axioms imply what is called the modal collapse (cf.~MC
+%   in Fig.~\ref{fig2}) $\phi\supset\Box\phi$, that is, contingent truth
+%   implies necessary truth (which can even be interpreted as an
+%   argument against free will; cf.~\cite{sobel2004logic}). MC is probably the most fundamental
+%   criticism put forward against G{\"o}del's argument.
+% \item For proving T1, only the $\supset$-direction of A1 is
+%   needed. However, the $\subset$-direction of A1 is required for
+%   proving T2. Some philosophers (e.g.~\cite{anderson90:_some_emend_of_goedel_ontol_proof}) try to avoid MC by
+%   eluding/replacing the $\supset$-direction of A1.
+% \item G{\"o}del's axioms imply a `flawless God', that is, an entity
+%   that can only have `positive' properties (cf.~FG in
+%   Fig.~\ref{fig2}). However, a comment by G{\"o}del in \cite{GoedelNotes}
+%   explains that `positive' is to be interpreted in a moral aesthetic
+%   sense only.
+
+% \footnote{In order to better understand G\"odel's notion of
+%     `positive' properties, we have reformulated G{\"o}del's theory
+%     and used `divine' instead of `positive'. Then we introduced
+%     orthogonal predicates `positive' and `negative' and we showed that
+%     God-like beings may well have positive and negative properties as
+%     long as all these properties are classified as divine properties. A respective
+%     formalization in Isabelle can be found at
+%     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/DivineVersion/GoedelGodDivine.thy}.}
+% \item Another implication of G{\"o}del's axioms is monotheism (see MT
+%   in Fig.~\ref{fig2}). MT can easily be proved by Satallax from FG and
+%   D1. It remains non-trivial to prove it directly from
+%   G{\"o}del's axioms.
+% \item \label{vary} All of the above findings hold for both constant domain semantics and varying domain semantics (for the domain of individuals).
+% \end{enumerate}
 
@@
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,fitting02:_types_tableaus_god,Hajek2002,Hajek2008},
-which claim to remedy some fundamental problem of G\"odel's and Scott's proofs, known as the modal collapse. The long-term goal
-is to work out a  landscape of the detailled logic parameters (e.g., constant
-vs. varying domains, rigid vs. non-rigid terms, logics KB vs.
-S4 vs.  S5, etc.) under which the proposed variants of the 
-modern ontological argument hold or fail.
+which claim to remedy some fundamental problem of G\"odel's and
+Scott's proofs, in particular, the modal collapse.  One interesting
+and very encouraging observation from these studies is, that the
+argumentation granularity as typically applied in these philosophy
+papers obviously already matches very well with the automation capabilities of our
+HOL-ATPs. The provides good evidence for the potential relevance of our
+approach (not only) wrt. other similar applications in metaphysics.
 
-There is little related work~\cite{oppenheimera11,rushby13}, and this focuses solely on the comparably simpler,
-original ontological argument by Anselm of Canterbury.  % These works do not achieve the close
-% correspondence between the original formulations and the formal
-% encodings that can be found in our approach and they do also not reach
-% the same of degree of proof automation.
+The long-term goal
+is to work out a landscape of the detailled logic parameters (e.g.,
+constant vs. varying domains, rigid vs. non-rigid terms, logics KB vs.
+S4 vs.  S5, etc.) under which the proposed variants of the modern
+ontological argument hold or fail.
+
+There is little related work~\cite{oppenheimera11,rushby13}, and this
+focuses solely on the comparably simpler, original ontological
+argument by Anselm of Canterbury.  These works do not achieve the
+close correspondence between the original formulations and the formal
+encodings that can be found in our approach and they do also not reach
+the same of degree of proof automation.
 
@@
 
+\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
+C.A.~Anderson.
+\newblock Some emendations of {G{\"o}del's} ontological proof.
+\newblock {\em Faith and Philosophy}, 7(3), 1990.
+
 \bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings.
+C.A.~Anderson and M.~Gettings.
 \newblock G\"odel ontological proof revisited.
@@
   1996.
-
-\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
-C.A.~Anderson.
-\newblock Some emendations of {G{\"o}del's} ontological proof.
-\newblock {\em Faith and Philosophy}, 7(3), 1990.
 
@@
 
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for
+  higher-order logic based on a relational model finder. In {\em Proc. of ITP
+  2010}, volume 6172 of LNCS, pp. 131--146. Springer, (2010).
+
 \bibitem{Satallax}
@@
 \newblock {\em Studia Logica: An International Journal for Symbolic Logic},
-  71(2):pp. 149--164, 2002.
+  71(2):pp. 149-164, 2002.
 
@@
 \newblock A theory of conditionals.
-\newblock In {\em Studies in Logical Theory}, pages 98--112. Blackwell, 1968.
+\newblock In {\em Studies in Logical Theory}, pages
+98--112. Blackwell, 1968.
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller.
+\newblock Automated reasoning in higher-order logic
+  using the {TPTP THF} infrastructure., {\em J. of Formalized Reasoning},
+  3(1),  1--27, (2010).
 

### 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
Fix typos and make minor sentence-level improvements.

@@
 \begin{abstract}
-  We have applied an elegant and flexible  logic embeddings
+  We have applied an elegant and flexible  logic embedding
   approach to verify and automate a prominent philosophical argument:
@@
   computer-assisted study, higher-order automated reasoning tools have
-  made some interesting observations, some of which were unknown so
-  far.
+  made some interesting observations, some of which were previously unknown.
 \end{abstract}
@@
 \cite{J23} and quantified conditional logics (QCL) \cite{C37}.  For
-many of the embedded logics few or none automated theorem provers did
+many of the embedded logics few or no automated theorem provers did
 exist before. HOL is exploited in this approach to encode the
@@
 
-The embeddings approach is flexible, because various modal logics (even
+The embedding approach is flexible, because various modal logics (even
 with multiple modalities or a mix of varying/cumulative domain
@@
 any TPTP THF compliant HOL-ATP in a reasoning tool for second-order modal logic. A
-Henkin-style semantics is thereby assumed for both sides, the side of HOL and
-for second-order modal logic.
+Henkin-style semantics is thereby assumed for both logics: HOL and
+second-order modal logic.
 
@@
                  Validity of lifted modal formulas is
-                 defined in an obvious ways (lines 20-21). Symmetry
+                 defined in the standard way (lines 20-21). Symmetry
                  of accessibility relation $r$ is postulated in lines
@@
 
-In recent work \cite{C40,J28} we have applied the embeddings approach
+In recent work \cite{C40,J28} we have applied the embedding approach
 to verify and automate a philosophical argument that has fascinated
-philosophers and theologists for about 1000 years: the ontological
+philosophers and theologians for about 1000 years: the ontological
 argument for the existence of God~\cite{sobel2004logic}. We have
-thereby concentrated on G\"odel's \cite{GoedelNotes}, respectively
-Scott's \cite{ScottNotes}, modern version of this argument, which
-employs a second-order modal logic (S5), for which, until now, no
+thereby concentrated on G\"odel's \cite{GoedelNotes} 
+modern version of this argument and on
+Scott's \cite{ScottNotes} modification, which
+employ a second-order modal logic (S5) for which, until now, no
 theorem provers were available.  In our computer-assisted study of the
@@
 Gödel's original, slightly different axioms are inconsistent. As far
-as we are aware of, this is a new result. As experiments with LEO-II
+as we are aware, this is a new result. As experiments with LEO-II
 revealed, the problem lies in a subtle difference in the definitions
 of the predicate \textit{essence} (characterizing the essential
-properties of an entity) between Gödel and Scott. In several philosophy papers, however, this
+properties of an entity) between Gödel and Scott. In several philosophy papers\marginpar{citation needed? weaken the "several" claim?}, however, this
 difference has been commented as a simple cosmetic change.  Moreover,
@@
 modal logic KB.  This is highly relevant since several philosophers
-have criticized G{\"o}del's argument for the use of logic S5.  Axiom
+have criticized G{\"o}del's argument for the use of logic S5\marginpar{citation needed? weaken the "several" claim?}.  Axiom
 B (symmetry), however, cannot be dropped, which in turn is confirmed by Nitpick.
-LEO-II and Stallax can also show that G{\"o}del's and Scott's axioms
+LEO-II and Satallax can also show that G{\"o}del's and Scott's axioms
 imply what is called the modal collapse: $\phi\supset\Box\phi$. This
-formula expresses that contingent truth implies necessary truth (which
+expresses that contingent truth implies necessary truth (which
 can even be interpreted as an argument against free will;
-cf.~\cite{sobel2004logic}). The modal collapse is probably the most
-fundamental criticism put forward against G{\"o}del's resp. Scott's
-versions of the argument. Further theorems that can be shown by LEO-II
+cf.~\cite{sobel2004logic}) and is probably the most
+fundamental criticism put forward against G{\"o}del's and Scott's
+versions of the argument. Other theorems that can be shown by LEO-II
 and Satallax include flawlessness of God and monotheism.
@@
 \cite{anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,fitting02:_types_tableaus_god,Hajek2002,Hajek2008},
-which claim to remedy some fundamental problem of G\"odel's and
-Scott's proofs, in particular, the modal collapse.  One interesting
+which claim to remedy some fundamental problems of G\"odel's and
+Scott's proofs, especially the modal collapse.  One interesting
 and very encouraging observation from these studies is, that the
-argumentation granularity as typically applied in these philosophy
-papers obviously already matches very well with the automation capabilities of our
-HOL-ATPs. The provides good evidence for the potential relevance of our
-approach (not only) wrt. other similar applications in metaphysics.
+argumentation granularity typical of these philosophy
+papers is already within reach of the capabilities 
+of our higher-order automated theorem provers.
+This provides good evidence for the potential 
+relevance of the embedding approach (not only) w.r.t. 
+other similar applications in metaphysics.
 
 The long-term goal
-is to work out a landscape of the detailled logic parameters (e.g.,
+is to methodically determine the range of logical parameters (e.g.,
 constant vs. varying domains, rigid vs. non-rigid terms, logics KB vs.
@@
 close correspondence between the original formulations and the formal
-encodings that can be found in our approach and they do also not reach
-the same of degree of proof automation.
+encodings that can be found in our approach and they also do not reach
+the same degree of proof automation.
 
@@
 embeddings-based approach. Most importantly, our work opens new
-perspectives towards a computational metaphysics.
+perspectives towards computational metaphysics.
 

### 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
Improve another sentence.

@@
 
-There is little related work~\cite{oppenheimera11,rushby13}, and this
-focuses solely on the comparably simpler, original ontological
+There have been few related works~\cite{oppenheimera11,rushby13}, 
+and they have
+focused solely on the comparably simpler, original ontological
 argument by Anselm of Canterbury. These works do not achieve the

### 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
Improve another sentence.

--- 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
@@
 
-There is little related work~\cite{oppenheimera11,rushby13}, and this
-focuses solely on the comparably simpler, original ontological
+There has been little related work~\cite{oppenheimera11,rushby13}, and this
+focused solely on the comparably simpler, original ontological
 argument by Anselm of Canterbury. These works do not achieve the

### 2015-03-02 Papers/2014/WADT/WADT-4pages.tex
slightly modified WADT paper

--- 2015-03-01 Papers/2014/WADT/WADT-4pages.tex
@@
 We briefly summarize some of these observations: Nitpick confirms that
-Scott's axioms are consistent, while LEO-II and Satallax demonstrate that
-Gödel's original, slightly different axioms are inconsistent. As far
-as we are aware, this is a new result. As experiments with LEO-II
+Scott's axioms are consistent, while LEO-II and Satallax demonstrate
+that Gödel's original, slightly different axioms are inconsistent. As
+far as we are aware, this is a new result. As experiments with LEO-II
 revealed, the problem lies in a subtle difference in the definitions
 of the predicate \textit{essence} (characterizing the essential
-properties of an entity) between Gödel and Scott. In several philosophy papers\marginpar{citation needed? weaken the "several" claim?}, however, this
-difference has been commented as a simple cosmetic change.  Moreover,
-instead of using modal logic S5, 
-LEO-II and Satallax can prove the final theorem (that is,
-$\square \exists x . G(x)$, necessarily there exists God) already for
-modal logic KB.  This is highly relevant since several philosophers
-have criticized G{\"o}del's argument for the use of logic S5\marginpar{citation needed? weaken the "several" claim?}.  Axiom
-B (symmetry), however, cannot be dropped, which in turn is confirmed by Nitpick.
-LEO-II and Satallax can also show that G{\"o}del's and Scott's axioms
-imply what is called the modal collapse: $\phi\supset\Box\phi$. This
-expresses that contingent truth implies necessary truth (which
-can even be interpreted as an argument against free will;
-cf.~\cite{sobel2004logic}) and is probably the most
-fundamental criticism put forward against G{\"o}del's and Scott's
-versions of the argument. Other theorems that can be shown by LEO-II
-and Satallax include flawlessness of God and monotheism.
+properties of an entity) between Gödel and Scott. In recent papers on
+the ontological argument (see e.g. below), some authors speak of an oversight/flaw by
+Gödel, some silently replace Gödel's definition without commenting and
+some simply stay with it.
+Moreover, instead of using modal logic S5, LEO-II and Satallax can
+prove the final theorem (that is, $\square \exists x . G(x)$,
+necessarily there exists God) already for modal logic KB.  This is
+highly relevant since some philosophers have criticized G{\"o}del's
+argument for the use of logic S5.  Axiom B (symmetry), however, cannot be dropped,
+which in turn is confirmed by Nitpick.  LEO-II and Satallax can also
+show that G{\"o}del's and Scott's axioms imply what is called the
+modal collapse: $\phi\supset\Box\phi$. This expresses that contingent
+truth implies necessary truth (which can even be interpreted as an
+argument against free will; cf.~\cite{sobel2004logic}) and is probably
+the most fundamental criticism put forward against G{\"o}del's and
+Scott's versions of the argument. Other theorems that can be shown by
+LEO-II and Satallax include flawlessness of God and 
+monotheism.
 

### 2015-03-31 Papers/InProgress/HajekAndersonControversy/main.tex

--- 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
@@
 \usepackage{modallogics}
-\usepackage[backend=bibtex,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
+\usepackage[backend=biber,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
 \addbibresource[label=bibs]{Bibliography.bib}
@@
 and T3' can be quickly automated (in logics \K, \K and \KB,
-respectively); the axioms A4 and A5 are proven redundant\footnote{
+respectively); the axioms A4 and A5' are proven redundant\footnote{
 An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
@@
 logic \KT (i.e. assuming reflexivity of the accessibility relation).
-For constant domain semantics, proving D1 is possible in logic \KFour,
+For constant domain semantics, proving D1' is possible in logic \KFour,
 whereas for varying domain semantics, a counter model can be found
@@
 claimed results, but also exposed a few mistakes and novel insights
-in a long standing controversy.  We believe the technology employed in
+in a long standing controversy. 
+We believe the technology employed in
 this work is ready to be  fruitfully adopted in larger scale by
@@
 % - Investigate Magari's proof
+
+% - Hájek actually uses multi-sorted first-order formalizations. We should check that our results still hold in a multi-sorted first-order setting. They should.
+
 % - Add apendix for Anderson's simplification
@@
 
+\section{Summary}
+
+\newcommand{\R}{Red.}
+\newcommand{\S}{Sup.}
+\newcommand{\P}{Prov.}
+\newcommand{\CS}{Countersat.}
+\newcommand{\U}{--}
+
+\begin{tabular}{lccc}
+Proof               & A4 & H:A4 & A5 & A5' & H:A5 & T3 & T3'  &  MC \\
+\hline
+Scott (const)       & 
+Scott (var)
+Anderson (const)    &     &      
+Anderson (var)      
+\AOE
+\AOEH
+\AOEHH
+\AOEHHH
+\end{tabular}
+
 \end{appendix}

### 2015-04-04 Papers/2014/WADT/WADT-4pages.tex
improving paper: adding reference, remove typo

@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \begin{flushright}
@@
 which claim to remedy some fundamental problems of G\"odel's and
-Scott's proofs, especially the modal collapse.  One interesting
+Scott's proofs, especially the modal collapse \cite{C41}.  One interesting
 and very encouraging observation from these studies is, that the
@@
   Applications}, pages 163 -- 168. IOS Press, 2014.
+
+\bibitem{C41}
+C. Benzm{\"u}ller and L. Weber and B. Woltzenlogel
+Paleo. Computer-Assisted Analysis of the {Anderson-H\'{a}jek}
+Ontological Controversy. Handbook of the 1st World Congress on Logic
+and Religion, Jo\~ao Pessoa, Brazil, 2015.
 

### 2015-04-05 Papers/2014/WADT/WADT-4pages.tex
Add bibitem for CSR paper

@@
 \newblock Automated reasoning in higher-order logic
-  using the {TPTP THF} infrastructure., {\em J. of Formalized Reasoning},
+  using the {TPTP THF} infrastructure, {\em J. of Formalized Reasoning},
   3(1),  1--27, (2010).
+
+\bibitem{CSR}
+C.~Benzm{\"u}ller and B. Woltzenlogel Paleo.
+\newblock Interacting with Modal Logics in the Coq Proof Assistant, {\em 10th International Computer Science Symposium in Russia (CSR)}, 2015.
 

### 2015-04-05 Papers/2014/WADT/WADT-4pages.tex
Add citation of CSR paper

@@
 
-In recent work \cite{C40,J28} we have applied the embedding approach
+In recent work \cite{C40,J28,CSR} we have applied the embedding approach
 to verify and automate a philosophical argument that has fascinated

### 2015-07-05 Papers/InProgress/HajekAndersonControversy/main.tex
Remove appendix and insert the formalizations directly into the text.

@@
 \noindent The axioms in G\"odel's ontological proof
-\citep{GoedelNotes,ScottNotes} (cf. Appendix \ref{apx:Goedel}) entail
-what is called \emph{modal collapse}
+\citep{GoedelNotes,ScottNotes} (cf. Fig. \ref{fig:Goedel}) entail what is called \emph{modal collapse}
 \citep{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
 \varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
-just for $\exists x. \mathit{God}(x)$ as intended. This fact, which
+just for $\exists x. \mathit{God}(x)$ as intended.
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A1] Either a property or its negation is positive, but not
+  both:
+  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified:
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$
+\item[D1] A \emph{God-like} being possesses all positive properties:
+  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$
+\item[A3]  The property of being God-like is positive:
+  $$\hol{P(G)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$
+\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
+  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$
+\item[T2]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
+\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$
+\item[L1] If a god-like being exists, then necessarily a god-like being exists:
+  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
+\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists:
+  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
+%
+\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Scott's version of G\"odel's ontological argument\label{fig:Goedel}}
+\end{figure}
+
+This fact, which
 has recently been confirmed with higher-order automated theorem
@@
 semantics}, the following results hold for \emph{Anderson's
-Emendation} (cf. Appendix \ref{apx:Anderson}): T1, C
+Emendation} (cf. Fig. \ref{fig:Anderson}): T1, C
 and T3' can be quickly automated (in logics \K, \K and \KB,
@@
 consistent.
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[A:A1] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$
+\item[A2] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P(\psi)]}$$
+\item[T1] Positive properties are possibly exemplified:
+  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive:
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$
+\item[A3']  The property of being God-like is positive:
+  $$\hol{P(G_A)}$$
+\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
+\item[A4]  Positive properties are necessarily positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
+\item[T2']  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y \varphi(y)]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
+\item[L1'] If a god-like being exists, then necessarily a god-like being exists:
+  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
+\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists:
+  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
+%
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Anderson's Emendation \label{fig:Anderson}}
+\end{figure}
 
@@
 of A4 and A5', as well as varying domain semantics, in his new
-emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
-\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
+emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Fig.
+\ref{fig:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
 axiom H:A12. Surprisingly, the computer-assisted formalization of
@@
 
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
+  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
+\item[A3']  The property of being God-like is positive:
+  $$\hol{P(G_H)}$$
+\item[A4]  Positive properties are necessarily positive:
+  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y [\varphi(y)]]}$$
+\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
+
+\item[L3]
+  \begin{itemize}
+  \item[(1)] The negation of a positive property is not positive:
+$$\hol{\allq \varphi[P(\varphi) \imp \neg P(\neg \varphi)] }$$
+  \item[(2)] Positive properties are possibly exemplified:
+$$\hol{\allq \varphi[P(\varphi) \imp \pos \exq x \varphi(x)] }$$
+  \item[(3)] If a god-like being exists, then necessarily a god-like being exists:
+$$\hol{\allq x[G_H(x) \imp \nec G_H(x)))] }$$
+  \item[(4)] All positive properties are necessarily implied by the property of being god-like:
+$$\hol{\allq \varphi[P(\varphi) \imp \nec \allq x[G_H(x) \imp \varphi(x)]] }$$
+  \end{itemize}
+\item[L4]  Being God-like is an essence of any
+  God-like being: $$\hol{\allq x [G_H(x) \imp \ess{G_H}{x}]}$$
+
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
+%
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Hájek's First Emendation \AOEH} \label{fig:Hajek1}
+\end{figure}
+
 Although Hájek did not notice the superfluousness of A4 and A5' in his
 \AOEH, he did describe yet another emendation (his \AOEHH, cf.
-Appendix \ref{apx:Hajek2}) where A4 and A5' are superfluous (though no
+Fig. \ref{fig:Hajek2}) where A4 and A5' are superfluous (though no
 claim is made w.r.t. to their redundancy), if A3' is replaced by a
@@
 
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive:
+  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$
+\item[H:A3]  The property of being God-like and existing actually is positive:
+  $$\hol{P(G_A \wedge E)}$$
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
+%
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Hájek's Second Emendation \AOEHH \label{fig:Hajek2}}
+\end{figure}
+
 Nevertheless, \AOEHH\ is explicitly regarded by
@@
 natural theory, based on a more sophisticated notion of positiveness.
-That is his final emendation (\AOEHHH, cf. \ref{apx:Hajek3}), which
+That is his final emendation (\AOEHHH, cf. Fig. \ref{fig:Hajek3}), which
 restores A3' and does use A4 and A5', albeit in a modified form (i.e.
-H:A4 and H:A5). The formalization of \AOEHHH\ shows that H:A4 is
+H:A4 and H:A5).
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+\begin{itemize}
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property:
+  $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive\textsuperscript{\#}:
+  $$\hol{G_H(x) \equiv \forall \varphi [P^\#(\varphi) \biimp \nec \varphi(x)]}$$
+\item[A3']  The property of being God-like is positive:
+  $$\hol{P(G_H)}$$
+\item[H:A4]  Positive\textsuperscript{\#} properties are necessarily positive\textsuperscript{\#}:
+  $$\hol{\allq \varphi [P^\#(\varphi) \imp \Box \; P^\#(\varphi)]}$$
+\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
+  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
+\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
+  \exq y [\varphi(y)]]}$$
+\item[H:A5] Necessary existence is a positive\textsuperscript{\#} property: $$\hol{P^\#(\NE_A)}$$
+%
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Hájek's Third Emendation \AOEHHH \label{fig:Hajek3}}
+\end{figure}
+
+The formalization of \AOEHHH\ shows that H:A4 is
 independent. The old A5' is independent as well, and both H:A4 and H:A5
@@
 provers (in logic \KFourB). Within the same trend, the alternative
-proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal})
-achieves a high level of minimality. He takes the property of being
-God-like as a primitive and defines (B:D1) the positive properties as
+proposed by \citet{bjordal99} (cf. Fig. \ref{fig:Bjordal})
+achieves a high level of minimality.
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+$G_B$ (God-like) is taken as primitive and $P_B$ (Positive) is defined.
+
+\bigskip
+
+\begin{itemize}
+\item[B:D1] A property is positive iff it is necessarily possessed by every God-like being.
+  $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$
+\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
+  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
+%  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
+\item[B:D2] a \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
+  $$\hol{\MCP(\phi,x) \equiv (\phi(x) \wedge P_B(\phi)) \wedge \allq \psi ((\psi(x) \wedge P_B(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
+\item[B:D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
+  $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
+\item[A:A1'] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$
+\item[A5'] Necessary existence is a positive property.
+ $$\hol{P_B(\NE_B)}$$
+\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$
+\end{itemize}
+\end{minipage}
+} %\vskip-.5em
+\caption{Bjordal's Alternative \label{fig:Bjordal}}
+\end{figure}
+
+He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as
 those properties necessarily possessed by every God-like being. He
@@
 \newpage
-\begin{appendix}
-\noindent
-
-\small
-
-\section{Scott's version of G\"odel's ontological argument} \label{apx:Goedel}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A1] Either a property or its negation is positive, but not
-  both:
-  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[D1] A \emph{God-like} being possesses all positive properties: 
-  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
-\item[A3]  The property of being God-like is positive: 
-  $$\hol{P(G)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
-  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
-\item[T2]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
-\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
-\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
-\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
-%
-\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
-\clearpage
-
-\section{Anderson's Emendation} \label{apx:Anderson}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[A:A1] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
-\item[A2] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P(\psi)]}$$
-\item[T1] Positive properties are possibly exemplified: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_A)}$$
-\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[T2']  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
-  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
-%
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
-\clearpage
-
-\section{Hájek's First Emendation \AOEH} \label{apx:Hajek1}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[H:A12] The negation of a property necessarily implied by a
-  positive property is not positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
-\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
-  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_H)}$$
-\item[A4]  Positive properties are necessarily positive: 
-  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y [\varphi(y)]]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
-
-\item[L3] 
-  \begin{itemize}
-  \item[(1)] The negation of a positive property is not positive:
-$$\hol{\allq \varphi[P(\varphi) \imp \neg P(\neg \varphi)] }$$
-  \item[(2)] Positive properties are possibly exemplified:
-$$\hol{\allq \varphi[P(\varphi) \imp \pos \exq x \varphi(x)] }$$
-  \item[(3)] If a god-like being exists, then necessarily a god-like being exists:
-$$\hol{\allq x[G_H(x) \imp \nec G_H(x)))] }$$  
-  \item[(4)] All positive properties are necessarily implied by the property of being god-like:
-$$\hol{\allq \varphi[P(\varphi) \imp \nec \allq x[G_H(x) \imp \varphi(x)]] }$$
-  \end{itemize}
-\item[L4]  Being God-like is an essence of any
-  God-like being: $$\hol{\allq x [G_H(x) \imp \ess{G_H}{x}]}$$
-
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
-% 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
-\clearpage
-
-\section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[H:A12] The negation of a property necessarily implied by a
-  positive property is not positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
-\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
-  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[H:A3]  The property of being God-like and existing actually is positive: 
-  $$\hol{P(G_A \wedge E)}$$
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
-% 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
-\section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-\begin{itemize}
-\item[H:A12] The negation of a property necessarily implied by a
-  positive property is not positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
-\item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property: 
-  $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
-\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive\textsuperscript{\#}: 
-  $$\hol{G_H(x) \equiv \forall \varphi [P^\#(\varphi) \biimp \nec \varphi(x)]}$$ 
-\item[A3']  The property of being God-like is positive: 
-  $$\hol{P(G_H)}$$
-\item[H:A4]  Positive\textsuperscript{\#} properties are necessarily positive\textsuperscript{\#}: 
-  $$\hol{\allq \varphi [P^\#(\varphi) \imp \Box \; P^\#(\varphi)]}$$ 
-\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
-  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
-  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
-  \exq y [\varphi(y)]]}$$
-\item[H:A5] Necessary existence is a positive\textsuperscript{\#} property: $$\hol{P^\#(\NE_A)}$$ 
-% 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
-\clearpage
-
-\section{Bjordal's Alternative} \label{apx:Bjordal}
-
-\framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-$G_B$ (God-like) is taken as primitive and $P_B$ (Positive) is defined.
-
-\bigskip
-
-\begin{itemize}
-\item[B:D1] A property is positive iff it is necessarily possessed by every God-like being.
-  $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$ 
-\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
-  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
-%  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
-\item[B:D2] a \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
-  $$\hol{\MCP(\phi,x) \equiv (\phi(x) \wedge P_B(\phi)) \wedge \allq \psi ((\psi(x) \wedge P_B(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
-\item[B:D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
-  $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
-\item[A:A1'] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$ 
-\item[A5'] Necessary existence is a positive property.
- $$\hol{P_B(\NE_B)}$$
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$ 
-\end{itemize}
-\end{minipage}
-} %\vskip-.5em
-
 \vskip 8em

### 2015-07-13 Papers/InProgress/HajekAndersonControversy/main.tex
Fix typo

@@
 For constant domain semantics, proving D1' is possible in logic \KFour,
-whereas for varying domain semantics, a counter model can be found
+whereas for varying domain semantics, a countermodel can be found
 even in logic \SFive. Conversely, the proof that B:D1 is entailed by

### 2015-07-13 Papers/InProgress/HajekAndersonControversy/main.tex
Add summary table to paper

@@
 
+Figure \ref{fig:summary} summarizes the results obtained by the automated reasoning tools.
+The following abbreviations are used: S/I = superfluous and independent; R = superfluous and redundant; S/U = superfluous and unknown whether redundant or independent; N/I = non-superfluous and independent; P = provable; CS = counter-satisfiable.
+The weakest logic required to show redundancy or provability is indicated in parentheses (\K is the default).
+Cells highlighted in red contain results that differ from what had been claimed by either Magari, Anderson or Hájek.
+Cells highlighted in yellow contain results that are surprising, albeit not contradicting any claims.
+Cells highlighted in green contain results where the tools were able to obtain the same results as humans, but using weaker modal logics.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{Summary.png}
+\caption{Summary of Results}
+\label{fig:summary}
+\end{figure}
+
 Using our approach, the formalization and (partly) automated analysis

### 2015-07-13 Papers/InProgress/HajekAndersonControversy/main.tex
Remove dead code
Ommit authors for double blind review

@@
 
-%----------AUTHOR 1
-\author[Benzm\"uller]{C. Benzm\"uller}
+% %----------AUTHOR 1
+% \author[Benzm\"uller]{C. Benzm\"uller}
 
-\address{%
-Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
-}
-\email{c.benzmueller@fu-berlin.com}
+% \address{%
+% Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
+% }
+% \email{c.benzmueller@fu-berlin.com}
 
-\thanks{This work was supported by German National Research Foundation (DFG) under
- grants BE 2501/9-1 and BE 2501/11-1.}
+% \thanks{This work was supported by German National Research Foundation (DFG) under
+%  grants BE 2501/9-1 and BE 2501/11-1.}
 
-%----------Author 2
-\author[Weber]{L. Weber}
-\address{
-Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
-}
-\email{leon.weber@fu-berlin.de}
+% %----------Author 2
+% \author[Weber]{L. Weber}
+% \address{
+% Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
+% }
+% \email{leon.weber@fu-berlin.de}
 
-%----------Author 3
-\author[Woltzenlogel-Paleo]{B. Woltzenlogel Paleo}
-\address{
-Room HA0402, Favoritenstra{\ss}e 9, 1040 Wien, Austria
-}
-\email{bruno.wp@gmail.com}
+% %----------Author 3
+% \author[Woltzenlogel-Paleo]{B. Woltzenlogel Paleo}
+% \address{
+% Room HA0402, Favoritenstra{\ss}e 9, 1040 Wien, Austria
+% }
+% \email{bruno.wp@gmail.com}
 
@@
 
-\newpage
-\vskip 8em
-
-\section{Summary}
-
-\newcommand{\R}{Red.}
-\newcommand{\S}{Sup.}
-\newcommand{\P}{Prov.}
-\newcommand{\CS}{Countersat.}
-\newcommand{\U}{--}
-
-\begin{tabular}{lccc}
-Proof               & A4 & H:A4 & A5 & A5' & H:A5 & T3 & T3'  &  MC \\
-\hline
-Scott (const)       &
-Scott (var)
-Anderson (const)    &     &
-Anderson (var)
-\AOE
-\AOEH
-\AOEHH
-\AOEHHH
-\end{tabular}
-
-\end{appendix}
-
 % ------------------------------------------------------------------------

### 2015-07-13 Papers/InProgress/HajekAndersonControversy/main.tex
Resolve inconsistencies between result table and text.

@@
 \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
-not only superflous, but also redundant. For A5', no conclusive results
-were achieved; neither a proof nor a countermodel could be
-automatically generated. Surprisingly, a countermodel for the weaker
+not only superflous, but also redundant. In addition, A5' becomes superfluous
+and independent. Surprisingly, a countermodel for the weaker
 A3' was succesfully generated. This is somewhat unsatisfactory (for
@@
 
-The formalization of \AOEHHH\ shows that H:A4 is
-independent. The old A5' is independent as well, and both H:A4 and H:A5
-are superfluous, but no conclusive results were achieved regarding
-independence or redundancy of H:A5.
+The formalization of \AOEHHH\ shows that both H:A4 and H:A5 are superfluous
+aswell as independent. For the old A5', no conclusive results were achieved.
 
@@
 quantifiers. Our computer-assisted study of this mixed variant shows
-that A4 is still redundant, but A5' becomes independent (hence not
+that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
 redundant). Unfortunately, a countermodel for T3 can then be found.
@@
 to eliminate A5' from his theory as well. Without any additional
-modification, the automated reasoners show that A5' is actually not
-superfluous. All these results, with the exception of the
+modification, the automated reasoners show that A5' is independent, but actually
+not superfluous. All these results, with the exception of the
 aforementioned countermodel for D1', hold for both constant and

### 2015-07-14 Papers/InProgress/HajekAndersonControversy/main.tex
Add formalization of Anderson's simplification.

@@
 \begin{itemize}
-\item[H:A12] The negation of a property necessarily implied by a
-  positive property is not positive:
-  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
-\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
-  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
+\item[H:DP] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
+  $$\hol{P(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
 \item[A3']  The property of being God-like is positive:
@@
 Additionally, Anderson \citep[footnote
-14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
+14]{anderson90:_some_emend_of_goedel_ontol_proof} (cf. Fig. \ref{fig:Anderson_simp}) remarks that only
 the quantifiers in T3' and in A:D2 need to be interpreted as
@@
 the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
-derivable. These claims have been confirmed by the automated theorem
+derivable.
+
+\begin{figure}[t]
+\noindent \framebox[\columnwidth][r]{
+\begin{minipage}{.94\columnwidth}\small
+$D$ (Defective) is taken as primitive and $P_{AS}$ (Positive) is defined.
+
+\bigskip
+
+\begin{itemize}
+\item[AS:D1] A property is positive iff its absence necessarily renders an individual defective and it is possible
+  that an individual has the property without being defective.
+   $$\hol{P_{AS}(\varphi) \equiv \nec (\allq x (\neg \varphi(x) \imp D(x)) \wedge (\neg \nec \allq x (\varphi(x) \imp D(x))))}$$
+\item[A:A1'] If a property is positive, its negation is not positive:
+  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \neg P_{AS}(\neg \varphi)]}$$
+\item[A2'] A property necessarily implied by a
+  positive property is positive:
+  $$\hol{\allq \varphi \allq \psi [(P_{AS}(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp P_{AS}(\psi)]}$$
+\item[A4']  Positive properties are necessarily positive:
+  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \nec P_{AS}(\varphi)]}$$
+\end{itemize}
+\end{minipage}
+} \vskip-.5em
+\caption{Anderson's Simplification \label{fig:Anderson_simp}}
+\end{figure}
+
+These claims have been confirmed by the automated theorem
 provers (in logic \KFourB). Within the same trend, the alternative
@@
 \end{minipage}
-} %\vskip-.5em
+} \vskip-.5em
 \caption{Bjordal's Alternative \label{fig:Bjordal}}

### 2015-07-14 Papers/InProgress/HajekAndersonControversy/main.tex
Move definitions from footnotes to main text.

@@
 
+\begin{definition}
+An axiom $A$ is \textbf{redundant} w.r.t. a set of axioms $S$ iff $A$ is
+entailed by $S$. \end{definition}
+
 For both \textbf{constant domain semantics} and \textbf{varying domain
@@
 and T3' can be quickly automated (in logics \K, \K and \KB,
-respectively); the axioms A4 and A5' are proven redundant\footnote{
-An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
-entailed by $S$.  } (the former in logic \KFourB and the latter
+respectively); the axioms A4 and A5' are proven redundant (the former in logic \KFourB and the latter
 already in \K); a trivial countermodel (with two worlds and  two
@@
 
+\begin{definition}
+An axiom $A$ is \textbf{superfluous} w.r.t. a set of axioms $S$ iff T3 is
+entailed by $S \setminus \{ A \}$.
+\end{definition}
+
 The redundancy of A4 and A5 is particularly controversial. Magari
-\cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
-An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is
-entailed by $S \setminus \{ A \}$. }, arguing that T3 is true in all
+\cite{Magari1988} claimed that A4 and A5 are superfluous, arguing that T3 is true in all
 models of the other axioms and definitions by Gödel.

### 2015-07-14 Papers/InProgress/HajekAndersonControversy/main.tex
Move definition of independency from footnotes to main text.

@@
 \AOEH\ shows that A4 and A5' are still superfluous. Moreover,
-A4 and A5' are independent\footnote{
+A4 and A5' are independent
+of the other axioms and definitions.
+Therefore, A4 and A5' are not redundant, despite their superfluousness.
+
+\begin{definition}
   An axiom $A$ is \emph{independent} of a set of axioms $S$ iff
@@
   $S$ where $A$ if false.
-}
-of the other axioms and definitions.
-Therefore, A4 and A5' are not redundant, despite their superfluousness.
+\end{definition}
 

### 2015-07-16 Papers/InProgress/HajekAndersonControversy/main.tex
minor changes

--- 2015-07-14 Papers/InProgress/HajekAndersonControversy/main.tex
@@
 technically simpler to analyze with computer support. We have
-formalized them using the proof assistant Isabelle/HOL \citep{Isabelle}
+formalized those using the proof assistant Isabelle/HOL \citep{Isabelle}
 together with the automated higher-order reasoners Leo-II \citep{C26},
@@
 of A4 and A5', as well as varying domain semantics, in his new
-emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Fig.
+emendation (named \AOEH\/ \citep[sec.~4]{Hajek2002}, cf. Fig.
 \ref{fig:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
@@
 \begin{itemize}
-\item[H:DP] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
-  $$\hol{P(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
+\item[H:A12] The negation of a property necessarily implied by a
+  positive property is not positive:
+  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
+  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
+\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
+  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
 \item[A3']  The property of being God-like is positive:
@@
 \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
-not only superflous, but also redundant. In addition, A5' becomes superfluous
-and independent. Surprisingly, a countermodel for the weaker
+not only superflous, but also redundant. For A5', no conclusive results
+were achieved; neither a proof nor a countermodel could be
+automatically generated. Surprisingly, a countermodel for the weaker
 A3' was succesfully generated. This is somewhat unsatisfactory (for
@@
 
-The formalization of \AOEHHH\ shows that both H:A4 and H:A5 are superfluous
-aswell as independent. For the old A5', no conclusive results were achieved.
+The formalization of \AOEHHH\ shows that H:A4 is
+independent. The old A5' is independent as well, and both H:A4 and H:A5
+are superfluous, but no conclusive results were achieved regarding
+independence or redundancy of H:A5.
 
 Additionally, Anderson \citep[footnote
-14]{anderson90:_some_emend_of_goedel_ontol_proof} (cf. Fig. \ref{fig:Anderson_simp}) remarks that only
+14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
 the quantifiers in T3' and in A:D2 need to be interpreted as
@@
 quantifiers. Our computer-assisted study of this mixed variant shows
-that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
+that A4 is still redundant, but A5' becomes independent (hence not
 redundant). Unfortunately, a countermodel for T3 can then be found.
@@
 the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
-derivable.
-
-\begin{figure}[t]
-\noindent \framebox[\columnwidth][r]{
-\begin{minipage}{.94\columnwidth}\small
-$D$ (Defective) is taken as primitive and $P_{AS}$ (Positive) is defined.
-
-\bigskip
-
-\begin{itemize}
-\item[AS:D1] A property is positive iff its absence necessarily renders an individual defective and it is possible
-  that an individual has the property without being defective.
-   $$\hol{P_{AS}(\varphi) \equiv \nec (\allq x (\neg \varphi(x) \imp D(x)) \wedge (\neg \nec \allq x (\varphi(x) \imp D(x))))}$$
-\item[A:A1'] If a property is positive, its negation is not positive:
-  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \neg P_{AS}(\neg \varphi)]}$$
-\item[A2'] A property necessarily implied by a
-  positive property is positive:
-  $$\hol{\allq \varphi \allq \psi [(P_{AS}(\varphi) \wedge \nec \allq x [\varphi(x)
-  \imp \psi(x)]) \imp P_{AS}(\psi)]}$$
-\item[A4']  Positive properties are necessarily positive:
-  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \nec P_{AS}(\varphi)]}$$
-\end{itemize}
-\end{minipage}
-} \vskip-.5em
-\caption{Anderson's Simplification \label{fig:Anderson_simp}}
-\end{figure}
-
-These claims have been confirmed by the automated theorem
+derivable. These claims have been confirmed by the automated theorem
 provers (in logic \KFourB). Within the same trend, the alternative
@@
 \end{minipage}
-} \vskip-.5em
+} %\vskip-.5em
 \caption{Bjordal's Alternative \label{fig:Bjordal}}
@@
 to eliminate A5' from his theory as well. Without any additional
-modification, the automated reasoners show that A5' is independent, but actually
-not superfluous. All these results, with the exception of the
+modification, the automated reasoners show that A5' is actually not
+superfluous. All these results, with the exception of the
 aforementioned countermodel for D1', hold for both constant and

### 2015-07-16 Talks/2015/CSR/talk-part-1.tex
Add CSR Talk

% \begin{transitionframe}{Images/Transitions/CartoucheThutmosisReverse(WaltersArtMuseum)(PublicDomain)}{black} \centering
% \textbf{Embedding Modal Logic into Higher-Order Logic}
% \end{transitionframe}

\begin{frame}{Quick Introduction to Modal Logics}{Modal Operators} \large
\centering

$$\nec P$$
P is necessary, P is obligatory, P is known, \\ 
P is believed, always P \ldots

\bigskip
\pause

$$\pos P$$
P is possible, P is permissible, P is epistemically possible, \\ 
P is doxastically possible, eventually P \ldots

\end{frame}

\begin{frame}{Quick Introduction to Modal Logics}{Kripke Semantics - Possible Worlds} \large

\includegraphics[width=0.8\textwidth]{Images/Kripke}

\end{frame}

\begin{frame}{Quick Introduction to Higher-Order Logics}{Higher-Order Quantification} \large

\begin{Huge}
$$
\forall x. (G(x) \equiv \alert{\forall \varphi}. P(\varphi) \imp \alert{\varphi}(x))
$$
\end{Huge}

\end{frame}

\begin{frame}{Quick Introduction to Higher-Order Logics}{Lambda Calculus: Types and Beta-Reduction} \large

\begin{Huge}
$$
(\lambda \varphi_{\alert{\iota \imp o}}. (P ~ \varphi) ) ~ G  \qquad  \leadsto_{\beta} \qquad (P ~ G)
$$
\end{Huge}
\end{frame}

\begin{frame}{Motivation, Challenge, Goals, Approach} \large

\visible<+->{
\hskip-1em Motivation : Applications (Ontologies, Paraconsistency, \emph{Philosophy}, \ldots) \\[1em]
}

\visible<+->{
\hskip-1em Challenge : No provers for \emph{Higher-order
  Quantified Modal
  Logics\/} (\textcolor{red}{QML}) \\[1em]
}

\visible<+->{
\hskip-1em Our goals : \\
  \qquad efficient automated provers for \textcolor{red}{QML} \\
  \qquad user-friendly interactive provers for \textcolor{red}{QML} \\[1em]
}

\visible<+->{
\hskip-1em Approach : Embed \textcolor{red}{QML} in \emph{Higher-Order Classical
  Logic\/} (\textcolor{blue}{HOL}) \\
\, \hfill Then use existing \textcolor{blue}{HOL} theorem provers for reasoning in \textcolor{red}{QML} \\
  \qquad\qquad\qquad interactive: \hfill Isabelle, HOL4, Hol Light, Coq, PVS, \ldots \\
  \qquad\qquad\qquad automated: \hfill TPS, LEO-II, Satallax, Nitpick, \ldots \\
}
\end{frame}

\begin{frame}{Embedding QML in HOL}\large

\hskip-1em\textcolor{red}{QML} \hfill
$\begin{array}{lll}\textcolor{red}{\varphi,\psi} & ::= &
  \textcolor{red}{\ldots}  \mid \textcolor{red}{\neg
    \varphi} \mid \textcolor{red}{\varphi \wedge \psi} \mid
  \textcolor{red}{\varphi \imp \psi}  \mid \textcolor{red}{\Box
    \varphi} \mid \textcolor{red}{\Diamond \varphi}  \mid
  \textcolor{red}{\forall {x}\, \varphi} \mid
  \textcolor{red}{\exists {x}\, \varphi} 
\mid \textcolor{red}{\forall {P}\, \varphi} \end{array}$ \\[1em]

\hskip-1em\textcolor{blue}{HOL}\hfill 
$\begin{array}{lll}
\textcolor{blue}{s,t} & ::= & \textcolor{blue}{C}  \mid
\textcolor{blue}{x \mid \lambda{x} s} \mid \textcolor{blue}{s\, t}
\mid \textcolor{blue}{\neg s} \mid \textcolor{blue}{s \vee t} \mid
\textcolor{blue}{\forall {x}\, t} 
\end{array}$ \\[1em]

\pause

\hskip-1em\textcolor{red}{QML} in \textcolor{blue}{HOL}: \quad \textcolor{red}{QML}
formulas $\textcolor{red}{\varphi}$ are mapped to
\textcolor{blue}{HOL} predicates $\textcolor{red}{\varphi_{\worldtype\typearrow o}}$

\begin{center}
\fcolorbox{blue}{white}{
$\begin{array}{lcl} 
    \textcolor{red}{\mnot} & = & \textcolor{blue}{
      \lambda{\varphi_{\worldtype\typearrow o}}\lambda{w_\worldtype}\neg \varphi w} \\ 
    \textcolor{red}{\mand} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}}
      \lambda{\psi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      (\varphi w \wedge \psi w)} \\ 
    \textcolor{red}{\imp} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}}
      \lambda{\psi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      (\neg \varphi w \vee \psi w)} \\ 
    \textcolor{red}{\forall} & = & \textcolor{blue}{ 
      \lambda{h_{\gamma\typearrow(\worldtype \typearrow o)}}
      \lambda{w_\worldtype} \forall {d_\gamma} \, h d w} \\
    \textcolor{red}{\exists} & = & \textcolor{blue}{ 
      \lambda{h_{\gamma\typearrow(\worldtype \typearrow o)}}
      \lambda{w_\worldtype} \exists {d_\gamma} \, h d w} \\
    \\
    \textcolor{red}{\Box} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      \forall {u_\worldtype}\, (\neg r w u \vee
      \varphi u)} \\ 
    \textcolor{red}{\Diamond} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      \exists {u_\worldtype}\, (r w u \wedge
      \varphi u)} \\ 
    \\
    \text{\textcolor{brown}{valid}} & = & \textcolor{blue}{
        \lambda{\varphi_{\worldtype\typearrow o}} \all{w_\worldtype}
        \varphi w}
\end{array}$
}  \quad \textcolor{blue}{Ax} 
\vskip1em
\end{center}

\pause

\quad The equations in \textcolor{blue}{Ax} are given as axioms to the \textcolor{blue}{HOL} provers. \\

\end{frame}

\begin{frame}{Embedding QML in HOL} \large

\hskip-1em Example \\[.5em]

 \textcolor{red}{QML} formula  \hfill \textcolor{red}{$\Diamond \exists x G(x)$}

\pause

 \textcolor{red}{QML} formula in \textcolor{blue}{HOL}  \hfill $\text{\textcolor{brown}{valid}}\, \textcolor{red}{(\Diamond \exists x G(x))_{\worldtype\typearrow o}}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype\textcolor{red}{(\Diamond \exists x
    G(x))_{\worldtype\typearrow o}}\, w}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \textcolor{red}{(\exists x G(x))_{\worldtype\typearrow o}} u)}$

% expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
%   w_\worldtype \exists {u_\worldtype} (r w u \wedge
%       \exists x \textcolor{red}{G(x)_{\worldtype\typearrow o}} u)}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \exists x G x u)}$ \\[1em]

\pause

\visible<6->{
\vfill
\begin{block}{What do we do?}
\vskip.5em
In order to prove that $\textcolor{red}{\varphi}$ is valid in \textcolor{red}{QML}, \\
we instead prove that 
$\text{\textcolor{brown}{valid}}\,
\textcolor{red}{\varphi_{\worldtype\typearrow o}}$ can be derived
from \textcolor{blue}{Ax} in \textcolor{blue}{HOL}. \\[1em]

This can be done with interactive or automated \textcolor{blue}{HOL} theorem provers.
\end{block}
}

\end{frame}

\begin{frame}{Embedding QML in HOL} \large

\hskip-1em Example \\[.5em]

 \textcolor{red}{QML} formula  \hfill \textcolor{red}{$\Diamond \exists x G(x)$}

 \textcolor{red}{QML} formula in \textcolor{blue}{HOL}  \hfill $\text{\textcolor{brown}{valid}}\, \textcolor{red}{(\Diamond \exists x G(x))_{\worldtype\typearrow o}}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype\textcolor{red}{(\Diamond \exists x
    G(x))_{\worldtype\typearrow o}}\, w}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \textcolor{red}{(\exists x G(x))_{\worldtype\typearrow o}} u)}$

% expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
%   w_\worldtype \exists {u_\worldtype} (r w u \wedge
%       \exists x \textcolor{red}{G(x)_{\worldtype\typearrow o}} u)}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \exists x G x u)}$ \\[1em]

\vfill
\begin{block}{Does this actually work in practice?}
\vskip.5em
\centering
Is it efficient ??

\bigskip

Is it user-friendly ??

\bigskip

\end{block}

\end{frame}

### 2015-07-16 Talks/2015/CSR/talk-part-3.tex
Add CSR Talk

\begin{frame}{Results Obtained with Fully Automated Reasoners}
\begin{changemargin}{-0.8cm}{-1.1cm}
\includegraphics[width=12.4cm]{Images/Results.png}
\end{changemargin}
\end{frame}

\begin{frame}{Results Obtained with Fully Automated Reasoners}
{A controversy between Magari, H\'ajek and Anderson regarding the redundancy of some axioms}
\begin{changemargin}{-0.8cm}{-1.1cm}
\includegraphics[width=12.4cm]{Images/ControversySummary.png}

\visible<2>{
\begin{minipage}{2.5cm}
\includegraphics[width=2.5cm]{./Images/Leibniz.png} 
\end{minipage}
\begin{minipage}{9cm}
\textbf{Leibniz (1646--1716)} \\[1em]
\emph{\textbf{characteristica universalis} and  \textbf{calculus ratiocinator}} \\[0.5em]
\footnotesize
\emph{
If controversies were to arise, there would be no more need of
disputation between two philosophers than between two
accountants. For it would suffice to take their pencils in their
hands, to sit down to their slates, and to say to each other \ldots :
Let us calculate.
}
\end{minipage}
}

\end{changemargin}
\end{frame}

% \begin{frame}{The Vision of Leibniz (1646--1716): \textit{Calculemus!}}
% \begin{changemargin}{-.5cm}{-.5cm}
% \begin{minipage}{4cm}
% \includegraphics[width=4cm]{./Images/Leibniz.png}

% \vspace*{1em}
% \color{gray}\footnotesize
% If controversies were to arise, there would be no more need of
% disputation between two philosophers than between two
% accountants. For it would suffice to take their pencils in their
% hands, to sit down to their slates, and to say to each other \ldots :
% Let us calculate. \\ \phantom{bla} \, \hfill (Translation by Russell)
% \end{minipage} \hfill
% \begin{minipage}{7cm} \small
% Quo facto, quando orientur controversiae, non magis disputatione opus erit inter
% duos philosophos, quam inter duos Computistas. Sufficiet enim calamos in
% manus sumere sedereque ad abacos, et sibi mutuo \ldots dicere: calculemus.
% \hfill (Leibniz, 1684)
% \vspace*{1em}

% \includegraphics[width=7cm,height=5cm]{./Images/Dispute.jpg}

% \vspace*{1em}

% \, \hfil \textbf{characteristica universalis} and  \textbf{calculus ratiocinator}
% \end{minipage}
% \end{changemargin}
% \end{frame}

\begin{frame}{Issues with Fully Automated Reasoning}{Proofs are hard to read and do not necessarily correspond to the informal proofs being verified}
\begin{changemargin}{-.5cm}{-1cm}
\colorbox{gray!20}{
\includegraphics[width=11.5cm,height=8cm]{./Images/LeoProof.png}
}
\end{changemargin}
\end{frame}

% \begin{transitionframe}{Images/Transitions/RussianChurches}{black}
% \textbf{Formalization and Verification in Coq}
% \end{transitionframe}

\begin{frame}{The Coq Proof Assistant}
{Winner of the ACM Software System Award in 2013}
\begin{changemargin}{-1cm}{-1cm}
\colorbox{gray}{\includegraphics[width=1.17\textwidth]{Images/Demos/CoqDemo.png}}
\end{changemargin}
\end{frame}

\begin{frame}{The Proof/Type System of Coq}
\begin{itemize}
\item Calculus of Inductive Constructions (CIC)
\item Related to CC and $\lambda C$ (cf. previous talk).
\item A minimalistic higher-order natural deduction calculus.
\item Typical natural deduction rules are admissible.
\end{itemize}
\end{frame}

\begin{frame}[shrink]{Typical Natural Deduction Rules}{Interactive proving using basic tactics in Coq feels roughly like constructing a natural deduction proof}

\begin{unnamedCalculus}

\vspace{1em}

\s\s
\infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
\s\s
\infer[\wedge_I]{A \wedge B}{A & B}
\s\s
\infer[\imp_I^h]{A \imp B}{ \infer*{B}{\infer[h]{A}{}} }

\vspace{2em}

\s\s
\infer[\vee_{I_1}]{A \vee B}{A}
\s\s
\infer[\wedge_{E_1}]{A}{A \wedge B}
\s\s
\infer[\imp_I]{A \imp B}{ B }

\vspace{2em}

\s\s
\infer[\vee_{I_2}]{A \vee B}{B}
\s\s
\infer[\wedge_{E_2}]{B}{A \wedge B}
\s\s
\infer[\imp_E]{B}{A & A \imp B}

\vspace{2em}

\s
\infer[\all_I]{\all x. A[x]}{ A[\alpha] }
\s
\infer[\all_E]{A[t]}{ \all x. A[x] }
\s\s
\infer[\ex_I]{\ex x. A[x]}{ A[t] }
\s
\infer[\ex_E]{A[\beta]}{ \ex x. A[x] }

\vspace{1em}

\s\s\s\s
$\neg A \equiv A \imp \bot$ 
\s\s\s 
\alert{\infer[\neg\neg_E]{A}{\neg\neg A}}

\vspace{1em}

\end{unnamedCalculus}

\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\begin{itemize}
\item \begin{LARGE} Challenges: \end{LARGE} \\[0.7em]
\begin{itemize}
\item \begin{large} Can we hide the semantic embedding from the user? \end{large}\\[0.5em]
\pause
\item \begin{large} Can we provide an interaction experience to the user that differs as little as possible from what he is already used to? \end{large}\\[0.5em]
\pause
\item \begin{large} Can we reconstruct, step-by-step in Coq, precisely Scott's formulation of G\"odel's argument?  \end{large}\\[0.5em]
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/1.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/2.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/3.png}
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/4.png}
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/5.png}\\
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/10.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/11.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/12.png}
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=0.75\textwidth]{Images/CoqCode/6.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/7.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/8.png}\\
\includegraphics[width=0.8\textwidth]{Images/CoqCode/9.png}
\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/10.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/11.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/12.png}

\bigskip

\includegraphics[width=\textwidth]{Images/CoqCode/10.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/13.png}
\end{frame}

\begin{frame}[shrink]{Natural Deduction Calculus}{Rules for Modalities}

\begin{unnamedCalculus}

\vspace{1em}

\s\s\s
\infer[\nec_I]{\nec A}{\alpha: \fbox{\infer*{A}{}} }
\s\s\s\s\s
\infer[\nec_E]{t: \fbox{ \infer*{}{A} }  }{\nec A}

\vspace{2em}

\s\s\s
\infer[\pos_I]{\pos A}{t: \fbox{\infer*{A}{}} }
\s\s\s\s\s
\infer[\pos_E]{\beta: \fbox{ \infer*{}{A} }  }{\pos A}

\vspace{1em}

\begin{center}
\textbf{eigen-box condition:}\\ 
$\nec_I$ and $\pos_E$ are \emph{strong} modal rules: \\
$\alpha$ and $\beta$ must be fresh names for the boxes they access \\ 
(in analogy to the eigen-variable condition for strong quantifier rules). \\
Every box must be accessed by \emph{exactly one} strong modal inference. \\
\vspace{0.5em}
\textbf{boxed assumption condition:} \\
assumptions should be discharged within \\
the box where they are created.
\end{center}

\end{unnamedCalculus}

\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}
\includegraphics[width=\textwidth]{Images/CoqCode/10.png}\\
\includegraphics[width=\textwidth]{Images/CoqCode/13.png}

\bigskip

\begin{center}
\includegraphics[width=0.8\textwidth]{Images/CoqCode/MP_Dia.png}
\end{center}

\end{frame}

\begin{frame}{Modal Logics in the Coq Proof Assistant}{Part of Scott's Formulation of G\"odel's Proof}
\includegraphics[width=\textwidth]{Images/CoqCode/14.png}
\end{frame}

% \begin{frame}[shrink]{Natural Deduction Calculus}{Three tiny examples}

% \vspace{1em}

% \infer[\imp_I]{A \imp A}{\infer[n]{A}{}}

% \infer[\imp_I]{A \imp A}{\infer[n]{A}{}}

% \vspace{1em}

% \end{frame}

% \begin{frame}{Natural Deduction Proofs}{T1 and C1}
% \begin{prooftree}
%         \AXC{\textbf{A2}} \dashedLine
%         \UIC{$ \all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
%         \UIC{$ \all \psi.[(P(\rho) \wedge \nec \all x.[\rho(x) \imp \psi(x)]) \imp P(\psi)]$} \RightLabel{$\all_E$}
%         \UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \doubleLine
%         \UIC{$(P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp P(\neg \rho)$}
%                         \AXC{\textbf{A1a}} \dashedLine
%                         \UIC{$\all \varphi.[ P(\neg \varphi) \imp \neg P(\varphi) ]$} \RightLabel{$\all_E$}
%                         \UIC{$ P(\neg \rho) \imp \neg P(\rho) $} \doubleLine
%                  \BIC{$ (P(\rho) \wedge \nec \all x.[\neg \rho(x)]) \imp \neg P(\rho) $} \doubleLine
%                  \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
%                  \UIC{\textbf{T1: }$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ] $}
% \end{prooftree}

% \begin{prooftree}
% \AXC{\textbf{A3}} \dashedLine
% \UIC{$P(G)$}
%                  \AXC{\textbf{T1}} \dashedLine
%                  \UIC{$\all \varphi.[ P(\varphi) \imp \pos \ex x.\varphi(x) ]$} \RightLabel{$\all_E $}
%                  \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
%     \BIC{$\pos \ex x. G(x)$}
% \end{prooftree}
% \end{frame}

% \begin{frame}{Natural Deduction Proofs}{T2 (Partial)}
% \includegraphics[scale=0.22]{Images/ProofOfT2Boxed.pdf}
% \end{frame}

### 2015-07-16 Talks/2015/CSR/talk-part-4.tex
Add CSR Talk

% \begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}{black}
% \textbf{Conclusions and Future Work}
% \end{transitionframe}

\begin{frame}{Conclusions} \large

\begin{itemize}
\item Contributions: \\[.5em]
\begin{itemize}
\item Efficient automated reasoning for QML \\[.3em]
\item \alert{User-friendly interactive reasoning with QML in Coq} \\[.3em]
\item \alert{A new natural deduction calculus for higher-order modal logics} \\[.3em]
\item Non-trivial new benchmark problems for HOL provers \\[.3em]
\item Verification of existing results about G\"odel's proof (e.g. Modal Collapse) \\[.3em]
\item New results about G\"odel's proof (e.g. Inconsistency) \\[.3em]
\item Resolution of philosophical controversies (Leibniz's dream) \\[.3em]
\item Computer-Assisted Theoretical Philosophy \\[.3em]
 \begin{itemize}
  \item Ed Zalta's \emph{Computational Metaphysics} project at Stanford University \\[.2em]
  \item John Rushby's formalization of Anselm's proof using PVS
  \end{itemize}
\end{itemize}

\end{itemize}
\end{frame}

### 2015-07-16 Papers/InProgress/HajekAndersonControversy/main.tex
Move definitions

@@
 
+\begin{samepage}
+The following definitions of redundancy, superfluousness and independence are used throughout this paper:
+
 \begin{definition}
-An axiom $A$ is \textbf{redundant} w.r.t. a set of axioms $S$ iff $A$ is
-entailed by $S$. \end{definition}
+An axiom $A$ is \textbf{redundant} w.r.t. a set of axioms $S$ iff $S$ entails $A$. 
+\end{definition}
+
+\begin{definition}
+An axiom $A$ is \textbf{superfluous} w.r.t. axiom set $S$ iff $S \setminus \{ A \}$ entails T3.
+\end{definition}
+
+\begin{definition}
+  An axiom $A$ is \textbf{independent} of a set of axioms $S$ iff
+  there are models of $S$ where $A$ is true and other models of
+  $S$ where $A$ if false.
+\end{definition}
+\end{samepage}
+
+\clearpage
 
@@
 
-\begin{definition}
-An axiom $A$ is \textbf{superfluous} w.r.t. a set of axioms $S$ iff T3 is
-entailed by $S \setminus \{ A \}$.
-\end{definition}
-
 The redundancy of A4 and A5 is particularly controversial. Magari
@@
 Therefore, A4 and A5' are not redundant, despite their superfluousness.
-
-\begin{definition}
-  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff
-  there are models of $S$ where $A$ is true and other models of
-  $S$ where $A$ if false.
-\end{definition}
 

### 2015-09-05 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Derivation of Brouwer Reduction Principle added
Merge Annika's contributios and include Scott's proof

@@
 \title{A Variant of G\"{o}del's Ontological Proof \\
-in a Natural Deduction Calculus \\
-(Draft)}
+in a Natural Deduction Calculus}
 
@@
   \and 
-  Vienna, Austria \\
-  \email{bruno@logic.at}
+  Vienna, Austria ; Canberra, Australia \\
+  \email{bruno.wp@gmail.com}
 }
@@
 \noindent
+\begin{footnotesize}
+\begin{center}
 ``There is a scientific (exact) philosophy and theology,
-which deals with concepts of the highest abstractness; and this is also most highly fruitful for science. [\ldots] Religions are, for the most part, bad; but religion is not.'' - Kurt G\"{o}del
+which deals with concepts of the highest abstractness; and this is also most highly fruitful for science. [\ldots] \\Religions are, for the most part, bad; but religion is not.'' \\
+- Kurt G\"{o}del \citep{Wang1996}[p. 316]
+\end{center}
+\end{footnotesize}
+
+%Begin with a focus on more recent stuff such as Gödel, not Anselm and so on. Focus on the improved level of formalization of the argument through Leibniz/Descartes/Kant and so on. 
+%Think about how much can be claimed about Russell, Descartes, Kant and so on... 
+%Russell eliminated or expained as a valuable contribution to the formal proof. 
+%check the Solmon paper to se about proofs in KB. Other ontological proofs. 
+%Relate introduction to the beginning quote
+%spell check
+%How to use the words God, God-like, god-like and greatest conceivable being. 
+%Theorem reference missing in the proof of modal collapse
+%Write the section on axiom B and the strength of the modal system that is necessary.
 
@@
 Anselm (1033-1109).
-His argument considered a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
-
-A major critique of this argument is that we do not know whether the concept greatest conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
+His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
+
+A major critique of this argument is that we do not know whether the concept of greatest conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
 %The question is if the argument is sound. 
-As Bertrand Russell pointed out, the definition of greatest allows us to define properties, like having boots, which the greatest being then also must have \citep{citation needed}.
+As Bertrand Russell pointed out, the definition of greatest conceivable being allows us to define properties, like having boots, which the being then also must have \citep{citation needed}.
 
@@
 This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
-All that we can say is that if God exists, then God necessarily exists \citep{citation needed}.
+The strongest claim that can be proven by the ontological argument is a conditional claim of necessary existence: if God exists, then God necessarily exists \citep{citation needed}.
 
 St.
-Anselm's argument was elaborated further by Descartes and Leibniz.
-Leibniz identified that establishing the possible existence of God is a critical missing step in St.
+Anselm's argument was further elaborated by Descartes and Leibniz.
+Leibniz identified the possible existence of God as a critical missing step in St.
 Anselm's argument.
 To fill this gap, he argued that the properties of God, the perfections, are compatible.
-This implies that it is possible to have all perfections at once and therefore the existence of a greatest being with all these properties is possible.
+This means that it is possible to satisfy all perfections at once which implies that
+the existence of a greatest conceivable being with all these properties is possible.
 
@@
 
-G\"odel's notion of positive property and Leibniz's notion of perfection are not exactly the same \citep{Adams}.
-On a technical level, the main distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim).
-In particular, one of G\"odel's axioms states that the conjunction of any set of positive properties is positive.
-And from this axiom, it follows immediately that the property of being God-like is positive.
-While this step is intuitively and informally clear, it is not easily formalizable in a typical logical calculus, because it requires inferring that being god-like is a (possibly infinite) conjuntion of positive properties, while it has only been defined as the property holding for individuals who have all positive properties.
-This interplay between universal quantification (in the definition of a God-like being) and conjunction (in G\"odel's mentioned axiom) is a technical difficulty that probably explains why, starting with Scott \cite{scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another one that simply assumes the positivity of the property of being God-like.
-
-The main criticism against G\"odel's ontological argument is the \emph{modal collapse}, an undesirable consequence of the stipulated axioms.
-Most recent works \cite{todo} on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
-This is discussed in more detail in Section \ref{sec:collapse}.
+G\"odel's notion of positive property and Leibniz's notion of perfection differ \citep{Adams}.
+The main formal distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim).
+In particular, one of G\"odel's axioms states that any conjunction of positive properties is itself positive.
+From this axiom, it is immediately deduced that the property of being God-like is positive.
+While this step is intuitively and informally clear, it is not easily formalizable in a standard logical calculus, because it requires inferring that being God-like is a (possibly infinite) conjunction of positive properties, from the universal definition of God-likeness as the property satisfied by individuals who have all positive properties.
+This interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) is a technical difficulty that probably explains why, starting with Scott \cite{Scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another that simply assumes the positivity of the property of being God-like.
+
+The main criticism against G\"odel's ontological argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
+Many recent works \cite{todo} on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
+This is discussed in greater detail in Section \ref{sec:collapse}, but possible solutions are not implemented in this paper.
 
@@
 
-The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
-A natural deduction style \citep{gentzen,prawitz} was chosen mainly for three reasons.
+The aim of this paper is to present two detailed formal proofs of G\"odel's ontological argument \cite{Goedel} in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
+The natural deduction style \citep{gentzen,prawitz} was chosen for three reasons.
 Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of calculi (e.g.
@@
 Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
-Finally, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.
-
-The main contribution of the paper, however, is the presentation of new proofs (also in natural deduction style) of the lemmas, theorems and corollaries in G\"odel's manuscript.
+Thirdly, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.
+
+The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus.
+The second contribution of the paper is a new proof (also in natural deduction style) of the lemmas, theorems and corollaries in G\"odel's manuscript.
 In contrast to Scott's proofs \cite{Scott}, the proofs presented here are simpler and shorter, as discussed in Section \ref{sec:conclusion}.
@@
 
-Adding the modal rules results in a calculus that is suitable for the basic normal modal logic \textbf{K}.
+Adding the modal rules results in a calculus that is suitable for the basic modal logic \textbf{K}.
 In other words, {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
@@
 \begin{prooftree}
-\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec_E$}
+\AXC{$ $}\RightLabel{2}
+\UIC{$\nec(A\imp B)$}\RightLabel{$\nec_E$}
 \UIC{$\omega: A\imp B$}
-      \AXC{$\nec A ^1 $}\RightLabel{$\nec_E$}
+      \AXC{$ $}\RightLabel{1}
+      \UIC{$\nec A $}\RightLabel{$\nec_E$}
       \UIC{$\omega: A$} \RightLabel{$\imp_E$}
@@
 
-ToDo: show example of unsound derivation.
+\begin{prooftree}
+  \AXC{$ $} \RightLabel{1}
+  \UIC{$\nec \psi$} \RightLabel{$\nec_E$}
+        \UIC{$\omega: \psi$} \RightLabel{$\pos_I$}
+        \UIC{$\pos \psi$} \RightLabel{$\imp_I^1$}
+    \UIC{$\nec \psi \imp \pos \psi$} \RightLabel{$\forall_I$}
+    \UIC{$\all \psi. (\nec \psi \imp \pos \psi)$}
+\end{prooftree}
+
+This example of an unsound derivation is prevented by the eigen-box condition because the box labelled by $\omega$ is not accessed by any strong inference.
 
@@
 
-\section{A distribution principle and axiom B}
-
-To prove G\"odel's ontological proof in natural deduction we will prove that a distribution principle is derivable and state Brouwer's reduction theorem.
+\section{Some Derivable Modal Principles}
+\label{sec:AxiomB}.
+
+As mentioned in \citep{and90}[Footnote 5], there are sources that discuss what modal principles are needed for the ontological argument.
+The standard system for the argument is S5.
+Through an analysis of the proofs weaker systems, in particular KB, can be shown to be sufficient for versions of the ontological argument \citep{sobel2}[p.
+152].
+The proof presented in Section \ref{sec:newproof} is such a version and explicitly shows a single use of Brouwer Reduction Principle, which, in turn, is provable from axiom B.
+All other proof steps are formalizable in a basic second-order modal logic K.
+
+In standard first-order modal logics the modal axioms correspond to geometric conditions on the frames of Kripke models \citep{negri}.
+The modal axiom {\bf T} corresponds to reflexivity of the frame and {\bf B} corresponds to symmetry of the frame.
+The frames of the modal logic {\bf S5} are in addition transitive.
+
+%G\"odel mentions in the notes which are saved of his proof that the modal system required for proving his theorem is a second order version of the modal system {\bf S5}.
+In \citep{and90} it is conjectured that the principles of proof could be restricted to the weaker modal logic {\bf KB}.
+The modal logic {\bf KB} consists of the basic modal system {\bf K} and the modal axioms {\bf T} and {\bf B}.
+
+ToDo: Sobel Anderson on KB: Sobel page 152
+and
+Anderson's footnote 5 on KB %Nathan Salmon: On the Logic of What Might Have Been,"
+
+\subsection{A modal distribution principle in K}
+
+To prove G\"odel's ontological proof in natural deduction we use a modal distribution principle derivable in K and  Brouwer's reduction theorem derivable from the modal Axiom B.
 
@@
 \begin{prooftree}
-\AXC{$ $} \dashedLine\RightLabel{2}
+\AXC{$ $} \RightLabel{2}
 \UIC{$\nec (A\imp B)$}\RightLabel{$\nec_E$}
 \UIC{$\omega: A\imp B$}
-\AXC{$ $} \dashedLine\RightLabel{1}
+\AXC{$ $} \RightLabel{1}
 \UIC{$\pos A$}\RightLabel{$\pos_E$}
@@
 
-Brouwer's reduction theorem is a version of axiom B for modal logic and a property that enforces symmetry as a geometric property for the frame of a Kripke sematics for modal logic. \citep{negri}
-%See the following for a discussion about axiom B: http://plato.stanford.edu/entries/logic-modal/
-
-\section{Scott's Proof in Natural Deduction}
+\subsection{Axiom B and Brouwer's Reduction Principle}
+
+Brouwer's reduction theorem is a version of axiom B for modal logic and a property that enforces symmetry as a geometric property for the frame of a Kripke sematics.
+
+\begin{lemma}\label{BRP}
+The Brouwer Reduction Principle $\pos\nec A\imp A$ is derivable from axiom B, $A\imp \nec \pos A$. 
+\end{lemma}
+
+\begin{proof}\hfill
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $}\RightLabel{$2$}
+\UIC{$\neg\nec \neg \nec A$}
+\AXC{$ $}\RightLabel{$ 1$}
+\UIC{$\neg A$}
+\AXC{Axiom B}\dashedLine
+\UIC{$\neg A\imp \nec\neg\nec\neg\neg A$}\RightLabel{$\imp_E$}
+\BIC{$\nec\neg\nec\neg\neg A$}\doubleLine
+\UIC{$\nec\neg\nec A$}
+\BIC{$(\neg\nec \neg \nec A)\&(\nec\neg \nec A)$}\RightLabel{$\neg_I^1$}
+\UIC{$\neg\neg A$} \RightLabel{$\neg\neg_E$}
+\UIC{$A$} \RightLabel{$\imp_I^2$}
+\UIC{$\neg\nec \neg \nec A\imp A$}
+\end{prooftree}
+\end{small}
+
+\end{proof}
+
+\subsection{S5 and its Iteration Principle}
+
+ToDo: Is this principle really commonly called ``S5 Iteration Principle''?
+
+\begin{lemma}
+What is possibly necessary is necessary:
+$$
+\pos \nec A \imp \nec A
+$$
+\end{lemma}
+
+\begin{proof}
+
+\begin{prooftree}
+\AXC{ToDo}
+\end{prooftree}
+
+\end{proof}
+
+\section{Scott's Proof in Natural Deduction}\label{sec:scott}
 
@@
 \setcounter{corollary}{0}
+\setcounter{definition}{0}
 
@@
 The formal argument of G\"odel is based on Leibniz proof, which in turn is based on Descartes proof.
-These proofs have two parts; a proof that if God's existence is possible, then it is necessary and a proof that God's existence is in fact possible.
+These proofs have two parts; a proof that if the exemplification of the god-likeness property is possible, then it is necessary and a proof that the exemplification is in fact possible.
 
@@
 
-\section{New Proof}
+\begin{theorem}
+\label{T1}
+Positive properties are possibly exemplified:
+$$
+P(\varphi) \imp \pos \ex x.\varphi(x)
+$$
+\end{theorem}
+\begin{proof} \hfill
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{3}
+\UIC{$P(\rho)$}
+		\AXC{Reflexivity} \dashedLine
+		\UIC{$\omega: \gamma = \gamma$} \RightLabel{$\imp_I$}
+		\UIC{$\omega: \rho{\gamma} \imp \gamma = \gamma$} \RightLabel{$\all_I$}
+		\UIC{$\omega: \all x. \rho(x) \imp x = x$} \RightLabel{$\nec_I$}
+		\UIC{$\nec (\all x. \rho(x) \imp x = x) $} \RightLabel{$\wedge_I$}
+	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) $} 
+			\AXC{Axiom 2 for $\rho$ and $\lambda x. x = x$} \dashedLine
+	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) \imp P(\lambda x. x = x)$}
+
+		\BIC{$P(\lambda x. x = x)$} 
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{3}
+\UIC{$P(\rho)$}
+	\AXC{$ $} \RightLabel{8}
+	\UIC{$\omega: \rho(\beta)$}
+			\AXC{$ $} \RightLabel{5}
+			\UIC{$\nec \all x. \neg \rho(x)$} \RightLabel{$\nec_E$}
+			\UIC{$\omega: \all x. \neg \rho(x) $} \RightLabel{$\all_E$}
+			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\neg_E$}
+		\BIC{$\omega: \bot$} \RightLabel{$\neg_I^9$}
+		\UIC{$\omega: \neg (\beta = \beta)$} \dottedLine
+		\UIC{$\omega: (\beta \neq \beta)$} \RightLabel{$\imp_I^8$}
+		\UIC{$\omega: (\rho(\beta) \imp (\beta \neq \beta))$} \RightLabel{$\all_I$}
+		\UIC{$\omega: (\all x. \rho(x) \imp (x \neq x))$} \RightLabel{$\nec_I$}
+		\UIC{$\nec (\all x. \rho(x) \imp (x \neq x))$}  \RightLabel{$\wedge_I$}
+	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) $} 
+			\AXC{Axiom 2 for $\rho$ and $\lambda x. x \neq x$} \dashedLine
+	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) \imp P(\lambda x. x \neq x)$} 
+		\BIC{$P(\lambda x. x \neq x)$} 
+\end{prooftree}
+
+\begin{prooftree}
+		\AXC{$ $} \dashedLine
+		\UIC{$P(\lambda x. x \neq x)$}    
+    		\AXC{$ $} \dashedLine
+			\UIC{$P(\lambda x. x = x)$} 
+					\AXC{Half of Axiom 1} \dashedLine
+					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
+				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
+			\BIC{$\bot$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{2}
+\UIC{$\box \neg \ex x. \rho(x)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: \neg \ex x. \rho(x)$}
+        \AXC{$ $} \RightLabel{4}
+        \UIC{$\rho(\alpha)$} \RightLabel{$\ex_I$}
+		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
+    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^4$}
+	\UIC{$\omega: \neg \rho(\alpha)$}
+	\UIC{$\omega: \all x. \neg \rho(x)$}
+	\UIC{$\nec \all x. \neg \rho(x)$}
+	        \AXC{$ $} \dashedLine
+	        \UIC{$\bot$} \RightLabel{$\neg_I^5$}
+    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
+  		\BIC{$\bot$} \RightLabel{$\neg_I^2$}
+  		\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
+  		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^3$}
+  		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
+
+\end{prooftree}
+
+\end{proof}
+
+\begin{definition}
+\label{D1}
+A \emph{god-like} being possesses all positive properties:
+$$
+G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+$$
+\end{definition}
+
+\begin{axiom}
+\label{A3}
+The property of being god-like is positive:
+$$
+P(G)
+$$
+\end{axiom}
+\begin{corollary}
+\label{C1}
+Possibly, God exists:
+$$
+\pos \ex x. G(x)
+$$
+\end{corollary}
+\begin{proof} \hfill
+\begin{prooftree}
+\AXC{Axiom \ref{A3}} \dashedLine
+\UIC{$P(G)$}
+\AXC{Theorem \ref{T1} for G} \dashedLine
+\UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
+\BIC{$\pos \ex x. G(x)$}
+\end{prooftree}
+\end{proof}
+
+\begin{axiom}
+\label{A4}
+Positive properties are necessarily positive:
+$$
+\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+$$
+\end{axiom}
+
+\begin{definition}
+\label{D2}
+An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties:
+$$
+\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
+$$
+\end{definition}
+
+\begin{theorem}
+\label{T2}
+Being God-like is an essence of any God-like being:
+$$
+\all x. G(x) \imp \ess{G}{x}
+$$
+\end{theorem}
+\begin{proof}
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$G(\gamma)$}
+
+	\AXC{$ $} \RightLabel{2}
+	\UIC{$\rho(\gamma)$}
+
+		\AXC{$ $} \RightLabel{4}
+		\UIC{$\neg P(\rho)$} 
+				\AXC{Axiom 1 for $\rho$} \RightLabel{$\imp_E$}
+			\BIC{$P(\lambda x. \neg \rho(x))$}
+
+					\AXC{$ $} \RightLabel{1}
+					\UIC{$G(\gamma)$} \dottedLine
+					\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
+					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
+				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
+	   \BIC{$\bot$} \RightLabel{$\neg_I^4$}
+	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
+	   \UIC{$P(\rho)$}
+	      \AXC{Axiom 4} \dashedLine
+	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
+		 \BIC{$\nec P(\rho)$}
+
+			\AXC{$ $} \RightLabel{5}
+			\UIC{$\omega: P(\rho)$}
+
+					\AXC{$ $} \RightLabel{6}
+					\UIC{$\omega: G(\delta) $} \dottedLine
+					\UIC{$\omega: \all \varphi. P(\varphi) \imp \varphi(\delta)$} \RightLabel{$\all_E$} 
+					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
+				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^6$}
+				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
+				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^5$}
+				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$}
+				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$}
+						\AXC{K} \RightLabel{$\imp_E$}
+					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$}
+			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I$}
+			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
+			\UIC{$\all q. q(\gamma) \imp \nec (\all y. G(y) \imp q(y) )$} \RightLabel{$\wedge_I$}
+		\BIC{$ G(\gamma) \wedge (\all q. q(\gamma) \imp \nec (\all y. G(y) \imp q(y) ))$} \dottedLine
+		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^1$}
+		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
+		\UIC{$\all x. G(x) \imp \ess{G}{x}$}
+\end{prooftree}
+
+\end{proof}
+
+\begin{definition}
+\label{D3}
+\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+$$
+E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
+$$
+\end{definition}
+
+\begin{axiom}
+\label{A5}
+Necessary existence is a positive property:
+$$
+P(E)
+$$
+\end{axiom}
+
+\begin{lemma}
+\label{L1}
+Exemplification of the god-likeness property implies the exemplification is necessary:
+$$
+\ex z. G(z) \imp \nec \ex x. G(x)
+$$
+\end{lemma}
+
+\begin{proof}
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
+\UIC{$G(\gamma)$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$ $} \dashedLine
+\UIC{$G(\gamma)$}
+		\AXC{Theorem 2} \dashedLine 
+		\UIC{$\all x. G(x) \imp \ess{G}{x}$} \RightLabel{$\all_E$}
+		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}
+	\BIC{$\ess{G}{\gamma}$}
+				\AXC{Axiom 5} \dashedLine
+				\UIC{$P(E)$}
+						\AXC{$ $} \dashedLine
+						\UIC{$G(\gamma)$} \dottedLine
+						\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
+						\UIC{$P(E) \imp E(\gamma)$} 
+					\BIC{$E(\gamma)$} \dottedLine
+					\UIC{$\all \varphi. \ess{\varphi}{\gamma} \imp \nec \ex x. \varphi(x)$} \RightLabel{$\all_E$}
+					\UIC{$\ess{G}{\gamma} \imp \nec \ex x. G(x)$}
+		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
+		\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
+
+\end{prooftree}
+
+\end{proof}
+
+\begin{lemma}
+\label{L2}
+If the existence of a God-like being is possible, then it is necessary:
+$$
+\pos \ex z. G(z) \imp \nec \ex x. G(x)
+$$
+\end{lemma}
+
+\begin{proof}
+
+\begin{footnotesize}
+\begin{prooftree}
+
+\AXC{$ $} \RightLabel{1}
+\UIC{$\pos \ex x. G(x)$} 
+		\AXC{Lemma 1} \dashedLine
+		\UIC{$\omega: \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\nec_I$}
+		\UIC{$\nec (\ex x. G(x) \imp \nec \ex x. G(x)) $}
+				\AXC{Distribution Principle}
+			\BIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
+	\BIC{$\pos \nec \ex x. G(x)$}
+			\AXC{S5 Iteration Principle} \dashedLine
+			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}
+		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
+		\UIC{$\pos \ex z. G(z) \imp \nec \ex x. G(x)$}
+\end{prooftree}
+\end{footnotesize}
+
+\end{proof}
+
+\begin{theorem}
+\label{T3}
+Necessarily, there exists a God-like being:
+$$
+\nec \ex x. G(x)
+$$
+\end{theorem}
+
+\begin{proof}
+
+\begin{prooftree}
+	\AXC{Corollary 1} \dashedLine
+	\UIC{$\pos \ex x. G(x)$}
+			\AXC{Lamma 2} \dashedLine
+			\UIC{$\pos \ex x. G(x) \imp \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
+		\BIC{$\nec \ex x. G(x)$}
+\end{prooftree}
+
+\end{proof}
+
+\section{New Proof}\label{sec:newproof}
+
+A outline of the proof is that firstly all positive properties are proven to be possibly exemplified (Theorem \ref{T1}).
+Then this theorem is applied to the property of god-likeness, thus showing the possible existence of a god-like being (Corollary \ref{C1}).
+In Lemma \ref{L1} it is classically proven that the hypothetically assumed existence of a god-like being implies that the existence is necessary.
+Lastly, the necessary existence of a god-like being is proven (Theorem \ref{T2}) from Corollary \ref{C1} and Lemma \ref{L1} by properties of the modal system.
+The properties required for the proof are the distribution principle and Brouwer's reduction theorem mentioned in Section \ref{sec:AxiomB}.
+The proof of Theorem \ref{T2} follows the proof of Sobel \citep{sobel2}[p.
+126--127].
+
 \setcounter{axiom}{0}
@@
 \setcounter{corollary}{0}
+\setcounter{definition}{0}
+
 \subsection{Possibly, God Exists}
@@
 \label{T1}
+
 Positive properties are possibly exemplified:
@@
 \begin{prooftree}
-
-\AXC{$[P(\rho)]^3$} 
-\AXC{$[\nec \neg \ex x.\rho(x) ]^2$} \RightLabel{$\nec_E$}
-\UIC{$\neg \ex x.\rho(x) $} 
-\AXC{$[\rho(x)]^1$} \RightLabel{$\ex_I$}
-\UIC{$\ex x.\rho(x) $} \RightLabel{$\imp_E$}
-\BIC{$ \bot $}\RightLabel{$\imp_I, 1$}
-\UIC{$ \neg \rho(x) $} \RightLabel{$\imp_I$}
-\UIC{$ \rho(x) \imp \neg \rho(x) $} \RightLabel{$\all_I$}
-\UIC{$ \all x.(\rho(x) \imp \neg \rho(x)) $}  \RightLabel{$\nec_I$}
+\AXC{$ $} \RightLabel{3}
+\UIC{$P(\rho)$} 
+\AXC{$ $} \RightLabel{2}
+\UIC{$\nec \neg \ex x.\rho(x) $} \RightLabel{$\nec_E$}
+\UIC{$\omega: \neg \ex x.\rho(x) $} 
+\AXC{$ $} \RightLabel{$1$}
+\UIC{$\omega: \rho(x)$} \RightLabel{$\ex_I$}
+\UIC{$\omega: \ex x.\rho(x) $} \RightLabel{$\imp_E$}
+\BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
+\UIC{$\omega:  \neg \rho(x) $} \RightLabel{$\imp_I$}
+\UIC{$\omega:  \rho(x) \imp \neg \rho(x) $} \RightLabel{$\all_I$}
+\UIC{$\omega:  \all x.(\rho(x) \imp \neg \rho(x)) $}  \RightLabel{$\nec_I$}
 \UIC{$ \nec \all x.(\rho(x) \imp \neg \rho(x)) $}\RightLabel{$\wedge_I$}
@@
 \UIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
-\AXC{Half of axiom \ref{A1}} \dashedLine
+\AXC{Half of Axiom \ref{A1}} \dashedLine
 \UIC{$ P(\neg \rho) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
 \BIC{$\neg P(\rho) $} 
-\AXC{$[P(\rho)]^3$} 
-\BIC{$\bot$} \RightLabel{$\imp_I, 2$}
-\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \dashedLine \RightLabel{definition of $\pos$}
-\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I, 3$}
+\AXC{$ $} \RightLabel{3} 
+\UIC{$P(\rho)$} 
+\BIC{$\bot$} \RightLabel{$\imp_I^2$}
+\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
+\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^3$}
 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
@@
 \end{proof}
+
 \begin{definition}
 \label{D1}
-A \emph{God-like} being possesses all positive properties:
+A \emph{god-like} being possesses all positive properties:
 $$
@@
 \label{A3}
-The property of being God-like is positive:
+The property of being god-like is positive:
 $$
@@
 
+\begin{axiom}
+\label{A4}
+Positive properties are necessarily positive:
+$$
+\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+$$
+\end{axiom}
+
 \begin{definition}
@@
 \label{L1}
-If there is a God, then necessarily there exists a God:
+Exemplification of the god-likeness property implies the exemplification is necessary:
 $$
@@
 
+\begin{small}
 \begin{prooftree}
@@
 \BIC{$P(\neg \psi)$}
-\AXC{$G(g)$}  \dashedLine \RightLabel{D\ref{D1}}
+\AXC{$G(g)$}  \dottedLine \RightLabel{D\ref{D1}}
 \UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
@@
 \BIC{$\neg \psi(g)$}
-\AXC{$ $} \RightLabel{4}
+\AXC{$ $} \RightLabel{3}
 \UIC{$\psi(g)$} \RightLabel{$\imp_E$}
@@
 \UIC{$P(\psi)$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$P(\psi)$}
-\AXC{$ $} \RightLabel{3}
-\UIC{$G(y)$}  \dashedLine \RightLabel{D\ref{D1}}
-\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
-\UIC{$P(\psi) \imp \psi(y)$}\RightLabel{$\imp_E$}
-\BIC{$\psi(y)$} \RightLabel{$\imp_I^3$}
-\UIC{$G(y) \imp \psi(y)$}\RightLabel{$\forall_I$}
-\UIC{$\forall y.(G(y) \imp \psi(y))$}\RightLabel{$\nec_I$}
-\UIC{$\nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\imp_I^4$}
+\AXC{Axiom \ref{A4}} \dashedLine
+\UIC{$\all \varphi.[P(\psi) \imp \nec\; P(\psi)]$}\RightLabel{$\forall_E$}
+\UIC{$P(\psi) \imp \nec\; P(\psi)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\psi)$}\RightLabel{$\imp_I^3$}
+\UIC{$\psi(g) \imp \nec\; P(\psi)$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{5}
+\UIC{$\nec P(\psi)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: P(\psi)$}
+\AXC{$ $} \RightLabel{4}
+\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
+\UIC{$\omega: P(\psi) \imp \psi(y)$}\RightLabel{$\imp_E$}
+\BIC{$\omega: \psi(y)$} \RightLabel{$\imp_I^4$}
+\UIC{$\omega: G(y) \imp \psi(y)$}\RightLabel{$\forall_I$}
+\UIC{$\omega: \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\nec_I$}
+\UIC{$\nec \forall y.(G(y) \imp \psi(y))$} \RightLabel{$\imp_I^5$}
+\UIC{$\nec P(\psi)\imp \nec \forall y .(G(y)\imp \psi(y))$} \RightLabel{$\imp_E$}
+\AXC{$ $} \RightLabel{6}
+\UIC{$\psi(g)$} \RightLabel{}
+\AXC{$ $} \dashedLine
+\UIC{$\psi(g)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\psi)$} \RightLabel{$\imp_E$}
+\BIC{$\nec \forall y .(G(y)\imp \psi(y))$}\RightLabel{$\imp_I^6$}
 \UIC{$\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\forall_I$}
@@
 \AXC{$G(g)$}  
-\BIC{$G(g) \wedge \forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dashedLine \RightLabel{D\ref{D2}}
+\BIC{$G(g) \wedge \forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
+\UIC{$\ess{G}{g}$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{ } \dashedLine
 \UIC{$\ess{G}{g}$}
@@
 \UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
-\AXC{$G(g)$}\dashedLine \RightLabel{D\ref{D1}}
+\AXC{$G(g)$}\dottedLine \RightLabel{D\ref{D1}}
 \UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
@@
 \end{prooftree}
+\end{small}
 \end{proof}
@@
 
-ToDo: this is proven in a way that is slightly different from G\"odel's 1970.
+The exemplification of the god-likeness property is provable from its possible exemplification by the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP}.
+The necessary exemplification of the god-likeness property can then be deduced from lemma \ref{L1}.
+This single use of the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP} follows the presentation of in Sobel \citep{sobel2}[p.126--127].
+
+\setcounter{theorem}{2}
 \begin{theorem}
-\label{T4}
-God necessarily exists:
+\label{T3}
+The property of god-likeness is necessarily exemplified:
 $$
@@
 \end{theorem}
+
+%Is the distribution principle and brouwers Red thm better than using axiom B straight? Check previous versions... ***
+
+%From that it is possible to that it exists with modal inferences... **** and axioms...
+
 \begin{proof}
@@
 \AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
-\AXC{ Theorem \ref{T1} } \dashedLine
+\AXC{ Corollary \ref{C1} } \dashedLine
 \UIC{$\pos \ex x. G(x)$}\RightLabel{$\imp_E$}
 \BIC{$\pos\nec \ex x. G(x)$}
-\AXC{ Brouwer's reduction theorem \ref{***} } \dashedLine
+\AXC{ Brouwer's reduction principle \ref{BRP} } \dashedLine
 \UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
@@
 
-\section{Modal collapse}
+\section{Modal collapse} \label{sec:collapse}
 
@@
 4,  section 6, theorems 9 and 10].
-As this is an undesirable consequence, many solutions to the problem of the modal collapse have been proposed.
-
-Anderson's solution \citep{citation needed} modifies the definitions of God-like being and essence, and eliminates half of an axiom.
+That is: for all properties, $\varphi$,
+$$\varphi \biimp \pos \varphi \biimp \nec \varphi
+$$
+If G\"odel's ontological proof is abstractly analysed, then we are proving a restricted modal collapse, which applies to one specific formula, the exemplification of god-likeness.
+In addition we prove and that the definition of god-likeness is sound, which means that it is possibly exemplified.
+Thus, the necessity of the formula follows.
+The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for all formulas.
+Therefore, an improvement would be obtained if the modal collapse was limited to one property, namely the property of god-likeness.
+A number of solutions to the problem of the modal collapse have been proposed.
+
+Anderson's solution \citep{citation needed} modifies the definitions of god-like being and essence, and eliminates half of an axiom.
 This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
@@
 
-Anderson \citep{anderson1990}[p.
+Anderson \citep{and90}[p.
 292] and Sobel \citep{sobel2}[p.
@@
 
-%In his manuscript, G\"odel warned that ``positive means positive in the moral aesthetic sense [\ldots]. Only then [are] the axioms true''.
-
 It is beyond the scope of this paper to analyze these solutions in detail or propose new solutions.
@@
 \begin{theorem}
-\label{T5}
-For all constant fomulas (without free variables), $A$, we have:
+\label{T3}
+For all constant formulas, $A$, (without free variables) a modal collapse 
 $$
@@
 $$
+is provable. 
 \end{theorem}
 
-Note that in intuitionistic predicate logic we have $\all y.[B \imp C] \biimp [\ex y.B \imp C]$ if $y$ is not free in $C$.
-
 \begin{small}
 \begin{prooftree}
-\AXC{Theorem \ref{T2}}\dashedLine
-\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }
-\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \RightLabel{Prop. logic }
-\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \RightLabel{Second order quantifier elimination and logical steps }
+\AXC{Theorem \ref{***}}\dashedLine
+\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
+\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}
+\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
 \UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
-\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \RightLabel{intuitionistic predicate logic}
+\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
@@
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\AXC{Theorem \ref{T4}}\dashedLine
+\AXC{Theorem \ref{T2}}\dashedLine
+\UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
 \UIC{$\ex y.G(y)$} \RightLabel{$\imp_E$}
-\BIC{$ A \imp \nec \all x. (G(x) \imp A)$} \RightLabel{intuitionistic predicate logic}
+\BIC{$ A \imp \nec \all x. (G(x) \imp A)$} \doubleLine
 \UIC{$ A \imp \nec (\ex x. G(x) \imp A)$}
@@
 \BIC{$ \nec (\ex x. G(x) \imp A)$} \RightLabel{$\nec_E$}
-\UIC{$ \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
-\AXC{Theorem \ref{T4}}\dashedLine
-\UIC{$\ex x.G(x)$} \RightLabel{$\imp_E$}
-\BIC{$A$} \RightLabel{$\nec_I$}
+\UIC{$ \omega:  \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
+\AXC{Theorem \ref{T2}}\dashedLine
+\UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: \ex x.G(x)$} \RightLabel{$\imp_E$}
+\BIC{$\omega: A$} \RightLabel{$\nec_I$}
 \UIC{$ \nec A$} \RightLabel{$\imp_I^1$}
@@
 \end{small}
-
-% This collapse of the system was first proved in \citep{sobel} and the argument is also given in \citep{fitting}[ch.
-11, sections 8-11] who provides summaries of two solutions by the modifications of the axioms.
-The argument may require that one doesn't instantiate the second order quantifier with $A$ directly but with $\lambda x.A$ to make $A$ a property.
-But still we have $(\lambda x.A)(y)\equiv A$ because $A$ is constant.
 
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusion}
 
@@
 
+\bibitem[{\itshape Goedel(***)}]{Goedel} 
+G\"odel, K. {\itshape Appendix A. Notes in Kurt G\"odel's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
+
 \bibitem[{\itshape Kant(1781)}]{kant}
@@
 
+\bibitem[{\itshape Scott (***)}]{Scott}
+Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
+
 \bibitem[{\itshape Sobel(1987)}]{sobel}
@@
 
+\bibitem[{\itshape Wang(1996)}]{Wang1996}
+Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
+
+***Gentzen and Prawitz: references to Natural Deduction. 
+Adams, Fuhrmann, Hajek and Bjordal: References to amendations. 
+Church: Reference to simple type theory.
+
 \end{thebibliography}

### 2015-09-14 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Commit Annika's improvements

@@
 While this step is intuitively and informally clear, it is not easily formalizable in a standard logical calculus, because it requires inferring that being God-like is a (possibly infinite) conjunction of positive properties, from the universal definition of God-likeness as the property satisfied by individuals who have all positive properties.
-This interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) is a technical difficulty that probably explains why, starting with Scott \cite{Scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another that simply assumes the positivity of the property of being God-like.
+This interplay between universal quantification (in the definition of a God-like being) and in!
+ finite conjunctions (in G\"odel's axiom for positive properties) is a technical difficulty that probably explains why, starting with Scott \cite{Scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another that simply assumes the positivity of the property of being God-like.
 
@@
 \begin{prooftree}
-\AXC{$ $} \RightLabel{3}
+\AXC{$ $} \RightLabel{5}
 \UIC{$P(\rho)$}
@@
 		\UIC{$\omega: \gamma = \gamma$} \RightLabel{$\imp_I$}
-		\UIC{$\omega: \rho{\gamma} \imp \gamma = \gamma$} \RightLabel{$\all_I$}
+		\UIC{$\omega: \rho(\gamma) \imp \gamma = \gamma$} \RightLabel{$\all_I$}
 		\UIC{$\omega: \all x. \rho(x) \imp x = x$} \RightLabel{$\nec_I$}
@@
 \begin{prooftree}
+\AXC{$ $} \RightLabel{5}
+\UIC{$P(\rho)$}
+	\AXC{$ $} \RightLabel{1}
+	\UIC{$\omega: \rho(\beta)$}
 			\AXC{$ $} \RightLabel{3}
-\UIC{$P(\rho)$}
-	\AXC{$ $} \RightLabel{8}
-	\UIC{$\omega: \rho(\beta)$}
-			\AXC{$ $} \RightLabel{5}
 			\UIC{$\nec \all x. \neg \rho(x)$} \RightLabel{$\nec_E$}
@@
 		\UIC{$\omega: \neg (\beta = \beta)$} \dottedLine
-		\UIC{$\omega: (\beta \neq \beta)$} \RightLabel{$\imp_I^8$}
+		\UIC{$\omega: (\beta \neq \beta)$} \RightLabel{$\imp_I^1$}
 		\UIC{$\omega: (\rho(\beta) \imp (\beta \neq \beta))$} \RightLabel{$\all_I$}
@@
 \begin{prooftree}
+\AXC{$ $} \RightLabel{4}
+\UIC{$\nec \neg \ex x. \rho(x)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: \neg \ex x. \rho(x)$}
         \AXC{$ $} \RightLabel{2}
-\UIC{$\box \neg \ex x. \rho(x)$} \RightLabel{$\nec_E$}
-\UIC{$\omega: \neg \ex x. \rho(x)$}
-        \AXC{$ $} \RightLabel{4}
         \UIC{$\rho(\alpha)$} \RightLabel{$\ex_I$}
 		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
-    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^4$}
-	\UIC{$\omega: \neg \rho(\alpha)$}
-	\UIC{$\omega: \all x. \neg \rho(x)$}
+    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^2$}
+	\UIC{$\omega: \neg \rho(\alpha)$} \RightLabel{$\all_I$}
+	\UIC{$\omega: \all x. \neg \rho(x)$} \RightLabel{$\nec_I$}
 	\UIC{$\nec \all x. \neg \rho(x)$}
 	        \AXC{$ $} \dashedLine
-	        \UIC{$\bot$} \RightLabel{$\neg_I^5$}
+	        \UIC{$\bot$} \RightLabel{$\neg_I^3$}
     		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
-  		\BIC{$\bot$} \RightLabel{$\neg_I^2$}
+  		\BIC{$\bot$} \RightLabel{$\neg_I^4$}
   		\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
-  		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^3$}
+  		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^5$}
   		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
@@
 
-\begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$G(\gamma)$}
+\begin{small}
+\begin{prooftree}
+
+	\AXC{$ $} \RightLabel{5}
+	\UIC{$\rho(\gamma)$}
 
 		\AXC{$ $} \RightLabel{2}
-	\UIC{$\rho(\gamma)$}
-
-		\AXC{$ $} \RightLabel{4}
 		\UIC{$\neg P(\rho)$} 
@@
 				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
-	   \BIC{$\bot$} \RightLabel{$\neg_I^4$}
+	   \BIC{$\bot$} \RightLabel{$\neg_I^2$}
 	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
+	   \UIC{$P(\rho)$}
+	   \end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$G(\gamma)$}
+
+ \AXC{$ $}\dashedLine
   \UIC{$P(\rho)$}
@@
 		 \BIC{$\nec P(\rho)$}
-
-			\AXC{$ $} \RightLabel{5}
+			\AXC{$ $} \RightLabel{4}
 			\UIC{$\omega: P(\rho)$}
-
-					\AXC{$ $} \RightLabel{6}
+					\AXC{$ $} \RightLabel{3}
 					\UIC{$\omega: G(\delta) $} \dottedLine
@@
 					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
-				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^6$}
+				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^3$}
 				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
-				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^5$}
-				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$}
-				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$}
-						\AXC{K} \RightLabel{$\imp_E$}
+				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^4$}
+				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
+				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
+						\AXC{ K} \RightLabel{$\imp_E$}
 					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$}
-			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I$}
+			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^5$}
 			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
-			\UIC{$\all q. q(\gamma) \imp \nec (\all y. G(y) \imp q(y) )$} \RightLabel{$\wedge_I$}
-		\BIC{$ G(\gamma) \wedge (\all q. q(\gamma) \imp \nec (\all y. G(y) \imp q(y) ))$} \dottedLine
+			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
+		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
 		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^1$}
@@
 \end{prooftree}
+\end{small}
 
@@
 	\UIC{$\pos \ex x. G(x)$}
-			\AXC{Lamma 2} \dashedLine
+			\AXC{Lemma 2} \dashedLine
 			\UIC{$\pos \ex x. G(x) \imp \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
@@
 \begin{prooftree}
-\AXC{Theorem \ref{***}}\dashedLine
+\AXC{Theorem 2 of Scott}\dashedLine
 \UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
@@
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\AXC{Theorem \ref{T2}}\dashedLine
+\AXC{Theorem 3 of Scott}\dashedLine
 \UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
@@
 \UIC{$ \omega:  \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
-\AXC{Theorem \ref{T2}}\dashedLine
+\AXC{Theorem 3 of Scott}\dashedLine
 \UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}

### 2015-09-25 Talks/2015/Tableaux/DemoMaterial/DEMO-THF/RemoteSOT.pl
added Tableaux invited talk

@@ -1,0 +1,195 @@
+#! /usr/bin/perl -w
+
+use strict "vars";
+use Getopt::Std;
+use HTTP::Request::Common;
+use LWP;
+use File::Temp qw/ mkstemp /;
+#------------------------------------------------------------------------------
+my $SystemOnTPTPFormReplyURL = "http://www.cs.miami.edu/~tptp/cgi-bin/SystemOnTPTPFormReply";
+
+# my $ProxyAndPort = "http://proxy.math.uni-bonn.de:3128";
+my $ProxyAndPort = "";
+
+my $TemporaryFile = "/tmp/RemoteSOT_XXXXXX";
+
+my %URLParameters = (
+    "NoHTML" => 1,
+    "QuietFlag" => "-q2",
+    "SubmitButton" => "RunParallel",
+    "ProblemSource" => "UPLOAD",
+    "AutoModeTimeLimit" => 300,
+    "AutoMode" => "-cE",
+    "AutoModeSystemsLimit" => 3,
+    "X2TPTP" => "",
+    );
+#------------------------------------------------------------------------------
+    my $URL;
+    my %Options;
+    my @Content;
+    my $Key;
+    my $ReportInfo;
+    my $ReportLevel;
+    my $MyAgent;
+    my $Request;
+    my $Response;
+    my $TempHandle;
+    my $Line;
+
+#----Get format and transform options if specified
+    getopts("hfw:r:q:t:c:l:s:SPp:a:y:",\%Options);
+
+#----Help
+    if (exists($Options{'h'})) {
+        print("Usage: RemoteSOT <options> [<File name>]\n");
+        print("    <options> are ...\n");
+        print("    -h              - print this help\n");
+        print("    -w[<status>]    - list available ATP systems\n");
+        print("    -rr             - recommend ATP systems at level\n");
+        print("    -r<type><level> - report type (i,c,s,t) at level (0,2)\n");
+        print("    -q<quietness>   - control amount of output\n");
+        print("    -t<timelimit>   - CPU time limit for system\n");
+        print("    -c<automode>    - one of N, E, S\n");
+        print("    -l<syslimit>    - maximal systems for automode\n");
+        print("    -s<system>      - specified system to use\n");
+        print("    -f              - force use of inappropriate system\n");
+        print("    -S              - TSTP format output\n");
+        print("    -p<filename>    - TPTP problem name\n");
[…]

### 2015-09-25 Talks/2015/Tableaux/DemoMaterial/DEMO-THF/call_tptp.sh

@@ -1,0 +1,47 @@
+tptp4X=./tptp4X
+call=./RemoteSOT.pl
+# if the above programs don't work for you download the SystemOnTPTP package,
+# install it and change the above paths accordingly
+
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='TPS--- LEO-II--- Satallax--- Isabelle--- cocATP--- Nitpick---'
+#proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
+
+timeout=200
+
+suffix=$(date "+%Y-%m-%d--%H:%M:%S")
+resultsfile=RESULTS-$suffix
+
+touch $resultsfile
+rm $resultsfile
+
+echo
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\) >> $resultsfile
+
+for file in "$@"
+do
+echo
+echo >> $resultsfile
+    for prover in $proverlist 
+    do
+    function_to_fork () { 
+      res=""
+      while ! (echo $res | grep --quiet RESULT); do
+          res=`$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover |grep RESULT` 
+	  echo ${file}---${prover}---${res}
+	  echo ${file}---${prover}---${res} >> $resultsfile
+      done   
+    }
+    function_to_fork &
+    done
+    wait
+    echo Done 
+    echo Done >> $resultsfile
+done
+
+
+
+
+

### 2015-09-29 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Commit Annika's changes

@@
 
-%Begin with a focus on more recent stuff such as Gödel, not Anselm and so on. Focus on the improved level of formalization of the argument through Leibniz/Descartes/Kant and so on. 
-%Think about how much can be claimed about Russell, Descartes, Kant and so on... 
-%Russell eliminated or expained as a valuable contribution to the formal proof. 
-%check the Solmon paper to se about proofs in KB. Other ontological proofs. 
-%Relate introduction to the beginning quote
 %spell check
-%How to use the words God, God-like, god-like and greatest conceivable being. 
-%Theorem reference missing in the proof of modal collapse
-%Write the section on axiom B and the strength of the modal system that is necessary.
-
-\section{Introduction}
-
-Ontological arguments for the existence of God can be traced back at least to St.
-Anselm (1033-1109).
-His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
-
-A major critique of this argument is that we do not know whether the concept of greatest conceivable being in fact designates anything or if it is inconsistent, like a round square \citep{fitting}[p.134].
-%The question is if the argument is sound. 
-As Bertrand Russell pointed out, the definition of greatest conceivable being allows us to define properties, like having boots, which the being then also must have \citep{citation needed}.
-
-Kant argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
+
+%Extra:
+
+%Difference in the formalizations of Leibniz and Gödel? Why did Gödel change to conjunction?
+
+%Kant argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
 This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
 The strongest claim that can be proven by the ontological argument is a conditional claim of necessary existence: if God exists, then God necessarily exists \citep{citation needed}.
-
-St.
-Anselm's argument was further elaborated by Descartes and Leibniz.
-Leibniz identified the possible existence of God as a critical missing step in St.
-Anselm's argument.
-To fill this gap, he argued that the properties of God, the perfections, are compatible.
-This means that it is possible to satisfy all perfections at once which implies that
-the existence of a greatest conceivable being with all these properties is possible.
-
-G\"odel studied Leibniz's work \citep{Adams} and brought his ontological argument to a modern form using a modal logic with higher-order quantification over properties.
-In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being having all positive properties.
-
-G\"odel's notion of positive property and Leibniz's notion of perfection differ \citep{Adams}.
-The main formal distinction seems to be that G\"odel's positive properties are not just atomic properties, like Leibniz's perfections, but can also consist of complex combinations of atomic properties \citep{fitting}[p.139.] (TODO: Check this claim).
-In particular, one of G\"odel's axioms states that any conjunction of positive properties is itself positive.
-From this axiom, it is immediately deduced that the property of being God-like is positive.
-While this step is intuitively and informally clear, it is not easily formalizable in a standard logical calculus, because it requires inferring that being God-like is a (possibly infinite) conjunction of positive properties, from the universal definition of God-likeness as the property satisfied by individuals who have all positive properties.
-This interplay between universal quantification (in the definition of a God-like being) and in!
- finite conjunctions (in G\"odel's axiom for positive properties) is a technical difficulty that probably explains why, starting with Scott \cite{Scott}, to whom G\"odel confided his manuscript, this axiom of G\"odel's has been replaced by another that simply assumes the positivity of the property of being God-like.
-
-The main criticism against G\"odel's ontological argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
-Many recent works \cite{todo} on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
-This is discussed in greater detail in Section \ref{sec:collapse}, but possible solutions are not implemented in this paper.
-
-%This axiom is unfortunately equivalent to the possibility of God's existence given the other axioms, leading to criticisms that it might be too strong \citep{sobel2}\citep{fitting}[Section 11.4.
-Objections.].
-
-%The possibility  God's existence in turn is equivalent to  God's necessary existence as well as  God's existence itself. Thus, we have an axiom which seems to be equivalent to the conclusion.
-
- % This axiom of the positiveness of conjunctions can be given a third order formulation which is the reason for the claim that third-order quantification may be required \citep{and}.
-%\citep{fitting}[p.
-148.]
 
@@
 %\citep{fitting}[The proof is Exercise 7.1, p.163].
-
-The aim of this paper is to present two detailed formal proofs of G\"odel's ontological argument \cite{Goedel} in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
-The natural deduction style \citep{gentzen,prawitz} was chosen for three reasons.
+Inclusion of identity relation gives monothemism?*** Reference missing?
+
+% The question arises what kind of second order quantification is allowed.
+The instantiations that are used in our proof are instantiated with predicate variables that later work as eigenvariables for the second order universal introduction.
+In the proof of corollary \ref{C1} we instantiate with $G$ and in the proof of lemma \ref{L1}.
+we instantiate with $G$ and $E$.
+%In Sobel on G\"{o}del's ontological proof by Robert E.
+Koons this question is raised.
+
+% One can also discuss if a second order modal logic is sound and complete.
+With Henkin models, which restrict the second order quantification with comprehension, we have that normal second order logic is sound and complete.
+There is a completeness proof for a second order modal logic found in A completeness theorem in second order modal logic by Cocchiarella.
+
+% ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
+
+\section{Introduction}
+
+Ontological arguments for the existence of God can be traced back at least to St.
+Anselm (1033-1109).
+His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
+St.
+Anselm's argument was further elaborated by Descartes, Leibniz and Kant.
+
+Leibniz identified the possible existence of God as a critical missing step in St.
+Anselm's argument.
+To fill this gap, he argued that the properties of God, the perfections, are compatible.
+This means that it is possible to satisfy all perfections at once which implies that the existence of a greatest conceivable being with all these properties is possible.
+
+G\"odel bulit on Leibniz's work \cite{adams} and brought the ontological argument to a modern form using a modal logic with higher-order quantification over properties.
+In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
+G\"odel's work was saved in his own notes \cite{Goedel} as well as notes by Scott \cite{Scott}, in whom he confided his proof.
+
+The increase in formality of the ontological argument has required a development of the basic notions.
+G\"odel's notion of positive property and Leibniz's notion of perfection differ.
+A formal distinction is that Leibniz's perfections are atomic whereas G\"odel's positive properties can consist of combinations of atomic properties \citep{fitting}[p.139].
+In particular, one of G\"odel's axioms states that any conjunction of positive properties is itself positive.
+From this axiom, it is immediately deduced that the property of being god-like is positive.
+Intuitively, a (possibly infinite) conjunction of positive properties is deduced from the universal definition of God-likeness.
+This deductive inference is not formalizable in a finite first-order calculus.
+The interplay between universal quantification (in the definition of a god-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{Scott}, this axiom o!
+ f G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+
+The main criticism against the ontological argument is that its soundness has been doubted.
+As Fitting \citep{fitting}[p.134] described the problem ''we do not know whether the concept of greatest conceivable being in fact designates anything or if it is inconsistent, like a round square''.
+
+The main criticism against G\"odel's formal argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
+Many recent works on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
+This is discussed in greater detail in Section \ref{sec:collapse}, but possible solutions are not implemented in this paper.
+
+The aim of this paper is to present two detailed formal proofs of G\"odel's ontological argument in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
+For a comprehensive introduction to natural deduction the reader can consult \citep{prawitz}.
+The natural deduction style was chosen for three reasons.
 Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of calculi (e.g.
@@
 The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus.
-The second contribution of the paper is a new proof (also in natural deduction style) of the lemmas, theorems and corollaries in G\"odel's manuscript.
-In contrast to Scott's proofs \cite{Scott}, the proofs presented here are simpler and shorter, as discussed in Section \ref{sec:conclusion}.
+In section \ref{sec:compl-sound} we prove that the calculus used is sound and complete relative to an axiomatic modal calculus and this detailed formalization of Scott's version has as been formally verified.
+The second contribution of the paper is a new proof (also in natural deduction style) in a natural deduction calculus.
+In contrast to Scott's proof \cite{Scott} formalized in the modal calculus S5, the latter proof is formalized in KB, as discussed in Section \ref{sec:AxiomB}.
 
@@
 \end{definition}
-
-ToDo: decide which connectives and corresponding inference rules will be primitive.
 
@@
 
-ToDo: decide if rules for diamond should be added.
-
 The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
@@
 \newcommand{\s}{\qquad}
-
-% ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
 
@@
 
-\subsection{Suitability for Rigid Higher-Order Modal Logic K}
+\subsection{Suitability for Rigid Higher-Order Modal Logic K}\label{sec:compl-sound}
 
@@
 
-ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
-
-\section{Some Derivable Modal Principles}
+\section{Derivable Modal Principles}
 \label{sec:AxiomB}.
 
-As mentioned in \citep{and90}[Footnote 5], there are sources that discuss what modal principles are needed for the ontological argument.
-The standard system for the argument is S5.
+In standard first-order modal logics the modal axioms correspond to geometric frame properties for Kripke models \cite{negri}.
+The modal axiom {\bf T} corresponds to reflexivity of the frame and {\bf B} corresponds to symmetry of the frame.
+The frames of the modal logic {\bf S5} are in addition transitive.
+
+There are discussion of what modal principles are needed for the ontological argument. 
+G\"odel mentions in his notes that the modal system required for proving his theorem is a second order version of the modal system {\bf S5}.
+In \cite{and90} it is conjectured that the principles of proof could be restricted to the weaker modal logic {\bf KB}.
+The modal logic {\bf KB} consists of the basic modal system {\bf K} and the modal axioms {\bf T} and {\bf B}.
+
 Through an analysis of the proofs weaker systems, in particular KB, can be shown to be sufficient for versions of the ontological argument \citep{sobel2}[p.
@@
 
-In standard first-order modal logics the modal axioms correspond to geometric conditions on the frames of Kripke models \citep{negri}.
-The modal axiom {\bf T} corresponds to reflexivity of the frame and {\bf B} corresponds to symmetry of the frame.
-The frames of the modal logic {\bf S5} are in addition transitive.
-
-%G\"odel mentions in the notes which are saved of his proof that the modal system required for proving his theorem is a second order version of the modal system {\bf S5}.
-In \citep{and90} it is conjectured that the principles of proof could be restricted to the weaker modal logic {\bf KB}.
-The modal logic {\bf KB} consists of the basic modal system {\bf K} and the modal axioms {\bf T} and {\bf B}.
-
-ToDo: Sobel Anderson on KB: Sobel page 152
-and
-Anderson's footnote 5 on KB %Nathan Salmon: On the Logic of What Might Have Been,"
+We will prove the necessary modal principles as two lemmas; a modal distribution principle derivable in K and Brouwer's reduction principle derivable in KB.
+In addition we prove an iteration principle, which is valid in S5, for the proof in section \ref{sec:scott}.
+The iteration principle is derived in KB with axiom 5.
 
 \subsection{A modal distribution principle in K}
-
-To prove G\"odel's ontological proof in natural deduction we use a modal distribution principle derivable in K and  Brouwer's reduction theorem derivable from the modal Axiom B.
 
@@
 
-\subsection{Axiom B and Brouwer's Reduction Principle}
-
-Brouwer's reduction theorem is a version of axiom B for modal logic and a property that enforces symmetry as a geometric property for the frame of a Kripke sematics.
+\subsection{Brouwer's Reduction Principle}
 
@@
 
-ToDo: Is this principle really commonly called ``S5 Iteration Principle''?
-
 \begin{lemma}
@@
 
-\begin{prooftree}
-\AXC{ToDo}
-\end{prooftree}
+Axiom 5 is $\pos A\imp \nec \pos A$.
+
+\begin{small}
+\begin{prooftree}
+
+\AXC{Brouwer's reduction principle}\dashedLine
+\UIC{$\pos \nec A\imp A$}\RightLabel{$\nec_I$}
+\UIC{$\nec(\pos \nec A\imp A)$}
+\AXC{Axiom K}\dashedLine
+\UIC{$\nec(\pos \nec A\imp A)\imp (\nec\pos \nec A\imp \nec A)$}\RightLabel{$\imp_E$}
+\BIC{$\nec\pos \nec A\imp \nec A$}
+
+\AXC{Axiom 5 on $\nec A$}\dashedLine
+\UIC{$\pos \nec A\imp \nec \pos \nec A$}
+
+\AXC{$ $}\RightLabel{$1$}
+\UIC{$\pos \nec A$}\RightLabel{$\imp_E$}
+\BIC{$\nec \pos \nec A$}\RightLabel{$\imp_E$}
+\BIC{$\nec A$}\RightLabel{$\imp_I^1$}
+\UIC{$\pos \nec A \imp  \nec A$}
+\end{prooftree}
+\end{small}
 
@@
 
-ToDo: natural language explanation of the proof in Anderson.
-
 The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
-The formal argument of G\"odel is based on Leibniz proof, which in turn is based on Descartes proof.
-These proofs have two parts; a proof that if the exemplification of the god-likeness property is possible, then it is necessary and a proof that the exemplification is in fact possible.
+The following proof is a detailed formalization of Scott's proof that has been verified in Coq \cite{***}.
+The formal argument has  two parts; a proof that if the exemplification of the god-likeness property is possible, then it is necessary and a proof that the exemplification is in fact possible.
 
@@
 \label{T2}
-Being God-like is an essence of any God-like being:
+Being god-like is an essence of any god-like being:
 $$
@@
 \label{L2}
-If the existence of a God-like being is possible, then it is necessary:
+If the existence of a god-like being is possible, then it is necessary:
 $$
@@
 \label{T3}
-Necessarily, there exists a God-like being:
+Necessarily, there exists a god-like being:
 $$
@@
 
-A outline of the proof is that firstly all positive properties are proven to be possibly exemplified (Theorem \ref{T1}).
+In this section we present a new proof in KB. The outline of the proof is: 
+Firstly all positive properties are proven to be possibly exemplified (Theorem \ref{T1}).
 Then this theorem is applied to the property of god-likeness, thus showing the possible existence of a god-like being (Corollary \ref{C1}).
@@
 The necessary exemplification of the god-likeness property can then be deduced from lemma \ref{L1}.
-This single use of the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP} follows the presentation of in Sobel \citep{sobel2}[p.126--127].
+This single use of the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP} follows the presentation of Sobel \citep{sobel2}[p.126--127].
 
@@
 \end{theorem}
-
-%Is the distribution principle and brouwers Red thm better than using axiom B straight? Check previous versions... ***
-
-%From that it is possible to that it exists with modal inferences... **** and axioms...
 
@@
 
-Anderson's solution \citep{citation needed} modifies the definitions of god-like being and essence, and eliminates half of an axiom.
+Anderson's solution \citep{and***} modifies the definitions of god-like being and essence, and eliminates half of an axiom.
 This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
-Another solution involving more substantial modifications is that of Bj{\o}rdal \citep{bjordal,fuhrmann}.
+Another solution involving more substantial modifications is that of Bj{\o}rdal \cite{bjordal,fuhrmann}.
 
@@
 
-% The question arises what kind of second order quantification is allowed.
-The instantiations that are used in our proof are instantiated with predicate variables that later work as eigenvariables for the second order universal introduction.
-In the proof of corollary \ref{C1} we instantiate with $G$ and in the proof of lemma \ref{L1}.
-we instantiate with $G$ and $E$.
-%In Sobel on G\"{o}del's ontological proof by Robert E.
-Koons this question is raised.
-
-% One can also discuss if a second order modal logic is sound and complete.
-With Henkin models, which restrict the second order quantification with comprehension, we have that normal second order logic is sound and complete.
-There is a completeness proof for a second order modal logic found in A completeness theorem in second order modal logic by Cocchiarella.
-
 \section{Conclusions}\label{sec:conclusion}
@@
 
+ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
+
+ToDo: natural language explanation of the proof in Anderson.
+
+ToDo: Sobel Anderson on KB: Sobel page 152
+and
+Anderson's footnote 5 on KB %Nathan Salmon: On the Logic of What Might Have Been,"
+
 \begin{thebibliography}{9}
+
+\bibitem[{\itshape Adams(1995)}]{adams}
+Adams, R. M. 1995. Introductory note to [G\"odel 1970].
 
@@
 
+\bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
+Bj{\o}rdal, F. 1999. 
+Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
+
+\bibitem[{\itshape Church(1940)}]{church}
+Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56?68.
+
 \bibitem[{\itshape Fitting(2002)}]{fitting}
@@
 
-\bibitem[{\itshape Goedel(***)}]{Goedel} 
-G\"odel, K. {\itshape Appendix A. Notes in Kurt G\"odel's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
+\bibitem[{\itshape Fuhrmann(2005)}]{fuhrmann}
+Fuhrmann, A. 2005.
+Existenz und Notwendigkeit -- Kurt G\"odel's Axiomatische Theologie, in Olsson e.J., Schr\"oder-Heister, P. and Spohn, W. (eds.) Logik in der Philosophie, pp. 349-374, Synchr.-Wissenschafts-Verlag.
+
+\bibitem[{\itshape Goedel(1970)}]{Goedel} 
+G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
+
+\bibitem[{\itshape Hajek(1996)}]{hajek}
+Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
+(eds.) Logic and Logical Algebra, pp. 125-136, Marcel Dekker.
 
@@
 
+\bibitem[{\itshape Prawitz (2006)}]{prawitz}
+Prawitz, D. 2006 (1st ed. 1965). {\itshape Natural deduction: a proof-theoretical study}. Mineola, New York: Dover publications.
+
 \bibitem[{\itshape Scott (***)}]{Scott}
@@
 \bibitem[{\itshape Sobel(1987)}]{sobel}
-Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: edited by J. J. Thompson. {\itshape On being and saying : essays for Richard Cartwright},  MIT Press.
+Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying : essays for Richard Cartwright},  MIT Press.
 
@@
 
-***Gentzen and Prawitz: references to Natural Deduction. 
-Adams, Fuhrmann, Hajek and Bjordal: References to amendations. 
-Church: Reference to simple type theory.
+***Adams, Fuhrmann, Hajek and Bjordal: Confirm references to ammendations and Church. 
+****References for \cite{coq} and \cite{isabelle} and other papers checked step-by-step in Coq \cite{BenzmullerWolzenlogel-Paleo}. Reference to the inspiration for the modal rules. 
+***One Anderson reference incomplete***
+***Scott reference: the year is missing***
 

### 2015-09-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Add keywords and explanation about co-authorship

@@
 
-\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}}
+\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}\thanks{Both co-authors contributed equally to this paper.
+Author order is alphabetical by surname.
+Bruno Woltzenlogel Paleo is the corresponding author.}}
 
@@
   \and 
-  Vienna, Austria ; Canberra, Australia \\
+  Vienna University of Technology, Austria ; Australian National University \\
   \email{bruno.wp@gmail.com} 
@@
 \maketitle
+
+\begin{abstract}
+ToDo: 200 words
+\end{abstract}
+
+\noindent
+\textbf{Keywords: } Higher-Order Modal Logics, Ontological Proof, Natural Deduction
 

### 2015-10-05 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve ND paper ina major way

--- 2015-09-29 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
@@
 
+\usepackage[utf8]{inputenc}
 \usepackage{fancybox}
 \usepackage{latexsym}
+
 \usepackage{proof}
 \usepackage{bussproofs}
-\usepackage{amsmath}
 \EnableBpAbbreviations
 \newcommand{\rl}[1]{\RightLabel{#1}}
+
+\usepackage{amsmath}
+\usepackage{ntheorem}
+
+\usepackage[margin=3.5cm]{geometry}
 
@@
 \usepackage[numbers]{natbib}
+
+\pagestyle{plain}
 
@@
 
+\newtheorem*{lemma*}{Lemma}
+
 \title{A Variant of G\"{o}del's Ontological Proof \\
@@
 
-\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}}
-
-\authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
-
-\institute{
-  Helsinki, Finland\\
-  \email{annika.siders@helsinki.fi}
-  \and 
-  Vienna, Austria ; Canberra, Australia \\
-  \email{bruno.wp@gmail.com}
-}
+% ToDo: uncomment this post-acceptance: the journal uses double blind reviewing.
+% \author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}}
+
+% \authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
+
+% \institute{
+%   Helsinki, Finland\\
+%   \email{annika.siders@helsinki.fi}
+%   \and 
+%   Vienna University of Technology, Austria ; Australian National University \\
+%   \email{bruno.wp@gmail.com} 
+% }
+
+\author{}
+\authorrunning{}
+\institute{}
 
@@
 \maketitle
+
+\begin{abstract}
+This paper presents two detailed formalizations of ontological arguments in a simple natural deduction calculus.
+The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills its gaps, thus verifying its correctness.
+The second formal proof improves the first one, by: firstly, relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5} in the proof of the last theorem;  and, secondly, not using equality to prove the first theorem.
+The second proof is also technically shorter than the first one, because it avoids unnecessary detours and uses Axiom 1 only once.
+\end{abstract}
+
+\noindent
+\textbf{Keywords: } G\"odel's Ontological Proof, Higher-Order Modal Logics, Natural Deduction
 
@@
 
-%spell check
-
-%Extra:
-
-%Difference in the formalizations of Leibniz and Gödel? Why did Gödel change to conjunction?
-
-%Kant argued against the ontological argument on the basis that existence is not an analytic property \citep{kant}.
-This means that existence cannot be contained in the definition of a concept, because it is generally synthetic.
-The strongest claim that can be proven by the ontological argument is a conditional claim of necessary existence: if God exists, then God necessarily exists \citep{citation needed}.
-
-% Another, possibly undesirable, consequence of the formalism is that there can only be one God, which is defined by the axioms.
-Namely, if one assumes that the theory contains the equality relation then a God has the property of being identical to itself.
-Since any other god-like being would have at least the same properties it shares the property of being identical to the one God.
-Thus, we have proved monotheism \citep{sobel2}[Ch.
-4, section 3.3.2].
-%\citep{fitting}[The proof is Exercise 7.1, p.163].
-Inclusion of identity relation gives monothemism?*** Reference missing?
-
-% The question arises what kind of second order quantification is allowed.
-The instantiations that are used in our proof are instantiated with predicate variables that later work as eigenvariables for the second order universal introduction.
-In the proof of corollary \ref{C1} we instantiate with $G$ and in the proof of lemma \ref{L1}.
-we instantiate with $G$ and $E$.
-%In Sobel on G\"{o}del's ontological proof by Robert E.
-Koons this question is raised.
-
-% One can also discuss if a second order modal logic is sound and complete.
-With Henkin models, which restrict the second order quantification with comprehension, we have that normal second order logic is sound and complete.
-There is a completeness proof for a second order modal logic found in A completeness theorem in second order modal logic by Cocchiarella.
-
-% ToDo: check if disjunction rules are used in Goedel's proof. If not, we could define disjunction as we already do for negation and equivalence.
-
 \section{Introduction}
@@
 To fill this gap, he argued that the properties of God, the perfections, are compatible.
-This means that it is possible to satisfy all perfections at once which implies that the existence of a greatest conceivable being with all these properties is possible.
+This means that it is possible to satisfy all perfections at once, which implies that the existence of a greatest conceivable being with all these properties is possible.
 
@@
 In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
-G\"odel's work was saved in his own notes \cite{Goedel} as well as notes by Scott \cite{Scott}, in whom he confided his proof.
-
-The increase in formality of the ontological argument has required a development of the basic notions.
+G\"odel's work was saved in his own notes \cite{Goedel} as well as notes by Scott \cite{scott}, in whom he confided his proof.
+
+The increase in formality of the ontological argument has required a development of its basic notions.
 G\"odel's notion of positive property and Leibniz's notion of perfection differ.
@@
 This deductive inference is not formalizable in a finite first-order calculus.
-The interplay between universal quantification (in the definition of a god-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{Scott}, this axiom o!
- f G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
-
-The main criticism against the ontological argument is that its soundness has been doubted.
-As Fitting \citep{fitting}[p.134] described the problem ''we do not know whether the concept of greatest conceivable being in fact designates anything or if it is inconsistent, like a round square''.
-
-The main criticism against G\"odel's formal argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
-Many recent works on the ontological argument have focused on proposing modifications of the argument that would not entail a modal collapse.
-This is discussed in greater detail in Section \ref{sec:collapse}, but possible solutions are not implemented in this paper.
-
-The aim of this paper is to present two detailed formal proofs of G\"odel's ontological argument in a natural deduction calculus (as defined in Section \ref{sec:calculus}).
-For a comprehensive introduction to natural deduction the reader can consult \citep{prawitz}.
+The interplay between universal quantification (in the definition of a god-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+
+The aim of this paper is to present two detailed formalizations of ontological arguments in a natural deduction calculus.
+For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz}.
+The higher-order modal natural deduction calculus proposed here has special introduction and elimination rules for modalities, as defined in Section \ref{sec:calculus}, and is sound and complete relative to an axiomatic modal calculus, as proven in Section \ref{sec:compl-sound}.
+
 The natural deduction style was chosen for three reasons.
-Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of calculi (e.g.
+Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of proof calculi (e.g.
 Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}).
@@
 Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
-Thirdly, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, the natural deduction formalizations have been easily checked step-by-step in Coq \cite{coqformalizations}.
-
-The first contribution of this paper is a detailed formalization of Scott's version \cite{Scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in a natural deduction calculus.
-In section \ref{sec:compl-sound} we prove that the calculus used is sound and complete relative to an axiomatic modal calculus and this detailed formalization of Scott's version has as been formally verified.
-The second contribution of the paper is a new proof (also in natural deduction style) in a natural deduction calculus.
-In contrast to Scott's proof \cite{Scott} formalized in the modal calculus S5, the latter proof is formalized in KB, as discussed in Section \ref{sec:AxiomB}.
+Thirdly, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this.
+
+The first contribution of this paper is a detailed formalization of Scott's version \cite{scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
+The second contribution of the paper is a new proof (also in natural deduction style).
+In contrast to Scott's proof \cite{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
+The new proof also does not rely on the equality predicate and is much shorter.
+
+A major criticism against G\"odel's formal argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
+This is discussed in greater detail in Section \ref{sec:collapse}, where a natural deduction derivation of the collapse is presented.
+Although many recent works on the ontological argument have proposed modifications of the argument that do not entail a modal collapse, these solutions are beyond the scope of this paper.
 
@@
 Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
-Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}.
-Beta-reduction is implicit; all rules are assumed modulo beta-reduction.
+%Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}. 
+Beta-reduction is implicit; all rules are assumed to operate modulo beta-reduction.
 A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules.
@@
 
+\newcommand{\subproof}{\star}
+
 Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
 Dashed lines are used to refer to an axiom or theorem with proof shown elsewhere.
+When proof trees are too large to fit in the page, some branches may be displayed separately.
+In such cases, the conclusion of the branch and the location in the main proof tree where the branch belongs are annotated with the same symbol (a subscripted $\subproof$).
 Dotted lines are used to indicate folding and unfolding of definitions.
@@
 Therefore, the boxes can be omitted without loss of information.
+
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
 
@@
 \end{calculus}
-
-%\clearpage
-
-\noindent
-The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND.
-The calculus with the additional rules shown in Figure \ref{fig:NDK} is called \NDK.
-
-%\clearpage
 
@@
 
-Without the restriction that every box must be accessed by exactly one strong modal inference, the calculus would be unsound for the modal logic \textbf{K}, because the formula $\all \psi.
-(\nec \psi \imp \pos \psi)$ would be derivable although it is not valid in \textbf{K}.
+Without the restriction that every box must be accessed by exactly one strong modal inference, the calculus would be unsound for the modal logic \textbf{K}.
+The formula $\all \psi.
+(\nec \psi \imp \pos \psi)$, for example, is not valid in \textbf{K} but would be provable according to the following proof:
 
@@
 
-This example of an unsound derivation is prevented by the eigen-box condition because the box labelled by $\omega$ is not accessed by any strong inference.
+The example proof above is unsound according to the {\NDK} calculus, because the eigen-box condition is violated: the box labelled by $\omega$ is not accessed by any strong inference.
 
@@
 Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
-Nevertheless, it should be noted that its adequacy has already been contested \cite{fitting}.
-
-\section{Derivable Modal Principles}
-\label{sec:AxiomB}.
-
-In standard first-order modal logics the modal axioms correspond to geometric frame properties for Kripke models \cite{negri}.
-The modal axiom {\bf T} corresponds to reflexivity of the frame and {\bf B} corresponds to symmetry of the frame.
-The frames of the modal logic {\bf S5} are in addition transitive.
-
-There are discussion of what modal principles are needed for the ontological argument. 
-G\"odel mentions in his notes that the modal system required for proving his theorem is a second order version of the modal system {\bf S5}.
-In \cite{and90} it is conjectured that the principles of proof could be restricted to the weaker modal logic {\bf KB}.
-The modal logic {\bf KB} consists of the basic modal system {\bf K} and the modal axioms {\bf T} and {\bf B}.
-
-Through an analysis of the proofs weaker systems, in particular KB, can be shown to be sufficient for versions of the ontological argument \citep{sobel2}[p.
-152].
-The proof presented in Section \ref{sec:newproof} is such a version and explicitly shows a single use of Brouwer Reduction Principle, which, in turn, is provable from axiom B.
-All other proof steps are formalizable in a basic second-order modal logic K.
-
-We will prove the necessary modal principles as two lemmas; a modal distribution principle derivable in K and Brouwer's reduction principle derivable in KB.
-In addition we prove an iteration principle, which is valid in S5, for the proof in section \ref{sec:scott}.
-The iteration principle is derived in KB with axiom 5.
-
-\subsection{A modal distribution principle in K}
-
-\begin{lemma}\label{DP}
-The distribution principle 
+Nevertheless, it should be noted that its adequacy for the ontological argument has already been contested \cite{fitting}.
+Another assumption made here is that the quantification domains are constant (i.e.
+independent of the possible worlds).
+Neither G\"odel's manuscript nor Scott's manuscript reveal whether they use constant or varying domains, and this is also the case for many variants (e.g.
+\cite{hajek}).
+Nevertheless, some authors of variants \cite[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
+Our choice of assuming constant domains is motivated by simplicity.
+
+\section{Some Useful Derivable Modal Principles}
+\label{sec:ModalLogics}
+
+From an axiomatic point of view, modal logics differ with respect to which additional axiom schemas they admit.
+Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accesibility relation in the Kripke models \cite{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
+
+In this section, natural deduction proofs of a three convenient derived modal principles are presented.
+These principles are used in the ontological proofs formalized in the next sections.
+Although they are well-known among modal logicians, we include their proofs here in order to have a self-contained and fully formal presentation of the ontological proofs, without any gaps.
+
+\bigskip
+
+\noindent
+The \emph{distribution principle} can be seen as a form of modus ponens within the scope of modalities: if $A \imp B$ holds in all accessible worlds and $A$ holds in an accessible world, then $B$ holds in an accessible world.
+This principle is provable in the modal logic {\bf K}.
+
+\begin{lemma*}[Distribution Principle]
+\label{DP} 
 $$\nec (A\imp B)\imp(\pos A\imp \pos B)$$
-is provable in our system K of modal logic. 
-\end{lemma}
+\end{lemma*}
 
@@
 
-\subsection{Brouwer's Reduction Principle}
-
-\begin{lemma}\label{BRP}
-The Brouwer Reduction Principle $\pos\nec A\imp A$ is derivable from axiom B, $A\imp \nec \pos A$. 
-\end{lemma}
+\bigskip
+
+\noindent
+\emph{Brouwer's reduction principle} is derivable in modal logic {\bf KB}, using axiom B.
+The proof of G\"odel and Scott (Section \ref{sec:scott}) do not make direct use of this principle, but the new proof presented in Section \ref{sec:newproof} does.
+
+\begin{lemma*}[Brouwer's Reduction Principle]
+\label{BRP}
+$$
+\pos\nec A\imp A
+$$ 
+\end{lemma*}
 
@@
 \AXC{$ $}\RightLabel{$2$}
+\UIC{$\pos \nec A $} \doubleLine
 \UIC{$\neg\nec \neg \nec A$}
@@
 \AXC{Axiom B}\dashedLine
+\UIC{$\neg A\imp \nec\pos\neg A$} \doubleLine
 \UIC{$\neg A\imp \nec\neg\nec\neg\neg A$}\RightLabel{$\imp_E$}
 \BIC{$\nec\neg\nec\neg\neg A$}\doubleLine
-\UIC{$\nec\neg\nec A$}
-\BIC{$(\neg\nec \neg \nec A)\&(\nec\neg \nec A)$}\RightLabel{$\neg_I^1$}
+\UIC{$\nec\neg\nec A$} \RightLabel{$\neg_E$}
+\BIC{$\bot$}\RightLabel{$\neg_I^1$}
 \UIC{$\neg\neg A$} \RightLabel{$\neg\neg_E$}
 \UIC{$A$} \RightLabel{$\imp_I^2$}
-\UIC{$\neg\nec \neg \nec A\imp A$}
+\UIC{$\pos \nec A\imp A$}
 \end{prooftree}
@@
 
-\subsection{S5 and its Iteration Principle}
-
-\begin{lemma}
-What is possibly necessary is necessary:
+\bigskip
+
+\noindent
+In the modal logic {\bf S5}, a sequential iteration of modalities can be collapsed to the last modality in the sequence. In other words, the following principles hold in {\bf S5}:
+\begin{itemize}
+\item $\pos^n \nec A \imp \nec A$
+\item $\nec^n \pos A \imp \pos A$
+\item $\pos^n \pos A \imp \pos A$
+\item $\nec^n \nec A \imp \nec A$
+\end{itemize}
+
+\noindent
+Below we exhibit a natural deduction proof for the first principle listed above  when $n = 1$.
+Informally, this particular case can be read as the claim that anything that is possibly necessary is in fact necessary, and it is sufficient for the ontological proofs of G\"odel and Scott.
+The proof of this principle depends on the previously proven Brouwer's reduction principle, on the Modal Axiom 5, and on K, whose derivation is shown in the proof of Theorem \ref{theorem:completeness}.
+
+\begin{lemma*}[Iteration Principle]
 $$
@@
 $$
-\end{lemma}
-
-\begin{proof}
-
-Axiom 5 is $\pos A\imp \nec \pos A$.
-
-\begin{small}
-\begin{prooftree}
-
-\AXC{Brouwer's reduction principle}\dashedLine
+\end{lemma*}
+
+\begin{proof}\hfill
+
+\begin{footnotesize}
+\begin{prooftree}
+
+\AXC{Brouwer's Reduction}\dashedLine
 \UIC{$\pos \nec A\imp A$}\RightLabel{$\nec_I$}
 \UIC{$\nec(\pos \nec A\imp A)$}
-\AXC{Axiom K}\dashedLine
+\AXC{K}\dashedLine
 \UIC{$\nec(\pos \nec A\imp A)\imp (\nec\pos \nec A\imp \nec A)$}\RightLabel{$\imp_E$}
@@
 
-\AXC{Axiom 5 on $\nec A$}\dashedLine
+\AXC{Axiom 5 for $\nec A$}\dashedLine
 \UIC{$\pos \nec A\imp \nec \pos \nec A$}
@@
 \end{prooftree}
-\end{small}
+\end{footnotesize}
 
@@
 
-The acceptance of the correctness of the ontological argument by G\"odel's work boils down to the intuitive correctness of the axioms and definitions and the belief in the soundness of the deductive system.
-The following proof is a detailed formalization of Scott's proof that has been verified in Coq \cite{***}.
-The formal argument has  two parts; a proof that if the exemplification of the god-likeness property is possible, then it is necessary and a proof that the exemplification is in fact possible.
+\newcommand{\scott}{^{\dagger}}
+
+In this section we present a detailed formalization of Scott's proof in the natural deduction calculus defined in Section \ref{sec:calculus}.
+The reasoning in Scott's manuscript was reproduced step-by-step and any reasoning gaps were filled in ways that we considered most natural.
+All derived formulas that do appear in Scott's manuscript are marked with a $\dagger$ here.
+Unmarked formulas were derived in the process of filling the gaps between the marked formulas mentioned in Scott's manuscript.
+
+% ToDo: uncomment this post-acceptance: the journal uses double blind reviewing.
+%The formal natural deduction proof presented here was verified in the Coq proof assistant\footnote{\url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Coq/GoedelGod_Scott.v}}.
+
+Scott's version of G\"odel's proof depends on 5 axioms that circumscribe the notion of \emph{positive} property, with positivity being denoted by the undefined second-order predicate symbol $P$.
+Additionally, 3 definitions are used for defining the notions of \emph{God-like}, \emph{essence} and \emph{necessary existence}.
+Technically, these definitions only abbreviate certain complex formulas.
+The argument would still go through if all defined symbols were replaced by the complex formulas they define.
+This observation is particularly relevant in the case of \emph{necessary existence}.
+Since this notion of ``existence'' is just an abbreviation, G\"odel's argument is not susceptible to Kant's criticism against Anselm's argument (that existence should not be treated as a predicate).
+In G\"odel's proof, existence is properly denoted by the existential quantifier.
+It is, therefore, unfortunately misleading to refer to the defined predicate symbol $E$ as ``necessary existence'', when in fact it is just a convenient abbreviation.
+
+Besides the 5 axioms and 3 definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
+The use of reflexivity could be avoided, but it gives a shorter proof.
 
@@
 $$
-P(\varphi) \imp \pos \ex x.\varphi(x)
+\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
 $$
@@
 		\UIC{$\omega: \all x. \rho(x) \imp x = x$} \RightLabel{$\nec_I$}
-		\UIC{$\nec (\all x. \rho(x) \imp x = x) $} \RightLabel{$\wedge_I$}
+		\UIC{$\nec (\all x. \rho(x) \imp x = x)\scott $} \RightLabel{$\wedge_I$}
 	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) $} 
@@
 	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) \imp P(\lambda x. x = x)$} 
-
-		\BIC{$P(\lambda x. x = x)$} 
+		\BIC{$[\subproof_1] \qquad P(\lambda x. x = x)\scott$} 
 \end{prooftree}
@@
 			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\neg_E$}
-		\BIC{$\omega: \bot$} \RightLabel{$\neg_I^9$}
+		\BIC{$\omega: \bot$} \RightLabel{$\bot_E$}
 		\UIC{$\omega: \neg (\beta = \beta)$} \dottedLine
@@
 		\UIC{$\omega: (\all x. \rho(x) \imp (x \neq x))$} \RightLabel{$\nec_I$}
-		\UIC{$\nec (\all x. \rho(x) \imp (x \neq x))$}  \RightLabel{$\wedge_I$}
+		\UIC{$\nec (\all x. \rho(x) \imp (x \neq x))\scott$}  \RightLabel{$\wedge_I$}
 	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) $} 
@@
 	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) \imp P(\lambda x. x \neq x)$} 
-		\BIC{$P(\lambda x. x \neq x)$} 
-\end{prooftree}
-
-\begin{prooftree}
-		\AXC{$ $} \dashedLine
-		\UIC{$P(\lambda x. x \neq x)$}    
-    		\AXC{$ $} \dashedLine
-			\UIC{$P(\lambda x. x = x)$} 
+		\BIC{$[\subproof_2] \qquad P(\lambda x. x \neq x)\scott$} 
+\end{prooftree}
+
+\begin{prooftree}
+		\AXC{$\subproof_2$} \dashedLine
+		\UIC{$P(\lambda x. x \neq x)\scott$}    
+    		\AXC{$\subproof_1$} \dashedLine
+			  \UIC{$P(\lambda x. x = x)\scott$} 
 					\AXC{Half of Axiom 1} \dashedLine
@@
 	\UIC{$\omega: \all x. \neg \rho(x)$} \RightLabel{$\nec_I$}
-	\UIC{$\nec \all x. \neg \rho(x)$}
+	\UIC{$\nec \all x. \neg \rho(x) \scott$}
 	        \AXC{$ $} \dashedLine
@@
   		\UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^5$}
-  		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
-
+  		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x)$} \RightLabel{$\all_I$}
+      \UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
 \end{prooftree}
@@
 \end{theorem}
-\begin{proof}
+\begin{proof}\hfill
 
@@
 		\UIC{$\neg P(\rho)$} 
-				\AXC{Axiom 1 for $\rho$} \RightLabel{$\imp_E$}
+				\AXC{Axiom 1} \dashedLine
+        \UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
+        \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
+        \UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
 			\BIC{$P(\lambda x. \neg \rho(x))$}
@@
 	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
-	   \UIC{$P(\rho)$}
+	   \UIC{$[\subproof_3] \qquad P(\rho)$}
 	   \end{prooftree}
@@
 
- \AXC{$ $}\dashedLine
+ \AXC{$\subproof_3$}\dashedLine
  \UIC{$P(\rho)$}
 	      \AXC{Axiom 4} \dashedLine
+        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
 	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
-		 \BIC{$\nec P(\rho)$}
+		 \BIC{$\nec P(\rho)\scott$}
 			\AXC{$ $} \RightLabel{4}
@@
 \label{L1}
-Exemplification of the god-likeness property implies the exemplification is necessary:
+If there is a god-like being, then there is a god-like being necessarily:
 $$
@@
 
-\begin{proof}
+\begin{proof}\hfill
 
@@
 \UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
-\UIC{$G(\gamma)$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$ $} \dashedLine
+\UIC{$[\subproof_4] \qquad G(\gamma)$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$\subproof_4$} \dashedLine
 \UIC{$G(\gamma)$}
@@
 		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}
-	\BIC{$\ess{G}{\gamma}$}
+	\BIC{$\ess{G}{\gamma}\scott$}
 				\AXC{Axiom 5} \dashedLine
 				\UIC{$P(E)$}
-						\AXC{$ $} \dashedLine
+						\AXC{$\subproof_4$} \dashedLine
 						\UIC{$G(\gamma)$} \dottedLine
@@
 						\UIC{$P(E) \imp E(\gamma)$} 
-					\BIC{$E(\gamma)$} \dottedLine
+					\BIC{$E(\gamma)\scott$} \dottedLine
 					\UIC{$\all \varphi. \ess{\varphi}{\gamma} \imp \nec \ex x. \varphi(x)$} \RightLabel{$\all_E$}
@@
 
-\begin{proof}
-
-\begin{footnotesize}
+\begin{proof}\hfill
+
+\begin{scriptsize}
 \begin{prooftree}
@@
 			\BIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
-	\BIC{$\pos \nec \ex x. G(x)$}
+	\BIC{$\pos \nec \ex x. G(x)\scott$}
 			\AXC{S5 Iteration Principle} \dashedLine
@@
 \end{prooftree}
-\end{footnotesize}
+\end{scriptsize}
 
@@
 
-In this section we present a new proof in KB. The outline of the proof is: 
-Firstly all positive properties are proven to be possibly exemplified (Theorem \ref{T1}).
-Then this theorem is applied to the property of god-likeness, thus showing the possible existence of a god-like being (Corollary \ref{C1}).
-In Lemma \ref{L1} it is classically proven that the hypothetically assumed existence of a god-like being implies that the existence is necessary.
-Lastly, the necessary existence of a god-like being is proven (Theorem \ref{T2}) from Corollary \ref{C1} and Lemma \ref{L1} by properties of the modal system.
-The properties required for the proof are the distribution principle and Brouwer's reduction theorem mentioned in Section \ref{sec:AxiomB}.
-The proof of Theorem \ref{T2} follows the proof of Sobel \citep{sobel2}[p.
-126--127].
+In this section we present a new proof in {\bf KB}.
+This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the equality axiom of reflexivity nor {\bf S5}'s modal iteration principle.
+Instead, it relies only on Brouwer's reduction principle (in the proof of Theorem 3).
+The new proof is also shorter than Scott's proof.
 
@@
 
-\subsection{Possibly, God Exists}
 \begin{axiom}
@@
 Either a property or its negation is positive, but not both:
-Axiom schema, for all $\varphi$:
-$$
-P(\neg \varphi) \biimp \neg P(\varphi)
+$$
+\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
 $$
 \end{axiom}
+
 \begin{axiom}
@@
 A property necessarily implied by a positive property is positive:
-Axiom schema, for all $\varphi$ and $\psi$: 
-$$
-(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)
+$$
+\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
 $$
@@
 \label{T1}
-
 Positive properties are possibly exemplified:
-Axiom schema, for all $\varphi$:
-$$
-P(\varphi) \imp \pos \ex x.\varphi(x)
+$$
+\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
 $$
@@
 \BIC{$ P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]$}
-\AXC{Axiom \ref{A2} for $\rho$ and $\neg\rho$} \dashedLine
-\UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\neg \rho)$} \RightLabel{$\imp_E$}
-\BIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$ $} \dashedLine
-\UIC{$P(\neg \rho)$} \RightLabel{$\imp_E$}
-\AXC{Half of Axiom \ref{A1}} \dashedLine
-\UIC{$ P(\neg \rho) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
+\AXC{Axiom \ref{A2} for $\rho$ and $\lambda x. \neg \rho(x)$} \dashedLine
+\UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\BIC{$[\subproof_5]\qquad P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$\subproof_5$} \dashedLine
+\UIC{$P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\AXC{Axiom 1} \dashedLine
+\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
+\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
+\UIC{$ P(\lambda x. \neg \rho(x)) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
 \BIC{$\neg P(\rho) $} 
@@
 \UIC{$ \pos \ex x.\rho(x)$}  \RightLabel{$\imp_I^3$}
-\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $}
-
-\end{prooftree}
-
-\end{proof}
+\UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
+\UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
+
+\end{prooftree}
+
+\end{proof}
+
+Comparing this new proof of Theorem 1 with the proof of Theorem 1 indicated by G\"odel and Scott, it is noticeable that new proof is significantly shorter.
+The key idea to obtain a shorter proof is to instantiate the second universally quantified property of Axiom 2 by $\lambda x.
+\neg \rho(x)$ instead of using $=$ or $\neq$ as G\"odel and Scott did.
+This eliminates the reliance of the ontological argument on the presence of $=$ and $\neq$ in the logical language.
+Furthermore, Axiom 2 is then needed only once.
+
+Although the simpler proof of Theorem 1 presented here was developed independently, it seems conceptually related to a possible simplification briefly and informally described by Anderson \cite[footnote 2]{and90} and attributed by him to an anonymous referee.
+The simpler proof of Theorem 1 is not particularly special.
+In fact, it is probably the most likely proof that any well-trained logician would construct, if asked to derive Theorem 1 from Axioms 1 and 2 using methodical formal proof search techniques.
+Modern techniques usually rely on unification to compute or guess good instances for weakly quantified variables, and therefore they would never result in instantiations containing $=$ and $\neq$, which are symbols that do not appear anywhere in the axioms or in the theorems to be proved.
+We may only wonder why G\"odel preferred to use those unusual instantiations.
+It might be an interesting question from a historical point of view, if this suggests that there might have been a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
 
@@
 \label{D1}
-A \emph{god-like} being possesses all positive properties:
+A \emph{God-like} being possesses all positive properties:
 $$
@@
 \end{axiom}
+
 \begin{corollary}
 \label{C1}
-Possibly, God exists:
+Possibly, a God-like being exists:
 $$
@@
 
-\subsection{If God exists, then God necessarily exists}
-
 \begin{axiom}
@@
 
+At this point, G\"odel and Scott proceed to prove Theorem 2, which is then used to prove Lemma 1.
+However, a close inspection of Scott's proof of Lemma 1 reveals that Theorem 2 is an unnecessary detour.
+Therefore, we do not include it in the new proof.
+
 \begin{definition}
@@
 \end{definition}
+
 \begin{axiom}
@@
 \end{axiom}
+
+The new proof of Lemma 1 does not rely on Theorem 2.
+Overall, it is slightly shorter than Scott's proofs of Theorem 2 and Lemma 1 combined, but a large part of the new proof of Lemma 1 is structurally very similar to Scott's proof of Theorem 2, and their underlying intuitive ideas are still essentially the same.
+However, despite being an unnecessary detour from a technical point of view, Theorem 2 is very interesting from a philosophical perspective.
+It breaks an otherwise long proof of Lemma 1 in a point that facilitates comprehension by humans at an intuitive and informal level.
+
+This phenomenon is very intriguing: the new proof of Lemma 1 is technically simpler (because it has fewer inferences), but Scott's proof of Lemma 1 with Theorem 2 can be considered intuitively simpler (because it is arguably easier to understand).
+This constitutes an interesting case for Hilbert's 24th Problem \cite{hilbert}, which asks for criteria to properly compare the simplicity of proofs.
+While Hilbert had mathematical proofs in mind, insights into the 24th problem could be gained by analyzing philosophical proofs as well.
+As G\"odel suggested, ``a scientific (exact) philosophy and theology [\ldots] is also most highly fruitful for science''.
+
 \begin{lemma}
 \label{L1}
-Exemplification of the god-likeness property implies the exemplification is necessary:
+If there is a god-like being, then there is a god-like being necessarily:
 $$
@@
 \UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
-\UIC{$G(g)$}
-\end{prooftree}
+\UIC{$[\subproof_6] \qquad G(\gamma)$}
+\end{prooftree}
+
+\begin{scriptsize}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{2}
+\UIC{$\neg P(\rho)$} 
+\AXC{Axiom 1} \dashedLine
+\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
+\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
+\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
+\BIC{$P(\lambda x. \neg \rho(x))$}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
+\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$}\RightLabel{$\imp_E$}
+\BIC{$\neg \rho(\gamma)$}
+\AXC{$ $} \RightLabel{3}
+\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
+\BIC{$\bot$}\RightLabel{$\imp_I^2$}
+\UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
+\UIC{$P(\rho)$}
+\AXC{Axiom \ref{A4}} \dashedLine
+\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
+\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
+\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
+\end{prooftree}
+\end{scriptsize}
 
@@
 \begin{prooftree}
-
-\AXC{$ $} \RightLabel{2}
-\UIC{$\neg P(\psi)$} 
-\AXC{Axiom \ref{A1}} \dashedLine
-\UIC{$\neg P(\psi) \imp P(\neg \psi)$}\RightLabel{$\imp_E$}
-\BIC{$P(\neg \psi)$}
-\AXC{$G(g)$}  \dottedLine \RightLabel{D\ref{D1}}
-\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
-\UIC{$P(\neg \psi) \imp \neg \psi(g)$}\RightLabel{$\imp_E$}
-\BIC{$\neg \psi(g)$}
-\AXC{$ $} \RightLabel{3}
-\UIC{$\psi(g)$} \RightLabel{$\imp_E$}
-\BIC{$\bot$}\RightLabel{$\imp_I^2$}
-\UIC{$\neg\neg P(\psi)$}\RightLabel{$\neg\neg_E$}
-\UIC{$P(\psi)$}
-\AXC{Axiom \ref{A4}} \dashedLine
-\UIC{$\all \varphi.[P(\psi) \imp \nec\; P(\psi)]$}\RightLabel{$\forall_E$}
-\UIC{$P(\psi) \imp \nec\; P(\psi)$}\RightLabel{$\imp_E$}
-\BIC{$\nec P(\psi)$}\RightLabel{$\imp_I^3$}
-\UIC{$\psi(g) \imp \nec\; P(\psi)$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
 \AXC{$ $} \RightLabel{5}
-\UIC{$\nec P(\psi)$} \RightLabel{$\nec_E$}
-\UIC{$\omega: P(\psi)$}
+\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: P(\rho)$}
 \AXC{$ $} \RightLabel{4}
@@
 \UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
-\UIC{$\omega: P(\psi) \imp \psi(y)$}\RightLabel{$\imp_E$}
-\BIC{$\omega: \psi(y)$} \RightLabel{$\imp_I^4$}
-\UIC{$\omega: G(y) \imp \psi(y)$}\RightLabel{$\forall_I$}
-\UIC{$\omega: \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\nec_I$}
-\UIC{$\nec \forall y.(G(y) \imp \psi(y))$} \RightLabel{$\imp_I^5$}
-\UIC{$\nec P(\psi)\imp \nec \forall y .(G(y)\imp \psi(y))$} \RightLabel{$\imp_E$}
+\UIC{$\omega: P(\rho) \imp \rho(y)$}\RightLabel{$\imp_E$}
+\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
+\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
+\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
+\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
+\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
 \AXC{$ $} \RightLabel{6}
-\UIC{$\psi(g)$} \RightLabel{}
-\AXC{$ $} \dashedLine
-\UIC{$\psi(g)\imp\Box P(\psi)$}\RightLabel{$\imp_E$}
-\BIC{$\nec P(\psi)$} \RightLabel{$\imp_E$}
-\BIC{$\nec \forall y .(G(y)\imp \psi(y))$}\RightLabel{$\imp_I^6$}
-\UIC{$\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y))$}\RightLabel{$\forall_I$}
-\UIC{$\forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\RightLabel{$\wedge_I$}
-\AXC{$G(g)$}  
-\BIC{$G(g) \wedge \forall \psi.(\psi(g) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
-\UIC{$\ess{G}{g}$}
+\UIC{$\rho(\gamma)$} \RightLabel{}
+\AXC{$\subproof_7$} \dashedLine
+\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
+\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
+\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}  
+\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
+\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
 \end{prooftree}
@@
 \begin{prooftree}
-\AXC{ } \dashedLine
-\UIC{$\ess{G}{g}$}
+\AXC{$\subproof_8$} \dashedLine
+\UIC{$\ess{G}{\gamma}$}
 \AXC{Axiom \ref{A5}} \dashedLine
 \UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
-\AXC{$G(g)$}\dottedLine \RightLabel{D\ref{D1}}
-\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(g))$}\RightLabel{$\forall_E$}
-\UIC{$P(E) \imp E(g)$}\RightLabel{$\imp_E$}
-\BIC{$E(g)$}\dottedLine \RightLabel{D\ref{D3}}
-\UIC{$ \all \varphi.[\ess{\varphi}{g} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
-\UIC{$ \ess{G}{g} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
+\UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
+\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{D3}}
+\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
+\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
 \BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
@@
 
-\subsection{God necessarily exists}
-
-The exemplification of the god-likeness property is provable from its possible exemplification by the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP}.
-The necessary exemplification of the god-likeness property can then be deduced from lemma \ref{L1}.
-This single use of the modal distribution principle \ref{DP} and Brouwer's reduction principle \ref{BRP} follows the presentation of Sobel \citep{sobel2}[p.126--127].
+Lemma 2 also turns out to be a superfluous detour and therefore the new proof does not include it.
+Inspecting Scott's proof of Lemma 2, we see that an important step is the derivation of $\nec \ex x.
+G(x)$ from $\pos \nec \ex x.
+G(x)$ and {\bf S5}'s iteration principle.
+Instead, we can derive $\ex x.
+G(x)$ from $\pos \nec \ex x.
+G(x)$ and Brouwer's reduction theorem and then derive $\nec \ex x.
+G(x)$ from $\ex x.
+G(x)$ by the $\nec_I$ (necessitation) inference rule.
+This is precisely what the new proof of Theorem 3 does.
 
@@
 \label{T3}
-The property of god-likeness is necessarily exemplified:
+Necessarily, there exists a god-like being:
 $$
@@
 \begin{prooftree}
-\AXC{ Distribution principle \ref{DP} } \dashedLine
+\AXC{ Distribution Principle} \dashedLine
 \UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
@@
 \BIC{$\pos\nec \ex x. G(x)$}
-\AXC{ Brouwer's reduction principle \ref{BRP} } \dashedLine
+\AXC{ Brouwer's Reduction Principle } \dashedLine
 \UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
@@
 
-\section{Modal collapse} \label{sec:collapse}
+Since the iteration principle requires the strong modal logic {\bf S5} while Browser's reduction principle relies only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
+There have been many discussions of what modal axioms are needed for the ontological argument.
+In their manuscripts, G\"odel and Scott do not write explicitly which logic they use.
+However, a proof step in their notes clearly relies on {\bf S5}'s \emph{iteration principle}.
+Anderson \cite[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}.
+Sobel \citep{sobel2}[p.
+152] acknowledges Anderson's conjecture, claims that it is correct and tries to informally explain how his formal proofs could be modified to rely on {\bf B} instead of {\bf S5}.
+Nevertheless, neither Anderson nor Sobel presented any formal proof relying only on {\bf B}.
+The new formal proof presented in Section \ref{sec:newproof} relies on {\bf KB}, which is weaker than {\bf B}.
+
+Curiously, in the new proof of Theorem 3, the actual existence of a God-like being ($\ex x.
+G(x)$) is proven \emph{before} its necessity ($\nec \ex x.
+G(x)$).
+This contrasts with G\"odel's proof, where the actual existence is never actually proven, although it can be derived as a trivial corollary of its necessity by applying the modal axiom T.
+
+\section{Modal Collapse} \label{sec:collapse}
 
@@
 4, section 6, theorems 9 and 10].
-That is: for all properties, $\varphi$,
-$$\varphi \biimp \pos \varphi \biimp \nec \varphi
-$$
-If G\"odel's ontological proof is abstractly analysed, then we are proving a restricted modal collapse, which applies to one specific formula, the exemplification of god-likeness.
-In addition we prove and that the definition of god-likeness is sound, which means that it is possibly exemplified.
-Thus, the necessity of the formula follows.
-The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for all formulas.
-Therefore, an improvement would be obtained if the modal collapse was limited to one property, namely the property of god-likeness.
-A number of solutions to the problem of the modal collapse have been proposed.
-
-Anderson's solution \citep{and***} modifies the definitions of god-like being and essence, and eliminates half of an axiom.
-This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
-Another solution involving more substantial modifications is that of Bj{\o}rdal \cite{bjordal,fuhrmann}.
-
-On another track, Fitting has argued that greater care has to be taken with the semantics of higher-order modal logics.
-Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional.
-Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
-
-Anderson \citep{and90}[p.
-292] and Sobel \citep{sobel2}[p.
-133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
-
-It is beyond the scope of this paper to analyze these solutions in detail or propose new solutions.
-The purpose of this section is simply to show natural deduction derivations of the modal collapse, thus confirming that it holds for the axioms used in the previous sections.
+That is: for all propositions $A$,
+$$
+A \biimp \pos A \biimp \nec A
+$$
+
+Below we show natural deduction derivations of the modal collapse, thereby confirming that it holds\footnote{It is well-known and uncontroversial that the modal collapse holds.
+We show a natural deduction proof here merely for the sake of self-containment and so that the proof becomes more accessible to readers who might be more familiarized with natural deduction than with Sobel's proof system.} for the axioms used in the previous sections.
+We show only the strongest implication above.
+The others are easy corollaries in modal logics where T holds.
+% ToDo: do we really need T?
 
 \begin{theorem}
-\label{T3}
-For all constant formulas, $A$, (without free variables) a modal collapse 
+\label{Collapse}
+For all propositions $A$, the following \emph{modal collapse} proposition is provable:
 $$
@@
 $$ 
-is provable. 
 \end{theorem}
@@
 \begin{prooftree}
-\AXC{Theorem 2 of Scott}\dashedLine
+\AXC{Theorem 2}\dashedLine
 \UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
@@
 \UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
+\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_9 $}\dashedLine
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-\begin{small}
-\begin{prooftree}
-\AXC{}\dashedLine
-\UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\AXC{Theorem 3 of Scott}\dashedLine
+\AXC{Theorem 3}\dashedLine
 \UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
@@
 \UIC{$ \omega:  \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
-\AXC{Theorem 3 of Scott}\dashedLine
+\AXC{Theorem 3}\dashedLine
 \UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
@@
 
-\section{Conclusions}\label{sec:conclusion}
-
-The proofs of theorem T1 and corollary C1 ($\pos \ex x.
-G(x)$) do not rely on equality and use axiom A2 ($\all \varphi.
-\all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]$) only once.
-Corollary C2 ($\ex x.
-G(x)$), which is usually regarded as a trivial corollary of main theorem T3 ($\nec \ex x.
-G(x)$) using the modal logic axiom T, is here derived directly from lemma L1 and corollary C1, not relying on T.
-The main theorem T3 becomes derivable from C2 by a single application of the necessitation rule.
-Furthermore, all proofs are done in the modal logic K, except for the proof of corollary C2, which requires one use of the axiom B of modal logic KB.
-
-ToDo: Sobel Anderson on KB: Sobel page 152
-
-Anderson's footnote 5 on KB
-
-ToDo: talk about anderson's footnote 2, where he argues that an anonymous referee knew the simpler proof of T1.
-
-Hajek (magari and others) on non-provability of ex x. G(x) in KD45.
-
-ToDo: Talk about constant domains and varying domains, mention Conchiarella's semantics and Anderson's footnote 14.
-
-ToDo: natural language explanation of the proof in Anderson.
-
-ToDo: Sobel Anderson on KB: Sobel page 152
-and
-Anderson's footnote 5 on KB %Nathan Salmon: On the Logic of What Might Have Been,"
+If G\"odel's ontological proof is abstractly analyzed, then we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being.
+The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas.
+Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being God-like or at least to a restricted collection of properties.
+Several solutions to the problem of modal collapse have been proposed.
+
+Anderson's solution \citep{and90} modifies the definitions of God-like being and essence, and eliminates half of an axiom.
+This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
+Another solution involving more substantial modifications is that of Bj{\o}rdal \cite{bjordal,fuhrmann}.
+
+On another track, Fitting has argued that greater care has to be taken with the semantics of higher-order modal logics.
+Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional.
+Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
+
+Anderson \citep{and90}[p.
+292] and Sobel \citep{sobel2}[p.
+133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
+
+It is likely that a reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.
 
@@
 
-\bibitem[{\itshape Adams(1995)}]{adams}
+\bibitem[{\itshape Adams (1995)}]{adams}
 Adams, R. M. 1995. Introductory note to [G\"odel 1970].
@@
 
-\bibitem[{\itshape Anderson \& Gettings(1996)}]{and}
+\bibitem[{\itshape Anderson \& Gettings (1996)}]{and}
 Anderson, C. A.\& Gettings, M. 1996.  {\itshape G\"odel's ontological proof revisited}. In: edited by Hajek P. {\itshape G\"odel '96},  Springer.
 
-\bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
+\bibitem[{\itshape Bj{\o}rdal (1999)}]{bjordal}
 Bj{\o}rdal, F. 1999. 
@@
 
-\bibitem[{\itshape Church(1940)}]{church}
+\bibitem[{\itshape Church (1940)}]{church}
 Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56?68.
 
-\bibitem[{\itshape Fitting(2002)}]{fitting}
+\bibitem[{\itshape Fitting (2002)}]{fitting}
 Fitting, M. 2002.  {\itshape Types, Tableaus, and G\"odel's God}, Kluwer Academic Publishers.
 
-\bibitem[{\itshape Fuhrmann(2005)}]{fuhrmann}
+\bibitem[{\itshape Fuhrmann (2005)}]{fuhrmann}
 Fuhrmann, A. 2005.
@@
 
-\bibitem[{\itshape Goedel(1970)}]{Goedel} 
+\bibitem[{\itshape Goedel (1970)}]{Goedel} 
 G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
 
-\bibitem[{\itshape Hajek(1996)}]{hajek}
+\bibitem[{\itshape Hajek (1996)}]{hajek}
 Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
@@
 
-\bibitem[{\itshape Kant(1781)}]{kant}
+\bibitem[{\itshape Kant (1781)}]{kant}
 Kant, I.  original 1781. {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
@@
 
-\bibitem[{\itshape Scott (***)}]{Scott}
+\bibitem[{\itshape Scott (2001)}]{scott}
 Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
 
-\bibitem[{\itshape Sobel(1987)}]{sobel}
+\bibitem[{\itshape Sobel (1987)}]{sobel}
 Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying : essays for Richard Cartwright},  MIT Press.
 
-\bibitem[{\itshape Sobel(2001)}]{sobel2}
+\bibitem[{\itshape Sobel (2001)}]{sobel2}
 Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
 
-\bibitem[{\itshape Wang(1996)}]{Wang1996}
+\bibitem[{\itshape Wang (1996)}]{Wang1996}
 Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
 
-***Adams, Fuhrmann, Hajek and Bjordal: Confirm references to ammendations and Church. 
-****References for \cite{coq} and \cite{isabelle} and other papers checked step-by-step in Coq \cite{BenzmullerWolzenlogel-Paleo}. Reference to the inspiration for the modal rules. 
-***One Anderson reference incomplete***
-***Scott reference: the year is missing***
+\bibitem[{\itshape Paulin-Mohring (2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
+All about Proofs, Proofs for All,
+Mathematical Logic and Foundations. College Publications, London.
+
+\bibitem[{\itshape Nipkow \& Paulson \& Wenzel (2002)}]{isabelle} Nipkow, T. \& Paulson, L.C. \& Wenzel, M. 2002. {\itshape Isabelle/HOL: A Proof Assistant for
+Higher-Order Logic}. Number 2283 in LNCS, Springer.
+
+\bibitem[{\itshape Thiele (2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
 

### 2015-10-05 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Commit Annika's corrections

@@
 % \institute{
-%   Helsinki, Finland\\
+%   University of Helsinki, Finland\\
 %   \email{annika.siders@helsinki.fi}
@@
 This paper presents two detailed formalizations of ontological arguments in a simple natural deduction calculus.
-The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills its gaps, thus verifying its correctness.
-The second formal proof improves the first one, by: firstly, relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5} in the proof of the last theorem;  and, secondly, not using equality to prove the first theorem.
-The second proof is also technically shorter than the first one, because it avoids unnecessary detours and uses Axiom 1 only once.
+The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills in the gaps, thus verifying its correctness.
+The second formal proof improves the first one, by: relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5};  and by avoiding the introduction of the equality predicate.
+The second proof is also technically shorter than the first one, because it eliminates unnecessary detours and uses Axiom 1 for the positivity of properties only once.
 \end{abstract}
@@
 Anselm (1033-1109).
-His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also has the property of existence.
+His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also had the property of existence.
 St.
@@
 In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
-G\"odel's work was saved in his own notes \cite{Goedel} as well as notes by Scott \cite{scott}, in whom he confided his proof.
+G\"odel's work was saved in his own notes \cite{Goedel} as well as in notes by Scott \cite{scott}, in whom he confided his proof.
 
@@
 In particular, one of G\"odel's axioms states that any conjunction of positive properties is itself positive.
-From this axiom, it is immediately deduced that the property of being god-like is positive.
+From this axiom, it is immediately deduced that the property of being God-like is positive.
 Intuitively, a (possibly infinite) conjunction of positive properties is deduced from the universal definition of God-likeness.
 This deductive inference is not formalizable in a finite first-order calculus.
-The interplay between universal quantification (in the definition of a god-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom o!
+ f G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
 
@@
 Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
-Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
+Secondly, it makes the ontological proof accessible to people who are familiar with a natural deduction style.
 Thirdly, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this.
@@
 Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
-Dashed lines are used to refer to an axiom or theorem with proof shown elsewhere.
-When proof trees are too large to fit in the page, some branches may be displayed separately.
+Dashed lines are used to refer to an axiom or theorem with the proof shown elsewhere.
+When proof trees are too large to fit on the page, some branches may be displayed separately.
 In such cases, the conclusion of the branch and the location in the main proof tree where the branch belongs are annotated with the same symbol (a subscripted $\subproof$).
@@
 \begin{calculus}
-{propositional rules}
+{Propositional rules}
 {fig:PropositionalRules}
@@
 \begin{calculus}
-{double negation elimination}
+{Double negation elimination}
 {fig:Classical}
@@
 \begin{calculus}
-{quantifier rules}
+{Quantifier rules}
 {fig:QuantifierRules}
@@
 \begin{calculus}
-{Rules for Modal Operators}
+{Rules for modal operators}
 {fig:NDK}
@@
 Without the restriction that every box must be accessed by exactly one strong modal inference, the calculus would be unsound for the modal logic \textbf{K}.
-The formula $\all \psi.
-(\nec \psi \imp \pos \psi)$, for example, is not valid in \textbf{K} but would be provable according to the following proof:
+For example, the formula $\all \psi.
+(\nec \psi \imp \pos \psi)$ is not valid in \textbf{K} but would be provable without this restriction:
 
@@
 
-The example proof above is unsound according to the {\NDK} calculus, because the eigen-box condition is violated: the box labelled by $\omega$ is not accessed by any strong inference.
+This example proof is unsound according to the {\NDK} calculus, because the eigen-box condition is violated: the box labelled by $\omega$ is not accessed by any strong inference.
 
 The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
-From the point of view of a \emph{possible worlds} semantics, rigidity means that their interpretation is independent of the world in which they are being interpreted.
+From the point of view of a \emph{possible world semantics}, rigidity means that their interpretation is independent of the world in which they are being interpreted.
 Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
@@
 \cite{hajek}).
-Nevertheless, some authors of variants \cite[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
-Our choice of assuming constant domains is motivated by simplicity.
+Nevertheless, some authors of variants \cite[footnotes !
+ 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains. Our choice of assuming constant domains is motivated by simplicity.
 
@@
 
-In this section, natural deduction proofs of a three convenient derived modal principles are presented.
+In this section, natural deduction proofs of three convenient derived modal principles are presented.
 These principles are used in the ontological proofs formalized in the next sections.
@@
 Informally, this particular case can be read as the claim that anything that is possibly necessary is in fact necessary, and it is sufficient for the ontological proofs of G\"odel and Scott.
-The proof of this principle depends on the previously proven Brouwer's reduction principle, on the Modal Axiom 5, and on K, whose derivation is shown in the proof of Theorem \ref{theorem:completeness}.
+The proof of this principle depends on the previously proven Brouwer's reduction principle, on the Modal Axiom 5, and on K, which was shown to be derivable in the proof of Theorem \ref{theorem:completeness}.
 
@@
 In this section we present a detailed formalization of Scott's proof in the natural deduction calculus defined in Section \ref{sec:calculus}.
-The reasoning in Scott's manuscript was reproduced step-by-step and any reasoning gaps were filled in ways that we considered most natural.
+The reasoning in Scott's manuscript has been reproduced step-by-step and all reasoning gaps are filled in by using the deduction considered most natural.
 All derived formulas that do appear in Scott's manuscript are marked with a $\dagger$ here.
@@
 In G\"odel's proof, existence is properly denoted by the existential quantifier.
-It is, therefore, unfortunately misleading to refer to the defined predicate symbol $E$ as ``necessary existence'', when in fact it is just a convenient abbreviation.
-
-Besides the 5 axioms and 3 definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
-The use of reflexivity could be avoided, but it gives a shorter proof.
+It is, therefore, unfortunately misleading to refer to the defined predicate symbol $E$ as ``necessary existence'', w!
+ hen in fact it is just a convenient abbreviation.
+
+Beside the 5 axioms and 3 definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
+The use of reflexivity could be avoided, but allows a shorter proof.
 
@@
 \label{T2}
-Being god-like is an essence of any god-like being:
+Being god-like is an essence of any God-like being:
 $$
@@
 \label{L1}
-If there is a god-like being, then there is a god-like being necessarily:
+If there is a God-like being, then there is a God-like being necessarily:
 $$
@@
 \label{L2}
-If the existence of a god-like being is possible, then it is necessary:
+If the existence of a God-like being is possible, then it is necessary:
 $$
@@
 \label{T3}
-Necessarily, there exists a god-like being:
+Necessarily, there exists a God-like being:
 $$
@@
 The key idea to obtain a shorter proof is to instantiate the second universally quantified property of Axiom 2 by $\lambda x.
-\neg \rho(x)$ instead of using $=$ or $\neq$ as G\"odel and Scott did.
+\neg \rho(x)$ instead of using the predicate symbols $=$ or $\neq$ as G\"odel and Scott did.
 This eliminates the reliance of the ontological argument on the presence of $=$ and $\neq$ in the logical language.
@@
 We may only wonder why G\"odel preferred to use those unusual instantiations.
-It might be an interesting question from a historical point of view, if this suggests that there might have been a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
+It might be an interesting question from a historical point of view, if this suggests that there might have b!
+ een a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
 
@@
 
-Since the iteration principle requires the strong modal logic {\bf S5} while Browser's reduction principle relies only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
+Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle relies only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
 There have been many discussions of what modal axioms are needed for the ontological argument.
@@
 The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas.
-Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being God-like or at least to a restricted collection of properties.
+Therefore, an improvement would be obtained if the modal collapse was limited only to the property of being god-like or at least to a restricted collection of properties.
 Several solutions to the problem of modal collapse have been proposed.
@@
 \bibitem[{\itshape Church (1940)}]{church}
-Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56?68.
+Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56--68.
 
@@
 Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
-(eds.) Logic and Logical Algebra, pp. 125-136, Marcel Dekker.
+(eds.) Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
 
@@
 \bibitem[{\itshape Sobel (1987)}]{sobel}
-Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying : essays for Richard Cartwright},  MIT Press.
+Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying: essays for Richard Cartwright},  MIT Press.
 

### 2015-10-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix broken line breaks

@@
 This deductive inference is not formalizable in a finite first-order calculus.
-The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom o!
- f G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
 
@@
 \cite{hajek}).
-Nevertheless, some authors of variants \cite[footnotes !
- 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains. Our choice of assuming constant domains is motivated by simplicity.
+Nevertheless, some authors of variants \cite[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
+Our choice of assuming constant domains is motivated by simplicity.
 
@@
 In G\"odel's proof, existence is properly denoted by the existential quantifier.
-It is, therefore, unfortunately misleading to refer to the defined predicate symbol $E$ as ``necessary existence'', w!
- hen in fact it is just a convenient abbreviation.
+It is, therefore, unfortunately misleading to refer to the defined predicate symbol $E$ as ``necessary existence'', when in fact it is just a convenient abbreviation.
 
@@
 We may only wonder why G\"odel preferred to use those unusual instantiations.
-It might be an interesting question from a historical point of view, if this suggests that there might have b!
- een a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
+It might be an interesting question from a historical point of view, if this suggests that there might have been a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
 

### 2015-10-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve a discussion sentence

@@
 
-If G\"odel's ontological proof is abstractly analyzed, then we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being.
+In G\"odel's ontological proof, we are proving a restricted modal collapse, which applies to one specific formula, the existence of a God-like being.
 The interest in the proof naturally decreases if a consequence of the axiomatization is a modal collapse for \emph{all} formulas.

### 2015-10-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Use double spacing, as required by journal
Create title page, as required by journal

@@
 \documentclass{llncs}
+
+\usepackage{setspace}
+\doublespacing
 
@@
 
-% ToDo: uncomment this post-acceptance: the journal uses double blind reviewing.
-% \author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2}}
-
-% \authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
-
-% \institute{
-%   University of Helsinki, Finland\\
-%   \email{annika.siders@helsinki.fi}
-%   \and 
-%   Vienna University of Technology, Austria ; Australian National University \\
-%   \email{bruno.wp@gmail.com} 
-% }
-
+\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2,3}}
+
+\authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
+
+\institute{
+  \email{annika.siders@helsinki.fi} \\
+  +358 2941 29241
+  Department of Philosophy \\
+  University of Helsinki\\
+  P.O. Box 24 (Unioninkatu 40 A) \\
+  Finland \\[2em]
+  \and 
+  \email{bruno@logic.at}
+  Theory and Logic Group \\ 
+  Vienna University of Technology \\ 
+  Austria \\[2em]
+  \and
+  \email{bruno.woltzenlogel.paleo@anu.edu.au}
+  Logic and Computation Group \\
+  College of Engineering and Computer Science \\
+  Australian National University \\
+  Australia
+}
+
+\begin{document}
+
+\maketitle
+
+\clearpage
+
+\title{A Variant of G\"{o}del's Ontological Proof \\
+in a Natural Deduction Calculus}
 \author{}
@@
 \institute{}
-
-\begin{document}
-
 \maketitle
@@
 The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills in the gaps, thus verifying its correctness.
-The second formal proof improves the first one, by: relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5};  and by avoiding the introduction of the equality predicate.
+The second formal proof improves the first one, by relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5}; and by avoiding the introduction of the equality predicate.
 The second proof is also technically shorter than the first one, because it eliminates unnecessary detours and uses Axiom 1 for the positivity of properties only once.

### 2015-12-11 Papers/InProgress/IJCAI/ijcai15.tex
IJCAI draft/plan added

%%%% ijcai15.tex

\typeout{IJCAI-15 Instructions for Authors}

% These are the instructions for authors for IJCAI-15.
% They are the same as the ones for IJCAI-11 with superficical wording
%   changes only.

\documentclass{article}
% The file ijcai15.sty is the style file for IJCAI-15 (same as ijcai07.sty).
\usepackage{ijcai15}

% Use the postscript times font!
\usepackage{times}

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

\title{IJCAI--15 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
\author{Qiang Yang \\
Hong Kong University of Science and Technology\\
Hong Kong, China \\
pcchair15@ijcai.org}

\begin{document}

\maketitle

\begin{abstract}
  The {\it IJCAI--15 Proceedings} will be printed from electronic
  manuscripts submitted by the authors. The electronic manuscript will
  also be included in the online version of the proceedings. This paper
  provides the style instructions.
\end{abstract}

\section{Introduction}

The {\it IJCAI--15 Proceedings} will be printed from electronic
manuscripts submitted by the authors. These must be PDF ({\em Portable
Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.

\subsection{Length of Papers}

Each accepted full paper is allocated six pages in the conference 
proceedings, excluded references. References can take up to one page.
Up to two additional pages may be purchased at a price 
of \$275 per page for any accepted paper. However, all 
{\em submissions} must 
be a maximum of six pages, plus at most one for references, in length.

\subsection{Word Processing Software}

As detailed below, IJCAI has prepared and made available a set of
\LaTeX{} macros and a Microsoft Word template for use in formatting
your paper. If you are using some other word processing software (such
as WordPerfect, etc.), please follow the format instructions given
below and ensure that your final paper looks as much like this sample
as possible.

\section{Style and Format}

\LaTeX{} and Word style files that implement these instructions
can be retrieved electronically. (See Appendix~\ref{stylefiles} for
instructions on how to obtain these files.)

\subsection{Layout}

Print manuscripts two columns to a page, in the manner in which these
instructions are printed. The exact dimensions for pages are:
\begin{itemize}
\item left and right margins: .75$''$
\item column width: 3.375$''$
\item gap between columns: .25$''$
\item top margin---first page: 1.375$''$
\item top margin---other pages: .75$''$
\item bottom margin: 1.25$''$
\item column height---first page: 6.625$''$
\item column height---other pages: 9$''$
\end{itemize}

All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
A4-size paper, use the given top and left margins, column width,
height, and gap, and modify the bottom and right margins as necessary.

\subsection{Format of Electronic Manuscript}

For the production of the electronic manuscript, you must use Adobe's
{\em Portable Document Format} (PDF). A PDF file can be generated, for
instance, on Unix systems using {\tt ps2pdf} or on Windows systems
using Adobe's Distiller. There is also a website with free software
and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
\LaTeX2e{}, this is accomplished by putting
\begin{quote} 
\mbox{\tt $\backslash$usepackage\{times\}}
\end{quote}
in the preamble.\footnote{You may want also to use the package {\tt
latexsym}, which defines all symbols known from the old \LaTeX{}
version.}

Additionally, it is of utmost importance to specify the American {\bf
letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
formatting the paper. When working with {\tt dvips}, for instance, one
should specify {\tt -t letter}.

\subsection{Title and Author Information}

Center the title on the entire width of the page in a 14-point bold
font. Below it, center the author name(s) in a 12-point bold font, and
then center the address(es) in a 12-point regular font. Credit to a
sponsoring agency can appear on the first page as a footnote.

\subsubsection{Blind Review}

In order to make blind reviewing possible, authors must omit their
names and affiliations when submitting the paper for review. In place
of names and affiliations, provide a list of content areas. When
referring to one's own work, use the third person rather than the
first person. For example, say, ``Previously,
Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
that\ldots'' Try to avoid including any information in the body of the
paper or references that would identify the authors or their
institutions. Such information can be added to the final camera-ready
version for publication.

\subsection{Abstract}

Place the abstract at the beginning of the first column 3$''$ from the
top of the page, unless that does not leave enough room for the title
and author information. Use a slightly smaller width than in the body
of the paper. Head the abstract with ``Abstract'' centered above the
body of the abstract in a 12-point bold font. The body of the abstract
should be in the same font as the body of the paper.

The abstract should be a concise, one-paragraph summary describing the
general thesis and conclusion of your paper. A reader should be able
to learn the purpose of the paper and the reason for its importance
from the abstract. The abstract should be no more than 200 words long.

\subsection{Text}

The main body of the text immediately follows the abstract. Use
10-point type in a clear, readable font with 1-point leading (10 on
11).

Indent when starting a new paragraph, except after major headings.

\subsection{Headings and Sections}

When necessary, headings should be used to separate major sections of
your paper. (These instructions use many headings to demonstrate their
appearance; your paper should have fewer headings.)

\subsubsection{Section Headings}

Print section headings in 12-point bold type in the style shown in
these instructions. Leave a blank space of approximately 10 points
above and 4 points below section headings.  Number sections with
arabic numerals.

\subsubsection{Subsection Headings}

Print subsection headings in 11-point bold type. Leave a blank space
of approximately 8 points above and 3 points below subsection
headings. Number subsections with the section number and the
subsection number (in arabic numerals) separated by a
period.

\subsubsection{Subsubsection Headings}

Print subsubsection headings in 10-point bold type. Leave a blank
space of approximately 6 points above subsubsection headings. Do not
number subsubsections.

\subsubsection{Special Sections}

You may include an unnumbered acknowledgments section, including
acknowledgments of help from colleagues, financial support, and
permission to publish.

Any appendices directly follow the text and look like sections, except
that they are numbered with capital letters instead of arabic
numerals.

The references section is headed ``References,'' printed in the same
style as a section heading but without a number. A sample list of
references is given at the end of these instructions. Use a consistent
format for references, such as that provided by Bib\TeX{}. The reference
list should not include unpublished work.

\subsection{Citations}

Citations within the text should include the author's last name and
the year of publication, for example~\cite{gottlob:nonmon}.  Append
lowercase letters to the year in cases of ambiguity.  Treat multiple
authors as in the following examples:~\cite{abelson-et-al:scheme}
or~\cite{bgf:Lixto} (for more than two authors) and
\cite{brachman-schmolze:kl-one} (for two authors).  If the author
portion of a citation is obvious, omit it, e.g.,
Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
\nocite{abelson-et-al:scheme}
\nocite{bgf:Lixto}
\nocite{brachman-schmolze:kl-one}
\nocite{gottlob:nonmon}
\nocite{gls:hypertrees}
\nocite{levesque:functional-foundations}
\nocite{levesque:belief}
\nocite{nebel:jair-2000}

\subsection{Footnotes}

Place footnotes at the bottom of the page in a 9-point font.  Refer to
them with superscript numbers.\footnote{This is how your footnotes
should appear.} Separate them from the text by a short
line.\footnote{Note the line separating these footnotes from the
text.} Avoid footnotes as much as possible; they interrupt the flow of
the text.

\section{Illustrations}

Place all illustrations (figures, drawings, tables, and photographs)
throughout the paper at the places where they are first discussed,
rather than at the end of the paper. If placed at the bottom or top of
a page, illustrations may run across both columns.

Illustrations must be rendered electronically or scanned and placed
directly in your document. All illustrations should be in black and
white, as color illustrations may cause problems. Line weights should
be 1/2-point or thicker. Avoid screens and superimposing type on
patterns as these effects may not reproduce well.

Number illustrations sequentially. Use references of the following
form: Figure 1, Table 2, etc. Place illustration numbers and captions
under illustrations. Leave a margin of 1/4-inch around the area
covered by the illustration and caption.  Use 9-point type for
captions, labels, and other text in illustrations.

\section*{Acknowledgments}

The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
files that implement them was supported by Schlumberger Palo Alto
Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
Preparation of the Microsoft Word file was supported by IJCAI.  An
early version of this document was created by Shirley Jowell and Peter
F. Patel-Schneider.  It was subsequently modified by Jennifer
Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
These instructions are the same as the ones for IJCAI--05, prepared by
Kurt Steinkraus, Massachusetts Institute of Technology, Computer
Science and Artificial Intelligence Lab.

\appendix

\section{\LaTeX{} and Word Style Files}\label{stylefiles}

The \LaTeX{} and Word style files are available on the IJCAI--15
website, {\tt http://www.ijcai-15.org/}.
These style files implement the formatting instructions in this
document.

The \LaTeX{} files are {\tt ijcai15.sty} and {\tt ijcai15.tex}, and
the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai15.bib}. The
\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
style file is for version 0.99c of Bib\TeX{} ({\em not} version
0.98i). The {\tt ijcai15.sty} file is the same as the {\tt
ijcai07.sty} file used for IJCAI--07.

The Microsoft Word style file consists of a single file, {\tt
ijcai15.doc}. This template is the same as the one used for
IJCAI--07.

These Microsoft Word and \LaTeX{} files contain the source of the
present document and may serve as a formatting sample.

Further information on using these styles for the preparation of
papers for IJCAI--15 can be obtained by contacting {\tt
pcchair15@ijcai.org}.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai15}

\end{document}

### 2015-12-11 Papers/InProgress/IJCAI/paper.tex
IJCAI draft/plan added

--- 2015-12-11 Papers/InProgress/IJCAI/ijcai15.tex
@@
 
-\typeout{IJCAI-15 Instructions for Authors}
+\typeout{Automating the Ontological Argument}
 
@@
 
-\title{IJCAI--15 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
-\author{Qiang Yang \\
-Hong Kong University of Science and Technology\\
-Hong Kong, China \\
-pcchair15@ijcai.org}
+\title{Automating the Ontological Argument}
+\author{Christoph Benzmüller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
+\author{}
 
@@
 \begin{abstract}
-  The {\it IJCAI--15 Proceedings} will be printed from electronic
-  manuscripts submitted by the authors. The electronic manuscript will
-  also be included in the online version of the proceedings. This paper
-  provides the style instructions.
+  Higher-order automated theorem provers have revealed some
+  philosophically profound new knowledge in metaphysics.
 \end{abstract}
@@
 \section{Introduction}
+G\"{o}del's ontological argument for the existence of God is frequently
+presented as a masterpiece argument in metaphysics. A huge body of
+literature is available which discusses G\"{o}del's original work and/or
+subsequent variants of it.
 
-The {\it IJCAI--15 Proceedings} will be printed from electronic
-manuscripts submitted by the authors. These must be PDF ({\em Portable
-Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.
+This paper presents an in depth analysis of G\"{o}del's original version
+of the ontological argument.
 
-\subsection{Length of Papers}
+The analysis has been conducted with automated theorem provers for
+classical higher-order logic (HOL). However, since higher-order modal
+logic is actually required for the encoding of G\"{o}del's argument, we
+utilize an semantic embedding of HOML in HOL.
 
-Each accepted full paper is allocated six pages in the conference 
-proceedings, excluded references. References can take up to one page.
-Up to two additional pages may be purchased at a price 
-of \$275 per page for any accepted paper. However, all 
-{\em submissions} must 
-be a maximum of six pages, plus at most one for references, in length.
+First applications of theorem proving technology in metaphysics
+appeared in 199? in the work Zalta and Oppenheimer.  Further work also
+includes Rushby \ldots. Different to what is presented here, none of
+these previous formalises and automates variants of HOML, which is
+however crucial for G\"{o}del's ontological argument.
 
-\subsection{Word Processing Software}
+The work presented here substantially extends previous work \cite{}.  
+The novel contributions of this paper include:
 
-As detailed below, IJCAI has prepared and made available a set of
-\LaTeX{} macros and a Microsoft Word template for use in formatting
-your paper. If you are using some other word processing software (such
-as WordPerfect, etc.), please follow the format instructions given
-below and ensure that your final paper looks as much like this sample
-as possible.
+\begin{itemize}
+\item Detailed analysis of the inconsistency of G\"{o}del's axioms. The
+inconsistency was detected before, but for a long time we were not
+able to extract an intuitive argument from the proof of the ATPs.
+Further experiments now led to discovery of a surprisingly accessible
+understanding of this inconsistency which is philosophically profound
+and which has not been presented yet in the respective literature.
 
-\section{Style and Format}
+\item Fully automated proof of Scott's theorem T3 from the axioms, that
+   is, without intermediate argumentation steps as used in previous
+   work.
 
-\LaTeX{} and Word style files that implement these instructions
-can be retrieved electronically. (See Appendix~\ref{stylefiles} for
-instructions on how to obtain these files.)
+\item Direct verification of modal collapse with Meson in Isabelle/HOL,
+which did not work for the previous.
 
-\subsection{Layout}
-
-Print manuscripts two columns to a page, in the manner in which these
-instructions are printed. The exact dimensions for pages are:
-\begin{itemize}
-\item left and right margins: .75$''$
-\item column width: 3.375$''$
-\item gap between columns: .25$''$
-\item top margin---first page: 1.375$''$
-\item top margin---other pages: .75$''$
-\item bottom margin: 1.25$''$
-\item column height---first page: 6.625$''$
-\item column height---other pages: 9$''$
+\item On a technical level we show that the novel embedding of S5 is more
+    effective than the previous one.
 \end{itemize}
 
-All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
-A4-size paper, use the given top and left margins, column width,
-height, and gap, and modify the bottom and right margins as necessary.
+These novel contributions are presented in Sections ? and ?.
 
-\subsection{Format of Electronic Manuscript}
+In Section ?, novel and previous findings are summarised and discussed 
+with regard to their relevance in metaphysics. In this section also 
+some further details and improvements on earlier 
+findings are presented.
 
-For the production of the electronic manuscript, you must use Adobe's
-{\em Portable Document Format} (PDF). A PDF file can be generated, for
-instance, on Unix systems using {\tt ps2pdf} or on Windows systems
-using Adobe's Distiller. There is also a website with free software
-and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
-uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
-\LaTeX2e{}, this is accomplished by putting
-\begin{quote} 
-\mbox{\tt $\backslash$usepackage\{times\}}
-\end{quote}
-in the preamble.\footnote{You may want also to use the package {\tt
-latexsym}, which defines all symbols known from the old \LaTeX{}
-version.}
+Section ? concludes the paper.
 
-Additionally, it is of utmost importance to specify the American {\bf
-letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
-formatting the paper. When working with {\tt dvips}, for instance, one
-should specify {\tt -t letter}.
+\section{Variants of the Ontological Argument}
+Short history, Anselm, Leibniz, etc.
+\subsection{G\"{o}del's Proof Script}
+\subsection{Scott's Variant}
+\subsection{Recent Variants}
 
-\subsection{Title and Author Information}
+\section{Automating HOML in HOL}
+\subsection{Outline of the Embedding}
+\subsection{Improved Embedding for S5}
 
-Center the title on the entire width of the page in a 14-point bold
-font. Below it, center the author name(s) in a 12-point bold font, and
-then center the address(es) in a 12-point regular font. Credit to a
-sponsoring agency can appear on the first page as a footnote.
+\section{Intuitive Inconsistency Argument}
+\subsection{LEO-II's Inaccessible Inconsistency Proof}
+Tell the story (here or earlier). Show parts of LEO proof.  Inaccessible, but relevant knowledge
+is contained.
 
-\subsubsection{Blind Review}
+\subsection{Argument Reconstruction in Isabelle}
+Present the argument informal and in Isabelle. Easy to understand in 
+in KB and KT, slightly harder in K.
 
-In order to make blind reviewing possible, authors must omit their
-names and affiliations when submitting the paper for review. In place
-of names and affiliations, provide a list of content areas. When
-referring to one's own work, use the third person rather than the
-first person. For example, say, ``Previously,
-Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
-than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
-that\ldots'' Try to avoid including any information in the body of the
-paper or references that would identify the authors or their
-institutions. Such information can be added to the final camera-ready
-version for publication.
+\section{Full Automation of Scott's Proof}
+Within the modified logic S5 Scott's proof can be fully automated. The
+intermediate argumentation steps as not needed anymore to support the
+prover.
 
-\subsection{Abstract}
+In the modified S5 setting also further results can be automatically verified, which 
+failed in previous work: verification of  modal collapse in
+Isabelle/HOL with Meson.
 
-Place the abstract at the beginning of the first column 3$''$ from the
-top of the page, unless that does not leave enough room for the title
-and author information. Use a slightly smaller width than in the body
-of the paper. Head the abstract with ``Abstract'' centered above the
-body of the abstract in a 12-point bold font. The body of the abstract
-should be in the same font as the body of the paper.
+\section{Conclusion}
 
-The abstract should be a concise, one-paragraph summary describing the
-general thesis and conclusion of your paper. A reader should be able
-to learn the purpose of the paper and the reason for its importance
-from the abstract. The abstract should be no more than 200 words long.
-
-\subsection{Text}
-
-The main body of the text immediately follows the abstract. Use
-10-point type in a clear, readable font with 1-point leading (10 on
-11).
-
-Indent when starting a new paragraph, except after major headings.
-
-\subsection{Headings and Sections}
-
-When necessary, headings should be used to separate major sections of
-your paper. (These instructions use many headings to demonstrate their
-appearance; your paper should have fewer headings.)
-
-\subsubsection{Section Headings}
-
-Print section headings in 12-point bold type in the style shown in
-these instructions. Leave a blank space of approximately 10 points
-above and 4 points below section headings.  Number sections with
-arabic numerals.
-
-\subsubsection{Subsection Headings}
-
-Print subsection headings in 11-point bold type. Leave a blank space
-of approximately 8 points above and 3 points below subsection
-headings. Number subsections with the section number and the
-subsection number (in arabic numerals) separated by a
-period.
-
-\subsubsection{Subsubsection Headings}
-
-Print subsubsection headings in 10-point bold type. Leave a blank
-space of approximately 6 points above subsubsection headings. Do not
-number subsubsections.
-
-\subsubsection{Special Sections}
-
-You may include an unnumbered acknowledgments section, including
-acknowledgments of help from colleagues, financial support, and
-permission to publish.
-
-Any appendices directly follow the text and look like sections, except
-that they are numbered with capital letters instead of arabic
-numerals.
-
-The references section is headed ``References,'' printed in the same
-style as a section heading but without a number. A sample list of
-references is given at the end of these instructions. Use a consistent
-format for references, such as that provided by Bib\TeX{}. The reference
-list should not include unpublished work.
-
-\subsection{Citations}
-
-Citations within the text should include the author's last name and
-the year of publication, for example~\cite{gottlob:nonmon}.  Append
-lowercase letters to the year in cases of ambiguity.  Treat multiple
-authors as in the following examples:~\cite{abelson-et-al:scheme}
-or~\cite{bgf:Lixto} (for more than two authors) and
-\cite{brachman-schmolze:kl-one} (for two authors).  If the author
-portion of a citation is obvious, omit it, e.g.,
-Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
-follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
-\nocite{abelson-et-al:scheme}
-\nocite{bgf:Lixto}
-\nocite{brachman-schmolze:kl-one}
-\nocite{gottlob:nonmon}
-\nocite{gls:hypertrees}
-\nocite{levesque:functional-foundations}
-\nocite{levesque:belief}
-\nocite{nebel:jair-2000}
-
-\subsection{Footnotes}
-
-Place footnotes at the bottom of the page in a 9-point font.  Refer to
-them with superscript numbers.\footnote{This is how your footnotes
-should appear.} Separate them from the text by a short
-line.\footnote{Note the line separating these footnotes from the
-text.} Avoid footnotes as much as possible; they interrupt the flow of
-the text.
-
-\section{Illustrations}
-
-Place all illustrations (figures, drawings, tables, and photographs)
-throughout the paper at the places where they are first discussed,
-rather than at the end of the paper. If placed at the bottom or top of
-a page, illustrations may run across both columns.
-
-Illustrations must be rendered electronically or scanned and placed
-directly in your document. All illustrations should be in black and
-white, as color illustrations may cause problems. Line weights should
-be 1/2-point or thicker. Avoid screens and superimposing type on
-patterns as these effects may not reproduce well.
-
-Number illustrations sequentially. Use references of the following
-form: Figure 1, Table 2, etc. Place illustration numbers and captions
-under illustrations. Leave a margin of 1/4-inch around the area
-covered by the illustration and caption.  Use 9-point type for
-captions, labels, and other text in illustrations.
+Good improvement (technical sense), Sledgehammer can still be improved
+(LEO-II finds inconsistency when call directly in THF Syntax, but not
+from within Isabelle). Without independent experiments directly in THF 
+the inconsistency would thus not have been detected.
 
@@
 
-The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
-files that implement them was supported by Schlumberger Palo Alto
-Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-Preparation of the Microsoft Word file was supported by IJCAI.  An
-early version of this document was created by Shirley Jowell and Peter
-F. Patel-Schneider.  It was subsequently modified by Jennifer
-Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
-These instructions are the same as the ones for IJCAI--05, prepared by
-Kurt Steinkraus, Massachusetts Institute of Technology, Computer
-Science and Artificial Intelligence Lab.
+Will be added at a later point.
+
+%German Research Foundation DFG, Chad Brown
 

### 2015-12-11 Papers/InProgress/IJCAI/paper.tex
Figures added
added formalization with improved notation

@@
 
+\usepackage{graphicx}
+
 \title{Automating the Ontological Argument}
@@
 G\"{o}del's ontological argument for the existence of God is frequently
-presented as a masterpiece argument in metaphysics. A huge body of
+presented as a masterpiece argument in metaphysics. A large body of
 literature is available which discusses G\"{o}del's original work and/or
@@
 is contained.
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
+\caption{Primitive substitution in LEO-II generates candidates for the
+empty property.} \label{LEO-Proof}
+\end{figure*}
+LEO-II's resolution proof is human unintuitive. However, it contained
+relevant hints to the empty property (cf. Fig.~\ref{LEO-Proof}).
 
@@
 in KB and KT, slightly harder in K.
+
+\subsection{Mapping to Gödel's Script}
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
+\caption{Sources of Inconsistence in G\"{o}del's proof script.} \label{GoedelScript}
+\end{figure*}
 

### 2015-12-11 Papers/InProgress/IJCAI/paper.tex
added graphics

@@
 \subsection{Argument Reconstruction in Isabelle}
-Present the argument informal and in Isabelle. Easy to understand in 
+Present the argument informal and in Isabelle (see
+Fig.~\ref{InconsistencyIsabelleK}). Easy to understand in 
 in KB and KT, slightly harder in K.
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/InconsistencyIsabelleK.png}}
+\caption{Reconstruction of inconsistency in Isabelle/HOl..} \label{InconsistencyIsabelleK}
+\end{figure*}
 

### 2015-12-12 Papers/InProgress/IJCAI/Formalizations/call_tptp.sh
improvements on syntax pictures in paper

@@ -1,0 +1,47 @@
+tptp4X=./tptp4X
+call=./RemoteSOT.pl
+# if the above programs don't work for you download the SystemOnTPTP package,
+# install it and change the above paths accordingly
+
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
+#proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
+proverlist='LEO-II--- Satallax--- Nitpick--- '
+#proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
+
+timeout=200
+
+suffix=$(date "+%Y-%m-%d--%H:%M:%S")
+resultsfile=RESULTS-$suffix
+
+touch $resultsfile
+rm $resultsfile
+
+echo
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\)
+echo Asking various HOL-ATPs in Miami remotely \(thanks to Geoff Sutcliffe\) >> $resultsfile
+
+for file in "$@"
+do
+echo
+echo >> $resultsfile
+    for prover in $proverlist 
+    do
+    function_to_fork () { 
+      res=""
+      while ! (echo $res | grep --quiet RESULT); do
+          res=`$tptp4X -t $timeout -x $file | $call -t $timeout -s $prover |grep RESULT` 
+	  echo ${file}---${prover}---${res}
+	  echo ${file}---${prover}---${res} >> $resultsfile
+      done   
+    }
+    function_to_fork &
+    done
+    wait
+    echo Done 
+    echo Done >> $resultsfile
+done
+
+
+
+
+

### 2015-12-12 Papers/InProgress/IJCAI/paper.tex
minor improvements
Further updates types reinserted in QML_S5 since without types Modal Collapse has a counter example!

@@
     effective than the previous one.
+
+\item Much improved syntax presentation.
+
+\item Technical issues detected within Isabelle-Sledgehammer; room for improvements
 \end{itemize}
@@
 \subsection{Improved Embedding for S5}
+We present an improved embedding (syntax and defintion of box) in  Fig.~\ref{QMLS5}).
+\begin{figure}
+\centerline{\includegraphics[width=\columnwidth]{./Images/QMLS5.png}}
+\caption{Improved Embedding of S5} \label{QMLS5}
+\end{figure}
+
+\section{Full Automation of Scott's Proof}
+Within the modified logic S5 Scott's proof can be fully automated. The
+intermediate argumentation steps as not needed anymore to support the
+prover, see Fig.~\ref{ScottS5}).
+\begin{figure}
+\centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
+\caption{Full Automation of the Ontological Argument in S5} \label{ScottS5}
+\end{figure}
+
+In the modified S5 setting also further results can be automatically verified, which 
+failed in previous work: verification of  modal collapse in
+Isabelle/HOL with Meson.
 
@@
 in KB and KT, slightly harder in K.
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/InconsistencyIsabelleK.png}}
+\begin{figure}
+\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
 \caption{Reconstruction of inconsistency in Isabelle/HOl..} \label{InconsistencyIsabelleK}
-\end{figure*}
+\end{figure}
 
@@
 \end{figure*}
-
-\section{Full Automation of Scott's Proof}
-Within the modified logic S5 Scott's proof can be fully automated. The
-intermediate argumentation steps as not needed anymore to support the
-prover.
-
-In the modified S5 setting also further results can be automatically verified, which 
-failed in previous work: verification of  modal collapse in
-Isabelle/HOL with Meson.
 

### 2015-12-22 Papers/InProgress/IJCAI/paper.tex
Types reinserted in QML_S5 -- without types Modal Collapse has a Countermodel! This is interesting for further studies.
references added (essence)

@@
 \subsection{Recent Variants}
+Dozens of contributions were published in the last four decades which
+present variants of G\"odel's/Scott's proofs. In many cases the
+motivation has been to remedy the modal collapse. Other authors are
+simply interested in a critical assessment. Particularly interesting
+from the perspective of this paper is that many literature
+contributions unfortunately work with or refer to G''odel's original
+definition of essence
+$$ todo $$
+This original version avoids the conjunct added by Scott expressing
+that essential properties of an individual should be possessed by the
+individual. As we show in this paper using this notion of essence
+leads to an inconsistent axiom system, which in a classical setting of
+course deserves no further attention, since from an inconsistent
+system everything can be concluded, including God's existence and
+non-existence simultaneously. In other words, these literature
+contributions discuss variants of the argument which from a classical
+logical perspective do not deserve any further attention.
+
+We give here a list of examples:
+\begin{itemize}
+\item Anderson and Getting's 1996, p. 168 \cite[p.168]{AndersonGettings1968}
+  use essence without conjunct.
+\item  Hazen 1998, p.365 \cite[p.365]{Hazen1998} states: ``G\"odel left this
+  clause out in [11], but this appears to have been an oversight--it
+  is included in related manuscripts''. There is not mentioning of an
+  inconsistency though.
+\item Look???, p. 514
+\item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
+  \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
+  Oppy uses: ``A is an essence of x iff for every property B, x has B
+  neces- sarily iff A entails B'' (this is from Anderson's
+  emendation). Check is this leads to inconsistency as well. In this
+  case we may write something like:
+
+ Oppy presents an critical
+  assessment in order to conclude that G\"odel's argument fails to
+  convince. Oppy apparently does not see that he is assessing an
+  inconsistent axiom system (and he in fact creates adaptations that
+  suffer the same problems).
+\item 
+\end{itemize}
+
+\footnote{Fuhrmann2005: Gödel vermerkt diese Konsequenz [dass aus der Definition unmittelbar folgt, daß alle wesentlichen
+Eigenschaften notwendig äquivalent sind] in einer Fußnote zur Definition.
+Die Definition selbst aber läßt im Definiens das Konjunkt Xx aus.
+Ohne das Konjunkt folgt jedoch die Konsequenz nicht.
+Es ist deshalb naheliegend anzunehmen, daß Gödel die Definition so beabsichtigte, wie sie hier notiert ist und wie Gödel sie selbst in früheren Notizbucheintragungen formuliert hat.}
 
@@
 \centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
-\caption{Full Automation of the Ontological Argument in S5} \label{ScottS5}
+\caption{Full Automation of the Ontological Argument in S5 } \label{ScottS5}
 \end{figure}
@@
 \begin{figure}
-\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+%\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
 \caption{Reconstruction of inconsistency in Isabelle/HOl..} \label{InconsistencyIsabelleK}

### 2015-12-22 Formalizations/THF/call_tptp.sh
nitpick instead of nitrox

@@ -5,7 +5,7 @@ call=./RemoteSOT.pl
 
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Isabelle---2013 Nitrox---2013 agsyHOL---1.0 TPS---3.120601S1b'
 #proverlist='LEO-II---1.6.2 Satallax---2.7 Nitrox---2013'
-proverlist='TPS---3.120601S1b LEO-II--- Satallax--- Nitrox--- '
+proverlist='TPS---3.120601S1b LEO-II--- Satallax--- Nitpick--- '
 #proverlist='LEO-II---1.6.2 Satallax-MaLeS---1.2 Nitrox---2013'
 
 timeout=200

### 2015-12-23 Papers/InProgress/IJCAI/paper.tex
minor updates

@@
 \item Look???, p. 514
-\item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
-  \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
-  Oppy uses: ``A is an essence of x iff for every property B, x has B
-  neces- sarily iff A entails B'' (this is from Anderson's
-  emendation). Check is this leads to inconsistency as well. In this
-  case we may write something like:
+% \item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
+%   \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
+%   Oppy uses: ``A is an essence of x iff for every property B, x has B
+%   neces- sarily iff A entails B'' (this is from Anderson's
+%   emendation). Check is this leads to inconsistency as well. In this
+%   case we may write something like:
 
- Oppy presents an critical
-  assessment in order to conclude that G\"odel's argument fails to
-  convince. Oppy apparently does not see that he is assessing an
-  inconsistent axiom system (and he in fact creates adaptations that
-  suffer the same problems).
-\item 
+%  Oppy presents an critical
+%   assessment in order to conclude that G\"odel's argument fails to
+%   convince. Oppy apparently does not see that he is assessing an
+%   inconsistent axiom system (and he in fact creates adaptations that
+%   suffer the same problems).
+% \item 
 \end{itemize}
@@
 \subsection{Outline of the Embedding}
+Logic textbooks commonly utilise classical higher-order logic (HOL)
+\cite{Church} as a meta-language to introduce the syntax and the
+semantics of object logics of interest, in which then reasoning
+problems in concrete application domains can be modeled and solved
+with pen and paper.  In fact, this approach can also be followed to
+enable interactive and automated proof for even very challenging
+objects logics with existing theorem provers for classical
+higher-order logic.
+
+% Just as commonly the case in logic textbooks, in the embeddings
+% approach classical higher-order logic is utilised as a meta-language
+% to encode the syntax and the semantics of object logics in which the
+% proof problems of interest are then modeled in. Modulo the embedding
+% state of the art automated theorem provers for classical higher-order
+% logic can then be utilised as object logic reasoning tools.
+
+For a computational analysis of G\"odel's ontological argument in this
+approach the embedding of modal higher-order logics (MHOL) such as K,
+KB and S5 for different domain conditions (possibilist and actualist)
+is required. This idea has been successfully followed in related work
+\cite{C40}. The embedding of modal higher-order logic is in fact
+straightforward. Formulas in MHOL such as are lifted to predicates
+over worlds, which are explicitly represented in the approach as
+terms. Propositional MHOL such as $\neg \varphi$, $\varphi\vee\psi$
+and $\box \varphi$ are then mapped to HOL formulas \ldots.
+
+It is easy
+to see that this captures the standard translation \cite{Ohlbach},
+which is here (intra-logically) realised in HOL by stating a set of
+equations; see also Fig. ~\ref{QMLS5}. New is that also quantifiers 
+can be mapped \ldots
+
+The approach is very flexible, KB, S5, domain conditions\ldots
+
 \subsection{Improved Embedding for S5}

### 2015-12-23 Papers/InProgress/IJCAI/paper.tex
wrote introduction

@@
 
-\usepackage{graphicx}
+\usepackage{graphicx,url}
 
@@
 \section{Introduction}
-G\"{o}del's ontological argument for the existence of God is frequently
-presented as a masterpiece argument in metaphysics. A large body of
-literature is available which discusses G\"{o}del's original work and/or
-subsequent variants of it.
+Without exaggeration one may claim that Kurt G\"{o}del's ontological
+argument for the existence of God \cite{GoedelNotes,ScottNotes} is
+amongst the most discussed formal proofs in modern literature. A rich
+body of publications -- including very recent ones -- present,
+discuss, assess, criticise, modify and improve G\"{o}del's original
+work (see e.g.~Sobel~\shortcite{Sobel} and Oppy~\shortcite{sep-ontological-arguments} and the
+references therein).  In philosophy lectures at universities the
+argument is regularly presented as a masterpiece argument in
+metaphysics. Since summer 2013, when Benzm\"uller and Woltzenlogel-Paleo~\shortcite{J30,C40} first
+reported on their successful initial computational
+analysis of the G\"odel's respectively Scott's variants of the proof,
+their contribution received a media repercussion on a global scale\footnote{A
+  small collection of these reports is available at {\scriptsize
+    \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
+and numerous bloggers commented on the proof
+\cite{fuhrmann15:_blogg_goedel}.
 
-This paper presents an in depth analysis of G\"{o}del's original version
-of the ontological argument.
+The in-depth computational analysis presented here substantially
+extends previous computer-assisted studies on G\"odel's ontological
+argument. Similar to the related work \cite{J30,C40} the analysis has
+been conducted with automated theorem provers for classical
+higher-order logic (HOL) even though G\"odel's proof is actually
+formulated in modal higher-order logic. To bridge between the two
+logic formalisms we utilise and further improve the logic embeddings
+approach \cite{J23,C40} as also employed in preceding related work.
 
-The analysis has been conducted with automated theorem provers for
-classical higher-order logic (HOL). However, since higher-order modal
-logic is actually required for the encoding of G\"{o}del's argument, we
-utilize an semantic embedding of HOML in HOL.
-
-First applications of theorem proving technology in metaphysics
-appeared in 199? in the work Zalta and Oppenheimer.  Further work also
-includes Rushby \ldots. Different to what is presented here, none of
-these previous formalises and automates variants of HOML, which is
-however crucial for G\"{o}del's ontological argument.
-
-The work presented here substantially extends previous work \cite{}.  
-The novel contributions of this paper include:
+The novel contributions of this paper include (but are not limited to):
 
 \begin{itemize}
-\item Detailed analysis of the inconsistency of G\"{o}del's axioms. The
-inconsistency was detected before, but for a long time we were not
-able to extract an intuitive argument from the proof of the ATPs.
-Further experiments now led to discovery of a surprisingly accessible
-understanding of this inconsistency which is philosophically profound
-and which has not been presented yet in the respective literature.
-
-\item Fully automated proof of Scott's theorem T3 from the axioms, that
-   is, without intermediate argumentation steps as used in previous
-   work.
-
-\item Direct verification of modal collapse with Meson in Isabelle/HOL,
-which did not work for the previous.
-
-\item On a technical level we show that the novel embedding of S5 is more
-    effective than the previous one.
-
-\item Much improved syntax presentation.
-
-\item Technical issues detected within Isabelle-Sledgehammer; room for improvements
+\item A detailed analysis of the inconsistency of G\"{o}del's
+  \shortcite{GoedelNotes} original version of the proof is presented
+  for higher-order modal logics KB and K (with possibilist and
+  actualist first-order quantifiers). The extraction, reconstruction
+  and verification of an informal, human intuitive argument has been
+  an open problem since the first detection of this inconsistency by
+  Benzmüller and Woltzenlogel-Paleo \shortcite{C40}. The experiments
+  reported here led to the discovery of a surprisingly accessible
+  understanding of this inconsistency which is philosophically
+  profound and which has not been presented yet in the respective
+  literature. The detection of this inconsistency in combination with
+  the work reported here thus demonstrates that higher-order automated
+  theorem provers (and thus AI systems) are capable of contributing
+  new and relevant knowledge in philosophy. 
+\item Another open question has been whether the G\"odel's
+  respectively Scott's main theorem T3 \textit{``Necessarily, there
+    exists God''} can eventually be proved fully automatically
+  directly from the axioms, that is, without the need for intermediate
+  argumentation steps (lemmata). Benzm\"uller and Woltzenlogel-Paleo
+  \shortcite{C40} still failed in achieving this. We here report
+  success for Scott's (consistent) variant of the proof in
+  higher-order modal logic S5. The fully automated proof has been
+  contributed by the theorem prover \textsc{Leo-II}~\cite{C26} and
+  subsequently verified in the proof assistant
+  Isabelle/HOL~\cite{NPW02}. On a technical level we show that the
+  novel embedding of S5 is more effective than the previous one.
+\item A third open issue has been whether the modal
+  collapse \cite{Sobel}, which is one of the most strongly criticised
+  `side-effects' of G\"odel's and Scott's variants of the proof, can
+  eventually be directly verified in Isabelle/HOL. Indeed, in previous work
+  the modal collapse has been confirmed by the provers
+  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a
+  direct verification in the highly trusted Isabelle/HOL still failed.
+\item On the technical side, we present a significantly improved
+  syntax representation of G\"odel's resp.~Scott's proof in
+  Isabelle/HOL. In fact, we demonstrate that a nearly perfect match
+  between the original pen and paper presentations and our encoding in
+  Isabelle/HOl is feasible. This is clearly an important prerequisite
+  for promoting the here employed theorem proving technology to
+  non-expert philosophers.
+\item On the other hand, we will
+  also point to several technical challenges which, at least for the
+  moment, still restrain non-expert users of these systems from
+  independently conducting similar experiments in computational
+  metaphysics as reported here and in previous work. 
+% Technical issues detected within Isabelle-Sledgehammer; room for improvements
 \end{itemize}
 
-These novel contributions are presented in Sections ? and ?.
+% These novel contributions are presented in Sections ? and ?.
+% In Section ?, novel and previous findings are summarised and discussed 
+% with regard to their relevance in metaphysics. In this section also 
+% some further details and improvements on earlier 
+% findings are presented.
+% Section ? concludes the paper.
 
-In Section ?, novel and previous findings are summarised and discussed 
-with regard to their relevance in metaphysics. In this section also 
-some further details and improvements on earlier 
-findings are presented.
+Give paper structure (maybe not needed) ...
 
-Section ? concludes the paper.
+First successful applications of theorem proving technology in
+metaphysics were reported by Oppenheimer and
+Zalta~\shortcite{oppenheimera11}. They employed the first-order
+\textsc{Prover9} \cite{Prover9} in their experiments. It was also
+Zalta and Oppenheimer who first proposed the term
+\textit{Computational Metaphysics} for this new research area. Further
+related work includes Rushby's~\shortcite{rushby13} work with the
+higher-order proof assistant \textsc{PVS} \cite{PVS}. Common to both
+works is a significant amount of proof-hand-coding work as well as their
+focus on St. Anselm's~\shortcite{Anselm} simpler, non-modal historical
+version of the ontological argument.
+
+\ldots. Different to what is presented here, none of these previous
+formalises and automates variants of HOML, which is however crucial
+for G\"{o}del's ontological argument.
 
@@
 \begin{figure}
-%\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
 \caption{Reconstruction of inconsistency in Isabelle/HOl..} \label{InconsistencyIsabelleK}
@@
 \bibliographystyle{named}
-\bibliography{ijcai15}
+\bibliography{Bibliography}
 

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Write abstract

@@
-%%%% ijcai15.tex
-
-\typeout{Automating the Ontological Argument}
-
-% These are the instructions for authors for IJCAI-15.
-% They are the same as the ones for IJCAI-11 with superficical wording
-%   changes only.
+\typeout{Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
 
 \documentclass{article}
-% The file ijcai15.sty is the style file for IJCAI-15 (same as ijcai07.sty).
 \usepackage{ijcai15}
@@
 
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
 \usepackage{graphicx,url}
 
-\title{Automating the Ontological Argument}
-\author{Christoph Benzmüller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
+\title{Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
+\author{Christoph Benzm\"uller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
 \author{}
@@
 \begin{abstract}
-  Higher-order automated theorem provers have revealed some
-  philosophically profound new knowledge in metaphysics.
+  This paper discusses the discovery of the inconsistency in G\"odel's ontological argument as a success story for artificial intelligence.
+Despite the popularity of the argument since the appearance of G\"odel's manuscript in 1970, the inconsistency of the axioms used in the argument remained unnoticed until 2013, when it was detected automatically by the higher-order theorem prover LEO-II.
+Understanding and verifying the refutation generated by the prover turned out to be a time-consuming task, and its completion, as reported here, required the reconstruction of the refutation in the Isabelle proof assistant.
+The development of an improved syntactical hiding for the embedding technique, utilizing Isabelle's binding notation mechanism, allows the refutation to be presented in a human-friendly way, suitable for non-experts in the technicalities of the higher-order theorem proving.
+This brings us a step closer to wider adoption of logic-based artificial intelligence tools by philosophers.
 \end{abstract}
@@
 \section{Introduction}
-Without exaggeration one may claim that Kurt G\"{o}del's ontological
+Without exaggeration Kurt G\"{o}del's ontological
 argument for the existence of God \cite{GoedelNotes,ScottNotes} is

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Improve introduction

@@
 argument is regularly presented as a masterpiece argument in
-metaphysics. Since summer 2013, when Benzm\"uller and Woltzenlogel-Paleo~\shortcite{J30,C40} first
-reported on their successful initial computational
-analysis of the G\"odel's respectively Scott's variants of the proof,
-their contribution received a media repercussion on a global scale\footnote{A
-  small collection of these reports is available at {\scriptsize
+metaphysics. Since 2013, when Benzm\"uller and Woltzenlogel-Paleo~\shortcite{J30,C40} first
+reported their successful initial computer-assisted
+analysis of G\"odel's proof and Scott's variant,
+their work has received a media repercussion on a global scale\footnote{A
+  small collection of news articles is available at {\scriptsize
     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
@@
 The in-depth computational analysis presented here substantially
-extends previous computer-assisted studies on G\"odel's ontological
+extends previous computer-assisted studies of G\"odel's ontological
 argument. Similar to the related work \cite{J30,C40} the analysis has
@@
 higher-order logic (HOL) even though G\"odel's proof is actually
-formulated in modal higher-order logic. To bridge between the two
-logic formalisms we utilise and further improve the logic embeddings
-approach \cite{J23,C40} as also employed in preceding related work.
+formulated in \emph{modal} higher-order logic. To bridge between the two
+logics we utilise and further improve the logic embedding
+approach \cite{J23,C40}, which has already been employed succesfully in preceding related work.
 
-The novel contributions of this paper include (but are not limited to):
+The novel contributions reported in this paper include the following:
 
@@
   an open problem since the first detection of this inconsistency by
-  Benzmüller and Woltzenlogel-Paleo \shortcite{C40}. The experiments
-  reported here led to the discovery of a surprisingly accessible
+  Benzmüller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
+  The experiments reported here led to the discovery of a surprisingly accessible
   understanding of this inconsistency which is philosophically
-  profound and which has not been presented yet in the respective
+  profound and which has not been presented yet in the
   literature. The detection of this inconsistency in combination with
-  the work reported here thus demonstrates that higher-order automated
-  theorem provers (and thus AI systems) are capable of contributing
-  new and relevant knowledge in philosophy. 
-\item Another open question has been whether the G\"odel's
-  respectively Scott's main theorem T3 \textit{``Necessarily, there
-    exists God''} can eventually be proved fully automatically
-  directly from the axioms, that is, without the need for intermediate
-  argumentation steps (lemmata). Benzm\"uller and Woltzenlogel-Paleo
-  \shortcite{C40} still failed in achieving this. We here report
-  success for Scott's (consistent) variant of the proof in
-  higher-order modal logic S5. The fully automated proof has been
-  contributed by the theorem prover \textsc{Leo-II}~\cite{C26} and
+  the work reported here thus demonstrates that artificial intelligence systems -- 
+  particularly higher-order automated
+  theorem provers -- are capable of assisting in the discovery and ellucidation of
+  \emph{new} and philosophically relevant knowledge. 
+\item Another open question has been whether the final theorem T3 \textit{``Necessarily, there
+    exists God''} could eventually be proved fully automatically
+  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
+  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
+  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
+  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
   subsequently verified in the proof assistant
-  Isabelle/HOL~\cite{NPW02}. On a technical level we show that the
-  novel embedding of S5 is more effective than the previous one.
-\item A third open issue has been whether the modal
+  Isabelle/HOL~\cite{NPW02}.
+\item A third open challenge the verification in Isabelle/HOL of the modal
   collapse \cite{Sobel}, which is one of the most strongly criticised
-  `side-effects' of G\"odel's and Scott's variants of the proof, can
-  eventually be directly verified in Isabelle/HOL. Indeed, in previous work
-  the modal collapse has been confirmed by the provers
-  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a
-  direct verification in the highly trusted Isabelle/HOL still failed.
+  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
+  the modal collapse has been derived by the provers
+  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
+  verification in the highly trusted Isabelle/HOL still failed. 
+  Now, with the more efficient embedding for S5, this verification has been done. 
 \item On the technical side, we present a significantly improved
@@
   Isabelle/HOl is feasible. This is clearly an important prerequisite
-  for promoting the here employed theorem proving technology to
-  non-expert philosophers.
-\item On the other hand, we will
+  for promoting the theorem proving technology employed here to a wide community of
+  philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
+\item On the other hand, we
   also point to several technical challenges which, at least for the
@@
 metaphysics were reported by Oppenheimer and
-Zalta~\shortcite{oppenheimera11}. They employed the first-order
-\textsc{Prover9} \cite{Prover9} in their experiments. It was also
-Zalta and Oppenheimer who first proposed the term
-\textit{Computational Metaphysics} for this new research area. Further
-related work includes Rushby's~\shortcite{rushby13} work with the
-higher-order proof assistant \textsc{PVS} \cite{PVS}. Common to both
+Zalta~\shortcite{oppenheimera11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
+\textsc{Prover9} \cite{Prover9} in their experiments. Later on, Rushby~\shortcite{rushby13} used the proof assistant \textsc{PVS} \cite{PVS}. Common to both
 works is a significant amount of proof-hand-coding work as well as their
-focus on St. Anselm's~\shortcite{Anselm} simpler, non-modal historical
-version of the ontological argument.
-
-\ldots. Different to what is presented here, none of these previous
-formalises and automates variants of HOML, which is however crucial
-for G\"{o}del's ontological argument.
+focus on a non-modal formalization of St. Anselm's~\shortcite{Anselm} simpler 
+and older ontological argument. 
+None of these previous works formalises and automates variants of \emph{higher-order} and \emph{modal} logics, which are, however, crucial
+for G\"{o}del's more complex ontological argument.
+x
 

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Add packages and commands
Minor changes

@@
 % the following package is optional:
-%\usepackage{latexsym}
+\usepackage{latexsym}
 
+\usepackage{amsmath}
+\usepackage{modallogics}
 \usepackage{graphicx,url}
+
+\newcommand{\imp}{\rightarrow}
+\newcommand{\biimp}{\leftrightarrow}
+\newcommand{\allq}{\forall}
+\newcommand{\exq}{\exists}
+\newcommand{\seq}{\vdash}
+\newcommand{\nec}{\Box} % necessarily
+\newcommand{\pos}{\Diamond} % possibly
 
@@
   an open problem since the first detection of this inconsistency by
-  Benzmüller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
+  Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
   The experiments reported here led to the discovery of a surprisingly accessible
@@
 \end{itemize}
-
-% These novel contributions are presented in Sections ? and ?.
-% In Section ?, novel and previous findings are summarised and discussed 
-% with regard to their relevance in metaphysics. In this section also 
-% some further details and improvements on earlier 
-% findings are presented.
-% Section ? concludes the paper.
-
-Give paper structure (maybe not needed) ...
 

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Write section about the history of the ontological argument

@@
 for G\"{o}del's more complex ontological argument.
-x
 
-\section{Variants of the Ontological Argument}
-Short history, Anselm, Leibniz, etc.
-\subsection{G\"{o}del's Proof Script}
-\subsection{Scott's Variant}
-\subsection{Recent Variants}
-Dozens of contributions were published in the last four decades which
-present variants of G\"odel's/Scott's proofs. In many cases the
-motivation has been to remedy the modal collapse. Other authors are
-simply interested in a critical assessment. Particularly interesting
-from the perspective of this paper is that many literature
-contributions unfortunately work with or refer to G''odel's original
-definition of essence
-$$ todo $$
-This original version avoids the conjunct added by Scott expressing
-that essential properties of an individual should be possessed by the
-individual. As we show in this paper using this notion of essence
-leads to an inconsistent axiom system, which in a classical setting of
-course deserves no further attention, since from an inconsistent
-system everything can be concluded, including God's existence and
-non-existence simultaneously. In other words, these literature
-contributions discuss variants of the argument which from a classical
-logical perspective do not deserve any further attention.
+\section{A Brief History of the Argument}
 
-We give here a list of examples:
-\begin{itemize}
-\item Anderson and Getting's 1996, p. 168 \cite[p.168]{AndersonGettings1968}
-  use essence without conjunct.
-\item  Hazen 1998, p.365 \cite[p.365]{Hazen1998} states: ``G\"odel left this
-  clause out in [11], but this appears to have been an oversight--it
-  is included in related manuscripts''. There is not mentioning of an
-  inconsistency though.
-\item Look???, p. 514
-% \item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
-%   \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
-%   Oppy uses: ``A is an essence of x iff for every property B, x has B
-%   neces- sarily iff A entails B'' (this is from Anderson's
-%   emendation). Check is this leads to inconsistency as well. In this
-%   case we may write something like:
+St.
+Anselm's ontological argument \cite{Proslogion} can be regarded as the ancestor of modern ontological arguments such as G\"odel's.
+In the millenium between Anselm and G\"odel, many philosophers modified and arguably improved Anselm's argument.
+Of particular importance to G\"odel was the work of Leibniz \cite{ToDo}.
+G\"odel's manuscript (Figure \ref{GoedelScript}) can be considered a translation of Leibniz's presentation of the argument into modern modal logic.
+G\"odel shared his manuscript with Scott, who shared a slightly different version with a larger public.
+Scott's version of the axioms and definitions, formalized in Isabelle, is shown in Figure \ref{ScottS5}.
+The main difference to G\"odel's version is the addition of a conjunct in the definition of \emph{essence}.
+G\"odel's different definition of essence can be seen either in his manuscript (Figure \ref{GoedelScript}) or, in more modern notation, in the Isabelle formalization shown in Figure \ref{nconsistencyIsabelleK}.
+For Scott, an essential property of an individual must be possessed by him/her.
+For G\"odel, this is not required.
+This difference has been considered inessential and merely an oversight by many.
+For instance, Hazen 1998, p.365 \cite[p.365]{Hazen1998} states that ``G\"odel left this clause out in [11], but this appears to have been an oversight--it is included in related manuscripts''.
+However, the omitted conjunct is in fact crucial.
+Without it, G\"odel's original axioms are inconsistent.
+With it, Scott's axioms are consistent.
 
-%  Oppy presents an critical
-%   assessment in order to conclude that G\"odel's argument fails to
-%   convince. Oppy apparently does not see that he is assessing an
-%   inconsistent axiom system (and he in fact creates adaptations that
-%   suffer the same problems).
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
+\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
+\end{figure*}
+
+% There is reason to believe that the omission was more than just an oversight.
+As pointed out by Fuhrmann \cite{Fuhrmann2005}, ``G\"odel vermerkt diese Konsequenz [dass aus der Definition unmittelbar folgt, da\ss alle wesentlichen Eigenschaften notwendig äquivalent sind] in einer Fußnote zur Definition.
+Die Definition selbst aber l\"a{\ss}t im Definiens das Konjunkt $Xx$ aus.
+Ohne das Konjunkt folgt jedoch die Konsequenz nicht.
+Es ist deshalb naheliegend anzunehmen, da\ss G\"odel die Definition so beabsichtigte, wie sie hier notiert ist und wie G\"odel sie selbst in fr\"uheren Notizbucheintragungen formuliert hat.''\footnote{Our translation: ``G\"odel remarks this consequence [that the definition of essence entails that all essences are necessarily equivalent] in a footnote to the definition.
+The definition, however, omits the conjunct.
+...
+ToDo''}
+% \marginpar{ToDo: check Fuhrmann's claim. He may be pointing out a second mistake by Gödel.}
+
+For more than four decades, the serious consequences of G\"odel's mistake remained unnoticed, despite intense activity focused on criticizing his argument.
+Especially since the discovery by Sobel \cite{Sobel} that a modal collapse is entailed by G\"odel's (or also Scott's) axioms, several variants have been proposed \cite{Anderson,AndersonGettings1968,Hajek,Hajek2,Hajek3,Bjordal} attempting to avoid the modal collapse.
+Many of these variants omit the crucial conjunct in the definition of essence as well\footnote{As these variants also change other axioms, on which the inconsistency of G\"odel's axioms depends, it is not necessarily the case that these variants are also inconsistent.}.
+Opponents of the argument (e.g.
+Oppy \cite[p.226/227]{Oppy1996} \cite[p.364]{Oppy2000} \cite[p.1068]{Oppy2008}) have also proposed parodies and other criticisms, referring to variants where the conjunct is omitted.
+
+%  Particularly interesting
+% from the perspective of this paper is that many literature
+% contributions unfortunately work with or refer to G\"odel's original
+% definition of essence
+% $$ todo $$
+% This original version avoids the conjunct added by Scott expressing
+% that essential properties of an individual should be possessed by the
+% individual.
+
+% We give here a list of examples:
+% \begin{itemize}
+% \item Anderson and Getting's 1996, p. 168 \cite[p.168]{AndersonGettings1968}
+%   use essence without conjunct.
 % \item  
-\end{itemize}
+% \item Look???, p. 514
+% % \item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
+% %   \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
+% %   Oppy uses: ``A is an essence of x iff for every property B, x has B
+% %   neces- sarily iff A entails B'' (this is from Anderson's
+% %   emendation). Check is this leads to inconsistency as well. In this
+% %   case we may write something like:
 
-\footnote{Fuhrmann2005: Gödel vermerkt diese Konsequenz [dass aus der Definition unmittelbar folgt, daß alle wesentlichen
-Eigenschaften notwendig äquivalent sind] in einer Fußnote zur Definition.
-Die Definition selbst aber läßt im Definiens das Konjunkt Xx aus.
-Ohne das Konjunkt folgt jedoch die Konsequenz nicht.
-Es ist deshalb naheliegend anzunehmen, daß Gödel die Definition so beabsichtigte, wie sie hier notiert ist und wie Gödel sie selbst in früheren Notizbucheintragungen formuliert hat.}
+% %  Oppy presents an critical
+% %   assessment in order to conclude that G\"odel's argument fails to
+% %   convince. Oppy apparently does not see that he is assessing an
+% %   inconsistent axiom system (and he in fact creates adaptations that
+% %   suffer the same problems).
+% % \item 
+% \end{itemize}
 
 \section{Automating HOML in HOL}
-\subsection{Outline of the Embedding}
-Logic textbooks commonly utilise classical higher-order logic (HOL)
+
+Logic textbooks \cite{ToDo: which} commonly utilize classical higher-order logic (HOL)
 \cite{Church} as a meta-language to introduce the syntax and the
-semantics of object logics of interest, in which then reasoning
+semantics of object logics of interest, in which reasoning
 problems in concrete application domains can be modeled and solved
-with pen and paper.  In fact, this approach can also be followed to
-enable interactive and automated proof for even very challenging
-objects logics with existing theorem provers for classical
+with pen and paper. In fact, this approach can also be followed for even very challenging
+object logics to
+enable interactive and automated theorem proving with existing theorem provers for classical
 higher-order logic.

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Write section about embeddings and embedding of S5+

@@
 For a computational analysis of G\"odel's ontological argument in this
-approach the embedding of modal higher-order logics (MHOL) such as K,
-KB and S5 for different domain conditions (possibilist and actualist)
+approach the embedding of higher-order modal logics (HOML) such as K,
+KB and S5 with various domain conditions (possibilist and actualist)
 is required. This idea has been successfully followed in related work
-\cite{C40}. The embedding of modal higher-order logic is in fact
-straightforward. Formulas in MHOL such as are lifted to predicates
-over worlds, which are explicitly represented in the approach as
-terms. Propositional MHOL such as $\neg \varphi$, $\varphi\vee\psi$
-and $\box \varphi$ are then mapped to HOL formulas \ldots.
+\cite{C40}. The embedding of higher-order modal logic is in fact
+straightforward. Formulas in HMOL are \emph{lifted}, translated to predicates
+over worlds, which are themselves explicitly represented as
+terms. The logical constants of HMOL are translated to HOL terms in such a way that, for instance, 
+%$\neg \varphi$, $\varphi\vee\psi$
+$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the HOL formulas $\forall w.
+(r w_0 w) \imp (\varphi w)$ and $\exists w.
+(r w_0 w) \wedge (\varphi w)$.
+This form of embedding is precisely the well-known standard translation \cite{Ohlbach,ModalLogicPatrickBlackburn},
+which is here (intra-logically) realized in HOL by stating a set of
+equations defining the logical constants (Fig.
+~\ref{QMLS5}).
+The resulting object logic is the HOML K with rigid terms and constant domains (possibilist quantifiers).
+Other logics (e.g.
+KB, S5) can be embedded by adding axioms that restrict the accessibility relation $r$.
+Varying domains and actualist quantifiers can be simulated by means of an existence predicate, which can be used to guard the quantifiers.
+Therefore, the embedding approach is very flexible.
 
-It is easy
-to see that this captures the standard translation \cite{Ohlbach},
-which is here (intra-logically) realised in HOL by stating a set of
-equations; see also Fig. ~\ref{QMLS5}. New is that also quantifiers 
-can be mapped \ldots
+\subsection{Improved Embedding for S5+}
 
-The approach is very flexible, KB, S5, domain conditions\ldots
+The modal logic S5 requires that the accessibility relation be reflexive, symmetric and transitive.
+The usual approach to embed S5 would be to use the standard translation for K described above and state these additional conditions as three HOL axioms:
+\begin{itemize}
+\item Reflexivity: $\forall x. (r~x~x)$
+\item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
+\item Transitivity: $\forall x. \forall y, \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
+\end{itemize}
 
-\subsection{Improved Embedding for S5}
-We present an improved embedding (syntax and defintion of box) in  Fig.~\ref{QMLS5}).
+We consider here a modal logic that we call S5+, which is characterized by the following condition on the accessibility relation:
+\begin{itemize}
+\item Omni-accessibility: $\forall x. \forall y. (r~x~y)$
+\end{itemize}
+
+It is easy to see that S5+ is at least as strong as S5: omni-accessibility entails reflexivity, symmetry and transitivity.
+S5+, in fact, \emph{strictly stronger} than S5.
+S5+ only admits complete\footnote{\emph{Complete} in the graph-theoretical sense: for any directed pair of vertices/worlds there is a directed edge connecting them.}} frames, whereas S5 admits non-complete frames as long as all their components are complete.
+
+For S5+, an improved embedding is possible.
+Omni-accessibility implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold.
+Therefore, they can be omitted, and the modal operators can be defined instead merely as:
+\begin{itemize}
+\item $\Box \varphi \equiv \forall w. (\varphi w)$ 
+\item $\Diamond \varphi \equiv \exists w. (\varphi w)$
+\end{itemize}
+
+This improved embedding for S5+ is shown in Fig.~\ref{QMLS5}).
 \begin{figure}
@@
 \end{figure}
+
+\section{Adequacy of S5+ for Metaphysics}
+
+S5+ is only slightly stronger than S5.  ToDo
 
@@
 \end{figure}
-
-\subsection{Mapping to Gödel's Script}
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
-\caption{Sources of Inconsistence in G\"{o}del's proof script.} \label{GoedelScript}
-\end{figure*}
 

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Move paragraph to conclusion

@@
 
+As we show in this paper using this notion of essence
+leads to an inconsistent axiom system, which in a classical setting of
+course deserves no further attention, since from an inconsistent
+system everything can be concluded, including God's existence and
+non-existence simultaneously. In other words, these literature
+contributions discuss variants of the argument which from a classical
+logical perspective do not deserve any further attention.
+
 \section*{Acknowledgments}

### 2016-01-07 Papers/InProgress/IJCAI/paper.tex
Remove useless instructions

@@
 
-\section{\LaTeX{} and Word Style Files}\label{stylefiles}
-
-The \LaTeX{} and Word style files are available on the IJCAI--15
-website, {\tt http://www.ijcai-15.org/}.
-These style files implement the formatting instructions in this
-document.
-
-The \LaTeX{} files are {\tt ijcai15.sty} and {\tt ijcai15.tex}, and
-the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai15.bib}. The
-\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
-style file is for version 0.99c of Bib\TeX{} ({\em not} version
-0.98i). The {\tt ijcai15.sty} file is the same as the {\tt
-ijcai07.sty} file used for IJCAI--07.
-
-The Microsoft Word style file consists of a single file, {\tt
-ijcai15.doc}. This template is the same as the one used for
-IJCAI--07.
-
-These Microsoft Word and \LaTeX{} files contain the source of the
-present document and may serve as a formatting sample.
-
-Further information on using these styles for the preparation of
-papers for IJCAI--15 can be obtained by contacting {\tt
-pcchair15@ijcai.org}.
-
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2016-01-08 Papers/InProgress/IJCAI/paper.tex
Improve section about embedding of S5

@@
 \usepackage{amsmath}
+\usepackage{xspace}
 \usepackage{modallogics}
 \usepackage{graphicx,url}
+\usepackage{txfonts} % needed for \Diamondblack
+\usepackage{float}
 
@@
 \newcommand{\seq}{\vdash}
-\newcommand{\nec}{\Box} % necessarily
-\newcommand{\pos}{\Diamond} % possibly
+
+\newcommand{\Dia}{\Diamond} % possibly
+\newcommand{\BlackBox}{\blacksquare}
+\newcommand{\BlackDia}{\Diamondblack}
 
@@
 
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
-\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
-\end{figure*}
-
 % There is reason to believe that the omission was more than just an oversight.
@@
 
-\subsection{Improved Embedding for S5+}
+\subsection{Improved Embedding for \SFiveU}
 
-The modal logic S5 requires that the accessibility relation be reflexive, symmetric and transitive.
-The usual approach to embed S5 would be to use the standard translation for K described above and state these additional conditions as three HOL axioms:
+The modal logic \SFive requires that the accessibility relation be reflexive, symmetric and transitive.
+The usual approach to embed \SFive would be to use the standard translation for K described above and state these additional conditions as three HOL axioms:
 \begin{itemize}
@@
 
-We consider here a modal logic that we call S5+, which is characterized by the following condition on the accessibility relation:
+We consider here a modal logic that we call \SFiveU, which is characterized by the following condition on the accessibility relation:
 \begin{itemize}
-\item Omni-accessibility: $\forall x. \forall y. (r~x~y)$
+\item Universality: $\forall x. \forall y. (r~x~y)$
 \end{itemize}
 
-It is easy to see that S5+ is at least as strong as S5: omni-accessibility entails reflexivity, symmetry and transitivity.
-S5+, in fact, \emph{strictly stronger} than S5.
-S5+ only admits complete\footnote{\emph{Complete} in the graph-theoretical sense: for any directed pair of vertices/worlds there is a directed edge connecting them.}} frames, whereas S5 admits non-complete frames as long as all their components are complete.
+It is easy to see that \SFiveU is at least as strong as \SFive: universality entails reflexivity, symmetry and transitivity.
+\SFiveU is, in fact, \emph{strictly stronger} than \SFive.
+\SFiveU only admits complete\footnote{A graph is \emph{complete} iff there is a directed edge connecting every ordered pair of vertices.} frames, whereas \SFive admits non-complete frames as long as all their components are complete.
 
-For S5+, an improved embedding is possible.
-Omni-accessibility implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold.
-Therefore, they can be omitted, and the modal operators can be defined instead merely as:
+For \SFiveU, an improved embedding is possible.
+Universality implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold.
+Therefore, they can be omitted and the accessibility relation can be dispensed altogether.
+The modal operators can then be defined merely as:
 \begin{itemize}
-\item $\Box \varphi \equiv \forall w. (\varphi w)$ 
-\item $\Diamond \varphi \equiv \exists w. (\varphi w)$
+\item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
+\item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
 \end{itemize}
 
-This improved embedding for S5+ is shown in Fig.~\ref{QMLS5}).
+The improved embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QMLS5}.
 \begin{figure}
 \centerline{\includegraphics[width=\columnwidth]{./Images/QMLS5.png}}
-\caption{Improved Embedding of S5} \label{QMLS5}
+\caption{Improved Embedding of \SFiveU} \label{QMLS5}
 \end{figure}
 
-\section{Adequacy of S5+ for Metaphysics}
+\SFiveU is only slightly stronger than \SFive.
+Most importantly, $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi^{U}$, where $\varphi^U$ is obtained from $\varphi$ by replacing all $\Box$ and $\Dia$ by, respectively, $\BlackBox$ and $\BlackDia$.
+Therefore, \SFiveU can be considered as adequate for metaphysics as \SFive (cf.
+\cite{Mattey}, \cite[p.
+127]{WilliamsonModalLogicAsMetaphysics}, \cite[ToDo]{DunnHardegree}).
 
-S5+ is only slightly stronger than S5.  ToDo
-
-\section{Full Automation of Scott's Proof}
-Within the modified logic S5 Scott's proof can be fully automated. The
+With the improved embedding for \SFiveU, the main theorem T3 can be derived from Scott's version of the axioms fully automatically. The
 intermediate argumentation steps as not needed anymore to support the
-prover, see Fig.~\ref{ScottS5}).
+prover.  This is shown in Fig.~\ref{ScottS5}. There it is also shown that the modal collapse can now be verified in
+Isabelle/HOL with Meson. In previous work, with a less efficient embedding for \SFive, this failed.
 \begin{figure}
 \centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
-\caption{Full Automation of the Ontological Argument in S5 } \label{ScottS5}
+\caption{Full Automation of T3 and MC in \SFiveU } \label{ScottS5}
 \end{figure}
+\marginpar{ToDo: this picture has to be replaced by one without Isabelle errors}
 
-In the modified S5 setting also further results can be automatically verified, which 
-failed in previous work: verification of  modal collapse in
-Isabelle/HOL with Meson.
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
+\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
+\end{figure*}
 

### 2016-01-08 Papers/InProgress/IJCAI/paper.tex
Move figure
Improve introduction

@@
 Understanding and verifying the refutation generated by the prover turned out to be a time-consuming task, and its completion, as reported here, required the reconstruction of the refutation in the Isabelle proof assistant.
-The development of an improved syntactical hiding for the embedding technique, utilizing Isabelle's binding notation mechanism, allows the refutation to be presented in a human-friendly way, suitable for non-experts in the technicalities of the higher-order theorem proving.
+The development of an improved syntactical hiding for the embedding technique, utilizing Isabelle's binding notation mechanism, allows the refutation to be presented in a human-friendly way, suitable for non-experts in the technicalities of higher-order theorem proving.
 This brings us a step closer to wider adoption of logic-based artificial intelligence tools by philosophers.
@@
 
-The novel contributions reported in this paper include the following:
+The main novel contribution reported in this paper is a detailed analysis (in various modal logics) 
+of the inconsistency of G\"{o}del's original version of the axioms used in his proof 
+\shortcite{GoedelNotes}. The extraction, reconstruction and verification of an informal, 
+human intuitive argument has been an open problem since the first detection of this inconsistency 
+by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
+The verified refutation (discussed in Section \ref{ToDo}) displays a surprisingly accessible
+explanation of the inconsistency, which is philosophically
+profound and never presented in the
+literature. The detection of this inconsistency in combination with
+the work reported here thus demonstrates that artificial intelligence systems 
+-- particularly higher-order automated theorem provers -- 
+are capable of assisting in the discovery and elucidation of
+\emph{new} and philosophically relevant knowledge.
 
+On the technical side, the quest for constructing a compelling refutation, 
+capable of convincing also human non-experts, led us to improve the 
+syntax of the embedding of modal logics in
+Isabelle/HOL (as discussed in Section  \ref{ToDo}). With the new syntax, a nearly perfect match
+between the original pen and paper presentations and our encoding in
+Isabelle/HOl is feasible. A more user-friendly syntax, as reported here, is clearly an important prerequisite
+for promoting the theorem proving technology employed here to a wider community of
+philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
+
+Another novel contribution reported here (in Section \ref{ToDo}) is the implementation of an alternative embedding for the only slightly stronger modal logic \SFiveU.
+Our experiments have shown that the new embedding is more efficient, as the following two open problems can now be solved:
 \begin{itemize}
-\item A detailed analysis of the inconsistency of G\"{o}del's
-  \shortcite{GoedelNotes} original version of the proof is presented
-  for higher-order modal logics KB and K (with possibilist and
-  actualist first-order quantifiers). The extraction, reconstruction
-  and verification of an informal, human intuitive argument has been
-  an open problem since the first detection of this inconsistency by
-  Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
-  The experiments reported here led to the discovery of a surprisingly accessible
-  understanding of this inconsistency which is philosophically
-  profound and which has not been presented yet in the
-  literature. The detection of this inconsistency in combination with
-  the work reported here thus demonstrates that artificial intelligence systems -- 
-  particularly higher-order automated
-  theorem provers -- are capable of assisting in the discovery and ellucidation of
-  \emph{new} and philosophically relevant knowledge. 
-\item Another open question has been whether the final theorem T3 \textit{``Necessarily, there
-    exists God''} could eventually be proved fully automatically
-  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
-  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
-  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
-  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
-  subsequently verified in the proof assistant
-  Isabelle/HOL~\cite{NPW02}.
-\item A third open challenge the verification in Isabelle/HOL of the modal
-  collapse \cite{Sobel}, which is one of the most strongly criticised
-  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
-  the modal collapse has been derived by the provers
-  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-  verification in the highly trusted Isabelle/HOL still failed. 
-  Now, with the more efficient embedding for S5, this verification has been done. 
-\item On the technical side, we present a significantly improved
-  syntax representation of G\"odel's resp.~Scott's proof in
-  Isabelle/HOL. In fact, we demonstrate that a nearly perfect match
-  between the original pen and paper presentations and our encoding in
-  Isabelle/HOl is feasible. This is clearly an important prerequisite
-  for promoting the theorem proving technology employed here to a wide community of
-  philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
-\item On the other hand, we
-  also point to several technical challenges which, at least for the
-  moment, still restrain non-expert users of these systems from
-  independently conducting similar experiments in computational
-  metaphysics as reported here and in previous work. 
-% Technical issues detected within Isabelle-Sledgehammer; room for improvements
+\item Proving the final theorem T3 \textit{``Necessarily, there
+    exists God''} fully automatically and directly from the axioms alone, without relying on the argument's lemmas.
+\item Verifying automatically the proof of the modal
+  collapse \cite{Sobel}, which is one of the most strongly criticized
+  `side-effects' of the argument's axioms.
 \end{itemize}
+
+\subsection{Related Work}
 
@@
 
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
-\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
-\end{figure*}
+Another open question has been whether the final theorem T3 \textit{``Necessarily, there
+    exists God''} could eventually be proved fully automatically
+  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
+  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
+  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
+  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
+  subsequently verified in the proof assistant
+  Isabelle/HOL~\cite{NPW02}.
+
+A third open challenge the verification in Isabelle/HOL of the modal
+  collapse \cite{Sobel}, which is one of the most strongly criticised
+  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
+  the modal collapse has been derived by the provers
+  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
+  verification in the highly trusted Isabelle/HOL still failed. 
+  Now, with the more efficient embedding for S5, this verification has been done.
+
+\subsection{Improved Syntax in Isabelle}
+
+use of binder notation, priorities
 
@@
 relevant hints to the empty property (cf. Fig.~\ref{LEO-Proof}).
+
+is presented
+  for higher-order modal logics KB and K (with possibilist and
+  actualist first-order quantifiers)
 
@@
 
+We also point to several technical challenges which, at least for the
+moment, still restrain non-expert users of these systems from
+independently conducting similar experiments in computational
+metaphysics as reported here and in previous work.
+% Technical issues detected within Isabelle-Sledgehammer; room for improvements
+
 \section*{Acknowledgments}
@@
 %German Research Foundation DFG, Chad Brown
+
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
+\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
+\end{figure*}
 

### 2016-01-08 Papers/InProgress/IJCAI/paper.tex
Improve paragraphs about better performance with improved embedding

@@
 
-With the improved embedding for \SFiveU, the main theorem T3 can be derived from Scott's version of the axioms fully automatically. The
-intermediate argumentation steps as not needed anymore to support the
-prover.  This is shown in Fig.~\ref{ScottS5}. There it is also shown that the modal collapse can now be verified in
-Isabelle/HOL with Meson. In previous work, with a less efficient embedding for \SFive, this failed.
+With the improved embedding for \SFiveU, the final theorem T3 (\textit{``Necessarily, there
+exists God''}) can be derived from Scott's version of the axioms fully automatically. 
+The fully automatic proof has been generated by the theorem prover \textsc{Leo-II}~\cite{C26} 
+and subsequently verified in the proof assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{ScottS5}. 
+On the other hand, with the embedding used in \shortcite{C40}, 
+the provers had to be given the intermediate theorem T2 and the corollary C in order to manage to prove T3.
+
+Another evidence that the new embedding provides a significant performance boost 
+is the successful automatic verification in Isabelle/HOL (with its automatic tactic Meson) of the modal
+collapse \cite{Sobel}, which is one of the most strongly criticized
+`side-effects' of G\"odel's and Scott's variants of the proof. In previous work \shortcite{C40}
+the modal collapse has been proven by the higher-order provers
+\textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
+verification in the highly trusted Isabelle/HOL still failed \cite{ArchiveFormalProofsGoedelGod}. The success with the new embedding can be seen in Fig.~\ref{ScottS5}.
+
 \begin{figure}

### 2016-01-08 Papers/InProgress/IJCAI/paper.tex
Write section about syntax improvements

@@
 
-Another open question has been whether the final theorem T3 \textit{``Necessarily, there
-    exists God''} could eventually be proved fully automatically
-  directly from the axioms alone. In \shortcite{C40}, the provers had only succeeded in proving 
-  T3 from intermediate argumentation steps (lemmata). Here we report that, by using a more efficient new
-  embedding for the higher-order modal logic S5, T3 can be directly proven from Scott's version of the axioms, which are consistent. The fully automatic proof has been
-  generated by the theorem prover \textsc{Leo-II}~\cite{C26} and
-  subsequently verified in the proof assistant
-  Isabelle/HOL~\cite{NPW02}.
-
-A third open challenge the verification in Isabelle/HOL of the modal
-  collapse \cite{Sobel}, which is one of the most strongly criticised
-  `side-effects' of G\"odel's and Scott's variants of the proof. In previous work
-  the modal collapse has been derived by the provers
-  \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-  verification in the highly trusted Isabelle/HOL still failed. 
-  Now, with the more efficient embedding for S5, this verification has been done.
-
 \subsection{Improved Syntax in Isabelle}
 
-use of binder notation, priorities
+Wider adoption of higher-order theorem proving technology for reasoning about and within embedded object logics, especially among non-expert users, is still hindered by the gap between the syntax used by people, when they write logical formulas with pen and paper, and the syntax used by higher-order logic theorem provers.
+Even when the syntax of the underlying higher-order system is elegant (as is the case in Isabelle/HOL), the embedding of HOML into HOL may easily expose details of HOL that may be uncommon to the user, disturbing his/her experience while using the system.
+To illustrate this point, Fig.~\ref{UglyEssence} shows how the definition of essence looked like in previous work \cite{ArchiveFormalProofsGoedelGod}, where advanced syntax-sugaring features were not used.
+It looks notably higher-order, and its style differs significantly from the common style seen in works on modal logics and the ontological argument.
+The following specific issues can be enumerated:
+\begin{enumerate}
+\item Lambda abstractions (which are a typical higher-order feature) appear explicitly in places where it didn't need to appear in a pure HOML formulation (cf.
+G\"odel's manuscript, Fig.~\ref{GoedelScript}).
+\item Quantifiers appear as higher-order defined constants, and not as binders.
+This forces the user to read (and write) formulas of the form $\forall (\lambda x.
+A(x))$ instead of the more common $\forall x.
+A(x)$.
+\item The lifted modal connectives are represented by prefixing the letter ``m'' (e.g.
+$m\wedge$ and $m\imp$).
+The prefix disturbs the user, as it constantly reminds him/her that there is something unusual about the modal connectives.
+\item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
+Instead of reading $\psi(y)$, as he would expect, he has to read $(\psi y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
+\end{enumerate}
+
+\begin{figure}
+\centerline{\includegraphics[width=\columnwidth]{./Images/UglyEssence.png}}
+\caption{ Definition of Essence using Old Syntax } \label{UglyEssence}
+\end{figure}
+
+In the embedding presented here, in Fig.~\ref{QMLS5}, advanced syntax-sugaring effects provided by Isabelle were used to prevent the issues enumerated above.
+The possibility to define boldface connectives allow us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
+As desired, the definition of essence in Fig.~\ref{QMLS5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
+The embedding technique is now completely transparent to the user.
+
+We hope that the syntax improvements described here will render the computer-assisted analysis of ontological arguments (as well as future works utilizing the embedding of modal logics into HOL) accessible to a wider audience and encourage philosophers to adopt logic-based artificial intelligence tools in order to conduct similar experiments in computational metaphysics.
 

### 2016-01-09 Papers/InProgress/IJCAI/paper.tex
Add footnote about personal communication with Dana Scott.
Write section about the inconsistency

@@
 \usepackage{txfonts} % needed for \Diamondblack
-\usepackage{float}
+\usepackage{color}
+\usepackage{xcolor}
 
@@
 \newcommand{\BlackDia}{\Diamondblack}
+
+\newcommand{\NE}{\mathit{NE}}
+\newcommand{\ess}[2]{#1 \mathit{ess} #2}
+\newcommand{\nec}{\Box}
+\newcommand{\pos}{\Dia}
 
@@
 Without it, G\"odel's original axioms are inconsistent.
-With it, Scott's axioms are consistent.
+With it, Scott's axioms are consistent\footnote{In personal communication, Dana Scott confirmed that he was unaware G\"odel's axioms were inconsistent.}.
 
@@
 \section{Intuitive Inconsistency Argument}
-\subsection{LEO-II's Inaccessible Inconsistency Proof}
-Tell the story (here or earlier). Show parts of LEO proof.  Inaccessible, but relevant knowledge
-is contained.
+
+In the typical workflow of trying to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first.
+For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}).
+Surprisingly, when this routine check was performed on G\"odel's axioms \cite{ECAI}, the LEO-II prover claimed that the axioms were inconsistent.
+Unfortunately, the refutation generated by LEO-II was barely human-readable.
+The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and machine-oriented proof calculus (higher-order resolution \cite{}) and syntax (TPTP THF \cite{}) were used.
+Part of the file is displayed in Fig.~\ref{LEO-Proof}.
+
 \begin{figure*}
 \centerline{\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
-\caption{Primitive substitution in LEO-II generates candidates for the
-empty property.} \label{LEO-Proof}
+\caption{Lines 115--120 of LEO-II's Refutation. Primitive Substitutions with Empty Property are Highlighted.} \label{LEO-Proof}
 \end{figure*}
-LEO-II's resolution proof is human unintuitive. However, it contained
-relevant hints to the empty property (cf. Fig.~\ref{LEO-Proof}).
 
-is presented
-  for higher-order modal logics KB and K (with possibilist and
-  actualist first-order quantifiers)
+Although LEO-II's resolution refutation is not easy to read for humans, it did contain
+relevant hints to the importance of the empty property (cf. Fig.~\ref{LEO-Proof}).
+
+\subsection{Informal Refutation}
+
+Based on the hints found in LEO-II's refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
+
+\begin{enumerate}
+\item From G\"odel's definition of essence 
+(${\ess{\phi}{x} \biimp {\allq \psi} (\psi(x)
+\imp {\nec} \allq y (\phi(y) \imp \psi(y)))}$) it follows that the empty property is an essence of every individual (\textbf{Empty Essence Lemma}): 
+$$\allq x\,(\ess{\emptyset}{x})$$.
+
+\item From theorem T1 (positive Properties are possibly
+  exemplified: ${\allq \phi} [P(\phi) \imp {\pos}  \exq x
+  \phi(x)]$) and axiom A5 (``necessary existence'' is a positive property: $P(\NE)$ ), it follows that $\NE$ is possibly exemplified:
+  $$
+  \pos \exq x [\NE(x)]
+  $$
+
+\item Expanding the definition of ``necessary existence'' ( 
+  ${\NE(x) \equiv \allq
+  \phi [\ess{\phi}{x} \imp \nec
+  \exq y \phi(y)]}$), the following is obtained:
+  $$
+  \pos \exq x [\allq \varphi [ \ess{\varphi}{x} \imp \nec \exq y [\varphi(y)] ] ]
+  $$
+
+\item The sentence above holds for all $\varphi$ and thus, in particular, for the empty property:
+$$
+\pos \exq x [ \ess{\emptyset}{x} \imp \nec \exq y [\emptyset(y)] ]
+$$
+
+\item By the Empty Essence Lemma, the antecedent of the implication above is valid. Therefore, the sentence above entails:
+$$
+\pos \exq x [ \nec \exq y [\emptyset(y)] ]
+$$
+
+\item By definition of $\emptyset$: 
+$$
+\pos \exq x [ \nec \bot ]
+$$
+
+\item As the existential quantifier is binding no variable within its scope, the sentence is equi-valid with:
+$$\pos \nec \bot $$
+
+\item To see that the sentence above is contradictory, we may reason semantically, in terms of possible worlds.
+If $w_0$ is the arbitrary current world, the $\pos$ operator forces the existence of a world $w$ accessible from $w_0$ such that $\nec \bot$ is true in $w$.
+But $\nec \bot$ can only be true in $w$, if there is no world $w'$ accessible from $w$.
+In logics with a reflexive or symmetric accessibility relation (e.g.
+\KB), it is easy to see that there must be a world $w'$ accessible from $w$: either $w'$ itself, in case of a reflexive relation, or $w_0$, in case of a symmetric relation.
+In fact, even in \K, with no accessibility condition, there must be a world $w'$ accessible from $w$.
+The reason is that $\pos \nec \bot$ should be \emph{valid} (true in all worlds).
+Therefore, it is true in $w$ as well, where the existence of an accessible world $w'$ is forced by the $\pos$ operator.
+As a model for $\pos \nec \bot$ (which is a consequence of G\"odel's axioms) cannot be built, G\"odel's axioms are inconsistent.
+\end{enumerate}
+
+Interestingly, the refutation automatically generated by LEO-II uses a symmetric accessibility relation, and thus requires the modal logic KB.
+The informal refutation described above, on the other hand, requires only the weaker modal logic K.
 
 \subsection{Argument Reconstruction in Isabelle}
-Present the argument informal and in Isabelle (see
-Fig.~\ref{InconsistencyIsabelleK}). Easy to understand in 
-in KB and KT, slightly harder in K.
+
+To verify the correctness of the informal argument explained above, it was reconstructed in Isabelle/HOL, using Metis to automate the inessential parts.
+The essential use of the Empty Essence Lemma, on the other hand, is explicitly stated, to ensure that Isabelle is reconstructing the same argument.
+In fact, without the help of this lemma, Metis is not strong enough to refute G\"odel's axioms.
+\marginpar{ToDo: check this claim about Metis.}
+
 \begin{figure}
 \centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
-\caption{Reconstruction of inconsistency in Isabelle/HOl..} \label{InconsistencyIsabelleK}
+\caption{Refutation of G\"odel's Axioms in Isabelle/HOl..} \label{InconsistencyIsabelleK}
 \end{figure}

### 2016-01-11 Papers/InProgress/IJCAI/paper.tex

@@
-\typeout{Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
+\typeout{An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
 
@@
 
-\title{Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
+\title{An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
 \author{Christoph Benzm\"uller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
@@
 syntax of the embedding of modal logics in
-Isabelle/HOL (as discussed in Section  \ref{ToDo}). With the new syntax, a nearly perfect match
+Isabelle/HOL (as discussed in Section  \ref{ToDo}). With the new
+syntax, a (nearly) perfect match
 between the original pen and paper presentations and our encoding in
@@
 
-Another novel contribution reported here (in Section \ref{ToDo}) is the implementation of an alternative embedding for the only slightly stronger modal logic \SFiveU.
-Our experiments have shown that the new embedding is more efficient, as the following two open problems can now be solved:
+Another novel contribution reported here (in Section \ref{ToDo}) is
+the implementation of an alternative embedding for the only slightly
+stronger modal logic \SFiveU. Our experiments have shown that the new
+embedding is more efficient, as the following two, previously open problems can now be solved:
 \begin{itemize}
 \item Proving the final theorem T3 \textit{``Necessarily, there
-    exists God''} fully automatically and directly from the axioms alone, without relying on the argument's lemmas.
-\item Verifying automatically the proof of the modal
+    exists God''} fully automatically and directly from the axioms
+  alone, without relying on the argument's lemmas (i.e.~without
+  relying on intermediate argumentation steps).
+\item Verifying automatically (in Isabelle/HOL) the proof of the modal
   collapse \cite{Sobel}, which is one of the most strongly criticized
@@
 metaphysics were reported by Oppenheimer and
-Zalta~\shortcite{oppenheimera11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
-\textsc{Prover9} \cite{Prover9} in their experiments. Later on, Rushby~\shortcite{rushby13} used the proof assistant \textsc{PVS} \cite{PVS}. Common to both
+Zalta~\shortcite{oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
+\textsc{Prover9} \cite{prover9-mace4} in their experiments. Later on, Rushby~\shortcite{rushby13} used the proof assistant \textsc{PVS} \cite{cade92-pvs}. Common to both
 works is a significant amount of proof-hand-coding work as well as their
-focus on a non-modal formalization of St. Anselm's~\shortcite{Anselm} simpler 
+focus on a non-modal formalization of St. Anselm's~\shortcite{Proslogion} simpler 
 and older ontological argument. 
@@
 G\"odel shared his manuscript with Scott, who shared a slightly different version with a larger public.
-Scott's version of the axioms and definitions, formalized in Isabelle, is shown in Figure \ref{ScottS5}.
+Scott's version of the axioms and definitions, formalized in Isabelle, is shown in Figure \ref{Scott_S5}.
 The main difference to G\"odel's version is the addition of a conjunct in the definition of \emph{essence}.
@@
 
-Logic textbooks \cite{ToDo: which} commonly utilize classical higher-order logic (HOL)
+Logic textbooks \cite{ToDo:which} commonly utilize classical higher-order logic (HOL)
 \cite{Church} as a meta-language to introduce the syntax and the
@@
 equations defining the logical constants (Fig.
-~\ref{QMLS5}).
+~\ref{QML_S5}).
 The resulting object logic is the HOML K with rigid terms and constant domains (possibilist quantifiers).
@@
 
-The improved embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QMLS5}.
+The improved embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QML_S5}.
 \begin{figure}
-\centerline{\includegraphics[width=\columnwidth]{./Images/QMLS5.png}}
-\caption{Improved Embedding of \SFiveU} \label{QMLS5}
+\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5.png}}
+\caption{Improved Embedding of \SFiveU} \label{QML_S5}
 \end{figure}
@@
 The fully automatic proof has been generated by the theorem prover \textsc{Leo-II}~\cite{C26} 
-and subsequently verified in the proof assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{ScottS5}. 
+and subsequently verified in the proof assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5}. 
 On the other hand, with the embedding used in \shortcite{C40}, 
@@
 \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-verification in the highly trusted Isabelle/HOL still failed \cite{ArchiveFormalProofsGoedelGod}. The success with the new embedding can be seen in Fig.~\ref{ScottS5}.
+verification in the highly trusted Isabelle/HOL still failed \cite{ArchiveFormalProofsGoedelGod}. The success with the new embedding can be seen in Fig.~\ref{Scott_S5}.
 
 \begin{figure}
-\centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
-\caption{Full Automation of T3 and MC in \SFiveU } \label{ScottS5}
+\centerline{\includegraphics[width=\columnwidth]{./Images/Scott_S5.png}}
+\caption{Full Automation of T3 in \SFiveU; Verification of MC} \label{Scott_S5}
 \end{figure}
@@
 
-In the embedding presented here, in Fig.~\ref{QMLS5}, advanced syntax-sugaring effects provided by Isabelle were used to prevent the issues enumerated above.
+In the embedding presented here, in Fig.~\ref{QML_S5}, advanced syntax-sugaring effects provided by Isabelle were used to prevent the issues enumerated above.
 The possibility to define boldface connectives allow us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
-As desired, the definition of essence in Fig.~\ref{QMLS5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
+As desired, the definition of essence in Fig.~\ref{QML_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
 The embedding technique is now completely transparent to the user.
@@
 For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}).
-Surprisingly, when this routine check was performed on G\"odel's axioms \cite{ECAI}, the LEO-II prover claimed that the axioms were inconsistent.
+Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the LEO-II prover claimed that the axioms were inconsistent.
 Unfortunately, the refutation generated by LEO-II was barely human-readable.

### 2016-01-11 Papers/InProgress/IJCAI/paper.tex
minor additions

@@
 Logic textbooks \cite{ToDo:which} commonly utilize classical higher-order logic (HOL)
-\cite{Church} as a meta-language to introduce the syntax and the
+\cite{Church40} as a meta-language to introduce the syntax and the
 semantics of object logics of interest, in which reasoning
@@
 %$\neg \varphi$, $\varphi\vee\psi$
-$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the HOL formulas $\forall w.
-(r w_0 w) \imp (\varphi w)$ and $\exists w.
-(r w_0 w) \wedge (\varphi w)$.
-This form of embedding is precisely the well-known standard translation \cite{Ohlbach,ModalLogicPatrickBlackburn},
-which is here (intra-logically) realized in HOL by stating a set of
-equations defining the logical constants (Fig.
-~\ref{QML_S5}).
-The resulting object logic is the HOML K with rigid terms and constant domains (possibilist quantifiers).
-Other logics (e.g.
-KB, S5) can be embedded by adding axioms that restrict the accessibility relation $r$.
-Varying domains and actualist quantifiers can be simulated by means of an existence predicate, which can be used to guard the quantifiers.
-Therefore, the embedding approach is very flexible.
+$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the
+HOL formulas $\forall w. (r w_0 w) \imp (\varphi w)$ and $\exists
+w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is precisely
+the well-known standard translation
+\cite{DBLP:journals/logcom/Ohlbach91}, which is here (intra-logically)
+realized in HOL by stating a set of equations defining the logical
+constants (Fig. ~\ref{QML_S5}). The resulting object logic is the HOML
+K with rigid terms and constant domains (possibilist
+quantifiers). Other logics (e.g. KB, S5) can be embedded by adding
+axioms that restrict the accessibility relation $r$. Varying domains
+and actualist quantifiers can be simulated by means of an existence
+predicate, which can be used to guard the quantifiers. Therefore, the
+embedding approach is very flexible.
 

### 2016-01-12 Papers/InProgress/IJCAI/paper.tex
paper improvements

@@
 \newcommand{\NE}{\mathit{NE}}
-\newcommand{\ess}[2]{#1 \mathit{ess} #2}
+\newcommand{\ess}[2]{#1\ \mathit{ess}\ #2}
 \newcommand{\nec}{\Box}
@@
   This paper discusses the discovery of the inconsistency in G\"odel's ontological argument as a success story for artificial intelligence.
-Despite the popularity of the argument since the appearance of G\"odel's manuscript in 1970, the inconsistency of the axioms used in the argument remained unnoticed until 2013, when it was detected automatically by the higher-order theorem prover LEO-II.
+Despite the popularity of the argument since the appearance of G\"odel's manuscript in 1970, the inconsistency of the axioms used in the argument remained unnoticed until 2013, when it was detected automatically by the higher-order theorem prover \textsc{Leo-II}.
 Understanding and verifying the refutation generated by the prover turned out to be a time-consuming task, and its completion, as reported here, required the reconstruction of the refutation in the Isabelle proof assistant.
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 Without exaggeration Kurt G\"{o}del's ontological
@@
 human intuitive argument has been an open problem since the first detection of this inconsistency 
-by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the LEO-II prover. 
-The verified refutation (discussed in Section \ref{ToDo}) displays a surprisingly accessible
+by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the \textsc{Leo-II} prover. 
+The verified refutation (discussed in \S\ref{sec:inconsistency}) displays a surprisingly accessible
 explanation of the inconsistency, which is philosophically
@@
 syntax of the embedding of modal logics in
-Isabelle/HOL (as discussed in Section  \ref{ToDo}). With the new
+Isabelle/HOL (as discussed in \S\ref{sec:improvedsyntax}). With the new
 syntax, a (nearly) perfect match
@@
 
-Another novel contribution reported here (in Section \ref{ToDo}) is
+Another novel contribution reported here (in \S\ref{sec:improvedembedding}) is
 the implementation of an alternative embedding for the only slightly
@@
     exists God''} fully automatically and directly from the axioms
-  alone, without relying on the argument's lemmas (i.e.~without
-  relying on intermediate argumentation steps).
+  alone, without relying on the argument's intermediate argumentation
+  steps (i.e., lemmata).
 \item Verifying automatically (in Isabelle/HOL) the proof of the modal
@@
 
-\section{A Brief History of the Argument}
+\section{A Brief History of the Argument}\label{sec:history}
 
-St.
-Anselm's ontological argument \cite{Proslogion} can be regarded as the ancestor of modern ontological arguments such as G\"odel's.
-In the millenium between Anselm and G\"odel, many philosophers modified and arguably improved Anselm's argument.
-Of particular importance to G\"odel was the work of Leibniz \cite{ToDo}.
-G\"odel's manuscript (Figure \ref{GoedelScript}) can be considered a translation of Leibniz's presentation of the argument into modern modal logic.
-G\"odel shared his manuscript with Scott, who shared a slightly different version with a larger public.
-Scott's version of the axioms and definitions, formalized in Isabelle, is shown in Figure \ref{Scott_S5}.
-The main difference to G\"odel's version is the addition of a conjunct in the definition of \emph{essence}.
-G\"odel's different definition of essence can be seen either in his manuscript (Figure \ref{GoedelScript}) or, in more modern notation, in the Isabelle formalization shown in Figure \ref{nconsistencyIsabelleK}.
-For Scott, an essential property of an individual must be possessed by him/her.
-For G\"odel, this is not required.
-This difference has been considered inessential and merely an oversight by many.
-For instance, Hazen 1998, p.365 \cite[p.365]{Hazen1998} states that ``G\"odel left this clause out in [11], but this appears to have been an oversight--it is included in related manuscripts''.
-However, the omitted conjunct is in fact crucial.
-Without it, G\"odel's original axioms are inconsistent.
-With it, Scott's axioms are consistent\footnote{In personal communication, Dana Scott confirmed that he was unaware G\"odel's axioms were inconsistent.}.
+St. Anselm's ontological argument \cite{Proslogion} can be regarded as
+the ancestor of modern ontological arguments such as G\"odel's. In the
+millenium between Anselm and G\"odel, many philosophers modified and
+arguably improved Anselm's argument. Of particular importance to
+G\"odel was the work of Leibniz \cite{ToDo:Leibniz}.  G\"odel's manuscript
+(Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's
+presentation of the argument into modern modal logic. G\"odel
+discussed his manuscript with Scott, who shared a slightly different
+version with a larger public. Scott's version of the axioms and
+definitions, formalized in Isabelle, is shown in
+Fig.~\ref{Scott_S5}. The main difference to G\"odel's version is the
+addition of a conjunct in the definition of \emph{essence}. G\"odel's
+different definition of essence can be seen either in his manuscript
+(Fig.~\ref{GoedelScript}) or, in more modern notation, in the Isabelle
+formalization shown in Fig.~\ref{InconsistencyIsabelleK}. For Scott,
+an essential property of an individual must be possessed by
+him/her. For G\"odel, this is not required. This difference has been
+considered inessential and merely an oversight by many. For instance,
+Hazen 1998, p.365 \cite[p.365]{Hazen} states that ``G\"odel left this
+clause out in [11], but this appears to have been an oversight--it is
+included in related manuscripts''. However, the omitted conjunct is in
+fact crucial. Without it, G\"odel's original axioms are
+inconsistent. With it, Scott's axioms are consistent\footnote{In
+  personal communication, Dana Scott confirmed that he was unaware at
+  the time G\"odel's axioms were inconsistent.}.
 
@@
 For more than four decades, the serious consequences of G\"odel's mistake remained unnoticed, despite intense activity focused on criticizing his argument.
-Especially since the discovery by Sobel \cite{Sobel} that a modal collapse is entailed by G\"odel's (or also Scott's) axioms, several variants have been proposed \cite{Anderson,AndersonGettings1968,Hajek,Hajek2,Hajek3,Bjordal} attempting to avoid the modal collapse.
+Especially since the discovery by Sobel \cite{Sobel} that a modal collapse is entailed by G\"odel's (or also Scott's) axioms, several variants have been proposed \cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal} attempting to avoid the modal collapse.
 Many of these variants omit the crucial conjunct in the definition of essence as well\footnote{As these variants also change other axioms, on which the inconsistency of G\"odel's axioms depends, it is not necessarily the case that these variants are also inconsistent.}.
 Opponents of the argument (e.g.
-Oppy \cite[p.226/227]{Oppy1996} \cite[p.364]{Oppy2000} \cite[p.1068]{Oppy2008}) have also proposed parodies and other criticisms, referring to variants where the conjunct is omitted.
+Oppy \cite[p.226/227]{oppy96:_goedel_ontol_argum} \cite[p.364]{oppy00:_respon_gettin} \cite[p.1068]{oppy08:_higher_order_ontol_argum}) have also proposed parodies and other criticisms, referring to variants where the conjunct is omitted.
 
@@
 
-\section{Automating HOML in HOL}
+\section{Automating HOML in HOL}\label{sec:homlinhol}
 
-Logic textbooks \cite{ToDo:which} commonly utilize classical higher-order logic (HOL)
+Logic textbooks % \cite{ToDo:which} 
+commonly utilize classical higher-order logic (HOL)
 \cite{Church40} as a meta-language to introduce the syntax and the
@@
 
-\subsection{Improved Embedding for \SFiveU}
+\subsection{Improved Embedding \SFiveU}\label{sec:improvedembedding}
 
-The modal logic \SFive requires that the accessibility relation be reflexive, symmetric and transitive.
-The usual approach to embed \SFive would be to use the standard translation for K described above and state these additional conditions as three HOL axioms:
+The modal logic \SFive requires that the accessibility relation be
+reflexive, symmetric and transitive. The usual approach to embed
+\SFive would be to use the standard translation for K described above
+and to state that $r$ is an equivalence relation, e.g., by postulating
+the following axioms:
 \begin{itemize}
@@
 \end{itemize}
-
-We consider here a modal logic that we call \SFiveU, which is characterized by the following condition on the accessibility relation:
+Instead, we consider here a modal logic that we call \SFiveU, which is
+characterized by the following condition on $r$:
 \begin{itemize}
@@
 
-It is easy to see that \SFiveU is at least as strong as \SFive: universality entails reflexivity, symmetry and transitivity.
-\SFiveU is, in fact, \emph{strictly stronger} than \SFive.
-\SFiveU only admits complete\footnote{A graph is \emph{complete} iff there is a directed edge connecting every ordered pair of vertices.} frames, whereas \SFive admits non-complete frames as long as all their components are complete.
+It is easy to see that \SFiveU is at least as strong as \SFive:
+universality entails that $r$ is an equivalence relation. \SFiveU is,
+in fact, \emph{strictly stronger} than \SFive. \SFiveU only admits
+complete\footnote{A graph is \emph{complete} iff there is a directed
+  edge connecting every ordered pair of vertices.} frames, whereas
+\SFive admits non-complete frames as long as all their components are
+complete. In other words, in \SFiveU we face one single equivalence
+class of possible worlds, while in \SFive we may deal with several
+disconnected equivalence classes. For this reason, \SFiveU considered
+as metaphysically more appropriate by many philosophers;
+cf. \cite[p.~127]{williamson13}.
 
-For \SFiveU, an improved embedding is possible.
-Universality implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold.
-Therefore, they can be omitted and the accessibility relation can be dispensed altogether.
-The modal operators can then be defined merely as:
-\begin{itemize}
-\item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
-\item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
-\end{itemize}
+For \SFiveU, an improved embedding is possible. Universality implies
+that the guarding predicates in the definitions of $\Box$ and
+$\Diamond$ always hold. Therefore, they can be omitted and the
+accessibility relation can be dispensed altogether. The modal
+operators can then be defined merely as:
+% \begin{itemize}
+% \item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
+% \item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
+% \end{itemize}
+\[\Box \varphi \equiv \forall w. (\varphi w) \qquad \text{and}
+\qquad \Dia \varphi \equiv \exists w. (\varphi w)\]
 
-The improved embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QML_S5}.
+The new embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QML_S5}.
 \begin{figure}
@@
 \end{figure}
-
-\SFiveU is only slightly stronger than \SFive.
-Most importantly, $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi^{U}$, where $\varphi^U$ is obtained from $\varphi$ by replacing all $\Box$ and $\Dia$ by, respectively, $\BlackBox$ and $\BlackDia$.
-Therefore, \SFiveU can be considered as adequate for metaphysics as \SFive (cf.
-\cite{Mattey}, \cite[p.
-127]{WilliamsonModalLogicAsMetaphysics}, \cite[ToDo]{DunnHardegree}).
-
-With the improved embedding for \SFiveU, the final theorem T3 (\textit{``Necessarily, there
-exists God''}) can be derived from Scott's version of the axioms fully automatically. 
-The fully automatic proof has been generated by the theorem prover \textsc{Leo-II}~\cite{C26} 
-and subsequently verified in the proof assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5}. 
-On the other hand, with the embedding used in \shortcite{C40}, 
-the provers had to be given the intermediate theorem T2 and the corollary C in order to manage to prove T3.
+%
+\marginpar{@Bruno: I commented you paragraph here, since I am not
+  convinced (yet) that e.g. BF and CBF are equally valid in \SFiveU
+  and \SFive}
+% \SFiveU is only slightly stronger than \SFive. Most importantly,
+% $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi^{U}$, where
+% $\varphi^U$ is obtained from $\varphi$ by replacing all $\Box$ and
+% $\Dia$ by, respectively, $\BlackBox$ and $\BlackDia$. Therefore,
+% \SFiveU can be considered as adequate for metaphysics as \SFive
+% (cf. \cite[p. 127]{WilliamsonModalLogicAsMetaphysics}). 
+% % \cite{Mattey}, \cite[ToDo]{DunnHardegree}
+%
+With this improved embedding, the final theorem T3
+(\textit{``Necessarily, there exists God''}) can be derived from
+Scott's version of the axioms fully automatically.  The fully
+automatic proof has been generated (in about 2,5 seconds) by the theorem prover
+\textsc{Leo-II}~\cite{C26} and subsequently verified in the proof
+assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5}.
+With the embedding used by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40}, the
+provers still had to be given the intermediate theorem T2 and the corollary
+C in order to manage to prove T3.
 
@@
 \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-verification in the highly trusted Isabelle/HOL still failed \cite{ArchiveFormalProofsGoedelGod}. The success with the new embedding can be seen in Fig.~\ref{Scott_S5}.
+verification in the highly trusted Isabelle/HOL still failed
+\cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5}.
 
@@
 \end{figure}
-\marginpar{ToDo: this picture has to be replaced by one without Isabelle errors}
+%\marginpar{ToDo: this picture has to be replaced by one without Isabelle errors}
 
-\subsection{Improved Syntax in Isabelle}
+\subsection{Improved Syntax in Isabelle}\label{sec:improvedsyntax}
 
@@
 Even when the syntax of the underlying higher-order system is elegant (as is the case in Isabelle/HOL), the embedding of HOML into HOL may easily expose details of HOL that may be uncommon to the user, disturbing his/her experience while using the system.
-To illustrate this point, Fig.~\ref{UglyEssence} shows how the definition of essence looked like in previous work \cite{ArchiveFormalProofsGoedelGod}, where advanced syntax-sugaring features were not used.
+To illustrate this point, Fig.~\ref{UglyEssence} shows how the definition of essence looked like in previous work \cite{J28}, where advanced syntax-sugaring features were not used.
 It looks notably higher-order, and its style differs significantly from the common style seen in works on modal logics and the ontological argument.
@@
 
-In the embedding presented here, in Fig.~\ref{QML_S5}, advanced syntax-sugaring effects provided by Isabelle were used to prevent the issues enumerated above.
-The possibility to define boldface connectives allow us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
-As desired, the definition of essence in Fig.~\ref{QML_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
+In the embedding presented here, in Fig.~\ref{QML_S5}, advanced
+syntax-sugaring effects provided by Isabelle were used to prevent
+issues as enumerated above.
+The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
+As desired, the definition of essence in Fig.~\ref{QML_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{Scott_S5}.
 The embedding technique is now completely transparent to the user.
@@
 
-\section{Intuitive Inconsistency Argument}
+\section{Intuitive Inconsistency Argument} \label{sec:inconsistency}
 
@@
 For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}).
-Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the LEO-II prover claimed that the axioms were inconsistent.
-Unfortunately, the refutation generated by LEO-II was barely human-readable.
-The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and machine-oriented proof calculus (higher-order resolution \cite{}) and syntax (TPTP THF \cite{}) were used.
+Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the \textsc{Leo-II} prover claimed that the axioms were inconsistent.
+Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable.
+The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and machine-oriented proof calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}) were used.
 Part of the file is displayed in Fig.~\ref{LEO-Proof}.
@@
 \centerline{\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
-\caption{Lines 115--120 of LEO-II's Refutation. Primitive Substitutions with Empty Property are Highlighted.} \label{LEO-Proof}
+\caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
+  substitutions (e.g. with the empty property) are highlighted.} \label{LEO-Proof}
 \end{figure*}
 
-Although LEO-II's resolution refutation is not easy to read for humans, it did contain
-relevant hints to the importance of the empty property (cf. Fig.~\ref{LEO-Proof}).
+Although \textsc{Leo-II}'s resolution refutation is not easy to read
+for humans, it did contain relevant hints to the importance of the
+empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof})
+respectively self-difference $\lambda x.  x\not=x$. Philosophers
+might actually prefer using the latter variant over the
+former. However, on the given logic setting, with functional and
+Boolean extensionality, both terms have identical
+denotations. Consequently, they are mutually replacable below and the
+reader may simply use the version he prefers.
 
@@
 
-Based on the hints found in LEO-II's refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
+Based on the hints found in \textsc{Leo-II}'s refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
 
@@
 \imp {\nec} \allq y (\phi(y) \imp \psi(y)))}$) it follows that the empty property is an essence of every individual (\textbf{Empty Essence Lemma}): 
-$$\allq x\,(\ess{\emptyset}{x})$$.
+$$\allq x\,(\ess{\emptyset}{x})$$
 
@@
 
-Interestingly, the refutation automatically generated by LEO-II uses a symmetric accessibility relation, and thus requires the modal logic KB.
+Interestingly, the refutation automatically generated by \textsc{Leo-II} uses a symmetric accessibility relation, and thus requires the modal logic KB.
 The informal refutation described above, on the other hand, requires only the weaker modal logic K.
@@
 
-\section{Conclusion}
+\begin{figure*}
+\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
+\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
+\end{figure*}
+
+\section{Conclusion}\label{sec:conclusion}
 
 Good improvement (technical sense), Sledgehammer can still be improved
-(LEO-II finds inconsistency when call directly in THF Syntax, but not
+(\textsc{Leo-II} finds inconsistency when call directly in THF Syntax, but not
 from within Isabelle). Without independent experiments directly in THF 
@@
 
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
-\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
-\end{figure*}
+\vfill
+ bla
 
-\appendix
+\newpage
+
+*
+
+\newpage
 

### 2016-01-12 Papers/InProgress/IJCAI/paper.tex
paper improvements

@@
 discuss, assess, criticise, modify and improve G\"{o}del's original
-work (see e.g.~Sobel~\shortcite{Sobel} and Oppy~\shortcite{sep-ontological-arguments} and the
+work (see e.g.~Sobel~\shortcite{sobel2004logic} and Oppy~\shortcite{sep-ontological-arguments} and the
 references therein).  In philosophy lectures at universities the
@@
 higher-order logic (HOL) even though G\"odel's proof is actually
-formulated in \emph{modal} higher-order logic. To bridge between the two
+formulated in \emph{modal} higher-order logic (HOML). To bridge between the two
 logics we utilise and further improve the logic embedding
@@
 
-For more than four decades, the serious consequences of G\"odel's mistake remained unnoticed, despite intense activity focused on criticizing his argument.
-Especially since the discovery by Sobel \cite{Sobel} that a modal collapse is entailed by G\"odel's (or also Scott's) axioms, several variants have been proposed \cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal} attempting to avoid the modal collapse.
-Many of these variants omit the crucial conjunct in the definition of essence as well\footnote{As these variants also change other axioms, on which the inconsistency of G\"odel's axioms depends, it is not necessarily the case that these variants are also inconsistent.}.
-Opponents of the argument (e.g.
-Oppy \cite[p.226/227]{oppy96:_goedel_ontol_argum} \cite[p.364]{oppy00:_respon_gettin} \cite[p.1068]{oppy08:_higher_order_ontol_argum}) have also proposed parodies and other criticisms, referring to variants where the conjunct is omitted.
+For more than four decades, the serious consequences of G\"odel's
+mistake remained unnoticed, despite intense activity focused on
+criticizing his argument. Especially since the discovery by Sobel
+\cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
+  $\phi\rightarrow \Box \phi$, states that contingent truth implies
+  necessary truth; it can be interpreted as \textit{everything is
+    pre-determined} or even \textit{there is no free will}. } is
+entailed by G\"odel's (or also Scott's) axioms, several variants have
+been proposed
+\cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
+attempting to avoid the modal collapse. Many of these variants omit
+the crucial conjunct in the definition of essence as well\footnote{As
+  these variants also change other axioms, on which the inconsistency
+  of G\"odel's axioms depends, it is not necessarily the case that
+  these variants are also inconsistent.}. Opponents of the argument
+(e.g. Oppy \cite[p.226/227]{oppy96:_goedel_ontol_argum}
+\cite[p.364]{oppy00:_respon_gettin}
+\cite[p.1068]{oppy08:_higher_order_ontol_argum}) have also proposed
+parodies and other criticisms, referring to variants where the
+conjunct is omitted.
 
@@
 \subsection{Improved Embedding \SFiveU}\label{sec:improvedembedding}
+\begin{figure}[t]
+\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5.png}}
+\caption{Improved Embedding of \SFiveU} \label{QML_S5}
+\end{figure}
 
@@
 class of possible worlds, while in \SFive we may deal with several
-disconnected equivalence classes. For this reason, \SFiveU considered
-as metaphysically more appropriate by many philosophers;
+disconnected equivalence classes. In fact, for this reason, \SFiveU is
+considered as metaphysically more appropriate by some philosophers;
 cf. \cite[p.~127]{williamson13}.
@@
 The new embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QML_S5}.
-\begin{figure}
-\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5.png}}
-\caption{Improved Embedding of \SFiveU} \label{QML_S5}
-\end{figure}
 %
@@
 The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
-As desired, the definition of essence in Fig.~\ref{QML_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{Scott_S5}.
+As desired, the definition of essence in Fig.~\ref{Scott_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
 The embedding technique is now completely transparent to the user.
@@
 
-Interestingly, the refutation automatically generated by \textsc{Leo-II} uses a symmetric accessibility relation, and thus requires the modal logic KB.
-The informal refutation described above, on the other hand, requires only the weaker modal logic K.
+Interestingly, the refutation automatically generated by
+\textsc{Leo-II} uses a symmetric accessibility relation, and thus
+requires the modal logic KB. The informal, human-constructed
+refutation described above, on the other hand, requires only the
+weaker modal logic K. In our experiments \textsc{Leo-II} (like all
+other HOL provers) was still to weak to automatically prove the
+inconsistency already in logic K. Hence, this remains an open problem for automated
+theorem provers.
 
@@
 \marginpar{ToDo: check this claim about Metis.}
-
-\begin{figure}
+\begin{figure}[t]
 \centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
@@
 \end{figure}
+
+\section{Conclusion}\label{sec:conclusion}
+The inconsistency of G\"odel's original axioms in his ontological
+argument has been detected automatically by the prover
+\textsc{Leo-II}. Here we have presented a rational reconstruction and
+verification of the inconsistency proof in Isabelle/HOL. Interestingly,
+the inconsistency argument applies to all normal modal logics
+including K.
 
@@
 
-\section{Conclusion}\label{sec:conclusion}
+Moreover, we have reported on several technical improvements regarding
+the semantic embedding approach. In fact, we have a achieved a nearly
+perfect match between pen and paper presentations in HOML and the
+syntax in Isabelle/HOL. As a result, the embedding of HOML in
+HOL is now fully transparent and, as we think, ready for wider adoption by
+philosophers.
 
-Good improvement (technical sense), Sledgehammer can still be improved
-(\textsc{Leo-II} finds inconsistency when call directly in THF Syntax, but not
-from within Isabelle). Without independent experiments directly in THF 
-the inconsistency would thus not have been detected.
+On the other hand, there is still room for improvements in
+Isabelle/HOL. For example, Isabelle/HOL's Sledgehammer tool, which
+realises calls to external automated theorem provers, such as
+\textsc{Leo-II}, still fails to detect the inconsistency of G\"odel's
+axioms, while a direct modeling of the problem in TPTP THF syntax in
+combination with a direct call of \textsc{Leo-II} succeeds. In other
+words, without independent experiments directly in TPTP THF the
+inconsistency would thus not have been detected. There are further
+technical issues regarding Sledgehammer that motivate
+improvements. For example, in default setting, Sledgehammer does not
+inform the user early that a proof has been found and instead silently
+applies first a series of time-consuming analysis tools (e.g. it's
+minimizer), before it eventually informs the user of success. For the
+involved proof problems presented here (e.g. T3), this phase of
+silence may take several minutes (during which the user might actually
+give up on the proof attempt), while \textsc{Leo-II} already reported
+success to Sledgehammer much earlier (e.g. after 2,5 seconds for T3).
 
-As we show in this paper using this notion of essence
-leads to an inconsistent axiom system, which in a classical setting of
-course deserves no further attention, since from an inconsistent
-system everything can be concluded, including God's existence and
-non-existence simultaneously. In other words, these literature
-contributions discuss variants of the argument which from a classical
-logical perspective do not deserve any further attention.
+Moreover, our work points to a challenge for proof transformation:
+even if \textsc{Leo-II}'s proof could be automatically verified
+in Isabelle, this would clearly not entail that Isabelle/HOL could
+also extract from this verification a human-intuitive argument, for
+example, an argument as has been presented here. For philosophers,
+however, the reconstruction of such human-intuitive arguments from
+machine found proofs will likely be crucial for accepting any novel results.
 
-We also point to several technical challenges which, at least for the
-moment, still restrain non-expert users of these systems from
-independently conducting similar experiments in computational
-metaphysics as reported here and in previous work.
+% Good improvement (technical sense), Sledgehammer can still be improved
+% (\textsc{Leo-II} finds inconsistency when call directly in THF Syntax, but not
+% from within Isabelle).
+
+% We also point to several technical challenges which, at least for the
+% moment, still restrain non-expert users of these systems from
+% independently conducting similar experiments in computational
+% metaphysics as reported here and in previous work.
+
 % Technical issues detected within Isabelle-Sledgehammer; room for improvements
@@
 %German Research Foundation DFG, Chad Brown
-
-\vfill
- bla
 

### 2016-01-12 Papers/InProgress/IJCAI/paper.tex
paper improvements

@@
 \textsc{Leo-II}. Here we have presented a rational reconstruction and
-verification of the inconsistency proof in Isabelle/HOL. Interestingly,
-the inconsistency argument applies to all normal modal logics
-including K.
-
+verification of the inconsistency proof in Isabelle/HOL. This
+inconsistency argument applies to all normal modal logics including K.
 \begin{figure*}
@@
 minimizer), before it eventually informs the user of success. For the
-involved proof problems presented here (e.g. T3), this phase of
-silence may take several minutes (during which the user might actually
-give up on the proof attempt), while \textsc{Leo-II} already reported
-success to Sledgehammer much earlier (e.g. after 2,5 seconds for T3).
+involved proof problems presented here (e.g. G\"odel's theorem T3:
+\textit{Necessarily, there exists God}), this phase of silence may
+take several minutes (during which the user might actually give up on
+the proof attempt), while \textsc{Leo-II} already reported success to
+Sledgehammer much earlier (e.g. after 2,5 seconds for T3).
 
@@
 machine found proofs will likely be crucial for accepting any novel results.
+
+Another open problem that we solved in this paper is a fully automatic
+proof of T3 directly from Scott's axioms. Again, this proof was
+contributed by \textsc{Leo-II}. However, this has been possible only
+after providing a more efficient embedding, based on a universal
+accessibility relation, for HOML S5 in HOL.
+
+Both, the automated detection of the inconsistency in G\"odel's axioms
+and the fully automatic proof of T3 from Scott's axioms demonstrates
+the potential of our AI technology for philosophy: this technology is
+already in its current state able to contribute novel results to
+metaphysics and to conduct reasoning steps at granularity-levels
+beyond common human capabilities.
 
@@
 
-*
-
-\newpage
-
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2016-01-13 Papers/InProgress/IJCAI/paper.tex
paper improvements

@@
 been conducted with automated theorem provers for classical
-higher-order logic (HOL) even though G\"odel's proof is actually
-formulated in \emph{modal} higher-order logic (HOML). To bridge between the two
-logics we utilise and further improve the logic embedding
-approach \cite{J23,C40}, which has already been employed succesfully in preceding related work.
+higher-order logic (HOL; cf. \cite{andrewsSEP} and the references
+therein) even though G\"odel's proof is actually formulated in
+higher-order \emph{modal} logic (HOML; cf. \cite{homl} and the
+references therein). To bridge between the two logics we utilise and
+further improve the logic embedding approach \cite{J23,C40}, which has
+already been employed successfully in preceding related work.
 
-The main novel contribution reported in this paper is a detailed analysis (in various modal logics) 
-of the inconsistency of G\"{o}del's original version of the axioms used in his proof 
-\shortcite{GoedelNotes}. The extraction, reconstruction and verification of an informal, 
-human intuitive argument has been an open problem since the first detection of this inconsistency 
-by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the \textsc{Leo-II} prover. 
-The verified refutation (discussed in \S\ref{sec:inconsistency}) displays a surprisingly accessible
-explanation of the inconsistency, which is philosophically
-profound and never presented in the
-literature. The detection of this inconsistency in combination with
-the work reported here thus demonstrates that artificial intelligence systems 
--- particularly higher-order automated theorem provers -- 
-are capable of assisting in the discovery and elucidation of
-\emph{new} and philosophically relevant knowledge.
+The main novel contribution reported in this paper is a detailed
+analysis (in various modal logics) of the inconsistency of G\"{o}del's
+original version of the axioms used in his proof
+\shortcite{GoedelNotes}. The extraction, reconstruction and
+verification of an informal, human intuitive argument has been an open
+problem since the first detection of this inconsistency by
+Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the
+\textsc{Leo-II} prover.  The verified refutation (discussed in
+\S\ref{sec:inconsistency}) displays a surprisingly accessible
+explanation of the inconsistency, which is philosophically profound
+and never presented in the literature. The detection of this
+inconsistency in combination with the work reported here thus
+demonstrates that artificial intelligence systems -- particularly
+higher-order automated theorem provers -- are capable of assisting in
+the discovery and elucidation of \emph{new} and philosophically
+relevant knowledge.
 
-On the technical side, the quest for constructing a compelling refutation, 
-capable of convincing also human non-experts, led us to improve the 
-syntax of the embedding of modal logics in
-Isabelle/HOL (as discussed in \S\ref{sec:improvedsyntax}). With the new
-syntax, a (nearly) perfect match
-between the original pen and paper presentations and our encoding in
-Isabelle/HOl is feasible. A more user-friendly syntax, as reported here, is clearly an important prerequisite
-for promoting the theorem proving technology employed here to a wider community of
-philosophers, who are not necessarily experts in automated reasoning or higher-order logic.
+On the technical side, the quest for constructing a compelling
+refutation, capable of convincing also human non-experts, led us to
+improve the syntax of the embedding of modal logics in Isabelle/HOL
+(as discussed in \S\ref{sec:improvedsyntax}). With the new syntax, a
+(nearly) perfect match between the original pen and paper
+presentations and our encoding in Isabelle/HOl is feasible. A more
+user-friendly syntax, as reported here, is clearly an important
+prerequisite for promoting the theorem proving technology employed
+here to a wider community of philosophers, who are not necessarily
+experts in automated reasoning or higher-order logic.
 
@@
 \item Proving the final theorem T3 \textit{``Necessarily, there
-    exists God''} fully automatically and directly from the axioms
+    exists God''} (in Scott's \shortcite{ScottNotes} consistent variant) fully automatically and directly from the axioms
   alone, without relying on the argument's intermediate argumentation
@@
 fact crucial. Without it, G\"odel's original axioms are
-inconsistent. With it, Scott's axioms are consistent\footnote{In
-  personal communication, Dana Scott confirmed that he was unaware at
-  the time G\"odel's axioms were inconsistent.}.
+inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5}
+where the model finder Nitpick confirms consistency).\footnote{In
+  personal communication, Dana Scott confirmed that he was unaware
+  that G\"odel's axioms were inconsistent.}
 
@@
   of G\"odel's axioms depends, it is not necessarily the case that
-  these variants are also inconsistent.}. Opponents of the argument
+  these variants are also inconsistent; each case needs to be analysed
+  separately.}. Opponents of the argument
 (e.g. Oppy \cite[p.226/227]{oppy96:_goedel_ontol_argum}
@@
 Logic textbooks % \cite{ToDo:which}
-commonly utilize classical higher-order logic (HOL)
-\cite{Church40} as a meta-language to introduce the syntax and the
-semantics of object logics of interest, in which reasoning
-problems in concrete application domains can be modeled and solved
-with pen and paper. In fact, this approach can also be followed for even very challenging
-object logics to
-enable interactive and automated theorem proving with existing theorem provers for classical
-higher-order logic.
+commonly utilize higher-order logic in an informal/semi-formal way as
+a meta-language to introduce the syntax and the semantics of object
+logics of interest, in which reasoning problems in concrete
+application domains can be modeled and solved with pen and paper. In
+fact, this approach can also be followed on the computer (using HOL as
+a formal meta-language) for even very challenging object logics (such
+as HOML) to enable interactive and automated theorem proving with
+existing theorem provers for HOL.
 
@@
 For a computational analysis of G\"odel's ontological argument in this
-approach the embedding of higher-order modal logics (HOML) such as K,
+approach the embedding of HOMLs such as K,
 KB and S5 with various domain conditions (possibilist and actualist)
 is required. This idea has been successfully followed in related work
-\cite{C40}. The embedding of higher-order modal logic is in fact
-straightforward. Formulas in HMOL are \emph{lifted}, translated to predicates
+\cite{C40}. The embedding of HOML is in fact
+straightforward. Formulas in HOML are \emph{lifted}, translated to predicates
 over worlds, which are themselves explicitly represented as
-terms. The logical constants of HMOL are translated to HOL terms in such a way that, for instance, 
+terms. The logical constants of HOML are translated to HOL terms in such a way that, for instance, 
 %$\neg \varphi$, $\varphi\vee\psi$
-$\Box \varphi$ and $\Diamond \varphi$ are mapped, respectively, to the
-HOL formulas $\forall w. (r w_0 w) \imp (\varphi w)$ and $\exists
-w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is precisely
-the well-known standard translation
-\cite{DBLP:journals/logcom/Ohlbach91}, which is here (intra-logically)
-realized in HOL by stating a set of equations defining the logical
-constants (Fig. ~\ref{QML_S5}). The resulting object logic is the HOML
-K with rigid terms and constant domains (possibilist
-quantifiers). Other logics (e.g. KB, S5) can be embedded by adding
-axioms that restrict the accessibility relation $r$. Varying domains
-and actualist quantifiers can be simulated by means of an existence
-predicate, which can be used to guard the quantifiers. Therefore, the
-embedding approach is very flexible.
+$\Box \varphi$ and $\Diamond \varphi$ (relative to a current world
+$w_o$) are mapped, respectively, to the HOL formulas
+$\forall w. (r w_0 w) \imp (\varphi w)$ and
+$\exists w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is
+precisely the well-known standard translation
+\cite{DBLP:journals/logcom/Ohlbach91}, which is here intra-logically
+realized --- and extended for quantifiers --- in HOL by stating a set
+of equations defining the logical constants (Fig.~\ref{QML_S5}). The
+resulting object logic is the HOML K with rigid terms and constant
+domains (possibilist quantifiers). Other logics (e.g. KB, S5) can be
+embedded by adding axioms that restrict the accessibility relation
+$r$. Varying domains and actualist quantifiers can be simulated by
+means of an existence predicate, which can be used to guard the
+quantifiers. Therefore, the embedding approach is very flexible.
 
@@
 \item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
-\item Transitivity: $\forall x. \forall y, \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
+\item Transitivity: $\forall x. \forall y. \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
 \end{itemize}
@@
 complete. In other words, in \SFiveU we face one single equivalence
-class of possible worlds, while in \SFive we may deal with several
+class of possible worlds, while in \SFive we may face several
 disconnected equivalence classes. In fact, for this reason, \SFiveU is
@@
 % \end{itemize}
-\[\Box \varphi \equiv \forall w. (\varphi w) \qquad \text{and}
-\qquad \Dia \varphi \equiv \exists w. (\varphi w)\]
+\[\Box \varphi \equiv \lambda w. \forall v.  \varphi(v) \qquad \text{and}
+\qquad \Dia \varphi \equiv \lambda w.\exists v. \varphi(v)\]
 
@@
 (\textit{``Necessarily, there exists God''}) can be derived from
-Scott's version of the axioms fully automatically.  The fully
+Scott's consistent version of the axioms fully automatically.  The fully
 automatic proof has been generated (in about 2,5 seconds) by the theorem prover
@@
 \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-verification in the highly trusted Isabelle/HOL still failed
+verification in the highly trusted Isabelle/HOL system still failed
 \cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5}.
 
-\begin{figure}
+\begin{figure}[t]
 \centerline{\includegraphics[width=\columnwidth]{./Images/Scott_S5.png}}
-\caption{Full Automation of T3 in \SFiveU; Verification of MC} \label{Scott_S5}
+\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5}
 \end{figure}
@@
 \item The lifted modal connectives are represented by prefixing the letter ``m'' (e.g.
-$m\wedge$ and $m\imp$).
+$m\wedge$ and $m\rightarrow$).
 The prefix disturbs the user, as it constantly reminds him/her that there is something unusual about the modal connectives.
@@
 for humans, it did contain relevant hints to the importance of the
-empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof})
-respectively self-difference $\lambda x.  x\not=x$. Philosophers
-might actually prefer using the latter variant over the
-former. However, on the given logic setting, with functional and
-Boolean extensionality, both terms have identical
-denotations. Consequently, they are mutually replacable below and the
-reader may simply use the version he prefers.
+empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof}),
+respectively, to self-difference $\lambda x.  x\not=x$.
+
+Note that both terms have identical denotations in a logic setting
+with full functional and Boolean extensionality as given
+here. Nevertheless, some philosophers may actually prefer the use of
+self-difference over the empty property\footnote{This occurred e.g. in
+  a private communication of the authors with Andr\'e Fuhrmann.} in
+the analysis below. However, for the proof to go through it is
+irrelevant which notion we use and the reader may simply replace the
+empty property by self-difference.
+
+% Philosophers
+% might actually prefer using the latter variant over the
+% former. However, on the given logic setting, with functional and
+% Boolean extensionality, both terms have identical
+% denotations. Consequently, they are mutually replacable below and the
+% reader may simply use the version he prefers.
 
@@
 (${\ess{\phi}{x} \biimp {\allq \psi} (\psi(x)
-\imp {\nec} \allq y (\phi(y) \imp \psi(y)))}$) it follows that the empty property is an essence of every individual (\textbf{Empty Essence Lemma}): 
+\imp {\nec} \allq y (\phi(y) \imp \psi(y)))}$) it follows that the
+empty property (or self-difference) is an essence of every individual (\textbf{Empty Essence Lemma}): 
 $$\allq x\,(\ess{\emptyset}{x})$$
@@
 
-\item Expanding the definition of ``necessary existence'' ( 
-  ${\NE(x) \equiv \allq
-  \phi [\ess{\phi}{x} \imp \nec
-  \exq y \phi(y)]}$), the following is obtained:
+\item Expanding the definition of ``necessary existence''
+  (${\NE(x) \equiv \allq \phi [\ess{\phi}{x} \imp \nec \exq y
+    \phi(y)]}$), the following is obtained:
   $$
@@
 
-\item The sentence above holds for all $\varphi$ and thus, in particular, for the empty property:
+\item The sentence above holds for all $\varphi$ and thus, in
+  particular, for the empty property (or self-difference):
 $$
@@
 \end{figure}
+
+\subsection{Mapping the Inconsistency to G\"odel}
+The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
+essence ($ess$), the definition of necessary existence ($NE$), the
+axioms A1(a) and A2 (which entail theorem T1), and axiom A5. It remains to demonstrate that
+these ingredients are actually present in G\"odel's manuscript in
+Fig.~\ref{GoedelScript}.
+
+This can easily be seen: Axioms A1(a) in
+Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom A2 and the
+highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
+A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to A4 and A3 in
+Fig.~\ref{GoedelScript}. The definitions of essence and necessary
+existence are easy to identify. Thus, we conclude that the verified
+inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
+G''odel's proof script in Fig.~\ref{GoedelScript}.
 
@@
 
-\newpage
-
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2016-01-13 Papers/InProgress/IJCAI/paper.tex
paper improvements

@@
 
-\newcommand{\imp}{\rightarrow}
+\newcommand{\imp}{{\rightarrow}}
 \newcommand{\biimp}{\leftrightarrow}
@@
 inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5}
-where the model finder Nitpick confirms consistency).\footnote{In
+where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
   personal communication, Dana Scott confirmed that he was unaware
@@
 
-\subsection{Improved Embedding \SFiveU}\label{sec:improvedembedding}
+\subsection{Improved Embedding}\label{sec:improvedembedding}
 \begin{figure}[t]
@@
 assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5}.
+The collaboration between the two systems has been orchestrated by
+Isabelle's Sledgehammer tool \cite{Sledgehammer}.
+
 With the embedding used by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40}, the
@@
 G\"odel's manuscript, Fig.~\ref{GoedelScript}).
-\item Quantifiers appear as higher-order defined constants, and not as binders.
+\item Quantifiers appear as higher-order dqefined constants, and not as binders.
 This forces the user to read (and write) formulas of the form $\forall (\lambda x.
@@
 A(x)$.
-\item The lifted modal connectives are represented by prefixing the letter ``m'' (e.g.
-$m\wedge$ and $m\rightarrow$).
-The prefix disturbs the user, as it constantly reminds him/her that there is something unusual about the modal connectives.
+\item The lifted modal connectives are represented by prefixing the
+  letter ``m'' (e.g. $m\wedge$ and $m\imp$). The prefix disturbs the
+  user, as it constantly reminds him/her that there is something
+  unusual about the modal connectives.
 \item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
@@
 
-\subsection{Informal Refutation}
+\subsection{Informal Argument}
 
@@
 The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
-essence ($ess$), the definition of necessary existence ($NE$), the
+essence (ess), the definition of necessary existence (NE), the
 axioms A1(a) and A2 (which entail theorem T1), and axiom A5. It remains to demonstrate that
@@
 
-This can easily be seen: Axioms A1(a) in
-Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom A2 and the
+This can easily be seen: Axioms A1a in
+Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
 highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
-A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to A4 and A3 in
+A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
 Fig.~\ref{GoedelScript}. The definitions of essence and necessary
@@
 inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
-G''odel's proof script in Fig.~\ref{GoedelScript}.
+G\"odel's original manuscript.
 
 \section{Conclusion}\label{sec:conclusion}
-The inconsistency of G\"odel's original axioms in his ontological
-argument has been detected automatically by the prover
+The axioms and definitions in G\"odel's manuscript are inconsistent;
+this has been detected automatically by the prover
 \textsc{Leo-II}. Here we have presented a rational reconstruction and
-verification of the inconsistency proof in Isabelle/HOL. This
-inconsistency argument applies to all normal modal logics including K.
+verification of the inconsistency argument in Isabelle/HOL. This
+argument is valid in all normal HOMLs including base logic K.
 \begin{figure*}
@@
 
-Moreover, we have reported on several technical improvements regarding
-the semantic embedding approach. In fact, we have a achieved a nearly
-perfect match between pen and paper presentations in HOML and the
-syntax in Isabelle/HOL. As a result, the embedding of HOML in
-HOL is now fully transparent and, as we think, ready for wider adoption by
+We have also presented several technical improvements regarding the
+semantic embedding approach. In particular, we have a achieved a
+nearly perfect match between pen and paper presentations in HOML and
+the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
+is now fully transparent and, as we think, ready for wider adoption by
 philosophers.
@@
 Isabelle/HOL. For example, Isabelle/HOL's Sledgehammer tool, which
-realises calls to external automated theorem provers, such as
+orchestrates the calls to external automated theorem provers, such as
 \textsc{Leo-II}, still fails to detect the inconsistency of G\"odel's
@@
 improvements. For example, in default setting, Sledgehammer does not
-inform the user early that a proof has been found and instead silently
-applies first a series of time-consuming analysis tools (e.g. it's
-minimizer), before it eventually informs the user of success. For the
+immediately inform the user when a proof has been found and instead silently
+ first executes a series of time-consuming proof analysis processes (e.g. it's
+dependency minimizer), before it eventually informs the user of success. For the
 involved proof problems presented here (e.g. G\"odel's theorem T3:
@@
 Moreover, our work points to a challenge for proof transformation:
-even if \textsc{Leo-II}'s proof could be automatically verified
-in Isabelle, this would clearly not entail that Isabelle/HOL could
-also extract from this verification a human-intuitive argument, for
-example, an argument as has been presented here. For philosophers,
-however, the reconstruction of such human-intuitive arguments from
-machine found proofs will likely be crucial for accepting any novel results.
+even if \textsc{Leo-II}'s proof can eventually be automatically
+verified in Isabelle, without a human-intuitive explanation of the
+proof, this will presumably be of limited value for philosophers.  The
+challenge is to extract from such a verification process an abstract
+level argument, similar to what we have presented in this paper: for
+philosophers, the reconstruction of such human-intuitive arguments
+from machine found proofs will likely be crucial for accepting any
+novel results.
 
@@
 contributed by \textsc{Leo-II}. However, this has been possible only
-after providing a more efficient embedding, based on a universal
-accessibility relation, for HOML S5 in HOL.
+after providing a more efficient embedding for HOML S5 in HOL.
 
@@
 the potential of our AI technology for philosophy: this technology is
-already in its current state able to contribute novel results to
+in its current state of development already capable of contributing novel results to
 metaphysics and to conduct reasoning steps at granularity-levels
@@
 
+\small 
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2016-01-15 Papers/InProgress/IJCAI/paper.tex
paper update

@@
 definitions, formalized in Isabelle, is shown in
-Fig.~\ref{Scott_S5}. The main difference to G\"odel's version is the
+Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is the
 addition of a conjunct in the definition of \emph{essence}. G\"odel's
@@
 fact crucial. Without it, G\"odel's original axioms are
-inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5}
+inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5U}
 where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
@@
 realized --- and extended for quantifiers --- in HOL by stating a set
-of equations defining the logical constants (Fig.~\ref{QML_S5}). The
+of equations defining the logical constants (Fig.~\ref{QML_S5U}). The
 resulting object logic is the HOML K with rigid terms and constant
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5.png}}
-\caption{Improved Embedding of \SFiveU} \label{QML_S5}
+\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5U.png}}
+\caption{Improved Embedding of \SFiveU} \label{QML_S5U}
 \end{figure}
@@
 
-The new embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QML_S5}.
+The new embedding of \SFiveU in Isabelle is shown in
+Fig.~\ref{QML_S5U}. It is important to note that we 
+have  $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$,
+despite the differences in the underlying frame structures.
+
 %
-\marginpar{@Bruno: I commented you paragraph here, since I am not
-  convinced (yet) that e.g. BF and CBF are equally valid in \SFiveU
-  and \SFive}
+% \marginpar{@Bruno: I commented you paragraph here, since I am not
+%   convinced (yet) that e.g. BF and CBF are equally valid in \SFiveU
+%   and \SFive}
 % \SFiveU is only slightly stronger than \SFive. Most importantly,
@@
 \textsc{Leo-II}~\cite{C26} and subsequently verified in the proof
-assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5}.
+assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5U}.
 The collaboration between the two systems has been orchestrated by
@@
 verification in the highly trusted Isabelle/HOL system still failed
-\cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5}.
+\cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5U}.
 
 \begin{figure}[t]
-\centerline{\includegraphics[width=\columnwidth]{./Images/Scott_S5.png}}
+\centerline{\includegraphics[width=\columnwidth]{./Images/Scott_S5U.png}}
 \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5}
+  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
 \end{figure}
@@
 
-In the embedding presented here, in Fig.~\ref{QML_S5}, advanced
+In the embedding presented here, in Fig.~\ref{QML_S5U}, advanced
 syntax-sugaring effects provided by Isabelle were used to prevent
@@
 The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
-As desired, the definition of essence in Fig.~\ref{Scott_S5} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
+As desired, the definition of essence in Fig.~\ref{Scott_S5U} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
 The embedding technique is now completely transparent to the user.
@@
 
-On the other hand, there is still room for improvements in
-Isabelle/HOL. For example, Isabelle/HOL's Sledgehammer tool, which
+On the other hand, there is still room for improvements. For example, Isabelle/HOL's Sledgehammer tool, which
 orchestrates the calls to external automated theorem provers, such as

### 2016-01-21 Papers/InProgress/IJCAI/paper.tex
picture added
paper improvements/additions

@@
 (nearly) perfect match between the original pen and paper
-presentations and our encoding in Isabelle/HOl is feasible. A more
+presentations and our encoding in Isabelle/HOL is feasible. A more
 user-friendly syntax, as reported here, is clearly an important
@@
 G\"odel was the work of Leibniz \cite{ToDo:Leibniz}.  G\"odel's manuscript
-(Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's
+(Fig.~\ref{GoedelScript}(a)) can be considered a translation of Leibniz's
 presentation of the argument into modern modal logic. G\"odel
@@
 different definition of essence can be seen either in his manuscript
-(Fig.~\ref{GoedelScript}) or, in more modern notation, in the Isabelle
+(Fig.~\ref{GoedelScript}(a)) or, in more modern notation, in the Isabelle
 formalization shown in Fig.~\ref{InconsistencyIsabelleK}. For Scott,
@@
 \item Lambda abstractions (which are a typical higher-order feature) appear explicitly in places where it didn't need to appear in a pure HOML formulation (cf.
-G\"odel's manuscript, Fig.~\ref{GoedelScript}).
+G\"odel's manuscript, Fig.~\ref{GoedelScript}(a)).
 \item Quantifiers appear as higher-order dqefined constants, and not as binders.
@@
 
-\subsection{Informal Argument}
-
+\subsection{Informal Argument} \label{sec:arg1}
 Based on the hints found in \textsc{Leo-II}'s refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
@@
 
-\subsection{Argument Reconstruction in Isabelle}
-
-To verify the correctness of the informal argument explained above, it was reconstructed in Isabelle/HOL, using Metis to automate the inessential parts.
-The essential use of the Empty Essence Lemma, on the other hand, is explicitly stated, to ensure that Isabelle is reconstructing the same argument.
-In fact, without the help of this lemma, Metis is not strong enough to refute G\"odel's axioms.
-\marginpar{ToDo: check this claim about Metis.}
+\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
+To verify the correctness of the informal argument explained above, it
+was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
+  external provers such as \textsc{Leo-II} or Satallax, 
+  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
+inessential parts. The essential use of the Empty Essence Lemma, on
+the other hand, is explicitly stated, to ensure that Isabelle is
+reconstructing the same argument. In fact, without the help of this
+lemma, Metis is not strong enough to refute G\"odel's
+axioms. \marginpar{ToDo: check this claim about Metis.}
 \begin{figure}[t]
 \centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
-\caption{Refutation of G\"odel's Axioms in Isabelle/HOl..} \label{InconsistencyIsabelleK}
+\caption{Refutation of G\"odel's Axioms in Isabelle/HOL.} \label{InconsistencyIsabelleK}
 \end{figure}
@@
 these ingredients are actually present in G\"odel's manuscript in
-Fig.~\ref{GoedelScript}.
+Fig.~\ref{GoedelScript}(a).
 
@@
 Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
-highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
+highlighted footnote remark in Fig.~\ref{GoedelScript}(a). Axioms A2 and
 A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
-Fig.~\ref{GoedelScript}. The definitions of essence and necessary
+Fig.~\ref{GoedelScript}(a). The definitions of essence and necessary
 existence are easy to identify. Thus, we conclude that the verified
@@
 G\"odel's original manuscript.
+
+\subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
+Isabelle/HOL's Sledgehammer tool, which orchestrates the calls to
+external automated theorem provers, such as \textsc{Leo-II}, still
+fails to detect the inconsistency of G\"odel's axioms in the standard
+embedding, while a direct modeling of the problem in TPTP THF syntax
+in combination with a direct call of \textsc{Leo-II} succeeds. In
+other words, without independent experiments directly in TPTP THF the
+inconsistency would thus not have been detected.
+\begin{figure*}
+\centerline{
+\includegraphics[width=.65\textwidth]{./Images/Manuscript2.png} \hfill
+\includegraphics[width=.35\textwidth]{./Images/Inconsistency_S5U_direct.png}
+}
+\caption{(a) G\"{o}del's manuscript, with mutually inconsistent axioms and
+  definitions highlighted; (b) Refutation of G\"odel's Axioms in
+\SFiveU} \label{Inconsistency_S5U} \label{GoedelScript} 
+\end{figure*}
+
+Recent experiments in Isabelle/HOL with the improved embedding \SFiveU were
+successful. Analogous to our findings in
+\$\ref{sec:improvedembedding}, the inconsistency can now be detected
+by \textsc{Leo-II} when called via
+Sledgehammer. Moreover, the result can subsequently be verified with
+Metis; cf. Fig.\ref{Inconsistency_S5U}. Interestingly, Metis generates an argument different to
+\$\ref{sec:arg1} and \$\ref{sec:arg2}, since axiom A1b is not used.
 
@@
 argument is valid in all normal HOMLs including base logic K.
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./Images/Manuscript2.png}}
-\caption{G\"{o}del's manuscript, with mutually inconsistent axioms and definitions highlighted.} \label{GoedelScript}
-\end{figure*}
 
@@
 
-On the other hand, there is still room for improvements. For example, Isabelle/HOL's Sledgehammer tool, which
-orchestrates the calls to external automated theorem provers, such as
-\textsc{Leo-II}, still fails to detect the inconsistency of G\"odel's
-axioms, while a direct modeling of the problem in TPTP THF syntax in
-combination with a direct call of \textsc{Leo-II} succeeds. In other
-words, without independent experiments directly in TPTP THF the
-inconsistency would thus not have been detected. There are further
-technical issues regarding Sledgehammer that motivate
-improvements. For example, in default setting, Sledgehammer does not
-immediately inform the user when a proof has been found and instead silently
- first executes a series of time-consuming proof analysis processes (e.g. it's
-dependency minimizer), before it eventually informs the user of success. For the
-involved proof problems presented here (e.g. G\"odel's theorem T3:
-\textit{Necessarily, there exists God}), this phase of silence may
-take several minutes (during which the user might actually give up on
-the proof attempt), while \textsc{Leo-II} already reported success to
-Sledgehammer much earlier (e.g. after 2,5 seconds for T3).
+On the other hand, there is still room for many pragmatical
+improvements in Isabelle; just one example: in default setting,
+Sledgehammer does not immediately inform the user when a proof has
+been found and instead silently first executes a series of
+time-consuming proof analysis processes (e.g. it's dependency
+minimizer), before it eventually reports success. For G\"odel's
+theorem T3 (\textit{Necessarily, there exists God}), for example, this
+phase of silence takes several minutes --- during which the user might
+actually give up on the proof attempt, while \textsc{Leo-II} already
+reported success to Sledgehammer after 2,5 seconds for T3
 
-Moreover, our work points to a challenge for proof transformation:
-even if \textsc{Leo-II}'s proof can eventually be automatically
-verified in Isabelle, without a human-intuitive explanation of the
-proof, this will presumably be of limited value for philosophers.  The
-challenge is to extract from such a verification process an abstract
-level argument, similar to what we have presented in this paper: for
-philosophers, the reconstruction of such human-intuitive arguments
-from machine found proofs will likely be crucial for accepting any
-novel results.
+More importantly, our work reveals a challenge for proof
+transformation: Without accompanying, human-intuitive explanations,
+the proofs generated by provers such as \textsc{Leo-II} or Metis, will
+presumably be only of limited value for philosophers.  The challenge
+is to extract abstract level arguments, similar to what we have
+presented in this paper. For philosophers, such human-intuitive
+arguments will likely be crucial for accepting any novel results.
 
@@
 
-\section*{Acknowledgments}
+%\section*{Acknowledgments} Will be added at a later point.
+\paragraph*{Acknowledgments} Will be added at a later point.
 
-Will be added at a later point.
+%\pagebreak
 
@@
 
-\small 
+%\small 
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2016-01-27 Papers/InProgress/IJCAI/submission.tex
Improve the paper at the sentence level

--- 2016-01-21 Papers/InProgress/IJCAI/paper.tex
@@
 
-\title{An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
+\title{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
 \author{Christoph Benzm\"uller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
@@
 extends previous computer-assisted studies of G\"odel's ontological
-argument. Similar to the related work \cite{J30,C40} the analysis has
+argument. Similarly to the related work \cite{J30,C40} the analysis has
 been conducted with automated theorem provers for classical
 higher-order logic (HOL; cf. \cite{andrewsSEP} and the references
-therein) even though G\"odel's proof is actually formulated in
+therein), even though G\"odel's proof is actually formulated in
 higher-order \emph{modal} logic (HOML; cf. \cite{homl} and the
@@
 analysis (in various modal logics) of the inconsistency of G\"{o}del's
-original version of the axioms used in his proof
+original version of the axioms used in his manuscript
 \shortcite{GoedelNotes}. The extraction, reconstruction and
@@
 On the technical side, the quest for constructing a compelling
-refutation, capable of convincing also human non-experts, led us to
-improve the syntax of the embedding of modal logics in Isabelle/HOL
+refutation, capable of convincing also human non-experts, led to an
+improvement of the syntax of the embedding of modal logics in Isabelle/HOL
 (as discussed in \S\ref{sec:improvedsyntax}). With the new syntax, a
@@
 Another novel contribution reported here (in \S\ref{sec:improvedembedding}) is
-the implementation of an alternative embedding for the only slightly
+the implementation of an alternative embedding for the slightly
 stronger modal logic \SFiveU. Our experiments have shown that the new
@@
   alone, without relying on the argument's intermediate argumentation
-  steps (i.e., lemmata).
+  steps (i.e., lemmata). \marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript. We need to fix this as I fixed the date for Goedel's manuscript.}
 \item Verifying automatically (in Isabelle/HOL) the proof of the modal
@@
 arguably improved Anselm's argument. Of particular importance to
-G\"odel was the work of Leibniz \cite{ToDo:Leibniz}.  G\"odel's manuscript
+G\"odel was the work of Leibniz. 
+%\cite{ToDo:Leibniz}.  
+G\"odel's manuscript
 (Fig.~\ref{GoedelScript}(a)) can be considered a translation of Leibniz's
@@
 definitions, formalized in Isabelle, is shown in
-Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is the
-addition of a conjunct in the definition of \emph{essence}. G\"odel's
+Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is an extra conjunct in the definition of \emph{essence}. G\"odel's
 different definition of essence can be seen either in his manuscript
@@
 considered inessential and merely an oversight by many. For instance,
-Hazen 1998, p.365 \cite[p.365]{Hazen} states that ``G\"odel left this
-clause out in [11], but this appears to have been an oversight--it is
-included in related manuscripts''. However, the omitted conjunct is in
+Hazen [\citeyear[p.365]{Hazen}] states that ``G\"odel left this
+clause out in [11], but this appears to have been an oversight''.
+% -- it is
+%included in related manuscripts''. 
+However, the extra conjunct is in
 fact crucial. Without it, G\"odel's original axioms are
@@
 criticizing his argument. Especially since the discovery by Sobel
-\cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
+\shortcite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
   $\phi\rightarrow \Box \phi$, states that contingent truth implies
@@
   of G\"odel's axioms depends, it is not necessarily the case that
-  these variants are also inconsistent; each case needs to be analysed
+  these variants are also inconsistent; they to be analyzed
   separately.}. Opponents of the argument
-(e.g. Oppy \cite[p.226/227]{oppy96:_goedel_ontol_argum}
-\cite[p.364]{oppy00:_respon_gettin}
-\cite[p.1068]{oppy08:_higher_order_ontol_argum}) have also proposed
+(e.g. Oppy [\citeyear[p.226--227]{oppy96:_goedel_ontol_argum};
+\citeyear[p.364]{oppy00:_respon_gettin};
+\citeyear[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
 parodies and other criticisms, referring to variants where the
@@
 
-For a computational analysis of G\"odel's ontological argument in this
-approach the embedding of HOMLs such as K,
-KB and S5 with various domain conditions (possibilist and actualist)
-is required. This idea has been successfully followed in related work
+For a computational analysis of G\"odel's ontological argument, the embedding of HOMLs such as K,
+KB and S5 with various domain conditions (possibilist and actualist quantification)
+is required. This idea has been successfully employed in related work
 \cite{C40}. The embedding of HOML is in fact
@@
 means of an existence predicate, which can be used to guard the
-quantifiers. Therefore, the embedding approach is very flexible.
+quantifiers. The embedding approach is, therefore, very flexible.
 
@@
 \begin{enumerate}
-\item Lambda abstractions (which are a typical higher-order feature) appear explicitly in places where it didn't need to appear in a pure HOML formulation (cf.
+\item Lambda abstractions (which are a typical higher-order feature) appear explicitly in places where they didn't need to appear in a pure HOML formulation (cf.
 G\"odel's manuscript, Fig.~\ref{GoedelScript}(a)).
-\item Quantifiers appear as higher-order dqefined constants, and not as binders.
+\item Quantifiers appear as higher-order defined constants, and not as binders.
 This forces the user to read (and write) formulas of the form $\forall (\lambda x.
@@
 \item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
-Instead of reading $\psi(y)$, as he would expect, he has to read $(\psi y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
+Instead of reading $\psi(y)$, as he would expect, he has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
 \end{enumerate}
@@
 
-In the typical workflow of trying to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first.
+In the typical workflow during an attempt to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first.
 For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}).
@@
 Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable.
-The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and machine-oriented proof calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}) were used.
+The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used a machine-oriented proof calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}).
 Part of the file is displayed in Fig.~\ref{LEO-Proof}.
@@
 for humans, it did contain relevant hints to the importance of the
-empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof}),
-respectively, to self-difference $\lambda x.  x\not=x$.
-
-Note that both terms have identical denotations in a logic setting
+empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof}).
+%
+Note that the terms for the empty property ($\lambda x. \bot$) and for the property of self-difference ($\lambda x.  x\not=x$) have identical denotations in a logic setting
 with full functional and Boolean extensionality as given
 here. Nevertheless, some philosophers may actually prefer the use of
-self-difference over the empty property\footnote{This occurred e.g. in
+self-difference over the empty property\footnote{This occurred in
   a private communication of the authors with Andr\'e Fuhrmann.} in
@@
 empty property by self-difference.
+In HOL it is customary to think of a set as a unary predicate that is true for (and only for) elements of the set.
+Therefore, the empty property (i.e.
+the unary predicate that is false for any individual) represents the empty set, and hence we use the empty set symbol ($\emptyset$) as an abbreviation for the empty property.
+\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?}
 
@@
 
-\item To see that the sentence above is contradictory, we may reason semantically, in terms of possible worlds.
+\item To see that the sentence above is contradictory, we may reason semantically, thinking of possible worlds.
 If $w_0$ is the arbitrary current world, the $\pos$ operator forces the existence of a world $w$ accessible from $w_0$ such that $\nec \bot$ is true in $w$.
@@
 weaker modal logic K. In our experiments \textsc{Leo-II} (like all
-other HOL provers) was still to weak to automatically prove the
+other HOL provers) was still too weak to automatically prove the
 inconsistency already in logic K. Hence, this remains an open problem for automated
@@
 reconstructing the same argument. In fact, without the help of this
-lemma, Metis is not strong enough to refute G\"odel's
+lemma, Metis is still not strong enough to refute G\"odel's
 axioms. \marginpar{ToDo: check this claim about Metis.}
@@
 essence (ess), the definition of necessary existence (NE), the
-axioms A1(a) and A2 (which entail theorem T1), and axiom A5. It remains to demonstrate that
+axioms A1(a) and A2 (which entail theorem T1), and axiom A5. It remains to show that
 these ingredients are actually present in G\"odel's manuscript in
@@
 \subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
+
 Isabelle/HOL's Sledgehammer tool, which orchestrates the calls to
@@
 \caption{(a) G\"{o}del's manuscript, with mutually inconsistent axioms and
-  definitions highlighted; (b) Refutation of G\"odel's Axioms in
+  definitions highlighted; (b) Refutation in
 \SFiveU} \label{Inconsistency_S5U} \label{GoedelScript} 
@@
 
-Recent experiments in Isabelle/HOL with the improved embedding \SFiveU were
-successful. Analogous to our findings in
-\$\ref{sec:improvedembedding}, the inconsistency can now be detected
-by \textsc{Leo-II} when called via
-Sledgehammer. Moreover, the result can subsequently be verified with
-Metis; cf. Fig.\ref{Inconsistency_S5U}. Interestingly, Metis generates an argument different to
+\marginpar{ToDo: Are we really supposed to use \$ instead of ``Sec.''?}
+On the other hand (and further confirming the claims from \$\ref{sec:improvedembedding}), 
+the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
+the inconsistency could be detected
+by \textsc{Leo-II} also when called via
+Sledgehammer. Moreover, the result could subsequently be verified with
+Metis even without the Empty Essence Lemma; cf. Fig.\ref{Inconsistency_S5U}. 
+Interestingly, Metis' argument is different from that of
 \$\ref{sec:arg1} and \$\ref{sec:arg2}, since axiom A1b is not used.
@@
 the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
-is now fully transparent and, as we think, ready for wider adoption by
+is now fully transparent and ready for wider adoption by
 philosophers.
@@
 been found and instead silently first executes a series of
-time-consuming proof analysis processes (e.g. it's dependency
-minimizer), before it eventually reports success. For G\"odel's
+time-consuming proof analysis processes (e.g. its dependency
+minimization), before it eventually reports success. For G\"odel's
 theorem T3 (\textit{Necessarily, there exists God}), for example, this
 phase of silence takes several minutes --- during which the user might
-actually give up on the proof attempt, while \textsc{Leo-II} already
-reported success to Sledgehammer after 2,5 seconds for T3
+actually give up on the proof attempt --- even though \textsc{Leo-II} already
+reported success to Sledgehammer after 2,5 seconds.
 
-More importantly, our work reveals a challenge for proof
-transformation: Without accompanying, human-intuitive explanations,
+More importantly, our work reveals a challenge for automated reasoning:
+the (so far partially manual) extraction of an informal argument from a formal proof. 
+Without accompanying human-understandable explanations,
 the proofs generated by provers such as \textsc{Leo-II} or Metis, will
-presumably be only of limited value for philosophers.  The challenge
-is to extract abstract level arguments, similar to what we have
-presented in this paper. For philosophers, such human-intuitive
-arguments will likely be crucial for accepting any novel results.
+presumably be only of limited value for philosophers, for whom intuitive
+arguments remain crucial for the acceptance of novel results.
 
@@
 
-Both, the automated detection of the inconsistency in G\"odel's axioms
-and the fully automatic proof of T3 from Scott's axioms demonstrates
+Both the automated detection of the inconsistency in G\"odel's axioms
+and the fully automatic proof of T3 from Scott's axioms demonstrate
 the potential of our AI technology for philosophy: this technology is
@@
 metaphysics and to conduct reasoning steps at granularity-levels
-beyond common human capabilities.
+beyond common human capabilities.  % ToDo: or even Gödel-level capabilities :-)
 
@@
 %\section*{Acknowledgments} Will be added at a later point.
-\paragraph*{Acknowledgments} Will be added at a later point.
+%\paragraph*{Acknowledgments} Will be added at a later point.
 
@@
 %German Research Foundation DFG, Chad Brown
+
+%ToDo: reduce paper length to 7 pages
+
+%ToDo: use \textsc for all systems. Not only for LEO-II.
 

### 2016-01-27 Papers/InProgress/IJCAI/submission.tex
Update marginpar comments.

@@
   alone, without relying on the argument's intermediate argumentation
-  steps (i.e., lemmata). \marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript. We need to fix this as I fixed the date for Goedel's manuscript.}
+  steps (i.e., lemmata).
+\marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript.
+We need to fix this as I fixed the date for Goedel's manuscript.
+But I don't know the correct date for Scott's manuscript.}
 \item Verifying automatically (in Isabelle/HOL) the proof of the modal
@@
 \end{figure}
-%\marginpar{ToDo: this picture has to be replaced by one without Isabelle errors}
 
@@
 the unary predicate that is false for any individual) represents the empty set, and hence we use the empty set symbol ($\emptyset$) as an abbreviation for the empty property.
-\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?}
+\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?
+Or should we instead stop using the empty set symbol in the informal argument (and use $\lambda x.
+\bot$ instead) and remove this convoluted commentary that I added here.}
 
@@
 lemma, Metis is still not strong enough to refute G\"odel's
-axioms. \marginpar{ToDo: check this claim about Metis.}
+axioms.
 \begin{figure}[t]

### 2016-01-28 Papers/InProgress/IJCAI/submission.tex
sentence added to abstract

@@
 \begin{abstract}
-  This paper discusses the discovery of the inconsistency in G\"odel's ontological argument as a success story for artificial intelligence.
-Despite the popularity of the argument since the appearance of G\"odel's manuscript in 1970, the inconsistency of the axioms used in the argument remained unnoticed until 2013, when it was detected automatically by the higher-order theorem prover \textsc{Leo-II}.
-Understanding and verifying the refutation generated by the prover turned out to be a time-consuming task, and its completion, as reported here, required the reconstruction of the refutation in the Isabelle proof assistant.
-The development of an improved syntactical hiding for the embedding technique, utilizing Isabelle's binding notation mechanism, allows the refutation to be presented in a human-friendly way, suitable for non-experts in the technicalities of higher-order theorem proving.
-This brings us a step closer to wider adoption of logic-based artificial intelligence tools by philosophers.
+  This paper discusses the discovery of the inconsistency in G\"odel's
+  ontological argument as a success story for artificial
+  intelligence. Despite the popularity of the argument since the
+  appearance of G\"odel's manuscript in 1970, the inconsistency of the
+  axioms used in the argument remained unnoticed until 2013, when it
+  was detected automatically by the higher-order theorem prover
+  \textsc{Leo-II}. Understanding and verifying the refutation
+  generated by the prover turned out to be a time-consuming task, and
+  its completion, as reported here, required the reconstruction of the
+  refutation in the Isabelle proof assistant. The development of an
+  improved syntactical hiding for the embedding technique, utilizing
+  Isabelle's binding notation mechanism, allows the refutation to be
+  presented in a human-friendly way, suitable for non-experts in the
+  technicalities of higher-order theorem proving. This brings us a
+  step closer to wider adoption of logic-based artificial intelligence
+  tools by philosophers.
+
+  The paper reports on further significant improvements regarding full
+  automation of the ontological argument in the context of a higher-order modal logic S5 with a
+  universal accessibility relation.
 \end{abstract}
@@
 %% The file named.bst is a bibliography style file for BibTeX 0.99c
+
+\small
 \bibliographystyle{named}

### 2016-01-28 Papers/InProgress/IJCAI/submission.tex
further improvements

@@
   refutation in the Isabelle proof assistant. The development of an
-  improved syntactical hiding for the embedding technique, utilizing
-  Isabelle's binding notation mechanism, allows the refutation to be
+  improved syntactical hiding for the embedding technique
+  % , utilizing
+  % Isabelle's binding notation mechanism, 
+  allows the refutation to be
   presented in a human-friendly way, suitable for non-experts in the
@@
 
-  The paper reports on further significant improvements regarding full
-  automation of the ontological argument in the context of a higher-order modal logic S5 with a
-  universal accessibility relation.
+  In addition, the paper reports on further improvements regarding the
+  full automation of the ontological argument in a higher-order modal
+  logic \textbf{S5} with a universal accessibility relation.
 \end{abstract}
@@
 their work has received a media repercussion on a global scale\footnote{A
-  small collection of news articles is available at {\scriptsize
+  collection of news articles is available at {\scriptsize
     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
@@
 here to a wider community of philosophers, who are not necessarily
-experts in automated reasoning or higher-order logic.
+experts in automated reasoning or HOL.
 
-Another novel contribution reported here (in \S\ref{sec:improvedembedding}) is
-the implementation of an alternative embedding for the slightly
-stronger modal logic \SFiveU. Our experiments have shown that the new
-embedding is more efficient, as the following two, previously open problems can now be solved:
+Another novel contribution reported here (in
+\S\ref{sec:improvedembedding}) is the implementation of an alternative
+embedding for the slightly stronger modal logic \SFiveU, which is
+based on a universal accessibility relation. Our experiments have
+shown that the new embedding is more efficient, as the following two,
+previously open problems can now be solved:
 \begin{itemize}
-\item Proving the final theorem T3 \textit{``Necessarily, there
-    exists God''} (in Scott's \shortcite{ScottNotes} consistent variant) fully automatically and directly from the axioms
+\item Proving automatically the final theorem T3 \textit{Necessarily, there
+    exists God} (in Scott's \shortcite{ScottNotes} consistent variant) directly from the axioms
   alone, without relying on the argument's intermediate argumentation
@@
   personal communication, Dana Scott confirmed that he was unaware
-  that G\"odel's axioms were inconsistent.}
+  at the time that G\"odel's axioms were inconsistent.}
+
+\begin{figure}[t]
+\centerline{\includegraphics[width=.95\columnwidth]{./Images/Scott_S5U.png}}
+\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
+\end{figure}
 
@@
 For more than four decades, the serious consequences of G\"odel's
-mistake remained unnoticed, despite intense activity focused on
+mistake remained unnoticed, despite intense activity on
 criticizing his argument. Especially since the discovery by Sobel
@@
 attempting to avoid the modal collapse. Many of these variants omit
-the crucial conjunct in the definition of essence as well\footnote{As
+the crucial conjunct in the definition of essence as well.\footnote{As
   these variants also change other axioms, on which the inconsistency
@@
   these variants are also inconsistent; they to be analyzed
-  separately.}. Opponents of the argument
+  separately.} Opponents of the argument
 (e.g. Oppy [\citeyear[p.226--227]{oppy96:_goedel_ontol_argum};
@@
 
-For a computational analysis of G\"odel's ontological argument, the embedding of HOMLs such as K,
-KB and S5 with various domain conditions (possibilist and actualist quantification)
+For a computational analysis of G\"odel's ontological argument, the embedding of HOMLs such as \textbf{K},
+\textbf{KB} and \textbf{S5} with various domain conditions (possibilist and actualist quantification)
 is required. This idea has been successfully employed in related work
 \cite{C40}. The embedding of HOML is in fact
-straightforward. Formulas in HOML are \emph{lifted}, translated to predicates
+straightforward. Formulas in HOML are \emph{lifted}, i.e., converted into predicates
 over worlds, which are themselves explicitly represented as
@@
 of equations defining the logical constants (Fig.~\ref{QML_S5U}). The
-resulting object logic is the HOML K with rigid terms and constant
-domains (possibilist quantifiers). Other logics (e.g. KB, S5) can be
+resulting object logic is the HOML \textbf{K} with rigid terms and constant
+domains (possibilist quantifiers). Other logics (e.g. \textbf{KB}, \textbf{S5}) can be
 embedded by adding axioms that restrict the accessibility relation
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=\columnwidth]{./Images/QML_S5U.png}}
+\centerline{\includegraphics[width=.95\columnwidth]{./Images/QML_S5U.png}}
 \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
@@
 reflexive, symmetric and transitive. The usual approach to embed
-\SFive would be to use the standard translation for K described above
+\SFive would be to use the standard translation for \textbf{K} described above
 and to state that $r$ is an equivalence relation, e.g., by postulating
@@
 
-The new embedding of \SFiveU in Isabelle is shown in
+The new embedding of \SFiveU in Isabelle/HOL is shown in
 Fig.~\ref{QML_S5U}. It is important to note that we 
@@
 With this improved embedding, the final theorem T3
-(\textit{``Necessarily, there exists God''}) can be derived from
+(\textit{Necessarily, there exists God}) can be derived from
 Scott's consistent version of the axioms fully automatically.  The fully
@@
 
-\begin{figure}[t]
-\centerline{\includegraphics[width=\columnwidth]{./Images/Scott_S5U.png}}
-\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
-\end{figure}
-
 \subsection{Improved Syntax in Isabelle}\label{sec:improvedsyntax}
 
-Wider adoption of higher-order theorem proving technology for reasoning about and within embedded object logics, especially among non-expert users, is still hindered by the gap between the syntax used by people, when they write logical formulas with pen and paper, and the syntax used by higher-order logic theorem provers.
-Even when the syntax of the underlying higher-order system is elegant (as is the case in Isabelle/HOL), the embedding of HOML into HOL may easily expose details of HOL that may be uncommon to the user, disturbing his/her experience while using the system.
-To illustrate this point, Fig.~\ref{UglyEssence} shows how the definition of essence looked like in previous work \cite{J28}, where advanced syntax-sugaring features were not used.
-It looks notably higher-order, and its style differs significantly from the common style seen in works on modal logics and the ontological argument.
-The following specific issues can be enumerated:
+Wider adoption of HOL theorem proving technology for
+reasoning about and within embedded object logics, especially among
+non-expert users, is still hindered by the gap between the syntax used
+by people, when they write logical formulas with pen and paper, and
+the syntax used by HOL theorem provers. Even when the syntax of the
+underlying higher-order system is elegant (as is the case in
+Isabelle/HOL), the embedding of HOML into HOL may easily expose
+details of HOL that may be uncommon to the user, disturbing his/her
+experience while using the system. To illustrate this point,
+Fig.~\ref{UglyEssence} shows how the definition of essence looked like
+in previous work \cite{J28}, where advanced syntax-sugaring features
+were not used. It looks notably higher-order, and its style differs
+significantly from the common style seen in works on modal logics and
+the ontological argument. The following specific issues can be
+enumerated:
 \begin{enumerate}
-\item Lambda abstractions (which are a typical higher-order feature) appear explicitly in places where they didn't need to appear in a pure HOML formulation (cf.
-G\"odel's manuscript, Fig.~\ref{GoedelScript}(a)).
+\item $\lambda$-abstractions, a typical higher-order feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}(a)).
 \item Quantifiers appear as higher-order defined constants, and not as binders.
@@
 \begin{figure}
-\centerline{\includegraphics[width=\columnwidth]{./Images/UglyEssence.png}}
+\centerline{\includegraphics[width=.95\columnwidth]{./Images/UglyEssence.png}}
 \caption{ Definition of Essence using Old Syntax } \label{UglyEssence}
@@
 \caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
-  substitutions (e.g. with the empty property) are highlighted.} \label{LEO-Proof}
+  substitutions (e.g. with the empty property) are highlighted. In the
+red part, property variable $\texttt{SV8}$ has been instantiated with the
+$\lambda \texttt{SV16}. \lambda \texttt{SV17}. \bot$, i.e., the empty property.
+} \label{LEO-Proof}
 \end{figure*}
@@
 for humans, it did contain relevant hints to the importance of the
-empty property $\lambda x. \bot$ (cf. Fig.~\ref{LEO-Proof}).
+empty property $\lambda x. \bot$ (respectively $\lambda x. \lambda
+y. \bot$ in the embedding approach, cf. Fig.~\ref{LEO-Proof}; the 
+additional ``$\lambda y.$'' is due to the formula lifting).
 %
@@
 irrelevant which notion we use and the reader may simply replace the
-empty property by self-difference.
-In HOL it is customary to think of a set as a unary predicate that is true for (and only for) elements of the set.
-Therefore, the empty property (i.e.
-the unary predicate that is false for any individual) represents the empty set, and hence we use the empty set symbol ($\emptyset$) as an abbreviation for the empty property.
+empty property by self-difference. In HOL it is customary to think of
+a set as a unary predicate that is true for (and only for) elements of
+the set. Therefore, the empty property (i.e. the unary predicate that
+is false for any individual) represents the empty set, and hence we
+use the empty set symbol $\emptyset$ as an abbreviation for the empty property.
 \marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?
@@
 empty property (or self-difference) is an essence of every individual (\textbf{Empty Essence Lemma}): 
-$$\allq x\,(\ess{\emptyset}{x})$$
+$$\allq x (\ess{\emptyset}{x})$$
 
-\item From theorem T1 (positive Properties are possibly
-  exemplified: ${\allq \phi} [P(\phi) \imp {\pos}  \exq x
+\item From theorem T1 (\textit{Positive properties are possibly
+  exemplified}: ${\allq \phi} [P(\phi) \imp {\pos}  \exq x
   \phi(x)]$) and axiom A5 (``necessary existence'' is a positive property: $P(\NE)$ ), it follows that $\NE$ is possibly exemplified:
@@
 \textsc{Leo-II} uses a symmetric accessibility relation, and thus
-requires the modal logic KB. The informal, human-constructed
+requires the modal logic \KB. The informal, human-constructed
 refutation described above, on the other hand, requires only the
-weaker modal logic K. In our experiments \textsc{Leo-II} (like all
+weaker modal logic \K. In our experiments \textsc{Leo-II} (like all
 other HOL provers) was still too weak to automatically prove the
-inconsistency already in logic K. Hence, this remains an open problem for automated
+inconsistency already in logic \K. Hence, this remains an open problem for automated
 theorem provers.
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+\centerline{\includegraphics[width=.95\columnwidth]{./Images/InconsistencyIsabelleK.png}}
 \caption{Refutation of G\"odel's Axioms in Isabelle/HOL.} \label{InconsistencyIsabelleK}
@@
 
-Isabelle/HOL's Sledgehammer tool, which orchestrates the calls to
-external automated theorem provers, such as \textsc{Leo-II}, still
-fails to detect the inconsistency of G\"odel's axioms in the standard
-embedding, while a direct modeling of the problem in TPTP THF syntax
-in combination with a direct call of \textsc{Leo-II} succeeds. In
+Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
+external provers such as \textsc{Leo-II}, still
+fails to detect the inconsistency of G\"odel's axioms (in the standard
+embedding), while a direct modeling of the problem in TPTP THF syntax
+in combination with a direct call of \textsc{Leo-II} succeeded. In
 other words, without independent experiments directly in TPTP THF the
@@
 
-\marginpar{ToDo: Are we really supposed to use \$ instead of ``Sec.''?}
-On the other hand (and further confirming the claims from \$\ref{sec:improvedembedding}), 
+%\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
+On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
 the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
@@
 Interestingly, Metis' argument is different from that of
-\$\ref{sec:arg1} and \$\ref{sec:arg2}, since axiom A1b is not used.
+\S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
 
@@
 verification of the inconsistency argument in Isabelle/HOL. This
-argument is valid in all normal HOMLs including base logic K.
+argument is valid in all normal HOMLs including base logic \K.
 
@@
 contributed by \textsc{Leo-II}. However, this has been possible only
-after providing a more efficient embedding for HOML S5 in HOL.
+after providing a more efficient embedding for HOML \textbf{S5} in HOL.
 

### 2016-01-29 Papers/InProgress/IJCAI/submission.tex
further improvements

@@
 \usepackage{latexsym}
+
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
   at the time that G\"odel's axioms were inconsistent.}
+
+% \begin{figure*}
+%   \centering
+%   \begin{subfigure}[t]{0.4\textwidth}
+%     \includegraphics[width=\textwidth]{./Images/Scott_S5U.png}
+%     \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+%       Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
+%   \end{subfigure}
+%   ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+%   % (or a blank line to force the subfigure onto a new line)
+%   \begin{subfigure}[t]{0.4\textwidth}
+%     \includegraphics[width=\textwidth]{./Images/QML_S5U.png}
+%     \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
+%   \end{subfigure}
+%   ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+%   % (or a blank line to force the subfigure onto a new line)
+%   \begin{subfigure}[t]{0.4\textwidth}
+%     \includegraphics[width=\textwidth]{./Images/Inconsistency_S5U_direct.png}
+%     \caption{Inconsistency  in \SFiveU} \label{Inconsistency_S5U} 
+%   \end{subfigure}
+%   \begin{subfigure}[t]{0.4\textwidth}
+%     \centerline{\includegraphics[width=\textwidth]{./Images/InconsistencyIsabelleK.png}}
+%     \caption{Refutation of G\"odel's Axioms in Isabelle/HOL.} \label{InconsistencyIsabelleK}
+%   \end{subfigure}
+% \end{figure*}
 
@@
 \centerline{\includegraphics[width=.95\columnwidth]{./Images/UglyEssence.png}}
-\caption{ Definition of Essence using Old Syntax } \label{UglyEssence}
+\caption{Definition of Essence using Old Syntax} \label{UglyEssence}
 \end{figure}
@@
 \begin{figure*}
-\centerline{
-\includegraphics[width=.65\textwidth]{./Images/Manuscript2.png} \hfill
-\includegraphics[width=.35\textwidth]{./Images/Inconsistency_S5U_direct.png}
-}
-\caption{(a) G\"{o}del's manuscript, with mutually inconsistent axioms and
-  definitions highlighted; (b) Refutation in
-\SFiveU} \label{Inconsistency_S5U} \label{GoedelScript} 
+  \centering
+  \begin{subfigure}[t]{0.72\textwidth}
+    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
+    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
+      definitions highlighted} \label{GoedelScript} 
+  \end{subfigure}
+   \begin{subfigure}[t]{0.27\textwidth}
+     \includegraphics[width=\textwidth]{./Images/Inconsistency_S5U_direct.png}
+     \caption{ Inconsistency in \SFiveU} \label{Inconsistency_S5U} 
+   \end{subfigure}
+% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
+% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
+% \caption{(a) G\"{o}del's manuscript, with mutually inconsistent axioms and
+%   definitions highlighted; (b) Inconsistency in \SFiveU} \label{GoedelScript} \label{Inconsistency_S5U} 
 \end{figure*}

### 2016-01-29 Papers/InProgress/IJCAI/submission.tex
further improvements

@@
 G\"odel's manuscript
-(Fig.~\ref{GoedelScript}(a)) can be considered a translation of Leibniz's
+(Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's
 presentation of the argument into modern modal logic. G\"odel
@@
 different definition of essence can be seen either in his manuscript
-(Fig.~\ref{GoedelScript}(a)) or, in more modern notation, in the Isabelle
+(Fig.~\ref{GoedelScript}) or, in more modern notation, in the Isabelle
 formalization shown in Fig.~\ref{InconsistencyIsabelleK}. For Scott,
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=.95\columnwidth]{./Images/Scott_S5U.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./Images/Scott_S5U.png}}
 \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=.95\columnwidth]{./Images/QML_S5U.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./Images/QML_S5U.png}}
 \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
@@
 \begin{enumerate}
-\item $\lambda$-abstractions, a typical higher-order feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}(a)).
+\item $\lambda$-abstractions, a typical higher-order feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
 \item Quantifiers appear as higher-order defined constants, and not as binders.
@@
 \begin{figure}
-\centerline{\includegraphics[width=.95\columnwidth]{./Images/UglyEssence.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./Images/UglyEssence.png}}
 \caption{Definition of Essence using Old Syntax} \label{UglyEssence}
@@
 is false for any individual) represents the empty set, and hence we
-use the empty set symbol $\emptyset$ as an abbreviation for the empty property.
-\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?
+use the empty set symbol $\emptyset$ as an abbreviation for the empty
+property. 
+\marginpar{Isabelle proof now uses empty set; the commentary is still convoluted}
+%\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?
 Or should we instead stop using the empty set symbol in the informal argument (and use $\lambda x.
@@
 empty property (or self-difference) is an essence of every individual (\textbf{Empty Essence Lemma}): 
-$$\allq x (\ess{\emptyset}{x})$$
+$$\allq x\; (\ess{\emptyset}{x})$$
 
@@
   constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
-inessential parts. The essential use of the Empty Essence Lemma, on
+inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
 the other hand, is explicitly stated, to ensure that Isabelle is
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=.95\columnwidth]{./Images/InconsistencyIsabelleK.png}}
-\caption{Refutation of G\"odel's Axioms in Isabelle/HOL.} \label{InconsistencyIsabelleK}
+\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
 \end{figure}
@@
 these ingredients are actually present in G\"odel's manuscript in
-Fig.~\ref{GoedelScript}(a).
+Fig.~\ref{GoedelScript}.
 
@@
 Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
-highlighted footnote remark in Fig.~\ref{GoedelScript}(a). Axioms A2 and
+highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
 A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
-Fig.~\ref{GoedelScript}(a). The definitions of essence and necessary
+Fig.~\ref{GoedelScript}. The definitions of essence and necessary
 existence are easy to identify. Thus, we conclude that the verified
@@
 inconsistency would thus not have been detected.
-\begin{figure*}
+\begin{figure*}[t]
   \centering
-  \begin{subfigure}[t]{0.72\textwidth}
+  \begin{subfigure}[t]{0.715\textwidth}
     \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
@@
   \end{subfigure}
-   \begin{subfigure}[t]{0.27\textwidth}
+   \begin{subfigure}[t]{0.28\textwidth}
      \includegraphics[width=\textwidth]{./Images/Inconsistency_S5U_direct.png}
-     \caption{ Inconsistency in \SFiveU} \label{Inconsistency_S5U} 
+     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
    \end{subfigure}
@@
 Sledgehammer. Moreover, the result could subsequently be verified with
-Metis even without the Empty Essence Lemma; cf. Fig.\ref{Inconsistency_S5U}. 
-Interestingly, Metis' argument is different from that of
-\S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
+Metis even without the Empty Essence Lemma; cf. Fig.~\ref{Inconsistency_S5U}. 
+% Interestingly, Metis' argument is different from that of
+% \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
 

### 2016-01-29 Papers/InProgress/IJCAI/submission.tex
minor fixes

@@
   steps (i.e., lemmata). 
-\marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript.
+% \marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript.
 We need to fix this as I fixed the date for Goedel's manuscript.
@@
 mistake remained unnoticed, despite intense activity on
-criticizing his argument. Especially since the discovery by Sobel
+criticising his argument. Especially since the discovery by Sobel
 \shortcite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
@@
 \subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
+\begin{figure}[t]
+\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
+\end{figure}
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[t]{0.715\textwidth}
+    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
+    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
+      definitions highlighted} \label{GoedelScript} 
+  \end{subfigure}
+   \begin{subfigure}[t]{0.28\textwidth}
+     \includegraphics[width=\textwidth,height=7.2cm]{./Images/Inconsistency_S5U_direct.png}
+     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
+   \end{subfigure}
+% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
+% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
+ \caption{The inconsistency in G\"{o}del's manuscript has been
+   detected and verified by HOL ATPs} 
+\end{figure*}
 To verify the correctness of the informal argument explained above, it
@@
 axioms.
-\begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
-\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
-\end{figure}
 
@@
 \subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
-
 Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
@@
 inconsistency would thus not have been detected.
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[t]{0.715\textwidth}
-    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
-    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
-      definitions highlighted} \label{GoedelScript} 
-  \end{subfigure}
-   \begin{subfigure}[t]{0.28\textwidth}
-     \includegraphics[width=\textwidth]{./Images/Inconsistency_S5U_direct.png}
-     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
-   \end{subfigure}
-% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
-% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
-% \caption{(a) G\"{o}del's manuscript, with mutually inconsistent axioms and
-%   definitions highlighted; (b) Inconsistency in \SFiveU} \label{GoedelScript} \label{Inconsistency_S5U} 
-\end{figure*}
 

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve abstract and introduction

@@
-\typeout{An AI Success Story in Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument}
+\typeout{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
 
@@
   was detected automatically by the higher-order theorem prover
-  \textsc{Leo-II}. Understanding and verifying the refutation
+  \textsc{Leo-II}. 
+  Understanding and verifying the refutation
   generated by the prover turned out to be a time-consuming task, and
   its completion, as reported here, required the reconstruction of the
-  refutation in the Isabelle proof assistant. The development of an
+  refutation in the Isabelle proof assistant and eventually led to a novel and more efficient way of automating higher-order modal logic \textbf{S5} with a universal accessibility relation.
+Furthermore, the development of an
   improved syntactical hiding for the embedding technique
-  % , utilizing
-  % Isabelle's binding notation mechanism, 
   allows the refutation to be
@@
   tools by philosophers.
-
-  In addition, the paper reports on further improvements regarding the
-  full automation of the ontological argument in a higher-order modal
-  logic \textbf{S5} with a universal accessibility relation.
 \end{abstract}
@@
 
-The in-depth computational analysis presented here substantially
+The in-depth analysis presented here substantially
 extends previous computer-assisted studies of G\"odel's ontological
@@
 based on a universal accessibility relation. Our experiments have
-shown that the new embedding is more efficient, as the following two,
+shown that the new embedding is more efficient, as the following two
 previously open problems can now be solved:
 \begin{itemize}
-\item Proving automatically the final theorem T3 \textit{Necessarily, there
-    exists God} (in Scott's \shortcite{ScottNotes} consistent variant) directly from the axioms
+\item Automatically proving the final theorem T3 (\textit{Necessarily, there
+  exists God}), directly from Scott's \shortcite{ScottNotes} (consistent) axioms
   alone, without relying on the argument's intermediate argumentation
   steps (i.e., lemmata).
-% \marginpar{ToDo: 2004 is the year of Sobel's book, not of Scott's manuscript.
-We need to fix this as I fixed the date for Goedel's manuscript.
-But I don't know the correct date for Scott's manuscript.}
-\item Verifying automatically (in Isabelle/HOL) the proof of the modal
+\item Automatically verifying, in Isabelle/HOL, the proof of the modal
   collapse \cite{Sobel}, which is one of the most strongly criticized
-  `side-effects' of the argument's axioms.
+  logical consequences of the argument's axioms.
 \end{itemize}

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Fix convoluted paragraph about empty set notation

@@
 red part, property variable $\texttt{SV8}$ has been instantiated with the
-$\lambda \texttt{SV16}. \lambda \texttt{SV17}. \bot$, i.e., the empty property.
+$\lambda \texttt{SV16}_{\mu}. \lambda \texttt{SV17}_{\iota}. \bot$, i.e., the (lifted) empty property.
 } \label{LEO-Proof}
@@
 for humans, it did contain relevant hints to the importance of the
-empty property $\lambda x. \bot$ (respectively $\lambda x. \lambda
-y. \bot$ in the embedding approach, cf. Fig.~\ref{LEO-Proof}; the 
-additional ``$\lambda y.$'' is due to the formula lifting).
+empty property $\lambda x. \bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets).
 %
-Note that the terms for the empty property ($\lambda x. \bot$) and for the property of self-difference ($\lambda x.  x\not=x$) have identical denotations in a logic setting
+Note that the terms for the empty property\footnote{An additional lambda abstraction occurs in the empty property in LEO-II's proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.} ($\lambda x.
+\bot$) and for the property of self-difference ($\lambda x.
+x\not=x$) have identical denotations in a logic setting
 with full functional and Boolean extensionality as given
@@
 irrelevant which notion we use and the reader may simply replace the
-empty property by self-difference. In HOL it is customary to think of
-a set as a unary predicate that is true for (and only for) elements of
-the set. Therefore, the empty property (i.e. the unary predicate that
-is false for any individual) represents the empty set, and hence we
-use the empty set symbol $\emptyset$ as an abbreviation for the empty
-property. 
-\marginpar{Isabelle proof now uses empty set; the commentary is still convoluted}
-%\marginpar{ToDo: should we use the empty set symbol in the isabelle code in figure 5 as well?
-Or should we instead stop using the empty set symbol in the informal argument (and use $\lambda x.
-\bot$ instead) and remove this convoluted commentary that I added here.}
+empty property by self-difference.
 

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Add missing verb

@@
   of G\"odel's axioms depends, it is not necessarily the case that
-  these variants are also inconsistent; they to be analyzed
+  these variants are also inconsistent; they must be analyzed
   separately.} Opponents of the argument

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Add guard to weaken claim

@@
 %\cite{ToDo:Leibniz}.  
-G\"odel's manuscript
+Although G\"odel's notion of positive property is not exactly the same as Leibniz's notion of perfection, G\"odel's manuscript
 (Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Shorten and reposition footnote

@@
 with full functional and Boolean extensionality as given
-here. Nevertheless, some philosophers may actually prefer the use of
-self-difference over the empty property\footnote{This occurred in
-  a private communication of the authors with Andr\'e Fuhrmann.} in
+here. Nevertheless, some philosophers\footnote{Private communication with Andr\'e Fuhrmann.} may actually prefer the use of
+self-difference over the empty property in
 the analysis below. However, for the proof to go through it is

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve section 2

@@
 an essential property of an individual must be possessed by
-him/her. For G\"odel, this is not required. This difference has been
+him/her. For G\"odel, this is not required.
+
+G\"odel's omission has been
 considered inessential and merely an oversight by many. For instance,
@@
 %included in related manuscripts''. 
-However, the extra conjunct is in
-fact crucial. Without it, G\"odel's original axioms are
-inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5U}
-where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
-  personal communication, Dana Scott confirmed that he was unaware
-  at the time that G\"odel's axioms were inconsistent.}
+For more than four decades, its serious consequences remained unnoticed, 
+despite numerous analysis and criticisms of the
+argument. Especially since the discovery by Sobel
+\shortcite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
+  $\phi\rightarrow \Box \phi$, states that contingent truth implies
+  necessary truth; it can be interpreted as \textit{everything is
+    pre-determined} or even \textit{there is no free will}. } is
+entailed by G\"odel's (or also Scott's) axioms, several variants have
+been proposed
+\cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
+attempting to avoid the modal collapse. Many of these variants omit
+the crucial conjunct in the definition of essence as well.\footnote{As
+  these variants also change other axioms, on which the inconsistency
+  of G\"odel's axioms depends, it is not necessarily the case that
+  these variants are also inconsistent; they must be analyzed
+  separately.} Opponents of the argument
+(e.g. Oppy [\citeyear[p.226--227]{oppy96:_goedel_ontol_argum};
+\citeyear[p.364]{oppy00:_respon_gettin};
+\citeyear[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
+parodies and other criticisms, referring to variants where the
+conjunct is omitted.
 
@@
 
-For more than four decades, the serious consequences of G\"odel's
-mistake remained unnoticed, despite intense activity on
-criticising his argument. Especially since the discovery by Sobel
-\shortcite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
-  $\phi\rightarrow \Box \phi$, states that contingent truth implies
-  necessary truth; it can be interpreted as \textit{everything is
-    pre-determined} or even \textit{there is no free will}. } is
-entailed by G\"odel's (or also Scott's) axioms, several variants have
-been proposed
-\cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
-attempting to avoid the modal collapse. Many of these variants omit
-the crucial conjunct in the definition of essence as well.\footnote{As
-  these variants also change other axioms, on which the inconsistency
-  of G\"odel's axioms depends, it is not necessarily the case that
-  these variants are also inconsistent; they must be analyzed
-  separately.} Opponents of the argument
-(e.g. Oppy [\citeyear[p.226--227]{oppy96:_goedel_ontol_argum};
-\citeyear[p.364]{oppy00:_respon_gettin};
-\citeyear[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
-parodies and other criticisms, referring to variants where the
-conjunct is omitted.
+However, as explained here, the extra conjunct is in
+fact crucial. Without it, G\"odel's original axioms are
+inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5U}
+where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
+  personal communication, Dana Scott confirmed that he was unaware
+  at the time that G\"odel's axioms were inconsistent.}
 

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve section 1.1

@@
 focus on a non-modal formalization of St. Anselm's~\shortcite{Proslogion} simpler 
-and older ontological argument. 
-None of these previous works formalises and automates variants of \emph{higher-order} and \emph{modal} logics, which are, however, crucial
-for G\"{o}del's more complex ontological argument.
+and older ontological argument. In contrast, the greater complexity of G\"odel's argument requires the formalization and automation of variants of \emph{higher-order} and \emph{modal} logics.
 

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve section 3
Improve section 3.2

@@
 embedded by adding axioms that restrict the accessibility relation
-$r$. Varying domains and actualist quantifiers can be simulated by
-means of an existence predicate, which can be used to guard the
+$r$. Varying domains and actualist quantifiers can be simulated by using
+ an existence predicate to guard the
 quantifiers. The embedding approach is, therefore, very flexible.
@@
 \begin{enumerate}
-\item $\lambda$-abstractions, a typical higher-order feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
+\item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
 \item Quantifiers appear as higher-order defined constants, and not as binders.
@@
 \item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
-Instead of reading $\psi(y)$, as he would expect, he has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
+Instead of reading $\psi(y)$, as he/she would expect, he/she has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
 \end{enumerate}
@@
 syntax-sugaring effects provided by Isabelle were used to prevent
-issues as enumerated above.
+issues as those enumerated above.
 The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
@@
 
-We hope that the syntax improvements described here will render the computer-assisted analysis of ontological arguments (as well as future works utilizing the embedding of modal logics into HOL) accessible to a wider audience and encourage philosophers to adopt logic-based artificial intelligence tools in order to conduct similar experiments in computational metaphysics.
+The syntax improvements described here render the computer-assisted analysis of ontological arguments accessible to a wider audience and ease the adoption of logic-based artificial intelligence tools by philosophers interested in topics where modal logic reasoning is required.
 

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve Section 4

@@
 Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable.
-The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used a machine-oriented proof calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}).
+The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used machine-oriented calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}).
 Part of the file is displayed in Fig.~\ref{LEO-Proof}.
@@
 
-This can easily be seen: Axioms A1a in
+This can easily be seen: Axiom A1a in
 Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
@@
 \subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
+
 Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
 external provers such as \textsc{Leo-II}, still
-fails to detect the inconsistency of G\"odel's axioms (in the standard
-embedding), while a direct modeling of the problem in TPTP THF syntax
+fails to detect the inconsistency of G\"odel's axioms in the standard
+embedding of \SFive, while a direct modeling of the problem in TPTP THF syntax
 in combination with a direct call of \textsc{Leo-II} succeeded. In
-other words, without independent experiments directly in TPTP THF the
-inconsistency would thus not have been detected.
+other words, without independent experiments with no mediation through Sledgehammer, the
+inconsistency would not have been detected.
 
@@
 Sledgehammer. Moreover, the result could subsequently be verified with
-Metis even without the Empty Essence Lemma; cf. Fig.~\ref{Inconsistency_S5U}. 
+Metis even without the Empty Essence Lemma (cf. Fig.~\ref{Inconsistency_S5U}). 
 % Interestingly, Metis' argument is different from that of

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve conclusion

@@
 \section{Conclusion}\label{sec:conclusion}
+
 The axioms and definitions in G\"odel's manuscript are inconsistent;
-this has been detected automatically by the prover
-\textsc{Leo-II}. Here we have presented a rational reconstruction and
+this was detected automatically by the prover
+\textsc{Leo-II}. Here we presented a rational reconstruction and
 verification of the inconsistency argument in Isabelle/HOL. This
@@
 We have also presented several technical improvements regarding the
-semantic embedding approach. In particular, we have a achieved a
+semantic embedding approach. In particular, we have achieved a
 nearly perfect match between pen and paper presentations in HOML and
 the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
-is now fully transparent and ready for wider adoption by
-philosophers.
+is now fully transparent, more user-friendly and ready for wider adoption.
 
@@
 proof of T3 directly from Scott's axioms. Again, this proof was
-contributed by \textsc{Leo-II}. However, this has been possible only
-after providing a more efficient embedding for HOML \textbf{S5} in HOL.
+contributed by \textsc{Leo-II}. This has become possible only
+after we provided a more efficient embedding for HOML \SFiveU (instead of \SFive) in HOL.
 
@@
 and the fully automatic proof of T3 from Scott's axioms demonstrate
-the potential of our AI technology for philosophy: this technology is
-in its current state of development already capable of contributing novel results to
+the potential of our AI technology for philosophy: this technology is,
+in its current state of development, already capable of contributing novel results to
 metaphysics and to conduct reasoning steps at granularity-levels

### 2016-01-31 Papers/InProgress/IJCAI/submission.tex
Improve section 4.3

@@
 \subsection{Mapping the Inconsistency to G\"odel}
+
 The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
@@
 
-This can easily be seen: Axiom A1a in
+This can be easily seen: Axiom A1a in
 Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
@@
 Fig.~\ref{GoedelScript}. The definitions of essence and necessary
-existence are easy to identify. Thus, we conclude that the verified
+existence are easy to identify. Therefore, the verified
 inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 

### 2016-02-01 Papers/InProgress/IJCAI/submission.tex
minor modifications

@@
 definitions, formalized in Isabelle, is shown in
-Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is an extra conjunct in the definition of \emph{essence}. G\"odel's
+Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is an
+extra conjunct in the definition of \emph{essence} (\emph{ess}). G\"odel's
 different definition of essence can be seen either in his manuscript
@@
 essence (ess), the definition of necessary existence (NE), the
-axioms A1(a) and A2 (which entail theorem T1), and axiom A5. It remains to show that
+axioms A1a and A2 (which entail theorem T1), and axiom A5. It remains to show that
 these ingredients are actually present in G\"odel's manuscript in

### 2016-02-01 Papers/InProgress/IJCAI/submission.tex
slight rewording in the abstract; added permission quote in Fig. 6

@@
   intelligence. Despite the popularity of the argument since the
-  appearance of G\"odel's manuscript in 1970, the inconsistency of the
-  axioms used in the argument remained unnoticed until 2013, when it
-  was detected automatically by the higher-order theorem prover
-  \textsc{Leo-II}. 
-  Understanding and verifying the refutation
-  generated by the prover turned out to be a time-consuming task, and
-  its completion, as reported here, required the reconstruction of the
-  refutation in the Isabelle proof assistant and eventually led to a novel and more efficient way of automating higher-order modal logic \textbf{S5} with a universal accessibility relation.
-Furthermore, the development of an
-  improved syntactical hiding for the embedding technique
-  allows the refutation to be
+  appearance of G\"odel's manuscript in the early 70's, the
+  inconsistency of the axioms used in the argument remained unnoticed
+  until 2013, when it was detected automatically by the higher-order
+  theorem prover \textsc{Leo-II}.  Understanding and verifying the
+  refutation generated by the prover turned out to be a time-consuming
+  task. Its completion, as reported here, required the
+  reconstruction of the refutation in the Isabelle proof assistant, and it
+  also led to a novel and more efficient way of automating
+  higher-order modal logic \textbf{S5} with a universal accessibility
+  relation. Furthermore, the development of an improved syntactical
+  hiding for the utilized logic embedding technique allows the refutation to be
   presented in a human-friendly way, suitable for non-experts in the
@@
     \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
-      definitions highlighted} \label{GoedelScript} 
+      definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript} 
   \end{subfigure}

### 2016-02-02 Papers/InProgress/IJCAI/submission.tex
paper has now exactly 6 pages plus references as required

@@
 resulting object logic is the HOML \textbf{K} with rigid terms and constant
-domains (possibilist quantifiers). Other logics (e.g. \textbf{KB}, \textbf{S5}) can be
+domains (possibilist quantifiers). Other logics (e.g. \textbf{KB}, \textbf{S5}) are
 embedded by adding axioms that restrict the accessibility relation
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./Images/QML_S5U.png}}
+\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./Images/QML_S5U.png}}
 \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
@@
 %\section*{Acknowledgments} Will be added at a later point.
-%\paragraph*{Acknowledgments} Will be added at a later point.
+\paragraph*{Acknowledgments} Will be added at a later point.
 

### 2016-03-07 Papers/2014/Universalis/ModalCollapse.tex
started with preparation of final version

@@
-\documentclass{birkjour}
+\documentclass{birkmult}
 
@@
 \usepackage{url}
+\usepackage{hyperref}
+\usepackage[hyphenbreaks]{breakurl}
+
 \usepackage{commands}
@@
 
-All results announced in this note have been 
+\sloppy All results announced in this note have been 
 obtained experimentally using interactive and 
@@
 \url{github.com/FormalTheology/GoedelGod/}
-in the files \texttt{ModalCollapse.thy} and \texttt{ModalSquareOfOpposition.thy} inside the folder \texttt{Formalizations/Isabelle/Meta} as well as in files inside the folder \texttt{Formalizations/Isabelle/Anderson}.
+in the files \url{ModalCollapse.thy} and \texttt{ModalSquareOfOpposition.thy} inside the folder \url{Formalizations/Isabelle/Meta} as well as in files inside the folder \texttt{Formalizations/Isabelle/Anderson}.
 

### 2016-03-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex

@@
 
-It is likely that a reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.
+A reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.
+
+\section{A natural deduction proof based on Anderson's axiomatization}
+
+\setcounter{axiom}{0}
+\setcounter{lemma}{0}
+\setcounter{theorem}{0}
+\setcounter{corollary}{0}
+\setcounter{definition}{0}
+
+%Anderson modification of the axioms in A1 equivalence removed and definition of essence altered. 
+%Anderson's implication instead of equivalence. 
+\begin{axiom}
+\label{A1}
+Either a property or its negation is positive, but not both:
+$$
+\all \varphi. [P(\neg \varphi) \imp \neg P(\varphi)]
+$$
+\end{axiom}
+
+\begin{axiom}
+\label{A2}
+A property necessarily implied by a positive property is positive:
+$$
+\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
+$$
+\end{axiom}
+
+\begin{theorem}
+\label{T1}
+Positive properties are possibly exemplified:
+$$
+\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
+$$
+\end{theorem}
+
+The proof is otherwise identical to the proof of theorem \ref{T1} in section \ref{sec:newproof} but the equivalence of axiom \ref{A1} does not need to be reduced to an implication by implication elimination.
+
+\begin{definition}[Anderson's alternative Godlike]
+\label{D1}
+%A \emph{God-like} being possesses all positive properties:
+$$
+G(x) \biimp \forall \varphi. [P(\varphi) \biimp \nec \varphi(x)]
+$$
+\end{definition}
+
+\begin{axiom}
+\label{A3}
+The property of being god-like is positive:
+$$
+P(G)
+$$
+\end{axiom}
+
+\begin{corollary}
+\label{C1}
+Possibly, a God-like being exists:
+$$
+\pos \ex x. G(x)
+$$
+\end{corollary}
+
+\begin{axiom}
+\label{A4}
+Positive properties are necessarily positive:
+$$
+\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+$$
+\end{axiom}
+
+\begin{definition}[Anderson's alternative essence with Scott's additional conjunct]
+\label{D2}
+%sAn \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties:
+$$
+\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\nec \psi(x) \biimp \nec \all x. (\varphi(x) \imp \psi(x)))
+$$
+\end{definition}
+
+%At this point, G\"odel and Scott proceed to prove Theorem 2, which is then used to prove Lemma 1.
+However, a close inspection of Scott's proof of Lemma 1 reveals that Theorem 2 is an unnecessary detour.
+Therefore, we do not include it in the new proof.
+
+\begin{definition}[Anderson's alternative where G\"odel's definition of essence is replaced by def. \ref{D2}]
+\label{D3}
+\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+$$
+E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
+$$
+\end{definition}
+
+\begin{axiom}
+\label{A5}
+Necessary existence is a positive property:
+$$
+P(E)
+$$
+\end{axiom}
+
+\begin{lemma}
+\label{L1}
+If there is a God-like being, then there is a God-like being necessarily:
+$$
+\ex z. G(z) \imp \nec \ex x. G(x)
+$$
+\end{lemma}
+
+\begin{proof} \hfill
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
+\UIC{$[\subproof_6] \quad G(\gamma)$}
+\end{prooftree}
+
+\begin{scriptsize}
+\begin{prooftree}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
+\AXC{$ $} \RightLabel{5}
+\UIC{$\nec \rho(\gamma)$} \RightLabel{$\imp_E$}
+\BIC{$P(\lambda x.\rho(x))$} 
+\AXC{$ $} \RightLabel{6}
+\UIC{$G(y) $}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(y)]$}
+\UIC{$[P(\lambda x.\rho(x)) \biimp \nec \rho(y)]$}
+\BIC{$\nec \rho(y)$}\RightLabel{$\nec_E$}
+\UIC{$\omega':  \rho(y)$}\RightLabel{$\imp_I^6$}
+\UIC{$\omega':  G(y)\imp \rho(y)$}\RightLabel{$\forall_I$}
+\UIC{$\omega':  \forall y. (G(y)\imp \rho(y))$}\RightLabel{$\nec_I$}
+\UIC{$\nec \forall y. (G(y)\imp \rho(y))$}\RightLabel{$\imp_I^5$}
+\UIC{$\nec \rho(\gamma)\imp \nec \forall y. (G(y)\imp \rho(y))$}
+\end{prooftree}
+\end{scriptsize}
+
+\begin{scriptsize}
+\begin{prooftree}
+\AXC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
+\AXC{$ $} \RightLabel{6}
+\UIC{$\nec \rho(\gamma)$} \RightLabel{}
+\UIC{$\rho(\gamma)$} \RightLabel{$\nec_E$}
+\AXC{$\subproof_7$} \dashedLine
+\UIC{$\omega': \rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
+\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
+\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}  
+\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
+\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
+\end{prooftree}
+\end{scriptsize}
+
+%implication for equivalence in essence: 
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{7}
+\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\nec_E$}
+\UIC{$\omega'':   \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\forall_E$}
+\UIC{$\omega'':   G(\gamma)\imp \varphi(\gamma)$}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$\omega'':  G(\gamma)$}  \RightLabel{$\imp_E$}
+\BIC{$\omega'':   \varphi(\gamma)$}\RightLabel{$\nec_I$}
+\UIC{$\nec \varphi(\gamma)$}\RightLabel{$\imp_I^7$}
+\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\subproof_8$} \dashedLine
+\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $}
+\AXC{$ $} \RightLabel{6}
+\UIC{$G(x)$}\RightLabel{$\wedge_I$}
+\BIC{$\ess{G}{\gamma}$}
+\AXC{Axiom \ref{A5}} \dashedLine
+\UIC{$P(E)$}\dottedLine 
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
+\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\imp_E$}
+\BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$ (reflexive frame)}
+\UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{D3}}
+\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
+\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
+\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
+\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
+\end{prooftree}
+\end{small}
+\end{proof}
+
+\setcounter{theorem}{2}
+\begin{theorem}
+\label{T3}
+Necessarily, there exists a God-like being:
+$$
+\nec \ex x. G(x)
+$$
+\end{theorem}
+
+\subsection{Why is the proof of modal collapse prevented?}
+
+The alteration of the definition of essence in Anderson's proof prevents the standard proof of modal collapse \ref{Collapse} 
+$$\vdash A \imp \nec A
+$$ 
+by the additional necessitation, which converts the proof of modal collapse into a proof of a tautology.
+$$
+\vdash \nec A \imp \nec A
+$$
+
+%For all propositions $A$, the following \emph{modal collapse} proposition is not provable with Anderson's alternative definitions with the same proof as G\"odel or Scott's:
+
+\begin{small}
+\begin{prooftree}
+\AXC{Theorem 2}\dashedLine
+\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
+\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$}
+\UIC{$\all y.[G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
+\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
+\UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
+\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_9 $}\dashedLine
+\UIC{$\ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
+\AXC{Theorem 3}\dashedLine
+\UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
+\UIC{$\ex y.G(y)$} \RightLabel{$\imp_E$}
+\BIC{$ \nec A \biimp \nec \all x. (G(x) \imp A)$} \doubleLine
+\UIC{$ \nec A \biimp \nec (\ex x. G(x) \imp A)$}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\nec A$} \RightLabel{$\imp_E$}
+\BIC{$ \nec (\ex x. G(x) \imp A)$} \RightLabel{$\nec_E$}
+\UIC{$ \omega:  \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
+\AXC{Theorem 3}\dashedLine
+\UIC{$\nec \ex y.G(y)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: \ex x.G(x)$} \RightLabel{$\imp_E$}
+\BIC{$\omega: A$} \RightLabel{$\nec_I$}
+\UIC{$ \nec A$} \RightLabel{$\imp_I^1$}
+   \UIC{$\nec A\imp \nec A$}
+\end{prooftree}
+\end{small}
+
+The exchange of the implication in the G\"odel-Scott definition of essence for an equivalence and the additional  modal necessitation of the property $A$ prevents the implication of a necessary fact from its contingent counterpart as is seen in the second part of the proof above because the hypothetical assumption of the contingent property $A$ has to be exchanged for its necessitation for the proof structure to be preserved.
+
+\section{A natural deduction proof based on Bj\o rdal's axiomatization}
+
+In Bj\o rdal's axiomatization godlikeness is taken as primitive, $G_B$.
+
+\begin{definition}
+\label{B:D1}
+A property is positive iff it is necessarily possessed by every God-like being.
+$$
+P_B(\varphi) \equiv \nec \all x. [G_B(x) \imp \varphi(x)]
+$$
+\end{definition}
+
+\begin{definition}
+\label{B:D2}
+A maximal composite of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
+$$MCP(\varphi, x) \equiv (\varphi(x) \wedge P_B (\varphi)) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(\varphi(y) \imp \psi(y)))
+$$
+\end{definition}
+
+\begin{definition}
+\label{B:D3}
+Necessary existence of an individual is the necessary exemplification of all its maximal composites.
+$$
+NE_B (x) \equiv \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))
+$$
+\end{definition}
+
+\begin{axiom}
+\label{A:A1'} If a property is positive, its negation is not positive: 
+$$
+\all \varphi. [P_B (\varphi) \imp \neg P_B (\neg\varphi)]
+$$
+\end{axiom}
+
+\begin{axiom}
+\label{A:A5'} Necessary existence is a positive property. 
+$$
+P_B (NE_B )
+$$
+\end{axiom}
+
+The proof of the possible instantiation of $G_B$ is adjusted by proving Theorem 1 of section \ref{sec:newproof} directly for $G_B$ and not for properties in general.
+Corollary 1 of section \ref{sec:newproof} follows by proving the positivity of godlikeness as in lemma \ref{B:L1} below.
+
+\begin{theorem}
+\label{T1}
+Positive properties are possibly exemplified:
+$$
+\all \varphi. [P(G_B) \imp \pos \ex x.G_B(x)]
+$$
+\end{theorem}
+\begin{proof} \hfill
+
+\begin{prooftree}
+\AXC{$ $} \RightLabel{2}
+\UIC{$\nec \neg \ex x.G_B(x) $} \RightLabel{$\nec_E$}
+\UIC{$\omega: \neg \ex x.G_B(x) $} 
+\AXC{$ $} \RightLabel{$1$}
+\UIC{$\omega: G_B(x)$} \RightLabel{$\ex_I$}
+\UIC{$\omega: \ex x.G_B(x) $} \RightLabel{$\imp_E$}
+\BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
+\UIC{$\omega:  \neg G_B(x) $} \RightLabel{$\imp_I$}
+\UIC{$\omega:  G_B(x) \imp \neg G_B(x) $} \RightLabel{$\all_I$}
+\UIC{$\omega:  \all x.(G_B(x) \imp \neg G_B(x)) $}  \RightLabel{$\nec_I$}
+\UIC{$ \nec \all x.(G_B(x) \imp \neg G_B(x)) $}\RightLabel{$\wedge_I$}
+\UIC{$P(\neg G_B)$}
+\AXC{Axiom 6***} \dashedLine
+\UIC{$\all \varphi. [P(\neg G_B) \imp \neg P(G_B)]$} \RightLabel{$\all_E$}
+\UIC{$ P( \neg G_B(x)) \imp \neg P(G_B) $}  \RightLabel{$\imp_E$}
+\BIC{$\neg P(G_B) $} 
+\AXC{$ $} \RightLabel{3} 
+\UIC{$P(G_B)$} 
+\BIC{$\bot$} \RightLabel{$\imp_I^2$}
+\UIC{$ \neg \nec \neg \ex x.G_B(x)$} \doubleLine %\RightLabel{definition of $\pos$}
+\UIC{$ \pos \ex x.G_B(x)$}  \RightLabel{$\imp_I^3$}
+\UIC{$ P(\rho) \imp \pos \ex x.G_B(x) $} \RightLabel{$\all_I$}
+\UIC{$ \all \varphi. [P(G_B) \imp \pos \ex x.G_B(x)] $}
+
+\end{prooftree}
+
+\end{proof}
+
+\begin{lemma}
+\label{B:L1}
+If there is a God-like being, then there is a God-like being necessarily:
+$$
+\ex z. G_B(z) \imp \nec \ex x. G_B(x)
+$$
+\end{lemma}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $}
+\UIC{$G_B(x) $} 
+\AXC{$ $}\doubleLine
+\UIC{$\nec \all x. [G_B(x) \imp G_B(x)] $}\RightLabel{ D\ref{B:D1}}\dottedLine
+\UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
+\BIC{$G_B(x) \wedge P_B(G_B) $}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $}\dashedLine
+\UIC{$G_B(x) \wedge P_B(G_B) $}
+\AXC{$ $}\RightLabel{ $1$}
+\UIC{$(\psi(x) \wedge P_B (\psi))$} \RightLabel{ $\wedge_E$}
+\UIC{$P_B (\psi)$}\RightLabel{ D\ref{B:D1}}\dottedLine
+\UIC{$\nec\all y.(G_B(y) \imp \psi(y))$} \RightLabel{ $\imp_I^1$}
+\UIC{$(\psi(x) \wedge P_B (\psi))\imp \nec\all y.(G_B(y) \imp \psi(y))$}\RightLabel{ $\forall_I$}
+\UIC{$\all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y)))$} \RightLabel{ $\wedge_I$}
+\BIC{$MCP(G_B, x) $}
+
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{Axiom \ref{A:A5'}}\dashedLine
+\UIC{$P_B (NE_B)$} \RightLabel{ D\ref{B:D1}}\dottedLine
+\UIC{$\nec \all x. [G_B(x) \imp NE_B(x)]$}\RightLabel{ D\ref{B:D3}}\dottedLine
+\UIC{$\nec \all x. [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\nec_E$}
+\UIC{$\all x. [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\forall_E$}
+\UIC{$ [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}
+\AXC{$ $} \RightLabel{ $2$}
+\UIC{$G_B(x) $}  \RightLabel{ $\imp_E$}
+\BIC{$\all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$} \RightLabel{ $\forall_E$}
+\UIC{$(MCP(G_B, x) \imp \nec \ex y.G_B(y))]$}
+\AXC{$ $}\dashedLine
+\UIC{$MCP(G_B, x) $} \RightLabel{ $\imp_E$}
+\BIC{$\nec \ex y.G_B(y)$}\RightLabel{ $\imp_I^2$}
+\UIC{$G_B(x) \imp \nec \ex y.G_B(y)$}
+\end{prooftree}
+\end{small}
+
+%ToDo: acknowledgements
 

### 2016-03-24 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Journal layout

@@
-\documentclass{llncs}
-
-\usepackage{setspace}
-\doublespacing
+\documentclass [smallextended] {svjour3} [natbib]
+
+%\usepackage{setspace}
+%\doublespacing
 
@@
 \usepackage{latexsym}
+\usepackage{natbib}
 
@@
 
-\author{Annika Siders\inst{1} \and Bruno Woltzenlogel Paleo\inst{2,3}}
+\author{Annika Siders \and Bruno Woltzenlogel Paleo}
 
@@
 \institute{ 
-  \email{annika.siders@helsinki.fi} \\
-  +358 2941 29241
+ Annika Siders \at 
   Department of Philosophy \\
@@
   P.O. Box 24 (Unioninkatu 40 A) \\
-  Finland \\[2em]
+  Finland \\ 
+  \email{annika.siders@helsinki.fi} \\
+  +358 2941 29241[2em]
   \and 
-  \email{bruno@logic.at}
+  Bruno Woltzenlogel Paleo \at 
   Theory and Logic Group \\ 
   Vienna University of Technology \\ 
-  Austria \\[2em]
+  Austria \\
+   \email{bruno@logic.at} \\
   \and
-  \email{bruno.woltzenlogel.paleo@anu.edu.au}
+    Bruno Woltzenlogel Paleo \at 
   Logic and Computation Group \\
@@
   Australia
+    \email{bruno.woltzenlogel.paleo@anu.edu.au} 
 }
@@
 
-\maketitle
-
-\clearpage
-
-\title{A Variant of G\"{o}del's Ontological Proof \\
-in a Natural Deduction Calculus}
+%\maketitle
+
+%\clearpage
+
+%\title{A Variant of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus}
 \author{}
@@
 \noindent
-\textbf{Keywords: } G\"odel's Ontological Proof, Higher-Order Modal Logics, Natural Deduction
+\textbf{Keywords: } G\"odel's Ontological Proof, Higher-Order Modal Logics, Natural Deduction,
 
@@
 
-G\"odel bulit on Leibniz's work \cite{adams} and brought the ontological argument to a modern form using a modal logic with higher-order quantification over properties.
+G\"odel bulit on Leibniz's work \citep{adams} and brought the ontological argument to a modern form using a modal logic with higher-order quantification over properties.
 In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
-G\"odel's work was saved in his own notes \cite{Goedel} as well as in notes by Scott \cite{scott}, in whom he confided his proof.
+G\"odel's work was saved in his own notes \citep{Goedel} as well as in notes by Scott \citep{scott}, in whom he confided his proof.
 
@@
 This deductive inference is not formalizable in a finite first-order calculus.
-The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \cite{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \citep{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
 
@@
 Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of proof calculi (e.g.
-Fitting's tableaux \cite{fitting} or Sobel's sentential modal calculus \cite{sobel2}).
+Fitting's tableaux \citep{fitting} or Sobel's sentential modal calculus \citep{sobel2}).
 Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
 Secondly, it makes the ontological proof accessible to people who are familiar with a natural deduction style.
-Thirdly, as natural deduction is the style used by proof assistants such as Coq \cite{coq} and Isabelle \cite{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this.
-
-The first contribution of this paper is a detailed formalization of Scott's version \cite{scott} of G\"odel's ontological argument \cite{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
+Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this.
+
+The first contribution of this paper is a detailed formalization of Scott's version \citep{scott} of G\"odel's ontological argument \citep{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
 The second contribution of the paper is a new proof (also in natural deduction style).
-In contrast to Scott's proof \cite{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
+In contrast to Scott's proof \citep{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
 The new proof also does not rely on the equality predicate and is much shorter.
@@
 Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
-Nevertheless, it should be noted that its adequacy for the ontological argument has already been contested \cite{fitting}.
+Nevertheless, it should be noted that its adequacy for the ontological argument has already been contested \citep{fitting}.
 Another assumption made here is that the quantification domains are constant (i.e.
@@
 Neither G\"odel's manuscript nor Scott's manuscript reveal whether they use constant or varying domains, and this is also the case for many variants (e.g.
-\cite{hajek}).
-Nevertheless, some authors of variants \cite[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
+\citep{hajek}).
+Nevertheless, some authors of variants \citep[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
 Our choice of assuming constant domains is motivated by simplicity.
@@
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
-From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accesibility relation in the Kripke models \cite{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accesibility relation in the Kripke models \citep{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
 In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
@@
 
-Although the simpler proof of Theorem 1 presented here was developed independently, it seems conceptually related to a possible simplification briefly and informally described by Anderson \cite[footnote 2]{and90} and attributed by him to an anonymous referee.
+Although the simpler proof of Theorem 1 presented here was developed independently, it seems conceptually related to a possible simplification briefly and informally described by Anderson \citep[footnote 2]{and90} and attributed by him to an anonymous referee.
 The simpler proof of Theorem 1 is not particularly special.
@@
 This phenomenon is very intriguing: the new proof of Lemma 1 is technically simpler (because it has fewer inferences), but Scott's proof of Lemma 1 with Theorem 2 can be considered intuitively simpler (because it is arguably easier to understand).
-This constitutes an interesting case for Hilbert's 24th Problem \cite{hilbert}, which asks for criteria to properly compare the simplicity of proofs.
+This constitutes an interesting case for Hilbert's 24th Problem \citep{hilbert}, which asks for criteria to properly compare the simplicity of proofs.
 While Hilbert had mathematical proofs in mind, insights into the 24th problem could be gained by analyzing philosophical proofs as well.
@@
 However, a proof step in their notes clearly relies on {\bf S5}'s \emph{iteration principle}.
-Anderson \cite[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}.
+Anderson \citep[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}.
 Sobel \citep{sobel2}[p.
@@
 This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
-Another solution involving more substantial modifications is that of Bj{\o}rdal \cite{bjordal,fuhrmann}.
+Another solution involving more substantial modifications is that of Bj{\o}rdal \citep{bjordal,fuhrmann}.
 
@@
 
-\bibitem[{\itshape Adams (1995)}]{adams}
+\bibitem[{\itshape Adams(1995)}]{adams}
 Adams, R. M. 1995. Introductory note to [G\"odel 1970].
 
-\bibitem[{\itshape Anderson (1990)}]{and90}
+\bibitem[{\itshape Anderson(1990)}]{and90}
 Anderson, C. A. 1990. {\itshape Some Emendations of G\"odel's Ontological Proof}. Faith and Philosophy, Vol. 7, Issue 3, pp. 291-303.
@@
 
-\bibitem[{\itshape Bj{\o}rdal (1999)}]{bjordal}
-Bj{\o}rdal, F. 1999. 
-Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
-
-\bibitem[{\itshape Church (1940)}]{church}
+\bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
+Bj{\o}rdal, F. 1999. Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
+
+\bibitem[{\itshape Church(1940)}]{church}
 Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56--68.
 
-\bibitem[{\itshape Fitting (2002)}]{fitting}
+\bibitem[{\itshape Fitting(2002)}]{fitting}
 Fitting, M. 2002.  {\itshape Types, Tableaus, and G\"odel's God}, Kluwer Academic Publishers.
 
-\bibitem[{\itshape Fuhrmann (2005)}]{fuhrmann}
+\bibitem[{\itshape Fuhrmann(2005)}]{fuhrmann}
 Fuhrmann, A. 2005.
@@
 
-\bibitem[{\itshape Goedel (1970)}]{Goedel} 
+\bibitem[{\itshape Goedel(1970)}]{Goedel} 
 G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
 
-\bibitem[{\itshape Hajek (1996)}]{hajek}
+\bibitem[{\itshape Hajek(1996)}]{hajek}
 Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
@@
 
-\bibitem[{\itshape Kant (1781)}]{kant}
+\bibitem[{\itshape Kant(1781)}]{kant}
 Kant, I.  original 1781. {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
 
-\bibitem[{\itshape Negri (2005)}]{negri}
+\bibitem[{\itshape Negri(2005)}]{negri}
 Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
 
-\bibitem[{\itshape Prawitz (2006)}]{prawitz}
-Prawitz, D. 2006 (1st ed. 1965). {\itshape Natural deduction: a proof-theoretical study}. Mineola, New York: Dover publications.
-
-\bibitem[{\itshape Scott (2001)}]{scott}
-Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
-
-\bibitem[{\itshape Sobel (1987)}]{sobel}
-Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying: essays for Richard Cartwright},  MIT Press.
-
-\bibitem[{\itshape Sobel (2001)}]{sobel2}
-Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
-
-\bibitem[{\itshape Wang (1996)}]{Wang1996}
-Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
-
-\bibitem[{\itshape Paulin-Mohring (2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
+\bibitem[{\itshape Nipkow \& Paulson \& Wenzel(2002)}]{isabelle} Nipkow, T. \& Paulson, L.C. \& Wenzel, M. 2002. {\itshape Isabelle/HOL: A Proof Assistant for
+Higher-Order Logic}. Number 2283 in LNCS, Springer.
+
+\bibitem[{\itshape Paulin-Mohring(2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
 All about Proofs, Proofs for All,
@@
 
-\bibitem[{\itshape Nipkow \& Paulson \& Wenzel (2002)}]{isabelle} Nipkow, T. \& Paulson, L.C. \& Wenzel, M. 2002. {\itshape Isabelle/HOL: A Proof Assistant for
-Higher-Order Logic}. Number 2283 in LNCS, Springer.
-
-\bibitem[{\itshape Thiele (2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
+\bibitem[{\itshape Prawitz(2006)}]{prawitz}
+Prawitz, D. 2006 (1st ed. 1965). {\itshape Natural deduction: a proof-theoretical study}. Mineola, New York: Dover publications.
+
+\bibitem[{\itshape Scott(2001)}]{scott}
+Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
+
+\bibitem[{\itshape Sobel(1987)}]{sobel}
+Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying: essays for Richard Cartwright},  MIT Press.
+
+\bibitem[{\itshape Sobel(2001)}]{sobel2}
+Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
+
+\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
+
+\bibitem[{\itshape Wang(1996)}]{Wang1996}
+Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
 

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex

@@
 Sec. 68T02 % Artificial Intelligence
+\marginpar{Change?: 03B15, 03B45, 03B35, 68T15}
 }
 
-\keywords{Modal Logics, Higher-Order Logics, Ontological Argument}
+\keywords{Modal Logics, Higher-Order Logics, Ontological Argument,
+  Interactive and Automated Theorem Proving}
 
@@
 
-Nevertheless, G{\"o}del's work still leaves room for criticism. 
-In particular, his axioms are so strong that they entail 
-a \emph{modal collapse} 
-\cite{Sobel1987,sobel2004logic}: 
-everything that is the case is so necessarily.  
-There has been an impressive body of recent and ongoing work
+Nevertheless, G{\"o}del's work still leaves room for criticism.  In
+particular, his axioms are so strong that, when assuming unrestricted
+comprehension principles\footnote{A possible direction to remedy modal
+  collapse, as studied e.g. by Koons
+  \cite{koons06:_sobel_goedel_ontol_proof} is to impose restrictions
+  on the domain of properties.}, 
+\marginpar{added footnote}
+they entail a \emph{modal collapse}
+\cite{Sobel1987,sobel2004logic}: everything that is the case is so
+necessarily.  There has been an impressive body of recent and ongoing
+work
 (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
@@
 collapse from the point of view of the modal square of opposition.
+\marginpar{@Bruno: Shall we add a sentence or two here?}
 
@@
 $\collapse{\phi'}$ may not be valid.
-\end{enumerate}
+\end{enumerate} \marginpar{@Bruno: Shall we respond here to the
+  rewievers criticism?}
 

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex
Address reviewer's criticisms and Christoph's marginpars

@@
 \subjclass{
-Prim. 03A02;  % Philosophical aspects of logic and foundations
-Sec. 68T02 % Artificial Intelligence
-\marginpar{Change?: 03B15, 03B45, 03B35, 68T15}
+Prim. 03B15; % Higher-order logic and type theory
+Sec. 68T15 % Theorem proving (deduction, resolution, etc.)
 }
+% \marginpar{Change?: 
+% 03A02 % Philosophical aspects of logic and foundations
+% 03B45 % Modal logic (including the logic of norms)
+% 03B35 % Mechanization of proofs and logical operations
+% 68T02 % Artificial Intelligence
+% }
 
@@
   on the domain of properties.}, 
-\marginpar{added footnote}
+%\marginpar{added footnote}
 they entail a \emph{modal collapse}
@@
 collapse from the point of view of the modal square of opposition. 
-\marginpar{@Bruno: Shall we add a sentence or two here?}
+Ontological arguments typically rely on an inversion of the normal direction of entailment
+in the modal square of opposition for one particular proposition (i.e. God's existence), 
+and the modal collapse shows that this inversion happens for all propositions, resulting in a total collapse of the modal square of opposition.
+
+%\marginpar{@Bruno: Shall we add a sentence or two here?}
 
@@
 
-Independently of the variant of the ontological argument 
-under consideration, the following can be said about 
-classes of collapsing propositions:
+In any modal logic at least as strong as \textbf{K}, and even without relying on axioms specific to 
+ontological arguments, it is easy to see (and even easier to check with an automated theorem prover) 
+the following facts about classes of collapsing propositions:
 \begin{enumerate}
@@
 $\collapse{\phi'}$ may not be valid.
-\end{enumerate} \marginpar{@Bruno: Shall we respond here to the
-  rewievers criticism?}
+\end{enumerate}
+
+An easy corollary of the second fact above is that any ontological argument relying on Lemma \textbf{L2} will necessarily lead to a modal collapse for all propositions that are logically equivalent to God's existence.
+The third and fourth facts above indicate that to prove that the modal collapse holds for a larger class of propositions would require using axioms specific to the ontological argument.
+
+% \marginpar{@Bruno: Shall we respond here to the
+%  rewievers criticism?}
 
@@
 
+\clearpage
+
 \end{document}

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex
some improvements for final version of modal collapse paper

@@
 properties and states a few reasonable (but debatable) axioms that
-such properties should satisfy.  
-The overall idea of G{\"o}del's proof
+such properties should satisfy.  The overall idea of G{\"o}del's proof
 is in the tradition of Anselm's argument, who defined God as some
-entity of which nothing greater can be conceived. 
-Anselm argued that
-existence in the actual world would make 
-such an assumed being even
-greater (more perfect); 
-hence, by definition, God must exist. However,
-for Anselm existence was treated as a predicate and 
-the possibility of God's existence was assumed as granted. 
-These issues were criticized by Kant and Leibniz, 
-respectively, and they were addressed in the work of G\"odel.
+entity of which nothing greater can be conceived.  Anselm argued that
+existence in the actual world would make such an assumed being even
+greater (more perfect), hence, by definition, God must exist. However,
+for Anselm existence was treated as a predicate and the possibility of
+God's existence was assumed as granted.  These issues were criticized
+by Kant and Leibniz, respectively, and they were addressed in the work
+of G\"odel.
 
@@
 in the modal square of opposition for one particular proposition (i.e. God's existence), 
-and the modal collapse shows that this inversion happens for all propositions, resulting in a total collapse of the modal square of opposition.
+and the modal collapse shows that this inversion in fact occurs for all propositions, resulting in a total collapse of the modal square of opposition.
 
@@
 verification and reconstruction of G\"odel's proof 
-\cite{J28,J30,W50,J29}, and a detailed 
+\cite{J28,J30,J29}, and a detailed 
 mathematical description is available in \cite{C40}.

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex
Make minor language improvements

@@
 Canterbury, Descartes and Leibniz.
-Kurt G{\"o}del and Dana Scott studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a second-order modal logic \cite{GoedelNotes,ScottNotes}, shown in Fig.
+Kurt G{\"o}del and Dana Scott studied and improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a second-order modal logic \cite{GoedelNotes,ScottNotes}, shown in Fig.
 \ref{fig:scott}.
@@
 such properties should satisfy.  The overall idea of G{\"o}del's proof
-is in the tradition of Anselm's argument, who defined God as some
+is in the tradition of Anselm's argument, who defined God as an
 entity of which nothing greater can be conceived.  Anselm argued that
@@
 and the references therein) proposing solutions for the modal
-collapse.  The goal of this contribution is to discuss the modal
+collapse.  The goal of this article is to discuss the modal
 collapse from the point of view of the modal square of opposition. 
@@
 semantics) under which the announced results hold must be 
-explicitly specified in the source code. This reduces the need 
-for interpretation and the danger of misunderstandings. 
+explicitly specified in the source code. Consequently, 
+the danger of misunderstandings is reduced. 
 Current technology is increasingly ready to be embraced by 

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex
Clarify statement about the need for ontological axioms for teh characterization of larger classes modally collapsing propositions.

@@
 
-An easy corollary of the second fact above is that any ontological argument relying on Lemma \textbf{L2} will necessarily lead to a modal collapse for all propositions that are logically equivalent to God's existence.
-The third and fourth facts above indicate that to prove that the modal collapse holds for a larger class of propositions would require using axioms specific to the ontological argument.
-
-% \marginpar{@Bruno: Shall we respond here to the
-%  rewievers criticism?}
+An easy corollary of the second fact above is that any ontological argument relying on Lemmas \textbf{L1} and \textbf{L2} will necessarily lead to a modal collapse for all propositions that are logically equivalent to God's existence.
+The third and fourth facts indicate that characterizations of larger classes of propositions for which the modal collapse holds require using axioms specific to the variant of the ontological argument under consideration, as in the case of \textbf{A:MC}.
 

### 2016-03-26 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix things that were preventing compilation
Update keywords
Change title to plural

@@
-\documentclass [smallextended] {svjour3} [natbib]
+%\documentclass[smallextended]{svjour3}[natbib]
+\documentclass{svjour3}
 
@@
 \usepackage{theorems}
-\usepackage[numbers]{natbib}
+%\usepackage[numbers]{natbib}
 
@@
 
-\title{A Variant of G\"{o}del's Ontological Proof \\
+\title{Variants of G\"{o}del's Ontological Proof \\
 in a Natural Deduction Calculus}
@@
 \noindent
-\textbf{Keywords: } G\"odel's Ontological Proof, Higher-Order Modal Logics, Natural Deduction,
+\textbf{Keywords: } Ontological Argument, Higher-Order Logics, Modal Logics, Natural Deduction.
 

### 2016-03-28 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve abstract and introduction

@@
-%\documentclass[smallextended]{svjour3}[natbib]
-\documentclass{svjour3}
+\documentclass[smallextended]{svjour3}
 
@@
 \usepackage{latexsym}
-\usepackage{natbib}
 
@@
 \usepackage{theorems}
-%\usepackage[numbers]{natbib}
+\usepackage[numbers]{natbib}
 
@@
 \begin{abstract}
-This paper presents two detailed formalizations of ontological arguments in a simple natural deduction calculus.
+This paper presents detailed formalizations of ontological arguments in a simple modal natural deduction calculus.
 The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills in the gaps, thus verifying its correctness.
@@
 The second proof is also technically shorter than the first one, because it eliminates unnecessary detours and uses Axiom 1 for the positivity of properties only once.
+The third and fourth proofs formalize, respectively, Anderson's and Bj{\o}rdal's variants of the ontological argument, which are known to be immune to modal collapse.
 \end{abstract}
@@
 Anselm (1033-1109).
-His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, apart from the other properties, also had the property of existence.
+His argument considers a greatest conceivable being, who must exist, because if it did not have the property of existence, then we could conceive of a greater being that, in addition to other properties, also had the property of existence.
 St.
-Anselm's argument was further elaborated by Descartes, Leibniz and Kant.
+Anselm's argument was criticized and further developed by Descartes, Leibniz and Kant, among many others.
 
@@
 
-The first contribution of this paper is a detailed formalization of Scott's version \citep{scott} of G\"odel's ontological argument \citep{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
-The second contribution of the paper is a new proof (also in natural deduction style).
+The first main contribution of this paper is a detailed formalization of Scott's version \citep{scott} of G\"odel's ontological argument \citep{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
+The second main contribution of the paper is a new proof (also in natural deduction style).
 In contrast to Scott's proof \citep{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
@@
 
-A major criticism against G\"odel's formal argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
+A major criticism against G\"odel's argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
 This is discussed in greater detail in Section \ref{sec:collapse}, where a natural deduction derivation of the collapse is presented.
-Although many recent works on the ontological argument have proposed modifications of the argument that do not entail a modal collapse, these solutions are beyond the scope of this paper.
+Many recent works on the ontological argument have proposed modifications of the argument that do not entail modal collapse.
+In this paper we present natural deduction formalizations of two of these solutions: Anderson's emendation (in Section \ref{sec:anderson}) and Bj{\o}rdal's variant (in Section \ref{sec:bjordal}).
 
@@
 \section{A natural deduction proof based on Anderson's axiomatization}
+\label{sec:anderson}
 
@@
 
-\section{A natural deduction proof based on Bj\o rdal's axiomatization}
+\section{A natural deduction proof based on Bj{\o}rdal's axiomatization}
+\label{sec:bjordal}
 

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Change \psi to \phi
Improve text

@@
 Without the restriction that every box must be accessed by exactly one strong modal inference, the calculus would be unsound for the modal logic \textbf{K}.
-For example, the formula $\all \psi.
-(\nec \psi \imp \pos \psi)$ is not valid in \textbf{K} but would be provable without this restriction:
+For example, the formula $\all \phi.
+(\nec \phi \imp \pos \phi)$ is not valid in \textbf{K} but would be provable without this restriction:
 
@@
   \AXC{$ $} \RightLabel{1}
-  \UIC{$\nec \psi$} \RightLabel{$\nec_E$}
-        \UIC{$\omega: \psi$} \RightLabel{$\pos_I$}
-        \UIC{$\pos \psi$} \RightLabel{$\imp_I^1$}
-    \UIC{$\nec \psi \imp \pos \psi$} \RightLabel{$\forall_I$}
-    \UIC{$\all \psi. (\nec \psi \imp \pos \psi)$}
+  \UIC{$\nec \phi$} \RightLabel{$\nec_E$}
+        \UIC{$\omega: \phi$} \RightLabel{$\pos_I$}
+        \UIC{$\pos \phi$} \RightLabel{$\imp_I^1$}
+    \UIC{$\nec \phi \imp \pos \phi$} \RightLabel{$\forall_I$}
+    \UIC{$\all \phi. (\nec \phi \imp \pos \phi)$}
 \end{prooftree}
@@
 From the point of view of a \emph{possible world semantics}, rigidity means that their interpretation is independent of the world in which they are being interpreted.
-Rigidity is silently assumed by most works investigating the ontological argument, and is explicitly assumed here.
+Rigidity is tacitly assumed by most works investigating the ontological argument, and is explicitly assumed here.
 Nevertheless, it should be noted that its adequacy for the ontological argument has already been contested \citep{fitting}.
@@
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
-From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accesibility relation in the Kripke models \citep{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models \citep{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
 In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
@@
 \emph{Brouwer's reduction principle} is derivable in modal logic {\bf KB}, using axiom B.
-The proof of G\"odel and Scott (Section \ref{sec:scott}) do not make direct use of this principle, but the new proof presented in Section \ref{sec:newproof} does.
+The proofs of G\"odel and Scott (Section \ref{sec:scott}) do not make direct use of this principle, but the new proof presented in Section \ref{sec:newproof} does.
 
@@
 
-Scott's version of G\"odel's proof depends on 5 axioms that circumscribe the notion of \emph{positive} property, with positivity being denoted by the undefined second-order predicate symbol $P$.
-Additionally, 3 definitions are used for defining the notions of \emph{God-like}, \emph{essence} and \emph{necessary existence}.
+Scott's version of G\"odel's proof depends on five axioms that circumscribe the notion of \emph{positive} property, with positivity being denoted by the undefined second-order predicate symbol $P$.
+Additionally, three definitions are used for establishing the notions of \emph{God-like}, \emph{essence} and \emph{necessary existence}.
 Technically, these definitions only abbreviate certain complex formulas.
@@
 
-Beside the 5 axioms and 3 definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
-The use of reflexivity could be avoided, but allows a shorter proof.
+In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
+The use of reflexivity could be avoided, but we chose to use it because it allows the construction of a shorter proof.
 
@@
 
-\section{New Proof}\label{sec:newproof}
+\section{A New Proof}\label{sec:newproof}
 
@@
 This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the equality axiom of reflexivity nor {\bf S5}'s modal iteration principle.
-Instead, it relies only on Brouwer's reduction principle (in the proof of Theorem 3).
+Instead, it relies only on Brouwer's reduction principle (for Theorem 3).
 The new proof is also shorter than Scott's proof.
@@
 
-Although the simpler proof of Theorem 1 presented here was developed independently, it seems conceptually related to a possible simplification briefly and informally described by Anderson \citep[footnote 2]{and90} and attributed by him to an anonymous referee.
+Although the simpler proof of Theorem 1 presented here was developed independently, it seems conceptually related to a possible simplification briefly described by Anderson \citep[footnote 2]{and90} and attributed by him to an anonymous referee.
 The simpler proof of Theorem 1 is not particularly special.
@@
 Modern techniques usually rely on unification to compute or guess good instances for weakly quantified variables, and therefore they would never result in instantiations containing $=$ and $\neq$, which are symbols that do not appear anywhere in the axioms or in the theorems to be proved.
-We may only wonder why G\"odel preferred to use those unusual instantiations.
-It might be an interesting question from a historical point of view, if this suggests that there might have been a change (perhaps evolution) in the way logicians construct formal proofs; or even from a philosophical point of view, if G\"odel's preference was philosophically motivated.
+We may wonder why G\"odel preferred to use those unusual instantiations.
+It might be an interesting question from a historical point of view, as it might suggest that there was an evolution in the way logicians construct formal proofs; or from a philosophical point of view, if G\"odel's preference was philosophically motivated.
 

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Remove duplicated proof
Improve text

@@
 \end{corollary}
-\begin{proof} \hfill
-\begin{prooftree}
-\AXC{Axiom \ref{A3}} \dashedLine
-\UIC{$P(G)$}
-\AXC{Theorem \ref{T1} for G} \dashedLine
-\UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
-\BIC{$\pos \ex x. G(x)$}
-\end{prooftree}
+\begin{proof} The proof of this corollary is identical to the proof of the same corollary shown in Section \ref{sec:scott}.
+% \begin{prooftree}
+% \AXC{Axiom \ref{A3}} \dashedLine
+% \UIC{$P(G)$}
+% \AXC{Theorem \ref{T1} for G} \dashedLine
+% \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
+% \BIC{$\pos \ex x. G(x)$}
+% \end{prooftree}
 \end{proof}
@@
 
-Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle relies only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
+Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle requires only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
 There have been many discussions of what modal axioms are needed for the ontological argument.
@@
 Nevertheless, neither Anderson nor Sobel presented any formal proof relying only on {\bf B}.
-The new formal proof presented in Section \ref{sec:newproof} relies on {\bf KB}, which is weaker than {\bf B}.
+The new formal proof presented in this section relies on {\bf KB}, which is weaker than {\bf B}.
 
@@
 G(x)$).
-This contrasts with G\"odel's proof, where the actual existence is never actually proven, although it can be derived as a trivial corollary of its necessity by applying the modal axiom T.
-
-\section{Modal Collapse} \label{sec:collapse}
+This contrasts with G\"odel's proof, where the actual existence is never actually proven.
+It could, nevertheless, be derived as a trivial corollary by applying the modal axiom T.
+
+\section{Modal Collapse} 
+\label{sec:collapse}
 

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Remove unnecessary statement that A is constant
Reorganize paragraph discussin solutions for the modal collapse

@@
 \UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
-\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
+%\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
 \UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
@@
 Several solutions to the problem of modal collapse have been proposed.
-
-Anderson's solution \citep{and90} modifies the definitions of God-like being and essence, and eliminates half of an axiom.
-This not only avoids the modal collapse, but also makes two of G\"odel's five axioms derivable from the others \citep{hajek} under some implicit additional assumptions \citep{fuhrmann}.
-Another solution involving more substantial modifications is that of Bj{\o}rdal \citep{bjordal,fuhrmann}.
-
-On another track, Fitting has argued that greater care has to be taken with the semantics of higher-order modal logics.
+Fitting, for instance, has argued that greater care has to be taken with the semantics of higher-order modal logics.
 Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional.
 Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
-
-Anderson \citep{and90}[p.
+Koon \citep{koon}, Anderson \citep{and90}[p.
 292] and Sobel \citep{sobel2}[p.
 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
-
-A reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.
+Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next section.
+
+%A reduction from {\bf S5} to {\bf KB} is possible for these variants as well, since the technique we used to replace {\bf S5}'s iteration principle by Brouwer's reduction principle depends only on the structure of the final steps of the proof of Theorem 3, but not on details of the axioms.
 
@@
 \UIC{$\all y.[G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
-\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
+%\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
 \UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve section about Anderson's variant

@@
 
-\section{A natural deduction proof based on Anderson's axiomatization}
+\section{Anderson's Emendation in Natural Deduction}
 \label{sec:anderson}
+
+Anderson's solution \citep{and90} modifies the definitions of God-like being and essence, and eliminates half of Axiom 1 (replacing the equivalence by an implication).
 
@@
 %Anderson's implication instead of equivalence. 
-\begin{axiom}
+\begin{axiom}[Emended]
 \label{A1}
-Either a property or its negation is positive, but not both:
+If the negation of a property is positive, the property is not positive:
 $$
@@
 
-\begin{axiom}
+\begin{axiom}[As in Scott's]
 \label{A2}
@@
 \end{theorem}
-
-The proof is otherwise identical to the proof of theorem \ref{T1} in section \ref{sec:newproof} but the equivalence of axiom \ref{A1} does not need to be reduced to an implication by implication elimination.
-
-\begin{definition}[Anderson's alternative Godlike]
+\begin{proof}
+This proof is omitted, because any of the proofs of Theorem 1 in Sections \ref{sec:scott} and \ref{sec:newproof} would suffice to prove this theorem with only small modifications: the equivalence of axiom \ref{A1} would not need to be reduced to an implication by conjunction elimination.
+\end{proof}
+
+\begin{definition}[Emended]
 \label{D1}
-%A \emph{God-like} being possesses all positive properties:
+A \emph{God-like} being necessarily possesses those and only those properties which are positive:
 $$
@@
 
-\begin{axiom}
+\begin{axiom}[As in Scott's]
 \label{A3}
@@
 \end{corollary}
-
-\begin{axiom}
+\begin{proof}
+This proof is omitted, because it is identical to the proof of the same corollary shown in Sections \ref{sec:scott} and \ref{sec:newproof}.
+\end{proof}
+
+\begin{axiom}[As in Scott's]
 \label{A4}
@@
 
-\begin{definition}[Anderson's alternative essence with Scott's additional conjunct]
+\begin{definition}[Emended]
 \label{D2}
-%sAn \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties:
+An \emph{essence} of an individual is a property possessed by it and necessarily implying those and only those properties necessarily possessed by it:
 $$
@@
 
-%At this point, G\"odel and Scott proceed to prove Theorem 2, which is then used to prove Lemma 1.
-However, a close inspection of Scott's proof of Lemma 1 reveals that Theorem 2 is an unnecessary detour.
-Therefore, we do not include it in the new proof.
-
-\begin{definition}[Anderson's alternative where G\"odel's definition of essence is replaced by def. \ref{D2}]
+\begin{definition}[As in Scott's]
 \label{D3}
@@
 
-\begin{axiom}
+\begin{axiom}[As in Scott's]
 \label{A5}
@@
 \end{axiom}
+
+Because of Anderson's emendations to Axiom 1 and Definition 1, the proof of Lemma 1 has to be modified as well.
+The proof shown below uses the modal axiom T to eliminate the extra box modality introduced by Anderson
+in his emended Definition 1.
 
@@
 \end{lemma}
-
 \begin{proof} \hfill
@@
 \UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\imp_E$}
-\BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$ (reflexive frame)}
-\UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{D3}}
+\BIC{$\nec E(\gamma)$}
+    \AXC{Axiom T} \dashedLine
+    \UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
+\BIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{D3}}
 \UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
@@
 \end{theorem}
-
-\subsection{Why is the proof of modal collapse prevented?}
-
-The alteration of the definition of essence in Anderson's proof prevents the standard proof of modal collapse \ref{Collapse} 
-$$\vdash A \imp \nec A
-$$ 
-by the additional necessitation, which converts the proof of modal collapse into a proof of a tautology.
-$$
-\vdash \nec A \imp \nec A
-$$
-
-%For all propositions $A$, the following \emph{modal collapse} proposition is not provable with Anderson's alternative definitions with the same proof as G\"odel or Scott's:
+\begin{proof}
+This proof is identical to the proof of Theorem 3 shown in Section \ref{sec:newproof}.
+\end{proof}
+
+We know from \citep{hajek} that Anderson's emendation also makes Axioms 4 and 5 derivable from the other axioms and definitions.
+Discussions about the redundancy of these two axioms started with an earlier (and wrong claim) claim by Magari \citep{magari} w.r.t.
+Scott's variant, and the redundancy was the topic of a long controversy \citep{controversy} between H\'ajek and Anderson.
+The proof presented above still uses Axioms 4 and 5, though, since the intention here was to show that Anderson's emendation requires only a minimal modification of the proofs presented in the previous sections.
+
+ToDo: should we try to show here also an alternative proof of T3 using only Anderson's axioms A1, A2 and A3 and definition D1??  I don't know how hard this would be. Hájek1996 could be helpful.
+
+\subsection{A Proof-Theoretical Explanation for the Immunity to Modal Collapse}
+
+Anderson's change in the definition of essence adds an extra box modality, which transforms the standard proof (shown in Section \ref{sec:collapse}) of modal collapse (i.e.
+$\vdash A \imp \nec A$) into a proof of a tautology ($\vdash \nec A \imp \nec A$), as shown below.
 
@@
 
-The exchange of the implication in the G\"odel-Scott definition of essence for an equivalence and the additional  modal necessitation of the property $A$ prevents the implication of a necessary fact from its contingent counterpart as is seen in the second part of the proof above because the hypothetical assumption of the contingent property $A$ has to be exchanged for its necessitation for the proof structure to be preserved.
+The exchange of the implication in Scott's definition of essence for an equivalence and the additional necessitation (box) modality before the property $\psi$ (which is instantiated by $\lambda x.
+A$) prevents the implication of a necessary fact from its contingent counterpart, as seen in the second part of the proof above, because the hypothetical assumption of the contingent property $A$ has to be exchanged for its necessitation $\nec A$ to preserve the proof structure.
 

### 2016-03-30 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve section about Bjordal's variant

@@
 
-\section{A natural deduction proof based on Bj{\o}rdal's axiomatization}
+\section{Bj{\o}rdal's Variant in Natural Deduction}
 \label{sec:bjordal}
 
-In Bj\o rdal's axiomatization godlikeness is taken as primitive, $G_B$.
-
-\begin{definition}
+Bj{\o}rdal's variant \citep{bjordal,fuhrmann} is another solution to the Modal Collapse.
+In striking contrast to other variants, Bj\o rdal's axiomatization takes god-likeness ($G_B$) as a primitive notion, upon which the notion of positive property is defined.
+This variant also significantly reduces the number of necessary axioms.
+Scott's variant of Axioms 2, 3 and 4 become unnecessary to prove the final theorem from Bj{\o}rdal's modified definitions 1, 2 and 3 and axioms 1 and 5.
+
+\setcounter{definition}{0}
+\begin{definition}[Changed]
 \label{B:D1}
-A property is positive iff it is necessarily possessed by every God-like being.
+A property is \emph{positive} iff it is necessarily possessed by every God-like being.
 $$
@@
 
-\begin{definition}
+Bj{\o}rdal refines the notion of essence (which he calls maximal composite) by restricting the attention to positive properties only. He calls this refined notion of essence a maximal composite.
+
+\begin{definition}[Changed]
 \label{B:D2}
-A maximal composite of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
+A \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
 $$MCP(\varphi, x) \equiv (\varphi(x) \wedge P_B (\varphi)) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(\varphi(y) \imp \psi(y)))
@@
 
-\begin{definition}
+\begin{definition}[Changed]
 \label{B:D3}
-Necessary existence of an individual is the necessary exemplification of all its maximal composites.
+\emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
 $$
@@
 
-\begin{axiom}
+\setcounter{axiom}{0}
+\begin{axiom}[Changed]
 \label{A:A1'} If a property is positive, its negation is not positive: 
@@
 
+\setcounter{axiom}{4}
 \begin{axiom}
@@
 
-The proof of the possible instantiation of $G_B$ is adjusted by proving Theorem 1 of section \ref{sec:newproof} directly for $G_B$ and not for properties in general.
-Corollary 1 of section \ref{sec:newproof} follows by proving the positivity of godlikeness as in lemma \ref{B:L1} below.
+The proof of the possible instantiation of $G_B$ is adjusted (as shown below) by proving Theorem 1 of section \ref{sec:newproof} directly for $G_B$ and not for positive properties in general.
 
@@
 \label{T1}
-Positive properties are possibly exemplified:
-$$
-\all \varphi. [P(G_B) \imp \pos \ex x.G_B(x)]
+If God-likeness is a positive property, then it is possibly exemplified:
+$$
+P(G_B) \imp \pos \ex x.G_B(x)
 $$
@@
 \UIC{$P(\neg G_B)$}
-\AXC{Axiom 6***} \dashedLine
+\AXC{Axiom 1} \dashedLine
 \UIC{$\all \varphi. [P(\neg G_B) \imp \neg P(G_B)]$} \RightLabel{$\all_E$}
@@
 \end{prooftree}
-
-\end{proof}
+\end{proof}
+
+ToDo: State corollary and prove it, instead of making a forward reference to lemma L1. Then use the corollary in lemma 1.
+
+Corollary 1 of section \ref{sec:newproof} follows by proving the positivity of godlikeness as in lemma \ref{B:L1} below.
 
@@
 \end{lemma}
-
+\begin{proof}
 \begin{small}
@@
 Anderson, C. A.\& Gettings, M. 1996.  {\itshape G\"odel's ontological proof revisited}. In: edited by Hajek P. {\itshape G\"odel '96},  Springer.
+
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo (2016)}]{controversy}
+Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}. In: {\itshape Logica Universalis} 10 (2016)
 
@@
 
+\bibitem[{\itshape Magari(1988)}]{magari}
+Magari, R. 1988. {\itshape Logica e Teofilia}. In: {\itshape Notizie di Logica} VII.4.
+
 \bibitem[{\itshape Negri(2005)}]{negri}

### 2016-04-01 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Add ToDo
Anderson proof update

@@
 \begin{axiom}[Emended]
-\label{A1}
+\label{A:A1}
 If the negation of a property is positive, the property is not positive:
@@
 \begin{axiom}[As in Scott's]
-\label{A2}
+\label{A:A2}
 A property necessarily implied by a positive property is positive:
@@
 \begin{theorem}
-\label{T1}
+\label{A:T1}
 Positive properties are possibly exemplified:
@@
 \begin{definition}[Emended]
-\label{D1}
+\label{A:D1}
 A \emph{God-like} being necessarily possesses those and only those properties which are positive:
@@
 \begin{axiom}[As in Scott's]
-\label{A3}
+\label{A:A3}
 The property of being god-like is positive:
@@
 \begin{corollary}
-\label{C1}
+\label{A:C1}
 Possibly, a God-like being exists:
@@
 \begin{axiom}[As in Scott's]
-\label{A4}
+\label{A:A4}
 Positive properties are necessarily positive:
@@
 \begin{definition}[Emended]
-\label{D2}
+\label{A:D2}
 An \emph{essence} of an individual is a property possessed by it and necessarily implying those and only those properties necessarily possessed by it:
@@
 \begin{definition}[As in Scott's]
-\label{D3}
+\label{A:D3}
 \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
@@
 \begin{axiom}[As in Scott's]
-\label{A5}
+\label{A:A5}
 Necessary existence is a positive property:
@@
 
+ToDo: Proof of $\star_7$ is missing in the proof of lemma 1 below.
+
 \begin{lemma}
-\label{L1}
+\label{A:L1}
 If there is a God-like being, then there is a God-like being necessarily:
@@
 \AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
@@
 \AXC{$ $} \RightLabel{6}
-\UIC{$G(y) $}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$G(y) $}\dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(y)]$}
@@
 \UIC{$\nec \forall y. (G(y)\imp \rho(y))$}\RightLabel{$\imp_I^5$}
-\UIC{$\nec \rho(\gamma)\imp \nec \forall y. (G(y)\imp \rho(y))$}
+\UIC{$[\subproof_7] \quad \nec \rho(\gamma)\imp \nec \forall y. (G(y)\imp \rho(y))$}
 \end{prooftree}
@@
 
-\begin{scriptsize}
-\begin{prooftree}
-\AXC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
-\AXC{$ $} \RightLabel{6}
-\UIC{$\nec \rho(\gamma)$} \RightLabel{}
-\UIC{$\rho(\gamma)$} \RightLabel{$\nec_E$}
-\AXC{$\subproof_7$} \dashedLine
-\UIC{$\omega': \rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
-\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
-\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
-\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
-\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
+%implication for equivalence in essence: 
+\begin{small}
+\begin{prooftree}
+\AXC{Axiom \ref{A:A3}} 
+\UIC{$P(G)$}
+\AXC{$ $} \RightLabel{7}
+\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\wedge_I$}
+\BIC{$P(G)\wedge \nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\imp_E$}
+\AXC{Axiom \ref{A:A2}} 
+\BIC{$P(\varphi)$}
 \AXC{$\subproof_6$} \dashedLine 
-\UIC{$G(\gamma)$}  
-\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
-\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
-\end{prooftree}
-\end{scriptsize}
-
-%implication for equivalence in essence: 
-\begin{small}
-\begin{prooftree}
-\AXC{$ $} \RightLabel{7}
-\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\nec_E$}
-\UIC{$\omega'':   \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\forall_E$}
-\UIC{$\omega'':   G(\gamma)\imp \varphi(\gamma)$}
-\AXC{$\subproof_6$} \dashedLine
-\UIC{$\omega'':  G(\gamma)$}  \RightLabel{$\imp_E$}
-\BIC{$\omega'':   \varphi(\gamma)$}\RightLabel{$\nec_I$}
-\UIC{$\nec \varphi(\gamma)$}\RightLabel{$\imp_I^7$}
-\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$\subproof_8$} \dashedLine
+\UIC{$  G(\gamma)$}  \RightLabel{D\ref{A:D1}}
+\UIC{$ \all \varphi. [P(\varphi)\biimp \nec \varphi(\gamma)]$}  \RightLabel{$\all_E$}
+\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
+\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\nec_I$}
+\UIC{$[\subproof_8] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\subproof_7$ and $\subproof_8$} \dashedLine
 \UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $}
+\UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)]) $}
 \AXC{$ $} \RightLabel{6}
@@
 \AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
@@
     \UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
-\BIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{D3}}
+\BIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
 \UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
@@
 \begin{theorem}
-\label{T3}
+\label{A:T3}
 Necessarily, there exists a God-like being:
@@
 
-ToDo: should we try to show here also an alternative proof of T3 using only Anderson's axioms A1, A2 and A3 and definition D1??  I don't know how hard this would be. Hájek1996 could be helpful.
+\subsection{A shorter proof without redundant axioms}
+
+\setcounter{axiom}{0}
+\setcounter{lemma}{0}
+\setcounter{theorem}{0}
+\setcounter{corollary}{0}
+\setcounter{definition}{0}
+
+There is an alternative proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as definition \ref{A:D1}. %Hájek1996 could be helpful.
+The proof requires only a modification of the proof of lemma \ref{A:L1}. The rest of the dependencies are as in the previous proof.
+
+Theorem \ref{A:T1} is derived from axiom \ref{A:A1} and axiom \ref{A:A2}.
+Corollary \ref{A:C1} is derived from axiom \ref{A:A3} and theorem \ref{A:T1}.
+The proof of lemma \ref{A:L1} can be modified, so as to be derived from axiom \ref{A:A3} using definition \ref{A:D1}.
+Finally, theorem \ref{A:T3} can be derived from lemma \ref{A:L1} and corollary \ref{A:C1}.
+
+\begin{lemma}
+\label{A2:L1}
+If there is a God-like being, then there is a God-like being necessarily:
+$$
+\ex z. G(z) \imp \nec \ex x. G(x)
+$$
+\end{lemma}
+
+\begin{proof} \hfill
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
+\UIC{$[\subproof_6] \quad G(\gamma)$}
+\end{prooftree}
+
+\begin{scriptsize}
+\begin{prooftree}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$P(G) \biimp \nec G(\gamma)$}
+\AXC{Axiom \ref{A:A3}} \RightLabel{}
+\UIC{$P(G)$} \RightLabel{$\biimp_E$}
+\BIC{$\nec G(\gamma)$}
+\AXC{$ $} \RightLabel{5}
+\UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
+\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^5$}
+\UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
+\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$}  \RightLabel{Axiom K}
+\UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
+\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^1$}
+\UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
+\end{prooftree}
+\end{scriptsize}
+
+\end{proof}
 
@@
 
+ToDo: Fix overfull hboxes
+
 %ToDo: acknowledgements

### 2016-04-01 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Bjordal proof fix
changed labels improves compiling

@@
 \begin{axiom}
-\label{A1}
+\label{N:A1}
 Either a property or its negation is positive, but not both:
@@
 \begin{axiom}
-\label{A2}
+\label{N:A2}
 A property necessarily implied by a positive property is positive:
@@
 \begin{theorem}
-\label{T1}
+\label{N:T1}
 Positive properties are possibly exemplified:
@@
 \begin{definition}
-\label{D1}
+\label{N:D1}
 A \emph{God-like} being possesses all positive properties:
@@
 \begin{axiom}
-\label{A3}
+\label{N:A3}
 The property of being god-like is positive:
@@
 \begin{corollary}
-\label{C1}
+\label{N:C1}
 Possibly, a God-like being exists:
@@
 \begin{axiom}
-\label{A4}
+\label{N:A4}
 Positive properties are necessarily positive:
@@
 \begin{definition}
-\label{D2}
+\label{N:D2}
 An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties:
@@
 \begin{definition}
-\label{D3}
+\label{N:D3}
 \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
@@
 \begin{axiom}
-\label{A5}
+\label{N:A5}
 Necessary existence is a positive property:
@@
 \begin{lemma}
-\label{L1}
+\label{N:L1}
 If there is a God-like being, then there is a God-like being necessarily:
@@
 \AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{N:D1}}
 \UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
@@
 \UIC{$P(\rho)$}
-\AXC{Axiom \ref{A4}} \dashedLine
+\AXC{Axiom \ref{N:A4}} \dashedLine
 \UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
@@
 \AXC{$ $} \RightLabel{4}
-\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{D1}}
+\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{N:D1}}
 \UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
@@
 \UIC{$G(\gamma)$}  
-\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
+\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
 \UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
@@
 \AXC{Axiom \ref{A5}} \dashedLine
-\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
+\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{N:D3}}
 \AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{D1}}
+\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{N:D1}}
 \UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
 \UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
-\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{D3}}
+\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{N:D3}}
 \UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
@@
 \begin{theorem}
-\label{T3}
+\label{N:T3}
 Necessarily, there exists a God-like being:
@@
 \UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
-\AXC{ Lemma \ref{L1} } \dashedLine
+\AXC{ Lemma \ref{N:L1} } \dashedLine
 \UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
@@
 \begin{theorem}
-\label{T1}
+\label{B:T1}
 If God-likeness is a positive property, then it is possibly exemplified:
@@
 \UIC{$ \pos \ex x.G_B(x)$}  \RightLabel{$\imp_I^3$}
-\UIC{$ P(\rho) \imp \pos \ex x.G_B(x) $} \RightLabel{$\all_I$}
-\UIC{$ \all \varphi. [P(G_B) \imp \pos \ex x.G_B(x)] $}
+\UIC{$ P(G_B) \imp \pos \ex x.G_B(x) $}
 

### 2016-04-02 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Bjordal proof corollary added
Anderson proof: axiom 4 added

@@
 
-The aim of this paper is to present two detailed formalizations of ontological arguments in a natural deduction calculus.
+The aim of this paper is to present detailed formalizations of variants of ontological arguments in a natural deduction calculus.
 For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz}.
@@
 133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
-Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next section.
+Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next sections.
 
@@
 
-ToDo: Proof of $\star_7$ is missing in the proof of lemma 1 below.
-
 \begin{lemma}
@@
 \begin{prooftree}
+\AXC{$ $} \RightLabel{5}
+\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
+\UIC{$\omega: P(\rho)$}
+\AXC{$ $} \RightLabel{4}
+\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{A:D1}}
+\UIC{$\omega: \forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\biimp_E$}
+\BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
+\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
+\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
+\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
+\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
+\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
+
 \AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{A:D1}}
+\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
 \UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
-\AXC{$ $} \RightLabel{5}
-\UIC{$\nec \rho(\gamma)$} \RightLabel{$\imp_E$}
-\BIC{$P(\lambda x.\rho(x))$} 
 \AXC{$ $} \RightLabel{6}
-\UIC{$G(y) $}\dottedLine \RightLabel{D\ref{A:D1}}
-\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(y)]$}
-\UIC{$[P(\lambda x.\rho(x)) \biimp \nec \rho(y)]$}
-\BIC{$\nec \rho(y)$}\RightLabel{$\nec_E$}
-\UIC{$\omega':  \rho(y)$}\RightLabel{$\imp_I^6$}
-\UIC{$\omega':  G(y)\imp \rho(y)$}\RightLabel{$\forall_I$}
-\UIC{$\omega':  \forall y. (G(y)\imp \rho(y))$}\RightLabel{$\nec_I$}
-\UIC{$\nec \forall y. (G(y)\imp \rho(y))$}\RightLabel{$\imp_I^5$}
-\UIC{$[\subproof_7] \quad \nec \rho(\gamma)\imp \nec \forall y. (G(y)\imp \rho(y))$}
+\UIC{$\nec\rho(\gamma)$} \RightLabel{}
+\BIC{$P(\rho)$} \RightLabel{$\imp_E$}
+
+\AXC{Axiom \ref{A:A4}} \dashedLine
+\UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
+\UIC{$[\subproof_7] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
 \end{prooftree}
@@
 \UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
-\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\nec_I$}
+\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^7$}
 \UIC{$[\subproof_8] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
@@
 \UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\imp_E$}
+\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
 \BIC{$\nec E(\gamma)$}
@@
 
+\setcounter{lemma}{0}
+\setcounter{theorem}{0}
+\setcounter{corollary}{0}
+
 Bj{\o}rdal's variant \citep{bjordal,fuhrmann} is another solution to the Modal Collapse.
@@
 \UIC{$\omega:  \all x.(G_B(x) \imp \neg G_B(x)) $}  \RightLabel{$\nec_I$}
-\UIC{$ \nec \all x.(G_B(x) \imp \neg G_B(x)) $}\RightLabel{$\wedge_I$}
+\UIC{$ \nec \all x.(G_B(x) \imp \neg G_B(x)) $}\dottedLine  \RightLabel{D\ref{B:D1}}
 \UIC{$P(\neg G_B)$}
@@
 
-ToDo: State corollary and prove it, instead of making a forward reference to lemma L1. Then use the corollary in lemma 1.
-
-Corollary 1 of section \ref{sec:newproof} follows by proving the positivity of godlikeness as in lemma \ref{B:L1} below.
+\begin{corollary}
+\label{B:C1}
+Possibly, a God-like being exists:
+$$
+\pos \ex x. G(x)
+$$
+\end{corollary}
+\begin{proof}
+
+\begin{prooftree}
+\AXC{$ $}\doubleLine
+\UIC{$\nec \all x. [G_B(x) \imp G_B(x)] $}\RightLabel{ D\ref{B:D1}}\dottedLine
+\UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
+\AXC{Theorem \ref{T1}}
+\UIC{$ P_B(G_B) \imp \pos \ex x.G_B(x) $} \RightLabel{ $\imp_E$}
+\BIC{$\pos \ex x.G_B(x) $}
+\end{prooftree}
+
+\end{proof}
 
@@
 \UIC{$G_B(x) $}  \RightLabel{ $\imp_E$}
-\BIC{$\all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$} \RightLabel{ $\forall_E$}
-\UIC{$(MCP(G_B, x) \imp \nec \ex y.G_B(y))]$}
+\BIC{$\all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))$} \RightLabel{ $\forall_E$}
+\UIC{$(MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
 \AXC{$ $}\dashedLine

### 2016-04-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Comment out ToDo.
Add missing citation
Make minor improvements in the text

@@
 
-In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the equality axiom of reflexivity (in the proof of Theorem 1).
+In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the axiom of reflexivity for equality (in the proof of Theorem 1).
 The use of reflexivity could be avoided, but we chose to use it because it allows the construction of a shorter proof.
@@
 In this section we present a new proof in {\bf KB}.
-This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the equality axiom of reflexivity nor {\bf S5}'s modal iteration principle.
+This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the axiom of reflexivity for equality nor {\bf S5}'s modal iteration principle.
 Instead, it relies only on Brouwer's reduction principle (for Theorem 3).
@@
 Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
-Koon \citep{koon}, Anderson \citep{and90}[p.
+Koon \citep{koons}, Anderson \citep{and90}[p.
 292] and Sobel \citep{sobel2}[p.
@@
 \begin{proof}
-This proof is omitted, because any of the proofs of Theorem 1 in Sections \ref{sec:scott} and \ref{sec:newproof} would suffice to prove this theorem with only small modifications: the equivalence of axiom \ref{A1} would not need to be reduced to an implication by conjunction elimination.
+This proof is omitted, because any of the proofs of Theorem 1 in Sections \ref{sec:scott} and \ref{sec:newproof} would suffice to prove this theorem with only a small modification: the equivalence of axiom \ref{A1} would not need to be reduced to an implication by conjunction elimination.
 \end{proof}
@@
 This variant also significantly reduces the number of necessary axioms.
-Scott's variant of Axioms 2, 3 and 4 become unnecessary to prove the final theorem from Bj{\o}rdal's modified definitions 1, 2 and 3 and axioms 1 and 5.
+Scott's variants of Axioms 2, 3 and 4 become unnecessary to prove the final theorem from Bj{\o}rdal's modified definitions 1, 2 and 3 and axioms 1 and 5.
 
@@
 
-Bj{\o}rdal refines the notion of essence (which he calls maximal composite) by restricting the attention to positive properties only. He calls this refined notion of essence a maximal composite.
+Bj{\o}rdal refines the notion of essence by restricting the attention to positive properties only. He calls this refined notion of essence a maximal composite.
 
@@
 
-ToDo: Fix overfull hboxes
+%ToDo: Fix overfull hboxes
 
@@
 
+\bibitem[{\itshape Koons(2006)}]{koons}
+Koons, R. 2006. Sobel on G\"odel's Ontological Proof. {\itshape Philosophia Christi}, 2:235-248.
+
 \bibitem[{\itshape Magari(1988)}]{magari}

### 2016-04-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Add numbers and references to subproofs of Bjordal's variant
Add missing definition folding in Bjordal's proof
Make MCP a new command, in order to avoid spurious spaces in math mode
Improve section about Anderson's proof without superfluous axioms

@@
 
-We know from \citep{hajek} that Anderson's emendation also makes Axioms 4 and 5 derivable from the other axioms and definitions.
+H\'ajek \citep{hajek} claimed that Anderson's emendation makes Axioms 4 and 5 derivable from the other axioms and definitions.
 Discussions about the redundancy of these two axioms started with an earlier (and wrong claim) claim by Magari \citep{magari} w.r.t.
 Scott's variant, and the redundancy was the topic of a long controversy \citep{controversy} between H\'ajek and Anderson.
-The proof presented above still uses Axioms 4 and 5, though, since the intention here was to show that Anderson's emendation requires only a minimal modification of the proofs presented in the previous sections.
-
-\subsection{A shorter proof without redundant axioms}
+As confirmed in Subsection \ref{sec:NonRedundantAnderson}, H\'ajek's claim is correct.
+Nevertheless, the proof presented above still uses Axioms 4 and 5, since this section's intention is to show that Anderson's emendation requires only a minimal modification of the proofs presented in the previous sections.
+
+\subsection{A Shorter Proof without Superfluous Axioms}
+\label{sec:NonRedundantAnderson}
 
@@
 
-There is an alternative proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as definition \ref{A:D1}. %Hájek1996 could be helpful.
-The proof requires only a modification of the proof of lemma \ref{A:L1}. The rest of the dependencies are as in the previous proof.
-
-Theorem \ref{A:T1} is derived from axiom \ref{A:A1} and axiom \ref{A:A2}.
-Corollary \ref{A:C1} is derived from axiom \ref{A:A3} and theorem \ref{A:T1}.
-The proof of lemma \ref{A:L1} can be modified, so as to be derived from axiom \ref{A:A3} using definition \ref{A:D1}.
-Finally, theorem \ref{A:T3} can be derived from lemma \ref{A:L1} and corollary \ref{A:C1}.
+In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify
+the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. Such a modified proof is shown below.
 
@@
 \UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$}  \RightLabel{Axiom K}
+\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$} \doubleLine  \RightLabel{(using Axiom K)}
 \UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
@@
 \end{proof}
+
+\noindent
+Proofs of other lemmas, theorems and corollaries remain the same as in the previous section.
+Theorem \ref{A:T1} is derived from axioms \ref{A:A1} and \ref{A:A2}.
+Corollary \ref{A:C1} is derived from Axiom \ref{A:A3} and Theorem \ref{A:T1}.
+Finally, Theorem \ref{A:T3} can be derived from Corollary \ref{A:C1} and Lemma \ref{A:L1}, which, as shown above, depends only on Axiom \ref{A:A3} and Definition \ref{A:D1}.
+Therefore, confirming H\'ajek's claim, axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3} are superfluous: they are not needed to prove Theorem \ref{A:T3}.
 
@@
 
+\newcommand{\MCP}{\mathit{MCP}}
+
 \begin{definition}[Changed]
@@
 A \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
-$$MCP(\varphi, x) \equiv (\varphi(x) \wedge P_B (\varphi)) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(\varphi(y) \imp \psi(y)))
+$$\MCP(\varphi, x) \equiv (\varphi(x) \wedge P_B (\varphi)) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(\varphi(y) \imp \psi(y)))
 $$
@@
 $$
-NE_B (x) \equiv \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))
+NE_B (x) \equiv \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))
 $$
@@
 \end{lemma}
-\begin{proof}
-\begin{small}
-\begin{prooftree}
-\AXC{$ $}
+\begin{proof}\hfill
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{$2$}
 \UIC{$G_B(x) $} 
@@
 \UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
-\BIC{$G_B(x) \wedge P_B(G_B) $}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ $}\dashedLine
+\BIC{$[\subproof_{10}] \qquad G_B(x) \wedge P_B(G_B) $}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{10} $}\dashedLine
 \UIC{$G_B(x) \wedge P_B(G_B) $}
@@
 \UIC{$\all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y)))$} \RightLabel{ $\wedge_I$}
-\BIC{$MCP(G_B, x) $}
+\BIC{$G_B(x) \wedge P_B(G_B) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y))) $} \RightLabel{ D\ref{B:D2} }\dottedLine
+\UIC{$[\subproof_{11}] \qquad \MCP(G_B, x) $}
 
@@
 \UIC{$\nec \all x. [G_B(x) \imp NE_B(x)]$}\RightLabel{ D\ref{B:D3}}\dottedLine
-\UIC{$\nec \all x. [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\nec_E$}
-\UIC{$\all x. [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\forall_E$}
-\UIC{$ [G_B(x) \imp \all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}
+\UIC{$\nec \all x. [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\nec_E$}
+\UIC{$\all x. [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\forall_E$}
+\UIC{$ [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}
 \AXC{$ $} \RightLabel{ $2$}
 \UIC{$G_B(x) $}  \RightLabel{ $\imp_E$}
-\BIC{$\all\varphi. (MCP(\varphi, x) \imp \nec \ex y.\varphi(y))$} \RightLabel{ $\forall_E$}
-\UIC{$(MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
-\AXC{$ $}\dashedLine
-\UIC{$MCP(G_B, x) $} \RightLabel{ $\imp_E$}
+\BIC{$\all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))$} \RightLabel{ $\forall_E$}
+\UIC{$(\MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
+\AXC{$ \subproof_{11} $}\dashedLine
+\UIC{$\MCP(G_B, x) $} \RightLabel{ $\imp_E$}
 \BIC{$\nec \ex y.G_B(y)$}\RightLabel{ $\imp_I^2$}

### 2016-04-12 Papers/InProgress/IJCAI/submission.tex
Add authors
Address reviewer's comments for IJCAI paper

@@
 \title{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
-\author{Christoph Benzm\"uller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
-\author{}
+\author{
+  \begin{minipage}{0.5\textwidth}\centering
+    Christoph Benzm\"uller\thanks{This work was supported by the 
+                                  German National Research Foundation (DFG) under
+                                  grants BE 2501/9-1 and BE 2501/11-1.} \\
+    Freie Universit\"at Berlin, Germany \\
+    c.benzmueller@fu-berlin.com
+  \end{minipage}
+  \begin{minipage}{0.5\textwidth}\centering
+    Bruno Woltzenlogel Paleo \\
+    Australian National University, Australia \\
+    bruno.wp@gmail.com
+  \end{minipage}
+}
 
@@
   intelligence. Despite the popularity of the argument since the
-  appearance of G\"odel's manuscript in the early 70's, the
+  appearance of G\"odel's manuscript in the early 1970's, the
   inconsistency of the axioms used in the argument remained unnoticed
@@
 body of publications -- including very recent ones -- present,
-discuss, assess, criticise, modify and improve G\"{o}del's original
+discuss, assess, criticize, modify and improve G\"{o}del's original
 work (see e.g.~Sobel~\shortcite{sobel2004logic} and Oppy~\shortcite{sep-ontological-arguments} and the
@@
 been conducted with automated theorem provers for classical
-higher-order logic (HOL; cf. \cite{andrewsSEP} and the references
+higher-order logic (HOL; cf.~\cite{andrewsSEP} and the references
 therein), even though G\"odel's proof is actually formulated in
-higher-order \emph{modal} logic (HOML; cf. \cite{homl} and the
+higher-order \emph{modal} logic (HOML; cf.~\cite{homl} and the
 references therein). To bridge between the two logics we utilise and
@@
 First successful applications of theorem proving technology in
-metaphysics were reported by Oppenheimer and
-Zalta~\shortcite{oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
+metaphysics were reported by Fitelson, Oppenheimer and
+Zalta~\shortcite{FitelsonZalta,oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
 \textsc{Prover9} \cite{prover9-mace4} in their experiments. Later on, Rushby~\shortcite{rushby13} used the proof assistant \textsc{PVS} \cite{cade92-pvs}. Common to both
@@
 arguably improved Anselm's argument. Of particular importance to
-G\"odel was the work of Leibniz. 
-%\cite{ToDo:Leibniz}.  
+G\"odel was the work of Leibniz~\cite{Adams}.  
 Although G\"odel's notion of positive property is not exactly the same as Leibniz's notion of perfection, G\"odel's manuscript
@@
 Hazen [\citeyear[p.365]{Hazen}] states that ``G\"odel left this
-clause out in [11], but this appears to have been an oversight''.
+clause out [\ldots] but this appears to have been an oversight''.
 % -- it is
@@
 \end{itemize}
-Instead, we consider here a modal logic that we call \SFiveU, which is
-characterized by the following condition on $r$:
+Instead, we consider here an alternative description, that we call \SFiveU, based on the following condition on $r$:
 \begin{itemize}
@@
 
-It is easy to see that \SFiveU is at least as strong as \SFive:
-universality entails that $r$ is an equivalence relation. \SFiveU is,
-in fact, \emph{strictly stronger} than \SFive. \SFiveU only admits
+It is important to note that $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$ 
+\cite{Blackburn}, and therefore \SFive and \SFiveU are traditionally considered to be 
+two different descriptions of the same modal logic. Nevertheless, \SFive and \SFiveU 
+differ in the shapes of frames they admit:
+%It is easy to see that \SFiveU is at least as strong as \SFive:
+%universality entails that $r$ is an equivalence relation. \SFiveU is,
+%in fact, \emph{strictly stronger} than \SFive. 
+\SFiveU only admits
 complete\footnote{A graph is \emph{complete} iff there is a directed
@@
 
-For \SFiveU, an improved embedding is possible. Universality implies
+Furthermore, for \SFiveU an improved embedding is possible. Universality implies
 that the guarding predicates in the definitions of $\Box$ and
@@
 The new embedding of \SFiveU in Isabelle/HOL is shown in
-Fig.~\ref{QML_S5U}. It is important to note that we 
-have  $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$,
-despite the differences in the underlying frame structures.
-
-%
-% \marginpar{@Bruno: I commented you paragraph here, since I am not
-%   convinced (yet) that e.g. BF and CBF are equally valid in \SFiveU
-%   and \SFive}
-% \SFiveU is only slightly stronger than \SFive. Most importantly,
-% $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi^{U}$, where
-% $\varphi^U$ is obtained from $\varphi$ by replacing all $\Box$ and
-% $\Dia$ by, respectively, $\BlackBox$ and $\BlackDia$. Therefore,
-% \SFiveU can be considered as adequate for metaphysics as \SFive
-% (cf. \cite[p. 127]{WilliamsonModalLogicAsMetaphysics}). 
-% % \cite{Mattey}, \cite[ToDo]{DunnHardegree}
-%
+Fig.~\ref{QML_S5U}.
 With this improved embedding, the final theorem T3
@@
 Scott's consistent version of the axioms fully automatically.  The fully
-automatic proof has been generated (in about 2,5 seconds) by the theorem prover
+automatic proof has been generated (in about 2.5 seconds) by the theorem prover
 \textsc{Leo-II}~\cite{C26} and subsequently verified in the proof
@@
 \begin{enumerate}
-\item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they didn't need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
+\item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they did not need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
 \item Quantifiers appear as higher-order defined constants, and not as binders.
@@
 actually give up on the proof attempt --- even though \textsc{Leo-II} already
-reported success to Sledgehammer after 2,5 seconds.
+reported success to Sledgehammer after 2.5 seconds.
 
@@
 
-%\section*{Acknowledgments} Will be added at a later point.
-\paragraph*{Acknowledgments} Will be added at a later point.
+%\paragraph*{Acknowledgments} Will be added at a later point. ToDo
 

### 2016-04-12 Papers/InProgress/IJCAI/submission.tex
some minor changes for the final version

@@
     Freie Universit\"at Berlin, Germany \\
-    c.benzmueller@fu-berlin.com
+    \& Stanford University, CSLI, USA \\
+    c.benzmueller@gmail.com
   \end{minipage}
@@
 \S\ref{sec:improvedembedding}) is the implementation of an alternative
-embedding for the slightly stronger modal logic \SFiveU, which is
+embedding for the 
+% slightly stronger 
+more effective modal logic \SFiveU, which is
 based on a universal accessibility relation. Our experiments have
@@
 automatic proof has been generated (in about 2.5 seconds) by the theorem prover
-\textsc{Leo-II}~\cite{C26} and subsequently verified in the proof
+\textsc{Leo-II}~\cite{leo2} and subsequently verified in the proof
 assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5U}.
@@
 \begin{figure*}
+% \begin{minipage}{.01\textwidth}\small
+% \vskip2cm * 
+% \end{minipage}
+% \begin{minipage}{.99\textwidth}
+% {\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
+% \end{minipage}
 \centerline{\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
@@
   substitutions (e.g. with the empty property) are highlighted. In the
-red part, property variable $\texttt{SV8}$ has been instantiated with the
+red part (see $\boldsymbol{\leftarrow}$), property variable $\texttt{SV8}$ has been instantiated with the
 $\lambda \texttt{SV16}_{\mu}. \lambda \texttt{SV17}_{\iota}. \bot$, i.e., the (lifted) empty property.
@@
 %
-Note that the terms for the empty property\footnote{An additional lambda abstraction occurs in the empty property in LEO-II's proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.} ($\lambda x.
+Note that the terms for the empty property\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.} ($\lambda x.
 \bot$) and for the property of self-difference ($\lambda x.
@@
 metaphysics and to conduct reasoning steps at granularity-levels
-beyond common human capabilities.  % ToDo: or even Gödel-level capabilities :-)
+beyond common human capabilities.  
+\vfill
 
-% Good improvement (technical sense), Sledgehammer can still be improved
-% (\textsc{Leo-II} finds inconsistency when call directly in THF Syntax, but not
-% from within Isabelle).
+\noindent\textbf{Acknowledgments:} We thank Chad Brown, who contributed to the rational  reconstruction
+of the inconsistency argument.
 
-% We also point to several technical challenges which, at least for the
-% moment, still restrain non-expert users of these systems from
-% independently conducting similar experiments in computational
-% metaphysics as reported here and in previous work.
-
-% Technical issues detected within Isabelle-Sledgehammer; room for improvements
-
-%\paragraph*{Acknowledgments} Will be added at a later point. ToDo
+\pagebreak
+%A discussion between Chad Brown and Benzm\"uller
+% significantly influenced the rational reconstruction of the inconsistency reported by \textsc{LEO-II}.
 
@@
 \bibliographystyle{named}
-\bibliography{Bibliography}
+%\bibliography{Bibliography}
+
+\begin{thebibliography}{}
+
+\bibitem[\protect\citeauthoryear{Adams}{1995}]{Adams}
+R.M. Adams.
+\newblock Introductory note to *1970.
+\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpubl. Essays and
+  Letters}}. Oxford Univ. Press, 1995.
+
+\bibitem[\protect\citeauthoryear{Anderson and
+  Gettings}{1996}]{AndersonGettings}
+A.C. Anderson and M.~Gettings.
+\newblock {G\"odel} ontological proof revisited.
+\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
+  Science, and Physics: Lecture Notes in Logic 6}}, pages 167--172. {Springer},
+  1996.
+
+\bibitem[\protect\citeauthoryear{Anderson}{1990}]{Anderson}
+C.A. Anderson.
+\newblock Some emendations of {G{\"o}del's} ontological proof.
+\newblock {\em Faith and Philosophy}, 7(3), 1990.
+
+\bibitem[\protect\citeauthoryear{Andrews}{2014}]{andrewsSEP}
+P.B. Andrews.
+\newblock Church's type theory.
+\newblock In E.N. Zalta, editor, {\em The Stanford Encyclopedia of Philosophy}.
+  Spring 2014 edition, 2014.
+
+\bibitem[\protect\citeauthoryear{Anselm}{1078}]{Proslogion}
+St. Anselm.
+\newblock Proslogion.
+\newblock In M.~Charlesworth, editor, {\em St.~Anselm's Proslogion}.
+  Oxford:OUP, 1078.
+\newblock Republished in 1965.
+
+\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and Paulson}{2013}]{J23}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Quantified multimodal logics in simple type theory.
+\newblock {\em Logica Universalis}, 7(1):7--20, 2013.
+
+\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and
+  Woltzenlogel-Paleo}{2013a}]{J30}
+C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo.
+\newblock Formalization, mechanization and automation of {G{\"o}del's} proof of
+  {God's} existence.
+\newblock {\em arXiv:1308.4526}, 2013.
+\newblock Preprint available as arXiv:1308.4526.
+
+\bibitem[\protect\citeauthoryear{Benzm\"uller and
+  Woltzenlogel-Paleo}{2013b}]{J28}
+C.~Benzm\"uller and B.~Woltzenlogel-Paleo.
+\newblock {G{\"o}del's God in Isabelle/HOL}.
+\newblock {\em Archive of Formal Proofs}, 2013, 2013.
+
+\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and
+  Woltzenlogel-Paleo}{2014}]{C40}
+C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo.
+\newblock Automating {G\"{o}del's} ontological proof of {God}'s existence with
+  higher-order automated theorem provers.
+\newblock In 
+  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
+  Applications}, pages 93 -- 98. IOS Press, 2014.
+
+\bibitem[\protect\citeauthoryear{Benzm{\"u}ller \bgroup \em et al.\egroup
+  }{2015}]{leo2}
+Christoph Benzm{\"u}ller, Lawrence~C. Paulson, Nik Sultana, and Frank
+  Thei{\ss}.
+\newblock The higher-order prover {LEO-II}.
+\newblock {\em J. of Automated Reasoning}, 55(4):389--404, 2015.
+
+\bibitem[\protect\citeauthoryear{Bj{\o}rdal}{1999}]{Bjordal}
+F.~Bj{\o}rdal.
+\newblock Understanding {G\"{o}del’s} ontological argument.
+\newblock In T.~Childers, editor, {\em The Logica Yearbook 1998}. Filosofia,
+  1999.
+
+\bibitem[\protect\citeauthoryear{Blackburn \bgroup \em et al.\egroup
+  }{2001}]{Blackburn}
+P.~Blackburn, M.~de~Rijke, and Y.~Venema.
+\newblock {\em Modal Logic}.
+\newblock Cambridge University Press, 2001.
+
+\bibitem[\protect\citeauthoryear{Blanchette and Nipkow}{2010}]{Nitpick}
+J.C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for higher-order logic based on a
+  relational model finder.
+\newblock In {\em ITP 2010}, number 6172 in LNCS, pages 131--146. Springer,
+  2010.
+
+\bibitem[\protect\citeauthoryear{Blanchette \bgroup \em et al.\egroup
+  }{2013}]{Sledgehammer}
+J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+\newblock Extending {Sledgehammer} with {SMT} solvers.
+\newblock {\em J. of Automated Reasoning}, 51(1):109--128, 2013.
+
+\bibitem[\protect\citeauthoryear{Brown}{2012}]{Satallax}
+C.E. Brown.
+\newblock Satallax: An automated higher-order prover.
+\newblock In {\em IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117. Springer,
+  2012.
+
+\bibitem[\protect\citeauthoryear{Fitelson and Zalta}{2007}]{FitelsonZalta}
+Branden Fitelson and Edward~N. Zalta.
+\newblock Steps toward a computational metaphysics.
+\newblock {\em J. Philosophical Logic}, 36(2):227--247, 2007.
+
+\bibitem[\protect\citeauthoryear{Fuhrmann}{2016}]{fuhrmann15:_blogg_goedel}
+A.~Fuhrmann.
+\newblock Blogging {G\"odel}: His ontological argument in the public eye.
+\newblock In K.~\'{S}wi{e}torzecka, editor, {\em Forthcoming in
+  \textit{G\"odel’s Ontological Argument - History, Modifications, and
+  Controversies}}. 2016.
+
+\bibitem[\protect\citeauthoryear{G\"odel}{1970}]{GoedelNotes}
+K.~G\"odel.
+\newblock {\em Appx. A: Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock In Sobel \shortcite{sobel2004logic}, 1970.
+
+\bibitem[\protect\citeauthoryear{H\'ajek}{1996}]{Hajek1}
+P.~H\'ajek.
+\newblock Magari and others on {G\"odel’s} ontological proof.
+\newblock In A.~Ursini and P.~Agliano, editors, {\em Logic and algebra}, page
+  125–135. Dekker, New York etc., 1996.
+
+\bibitem[\protect\citeauthoryear{H\'ajek}{2001}]{Hajek2}
+P.~H\'ajek.
+\newblock {Der Mathematiker und die Frage der Existenz Gottes}.
+\newblock In B.~Buldt~et al., editor, {\em {Kurt G{\"o}del. Wahrheit und
+  Beweisbarkeit}}, pages 325--336. öbv \& hpt, Wien, 2001.
+\newblock ISBN 3-209-03835-X.
+
+\bibitem[\protect\citeauthoryear{H{\'{a}}jek}{2002}]{Hajek3}
+P.~H{\'{a}}jek.
+\newblock A new small emendation of {G{\"{o}}del's} ontological proof.
+\newblock {\em Studia Logica}, 71(2):149--164, 2002.
+
+\bibitem[\protect\citeauthoryear{Hazen}{1998}]{Hazen}
+A.P. Hazen.
+\newblock On {G\"odel's} ontological proof.
+\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+
+\bibitem[\protect\citeauthoryear{McCune}{2010}]{prover9-mace4}
+W.~McCune.
+\newblock {Prover9} and {Mace4} (2005--2010).
+\newblock {\small \verb|http://www.cs.unm.edu/~mccune/prover9/|}, 2010.
+
+\bibitem[\protect\citeauthoryear{Muskens}{2006}]{homl}
+R.~Muskens.
+\newblock {Higher Order Modal Logic}.
+\newblock In P.~Blackburn~et al., editor, {\em Handbook of Modal Logic},
+  Studies in Logic and Practical Reasoning, pages 621--653. Elsevier,
+  Dordrecht, 2006.
+
+\bibitem[\protect\citeauthoryear{Nipkow \bgroup \em et al.\egroup
+  }{2002}]{NPW02}
+T.~Nipkow, L.~Paulson, and M.~Wenzel.
+\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
+\newblock Number 2283 in LNCS. Springer, 2002.
+
+\bibitem[\protect\citeauthoryear{Ohlbach}{1991}]{DBLP:journals/logcom/Ohlbach91}
+H.J. Ohlbach.
+\newblock Semantics-based translation methods for modal logics.
+\newblock {\em J. Log. Comput.}, 1(5):691--746, 1991.
+
+\bibitem[\protect\citeauthoryear{Oppenheimer and Zalta}{2011}]{oppenheimer11}
+P.E. Oppenheimer and E.N. Zalta.
+\newblock A computationally-discovered simplification of the ontological
+  argument.
+\newblock {\em Australasian J. of Philosophy}, 89(2):333--349, 2011.
+
+\bibitem[\protect\citeauthoryear{Oppy}{1996}]{oppy96:_goedel_ontol_argum}
+G.~Oppy.
+\newblock G{\"o}delian ontological arguments.
+\newblock {\em Analysis}, 56(4):226--230, 1996.
+
+\bibitem[\protect\citeauthoryear{Oppy}{2000}]{oppy00:_respon_gettin}
+G.~Oppy.
+\newblock Response to {Gettings}.
+\newblock {\em Analysis}, 60(4):363--367, 2000.
+
+\bibitem[\protect\citeauthoryear{Oppy}{2008}]{oppy08:_higher_order_ontol_argum}
+G.~Oppy.
+\newblock Higher-order ontological arguments.
+\newblock {\em Philosophy Compass}, 3(5):1066--1078, 2008.
+
+\bibitem[\protect\citeauthoryear{Oppy}{2015}]{sep-ontological-arguments}
+G.~Oppy.
+\newblock Ontological arguments.
+\newblock In Edward~N. Zalta, editor, {\em The Stanford Encyclopedia of
+  Philosophy}. Spring 2015 edition, 2015.
+
+\bibitem[\protect\citeauthoryear{Owre \bgroup \em et al.\egroup
+  }{1992}]{cade92-pvs}
+{S.} Owre, {J.}~{M.} Rushby, and {N.} Shankar.
+\newblock {PVS:} {A} prototype verification system.
+\newblock In Deepak Kapur, editor, {\em CADE}, volume 607 of {\em LNAI}, pages
+  748--752, Saratoga, {NY}, jun 1992. Springer.
+
+\bibitem[\protect\citeauthoryear{Rushby}{2013}]{rushby13}
+J.~Rushby.
+\newblock The ontological argument in {PVS}.
+\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St.
+  Petersburg, Russia, 2013.
+
+\bibitem[\protect\citeauthoryear{Scott}{1972}]{ScottNotes}
+D.~Scott.
+\newblock {\em Appx. B: Notes in Dana Scott's Hand}, pages 145--146.
+\newblock In Sobel \shortcite{sobel2004logic}, 1972.
+
+\bibitem[\protect\citeauthoryear{Sobel}{1987}]{Sobel}
+J.H. Sobel.
+\newblock G\"odel's ontological proof.
+\newblock In {\em {On Being and Saying. Essays for Richard Cartwright}}, pages
+  241--261. {MIT Press}, 1987.
+
+\bibitem[\protect\citeauthoryear{Sobel}{2004}]{sobel2004logic}
+J.H. Sobel.
+\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
+\newblock Cambridge U. Press, 2004.
+
+\bibitem[\protect\citeauthoryear{Sultana and Benzm{\"u}ller}{2013}]{W47}
+N.~Sultana and C.~Benzm{\"u}ller.
+\newblock Understanding {LEO-II's} proofs.
+\newblock In Konstantin Korovin, Stephan Schulz, and Eugenia Ternovska,
+  editors, {\em IWIL 2012}, volume~22 of {\em EPiC Series}, pages 33--52,
+  Merida, Venezuela, 2013. EasyChair.
+
+\bibitem[\protect\citeauthoryear{Sutcliffe and Benzm{\"u}ller}{2010}]{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller.
+\newblock Automated reasoning in higher-order logic using the {TPTP THF}
+  infrastructure.
+\newblock {\em J. of Formalized Reasoning}, 3(1):1--27, 2010.
+
+\bibitem[\protect\citeauthoryear{Williamson}{2013}]{williamson13}
+T.~Williamson.
+\newblock {\em Modal Logic as Metaphysics}.
+\newblock Oxford:OUP, 2013.
+
+\end{thebibliography}
 

### 2016-04-12 Papers/InProgress/IJCAI/submission.tex
uodate of affiliations

@@
                                   grants BE 2501/9-1 and BE 2501/11-1.} \\
-    Freie Universit\"at Berlin, Germany \\
-    \& Stanford University, CSLI, USA \\
+    \normalfont Freie Universit\"at Berlin \& Stanford University\\
     c.benzmueller@gmail.com
@@
     Bruno Woltzenlogel Paleo \\
-    Australian National University, Australia \\
+    \normalfont Australian National University\\
     bruno.wp@gmail.com

### 2016-04-13 Papers/InProgress/IJCAI/main.tex
moved file

--- 2016-04-12 Papers/InProgress/IJCAI/submission.tex
@@
                                   German National Research Foundation (DFG) under
-                                  grants BE 2501/9-1 and BE 2501/11-1.} \\
+                                  grants BE 2501/9-2 and BE 2501/11-1.} \\
     \normalfont Freie Universit\"at Berlin \& Stanford University\\

### 2016-04-17 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
fix of overfull hboxes
Gabbay reference added
Remove completed ToDo
Fix subproof numbers
conclusion section

@@
 \usepackage{theorems}
-\usepackage[numbers]{natbib}
+%\usepackage[numbers]{natbib}
+\usepackage{natbib}
 
@@
 The aim of this paper is to present detailed formalizations of variants of ontological arguments in a natural deduction calculus.
-For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz}.
+For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz} or \citep{gabbay}.
 The higher-order modal natural deduction calculus proposed here has special introduction and elimination rules for modalities, as defined in Section \ref{sec:calculus}, and is sound and complete relative to an axiomatic modal calculus, as proven in Section \ref{sec:compl-sound}.
@@
 \infer[\bot_E]{A}{ \bot }
-\s
+\s\s
 \infer[\imp_I]{A \imp B}{ B }
-\s
+\s\s
 \infer[\imp_I^n]{A \imp B}{ \infer*{B}{\infer[n]{A}{}} }
-\s
+\s\s
 \infer[\imp_E]{B}{A & A \imp B}
@@
 
-\s
+\s\s
 \infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] }
-\s
+\s\s
 \infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }
@@
 \infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
-\s
+\s\s
 \infer[\ex_E]{A[\beta]}{ \ex x_{\tau}. A[x] }
@@
 
-\s\s\s\s
+\s\s\s\s\s\s
 \infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
@@
 
-\s\s\s\s
+\s\s\s\s\s\s
 \infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
@@
 \end{prooftree}
-
+\end{small}
+
+\begin{small}
 \begin{prooftree}
@@
 \end{prooftree}
-
+\end{small}
 \end{proof}
@@
 
-\begin{small}
+\begin{scriptsize}
 \begin{prooftree}
@@
 	   \end{prooftree}
-\end{small}
-
-\begin{small}
+\end{scriptsize}
+
+\begin{scriptsize}
 \begin{prooftree}
@@
 \end{prooftree}
-\end{small}
+\end{scriptsize}
 
@@
 
+\begin{scriptsize}
 \begin{prooftree}
@@
 \end{prooftree}
-
+\end{scriptsize}
+
+\begin{small}
 \begin{prooftree}
@@
 
-\begin{scriptsize}
+\begin{tiny}
 \begin{prooftree}
@@
 \end{prooftree}
-\end{scriptsize}
+\end{tiny}
 
@@
 
-\begin{scriptsize}
+\begin{prooftree}
+\AXC{$\subproof_6$} \dashedLine
+\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
+\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
+\AXC{$ $} \RightLabel{6}
+\UIC{$\nec\rho(\gamma)$} \RightLabel{}
+\BIC{$P(\rho)$} \RightLabel{$\imp_E$}
+
+\AXC{Axiom \ref{A:A4}} \dashedLine
+\UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$[\subproof_7] \quad \nec P(\rho)$} \RightLabel{$\imp_E$}
+\end{prooftree}
+
 \begin{prooftree}
@@
 
-\AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
-\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
-\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
-\AXC{$ $} \RightLabel{6}
-\UIC{$\nec\rho(\gamma)$} \RightLabel{}
-\BIC{$P(\rho)$} \RightLabel{$\imp_E$}
-
-\AXC{Axiom \ref{A:A4}} \dashedLine
-\UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
-\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+\AXC{$\subproof_7$} \dashedLine
+\UIC{$\nec P(\rho)$}
 \BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
-\UIC{$[\subproof_7] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
-\end{prooftree}
-\end{scriptsize}
+\UIC{$[\subproof_8] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
+\end{prooftree}
 
 %implication for equivalence in essence: 
-\begin{small}
+%\begin{small}
 \begin{prooftree}
@@
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^7$}
-\UIC{$[\subproof_8] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$\subproof_7$ and $\subproof_8$} \dashedLine
+\UIC{$[\subproof_9] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
+\end{prooftree}
+%\end{small}
+
+\begin{tiny}
+\begin{prooftree}
+\AXC{$\subproof_8$ and $\subproof_9$} \dashedLine
 \UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $}
@@
 \end{prooftree}
-\end{small}
+\end{tiny}
 \end{proof}
@@
 \UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
-\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_9 $}\dashedLine
+\UIC{$[\subproof_{10}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{10} $}\dashedLine
 \UIC{$\ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
@@
 \UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
-\BIC{$[\subproof_{10}] \qquad G_B(x) \wedge P_B(G_B) $}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_{10} $}\dashedLine
+\BIC{$[\subproof_{11}] \qquad G_B(x) \wedge P_B(G_B) $}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{11} $}\dashedLine
 \UIC{$G_B(x) \wedge P_B(G_B) $}
@@
 \BIC{$G_B(x) \wedge P_B(G_B) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y))) $} \RightLabel{ D\ref{B:D2} }\dottedLine
-\UIC{$[\subproof_{11}] \qquad \MCP(G_B, x) $}
+\UIC{$[\subproof_{12}] \qquad \MCP(G_B, x) $}
 
@@
 \UIC{$(\MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
-\AXC{$ \subproof_{11} $}\dashedLine
+\AXC{$ \subproof_{12} $}\dashedLine
 \UIC{$\MCP(G_B, x) $} \RightLabel{ $\imp_E$}
@@
 
-%ToDo: Fix overfull hboxes
-
 %ToDo: acknowledgements
+
+\section{Conclusions}
+
+The proof of section \ref{sec:newproof} shows explicitly that the weak modal system {\bf KB} is sufficient for the ontological argument as opposed to the standard modal calculus {\bf S5}.
+This weakening of the system is based on an explicit formalization and confirms the result briefly described by Anderson \citep[footnote 2]{and90}.
+The weakening also applies to the emendation and variant of Anderson and Bj\o rdal respectively.
+The proofs are short and accessible to the general reader because of the choice of a natural deduction calculus.
+In addition to weakening the modal system the proof of section \ref{sec:newproof} shows that the equality relation is not needed to prove the main theorem.
+In the second proof of section \ref{sec:anderson} the axioms 4 and 5 as well as definitions 2 and 3 are explicitly shown to be redundant in Anderson's emendation.
 
@@
 
-\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo (2016)}]{controversy}
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{controversy}
 Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}. In: {\itshape Logica Universalis} 10 (2016)
@@
 
+\bibitem[{\itshape Gabbay(1996)}]{gabbay}
+Gabbay, D. 1996. {\itshape Labelled Deductive Systems: volume I}. New York: Oxford Science publications.
+
 \bibitem[{\itshape Goedel(1970)}]{Goedel} 

### 2016-06-14 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Revert "conclusion section"
This reverts commit 97ce5bd05b1cd541a4fbbb8b14bded7627d879d0.

@@
 The proofs are short and accessible to the general reader because of the choice of a natural deduction calculus.
-In addition to weakening the modal system the proof of section \ref{sec:newproof} shows that the equality relation is not needed to prove the main theorem.
-In the second proof of section \ref{sec:anderson} the axioms 4 and 5 as well as definitions 2 and 3 are explicitly shown to be redundant in Anderson's emendation.
 

### 2016-06-14 Papers/2016/IJCAI-Poster/poster.tex

\documentclass[final,hyperref={pdfpagelabels=false},xcolor=dvipsnames]{beamer}
\mode<presentation>
  {
  %  \usetheme{Berlin}
  \usetheme{FUBerlin}
  \usecolortheme{orchid}
%\usetheme{TUGraz}
  }
  \usepackage{times}
  \usepackage{amsmath,amsthm, amssymb, latexsym, fancyvrb}
  \boldmath
  \usepackage[english]{babel}
  \usepackage[latin1]{inputenc}
  \usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}

  \usepackage{stmaryrd}
  \newcommand{\typearrow}{\shortrightarrow}
  \newcommand{\itype}{\iota}
  \newcommand{\utype}{u}
  \newcommand{\proptype}{\tau}
  \newcommand{\exInW}{eiw}
  \newcommand{\chriscite}[1]{\textcolor{darkgray}{\textit{\scriptsize Reading:
        #1}}}
  \newcommand{\qcl}{QCL}

  \newcommand{\holpp}{\textcolor{brown}{H}}
  \newcommand{\holp}{\textcolor{brown}{$\holpp$}}
  \newcommand{\holpLeo}{\textcolor{brown}{$\holpp_{L}$}}
  \newcommand{\holpSatallax}{\textcolor{brown}{$\holpp_{S}$}}
  \newcommand{\holpIsabelle}{\textcolor{brown}{$\holpp_{I}$}}
  \newcommand{\holpNitpick}{\textcolor{brown}{$\holpp_{N}$}}
  \newcommand{\holpAgsyhol}{\textcolor{brown}{$\holpp_{A}$}}
  \newcommand{\holpLSA}{\textcolor{brown}{$\holpp_{A,L,S}$}}
  \newcommand{\holpLS}{\textcolor{brown}{$\holpp_{L,S}$}}
  \newcommand{\holpLA}{\textcolor{brown}{$\holpp_{A,L}$}}
  \newcommand{\holpSA}{\textcolor{brown}{$\holpp_{A,S}$}}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
  \graphicspath{{figures/}}
  \title[Automating Quantified Conditional Logics in HOL]{Automating Quantified Conditional Logics in HOL}
  \author[Benzmüller]{Christoph Benzmüller}
%  \institute[Freie Universität Berlin]{Freie Universität Berlin,   Germany}
  %\date{Jul. 31th, 2007}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
  \begin{document}
  \begin{frame}{} 
  \vskip-3ex
    \begin{columns}[t]
      \begin{column}{.32\linewidth}
        \begin{block}{\large Abstract} \footnotesize
          A notion of \textbf{quantified conditional logics (QCLs)} is
          provided that includes quantification over
          \textbf{individual and propositional variables}. The former
          is supported with respect to \textbf{constant and variable
            domain semantics}. In addition, a \textbf{sound and
            complete embedding of this framework in classical
            higher-order logic (HOL)} is presented. Using prominent examples
          from the literature it is demonstrated how this embedding
          enables \textbf{effective automation of reasoning} within
          (object-level) and about (meta-level) quantified conditional
          logics with \textbf{off-the-shelf higher-order theorem
            provers and model finders}.
        \end{block}
        \begin{block}{\large Overall Motivation and Contribution}
          \footnotesize QCLs are very expressive non-classical logics;
          they have many applications; no provers have been
          available so far.  However,
          \begin{itemize}
           \item \textbf{\textcolor{blue}{QCLs are fragments of HOL (with Henkin semantics)}}
           \item \textbf{\textcolor{blue}{and they can easily be automated as such,}}
           \item \textbf{\textcolor{blue}{they inherit important meta- resp. proof-theoretical
             properties (cut-elimination, compactness, etc.), and}}
           \item \textbf{\textcolor{blue}{they can easily be combined with other logics in 
             HOL.}}
           \end{itemize}
           {This reserach is part of a larger project which
             takes HOL as starting point for studying classical and
             non-classical logics and their combinations.} \\[-1.5ex] \ \hfill
 \chriscite{[Benzm\"uller'13]}
        \vskip.1ex
        \end{block}
      \end{column}
      \hfill
      \begin{column}{.32\linewidth}
        \begin{block}{\large Quantified Conditional Logics (QCLs)}
          \footnotesize  
          \colorbox{GreenYellow}{
            \begin{minipage}{.93\textwidth} 
              \begin{align*}
                \textcolor{red}{\varphi},\textcolor{red}{\psi} ::= \quad & \textcolor{red}{P} \mid \textcolor{red}{k(X^1,\ldots,X^n)} \mid \textcolor{red}{\neg \varphi} \mid \textcolor{red}{\varphi \vee \psi} \mid \textcolor{red}{\varphi \Rightarrow \psi} \mid \\
                & \textcolor{red}{\forall^{co} X \varphi} \mid \textcolor{red}{\forall^{va} X \varphi} \mid  \textcolor{red}{\forall P \varphi}  \\[-1em]
              \end{align*}
            \end{minipage}
          }
          \vskip1.5ex
          \textbf{Interpretation}: $M = \langle S,f,D,D',Q,I
          \rangle$ where $S$ is a set of 'worlds',
          $f: S \times 2^{S} \mapsto 2^{S}$ is the selection function,
          $D\not=\emptyset$ is a set of \emph{individuals} (constant
          domain), $D'$ is a function that assigns a
          subset $D'(w)\not=\emptyset$ of $D$ to each world $w$
          (varying domains), $Q\not=\emptyset$ is a collection of subsets
          of $W$ (prop.~domain), and $I$ is an interpretation function
          s.t. for each predicate symbol $k$, $I(k,w) \subseteq
          D^n$. \\
          \vskip1ex
          \textbf{Satisfiability} of $\textcolor{red}{\varphi}$ (denoted as $M,g,s
          \models \textcolor{red}{\varphi}$) for an interpretation $M$, a world $s \in
          S$, and a variable assignment $g = (g^{i},g^{p})$:
          \begin{align*}
            M,g,s \models\;& \textcolor{red}{k(X^1,...\,,X^n)} \text{\;iff\;} \langle g^{i}(\textcolor{red}{X^1}),...\,, g^{i}(\textcolor{red}{X^n}) \rangle \in I(k,w)\\
            M,g,s \models\;& \textcolor{red}{P}  \text{\;iff\;} s \in g^{p}(\textcolor{red}{P})
            \\
            M,g,s \models\;& \textcolor{red}{\neg \varphi} \text{\;iff\;} M,g,s \not\models \textcolor{red}{\varphi} \, (\text{that is, not } M,g,s \models \textcolor{red}{\varphi})
            \\
            M,g,s \models\;& \textcolor{red}{\varphi \vee \psi} \text{\;iff\;} M,g,s \models \textcolor{red}{\varphi} \text{ or } M,g,s \models \textcolor{red}{\psi}
            \\
            M,g,s \models\;& \textcolor{red}{\forall^{co} X \varphi} \text{\;iff\;} M,([d/X]g^{i},g^{p}), s \models \textcolor{red}{\varphi} \text{ for all } d \in D
            \\
            M,g,s \models\;& \textcolor{red}{\forall^{va} X \varphi} \text{\;iff\;} M,([d/X]g^{i},g^{p}), s \models \textcolor{red}{\varphi} \text{ for all } d \in D'(s)
            \\
            M,g,s \models\;& \textcolor{red}{\forall P \varphi} \text{\;iff\;} M,(g^{i},[p/P]g^{p}),s \models \textcolor{red}{\varphi} \text{ for all } p \in Q
            \\
            M,g,s \models\;& \textcolor{red}{\varphi \Rightarrow \psi} \text{\;iff\;}
            M,g,t \models \textcolor{red}{\psi} \text{ for all } t \in S \text{ s.t.}  t \in f(s,[\textcolor{red}{\varphi}]) \\ & \text{ where } [\textcolor{red}{\varphi}] = \{u \mid M,g,u \models \textcolor{red}{\varphi}\}
          \end{align*}
          $M \models^{\qcl} \textcolor{red}{\varphi}$ iff $M,g,s \models \textcolor{red}{\varphi}$ for all $s$, $g$. $\models \textcolor{red}{\varphi}$
          iff $M \models^{\qcl} \textcolor{red}{\varphi}$ for all
          $M$.\\
          \vskip1ex
          \chriscite{[Stalnaker'68],[Delgrande'98]}
        \end{block}
      \end{column}
      \hfill
      \begin{column}{.32\linewidth}
        \begin{block}{\large Classical Higher-order Logic (HOL)}
          \footnotesize

          Types \hfill ${\alpha,\beta} ::= {\iota}
          \textcolor{gray}{\,(worlds)}\mid
          {\mu} \textcolor{gray}{\,(indiv.)} \mid {o}
          \textcolor{gray}{\,(Booleans)}\mid {\alpha \rightarrow
            \beta}$ \\[1em]
          \colorbox{GreenYellow}{
          \begin{minipage}{.93\textwidth} 
          \begin{align*}
            \textcolor{blue}{s},\textcolor{blue}{t} ::=  & \textcolor{blue}{c_{\alpha}} \mid \textcolor{blue}{X_{\alpha}} \mid \textcolor{blue}{(\lambda X_{\alpha}s_{\beta})_{\alpha \typearrow \beta}} \mid 
            \textcolor{blue}{(s_{\alpha \typearrow \beta}\, t_\alpha)_{\beta}} \mid \\
            & \textcolor{blue}{(\neg_{o \typearrow o}\;s_{o})_o} \mid \textcolor{blue}{(s_o \vee_{o\typearrow o \typearrow o} t_o)_o}
            \mid \textcolor{blue}{(\Pi_{(\alpha \typearrow o)\typearrow o}\; s_{\alpha
              \typearrow o})_o}  \\[-1em]
          \end{align*}
          Note:~Binder~notation~\textcolor{blue}{$\forall X_\alpha
            t_o$}~as~syntactic~sugar~for~\textcolor{blue}{$\Pi_{(\alpha
              \typearrow o)\typearrow o} \lambda X_\alpha t_o$}\, 
        \end{minipage}} \vskip2.5ex \textbf{Frame}: collection
      $\{D_{\alpha}\}_{\alpha \in {T}}$ s.t. $D_{o} = \{T,F\}$,
      $D_\itype \not= \emptyset$ and $D_\utype \not= \emptyset$
      arbitrary, and $D_{\alpha \typearrow \beta}$ are collections of
      total functions from
      $D_{\alpha}$ to $D_{\beta}$.\\
      \vskip1.5ex \textbf{Interpretation}: Tuple $\langle
      \{D_{\alpha}\}_{\alpha \in {T}}, I \rangle$ where
      $\{D_{\alpha}\}_{\alpha \in {T}}$ is a frame and where function
      $I$ maps each typed constant symbol $\textcolor{blue}{c_{\alpha}}$ to an
      appropriate element of $D_{\alpha}$, called the
      \emph{denotation} of $c_{\alpha}$. The denotations of
      $\textcolor{blue}{\neg},\textcolor{blue}{\vee}$ and $\textcolor{blue}{\Pi_{(\alpha \typearrow o)\typearrow o}}$ are
      always chosen as usual.\\
      \vskip1.5ex 
      An interpretation is a \textbf{Henkin model} iff there
      is a valuation function ${V}$ s.t.  ${V}(\phi,\textcolor{blue}{s_{\alpha}}) \in
      D_{\alpha}$ for each variable assignment $\phi$ and term
      $\textcolor{blue}{s_{\alpha}}$, and the following conditions are satisfied:
      ${V}(\phi,\textcolor{blue}{X_{\alpha}}) = \phi(\textcolor{blue}{X_\alpha})$, ${V}(\phi,\textcolor{blue}{c_{\alpha}}) =
      I(\textcolor{blue}{c_{\alpha}})$, ${V}(\phi,\textcolor{blue}{l_{\alpha \typearrow \beta}\,
      r_{\alpha}}) = ({V}(\phi,\textcolor{blue}{l_{\alpha \typearrow \beta}})
      {V}(\phi,\textcolor{blue}{r_{\alpha}}))$, and ${V}(\phi,\textcolor{blue}{\lambda
      X_{\alpha}s_{\beta}})$ represents the function from $D_{\alpha}$
      into $D_{\beta}$ whose value for each argument $z \in
      D_{\alpha}$ is ${V}(\phi[z/\textcolor{blue}{X_{\alpha}}], \textcolor{blue}{s_\beta})$. % , where
%       $\phi[z/X_\alpha]$ is that assignment
%       s.t. $\phi[z/X_\alpha](X_\alpha) = z$ and
%       $\phi[z/X_\alpha]Y_\beta = \phi Y_\beta$ when $Y_\beta
%       \not=X_\alpha$.
      If an interpretation is an {Henkin model} the function ${V}$ is
      uniquely determined. \\
      \vskip1ex
      ${H}\models^{HOL} \textcolor{blue}{s}$ iff ${V}(\phi,\textcolor{blue}{s}) = T$ for all $\phi$.
      $\models \textcolor{blue}{s}$ iff ${H}\models^{HOL} \textcolor{blue}{s}$ for all
      ${H}$. \\
      \vskip2.7ex
        \chriscite{[Church'40],[Andrews'72a/b],[Benzm\"ullerEtAl'04]
          \vskip.5ex
        }
        \end{block}
      \end{column}
    \end{columns}
    \vskip2ex
    \begin{block}{\large Embedding QCLs in HOL --- In other words: QCLs are simple Fragments of HOL!} \footnotesize
      \begin{columns}[t]
        \begin{column}{.31\textwidth}
          The mapping $\textcolor{blue}{\lfloor \textcolor{red}{\cdot} \rfloor}$ identifies
          \qcl\ formulas $\textcolor{red}{\varphi}$ with HOL terms $\textcolor{blue}{\lfloor
          \textcolor{red}{\varphi} \rfloor}$ of type $\proptype:=\itype \typearrow o$. The mapping is recursively defined:
          \[
          \begin{array}{lcl}
            \textcolor{blue}{\lfloor \textcolor{red}{P} \rfloor} &=& \textcolor{blue}{P_\proptype} \\
            \textcolor{blue}{\lfloor \textcolor{red}{k(X^1,\ldots,X^n)} \rfloor} % &=& \lfloor k \rfloor \lfloor X^1 \rfloor \ldots \lfloor X^n \rfloor \\
            &=& \textcolor{blue}{k_{u^n \typearrow \proptype}\,X^1_\utype \ldots X^n_\utype}\\
            \textcolor{blue}{\lfloor \textcolor{red}{\neg \varphi} \rfloor} &=&  \textcolor{blue}{\neg_{\proptype\typearrow \proptype}}\,\lfloor \textcolor{red}{\varphi} \rfloor \\
            \textcolor{blue}{\lfloor \textcolor{red}{\varphi\vee \psi} \rfloor} &=&  \textcolor{blue}{\vee_{\proptype \typearrow \proptype \typearrow \proptype}}\,\lfloor\textcolor{red}{\varphi} \rfloor \lfloor \textcolor{red}{\psi} \rfloor\\
            \textcolor{blue}{\lfloor \textcolor{red}{\varphi\Rightarrow \psi} \rfloor} &=&  \textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}}\,\lfloor\textcolor{red}{\varphi} \rfloor \lfloor \textcolor{red}{\psi} \rfloor \\
            \textcolor{blue}{\lfloor \textcolor{red}{\forall^{co} X \varphi} \rfloor} &=&   \textcolor{blue}{\Pi^{co}_{(\utype\typearrow\proptype)\typearrow\proptype}\,\lambda X_\utype} \lfloor\textcolor{red}{\varphi} \rfloor \\
            \textcolor{blue}{\lfloor \textcolor{red}{\forall^{va} X \varphi} \rfloor} &=&  \textcolor{blue}{\Pi^{va}_{(\utype\typearrow\proptype)\typearrow\proptype}\,\lambda X_\utype} \lfloor\textcolor{red}{\varphi} \rfloor \\
            % \end{array}\]
            % \[
            % \begin{array}{lcl}
            \textcolor{blue}{\lfloor \textcolor{red}{\forall P \varphi} \rfloor} &=&  \textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}\,\lambda P_\proptype} \lfloor\textcolor{red}{\varphi} \rfloor
          \end{array}
          \]
          $\textcolor{blue}{P_\proptype}$ and $\textcolor{blue}{X^1_\utype},\ldots,\textcolor{blue}{X^n_\utype}$ are variables and
          $\textcolor{blue}{k_{u^n \typearrow \proptype}}$ is a 
          constant symbol.  
        \end{column}
        \hfill
        \begin{column}{.31\textwidth}
          $\textcolor{blue}{\neg_{\proptype\typearrow \proptype}}$, $\textcolor{blue}{\vee_{\proptype \typearrow
            \proptype \typearrow \proptype}}$, $\textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}}$, $\textcolor{blue}{\Pi^{co,va}_{(\utype\typearrow\proptype)\typearrow\proptype}}$
          and $\textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}}$ realize the \qcl\ connectives in
          HOL. They abbreviate the following HOL terms:% \footnote{Note the predicate argument $A$ of $f$ in the term for $\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}$ and the second-order quantifier $\forall P_{\proptype}$ in the term for $\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}$. FOL encodings of both constructs, if feasible, will be less natural.}
          \[
          \begin{array}{ll}
            \textcolor{blue}{\neg_{\proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda X_\itype \neg(A\,X)}\\
             \textcolor{blue}{\vee_{\proptype \typearrow \proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda B_\proptype \lambda X_\itype (A\,X \vee B\,X)}\\
             \textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda B_\proptype \lambda X_\itype \forall V_\itype (f\, X\, A\, V \rightarrow B\, V)}\\
             \textcolor{blue}{\Pi^{co}_{(\utype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {Q_{\utype\typearrow\proptype}}  \lambda {V_\itype}  \forall{X_\utype}  (Q\,X\,V)} \\
             \textcolor{blue}{\Pi^{va}_{(\utype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {Q_{\utype\typearrow\proptype}}  \lambda {V_\itype}  \forall{X_\utype}  (\exInW\,V\,X \rightarrow Q\,X\,V)} \\
             \textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {R_{\proptype\typearrow\proptype}}  \lambda {V_\itype}  \forall {P_\proptype}  (R\,P\,V)}
          \end{array}
          \]
          % Constant symbol $f$ in the mapping of $\Rightarrow$ is of type ${i
%             \typearrow \proptype \typearrow \proptype}$. It realizes the
%           selection function. Constant symbol $\exInW$ (for 'exists in world'),
%           which is of type ${\itype\typearrow\utype\typearrow o}$,
%           is associated with the varying domains. 
          The interpretations of $\textcolor{blue}{f}$ and $\textcolor{blue}{\exInW}$ are chosen
          appropriately. %  below,
%           cf. Def.~\ref{def:hm}. Moreover, f
          For the varying domains
          non-emptiness is postulated:
          % \begin{align*}
             $\textcolor{blue}{\forall W_\itype \exists X_\utype (\exInW\,W\,X)}$ \\[1em]
          % \end{align*}
          Meta-level notion of validity defined as $
          \textcolor{blue}{\text{vld}_{\proptype\typearrow o}} = \textcolor{blue}{\lambda A_\proptype
          \forall S_\itype (A\, S)}$. \\
        \end{column}
        \hfill
        \begin{column}{.31\textwidth}
          \vskip-2ex
          \colorbox{GreenYellow}{
            \begin{minipage}{.96\textwidth}
              \vskip1ex
              \textbf{Theorem: Soundness and Completeness}
                $$\models^{\qcl}
                \textcolor{red}{\varphi} \,\, \text{ iff } \,\, \{\text{NE}\}\models^{HOL} \textcolor{blue}{\text{vld}_{\proptype\typearrow o}}\, \textcolor{blue}{\lfloor \textcolor{red}{\varphi}
                \rfloor} \text{\,\,(wrt Henkin semantics)}$$
                (Proof: By relating Kripke structures to Henkin models.)
                \vskip1ex
            \end{minipage}
          } 
          \vskip2ex
          \colorbox{GreenYellow}{
            \begin{minipage}{.96\textwidth}
              \vskip1ex
              \textbf{Corollary: Cut-elimination for QCL}
              $$\text{There are cut-free calculi for QCL.}$$ 
              (Proof: Take any cut-free calculus for HOL, e.g. the cut-free sequent calculus from [Benzm\"ullerEtAl'09]. Note, however, the potential impact of cut-simulation.) 
                \vskip2ex
            \end{minipage}        
          } 
          \vskip2ex
          \chriscite{Earlier work is reported in [Benzm.Genovese'11]}   
        \end{column}
      \end{columns}
      % \vskip2ex      \chriscite{[Benzm.Genovese'11]}
    \end{block}
    \vskip0ex
    \begin{columns}[t]
      \begin{column}{.27\textwidth}
        \begin{block}{\large The Encoding in THF0-Syntax} 
          \vskip1.2ex 
          \begin{minipage}{.92\textwidth}
               \VerbatimInput[frame=single,%
               commandchars=\\\{\},%
               fontfamily=courier,fontseries=b,%
               fontsize=\tiny,%
               rulecolor=\color{green},%
               fillcolor=\color{YellowGreen},%
               formatcom=\color{blue},%
               framerule=0pt,%
               framesep=0pt,numbers=right]%
               {Axioms_annotated.ax}
%            \textcolor{blue}{\bf\VerbatimInput[fontsize=\tiny]{Axioms.ax}}
          \end{minipage}
          \vskip2ex
          \chriscite{Introduction to THF0-Syntax [SutcliffeBenzm.'10]}
        \end{block}
      \end{column}
      \hfill
      \begin{column}{.70\textwidth}
        \begin{block}{\large Automating Prominent Examples from the Literature (in \qcl+ID+MP)}
          \begin{columns}[t]
            \begin{column}{.30\textwidth} \footnotesize
              \textbf{Example: Pegasus, the winged horse} \\[1ex]

              It can be consistently stated (in \qcl+ID+MP) that: \textcolor{green}{\textit{``Horses (h) contingently do not
              have wings (w) but Pegasus (p) is a winged horse.''}}
              \textcolor{red}{\begin{align*}
                & \textcolor{red}{\forall^{va} X (h(X) \rightarrow \neg w(X)),\quad h(p),\quad w(p)}
              \end{align*}}
              THF0 encoding of this example:
              \vskip1ex
              \begin{minipage}{.92\textwidth}
                \VerbatimInput[frame=single,%
                commandchars=\\\{\},%
                fontfamily=courier,fontseries=b,%
                fontsize=\tiny,%
                rulecolor=\color{green},%
                fillcolor=\color{YellowGreen},%
                formatcom=\color{blue},%
                framerule=0pt,%
                framesep=0pt,numbers=right]%
                {E1_annotated.thf}
                \end{minipage}
              %\textcolor{blue}{\bf \VerbatimInput[fontsize=\tiny]{E1.thf}}
              \vskip1ex
              \holp\ confirms the satisfiability of these formulas (with
              \holpNitpick=7.7). The finite model generated by Nitpick tells us that
              Pegasus is not 'actual', i.e., does not exist (cf. $\textcolor{blue}{eiw}$) in any world. As expected, when the example problem is formulated
              with $\textcolor{red}{\forall^{co}}$ instead of\, $\textcolor{red}{\forall^{va}}$ then \holp\ reports
              unsatisfiability (\holpLS=0.0, \holpIsabelle=5.8).\\
              %\chriscite{These examples have been discussed (but not automated!) in [Delgrande'98]}
            \end{column}
            \begin{column}{.61\textwidth} \footnotesize
              \vskip-3.6ex

              \, \hfill Notation: \textcolor{red}{$\phi \Rightarrow_X \psi$} := \textcolor{red}{$(\exists^{va} X \phi) \Rightarrow \forall^{va} X
              (\phi \rightarrow \psi)$} \,
              \vskip1ex

              \textbf{Example: Opus, the penguin} \\[1ex]

                 \textcolor{green}{\textit{``Birds (b) normally fly (f), but Opus (o) is a bird that normally does not fly.''}}
                \begin{align*}
                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad b(o), \quad  b(o) \Rightarrow \neg f(o)}
                \end{align*}
                \holp\ reports a finite model (\holpNitpick=8.6). When $\textcolor{red}{\forall^{co}}$ is used:  \holp\  says unsatisfiable (\holpSatallax=0.0, \holpIsabelle=7.9).\\[.5em]

                 \textcolor{green}{\textit{``Birds normally fly and 
                necessarily Opus the bird does not fly.''}}
                \begin{align*}
                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad \Box(b(o) \wedge \neg f(o))}
                \end{align*}
                \holp\ reports a finite model (\holpNitpick=8.7).  When $\textcolor{red}{\forall^{co}}$ is used:
                \holp\  says  unsatisfiable
                (\holpSatallax=0.0, \holpIsabelle=7.6).\\[.5em]

                \textcolor{green}{\textit{``Birds normally fly and necessarily there is a non-flying bird.''}}
                \begin{align*}
                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad \Box\exists^{va}(b(X) \wedge \neg f(X))}
                \end{align*}
                \holp\ reports unsatisfiability  (\holpSatallax=0.0, \holpIsabelle=8.7), also when $\textcolor{red}{\forall^{co}}$ is used (\holpSatallax=0.0, \holpIsabelle=8.8).\\[.5em]

                \textcolor{green}{\textit{``Birds normally fly, penguins normally do not fly and that all penguins are necessarily birds.''}}
                \begin{align*}
                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))} 
                \end{align*}
                \holp\ generates a finite model ($\textcolor{red}{\forall^{va}}$: \holpNitpick=8.8; $\textcolor{red}{\forall^{co}}$: \holpNitpick=7.9).\\[.5em]

 Moreover, \holp\ can conclude from the
                statements above  that \textcolor{green}{\textit{``Birds are normally
                not penguins.''}} ($\textcolor{red}{\forall^{va}}$: \holpSatallax=0.9, \holpLeo=10.2, \holpAgsyhol=9.4; $\textcolor{red}{\forall^{co}}$: \holpSatallax=0.8, \holpLeo=10.1, \holpAgsyhol=0.3):
                \begin{align*}
                  \textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))  \quad \vdash \quad b(X) \Rightarrow_X \neg p(X)}
                \end{align*}
                In line with Delgrande, \holp\ reports a
                countermodel for the following statement (\holpNitpick=8.7):
                \begin{align*}
                  \textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))  \quad \vdash \quad  b(o) \Rightarrow \neg p(o)}
                \end{align*}
                However, when $\textcolor{red}{\forall^{co}}$ is used, \holp\ reports a  theorem (\holpSatallax=0.8, \holpAgsyhol=0.4).
 \vskip2ex
        \chriscite{These examples have been discussed (but not automated) in [Delgrande'98]}
            \end{column}
          \end{columns}         
        \end{block}
      \end{column}
    \end{columns}
    \vskip2ex
    \begin{columns}[t]
      \begin{column}{.42\linewidth}
        \begin{block}{The HOL Metaprover $H$
          %  \holp=(\holpLeo,\holpSatallax,\holpIsabelle,\holpNitpick,\holpAgsyhol)
          }
          \footnotesize
          The \holp\ metaprover for HOL sequentially calls the following prover and model finders: \\[.5em]
          \quad\textbf{\holpLeo} LEO-II (Benzm\"uller/Sultana/Theiss): \url{http://www.leoprover.org} \\
          \quad\textbf{\holpSatallax} Satallax (Brown): \url{http://www.ps.uni-saarland.de/~cebrown/satallax/}\\
          \quad\textbf{\holpIsabelle} Isabelle (Blanchette/Paulson/Nipkow/\ldots): \url{http://isabelle.in.tum.de/} \\
          \quad\textbf{\holpNitpick} Nitpick (Blanchette): \url{http://www4.in.tum.de/~blanchet/nitpick.html} \\
          \quad\textbf{\holpAgsyhol} agsyHol (Lindblatt): \url{https://github.com/frelindb/agsyHOL}\\[.6em]
          These systems support THF0 syntax. These provers are remotely
          available via \\ SystemOnTPTP:
          \url{http://www.cs.miami.edu/~tptp/cgi-bin/SystemOnTPTP}
        \end{block}
      \end{column}
      \hfill
      \begin{column}{.56\linewidth}
        \begin{block}{References and Further Reading} \scriptsize 
         \vskip1ex
          \begin{tabular}{ll}
            [Andrews'72a] & P.B. Andrews.
            General models, descriptions, and choice in type theory.
            JSL, 37(2):385-394, 1972. \\

            [Andrews'72b] & P.B. Andrews.
            General models and extensionality.
            JSL, 37(2):395--397, 1972. \\

            [Benzm\"uller'13] & C. Benzmüller. A top-down approach to 
            combining logics, Proc. of ICAART 2013, Barcelona, Spain, 2013. \\

            [Benzm\"ullerEtAl'04] & C. Benzmüller, C. E. Brown, and M. 
            Kohlhase, Higher order semantics and extensionality. JSL, 69(4):1027-1088, 2004. \\

            [Benzm\"ullerEtAl'09] & C. Benzm{\"u}ller, C. E. Brown, and M. Kohlhase.
            Cut-simulation and impredicativity. LMCS, 5(1:6):1--21, 2009. \\

            [Benzm.Genovese'11] & C.~Benzm{\"u}ller and V.~Genovese.
            Quantified conditional logics are fragments of {HOL}. NCMPL 2011. arXiv:1204.5920v1\\

            [Church'40] & A.~Church. A formulation of the simple theory of types.  
            JSL, 5:56--68, 1940. \\

            [Delgrande'98] & J.P. Delgrande. On first-order conditional logics.

            Artificial Intelligence, 105(1-2):105--137, 1998. \\

 %           [Henkin'50] & \\

            [Stalnaker'68] & R.C. Stalnaker.
            A theory of conditionals.
            In Studies in Logical Theory, pp. 98--112. Blackwell, 1968.\\

            [SutcliffeBenzm.'10] & G.~Sutcliffe and C.~Benzm. 
            Automated reasoning in HOL using the {TPTP THF}
            infrastructure.
            {\em JFR}, 3(1):1--27, 2010. \\[.9em]
           \end{tabular}
        \end{block}
      \end{column}
    \end{columns}

 \end{frame}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2016-06-21 Papers/2016/IJCAI-Poster/poster.tex
added/update

@@
   \graphicspath{{figures/}}
-  \title[Automating Quantified Conditional Logics in HOL]{Automating Quantified Conditional Logics in HOL}
-  \author[Benzmüller]{Christoph Benzmüller}
-%  \institute[Freie Universität Berlin]{Freie Universität Berlin,   Germany}
+  \title[Automating Quantified Conditional Logics in HOL]{The
+    Inconsistency in \\ G{\"o}del's Ontological Argument \\
+--- A Success Story for AI in Metaphysics ---}
+  \author[]{Christoph Benzm{\"u}ller$^1$ and Bruno Woltzenlogel Paleo}
+  \institute[]{{\footnotesize $^1$Supported by the German Research Foundation   (grant BE2501/12-1,2)}}
   %\date{Jul. 31th, 2007}
@@
     \begin{columns}[t]
-      \begin{column}{.32\linewidth}
-        \begin{block}{\large Abstract} \footnotesize
-          A notion of \textbf{quantified conditional logics (QCLs)} is
-          provided that includes quantification over
-          \textbf{individual and propositional variables}. The former
-          is supported with respect to \textbf{constant and variable
-            domain semantics}. In addition, a \textbf{sound and
-            complete embedding of this framework in classical
-            higher-order logic (HOL)} is presented. Using prominent examples
-          from the literature it is demonstrated how this embedding
-          enables \textbf{effective automation of reasoning} within
-          (object-level) and about (meta-level) quantified conditional
-          logics with \textbf{off-the-shelf higher-order theorem
-            provers and model finders}.
-        \end{block}
-        \begin{block}{\large Overall Motivation and Contribution}
-          \footnotesize QCLs are very expressive non-classical logics;
-          they have many applications; no provers have been
-          available so far.  However,
-          \begin{itemize}
-           \item \textbf{\textcolor{blue}{QCLs are fragments of HOL (with Henkin semantics)}}
-           \item \textbf{\textcolor{blue}{and they can easily be automated as such,}}
-           \item \textbf{\textcolor{blue}{they inherit important meta- resp. proof-theoretical
-             properties (cut-elimination, compactness, etc.), and}}
-           \item \textbf{\textcolor{blue}{they can easily be combined with other logics in 
-             HOL.}}
-           \end{itemize}
-           {This reserach is part of a larger project which
-             takes HOL as starting point for studying classical and
-             non-classical logics and their combinations.} \\[-1.5ex] \ \hfill
- \chriscite{[Benzm\"uller'13]}
-        \vskip.1ex
+      \begin{column}{.49\linewidth}
+        \begin{block}{\large Motivation} 
+          \includegraphics[width=\textwidth]{VisionLeibniz}
         \end{block}
@@
       \hfill
-      \begin{column}{.32\linewidth}
-        \begin{block}{\large Quantified Conditional Logics (QCLs)}
-          \footnotesize  
-          \colorbox{GreenYellow}{
-            \begin{minipage}{.93\textwidth} 
-              \begin{align*}
-                \textcolor{red}{\varphi},\textcolor{red}{\psi} ::= \quad & \textcolor{red}{P} \mid \textcolor{red}{k(X^1,\ldots,X^n)} \mid \textcolor{red}{\neg \varphi} \mid \textcolor{red}{\varphi \vee \psi} \mid \textcolor{red}{\varphi \Rightarrow \psi} \mid \\
-                & \textcolor{red}{\forall^{co} X \varphi} \mid \textcolor{red}{\forall^{va} X \varphi} \mid  \textcolor{red}{\forall P \varphi}  \\[-1em]
-              \end{align*}
-            \end{minipage}
-          }
-          \vskip1.5ex
-          \textbf{Interpretation}: $M = \langle S,f,D,D',Q,I
-          \rangle$ where $S$ is a set of 'worlds',
-          $f: S \times 2^{S} \mapsto 2^{S}$ is the selection function,
-          $D\not=\emptyset$ is a set of \emph{individuals} (constant
-          domain), $D'$ is a function that assigns a
-          subset $D'(w)\not=\emptyset$ of $D$ to each world $w$
-          (varying domains), $Q\not=\emptyset$ is a collection of subsets
-          of $W$ (prop.~domain), and $I$ is an interpretation function
-          s.t. for each predicate symbol $k$, $I(k,w) \subseteq
-          D^n$. \\
-          \vskip1ex
-          \textbf{Satisfiability} of $\textcolor{red}{\varphi}$ (denoted as $M,g,s
-          \models \textcolor{red}{\varphi}$) for an interpretation $M$, a world $s \in
-          S$, and a variable assignment $g = (g^{i},g^{p})$:
-          \begin{align*}
-            M,g,s \models\;& \textcolor{red}{k(X^1,...\,,X^n)} \text{\;iff\;} \langle g^{i}(\textcolor{red}{X^1}),...\,, g^{i}(\textcolor{red}{X^n}) \rangle \in I(k,w)\\
-            M,g,s \models\;& \textcolor{red}{P}  \text{\;iff\;} s \in g^{p}(\textcolor{red}{P})
-            \\
-            M,g,s \models\;& \textcolor{red}{\neg \varphi} \text{\;iff\;} M,g,s \not\models \textcolor{red}{\varphi} \, (\text{that is, not } M,g,s \models \textcolor{red}{\varphi})
-            \\
-            M,g,s \models\;& \textcolor{red}{\varphi \vee \psi} \text{\;iff\;} M,g,s \models \textcolor{red}{\varphi} \text{ or } M,g,s \models \textcolor{red}{\psi}
-            \\
-            M,g,s \models\;& \textcolor{red}{\forall^{co} X \varphi} \text{\;iff\;} M,([d/X]g^{i},g^{p}), s \models \textcolor{red}{\varphi} \text{ for all } d \in D
-            \\
-            M,g,s \models\;& \textcolor{red}{\forall^{va} X \varphi} \text{\;iff\;} M,([d/X]g^{i},g^{p}), s \models \textcolor{red}{\varphi} \text{ for all } d \in D'(s)
-            \\
-            M,g,s \models\;& \textcolor{red}{\forall P \varphi} \text{\;iff\;} M,(g^{i},[p/P]g^{p}),s \models \textcolor{red}{\varphi} \text{ for all } p \in Q
-            \\
-            M,g,s \models\;& \textcolor{red}{\varphi \Rightarrow \psi} \text{\;iff\;}
-            M,g,t \models \textcolor{red}{\psi} \text{ for all } t \in S \text{ s.t.}  t \in f(s,[\textcolor{red}{\varphi}]) \\ & \text{ where } [\textcolor{red}{\varphi}] = \{u \mid M,g,u \models \textcolor{red}{\varphi}\}
-          \end{align*}
-          $M \models^{\qcl} \textcolor{red}{\varphi}$ iff $M,g,s \models \textcolor{red}{\varphi}$ for all $s$, $g$. $\models \textcolor{red}{\varphi}$
-          iff $M \models^{\qcl} \textcolor{red}{\varphi}$ for all
-          $M$.\\
-          \vskip1ex
-          \chriscite{[Stalnaker'68],[Delgrande'98]}
+      \begin{column}{.48\linewidth}
+        \begin{block}{\large Application: Gödel's Ontological Argument
+          for the Existence of God}
+           \includegraphics[width=\textwidth]{GoedelInconsistency}
+        \end{block}
+      \end{column}
+    \end{columns} \vskip-3ex
+    \begin{columns}[t]
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Scott's vs. Gödel's Version} 
+          \includegraphics[width=\textwidth]{VisionLeibniz}
         \end{block}
@@
       \hfill
-      \begin{column}{.32\linewidth}
-        \begin{block}{\large Classical Higher-order Logic (HOL)}
-          \footnotesize
-
-          Types \hfill ${\alpha,\beta} ::= {\iota}
-          \textcolor{gray}{\,(worlds)}\mid
-          {\mu} \textcolor{gray}{\,(indiv.)} \mid {o}
-          \textcolor{gray}{\,(Booleans)}\mid {\alpha \rightarrow
-            \beta}$ \\[1em]
-          \colorbox{GreenYellow}{
-          \begin{minipage}{.93\textwidth} 
-          \begin{align*}
-            \textcolor{blue}{s},\textcolor{blue}{t} ::=  & \textcolor{blue}{c_{\alpha}} \mid \textcolor{blue}{X_{\alpha}} \mid \textcolor{blue}{(\lambda X_{\alpha}s_{\beta})_{\alpha \typearrow \beta}} \mid
-            \textcolor{blue}{(s_{\alpha \typearrow \beta}\, t_\alpha)_{\beta}} \mid \\
-            & \textcolor{blue}{(\neg_{o \typearrow o}\;s_{o})_o} \mid \textcolor{blue}{(s_o \vee_{o\typearrow o \typearrow o} t_o)_o}
-            \mid \textcolor{blue}{(\Pi_{(\alpha \typearrow o)\typearrow o}\; s_{\alpha
-              \typearrow o})_o}  \\[-1em]
-          \end{align*}
-          Note:~Binder~notation~\textcolor{blue}{$\forall X_\alpha
-            t_o$}~as~syntactic~sugar~for~\textcolor{blue}{$\Pi_{(\alpha
-              \typearrow o)\typearrow o} \lambda X_\alpha t_o$}\, 
-        \end{minipage}} \vskip2.5ex \textbf{Frame}: collection
-      $\{D_{\alpha}\}_{\alpha \in {T}}$ s.t. $D_{o} = \{T,F\}$,
-      $D_\itype \not= \emptyset$ and $D_\utype \not= \emptyset$
-      arbitrary, and $D_{\alpha \typearrow \beta}$ are collections of
-      total functions from
-      $D_{\alpha}$ to $D_{\beta}$.\\
-      \vskip1.5ex \textbf{Interpretation}: Tuple $\langle
-      \{D_{\alpha}\}_{\alpha \in {T}}, I \rangle$ where
-      $\{D_{\alpha}\}_{\alpha \in {T}}$ is a frame and where function
-      $I$ maps each typed constant symbol $\textcolor{blue}{c_{\alpha}}$ to an
-      appropriate element of $D_{\alpha}$, called the
-      \emph{denotation} of $c_{\alpha}$. The denotations of
-      $\textcolor{blue}{\neg},\textcolor{blue}{\vee}$ and $\textcolor{blue}{\Pi_{(\alpha \typearrow o)\typearrow o}}$ are
-      always chosen as usual.\\
-      \vskip1.5ex 
-      An interpretation is a \textbf{Henkin model} iff there
-      is a valuation function ${V}$ s.t.  ${V}(\phi,\textcolor{blue}{s_{\alpha}}) \in
-      D_{\alpha}$ for each variable assignment $\phi$ and term
-      $\textcolor{blue}{s_{\alpha}}$, and the following conditions are satisfied:
-      ${V}(\phi,\textcolor{blue}{X_{\alpha}}) = \phi(\textcolor{blue}{X_\alpha})$, ${V}(\phi,\textcolor{blue}{c_{\alpha}}) =
-      I(\textcolor{blue}{c_{\alpha}})$, ${V}(\phi,\textcolor{blue}{l_{\alpha \typearrow \beta}\,
-      r_{\alpha}}) = ({V}(\phi,\textcolor{blue}{l_{\alpha \typearrow \beta}})
-      {V}(\phi,\textcolor{blue}{r_{\alpha}}))$, and ${V}(\phi,\textcolor{blue}{\lambda
-      X_{\alpha}s_{\beta}})$ represents the function from $D_{\alpha}$
-      into $D_{\beta}$ whose value for each argument $z \in
-      D_{\alpha}$ is ${V}(\phi[z/\textcolor{blue}{X_{\alpha}}], \textcolor{blue}{s_\beta})$. % , where
-%       $\phi[z/X_\alpha]$ is that assignment
-%       s.t. $\phi[z/X_\alpha](X_\alpha) = z$ and
-%       $\phi[z/X_\alpha]Y_\beta = \phi Y_\beta$ when $Y_\beta
-%       \not=X_\alpha$.
-      If an interpretation is an {Henkin model} the function ${V}$ is
-      uniquely determined. \\
-      \vskip1ex
-      ${H}\models^{HOL} \textcolor{blue}{s}$ iff ${V}(\phi,\textcolor{blue}{s}) = T$ for all $\phi$.
-      $\models \textcolor{blue}{s}$ iff ${H}\models^{HOL} \textcolor{blue}{s}$ for all
-      ${H}$. \\
-      \vskip2.7ex
-        \chriscite{[Church'40],[Andrews'72a/b],[Benzm\"ullerEtAl'04]
-          \vskip.5ex
-        }
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Technique: Semantic Embedding in HOL}
+          \includegraphics[width=\textwidth]{SemanticEmbedding}
         \end{block}
       \end{column}
-    \end{columns}
-    \vskip2ex
-    \begin{block}{\large Embedding QCLs in HOL --- In other words: QCLs are simple Fragments of HOL!} \footnotesize
-      \begin{columns}[t]
-        \begin{column}{.31\textwidth}
-          The mapping $\textcolor{blue}{\lfloor \textcolor{red}{\cdot} \rfloor}$ identifies
-          \qcl\ formulas $\textcolor{red}{\varphi}$ with HOL terms $\textcolor{blue}{\lfloor
-          \textcolor{red}{\varphi} \rfloor}$ of type $\proptype:=\itype \typearrow o$. The mapping is recursively defined:
-          \[
-          \begin{array}{lcl}
-            \textcolor{blue}{\lfloor \textcolor{red}{P} \rfloor} &=& \textcolor{blue}{P_\proptype} \\
-            \textcolor{blue}{\lfloor \textcolor{red}{k(X^1,\ldots,X^n)} \rfloor} % &=& \lfloor k \rfloor \lfloor X^1 \rfloor \ldots \lfloor X^n \rfloor \\
-            &=& \textcolor{blue}{k_{u^n \typearrow \proptype}\,X^1_\utype \ldots X^n_\utype}\\
-            \textcolor{blue}{\lfloor \textcolor{red}{\neg \varphi} \rfloor} &=&  \textcolor{blue}{\neg_{\proptype\typearrow \proptype}}\,\lfloor \textcolor{red}{\varphi} \rfloor \\
-            \textcolor{blue}{\lfloor \textcolor{red}{\varphi\vee \psi} \rfloor} &=&  \textcolor{blue}{\vee_{\proptype \typearrow \proptype \typearrow \proptype}}\,\lfloor\textcolor{red}{\varphi} \rfloor \lfloor \textcolor{red}{\psi} \rfloor\\
-            \textcolor{blue}{\lfloor \textcolor{red}{\varphi\Rightarrow \psi} \rfloor} &=&  \textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}}\,\lfloor\textcolor{red}{\varphi} \rfloor \lfloor \textcolor{red}{\psi} \rfloor \\
-            \textcolor{blue}{\lfloor \textcolor{red}{\forall^{co} X \varphi} \rfloor} &=&   \textcolor{blue}{\Pi^{co}_{(\utype\typearrow\proptype)\typearrow\proptype}\,\lambda X_\utype} \lfloor\textcolor{red}{\varphi} \rfloor \\
-            \textcolor{blue}{\lfloor \textcolor{red}{\forall^{va} X \varphi} \rfloor} &=&  \textcolor{blue}{\Pi^{va}_{(\utype\typearrow\proptype)\typearrow\proptype}\,\lambda X_\utype} \lfloor\textcolor{red}{\varphi} \rfloor \\
-            % \end{array}\]
-            % \[
-            % \begin{array}{lcl}
-            \textcolor{blue}{\lfloor \textcolor{red}{\forall P \varphi} \rfloor} &=&  \textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}\,\lambda P_\proptype} \lfloor\textcolor{red}{\varphi} \rfloor
-          \end{array}
-          \]
-          $\textcolor{blue}{P_\proptype}$ and $\textcolor{blue}{X^1_\utype},\ldots,\textcolor{blue}{X^n_\utype}$ are variables and
-          $\textcolor{blue}{k_{u^n \typearrow \proptype}}$ is a 
-          constant symbol.  
-        \end{column}
-        \hfill
-        \begin{column}{.31\textwidth}
-          $\textcolor{blue}{\neg_{\proptype\typearrow \proptype}}$, $\textcolor{blue}{\vee_{\proptype \typearrow
-            \proptype \typearrow \proptype}}$, $\textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}}$, $\textcolor{blue}{\Pi^{co,va}_{(\utype\typearrow\proptype)\typearrow\proptype}}$
-          and $\textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}}$ realize the \qcl\ connectives in
-          HOL.
-They abbreviate the following HOL terms:% \footnote{Note the predicate argument $A$ of $f$ in the term for $\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}$ and the second-order quantifier $\forall P_{\proptype}$ in the term for $\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}$.
-FOL encodings of both constructs, if feasible, will be less natural.}
-          \[
-          \begin{array}{ll}
-            \textcolor{blue}{\neg_{\proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda X_\itype \neg(A\,X)}\\
-             \textcolor{blue}{\vee_{\proptype \typearrow \proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda B_\proptype \lambda X_\itype (A\,X \vee B\,X)}\\
-             \textcolor{blue}{\Rightarrow_{\proptype \typearrow \proptype \typearrow \proptype}} &=  \textcolor{blue}{\lambda A_\proptype \lambda B_\proptype \lambda X_\itype \forall V_\itype (f\, X\, A\, V \rightarrow B\, V)}\\
-             \textcolor{blue}{\Pi^{co}_{(\utype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {Q_{\utype\typearrow\proptype}}  \lambda {V_\itype}  \forall{X_\utype}  (Q\,X\,V)} \\
-             \textcolor{blue}{\Pi^{va}_{(\utype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {Q_{\utype\typearrow\proptype}}  \lambda {V_\itype}  \forall{X_\utype}  (\exInW\,V\,X \rightarrow Q\,X\,V)} \\
-             \textcolor{blue}{\Pi_{(\proptype\typearrow\proptype)\typearrow\proptype}} & =  \textcolor{blue}{\lambda {R_{\proptype\typearrow\proptype}}  \lambda {V_\itype}  \forall {P_\proptype}  (R\,P\,V)}
-          \end{array}
-          \]
-          % Constant symbol $f$ in the mapping of $\Rightarrow$ is of type ${i
-%             \typearrow \proptype \typearrow \proptype}$. It realizes the
-%           selection function. Constant symbol $\exInW$ (for 'exists in world'),
-%           which is of type ${\itype\typearrow\utype\typearrow o}$,
-%           is associated with the varying domains. 
-          The interpretations of $\textcolor{blue}{f}$ and $\textcolor{blue}{\exInW}$ are chosen
-          appropriately. %  below,
-%           cf. Def.~\ref{def:hm}. Moreover, f
-          For the varying domains
-          non-emptiness is postulated:
-          % \begin{align*}
-             $\textcolor{blue}{\forall W_\itype \exists X_\utype (\exInW\,W\,X)}$ \\[1em]
-          % \end{align*}
-          Meta-level notion of validity defined as $
-          \textcolor{blue}{\text{vld}_{\proptype\typearrow o}} = \textcolor{blue}{\lambda A_\proptype
-          \forall S_\itype (A\, S)}$. \\
-        \end{column}
-        \hfill
-        \begin{column}{.31\textwidth}
-          \vskip-2ex
-          \colorbox{GreenYellow}{
-            \begin{minipage}{.96\textwidth}
-              \vskip1ex
-              \textbf{Theorem: Soundness and Completeness}
-                $$\models^{\qcl}
-                \textcolor{red}{\varphi} \,\, \text{ iff } \,\, \{\text{NE}\}\models^{HOL} \textcolor{blue}{\text{vld}_{\proptype\typearrow o}}\, \textcolor{blue}{\lfloor \textcolor{red}{\varphi}
-                \rfloor} \text{\,\,(wrt Henkin semantics)}$$
-                (Proof: By relating Kripke structures to Henkin models.)
-                \vskip1ex
-            \end{minipage}
-          } 
-          \vskip2ex
-          \colorbox{GreenYellow}{
-            \begin{minipage}{.96\textwidth}
-              \vskip1ex
-              \textbf{Corollary: Cut-elimination for QCL}
-              $$\text{There are cut-free calculi for QCL.}$$ 
-              (Proof: Take any cut-free calculus for HOL, e.g. the cut-free sequent calculus from [Benzm\"ullerEtAl'09]. Note, however, the potential impact of cut-simulation.) 
-                \vskip2ex
-            \end{minipage}        
-          } 
-          \vskip2ex
-          \chriscite{Earlier work is reported in [Benzm.Genovese'11]}   
-        \end{column}
-      \end{columns}
-      % \vskip2ex      \chriscite{[Benzm.Genovese'11]}
-    \end{block}
-    \vskip0ex
-    \begin{columns}[t]
-      \begin{column}{.27\textwidth}
-        \begin{block}{\large The Encoding in THF0-Syntax} 
-          \vskip1.2ex 
-          \begin{minipage}{.92\textwidth}
-               \VerbatimInput[frame=single,%
-               commandchars=\\\{\},%
-               fontfamily=courier,fontseries=b,%
-               fontsize=\tiny,%
-               rulecolor=\color{green},%
-               fillcolor=\color{YellowGreen},%
-               formatcom=\color{blue},%
-               framerule=0pt,%
-               framesep=0pt,numbers=right]%
-               {Axioms_annotated.ax}
-%            \textcolor{blue}{\bf\VerbatimInput[fontsize=\tiny]{Axioms.ax}}
-          \end{minipage}
-          \vskip2ex
-          \chriscite{Introduction to THF0-Syntax [SutcliffeBenzm.'10]}
+     \begin{column}{.33\linewidth}
+        \begin{block}{\large Application: Gödel's Ontological Argument
+          for the Existence of God}
+           \includegraphics[width=\textwidth]{GoedelInconsistency}
         \end{block}
       \end{column}
-      \hfill
-      \begin{column}{.70\textwidth}
-        \begin{block}{\large Automating Prominent Examples from the Literature (in \qcl+ID+MP)}
-          \begin{columns}[t]
-            \begin{column}{.30\textwidth} \footnotesize
-              \textbf{Example: Pegasus, the winged horse} \\[1ex]
-
-              It can be consistently stated (in \qcl+ID+MP) that: \textcolor{green}{\textit{``Horses (h) contingently do not
-              have wings (w) but Pegasus (p) is a winged horse.''}}
-              \textcolor{red}{\begin{align*}
-                & \textcolor{red}{\forall^{va} X (h(X) \rightarrow \neg w(X)),\quad h(p),\quad w(p)}
-              \end{align*}}
-              THF0 encoding of this example:
-              \vskip1ex
-              \begin{minipage}{.92\textwidth}
-                \VerbatimInput[frame=single,%
-                commandchars=\\\{\},%
-                fontfamily=courier,fontseries=b,%
-                fontsize=\tiny,%
-                rulecolor=\color{green},%
-                fillcolor=\color{YellowGreen},%
-                formatcom=\color{blue},%
-                framerule=0pt,%
-                framesep=0pt,numbers=right]%
-                {E1_annotated.thf}
-                \end{minipage}
-              %\textcolor{blue}{\bf \VerbatimInput[fontsize=\tiny]{E1.thf}}
-              \vskip1ex
-              \holp\ confirms the satisfiability of these formulas (with
-              \holpNitpick=7.7). The finite model generated by Nitpick tells us that
-              Pegasus is not 'actual', i.e., does not exist (cf. $\textcolor{blue}{eiw}$) in any world. As expected, when the example problem is formulated
-              with $\textcolor{red}{\forall^{co}}$ instead of\, $\textcolor{red}{\forall^{va}}$ then \holp\ reports
-              unsatisfiability (\holpLS=0.0, \holpIsabelle=5.8).\\
-              %\chriscite{These examples have been discussed (but not automated!) in [Delgrande'98]}
-            \end{column}
-            \begin{column}{.61\textwidth} \footnotesize
-              \vskip-3.6ex
-
-              \, \hfill Notation: \textcolor{red}{$\phi \Rightarrow_X \psi$} := \textcolor{red}{$(\exists^{va} X \phi) \Rightarrow \forall^{va} X
-              (\phi \rightarrow \psi)$} \,
-              \vskip1ex
-
-              \textbf{Example: Opus, the penguin} \\[1ex]
-
-                 \textcolor{green}{\textit{``Birds (b) normally fly (f), but Opus (o) is a bird that normally does not fly.''}}
-                \begin{align*}
-                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad b(o), \quad  b(o) \Rightarrow \neg f(o)}
-                \end{align*}
-                \holp\ reports a finite model (\holpNitpick=8.6). When $\textcolor{red}{\forall^{co}}$ is used:  \holp\  says unsatisfiable (\holpSatallax=0.0, \holpIsabelle=7.9).\\[.5em]
-
-                 \textcolor{green}{\textit{``Birds normally fly and 
-                necessarily Opus the bird does not fly.''}}
-                \begin{align*}
-                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad \Box(b(o) \wedge \neg f(o))}
-                \end{align*}
-                \holp\ reports a finite model (\holpNitpick=8.7).  When $\textcolor{red}{\forall^{co}}$ is used:
-                \holp\  says  unsatisfiable
-                (\holpSatallax=0.0, \holpIsabelle=7.6).\\[.5em]
-
-                \textcolor{green}{\textit{``Birds normally fly and necessarily there is a non-flying bird.''}}
-                \begin{align*}
-                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad \Box\exists^{va}(b(X) \wedge \neg f(X))}
-                \end{align*}
-                \holp\ reports unsatisfiability  (\holpSatallax=0.0, \holpIsabelle=8.7), also when $\textcolor{red}{\forall^{co}}$ is used (\holpSatallax=0.0, \holpIsabelle=8.8).\\[.5em]
-
-                \textcolor{green}{\textit{``Birds normally fly, penguins normally do not fly and that all penguins are necessarily birds.''}}
-                \begin{align*}
-                  &\textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))} 
-                \end{align*}
-                \holp\ generates a finite model ($\textcolor{red}{\forall^{va}}$: \holpNitpick=8.8; $\textcolor{red}{\forall^{co}}$: \holpNitpick=7.9).\\[.5em]
-
- Moreover, \holp\ can conclude from the
-                statements above  that \textcolor{green}{\textit{``Birds are normally
-                not penguins.''}} ($\textcolor{red}{\forall^{va}}$: \holpSatallax=0.9, \holpLeo=10.2, \holpAgsyhol=9.4; $\textcolor{red}{\forall^{co}}$: \holpSatallax=0.8, \holpLeo=10.1, \holpAgsyhol=0.3):
-                \begin{align*}
-                  \textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))  \quad \vdash \quad b(X) \Rightarrow_X \neg p(X)}
-                \end{align*}
-                In line with Delgrande, \holp\ reports a
-                countermodel for the following statement (\holpNitpick=8.7):
-                \begin{align*}
-                  \textcolor{red}{b(X) \Rightarrow_X f(X), \quad p(X) \Rightarrow_X \neg f(X), \quad \forall^{va} \Box (p(X) \rightarrow b(X))  \quad \vdash \quad  b(o) \Rightarrow \neg p(o)}
-                \end{align*}
-                However, when $\textcolor{red}{\forall^{co}}$ is used, \holp\ reports a  theorem (\holpSatallax=0.8, \holpAgsyhol=0.4).
- \vskip2ex
-        \chriscite{These examples have been discussed (but not automated) in [Delgrande'98]}
-            \end{column}
-          \end{columns}         
-        \end{block}
-      \end{column}
-    \end{columns}
-    \vskip2ex
-    \begin{columns}[t]
-      \begin{column}{.42\linewidth}
-        \begin{block}{The HOL Metaprover $H$
-          %  \holp=(\holpLeo,\holpSatallax,\holpIsabelle,\holpNitpick,\holpAgsyhol)
-          }
-          \footnotesize
-          The \holp\ metaprover for HOL sequentially calls the following prover and model finders: \\[.5em]
-          \quad\textbf{\holpLeo} LEO-II (Benzm\"uller/Sultana/Theiss): \url{http://www.leoprover.org} \\
-          \quad\textbf{\holpSatallax} Satallax (Brown): \url{http://www.ps.uni-saarland.de/~cebrown/satallax/}\\
-          \quad\textbf{\holpIsabelle} Isabelle (Blanchette/Paulson/Nipkow/\ldots): \url{http://isabelle.in.tum.de/} \\
-          \quad\textbf{\holpNitpick} Nitpick (Blanchette): \url{http://www4.in.tum.de/~blanchet/nitpick.html} \\
-          \quad\textbf{\holpAgsyhol} agsyHol (Lindblatt): \url{https://github.com/frelindb/agsyHOL}\\[.6em]
-          These systems support THF0 syntax. These provers are remotely
-          available via \\ SystemOnTPTP:
-          \url{http://www.cs.miami.edu/~tptp/cgi-bin/SystemOnTPTP}
-        \end{block}
-      \end{column}
-      \hfill
-      \begin{column}{.56\linewidth}
-        \begin{block}{References and Further Reading} \scriptsize 
-         \vskip1ex
-          \begin{tabular}{ll}
-            [Andrews'72a] & P.B. Andrews.
-            General models, descriptions, and choice in type theory.
-            JSL, 37(2):385-394, 1972. \\
-
-            [Andrews'72b] & P.B. Andrews.
-            General models and extensionality.
-            JSL, 37(2):395--397, 1972. \\
-
-            [Benzm\"uller'13] & C. Benzmüller. A top-down approach to 
-            combining logics, Proc. of ICAART 2013, Barcelona, Spain, 2013. \\
-
-            [Benzm\"ullerEtAl'04] & C. Benzmüller, C. E. Brown, and M. 
-            Kohlhase, Higher order semantics and extensionality. JSL, 69(4):1027-1088, 2004. \\
-
-            [Benzm\"ullerEtAl'09] & C. Benzm{\"u}ller, C. E. Brown, and M. Kohlhase.
-            Cut-simulation and impredicativity. LMCS, 5(1:6):1--21, 2009. \\
-
-            [Benzm.Genovese'11] & C.~Benzm{\"u}ller and V.~Genovese.
-            Quantified conditional logics are fragments of {HOL}. NCMPL 2011. arXiv:1204.5920v1\\
-
-            [Church'40] & A.~Church. A formulation of the simple theory of types.  
-            JSL, 5:56--68, 1940. \\
-
-            [Delgrande'98] & J.P. Delgrande. On first-order conditional logics.
-
-            Artificial Intelligence, 105(1-2):105--137, 1998. \\
-
- %           [Henkin'50] & \\
-
-            [Stalnaker'68] & R.C. Stalnaker.
-            A theory of conditionals.
-            In Studies in Logical Theory, pp. 98--112. Blackwell, 1968.\\
-
-            [SutcliffeBenzm.'10] & G.~Sutcliffe and C.~Benzm. 
-            Automated reasoning in HOL using the {TPTP THF}
-            infrastructure.
-            {\em JFR}, 3(1):1--27, 2010. \\[.9em]
-           \end{tabular}
-        \end{block}
-      \end{column}
-    \end{columns}
 

### 2016-06-22 Papers/2016/IJCAI-Poster/poster.tex
Fix textwidths
Add more rows and fix dimensions
Add screenshots of slides
Adjust horizontal spacing
Adjust vertical spacing
Add summary of results

@@
     \begin{columns}[t]
-      \begin{column}{.49\linewidth}
+      \begin{column}{.5\linewidth}
         \begin{block}{\large Motivation} 
-          \includegraphics[width=\textwidth]{VisionLeibniz}
+          \includegraphics[width=0.98\textwidth]{./Images/VisionLeibniz}
         \end{block}
@@
       \begin{column}{.48\linewidth}
-        \begin{block}{\large Application: Gödel's Ontological Argument
-          for the Existence of God}
-           \includegraphics[width=\textwidth]{GoedelInconsistency}
+        \begin{block}{\large Application: Gödel's Ontological
+            Argument} \centering
+           \includegraphics[width=0.98\textwidth]{./Images/GoedelInconsistency}
         \end{block}
       \end{column}
-    \end{columns} \vskip-3ex
+    \end{columns} 
     \begin{columns}[t]
       \begin{column}{.33\linewidth}
-        \begin{block}{\large Scott's vs. Gödel's Version} 
-          \includegraphics[width=\textwidth]{VisionLeibniz}
+        \begin{block}{\large Scott's and Gödel's Versions} 
+          \includegraphics[width=0.98\textwidth]{./Images/Scott_vs_Goedel}
         \end{block}
       \end{column}
-      \hfill
+      %\hfill
       \begin{column}{.33\linewidth}
-        \begin{block}{\large Technique: Semantic Embedding in HOL}
-          \includegraphics[width=\textwidth]{SemanticEmbedding}
+        \begin{block}{\large Approach: Semantic Embedding}
+          \includegraphics[width=0.98\textwidth]{./Images/SemanticEmbedding}
         \end{block}
@@
      \begin{column}{.33\linewidth}
-        \begin{block}{\large Application: Gödel's Ontological Argument
-          for the Existence of God}
-           \includegraphics[width=\textwidth]{GoedelInconsistency}
+        \begin{block}{\large Standard Translation of Modal Logic}
+           \includegraphics[width=0.98\textwidth]{./Images/StandardTranslation}
         \end{block}
       \end{column}
+    \end{columns}
+    \begin{columns}[t]
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Modal Logic in Isabelle} 
+          \includegraphics[width=0.98\textwidth]{./Images/ModalLogicInIsabelle}
+        \end{block}
+      \end{column}
+      %\hfill
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Scott's Version in Isabelle}
+          \vspace{25pt}
+          \includegraphics[width=0.98\textwidth]{./Images/ScottInIsabelle}
+          \vspace{35pt}
+        \end{block}
+      \end{column}
+     \begin{column}{.33\linewidth}
+        \begin{block}{\large LEO-II's Refutation of Gödel's Axioms}
+           \includegraphics[width=0.98\textwidth]{./Images/Refutation}
+        \end{block}
+      \end{column}
+    \end{columns}
+    \begin{columns}[t]
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Refutation Reconstruction in Isabelle} 
+          \includegraphics[width=0.98\textwidth]{./Images/IsabelleRefutation}
+        \end{block}
+      \end{column}
+      %\hfill
+      \begin{column}{.33\linewidth}
+        \begin{block}{\large Explaining the Inconsistency}
+          \includegraphics[width=0.98\textwidth]{./Images/Explanation}
+          \vspace{75pt}
+        \end{block}
+      \end{column}
+     \begin{column}{.33\linewidth}
+        \begin{block}{\large Summary of Results}
+           \begin{itemize}
+              \item Inconsistency of G\"odel's original axioms has now been verified
+              \item Reason for inconsistency is finally well understood and explained
+              \item Reconstruction of the refutation in Isabelle led to various user interface and performence improvements for the embedding of modal logic in Isabelle
+              \item Higher-order interactive and automated reasoning technology is ready for applications in philosophy
+           \end{itemize}
+           \vspace{25pt}
+        \end{block}
+      \end{column}
+    \end{columns}
 

### 2016-06-23 Papers/2016/KI/main.tex
first draft for KI paper

--- 2016-04-13 Papers/InProgress/IJCAI/main.tex
@@
-\typeout{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
+\documentclass{llncs}
 
-\documentclass{article}
-\usepackage{ijcai15}
-
-% Use the postscript times font!
-\usepackage{times}
+\usepackage[utf8]{inputenc}
+%\usepackage{fancyvrb,amssymb,graphicx}
+%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 \usepackage{graphicx}
-\usepackage{caption}
+% \usepackage{caption}
 \usepackage{subcaption}
+\captionsetup{compatibility=false}
+
+\usepackage{wrapfig}
 
@@
 \usepackage{xspace}
-\usepackage{modallogics}
 \usepackage{graphicx,url}
@@
 \usepackage{xcolor}
+
+%\usepackage{modallogics}
+\newcommand{\logic}[1]{\textbf{#1}\xspace}
+\newcommand{\KB}{\logic{KB}}
+\newcommand{\KFour}{\logic{K4}}
+\newcommand{\KFourB}{\logic{K4B}}
+\newcommand{\K}{\logic{K}}
+\newcommand{\KT}{\logic{KT}}
+\newcommand{\SFour}{\logic{S4}}
+\newcommand{\SFive}{\logic{S5}}
+\newcommand{\SFiveU}{\logic{S5\textsuperscript{U}}}
 
@@
 
+\begin{document}
+
 \title{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
-\author{
-  \begin{minipage}{0.5\textwidth}\centering
-    Christoph Benzm\"uller\thanks{This work was supported by the 
-                                  German National Research Foundation (DFG) under
-                                  grants BE 2501/9-2 and BE 2501/11-1.} \\
-    \normalfont Freie Universit\"at Berlin \& Stanford University\\
-    c.benzmueller@gmail.com
-  \end{minipage}
-  \begin{minipage}{0.5\textwidth}\centering
-    Bruno Woltzenlogel Paleo \\
-    \normalfont Australian National University\\
-    bruno.wp@gmail.com
-  \end{minipage}
+\author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
+    the German Research Foundation DFG under grants BE2501/9-1,2} \and Bruno Woltzenlogel Paleo\inst{2} 
 }
 
-\begin{document}
+\institute{
+ Freie Universit\"at Berlin, Berlin, Germany \\ \url{c.benzmueller@fu-berlin.de}
+\and
+ Australien National University, Canberra,  Australia \\ \url{bruno.wp@gmail.com}
+}
 
@@
   task. Its completion, as reported here, required the
-  reconstruction of the refutation in the Isabelle proof assistant, and it
-  also led to a novel and more efficient way of automating
-  higher-order modal logic \textbf{S5} with a universal accessibility
-  relation. Furthermore, the development of an improved syntactical
-  hiding for the utilized logic embedding technique allows the refutation to be
-  presented in a human-friendly way, suitable for non-experts in the
-  technicalities of higher-order theorem proving. This brings us a
-  step closer to wider adoption of logic-based artificial intelligence
-  tools by philosophers.
+  reconstruction of the refutation in the Isabelle proof assistant.% , and it
+  % also led to a novel and more efficient way of automating
+  % higher-order modal logic \textbf{S5} with a universal accessibility
+  % relation. Furthermore, the development of an improved syntactical
+  % hiding for the utilized logic embedding technique allows the refutation to be
+  % presented in a human-friendly way, suitable for non-experts in the
+  % technicalities of higher-order theorem proving. This brings us a
+  % step closer to wider adoption of logic-based artificial intelligence
+  % tools by philosophers.
+  This paper is short version our contribution to IJCAI 2016.
 \end{abstract}
@@
 discuss, assess, criticize, modify and improve G\"{o}del's original
-work (see e.g.~Sobel~\shortcite{sobel2004logic} and Oppy~\shortcite{sep-ontological-arguments} and the
+work (see e.g.~Sobel~\cite{sobel2004logic} and Oppy~\cite{sep-ontological-arguments} and the
 references therein).  In philosophy lectures at universities the
 argument is regularly presented as a masterpiece argument in
-metaphysics. Since 2013, when Benzm\"uller and Woltzenlogel-Paleo~\shortcite{J30,C40} first
-reported their successful initial computer-assisted
+metaphysics. Since 2013, when we~\cite{J30,C40} first
+reported our successful initial computer-assisted
 analysis of G\"odel's proof and Scott's variant,
-their work has received a media repercussion on a global scale\footnote{A
+our work has received a media repercussion on a global scale\footnote{A
   collection of news articles is available at {\scriptsize
@@
 further improve the logic embedding approach \cite{J23,C40}, which has
-already been employed successfully in preceding related work.
+already been employed successfully in our preceding work.
 
-The main novel contribution reported in this paper is a detailed
+The main novel contribution reported here is a detailed
 analysis (in various modal logics) of the inconsistency of G\"{o}del's
 original version of the axioms used in his manuscript
-\shortcite{GoedelNotes}. The extraction, reconstruction and
+\cite{GoedelNotes}. The extraction, reconstruction and
 verification of an informal, human intuitive argument has been an open
 problem since the first detection of this inconsistency by
-Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40} with the
+Benzm\"uller and Woltzenlogel-Paleo \cite{C40} with the
 \textsc{Leo-II} prover.  The verified refutation (discussed in
@@
 
-On the technical side, the quest for constructing a compelling
-refutation, capable of convincing also human non-experts, led to an
-improvement of the syntax of the embedding of modal logics in Isabelle/HOL
-(as discussed in \S\ref{sec:improvedsyntax}). With the new syntax, a
-(nearly) perfect match between the original pen and paper
-presentations and our encoding in Isabelle/HOL is feasible. A more
-user-friendly syntax, as reported here, is clearly an important
-prerequisite for promoting the theorem proving technology employed
-here to a wider community of philosophers, who are not necessarily
-experts in automated reasoning or HOL.
+% On the technical side, the quest for constructing a compelling
+% refutation, capable of convincing also human non-experts, led to an
+% improvement of the syntax of the embedding of modal logics in Isabelle/HOL
+% (as discussed in \S\ref{sec:improvedsyntax}). With the new syntax, a
+% (nearly) perfect match between the original pen and paper
+% presentations and our encoding in Isabelle/HOL is feasible. A more
+% user-friendly syntax, as reported here, is clearly an important
+% prerequisite for promoting the theorem proving technology employed
+% here to a wider community of philosophers, who are not necessarily
+% experts in automated reasoning or HOL.
 
-Another novel contribution reported here (in
-\S\ref{sec:improvedembedding}) is the implementation of an alternative
-embedding for the 
-% slightly stronger 
-more effective modal logic \SFiveU, which is
-based on a universal accessibility relation. Our experiments have
-shown that the new embedding is more efficient, as the following two
-previously open problems can now be solved:
-\begin{itemize}
-\item Automatically proving the final theorem T3 (\textit{Necessarily, there
-  exists God}), directly from Scott's \shortcite{ScottNotes} (consistent) axioms
-  alone, without relying on the argument's intermediate argumentation
-  steps (i.e., lemmata).
-\item Automatically verifying, in Isabelle/HOL, the proof of the modal
-  collapse \cite{Sobel}, which is one of the most strongly criticized
-  logical consequences of the argument's axioms.
-\end{itemize}
+% Another novel contribution reported here (in
+% \S\ref{sec:improvedembedding}) is the implementation of an alternative
+% embedding for the 
+% % slightly stronger 
+% more effective modal logic \SFiveU, which is
+% based on a universal accessibility relation. Our experiments have
+% shown that the new embedding is more efficient, as the following two
+% previously open problems can now be solved:
+% \begin{itemize}
+% \item Automatically proving the final theorem T3 (\textit{Necessarily, there
+%   exists God}), directly from Scott's \cite{ScottNotes} (consistent) axioms
+%   alone, without relying on the argument's intermediate argumentation
+%   steps (i.e., lemmata).
+% \item Automatically verifying, in Isabelle/HOL, the proof of the modal
+%   collapse \cite{Sobel}, which is one of the most strongly criticized
+%   logical consequences of the argument's axioms.
+% \end{itemize}
 
@@
 metaphysics were reported by Fitelson, Oppenheimer and
-Zalta~\shortcite{FitelsonZalta,oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
-\textsc{Prover9} \cite{prover9-mace4} in their experiments. Later on, Rushby~\shortcite{rushby13} used the proof assistant \textsc{PVS} \cite{cade92-pvs}. Common to both
+Zalta~\cite{FitelsonZalta,oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
+\textsc{Prover9} \cite{prover9-mace4} in their experiments. Later on,
+Rushby~\cite{rushby13} used the proof assistant \textsc{PVS} \cite{cade92-pvs}. Common to both
 works is a significant amount of proof-hand-coding work as well as their
-focus on a non-modal formalization of St. Anselm's~\shortcite{Proslogion} simpler 
+focus on a non-modal formalization of St. Anselm's~\cite{Proslogion} simpler 
 and older ontological argument. In contrast, the greater complexity of G\"odel's argument requires the formalization and automation of variants of \emph{higher-order} and \emph{modal} logics.
 
-\section{A Brief History of the Argument}\label{sec:history}
-
+\begin{wrapfigure}[22]{r}{0.55\textwidth} \centering \vskip-3em
+  \includegraphics[width=.52\textwidth]{./Scott_S5U.png}
+  \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
+\end{wrapfigure} 
+\section{A Brief History}\label{sec:history}
 St. Anselm's ontological argument \cite{Proslogion} can be regarded as
@@
 G\"odel was the work of Leibniz~\cite{Adams}.  
-Although G\"odel's notion of positive property is not exactly the same as Leibniz's notion of perfection, G\"odel's manuscript
+Although G\"odel's notion of positive property is not exactly the same
+as Leibniz's notion of perfection, G\"odel's manuscript
 (Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's
@@
 definitions, formalized in Isabelle, is shown in
-Fig.~\ref{Scott_S5U}. The main difference to G\"odel's version is an
+Fig.~\ref{Scott_S5U}. 
+The main difference to G\"odel's version is an
 extra conjunct in the definition of \emph{essence} (\emph{ess}). G\"odel's
@@
 
+\begin{wrapfigure}[20]{r}{0.55\textwidth}
+    \centering \vskip-1em
+    \centerline{\includegraphics[width=0.52\textwidth]{./InconsistencyIsabelleK.png}}
+ \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
+\end{wrapfigure}
 G\"odel's omission has been
 considered inessential and merely an oversight by many. For instance,
-Hazen [\citeyear[p.365]{Hazen}] states that ``G\"odel left this
+Hazen [\cite[p.365]{Hazen}] states that ``G\"odel left this
 clause out [\ldots] but this appears to have been an oversight''.
@@
 argument. Especially since the discovery by Sobel
-\shortcite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
+\cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
   $\phi\rightarrow \Box \phi$, states that contingent truth implies
@@
   separately.} Opponents of the argument
-(e.g. Oppy [\citeyear[p.226--227]{oppy96:_goedel_ontol_argum};
-\citeyear[p.364]{oppy00:_respon_gettin};
-\citeyear[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
+(e.g. Oppy [\cite[p.226--227]{oppy96:_goedel_ontol_argum};
+\cite[p.364]{oppy00:_respon_gettin};
+\cite[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
 parodies and other criticisms, referring to variants where the
@@
 
-\begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./Images/Scott_S5U.png}}
-\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
-\end{figure}
+% \begin{figure}[t]{0.48\textwidth}
+% \centerline{\includegraphics[width=\textwidth]{./InconsistencyIsabelleK.png}}
+% \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
+% \end{figure}
+
+% \begin{figure}[t]{0.48\textwidth}
+% \centerline{\includegraphics[width=\textwidth]{./Scott_S5U.png}}
+% \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+%   Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
+% \end{figure}
 
@@
 
-\subsection{Improved Embedding}\label{sec:improvedembedding}
-\begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./Images/QML_S5U.png}}
-\caption{Improved Embedding of \SFiveU} \label{QML_S5U}
-\end{figure}
+% \subsection{Improved Embedding}\label{sec:improvedembedding}
+% \begin{figure}[t]
+% \centerline{\includegraphics[width=1\columnwidth,height=10cm]{./QML_S5U.png}}
+% \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
+% \end{figure}
 
-The modal logic \SFive requires that the accessibility relation be
-reflexive, symmetric and transitive. The usual approach to embed
-\SFive would be to use the standard translation for \textbf{K} described above
-and to state that $r$ is an equivalence relation, e.g., by postulating
-the following axioms:
-\begin{itemize}
-\item Reflexivity: $\forall x. (r~x~x)$
-\item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
-\item Transitivity: $\forall x. \forall y. \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
-\end{itemize}
-Instead, we consider here an alternative description, that we call \SFiveU, based on the following condition on $r$:
-\begin{itemize}
-\item Universality: $\forall x. \forall y. (r~x~y)$
-\end{itemize}
+% The modal logic \SFive requires that the accessibility relation be
+% reflexive, symmetric and transitive. The usual approach to embed
+% \SFive would be to use the standard translation for \textbf{K} described above
+% and to state that $r$ is an equivalence relation, e.g., by postulating
+% the following axioms:
+% \begin{itemize}
+% \item Reflexivity: $\forall x. (r~x~x)$
+% \item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
+% \item Transitivity: $\forall x. \forall y. \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
+% \end{itemize}
+% Instead, we consider here an alternative description, that we call \SFiveU, based on the following condition on $r$:
+% \begin{itemize}
+% \item Universality: $\forall x. \forall y. (r~x~y)$
+% \end{itemize}
 
-It is important to note that $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$ 
-\cite{Blackburn}, and therefore \SFive and \SFiveU are traditionally considered to be 
-two different descriptions of the same modal logic. Nevertheless, \SFive and \SFiveU 
-differ in the shapes of frames they admit:
-%It is easy to see that \SFiveU is at least as strong as \SFive:
-%universality entails that $r$ is an equivalence relation. \SFiveU is,
-%in fact, \emph{strictly stronger} than \SFive. 
-\SFiveU only admits
-complete\footnote{A graph is \emph{complete} iff there is a directed
-  edge connecting every ordered pair of vertices.} frames, whereas
-\SFive admits non-complete frames as long as all their components are
-complete. In other words, in \SFiveU we face one single equivalence
-class of possible worlds, while in \SFive we may face several
-disconnected equivalence classes. In fact, for this reason, \SFiveU is
-considered as metaphysically more appropriate by some philosophers;
-cf. \cite[p.~127]{williamson13}.
+% It is important to note that $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$ 
+% \cite{Blackburn}, and therefore \SFive and \SFiveU are traditionally considered to be 
+% two different descriptions of the same modal logic. Nevertheless, \SFive and \SFiveU 
+% differ in the shapes of frames they admit:
+% %It is easy to see that \SFiveU is at least as strong as \SFive:
+% %universality entails that $r$ is an equivalence relation. \SFiveU is,
+% %in fact, \emph{strictly stronger} than \SFive. 
+% \SFiveU only admits
+% complete\footnote{A graph is \emph{complete} iff there is a directed
+%   edge connecting every ordered pair of vertices.} frames, whereas
+% \SFive admits non-complete frames as long as all their components are
+% complete. In other words, in \SFiveU we face one single equivalence
+% class of possible worlds, while in \SFive we may face several
+% disconnected equivalence classes. In fact, for this reason, \SFiveU is
+% considered as metaphysically more appropriate by some philosophers;
+% cf. \cite[p.~127]{williamson13}.
 
-Furthermore, for \SFiveU an improved embedding is possible. Universality implies
-that the guarding predicates in the definitions of $\Box$ and
-$\Diamond$ always hold. Therefore, they can be omitted and the
-accessibility relation can be dispensed altogether. The modal
-operators can then be defined merely as:
-% \begin{itemize}
-% \item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
-% \item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
-% \end{itemize}
-\[\Box \varphi \equiv \lambda w. \forall v.  \varphi(v) \qquad \text{and}
-\qquad \Dia \varphi \equiv \lambda w.\exists v. \varphi(v)\]
+% Furthermore, for \SFiveU an improved embedding is possible. Universality implies
+% that the guarding predicates in the definitions of $\Box$ and
+% $\Diamond$ always hold. Therefore, they can be omitted and the
+% accessibility relation can be dispensed altogether. The modal
+% operators can then be defined merely as:
+% % \begin{itemize}
+% % \item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
+% % \item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
+% % \end{itemize}
+% \[\Box \varphi \equiv \lambda w. \forall v.  \varphi(v) \qquad \text{and}
+% \qquad \Dia \varphi \equiv \lambda w.\exists v. \varphi(v)\]
 
-The new embedding of \SFiveU in Isabelle/HOL is shown in
-Fig.~\ref{QML_S5U}.
-With this improved embedding, the final theorem T3
-(\textit{Necessarily, there exists God}) can be derived from
-Scott's consistent version of the axioms fully automatically.  The fully
-automatic proof has been generated (in about 2.5 seconds) by the theorem prover
-\textsc{Leo-II}~\cite{leo2} and subsequently verified in the proof
-assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5U}.
-The collaboration between the two systems has been orchestrated by
-Isabelle's Sledgehammer tool \cite{Sledgehammer}.
+% The new embedding of \SFiveU in Isabelle/HOL is shown in
+% Fig.~\ref{QML_S5U}.
+% With this improved embedding, the final theorem T3
+% (\textit{Necessarily, there exists God}) can be derived from
+% Scott's consistent version of the axioms fully automatically.  The fully
+% automatic proof has been generated (in about 2.5 seconds) by the theorem prover
+% \textsc{Leo-II}~\cite{leo2} and subsequently verified in the proof
+% assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5U}.
+% The collaboration between the two systems has been orchestrated by
+% Isabelle's Sledgehammer tool \cite{Sledgehammer}.
 
-With the embedding used by Benzm\"uller and Woltzenlogel-Paleo \shortcite{C40}, the
-provers still had to be given the intermediate theorem T2 and the corollary
-C in order to manage to prove T3.
+% With the embedding used by Benzm\"uller and Woltzenlogel-Paleo \cite{C40}, the
+% provers still had to be given the intermediate theorem T2 and the corollary
+% C in order to manage to prove T3.
 
-Another evidence that the new embedding provides a significant performance boost 
-is the successful automatic verification in Isabelle/HOL (with its automatic tactic Meson) of the modal
-collapse \cite{Sobel}, which is one of the most strongly criticized
-`side-effects' of G\"odel's and Scott's variants of the proof. In previous work \shortcite{C40}
-the modal collapse has been proven by the higher-order provers
-\textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-verification in the highly trusted Isabelle/HOL system still failed
-\cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5U}.
+% Another evidence that the new embedding provides a significant performance boost 
+% is the successful automatic verification in Isabelle/HOL (with its automatic tactic Meson) of the modal
+% collapse \cite{Sobel}, which is one of the most strongly criticized
+% `side-effects' of G\"odel's and Scott's variants of the proof. In previous work \cite{C40}
+% the modal collapse has been proven by the higher-order provers
+% \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
+% verification in the highly trusted Isabelle/HOL system still failed
+% \cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5U}.
 
-\subsection{Improved Syntax in Isabelle}\label{sec:improvedsyntax}
+% \subsection{Improved Syntax in Isabelle}\label{sec:improvedsyntax}
 
-Wider adoption of HOL theorem proving technology for
-reasoning about and within embedded object logics, especially among
-non-expert users, is still hindered by the gap between the syntax used
-by people, when they write logical formulas with pen and paper, and
-the syntax used by HOL theorem provers. Even when the syntax of the
-underlying higher-order system is elegant (as is the case in
-Isabelle/HOL), the embedding of HOML into HOL may easily expose
-details of HOL that may be uncommon to the user, disturbing his/her
-experience while using the system. To illustrate this point,
-Fig.~\ref{UglyEssence} shows how the definition of essence looked like
-in previous work \cite{J28}, where advanced syntax-sugaring features
-were not used. It looks notably higher-order, and its style differs
-significantly from the common style seen in works on modal logics and
-the ontological argument. The following specific issues can be
-enumerated:
-\begin{enumerate}
-\item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they did not need to in a pure HOML formulation (cf. G\"odel's manuscript, Fig.~\ref{GoedelScript}).
-\item Quantifiers appear as higher-order defined constants, and not as binders.
+% Wider adoption of HOL theorem proving technology for
+% reasoning about and within embedded object logics, especially among
+% non-expert users, is still hindered by the gap between the syntax used
+% by people, when they write logical formulas with pen and paper, and
+% the syntax used by HOL theorem provers. Even when the syntax of the
+% underlying higher-order system is elegant (as is the case in
+% Isabelle/HOL), the embedding of HOML into HOL may easily expose
+% details of HOL that may be uncommon to the user, disturbing his/her
+% experience while using the system. To illustrate this point,
+% Fig.~\ref{UglyEssence} shows how the definition of essence looked like
+% in previous work \cite{J28}, where advanced syntax-sugaring features
+% were not used. It looks notably higher-order, and its style differs
+% significantly from the common style seen in works on modal logics and
+% the ontological argument. The following specific issues can be
+% enumerated:
+% \begin{enumerate}
+% \item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they did not need to in a pure HOML formulation (cf.
+G\"odel's manuscript, Fig.~\ref{GoedelScript}).
+% \item Quantifiers appear as higher-order defined constants, and not as binders.
 This forces the user to read (and write) formulas of the form $\forall (\lambda x.
@@
 A(x)$.
-\item The lifted modal connectives are represented by prefixing the
-  letter ``m'' (e.g. $m\wedge$ and $m\imp$). The prefix disturbs the
-  user, as it constantly reminds him/her that there is something
-  unusual about the modal connectives.
-\item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
-Instead of reading $\psi(y)$, as he/she would expect, he/she has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
-\end{enumerate}
+% \item The lifted modal connectives are represented by prefixing the
+%   letter ``m'' (e.g. $m\wedge$ and $m\imp$). The prefix disturbs the
+%   user, as it constantly reminds him/her that there is something
+%   unusual about the modal connectives.
+% \item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
+% Instead of reading $\psi(y)$, as he/she would expect, he/she has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
+% \end{enumerate}
 
-\begin{figure}
-\centerline{\includegraphics[width=1\columnwidth]{./Images/UglyEssence.png}}
-\caption{Definition of Essence using Old Syntax} \label{UglyEssence}
-\end{figure}
+% \begin{figure}
+% \centerline{\includegraphics[width=1\columnwidth]{./UglyEssence.png}}
+% \caption{Definition of Essence using Old Syntax} \label{UglyEssence}
+% \end{figure}
 
-In the embedding presented here, in Fig.~\ref{QML_S5U}, advanced
-syntax-sugaring effects provided by Isabelle were used to prevent
-issues as those enumerated above.
+% In the embedding presented here, in Fig.~\ref{QML_S5U}, advanced
+% syntax-sugaring effects provided by Isabelle were used to prevent
+% issues as those enumerated above.
 The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
@@
 
-The syntax improvements described here render the computer-assisted analysis of ontological arguments accessible to a wider audience and ease the adoption of logic-based artificial intelligence tools by philosophers interested in topics where modal logic reasoning is required.
+% The syntax improvements described here render the computer-assisted analysis of ontological arguments accessible to a wider audience and ease the adoption of logic-based artificial intelligence tools by philosophers interested in topics where modal logic reasoning is required.
 
@@
 \begin{figure*}
-% \begin{minipage}{.01\textwidth}\small
-% \vskip2cm * 
-% \end{minipage}
-% \begin{minipage}{.99\textwidth}
-% {\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
-% \end{minipage}
-\centerline{\includegraphics[width=\textwidth]{./Images/LEO-Proof.png}}
+\centerline{\includegraphics[width=\textwidth]{./LEO-Proof.png}}
 \caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./InconsistencyIsabelleK.png}}
 \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
@@
   \begin{subfigure}[t]{0.715\textwidth}
-    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
+    \includegraphics[width=\textwidth]{./Manuscript2.png}
     \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
@@
    \begin{subfigure}[t]{0.28\textwidth}
-     \includegraphics[width=\textwidth,height=7.2cm]{./Images/Inconsistency_S5U_direct.png}
+     \includegraphics[width=\textwidth,height=7.2cm]{./Inconsistency_S5U_direct.png}
      \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
@@
 \small
-\bibliographystyle{named}
-%\bibliography{Bibliography}
-
-\begin{thebibliography}{}
-
-\bibitem[\protect\citeauthoryear{Adams}{1995}]{Adams}
-R.M. Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpubl. Essays and
-  Letters}}. Oxford Univ. Press, 1995.
-
-\bibitem[\protect\citeauthoryear{Anderson and
-  Gettings}{1996}]{AndersonGettings}
-A.C. Anderson and M.~Gettings.
-\newblock {G\"odel} ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}, pages 167--172. {Springer},
-  1996.
-
-\bibitem[\protect\citeauthoryear{Anderson}{1990}]{Anderson}
-C.A. Anderson.
-\newblock Some emendations of {G{\"o}del's} ontological proof.
-\newblock {\em Faith and Philosophy}, 7(3), 1990.
-
-\bibitem[\protect\citeauthoryear{Andrews}{2014}]{andrewsSEP}
-P.B. Andrews.
-\newblock Church's type theory.
-\newblock In E.N. Zalta, editor, {\em The Stanford Encyclopedia of Philosophy}.
-  Spring 2014 edition, 2014.
-
-\bibitem[\protect\citeauthoryear{Anselm}{1078}]{Proslogion}
-St. Anselm.
-\newblock Proslogion.
-\newblock In M.~Charlesworth, editor, {\em St.~Anselm's Proslogion}.
-  Oxford:OUP, 1078.
-\newblock Republished in 1965.
-
-\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and Paulson}{2013}]{J23}
-C.~Benzm{\"u}ller and L.C. Paulson.
-\newblock Quantified multimodal logics in simple type theory.
-\newblock {\em Logica Universalis}, 7(1):7--20, 2013.
-
-\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and
-  Woltzenlogel-Paleo}{2013a}]{J30}
-C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo.
-\newblock Formalization, mechanization and automation of {G{\"o}del's} proof of
-  {God's} existence.
-\newblock {\em arXiv:1308.4526}, 2013.
-\newblock Preprint available as arXiv:1308.4526.
-
-\bibitem[\protect\citeauthoryear{Benzm\"uller and
-  Woltzenlogel-Paleo}{2013b}]{J28}
-C.~Benzm\"uller and B.~Woltzenlogel-Paleo.
-\newblock {G{\"o}del's God in Isabelle/HOL}.
-\newblock {\em Archive of Formal Proofs}, 2013, 2013.
-
-\bibitem[\protect\citeauthoryear{Benzm{\"u}ller and
-  Woltzenlogel-Paleo}{2014}]{C40}
-C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo.
-\newblock Automating {G\"{o}del's} ontological proof of {God}'s existence with
-  higher-order automated theorem provers.
-\newblock In 
-  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
-  Applications}, pages 93 -- 98. IOS Press, 2014.
-
-\bibitem[\protect\citeauthoryear{Benzm{\"u}ller \bgroup \em et al.\egroup
-  }{2015}]{leo2}
-Christoph Benzm{\"u}ller, Lawrence~C. Paulson, Nik Sultana, and Frank
-  Thei{\ss}.
-\newblock The higher-order prover {LEO-II}.
-\newblock {\em J. of Automated Reasoning}, 55(4):389--404, 2015.
-
-\bibitem[\protect\citeauthoryear{Bj{\o}rdal}{1999}]{Bjordal}
-F.~Bj{\o}rdal.
-\newblock Understanding {G\"{o}del’s} ontological argument.
-\newblock In T.~Childers, editor, {\em The Logica Yearbook 1998}. Filosofia,
-  1999.
-
-\bibitem[\protect\citeauthoryear{Blackburn \bgroup \em et al.\egroup
-  }{2001}]{Blackburn}
-P.~Blackburn, M.~de~Rijke, and Y.~Venema.
-\newblock {\em Modal Logic}.
-\newblock Cambridge University Press, 2001.
-
-\bibitem[\protect\citeauthoryear{Blanchette and Nipkow}{2010}]{Nitpick}
-J.C. Blanchette and T.~Nipkow.
-\newblock Nitpick: A counterexample generator for higher-order logic based on a
-  relational model finder.
-\newblock In {\em ITP 2010}, number 6172 in LNCS, pages 131--146. Springer,
-  2010.
-
-\bibitem[\protect\citeauthoryear{Blanchette \bgroup \em et al.\egroup
-  }{2013}]{Sledgehammer}
-J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
-\newblock Extending {Sledgehammer} with {SMT} solvers.
-\newblock {\em J. of Automated Reasoning}, 51(1):109--128, 2013.
-
-\bibitem[\protect\citeauthoryear{Brown}{2012}]{Satallax}
-C.E. Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117. Springer,
-  2012.
-
-\bibitem[\protect\citeauthoryear{Fitelson and Zalta}{2007}]{FitelsonZalta}
-Branden Fitelson and Edward~N. Zalta.
-\newblock Steps toward a computational metaphysics.
-\newblock {\em J. Philosophical Logic}, 36(2):227--247, 2007.
-
-\bibitem[\protect\citeauthoryear{Fuhrmann}{2016}]{fuhrmann15:_blogg_goedel}
-A.~Fuhrmann.
-\newblock Blogging {G\"odel}: His ontological argument in the public eye.
-\newblock In K.~\'{S}wi{e}torzecka, editor, {\em Forthcoming in
-  \textit{G\"odel’s Ontological Argument - History, Modifications, and
-  Controversies}}. 2016.
-
-\bibitem[\protect\citeauthoryear{G\"odel}{1970}]{GoedelNotes}
-K.~G\"odel.
-\newblock {\em Appx. A: Notes in Kurt G\"odel's Hand}, pages 144--145.
-\newblock In Sobel \shortcite{sobel2004logic}, 1970.
-
-\bibitem[\protect\citeauthoryear{H\'ajek}{1996}]{Hajek1}
-P.~H\'ajek.
-\newblock Magari and others on {G\"odel’s} ontological proof.
-\newblock In A.~Ursini and P.~Agliano, editors, {\em Logic and algebra}, page
-  125–135. Dekker, New York etc., 1996.
-
-\bibitem[\protect\citeauthoryear{H\'ajek}{2001}]{Hajek2}
-P.~H\'ajek.
-\newblock {Der Mathematiker und die Frage der Existenz Gottes}.
-\newblock In B.~Buldt~et al., editor, {\em {Kurt G{\"o}del. Wahrheit und
-  Beweisbarkeit}}, pages 325--336. öbv \& hpt, Wien, 2001.
-\newblock ISBN 3-209-03835-X.
-
-\bibitem[\protect\citeauthoryear{H{\'{a}}jek}{2002}]{Hajek3}
-P.~H{\'{a}}jek.
-\newblock A new small emendation of {G{\"{o}}del's} ontological proof.
-\newblock {\em Studia Logica}, 71(2):149--164, 2002.
-
-\bibitem[\protect\citeauthoryear{Hazen}{1998}]{Hazen}
-A.P. Hazen.
-\newblock On {G\"odel's} ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
-
-\bibitem[\protect\citeauthoryear{McCune}{2010}]{prover9-mace4}
-W.~McCune.
-\newblock {Prover9} and {Mace4} (2005--2010).
-\newblock {\small \verb|http://www.cs.unm.edu/~mccune/prover9/|}, 2010.
-
-\bibitem[\protect\citeauthoryear{Muskens}{2006}]{homl}
-R.~Muskens.
-\newblock {Higher Order Modal Logic}.
-\newblock In P.~Blackburn~et al., editor, {\em Handbook of Modal Logic},
-  Studies in Logic and Practical Reasoning, pages 621--653. Elsevier,
-  Dordrecht, 2006.
-
-\bibitem[\protect\citeauthoryear{Nipkow \bgroup \em et al.\egroup
-  }{2002}]{NPW02}
-T.~Nipkow, L.~Paulson, and M.~Wenzel.
-\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
-
-\bibitem[\protect\citeauthoryear{Ohlbach}{1991}]{DBLP:journals/logcom/Ohlbach91}
-H.J. Ohlbach.
-\newblock Semantics-based translation methods for modal logics.
-\newblock {\em J. Log. Comput.}, 1(5):691--746, 1991.
-
-\bibitem[\protect\citeauthoryear{Oppenheimer and Zalta}{2011}]{oppenheimer11}
-P.E. Oppenheimer and E.N. Zalta.
-\newblock A computationally-discovered simplification of the ontological
-  argument.
-\newblock {\em Australasian J. of Philosophy}, 89(2):333--349, 2011.
-
-\bibitem[\protect\citeauthoryear{Oppy}{1996}]{oppy96:_goedel_ontol_argum}
-G.~Oppy.
-\newblock G{\"o}delian ontological arguments.
-\newblock {\em Analysis}, 56(4):226--230, 1996.
-
-\bibitem[\protect\citeauthoryear{Oppy}{2000}]{oppy00:_respon_gettin}
-G.~Oppy.
-\newblock Response to {Gettings}.
-\newblock {\em Analysis}, 60(4):363--367, 2000.
-
-\bibitem[\protect\citeauthoryear{Oppy}{2008}]{oppy08:_higher_order_ontol_argum}
-G.~Oppy.
-\newblock Higher-order ontological arguments.
-\newblock {\em Philosophy Compass}, 3(5):1066--1078, 2008.
-
-\bibitem[\protect\citeauthoryear{Oppy}{2015}]{sep-ontological-arguments}
-G.~Oppy.
-\newblock Ontological arguments.
-\newblock In Edward~N. Zalta, editor, {\em The Stanford Encyclopedia of
-  Philosophy}. Spring 2015 edition, 2015.
-
-\bibitem[\protect\citeauthoryear{Owre \bgroup \em et al.\egroup
-  }{1992}]{cade92-pvs}
-{S.} Owre, {J.}~{M.} Rushby, and {N.} Shankar.
-\newblock {PVS:} {A} prototype verification system.
-\newblock In Deepak Kapur, editor, {\em CADE}, volume 607 of {\em LNAI}, pages
-  748--752, Saratoga, {NY}, jun 1992. Springer.
-
-\bibitem[\protect\citeauthoryear{Rushby}{2013}]{rushby13}
-J.~Rushby.
-\newblock The ontological argument in {PVS}.
-\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St.
-  Petersburg, Russia, 2013.
-
-\bibitem[\protect\citeauthoryear{Scott}{1972}]{ScottNotes}
-D.~Scott.
-\newblock {\em Appx. B: Notes in Dana Scott's Hand}, pages 145--146.
-\newblock In Sobel \shortcite{sobel2004logic}, 1972.
-
-\bibitem[\protect\citeauthoryear{Sobel}{1987}]{Sobel}
-J.H. Sobel.
-\newblock G\"odel's ontological proof.
-\newblock In {\em {On Being and Saying. Essays for Richard Cartwright}}, pages
-  241--261. {MIT Press}, 1987.
-
-\bibitem[\protect\citeauthoryear{Sobel}{2004}]{sobel2004logic}
-J.H. Sobel.
-\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambridge U. Press, 2004.
-
-\bibitem[\protect\citeauthoryear{Sultana and Benzm{\"u}ller}{2013}]{W47}
-N.~Sultana and C.~Benzm{\"u}ller.
-\newblock Understanding {LEO-II's} proofs.
-\newblock In Konstantin Korovin, Stephan Schulz, and Eugenia Ternovska,
-  editors, {\em IWIL 2012}, volume~22 of {\em EPiC Series}, pages 33--52,
-  Merida, Venezuela, 2013. EasyChair.
-
-\bibitem[\protect\citeauthoryear{Sutcliffe and Benzm{\"u}ller}{2010}]{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller.
-\newblock Automated reasoning in higher-order logic using the {TPTP THF}
-  infrastructure.
-\newblock {\em J. of Formalized Reasoning}, 3(1):1--27, 2010.
-
-\bibitem[\protect\citeauthoryear{Williamson}{2013}]{williamson13}
-T.~Williamson.
-\newblock {\em Modal Logic as Metaphysics}.
-\newblock Oxford:OUP, 2013.
-
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{Bibliography}
 

### 2016-06-24 Papers/2016/KI/main.tex
update

@@
 
-\title{The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics}
+\title{The Inconsistency in G\"odel's Ontological Argument: A Success
+  Story for Automated Theorem Proving in Metaphysics}
 \author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
@@
 \begin{abstract}
-  This paper discusses the discovery of the inconsistency in G\"odel's
+  We discusses the discovery of the inconsistency in G\"odel's
   ontological argument as a success story for artificial
@@
   % tools by philosophers.
-  This paper is short version our contribution to IJCAI 2016.
+
+  (This paper is a short version of our contribution to IJCAI 2016.)
 \end{abstract}
@@
 
-The main novel contribution reported here is a detailed
+The main novel contribution is a detailed
 analysis (in various modal logics) of the inconsistency of G\"{o}del's
@@
 problem since the first detection of this inconsistency by
-Benzm\"uller and Woltzenlogel-Paleo \cite{C40} with the
-\textsc{Leo-II} prover.  The verified refutation (discussed in
+the
+\textsc{Leo-II} prover \cite{C40}.  The verified refutation (discussed in
 \S\ref{sec:inconsistency}) displays a surprisingly accessible
@@
 inconsistency in combination with the work reported here thus
-demonstrates that artificial intelligence systems -- particularly
-higher-order automated theorem provers -- are capable of assisting in
+demonstrates that 
+higher-order automated theorem provers are capable of assisting in
 the discovery and elucidation of \emph{new} and philosophically
@@
 
-\subsection{Related Work}
-
+\paragraph{Related Work.}
 First successful applications of theorem proving technology in
@@
 extra conjunct in the definition of \emph{essence} (\emph{ess}). G\"odel's
-different definition of essence can be seen either in his manuscript
-(Fig.~\ref{GoedelScript}) or, in more modern notation, in the Isabelle
+different definition of essence can be seen 
+% either in his manuscript
+% (Fig.~\ref{GoedelScript}) or, in more modern notation,
+in the Isabelle
 formalization shown in Fig.~\ref{InconsistencyIsabelleK}. For Scott,
@@
 G\"odel's omission has been
-considered inessential and merely an oversight by many. For instance,
-Hazen [\cite[p.365]{Hazen}] states that ``G\"odel left this
-clause out [\ldots] but this appears to have been an oversight''.
+considered inessential and merely an oversight by many. 
+% For instance,
+% Hazen [\cite[p.365]{Hazen}] states that ``G\"odel left this
+% clause out [\ldots] but this appears to have been an oversight''.
 % -- it is
@@
 
-Logic textbooks % \cite{ToDo:which}
-commonly utilize higher-order logic in an informal/semi-formal way as
-a meta-language to introduce the syntax and the semantics of object
-logics of interest, in which reasoning problems in concrete
-application domains can be modeled and solved with pen and paper. In
-fact, this approach can also be followed on the computer (using HOL as
-a formal meta-language) for even very challenging object logics (such
-as HOML) to enable interactive and automated theorem proving with
-existing theorem provers for HOL.
+% Logic textbooks % \cite{ToDo:which}
+% commonly utilize higher-order logic in an informal/semi-formal way as
+% a meta-language to introduce the syntax and the semantics of object
+% logics of interest, in which reasoning problems in concrete
+% application domains can be modeled and solved with pen and paper. In
+% fact, this approach can also be followed on the computer (using HOL as
+% a formal meta-language) for even very challenging object logics (such
+% as HOML) to enable interactive and automated theorem proving with
+% existing theorem provers for HOL.
 
@@
 
-For a computational analysis of G\"odel's ontological argument, the embedding of HOMLs such as \textbf{K},
-\textbf{KB} and \textbf{S5} with various domain conditions (possibilist and actualist quantification)
-is required. This idea has been successfully employed in related work
-\cite{C40}. The embedding of HOML is in fact
-straightforward. Formulas in HOML are \emph{lifted}, i.e., converted into predicates
-over worlds, which are themselves explicitly represented as
-terms. The logical constants of HOML are translated to HOL terms in such a way that, for instance, 
+For a computational analysis of G\"odel's ontological argument, we
+utilised an embedding of HOMLs, such as \textbf{K}, \textbf{KB} and
+\textbf{S5} with various domain conditions (possibilist and actualist
+quantification), in HOL. This idea has already been successfully
+employed in earlier work \cite{C40}. The embedding of HOML is in fact
+straightforward. Formulas in HOML are \emph{lifted}, i.e., converted
+into predicates over worlds, which are themselves explicitly
+represented as terms. The logical constants of HOML are translated to
+HOL terms in such a way that, for instance,
 %$\neg \varphi$, $\varphi\vee\psi$
@@
 $\exists w. (r w_0 w) \wedge (\varphi w)$. This form of embedding is
-precisely the well-known standard translation
-\cite{DBLP:journals/logcom/Ohlbach91}, which is here intra-logically
-realized --- and extended for quantifiers --- in HOL by stating a set
-of equations defining the logical constants (Fig.~\ref{QML_S5U}). The
-resulting object logic is the HOML \textbf{K} with rigid terms and constant
-domains (possibilist quantifiers). Other logics (e.g. \textbf{KB}, \textbf{S5}) are
+precisely the well-known standard translation,
+%\cite{DBLP:journals/logcom/Ohlbach91}, 
+which is here intra-logically realized --- and extended for
+quantifiers --- in HOL by stating a set of equations defining the
+logical constants. The resulting object logic is
+the HOML \textbf{K} with rigid terms and constant domains (possibilist
+quantifiers). Other logics (e.g. \textbf{KB}, \textbf{S5}) are
 embedded by adding axioms that restrict the accessibility relation
-$r$. Varying domains and actualist quantifiers can be simulated by using
- an existence predicate to guard the
-quantifiers. The embedding approach is, therefore, very flexible.
+$r$. Varying domains and actualist quantifiers can be simulated by
+using an existence predicate to guard the quantifiers. The embedding
+approach is, therefore, very flexible.
 
@@
 
-In the typical workflow during an attempt to prove a conjecture with a theorem prover, it is customary to check the consistency of the axioms first.
-For if the axioms are inconsistent, anything (including the conjecture) would be trivially derivable in classical logic (\emph{ex falso quodlibet}).
-Surprisingly, when this routine check was performed on G\"odel's axioms \cite{C40}, the \textsc{Leo-II} prover claimed that the axioms were inconsistent.
-Unfortunately, the refutation generated by \textsc{Leo-II} was barely human-readable.
-The text file was 153 lines\footnote{Long lines with an average of 184 characters per line.} long and used machine-oriented calculus (higher-order resolution \cite{W47}) and syntax (TPTP THF \cite{J22}).
-Part of the file is displayed in Fig.~\ref{LEO-Proof}.
+In the typical workflow during an attempt to prove a conjecture with a
+theorem prover, it is customary to check the consistency of the axioms
+first. For if the axioms are inconsistent, anything (including the
+conjecture) would be trivially derivable in classical logic (\emph{ex
+  falso quodlibet}). Surprisingly, when this routine check was
+performed on G\"odel's axioms \cite{C40}, the \textsc{Leo-II} prover
+claimed that the axioms were inconsistent. Unfortunately, the
+refutation generated by \textsc{Leo-II} was barely human-readable. The
+text file was 153 lines\footnote{Long lines with an average of 184
+  characters per line.} long and used machine-oriented calculus
+(higher-order resolution \cite{W47}) and syntax (TPTP THF
+\cite{J22}).
 
-\begin{figure*}
-\centerline{\includegraphics[width=\textwidth]{./LEO-Proof.png}}
-\caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
-  substitutions (e.g. with the empty property) are highlighted. In the
-red part (see $\boldsymbol{\leftarrow}$), property variable $\texttt{SV8}$ has been instantiated with the
-$\lambda \texttt{SV16}_{\mu}. \lambda \texttt{SV17}_{\iota}. \bot$, i.e., the (lifted) empty property.
-} \label{LEO-Proof}
-\end{figure*}
+% Part of the file is displayed in Fig.~\ref{LEO-Proof}.
+
+% \begin{figure*}
+% \centerline{\includegraphics[width=\textwidth]{./LEO-Proof.png}}
+% \caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
+%   substitutions (e.g. with the empty property) are highlighted. In the
+% red part (see $\boldsymbol{\leftarrow}$), property variable $\texttt{SV8}$ has been instantiated with the
+% $\lambda \texttt{SV16}_{\mu}. \lambda \texttt{SV17}_{\iota}. \bot$, i.e., the (lifted) empty property.
+% } \label{LEO-Proof}
+% \end{figure*}
 
@@
 
-\subsection{Informal Argument} \label{sec:arg1}
+\paragraph{Informal Argument.} \label{sec:arg1}
 Based on the hints found in \textsc{Leo-II}'s refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
@@
 \imp {\nec} \allq y (\phi(y) \imp \psi(y)))}$) it follows that the
-empty property (or self-difference) is an essence of every individual (\textbf{Empty Essence Lemma}): 
-$$\allq x\; (\ess{\emptyset}{x})$$
+empty property (or self-difference) is an essence of every individual
+(\textbf{Empty Essence Lemma}): \hfill $\allq x\; (\ess{\emptyset}{x})$
 
@@
   exemplified}: ${\allq \phi} [P(\phi) \imp {\pos}  \exq x
-  \phi(x)]$) and axiom A5 (``necessary existence'' is a positive property: $P(\NE)$ ), it follows that $\NE$ is possibly exemplified:
-  $$
-  \pos \exq x [\NE(x)]
-  $$
+  \phi(x)]$) and axiom A5 (``necessary existence'' is a positive
+  property: $P(\NE)$ ), it follows that $\NE$ is possibly exemplified:
+  \hfill $  \pos \exq x [\NE(x)] $
 
@@
   (${\NE(x) \equiv \allq \phi [\ess{\phi}{x} \imp \nec \exq y
-    \phi(y)]}$), the following is obtained:
-  $$
-  \pos \exq x [\allq \varphi [ \ess{\varphi}{x} \imp \nec \exq y [\varphi(y)] ] ]
-  $$
+    \phi(y)]}$), the following is obtained: \hfill $  \pos \exq x
+  [\allq \varphi [ \ess{\varphi}{x} \imp \nec \exq y [\varphi(y)] ] ] $
 
 \item The sentence above holds for all $\varphi$ and thus, in
-  particular, for the empty property (or self-difference):
-$$
-\pos \exq x [ \ess{\emptyset}{x} \imp \nec \exq y [\emptyset(y)] ]
-$$
+  particular, for the empty property (or self-difference): \hfill $\pos \exq x [ \ess{\emptyset}{x} \imp \nec \exq y [\emptyset(y)] ]$
 
-\item By the Empty Essence Lemma, the antecedent of the implication above is valid. Therefore, the sentence above entails:
-$$
-\pos \exq x [ \nec \exq y [\emptyset(y)] ]
-$$
+\item By the Empty Essence Lemma, the antecedent of the implication
+  above is valid. Therefore, the sentence above entails: \hfill $\pos \exq x [ \nec \exq y [\emptyset(y)] ]$
 
-\item By definition of $\emptyset$: 
-$$
-\pos \exq x [ \nec \bot ]
-$$
+\item By definition of $\emptyset$: \hfill $\pos \exq x [ \nec \bot ]$
 
-\item As the existential quantifier is binding no variable within its scope, the sentence is equi-valid with:
-$$\pos \nec \bot $$
+\item As the existential quantifier is binding no variable within its
+  scope, the sentence is equi-valid with: \hfill $\pos \nec \bot $
 
@@
 
-\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
-\begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./InconsistencyIsabelleK.png}}
-\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
-\end{figure}
+\paragraph{Argument Reconstruction in Isabelle.}  \label{sec:arg2}
 \begin{figure*}[t]
@@
 
-\subsection{Mapping the Inconsistency to G\"odel}
-
+\paragraph{Mapping the Inconsistency to G\"odel.}
 The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
@@
 
-\subsection{Inconsistency of G\"odel's Axioms in \SFiveU}
-
+\paragraph{Inconsistency of G\"odel's Axioms in \SFiveU.}
 Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
@@
 
-\pagebreak
+%\pagebreak
 %A discussion between Chad Brown and Benzm\"uller

### 2016-07-08 Papers/2016/KI/main.tex
update

@@
 \title{The Inconsistency in G\"odel's Ontological Argument: A Success
-  Story for Automated Theorem Proving in Metaphysics}
+  Story for Artificial Intelligence in Metaphysics (Extended Abstract)}
 \author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
@@
   ontological argument as a success story for artificial
-  intelligence. Despite the popularity of the argument since the
-  appearance of G\"odel's manuscript in the early 1970's, the
-  inconsistency of the axioms used in the argument remained unnoticed
-  until 2013, when it was detected automatically by the higher-order
-  theorem prover \textsc{Leo-II}.  Understanding and verifying the
-  refutation generated by the prover turned out to be a time-consuming
-  task. Its completion, as reported here, required the
-  reconstruction of the refutation in the Isabelle proof assistant.% , and it
+  intelligence. Despite the popularity of G\"odel's argument this
+  inconsistency remained unnoticed until 2013, when it was detected
+  automatically by the higher-order theorem prover
+  \textsc{Leo-II}.  % Understanding and verifying the
+  % refutation generated by the prover turned out to be a time-consuming
+  % task. Its completion, as reported here, required the
+  % reconstruction of the refutation in the Isabelle proof assistant.% , and it
   % also led to a novel and more efficient way of automating
@@
 
-  (This paper is a short version of our contribution to IJCAI 2016.)
+  (This text is a significantly shortened version of our contribution to IJCAI 2016.)
 \end{abstract}
@@
 \section{Introduction}\label{sec:introduction}
-Without exaggeration Kurt G\"{o}del's ontological
+% Without exaggeration 
+Kurt G\"{o}del's ontological
 argument for the existence of God \cite{GoedelNotes,ScottNotes} is
@@
 work (see e.g.~Sobel~\cite{sobel2004logic} and Oppy~\cite{sep-ontological-arguments} and the
-references therein).  In philosophy lectures at universities the
-argument is regularly presented as a masterpiece argument in
-metaphysics. Since 2013, when we~\cite{J30,C40} first
-reported our successful initial computer-assisted
-analysis of G\"odel's proof and Scott's variant,
-our work has received a media repercussion on a global scale\footnote{A
-  collection of news articles is available at {\scriptsize
-    \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
-and numerous bloggers commented on the proof
-\cite{fuhrmann15:_blogg_goedel}.
+references therein).
 
-The in-depth analysis presented here substantially
-extends previous computer-assisted studies of G\"odel's ontological
-argument. Similarly to the related work \cite{J30,C40} the analysis has
-been conducted with automated theorem provers for classical
-higher-order logic (HOL; cf.~\cite{andrewsSEP} and the references
-therein), even though G\"odel's proof is actually formulated in
-higher-order \emph{modal} logic (HOML; cf.~\cite{homl} and the
-references therein). To bridge between the two logics we utilise and
-further improve the logic embedding approach \cite{J23,C40}, which has
-already been employed successfully in our preceding work.
+ % In philosophy lectures at universities the
+% argument is regularly presented as a masterpiece argument in
+% metaphysics. Since 2013, when we~\cite{J30,C40} first
+% reported our successful initial computer-assisted
+% analysis of G\"odel's proof and Scott's variant,
+% our work has received a media repercussion on a global scale\footnote{A
+%   collection of news articles is available at {\scriptsize
+%     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
+% and numerous bloggers commented on the proof
+% \cite{fuhrmann15:_blogg_goedel}.
 
-The main novel contribution is a detailed
-analysis (in various modal logics) of the inconsistency of G\"{o}del's
-original version of the axioms used in his manuscript
-\cite{GoedelNotes}. The extraction, reconstruction and
-verification of an informal, human intuitive argument has been an open
-problem since the first detection of this inconsistency by
-the
-\textsc{Leo-II} prover \cite{C40}.  The verified refutation (discussed in
-\S\ref{sec:inconsistency}) displays a surprisingly accessible
-explanation of the inconsistency, which is philosophically profound
-and never presented in the literature. The detection of this
-inconsistency in combination with the work reported here thus
-demonstrates that 
-higher-order automated theorem provers are capable of assisting in
-the discovery and elucidation of \emph{new} and philosophically
-relevant knowledge.
+However, G\"odel's original version of the axioms used in his
+manuscript \cite{GoedelNotes} are inconsistent, and this fact has
+remained unnoticed to philosophers until 2013, when it was detected
+automatically by the higher-order theorem prover \textsc{Leo-II}
+\cite{C40}.  We here report on the extraction, reconstruction and
+verification of an informal, human intuitive argument for this
+inconsistency in the Isabelle/HOL proof assistant.  The respective
+presentation
+of this intuitive argument further extends and improves our contribution 
+to IJCAI 2016 \cite{C55}.
+
+% The verified refutation (discussed
+% in \S\ref{sec:inconsistency}) displays a surprisingly accessible
+% explanation of the inconsistency, which is philosophically profound
+% and never presented in the literature. 
+% The detection of this
+% inconsistency in combination with the work reported here thus
+% demonstrates that higher-order automated theorem provers are capable
+% of assisting in the discovery and elucidation of \emph{new} and
+% philosophically relevant knowledge.
 
@@
 
-\paragraph{Related Work.}
-First successful applications of theorem proving technology in
-metaphysics were reported by Fitelson, Oppenheimer and
-Zalta~\cite{FitelsonZalta,oppenheimer11}, who coined the term \textit{Computational Metaphysics} for this new research area and employed the first-order
-\textsc{Prover9} \cite{prover9-mace4} in their experiments. Later on,
-Rushby~\cite{rushby13} used the proof assistant \textsc{PVS} \cite{cade92-pvs}. Common to both
-works is a significant amount of proof-hand-coding work as well as their
-focus on a non-modal formalization of St. Anselm's~\cite{Proslogion} simpler 
-and older ontological argument. In contrast, the greater complexity of G\"odel's argument requires the formalization and automation of variants of \emph{higher-order} and \emph{modal} logics.
+In our experiments we have employed theorem provers for classical
+higher-order logic (HOL; cf.~\cite{andrewsSEP} and the references
+therein), even though G\"odel's proof is actually formulated in
+higher-order \emph{modal} logic (HOML; cf.~\cite{homl} and the
+references therein). To bridge between the two logics we adopt the
+logic embedding approach \cite{J23,C40}.
 
-\begin{wrapfigure}[22]{r}{0.55\textwidth} \centering \vskip-3em
-  \includegraphics[width=.52\textwidth]{./Scott_S5U.png}
-  \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
-\end{wrapfigure} 
-\section{A Brief History}\label{sec:history}
-St. Anselm's ontological argument \cite{Proslogion} can be regarded as
-the ancestor of modern ontological arguments such as G\"odel's. In the
-millenium between Anselm and G\"odel, many philosophers modified and
-arguably improved Anselm's argument. Of particular importance to
-G\"odel was the work of Leibniz~\cite{Adams}.  
-Although G\"odel's notion of positive property is not exactly the same
-as Leibniz's notion of perfection, G\"odel's manuscript
-(Fig.~\ref{GoedelScript}) can be considered a translation of Leibniz's
-presentation of the argument into modern modal logic. G\"odel
+%\paragraph{Related Work.}
+Applications of (first-order) theorem proving technology in
+metaphysics were first reported by Fitelson, Oppenheimer and
+Zalta~\cite{FitelsonZalta,oppenheimer11}. Zalta also coined the term
+\textit{Computational Metaphysics}. Later on, Rushby~\cite{rushby13} used the proof
+assistant \textsc{PVS}. Common to both works is a
+significant amount of proof-hand-coding work as well as their focus on
+a non-modal formalization of St. Anselm's simpler
+and older ontological argument. % In contrast, the greater complexity of
+% G\"odel's argument requires the formalization and automation of
+% variants of \emph{higher-order} and \emph{modal} logics.
+
+\begin{figure}\centering 
+  \includegraphics[width=.495\textwidth]{./Scott_S5U.png}
+  \includegraphics[width=.495\textwidth]{./InconsistencyIsabelleK.png}
+  % \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
+  % Axioms;  Automatic Verification of Modal Collapse} 
+\caption{Scott's consistent axioms (left) and proof of the
+  inconsistency of (a subset of) G\"odel's original  axioms (right)}
+  % Axioms;  Automatic Verification of Modal Collapse} 
+\label{Scott_Goedel}
+\end{figure} 
+\paragraph{A Brief History.}\label{sec:history}
+% St. Anselm's ontological argument \cite{Proslogion} can be regarded as
+% the ancestor of modern ontological arguments such as G\"odel's. In the
+% millenium between Anselm and G\"odel, many philosophers modified and
+% arguably improved Anselm's argument. Of particular importance to
+% G\"odel was the work of Leibniz~\cite{Adams}.  
+% Although G\"odel's notion of positive property is not exactly the same
+% as Leibniz's notion of perfection, 
+G\"odel's manuscript can be considered a translation of Leibniz's
+ideas on the argument into modern modal logic. G\"odel
 discussed his manuscript with Scott, who shared a slightly different
@@
 definitions, formalized in Isabelle, is shown in
-Fig.~\ref{Scott_S5U}. 
+Fig.~\ref{Scott_Goedel}. 
 The main difference to G\"odel's version is an
-extra conjunct in the definition of \emph{essence} (\emph{ess}). G\"odel's
-different definition of essence can be seen 
-% either in his manuscript
-% (Fig.~\ref{GoedelScript}) or, in more modern notation,
-in the Isabelle
-formalization shown in Fig.~\ref{InconsistencyIsabelleK}. For Scott,
+extra conjunct in the definition of \emph{essence} (\emph{ess}); cf. Fig.~\ref{Scott_Goedel}. For Scott,
 an essential property of an individual must be possessed by
@@
 
-\begin{wrapfigure}[20]{r}{0.55\textwidth}
-    \centering \vskip-1em
-    \centerline{\includegraphics[width=0.52\textwidth]{./InconsistencyIsabelleK.png}}
- \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
-\end{wrapfigure}
+% \begin{wrapfigure}[20]{r}{0.55\textwidth}
+%     \centering \vskip-1em
+%     \centerline{\includegraphics[width=0.52\textwidth]{./InconsistencyIsabelleK.png}}
+%  \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
+% \end{wrapfigure}
 G\"odel's omission has been
@@
 despite numerous analysis and criticisms of the
-argument. Especially since the discovery by Sobel
-\cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
-  $\phi\rightarrow \Box \phi$, states that contingent truth implies
-  necessary truth; it can be interpreted as \textit{everything is
-    pre-determined} or even \textit{there is no free will}. } is
-entailed by G\"odel's (or also Scott's) axioms, several variants have
-been proposed
-\cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
-attempting to avoid the modal collapse. Many of these variants omit
-the crucial conjunct in the definition of essence as well.\footnote{As
-  these variants also change other axioms, on which the inconsistency
-  of G\"odel's axioms depends, it is not necessarily the case that
-  these variants are also inconsistent; they must be analyzed
-  separately.} Opponents of the argument
-(e.g. Oppy [\cite[p.226--227]{oppy96:_goedel_ontol_argum};
-\cite[p.364]{oppy00:_respon_gettin};
-\cite[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
-parodies and other criticisms, referring to variants where the
-conjunct is omitted.
+argument.%  Especially since the discovery by Sobel
+% \cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
+%   $\phi\rightarrow \Box \phi$, states that contingent truth implies
+%   necessary truth; it can be interpreted as \textit{everything is
+%     pre-determined} or even \textit{there is no free will}. } is
+% entailed by G\"odel's (or also Scott's) axioms, several variants have
+% been proposed
+% \cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
+% attempting to avoid the modal collapse. Many of these variants omit
+% the crucial conjunct in the definition of essence as well.\footnote{As
+%   these variants also change other axioms, on which the inconsistency
+%   of G\"odel's axioms depends, it is not necessarily the case that
+%   these variants are also inconsistent; they must be analyzed
+%   separately.} Opponents of the argument
+% (e.g. Oppy [\cite[p.226--227]{oppy96:_goedel_ontol_argum};
+% \cite[p.364]{oppy00:_respon_gettin};
+% \cite[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
+% parodies and other criticisms, referring to variants where the
+% conjunct is omitted.
 
@@
 fact crucial. Without it, G\"odel's original axioms are
-inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_S5U}
+inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_Goedel}
 where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
@@
 
-For a computational analysis of G\"odel's ontological argument, we
-utilised an embedding of HOMLs, such as \textbf{K}, \textbf{KB} and
+In our experiments in this branch of metaphysics we
+utilise an embedding of HOMLs, such as \textbf{K}, \textbf{KB} and
 \textbf{S5} with various domain conditions (possibilist and actualist
-quantification), in HOL. This idea has already been successfully
-employed in earlier work \cite{C40}. The embedding of HOML is in fact
-straightforward. Formulas in HOML are \emph{lifted}, i.e., converted
+quantification), in HOL. More precisely, formulas in HOML are \emph{lifted}, i.e., converted
 into predicates over worlds, which are themselves explicitly
@@
 
+\textbf{Bruno: Here we should present the syntactical argument for the
+  inconsistency of $\pos \nec \bot $.}
+
+\section{Conclusion}\label{sec:conclusion}
+
+The axioms and definitions in G\"odel's manuscript are inconsistent;
+this was detected automatically by the prover
+\textsc{Leo-II}. We presented a rational reconstruction and
+verification of the inconsistency argument in Isabelle/HOL. This
+argument is valid in all normal HOMLs including base logic \K.
+
+Extending the presentation in our IJCAI 2016 paper \cite{C55}, we have added
+a syntactical proof for the inconsistency of  $\pos \nec \bot$.  Previously we had reduced the inconsistency
+analysis to exactly this proposition and we had then argued
+semantically that this proposition must be false already in modal
+logic \K.
+
+% We have also presented several technical improvements regarding the
+% semantic embedding approach. In particular, we have achieved a
+% nearly perfect match between pen and paper presentations in HOML and
+% the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
+% is now fully transparent, more user-friendly and ready for wider adoption.
+
+% On the other hand, there is still room for many pragmatical
+% improvements in Isabelle; just one example: in default setting,
+% Sledgehammer does not immediately inform the user when a proof has
+% been found and instead silently first executes a series of
+% time-consuming proof analysis processes (e.g. its dependency
+% minimization), before it eventually reports success. For G\"odel's
+% theorem T3 (\textit{Necessarily, there exists God}), for example, this
+% phase of silence takes several minutes --- during which the user might
+% actually give up on the proof attempt --- even though \textsc{Leo-II} already
+% reported success to Sledgehammer after 2.5 seconds.
+
+Our work reveals a challenge for automated reasoning:
+the (so far partially manual) extraction of an informal argument from a formal proof. 
+Without accompanying human-understandable explanations,
+the proofs generated by provers such as \textsc{Leo-II} or Metis, will
+presumably be only of limited value for philosophers, for whom intuitive
+arguments remain crucial for the acceptance of novel results.
+
+% Another open problem that we solved in this paper is a fully automatic
+% proof of T3 directly from Scott's axioms. Again, this proof was
+% contributed by \textsc{Leo-II}. This has become possible only
+% after we provided a more efficient embedding for HOML \SFiveU (instead of \SFive) in HOL.
+
+% Both the automated detection of the inconsistency in G\"odel's axioms
+% and the fully automatic proof of T3 from Scott's axioms
+Our work 
+demonstrates
+the potential of higher-order automated theorem proving and the
+semantic embeddings approach for philosophy: this technology is,
+in its current state of development, already capable of contributing novel results to
+metaphysics and to conduct reasoning steps at granularity-levels
+beyond common human capabilities.  
+\vfill
+
+% \noindent\textbf{Acknowledgments:} We thank Chad Brown, who contributed to the rational  reconstruction
+% of the inconsistency argument.
+
+%\pagebreak
+%A discussion between Chad Brown and Benzm\"uller
+% significantly influenced the rational reconstruction of the inconsistency reported by \textsc{LEO-II}.
+
+%\pagebreak
+
+%German Research Foundation DFG, Chad Brown
+
+%ToDo: reduce paper length to 7 pages
+
+%ToDo: use \textsc for all systems. Not only for LEO-II.
+
+%\small 
+%% The file named.bst is a bibliography style file for BibTeX 0.99c
+
+\bibliographystyle{plain}
+\bibliography{Bibliography}
+
+\pagebreak
+
 \paragraph{Argument Reconstruction in Isabelle.}  \label{sec:arg2}
@@
 
-\section{Conclusion}\label{sec:conclusion}
-
-The axioms and definitions in G\"odel's manuscript are inconsistent;
-this was detected automatically by the prover
-\textsc{Leo-II}. Here we presented a rational reconstruction and
-verification of the inconsistency argument in Isabelle/HOL. This
-argument is valid in all normal HOMLs including base logic \K.
-
-We have also presented several technical improvements regarding the
-semantic embedding approach. In particular, we have achieved a
-nearly perfect match between pen and paper presentations in HOML and
-the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
-is now fully transparent, more user-friendly and ready for wider adoption.
-
-On the other hand, there is still room for many pragmatical
-improvements in Isabelle; just one example: in default setting,
-Sledgehammer does not immediately inform the user when a proof has
-been found and instead silently first executes a series of
-time-consuming proof analysis processes (e.g. its dependency
-minimization), before it eventually reports success. For G\"odel's
-theorem T3 (\textit{Necessarily, there exists God}), for example, this
-phase of silence takes several minutes --- during which the user might
-actually give up on the proof attempt --- even though \textsc{Leo-II} already
-reported success to Sledgehammer after 2.5 seconds.
-
-More importantly, our work reveals a challenge for automated reasoning:
-the (so far partially manual) extraction of an informal argument from a formal proof. 
-Without accompanying human-understandable explanations,
-the proofs generated by provers such as \textsc{Leo-II} or Metis, will
-presumably be only of limited value for philosophers, for whom intuitive
-arguments remain crucial for the acceptance of novel results.
-
-Another open problem that we solved in this paper is a fully automatic
-proof of T3 directly from Scott's axioms. Again, this proof was
-contributed by \textsc{Leo-II}. This has become possible only
-after we provided a more efficient embedding for HOML \SFiveU (instead of \SFive) in HOL.
-
-Both the automated detection of the inconsistency in G\"odel's axioms
-and the fully automatic proof of T3 from Scott's axioms demonstrate
-the potential of our AI technology for philosophy: this technology is,
-in its current state of development, already capable of contributing novel results to
-metaphysics and to conduct reasoning steps at granularity-levels
-beyond common human capabilities.  
-\vfill
-
-\noindent\textbf{Acknowledgments:} We thank Chad Brown, who contributed to the rational  reconstruction
-of the inconsistency argument.
-
-%\pagebreak
-%A discussion between Chad Brown and Benzm\"uller
-% significantly influenced the rational reconstruction of the inconsistency reported by \textsc{LEO-II}.
-
-%\pagebreak
-
-%German Research Foundation DFG, Chad Brown
-
-%ToDo: reduce paper length to 7 pages
-
-%ToDo: use \textsc for all systems. Not only for LEO-II.
-
-%\small 
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-
-\small
-\bibliographystyle{plain}
-\bibliography{Bibliography}
-
 \end{document}

### 2016-07-09 Papers/2016/KI/main.tex
Improve introduction

@@
 
-\title{The Inconsistency in G\"odel's Ontological Argument: A Success
-  Story for Artificial Intelligence in Metaphysics (Extended Abstract)}
+\title{An Object-Logic Explanation for the Inconsistency in G\"odel's Ontological Theory}
 \author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
@@
 \begin{abstract}
-  We discusses the discovery of the inconsistency in G\"odel's
+  This paper discusses the discovery of the 
+  inconsistency in G\"odel's
   ontological argument as a success story for artificial
-  intelligence. Despite the popularity of G\"odel's argument this
-  inconsistency remained unnoticed until 2013, when it was detected
-  automatically by the higher-order theorem prover
-  \textsc{Leo-II}.  % Understanding and verifying the
-  % refutation generated by the prover turned out to be a time-consuming
-  % task. Its completion, as reported here, required the
-  % reconstruction of the refutation in the Isabelle proof assistant.% , and it
-  % also led to a novel and more efficient way of automating
-  % higher-order modal logic \textbf{S5} with a universal accessibility
-  % relation. Furthermore, the development of an improved syntactical
-  % hiding for the utilized logic embedding technique allows the refutation to be
-  % presented in a human-friendly way, suitable for non-experts in the
-  % technicalities of higher-order theorem proving. This brings us a
-  % step closer to wider adoption of logic-based artificial intelligence
-  % tools by philosophers.
+  intelligence. Despite the popularity of G\"odel's argument, this
+  inconsistency remained unnoticed until 2013, 
+  when it was detected automatically by the 
+  higher-order theorem prover \textsc{Leo-II} \cite{ECAI2014}.
+Complementing the meta-logic explanation for the inconsistency available in our IJCAI 2016 paper \cite{C55}, we present here a new purely object-logic explanation that does not rely on semantic argumentation.
 
-  (This text is a significantly shortened version of our contribution to IJCAI 2016.)
+  % (This extended abstract is mostly a shortened version of our IJCAI 2016 paper \cite{C55}, but the explanation of the inconsistency is more detailed here.)
 \end{abstract}
+
+% ToDo: Use \textsc for all system names
 
@@
 
- % In philosophy lectures at universities the
-% argument is regularly presented as a masterpiece argument in
-% metaphysics. Since 2013, when we~\cite{J30,C40} first
-% reported our successful initial computer-assisted
-% analysis of G\"odel's proof and Scott's variant,
-% our work has received a media repercussion on a global scale\footnote{A
-%   collection of news articles is available at {\scriptsize
-%     \url{https://github.com/FormalTheology/GoedelGod/blob/master/Press/LinksToNews.md}}},
-% and numerous bloggers commented on the proof
-% \cite{fuhrmann15:_blogg_goedel}.
+Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the Coq proof assistant \cite{CSR}.
+To bridge the gap between higher-order logics (HOL; cf.~\cite{andrewsSEP} and the references
+therein), as used by these systems, and higher-order \emph{modal} logics (HOML; cf.~\cite{homl} and the
+references therein), on which the ontological argument argument, the logic embedding approach \cite{J23,C40} was used.
 
-However, G\"odel's original version of the axioms used in his
-manuscript \cite{GoedelNotes} are inconsistent, and this fact has
-remained unnoticed to philosophers until 2013, when it was detected
-automatically by the higher-order theorem prover \textsc{Leo-II}
-\cite{C40}.  We here report on the extraction, reconstruction and
-verification of an informal, human intuitive argument for this
-inconsistency in the Isabelle/HOL proof assistant.  The respective
-presentation
-of this intuitive argument further extends and improves our contribution 
-to IJCAI 2016 \cite{C55}.
+However, G\"odel's original axioms, as used in his
+manuscript \cite{GoedelNotes}, are inconsistent. This fact has
+remained unnoticed to philosophers until 2013, 
+when \textsc{Leo-II} found a surprising refutation of the axioms.
 
-% The verified refutation (discussed
-% in \S\ref{sec:inconsistency}) displays a surprisingly accessible
-% explanation of the inconsistency, which is philosophically profound
-% and never presented in the literature. 
-% The detection of this
-% inconsistency in combination with the work reported here thus
-% demonstrates that higher-order automated theorem provers are capable
-% of assisting in the discovery and elucidation of \emph{new} and
-% philosophically relevant knowledge.
+In \cite{C55}, we have extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we have reconstructed and
+verified it in the Isabelle proof assistant.
+However, that explanation relied on reasoning at the meta-logic (HOL) level, which was only possible because of the embedding.
+Here we complement that work with a purely object-logic (HOML) explanation.
 
-% On the technical side, the quest for constructing a compelling
-% refutation, capable of convincing also human non-experts, led to an
-% improvement of the syntax of the embedding of modal logics in Isabelle/HOL
-% (as discussed in \S\ref{sec:improvedsyntax}). With the new syntax, a
-% (nearly) perfect match between the original pen and paper
-% presentations and our encoding in Isabelle/HOL is feasible. A more
-% user-friendly syntax, as reported here, is clearly an important
-% prerequisite for promoting the theorem proving technology employed
-% here to a wider community of philosophers, who are not necessarily
-% experts in automated reasoning or HOL.
-
-% Another novel contribution reported here (in
-% \S\ref{sec:improvedembedding}) is the implementation of an alternative
-% embedding for the 
-% % slightly stronger 
-% more effective modal logic \SFiveU, which is
-% based on a universal accessibility relation. Our experiments have
-% shown that the new embedding is more efficient, as the following two
-% previously open problems can now be solved:
-% \begin{itemize}
-% \item Automatically proving the final theorem T3 (\textit{Necessarily, there
-%   exists God}), directly from Scott's \cite{ScottNotes} (consistent) axioms
-%   alone, without relying on the argument's intermediate argumentation
-%   steps (i.e., lemmata).
-% \item Automatically verifying, in Isabelle/HOL, the proof of the modal
-%   collapse \cite{Sobel}, which is one of the most strongly criticized
-%   logical consequences of the argument's axioms.
-% \end{itemize}
-
-In our experiments we have employed theorem provers for classical
-higher-order logic (HOL; cf.~\cite{andrewsSEP} and the references
-therein), even though G\"odel's proof is actually formulated in
-higher-order \emph{modal} logic (HOML; cf.~\cite{homl} and the
-references therein). To bridge between the two logics we adopt the
-logic embedding approach \cite{J23,C40}.
-
-%\paragraph{Related Work.}
 Applications of (first-order) theorem proving technology in
 metaphysics were first reported by Fitelson, Oppenheimer and
-Zalta~\cite{FitelsonZalta,oppenheimer11}. Zalta also coined the term
-\textit{Computational Metaphysics}. Later on, Rushby~\cite{rushby13} used the proof
-assistant \textsc{PVS}. Common to both works is a
+Zalta~\cite{FitelsonZalta,oppenheimer11}. 
+%Zalta also coined the term
+%\textit{Computational Metaphysics}. 
+Later on, Rushby~\cite{rushby13} used the \textsc{PVS} proof
+assistant. Common to both works is a
 significant amount of proof-hand-coding work as well as their focus on
 a non-modal formalization of St. Anselm's simpler
-and older ontological argument. % In contrast, the greater complexity of
-% G\"odel's argument requires the formalization and automation of
-% variants of \emph{higher-order} and \emph{modal} logics.
+and older ontological argument.
 
@@
 \end{figure}
+
 \paragraph{A Brief History.}\label{sec:history}
@@
 
-% \begin{wrapfigure}[20]{r}{0.55\textwidth}
-%     \centering \vskip-1em
-%     \centerline{\includegraphics[width=0.52\textwidth]{./InconsistencyIsabelleK.png}}
-%  \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
-% \end{wrapfigure}
 G\"odel's omission has been
 considered inessential and merely an oversight by many. 
-% For instance,
-% Hazen [\cite[p.365]{Hazen}] states that ``G\"odel left this
-% clause out [\ldots] but this appears to have been an oversight''.
-% -- it is
-%included in related manuscripts''. 
 For more than four decades, its serious consequences remained unnoticed, 

### 2016-07-09 Papers/2016/KI/main.tex
Improve history and embedding sections

@@
 
-\paragraph{A Brief History.}\label{sec:history}
-% St. Anselm's ontological argument \cite{Proslogion} can be regarded as
-% the ancestor of modern ontological arguments such as G\"odel's. In the
-% millenium between Anselm and G\"odel, many philosophers modified and
-% arguably improved Anselm's argument. Of particular importance to
-% G\"odel was the work of Leibniz~\cite{Adams}.  
-% Although G\"odel's notion of positive property is not exactly the same
-% as Leibniz's notion of perfection, 
+\section{An Essential Difference in the Definitions of Essence.}
+\label{sec:history}
 G\"odel's manuscript can be considered a translation of Leibniz's
@@
 The main difference to G\"odel's version is an
-extra conjunct in the definition of \emph{essence} (\emph{ess}); cf. Fig.~\ref{Scott_Goedel}. For Scott,
+extra conjunct in the definition of \emph{essence} (\emph{ess}). For Scott,
 an essential property of an individual must be possessed by
@@
 For more than four decades, its serious consequences remained unnoticed, 
-despite numerous analysis and criticisms of the
-argument.%  Especially since the discovery by Sobel
-% \cite{Sobel} that modal collapse (MC)\footnote{The modal collapse,
-%   $\phi\rightarrow \Box \phi$, states that contingent truth implies
-%   necessary truth; it can be interpreted as \textit{everything is
-%     pre-determined} or even \textit{there is no free will}. } is
-% entailed by G\"odel's (or also Scott's) axioms, several variants have
-% been proposed
-% \cite{Anderson,AndersonGettings,Hajek1,Hajek2,Hajek3,Bjordal}
-% attempting to avoid the modal collapse. Many of these variants omit
-% the crucial conjunct in the definition of essence as well.\footnote{As
-%   these variants also change other axioms, on which the inconsistency
-%   of G\"odel's axioms depends, it is not necessarily the case that
-%   these variants are also inconsistent; they must be analyzed
-%   separately.} Opponents of the argument
-% (e.g. Oppy [\cite[p.226--227]{oppy96:_goedel_ontol_argum};
-% \cite[p.364]{oppy00:_respon_gettin};
-% \cite[p.1068]{oppy08:_higher_order_ontol_argum}]) have also proposed
-% parodies and other criticisms, referring to variants where the
-% conjunct is omitted.
-
-% \begin{figure*}
-%   \centering
-%   \begin{subfigure}[t]{0.4\textwidth}
-%     \includegraphics[width=\textwidth]{./Images/Scott_S5U.png}
-%     \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-%       Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
-%   \end{subfigure}
-%   ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-%   % (or a blank line to force the subfigure onto a new line)
-%   \begin{subfigure}[t]{0.4\textwidth}
-%     \includegraphics[width=\textwidth]{./Images/QML_S5U.png}
-%     \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
-%   \end{subfigure}
-%   ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-%   % (or a blank line to force the subfigure onto a new line)
-%   \begin{subfigure}[t]{0.4\textwidth}
-%     \includegraphics[width=\textwidth]{./Images/Inconsistency_S5U_direct.png}
-%     \caption{Inconsistency  in \SFiveU} \label{Inconsistency_S5U} 
-%   \end{subfigure}
-%   \begin{subfigure}[t]{0.4\textwidth}
-%     \centerline{\includegraphics[width=\textwidth]{./Images/InconsistencyIsabelleK.png}}
-%     \caption{Refutation of G\"odel's Axioms in Isabelle/HOL.} \label{InconsistencyIsabelleK}
-%   \end{subfigure}
-% \end{figure*}
-
-% \begin{figure}[t]{0.48\textwidth}
-% \centerline{\includegraphics[width=\textwidth]{./InconsistencyIsabelleK.png}}
-% \caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
-% \end{figure}
-
-% \begin{figure}[t]{0.48\textwidth}
-% \centerline{\includegraphics[width=\textwidth]{./Scott_S5U.png}}
-% \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
-%   Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
-% \end{figure}
-
-% There is reason to believe that the omission was more than just an oversight.
-As pointed out by Fuhrmann \cite{Fuhrmann2005}, ``G\"odel vermerkt diese Konsequenz [dass aus der Definition unmittelbar folgt, da\ss alle wesentlichen Eigenschaften notwendig äquivalent sind] in einer Fußnote zur Definition.
-Die Definition selbst aber l\"a{\ss}t im Definiens das Konjunkt $Xx$ aus.
-Ohne das Konjunkt folgt jedoch die Konsequenz nicht.
-Es ist deshalb naheliegend anzunehmen, da\ss G\"odel die Definition so beabsichtigte, wie sie hier notiert ist und wie G\"odel sie selbst in fr\"uheren Notizbucheintragungen formuliert hat.''\footnote{Our translation: ``G\"odel remarks this consequence [that the definition of essence entails that all essences are necessarily equivalent] in a footnote to the definition.
-The definition, however, omits the conjunct.
-...
-ToDo''}
-% \marginpar{ToDo: check Fuhrmann's claim. He may be pointing out a second mistake by Gödel.}
-
+despite numerous analyses and criticisms of the
+argument.
 However, as explained here, the extra conjunct is in
@@
 inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_Goedel}
-where the model finder Nitpick \cite{Nitpick} confirms consistency).\footnote{In
+where the model finder Nitpick \cite{Nitpick} confirms consistency). In
 personal communication, Dana Scott confirmed that he was unaware
-  at the time that G\"odel's axioms were inconsistent.}
+that G\"odel's original axioms were inconsistent.
 
@@
 $r$. Varying domains and actualist quantifiers can be simulated by
-using an existence predicate to guard the quantifiers. The embedding
-approach is, therefore, very flexible.
+using an existence predicate to guard the quantifiers. 
+%The embedding
+%approach is, therefore, very flexible.
 

### 2016-07-09 Papers/2016/KI/main.tex
Improve text about the inconsistency

@@
 claimed that the axioms were inconsistent. Unfortunately, the
-refutation generated by \textsc{Leo-II} was barely human-readable. The
-text file was 153 lines\footnote{Long lines with an average of 184
-  characters per line.} long and used machine-oriented calculus
-(higher-order resolution \cite{W47}) and syntax (TPTP THF
+refutation generated by \textsc{Leo-II} was barely human-readable. The refutation was based on machine-oriented inference rules (the higher-order resolution calculus \cite{W47}), and the
+text file had 153 lines (with an average of 184
+  characters per line) and used a machine-oriented syntax (TPTP THF
 \cite{J22}).
-
-% Part of the file is displayed in Fig.~\ref{LEO-Proof}.
-
-% \begin{figure*}
-% \centerline{\includegraphics[width=\textwidth]{./LEO-Proof.png}}
-% \caption{Lines 115--120 of \textsc{Leo-II}'s refutation. Primitive
-%   substitutions (e.g. with the empty property) are highlighted. In the
-% red part (see $\boldsymbol{\leftarrow}$), property variable $\texttt{SV8}$ has been instantiated with the
-% $\lambda \texttt{SV16}_{\mu}. \lambda \texttt{SV17}_{\iota}. \bot$, i.e., the (lifted) empty property.
-% } \label{LEO-Proof}
-% \end{figure*}
 
@@
 x\not=x$) have identical denotations in a logic setting
-with full functional and Boolean extensionality as given
+with functional and Boolean extensionality as assumed
 here. Nevertheless, some philosophers\footnote{Private communication with Andr\'e Fuhrmann.} may actually prefer the use of

### 2016-07-10 Papers/2016/KI/main.tex
Remove commented lines
Comment out appendix

@@
 
-%  Particularly interesting
-% from the perspective of this paper is that many literature
-% contributions unfortunately work with or refer to G\"odel's original
-% definition of essence
-% $$ todo $$
-% This original version avoids the conjunct added by Scott expressing
-% that essential properties of an individual should be possessed by the
-% individual.
-
-% We give here a list of examples:
-% \begin{itemize}
-% \item Anderson and Getting's 1996, p. 168 \cite[p.168]{AndersonGettings1968}
-%   use essence without conjunct.
-% \item  
-% \item Look???, p. 514
-% % \item Oppy 1996, p.226-227 \cite[p.226/227]{Oppy1996}, Oppy 2000, p. 364
-% %   \cite[p.364]{Oppy200}, Oppy 2008, p. 1068 \cite[p.1068]{Oppy2008}:
-% %   Oppy uses: ``A is an essence of x iff for every property B, x has B
-% %   neces- sarily iff A entails B'' (this is from Anderson's
-% %   emendation). Check is this leads to inconsistency as well. In this
-% %   case we may write something like:
-
-% %  Oppy presents an critical
-% %   assessment in order to conclude that G\"odel's argument fails to
-% %   convince. Oppy apparently does not see that he is assessing an
-% %   inconsistent axiom system (and he in fact creates adaptations that
-% %   suffer the same problems).
-% % \item 
-% \end{itemize}
-
 \section{Automating HOML in HOL}\label{sec:homlinhol}
-
-% Logic textbooks % \cite{ToDo:which}
-% commonly utilize higher-order logic in an informal/semi-formal way as
-% a meta-language to introduce the syntax and the semantics of object
-% logics of interest, in which reasoning problems in concrete
-% application domains can be modeled and solved with pen and paper. In
-% fact, this approach can also be followed on the computer (using HOL as
-% a formal meta-language) for even very challenging object logics (such
-% as HOML) to enable interactive and automated theorem proving with
-% existing theorem provers for HOL.
-
-% Just as commonly the case in logic textbooks, in the embeddings
-% approach classical higher-order logic is utilised as a meta-language
-% to encode the syntax and the semantics of object logics in which the
-% proof problems of interest are then modeled in. Modulo the embedding
-% state of the art automated theorem provers for classical higher-order
-% logic can then be utilised as object logic reasoning tools.
 
@@
 using an existence predicate to guard the quantifiers.
-%The embedding
-%approach is, therefore, very flexible.
-
-% \subsection{Improved Embedding}\label{sec:improvedembedding}
-% \begin{figure}[t]
-% \centerline{\includegraphics[width=1\columnwidth,height=10cm]{./QML_S5U.png}}
-% \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
-% \end{figure}
-
-% The modal logic \SFive requires that the accessibility relation be
-% reflexive, symmetric and transitive. The usual approach to embed
-% \SFive would be to use the standard translation for \textbf{K} described above
-% and to state that $r$ is an equivalence relation, e.g., by postulating
-% the following axioms:
-% \begin{itemize}
-% \item Reflexivity: $\forall x. (r~x~x)$
-% \item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
-% \item Transitivity: $\forall x. \forall y. \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
-% \end{itemize}
-% Instead, we consider here an alternative description, that we call \SFiveU, based on the following condition on $r$:
-% \begin{itemize}
-% \item Universality: $\forall x. \forall y. (r~x~y)$
-% \end{itemize}
-
-% It is important to note that $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi$ 
-% \cite{Blackburn}, and therefore \SFive and \SFiveU are traditionally considered to be 
-% two different descriptions of the same modal logic. Nevertheless, \SFive and \SFiveU 
-% differ in the shapes of frames they admit:
-% %It is easy to see that \SFiveU is at least as strong as \SFive:
-% %universality entails that $r$ is an equivalence relation. \SFiveU is,
-% %in fact, \emph{strictly stronger} than \SFive. 
-% \SFiveU only admits
-% complete\footnote{A graph is \emph{complete} iff there is a directed
-%   edge connecting every ordered pair of vertices.} frames, whereas
-% \SFive admits non-complete frames as long as all their components are
-% complete. In other words, in \SFiveU we face one single equivalence
-% class of possible worlds, while in \SFive we may face several
-% disconnected equivalence classes. In fact, for this reason, \SFiveU is
-% considered as metaphysically more appropriate by some philosophers;
-% cf. \cite[p.~127]{williamson13}.
-
-% Furthermore, for \SFiveU an improved embedding is possible. Universality implies
-% that the guarding predicates in the definitions of $\Box$ and
-% $\Diamond$ always hold. Therefore, they can be omitted and the
-% accessibility relation can be dispensed altogether. The modal
-% operators can then be defined merely as:
-% % \begin{itemize}
-% % \item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
-% % \item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
-% % \end{itemize}
-% \[\Box \varphi \equiv \lambda w. \forall v.  \varphi(v) \qquad \text{and}
-% \qquad \Dia \varphi \equiv \lambda w.\exists v. \varphi(v)\]
-
-% The new embedding of \SFiveU in Isabelle/HOL is shown in
-% Fig.~\ref{QML_S5U}.
-% With this improved embedding, the final theorem T3
-% (\textit{Necessarily, there exists God}) can be derived from
-% Scott's consistent version of the axioms fully automatically.  The fully
-% automatic proof has been generated (in about 2.5 seconds) by the theorem prover
-% \textsc{Leo-II}~\cite{leo2} and subsequently verified in the proof
-% assistant Isabelle/HOL~\cite{NPW02}, as shown in Fig.~\ref{Scott_S5U}.
-% The collaboration between the two systems has been orchestrated by
-% Isabelle's Sledgehammer tool \cite{Sledgehammer}.
-
-% With the embedding used by Benzm\"uller and Woltzenlogel-Paleo \cite{C40}, the
-% provers still had to be given the intermediate theorem T2 and the corollary
-% C in order to manage to prove T3.
-
-% Another evidence that the new embedding provides a significant performance boost 
-% is the successful automatic verification in Isabelle/HOL (with its automatic tactic Meson) of the modal
-% collapse \cite{Sobel}, which is one of the most strongly criticized
-% `side-effects' of G\"odel's and Scott's variants of the proof. In previous work \cite{C40}
-% the modal collapse has been proven by the higher-order provers
-% \textsc{Satallax} \cite{Satallax} and \textsc{Leo-II}, but a fully automatic
-% verification in the highly trusted Isabelle/HOL system still failed
-% \cite{J28}.  The success with the new embedding can be seen in Fig.~\ref{Scott_S5U}.
-
-% \subsection{Improved Syntax in Isabelle}\label{sec:improvedsyntax}
-
-% Wider adoption of HOL theorem proving technology for
-% reasoning about and within embedded object logics, especially among
-% non-expert users, is still hindered by the gap between the syntax used
-% by people, when they write logical formulas with pen and paper, and
-% the syntax used by HOL theorem provers. Even when the syntax of the
-% underlying higher-order system is elegant (as is the case in
-% Isabelle/HOL), the embedding of HOML into HOL may easily expose
-% details of HOL that may be uncommon to the user, disturbing his/her
-% experience while using the system. To illustrate this point,
-% Fig.~\ref{UglyEssence} shows how the definition of essence looked like
-% in previous work \cite{J28}, where advanced syntax-sugaring features
-% were not used. It looks notably higher-order, and its style differs
-% significantly from the common style seen in works on modal logics and
-% the ontological argument. The following specific issues can be
-% enumerated:
-% \begin{enumerate}
-% \item $\lambda$-abstractions, which are typically a HOL feature, appear explicitly in places where they did not need to in a pure HOML formulation (cf.
-G\"odel's manuscript, Fig.~\ref{GoedelScript}).
-% \item Quantifiers appear as higher-order defined constants, and not as binders.
-This forces the user to read (and write) formulas of the form $\forall (\lambda x.
-A(x))$ instead of the more common $\forall x.
-A(x)$.
-% \item The lifted modal connectives are represented by prefixing the
-%   letter ``m'' (e.g. $m\wedge$ and $m\imp$). The prefix disturbs the
-%   user, as it constantly reminds him/her that there is something
-%   unusual about the modal connectives.
-% \item Higher-order parenthesis conventions for the application of a predicate to a term are used. 
-% Instead of reading $\psi(y)$, as he/she would expect, he/she has to read $(\psi ~ y)$. Outside niche areas in computer science, the former syntax is more widely known than the latter.
-% \end{enumerate}
-
-% \begin{figure}
-% \centerline{\includegraphics[width=1\columnwidth]{./UglyEssence.png}}
-% \caption{Definition of Essence using Old Syntax} \label{UglyEssence}
-% \end{figure}
-
-% In the embedding presented here, in Fig.~\ref{QML_S5U}, advanced
-% syntax-sugaring effects provided by Isabelle were used to prevent
-% issues as those enumerated above.
-The possibility to define boldface connectives allows us to drop the prefix; ``\texttt{binder}'' annotations enable modal quantifiers to be used in the standard binding way and reduce the need for explicit lambda abstractions; and a careful choice of priorities for infix connectives gives the parenthesis conventions that are more familiar to the user.
-As desired, the definition of essence in Fig.~\ref{Scott_S5U} is undeniably more immediately recognizable and comprehensible than the definition in Fig.~\ref{UglyEssence}.
-The embedding technique is now completely transparent to the user.
-
-% The syntax improvements described here render the computer-assisted analysis of ontological arguments accessible to a wider audience and ease the adoption of logic-based artificial intelligence tools by philosophers interested in topics where modal logic reasoning is required.
 
@@
 
-\pagebreak
+% \pagebreak
 
-\paragraph{Argument Reconstruction in Isabelle.}  \label{sec:arg2}
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[t]{0.715\textwidth}
-    \includegraphics[width=\textwidth]{./Manuscript2.png}
-    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
-      definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript} 
-  \end{subfigure}
-   \begin{subfigure}[t]{0.28\textwidth}
-     \includegraphics[width=\textwidth,height=7.2cm]{./Inconsistency_S5U_direct.png}
-     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
-   \end{subfigure}
-% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
-% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
- \caption{The inconsistency in G\"{o}del's manuscript has been
-   detected and verified by HOL ATPs} 
-\end{figure*}
-To verify the correctness of the informal argument explained above, it
-was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
-  external provers such as \textsc{Leo-II} or Satallax, 
-  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
-inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
-the other hand, is explicitly stated, to ensure that Isabelle is
-reconstructing the same argument. In fact, without the help of this
-lemma, Metis is still not strong enough to refute G\"odel's
-axioms.
+% \paragraph{Argument Reconstruction in Isabelle.}  \label{sec:arg2}
+% \begin{figure*}[t]
+%   \centering
+%   \begin{subfigure}[t]{0.715\textwidth}
+%     \includegraphics[width=\textwidth]{./Manuscript2.png}
+%     \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
+%       definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript}
+%   \end{subfigure}
+%    \begin{subfigure}[t]{0.28\textwidth}
+%      \includegraphics[width=\textwidth,height=7.2cm]{./Inconsistency_S5U_direct.png}
+%      \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
+%    \end{subfigure}
+% % \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
+% % \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
+%  \caption{The inconsistency in G\"{o}del's manuscript has been
+%    detected and verified by HOL ATPs} 
+% \end{figure*}
+% To verify the correctness of the informal argument explained above, it
+% was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
+%   external provers such as \textsc{Leo-II} or Satallax, 
+%   constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
+% inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
+% the other hand, is explicitly stated, to ensure that Isabelle is
+% reconstructing the same argument. In fact, without the help of this
+% lemma, Metis is still not strong enough to refute G\"odel's
+% axioms.
 
-\paragraph{Mapping the Inconsistency to G\"odel.}
-The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
-essence (ess), the definition of necessary existence (NE), the
-axioms A1a and A2 (which entail theorem T1), and axiom A5. It remains to show that
-these ingredients are actually present in G\"odel's manuscript in
-Fig.~\ref{GoedelScript}.
+% \paragraph{Mapping the Inconsistency to G\"odel.}
+% The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
+% essence (ess), the definition of necessary existence (NE), the
+% axioms A1a and A2 (which entail theorem T1), and axiom A5. It remains to show that
+% these ingredients are actually present in G\"odel's manuscript in
+% Fig.~\ref{GoedelScript}.
 
-This can be easily seen: Axiom A1a in
-Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
-highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
-A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
-Fig.~\ref{GoedelScript}. The definitions of essence and necessary
-existence are easy to identify. Therefore, the verified
-inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
-G\"odel's original manuscript.
+% This can be easily seen: Axiom A1a in
+% Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
+% highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
+% A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
+% Fig.~\ref{GoedelScript}. The definitions of essence and necessary
+% existence are easy to identify. Therefore, the verified
+% inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
+% G\"odel's original manuscript.
 
-\paragraph{Inconsistency of G\"odel's Axioms in \SFiveU.}
-Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
-external provers such as \textsc{Leo-II}, still
-fails to detect the inconsistency of G\"odel's axioms in the standard
-embedding of \SFive, while a direct modeling of the problem in TPTP THF syntax
-in combination with a direct call of \textsc{Leo-II} succeeded. In
-other words, without independent experiments with no mediation through Sledgehammer, the
-inconsistency would not have been detected.
+% \paragraph{Inconsistency of G\"odel's Axioms in \SFiveU.}
+% Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
+% external provers such as \textsc{Leo-II}, still
+% fails to detect the inconsistency of G\"odel's axioms in the standard
+% embedding of \SFive, while a direct modeling of the problem in TPTP THF syntax
+% in combination with a direct call of \textsc{Leo-II} succeeded. In
+% other words, without independent experiments with no mediation through Sledgehammer, the
+% inconsistency would not have been detected.
 
-%\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
-On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
-the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
-the inconsistency could be detected
-by \textsc{Leo-II} also when called via
-Sledgehammer. Moreover, the result could subsequently be verified with
-Metis even without the Empty Essence Lemma (cf. Fig.~\ref{Inconsistency_S5U}). 
-% Interestingly, Metis' argument is different from that of
-% \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
+% %\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
+% On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
+% the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
+% the inconsistency could be detected
+% by \textsc{Leo-II} also when called via
+% Sledgehammer. Moreover, the result could subsequently be verified with
+% Metis even without the Empty Essence Lemma (cf. Fig.~\ref{Inconsistency_S5U}). 
+% % Interestingly, Metis' argument is different from that of
+% % \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
 

### 2016-07-10 Papers/2016/KI/main.tex
Improve section about explanation

@@
 
-\section{Intuitive Inconsistency Argument} \label{sec:inconsistency}
+\section{Intuitive Explanations for the Inconsistency} 
+\label{sec:inconsistency}
 
@@
 for humans, it did contain relevant hints to the importance of the
-empty property $\lambda x. \bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets).
+empty property\footnote{
+  Note that the terms for the empty property ($\lambda x. \bot$) and for the property of self-difference ($\lambda x.  x\not=x$) have identical denotations in the logic setting
+  with functional and Boolean extensionality assumed
+  here. 
+  %Nevertheless, some philosophers may actually prefer the use of
+  %self-difference over the empty property in
+  %the analysis below. 
+  For the proof to go through it is
+  irrelevant which property is used.
+} $\lambda x.
+\bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets)\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.}.
+Based on this hint, we conceived the following informal explanation for the inconsistency of G\"odel's axioms (reproduced without change from \cite{C55}):
 %
-Note that the terms for the empty property\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.} ($\lambda x.
-\bot$) and for the property of self-difference ($\lambda x.
-x\not=x$) have identical denotations in a logic setting
-with functional and Boolean extensionality as assumed
-here. Nevertheless, some philosophers\footnote{Private communication with Andr\'e Fuhrmann.} may actually prefer the use of
-self-difference over the empty property in
-the analysis below. However, for the proof to go through it is
-irrelevant which notion we use and the reader may simply replace the
-empty property by self-difference.
-
-% Philosophers
-% might actually prefer using the latter variant over the
-% former. However, on the given logic setting, with functional and
-% Boolean extensionality, both terms have identical
-% denotations. Consequently, they are mutually replacable below and the
-% reader may simply use the version he prefers.
-
-\paragraph{Informal Argument.} \label{sec:arg1}
-Based on the hints found in \textsc{Leo-II}'s refutation, we conceived the following informal explanation for the inconsistency of G\"odel's axioms:
 
@@
 But $\nec \bot$ can only be true in $w$, if there is no world $w'$ accessible from $w$.
-In logics with a reflexive or symmetric accessibility relation (e.g.
+In logics\footnote{
+  Interestingly, the refutation automatically generated by
+  \textsc{Leo-II} uses a symmetric accessibility relation, and thus
+  requires the modal logic \KB. The informal, human-constructed
+  refutations described here, on the other hand, requires only the
+  weaker modal logic \K. In our experiments \textsc{Leo-II} (like all
+  other HOL provers) was still too weak to automatically prove the
+  inconsistency already in logic \K. Hence, this remains an open problem for automated
+  theorem provers.
+} with a reflexive or symmetric accessibility relation (e.g.
 \KB), it is easy to see that there must be a world $w'$ accessible from $w$: either $w'$ itself, in case of a reflexive relation, or $w_0$, in case of a symmetric relation.
@@
 \end{enumerate}
+
+\textbf{Bruno: Here we should present the syntactical argument for the
+  inconsistency of $\pos \nec \bot $.}
+
+\section{Conclusion}\label{sec:conclusion}
 
@@
 theorem provers.
-
-\textbf{Bruno: Here we should present the syntactical argument for the
-  inconsistency of $\pos \nec \bot $.}
-
-\section{Conclusion}\label{sec:conclusion}
 

### 2016-07-10 Papers/2016/KI/main.tex
Simplify abstract
Remove duplicate paragraph

@@
 \begin{abstract}
-  This paper discusses the discovery of the 
+  This paper discusses the 
   inconsistency in G\"odel's
-  ontological argument as a success story for artificial
-  intelligence. Despite the popularity of G\"odel's argument, this
+  ontological argument. Despite the popularity of G\"odel's argument, this
   inconsistency remained unnoticed until 2013, 
@@
 
-Interestingly, the refutation automatically generated by
-\textsc{Leo-II} uses a symmetric accessibility relation, and thus
-requires the modal logic \KB. The informal, human-constructed
-refutation described above, on the other hand, requires only the
-weaker modal logic \K. In our experiments \textsc{Leo-II} (like all
-other HOL provers) was still too weak to automatically prove the
-inconsistency already in logic \K. Hence, this remains an open problem for automated
-theorem provers.
-
 The axioms and definitions in G\"odel's manuscript are inconsistent;

### 2016-07-10 Papers/2016/KI/main.tex
Fix typo
Add derivation figure and Coq scripts

@@
 \usepackage{xcolor}
+
+\usepackage{proof}
+
+\usepackage{fancyvrb}
+\newcommand{\red}[1]{\textcolor[rgb]{1,0,0}{#1}}
+\newcommand{\blue}[1]{\textcolor[rgb]{0,0,1}{#1}}
+\newcommand{\brown}[1]{\textcolor[rgb]{0.8,0.6,0.4}{#1}}
+
+\newcommand{\Parameter}{\red{Parameter}}
+\newcommand{\Ltac}{\red{Ltac}}
+\newcommand{\Axiom}{\red{Axiom}}
+\newcommand{\Lemma}{\red{Lemma}}
+\newcommand{\Theorem}{\red{Theorem}}
+\newcommand{\Definition}{\red{Definition}}
+\newcommand{\Notation}{\blue{Notation}}
+\newcommand{\Prop}{\blue{Prop}}
+\newcommand{\Type}{\blue{Type}}
+\newcommand{\llet}{\blue{let}}
+\newcommand{\match}{\blue{match}}
+\newcommand{\with}{\blue{with}}
+\newcommand{\eend}{\blue{end}}
+\newcommand{\iin}{\blue{in}}
+\newcommand{\as}{\blue{as}}
+\newcommand{\Require}{\blue{Require}}
+\newcommand{\Import}{\blue{Import}}
+\newcommand{\fforall}{\blue{forall}}
+\newcommand{\fun}{\blue{fun}}
+\newcommand{\Proof}{\blue{Proof}}
+\newcommand{\Qed}{\blue{Qed}}
+\newcommand{\Hint}{\blue{Hint}}
+\newcommand{\com}[1]{\brown{#1}}
+\newcommand{\bslash}{\symbol{92}}
 
@@
 \and
- Australien National University, Canberra,  Australia \\ \url{bruno.wp@gmail.com}
+ Australian National University, Canberra,  Australia \\ \url{bruno.wp@gmail.com}
 }
@@
 
-\textbf{Bruno: Here we should present the syntactical argument for the
-  inconsistency of $\pos \nec \bot $.}
+If we were to convert the informal proof above to a formal proof, the semantic reasoning in step 8 would require a leap to the meta-logic (HOL), in order to expand the definitions of the modal operators and reason directly about possible worlds.
+The following alternative derivation avoids this leap and remains purely within the object logic (HOML \K):
+
+\begin{enumerate}
+\item[8\textsuperscript{*}.] We must derive $\bot$ from $\pos \nec \bot$.
+In order to derive $\bot$, it suffices to show that there exists a derivable proposition such that its negation is also derivable.
+We choose $\pos \pos \bot$ as a candidate proposition, and hence we must show that:
+\begin{itemize}
+\item $\neg \pos \pos \bot$ \emph{is derivable:} this proposition is logically equivalent to $\nec \nec \top$, which is trivially derivable from $\top$ by two applications of the necessitation inference rule.
+
+\item $\pos \pos \bot$ \emph{is derivable:} and indeed, it can be derived (using a recently developed natural deduction calculus for modal logic \K \cite{CSR}) as follows:
+
+\includegraphics[width=0.7\textwidth]{Derivation}
+
+% $$
+% \infer[\pos_I]{\pos \pos \bot}{
+% \infer[\pos_I]{\pos \bot}{  
+% \infer[\nec_E]{\bot}{
+% \infer[\pos_E]{\nec \bot}{
+%   \pos \nec \bot
+% }
+% }
+% }
+% }
+% $$
+\end{itemize}
+
+\end{enumerate}
+
+An interesting, unusual and noticeable feature of the derivation shown above is that the leftmost $\pos_E$ (diamond elimination) inference derives a formula ($\nec \bot$) that is never used as a premise.
+This is necessary because of the \emph{eigen-box} condition, which requires that every box must be accessed by exactly one \emph{strong} modal rule.
+The purpose of the \emph{strong} $\pos_E$ inference is merely to create and access the innermost box that is needed by the \emph{weak} $\nec_E$ and $\pos_I$ inferences inside the outermost box.
+
+\newcommand{\verbsize}{\small}
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Theorem dia_box_false_to_false_meta: [(dia (box mFalse))] -> [mFalse].
+\Proof. 
+intro H. intro w.
+destruct (H w) as [w0 [R0 H0]]. destruct (H w0) as [w1 [R1 H1]].
+box_elim H0 w1 HF.
+unfold mFalse in HF. destruct HF as [p [HF1 HF2]].
+contradiction.
+\Qed.
+\end{Verbatim}
+
+\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\Lemma mimplies_to_mnot: [mforall p:o, (p m-> mFalse) m-> (m~ p)].
+\Proof. mv.
+intro p. intro H. intro H0.
+destruct (H H0) as [p0 [H1 H2]].
+apply H2. exact H1.
+\Qed.
+
+\Lemma dia_not_not_box: [ mforall p, (dia (m~ p)) m-> (m~ (box p)) ].
+\Proof. mv.
+intro p. intro H1. intro H2. dia_e H1. apply H. box_e H2 H3. exact H3.
+\Qed.
+
+\Theorem dia_box_false_to_false_object: [(dia (box mFalse))] -> [mFalse].
+\Proof.
+intro H. intro w.
+exists (dia (dia mFalse)).
+split.
+  dia_e (H w). dia_e (H w0). dia_i w0. dia_i w1. box_e H0 H3. exact H3.
+
+  apply box_not_not_dia. box_i. apply box_not_not_dia. box_i.
+  apply mimplies_to_mnot. intro H4. exact H4.
+\Qed.
+
+\end{Verbatim}
 

### 2016-07-11 Papers/2016/KI/main.tex
Adjust size of derivation
improve breakline in coq script
Compare Coq scripts and write conclusion

@@
 \newcommand{\bslash}{\symbol{92}}
+
+\newcommand{\verbsize}{\small}
 
@@
 
-\item From theorem T1 (\textit{Positive properties are possibly
+\item From axiom A5 (`necessary existence' is a positive
+  property: $P(\NE)$ ) and theorem T1 (\textit{Positive properties are possibly
   exemplified}: ${\allq \phi} [P(\phi) \imp {\pos}  \exq x
-  \phi(x)]$) and axiom A5 (``necessary existence'' is a positive
-  property: $P(\NE)$ ), it follows that $\NE$ is possibly exemplified:
+  \phi(x)]$), it follows that $\NE$ is possibly exemplified:
   \hfill $  \pos \exq x [\NE(x)] $
 
-\item Expanding the definition of ``necessary existence''
+\item Expanding the definition of `necessary existence'
   (${\NE(x) \equiv \allq \phi [\ess{\phi}{x} \imp \nec \exq y
@@
 If we were to convert the informal proof above to a formal proof, the semantic reasoning in step 8 would require a leap to the meta-logic (HOL), in order to expand the definitions of the modal operators and reason directly about possible worlds.
-The following alternative derivation avoids this leap and remains purely within the object logic (HOML \K):
+The following alternative proof avoids this leap and remains purely within the object logic (HOML \K):
 
@@
 We choose $\pos \pos \bot$ as a candidate proposition, and hence we must show that:
-\begin{itemize}
-\item $\neg \pos \pos \bot$ \emph{is derivable:} this proposition is logically equivalent to $\nec \nec \top$, which is trivially derivable from $\top$ by two applications of the necessitation inference rule.
+\begin{enumerate}
+\item $\neg \pos \pos \bot$ \emph{is derivable:} this proposition is equi-valid to $\nec \nec \top$, which is trivially derivable from $\top$ by two applications of the necessitation inference rule.
 
 \item $\pos \pos \bot$ \emph{is derivable:} and indeed, it can be derived (using a recently developed natural deduction calculus for modal logic \K \cite{CSR}) as follows:
-
-\includegraphics[width=0.7\textwidth]{Derivation}
+\smallskip
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Derivation}
+\end{center}
+\smallskip
 
@@
 % $$
-\end{itemize}
+\end{enumerate}
 
@@
 
-An interesting, unusual and noticeable feature of the derivation shown above is that the leftmost $\pos_E$ (diamond elimination) inference derives a formula ($\nec \bot$) that is never used as a premise.
+An interesting and unusual feature of the derivation shown above is that the leftmost $\pos_E$ (diamond elimination) inference derives a formula ($\nec \bot$) that is never used as a premise.
 This is necessary because of the \emph{eigen-box} condition, which requires that every box must be accessed by exactly one \emph{strong} modal rule.
@@
 
-\newcommand{\verbsize}{\small}
+The proofs above have been formalized and verified step-by-step in \textsc{Coq}\footnote{The complete formal proofs can be found in: \url{https://github.com/FormalTheology/GoedelGod}.}.
+The following \textsc{Coq} script shows the formalization of step 8 of the meta-logic proof.
 
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
-\Theorem dia_box_false_to_false_meta: [(dia (box mFalse))] -> [mFalse].
+\Lemma dia_box_false_to_false_meta: [(dia (box mFalse))] -> [mFalse].
 \Proof. 
@@
 
+\noindent
+The other \textsc{Coq} scripts below show the formalization of step 8\textsuperscript{*} of the object-logic proof.
+
 \begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
 \Lemma mimplies_to_mnot: [mforall p:o, (p m-> mFalse) m-> (m~ p)].
-\Proof. mv.
-intro p. intro H. intro H0.
-destruct (H H0) as [p0 [H1 H2]].
-apply H2. exact H1.
+\Proof. mv. intro p. intro H. intro H0. 
+destruct (H H0) as [p0 [H1 H2]]. apply H2. exact H1.
 \Qed.
@@
 
-\Theorem dia_box_false_to_false_object: [(dia (box mFalse))] -> [mFalse].
+\Lemma dia_box_false_to_false_object: [(dia (box mFalse))] -> [mFalse].
 \Proof.
@@
 \Qed.
+\end{Verbatim}
 
-\end{Verbatim}
+Due to a deliberate and disciplined use of only 
+the simplest (and non-automatic) \textsc{Coq} tactics, 
+there is a straightforward correspondence between the tactics used 
+in the scripts above and the inference rules of the modal natural 
+deduction calculus \cite{CSR}. Therefore, the lengths of the proof 
+scripts (in number of tactic applications) can serve as estimations 
+for the lengths of the corresponding natural deduction proof. 
+It is noticeable that the meta-logic proof is significantly shorter 
+than the pure object-logic proof. An in-depth analysis 
+reveals that the reasoning about the possible worlds semantics in 
+the meta-logic proof acts as a short-cut: when it becomes impossible (in step 8) to build the third world $w'$ (because $\bot$ would have to hold in it, and thus $w'$ would be contradictory), a contradiction at the HOL level can be immediately derived, completing the proof.
+In the object-level proof, on the other hand, such a contradiction had to be found in the arbitrary initial world $w$.
+This not only requires additional tedious logical inferences (cf.
+the proofs of the lemmas \texttt{mimplies\_to\_mnot} and \texttt{dia\_not\_not\_box}), but also a non-trivial guessing of the instantiating contradictory proposition $\pos \pos \bot$, whose purpose is precisely to carry over the contradiction from $w'$ back to $w$.
 
@@
 
-The axioms and definitions in G\"odel's manuscript are inconsistent;
+The axioms and definitions in G\"odel's manuscript are inconsistent (even in the weakest modal logic \K);
 this was detected automatically by the prover
-\textsc{Leo-II}. We presented a rational reconstruction and
-verification of the inconsistency argument in Isabelle/HOL. This
-argument is valid in all normal HOMLs including base logic \K.
+\textsc{Leo-II}.
+In our previous work \cite{C55}, we presented a human-readable and intuitive meta-logic explanation for the inconsistency, and we formalized and semi-automatically reconstructed it in the \textsc{Isabelle} proof assistant.
+Here this work was extended with an object-level explanation, and both explanations were formalized step-by-step in \textsc{Coq} proof assistant.
+A comparison of the formal \textsc{Coq} proofs of both explanations revealed that the meta-logic reasoning is more powerful, because it enables shortcuts and, therefore, requires fewer inferences and guesses.
+We conjecture that this is not an accident, but rather a fundamental reason why the embedding approach is effective in practice.
 
-Extending the presentation in our IJCAI 2016 paper \cite{C55}, we have added
-a syntactical proof for the inconsistency of  $\pos \nec \bot$.  Previously we had reduced the inconsistency
-analysis to exactly this proposition and we had then argued
-semantically that this proposition must be false already in modal
-logic \K.
+It is entertaining to notice that, although our work reveals a mistake in G\"odel's manuscript, it does substantiate G\"odel's belief that 
+``there is a scientific (exact) philosophy and theology,
+which deals with concepts of the highest abstractness; 
+and this is also most highly fruitful for science.'' \cite{Wang1996}[p.
+316].
+Indeed, through the investigation of G\"odel's mistake, we have been led to a fruitful conjecture in the fields of automated reasoning and proof theory.
 
@@
 
-Our work reveals a challenge for automated reasoning:
-the (so far partially manual) extraction of an informal argument from a formal proof. 
-Without accompanying human-understandable explanations,
-the proofs generated by provers such as \textsc{Leo-II} or Metis, will
-presumably be only of limited value for philosophers, for whom intuitive
-arguments remain crucial for the acceptance of novel results.
+% Our work reveals a challenge for automated reasoning:
+% the (so far partially manual) extraction of an informal argument from a formal proof. 
+% Without accompanying human-understandable explanations,
+% the proofs generated by provers such as \textsc{Leo-II} or Metis, will
+% presumably be only of limited value for philosophers, for whom intuitive
+% arguments remain crucial for the acceptance of novel results.
 
@@
 % and the fully automatic proof of T3 from Scott's axioms
-Our work 
-demonstrates
-the potential of higher-order automated theorem proving and the
-semantic embeddings approach for philosophy: this technology is,
-in its current state of development, already capable of contributing novel results to
-metaphysics and to conduct reasoning steps at granularity-levels
-beyond common human capabilities.  
-\vfill
+% Our work 
+% demonstrates
+% the potential of higher-order automated theorem proving and the
+% semantic embeddings approach for philosophy: this technology is,
+% in its current state of development, already capable of contributing novel results to
+% metaphysics and to conduct reasoning steps at granularity-levels
+% beyond common human capabilities.  
+% \vfill
 

### 2016-07-11 Papers/2016/KI/main.tex
Compress Coq scripts and remove footnote.

@@
 
-The proofs above have been formalized and verified step-by-step in \textsc{Coq}\footnote{The complete formal proofs can be found in: \url{https://github.com/FormalTheology/GoedelGod}.}.
+The proofs above have been formalized and verified step-by-step in \textsc{Coq}.
+The complete proofs can be found in \url{https://github.com/FormalTheology/GoedelGod}.
 The following \textsc{Coq} script shows the formalization of step 8 of the meta-logic proof.
@@
 \Lemma dia_box_false_to_false_meta: [(dia (box mFalse))] -> [mFalse].
-\Proof. 
-intro H. intro w.
+\Proof. intro H. intro w.
 destruct (H w) as [w0 [R0 H0]]. destruct (H w0) as [w1 [R1 H1]].
-box_elim H0 w1 HF.
-unfold mFalse in HF. destruct HF as [p [HF1 HF2]].
-contradiction.
-\Qed.
+box_elim H0 w1 HF. unfold mFalse in HF. destruct HF as [p [HF1 HF2]].
+contradiction. \Qed.
 \end{Verbatim}
@@
 \Proof. mv. intro p. intro H. intro H0. 
-destruct (H H0) as [p0 [H1 H2]]. apply H2. exact H1.
-\Qed.
+destruct (H H0) as [p0 [H1 H2]]. apply H2. exact H1. \Qed.
 
 \Lemma dia_not_not_box: [ mforall p, (dia (m~ p)) m-> (m~ (box p)) ].
-\Proof. mv.
-intro p. intro H1. intro H2. dia_e H1. apply H. box_e H2 H3. exact H3.
-\Qed.
+\Proof. mv. intro p. intro H1. intro H2. 
+dia_e H1. apply H. box_e H2 H3. exact H3. \Qed.
 
 \Lemma dia_box_false_to_false_object: [(dia (box mFalse))] -> [mFalse].
-\Proof.
-intro H. intro w.
-exists (dia (dia mFalse)).
+\Proof. intro H. intro w. exists (dia (dia mFalse)).
 split.
@@
 
+\noindent
 Due to a deliberate and disciplined use of only 

### 2016-07-11 Papers/2016/KI/main.tex
Typeset system names

@@
 
-% ToDo: Use \textsc for all system names
-
 \section{Introduction}\label{sec:introduction}
@@
 
-Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the Coq proof assistant \cite{CSR}.
+Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the \textsc{Coq} proof assistant \cite{CSR}.
 To bridge the gap between higher-order logics (HOL; cf.~\cite{andrewsSEP} and the references
@@
 In \cite{C55}, we have extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we have reconstructed and
-verified it in the Isabelle proof assistant.
+verified it in the \textsc{Isabelle} proof assistant.
 However, that explanation relied on reasoning at the meta-logic (HOL) level, which was only possible because of the embedding.
@@
 version with a larger public. Scott's version of the axioms and
-definitions, formalized in Isabelle, is shown in
+definitions, formalized in \textsc{Isabelle}, is shown in
 Fig.~\ref{Scott_Goedel}. 
@@
   here. 
-  %Nevertheless, some philosophers may actually prefer the use of
-  %self-difference over the empty property in
-  %the analysis below. 
   For the proof to go through it is
@@
 } $\lambda x.
-\bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets)\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in Isabelle) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.}.
+\bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets)\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in \textsc{Isabelle}) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.}.
 Based on this hint, we conceived the following informal explanation for the inconsistency of G\"odel's axioms (reproduced without change from \cite{C55}):
@@
 
-% We have also presented several technical improvements regarding the
-% semantic embedding approach. In particular, we have achieved a
-% nearly perfect match between pen and paper presentations in HOML and
-% the syntax in Isabelle/HOL. As a result, the embedding of HOML in HOL
-% is now fully transparent, more user-friendly and ready for wider adoption.
-
-% On the other hand, there is still room for many pragmatical
-% improvements in Isabelle; just one example: in default setting,
-% Sledgehammer does not immediately inform the user when a proof has
-% been found and instead silently first executes a series of
-% time-consuming proof analysis processes (e.g. its dependency
-% minimization), before it eventually reports success. For G\"odel's
-% theorem T3 (\textit{Necessarily, there exists God}), for example, this
-% phase of silence takes several minutes --- during which the user might
-% actually give up on the proof attempt --- even though \textsc{Leo-II} already
-% reported success to Sledgehammer after 2.5 seconds.
-
-% Our work reveals a challenge for automated reasoning:
-% the (so far partially manual) extraction of an informal argument from a formal proof. 
-% Without accompanying human-understandable explanations,
-% the proofs generated by provers such as \textsc{Leo-II} or Metis, will
-% presumably be only of limited value for philosophers, for whom intuitive
-% arguments remain crucial for the acceptance of novel results.
-
-% Another open problem that we solved in this paper is a fully automatic
-% proof of T3 directly from Scott's axioms. Again, this proof was
-% contributed by \textsc{Leo-II}. This has become possible only
-% after we provided a more efficient embedding for HOML \SFiveU (instead of \SFive) in HOL.
-
-% Both the automated detection of the inconsistency in G\"odel's axioms
-% and the fully automatic proof of T3 from Scott's axioms
-% Our work 
-% demonstrates
-% the potential of higher-order automated theorem proving and the
-% semantic embeddings approach for philosophy: this technology is,
-% in its current state of development, already capable of contributing novel results to
-% metaphysics and to conduct reasoning steps at granularity-levels
-% beyond common human capabilities.  
-% \vfill
-
-% \noindent\textbf{Acknowledgments:} We thank Chad Brown, who contributed to the rational  reconstruction
-% of the inconsistency argument.
-
-%\pagebreak
-%A discussion between Chad Brown and Benzm\"uller
-% significantly influenced the rational reconstruction of the inconsistency reported by \textsc{LEO-II}.
-
-%\pagebreak
-
-%German Research Foundation DFG, Chad Brown
-
-%ToDo: reduce paper length to 7 pages
-
-%ToDo: use \textsc for all systems. Not only for LEO-II.
-
-%\small 
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-
 \bibliographystyle{plain}
@@
 
-% \pagebreak
-
-% \paragraph{Argument Reconstruction in Isabelle.}  \label{sec:arg2}
-% \begin{figure*}[t]
-%   \centering
-%   \begin{subfigure}[t]{0.715\textwidth}
-%     \includegraphics[width=\textwidth]{./Manuscript2.png}
-%     \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
-%       definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript}
-%   \end{subfigure}
-%    \begin{subfigure}[t]{0.28\textwidth}
-%      \includegraphics[width=\textwidth,height=7.2cm]{./Inconsistency_S5U_direct.png}
-%      \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
-%    \end{subfigure}
-% % \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
-% % \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
-%  \caption{The inconsistency in G\"{o}del's manuscript has been
-%    detected and verified by HOL ATPs} 
-% \end{figure*}
-% To verify the correctness of the informal argument explained above, it
-% was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
-%   external provers such as \textsc{Leo-II} or Satallax, 
-%   constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
-% inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
-% the other hand, is explicitly stated, to ensure that Isabelle is
-% reconstructing the same argument. In fact, without the help of this
-% lemma, Metis is still not strong enough to refute G\"odel's
-% axioms.
-
-% \paragraph{Mapping the Inconsistency to G\"odel.}
-% The inconsistency verified in Fig.~\ref{InconsistencyIsabelleK} follows from the definition of
-% essence (ess), the definition of necessary existence (NE), the
-% axioms A1a and A2 (which entail theorem T1), and axiom A5. It remains to show that
-% these ingredients are actually present in G\"odel's manuscript in
-% Fig.~\ref{GoedelScript}.
-
-% This can be easily seen: Axiom A1a in
-% Fig.~\ref{InconsistencyIsabelleK} is implied by Axiom Ax2 and the
-% highlighted footnote remark in Fig.~\ref{GoedelScript}. Axioms A2 and
-% A5 in Fig.~\ref{InconsistencyIsabelleK} correspond to Ax4 and Ax3 in
-% Fig.~\ref{GoedelScript}. The definitions of essence and necessary
-% existence are easy to identify. Therefore, the verified
-% inconsistency from Fig.~\ref{InconsistencyIsabelleK} does apply to 
-% G\"odel's original manuscript.
-
-% \paragraph{Inconsistency of G\"odel's Axioms in \SFiveU.}
-% Isabelle/HOL's Sledgehammer tool, which orchestrates calls to
-% external provers such as \textsc{Leo-II}, still
-% fails to detect the inconsistency of G\"odel's axioms in the standard
-% embedding of \SFive, while a direct modeling of the problem in TPTP THF syntax
-% in combination with a direct call of \textsc{Leo-II} succeeded. In
-% other words, without independent experiments with no mediation through Sledgehammer, the
-% inconsistency would not have been detected.
-
-% %\marginpar{ToDo: Are we really supposed to use \S instead of ``Sec.''?}
-% On the other hand (and further confirming the claims from \S\ref{sec:improvedembedding}), 
-% the reconstruction in Isabelle/HOL with the improved embedding for \SFiveU was more efficient: 
-% the inconsistency could be detected
-% by \textsc{Leo-II} also when called via
-% Sledgehammer. Moreover, the result could subsequently be verified with
-% Metis even without the Empty Essence Lemma (cf. Fig.~\ref{Inconsistency_S5U}). 
-% % Interestingly, Metis' argument is different from that of
-% % \S\ref{sec:arg1} and \S\ref{sec:arg2}, since axiom A1b is not used.
-
 \end{document}

### 2016-07-11 Papers/2016/KI/main.tex
Compress proof note

@@
 \title{An Object-Logic Explanation for the Inconsistency in G\"odel's Ontological Theory}
-\author{Christoph Benzm\"uller\inst{1}\thanks{This work has been supported by
-    the German Research Foundation DFG under grants BE2501/9-1,2} \and Bruno Woltzenlogel Paleo\inst{2} 
+\author{Christoph Benzm\"uller\inst{1}\thanks{This work was supported by
+    the German Research Foundation DFG grant BE2501/9-1,2} \and Bruno Woltzenlogel Paleo\inst{2} 
 }

### 2016-07-11 Papers/2016/KI/main.tex
Compress

@@
 when \textsc{Leo-II} found a surprising refutation of the axioms.
-
-In \cite{C55}, we have extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we have reconstructed and
+In \cite{C55} we extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we reconstructed and
 verified it in the \textsc{Isabelle} proof assistant.
-However, that explanation relied on reasoning at the meta-logic (HOL) level, which was only possible because of the embedding.
-Here we complement that work with a purely object-logic (HOML) explanation.
+But that explanation relied on reasoning at the meta-logic (HOL) level, which was only possible because of the embedding.
+Here we complement that work with a purely object-logic (HOML) explanation, and we compare and formalize both explanations in the \textsc{Coq} proof assistant.
 
@@
 } $\lambda x.
-\bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets)\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and also in the reconstruction in \textsc{Isabelle}) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.}.
+\bot$ (also denoted $\emptyset$, as in HOL it is customary to think of unary predicates as sets)\footnote{An additional lambda abstraction occurs in the empty property in \textsc{Leo-II}'s proof (and in the reconstruction in \textsc{Isabelle}) because the embedding approach lifts the boolean type $o$ to $\iota \imp o$.}.
 Based on this hint, we conceived the following informal explanation for the inconsistency of G\"odel's axioms (reproduced without change from \cite{C55}):
@@
 \item $\pos \pos \bot$ \emph{is derivable:} and indeed, it can be derived (using a recently developed natural deduction calculus for modal logic \K \cite{CSR}) as follows:
-\smallskip
+%\smallskip
 \begin{center}
@@
 \end{center}
-\smallskip
+%\smallskip
 
@@
   apply box_not_not_dia. box_i. apply box_not_not_dia. box_i.
-  apply mimplies_to_mnot. intro H4. exact H4.
-\Qed.
+  apply mimplies_to_mnot. intro H4. exact H4. \Qed.
 \end{Verbatim}

### 2016-07-11 Papers/2016/KI/main.tex
update from Christoph

@@
 
-\newcommand{\verbsize}{\small}
+\newcommand{\verbsize}{\scriptsize}
 
@@
 
-\title{An Object-Logic Explanation for the Inconsistency in G\"odel's Ontological Theory}
+\title{An Object-Logic Explanation for the Inconsistency in G\"odel's
+  Ontological Theory (Extended Abstract)}
 \author{Christoph Benzm\"uller\inst{1}\thanks{This work was supported by
@@
   when it was detected automatically by the 
-  higher-order theorem prover \textsc{Leo-II} \cite{ECAI2014}.
+  higher-order theorem prover \textsc{Leo-II}.
 Complementing the meta-logic explanation for the inconsistency available in our IJCAI 2016 paper \cite{C55}, we present here a new purely object-logic explanation that does not rely on semantic argumentation.
@@
 fact crucial. Without it, G\"odel's original axioms are
-inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_Goedel}
+inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_Goedel},
 where the model finder Nitpick \cite{Nitpick} confirms consistency). In
@@
 $\Box \varphi$ and $\Diamond \varphi$ (relative to a current world
-$w_o$) are mapped, respectively, to the HOL formulas
+$w_0$) are mapped, respectively, to the HOL formulas
 $\forall w. (r w_0 w) \imp (\varphi w)$ and
@@
 claimed that the axioms were inconsistent. Unfortunately, the
-refutation generated by \textsc{Leo-II} was barely human-readable. The refutation was based on machine-oriented inference rules (the higher-order resolution calculus \cite{W47}), and the
+refutation generated by \textsc{Leo-II} was barely human-readable. The
+refutation was based on machine-oriented inference rules (a higher-order resolution calculus \cite{J30}), and the
 text file had 153 lines (with an average of 184
@@
 
-\item $\pos \pos \bot$ \emph{is derivable:} and indeed, it can be derived (using a recently developed natural deduction calculus for modal logic \K \cite{CSR}) as follows:
+\item $\pos \pos \bot$ \emph{is derivable:} and indeed, it can be
+  derived (using a recently developed natural deduction calculus for
+  modal logic \K \cite{CSR}) as follows: \\
 %\smallskip
+
 \begin{center}
-\includegraphics[width=0.5\textwidth]{Derivation}
+\includegraphics[width=0.4\textwidth]{Derivation}
 \end{center}
@@
 
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\begin{Verbatim}[frame=single,commandchars=\\\{\},fontsize=\verbsize]
 \Lemma dia_box_false_to_false_meta: [(dia (box mFalse))] -> [mFalse].
@@
 
-\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
+\begin{Verbatim}[frame=single,commandchars=\\\{\},fontsize=\verbsize]
 \Lemma mimplies_to_mnot: [mforall p:o, (p m-> mFalse) m-> (m~ p)].

### 2016-07-11 Papers/2016/KI/main.tex
Format "Nitpick" with \textsc
update from Christoph

@@
 
-Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the \textsc{Coq} proof assistant \cite{CSR}.
-To bridge the gap between higher-order logics (HOL; cf.~\cite{andrewsSEP} and the references
-therein), as used by these systems, and higher-order \emph{modal} logics (HOML; cf.~\cite{homl} and the
-references therein), on which the ontological argument argument, the logic embedding approach \cite{J23,C40} was used.
+Scott's version of G\"odel's argument was automatically reconstructed
+by higher-order automated theorem provers \cite{C40} and its
+correctness was verified step-by-step in the \textsc{Coq} proof
+assistant \cite{CSR}. To bridge the gap between higher-order logics
+(HOL; cf.~\cite{andrewsSEP} and the references therein), as used by
+these systems, and higher-order \emph{modal} logics (HOML;
+cf.~\cite{homl} and the references therein), on which the ontological
+argument argument, the logic embedding approach \cite{J23,C40} was
+used.
 
@@
 remained unnoticed to philosophers until 2013, 
-when \textsc{Leo-II} found a surprising refutation of the axioms.
+when \textsc{Leo-II} \cite{J30} found a surprising refutation of the axioms.
 In \cite{C55} we extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we reconstructed and
@@
 A comparison of the formal \textsc{Coq} proofs of both explanations revealed that the meta-logic reasoning is more powerful, because it enables shortcuts and, therefore, requires fewer inferences and guesses.
-We conjecture that this is not an accident, but rather a fundamental reason why the embedding approach is effective in practice.
+We conjecture that this is not accidental, but rather a fundamental reason why the embedding approach is effective in practice.
 
-It is entertaining to notice that, although our work reveals a mistake in G\"odel's manuscript, it does substantiate G\"odel's belief that 
-``there is a scientific (exact) philosophy and theology,
-which deals with concepts of the highest abstractness; 
-and this is also most highly fruitful for science.'' \cite{Wang1996}[p.
-316].
-Indeed, through the investigation of G\"odel's mistake, we have been led to a fruitful conjecture in the fields of automated reasoning and proof theory.
+It is kind of entertaining that our work reveals a mistake in
+G\"odel's manuscript and at same time further substantiates G\"odel's
+belief that ``there is a scientific (exact) philosophy and theology,
+which deals with concepts of the highest abstractness; and this is
+also most highly fruitful for science.''
+\cite{Wang1996}[p. 316]. Indeed, through the investigation of
+G\"odel's mistake, we have been led to an interesting little
+conjecture in automated reasoning and proof theory (the global axiom
+$\pos \nec \bot$ is inconsistent).
 

### 2016-07-11 Papers/2016/KI/main.tex
Fix typos and make minor grammatical improvements

--- 2016-07-11 Papers/2016/KI/main.tex
@@
 therein), as used by these systems, and higher-order \emph{modal} logics (HOML; cf.~\cite{homl} and the
-references therein), on which the ontological argument argument, the logic embedding approach \cite{J23,C40} was used.
+references therein), on which the ontological argument is based, the logic embedding approach \cite{J23,C40} was used.
 
@@
 verified it in the \textsc{Isabelle} proof assistant.
-But that explanation relied on reasoning at the meta-logic (HOL) level, which was only possible because of the embedding.
-Here we complement that work with a purely object-logic (HOML) explanation, and we compare and formalize both explanations in the \textsc{Coq} proof assistant.
+But that explanation relied on reasoning at the \emph{meta-logic} (HOL) level, which was only possible because of the embedding.
+Here we complement that work with a purely \emph{object-logic} (HOML) explanation, and we compare and formalize both explanations in the \textsc{Coq} proof assistant.
 
@@
 
-\section{An Essential Difference in the Definitions of Essence.}
+\section{An Essential Difference in the Definitions of Essence}
 \label{sec:history}
+
 G\"odel's manuscript can be considered a translation of Leibniz's
@@
 inconsistent. With it, Scott's axioms are consistent (cf.~Fig.~\ref{Scott_Goedel},
-where the model finder Nitpick \cite{Nitpick} confirms consistency). In
+where the model finder \textsc{Nitpick} \cite{Nitpick} confirms consistency). In
 personal communication, Dana Scott confirmed that he was unaware
@@
 In our experiments in this branch of metaphysics we
-utilise an embedding of HOMLs, such as \textbf{K}, \textbf{KB} and
+utilize an embedding of HOMLs, such as \textbf{K}, \textbf{KB} and
 \textbf{S5} with various domain conditions (possibilist and actualist
@@
 quantifiers --- in HOL by stating a set of equations defining the
-logical constants. The resulting object logic is
+logical constants. The resulting logic is
 the HOML \textbf{K} with rigid terms and constant domains (possibilist
@@
   here. 
-  For the proof to go through it is
+  For the proof to go through, it is
   irrelevant which property is used.
@@
 
-If we were to convert the informal proof above to a formal proof, the semantic reasoning in step 8 would require a leap to the meta-logic (HOL), in order to expand the definitions of the modal operators and reason directly about possible worlds.
-The following alternative proof avoids this leap and remains purely within the object logic (HOML \K):
+If we were to convert the informal proof above to a formal proof, the semantic reasoning in step 8 would require a leap to the meta-logic (HOL), in order to expand the definitions of modal operators and reason directly about possible worlds.
+The alternative proof below avoids this leap and remains purely within the object logic (HOML \K):
 
@@
 In order to derive $\bot$, it suffices to show that there exists a derivable proposition such that its negation is also derivable.
-We choose $\pos \pos \bot$ as a candidate proposition, and hence we must show that:
+We choose $\pos \pos \bot$ as the candidate proposition, and hence we must show that:
 \begin{enumerate}
@@
 the meta-logic proof acts as a short-cut: when it becomes impossible (in step 8) to build the third world $w'$ (because $\bot$ would have to hold in it, and thus $w'$ would be contradictory), a contradiction at the HOL level can be immediately derived, completing the proof.
-In the object-level proof, on the other hand, such a contradiction had to be found in the arbitrary initial world $w$.
-This not only requires additional tedious logical inferences (cf.
-the proofs of the lemmas \texttt{mimplies\_to\_mnot} and \texttt{dia\_not\_not\_box}), but also a non-trivial guessing of the instantiating contradictory proposition $\pos \pos \bot$, whose purpose is precisely to carry over the contradiction from $w'$ back to $w$.
+In the object-level proof, on the other hand, such a contradiction has to be found in the arbitrary initial world $w$.
+This requires not only additional tedious logical inferences (cf.
+the proofs of the lemmas \texttt{mimplies\_to\_mnot} and \texttt{dia\_not\_not\_box}), but also a non-trivial guessing of the contradictory proposition $\pos \pos \bot$, whose purpose is precisely to carry over the contradiction from $w'$ back to $w$.
 

### 2016-07-11 Papers/2016/KI/main.tex

@@
 
-Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the \textsc{Coq} proof assistant \cite{CSR}.
-To bridge the gap between higher-order logics (HOL; cf.~\cite{andrewsSEP} and the references
-therein), as used by these systems, and higher-order \emph{modal} logics (HOML; cf.~\cite{homl} and the
-references therein), on which the ontological argument is based, the logic embedding approach \cite{J23,C40} was used.
+Scott's version of G\"odel's argument was automatically reconstructed
+by higher-order automated theorem provers \cite{C40} and its
+correctness was verified step-by-step in the \textsc{Coq} proof
+assistant \cite{CSR}. To bridge the gap between higher-order logics
+(HOL; cf.~\cite{andrewsSEP} and the references therein), as used by
+these systems, and higher-order \emph{modal} logics (HOML;
+cf.~\cite{homl} and the references therein), on which the ontological argument is based, the logic embedding approach \cite{J23,C40} was used.
 

### 2016-07-15 Papers/2016/KI/main.tex

@@
 \bibliographystyle{plain}
-\bibliography{Bibliography}
+%\bibliography{Bibliography}
+\begin{thebibliography}{10}
+
+\bibitem{andrewsSEP}
+P.B. Andrews.
+\newblock Church's type theory.
+\newblock In E.N. Zalta, editor, {\em The Stanford Encyclopedia of Philosophy}.
+  Spring 2014 edition, 2014.
+
+\bibitem{J23}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Quantified multimodal logics in simple type theory.
+\newblock {\em Logica Universalis}, 7(1):7--20, 2013.
+
+\bibitem{J30}
+C.~Benzm{\"u}ller, L.C. Paulson, N.~Sultana, and F.~Thei{\ss}.
+\newblock The higher-order prover {LEO-II}.
+\newblock {\em Journal of Automated Reasoning}, 55(4):389--404, 2015.
+
+\bibitem{C40}
+C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo.
+\newblock Automating {G\"{o}del's} ontological proof of {God}'s existence with
+  higher-order automated theorem provers.
+\newblock In Torsten Schaub, Gerhard Friedrich, and Barry O'Sullivan, editors,
+  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
+  Applications}, pages 93 -- 98. IOS Press, 2014.
+
+\bibitem{CSR}
+C.~Benzm{\"{u}}ller and B.~Woltzenlogel~Paleo.
+\newblock Interacting with modal logics in the coq proof assistant.
+\newblock In Lev~D. Beklemishev and Daniil~V. Musatov, editors, {\em Computer
+  Science - Theory and Applications - 10th International Computer Science
+  Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July 13-17, 2015,
+  Proceedings}, volume 9139 of {\em LNCS}, pages 398--411. Springer, 2015.
+
+\bibitem{C55}
+C.~Benzm{\"u}ller and B.~Woltzenlogel~Paleo.
+\newblock The inconsistency in {G{\"o}del's} ontological argument: A success
+  story for {AI} in metaphysics.
+\newblock In {\em IJCAI 2016}, 2016.
+
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for higher-order logic based on a
+  relational model finder.
+\newblock In {\em ITP 2010}, number 6172 in LNCS, pages 131--146. Springer,
+  2010.
+
+\bibitem{FitelsonZalta}
+B.~Fitelson and E.~N. Zalta.
+\newblock Steps toward a computational metaphysics.
+\newblock {\em J. Philosophical Logic}, 36(2):227--247, 2007.
+
+\bibitem{GoedelNotes}
+K.~G\"odel.
+\newblock {\em Appx. A: Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock In Sobel \cite{sobel2004logic}, 1970.
+
+\bibitem{homl}
+R.~Muskens.
+\newblock {Higher Order Modal Logic}.
+\newblock In P.~Blackburn~et al., editor, {\em Handbook of Modal Logic},
+  Studies in Logic and Practical Reasoning, pages 621--653. Elsevier,
+  Dordrecht, 2006.
+
+\bibitem{oppenheimer11}
+P.E. Oppenheimer and E.N. Zalta.
+\newblock A computationally-discovered simplification of the ontological
+  argument.
+\newblock {\em Australasian J. of Philosophy}, 89(2):333--349, 2011.
+
+\bibitem{sep-ontological-arguments}
+G.~Oppy.
+\newblock Ontological arguments.
+\newblock In Edward~N. Zalta, editor, {\em The Stanford Encyclopedia of
+  Philosophy}. Spring 2015 edition, 2015.
+
+\bibitem{rushby13}
+J.~Rushby.
+\newblock The ontological argument in {PVS}.
+\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St.
+  Petersburg, Russia, 2013.
+
+\bibitem{ScottNotes}
+D.~Scott.
+\newblock {\em Appx. B: Notes in Dana Scott's Hand}, pages 145--146.
+\newblock In Sobel \cite{sobel2004logic}, 1972.
+
+\bibitem{sobel2004logic}
+J.H. Sobel.
+\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
+\newblock Cambridge U. Press, 2004.
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller.
+\newblock Automated reasoning in higher-order logic using the {TPTP THF}
+  infrastructure.
+\newblock {\em J. of Formalized Reasoning}, 3(1):1--27, 2010.
+
+\bibitem{Wang1996}
+H.~Wang.
+\newblock {\em A Logical Journey: From G\"odel to Philosophy}.
+\newblock MIT Press, 1996.
+
+\end{thebibliography}
 

### 2016-08-14 Papers/2014/Universalis/ModalCollapse.tex
added my DFG grant references
minor editing
Leibniz/Lenzen initial commit

@@
 
+\thanks{This work has been supported by the German National Research Foundation (DFG) under grant BE 2501/9-1,2.}
+
 %----------classification, keywords, date
@@
 Prim. 03B15; % Higher-order logic and type theory
-Sec. 68T15 % Theorem proving (deduction, resolution, etc.)
+Sec. 68T15% Theorem proving (deduction, resolution, etc.)
 }
-% \marginpar{Change?: 
 % 03A02 % Philosophical aspects of logic and foundations
@@
 % 68T02 % Artificial Intelligence
-% }
 
@@
 
-\dedicatory{ }
+\dedicatory{}
 
@@
   on the domain of properties.}, 
-%\marginpar{added footnote}
 they entail a \emph{modal collapse}
@@
 and the modal collapse shows that this inversion in fact occurs for all propositions, resulting in a total collapse of the modal square of opposition.
-
-%\marginpar{@Bruno: Shall we add a sentence or two here?}
 

### 2016-08-14 Papers/2016/IJCAI/FinalVersion/main.tex
Leibniz/Lenzen initial commit

--- 2016-04-13 Papers/InProgress/IJCAI/main.tex
@@
 \documentclass{article}
-\usepackage{ijcai15}
+\usepackage{ijcai16}
 
@@
 \usepackage{times}
+
+\pdfinfo{
+/Title (The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics)
+/Author (Christoph Benzm\"uller, Bruno Woltzenlogel Paleo) }
 
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth]{./Images/Scott_S5U.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./Scott_S5U.png}}
 \caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
@@
 \begin{figure}[t]
-\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./Images/QML_S5U.png}}
+\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./QML_S5U.png}}
 \caption{Improved Embedding of \SFiveU} \label{QML_S5U}
@@
 \begin{figure}
-\centerline{\includegraphics[width=1\columnwidth]{./Images/UglyEssence.png}}
+\centerline{\includegraphics[width=1\columnwidth]{./UglyEssence.png}}
 \caption{Definition of Essence using Old Syntax} \label{UglyEssence}

### 2016-08-14 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Revier 1. (1) Koon changed to Koons
Reviewer 1: (2) G(x) changed to G(\gamma)
Reviewer 2:  (D.7) Section 4, Proof of Theorem 2
(D.8) Section 4, Proof of Lemma 1
(D.9) Section 4, Proof of Lemma 2
(D.10) Section 5, Proof of Lemma 1
(D.11) Section 6, Proof of Theorem 1
(D.4) Section 3

@@
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
-From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models \citep{negri}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2}): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
 In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
@@
 
-	\AXC{$ $} \RightLabel{5}
+	\AXC{$ $} \RightLabel{4}
 	\UIC{$\rho(\gamma)$}
 
-		\AXC{$ $} \RightLabel{2}
+		\AXC{$ $} \RightLabel{1}
 		\UIC{$\neg P(\rho)$} 
@@
 
-					\AXC{$ $} \RightLabel{1}
+					\AXC{$ $} \RightLabel{5}
 					\UIC{$G(\gamma)$} \dottedLine
@@
 				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
-	   \BIC{$\bot$} \RightLabel{$\neg_I^2$}
+	   \BIC{$\bot$} \RightLabel{$\neg_I^1$}
 	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
@@
 \begin{prooftree}
-\AXC{$ $} \RightLabel{1}
+\AXC{$ $} \RightLabel{5}
 \UIC{$G(\gamma)$}
@@
 		 \BIC{$\nec P(\rho)\scott$}
-			\AXC{$ $} \RightLabel{4}
+			\AXC{$ $} \RightLabel{3}
 			\UIC{$\omega: P(\rho)$}
-					\AXC{$ $} \RightLabel{3}
+					\AXC{$ $} \RightLabel{2}
 					\UIC{$\omega: G(\delta) $} \dottedLine
@@
 					\UIC{$\omega: P(\rho) \imp \rho(\delta)$} \RightLabel{$\imp_E$}
-				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^3$}
+				\BIC{$\omega: \rho(\delta) $} \RightLabel{$\imp_I^2$}
 				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
-				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^4$}
+				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^3$}
 				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
 				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
-						\AXC{ K} \RightLabel{$\imp_E$}
-					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$}
-			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^5$}
+						\AXC{Axiom K} \RightLabel{$\imp_E$}
+					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
+			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
 			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
@@
 		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
-		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^1$}
+		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^5$}
 		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$} \RightLabel{$\all_I$}
@@
 \UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
-\UIC{$[\subproof_4] \qquad G(\gamma)$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$\subproof_4$} \dashedLine
 \UIC{$G(\gamma)$}
@@
 		\UIC{$\all x. G(x) \imp \ess{G}{x}$} \RightLabel{$\all_E$}
-		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}
+		\UIC{$G(\gamma) \imp \ess{G}{\gamma}$}  \RightLabel{$\imp_E$}
 	\BIC{$\ess{G}{\gamma}\scott$}
@@
 				\UIC{$P(E)$}
-						\AXC{$\subproof_4$} \dashedLine
+						\AXC{$ $} \RightLabel{1}
+						\UIC{$\ex z. G(z)$} \RightLabel{$\ex_E$}
 						\UIC{$G(\gamma)$} \dottedLine
 						\UIC{$\all \varphi. P(\varphi) \imp \varphi(\gamma)$} \RightLabel{$\all_E$}
-						\UIC{$P(E) \imp E(\gamma)$} 
+						\UIC{$P(E) \imp E(\gamma)$} \RightLabel{$\imp_E$}
 					\BIC{$E(\gamma)\scott$} \dottedLine
 					\UIC{$\all \varphi. \ess{\varphi}{\gamma} \imp \nec \ex x. \varphi(x)$} \RightLabel{$\all_E$}
-					\UIC{$\ess{G}{\gamma} \imp \nec \ex x. G(x)$}
+					\UIC{$\ess{G}{\gamma} \imp \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
 		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
@@
 		\AXC{Lemma 1} \dashedLine
-		\UIC{$\omega: \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\nec_I$}
+		\UIC{$\omega: \ex x. G(x) \imp \nec \ex x. G(x)$} \RightLabel{$\nec_I$}
 		\UIC{$\nec (\ex x. G(x) \imp \nec \ex x. G(x)) $}
@@
 \begin{proof} \hfill
-\begin{prooftree}
-\AXC{$ $} \RightLabel{1}
-\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
-\UIC{$[\subproof_6] \qquad G(\gamma)$}
-\end{prooftree}
-
-\begin{tiny}
+
+\begin{small}
 \begin{prooftree}
@@
 \BIC{$P(\lambda x. \neg \rho(x))$}
-\AXC{$\subproof_6$} \dashedLine
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
 \UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{N:D1}}
@@
 \end{prooftree}
-\end{tiny}
+\end{small}
 
@@
 \UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
-\AXC{$\subproof_6$} \dashedLine
-\UIC{$G(\gamma)$}  
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
+\UIC{$G(\gamma)$}  \RightLabel{$\wedge_I$}
 \BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
@@
 \AXC{Axiom \ref{A5}} \dashedLine
-\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{N:D3}}
-\AXC{$\subproof_6$} \dashedLine
+\UIC{$P(E)$}\dottedLine 
+\AXC{$ $} \RightLabel{1}
+\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
 \UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{N:D1}}
@@
 \AXC{Theorem 2}\dashedLine
-\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
-\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}
+\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\dottedLine
+\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}\doubleLine
 \UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
@@
 Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
-Koon \citep{koons}, Anderson \citep{and90}[p.
+Koons \citep{koons}, Anderson \citep{and90}[p.
 292] and Sobel \citep{sobel2}[p.
@@
 \UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
-\AXC{$ $} \RightLabel{6}
+\AXC{$ $} \RightLabel{4}
 \UIC{$\nec\rho(\gamma)$} \RightLabel{}
@@
 \begin{prooftree}
-\AXC{$ $} \RightLabel{5}
+\AXC{$ $} \RightLabel{3}
 \UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
 \UIC{$\omega: P(\rho)$}
-\AXC{$ $} \RightLabel{4}
+\AXC{$ $} \RightLabel{2}
 \UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{A:D1}}
@@
 \BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
-\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
+\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^2$}
 \UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
 \UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
-\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
+\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^3$}
 \UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
@@
 \UIC{$\nec P(\rho)$}
-\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
+\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^4$}
 \UIC{$[\subproof_8] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
@@
 \UIC{$P(G)$}
-\AXC{$ $} \RightLabel{7}
+\AXC{$ $} \RightLabel{5}
 \UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\wedge_I$}
@@
 \UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
-\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^7$}
+\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
 \UIC{$[\subproof_9] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
@@
 \UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)]) $}
-\AXC{$ $} \RightLabel{6}
-\UIC{$G(x)$}\RightLabel{$\wedge_I$}
+\AXC{$ \subproof_6 $} \dashedLine
+\UIC{$G(\gamma)$}\RightLabel{$\wedge_I$}
 \BIC{$\ess{G}{\gamma}$}
@@
 \BIC{$\nec G(\gamma)$}
-\AXC{$ $} \RightLabel{5}
+\AXC{$ $} \RightLabel{2}
 \UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
-\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^5$}
+\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^2$}
 \UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
@@
 
+\bibitem[{\itshape Fitting \& Mendelsohn(1998)}]{fitting2}
+Fitting, M. \& Mendelsohn, R. 1998.  {\itshape First-order Modal Logic}, Synthese Library vol. 277. Kluwer Academic Publishers.
+
 \bibitem[{\itshape Fitting(2002)}]{fitting}
@@
 
-\bibitem[{\itshape Negri(2005)}]{negri}
-Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
+%\bibitem[{\itshape Negri(2005)}]{negri}
+%Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
 

### 2016-08-15 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
(D.6) Section 4

@@
   (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
+  & s_{\tau} = t_{\tau} \ \mid \\
 & \bot_o \ \mid \
@@
 Furthermore, subscript types are omitted when they are clear from the context.
-Negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
+Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
 \end{definition}
@@
 				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
-			\BIC{$\bot$}
+			\BIC{$[\subproof_3] \qquad \bot$}
 \end{prooftree}
@@
         \AXC{$ $} \RightLabel{2}
-        \UIC{$\rho(\alpha)$} \RightLabel{$\ex_I$}
+        \UIC{$\omega: \rho(\alpha)$} \RightLabel{$\ex_I$}
 		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
@@
 	\UIC{$\nec \all x. \neg \rho(x) \scott$}
-	        \AXC{$ $} \dashedLine
+	        \AXC{$\subproof_3 $} \dashedLine
 	        \UIC{$\bot$} \RightLabel{$\neg_I^3$}
@@
 	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
-	   \UIC{$[\subproof_3] \qquad P(\rho)$}
+	   \UIC{$[\subproof_4] \qquad P(\rho)$}
 	   \end{prooftree}
@@
 
- \AXC{$\subproof_3$}\dashedLine
+ \AXC{$\subproof_4$}\dashedLine
  \UIC{$P(\rho)$}

### 2016-08-18 Papers/InProgress/NaturalDeduction/GodProof-ND.tex

@@
 
-Double lines are used to abbreviate tedious propositional reasoning steps in the derivations.
+Double lines are used to abbreviate tedious propositional reasoning steps in the derivations where the reasoning is otherwise clear.
 Dashed lines are used to refer to an axiom or theorem with the proof shown elsewhere.
@@
 \newcommand{\s}{\qquad}
+
+\begin{calculus}
+{Equality rules}
+{fig:EqualityRules}
+
+\vspace{1em}
+
+\s\s\s\s
+\infer[Ref]{t=t}{}
+\s\s
+\infer[Sym]{s=t}{ t=s }
+\s\s
+\infer[Trans]{s=t}{s=u & u=t}
+
+\vspace{1em}
+\end{calculus}
 
@@
 These principles are used in the ontological proofs formalized in the next sections.
-Although they are well-known among modal logicians, we include their proofs here in order to have a self-contained and fully formal presentation of the ontological proofs, without any gaps.
+Although they are well-known among modal logicians, we include their proofs here in order to have a self-contained and fully formal presentation of the ontological proofs.
 

### 2016-09-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2, Comment C, part 2

@@
 Secondly, it makes the ontological proof accessible to people who are familiar with a natural deduction style.
-Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this.
+Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicCoq}.
 
 The first main contribution of this paper is a detailed formalization of Scott's version \citep{scott} of G\"odel's ontological argument \citep{Goedel} (as shown in Section \ref{sec:scott}) in the proposed natural deduction calculus.
-The second main contribution of the paper is a new proof (also in natural deduction style).
+The second main contribution of the paper is a new proof (shown in Section \ref{sec:newproof}), also in natural deduction style.
 In contrast to Scott's proof \citep{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
 The new proof also does not rely on the equality predicate and is much shorter.
+Although there have been claims (as discussed in more detail in Section \ref{sec:newproof}) that \textbf{KB} is sufficent and the equality predicate is superfluous, formal proofs substantiating these claims were missing up to now.
 

### 2016-09-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2, Comment C

@@
 In this paper we present natural deduction formalizations of two of these solutions: Anderson's emendation (in Section \ref{sec:anderson}) and Bj{\o}rdal's variant (in Section \ref{sec:bjordal}).
+
+An early unpublished draft of Section \ref{newproof} of this paper,
+prepared by Woltzenlogel Paleo in March 2013, served as the starting
+point for two distinct but complementary research endeavours. Siders
+and Woltzenlogel Paleo transformed that early draft into this paper,
+aiming for human-readable and yet formal natural deduction proofs.
+Benzm\"uller and Woltzenlogel Paleo, on the other hand, focused on the
+verification of variants of G\"odel's ontological arguments using
+higher-order automated reasoning tools, which quickly led to many
+publications. The main automatically obtained results about G\"odel's
+and Scott's proofs appeared in the European Conference on Artificial
+Intelligence \citep{ECAI2014} and, two years later, a detailed
+analysis of an inconsistency found in G\"odel's axioms was published
+in the International Joint Conference on Artificial Intelligence
+\citep{IJCAI2016}. The automatic verification of variants by Anderson,
+Bjordal and H\'ajek, on the other hand were presented in the First
+World Congress on Logic and Religion
+\citep{WorldCongressLogicAndReligion} and published in
+\citep{AndersonHajekControversy} and results related to the modal
+collapse were reported in the 4th Word Congress on the Square of
+Opposition \citep{SquareOfOpposition} and later published in
+\citep{ModalCollapse}. More detailed reconstructions of Scott's
+variant within the interactive proof assistants Isabelle and Coq were
+relesed, respectively, in Isabellle's Archive of Formal Proofs
+\citep{IsabelleGod} and in the Symposium of Computer Science in Russia
+\citep{CSRModalLogicInCoq}. The popularity of the topic led to invited
+talks and corresponding invited abstracts in the proceedings of
+workshops \citep{WADT,WADTPreProceedings,IWIL} and summer schools
+\citep{SpiritualQuest,ReasoningWebSummerSchool} as well as articles
+targeting a more general audience \citep{AustrianComputerSociety}.
+More recently, this line of research ventured into other types of
+ontological arguments, and a computer-assisted reconstruction of an
+argument by Leibniz will appear as a chapter in a book dedicated to
+the 300th anniversary of Leibniz's death
+\citep{LeibnizOntologicalProof}. The automatic verification of Gödel's
+and Scott's variants \citep{ECAI2014} was preceded by an automatic
+verification of (a non-modal version of) Anselm's ontological argument
+\citep{ZaltaOppenheimer}. However, whereas the complexity of Gödel's
+and Scott's arguments required the embedding of higher-order modal
+logics into higher-order logics and the use of higher-order automated
+reasoners, Anselm's argument greater simplicity enabled it to be
+verified with a first-order prover. Anselm's argument has also been
+formalized within the interactive proof assistant PVS \citep{Rushby}
+and there has been a so far failed attempt to formalize Avicenna's
+argument in PVS as well.
 
@@
 
+\bibitem[{\itshape Oppenheimer \& Zalta(2011)}]{ZaltaOppenheimer}
+Oppenheimer, P. \& Zalta, E. 2011. {\itshape A Computationally-Discovered Simplification of the Ontological Argument}. Australasian Journal of Philosophy, 89(2), 333-349.
+
+\bibitem[{\itshape Rushby(2013)}]{Rushby}
+Rushby, J. 2013. {\itshape The Ontological Argument in PVS}. CAV Workshop ``Fun with Formal Methods''. St. Petersburg.
+
+% TODO: transform the following bibtex entries into bibitems
+
+% @incollection{LeibnizOntologicalProof,
+%   author      = {Bentert, Matthias and Benzm\"uller, Christoph and Streit, David and Woltzenlogel Paleo, Bruno},
+%   title       = {Analysis of an Ontological Proof Proposed by Leibniz},
+%   editor      = {Tandy, Charles},
+%   booktitle   = {[to appear in] Death and Anti-Death, Volume 14: Four Decades after Michael Polanyi, Three Centuries after G. W. Leibniz},
+%   publisher   = {Ria University Press},
+%   address     = {},
+%   isbn        = {978-1-934297-25-4},
+%   pages       = {},
+%   chapter     = {},
+%   year = {2016},
+%   note = {\textbf{[Invited]} }
+% }
+
+% @inproceedings{IJCAI2016,
+%   author = {Benzmüller, C. and Woltzenlogel Paleo, B.},
+%   title  = {The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics},
+%   booktitle = {International Joint Conference on Artificial Intelligence - IJCAI},
+%   keywords ={HOML,Bruno,COREA*},
+%   year = {2016}
+% }
+
+% @article{AndersonHajekControversy,
+%     author = {Benzmüller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
+%     title  = {Computer-Assisted Analysis of the Anderson-Hájek Ontological Controversy},
+%     journal = {Logica Universalis},
+%     publisher = {Springer},
+%     issn = {1661-8297},
+%     volume = {10},
+%     year = {2016},
+%     keywords = {Bruno,HOML},
+%     pages = {1--10},
+%     note = {\emph{Accepted for publication on the 28th of February 2016}}
+% }
+
+% @inproceedings{WorldCongressLogicAndReligion,
+%     author = {Benzmüller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
+%     title  = {Computer-Assisted Analysis of the Anderson-Hájek Ontological Controversy},
+%     booktitle = {1st World Congress on Logic and Religion},
+%     year = {2015},
+%     keywords = {Bruno,Abstract,HOML},
+%     pages = {49--50}
+% }
+
+% @inproceeding{WADT,
+%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {On Logic Embeddings and Gödel's God},
+%    booktitle = {Recent Trends in Algebraic Development Techniques},
+%    year = {2014},
+%    crossref = {ADT},
+%    keywords = {Bruno,Invited,HOML},
+%    note = {\textbf{[Invited]} }
+% }
+
+% @inproceedings{SpiritualQuest,
+%   author    = {Christoph Benzmüller and Bruno Woltzenlogel Paleo},
+%   title     = {Experiments in Computational Metaphysics: Gödel's Proof of God's Existence},
+%   booktitle = {Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November)},
+%   year      = {2015},
+%   crossref  = {SpiritualQuest},
+%   volume    = {9},
+%   keywords  = {Bruno,Invited,HOML},
+%   note = {\textbf{[Invited]} },
+%   pages = {23--40}
+% }
+
+% @incollection{ReasoningWebSummerSchool,
+%   author    = {Christoph Benzm{\"{u}}ller and
+%                Bruno Woltzenlogel Paleo},
+%   title     = {Higher-Order Modal Logics: Automation and Applications},
+%   booktitle = {Reasoning Web. Web Logic Rules - 11th International Summer School
+%                2015, Berlin, Germany, July 31 - August 4, 2015, Tutorial Lectures},
+%   pages     = {32--74},
+%   year      = {2015},
+%   crossref  = {DBLP:conf/rweb/2015},
+%   url       = {http://dx.doi.org/10.1007/978-3-319-21768-0_2},
+%   doi       = {10.1007/978-3-319-21768-0_2},
+%   timestamp = {Sat, 25 Jul 2015 15:58:28 +0200},
+%   biburl    = {http://dblp.uni-trier.de/rec/bib/conf/rweb/BenzmullerP15},
+%   bibsource = {dblp computer science bibliography, http://dblp.org},
+%   keywords = {Bruno,HOML,Invited,DBLP}
+% }
+
+% @inproceedings{CSRModalLogicInCoq,
+%   author    = {Christoph Benzm{\"{u}}ller and
+%                Bruno Woltzenlogel Paleo},
+%   title     = {Interacting with Modal Logics in the Coq Proof Assistant},
+%   booktitle = {Computer Science - Theory and Applications - 10th International Computer
+%                Science Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July
+%                13-17, 2015, Proceedings},
+%   pages     = {398--411},
+%   year      = {2015},
+%   crossref  = {DBLP:conf/csr/2015},
+%   url       = {http://dx.doi.org/10.1007/978-3-319-20297-6_25},
+%   doi       = {10.1007/978-3-319-20297-6_25},
+%   timestamp = {Tue, 23 Jun 2015 13:05:11 +0200},
+%   biburl    = {http://dblp.uni-trier.de/rec/bib/conf/csr/BenzmullerP15},
+%   bibsource = {dblp computer science bibliography, http://dblp.org},
+%   keywords  = {Bruno,HOML,COREC,DBLP}
+% }
+
+% @inproceedings{ECAI2014,
+%   Author = {Christoph Benzm{\"{u}}ller and Bruno Woltzenlogel Paleo},
+%   Bibsource = {dblp computer science bibliography, http://dblp.org},
+%   Biburl = {http://dblp.uni-trier.de/rec/bib/conf/ecai/BenzmullerP14},
+%   Booktitle = {{ECAI} 2014 - 21st European Conference on Artificial Intelligence, 18-22 August 2014, Prague, Czech Republic - Including Prestigious Applications of Intelligent Systems {(PAIS} 2014)},
+%   Crossref = {DBLP:conf/ecai/2014},
+%   Date-Modified = {2014-10-14 07:42:43 +0000},
+%   Doi = {10.3233/978-1-61499-419-0-93},
+%   Keywords = {Bruno,DBLP,COREA,HOML},
+%   Pages = {93--98},
+%   Timestamp = {Mon, 13 Oct 2014 13:39:42 +0200},
+%   Title = {Automating G{\"{o}}del's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers},
+%   Url = {http://dx.doi.org/10.3233/978-1-61499-419-0-93},
+%   Year = {2014},
+%     note = {\textbf{[Nominated for Best Paper]} }
+% }
+
+% @inproceedings{SquareOfOpposition,
+%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {Formalization and Automated Verification of Gödel's Proof of God's Existence},
+%    booktitle = {4th World Congress on the Square of Opposition},
+%    year = {2014},
+%    keywords = {Bruno,Abstract,HOML},
+%    pages = {24--25}
+% }
+
+% @inworkshopinvited{WADTPreProceedings,
+%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {On Logic Embeddings and Gödel's God},
+%    booktitle = {22nd International Workshop on Algebraic Development Techniques},
+%    year = {2014},
+%    keywords = {Bruno,Abstract,Invited,HOML},
+%    note = {\textbf{[Invited]} },
+%    pages = {8--9},
+%    url = {http://wadt2014.cs.ovgu.de/pre-proceedings.pdf}
+% }
+
+% @article{ModalCollapse,
+%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {The Modal Collapse as a Collapse of the Modal Square of Opposition},
+%    journal = {Studies in Universal Logic},
+%    publisher = {Springer},
+%    issn = {2297-0282},
+%    year = {2016},
+%    keywords = {Bruno,HOML},
+%    pages   =  {1--7},
+%    note    = {\emph{Submitted in November 2014. Accepted for publication on the 28th of February 2016}}
+% }
+
+% @article{IsabelleGod,
+%   author    = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%   title     = {G{\"{o}}del's God in Isabelle/HOL},
+%   journal   = {Archive of Formal Proofs},
+%   year      = {2013},
+%   volume    = {},
+%   keywords  = {Bruno,DBLP,HOML},
+%   url       = {http://afp.sourceforge.net/entries/GoedelGod.shtml},
+%   timestamp = {Wed, 15 Oct 2014 15:14:28 +0200},
+%   biburl    = {http://dblp.uni-trier.de/rec/bib/journals/afp/Benzmueller13},
+%   bibsource = {dblp computer science bibliography, http://dblp.org}
+% }
+
+% @inworkshopinvited{IWIL,
+%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {Gödel's God on the Computer},
+%    booktitle = {10th International Workshop on the Implementation of Logics},
+%    year = {2013},
+%    keywords = {Bruno,Abstract,Invited,HOML},
+%    note = {\textbf{[Invited]} },
+%    pages = {1--2}
+% }
+
+% @article{AustrianComputerSociety,
+%    author = {Woltzenlogel Paleo, Bruno},
+%    title  = {Automated Verification and Reconstruction of Gödel's Proof of God's Existence},
+%    journal = {Journal of the Austrian Computer Society},
+%    year = {2013},
+%    keywords = {Bruno,Invited,S,HOML},
+%    pages    = {4--6}
+% }
+
 \end{thebibliography}

### 2016-09-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
checking from section 6. Part 1.

@@
 the 300th anniversary of Leibniz's death
-\citep{LeibnizOntologicalProof}. The automatic verification of Gödel's
+\citep{LeibnizOntologicalProof}. The automatic verification of G\"odel's
 and Scott's variants \citep{ECAI2014} was preceded by an automatic
 verification of (a non-modal version of) Anselm's ontological argument
-\citep{ZaltaOppenheimer}. However, whereas the complexity of Gödel's
+\citep{ZaltaOppenheimer}. However, whereas the complexity of G\"odel's
 and Scott's arguments required the embedding of higher-order modal
@@
 
-\begin{scriptsize}
+\begin{small}
 \begin{prooftree}
@@
 	   \end{prooftree}
-\end{scriptsize}
-
-\begin{scriptsize}
+\end{small}
+
+\begin{small}
 \begin{prooftree}
@@
 				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
-				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \RightLabel{$\imp_E$}
-						\AXC{Axiom K} \RightLabel{$\imp_E$}
-					\BIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
+				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \doubleLine \RightLabel{Axiom K}
+
+					\UIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
 			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
@@
 \end{prooftree}
-\end{scriptsize}
+\end{small}
 
@@
 \UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
-\UIC{$P(\rho)$}
-\AXC{Axiom \ref{N:A4}} \dashedLine
-\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
-\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
-\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
-\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
+\UIC{$[\subproof_7] \qquad P(\rho)$}
 \end{prooftree}
@@
 \UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
-\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
-\AXC{$ $} \RightLabel{6}
-\UIC{$\rho(\gamma)$} \RightLabel{}
-\AXC{$\subproof_7$} \dashedLine
-\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
-\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$}
+
+%\AXC{$ $} \RightLabel{6}
+%\UIC{$\rho(\gamma)$} \RightLabel{}
+\AXC{$ \subproof_7$} \dashedLine
+\UIC{$P(\rho)$} %\RightLabel{$\nec_E$}
+\AXC{Axiom \ref{N:A4}} \dashedLine
+\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
+\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
+\BIC{$\nec P(\rho)$}\RightLabel{$\imp_E$}%\RightLabel{$\imp_I^3$}
+%\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
+%\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
+%\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
+
 \BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
@@
 \AXC{$ $} \RightLabel{4}
-\UIC{$\nec\rho(\gamma)$} \RightLabel{}
-\BIC{$P(\rho)$} \RightLabel{$\imp_E$}
+\UIC{$\nec\rho(\gamma)$} \RightLabel{$\biimp_E$}
+\BIC{$P(\rho)$} \RightLabel{$\forall_E$}
 
@@
 \UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^3$}
-\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
+\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$}
 
 \AXC{$\subproof_7$} \dashedLine
-\UIC{$\nec P(\rho)$}
+\UIC{$\nec P(\rho)$}\RightLabel{$\imp_E$}
 \BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^4$}
@@
 \begin{prooftree}
-\AXC{Axiom \ref{A:A3}} 
+\AXC{Axiom \ref{A:A3}} \dashedLine
 \UIC{$P(G)$}
@@
 
-\begin{tiny}
+\begin{small}
 \begin{prooftree}
 \AXC{$\subproof_8$ and $\subproof_9$} \dashedLine
-\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $}
+\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $} \RightLabel{$\forall_I$}
 \UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)]) $}
@@
 \UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
-\BIC{$\nec E(\gamma)$}
-    \AXC{Axiom T} \dashedLine
-    \UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
-\BIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
+\BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$}
+   % \AXC{Axiom T} \dashedLine
+    %\UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
+\UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
 \UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
@@
 \end{prooftree}
-\end{tiny}
+\end{small}
+%\end{tiny}
 \end{proof}
@@
 
-\begin{scriptsize}
+%\begin{scriptsize}
 \begin{prooftree}
@@
 \UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$} \doubleLine  \RightLabel{(using Axiom K)}
+\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$} \doubleLine  \RightLabel{(Axiom K)}
 \UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
@@
 \end{prooftree}
-\end{scriptsize}
+%\end{scriptsize}
 
@@
 \BIC{$ \nec (\ex x. G(x) \imp A)$} \RightLabel{$\nec_E$}
-\UIC{$ \omega:  \ex x. G(x) \imp A$} \RightLabel{$\imp_E$}
+\UIC{$ \omega:  \ex x. G(x) \imp A$} %\RightLabel{$\imp_E$}
 \AXC{Theorem 3}\dashedLine

### 2016-09-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2; Comment D3.

@@
 Nevertheless, some authors of variants \citep[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
-Our choice of assuming constant domains is motivated by simplicity.
+Our choice of constant domains is motivated by simplicity.
+As is well-known, varying domain quantifiers (also known as \emph{actualistic} quantifiers) can be simulated by constant domain quantifiers (also known as \emph{possibilistic} quantifiers) guarded by a primitive \emph{existence} predicate\footnote{Formally, $\forall^e x.
+F[x] \equiv \forall x.
+E(x) \imp F[x]$ and $\exists^e x.
+F[x] \equiv \exists x.
+E(x) \wedge F[x]$}; therefore, the formalization of ontological arguments with varying domain quantification would be feasible in the natural deduction calculus presented here as well.
 

### 2016-09-06 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
reviewer 2: (D.4) Blackburn reference added.

@@
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
-From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2}): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2} or \citep[Section 3.1]{blackburn} on Frame definability): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
 In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
@@
 
+\bibitem[{\itshape Blackburn \& al.(2001)}]{blackburn}
+Blackburn, P. \&  M. de Rijke \& Y. Venema. 2001. {\itshape Modal Logic}. Cambridge: Cambridge University Press.
+
 \bibitem[{\itshape Church(1940)}]{church}
@@
 % @inproceedings{IJCAI2016,
-%   author = {Benzmüller, C. and Woltzenlogel Paleo, B.},
+%   author = {Benzm?ller, C. and Woltzenlogel Paleo, B.},
 %   title  = {The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics},
@@
 % @article{AndersonHajekControversy,
-%     author = {Benzmüller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
-%     title  = {Computer-Assisted Analysis of the Anderson-Hájek Ontological Controversy},
+%     author = {Benzm?ller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
+%     title  = {Computer-Assisted Analysis of the Anderson-H?jek Ontological Controversy},
 %     journal = {Logica Universalis},
@@
 % @inproceedings{WorldCongressLogicAndReligion,
-%     author = {Benzmüller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
-%     title  = {Computer-Assisted Analysis of the Anderson-Hájek Ontological Controversy},
+%     author = {Benzm?ller, Christoph and Weber, Leon and Woltzenlogel Paleo, Bruno},
+%     title  = {Computer-Assisted Analysis of the Anderson-H?jek Ontological Controversy},
 %     booktitle = {1st World Congress on Logic and Religion},
@@
 % @inproceeding{WADT,
-%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
-%    title  = {On Logic Embeddings and Gödel's God},
+%    author = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {On Logic Embeddings and G?del's God},
 %    booktitle = {Recent Trends in Algebraic Development Techniques},
@@
 % @inproceedings{SpiritualQuest,
-%   author    = {Christoph Benzmüller and Bruno Woltzenlogel Paleo},
-%   title     = {Experiments in Computational Metaphysics: Gödel's Proof of God's Existence},
+%   author    = {Christoph Benzm?ller and Bruno Woltzenlogel Paleo},
+%   title     = {Experiments in Computational Metaphysics: G?del's Proof of God's Existence},
 %   booktitle = {Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November)},
@@
 % @inproceedings{SquareOfOpposition,
-%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
-%    title  = {Formalization and Automated Verification of Gödel's Proof of God's Existence},
+%    author = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {Formalization and Automated Verification of G?del's Proof of God's Existence},
 %    booktitle = {4th World Congress on the Square of Opposition},
@@
 % @inworkshopinvited{WADTPreProceedings,
-%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
-%    title  = {On Logic Embeddings and Gödel's God},
+%    author = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {On Logic Embeddings and G?del's God},
 %    booktitle = {22nd International Workshop on Algebraic Development Techniques},
@@
 % @article{ModalCollapse,
-%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%    author = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
 %    title  = {The Modal Collapse as a Collapse of the Modal Square of Opposition},
@@
 % @article{IsabelleGod,
-%   author    = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
+%   author    = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
 %   title     = {G{\"{o}}del's God in Isabelle/HOL},
@@
 % @inworkshopinvited{IWIL,
-%    author = {Benzmüller, Christoph and Woltzenlogel Paleo, Bruno},
-%    title  = {Gödel's God on the Computer},
+%    author = {Benzm?ller, Christoph and Woltzenlogel Paleo, Bruno},
+%    title  = {G?del's God on the Computer},
 %    booktitle = {10th International Workshop on the Implementation of Logics},
@@
 %    author = {Woltzenlogel Paleo, Bruno},
-%    title  = {Automated Verification and Reconstruction of Gödel's Proof of God's Existence},
+%    title  = {Automated Verification and Reconstruction of G?del's Proof of God's Existence},
 %    journal = {Journal of the Austrian Computer Society},

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
checking from section 6. Part 2.
Move equality rules

@@
 
-The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical}, \ref{fig:QuantifierRules} and \ref{fig:EqualityRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
 
 \newcommand{\s}{\qquad}
-
-\begin{calculus}
-{Equality rules}
-{fig:EqualityRules}
-
-\vspace{1em}
-
-\s\s\s\s
-\infer[Ref]{t=t}{}
-\s\s
-\infer[Sym]{s=t}{ t=s }
-\s\s
-\infer[Trans]{s=t}{s=u & u=t}
-
-\vspace{1em}
-\end{calculus}
 
@@
 
+\end{calculus}
+
+\begin{calculus}
+{Equality rules}
+{fig:EqualityRules}
+
+\vspace{1em}
+
+\s\s\s\s
+\infer[Ref]{t=t}{}
+\s\s
+\infer[Sym]{s=t}{ t=s }
+\s\s
+\infer[Trans]{s=t}{s=u & u=t}
+
+\vspace{1em}
 \end{calculus}
@@
 \UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
-\UIC{$[\subproof_7] \qquad P(\rho)$}
+\UIC{$[\subproof_6] \qquad P(\rho)$}
 \end{prooftree}
@@
 %\UIC{$\rho(\gamma)$} \RightLabel{}
-\AXC{$ \subproof_7$} \dashedLine
+\AXC{$ \subproof_6$} \dashedLine
 \UIC{$P(\rho)$} %\RightLabel{$\nec_E$}
@@
 \BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
-\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$\subproof_8$} \dashedLine
+\UIC{$[\subproof_7] \qquad \ess{G}{\gamma}$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\subproof_7$} \dashedLine
 \UIC{$\ess{G}{\gamma}$}
@@
 \UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
-\UIC{$[\subproof_9] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_9 $}\dashedLine
+\UIC{$[\subproof_8] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_8 $}\dashedLine
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
@@
 \UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
-\UIC{$[\subproof_6] \quad G(\gamma)$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$\subproof_6$} \dashedLine
+\UIC{$[\subproof_9] \quad G(\gamma)$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$\subproof_9$} \dashedLine
 \UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
@@
 \UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
-\BIC{$[\subproof_7] \quad \nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$[\subproof_{10}] \quad \nec P(\rho)$} \RightLabel{$\imp_E$}
 \end{prooftree}
@@
 
-\AXC{$\subproof_7$} \dashedLine
+\AXC{$\subproof_{10}$} \dashedLine
 \UIC{$\nec P(\rho)$}\RightLabel{$\imp_E$}
 \BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^4$}
-\UIC{$[\subproof_8] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
+\UIC{$[\subproof_{11}] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
 \end{prooftree}
@@
 \BIC{$P(\varphi)$}
-\AXC{$\subproof_6$} \dashedLine 
+\AXC{$\subproof_9$} \dashedLine 
 \UIC{$  G(\gamma)$}  \RightLabel{D\ref{A:D1}}
@@
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
-\UIC{$[\subproof_9] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
+\UIC{$[\subproof_{12}] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
 \end{prooftree}
@@
 \begin{prooftree}
-\AXC{$\subproof_8$ and $\subproof_9$} \dashedLine
+\AXC{$\subproof_{11}$ and $\subproof_{12}$} \dashedLine
 \UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $} \RightLabel{$\forall_I$}
 \UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)]) $}
-\AXC{$ \subproof_6 $} \dashedLine
+\AXC{$ \subproof_{9}$} \dashedLine
 \UIC{$G(\gamma)$}\RightLabel{$\wedge_I$}
@@
 \UIC{$P(E)$}\dottedLine 
-\AXC{$\subproof_6$} \dashedLine
+\AXC{$\subproof_9$} \dashedLine
 \UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{A:D1}}
@@
 \begin{prooftree}
-\AXC{$ $} \RightLabel{1}
+\AXC{$ $} \RightLabel{2}
 \UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
-\UIC{$[\subproof_6] \quad G(\gamma)$}
-\end{prooftree}
-
-%\begin{scriptsize}
-\begin{prooftree}
-\AXC{$\subproof_6$} \dashedLine
 \UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
@@
 \BIC{$\nec G(\gamma)$}
-\AXC{$ $} \RightLabel{2}
+\AXC{$ $} \RightLabel{1}
 \UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
-\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^2$}
+\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^1$}
 \UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
@@
 \UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
-\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^1$}
+\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^2$}
 \UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
@@
 \UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
-\UIC{$[\subproof_{10}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_{10} $}\dashedLine
+\UIC{$[\subproof_{13}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{13} $}\dashedLine
 \UIC{$\ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
@@
 \UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
-\BIC{$[\subproof_{11}] \qquad G_B(x) \wedge P_B(G_B) $}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_{11} $}\dashedLine
+\BIC{$[\subproof_{14}] \qquad G_B(x) \wedge P_B(G_B) $}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{14} $}\dashedLine
 \UIC{$G_B(x) \wedge P_B(G_B) $}
@@
 \BIC{$G_B(x) \wedge P_B(G_B) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y))) $} \RightLabel{ D\ref{B:D2} }\dottedLine
-\UIC{$[\subproof_{12}] \qquad \MCP(G_B, x) $}
+\UIC{$[\subproof_{15}] \qquad \MCP(G_B, x) $}
 
@@
 \UIC{$(\MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
-\AXC{$ \subproof_{12} $}\dashedLine
+\AXC{$ \subproof_{15} $}\dashedLine
 \UIC{$\MCP(G_B, x) $} \RightLabel{ $\imp_E$}

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2, Comment A, Part 1: Make the language of definition 2 match with the language of the calculus

@@
   \imp_{o\imp o\imp o} \ \mid \ 
-  \wedge_{o\imp o\imp o} \ \mid \
-  \vee_{o\imp o\imp o} \ \mid \\
+  \wedge_{o\imp o\imp o} \ \mid \\
 & \all_{(\tau\imp o)\imp o} \ \mid \ 
@@
 Furthermore, subscript types are omitted when they are clear from the context.
-Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
+Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
 \end{definition}

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2, comment A (i).
Reviewer 2: Leibniz's Equality

@@
   (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
-  & s_{\tau} = t_{\tau} \ \mid \\
 & \bot_o \ \mid \
@@
 Furthermore, subscript types are omitted when they are clear from the context.
-Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
+Negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$), equivalence ($\biimp_{o\imp o\imp o}$), equality ($=_{\tau\imp\tau\imp o}$), inequality ($\neq_{\tau\imp\tau\imp o}$), are defined by $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$, $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$, $s = t \equiv \forall P.
+(P \ s) \biimp (P \ t)$ (Leibniz's equality) and $s \neq t\equiv \neg (s=t)$, .
 \end{definition}
@@
 %Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}. 
-Beta-reduction is implicit; all rules are assumed to operate modulo beta-reduction.
+Beta-reduction and alpha-conversion are implicit: all rules are assumed to operate modulo these operations.
 A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules.
@@
 
-The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical}, \ref{fig:QuantifierRules} and \ref{fig:EqualityRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
 

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Remove primitive equality rules
Reviewer 2: alpha-conversion
Remove comment
Reviewer 2: Beta-reduction

@@
 Furthermore, subscript types are omitted when they are clear from the context.
+Implicit alpha-conversion is assumed: two expressions are considered to be the same modulo renaming of their bound variables.
 Negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$), equivalence ($\biimp_{o\imp o\imp o}$), equality ($=_{\tau\imp\tau\imp o}$), inequality ($\neq_{\tau\imp\tau\imp o}$), are defined by $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$, $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$, $s = t \equiv \forall P.
@@
 The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
-The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}.
+The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
 Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
-%Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}. 
-Beta-reduction and alpha-conversion are implicit: all rules are assumed to operate modulo these operations.
 A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules.
@@
 
+Applications of the $\beta$ rule are silently ommitted.
 Double lines are used to abbreviate tedious propositional reasoning steps in the derivations where the reasoning is otherwise clear.
@@
 \begin{calculus}
-{Equality rules}
-{fig:EqualityRules}
+{Reduction rules}
+{fig:ReductionRules}
 
@@
 
-\s\s\s\s
-\infer[Ref]{t=t}{}
-\s\s
-\infer[Sym]{s=t}{ t=s }
-\s\s
-\infer[Trans]{s=t}{s=u & u=t}
+\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad
+\infer[\beta]{B}{A}
 
 \vspace{1em}
+
+\begin{center}
+
+only if $B$ and $A$ are $\beta$-equivalent \\ (i.e. there is a $C$ such that both $B$ and $A$ $\beta$-reduce to $C$)
+
+\end{center}
+
+\vspace{1em}
+
 \end{calculus}

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: Ancestor/Descendant

@@
 
+In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas changes due to a substitution).
+For any expression occurring both in a formula in the premise and in its corresponding (sub-)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
+
 \newcommand{\subproof}{\star}

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Restrict condition

@@
 \textbf{eigen-variable conditions:} \\
-if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$; \\
-if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$.
+if $\rho$ is a $\all_I$ inference eliminating a variable $\alpha$, then any occurrence of $\alpha$ in the proof should be an ancestor of the occurrence of $\alpha$ eliminated by $\rho$ and not a descendant of a variable introduced by a $\ex_E$ inference; \\
+if $\rho$ is a $\ex_E$ inference introducing a variable $\beta$, then any occurrence of $\beta$ in the proof should be a descendant of the occurrence of $\beta$ introduced by $\rho$ and not an ancestor of a variable eliminated by a $\all_I$ inference.
 \end{center}

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: Define "Access"

@@
 \textbf{eigen-box condition:}\\ 
+An modal inference is said to \emph{access}\\
+ the box immediately above or below it. \\
 $\nec_I$ and $\pos_E$ are \emph{strong} modal rules: \\

### 2016-09-09 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix typo
Reviewer 2; comment D2: clarify novel aspects of the calculus in the introduction

@@
 For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz} or \citep{gabbay}.
+The natural deduction calculus proposed and used in this paper is an extension of a standard natural deduction calculus for (non-modal) higher-order logic \citep{BezmuellerBrownKohlhase} with new introduction and elimination rules for modal operators (as presented in Section \ref{sec:calculus}).
+This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}).
+
 The higher-order modal natural deduction calculus proposed here has special introduction and elimination rules for modalities, as defined in Section \ref{sec:calculus}, and is sound and complete relative to an axiomatic modal calculus, as proven in Section \ref{sec:compl-sound}.
@@
 Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
-Secondly, it makes the ontological proof accessible to people who are familiar with a natural deduction style.
+Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
 Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicCoq}.

### 2016-09-09 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2; Comment D2: Further details

@@
 This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}).
-
-The higher-order modal natural deduction calculus proposed here has special introduction and elimination rules for modalities, as defined in Section \ref{sec:calculus}, and is sound and complete relative to an axiomatic modal calculus, as proven in Section \ref{sec:compl-sound}.
+There are various higher-order logics \citep{BezmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}.
+The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles.
+With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient for achieving the stated aim.
 

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: B: Fix box introduction in proof of iteration lemma
Fix typo

@@
 
-G\"odel bulit on Leibniz's work \citep{adams} and brought the ontological argument to a modern form using a modal logic with higher-order quantification over properties.
+G\"odel built on Leibniz's work \citep{adams} and brought the ontological argument to a modern form using a modal logic with higher-order quantification over properties.
 In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
@@
 
-\AXC{Brouwer's Reduction}\dashedLine
-\UIC{$\pos \nec A\imp A$}\RightLabel{$\nec_I$}
+\AXC{$\omega:$ Brouwer's Reduction}\dashedLine
+\UIC{$\omega: \pos \nec A\imp A$}\RightLabel{$\nec_I$}
 \UIC{$\nec(\pos \nec A\imp A)$}

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Add a fourth reason for natural deduction style.

@@
 
-The natural deduction style was chosen for three reasons.
+The natural deduction style was chosen for four reasons.
 Firstly, presentations of G\"odel's proof are typically either informal or formalized in other styles of proof calculi (e.g.
@@
 Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicCoq}.
+And finally, in comparison to other kinds of proof systems (e.g.
+resolution, tableaux, sequent calculi), natural deduction can be considered less technical, more human-readable and more intuitive, because it relies on natural principles of reasoning such as modus ponens (implication elimination) and reasoning by assumption (implication introduction); these qualities are conducive to rendering the formalizations of ontological arguments accessible to a wider audience.
 

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Relate ND to the calculus of Benzmueller, Brown and Kohlhase

@@
 
-The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
+The natural deduction calculus proposed and used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
 The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
-Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
+Finally, modal operators are handled by special new rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
 A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules.
@@
 
-The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND.
+This calculus is essentially the same as the simplest natural deduction calculus in the family of calculi discussed in \citep{BenzmuellerBrownKohlhase}.
+The main difference is that they take disjunction as a primitive connective instead of conjunction, and they use a sequent notation, where the assumptions are listed in the left side of the sequent.
+The calculus with the new rules for modal operators shown in Figure \ref{fig:NDK} is named \NDK.
 
@@
 
+\bibitem[{\itshape Benzm\"uller \& Brown \& Kohlhase(2004)}]{BenzmuellerBrownKohlhase}
+Benzm\"uller, C. \& Brown C. \& Kohlhase, M. In: {\ishape Journal of Symbolic Logic}, Vol. 69, N. 4, pp. 1027--1088.
+
 \bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{controversy}

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: Clarify which HOML

@@
 
-Adding the modal rules results in a calculus that is suitable for the basic modal logic \textbf{K}.
-In other words, {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
+{\ND} is a calculus for a (non-modal) higher-order logic without 
+functional and boolean extensionality principles \citep{BenzmuellerBrownKohlhase}.
+Adding the modal rules results in a calculus that is suitable for a basic rigid modal logic \textbf{K}, in the sense that {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
 

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: eta-expansion

@@
 The natural deduction calculus proposed and used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
-The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
+The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction\footnote{Rules for $\eta$-expansion \citep{BenzmuellerBrownKohlhase} are not included here, because they are not needed in the ontological arguments.} is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
 Finally, modal operators are handled by special new rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: Binding notation
Fix typo

@@
 The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles.
-With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient for achieving the stated aim.
+With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient to achieve the stated aim.
 
@@
 where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$.
-Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed.
+Parenthesis conventions, infix notation for propositional connectives and binding notation\footnote{$\forall x.
+F[x]$ and $\exists x.
+F[x]$ are assumed to be abbreviations for, respectively, $\forall (\lambda x.
+F[x])$ and $\exists (\lambda x.
+F[x])$.} for quantifiers are assumed.
 Furthermore, subscript types are omitted when they are clear from the context.

### 2016-09-11 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
reflexivity proof added

@@
 
+\begin{lemma*}[Reflexivity]
+$$
+s=s
+$$
+\end{lemma*}
+
+\begin{proof}\hfill
+
+\begin{footnotesize}
+\begin{prooftree}
+\AXC{}\RightLabel{$1$}
+\UIC{$P s$}\RightLabel{$\imp_I^1$}
+\UIC{$P s\imp P s$}
+\AXC{}\RightLabel{$2$}
+\UIC{$P s$}\RightLabel{$\imp_I^2$}
+\UIC{$P s\imp P s$}\RightLabel{$\wedge_I$}
+\BIC{$P s\biimp P s$}\RightLabel{$\forall_I$}
+\UIC{$\forall P. P s\biimp P s$}\dottedLine
+\UIC{$s=s$}
+\end{prooftree}
+\end{footnotesize}
+
+\end{proof}
+
 \section{Scott's Proof in Natural Deduction}\label{sec:scott}
@@
 
-In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the axiom of reflexivity for equality (in the proof of Theorem 1).
+In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the lemma of reflexivity for equality (in the proof of Theorem 1).
 The use of reflexivity could be avoided, but we chose to use it because it allows the construction of a shorter proof.
@@
 \bibitem[{\itshape Benzm\"uller \& Brown \& Kohlhase(2004)}]{BenzmuellerBrownKohlhase}
-Benzm\"uller, C. \& Brown C. \& Kohlhase, M. In: {\ishape Journal of Symbolic Logic}, Vol. 69, N. 4, pp. 1027--1088.
+Benzm\"uller, C. \& Brown C. \& Kohlhase, M. In: {\itshape Journal of Symbolic Logic}, Vol. 69, N. 4, pp. 1027--1088.
 

### 2016-09-11 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Proofs checked in the whole paper.
Only rules that are defined in the calculus are used. Labels to the inference steps have been added where they where missing. Some parentheses changed.
fix of overfull hboxes
bibtex entries into bibitems at end of the paper.

@@
 \BIC{$\nec\neg\nec\neg\neg A$}\doubleLine
-\UIC{$\nec\neg\nec A$} \RightLabel{$\neg_E$}
-\BIC{$\bot$}\RightLabel{$\neg_I^1$}
+\UIC{$\nec\neg\nec A$} \RightLabel{$\imp_E$}
+\BIC{$\bot$}\RightLabel{$\imp_I^1$}
 \UIC{$\neg\neg A$} \RightLabel{$\neg\neg_E$}
@@
 
-\AXC{$\omega:$ Brouwer's Reduction}\dashedLine
+\AXC{Brouwer's Reduction}\dashedLine
 \UIC{$\omega: \pos \nec A\imp A$}\RightLabel{$\nec_I$}
@@
 		\UIC{$\omega: \rho(\gamma) \imp \gamma = \gamma$} \RightLabel{$\all_I$}
-		\UIC{$\omega: \all x. \rho(x) \imp x = x$} \RightLabel{$\nec_I$}
-		\UIC{$\nec (\all x. \rho(x) \imp x = x)\scott $} \RightLabel{$\wedge_I$}
-	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) $} 
+		\UIC{$\omega: \all x. (\rho(x) \imp x = x)$} \RightLabel{$\nec_I$}
+		\UIC{$\nec \all x. (\rho(x) \imp x = x)\scott $} \RightLabel{$\wedge_I$}
+	\BIC{$P(\rho) \wedge \nec \all x. (\rho(x) \imp x = x) $} 
 			\AXC{Axiom 2 for $\rho$ and $\lambda x. x = x$} \dashedLine
-	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x = x) \imp P(\lambda x. x = x)$} 
+	    	\UIC{$P(\rho) \wedge \nec \all x. (\rho(x) \imp x = x) \imp P(\lambda x. x = x)$} \RightLabel{$\imp_E$}
 		\BIC{$[\subproof_1] \qquad P(\lambda x. x = x)\scott$} 
@@
 			\UIC{$\omega: \all x. \neg \rho(x) $} \RightLabel{$\all_E$}
-			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\neg_E$}
+			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\imp_E$}
 		\BIC{$\omega: \bot$} \RightLabel{$\bot_E$}
 		\UIC{$\omega: \neg (\beta = \beta)$} \dottedLine
-		\UIC{$\omega: (\beta \neq \beta)$} \RightLabel{$\imp_I^1$}
-		\UIC{$\omega: (\rho(\beta) \imp (\beta \neq \beta))$} \RightLabel{$\all_I$}
-		\UIC{$\omega: (\all x. \rho(x) \imp (x \neq x))$} \RightLabel{$\nec_I$}
-		\UIC{$\nec (\all x. \rho(x) \imp (x \neq x))\scott$}  \RightLabel{$\wedge_I$}
-	\BIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) $} 
+		\UIC{$\omega: \beta \neq \beta$} \RightLabel{$\imp_I^1$}
+		\UIC{$\omega: \rho(\beta) \imp \beta \neq \beta$} \RightLabel{$\all_I$}
+		\UIC{$\omega: \all x. (\rho(x) \imp x \neq x)$} \RightLabel{$\nec_I$}
+		\UIC{$\nec \all x. (\rho(x) \imp x \neq x)\scott$}  \RightLabel{$\wedge_I$}
+	\BIC{$P(\rho) \wedge \nec \all x. (\rho(x) \imp x \neq x) $} 
 			\AXC{Axiom 2 for $\rho$ and $\lambda x. x \neq x$} \dashedLine
-	    	\UIC{$P(\rho) \wedge \nec (\all x. \rho(x) \imp x \neq x) \imp P(\lambda x. x \neq x)$} 
+	    	\UIC{$P(\rho) \wedge \nec \all x. (\rho(x) \imp x \neq x) \imp P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
 		\BIC{$[\subproof_2] \qquad P(\lambda x. x \neq x)\scott$} 
@@
 					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
-				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
+				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
 			\BIC{$[\subproof_3] \qquad \bot$}
@@
         \UIC{$\omega: \rho(\alpha)$} \RightLabel{$\ex_I$}
-		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
-    \BIC{$\omega: \bot$} \RightLabel{$\neg_I^2$}
+		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\imp_E$}
+    \BIC{$\omega: \bot$} \RightLabel{$\imp_I^2$}
 	\UIC{$\omega: \neg \rho(\alpha)$} \RightLabel{$\all_I$}
@@
 	        \AXC{$\subproof_3 $} \dashedLine
-	        \UIC{$\bot$} \RightLabel{$\neg_I^3$}
-    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
-  		\BIC{$\bot$} \RightLabel{$\neg_I^4$}
+	        \UIC{$\bot$} \RightLabel{$\imp_I^3$}
+    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\imp_E$}
+  		\BIC{$\bot$} \RightLabel{$\imp_I^4$}
   		\UIC{$ \neg \nec \neg \ex x.\rho(x)$} \doubleLine %\RightLabel{definition of $\pos$}
@@
 					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
-				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
-	   \BIC{$\bot$} \RightLabel{$\neg_I^1$}
+				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\imp_E$}      
+	   \BIC{$\bot$} \RightLabel{$\imp_I^1$}
 	   \UIC{$\neg\neg P(\rho)$} \RightLabel{$\neg\neg_E$}
@@
 				\UIC{$\omega: G(\delta) \imp \rho(\delta) $} \RightLabel{$\all_I$}
-				\UIC{$\omega: \all y. G(y) \imp \rho(y) $} \RightLabel{$\imp_I^3$}
-				\UIC{$\omega: P(\rho) \imp (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
-				\UIC{$\nec (P(\rho) \imp (\all y. G(y) \imp \rho(y) ))$} \doubleLine \RightLabel{Axiom K}
-
-					\UIC{$\nec P(\rho) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
-			\BIC{$\nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
-			\UIC{$\rho(\gamma) \imp \nec (\all y. G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
-			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
-		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec (\all y. G(y) \imp \varphi(y) ))$} \dottedLine
+				\UIC{$\omega: \all y. (G(y) \imp \rho(y)) $} \RightLabel{$\imp_I^3$}
+				\UIC{$\omega: P(\rho) \imp \all y. (G(y) \imp \rho(y) )$} \RightLabel{$\nec_I$}
+				\UIC{$\nec (P(\rho) \imp \all y. (G(y) \imp \rho(y) ))$} \doubleLine \RightLabel{Axiom K}
+
+					\UIC{$\nec P(\rho) \imp \nec \all y. (G(y) \imp \rho(y) )$} \RightLabel{$\imp_E$}
+			\BIC{$\nec \all y. (G(y) \imp \rho(y) )$} \RightLabel{$\imp_I^4$}
+			\UIC{$\rho(\gamma) \imp \nec \all y. (G(y) \imp \rho(y) )$} \RightLabel{$\all_I$}
+			\UIC{$\all \varphi. \varphi(\gamma) \imp \nec \all y. (G(y) \imp \varphi(y) )$} \RightLabel{$\wedge_I$}
+		\BIC{$ G(\gamma) \wedge (\all \varphi. \varphi(\gamma) \imp \nec \all y. (G(y) \imp \varphi(y) ))$} \dottedLine
 		\UIC{$\ess{G}{\gamma}$} \RightLabel{$ \imp_I^5$}
@@
 \AXC{$ $} \RightLabel{3} 
-\UIC{$P(\rho)$} 
+\UIC{$P(\rho)$} \RightLabel{$\imp_E$}
 \BIC{$\bot$} \RightLabel{$\imp_I^2$}
@@
 \UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
-\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
+\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}\RightLabel{$\wedge_E$}
+\UIC{$P(\lambda x.\rho(x)) \imp \nec \rho(\gamma)$}
 \AXC{$ $} \RightLabel{4}
-\UIC{$\nec\rho(\gamma)$} \RightLabel{$\biimp_E$}
+\UIC{$\nec\rho(\gamma)$} \RightLabel{$\imp_E$}
 \BIC{$P(\rho)$} \RightLabel{$\forall_E$}
@@
 \UIC{$\omega: \forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
-\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\biimp_E$}
+\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\wedge_E$}
+\UIC{$\omega: P(\rho) \imp \nec \rho(y)$}\RightLabel{$\imp_E$}
 \BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
@@
 \UIC{$ \all \varphi. [P(\varphi)\biimp \nec \varphi(\gamma)]$}  \RightLabel{$\all_E$}
-\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
+\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\wedge_E$}
+\UIC{$ P(\varphi)\imp \nec \varphi(\gamma)$}  \RightLabel{$\imp_E$}
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
@@
 \begin{prooftree}
-\AXC{$\subproof_{11}$ and $\subproof_{12}$} \dashedLine
+\AXC{$\subproof_{11}$ and $\subproof_{12}$} \dashedLine \RightLabel{$\wedge_I$}
 \UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $} \RightLabel{$\forall_I$}
@@
 \UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
+\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\wedge_E$}
+\UIC{$P(E) \imp \nec E(\gamma)$}\RightLabel{$\imp_E$}
 \BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$}
@@
 \UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
-\UIC{$P(G) \biimp \nec G(\gamma)$}
+\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\wedge_E$}
+\UIC{$P(G) \imp \nec G(\gamma)$}
 \AXC{Axiom \ref{A:A3}} \RightLabel{}
-\UIC{$P(G)$} \RightLabel{$\biimp_E$}
+\UIC{$P(G)$} \RightLabel{$\imp_E$}
 \BIC{$\nec G(\gamma)$}
@@
 \AXC{$ $} \RightLabel{3} 
-\UIC{$P(G_B)$} 
+\UIC{$P(G_B)$} \RightLabel{$\imp_E$}
 \BIC{$\bot$} \RightLabel{$\imp_I^2$}
@@
 
+\bibitem[{\itshape Bentert \& Benzm\"uller \& Streit \& Woltzenlogel Paleo(2016)}]{LeibnizOntologicalProof}
+Bentert, M.
+\& Benzm\"uller, C.
+\& Streit, D.
+\& Woltzenlogel Paleo, B.
+2016.
+{\itshape Analysis of an Ontological Proof Proposed by Leibniz}.
+[to appear in]  Death and Anti-Death, Volume 14: Four Decades after Michael Polanyi, Three Centuries after G.
+W.
+Leibniz.
+ed.
+Tandy, C.
+Ria University Press.
 % @incollection{LeibnizOntologicalProof,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016)}]{IJCAI2016}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2016.
+{\itshape The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics}.
+International Joint Conference on Artificial Intelligence - IJCAI.
 % @inproceedings{IJCAI2016,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{AndersonHajekControversy}
+Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. Logica Universalis. Springer. 
 % @article{AndersonHajekControversy,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2015)}]{WorldCongressLogicAndReligion}
+Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2015. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. 1st World Congress on Logic and Religion. 
 % @inproceedings{WorldCongressLogicAndReligion,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{WADT}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape On Logic Embeddings and G\"odel's God}. Recent Trends in Algebraic Development Techniques. 
 % @inproceeding{WADT,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{SpiritualQuest}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015.
+{\itshape Experiments in Computational Metaphysics: G\"odel's Proof of God's Existence}.
+Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November).
+vol.
+9.
+pp.
+23--40.
 % @inproceedings{SpiritualQuest,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{ReasoningWebSummerSchool}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015.
+{\itshape Higher-Order Modal Logics: Automation and Applications}.
+Reasoning Web.
+Web Logic Rules - 11th International Summer School 2015, Berlin, Germany, July 31 - August 4, 2015, Tutorial Lectures.
+pp.
+32--74.
 % @incollection{ReasoningWebSummerSchool,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{CSRModalLogicInCoq}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015.
+{\itshape Interacting with Modal Logics in the Coq Proof Assistant}.
+Computer Science - Theory and Applications - 10th International Computer Science Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July   13-17, 2015, Proceedings.
+pp.
+398--411.
 % @inproceedings{CSRModalLogicInCoq,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{ECAI2014}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2014.
+{\itshape Automating G{\"{o}}del's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers}.
+ECAI 2014 - 21st European Conference on Artificial Intelligence, 18-22 August 2014, Prague, Czech Republic - Including Prestigious Applications of Intelligent Systems (PAIS 2014), pp.
+93--98.
 % @inproceedings{ECAI2014,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{SquareOfOpposition}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape Formalization and Automated Verification of G\"odel's Proof of God's Existence}. 4th World Congress on the Square of Opposition, pp. 24--25. 
 % @inproceedings{SquareOfOpposition,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{WADTPreProceedings}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape On Logic Embeddings and G\"odel's God}. 22nd International Workshop on Algebraic Development Techniques, pp. 8--9. 
 % @inworkshopinvited{WADTPreProceedings,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016)}]{ModalCollapse}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Springer. pp. 1--7. 
 % @article{ModalCollapse,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013)}]{IsabelleGod}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape G{\"{o}}del's God in Isabelle/HOL}. Archive of Formal Proofs. 
 % @article{IsabelleGod,
@@
 
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013)}]{IWIL}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape G{\"{o}}del's God on the Computer}. 10th International Workshop on the Implementation of Logics. pp. 1--2.
 % @inworkshopinvited{IWIL,
@@
 
+\bibitem[{\itshape Woltzenlogel Paleo(2013)}]{AustrianComputerSociety}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape Automated Verification and Reconstruction of G\"odel's Proof of God's Existence}. Journal of the Austrian Computer Society. pp. 4--6.
 % @article{AustrianComputerSociety,

### 2016-09-11 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Citations fixed.

@@
 For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz} or \citep{gabbay}.
-The natural deduction calculus proposed and used in this paper is an extension of a standard natural deduction calculus for (non-modal) higher-order logic \citep{BezmuellerBrownKohlhase} with new introduction and elimination rules for modal operators (as presented in Section \ref{sec:calculus}).
+The natural deduction calculus proposed and used in this paper is an extension of a standard natural deduction calculus for (non-modal) higher-order logic \citep{BenzmuellerBrownKohlhase} with new introduction and elimination rules for modal operators (as presented in Section \ref{sec:calculus}).
 This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}).
-There are various higher-order logics \citep{BezmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}.
+There are various higher-order logics \citep{BenzmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}.
 The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles.
@@
 
-%\bibitem[{\itshape Negri(2005)}]{negri}
-%Negri, S. 2005. {\itshape Proof analysis in modal logic}. J. Philosophical Logic, vol. 34, pp. 507--544.
+\bibitem[{\itshape Muskens(2006)}]{Muskens}
+Muskens, R.
+2006.
+{\itshape Higher Order Modal Logic}.
+In P.
+Blackburn, J.F.A.K.
+van Benthem, and F.
+Wolter, editors, Handbook of Modal Logic, Studies in Logic and Practical Reasoning, pages 621-653.
+Elsevier, Dordrecht, 2006.
 

### 2016-09-12 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
citations and references are working

@@
 Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
-Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicCoq}.
+Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicInCoq}.
 And finally, in comparison to other kinds of proof systems (e.g.
@@
 
-An early unpublished draft of Section \ref{newproof} of this paper,
+An early unpublished draft of Section \ref{sec:newproof} of this paper,
 prepared by Woltzenlogel Paleo in March 2013, served as the starting
@@
 \citep{WorldCongressLogicAndReligion} and published in
-\citep{AndersonHajekControversy} and results related to the modal
+\citep{controversy} and results related to the modal
 collapse were reported in the 4th Word Congress on the Square of
@@
 
-\bibitem[{\itshape Benzm\"uller \& Brown \& Kohlhase(2004)}]{BenzmuellerBrownKohlhase}
-Benzm\"uller, C. \& Brown C. \& Kohlhase, M. In: {\itshape Journal of Symbolic Logic}, Vol. 69, N. 4, pp. 1027--1088.
-
-\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{controversy}
-Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}. In: {\itshape Logica Universalis} 10 (2016)
-
-\bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
-Bj{\o}rdal, F. 1999. Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
-
-\bibitem[{\itshape Blackburn \& al.(2001)}]{blackburn}
-Blackburn, P. \&  M. de Rijke \& Y. Venema. 2001. {\itshape Modal Logic}. Cambridge: Cambridge University Press.
-
-\bibitem[{\itshape Church(1940)}]{church}
-Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56--68.
-
-\bibitem[{\itshape Fitting \& Mendelsohn(1998)}]{fitting2}
-Fitting, M. \& Mendelsohn, R. 1998.  {\itshape First-order Modal Logic}, Synthese Library vol. 277. Kluwer Academic Publishers.
-
-\bibitem[{\itshape Fitting(2002)}]{fitting}
-Fitting, M. 2002.  {\itshape Types, Tableaus, and G\"odel's God}, Kluwer Academic Publishers.
-
-\bibitem[{\itshape Fuhrmann(2005)}]{fuhrmann}
-Fuhrmann, A. 2005.
-Existenz und Notwendigkeit -- Kurt G\"odel's Axiomatische Theologie, in Olsson e.J., Schr\"oder-Heister, P. and Spohn, W. (eds.) Logik in der Philosophie, pp. 349-374, Synchr.-Wissenschafts-Verlag.
-
-\bibitem[{\itshape Gabbay(1996)}]{gabbay}
-Gabbay, D. 1996. {\itshape Labelled Deductive Systems: volume I}. New York: Oxford Science publications.
-
-\bibitem[{\itshape Goedel(1970)}]{Goedel} 
-G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
-
-\bibitem[{\itshape Hajek(1996)}]{hajek}
-Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
-(eds.) Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
-
-\bibitem[{\itshape Kant(1781)}]{kant}
-Kant, I.  original 1781. {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
-
-\bibitem[{\itshape Koons(2006)}]{koons}
-Koons, R. 2006. Sobel on G\"odel's Ontological Proof. {\itshape Philosophia Christi}, 2:235-248.
-
-\bibitem[{\itshape Magari(1988)}]{magari}
-Magari, R. 1988. {\itshape Logica e Teofilia}. In: {\itshape Notizie di Logica} VII.4.
-
-\bibitem[{\itshape Muskens(2006)}]{Muskens}
-Muskens, R.
-2006.
-{\itshape Higher Order Modal Logic}.
-In P.
-Blackburn, J.F.A.K.
-van Benthem, and F.
-Wolter, editors, Handbook of Modal Logic, Studies in Logic and Practical Reasoning, pages 621-653.
-Elsevier, Dordrecht, 2006.
-
-\bibitem[{\itshape Nipkow \& Paulson \& Wenzel(2002)}]{isabelle} Nipkow, T. \& Paulson, L.C. \& Wenzel, M. 2002. {\itshape Isabelle/HOL: A Proof Assistant for
-Higher-Order Logic}. Number 2283 in LNCS, Springer.
-
-\bibitem[{\itshape Paulin-Mohring(2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
-All about Proofs, Proofs for All,
-Mathematical Logic and Foundations. College Publications, London.
-
-\bibitem[{\itshape Prawitz(2006)}]{prawitz}
-Prawitz, D. 2006 (1st ed. 1965). {\itshape Natural deduction: a proof-theoretical study}. Mineola, New York: Dover publications.
-
-\bibitem[{\itshape Scott(2001)}]{scott}
-Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
-
-\bibitem[{\itshape Sobel(1987)}]{sobel}
-Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying: essays for Richard Cartwright},  MIT Press.
-
-\bibitem[{\itshape Sobel(2001)}]{sobel2}
-Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
-
-\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
-
-\bibitem[{\itshape Wang(1996)}]{Wang1996}
-Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
-
-\bibitem[{\itshape Oppenheimer \& Zalta(2011)}]{ZaltaOppenheimer}
-Oppenheimer, P. \& Zalta, E. 2011. {\itshape A Computationally-Discovered Simplification of the Ontological Argument}. Australasian Journal of Philosophy, 89(2), 333-349.
-
-\bibitem[{\itshape Rushby(2013)}]{Rushby}
-Rushby, J. 2013. {\itshape The Ontological Argument in PVS}. CAV Workshop ``Fun with Formal Methods''. St. Petersburg.
-
-% TODO: transform the following bibtex entries into bibitems
-
 \bibitem[{\itshape Bentert \& Benzm\"uller \& Streit \& Woltzenlogel Paleo(2016)}]{LeibnizOntologicalProof}
@@
 Ria University Press.
+
+\bibitem[{\itshape Benzm\"uller \& Brown \& Kohlhase(2004)}]{BenzmuellerBrownKohlhase}
+Benzm\"uller, C. \& Brown C. \& Kohlhase, M. In: {\itshape Journal of Symbolic Logic}, Vol. 69, N. 4, pp. 1027--1088.
+
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2015)}]{WorldCongressLogicAndReligion}
+Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2015. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. 1st World Congress on Logic and Religion.
+
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{controversy}
+Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}. In: {\itshape Logica Universalis} 10 (2016)
+
+%\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{AndersonHajekControversy}
+%Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. Logica Universalis. Springer.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013a)}]{IWIL}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013a. {\itshape G{\"{o}}del's God on the Computer}. 10th International Workshop on the Implementation of Logics. pp. 1--2.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013b)}]{IsabelleGod}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013b. {\itshape G{\"{o}}del's God in Isabelle/HOL}. Archive of Formal Proofs.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014a)}]{WADT}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014a. {\itshape On Logic Embeddings and G\"odel's God}. Recent Trends in Algebraic Development Techniques.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014b)}]{WADTPreProceedings}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014b. {\itshape On Logic Embeddings and G\"odel's God}. 22nd International Workshop on Algebraic Development Techniques, pp. 8--9.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014c)}]{SquareOfOpposition}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014c. {\itshape Formalization and Automated Verification of G\"odel's Proof of God's Existence}. 4th World Congress on the Square of Opposition, pp. 24--25.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014d)}]{ECAI2014}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2014d.
+{\itshape Automating G{\"{o}}del's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers}.
+ECAI 2014 - 21st European Conference on Artificial Intelligence, 18-22 August 2014, Prague, Czech Republic - Including Prestigious Applications of Intelligent Systems (PAIS 2014), pp.
+93--98.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015a)}]{CSRModalLogicInCoq}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015a.
+{\itshape Interacting with Modal Logics in the Coq Proof Assistant}.
+Computer Science - Theory and Applications - 10th International Computer Science Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July   13-17, 2015, Proceedings.
+pp.
+398--411.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015b)}]{ReasoningWebSummerSchool}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015b.
+{\itshape Higher-Order Modal Logics: Automation and Applications}.
+Reasoning Web.
+Web Logic Rules - 11th International Summer School 2015, Berlin, Germany, July 31 - August 4, 2015, Tutorial Lectures.
+pp.
+32--74.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015c)}]{SpiritualQuest}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2015c.
+{\itshape Experiments in Computational Metaphysics: G\"odel's Proof of God's Existence}.
+Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November).
+vol.
+9.
+pp.
+23--40.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016a)}]{IJCAI2016}
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2016a.
+{\itshape The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics}.
+International Joint Conference on Artificial Intelligence - IJCAI.
+
+\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016b)}]{ModalCollapse}
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016b. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Springer. pp. 1--7.
+
+\bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
+Bj{\o}rdal, F. 1999. Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
+
+\bibitem[{\itshape Blackburn \& al.(2001)}]{blackburn}
+Blackburn, P. \&  M. de Rijke \& Y. Venema. 2001. {\itshape Modal Logic}. Cambridge: Cambridge University Press.
+
+\bibitem[{\itshape Church(1940)}]{church}
+Church, A. 1940. {\itshape A Formulation of the Simple Theory of Types}, Journal of Symbolic Logic, 5: 56--68.
+
+\bibitem[{\itshape Fitting \& Mendelsohn(1998)}]{fitting2}
+Fitting, M. \& Mendelsohn, R. 1998.  {\itshape First-order Modal Logic}, Synthese Library vol. 277. Kluwer Academic Publishers.
+
+\bibitem[{\itshape Fitting(2002)}]{fitting}
+Fitting, M. 2002.  {\itshape Types, Tableaus, and G\"odel's God}, Kluwer Academic Publishers.
+
+\bibitem[{\itshape Fuhrmann(2005)}]{fuhrmann}
+Fuhrmann, A. 2005.
+Existenz und Notwendigkeit -- Kurt G\"odel's Axiomatische Theologie, in Olsson e.J., Schr\"oder-Heister, P. and Spohn, W. (eds.) Logik in der Philosophie, pp. 349-374, Synchr.-Wissenschafts-Verlag.
+
+\bibitem[{\itshape Gabbay(1996)}]{gabbay}
+Gabbay, D. 1996. {\itshape Labelled Deductive Systems: volume I}. New York: Oxford Science publications.
+
+\bibitem[{\itshape G\"odel(1970)}]{Goedel} 
+G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
+
+\bibitem[{\itshape Hajek(1996)}]{hajek}
+Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
+(eds.) Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
+
+\bibitem[{\itshape Kant(1781)}]{kant}
+Kant, I.  original 1781. {\itshape Critique of Pure Reason}, J. M. Dent \& Sons LTD, edition from 1959.
+
+\bibitem[{\itshape Koons(2006)}]{koons}
+Koons, R. 2006. Sobel on G\"odel's Ontological Proof. {\itshape Philosophia Christi}, 2:235-248.
+
+\bibitem[{\itshape Magari(1988)}]{magari}
+Magari, R. 1988. {\itshape Logica e Teofilia}. In: {\itshape Notizie di Logica} VII.4.
+
+\bibitem[{\itshape Muskens(2006)}]{Muskens}
+Muskens, R.
+2006.
+{\itshape Higher Order Modal Logic}.
+In P.
+Blackburn, J.F.A.K.
+van Benthem, and F.
+Wolter, editors, Handbook of Modal Logic, Studies in Logic and Practical Reasoning, pages 621-653.
+Elsevier, Dordrecht, 2006.
+
+\bibitem[{\itshape Nipkow \& Paulson \& Wenzel(2002)}]{isabelle} Nipkow, T. \& Paulson, L.C. \& Wenzel, M. 2002. {\itshape Isabelle/HOL: A Proof Assistant for
+Higher-Order Logic}. Number 2283 in LNCS, Springer.
+
+\bibitem[{\itshape Oppenheimer \& Zalta(2011)}]{ZaltaOppenheimer}
+Oppenheimer, P. \& Zalta, E. 2011. {\itshape A Computationally-Discovered Simplification of the Ontological Argument}. Australasian Journal of Philosophy, 89(2), 333-349.
+
+\bibitem[{\itshape Paulin-Mohring(2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
+All about Proofs, Proofs for All,
+Mathematical Logic and Foundations. College Publications, London.
+
+\bibitem[{\itshape Prawitz(2006)}]{prawitz}
+Prawitz, D. 2006 (1st ed. 1965). {\itshape Natural deduction: a proof-theoretical study}. Mineola, New York: Dover publications.
+
+\bibitem[{\itshape Rushby(2013)}]{Rushby}
+Rushby, J. 2013. {\itshape The Ontological Argument in PVS}. CAV Workshop ``Fun with Formal Methods''. St. Petersburg.
+
+\bibitem[{\itshape Scott(2001)}]{scott}
+Scott, D. {\itshape Appendix B. Notes in Dana Scott's hand} in Logic and Theism: Arguments for and against Beliefs in God by Sobel, J. H.
+
+\bibitem[{\itshape Sobel(1987)}]{sobel}
+Sobel, J. H. 1987. {\itshape G\"odel's Ontological Proof}. In: J. J. Thompson (ed.). {\itshape On being and saying: essays for Richard Cartwright},  MIT Press.
+
+\bibitem[{\itshape Sobel(2001)}]{sobel2}
+Sobel, J. H. 2001. {\itshape Logic and Theism: Arguments for and against Beliefs in God}, Cambridge University Press.
+
+\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
+
+\bibitem[{\itshape Wang(1996)}]{Wang1996}
+Wang, H. 1996. {\itshape A Logical Journey: From G\"odel to Philosophy}, The MIT Press.
+
+\bibitem[{\itshape Woltzenlogel Paleo(2013)}]{AustrianComputerSociety}
+ Woltzenlogel Paleo, B. 2013. {\itshape Automated Verification and Reconstruction of G\"odel's Proof of God's Existence}. Journal of the Austrian Computer Society. pp. 4--6.
+
+% TODO: transform the following bibtex entries into bibitems
+
 % @incollection{LeibnizOntologicalProof,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016)}]{IJCAI2016}
-Benzm\"uller, C.
-\& Woltzenlogel Paleo, B.
-2016.
-{\itshape The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics}.
-International Joint Conference on Artificial Intelligence - IJCAI.
 % @inproceedings{IJCAI2016,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{AndersonHajekControversy}
-Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. Logica Universalis. Springer. 
 % @article{AndersonHajekControversy,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2015)}]{WorldCongressLogicAndReligion}
-Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2015. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. 1st World Congress on Logic and Religion. 
 % @inproceedings{WorldCongressLogicAndReligion,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{WADT}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape On Logic Embeddings and G\"odel's God}. Recent Trends in Algebraic Development Techniques. 
 % @inproceeding{WADT,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{SpiritualQuest}
-Benzm\"uller, C.
-\& Woltzenlogel Paleo, B.
-2015.
-{\itshape Experiments in Computational Metaphysics: G\"odel's Proof of God's Existence}.
-Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November).
-vol.
-9.
-pp.
-23--40.
 % @inproceedings{SpiritualQuest,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{ReasoningWebSummerSchool}
-Benzm\"uller, C.
-\& Woltzenlogel Paleo, B.
-2015.
-{\itshape Higher-Order Modal Logics: Automation and Applications}.
-Reasoning Web.
-Web Logic Rules - 11th International Summer School 2015, Berlin, Germany, July 31 - August 4, 2015, Tutorial Lectures.
-pp.
-32--74.
 % @incollection{ReasoningWebSummerSchool,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2015)}]{CSRModalLogicInCoq}
-Benzm\"uller, C.
-\& Woltzenlogel Paleo, B.
-2015.
-{\itshape Interacting with Modal Logics in the Coq Proof Assistant}.
-Computer Science - Theory and Applications - 10th International Computer Science Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July   13-17, 2015, Proceedings.
-pp.
-398--411.
 % @inproceedings{CSRModalLogicInCoq,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{ECAI2014}
-Benzm\"uller, C.
-\& Woltzenlogel Paleo, B.
-2014.
-{\itshape Automating G{\"{o}}del's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers}.
-ECAI 2014 - 21st European Conference on Artificial Intelligence, 18-22 August 2014, Prague, Czech Republic - Including Prestigious Applications of Intelligent Systems (PAIS 2014), pp.
-93--98.
 % @inproceedings{ECAI2014,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{SquareOfOpposition}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape Formalization and Automated Verification of G\"odel's Proof of God's Existence}. 4th World Congress on the Square of Opposition, pp. 24--25. 
 % @inproceedings{SquareOfOpposition,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014)}]{WADTPreProceedings}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014. {\itshape On Logic Embeddings and G\"odel's God}. 22nd International Workshop on Algebraic Development Techniques, pp. 8--9. 
 % @inworkshopinvited{WADTPreProceedings,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016)}]{ModalCollapse}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Springer. pp. 1--7. 
 % @article{ModalCollapse,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013)}]{IsabelleGod}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape G{\"{o}}del's God in Isabelle/HOL}. Archive of Formal Proofs. 
 % @article{IsabelleGod,
@@
 
-\bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013)}]{IWIL}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape G{\"{o}}del's God on the Computer}. 10th International Workshop on the Implementation of Logics. pp. 1--2.
 % @inworkshopinvited{IWIL,
@@
 
-\bibitem[{\itshape Woltzenlogel Paleo(2013)}]{AustrianComputerSociety}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013. {\itshape Automated Verification and Reconstruction of G\"odel's Proof of God's Existence}. Journal of the Austrian Computer Society. pp. 4--6.
 % @article{AustrianComputerSociety,

### 2016-09-16 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Make several minor improvements in language and proofs

@@
 The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills in the gaps, thus verifying its correctness.
-The second formal proof improves the first one, by relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5}; and by avoiding the introduction of the equality predicate.
+The second formal proof improves the first one, by relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5}; and by avoiding the equality relation.
 The second proof is also technically shorter than the first one, because it eliminates unnecessary detours and uses Axiom 1 for the positivity of properties only once.
@@
 In this setting, he gave precise axioms describing the notion of \emph{positive} property and defined God as a being that has all positive properties.
-G\"odel's work was saved in his own notes \citep{Goedel} as well as in notes by Scott \citep{scott}, in whom he confided his proof.
+G\"odel's work was saved in his own notes \citep{Goedel} as well as in notes by Scott \citep{scott}, to whom he confided his proof.
 
@@
 This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}).
-There are various higher-order logics \citep{BenzmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}.
+There are many higher-order logics \citep{BenzmuellerBrownKohlhase} and higher-order modal logics \citep{Muskens}.
 The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles.
@@
 Therefore, a formalization in natural deduction is a valuable complement to the existing presentations.
-Secondly, it makes the ontological proof accessible to people who are familiarized with a natural deduction style.
+Secondly, it makes the ontological proof accessible to people who are familiarized with the natural deduction style.
 Thirdly, as natural deduction is the style used by proof assistants such as Coq \citep{coq} and Isabelle \citep{isabelle}, natural deduction formalizations can be verified step-by-step in such proof assistants, and we have in fact done this \citep{CSRModalLogicInCoq}.
@@
 In contrast to Scott's proof \citep{scott}, which requires the modal logic \textbf{S5}, the new proof requires only the weaker modal logic \textbf{KB}.
-The new proof also does not rely on the equality predicate and is much shorter.
-Although there have been claims (as discussed in more detail in Section \ref{sec:newproof}) that \textbf{KB} is sufficent and the equality predicate is superfluous, formal proofs substantiating these claims were missing up to now.
-
-A major criticism against G\"odel's argument is an undesirable consequence of the stipulated axioms, called \emph{modal collapse}.
+The new proof also does not rely on the equality relation and is much shorter.
+Although there have been claims (as discussed in more detail in Section \ref{sec:newproof}) about the sufficiency of \textbf{KB} and the superfluousness of the equality relation, formal proofs substantiating these claims were missing up to now.
+
+A major criticism against G\"odel's argument is an undesirable consequence of the stipulated axioms called \emph{modal collapse}.
 This is discussed in greater detail in Section \ref{sec:collapse}, where a natural deduction derivation of the collapse is presented.
@@
 In this paper we present natural deduction formalizations of two of these solutions: Anderson's emendation (in Section \ref{sec:anderson}) and Bj{\o}rdal's variant (in Section \ref{sec:bjordal}).
+
+\subsection{Related Work}
 
@@
 \citep{IJCAI2016}. The automatic verification of variants by Anderson,
-Bjordal and H\'ajek, on the other hand were presented in the First
+Bjordal and H\'ajek were presented in the First
 World Congress on Logic and Religion
@@
 variant within the interactive proof assistants Isabelle and Coq were
-relesed, respectively, in Isabellle's Archive of Formal Proofs
+released, respectively, in Isabellle's Archive of Formal Proofs
 \citep{IsabelleGod} and in the Symposium of Computer Science in Russia
@@
 logics into higher-order logics and the use of higher-order automated
-reasoners, Anselm's argument greater simplicity enabled it to be
-verified with a first-order prover. Anselm's argument has also been
+reasoners, Anselm's argument greater simplicity enabled its
+verification with a first-order prover. Anselm's argument has also been
 formalized within the interactive proof assistant PVS \citep{Rushby}
-and there has been a so far failed attempt to formalize Avicenna's
+and there has been a failed attempt to formalize Avicenna's
 argument in PVS as well.
@@
 
-In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas changes due to a substitution).
+In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas differ due to substitution).
 For any expression occurring both in a formula in the premise and in its corresponding (sub-)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
@@
 Applications of the $\beta$ rule are silently ommitted.
-Double lines are used to abbreviate tedious propositional reasoning steps in the derivations where the reasoning is otherwise clear.
+Double lines are used to abbreviate propositional reasoning steps when they are tedious and too space-consuming.
 Dashed lines are used to refer to an axiom or theorem with the proof shown elsewhere.
@@
 
-The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND.
+The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical}, \ref{fig:QuantifierRules} and \ref{fig:ReductionRules} is named \ND.
 This calculus is essentially the same as the simplest natural deduction calculus in the family of calculi discussed in \citep{BenzmuellerBrownKohlhase}.
@@
 \label{theorem:completeness}
-{\NDK} is complete, relative to {\ND} extended with axiom K and the necessitation rule.
+{\NDK} is complete relative to {\ND} extended with axiom K and necessitation.
 \end{theorem}
@@
 \label{theorem:soundness}
-{\NDK} is sound, relative to {\ND} extended with axiom K and the necessitation rule.
+{\NDK} is sound relative to {\ND} extended with axiom K and necessitation.
 \end{theorem}
@@
 \citep{hajek}).
-Nevertheless, some authors of variants \citep[footnotes 11 and 14]{and90} of G\"odel's ontological argument do explicitly state a preference for varying domains.
-Our choice of constant domains is motivated by simplicity.
+Nevertheless, some authors of variants of G\"odel's ontological argument do explicitly state a preference for varying domains.
+Our choice of constant domains is motivated by simplicity \citep[footnotes 11 and 14]{and90}.
 As is well-known, varying domain quantifiers (also known as \emph{actualistic} quantifiers) can be simulated by constant domain quantifiers (also known as \emph{possibilistic} quantifiers) guarded by a primitive \emph{existence} predicate\footnote{Formally, $\forall^e x.
@@
 F[x] \equiv \exists x.
-E(x) \wedge F[x]$}; therefore, the formalization of ontological arguments with varying domain quantification would be feasible in the natural deduction calculus presented here as well.
-
-\section{Some Useful Derivable Modal Principles}
+E(x) \wedge F[x]$.}; therefore, the formalization of ontological arguments with varying domain quantification would be feasible in the natural deduction calculus presented here as well.
+
+\section{Some Useful Derivable Principles}
 \label{sec:ModalLogics}
 
-From an axiomatic point of view, modal logics differ with respect to which additional axiom schemas they admit.
+From an axiomatic point of view, modal logics stronger than {\bf K} differ from each other with respect to which axiom schemas they admit in addition to K (and necessitation).
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
 From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2} or \citep[Section 3.1]{blackburn} on Frame definability): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
-In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, K and B are satisfied; and in the models of logic {\bf B}, K, T and B are satisfied.
+In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, B is satisfied; and in the models of logic {\bf B}, T and B are satisfied.
 
@@
 
+The \emph{reflexivity} of equality is a theorem in higher-order logics where equality is defined as Leibniz's equality. A natural deduction proof can be easily constructed, as shown below.
+
 \begin{lemma*}[Reflexivity]
 $$
-s=s
+\forall x. x = x
 $$
@@
 \AXC{}\RightLabel{$1$}
-\UIC{$P s$}\RightLabel{$\imp_I^1$}
-\UIC{$P s\imp P s$}
+\UIC{$P(\alpha)$}\RightLabel{$\imp_I^1$}
+\UIC{$P(\alpha) \imp P(\alpha)$}
 \AXC{}\RightLabel{$2$}
-\UIC{$P s$}\RightLabel{$\imp_I^2$}
-\UIC{$P s\imp P s$}\RightLabel{$\wedge_I$}
-\BIC{$P s\biimp P s$}\RightLabel{$\forall_I$}
-\UIC{$\forall P. P s\biimp P s$}\dottedLine
-\UIC{$s=s$}
+\UIC{$P(\alpha)$}\RightLabel{$\imp_I^2$}
+\UIC{$P(\alpha) \imp P(\alpha)$}\RightLabel{$\wedge_I$}
+\UIC{$(P(\alpha) \imp P(\alpha)) \wedge (P(\alpha) \imp P(\alpha))$} \dottedLine
+\BIC{$P(\alpha) \biimp P(\alpha)$}\RightLabel{$\forall_I$}
+\UIC{$\forall P. P(\alpha) \biimp P(\alpha)$}\dottedLine
+\UIC{$\alpha=\alpha$} \RightLabel{$\forall_I$}
+\UIC{$\forall x. x=x$} 
 \end{prooftree}
@@
 In this section we present a detailed formalization of Scott's proof in the natural deduction calculus defined in Section \ref{sec:calculus}.
-The reasoning in Scott's manuscript has been reproduced step-by-step and all reasoning gaps are filled in by using the deduction considered most natural.
+The reasoning in Scott's manuscript has been reproduced step-by-step and reasoning gaps are completed by using the deduction steps considered most natural.
 All derived formulas that do appear in Scott's manuscript are marked with a $\dagger$ here.
-Unmarked formulas were derived in the process of filling the gaps between the marked formulas mentioned in Scott's manuscript.
+Unmarked formulas were derived in the process of filling the gaps between the marked formulas.
 
@@
 		\AXC{Reflexivity} \dashedLine
+    \UIC{$\omega: \forall x. x = x$} \RightLabel{$\forall_E$}
 		\UIC{$\omega: \gamma = \gamma$} \RightLabel{$\imp_I$}
@@
 
-Comparing this new proof of Theorem 1 with the proof of Theorem 1 indicated by G\"odel and Scott, it is noticeable that new proof is significantly shorter.
+Comparing this new proof of Theorem 1 with the proof of Theorem 1 indicated by G\"odel and Scott, it is noticeable that the new proof is significantly shorter.
 The key idea to obtain a shorter proof is to instantiate the second universally quantified property of Axiom 2 by $\lambda x.
-\neg \rho(x)$ instead of using the predicate symbols $=$ or $\neq$ as G\"odel and Scott did.
+\neg \rho(x)$ instead of using the relation symbols $=$ or $\neq$ as G\"odel and Scott did.
 This eliminates the reliance of the ontological argument on the presence of $=$ and $\neq$ in the logical language.
@@
 \end{corollary}
-\begin{proof} The proof of this corollary is identical to the proof of the same corollary shown in Section \ref{sec:scott}.
+\begin{proof} The proof of this corollary is identical to the proof shown in Section \ref{sec:scott}.
 % \begin{prooftree}
@@
 
-This phenomenon is very intriguing: the new proof of Lemma 1 is technically simpler (because it has fewer inferences), but Scott's proof of Lemma 1 with Theorem 2 can be considered intuitively simpler (because it is arguably easier to understand).
+This phenomenon is intriguing: the new proof of Lemma 1 is technically simpler (because it has fewer inferences), but Scott's proof of Lemma 1 with Theorem 2 can be considered intuitively simpler (because it is easier to understand).
 This constitutes an interesting case for Hilbert's 24th Problem \citep{hilbert}, which asks for criteria to properly compare the simplicity of proofs.
@@
 G(x)$).
-This contrasts with G\"odel's proof, where the actual existence is never actually proven.
-It could, nevertheless, be derived as a trivial corollary by applying the modal axiom T.
+This contrasts with G\"odel's proof, where the actual existence is never actually proven (although it could, nevertheless, be derived as a trivial corollary by applying the modal axiom T).
 

### 2016-09-16 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Make a few minor improvements

@@
 H\'ajek \citep{hajek} claimed that Anderson's emendation makes Axioms 4 and 5 derivable from the other axioms and definitions.
-Discussions about the redundancy of these two axioms started with an earlier (and wrong claim) claim by Magari \citep{magari} w.r.t.
-Scott's variant, and the redundancy was the topic of a long controversy \citep{controversy} between H\'ajek and Anderson.
+Discussions about the redundancy of these two axioms started with an earlier (and wrong) claim by Magari \citep{magari} w.r.t.
+Scott's variant, and the redundancy was the topic of a long controversy between H\'ajek and Anderson \citep{controversy}.
 As confirmed in Subsection \ref{sec:NonRedundantAnderson}, H\'ajek's claim is correct.
@@
 In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify
-the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. Such a modified proof is shown below.
+the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. The modified proof is shown below.
 
@@
 \UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
-\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
 \UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\wedge_E$}
@@
 \UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
-\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$} \doubleLine  \RightLabel{(Axiom K)}
+\UIC{$\nec (G(\gamma)\imp \ex x.G(x))$} \doubleLine  \RightLabel{(Axiom K)}
 \UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
@@
 
-The proof of section \ref{sec:newproof} shows explicitly that the weak modal system {\bf KB} is sufficient for the ontological argument as opposed to the standard modal calculus {\bf S5}.
-This weakening of the system is based on an explicit formalization and confirms the result briefly described by Anderson \citep[footnote 2]{and90}.
-The weakening also applies to the emendation and variant of Anderson and Bj\o rdal respectively.
-The proofs are short and accessible to the general reader because of the choice of a natural deduction calculus.
+The formal proofs presented here are short and accessible because of the choice of a natural deduction calculus.
+The new proof of section \ref{sec:newproof} shows explicitly that the weaker modal logic {\bf KB} is sufficient for the ontological argument.
+The stronger modal logic {\bf S5}, which is usually taken for granted in discussions of G\"odel's ontological argument, is not necessary.
+This weakening of the underlying logic confirms a claim briefly mentioned by Anderson \citep[footnote 2]{and90}.
+{\bf KB} is also sufficient for the variants of Anderson and Bj\o rdal, which address Sobel's criticism of the modal collapse.
 

### 2016-09-16 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Replace biimp by equiv in definitions

@@
 $$
-G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
+G(x) \equiv \forall \varphi. [P(\varphi) \to \varphi(x)]
 $$
@@
 $$
-\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
+\ess{\varphi}{x} \equiv \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
 $$
@@
 $$
-E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
+E(x) \equiv \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
 $$
@@
 $$
-G(x) \biimp \forall \varphi. [P(\varphi) \biimp \nec \varphi(x)]
+G(x) \equiv \forall \varphi. [P(\varphi) \biimp \nec \varphi(x)]
 $$
@@
 $$
-\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\nec \psi(x) \biimp \nec \all x. (\varphi(x) \imp \psi(x)))
+\ess{\varphi}{x} \equiv \varphi(x) \wedge \all \psi. (\nec \psi(x) \biimp \nec \all x. (\varphi(x) \imp \psi(x)))
 $$

### 2016-09-16 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Replace brackets by parentheses

@@
 $$
-\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
+\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))
 $$
@@
 $$
-\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
+\all \varphi. \all \psi.((P(\varphi) \wedge \nec \all x.(\varphi(x) \imp \psi(x))) \imp P(\psi))
 $$
@@
 $$
-\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
+\all \varphi. (P(\varphi) \imp \pos \ex x.\varphi(x))
 $$
@@
   		\UIC{$ P(\rho) \imp \pos \ex x.\rho(x)$} \RightLabel{$\all_I$}
-      \UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
+      \UIC{$ \all \varphi. (P(\varphi) \imp \pos \ex x.\varphi(x)) $}
 \end{prooftree}
@@
 $$
-G(x) \equiv \forall \varphi. [P(\varphi) \to \varphi(x)]
+G(x) \equiv \forall \varphi. (P(\varphi) \to \varphi(x))
 $$
@@
 $$
-\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
+\all \varphi.(P(\varphi) \to \Box \; P(\varphi))
 $$
@@
 				\AXC{Axiom 1} \dashedLine
-        \UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
+        \UIC{$\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))$} \RightLabel{$\all_E$}
         \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
@@
 	      \AXC{Axiom 4} \dashedLine
-        \UIC{$\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]$} \RightLabel{$\all_E$}
+        \UIC{$\all \varphi.(P(\varphi) \to \Box \; P(\varphi))$} \RightLabel{$\all_E$}
 	   	  \UIC{$P(\rho) \imp \nec P(\rho)$} \RightLabel{$\imp_E$}
@@
 $$
-E(x) \equiv \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
+E(x) \equiv \all \varphi.(\ess{\varphi}{x} \imp \nec \ex y.\varphi(y))
 $$
@@
 \UIC{$ \nec \all x.(\rho(x) \imp \neg \rho(x)) $}\RightLabel{$\wedge_I$}
-\BIC{$ P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]$}
+\BIC{$ P(\rho) \wedge \nec \all x.(\rho(x) \imp \neg \rho(x))$}
 \AXC{Axiom \ref{A2} for $\rho$ and $\lambda x. \neg \rho(x)$} \dashedLine
-\UIC{$(P(\rho) \wedge \nec \all x.[\rho(x) \imp \neg \rho(x)]) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\UIC{$(P(\rho) \wedge \nec \all x.(\rho(x) \imp \neg \rho(x))) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
 \BIC{$[\subproof_5]\qquad P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
@@
 \AXC{Axiom 1} \dashedLine
-\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
+\UIC{$\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))$} \RightLabel{$\all_E$}
 \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
@@
 \UIC{$ P(\rho) \imp \pos \ex x.\rho(x) $} \RightLabel{$\all_I$}
-\UIC{$ \all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)] $}
+\UIC{$ \all \varphi. (P(\varphi) \imp \pos \ex x.\varphi(x)) $}
 
@@
 \AXC{Axiom \ref{N:A4}} \dashedLine
-\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
+\UIC{$\all \varphi.(P(\varphi) \imp \nec\; P(\varphi))$}\RightLabel{$\forall_E$}
 \UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
@@
 \BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{N:D3}}
-\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
+\UIC{$ \all \varphi.(\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)) $}\RightLabel{$\all_E$}
 \UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
@@
 \AXC{ Distribution Principle} \dashedLine
-\UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
+\UIC{$\nec (\ex x. G(x)\imp \nec \ex x. G(x)) \imp (\pos \ex x. G(x)\imp \pos\nec \ex x. G(x))$}
 \AXC{ Lemma \ref{N:L1} } \dashedLine
 \UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
-\UIC{$\nec[\ex x. G(x)\imp \nec \ex x. G(x)]$}\RightLabel{$\imp_E$}
+\UIC{$\nec(\ex x. G(x)\imp \nec \ex x. G(x))$}\RightLabel{$\imp_E$}
 \BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
@@
 \begin{prooftree}
-\AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
+\AXC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
 \AXC{ Corollary \ref{C1} } \dashedLine
@@
 \AXC{Theorem 2}\dashedLine
-\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\dottedLine
-\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$}\doubleLine
-\UIC{$\all y.[G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
-%\UIC{$\all y.[G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
-\UIC{$\all y.[G(y) \imp (A \imp \nec \all x. (G(x) \imp A))]$} \doubleLine
+\UIC{$\all y.(G(y) \imp \ess{G}{y})$} \RightLabel{D2 }\dottedLine
+\UIC{$\all y.(G(y) \imp G(y) \wedge \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x))))$}\doubleLine
+\UIC{$\all y.(G(y) \imp \all \psi. (\psi(y) \imp \nec \all x. (G(x) \imp \psi(x))))$} \doubleLine
+%\UIC{$\all y.(G(y) \imp (A(y) \imp \nec \all x. (G(x) \imp A(x))))$} \RightLabel{$A$ is constant}
+\UIC{$\all y.(G(y) \imp (A \imp \nec \all x. (G(x) \imp A)))$} \doubleLine
 \UIC{$[\subproof_8] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
@@
 $$
-\all \varphi. [P(\neg \varphi) \imp \neg P(\varphi)]
+\all \varphi. (P(\neg \varphi) \imp \neg P(\varphi))
 $$
@@
 $$
-G(x) \equiv \forall \varphi. [P(\varphi) \biimp \nec \varphi(x)]
+G(x) \equiv \forall \varphi. (P(\varphi) \biimp \nec \varphi(x))
 $$
@@
 \UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
-\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
 \UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}\RightLabel{$\wedge_E$}
@@
 \UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{A:D1}}
-\UIC{$\omega: \forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
+\UIC{$\omega: \forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
 \UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\wedge_E$}
@@
 \AXC{$ $} \RightLabel{5}
-\UIC{$\nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\wedge_I$}
-\BIC{$P(G)\wedge \nec  \forall y. [G(y)\imp \varphi(y)]$}\RightLabel{$\imp_E$}
+\UIC{$\nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\wedge_I$}
+\BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\imp_E$}
 \AXC{Axiom \ref{A:A2}} 
@@
 \UIC{$  G(\gamma)$}  \RightLabel{D\ref{A:D1}}
-\UIC{$ \all \varphi. [P(\varphi)\biimp \nec \varphi(\gamma)]$}  \RightLabel{$\all_E$}
+\UIC{$ \all \varphi. (P(\varphi)\biimp \nec \varphi(\gamma))$}  \RightLabel{$\all_E$}
 \UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\wedge_E$}
@@
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
-\UIC{$[\subproof_{12}] \quad \nec  \forall y. [G(y)\imp \varphi(y)]\imp \nec \varphi(\gamma)$}
+\UIC{$[\subproof_{12}] \quad \nec  \forall y. (G(y)\imp \varphi(y))\imp \nec \varphi(\gamma)$}
 \end{prooftree}
@@
 \AXC{$\subproof_{11}$ and $\subproof_{12}$} \dashedLine \RightLabel{$\wedge_I$}
-\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)] $} \RightLabel{$\forall_I$}
-\UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. [G(y)\imp \varphi(y)]) $}
+\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
+\UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y))) $}
 \AXC{$ \subproof_{9}$} \dashedLine
@@
 \UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
-\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
+\UIC{$ \all \varphi.(\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)) $}\RightLabel{$\all_E$}
 \UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
@@
 \AXC{Theorem 2}\dashedLine
-\UIC{$\all y.[G(y) \imp \ess{G}{y}]$} \RightLabel{D2 }\doubleLine
-\UIC{$\all y.[G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$}
-\UIC{$\all y.[G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x)))]$} \doubleLine
-%\UIC{$\all y.[G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x)))]$} \RightLabel{$A$ is constant}
-\UIC{$\all y.[G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))]$} \doubleLine
+\UIC{$\all y.(G(y) \imp \ess{G}{y})$} \RightLabel{D2 }\doubleLine
+\UIC{$\all y.(G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x))))$}
+\UIC{$\all y.(G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x))))$} \doubleLine
+%\UIC{$\all y.(G(y) \imp (\nec A(y) \biimp \nec \all x. (G(x) \imp A(x))))$} \RightLabel{$A$ is constant}
+\UIC{$\all y.(G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A)))$} \doubleLine
 \UIC{$[\subproof_{13}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
@@
 $$
-P_B(\varphi) \equiv \nec \all x. [G_B(x) \imp \varphi(x)]
+P_B(\varphi) \equiv \nec \all x. (G_B(x) \imp \varphi(x))
 $$
@@
 $$
-\all \varphi. [P_B (\varphi) \imp \neg P_B (\neg\varphi)]
+\all \varphi. (P_B (\varphi) \imp \neg P_B (\neg\varphi))
 $$
@@
 \AXC{Axiom 1} \dashedLine
-\UIC{$\all \varphi. [P(\neg G_B) \imp \neg P(G_B)]$} \RightLabel{$\all_E$}
+\UIC{$\all \varphi. (P(\neg G_B) \imp \neg P(G_B))$} \RightLabel{$\all_E$}
 \UIC{$ P( \neg G_B(x)) \imp \neg P(G_B) $}  \RightLabel{$\imp_E$}
@@
 \AXC{$ $}\doubleLine
-\UIC{$\nec \all x. [G_B(x) \imp G_B(x)] $}\RightLabel{ D\ref{B:D1}}\dottedLine
+\UIC{$\nec \all x. (G_B(x) \imp G_B(x)) $}\RightLabel{ D\ref{B:D1}}\dottedLine
 \UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
@@
 \UIC{$P_B (NE_B)$} \RightLabel{ D\ref{B:D1}}\dottedLine
-\UIC{$\nec \all x. [G_B(x) \imp NE_B(x)]$}\RightLabel{ D\ref{B:D3}}\dottedLine
-\UIC{$\nec \all x. [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\nec_E$}
-\UIC{$\all x. [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}  \RightLabel{ $\forall_E$}
-\UIC{$ [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y))]$}
+\UIC{$\nec \all x. (G_B(x) \imp NE_B(x))$}\RightLabel{ D\ref{B:D3}}\dottedLine
+\UIC{$\nec \all x. (G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}  \RightLabel{ $\nec_E$}
+\UIC{$\all x. (G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}  \RightLabel{ $\forall_E$}
+\UIC{$ [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}
 \AXC{$ $} \RightLabel{ $2$}
@@
 This weakening of the underlying logic confirms a claim briefly mentioned by Anderson \citep[footnote 2]{and90}.
-{\bf KB} is also sufficient for the variants of Anderson and Bj\o rdal, which address Sobel's criticism of the modal collapse.
+{\bf KB} is also sufficient for the variants of Anderson and Bj{\o}rdal, which address Sobel's criticism of the modal collapse.
 

### 2016-09-30 Papers/2014/WADT/BenzmuellerWoltzenlogelPaleo/paper.tex

--- 2015-04-05 Papers/2014/WADT/WADT-4pages.tex
@@
 \newblock In 
-  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
+  {\em ECAI 2014}, volume 263 of {\em Frontiers in AI and
   Applications}, pages 163 -- 168. IOS Press, 2014.
@@
 and Religion, Jo\~ao Pessoa, Brazil, 2015.
+
+\bibitem{CSR}
+C.~Benzm{\"u}ller and B. Woltzenlogel Paleo.
+\newblock Interacting with Modal Logics in the Coq Proof Assistant,
+{\em 10th Intl. Computer Science Symp. in Russia (CSR)}, 2015.
 
@@
 
-\bibitem{CSR}
-C.~Benzm{\"u}ller and B. Woltzenlogel Paleo.
-\newblock Interacting with Modal Logics in the Coq Proof Assistant, {\em 10th International Computer Science Symposium in Russia (CSR)}, 2015.
-
 \end{thebibliography}

### 2016-09-30 Papers/2016/IJCAI/ijcai16.tex

--- 2015-12-11 Papers/InProgress/IJCAI/ijcai15.tex
@@
-%%%% ijcai15.tex
+%%%% ijcai16.tex
 
-\typeout{IJCAI-15 Instructions for Authors}
+\typeout{IJCAI-16 Instructions for Authors}
 
-% These are the instructions for authors for IJCAI-15.
+% These are the instructions for authors for IJCAI-16.
 % They are the same as the ones for IJCAI-11 with superficical wording
@@
 \documentclass{article}
-% The file ijcai15.sty is the style file for IJCAI-15 (same as ijcai07.sty).
-\usepackage{ijcai15}
+% The file ijcai16.sty is the style file for IJCAI-16 (same as ijcai07.sty).
+\usepackage{ijcai16}
 
@@
 \usepackage{times}
+%The following must be included in the document preamble (it cannot be included in a script), 
+%without ANY LaTeX coding — even an accent:
+\pdfinfo{
+/Title ()
+/Author () }
+%Here's an example of properly completed pdfinfo (note that there is no accent mark in 
+%Bessière, there are no LaTex codes, and the names are separated only by commas):
+%\pdfinfo{
+%/Title (Arc-Consistency and Arc-Consistency Again)
+%/Author (Christian Bessiere, Marie-Odile Cordier) }
 
@@
 
-\title{IJCAI--15 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
-\author{Qiang Yang \\
-Hong Kong University of Science and Technology\\
-Hong Kong, China \\
-pcchair15@ijcai.org}
+\title{IJCAI--16 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
+\author{Subbarao Kambhampati \\ 
+Arizona State University, Tempe Arizona  \\
+ijcai16programchair@gmail.com}
 
@@
 \begin{abstract}
-  The {\it IJCAI--15 Proceedings} will be printed from electronic
+  The {\it IJCAI--16 Proceedings} will be printed from electronic
   manuscripts submitted by the authors. The electronic manuscript will
@@
 
-The {\it IJCAI--15 Proceedings} will be printed from electronic
+The {\it IJCAI--16 Proceedings} will be printed from electronic
 manuscripts submitted by the authors. These must be PDF ({\em Portable
@@
 
-The \LaTeX{} and Word style files are available on the IJCAI--15
-website, {\tt http://www.ijcai-15.org/}.
+The \LaTeX{} and Word style files are available on the IJCAI--16
+website, {\tt http://www.ijcai-16.org/}.
 These style files implement the formatting instructions in this
@@
 
-The \LaTeX{} files are {\tt ijcai15.sty} and {\tt ijcai15.tex}, and
-the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai15.bib}. The
+The \LaTeX{} files are {\tt ijcai16.sty} and {\tt ijcai16.tex}, and
+the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai16.bib}. The
 \LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
 style file is for version 0.99c of Bib\TeX{} ({\em not} version
-0.98i). The {\tt ijcai15.sty} file is the same as the {\tt
+0.98i). The {\tt ijcai16.sty} file is the same as the {\tt
 ijcai07.sty} file used for IJCAI--07.
@@
 The Microsoft Word style file consists of a single file, {\tt
-ijcai15.doc}. This template is the same as the one used for
+ijcai16.doc}. This template is the same as the one used for
 IJCAI--07.
@@
 Further information on using these styles for the preparation of
-papers for IJCAI--15 can be obtained by contacting {\tt
-pcchair15@ijcai.org}.
+papers for IJCAI--16 can be obtained by contacting {\tt
+ijcai16programchair@gmail.com}.
 
@@
 \bibliographystyle{named}
-\bibliography{ijcai15}
+\bibliography{ijcai16}
 

### 2016-11-04 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Adding admissible rules in section 2
Reference layout corrected

@@
 \begin{calculus}
+{Admissible propositional rules}
+{fig:AdmissiblePropositionalRules}
+
+\vspace{1em}
+
+\s\s
+\infer[\neg_E]{\bot}{ \neg A & A }
+\s\s
+\infer[\biimp_E]{A}{A\biimp B & B  }
+\s\s
+\infer[\biimp_E]{B}{A\biimp B & A  }
+
+\vspace{1em}
+\end{calculus}
+
+\begin{calculus}
 {Double negation elimination}
@@
 independent of the possible worlds).
-Neither G\"odel's manuscript nor Scott's manuscript reveal whether they use constant or varying domains, and this is also the case for many variants (e.g.
-\citep{hajek}).
+Neither G\"odel's manuscript nor Scott's manuscript reveal whether they use constant or varying domains, and this is also the case for many variants \citetext{e.g.
+\citealt{hajek}}.
 Nevertheless, some authors of variants of G\"odel's ontological argument do explicitly state a preference for varying domains.
@@
 Some common axiom schemas of relevance to the ontological argument are T ($\nec A \imp A$), B ($A \imp \nec \pos A$), 4 ($\nec A \imp \nec \nec A $) and 5 ($\pos A \imp \nec \pos A$).
-From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models (See for example \citep{fitting2} or \citep[Section 3.1]{blackburn} on Frame definability): T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
+From a semantic point of view, these axioms correspond to geometric frame conditions that must be satisfied by the accessibility relation in Kripke models \citetext{see e.g.
+\citealt{fitting2} or \citealt[Section 3.1]{blackburn} on Frame definability}: T corresponds to reflexivity, B corresponds to symmetry, 4 corresponds to transitivity and 5 corresponds to euclidianity.
 In the models of logic {\bf S5}, all these axioms are satisfied; in the models of logic {\bf KB}, B is satisfied; and in the models of logic {\bf B}, T and B are satisfied.
@@
 \BIC{$\nec\neg\nec\neg\neg A$}\doubleLine
-\UIC{$\nec\neg\nec A$} \RightLabel{$\imp_E$}
+\UIC{$\nec\neg\nec A$} \RightLabel{$\neg_E$}
 \BIC{$\bot$}\RightLabel{$\imp_I^1$}
@@
 			\UIC{$\omega: \all x. \neg \rho(x) $} \RightLabel{$\all_E$}
-			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\imp_E$}
+			\UIC{$\omega: \neg \rho(\beta)$} \RightLabel{$\neg_E$}
 		\BIC{$\omega: \bot$} \RightLabel{$\bot_E$}
@@
 					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
-				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
+				\BIC{$\neg P(\lambda x. x \neq x)$} \RightLabel{$\neg_E$}
 			\BIC{$[\subproof_3] \qquad \bot$}
@@
         \UIC{$\omega: \rho(\alpha)$} \RightLabel{$\ex_I$}
-		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\imp_E$}
+		\UIC{$\omega: \ex x. \rho(x)$} \RightLabel{$\neg_E$}
     \BIC{$\omega: \bot$} \RightLabel{$\imp_I^2$}
@@
 	        \UIC{$\bot$} \RightLabel{$\imp_I^3$}
-    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\imp_E$}
+    		\UIC{$\neg \nec \all x. \neg \rho(x)$} \RightLabel{$\neg_E$}
   		\BIC{$\bot$} \RightLabel{$\imp_I^4$}
@@
         \UIC{$\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))$} \RightLabel{$\all_E$}
-        \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
-        \UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+        \UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\biimp_E$}
 			\BIC{$P(\lambda x. \neg \rho(x))$}
@@
 					\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$} \RightLabel{$\imp_E$}
-				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\imp_E$}      
+				\BIC{$\neg \rho(\gamma)$} \RightLabel{$\neg_E$}      
 	   \BIC{$\bot$} \RightLabel{$\imp_I^1$}
@@
 			\AXC{S5 Iteration Principle} \dashedLine
-			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}
+			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}\RightLabel{$\imp_E$}
 		\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
@@
 \UIC{$\omega: \rho(x)$} \RightLabel{$\ex_I$}
-\UIC{$\omega: \ex x.\rho(x) $} \RightLabel{$\imp_E$}
+\UIC{$\omega: \ex x.\rho(x) $} \RightLabel{$\neg_E$}
 \BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
@@
 \AXC{$\subproof_5$} \dashedLine
-\UIC{$P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\UIC{$P(\lambda x. \neg \rho(x))$} %\RightLabel{$\imp_E$}
 \AXC{Axiom 1} \dashedLine
 \UIC{$\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))$} \RightLabel{$\all_E$}
-\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
-\UIC{$ P(\lambda x. \neg \rho(x)) \imp \neg P(\rho) $}  \RightLabel{$\imp_E$}
+\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\biimp_E$}
 \BIC{$\neg P(\rho) $} 
 \AXC{$ $} \RightLabel{3} 
-\UIC{$P(\rho)$} \RightLabel{$\imp_E$}
+\UIC{$P(\rho)$} \RightLabel{$\neg_E$}
 \BIC{$\bot$} \RightLabel{$\imp_I^2$}
@@
 \UIC{$\all \varphi. (P(\neg \varphi) \biimp \neg P(\varphi))$} \RightLabel{$\all_E$}
-\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
-\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
+\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\biimp_E$}
 \BIC{$P(\lambda x. \neg \rho(x))$}
@@
 \AXC{$ $} \RightLabel{3}
-\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
+\UIC{$\rho(\gamma)$} \RightLabel{$\neg_E$}
 \BIC{$\bot$}\RightLabel{$\imp_I^2$}
@@
 Anderson \citep[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}.
-Sobel \citep{sobel2}[p.
-152] acknowledges Anderson's conjecture, claims that it is correct and tries to informally explain how his formal proofs could be modified to rely on {\bf B} instead of {\bf S5}.
+Sobel \citep[p.
+152]{sobel2} acknowledges Anderson's conjecture, claims that it is correct and tries to informally explain how his formal proofs could be modified to rely on {\bf B} instead of {\bf S5}.
 Nevertheless, neither Anderson nor Sobel presented any formal proof relying only on {\bf B}.
@@
 Quantified variables may be rigid or flexible; and properties may be treated as intensional or extensional.
-Making the right choices may prevent the modal collapse \citep{fitting}[Sections 11.9 and 11.10].
-Koons \citep{koons}, Anderson \citep{and90}[p.
-292] and Sobel \citep{sobel2}[p.
-133] also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
+Making the right choices may prevent the modal collapse \citep[Sections 11.9 and 11.10]{fitting}.
+Koons \citep{koons}, Anderson \citep[p.
+292]{and90} and Sobel \citep[p.
+133]{sobel2} also discuss the idea that the notion of property over which quantification is allowed might be too general and restrictions might be appropriate.
 Two alternative and more conservative solutions, requiring only modifications of the axioms and definitions but no change in the logic, are discussed in greater detail in the next sections.
@@
 \AXC{$\subproof_9$} \dashedLine
-\UIC{$G(\gamma)$} \RightLabel{}\dottedLine \RightLabel{D\ref{A:D1}}
+\UIC{$G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}\RightLabel{$\wedge_E$}
-\UIC{$P(\lambda x.\rho(x)) \imp \nec \rho(\gamma)$}
+\UIC{$P(\lambda x.\rho(x)) \biimp \nec \rho(\gamma)$}
 \AXC{$ $} \RightLabel{4}
-\UIC{$\nec\rho(\gamma)$} \RightLabel{$\imp_E$}
+\UIC{$\nec\rho(\gamma)$} \RightLabel{$\biimp_E$}
 \BIC{$P(\rho)$} \RightLabel{$\forall_E$}
@@
 \UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
-\BIC{$[\subproof_{10}] \quad \nec P(\rho)$} \RightLabel{$\imp_E$}
+\BIC{$[\subproof_{10}] \quad \nec P(\rho)$} %\RightLabel{$\imp_E$}
 \end{prooftree}
@@
 \UIC{$\omega: \forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\wedge_E$}
+\UIC{$\omega: P(\rho) \biimp \nec \rho(y)$}\RightLabel{$\biimp_E$}
 \UIC{$\omega: P(\rho) \imp \nec \rho(y)$}\RightLabel{$\imp_E$}
@@
 \UIC{$ \all \varphi. (P(\varphi)\biimp \nec \varphi(\gamma))$}  \RightLabel{$\all_E$}
-\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\wedge_E$}
-\UIC{$ P(\varphi)\imp \nec \varphi(\gamma)$}  \RightLabel{$\imp_E$}
+\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
@@
 \UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\wedge_E$}
-\UIC{$P(E) \imp \nec E(\gamma)$}\RightLabel{$\imp_E$}
+\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
 \BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$}
@@
 \UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
-\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\wedge_E$}
-\UIC{$P(G) \imp \nec G(\gamma)$}
+\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\biimp_E$}
 \AXC{Axiom \ref{A:A3}} \RightLabel{}
@@
 \AXC{$ $} \RightLabel{1}
-\UIC{$\nec A$} \RightLabel{$\imp_E$}
+\UIC{$\nec A$} \RightLabel{$\biimp_E$}
 \BIC{$ \nec (\ex x. G(x) \imp A)$} \RightLabel{$\nec_E$}
@@
 \UIC{$\omega: G_B(x)$} \RightLabel{$\ex_I$}
-\UIC{$\omega: \ex x.G_B(x) $} \RightLabel{$\imp_E$}
+\UIC{$\omega: \ex x.G_B(x) $} \RightLabel{$\neg_E$}
 \BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
@@
 \AXC{$ $} \RightLabel{3} 
-\UIC{$P(G_B)$} \RightLabel{$\imp_E$}
+\UIC{$P(G_B)$} \RightLabel{$\neg_E$}
 \BIC{$\bot$} \RightLabel{$\imp_I^2$}
@@
 \UIC{$\all x. (G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}  \RightLabel{ $\forall_E$}
-\UIC{$ [G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}
+\UIC{$ (G_B(x) \imp \all\varphi. (\MCP(\varphi, x) \imp \nec \ex y.\varphi(y)))$}
 \AXC{$ $} \RightLabel{ $2$}

### 2016-11-04 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Author names added
Bibliography edited

@@
 
-\author{Annika Siders \and Bruno Woltzenlogel Paleo}
-
-\authorrunning{A.\~Siders \and B.\~Woltzenlogel Paleo}
+\author{Annika Kanckos \and Bruno Woltzenlogel Paleo}
+
+\authorrunning{A.\~Kanckos \and B.\~Woltzenlogel Paleo}
 
 \institute{ 
- Annika Siders \at 
+ Annika Kanckos \at 
   Department of Philosophy \\
@@
 %\title{A Variant of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus}
-\author{}
-\authorrunning{}
-\institute{}
+%\author{}
+%\authorrunning{}
+%\institute{}
 \maketitle
@@
 Leibniz.
-ed.
+Ed.
 Tandy, C.
@@
 \bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2015)}]{WorldCongressLogicAndReligion}
-Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2015. {\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}. 1st World Congress on Logic and Religion.
+Benzm\"uller, C.
+\& Weber, L.
+\& Woltzenlogel Paleo, B.
+2015.
+{\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}.
+Handbook of the 1st World Congress on Logic and Religion.
+Eds.
+Silvestre, R.
+S.
+\& B\'{e}ziau, J-Y.
 
@@
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013a)}]{IWIL}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013a. {\itshape G{\"{o}}del's God on the Computer}. 10th International Workshop on the Implementation of Logics. pp. 1--2.
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013a. {\itshape G{\"{o}}del's God on the Computer}. Proceedings of the 10th International Workshop on the Implementation of Logics. pp. 1--2.
 
@@
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014a)}]{WADT}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014a. {\itshape On Logic Embeddings and G\"odel's God}. Recent Trends in Algebraic Development Techniques.
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2014a.
+{\itshape On Logic Embeddings and G\"odel's God}.
+Procedings WADT 2014: Revised Selected Papers of the 22nd International Workshop on Algebraic Development Techniques.
+Eds.
+Codescu, M.
+et al.
+Recent Trends in Algebraic Development Techniques.
+Springer-Verlag New York.
+pp.
+3--6.
 
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014b)}]{WADTPreProceedings}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014b. {\itshape On Logic Embeddings and G\"odel's God}. 22nd International Workshop on Algebraic Development Techniques, pp. 8--9.
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2014b.
+{\itshape On Logic Embeddings and G\"odel's God}.
+Preliminary Proceedings of the 22nd International Workshop on Algebraic Development Techniques.
+Eds.
+Diaconescu, R.
+et al.
+pp.
+8--9.
 
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2014c)}]{SquareOfOpposition}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2014c. {\itshape Formalization and Automated Verification of G\"odel's Proof of God's Existence}. 4th World Congress on the Square of Opposition, pp. 24--25.
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2014c.
+{\itshape Formalization and Automated Verification of G\"odel's Proof of God's Existence}.
+Handbook of the World Congress on the Square of Opposition IV.
+Eds.
+B\'{e}ziau, J-Y.
+\& Gan-Krzywoszyrinska, K.
+pp.
+24--25.
 
@@
 {\itshape Automating G{\"{o}}del's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers}.
-ECAI 2014 - 21st European Conference on Artificial Intelligence, 18-22 August 2014, Prague, Czech Republic - Including Prestigious Applications of Intelligent Systems (PAIS 2014), pp.
+Proceedings ECAI 2014 - 21st European Conference on Artificial Intelligence- Including Prestigious Applications of Intelligent Systems (PAIS 2014).
+Eds.
+Schaub, T.
+et al.
+pp.
 93--98.
@@
 {\itshape Interacting with Modal Logics in the Coq Proof Assistant}.
-Computer Science - Theory and Applications - 10th International Computer Science Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July   13-17, 2015, Proceedings.
+Computer Science - Theory and Applications - Proceedings of the 10th International Computer Science Symposium in Russia, {CSR} 2015.
+Lecture Notes in Computer Science.
+Eds.
+Beklemishev, L.
+D.
+\& Musatov, D.
+V.
+Springer.
 pp.
@@
 Reasoning Web.
-Web Logic Rules - 11th International Summer School 2015, Berlin, Germany, July 31 - August 4, 2015, Tutorial Lectures.
+Web Logic Rules - 11th International Summer School 2015.
+Tutorial Lectures.
+Lecture Notes in Computer Science.
+Springer.
 pp.
@@
 {\itshape Experiments in Computational Metaphysics: G\"odel's Proof of God's Existence}.
-Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest (IIT Kharagpur, 30th October -- 1st November).
+Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest.
 vol.
@@
 {\itshape The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics}.
-International Joint Conference on Artificial Intelligence - IJCAI.
+Proceedings of the Twenty-Fifth International Joint Conference on Artificial Intelligence - IJCAI.
+Ed.
+Kambhampati, S.
+AAAI Press.
 
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016b)}]{ModalCollapse}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016b. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Springer. pp. 1--7.
+Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016b. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Ed. B\'{e}ziau, J-Y. Springer. pp. 1--7.
 
@@
 \bibitem[{\itshape Rushby(2013)}]{Rushby}
-Rushby, J. 2013. {\itshape The Ontological Argument in PVS}. CAV Workshop ``Fun with Formal Methods''. St. Petersburg.
+Rushby, J. 2013. {\itshape The Ontological Argument in PVS}. Proceedings of CAV Workshop ``Fun with Formal Methods''. St. Petersburg. Ed. Shilov, N.
 
@@
 
-\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January.
+\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January 2003.
 

### 2016-11-04 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Corrected up to p. 9

@@
 variant within the interactive proof assistants Isabelle and Coq were
-released, respectively, in Isabellle's Archive of Formal Proofs
+released, respectively, in Isabelle's Archive of Formal Proofs
 \citep{IsabelleGod} and in the Symposium of Computer Science in Russia
@@
 The natural deduction calculus proposed and used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}.
-The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction\footnote{Rules for $\eta$-expansion \citep{BenzmuellerBrownKohlhase} are not included here, because they are not needed in the ontological arguments.} is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
+The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}, and $\beta$-reduction\footnote{Rules for $\eta$-expansion \citep{BenzmuellerBrownKohlhase} are not included here, because they are not needed in the ontological arguments studied here.} is handled by the $\beta$ rule shown in Figure \ref{fig:ReductionRules}.
 Finally, modal operators are handled by special new rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
@@
 In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas differ due to substitution).
-For any expression occurring both in a formula in the premise and in its corresponding (sub-)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
+For any expression occurring both in a formula in the premise and in its corresponding (sub-) formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
 
@@
 \textbf{eigen-box condition:}\\ 
-An modal inference is said to \emph{access}\\
+a modal inference is said to \emph{access}\\
  the box immediately above or below it. \\
@@
 \UIC{$P(\alpha) \imp P(\alpha)$}\RightLabel{$\wedge_I$}
-\UIC{$(P(\alpha) \imp P(\alpha)) \wedge (P(\alpha) \imp P(\alpha))$} \dottedLine
-\BIC{$P(\alpha) \biimp P(\alpha)$}\RightLabel{$\forall_I$}
+\BIC{$(P(\alpha) \imp P(\alpha)) \wedge (P(\alpha) \imp P(\alpha))$} \dottedLine
+\UIC{$P(\alpha) \biimp P(\alpha)$}\RightLabel{$\forall_I$}
 \UIC{$\forall P. P(\alpha) \biimp P(\alpha)$}\dottedLine

### 2016-11-07 Papers/InProgress/NaturalDeduction/GodProof-ND.tex

@@
 \bibitem[{\itshape Bj{\o}rdal(1999)}]{bjordal}
-Bj{\o}rdal, F. 1999. Understanding G\"odel's Ontological Argument, in Timothy Childers (ed.) The Logica Yearbook 1998, pp. 214-217, Filosofia.
+Bj{\o}rdal, F. 1999. Understanding G\"odel's Ontological Argument. Ed. Childers, T. The Logica Yearbook 1998, pp. 214-217, Filosofia.
 
@@
 Fuhrmann, A. 2005.
-Existenz und Notwendigkeit -- Kurt G\"odel's Axiomatische Theologie, in Olsson e.J., Schr\"oder-Heister, P. and Spohn, W. (eds.) Logik in der Philosophie, pp. 349-374, Synchr.-Wissenschafts-Verlag.
+Existenz und Notwendigkeit -- Kurt G\"odel's Axiomatische Theologie. Eds. Olsson e.J., Schr\"oder-Heister, P. and Spohn, W. Logik in der Philosophie, pp. 349-374, Synchr.-Wissenschafts-Verlag.
 
@@
 \bibitem[{\itshape G\"odel(1970)}]{Goedel} 
-G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
+G\"odel, K. Ontological Proof. In Kurt G\"odel Collected Works vol. III. Ed. Feferman, S. et al., pp. 403--404, 139, 145 or {\itshape Appendix A. Notes in Kurt G\"odel's hand} in [Sobel 2001].
 
 \bibitem[{\itshape Hajek(1996)}]{hajek}
-Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof, in Ursini et alii
-(eds.) Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
+Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof. Eds. Ursini et al. Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
 
@@
 {\itshape Higher Order Modal Logic}.
-In P.
-Blackburn, J.F.A.K.
-van Benthem, and F.
-Wolter, editors, Handbook of Modal Logic, Studies in Logic and Practical Reasoning, pages 621-653.
+Eds.
+Blackburn, P.
+\& van Benthem, J.F.A.K.
+\& Wolter, F.
+Handbook of Modal Logic, Studies in Logic and Practical Reasoning, pages 621-653.
 Elsevier, Dordrecht, 2006.
@@
 
-\bibitem[{\itshape Paulin-Mohring(2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. In %D. Delahaye and B. Woltzenlogel Paleo, editors, 
+\bibitem[{\itshape Paulin-Mohring(2015)}]{coq} Paulin-Mohring, C. 2015. {\itshape Introduction to the calculus of inductive constructions}. Eds. Delahaye, D. \& Woltzenlogel Paleo, B. 
 All about Proofs, Proofs for All,
@@
  Woltzenlogel Paleo, B. 2013. {\itshape Automated Verification and Reconstruction of G\"odel's Proof of God's Existence}. Journal of the Austrian Computer Society. pp. 4--6.
+
+TODO: Check paper published in Logica Universalis 2016. 
+TODO: 2013a lacks editors. Same for 2015b. 2015c.
 

### 2016-11-07 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix typo

@@
 In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas differ due to substitution).
-For any expression occurring both in a formula in the premise and in its corresponding (sub-) formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
+For any expression occurring both in a formula in the premise and in its corresponding (sub-)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
 
@@
 
-\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilberts twenty-fourth problem}, American Mathematical Monthly, January 2003.
+\bibitem[{\itshape Thiele(2003)}]{hilbert} Thiele, R. 2003. {\itshape Hilbert's twenty-fourth problem}, American Mathematical Monthly, January 2003.
 

### 2016-11-07 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Add missing information to references in bibliography

@@
 
-\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2016)}]{controversy}
-Benzm\"uller, C. \& Weber, L. \& Woltzenlogel Paleo, B. 2016. {\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}. In: {\itshape Logica Universalis} 10 (2016)
+\bibitem[{\itshape Benzm\"uller \& Weber \& Woltzenlogel Paleo(2017)}]{controversy}
+Benzm\"uller, C.
+\& Weber, L.
+\& Woltzenlogel Paleo, B.
+2017.
+{\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}.
+To appear in: {\itshape Logica Universalis} Volume 11, Issue 1.
+Eds.
+Silvestre, R.
+S.
+\& B\'{e}ziau, J-Y.
+Springer.
+pp.
+1--11.
 
@@
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2013a)}]{IWIL}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2013a. {\itshape G{\"{o}}del's God on the Computer}. Proceedings of the 10th International Workshop on the Implementation of Logics. pp. 1--2.
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2013a.
+{\itshape G{\"{o}}del's God on the Computer}.
+Proceedings of the 10th International Workshop on the Implementation of Logics.
+Eds.
+Schulz, S.; Sutcliffe, G.
+and Konev, B.
+pp.
+1--2.
 
@@
 Lecture Notes in Computer Science.
+Eds.
+Faber, W.
+and Paschke, A.
 Springer.
@@
 Science and Spiritual Quest: Proceedings of the 9th All India Students' Conference on Science and Spiritual Quest.
-vol.
+Vol.
 9.
+Eds.
+Mishra, S.C.; Ghosh, S.
+and Agarwal, V.
+Bhaktivedanta Institute.
 pp.
@@
 \bibitem[{\itshape Benzm\"uller \& Woltzenlogel Paleo(2016b)}]{ModalCollapse}
-Benzm\"uller, C. \& Woltzenlogel Paleo, B. 2016b. {\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}. Studies in Universal Logic, Ed. B\'{e}ziau, J-Y. Springer. pp. 1--7.
+Benzm\"uller, C.
+\& Woltzenlogel Paleo, B.
+2016b.
+{\itshape The Modal Collapse as a Collapse of the Modal Square of Opposition}.
+Studies in Universal Logic.
+The Square of Opposition: A Cornerstone of Thought.
+Eds.
+B\'{e}ziau, J-Y.
+and Basti, G.
+Springer.
+pp.
+1--7.
 
@@
  Woltzenlogel Paleo, B. 2013. {\itshape Automated Verification and Reconstruction of G\"odel's Proof of God's Existence}. Journal of the Austrian Computer Society. pp. 4--6.
-
-TODO: Check paper published in Logica Universalis 2016. 
-TODO: 2013a lacks editors. Same for 2015b. 2015c.
-
-% TODO: transform the following bibtex entries into bibitems
 

### 2016-11-07 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix several minor issues noticed by the reviewer

@@
 
-In every inference rule, (sub-)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas differ due to substitution).
-For any expression occurring both in a formula in the premise and in its corresponding (sub-)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
+In every inference rule, (sub)formulas in the conclusion correspond\footnote{This correspondence could be defined formally, but it would be tedious to do so.} to formulas in the premises (even in the case of quantifier rules, where the formulas differ due to substitution).
+For any expression occurring both in a formula in the premise and in its corresponding (sub)formula in the conclusion, the occurrence in the premise is said to be an \emph{ancestor} of the occurrence in the conclusion, and the occurrence in the conclusion is said to be a \emph{descendant} of the occurrence in the premise.
 
@@
 \label{A1}
-Either a property or its negation is positive, but not both:
+Either a property or its negation\footnote{Throughout the paper, $\neg \varphi$ is an abbreviation for $\lambda x. \neg \varphi(x)$} is positive, but not both:
 $$
@@
 			  \UIC{$P(\lambda x. x = x)\scott$} 
-					\AXC{Half of Axiom 1} \dashedLine
+					\AXC{Contraposition of Half of Axiom 1} \dashedLine
 					\UIC{$P(\lambda x. x = x) \imp \neg P(\lambda x. x \neq x)$} \RightLabel{$\imp_E$}
@@
 $$
-\all \varphi.(P(\varphi) \to \Box \; P(\varphi))
+\all \varphi.(P(\varphi) \to \Box P(\varphi))
 $$
@@
 \label{A:D2}
-An \emph{essence} of an individual is a property possessed by it and necessarily implying those and only those properties necessarily possessed by it:
-$$
-\ess{\varphi}{x} \equiv \varphi(x) \wedge \all \psi. (\nec \psi(x) \biimp \nec \all x. (\varphi(x) \imp \psi(x)))
+An \emph{essence} of an individual is a property necessarily implying those and only those properties necessarily possessed by it:
+$$
+\ess{\varphi}{x} \equiv \all \psi. (\nec \psi(x) \biimp \nec \all x. (\varphi(x) \imp \psi(x)))
 $$
@@
 \AXC{$\subproof_9$} \dashedLine 
-\UIC{$  G(\gamma)$}  \RightLabel{D\ref{A:D1}}
+\UIC{$  G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
 \UIC{$ \all \varphi. (P(\varphi)\biimp \nec \varphi(\gamma))$}  \RightLabel{$\all_E$}
@@
 \UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y))) $}
-\AXC{$ \subproof_{9}$} \dashedLine
-\UIC{$G(\gamma)$}\RightLabel{$\wedge_I$}
-\BIC{$\ess{G}{\gamma}$}
+\dottedLine \RightLabel{D\ref{A:D2}}
+\UIC{$\ess{G}{\gamma}$}
 \AXC{Axiom \ref{A5}} \dashedLine
@@
 \AXC{Axiom \ref{A:A3}} \RightLabel{}
-\UIC{$P(G)$} \RightLabel{$\imp_E$}
+\UIC{$P(G)$} \RightLabel{$\biimp_E$}
 \BIC{$\nec G(\gamma)$}
@@
 Proofs of other lemmas, theorems and corollaries remain the same as in the previous section.
-Theorem \ref{A:T1} is derived from axioms \ref{A:A1} and \ref{A:A2}.
+Theorem \ref{A:T1} is derived from Axioms \ref{A:A1} and \ref{A:A2}.
 Corollary \ref{A:C1} is derived from Axiom \ref{A:A3} and Theorem \ref{A:T1}.
@@
 \UIC{$\all y.(G(y) \imp \ess{G}{y})$} \RightLabel{D2 }\doubleLine
-\UIC{$\all y.(G(y) \imp G(y) \wedge \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x))))$}
 \UIC{$\all y.(G(y) \imp \all \psi. (\nec \psi(y) \biimp \nec \all x. (G(x) \imp \psi(x))))$} \doubleLine
@@
 \AXC{Axiom 1} \dashedLine
-\UIC{$\all \varphi. (P(\neg G_B) \imp \neg P(G_B))$} \RightLabel{$\all_E$}
+\UIC{$\all \varphi. (P(\neg \varphi) \imp \neg P(\varphi))$} \RightLabel{$\all_E$}
 \UIC{$ P( \neg G_B(x)) \imp \neg P(G_B) $}  \RightLabel{$\imp_E$}
@@
 $$
-\pos \ex x. G(x)
+\pos \ex x. G_B(x)
 $$
@@
 \UIC{$\nec \all x. (G_B(x) \imp G_B(x)) $}\RightLabel{ D\ref{B:D1}}\dottedLine
-\UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
-\AXC{Theorem \ref{T1}}
+\UIC{$P_B(G_B) $} 
+\AXC{Theorem \ref{T1}} \dashedLine
 \UIC{$ P_B(G_B) \imp \pos \ex x.G_B(x) $} \RightLabel{ $\imp_E$}
@@
 {\bf KB} is also sufficient for the variants of Anderson and Bj{\o}rdal, which address Sobel's criticism of the modal collapse.
+
+\paragraph{Acknowledgements:} This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
+We thank Christoph Benzm\"uller for collaborations on several related works.
 

### 2016-11-07 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fixed reviewer's comments.

@@
 functional and boolean extensionality principles \citep{BenzmuellerBrownKohlhase}.
-Adding the modal rules results in a calculus that is suitable for a basic rigid modal logic \textbf{K}, in the sense that {\NDK} is sound and complete relative to {\ND} extended with axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
+Adding the modal rules results in a calculus that is suitable for a basic rigid modal logic \textbf{K}, in the sense that {\NDK} is sound and complete relative to {\ND} extended with Axiom K ($\nec(A\imp B)\imp (\nec A\imp \nec B)$) and the necessitation rule (which establishes that $\nec A$ is a theorem if $A$ is a theorem).
 
@@
 \label{theorem:completeness}
-{\NDK} is complete relative to {\ND} extended with axiom K and necessitation.
+{\NDK} is complete relative to {\ND} extended with Axiom K and necessitation.
 \end{theorem}
@@
 \label{theorem:soundness}
-{\NDK} is sound relative to {\ND} extended with axiom K and necessitation.
+{\NDK} is sound relative to {\ND} extended with Axiom K and necessitation.
 \end{theorem}
 \begin{proof}
-It is necessary to show that {\NDK} proofs of the following form can be translated to proofs in {\ND} extended with the axiom K and the necessitation rule.
+It is necessary to show that {\NDK} proofs of the following form can be translated to proofs in {\ND} extended with the Axiom K and the necessitation rule.
 
@@
 \noindent
-A translation to {\ND} extended with axiom K and necessitation is shown below for the case when $n=1$:
+A translation to {\ND} extended with Axiom K and necessitation is shown below for the case when $n=1$:
 
@@
 
-The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
+The straightforward combinations of the quantifier rules of {\ND} either with the modal rules of {\NDK} or with Axiom K and the necessitation rule are suitable for a higher-order modal logic where constants and variables are \emph{rigid}.
 From the point of view of a \emph{possible world semantics}, rigidity means that their interpretation is independent of the world in which they are being interpreted.
@@
 \noindent
-\emph{Brouwer's reduction principle} is derivable in modal logic {\bf KB}, using axiom B.
+\emph{Brouwer's reduction principle} is derivable in modal logic {\bf KB}, using Axiom B.
 The proofs of G\"odel and Scott (Section \ref{sec:scott}) do not make direct use of this principle, but the new proof presented in Section \ref{sec:newproof} does.
@@
 
-\begin{scriptsize}
-\begin{prooftree}
-
-\AXC{$ $} \RightLabel{1}
-\UIC{$\pos \ex x. G(x)$} 
+\begin{small}
+\begin{prooftree}
+
 		\AXC{Lemma 1} \dashedLine
@@
 		\UIC{$\nec (\ex x. G(x) \imp \nec \ex x. G(x)) $} 
-				\AXC{Distribution Principle}
-			\BIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
+				\AXC{Distribution Principle}\dashedLine
+				\UIC{$ \nec (\ex x. G(x) \imp \nec \ex x. G(x))\imp (\pos \ex x. G(x) \imp \pos \nec \ex x. G(x))$} \RightLabel{$\imp_E$}
+			\BIC{$[\subproof_5] \qquad \pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
+
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ $} \RightLabel{1}
+\UIC{$\pos \ex x. G(x)$} 
+ \AXC{$\subproof_5$}\dashedLine
+ 	\UIC{$\pos \ex x. G(x) \imp \pos \nec \ex x. G(x)$} \RightLabel{$\imp_E$}
  \BIC{$\pos \nec \ex x. G(x)\scott$} 
@@
 In this section we present a new proof in {\bf KB}.
-This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the axiom of reflexivity for equality nor {\bf S5}'s modal iteration principle.
+This proof uses exactly the same axioms and definitions of Scott's proof shown in the previous sections, but it uses neither the Axiom of reflexivity for equality nor {\bf S5}'s Modal iteration principle.
 Instead, it relies only on Brouwer's reduction principle (for Theorem 3).
@@
 \UIC{$(P(\rho) \wedge \nec \all x.(\rho(x) \imp \neg \rho(x))) \imp P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
-\BIC{$[\subproof_5]\qquad P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
+\BIC{$[\subproof_6]\qquad P(\lambda x. \neg \rho(x))$} \RightLabel{$\imp_E$}
 \end{prooftree}
@@
 \begin{prooftree}
-\AXC{$\subproof_5$} \dashedLine
+\AXC{$\subproof_6$} \dashedLine
 \UIC{$P(\lambda x. \neg \rho(x))$} %\RightLabel{$\imp_E$}
@@
 \UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
-\UIC{$[\subproof_6] \qquad P(\rho)$}
+\UIC{$[\subproof_7] \qquad P(\rho)$}
 \end{prooftree}
@@
 %\UIC{$\rho(\gamma)$} \RightLabel{}
-\AXC{$ \subproof_6$} \dashedLine
+\AXC{$ \subproof_7$} \dashedLine
 \UIC{$P(\rho)$} %\RightLabel{$\nec_E$}
@@
 \BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
-\UIC{$[\subproof_7] \qquad \ess{G}{\gamma}$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$\subproof_7$} \dashedLine
+\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$\subproof_8$} \dashedLine
 \UIC{$\ess{G}{\gamma}$}
@@
 \UIC{$\nec(\ex x. G(x)\imp \nec \ex x. G(x))$}\RightLabel{$\imp_E$}
-\BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
-\end{prooftree}
-\end{small}
-\begin{small}
-\begin{prooftree}
-\AXC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
+\BIC{$[\subproof_9] \qquad \pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
+\end{prooftree}
+\end{small}
+\begin{small}
+\begin{prooftree}
+\AXC{$\subproof_9$}\dashedLine
+\UIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
 \AXC{ Corollary \ref{C1} } \dashedLine
@@
 G(x)$).
-This contrasts with G\"odel's proof, where the actual existence is never actually proven (although it could, nevertheless, be derived as a trivial corollary by applying the modal axiom T).
+This contrasts with G\"odel's proof, where the actual existence is never actually proven (although it could, nevertheless, be derived as a trivial corollary by applying the modal Axiom T).
 
@@
 \UIC{$\all y.(G(y) \imp (A \imp \nec \all x. (G(x) \imp A)))$} \doubleLine
-\UIC{$[\subproof_8] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_8 $}\dashedLine
+\UIC{$[\subproof_{10}] \qquad \ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{10} $}\dashedLine
 \UIC{$\ex y.G(y) \imp (A \imp \nec \all x. (G(x) \imp A))$}
@@
 \begin{proof}
-This proof is omitted, because any of the proofs of Theorem 1 in Sections \ref{sec:scott} and \ref{sec:newproof} would suffice to prove this theorem with only a small modification: the equivalence of axiom \ref{A1} would not need to be reduced to an implication by conjunction elimination.
+This proof is omitted, because any of the proofs of Theorem 1 in Sections \ref{sec:scott} and \ref{sec:newproof} would suffice to prove this theorem with only a small modification: the equivalence of Axiom \ref{A1} would not need to be reduced to an implication by conjunction elimination.
 \end{proof}
@@
 Because of Anderson's emendations to Axiom 1 and Definition 1, the proof of Lemma 1 has to be modified as well.
-The proof shown below uses the modal axiom T to eliminate the extra box modality introduced by Anderson
-in his emended Definition 1.
 
@@
 \UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
-\UIC{$[\subproof_9] \quad G(\gamma)$}
-\end{prooftree}
-
-\begin{prooftree}
-\AXC{$\subproof_9$} \dashedLine
+\UIC{$[\subproof_{11}] \quad G(\gamma)$}
+\end{prooftree}
+
+\begin{prooftree}
+\AXC{$\subproof_{11}$} \dashedLine
 \UIC{$G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
@@
 \UIC{$P(\rho)\imp \nec P(\rho)$} \RightLabel{$\imp_E$}
-\BIC{$[\subproof_{10}] \quad \nec P(\rho)$} %\RightLabel{$\imp_E$}
+\BIC{$[\subproof_{12}] \quad \nec P(\rho)$} %\RightLabel{$\imp_E$}
 \end{prooftree}
@@
 
-\AXC{$\subproof_{10}$} \dashedLine
+\AXC{$\subproof_{12}$} \dashedLine
 \UIC{$\nec P(\rho)$}\RightLabel{$\imp_E$}
 \BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^4$}
-\UIC{$[\subproof_{11}] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
+\UIC{$[\subproof_{13}] \qquad \nec\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
 \end{prooftree}
@@
 \BIC{$P(\varphi)$}
-\AXC{$\subproof_9$} \dashedLine 
+\AXC{$\subproof_{11}$} \dashedLine 
 \UIC{$  G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
@@
 \BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
-\UIC{$[\subproof_{12}] \quad \nec  \forall y. (G(y)\imp \varphi(y))\imp \nec \varphi(\gamma)$}
+\UIC{$[\subproof_{14}] \quad \nec  \forall y. (G(y)\imp \varphi(y))\imp \nec \varphi(\gamma)$}
 \end{prooftree}
@@
 \begin{prooftree}
-\AXC{$\subproof_{11}$ and $\subproof_{12}$} \dashedLine \RightLabel{$\wedge_I$}
+\AXC{$\subproof_{13}$ and $\subproof_{14}$} \dashedLine \RightLabel{$\wedge_I$}
 \UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
@@
 \UIC{$P(E)$}\dottedLine 
-\AXC{$\subproof_9$} \dashedLine
+\AXC{$\subproof_{11}$} \dashedLine
 \UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{A:D1}}
@@
 
-In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify
-the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. The modified proof is shown below.
+In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's Axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify
+the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on Axioms \ref{A:A4} and \ref{A:A5} and Definitions \ref{A:D2} and \ref{A:D3}. The modified proof is shown below.
 
@@
 Finally, Theorem \ref{A:T3} can be derived from Corollary \ref{A:C1} and Lemma \ref{A:L1}, which, as shown above, depends only on Axiom \ref{A:A3} and Definition \ref{A:D1}.
-Therefore, confirming H\'ajek's claim, axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3} are superfluous: they are not needed to prove Theorem \ref{A:T3}.
+Therefore, confirming H\'ajek's claim, Axioms \ref{A:A4} and \ref{A:A5} and Definitions \ref{A:D2} and \ref{A:D3} are superfluous: they are not needed to prove Theorem \ref{A:T3}.
 
@@
 \UIC{$\all y.(G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A)))$} \doubleLine
-\UIC{$[\subproof_{13}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_{13} $}\dashedLine
+\UIC{$[\subproof_{15}] \qquad \ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{15} $}\dashedLine
 \UIC{$\ex y.G(y) \imp (\nec A \biimp \nec \all x. (G(x) \imp A))$}
@@
 This variant also significantly reduces the number of necessary axioms.
-Scott's variants of Axioms 2, 3 and 4 become unnecessary to prove the final theorem from Bj{\o}rdal's modified definitions 1, 2 and 3 and axioms 1 and 5.
+Scott's variants of Axioms 2, 3 and 4 become unnecessary to prove the final theorem from Bj{\o}rdal's modified Definitions 1, 2 and 3 and Axioms 1 and 5.
 
@@
 \AXC{$ $} \RightLabel{$2$}
+\UIC{$\ex x.G_B(x) $} \RightLabel{ $\ex_E$}
 \UIC{$G_B(x) $} 
@@
 \UIC{$P_B(G_B) $}\RightLabel{ $\wedge_I$}
-\BIC{$[\subproof_{14}] \qquad G_B(x) \wedge P_B(G_B) $}
-\end{prooftree}
-\end{small}
-
-\begin{small}
-\begin{prooftree}
-\AXC{$ \subproof_{14} $}\dashedLine
+\BIC{$[\subproof_{16}] \qquad G_B(x) \wedge P_B(G_B) $}
+\end{prooftree}
+\end{small}
+
+\begin{small}
+\begin{prooftree}
+\AXC{$ \subproof_{16} $}\dashedLine
 \UIC{$G_B(x) \wedge P_B(G_B) $}
@@
 \BIC{$G_B(x) \wedge P_B(G_B) \wedge \all\psi.((\psi(x) \wedge P_B (\psi)) \imp \nec\all y.(G_B(y) \imp \psi(y))) $} \RightLabel{ D\ref{B:D2} }\dottedLine
-\UIC{$[\subproof_{15}] \qquad \MCP(G_B, x) $}
+\UIC{$[\subproof_{17}] \qquad \MCP(G_B, x) $}
 
@@
 \AXC{$ $} \RightLabel{ $2$}
+\UIC{$\ex x.G_B(x) $}\RightLabel{ $\ex_E$}
 \UIC{$G_B(x) $}  \RightLabel{ $\imp_E$}
@@
 \UIC{$(\MCP(G_B, x) \imp \nec \ex y.G_B(y))$}
-\AXC{$ \subproof_{15} $}\dashedLine
+\AXC{$ \subproof_{17} $}\dashedLine
 \UIC{$\MCP(G_B, x) $} \RightLabel{ $\imp_E$}
 \BIC{$\nec \ex y.G_B(y)$}\RightLabel{ $\imp_I^2$}
-\UIC{$G_B(x) \imp \nec \ex y.G_B(y)$}
+\UIC{$\ex x.G_B(x) \imp \nec \ex y.G_B(y)$}
 \end{prooftree}

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Improve acknowledgements

@@
 
-\paragraph{Acknowledgements:} This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
-We thank Christoph Benzm\"uller for collaborations on several related works.
+\paragraph{Acknowledgements:} We thank Padre Edvaldo (from Piracicaba, Brazil) for assisting Bruno and his family in a moment of need.
+His help triggered the beginning of this work in 2013.
+We are grateful to Christoph Benzm\"uller for collaborations on several related works.
+This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
 

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix subproof below Axiom 2 in page 18
Make minor improvements to proofs
Improve acknowledgements.

@@
 \BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\imp_E$}
-\AXC{Axiom \ref{A:A2}} 
+\AXC{Axiom \ref{A:A2}} \dashedLine
+\UIC{\ldots\footnote{The formulas are ommitted here due to lack of space.}} \doubleLine \RightLabel{$\all_E$ x 2}
+\UIC{\ldots}
 \BIC{$P(\varphi)$}
@@
 \begin{prooftree}
-\AXC{$\subproof_{13}$ and $\subproof_{14}$} \dashedLine \RightLabel{$\wedge_I$}
-\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
+\AXC{$\subproof_{13}$}
+\AXC{$\subproof_{14}$} \RightLabel{$\wedge_I$}
+\BIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
 \UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y))) $}
@@
 \UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\biimp_E$}
-\AXC{Axiom \ref{A:A3}} \RightLabel{}
+\AXC{Axiom \ref{A:A3}} \dashedLine \RightLabel{}
 \UIC{$P(G)$} \RightLabel{$\biimp_E$}
@@
 \paragraph{Acknowledgements:} We thank Padre Edvaldo (from Piracicaba, Brazil) for assisting Bruno and his family in a moment of need.
-His help triggered the beginning of this work in 2013.
-We are grateful to Christoph Benzm\"uller for collaborations on several related works.
+Gratitude for his assistance is what triggered the production of an initial draft of this article in early 2013.
+We are thankful to Christoph Benzm\"uller for collaborations on several related articles.
 This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix proof of Lemma 1. Use Modal Axiom T.

@@
 Because of Anderson's emendations to Axiom 1 and Definition 1, the proof of Lemma 1 has to be modified as well.
+The proof shown below has been modified as little as possible. Interestingly, in order to eliminate 
+the extra box modality in Anderson's emended Definition D\ref{A:D1}, the modified proof below uses the modal axiom T.
 
@@
 \UIC{$\omega: P(\rho) \imp \nec \rho(y)$}\RightLabel{$\imp_E$}
-\BIC{$\omega: \nec \rho(y)$}\RightLabel{$\nec_E$}
-\UIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^2$}
+\BIC{$\omega: \nec \rho(y)$}  
+    \AXC{$\omega:$ Axiom T} \dashedLine
+    \UIC{$\omega: \nec \rho(y) \imp \rho(y)$} \RightLabel{$\imp_E$}
+\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^2$}
 \UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
@@
 In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's Axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify
-the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on Axioms \ref{A:A4} and \ref{A:A5} and Definitions \ref{A:D2} and \ref{A:D3}. The modified proof is shown below.
+the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on Axioms \ref{A:A4} and \ref{A:A5} and Definitions \ref{A:D2} and \ref{A:D3}.
+The modified proof is shown below.
+In contrast to the previous proof, the shorter proof does not use the modal axiom T.
+Therefore, modal logic \textbf{KB} is sufficient for Anderson's emendation as well.
 
@@
 The stronger modal logic {\bf S5}, which is usually taken for granted in discussions of G\"odel's ontological argument, is not necessary.
-This weakening of the underlying logic confirms a claim briefly mentioned by Anderson \citep[footnote 2]{and90}.
+This weakening of the underlying logic confirms and strengthens a claim briefly mentioned by Anderson \citep[footnote 2]{and90}.
 {\bf KB} is also sufficient for the variants of Anderson and Bj{\o}rdal, which address Sobel's criticism of the modal collapse.

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Boldface logic names

@@
 There are many higher-order logics \citep{BenzmuellerBrownKohlhase} and higher-order modal logics \citep{Muskens}.
-The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles.
-With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient to achieve the stated aim.
+The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic {\bf K} without extensionality principles.
+With additional axioms for stronger modal logics, such as {\bf KB} and {\bf S5}, this calculus is sufficient to achieve the stated aim.
 
@@
  \BIC{$\pos \nec \ex x. G(x)\scott$} 
-			\AXC{S5 Iteration Principle} \dashedLine
+			\AXC{{\bf S5} Iteration Principle} \dashedLine
 			\UIC{$\pos \nec \ex x. G(x) \imp \nec \ex x. G(x)$}\RightLabel{$\imp_E$}

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix footnote about omitted Axiom 2 in proof of lemma 1 and write about the need for Axioms 2 and 3 in the proof of Anderson’s lemma 1.

@@
 
-Because of Anderson's emendations to Axiom 1 and Definition 1, the proof of Lemma 1 has to be modified as well.
+Because of Anderson's emendations to Axiom 1 and Definitions 1 and 2, the proof of Lemma 1 has to be modified\footnote{Due to lack of space, the instantiation of Axiom \ref{A:A2} are ommitted in the emended proof of Lemma \ref{A:L1}.} as well.
 The proof shown below has been modified as little as possible. Interestingly, in order to eliminate 
-the extra box modality in Anderson's emended Definition D\ref{A:D1}, the modified proof below uses the modal axiom T.
+the extra box modality in Anderson's emended Definition \ref{A:D1}, the modified proof below uses the modal axiom T. 
+Likewise, to obtain the extra box in the emended Definition \ref{A:D2}, Axioms 2 and 3 were used.
 
@@
 \AXC{Axiom \ref{A:A2}} \dashedLine
-\UIC{\ldots\footnote{The formulas are ommitted here due to lack of space.}} \doubleLine \RightLabel{$\all_E$ x 2}
-\UIC{\ldots}
+\UIC{.} \doubleLine \RightLabel{$\all_E$ x 2}
+\UIC{.}
 \BIC{$P(\varphi)$}

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Minor improvements

@@
 
-\author{Annika Kanckos \and Bruno Woltzenlogel Paleo}
+\author{Annika Kanckos \and Bruno Woltzenlogel Paleo\thanks{Author order is alphabetical by surname.}}
 
@@
   Austria \\
-   \email{bruno@logic.at} \\
+   \email{bruno@logic.at}
   \and
@@
 The first formal proof closely follows the hints in Scott's manuscript about G\"odel's argument and fills in the gaps, thus verifying its correctness.
-The second formal proof improves the first one, by relying on the weaker modal logic {\bf KB} instead of the stronger modal logic {\bf S5}; and by avoiding the equality relation.
+The second formal proof improves the first one, by relying on the weaker modal logic {\bf KB} instead of {\bf S5} and by avoiding the equality relation.
 The second proof is also technically shorter than the first one, because it eliminates unnecessary detours and uses Axiom 1 for the positivity of properties only once.
@@
 \end{abstract}
+
+\smallskip
 
@@
 This deductive inference is not formalizable in a finite first-order calculus.
-The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \citep{scott}, this axiom of G\"odel has been replaced by another that simply assumes the positivity of the property of being god-like.
+The interplay between universal quantification (in the definition of a God-like being) and infinite conjunctions (in G\"odel's axiom for positive properties) could explain why, starting with Scott \citep{scott}, this axiom of G\"odel has been replaced by another that simply states the positivity of the property of being god-like.
 
@@
 The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic {\bf K} without extensionality principles.
-With additional axioms for stronger modal logics, such as {\bf KB} and {\bf S5}, this calculus is sufficient to achieve the stated aim.
+By adding axioms for stronger modal logics (e.g.
+{\bf KB} and {\bf S5}), this calculus is sufficient to achieve this paper's aim.
 
@@
 The new proof also does not rely on the equality relation and is much shorter.
-Although there have been claims (as discussed in more detail in Section \ref{sec:newproof}) about the sufficiency of \textbf{KB} and the superfluousness of the equality relation, formal proofs substantiating these claims were missing up to now.
-
-A major criticism against G\"odel's argument is an undesirable consequence of the stipulated axioms called \emph{modal collapse}.
+Although there have been claims (as discussed in more detail in Section \ref{sec:newproof}) about the sufficiency of weaker modal logics and the superfluousness of the equality relation, formal proofs substantiating these claims were missing up to now.
+
+A major criticism against G\"odel's argument is an undesirable consequence of the stipulated axioms known as \emph{modal collapse}.
 This is discussed in greater detail in Section \ref{sec:collapse}, where a natural deduction derivation of the collapse is presented.
@@
 \citealt{hajek}}.
-Nevertheless, some authors of variants of G\"odel's ontological argument do explicitly state a preference for varying domains.
-Our choice of constant domains is motivated by simplicity \citep[footnotes 11 and 14]{and90}.
+Nevertheless, some authors of variants of G\"odel's ontological argument do explicitly state a preference for varying domains \citep[footnotes 11 and 14]{and90}.
+Our choice of constant domains is motivated by simplicity.
 As is well-known, varying domain quantifiers (also known as \emph{actualistic} quantifiers) can be simulated by constant domain quantifiers (also known as \emph{possibilistic} quantifiers) guarded by a primitive \emph{existence} predicate\footnote{Formally, $\forall^e x.
@@
 Unmarked formulas were derived in the process of filling the gaps between the marked formulas.
-
-% ToDo: uncomment this post-acceptance: the journal uses double blind reviewing.
-%The formal natural deduction proof presented here was verified in the Coq proof assistant\footnote{\url{https://github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Coq/GoedelGod_Scott.v}}.
 
@@
 \end{corollary}
+
+\clearpage
+
 \begin{proof} \hfill
@@
 The simpler proof of Theorem 1 is not particularly special.
-In fact, it is probably the most likely proof that any well-trained logician would construct, if asked to derive Theorem 1 from Axioms 1 and 2 using methodical formal proof search techniques.
+In fact, it is probably the most likely proof that any well-trained logician would construct, if asked to derive Theorem 1 from Axioms 1 and 2 using methodical proof search techniques.
 Modern techniques usually rely on unification to compute or guess good instances for weakly quantified variables, and therefore they would never result in instantiations containing $=$ and $\neq$, which are symbols that do not appear anywhere in the axioms or in the theorems to be proved.
 We may wonder why G\"odel preferred to use those unusual instantiations.
-It might be an interesting question from a historical point of view, as it might suggest that there was an evolution in the way logicians construct formal proofs; or from a philosophical point of view, if G\"odel's preference was philosophically motivated.
+It might be an interesting question from a historical point of view, suggesting that there may have been an evolution in the way logicians construct formal proofs, or from a philosophical point of view, if G\"odel's preference was philosophically motivated.
 
@@
 We show only the strongest implication above.
-The others are easy corollaries in modal logics where T holds.
-% ToDo: do we really need T?
+The others are easy corollaries in modal logics where, for instance, $\nec A \imp \pos A$ or $\pos \top$ holds.
 
@@
 
-Because of Anderson's emendations to Axiom 1 and Definitions 1 and 2, the proof of Lemma 1 has to be modified\footnote{Due to lack of space, the instantiation of Axiom \ref{A:A2} are ommitted in the emended proof of Lemma \ref{A:L1}.} as well.
+Because of Anderson's emendations to Axiom 1 and Definitions 1 and 2, the proof of Lemma 1 has to be modified\footnote{Due to lack of space, the instantiation of Axiom \ref{A:A2} is ommitted in the emended proof of Lemma \ref{A:L1}.} as well.
 The proof shown below has been modified as little as possible. Interestingly, in order to eliminate 
@@
 
-%ToDo: acknowledgements
-
 \section{Conclusions}
@@
 We are thankful to Christoph Benzm\"uller for collaborations on several related articles.
-This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
+This work was supported by a Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
 

### 2016-12-07 Papers/2016/Studia Logica - NaturalDeduction/GodProof-ND.tex
Move accepted Natural Deduction paper

--- 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
@@
 prepared by Woltzenlogel Paleo in March 2013, served as the starting
-point for two distinct but complementary research endeavours. Siders
+point for two distinct but complementary research endeavours. Kanckos
 and Woltzenlogel Paleo transformed that early draft into this paper,
@@
 \citep{IJCAI2016}. The automatic verification of variants by Anderson,
-Bjordal and H\'ajek were presented in the First
+Bj\o rdal and H\'ajek were presented in the First
 World Congress on Logic and Religion
@@
 
-Applications of the $\beta$ rule are silently ommitted.
+Applications of the $\beta$ rule are silently omitted.
 Double lines are used to abbreviate propositional reasoning steps when they are tedious and too space-consuming.
@@
 
-In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s iteration principle (in the proof of Lemma 2) and the lemma of reflexivity for equality (in the proof of Theorem 1).
+In addition to the five axioms and three definitions that pertain specifically to the argument, the proof also uses {\bf S5}'s Iteration principle (in the proof of Lemma 2) and the lemma of reflexivity for equality (in the proof of Theorem 1).
 The use of reflexivity could be avoided, but we chose to use it because it allows the construction of a shorter proof.
@@
 G(x)$ from $\pos \nec \ex x.
-G(x)$ and {\bf S5}'s iteration principle.
+G(x)$ and {\bf S5}'s Iteration principle.
 Instead, we can derive $\ex x.
@@
 
-Since the iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle requires only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s iteration principle is philosophically profound.
+Since the Iteration principle requires the strong modal logic {\bf S5} while Brouwer's reduction principle requires only the much weaker modal logic {\bf KB},  the possibility of proving Theorem 3 using Brouwer's reduction principle instead of {\bf S5}'s Iteration principle is philosophically profound.
 There have been many discussions of what modal axioms are needed for the ontological argument.
 In their manuscripts, G\"odel and Scott do not write explicitly which logic they use.
-However, a proof step in their notes clearly relies on {\bf S5}'s \emph{iteration principle}.
+However, a proof step in their notes clearly relies on {\bf S5}'s \emph{Iteration principle}.
 Anderson \citep[footnote 5]{and90} cites several articles questioning the adequacy of {\bf S5} for the ontological argument and conjectures that the ontological argument could be restricted to the weaker modal logic {\bf B}.
@@
 We show only the strongest implication above.
-The others are easy corollaries in modal logics where, for instance, $\nec A \imp \pos A$ or $\pos \top$ holds.
+The others are easy corollaries in sufficiently strong modal logics (e.g.
+where, for instance, $\nec A \imp \pos A$ or $\pos \top$ holds).
 
@@
 
-Because of Anderson's emendations to Axiom 1 and Definitions 1 and 2, the proof of Lemma 1 has to be modified\footnote{Due to lack of space, the instantiation of Axiom \ref{A:A2} is ommitted in the emended proof of Lemma \ref{A:L1}.} as well.
+Because of Anderson's emendations to Axiom 1 and Definitions 1 and 2, the proof of Lemma 1 has to be modified\footnote{Due to lack of space, the instantiation of Axiom \ref{A:A2} is omitted in the emended proof of Lemma \ref{A:L1}.} as well.
 The proof shown below has been modified as little as possible. Interestingly, in order to eliminate 
@@
 \UIC{$\nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\wedge_I$}
-\BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\imp_E$}
+\BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}
 \AXC{Axiom \ref{A:A2}} \dashedLine
 \UIC{.} \doubleLine \RightLabel{$\all_E$ x 2}
-\UIC{.}
+\UIC{.}\RightLabel{$\imp_E$}
 \BIC{$P(\varphi)$}
@@
 \bibitem[{\itshape Anderson \& Gettings (1996)}]{and}
-Anderson, C. A.\& Gettings, M. 1996.  {\itshape G\"odel's ontological proof revisited}. In: edited by Hajek P. {\itshape G\"odel '96},  Springer.
+Anderson, C. A.\& Gettings, M. 1996.  {\itshape G\"odel's ontological proof revisited}. In: edited by H\'ajek P. {\itshape G\"odel '96},  Springer.
 
@@
 2015.
-{\itshape Computer-Assisted Analysis of the Anderson-Hajek Ontological Controversy}.
+{\itshape Computer-Assisted Analysis of the Anderson-H\'ajek Ontological Controversy}.
 Handbook of the 1st World Congress on Logic and Religion.
@@
 
-\bibitem[{\itshape Hajek(1996)}]{hajek}
-Hajek, P. 1996. Magari and others on G\"odel's Ontological Proof. Eds. Ursini et al. Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
+\bibitem[{\itshape H\'ajek(1996)}]{hajek}
+H\'ajek, P. 1996. Magari and others on G\"odel's Ontological Proof. Eds. Ursini et al. Logic and Logical Algebra, pp. 125--136, Marcel Dekker.
 

### 2016-12-30 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
modifications requested by report 1

--- 2015-07-16 Papers/InProgress/HajekAndersonControversy/main.tex
@@
 \usepackage{url}
+\usepackage{rotating}
 \usepackage{commands}
+\usepackage{comment}
+\usepackage{booktabs}
+\usepackage{modallogics}
 \usepackage{xcolor}
-\usepackage{comment}
-\usepackage{modallogics}
 \usepackage[backend=biber,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
@@
 
-% %----------AUTHOR 1
-% \author[Benzm\"uller]{C. Benzm\"uller}
+%----------AUTHOR 1
+\author[Benzm\"uller]{C. Benzm\"uller}
 
-% \address{%
-% Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
-% }
-% \email{c.benzmueller@fu-berlin.com}
+\address{%
+Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
+}
+\email{c.benzmueller@fu-berlin.com}
 
-% \thanks{This work was supported by German National Research Foundation (DFG) under
-%  grants BE 2501/9-1 and BE 2501/11-1.}
+\thanks{This work was supported by German National Research Foundation (DFG) under
+ grants BE 2501/9-1 and BE 2501/11-1.}
 
-% %----------Author 2
-% \author[Weber]{L. Weber}
-% \address{
-% Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
-% }
-% \email{leon.weber@fu-berlin.de}
+%----------Author 2
+\author[Weber]{L. Weber}
+\address{
+Dep.~of Mathematics and Computer Science, Freie Universit\"at Berlin, Germany
+}
+\email{leon.weber@fu-berlin.de}
 
-% %----------Author 3
-% \author[Woltzenlogel-Paleo]{B. Woltzenlogel Paleo}
-% \address{
-% Room HA0402, Favoritenstra{\ss}e 9, 1040 Wien, Austria
-% }
-% \email{bruno.wp@gmail.com}
+%----------Author 3
+\author[Woltzenlogel-Paleo]{B. Woltzenlogel Paleo}
+\address{
+Room HA0402, Favoritenstra{\ss}e 9, 1040 Wien, Austria
+}
+\email{bruno.wp@gmail.com}
 
@@
 \item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$
-\item[L1] If a god-like being exists, then necessarily a god-like being exists:
+\item[L1] If a God-like being exists, then necessarily a God-like being exists:
   $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
-\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists:
+\item[L2] If possibly a God-like being exists, then necessarily a God-like being exists:
   $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
@@
 Emendation} (cf. Fig. \ref{fig:Anderson}): T1, C
-and T3' can be quickly automated (in logics \K, \K and \KB,
-respectively); the axioms A4 and A5' are proven redundant (the former in logic \KFourB and the latter
+and T3$'$ can be quickly automated (in logics \K, \K and \KB,
+respectively); the axioms A4 and A5$'$ are proven redundant (the former in logic \KFourB and the latter
 already in \K); a trivial countermodel (with two worlds and  two
@@
   $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$
-\item[A3']  The property of being God-like is positive:
+\item[A3$'$]  The property of being God-like is positive:
   $$\hol{P(G_A)}$$
@@
   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
-\item[T2']  Being God-like is an essence of any
+\item[T2$'$]  Being God-like is an essence of any
   God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+\item[D3$'$] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
   \exq y \varphi(y)]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
-\item[L1'] If a god-like being exists, then necessarily a god-like being exists:
+\item[A5$'$] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
+\item[L1$'$] If a God-like being exists, then necessarily a God-like being exists:
   $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
-\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists:
+\item[L2$'$] If possibly a God-like being exists, then necessarily a God-like being exists:
   $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
 %
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
+\item[T3$'$] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
 \end{itemize}
@@
 models of the other axioms and definitions by Gödel.
-\citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this
+\citet[pp.~5--6]{Hajek_Magari_and_others_1996} investigated this
 further, and claimed that Magari's claim is not valid, but is
@@
 }
-of A4 and A5', as well as varying domain semantics, in his new
+of A4 and A5$'$, as well as varying domain semantics, in his new
 emendation (named \AOEH\/ \citep[sec.~4]{Hajek2002}, cf. Fig.
@@
 axiom H:A12. Surprisingly, the computer-assisted formalization of
-\AOEH\ shows that A4 and A5' are still superfluous. Moreover,
-A4 and A5' are independent
+\AOEH\ shows that A4 and A5$'$ are still superfluous. Moreover,
+A4 and A5$'$ are independent
 of the other axioms and definitions.
-Therefore, A4 and A5' are not redundant, despite their superfluousness.
+Therefore, A4 and A5$'$ are not redundant, despite their superfluousness.
 
@@
   $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
-\item[A3']  The property of being God-like is positive:
+\item[A3$'$]  The property of being God-like is positive:
   $$\hol{P(G_H)}$$
@@
   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+\item[D3$'$] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
   \exq y [\varphi(y)]]}$$
-\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
+\item[A5$'$] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$
 
@@
 $$\hol{\allq \varphi[P(\varphi) \imp \pos \exq x \varphi(x)] }$$
-  \item[(3)] If a god-like being exists, then necessarily a god-like being exists:
+  \item[(3)] If a God-like being exists, then necessarily a God-like being exists:
 $$\hol{\allq x[G_H(x) \imp \nec G_H(x)))] }$$
-  \item[(4)] All positive properties are necessarily implied by the property of being god-like:
+  \item[(4)] All positive properties are necessarily implied by the property of being God-like:
 $$\hol{\allq \varphi[P(\varphi) \imp \nec \allq x[G_H(x) \imp \varphi(x)]] }$$
@@
 
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
+\item[T3$'$] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
 %
@@
 
-Although Hájek did not notice the superfluousness of A4 and A5' in his
+Although Hájek did not notice the superfluousness of A4 and A5$'$ in his
 \AOEH, he did describe yet another emendation (his \AOEHH, cf.
-Fig. \ref{fig:Hajek2}) where A4 and A5' are superfluous (though no
-claim is made w.r.t. to their redundancy), if A3' is replaced by a
+Fig. \ref{fig:Hajek2}) where A4 and A5$'$ are superfluous (though no
+claim is made w.r.t. to their redundancy), if A3$'$ is replaced by a
 stronger axiom (H:A3) additionally stating that the property of actual
@@
 \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
-not only superflous, but also redundant. In addition, A5' becomes superfluous
+not only superflous, but also redundant. In addition, A5$'$ becomes superfluous
 and independent. Surprisingly, a countermodel for the weaker
-A3' was succesfully generated. This is somewhat unsatisfactory (for
+A3$'$ was succesfully generated. This is somewhat unsatisfactory (for
 theistic goals), because it shows that \AOEHH\ does not entail the
@@
   $$\hol{P(G_A \wedge E)}$$
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
+\item[T3$'$] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
 %
@@
 That is his final emendation (\AOEHHH, cf. Fig. \ref{fig:Hajek3}), which
-restores A3' and does use A4 and A5', albeit in a modified form (i.e.
+restores A3$'$ and does use A4 and A5$'$, albeit in a modified form (i.e.
 H:A4 and H:A5).
@@
   $$\hol{G_H(x) \equiv \forall \varphi [P^\#(\varphi) \biimp \nec \varphi(x)]}$$
-\item[A3']  The property of being God-like is positive:
+\item[A3$'$]  The property of being God-like is positive:
   $$\hol{P(G_H)}$$
@@
   \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
-\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
+\item[D3$'$] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
   $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
@@
 The formalization of \AOEHHH\ shows that both H:A4 and H:A5 are superfluous
-aswell as independent. For the old A5', no conclusive results were achieved.
+as well as independent. For the old A5$'$, no conclusive results were achieved.
 
@@
 14]{anderson90:_some_emend_of_goedel_ontol_proof} (cf. Fig. \ref{fig:Anderson_simp}) remarks that only
-the quantifiers in T3' and in A:D2 need to be interpreted as
+the quantifiers in T3$'$ and in A:D2 need to be interpreted as
 actualistic quantifiers, while others may be taken as possibilistic
 quantifiers. Our computer-assisted study of this mixed variant shows
-that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
+that A4 is still redundant in logic \KFourB, but A5$'$ becomes independent (hence not
 redundant). Unfortunately, a countermodel for T3 can then be found.
@@
    $$\hol{P_{AS}(\varphi) \equiv \nec (\allq x (\neg \varphi(x) \imp D(x)) \wedge (\neg \nec \allq x (\varphi(x) \imp D(x))))}$$
-\item[A:A1'] If a property is positive, its negation is not positive:
+\item[A:A1$'$] If a property is positive, its negation is not positive:
   $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \neg P_{AS}(\neg \varphi)]}$$
-\item[A2'] A property necessarily implied by a
+\item[A2$'$] A property necessarily implied by a
   positive property is positive:
@@
   \imp \psi(x)]) \imp P_{AS}(\psi)]}$$
-\item[A4']  Positive properties are necessarily positive:
+\item[A4$'$]  Positive properties are necessarily positive:
   $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \nec P_{AS}(\varphi)]}$$
@@
   $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$
-\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
-  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
+\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1$'$ and axioms A2$'$, A3$'$ and A4$'$.
+  $$\hol{\textrm{B:D1} \biimp \textrm{D1$'$} \wedge \textrm{A2$'$} \wedge \textrm{A3$'$} \wedge \textrm{A4$'$} }$$
 %  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
@@
   $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
-\item[A:A1'] If a property is positive, its negation is not positive:
+\item[A:A1$'$] If a property is positive, its negation is not positive:
   $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$
-\item[A5'] Necessary existence is a positive property.
+\item[A5$'$] Necessary existence is a positive property.
  $$\hol{P_B(\NE_B)}$$
-\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$
+\item[T3$'$] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$
 \end{itemize}
@@
 } \vskip-.5em
-\caption{Bjordal's Alternative \label{fig:Bjordal}}
+\caption{Bjørdal's Alternative \label{fig:Bjordal}}
 \end{figure}
@@
 then briefly indicates (B:L1) that B:D1 is logically equivalent, under
-modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
-has been confirmed in the computer-assisted formalization: A2' and A3'
-can be quickly automatically derived in logic \K. A4' can be proved in
+modal logic \SFour, to the conjunction of D1$'$, A2$'$, A3$'$ and A4$'$. This
+has been confirmed in the computer-assisted formalization: A2$'$ and A3$'$
+can be quickly automatically derived in logic \K. A4$'$ can be proved in
 logic \KT (i.e. assuming reflexivity of the accessibility relation).
-For constant domain semantics, proving D1' is possible in logic \KFour,
+For constant domain semantics, proving D1$'$ is possible in logic \KFour,
 whereas for varying domain semantics, a countermodel can be found
 even in logic \SFive. Conversely, the proof that B:D1 is entailed by
-D1', A2', A3' and A4' is possible already in logic \K. The provers
-also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
-A5' already in logic \KB. Bjordal's last paragraph briefly mentions
-Hájek's ideas about the superfluousness of A5' and claims that it is
+D1$'$, A2$'$, A3$'$ and A4$'$ is possible already in logic \K. The provers
+also show that theorem T3$'$ follows from B:D1, B:D2, B:D3, A:A1$'$ and
+A5$'$ already in logic \KB. Bjørdal's last paragraph briefly mentions
+Hájek's ideas about the superfluousness of A5$'$ and claims that it is
 possible, with (unclear) additional modifications of the definitions,
-to eliminate A5' from his theory as well. Without any additional
-modification, the automated reasoners show that A5' is independent, but actually
+to eliminate A5$'$ from his theory as well. Without any additional
+modification, the automated reasoners show that A5$'$ is independent, but actually
 not superfluous. All these results, with the exception of the
-aforementioned countermodel for D1', hold for both constant and
+aforementioned countermodel for D1$'$, hold for both constant and
 varying domain semantics.
 
-Figure \ref{fig:summary} summarizes the results obtained by the automated reasoning tools.
-The following abbreviations are used: S/I = superfluous and independent; R = superfluous and redundant; S/U = superfluous and unknown whether redundant or independent; N/I = non-superfluous and independent; P = provable; CS = counter-satisfiable.
-The weakest logic required to show redundancy or provability is indicated in parentheses (\K is the default).
-Cells highlighted in red contain results that differ from what had been claimed by either Magari, Anderson or Hájek.
-Cells highlighted in yellow contain results that are surprising, albeit not contradicting any claims.
-Cells highlighted in green contain results where the tools were able to obtain the same results as humans, but using weaker modal logics.
+\def\tabred{red!50}
+\def\tabgreen{green!50}
+\def\tabyellow{yellow!50}
+\def\tabbox#1#2{\colorbox{#1}{\begin{minipage}[c][1em]{3.3em}
+      \centering #2 \end{minipage}}}
+\def\tabboxs#1#2{\colorbox{#1}{\begin{minipage}[c][1em]{1.5em}
+      \centering #2 \end{minipage}}}
 
 \begin{figure}
-\includegraphics[width=\textwidth]{Summary.png}
-\caption{Summary of Results}
+\begin{sideways}
+\includegraphics[width=1.2\textwidth]{Summary.png} (to be removed)
+\end{sideways}
+%
+\begin{sideways}
+\scalebox{.9}{
+\begin{tabular}{lccccccccccccc} 
+\toprule
+Investigated Variant & D1$'$ & A:A1$'$ & A2$'$ & A3$'$ & A4 & A4$'$ & H:A4 & A5 & A5$'$
+  & H:A5 & T3 & T3$'$ & MC \\  
+\midrule 
+Scott (const) & - & - & - & - & \tabboxs{\tabred}{N/I} & - & - & \tabboxs{\tabred}{N/I} & - & - & P & -  & P \\
+Scott (var) & - & - & - & - & \tabboxs{\tabred}{N/I} & - & - & \tabboxs{\tabred}{N/I} & - & - & P & -  & P \\
+Anderson (const) & - & - & - & - & \tabbox{\tabred}{R(K4B)} & - & - & - & \tabboxs{\tabred}{R}  & - & - & P  & CS \\
+Anderson (var) & - & - & - & - & \tabbox{\tabred}{R(K4B)} & - & - & - & \tabboxs{\tabred}{R}  & - & - & P  & CS \\
+Anderson (mix) & - & - & - & - & \tabbox{\tabred}{R(K4B)} & - & - & - & \tabboxs{\tabred}{I}  & - & - & \tabboxs{\tabred}{CS} & CS \\
+H\'ajek AOE$'$ (var) & - & - & \tabboxs{\tabyellow}{CS}  &  & \tabboxs{\tabred}{S/I} & - & - & - & \tabboxs{\tabred}{S/I}  & - & - & P(KB) & CS \\
+H\'ajek AOE$'$\_0 (var) & - & - & - & \tabboxs{\tabyellow}{CS} & \tabboxs{\tabgreen}{R} & - & - & - & S/U & - & - & P(KB) & CS \\
+H\'ajek AOE$''$ (var) & - & - &  &  & - & - & \tabboxs{\tabred}{S/I} & - & - & \tabboxs{\tabred}{S/I} & - & P(KB) & CS \\
+Anderson (simp) (var) & - & R & R  &  &  & R(K4B) & - & - & & - & - &  & \\
+Bjørdal (const) & \tabbox{\tabgreen}{R(K4)} & - & \tabboxs{\tabgreen}{R} & \tabboxs{\tabgreen}{R} & - & \tabbox{\tabgreen}{R(KT)} & - & - &\tabboxs{\tabyellow}{N/I} & - & - & \tabbox{\tabgreen}{P(KB)} &  CS \\
+Bjørdal (var) & \tabboxs{\tabyellow}{CS} & - & \tabboxs{\tabgreen}{R} & \tabboxs{\tabgreen}{R} & - & \tabbox{\tabgreen}{R(KT)} & - & - &\tabboxs{\tabyellow}{N/I} & - & - & \tabbox{\tabgreen}{P(KB)} &  CS \\
+\bottomrule
+\end{tabular}
+}
+\end{sideways}
+\caption{Summary of Results.}
 \label{fig:summary}
 \end{figure}
+
+Figure \ref{fig:summary} summarizes the results obtained by the
+automated reasoning tools. The following abbreviations are used: S/I =
+superfluous and independent; R = superfluous and redundant; S/U =
+superfluous and unknown whether redundant or independent; N/I =
+non-superfluous and independent; P = provable; CS =
+counter-satisfiable. The weakest logic required to show redundancy or
+provability is indicated in parentheses (\K is the default).  Cells
+highlighted in \colorbox{\tabred}{red} contain results that differ
+from what had been claimed by either Magari, Anderson or Hájek. Cells
+highlighted in \colorbox{\tabyellow}{yellow} contain results that are
+surprising, albeit not contradicting any claims. Cells highlighted in
+\colorbox{\tabgreen}{green} contain results where the tools were able
+to obtain the same results as humans, but using weaker modal logics.
+\footnote{EDNOTE: What is the difference between - and empty cells?}
 

### 2017-01-02 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
modifications requested by report 1

@@
 
-\noindent The axioms in G\"odel's ontological proof
-\citep{GoedelNotes,ScottNotes} (cf. Fig. \ref{fig:Goedel}) entail what is called \emph{modal collapse}
-\citep{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
-\varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
-just for $\exists x. \mathit{God}(x)$ as intended.
+\section{Introduction}
+We present an exemplary study in \emph{Computational Metaphysics},
+which is an emerging, interdisciplinary field aiming at the rigorous
+formalisation and deep logical assessment of rational, philosophical
+arguments on the computer.  The particular focus here is on the
+ontological argument for the existence of God. More precisely, we
+further expand our studies on modern (modal) variants of the
+ontological argument that are essentially emendations of Kurt Gödel's
+seminal contribution \citep{GoedelNotes}, respectively, Dana Scott's
+variant of it \cite{ScottNotes}. In related experiments
+\citep{C55,C40,J30}, we have already demonstrated that our technology
+is suited not only to formally assess the logical quality of the
+presented arguments, but it may even generate new knowledge about
+it. For example, the theorem prover Leo-II \cite{Leo-II} detected an
+inconsistency in Gödel's original variant \citep{GoedelNotes} of the
+argument, which is not the case for Scott \citep{ScottNotes}.  
+We here assume some familiarity with these related experiments.
 
+The particular motivation and context for work presented in this
+article is as follows: The axioms in G\"odel's ontological proof
+\citep{GoedelNotes,ScottNotes} (cf. Fig. \ref{fig:Goedel}) entail what
+is called \emph{modal collapse} \citep{Sobel1987,SobelBook2004}: the
+formula $\varphi \rightarrow \Box \varphi$, abbreviated as MC, holds
+for any formula $\varphi$ and not just for
+$\exists x. \mathit{God}(x)$ as intended.
 \begin{figure}[t]
@@
 \end{figure}
-
-This fact, which
-has recently been confirmed with higher-order automated theorem
-provers \citep{C40,J30}, has led to strong criticism of the argument
-and stimulated attempts to remedy the problem.
+This fact, which has been confirmed with higher-order automated
+theorem provers \citep{C40,J30} in our previous experiments, has led
+to strong criticism of the argument and stimulated attempts to remedy
+the problem.
 \citet{Hajek_Magari_and_others_1996,Hajek_der_Mathematiker_2001}
 proposed the use of cautious instead of full comprehension principles,
-and \citet{fitting02:_types_tableaus_god} took
-greater care of the semantics of higher-order quantifiers
-in the presence of modalities. Others, such as
-\citet{anderson90:_some_emend_of_goedel_ontol_proof}, \citet{Hajek2002}
-and \citet{bjordal99}, proposed emendations of G\"odel's
-axioms and definitions. They require neither comprehension
+and \citet{fitting02:_types_tableaus_god} took greater care of the
+semantics of higher-order quantifiers in the presence of
+modalities. Others, such as
+\citet{anderson90:_some_emend_of_goedel_ontol_proof},
+\citet{Hajek2002} and \citet{bjordal99}, proposed emendations of
+G\"odel's axioms and definitions. They require neither comprehension
 restrictions nor more complex semantics. Therefore, they are
 technically simpler to analyze with computer support. We have
-formalized those using the proof assistant Isabelle/HOL \citep{Isabelle}
-together with the automated higher-order reasoners Leo-II \citep{C26},
-Satallax \citep{brown2012satallax}, Metis
+formalized those using the proof assistant Isabelle/HOL
+\citep{Isabelle} together with the automated higher-order reasoners
+Leo-II \citep{Leo-II}, Satallax \citep{brown2012satallax}, Metis
 \citep{Hurd03first-orderproof}, and Nitpick \citep{Nitpick}.  Our
-formalizations\footnote{The   formalizations are available in the
-subdirectories \url{Anderson}, \url{Hajek} and \url{Bjordal} at \url{gi
-thub.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/
-}.} employ the embedding of higher-order modal logic (HOML) in
-classical higher-order logic (HOL) as introduced in previous work
-\citep{C40,J30,J23}. We explored the effect of different domain
-conditions on the provability of lemmas, theorems and even axioms.
-This was motivated by a controversy between Hájek and Anderson
-regarding the redundancy of some axioms in Anderson's emendation. In
-\emph{constant domain semantics}, the individual domains are the same
-in all possible worlds. In \emph{varying domain semantics}, the
-domains may vary from world to world. This variation is technically
-encoded with the help of an existence relation expressing which
-individuals actually exist in each world. Quantifiers are then
-uniformly formalized as \emph{actualistic quantifiers} (i.e. guarded by
-the existence relation). Our main results are summarized here.
+formalizations employ the embedding of higher-order modal logic (HOML)
+in classical higher-order logic (HOL) as introduced in previous work
+\citep{C40,J23}. We explored the effect of different domain conditions
+on the provability of lemmas, theorems and even axioms.  This was
+motivated by a controversy between Hájek and Anderson regarding the
+redundancy of some axioms in Anderson's emendation. In \emph{constant
+  domain semantics}, the individual domains are the same in all
+possible worlds. In \emph{varying domain semantics}, the domains may
+vary from world to world. This variation is technically encoded with
+the help of an existence relation expressing which individuals
+actually exist in each world. Quantifiers are then uniformly
+formalized as \emph{actualistic quantifiers} (i.e. guarded by the
+existence relation). \footnote{EDNOTE: This is misleading statement;
+  we need to reword}
+Our main results, which we have presented at the
+1st World Congress on Logic and Religion \citep{C41}, are summarized
+below. The corresponding formalizations, containing all the details,
+are available in the subdirectories \url{Anderson}, \url{Hajek} and
+\url{Bjordal} at
+\url{github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.
 
+\section{Results from our Experiments}
 For all emendations and variants discussed here, the axioms and
@@
 
-Using our approach, the formalization and (partly) automated analysis
-of several variants of G\"odel's ontological argument has been
-surprisingly straightforward. The provers not only confirmed many
-claimed results, but also exposed a few mistakes and novel insights
-in a long standing controversy.
-We believe the technology employed in
-this work is ready to be fruitfully adopted in larger scale by
-philosophers.
+\section{Conclusion}
+Using our approach (semantical embeddings of variants of higher-order
+modal logics in classical higher-order logic), the formalization and
+(partly) automated analysis of several variants of G\"odel's
+ontological argument has been surprisingly straightforward. The
+provers not only confirmed many claimed results, but also exposed a
+few mistakes and novel insights in a long standing controversy.  We
+believe the technology employed in this work is ready to be fruitfully
+adopted in larger scale by philosophers. Moreover, the technology is
+by now means restricted to metaphysics or philosophy, and further work
+includes similar applications in other areas, including, for example,
+law, politics and ethics.
 

### 2017-01-02 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
modifications requested by report 2

--- 2016-12-30 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
@@
 
-\noindent The axioms in G\"odel's ontological proof
-\citep{GoedelNotes,ScottNotes} (cf. Fig. \ref{fig:Goedel}) entail what is called \emph{modal collapse}
-\citep{Sobel1987,SobelBook2004}: the formula $\varphi \rightarrow \Box
-\varphi$, abbreviated as MC, holds for any formula $\varphi$ and not
-just for $\exists x. \mathit{God}(x)$ as intended.
+\section{Introduction}
+We present an exemplary study in \emph{Computational Metaphysics},
+which is an emerging, interdisciplinary field aiming at the rigorous
+formalisation and deep logical assessment of rational, philosophical
+arguments on the computer.  The particular focus here is on the
+ontological argument for the existence of God.  While related work
+\citep{oppenheimer11,rushby13} has concentrated on Anselm's simpler original
+version of the ontological argument and formalized it in classical
+predicate logic, we here focus on modern variants of it requiring
+higher-order modal logic. More precisely, the variants studied here are
+all emendations of Kurt Gödel's seminal contribution
+\citep{GoedelNotes}, respectively, Dana Scott's variant of it
+\cite{ScottNotes} (cf. Fig.~\ref{fig:Goedel}).
+
+In preceding experiments \citep{C55,C40,J30}, we have already
+demonstrated that the technology we employ is well suited for the
+task. In fact, computers may even contribute philosophically relevant
+new knowledge. For example, the theorem prover Leo-II \cite{Leo-II}
+detected an inconsistency in Gödel's original script
+\citep{GoedelNotes} of the argument, which was undetected by
+philosophers before.  This inconsistency obviously renders Gödel's
+original version of the argument as pointless.  However, as our
+preceding experiments also confirmed, Scott's variant
+\citep{ScottNotes} from Fig.~\ref{fig:Goedel} is immune to this
+issue.\footnote{Gödel avoids the first conjunct $\varphi(x)$ in the
+  definition of \emph{essence (D2)}, which Scott added for cosmetic
+  reason. However, as Leo-II detected, one can prove from Gödel's
+  definition that the empty property becomes an essence of every
+  individual, which, together with theorem T1, axiom A5 and the
+  defintion of \emph{necessary existence (D3)} causes the
+  inconsistency. For more details see \citet{C55}.} Hence, in the
+remainder of this article the term \emph{Gödel's ontological argument}
+actually refers to Scott's variant of it, whose soundness has been
+formally 
+verified with our technology (for logics entailing higher-order modal
+logic KB with constant domain or varying domain semantics). We here
+assume some familiarity with these related experiments and papers.
 
@@
 
-This fact, which
-has recently been confirmed with higher-order automated theorem
-provers \citep{C40,J30}, has led to strong criticism of the argument
-and stimulated attempts to remedy the problem.
+The particular motivation and context for the work presented in the given
+article is as follows: It is well known that the axioms in G\"odel's
+ontological proof (cf. Fig. \ref{fig:Goedel}) entail what is called
+\emph{modal collapse} \citep{Sobel1987,SobelBook2004}. This means that
+the formula $\varphi \rightarrow \Box \varphi$, abbreviated as MC,
+holds for any formula $\varphi$ and not just for
+$\exists x. \mathit{God}(x)$ as intended.  Hence, contingent truth
+implies necessary truth, or, in other words, there are no unnecessary
+contingent truths. In fact, one may even go as far and interpret MC as a result
+against free will.
+%
+The fact that MC follows from Gödel's axioms, which has also been
+confirmed with higher-order automated theorem provers in our previous
+experiments \citep{C40,J30}, has led to strong criticism of the
+argument and stimulated attempts to remedy the problem.
 \citet{Hajek_Magari_and_others_1996,Hajek_der_Mathematiker_2001}
 proposed the use of cautious instead of full comprehension principles,
-and \citet{fitting02:_types_tableaus_god} took
-greater care of the semantics of higher-order quantifiers
-in the presence of modalities. Others, such as
-\citet{anderson90:_some_emend_of_goedel_ontol_proof}, \citet{Hajek2002}
-and \citet{bjordal99}, proposed emendations of G\"odel's
-axioms and definitions. They require neither comprehension
-restrictions nor more complex semantics. Therefore, they are
-technically simpler to analyze with computer support. We have
-formalized those using the proof assistant Isabelle/HOL \citep{Isabelle}
-together with the automated higher-order reasoners Leo-II \citep{C26},
-Satallax \citep{brown2012satallax}, Metis
-\citep{Hurd03first-orderproof}, and Nitpick \citep{Nitpick}.   Our
-formalizations\footnote{The   formalizations are available in the
-subdirectories \url{Anderson}, \url{Hajek} and \url{Bjordal} at \url{gi
-thub.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/
-}.} employ the embedding of higher-order modal logic (HOML) in
-classical higher-order logic (HOL) as introduced in previous work
-\citep{C40,J30,J23}. We explored the effect of different domain
-conditions on the provability of lemmas, theorems and even axioms.
-This was motivated by a controversy between Hájek and Anderson
-regarding the redundancy of some axioms in Anderson's emendation. In
-\emph{constant domain semantics}, the individual domains are the same
-in all possible worlds. In \emph{varying domain semantics}, the
-domains may vary from world to world. This variation is technically
-encoded with the help of an existence relation expressing which
-individuals actually exist in each world. Quantifiers are then
-uniformly formalized as \emph{actualistic quantifiers} (i.e. guarded by
-the existence relation). Our main results are summarized here.
+and \citet{fitting02:_types_tableaus_god} took greater care of the
+semantics of higher-order quantifiers in the presence of
+modalities. Others, such as
+\citet{anderson90:_some_emend_of_goedel_ontol_proof},
+\citet{Hajek2002} and \citet{bjordal99}, proposed emendations of
+G\"odel's axioms and definitions. They require neither comprehension
+restrictions nor more complex semantics and are therefore technically
+simpler to analyze with computer support. Hence, we here concentrate
+on the latter emendations of Gödel's argument and leave the former
+ones for future work.  Again, familiarity with the papers by
+\citet{anderson90:_some_emend_of_goedel_ontol_proof},
+\citet{Hajek2002} and \citet{bjordal99} is assumed, since we here are only
+briefly commenting/assessing various claims made in there.
 
-For all emendations and variants discussed here, the axioms and
+Our formalizations employ the embedding of higher-order modal logic
+(HOML) in classical higher-order logic (HOL) as introduced in previous
+work \citep{C40,J23} as enabling technology.  We explored the effect
+of different domain conditions on the provability of lemmas, theorems
+and even axioms.  This was motivated by a controversy between Hájek
+and Anderson regarding the redundancy of some axioms in Anderson's
+emendation. In \emph{constant domain semantics} (\emph{possibilist
+  notion of quantification}), the individual domains are the same in
+all possible worlds. In \emph{varying domain semantics}
+(\emph{actualist notion of quantification}), the domains may vary from
+world to world. The latter notion is technically encoded in our
+approach with the help of an existence relation expressing which
+individuals actually exist in each world. In other words, actualist
+quantification is formalized as possibilist quantification guarded by
+the existence relation.\footnote{EDNOTE: This part was
+  misleading/wrong before; I reworded it.} This technical solution
+enables us to flexibly switch between these different notions of
+quantification and to even mix them. When we mention varying
+domain semantics (actualist quantification) below, we in fact mean
+varying domain semantics for the domain of individuals only, while for
+the other types we still assume the standard constant domain semantics
+(possibilist quantification).  This is, to our best knowledge, in line 
+with the intention of the authors (it would of course be easily
+possible in our approach to study other combinations).
+
+In our
+experiments we have utilized the proof assistant Isabelle/HOL
+\citep{Isabelle} together with the automated higher-order reasoners
+Leo-II \citep{Leo-II}, Satallax \citep{Satallax}, Metis
+\citep{Hurd03first-orderproof}, and Nitpick \citep{Nitpick}.
+
+Our main results, which we have
+presented at the 1st World Congress on Logic and Religion \citep{C41},
+are presented in more detail below. The corresponding formalizations,
+containing all the details, are available online: see the
+subdirectories \url{Anderson}, \url{Hajek} and \url{Bjordal} at
+\url{github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.
+\footnote{Ideally, Logica Universalis should provide means for
+  permanent storing of such additional material.}
+
+\section{Results of the Experiments}
+For \emph{all emendations and variants} discussed here, the axioms and
 definitions have been shown to be consistent and not to entail modal
@@
 
-\begin{samepage}
+Next, we address some controversially debated claims regarding the redundancy,
+superfluousness and independence of particular axioms in the
+investigated emendations and variants. Several novel findings are
+reported.
+
+%\begin{samepage}
 The following definitions of redundancy, superfluousness and independence are used throughout this paper:
@@
 \begin{definition}
-An axiom $A$ is \textbf{redundant} w.r.t. a set of axioms $S$ iff $S$ entails $A$. 
+An axiom $A$ is \emph{redundant} w.r.t. a set of axioms $S$ iff $S$ entails $A$. 
 \end{definition}
@@
 \begin{definition}
-An axiom $A$ is \textbf{superfluous} w.r.t. axiom set $S$ iff $S \setminus \{ A \}$ entails T3.
+An axiom $A$ is \emph{superfluous} w.r.t. axiom set $S$ iff $S \setminus \{ A \}$ entails T3.
 \end{definition}
@@
 \begin{definition}
-  An axiom $A$ is \textbf{independent} of a set of axioms $S$ iff
+  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff
   there are models of $S$ where $A$ is true and other models of
@@
 \end{definition}
-\end{samepage}
+%\end{samepage}
 
-\clearpage
+%\clearpage
 
-For both \textbf{constant domain semantics} and \textbf{varying domain
-semantics}, the following results hold for \emph{Anderson's
+For both {constant domain semantics} (possibilist quantification) and {varying domain
+semantics} (actualist quantification), the following results hold for \emph{Anderson's
 Emendation} (cf. Fig. \ref{fig:Anderson}): T1, C
@@
 not only superfluous, but also redundant. \citet[footnote 1 in
-p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim,
+p.~1]{AndersonGettings} rebutted Hájek's claim,
 arguing that the redundancy of A4 and A5 holds only under constant
@@
 semantics). Our results show that Hájek was originally right, under
-both constant and varying domain semantics.
+both constant and varying domain semantics (for the domain of individuals).
 
@@
 \begin{sideways}
-\includegraphics[width=1.2\textwidth]{Summary.png} (to be removed)
+%\includegraphics[width=1.2\textwidth]{Summary.png} \fbox{EDNOTE: to be removed}
 \end{sideways}
@@
 \begin{sideways}
-\scalebox{.9}{
+%\scalebox{.9}{
+\resizebox{19cm}{4.5cm}{
 \begin{tabular}{lccccccccccccc} 
@@
 
-Using our approach, the formalization and (partly) automated analysis
-of several variants of G\"odel's ontological argument has been
-surprisingly straightforward. The provers not only confirmed many
-claimed results, but also exposed a few mistakes and novel insights
-in a long standing controversy.
-We believe the technology employed in
-this work is ready to be fruitfully adopted in larger scale by
-philosophers.
+\section{Conclusion}
+Using our approach, that is, semantical embeddings of variants of higher-order
+modal logics in classical higher-order logic, the formalization and
+(partly) automated analysis of several variants of G\"odel's
+ontological argument has been surprisingly straightforward. The
+higher-order provers we employed not only confirmed many claimed results, but also exposed a
+few mistakes and novel insights in a long standing controversy.  We
+believe the technology employed in this work is ready to be fruitfully
+adopted in larger scale by philosophers.  Moreover, our approach is
+by no means restricted to metaphysics or philosophy, and further work
+includes similar applications in other areas, including, for example,
+law, politics and ethics. In fact, we claim, that our approach, at
+least to some degree, a realises Leibniz' dream of a characteristica universalis
+and calculus ratiocinator.
 

### 2017-01-04 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Fix a minor conceptual mistake about the notion of contingency.
Fix minor typos and mistakes.

@@
 holds for any formula $\varphi$ and not just for
-$\exists x. \mathit{God}(x)$ as intended.  Hence, contingent truth
-implies necessary truth, or, in other words, there are no unnecessary
-contingent truths. In fact, one may even go as far and interpret MC as a result
+$\exists x. \mathit{God}(x)$ as intended.  Hence, anything that is the case is necessarily the case, actual truth implies necessary truth, or, in other words, there are no
+contingent truths. In fact, one may even go further and interpret MC as a result
 against free will.
@@
 quantification is formalized as possibilist quantification guarded by
-the existence relation.\footnote{EDNOTE: This part was
-  misleading/wrong before; I reworded it.} This technical solution
+the existence relation. This technical solution
 enables us to flexibly switch between these different notions of
@@
 \url{github.com/FormalTheology/GoedelGod/blob/master/Formalizations/Isabelle/}.
-\footnote{Ideally, Logica Universalis should provide means for
-  permanent storing of such additional material.}
+\footnote{Ideally, Logica Universalis should provide
+  permanent storage of such additional material.}
 
@@
 \begin{sideways}
-%\includegraphics[width=1.2\textwidth]{Summary.png} \fbox{EDNOTE: to be removed}
+%\includegraphics[width=1.2\textwidth]{Summary.png} 
 \end{sideways}
@@
 law, politics and ethics. In fact, we claim, that our approach, at
-least to some degree, a realises Leibniz' dream of a characteristica universalis
+least to some degree, realizes Leibniz' dream of a characteristica universalis
 and calculus ratiocinator.

### 2017-01-06 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Clarify reference to metis

@@
 experiments we have utilized the proof assistant Isabelle/HOL
-\citep{Isabelle} together with the automated higher-order reasoners
-Leo-II \citep{Leo-II}, Satallax \citep{Satallax}, Metis
-\citep{Hurd03first-orderproof}, and Nitpick \citep{Nitpick}.
+\citep{Isabelle} together with the external automated higher-order provers
+Leo-II \citep{Leo-II} and Satallax \citep{Satallax}, called through Sledgehammer.
+Sledgehammer analyzes the proofs generated by these provers and tells which axioms or previously proven lemmas are need to reconstruct the proof inside Isabelle using Isabelle's \emph{metis} method \citep{PaulsonSusanto}, which relies on a Isabelle porting of the resolution-based first-order prover Metis
+\citep{Hurd03first-orderproof}. To find counter-examples, we used Nitpick \citep{Nitpick}, also through Isabelle.
 

### 2017-01-06 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Remove redundant sentence

@@
 version of the ontological argument and formalized it in classical
-predicate logic, we here focus on modern variants of it requiring
-higher-order modal logic. More precisely, the variants studied here are
-all emendations of Kurt Gödel's seminal contribution
-\citep{GoedelNotes}, respectively, Dana Scott's variant of it
-\cite{ScottNotes} (cf. Fig.~\ref{fig:Goedel}).
+predicate logic, we here focus on modern variants of Kurt Gödel's seminal contribution
+\citep{GoedelNotes} requiring
+higher-order modal logic.
+
+% More precisely, the variants studied here are
+% all emendations of Dana Scott's variant
+% \cite{ScottNotes} (cf. Fig.~\ref{fig:Goedel}).
 

### 2017-01-06 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Add missing dashes

@@
 H\'ajek AOE$'$\_0 (var) & - & - & - & \tabboxs{\tabyellow}{CS} & \tabboxs{\tabgreen}{R} & - & - & - & S/U & - & - & P(KB) & CS \\
-H\'ajek AOE$''$ (var) & - & - &  &  & - & - & \tabboxs{\tabred}{S/I} & - & - & \tabboxs{\tabred}{S/I} & - & P(KB) & CS \\
-Anderson (simp) (var) & - & R & R  &  &  & R(K4B) & - & - & & - & - &  & \\
+H\'ajek AOE$''$ (var) & - & - & - & - & - & - & \tabboxs{\tabred}{S/I} & - & - & \tabboxs{\tabred}{S/I} & - & P(KB) & CS \\
+Anderson (simp) (var) & - & R & R  & - & - & R(K4B) & - & - & - & - & - & - & - \\
 Bjørdal (const) & \tabbox{\tabgreen}{R(K4)} & - & \tabboxs{\tabgreen}{R} & \tabboxs{\tabgreen}{R} & - & \tabbox{\tabgreen}{R(KT)} & - & - &\tabboxs{\tabyellow}{N/I} & - & - & \tabbox{\tabgreen}{P(KB)} &  CS \\
@@
 to obtain the same results as humans, but using weaker modal logics.
-\footnote{EDNOTE: What is the difference between - and empty cells?}
 

### 2017-01-06 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Improve various sentences

@@
 \citep{GoedelNotes} of the argument, which was undetected by
-philosophers before.  This inconsistency obviously renders Gödel's
-original version of the argument as pointless.  However, as our
+philosophers before.  This inconsistency renders pointless Gödel's
+original version of the argument. Nevertheless, as our
 preceding experiments also confirmed, Scott's variant
@@
 remainder of this article the term \emph{Gödel's ontological argument}
-actually refers to Scott's variant of it, whose soundness has been
+actually refers to Scott's variant, whose logical validity has been
 formally 
-verified with our technology (for logics entailing higher-order modal
-logic KB with constant domain or varying domain semantics). We here
+verified with our technology (for higher-order modal
+logic KB (or stronger logics) with constant domain or varying domain semantics). Here we
 assume some familiarity with these related experiments and papers.
@@
 \citet{Hajek_Magari_and_others_1996,Hajek_der_Mathematiker_2001}
-proposed the use of cautious instead of full comprehension principles,
+proposed the use of cautious restrictions of comprehension principles,
 and \citet{fitting02:_types_tableaus_god} took greater care of the
@@
 restrictions nor more complex semantics and are therefore technically
-simpler to analyze with computer support. Hence, we here concentrate
+simpler to analyze with computer support. Therefore, we concentrate
 on the latter emendations of Gödel's argument and leave the former
@@
 \citet{anderson90:_some_emend_of_goedel_ontol_proof},
-\citet{Hajek2002} and \citet{bjordal99} is assumed, since we here are only
-briefly commenting/assessing various claims made in there.
+\citet{Hajek2002} and \citet{bjordal99} is assumed, since we only
+briefly assess various claims made there.
 
@@
 (HOML) in classical higher-order logic (HOL) as introduced in previous
-work \citep{C40,J23} as enabling technology.  We explored the effect
-of different domain conditions on the provability of lemmas, theorems
-and even axioms.  This was motivated by a controversy between Hájek
+work \citep{C40,J23} as an enabling technology to settle a long-standing
+controversy between Hájek
 and Anderson regarding the redundancy of some axioms in Anderson's
-emendation. In \emph{constant domain semantics} (\emph{possibilist
+emendation. A core issue in the controversy was whether Anderson's emendation
+should be interpreted with constant or varying domains and how the provability
+of lemmas, theorems and even axioms was affected by this choice.
+In \emph{constant domain semantics} (\emph{possibilist
   notion of quantification}), the individual domains are the same in
@@
 Leo-II \citep{Leo-II} and Satallax \citep{Satallax}, called through Sledgehammer.
-Sledgehammer analyzes the proofs generated by these provers and tells which axioms or previously proven lemmas are need to reconstruct the proof inside Isabelle using Isabelle's \emph{metis} method \citep{PaulsonSusanto}, which relies on a Isabelle porting of the resolution-based first-order prover Metis
+Sledgehammer analyzes the proofs generated by these provers and tells which axioms or previously proven lemmas are needed to reconstruct the proof inside Isabelle using its \emph{metis} method \citep{PaulsonSusanto}, which relies on a reimplementation of the resolution-based first-order prover Metis
 \citep{Hurd03first-orderproof}. To find counter-examples, we used Nitpick \citep{Nitpick}, also through Isabelle.
@@
 
-Next, we address some controversially debated claims regarding the redundancy,
+Next, we investigate some controversially debated claims regarding the redundancy,
 superfluousness and independence of particular axioms in the

### 2017-01-06 Papers/2017/LogicaUniversalis-HajekAndersonControversy/main.tex
Improve conclusion

@@
 \section{Conclusion}
-Using our approach, that is, semantical embeddings of variants of higher-order
+Using our approach of semantically embedding higher-order
 modal logics in classical higher-order logic, the formalization and
@@
 higher-order provers we employed not only confirmed many claimed results, but also exposed a
-few mistakes and novel insights in a long standing controversy.  We
+few mistakes and novel insights in a long-standing controversy.  We
 believe the technology employed in this work is ready to be fruitfully
@@
 by no means restricted to metaphysics or philosophy, and further work
-includes similar applications in other areas, including, for example,
-law, politics and ethics. In fact, we claim, that our approach, at
-least to some degree, realizes Leibniz' dream of a characteristica universalis
-and calculus ratiocinator.
+could explore similar applications in other areas, including, for example,
+law, politics and ethics. In fact, we claim that our approach, at
+least to some degree, realizes Leibniz's dream of a characteristica universalis
+and a calculus ratiocinator.
 

### 2017-05-30 Formalizations/Isabelle/Salamucha/Notation/TranslationManual.tex
Fix typo and include tex source for the Translation Manual

\documentclass[english,ngerman,paper=a4,]{scrartcl}%
\input{Praeambel}%
%
\usepackage%
%[hidelinks]%
{hyperref} %%%% HYPERREF SOLLTE BIS AUF AUSNAHMEN IMMER DAS LETZTE PAKET SEIN %%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}%
%
%
%
%
%
%
\renewcommand{\baselinestretch}{1.50}\normalsize%
%\tableofcontents\newpage%
%
%%%%%%%%%%%%%%%%%%%%%

\makeatletter
% Remove \@date and spacing following it from \@maketitle
\patchcmd{\@maketitle}% <cmd>
  {{\usekomafont{date}{\@date \par}}%
    \vskip \z@ \@plus 1em}% <search>
  {}% <replace>
  {}{}% <success><failure>
\makeatother
%
\title{Translation Manual for Salamuchas Formalization of the Ex Motu Argument}
\maketitle
\section{Logical Notation}
%%
%

\begin{tabular}{|c|c|c|c|}
\hline 
\textbf{Name of Operator} & \textbf{Samalamucha first variant} & \textbf{Second variant} & \textbf{\enquote{Standard notation}} \\ 
\hline 
Conjunction & $p . q$ & $p \wedge q$ & $p \wedge q$ \\ 
\hline 
Disjunction & $p \vee q$ & $p \vee q$  & $p \vee q$ \\ 
\hline 
Negation & $\sim p$ & $\sim p$ & $\neg p$ \\ 
\hline 
(Material) Implication & $p \supset q$ & $p \rightarrow q$ & $p \rightarrow q$ \\ 
\hline 
Biconditional & $p \equiv q$ & $p \equiv q$ & $p \leftrightarrow q$ \\ 
\hline 
Universal Quantifier & $[x]. \phi (x)$ & $\wedge x \phi (x)$ & $\forall x. \phi (x)$ \\ 
\hline 
Existential Quantifier & $[\exists x]. \phi (x)$& $\vee x \phi (x)$ & $\exists x. \phi (x)$ \\ 
\hline 
\end{tabular}

\section{Definitions of relations, predicates, etc.}
\begin{tabular}{|c|c|c|}
\hline 
\textbf{Notation} & \textbf{Name/Description} & \textbf{Definition} \\ 
\hline 
$x R y$ & Relation (at first!) & the usual \\ 
\hline 
$C'R$& Set of all elements of a relation & $x \in C'R \Leftrightarrow \exists t. (t R x \vee x R t)$ \\ 
\hline 
$K(R)$ & Ordering relation & transitive, irreflexive, connected relation \\ 
\hline 
$f x$ & x is in motion &  - \\ 
\hline 
$x R y$ & x moves y&  - \\ 
\hline 
$M_{x}(a)$ or $a M x$ & a is the proper part of x [sic] & -  \\ 
\hline
$xA_{s}y$ & x is in aspect S \emph{in actu} to y [emph. orig.] & -  \\ 
\hline 
$C x$ & x is a body & -  \\ 
\hline 
 $t_{i}Fx$ $t_i$ & $t_i$ is the duration of movement of x [sic] & -  \\ 
\hline 
 $F(t_i)$ or $Ht_i$ & $t_i$ is the finite period of time [sic]& -  \\ 
\hline 
\end{tabular}
%
%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{jurabib}
\bibliography{/home/knork/Dokumente/Latex/Philosophie.bib}

\end{document}

### 2017-05-31 Formalizations/Isabelle/Salamucha/Notation/TranslationManual.tex
Fix typo and add tex source to translation manual
Refactor Tex Source

@@
-\documentclass[english,ngerman,paper=a4,]{scrartcl}%
-\input{Praeambel}%
-%
-\usepackage%
-%[hidelinks]%
-{hyperref} %%%% HYPERREF SOLLTE BIS AUF AUSNAHMEN IMMER DAS LETZTE PAKET SEIN %%%%
+\documentclass[english,paper=a4,]{scrartcl}%
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage{kpfonts}
+\usepackage{microtype}
+\usepackage{babel}
+\usepackage{geometry}%Alternativ typearea (KOMA)
+\geometry{lmargin=2cm,tmargin=2cm,rmargin=1.5cm,bmargin=2cm}
+\usepackage{marvosym}
+\usepackage{tabularx}% Tabellen bestimmter Breite
+\usepackage{csquotes}
+\usepackage{amsmath}
+\usepackage{xspace}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}%
-%
-%
-%
-%
-%
-%
-\renewcommand{\baselinestretch}{1.50}\normalsize%
-%\tableofcontents\newpage%
-%
-%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\renewcommand{\baselinestretch}{1.50}\normalsize%%
 
@@
 %
-\title{Translation Manual for Salamuchas Formalization of the Ex Motu Argument}
+\title{Translation Manual for Salamucha's Formalization of the \emph{Ex Motu} Argument}
+%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%
 \maketitle
@@
 \hline 
-\textbf{Name of Operator} & \textbf{Samalamucha first variant} & \textbf{Second variant} & \textbf{\enquote{Standard notation}} \\ 
+\textbf{Name of Operator} & \textbf{Samalamucha's first version} & \textbf{Second version} & \textbf{\enquote{Standard notation}} \\ 
 \hline 
@@
 \hline 
- $t_{i}Fx$ $t_i$ & $t_i$ is the duration of movement of x [sic] & -  \\ 
+ $t_{i}Fx$  & $t_i$ is the duration of movement of x [sic] & -  \\ 
 \hline 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{jurabib}
-\bibliography{/home/knork/Dokumente/Latex/Philosophie.bib}
-
 \end{document}

### 2017-08-16 Papers/2013/ArXiv/GodProof-Abstract.tex
Update arXiv abstract

--- 2013-09-10 Abstract/ArXiv/GodProof-Abstract.tex
@@
 \maketitle
+
+\bigskip
+
+\noindent
+\textbf{Update (16/08/2017):} The abstract below, uploaded to arXiv on 21/08/2013, was the first communication of the computer-assisted formalization of G\"odel's ontological proof.
+Since then, the following longer papers have been published: \cite{B39,B40,B45,B47,B49,B52,B56,B57,B60,B61,B62,B63,B64,B65,B66}.
+
+\bigskip
+\bigskip
 
@@
 \bibliographystyle{plain}
-%\bibliography{Bibliography}
-\begin{thebibliography}{10}
+\bibliography{Bibliography}
+% \begin{thebibliography}{10}
 
-\bibitem{Adams}
-R.M. Adams.
-\newblock Introductory note to *1970.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1995.
+% \bibitem{Adams}
+% R.M. Adams.
+% \newblock Introductory note to *1970.
+% \newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+%   Letters}}. Oxford University Press, 1995.
 
-\bibitem{AndersonGettings}
-A.C. Anderson and M.~Gettings.
-\newblock G\"odel ontological proof revisited.
-\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
-  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
+% \bibitem{AndersonGettings}
+% A.C. Anderson and M.~Gettings.
+% \newblock G\"odel ontological proof revisited.
+% \newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
+%   Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.
 
-\bibitem{B9}
-C.~Benzm{\"u}ller and L.C. Paulson.
-\newblock Exploring properties of normal multimodal logics in simple type
-  theory with {LEO-II}.
-\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
-  Birthday}}, pages 386--406. College Publications.
+% \bibitem{B9}
+% C.~Benzm{\"u}ller and L.C. Paulson.
+% \newblock Exploring properties of normal multimodal logics in simple type
+%   theory with {LEO-II}.
+% \newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
+%   Birthday}}, pages 386--406. College Publications.
 
-\bibitem{J23}
-C.~Benzm{\"u}ller and L.C. Paulson.
-\newblock Quantified multimodal logics in simple type theory.
-\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
-  7(1):7--20, 2013.
+% \bibitem{J23}
+% C.~Benzm{\"u}ller and L.C. Paulson.
+% \newblock Quantified multimodal logics in simple type theory.
+% \newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
+%   7(1):7--20, 2013.
 
-\bibitem{LEO-II}
-C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
-\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
-  logic.
-\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
-  162--170. Springer, 2008.
+% \bibitem{LEO-II}
+% C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
+% \newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
+%   logic.
+% \newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
+%   162--170. Springer, 2008.
 
-\bibitem{Coq}
-Y.~Bertot and P.~Casteran.
-\newblock {\em {Interactive Theorem Proving and Program Development}}.
-\newblock Springer, 2004.
+% \bibitem{Coq}
+% Y.~Bertot and P.~Casteran.
+% \newblock {\em {Interactive Theorem Proving and Program Development}}.
+% \newblock Springer, 2004.
 
-\bibitem{Sledgehammer}
-J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
-\newblock Extending {Sledgehammer} with {SMT} solvers.
-\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
+% \bibitem{Sledgehammer}
+% J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+% \newblock Extending {Sledgehammer} with {SMT} solvers.
+% \newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
 
-\bibitem{Nitpick}
-J.C. Blanchette and T.~Nipkow.
-\newblock Nitpick: A counterexample generator for higher-order logic based on a
-  relational model finder.
-\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
-  Springer, 2010.
+% \bibitem{Nitpick}
+% J.C. Blanchette and T.~Nipkow.
+% \newblock Nitpick: A counterexample generator for higher-order logic based on a
+%   relational model finder.
+% \newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+%   Springer, 2010.
 
-\bibitem{Satallax}
-C.E. Brown.
-\newblock Satallax: An automated higher-order prover.
-\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
-  Springer, 2012.
+% \bibitem{Satallax}
+% C.E. Brown.
+% \newblock Satallax: An automated higher-order prover.
+% \newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+%   Springer, 2012.
 
-\bibitem{ContemporaryBibliography}
-R.~Corazzon.
-\newblock Contemporary bibligraphy on the ontological proof
-  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
+% \bibitem{ContemporaryBibliography}
+% R.~Corazzon.
+% \newblock Contemporary bibligraphy on the ontological proof
+%   (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
 
-\bibitem{Fitting}
-M.~Fitting.
-\newblock {\em {Types, Tableaux and G\"odel's God}}.
-\newblock Kluver Academic Press, 2002.
+% \bibitem{Fitting}
+% M.~Fitting.
+% \newblock {\em {Types, Tableaux and G\"odel's God}}.
+% \newblock Kluver Academic Press, 2002.
 
-\bibitem{Goedel1970}
-K.~G\"odel.
-\newblock Ontological proof.
-\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
-  Letters}}. Oxford University Press, 1970.
+% \bibitem{Goedel1970}
+% K.~G\"odel.
+% \newblock Ontological proof.
+% \newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+%   Letters}}. Oxford University Press, 1970.
 
-\bibitem{GoedelNotes}
-K.~G\"odel.
-\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{GoedelNotes}
+% K.~G\"odel.
+% \newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{Hazen}
-A.P. Hazen.
-\newblock On g\"odel's ontological proof.
-\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+% \bibitem{Hazen}
+% A.P. Hazen.
+% \newblock On g\"odel's ontological proof.
+% \newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
 
-\bibitem{Hurd03first-orderproof}
-J.~Hurd.
-\newblock First-order proof tactics in higher-order logic theorem provers.
-\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
-  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
+% \bibitem{Hurd03first-orderproof}
+% J.~Hurd.
+% \newblock First-order proof tactics in higher-order logic theorem provers.
+% \newblock In {\em Design and Application of Strategies/Tactics in Higher Order
+%   Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.
 
-\bibitem{Isabelle}
-T.~Nipkow, L.C. Paulson, and M.~Wenzel.
-\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
-\newblock Number 2283 in LNCS. Springer, 2002.
+% \bibitem{Isabelle}
+% T.~Nipkow, L.C. Paulson, and M.~Wenzel.
+% \newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
+% \newblock Number 2283 in LNCS. Springer, 2002.
 
-\bibitem{FormalTheologyRepository}
-B.~Woltzenlogel Paleo and C.~Benzm\"uller.
-\newblock Formal theology repository
-  (\url{https://github.com/FormalTheology/GoedelGod}).
+% \bibitem{FormalTheologyRepository}
+% B.~Woltzenlogel Paleo and C.~Benzm\"uller.
+% \newblock Formal theology repository
+%   (\url{https://github.com/FormalTheology/GoedelGod}).
 
-\bibitem{ScottNotes}
-D.~Scott.
-\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
-\newblock In  \cite{sobel2004logic}, 2004.
+% \bibitem{ScottNotes}
+% D.~Scott.
+% \newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
+% \newblock In  \cite{sobel2004logic}, 2004.
 
-\bibitem{sobel2004logic}
-J.H. Sobel.
-\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
-\newblock Cambridge U. Press, 2004.
+% \bibitem{sobel2004logic}
+% J.H. Sobel.
+% \newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
+% \newblock Cambridge U. Press, 2004.
 
-\bibitem{J22}
-G.~Sutcliffe and C.~Benzm{\"u}ller.
-\newblock Automated reasoning in higher-order logic using the {TPTP THF}
-  infrastructure.
-\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
+% \bibitem{J22}
+% G.~Sutcliffe and C.~Benzm{\"u}ller.
+% \newblock Automated reasoning in higher-order logic using the {TPTP THF}
+%   infrastructure.
+% \newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
 
-\end{thebibliography}
+% \end{thebibliography}
 

### 2017-09-01 Papers/2013/ArXiv/GodProof-Abstract.tex
Upload Christoph's Changes

@@
 \newcommand{\NE}{\mathit{NE}}
+
+\usepackage[
+    firstinits=true,
+     backend=biber, 
+     natbib=true,
+     isbn=false,
+     doi=true,
+     url=false,
+    style=numeric
+]{biblatex}
+\addbibresource[label=bibs]{Bibliography.bib}
+\addbibresource[label=bibs]{chris.bib}
 
@@
 \noindent
-\textbf{Update (16/08/2017):} The abstract below, uploaded to arXiv on 21/08/2013, was the first communication of the computer-assisted formalization of G\"odel's ontological proof.
-Since then, the following longer papers have been published: \cite{B39,B40,B45,B47,B49,B52,B56,B57,B60,B61,B62,B63,B64,B65,B66}.
+\textbf{Update (31/08/2017):} The abstract below, uploaded to arXiv on
+21/08/2013, was the first communication of the computer-assisted
+formalization of G\"odel's ontological proof. Since then, the
+following longer papers have been published:
+%\cite{B39,B40,B45,B47,B49,B52,B56,B57,B60,B61,B62,B63,B64,B65,B66}.
+\cite{W50,J28,B15,C39,C40,C41,C42,C44,C46,C52,W55,J32,C55,C60,B16,B66,,J35,J36,C65}.
 
@@
 
+\small
+
+\printbibliography
+
 %\footnotesize
-\bibliographystyle{plain}
-\bibliography{Bibliography}
+%\bibliographystyle{plain}
+%\bibliography{Bibliography,chris}
 % \begin{thebibliography}{10}

### 2017-09-01 Papers/2013/ArXiv/GodProof-Abstract.tex
Fix bibliography

@@
 %\cite{B39,B40,B45,B47,B49,B52,B56,B57,B60,B61,B62,B63,B64,B65,B66}.
-\cite{W50,J28,B15,C39,C40,C41,C42,C44,C46,C52,W55,J32,C55,C60,B16,B66,,J35,J36,C65}.
+\cite{W50,J28,B15,C39,C40,C41,C42,C44,C46,C52,W55,J32,C55,C60,B16,B66,J35,J36,C65}.
 
@@
 with an unprecedent degree of detail 
-and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository}. 
+and formality with the help of theorem provers; cf.~\cite{FormalTheologyRepository,ComputationalPhilosophyRepository}. 
 The following has been done (and in this order):

### 2017-09-03 Papers/2013/ArXiv/GodProof-Abstract.tex
bibtex instead of biber

@@
 
-\usepackage[
-    firstinits=true,
-     backend=biber, 
-     natbib=true,
-     isbn=false,
-     doi=true,
-     url=false,
-    style=numeric
-]{biblatex}
-\addbibresource[label=bibs]{Bibliography.bib}
-\addbibresource[label=bibs]{chris.bib}
+% \usepackage[
+%     firstinits=true,
+%      backend=bibtex, 
+%      natbib=true,
+%      isbn=false,
+%      doi=true,
+%      url=false,
+%     style=numeric
+% ]{biblatex}
+% \addbibresource[label=bibs]{Bibliography.bib}
+% \addbibresource[label=bibs]{chris.bib}
 
@@
 
-\small
+\footnotesize
 
-\printbibliography
+\begin{thebibliography}{10}
+
+\bibitem{Adams}
+R.M. Adams.
+\newblock Introductory note to *1970.
+\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+  Letters}}. Oxford University Press, 1995.
+
+\bibitem{AndersonGettings}
+A.C. Anderson and M.~Gettings.
+\newblock G\"odel ontological proof revisited.
+\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
+  Science, and Physics: Lecture Notes in Logic 6}}, pages 167--172. {Springer},
+  1996.
+
+\bibitem{B16}
+Matthias Bentert, Christoph Benzm{\"u}ller, David Streit, and Bruno
+  Woltzenlogel~Paleo.
+\newblock Analysis of an ontological proof proposed by {Leibniz}.
+\newblock In Charles Tandy, editor, {\em Death and Anti-Death, Volume 14: Four
+  Decades after Michael Polanyi, Three Centuries after G.W. Leibniz}. Ria
+  University Press, 2016.
+
+\bibitem{B9}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Exploring properties of normal multimodal logics in simple type
+  theory with {LEO-II}.
+\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
+  Birthday}}, pages 386--406. College Publications, 2008.
+
+\bibitem{J23}
+C.~Benzm{\"u}ller and L.C. Paulson.
+\newblock Quantified multimodal logics in simple type theory.
+\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
+  7(1):7--20, 2013.
+
+\bibitem{LEO-II}
+C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
+\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
+  logic.
+\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
+  162--170. Springer, 2008.
+
+\bibitem{C42}
+Christoph Benzm{\"u}ller.
+\newblock {G\"{o}del's} ontological argument revisited -- findings from a
+  computer-supported analysis (invited).
+\newblock In Ricardo~Souza Silvestre and Jean-Yves B\'eziau, editors, {\em
+  Handbook of the 1st World Congress on Logic and Religion, Jo\~ao Pessoa,
+  Brazil}, page~13, 2015.
+\newblock (Invited abstract).
+
+\bibitem{C41}
+Christoph Benzm{\"u}ller, Leon Weber, and Bruno Woltzenlogel~Paleo.
+\newblock Computer-assisted analysis of the {Anderson-H\'{a}jek} ontological
+  controversy.
+\newblock In Ricardo~Souza Silvestre and Jean-Yves B\'eziau, editors, {\em
+  Handbook of the 1st World Congress on Logic and Religion, Joao Pessoa,
+  Brasil}, pages 53--54, 2015.
+\newblock (superseded by 2016 article in Logica Universalis).
+
+\bibitem{J32}
+Christoph Benzm{\"u}ller, Leon Weber, and Bruno Woltzenlogel~Paleo.
+\newblock Computer-assisted analysis of the {Anderson-H\'{a}jek} controversy.
+\newblock {\em Logica Universalis}, 11(1):139--151, 2017.
+
+\bibitem{J28}
+Christoph Benzm\"uller and Bruno Woltzenlogel~Paleo.
+\newblock {G{\"o}del's God in Isabelle/HOL}.
+\newblock {\em Archive of Formal Proofs}, 2013.
+\newblock (Formally verified).
+
+\bibitem{W50}
+Christoph Benzm\"uller and Bruno Woltzenlogel~Paleo.
+\newblock {G\"odel's} {God} on the computer.
+\newblock In S.~Schulz, G.~Sutcliffe, and B.~Konev, editors, {\em Proceedings
+  of the 10th International Workshop on the Implementation of Logics}, 2013.
+\newblock (Invited paper).
+
+\bibitem{C40}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock Automating {G\"{o}del's} ontological proof of {God}'s existence with
+  higher-order automated theorem provers.
+\newblock In Torsten Schaub, Gerhard Friedrich, and Barry O'Sullivan, editors,
+  {\em ECAI 2014}, volume 263 of {\em Frontiers in Artificial Intelligence and
+  Applications}, pages 93 -- 98. IOS Press, 2014.
+
+\bibitem{C39}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock {G\"odel's} proof of {God's} existence.
+\newblock In Jean-Yves Beziau and Katarzyna Gan-Krzywoszynska, editors, {\em
+  Handbook of the World Congress on the Square of Opposition IV}, pages 22--23,
+  2014.
+\newblock (superseded by ECAI-2014 paper).
+
+\bibitem{C52}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock Experiments in computational metaphysics: {G\"odel's} proof of god's
+  existence.
+\newblock In Subhash~C. Mishram, Ramgopal Uppaluri, and Varun Agarwal, editors,
+  {\em Science \& Spiritual Quest, Proceedings of the 9th All India Students'
+  Conference, 30th October -- 1 November, 2015, IIT Kharagpur, India}, pages
+  23--40. Bhaktivedanta Institute, Kolkata, \url{www.binstitute.org}, 2015.
+\newblock (Invited paper, superseded by article in Savijnanam).
+
+\bibitem{C46}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock Higher-order modal logics: Automation and applications.
+\newblock In Adrian Paschke and Wolfgang Faber, editors, {\em Reasoning Web.
+  Web Logic Rules - 11th International Summer School 2015, Berlin, Germany,
+  July 31 - August 4, 2015, Tutorial Lectures}, number 9203 in LNCS, pages
+  32--74, Berlin, Germany, 2015. Springer.
+
+\bibitem{C44}
+Christoph Benzm{\"{u}}ller and Bruno Woltzenlogel~Paleo.
+\newblock Interacting with modal logics in the {Coq} proof assistant.
+\newblock In Lev~D. Beklemishev and Daniil~V. Musatov, editors, {\em Computer
+  Science - Theory and Applications - 10th International Computer Science
+  Symposium in Russia, {CSR} 2015, Listvyanka, Russia, July 13-17, 2015,
+  Proceedings}, volume 9139 of {\em LNCS}, pages 398--411. Springer, 2015.
+
+\bibitem{W55}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock On logic embeddings and {G\"odel's} {God}.
+\newblock In Mihai Codescu, Razvan Diaconescu, and Ionut Tutu, editors, {\em
+  Recent Trends in Algebraic Development Techniques: 22nd International
+  Workshop, WADT 2014, Sinaia, Romania, September 4-7, 2014, Revised Selected
+  Papers}, number 9563 in LNCS, pages 3--6, Sinaia, Romania, 2015. Springer.
+\newblock (Invited paper).
+
+\bibitem{C55}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock The inconsistency in {G{\"o}del's} ontological argument: A success
+  story for {AI} in metaphysics.
+\newblock In Subbarao Kambhampati, editor, {\em IJCAI 2016}, volume 1-3, pages
+  936--942. AAAI Press, 2016.
+
+\bibitem{B15}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock The modal collapse as a collapse of the modal square of opposition.
+\newblock In Jean-Yves B\'{e}ziau and Gianfranco Basti, editors, {\em The
+  Square of Opposition: A Cornerstone of Thought (Collection of papers related
+  to the World Congress on the Square of Opposition IV, Vatican, 2014),
+  \url{http://www.springer.com/us/book/9783319450612}}, Studies in Universal
+  Logic. Springer International Publishing Switzerland, 2016.
+
+\bibitem{C60}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock An object-logic explanation for the inconsistency in {G\"odel's}
+  ontological theory (extended abstract, sister conferences).
+\newblock In Malte Helmert and Franz Wotawa, editors, {\em KI 2016: Advances in
+  Artificial Intelligence, Proceedings}, LNCS, pages 244--250, Berlin, Germany,
+  2016. Springer.
+
+\bibitem{J36}
+Christoph Benzm{\"u}ller and Bruno Woltzenlogel~Paleo.
+\newblock Experiments in {Computational Metaphysics}: {G\"odel's} proof of
+  {God's} existence.
+\newblock {\em Savijnanam: scientific exploration for a spiritual paradigm.
+  Journal of the Bhaktivedanta Institute}, 9:43--57, 2017.
+
+\bibitem{Coq}
+Y.~Bertot and P.~Casteran.
+\newblock {\em {Interactive Theorem Proving and Program Development}}.
+\newblock Springer, 2004.
+
+\bibitem{Sledgehammer}
+J.C. Blanchette, S.~B\"ohme, and L.C. Paulson.
+\newblock Extending {Sledgehammer} with {SMT} solvers.
+\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.
+
+\bibitem{Nitpick}
+J.C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for higher-order logic based on a
+  relational model finder.
+\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pages 131--146.
+  Springer, 2010.
+
+\bibitem{Satallax}
+C.E. Brown.
+\newblock Satallax: An automated higher-order prover.
+\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
+  Springer, 2012.
+
+\bibitem{ContemporaryBibliography}
+R.~Corazzon.
+\newblock Contemporary bibligraphy on the ontological proof
+  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).
+
+\bibitem{Fitting}
+M.~Fitting.
+\newblock {\em {Types, Tableaux and G\"odel's God}}.
+\newblock Kluver Academic Press, 2002.
+
+\bibitem{C65}
+David Fuenmayor and Christoph Benzm{\"u}ller.
+\newblock Automating emendations of the ontological argument in intensional
+  higher-order modal logic.
+\newblock In {\em KI 2017: Advances in Artificial Intelligence 40th Annual
+  German Conference on AI}, LNAI. Springer, 2017.
+
+\bibitem{J35}
+David Fuenmayor and Christoph Benzm\"uller.
+\newblock {Types, Tableaus and G{\"o}del's God in Isabelle/HOL}.
+\newblock {\em Archive of Formal Proofs}, 2017.
+\newblock Formally verified with Isabelle/HOL.
+
+\bibitem{Goedel1970}
+K.~G\"odel.
+\newblock Ontological proof.
+\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
+  Letters}}. Oxford University Press, 1970.
+
+\bibitem{GoedelNotes}
+K.~G\"odel.
+\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{Hazen}
+A.P. Hazen.
+\newblock On g\"odel's ontological proof.
+\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.
+
+\bibitem{Hurd03first-orderproof}
+J.~Hurd.
+\newblock First-order proof tactics in higher-order logic theorem provers.
+\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
+  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, pages 56--68, 2003.
+
+\bibitem{Isabelle}
+T.~Nipkow, L.C. Paulson, and M.~Wenzel.
+\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
+\newblock Number 2283 in LNCS. Springer, 2002.
+
+\bibitem{ScottNotes}
+D.~Scott.
+\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
+\newblock In  \cite{sobel2004logic}, 2004.
+
+\bibitem{sobel2004logic}
+J.H. Sobel.
+\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
+\newblock Cambridge U. Press, 2004.
+
+\bibitem{J22}
+G.~Sutcliffe and C.~Benzm{\"u}ller.
+\newblock Automated reasoning in higher-order logic using the {TPTP THF}
+  infrastructure.
+\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.
+
+\bibitem{B66}
+B.~Woltzenlogel~Paleo.
+\newblock Leibniz's characteristica universalis and calculus ratiocinator
+  today.
+\newblock In Charles Tandy, editor, {\em Death and Anti-Death, Volume 14: Four
+  Decades after Michael Polanyi, Three Centuries after G. W. Leibniz}. Ria
+  University Press, 2016.
+
+\bibitem{ComputationalPhilosophyRepository}
+B.~Woltzenlogel~Paleo and C.~Benzm\"uller.
+\newblock Computational philosophy repository
+  (\url{https://gitlab.com/aossie/ComputationalPhilosophy/}).
+
+\bibitem{FormalTheologyRepository}
+B.~Woltzenlogel~Paleo and C.~Benzm\"uller.
+\newblock Formal theology repository
+  (\url{https://github.com/FormalTheology/GoedelGod}).
+
+\end{thebibliography}
+
+%\bibliographystyle{plain}
+%\bibliography{Bibliography,chris}
+
+%\printbibliography
 
